### 2012-10-17 paper.Rnw
reorg as a knitr project

\documentclass{article}
\usepackage[margin=1.5in]{geometry}
\usepackage{algorithm}
\usepackage{color}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{url}

\newcommand{\todo}[1]{
    \addcontentsline{tdo}{todo}{\protect{#1}}
    \textcolor{red}{#1}
}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Hippocampal segmentation with MAGeT}
\author{Jon Pipitone, }
\maketitle

\section{Introduction}

The hippocampus is of particular interest to many researchers because it is implicated in forms of brain dysfunction such as Alzheimer's disease and schizophrenia, and has functional significance in cognitive processes such as learning and memory.  For many research questions involving neuroimaging data, accurate identification of the hippocampus and its subregions in participant MR images is a necessary first step in order to analyse neuroanatomical differences and changes in  sample populations.  \todo{expand}

Currently, the gold standard for neuroanatomical segmentation is manual labelling by an expert human rater.  This is problematic for several reasons.  Practically, manual segmentation takes a significant investment of time and expertise (Hammers2003) which may not be readily available to researchers or clinicians.  <blah> estimates that expert segmentation can take anywhere between 45 minutes to several days per image depending on the image resolution and protocol.  The amount of data produced in neuroimaging experiments increasingly exceeds the capacity for identification of specific neuroanatomical structures by an expert manual rater.  It may be difficult for laboratories to find or train, someone with the expertise to do such segmentation.

Compounding this is that the hippocampus is a problematic structure for segmentation. Its very definition is disputed and varies widely in the literature (Geuze2004).  This has led to a variety of competing segmentation protocols in the literature which researchers can adhere to, including necessary ad-hoc modifications to adapt the definition in order to answer a particular research question.  Efforts are underway to create a unified protocol (Jack2011) for volumetry applications. But, even with a unified protocol, in order to investigate certain research questions it may be necessary to select from different hippocampal definitions. For example, (Poppenk2011) found that overall hippocampal volume change did not predict recollection memory performance, but dividing the hippocampus into anterior and posterior regions did.

\todo{this bit needs to be a little clearer - was it just the division of the HC that led to the improved understanding or was it that spatially localized differences in volumes are a better way to describe the difference).}

\todo{- Also, other situations: developmental studies -- need different definitions depending on the age range of subjects
-- additionally, may want to tweak definition to test certain hypotheses (e.g. include subiculum: yes/no, more or less 
 conservative WM/GM boundaries, compare with definitions from a different study,, etc...).}

One can also expect neuroanatomical variability in this structure as it loses volume through aging and in the time course of different diseases.  \todo{This bit isn't really "joined to the previous statement"  may need a little glue here}

Automated segmentation techniques are available and can minimize several of the confounds often observed in studies that involve manual volumetry, such as the time and expertise required for segmentation and intra- and inter-rater reliability.  A class of segmentation approaches, known as atlas-based segmentation (also, template-based segmentation), make use of expertly labeled neuroanatomical atlases. An atlas MR image is warped to fit the subject neuroanatomy using nonlinear registration techniques [4, 11], and then the atlas labels are propagated via the registration transformation from the atlas to match the unique neuroanatomy of a particular subject.  Segmentation accuracy is affected by nonlinear registration and resampling errors, and by the extent to which the atlas neuroanatomy differs from the underlying neuroanatomical variability of the population under study (citation needed).  To address these limitations, several atlases can be used (multi-atlas segmentation). Labels from each atlas are propagated to the subjects, by way of an nonlinear image registration, producing as many labellings for each subject as atlases. A label fusion technique is employed to merge the candidate labellings for a subject into a single segmentation using a decision rule (for instance, voxel-wise voting). Not all of the atlases need be used to segment each subject. Selecting only those atlases that are most similar to the subject image has been shown to significantly improve overall registration and segmentation accuracy (Aljabar 2008, Collins2010, Leung2010, Lotjonen2010).

\todo{briefly discuss other automated segmentation methods, and justify our focus on atlas-based methods}

In summary, whilst manual segmentation is the gold standard for neuroanatomical segmentation, it is time-consuming and requires an investment in expertise and training, and may be inflexible in the face of differing hippocampal definitions or changes. Atlas-based automated segmentation dramatically reduces the time needed for segmentation, and can be completed without any expert rater as long as an appropriate atlas library is available <be more specific (e.g. shares protocol, demographics, phenotypes of subject population)>.  In the situation where such an atlas library does not exist, then an expert rater is needed. Many automated approaches rely on an atlas library containing between 30 and 80 expertly labeled brains (Heckemann2006, Collins2010) which, even though producing such a library may require less manual work than segmenting the entire set of subjects, it still poses a substantial time commitment.

In this paper we extend our automated method for subcortical segmentation (striatum, globus pallidus, and thalamus; Chakravarty2012) to hippocampal segmentation. This method, known as MAGeT brain, is suited for situations where manual effort is at a premium, and flexibility is needed in atlas definitions.  Our strategy is to bootstrap the familiar multi-atlas-based segmentation process from only a very small number of atlases.  In our previous work we used only a single atlas, based on serial histological slices, and validated the algorithm on 20 adolescent MR images.  In this work, we extensively investigate MAGeT used with differently sized atlas and template libraries, registration algorithms, label fusion strategies, and hippocampal definitions.  We evaluate our algorithm on a subset of images from the ADNI dataset, and compare our segmentations to the manual segmentations provided by ADNI.  Additionally, we evaluate the segmentations produced with MAGeT using hippocampal definitions from five high-resolution T1 in-vivo atlases (Winterburn2012). <conclude with summary that this is detailed validation the algorithm in order to pick the best operating parameters, whereas our previous work showed the feasibility of the approach>

The essential insight of MAGeT brain, generating a template library, is not new. Heckemann2006 compared a multi-atlas-based segmentation method to that of generating a template library from a single atlas and found poor performance so did not investigate this approach any further. The LEAP algorithm (Wolz2010) begins with 30 atlases and labels the most similar unlabeled images. The labeled images from this step are included in the atlas set for the following step, and this proceeds until all images have been labeled. <expand; include other multi-atlas methods too? >

In this work we conduct a thorough validation of the MAGeT brain approach. Our aim is not to improve on segmentation accuracy beyond existing methods, but instead to provide a method that trades off manual segmentation expertise for computational processing time whilst providing useful accuracy for the research and clinical setting. <blah blah.. this sucks>

\section{Materials and Methods}
explain two registration methods, definitions, fusion methods. 
compare to each other, and then compare atlas-based, true multi-atlas, naive methods.
trying to make choices as to the best reg/fusion/

evaluations
described previously in Chakravarty2012.

\subsection{Segmentation Algorithm}
The segmentation approach we follow is an extension of traditional multi-atlas segmentation. In multi-atlas segmentation, the inputs are a template library consisting of a set of manually labelled MR images (atlases), and a set of unlabelled MR images to be labelled.  Each unlabelled MR image is segmented by first propagating to it the labels from the template library images via image registration, and then these candidate labels are combined by a label fusion method (e.g. voxel-wise majority voting). The label fusion method used may determine which templates are used to segment a particular subject (for instance, voxel-wise majority vote takes labels from all templates, whereas a weighted majority vote selects only those templates most similar to the subject).

Our extension to this approach is to add a preliminary stage in which we construct the template library using a much smaller set of atlases and a subset of the given unlabelled images.  To begin, we (possibly arbitrarily) select from the unlabelled images to form the template library.  Multiple labellings are given to each image in the template library via label propagation from each of the atlases (label fusion is not conducted).  Traditional multi-atlas segmentation is then used to produce segmentations for the entire set of unlabelled images (including those used in the template library).  Our algorithm is illustrated in pseudocode in Algorithm 1, and diagrammatically in Figure 1.

- Essentially, we add an extra layer of label propagation between the set of atlases 
- This allows the template library to better reflect the constitution of the images to be labelled.

Source code can be found at \url{http://github.com/pipitone/MAGeTbrain}.

\begin{algorithm}
\caption{Pseudocode for the MAGeT Brain algorithm}
\label{pseudocodesdf}
\begin{algorithmic}
\Function{MultiAtlas}{Templates, Subjects}
  \ForAll{$subject$}
    \ForAll{$template$}
      \State propagate all labels for template to subject space
      \State store subject labels
    \EndFor
    \State fuse subject labels
  \EndFor
\EndFunction
\\
\Function {MAGeTBrain}{Subjects, Atlases, n}
  \For{$i = 1 \to n$}
    \State choose a subject to be used as a template
    \State propagate labels from each atlas to template space
    \State store the template with all of its labels
  \EndFor
  \State MultiAtlas(Templates, Subjects)
\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{figure}[h]
  \centering
    \includegraphics[width=0.5\textwidth]{MAGeT-figure.png}
  \caption{Diagram of the MAGeT Brain algorithm}
\end{figure}

\subsection{Image Processing and Registration Methods}

Before images were registered, the N3 algorithm [16] is first used to minimize the intensity nonuniformity in each of the atlases and unlabeled subject images. In this work we investigated the performance of two registration methods.

\subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}

The ANIMAL algorithm carries out Image registration is in two phases. In the first, a 12-parameter linear transformation (3 translations, rotations, scales, shears) is estimated between images using an algorithm that maximizes the correlation between blurred MR intensities and gradient magnitude over the whole brain [3]. In the second phase, nonlinear registration is completed using the ANIMAL algorithm [4]: an iterative procedure that estimates a 3D deformation field between two MR images. At first, large deformations are estimated using blurred version of the input data. These larger deformations are then input to subsequent steps where the fit is refined by estimating smaller deformations on data blurred with a Gaussian kernel with a smaller FWHM. The final transformation is a set of local translations defined on a bed of equally spaced nodes that were estimated through the optimization of the correlation coefficient. 
\todo{should add the MNI website here; also we should add the parameters we used somewhere}

\subsubsection{Automatic Normalization Tools (ANTS)}

ANTs is a diffeomorphic registration algorithm which provides great flexibility over the choice of transformation model, objective function, and the consistency of the final transformation. The transformation is estimated in a hierarchical fashion where the MRI data is subsampled, allowing large deformations to be estimated and successively refined at later hierarchical stages (where the data is subsampled to a finer grid). The deformation field and the objective function are regularized with a Gaussian kernel at each level of the hierarchy. The ANTs algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the ANTS algorithm compatible with the MINC data format, mincANTS \url{https://github.com/vfonov/mincANTS} \todo{(same here we should add the parameters we used).}

\subsection{Label Fusion Methods}

Label fusion is term given to the process of combining the information from several candidate labellings for an MR image into a single labelling.  In this paper we explore the benefits of three different fusion methods.

\subsubsection{Voxel-wise Majority Vote}

Labels are propagated from all template library images to a subject.  Each output voxel is given the most frequent label at that voxel location amongst all candidate labellings.  Ties are broken arbitrarily.

\subsubsection{Cross-correlation Weighted Majority Vote}

Labels are propagated from only those template library images which most similar to the subject.  
A voxel-wise majority vote is carried out from the labels from the top n ranked template library 
images, where n is parameter set by the user.

Similarity between a subject and a template image is measured using normalized cross-correlation of intensity values, over a region of interest defined on each template, after the subject has been linearly registered to the template. The heuristic used for defining a meaningful region of interest for each template is to use the extent of all the propagated labellings from each atlas (after linear registration only) and then dilating this region by three voxels.

The utility to calculate the cross-correlation similarity measure is implemented as part of the ANIMAL toolkit. \todo{(ANIMAL or MINC)}

\subsubsection{Normalised Mutual Information Weighted Majority Vote}

This label fusion method is identical to the above process except a normalised mutual information score is used over the region of interest between a template and subject instead of cross-correlation.

The utility to calculate the normalised mutual information measure is implemented as part of the EZMinc package (based on ITK NMI routine).

\subsection{Goodness-of-fit}

Each segmentation was evaluated against the 'gold-standard' manual segmentation from the dataset using the Dice Kappa ($\kappa$) overlap metric:

\begin{equation*} 
\kappa=\frac{2a}{2a+b+c}
\end{equation*}

where $a$ is the number of voxels common to the candidate segmentation and the gold standard and $b+c$ is the sum of the voxels uniquely identified by either the automatically generated candidate segmentation or the gold-standard. \todo{I would do this after you define the experiments}

\subsection{Data and Experiments}

Three different experiments were conducted to evaluate the influence of registration method, atlas set size and resolution, template set size, and label fusion method and parameters.

\subsubsection{ADNI Validation Experiment}

In this experiment we performed repeated random sub-sampling cross-validation of the MAGeT algorithm with a pool of 69 images and labels from the ADNI dataset.  The images are T1-weighted \todo{<insert description of scan/machine types>}.  The hippocampal labels are provided as part of the ADNI dataset, and are produced \todo{<insert description of SNT labellings>}.  The parameters we varied, and the ranges over which we varied are listed in \todo{table}.  We performed 10 validations per parameter combination.  In each validation trial, a random sample of subjects were chosen from the pool without replacement as the atlas set, whereas the template set was chosen randomly with replacement.

\begin{table}
    \begin{tabular}{l|l}
        Parameter           & Values tested                                                      \\ \hline
        Number of Templates & 3 to 20                                                            \\ 
        Number of Atlases   & 3 to 9                                                             \\ 
        Registration Method & ANTS or ANIMAL                                                     \\ 
        Label Fusion Method & majority vote, cross-correlation weighted vote, NMI weighted vote. 
    \end{tabular}
\end{table}

We used the following parameters for ANTS, during registration: 
\begin{verbatim}
  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
   --number-of-affine-iterations 10000x10000x10000x10000x10000 
   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
   --use-Histogram-Matching --MI-option 32x16000
   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
   -o transformation.xfm
 \end{verbatim}
These settings were adapted from the "reasonable starting point" given in the ANTS manual. \todo{<rationale?>}

For the purposes of this work we used the regularization parameters optimized in Robbins et al.[15]. \todo{right? (but added a second layer).}

\todo{multi-atlas comparison, "naive" comparison}

\subsubsection{Winterburn High-resolution Hippocampal Atlas - ADNI Validation}

In this experiment we explored using the MAGeT brain algorithm with the Winterburn high-resolution atlases to segment 21 randomly chosen MR images from the ADNI dataset (seven each of healthy, MCI and AD subjects).  The Winterburn atlases are digital segmentations of the hippocampus in five  in-vivo 300u isotropic T1-weighted MR scans, and include subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.  Subjects in the Winterburn atlases range in age from 29-57 years (mean age of 37), and include two males and three females.

Since hippocampal segmentation protocols differ between the ADNI labels and Winterburn atlases, this poses a problem for direct similarity comparisons between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate the performance of MAGeT brain, we compared classification accuracy of subjects by diagnosis based on hippocampal volume using both the SMT labels and our produced labels.  \todo{description of validation -- rms {\tt validate} or t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}

\todo{For minctracc -- we registered to TAL}

\subsection{Results}
\subsubsection{A2A}

Kappa vs. Number of templates: 
Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))

<<a2a, cache=TRUE, fig = T, echo = F, dev='CairoPDF', warning=F, error=F>>=
library(ggplot2)
multiplot <- function(..., plotlist=NULL, cols) {
    require(grid)

    # Make a list from the ... arguments and plotlist
    plots <- c(list(...), plotlist)

    numPlots = length(plots)

    # Make the panel
    plotCols = cols                          # Number of columns of plots
    plotRows = ceiling(numPlots/plotCols) # Number of rows needed, calculated from # of cols

    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(plotRows, plotCols)))
    vplayout <- function(x, y)
        viewport(layout.pos.row = x, layout.pos.col = y)

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
        curRow = ceiling(i/plotCols)
        curCol = (i-1) %% plotCols + 1
        print(plots[[i]], vp = vplayout(curRow, curCol ))
    }

}

#####################################################
# TRACC
t_naive      = read.csv(file='data/a2a_tracc_naive.csv')
t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
t_maj        = subset(t_data, method == 'majvote')
t_xcorr      = subset(t_data, method == 'xcorr')
t_nmi        = subset(t_data, method == 'nmi')
t_mean_naive = mean(t_naive$k)
t_ylimits    = c(0.7,0.85)

t_multiatlas = read.csv(file='data/a2a_tracc_multiatlas_2012_08_14.csv')
t_ma_means   = aggregate( k ~ num_atlases , data=t_multiatlas, mean)
t_maj$k_minus_ma_mean   <- t_maj$k   - t_ma_means[ t_maj$num_atlases - 3, 2 ]
t_xcorr$k_minus_ma_mean <- t_xcorr$k - t_ma_means[ t_xcorr$num_atlases - 3, 2 ]
t_nmi$k_minus_ma_mean   <- t_nmi$k   - t_ma_means[ t_nmi$num_atlases - 3, 2 ]

# Majority Vote
# plots kappa by numbeer of templates used
#geom = c('smooth'), line=1.5, formula=y~poly(x,2), method='lm') + 
t_maj_plot <- qplot(x=num_templates, y=k, data=t_maj, colour=as.factor(num_atlases), 
  geom = c('smooth')) + 
geom_hline(aes(yintercept=t_mean_naive)) + 
geom_hline(aes(yintercept = k, colour = as.factor(num_atlases)), linetype="dotted", 
  data = t_ma_means ) + 
opts(title = "ANIMAL\nMajority vote") +
opts(legend.position = "None") + 
scale_colour_hue(name="Atlases") +
xlab( "Number of Templates" ) + 
ylab( "Kappa" ) + 
coord_cartesian(ylim=t_ylimits)

# X-Corr results with a full suite of templates
# plots kappa by top N templates choosen 
t_xcorr_plot <- qplot(x=top_n, y=k, data=t_xcorr,
colour=as.factor(num_atlases), geom=c('smooth')) + 
geom_hline(aes(yintercept=t_mean_naive))  +
coord_cartesian(ylim=t_ylimits) + 
opts(title = "X-Corr vote") +
opts(legend.position = "None") + 
scale_colour_hue(name="Atlases") +
xlab( "Number of Templates" ) + 
ylab( "Kappa" ) + 
coord_cartesian(ylim=t_ylimits)

# NMI results with a full suite of templates
# plots kappa by top N templates choosen 
t_nmi_plot <- qplot(x=top_n, y=k, data=t_nmi,
colour=as.factor(num_atlases), geom=c('smooth')) +
geom_hline(aes(yintercept=t_mean_naive))  +
opts(title = "NMI vote") + 
opts(legend.position = "None") + 
scale_colour_hue(name="Atlases") +
xlab( "Number of Templates" ) + 
ylab( "Kappa" ) + 
coord_cartesian(ylim=t_ylimits)

#####################################################
# ANTS

a_naive      = read.csv(file='data/a2a_ants_naive_2012_08_30.csv')
a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')
a_maj        = subset(a_data, method == 'majvote')
a_xcorr      = subset(a_data, method == 'xcorr')
a_nmi        = subset(a_data, method == 'nmi')
a_mean_naive = mean(a_naive$k)
a_ylimits    = c(0.7,0.85)

a_multiatlas = read.csv(file='data/a2a_ants_multiatlas_2012_08_14.csv')
a_ma_means   = aggregate( k ~ num_atlases , data=a_multiatlas, mean)
a_maj$k_minus_ma_mean   <- a_maj$k - a_ma_means[ a_maj$num_atlases - 3, 2 ]
a_xcorr$k_minus_ma_mean <- a_xcorr$k - a_ma_means[ a_xcorr$num_atlases - 3, 2 ]
a_nmi$k_minus_ma_mean   <- a_nmi$k   - a_ma_means[ a_nmi$num_atlases - 3, 2 ]

# Majority Vote
# plots kappa by numbeer of templates used

a_maj_plot <- qplot(x=num_templates, y=k, data=a_maj, colour=as.factor(num_atlases), 
  geom=c('smooth')) + 
geom_hline(aes(yintercept=a_mean_naive))  +
coord_cartesian(ylim=a_ylimits) + 
geom_hline(aes(yintercept = k, colour = as.factor(num_atlases)), linetype="dotted", 
  data = a_ma_means ) +
opts(title = "ANTS\nMajority vote")  +
opts(legend.position = "None") + 
scale_colour_hue(name="Atlases") +
xlab( "Number of Templates" ) + 
ylab( "Kappa" )

# X-Corr results with a full suite of templates
# plots kappa by top N templates choosen 
a_xcorr_plot <- qplot(x=top_n, y=k, data=a_xcorr,
colour=as.factor(num_atlases), geom=c('smooth')) + 
geom_hline(aes(yintercept=a_mean_naive))  +
opts(title = "X-Corr vote") +
opts(legend.direction = "horizontal") + 
opts(legend.position = "bottom") + 
coord_cartesian(ylim=a_ylimits)  +
scale_colour_hue(name="Atlases") +
xlab( "Number of Templates" ) + 
ylab( "Kappa" )

# NMI results with a full suite of templates
# plots kappa by top N templates choosen 
a_nmi_plot <- qplot(x=top_n, y=k, data=a_nmi,
colour=as.factor(num_atlases), geom=c('smooth')) +
geom_hline(aes(yintercept=a_mean_naive))  +
opts(title = "NMI vote") + 
opts(legend.position = "None") + 
coord_cartesian(ylim=a_ylimits) +
scale_colour_hue(name="Atlases") +
xlab( "Number of Templates" ) + 
ylab( "Kappa" )

#loess.control(trace.hat = c("approximate"), statistics = c("approximate"), surface = c("interpolate"))
pdf("a2a_kbytemplates.pdf", width=15, height=15)
multiplot(t_maj_plot  , a_maj_plot  ,
          t_xcorr_plot, a_xcorr_plot, 
          t_nmi_plot  , a_nmi_plot  , cols=2)
dev.off()

############################################
#  Kappa adjusted by Multi-atlas mean 
a_ylimits    = c(-0.03,0.04)
t_ylimits    = a_ylimits

t_maj_plot <- qplot(x=num_templates, y=k_minus_ma_mean, data=t_maj, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
opts(title = "ANIMAL\nMajority vote") + opts(legend.position = "None") + 
xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
coord_cartesian(ylim=t_ylimits)

t_xcorr_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=t_xcorr, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
opts(title = "X-Corr vote") + opts(legend.position = "None") + 
xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
coord_cartesian(ylim=t_ylimits)

t_nmi_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=t_nmi, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
opts(title = "NMI vote") + opts(legend.position = "None") + 
xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
coord_cartesian(ylim=t_ylimits)

a_maj_plot <- qplot(x=num_templates, y=k_minus_ma_mean, data=a_maj, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
opts(title = "ANTS\nMajority vote") + opts(legend.position = "None") + 
xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
coord_cartesian(ylim=a_ylimits)

a_xcorr_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=a_xcorr, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
opts(title = "X-Corr vote") + opts(legend.position = "None") + 
xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
coord_cartesian(ylim=a_ylimits)

a_nmi_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=a_nmi, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
opts(title = "NMI vote") + opts(legend.position = "None") + 
xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
coord_cartesian(ylim=a_ylimits)

#print(a_nmi_plot)
#pdf("a2a_kminusMAmean.pdf", width=15, height=15)
multiplot(t_maj_plot  , a_maj_plot  ,
          t_xcorr_plot, a_xcorr_plot, 
          t_nmi_plot  , a_nmi_plot  , cols=2)
#dev.off()
@

Kappa minus multi-atlas mean vs. Number of Templates

Multi-atlas means:

\begin{table}
    \begin{tabular}{c|c|c}
        Number of Atlases & ANIMAL Kappa (Jaccard) & ANTS Kappa (Jaccard) \\ \hline
        3                 & 0.76 (0.63)             & 0.80 (0.69)          \\ 
        4                 & 0.75 (0.62)             & 0.79 (0.67)          \\ 
        5                 & 0.79 (0.66)             & 0.82 (0.71)          \\ 
        6                 & 0.78 (0.65)             & 0.82 (0.70)          \\ 
        7                 & 0.80 (0.67)             & 0.83 (0.72)          \\ 
        8                 & 0.79 (0.67)             & 0.83 (0.72)          \\ 
        9                 & 0.80 (0.68)             & 0.84 (0.73)          \\
    \end{tabular}
\end{table}

- more atlases -> better performance
- larger template library -> better performance, but tails off around 10-15 templates
- no significant difference between majority or weighted vote methods (haven’t tested this statistically though). 
- consistently performs better than average naive performance by XXX
- using ANTS, with a large enough template library (>12) MAGeT brain performs better than the average multi-atlas approach with the same number of atlases.  using ANIMAL, 5 or more atlases needed before boost seen. 
- more atlases -> smaller template library required to improve on average multi-atlas performance
- discuss variance?  best/worst case? --how often do we expect random template library selection to work decently

\todo{cost (in registrations) / benefit trade off graph:  show number of registrations per Kappa?  or hours of manual labour per Kappa?)}

\subsubsection{Winterburn Atlas Segmentation of ADNI Baseline Images}

- A2A shows that if atlas population strongly(?) represents subject set variability, then free choice from atlas population will produce improvements (we know this b/c of extensive validation trials). 
- what about in the case where atlas population doesn’t strongly represent subject set variability (e.g. a priori atlas set)?  then, we can use atlas selection to refine atlas set?

\todo{Kappa against our manual rater is low}

<<ADNI-baseline-volumes-prep,echo=F,cache=T>>=
data = read.csv('data/ADNI_baseline_volumes.csv')
unscaled=subset(data,Scaled==0 & Baseline==1 & DX != '')
attach(unscaled)
totals = data.frame(LONIUID,L+R,UCSFFSL_L+UCSFSL_R,MAPER_L+MAPER_R,UCSFSNT_L + UCSFSNT_R,SNT_L + SNT_R,DX)
colnames(totals) <- c("ID","MAGeT","FSL","MAPER","UCSFSNT","SNT","DX")
@

\begin{figure}[h]
<<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
library(reshape2)
qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),colour=variable,geom="boxplot")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-boxplot}
\end{figure}

\begin{figure}[h]
<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
library(reshape2)
qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),measure.vars=c("FSL","MAPER","MAGeT")),colour=variable, geom="smooth", method="lm")
@
  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
  \label{ADNI-baseline-volumes-plot}
\end{figure}

\end{document}

### 2012-10-17 paper.tex
reorg as a knitr project

--- 2012-10-17 paper.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage{graphicx, color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
+\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
+\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
+\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
+\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
+\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
+\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
+\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
+\newcommand{\hlslot}[1]{\textit{#1}}%
+\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage[margin=1.5in]{geometry}
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
 
-<<a2a, cache=TRUE, fig = T, echo = F, dev='CairoPDF', warning=F, error=F>>=
-library(ggplot2)
-multiplot <- function(..., plotlist=NULL, cols) {
-    require(grid)
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-    # Make a list from the ... arguments and plotlist
-    plots <- c(list(...), plotlist)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-    numPlots = length(plots)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-    # Make the panel
-    plotCols = cols                          # Number of columns of plots
-    plotRows = ceiling(numPlots/plotCols) # Number of rows needed, calculated from # of cols
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(plotRows, plotCols)))
-    vplayout <- function(x, y)
-        viewport(layout.pos.row = x, layout.pos.col = y)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-        curRow = ceiling(i/plotCols)
-        curCol = (i-1) %% plotCols + 1
-        print(plots[[i]], vp = vplayout(curRow, curCol ))
-    }
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-}
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}\begin{verbatim}
+## pdf 
+##   2
+\end{verbatim}
 
-#####################################################
-# TRACC
-t_naive      = read.csv(file='data/a2a_tracc_naive.csv')
-t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
-t_maj        = subset(t_data, method == 'majvote')
-t_xcorr      = subset(t_data, method == 'xcorr')
-t_nmi        = subset(t_data, method == 'nmi')
-t_mean_naive = mean(t_naive$k)
-t_ylimits    = c(0.7,0.85)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-t_multiatlas = read.csv(file='data/a2a_tracc_multiatlas_2012_08_14.csv')
-t_ma_means   = aggregate( k ~ num_atlases , data=t_multiatlas, mean)
-t_maj$k_minus_ma_mean   <- t_maj$k   - t_ma_means[ t_maj$num_atlases - 3, 2 ]
-t_xcorr$k_minus_ma_mean <- t_xcorr$k - t_ma_means[ t_xcorr$num_atlases - 3, 2 ]
-t_nmi$k_minus_ma_mean   <- t_nmi$k   - t_ma_means[ t_nmi$num_atlases - 3, 2 ]
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-# Majority Vote
-# plots kappa by numbeer of templates used
-#geom = c('smooth'), line=1.5, formula=y~poly(x,2), method='lm') + 
-t_maj_plot <- qplot(x=num_templates, y=k, data=t_maj, colour=as.factor(num_atlases), 
-  geom = c('smooth')) + 
-geom_hline(aes(yintercept=t_mean_naive)) + 
-geom_hline(aes(yintercept = k, colour = as.factor(num_atlases)), linetype="dotted", 
-  data = t_ma_means ) + 
-opts(title = "ANIMAL\nMajority vote") +
-opts(legend.position = "None") + 
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" ) + 
-coord_cartesian(ylim=t_ylimits)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-# X-Corr results with a full suite of templates
-# plots kappa by top N templates choosen 
-t_xcorr_plot <- qplot(x=top_n, y=k, data=t_xcorr,
-colour=as.factor(num_atlases), geom=c('smooth')) + 
-geom_hline(aes(yintercept=t_mean_naive))  +
-coord_cartesian(ylim=t_ylimits) + 
-opts(title = "X-Corr vote") +
-opts(legend.position = "None") + 
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" ) + 
-coord_cartesian(ylim=t_ylimits)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-# NMI results with a full suite of templates
-# plots kappa by top N templates choosen 
-t_nmi_plot <- qplot(x=top_n, y=k, data=t_nmi,
-colour=as.factor(num_atlases), geom=c('smooth')) +
-geom_hline(aes(yintercept=t_mean_naive))  +
-opts(title = "NMI vote") + 
-opts(legend.position = "None") + 
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" ) + 
-coord_cartesian(ylim=t_ylimits)
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}
 
-#####################################################
-# ANTS
-
-a_naive      = read.csv(file='data/a2a_ants_naive_2012_08_30.csv')
-a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')
-a_maj        = subset(a_data, method == 'majvote')
-a_xcorr      = subset(a_data, method == 'xcorr')
-a_nmi        = subset(a_data, method == 'nmi')
-a_mean_naive = mean(a_naive$k)
-a_ylimits    = c(0.7,0.85)
-
-a_multiatlas = read.csv(file='data/a2a_ants_multiatlas_2012_08_14.csv')
-a_ma_means   = aggregate( k ~ num_atlases , data=a_multiatlas, mean)
-a_maj$k_minus_ma_mean   <- a_maj$k - a_ma_means[ a_maj$num_atlases - 3, 2 ]
-a_xcorr$k_minus_ma_mean <- a_xcorr$k - a_ma_means[ a_xcorr$num_atlases - 3, 2 ]
-a_nmi$k_minus_ma_mean   <- a_nmi$k   - a_ma_means[ a_nmi$num_atlases - 3, 2 ]
-
-# Majority Vote
-# plots kappa by numbeer of templates used
-
-a_maj_plot <- qplot(x=num_templates, y=k, data=a_maj, colour=as.factor(num_atlases), 
-  geom=c('smooth')) + 
-geom_hline(aes(yintercept=a_mean_naive))  +
-coord_cartesian(ylim=a_ylimits) + 
-geom_hline(aes(yintercept = k, colour = as.factor(num_atlases)), linetype="dotted", 
-  data = a_ma_means ) +
-opts(title = "ANTS\nMajority vote")  +
-opts(legend.position = "None") + 
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" )
-
-# X-Corr results with a full suite of templates
-# plots kappa by top N templates choosen 
-a_xcorr_plot <- qplot(x=top_n, y=k, data=a_xcorr,
-colour=as.factor(num_atlases), geom=c('smooth')) + 
-geom_hline(aes(yintercept=a_mean_naive))  +
-opts(title = "X-Corr vote") +
-opts(legend.direction = "horizontal") + 
-opts(legend.position = "bottom") + 
-coord_cartesian(ylim=a_ylimits)  +
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" )
-
-# NMI results with a full suite of templates
-# plots kappa by top N templates choosen 
-a_nmi_plot <- qplot(x=top_n, y=k, data=a_nmi,
-colour=as.factor(num_atlases), geom=c('smooth')) +
-geom_hline(aes(yintercept=a_mean_naive))  +
-opts(title = "NMI vote") + 
-opts(legend.position = "None") + 
-coord_cartesian(ylim=a_ylimits) +
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" )
-
-#loess.control(trace.hat = c("approximate"), statistics = c("approximate"), surface = c("interpolate"))
-pdf("a2a_kbytemplates.pdf", width=15, height=15)
-multiplot(t_maj_plot  , a_maj_plot  ,
-          t_xcorr_plot, a_xcorr_plot, 
-          t_nmi_plot  , a_nmi_plot  , cols=2)
-dev.off()
-
-############################################
-#  Kappa adjusted by Multi-atlas mean 
-a_ylimits    = c(-0.03,0.04)
-t_ylimits    = a_ylimits
-
-t_maj_plot <- qplot(x=num_templates, y=k_minus_ma_mean, data=t_maj, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "ANIMAL\nMajority vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-t_xcorr_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=t_xcorr, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "X-Corr vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-t_nmi_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=t_nmi, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "NMI vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-a_maj_plot <- qplot(x=num_templates, y=k_minus_ma_mean, data=a_maj, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "ANTS\nMajority vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=a_ylimits)
-
-a_xcorr_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=a_xcorr, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "X-Corr vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=a_ylimits)
-
-a_nmi_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=a_nmi, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "NMI vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=a_ylimits)
-
-#print(a_nmi_plot)
-#pdf("a2a_kminusMAmean.pdf", width=15, height=15)
-multiplot(t_maj_plot  , a_maj_plot  ,
-          t_xcorr_plot, a_xcorr_plot, 
-          t_nmi_plot  , a_nmi_plot  , cols=2)
-#dev.off()
-@
+{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
+Use 'method = x' to change the smoothing method.}}\end{kframe}\includegraphics[width=\maxwidth]{figure/a2a}
+\end{knitrout}
 
@@
 
-<<ADNI-baseline-volumes-prep,echo=F,cache=T>>=
-data = read.csv('data/ADNI_baseline_volumes.csv')
-unscaled=subset(data,Scaled==0 & Baseline==1 & DX != '')
-attach(unscaled)
-totals = data.frame(LONIUID,L+R,UCSFFSL_L+UCSFSL_R,MAPER_L+MAPER_R,UCSFSNT_L + UCSFSNT_R,SNT_L + SNT_R,DX)
-colnames(totals) <- c("ID","MAGeT","FSL","MAPER","UCSFSNT","SNT","DX")
-@
+\begin{figure}[h]
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
 
-\begin{figure}[h]
-<<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-library(reshape2)
-qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),colour=variable,geom="boxplot")
-@
+{\centering \includegraphics[width=\maxwidth]{figure/ADNI-baseline-volumes-boxplot}
+
+}
+
+\end{knitrout}
+
   \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
@@
 \begin{figure}[h]
-<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
-library(reshape2)
-qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),measure.vars=c("FSL","MAPER","MAGeT")),colour=variable, geom="smooth", method="lm")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 359 rows containing missing values (stat\_smooth).}}
+
+{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 355 rows containing missing values (stat\_smooth).}}
+
+{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 355 rows containing missing values (stat\_smooth).}}\end{kframe}
+
+{\centering \includegraphics[width=\maxwidth]{figure/ADNI-baseline-volumes-plot}
+
+}
+
+\end{knitrout}
+
   \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}

### 2012-10-17 r/a2a.r

@@ -1,0 +1,192 @@
+library(ggplot2)
+# Authors jon julie jason martin jens aristotle mallar
+#X11 = function (...) grDevices::x11(...,type='cairo')
+#X11 = CairoX11()
+
+multiplot <- function(..., plotlist=NULL, cols) {
+    require(grid)
+
+    # Make a list from the ... arguments and plotlist
+    plots <- c(list(...), plotlist)
+
+    numPlots = length(plots)
+
+    # Make the panel
+    plotCols = cols                          # Number of columns of plots
+    plotRows = ceiling(numPlots/plotCols) # Number of rows needed, calculated from # of cols
+
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(plotRows, plotCols)))
+    vplayout <- function(x, y)
+        viewport(layout.pos.row = x, layout.pos.col = y)
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+        curRow = ceiling(i/plotCols)
+        curCol = (i-1) %% plotCols + 1
+        print(plots[[i]], vp = vplayout(curRow, curCol ))
+    }
+
+}
+
+#####################################################
+# TRACC
+t_naive      = read.csv(file='a2a_tracc_naive.csv')
+t_data       = read.csv(file='a2a_tracc_results_2012_08_30.csv')
+t_maj        = subset(t_data, method == 'majvote')
+t_xcorr      = subset(t_data, method == 'xcorr')
+t_nmi        = subset(t_data, method == 'nmi')
+t_mean_naive = mean(t_naive$k)
+t_ylimits    = c(0.7,0.85)
+
+t_multiatlas = read.csv(file='a2a_tracc_multiatlas_2012_08_14.csv')
+t_ma_means   = aggregate( k ~ num_atlases , data=t_multiatlas, mean)
+t_maj$k_minus_ma_mean   <- t_maj$k   - t_ma_means[ t_maj$num_atlases - 3, 2 ]
+t_xcorr$k_minus_ma_mean <- t_xcorr$k - t_ma_means[ t_xcorr$num_atlases - 3, 2 ]
+t_nmi$k_minus_ma_mean   <- t_nmi$k   - t_ma_means[ t_nmi$num_atlases - 3, 2 ]
+
+# Majority Vote
[…]

### 2012-10-17 r/adni_classification.r

@@ -1,0 +1,9 @@
+# Classify the 69 ADNI subjects by hippocampal volume
+data = read.csv('adni_data.csv', sep=";")
+
+# defaults showing t-values for AD to NX 
+summary(lm((HC_left_vol + HC_right_vol) ~ DX + Sex + Age, data = data))
+
+# NX to AD/NX
+data$DX<-relevel(x=data$DX, ref="NX")
+summary(lm((HC_left_vol + HC_right_vol) ~ DX + Sex + Age, data = data))

### 2012-10-17 r/dan-model-validation.r

@@ -1,0 +1,35 @@
+library(rms)
+data = read.csv('adni_data.csv', sep=';')
+summary(data)
+dd <- datadist(data)
+options(datadist="dd")
+describe(data)
+mod <- lrm(data=data, DX ~ hc_left_vol + Age + Sex)
+mod
+modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex)
+modright <- lrm(data=data, DX ~ hc_right_vol + Age + Sex)
+modright <- lrm(data=data, DX ~ hc_right_vol + Age + Sex, subset=DX!=MCI)
+modright <- lrm(data=data, DX ~ hc_right_vol + Age + Sex, subset=DX!="MCI")
+modright
+modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="MCI")
+modleft
+
+summary(modleft)
+summary(modright)
+modleft
+modleft2 <- lrm(data=data, DX ~ Age + Sex, subset=DX!="MCI")
+lrtest(modleft, modleft2)
+validate(modleft, method=.632)
+modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="MCI", x=TRUE, y=TRUE)
+validate(modleft, method=.632)
+validate(modleft, method=".632")
+validate(modleft, method=boot)
+validate(modleft, method="boot")
+calibrate(modleft, method="boot")
+modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="MCI", x=TRUE, y=TRUE)
+modleft
+summary(modleft)
+print(modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="AD", x=TRUE, y=TRUE))
+summary(modleft)
+print(modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="NX", x=TRUE, y=TRUE))
+savehistory("dan-model-validation.R")

### 2012-11-14 paper.Rnw
Added the SZ dataset results -- MAGeT vs. Manual volumes.

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\SweaveOpts{echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}
+
+<<loadpackages, results="hide", echo=TRUE, cache=FALSE>>=
+library(ggplot2)
+library(Hmisc)
+library(reshape2)
+@
 
 \title{Hippocampal segmentation with MAGeT}
-\author{Jon Pipitone, }
+\author{Jon Pipitone}
 \maketitle
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{MAGeT-figure.png}
+    \includegraphics[width=0.5\textwidth]{figure/MAGeT-figure.png}
   \caption{Diagram of the MAGeT Brain algorithm}
@@
 <<a2a, cache=TRUE, fig = T, echo = F, dev='CairoPDF', warning=F, error=F>>=
-library(ggplot2)
 multiplot <- function(..., plotlist=NULL, cols) {
@@
 #loess.control(trace.hat = c("approximate"), statistics = c("approximate"), surface = c("interpolate"))
-pdf("a2a_kbytemplates.pdf", width=15, height=15)
+#pdf("a2a_kbytemplates.pdf", width=15, height=15)
 multiplot(t_maj_plot  , a_maj_plot  ,
@@
           t_nmi_plot  , a_nmi_plot  , cols=2)
-dev.off()
+#dev.off()
 
@@
 - larger template library -> better performance, but tails off around 10-15 templates
-- no significant difference between majority or weighted vote methods (haven’t tested this statistically though). 
+- no significant difference between majority or weighted vote methods (haven???t tested this statistically though). 
 - consistently performs better than average naive performance by XXX
@@
 - A2A shows that if atlas population strongly(?) represents subject set variability, then free choice from atlas population will produce improvements (we know this b/c of extensive validation trials).
-- what about in the case where atlas population doesn’t strongly represent subject set variability (e.g. a priori atlas set)?  then, we can use atlas selection to refine atlas set?
+- what about in the case where atlas population doesn???t strongly represent subject set variability (e.g. a priori atlas set)?  then, we can use atlas selection to refine atlas set?
 
@@
 <<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-library(reshape2)
 qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),colour=variable,geom="boxplot")
@@
 <<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
-library(reshape2)
 qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),measure.vars=c("FSL","MAPER","MAGeT")),colour=variable, geom="smooth", method="lm")
@@
 
+\subsubsection{First Episode Schizophrenic patients}
+
+<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
+fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
+fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
+tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ, data=fep_sz_demographics, method = "reverse", test = FALSE)
+latex(tab, file="", size="scriptsize", where="h")
+@
+
+\begin{figure}[h]
+<<SZ_volumes,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
+SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
+qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm")
+@
+  \caption{Comparison of total HC volumes for MAGeT against manually rated volumes}
+  \label{SZ_volumes}
+\end{figure}
 \end{document}

### 2012-11-14 paper.tex
Added the SZ dataset results -- MAGeT vs. Manual volumes.

@@
 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlfunctioncall{library}(ggplot2)
+\hlfunctioncall{library}(Hmisc)
+\hlfunctioncall{library}(reshape2)
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
 \title{Hippocampal segmentation with MAGeT}
-\author{Jon Pipitone, }
+\author{Jon Pipitone}
 \maketitle
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
 
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
+{\centering \includegraphics[width=\maxwidth]{figure/a2a1}
 
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
+}
 
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
+{\centering \includegraphics[width=\maxwidth]{figure/a2a2}
 
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
+}
 
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}\begin{verbatim}
-## pdf 
-##   2
-\end{verbatim}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}
-
-{\ttfamily\noindent\itshape\textcolor{messagecolor}{\#\# geom\_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y \textasciitilde{} s(x, bs = "cs").
-Use 'method = x' to change the smoothing method.}}\end{kframe}\includegraphics[width=\maxwidth]{figure/a2a}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 359 rows containing missing values (stat\_smooth).}}
-
-{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 355 rows containing missing values (stat\_smooth).}}
-
-{\ttfamily\noindent\textcolor{warningcolor}{\#\# Warning: Removed 355 rows containing missing values (stat\_smooth).}}\end{kframe}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
 
@@
 
+\subsubsection{First Episode Schizophrenic patients}
+
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+\begin{table}[h]
+\scriptsize
+\caption{Descriptive Statistics by DX\label{tab}} 
+\begin{center}
+\begin{tabular}{lrc}
+\hline\hline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{FEP}\tabularnewline
+&&\multicolumn{1}{c}{{\scriptsize $N=81$}}\tabularnewline
+\hline
+Age&80&{\scriptsize 21~}{23 }{\scriptsize 26} \tabularnewline
+Gender~:~M&81&63\%~{\scriptsize~(51)}\tabularnewline
+Handedness~:~ambi&81&~6\%~{\scriptsize~(~5)}\tabularnewline
+~~~~left&&~5\%~{\scriptsize~(~4)}\tabularnewline
+~~~~right&&89\%~{\scriptsize~(72)}\tabularnewline
+Education&81&{\scriptsize 11~}{13 }{\scriptsize 15} \tabularnewline
+SES~:~lower&81&31\%~{\scriptsize~(25)}\tabularnewline
+~~~~middle&&54\%~{\scriptsize~(44)}\tabularnewline
+~~~~upper&&15\%~{\scriptsize~(12)}\tabularnewline
+FSIQ&79&{\scriptsize  88~}{102 }{\scriptsize 109} \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table}
+
+\begin{figure}[h]
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/SZ_volumes}
+
+}
+
+\end{knitrout}
+
+  \caption{Comparison of total HC volumes for MAGeT against manually rated volumes}
+  \label{SZ_volumes}
+\end{figure}
 \end{document}

### 2012-11-14 paper.Rnw
Moved a2a figures to using facets for more better prettier.
Also included ADNI and SZ demographic data.

@@
 \usepackage[margin=1.5in]{geometry}
-\usepackage{algorithm}
-\usepackage{color}
-\usepackage{algpseudocode}
-\usepackage{amsmath}
-\usepackage{url}
+\usepackage{algorithm, color,algpseudocode, amsmath, url, geometry, ctable }
 
@@
 \begin{document}
-\SweaveOpts{echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}
+\SweaveOpts{echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}
 
-<<loadpackages, results="hide", echo=TRUE, cache=FALSE>>=
+<<loadpackages, results="hide", echo=FALSE, cache=FALSE>>=
 library(ggplot2)
@@
 library(reshape2)
+library(ADNIMERGE)
 @
@@
 
+<<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
+adni_69_RIDs = c("295","413","619","685","729","782","938","954",
+                 "1018","1155","907","981","222","324","448","546",
+                 "553","572","602","610","814","1341","675","681",
+                 "731","1130","41","68","70","101","249","293","316",
+                 "414","698","1206","1222","1339","751","1030","3",
+                 "5","10","16","22","53","168","183","241","1205",
+                 "328","1095","991","213","159","337","343","642",
+                 "753","1109","1217","29","56","1188","1293","149",
+                 "382","431","1202")
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTMARRY + PTETHCAT + PTRACCAT + CDRSB + ADAS13 + MMSE, 
+               data = adnimerge, subset = VISCODE == 'bl' & RID %in% adni_69_RIDs, 
+               method = "reverse", test=TRUE, overall=TRUE)
+
+latex(tab, file="", size="scriptsize", where="h",landscape=TRUE, ctable=TRUE)
+@
+
 \todo{For minctracc -- we registered to TAL}
@@
 Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))
-
-<<a2a, cache=TRUE, fig = T, echo = F, dev='CairoPDF', warning=F, error=F>>=
-multiplot <- function(..., plotlist=NULL, cols) {
-    require(grid)
-
-    # Make a list from the ... arguments and plotlist
-    plots <- c(list(...), plotlist)
-
-    numPlots = length(plots)
-
-    # Make the panel
-    plotCols = cols                          # Number of columns of plots
-    plotRows = ceiling(numPlots/plotCols) # Number of rows needed, calculated from # of cols
-
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(plotRows, plotCols)))
-    vplayout <- function(x, y)
-        viewport(layout.pos.row = x, layout.pos.col = y)
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-        curRow = ceiling(i/plotCols)
-        curCol = (i-1) %% plotCols + 1
-        print(plots[[i]], vp = vplayout(curRow, curCol ))
-    }
-
-}
-
-#####################################################
-# TRACC
-t_naive      = read.csv(file='data/a2a_tracc_naive.csv')
+<<a2a_facet, cache=TRUE, fig=TRUE>>=
 t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
-t_maj        = subset(t_data, method == 'majvote')
-t_xcorr      = subset(t_data, method == 'xcorr')
-t_nmi        = subset(t_data, method == 'nmi')
-t_mean_naive = mean(t_naive$k)
-t_ylimits    = c(0.7,0.85)
+a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')
 
 t_multiatlas = read.csv(file='data/a2a_tracc_multiatlas_2012_08_14.csv')
+a_multiatlas = read.csv(file='data/a2a_ants_multiatlas_2012_08_14.csv')
 t_ma_means   = aggregate( k ~ num_atlases , data=t_multiatlas, mean)
-t_maj$k_minus_ma_mean   <- t_maj$k   - t_ma_means[ t_maj$num_atlases - 3, 2 ]
-t_xcorr$k_minus_ma_mean <- t_xcorr$k - t_ma_means[ t_xcorr$num_atlases - 3, 2 ]
-t_nmi$k_minus_ma_mean   <- t_nmi$k   - t_ma_means[ t_nmi$num_atlases - 3, 2 ]
+a_ma_means   = aggregate( k ~ num_atlases , data=a_multiatlas, mean)
 
-# Majority Vote
-# plots kappa by numbeer of templates used
-#geom = c('smooth'), line=1.5, formula=y~poly(x,2), method='lm') + 
-t_maj_plot <- qplot(x=num_templates, y=k, data=t_maj, colour=as.factor(num_atlases), 
-  geom = c('smooth')) + 
-geom_hline(aes(yintercept=t_mean_naive)) + 
-geom_hline(aes(yintercept = k, colour = as.factor(num_atlases)), linetype="dotted", 
-  data = t_ma_means ) + 
-opts(title = "ANIMAL\nMajority vote") +
-opts(legend.position = "None") + 
+# add registration method to data frames (before joining)
+t_data       = cbind(t_data, reg_method = "ANIMAL")
+a_data       = cbind(a_data, reg_method = "ANTS")
+
+# add column with kappa - multiatlas mean
+# num_atlases - 2 is a cheap trick to perform the necessary lookup into the ma_means tables
+t_data$k_minus_ma = t_data$k - t_ma_means[ t_data$num_atlases - 2, 2 ]
+a_data$k_minus_ma = a_data$k - a_ma_means[ a_data$num_atlases - 2, 2 ]
+
+all_data     = rbind(t_data,a_data)
+
+# use top_n as the number of templates for weighted voting schemes
+all_data[ all_data$method != "majvote", ]$num_templates = all_data[ all_data$method != "majvote", ]$top_n
+
+t_naive      = read.csv(file='data/a2a_tracc_naive.csv')
+a_naive      = read.csv(file='data/a2a_ants_naive_2012_08_30.csv')
+t_mean_naive = mean(t_naive$k)
+a_mean_naive = mean(a_naive$k)
+df_naive     = data.frame(
+                value = c(rep(t_mean_naive,3),rep(a_mean_naive,3)),
+                method = rep(c("majvote", "nmi", "xcorr"),2), 
+                reg_method = c(rep("ANIMAL",3),rep("ANTS",3)))
+
+method_labels <- list("Majority Vote" = "majvote", "NMI Vote" = "nmi", "Cross-correlation Vote" = "xcorr")
+levels(all_data$method) <- method_labels
+levels(df_naive$method) <- method_labels
+
+ggplot(all_data, aes(x=num_templates, y=k, colour=as.factor(num_atlases))) + 
+  facet_grid(method ~ reg_method) + 
+  geom_smooth() + 
+  geom_hline(data = df_naive, aes(yintercept = value, linetype="dashed")) + 
   scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" ) + 
-coord_cartesian(ylim=t_ylimits)
+  xlab( "Number of Templates" ) + ylab( "Mean Kappa" ) 
+@
 
-# X-Corr results with a full suite of templates
-# plots kappa by top N templates choosen 
-t_xcorr_plot <- qplot(x=top_n, y=k, data=t_xcorr,
-colour=as.factor(num_atlases), geom=c('smooth')) + 
-geom_hline(aes(yintercept=t_mean_naive))  +
-coord_cartesian(ylim=t_ylimits) + 
-opts(title = "X-Corr vote") +
-opts(legend.position = "None") + 
+<<a2a_facet_mameans, cache=TRUE, fig=TRUE>>=
+ggplot(all_data, aes(x=num_templates, y=k_minus_ma, colour=as.factor(num_atlases))) + 
+  facet_grid(method ~ reg_method) + 
+  geom_smooth() + 
   scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-# NMI results with a full suite of templates
-# plots kappa by top N templates choosen 
-t_nmi_plot <- qplot(x=top_n, y=k, data=t_nmi,
-colour=as.factor(num_atlases), geom=c('smooth')) +
-geom_hline(aes(yintercept=t_mean_naive))  +
-opts(title = "NMI vote") + 
-opts(legend.position = "None") + 
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-#####################################################
-# ANTS
-
-a_naive      = read.csv(file='data/a2a_ants_naive_2012_08_30.csv')
-a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')
-a_maj        = subset(a_data, method == 'majvote')
-a_xcorr      = subset(a_data, method == 'xcorr')
-a_nmi        = subset(a_data, method == 'nmi')
-a_mean_naive = mean(a_naive$k)
-a_ylimits    = c(0.7,0.85)
-
-a_multiatlas = read.csv(file='data/a2a_ants_multiatlas_2012_08_14.csv')
-a_ma_means   = aggregate( k ~ num_atlases , data=a_multiatlas, mean)
-a_maj$k_minus_ma_mean   <- a_maj$k - a_ma_means[ a_maj$num_atlases - 3, 2 ]
-a_xcorr$k_minus_ma_mean <- a_xcorr$k - a_ma_means[ a_xcorr$num_atlases - 3, 2 ]
-a_nmi$k_minus_ma_mean   <- a_nmi$k   - a_ma_means[ a_nmi$num_atlases - 3, 2 ]
-
-# Majority Vote
-# plots kappa by numbeer of templates used
-
-a_maj_plot <- qplot(x=num_templates, y=k, data=a_maj, colour=as.factor(num_atlases), 
-  geom=c('smooth')) + 
-geom_hline(aes(yintercept=a_mean_naive))  +
-coord_cartesian(ylim=a_ylimits) + 
-geom_hline(aes(yintercept = k, colour = as.factor(num_atlases)), linetype="dotted", 
-  data = a_ma_means ) +
-opts(title = "ANTS\nMajority vote")  +
-opts(legend.position = "None") + 
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" )
-
-# X-Corr results with a full suite of templates
-# plots kappa by top N templates choosen 
-a_xcorr_plot <- qplot(x=top_n, y=k, data=a_xcorr,
-colour=as.factor(num_atlases), geom=c('smooth')) + 
-geom_hline(aes(yintercept=a_mean_naive))  +
-opts(title = "X-Corr vote") +
-opts(legend.direction = "horizontal") + 
-opts(legend.position = "bottom") + 
-coord_cartesian(ylim=a_ylimits)  +
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" )
-
-# NMI results with a full suite of templates
-# plots kappa by top N templates choosen 
-a_nmi_plot <- qplot(x=top_n, y=k, data=a_nmi,
-colour=as.factor(num_atlases), geom=c('smooth')) +
-geom_hline(aes(yintercept=a_mean_naive))  +
-opts(title = "NMI vote") + 
-opts(legend.position = "None") + 
-coord_cartesian(ylim=a_ylimits) +
-scale_colour_hue(name="Atlases") +
-xlab( "Number of Templates" ) + 
-ylab( "Kappa" )
-
-#loess.control(trace.hat = c("approximate"), statistics = c("approximate"), surface = c("interpolate"))
-#pdf("a2a_kbytemplates.pdf", width=15, height=15)
-multiplot(t_maj_plot  , a_maj_plot  ,
-          t_xcorr_plot, a_xcorr_plot, 
-          t_nmi_plot  , a_nmi_plot  , cols=2)
-#dev.off()
-
-############################################
-#  Kappa adjusted by Multi-atlas mean 
-a_ylimits    = c(-0.03,0.04)
-t_ylimits    = a_ylimits
-
-t_maj_plot <- qplot(x=num_templates, y=k_minus_ma_mean, data=t_maj, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "ANIMAL\nMajority vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-t_xcorr_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=t_xcorr, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "X-Corr vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-t_nmi_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=t_nmi, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "NMI vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=t_ylimits)
-
-a_maj_plot <- qplot(x=num_templates, y=k_minus_ma_mean, data=a_maj, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "ANTS\nMajority vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=a_ylimits)
-
-a_xcorr_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=a_xcorr, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "X-Corr vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=a_ylimits)
-
-a_nmi_plot <- qplot(x=top_n, y=k_minus_ma_mean, data=a_nmi, colour=as.factor(num_atlases), geom = c('smooth'), se=FALSE) + 
-opts(title = "NMI vote") + opts(legend.position = "None") + 
-xlab( "Number of Templates" ) + ylab( "Kappa - MA means" ) + 
-coord_cartesian(ylim=a_ylimits)
-
-#print(a_nmi_plot)
-#pdf("a2a_kminusMAmean.pdf", width=15, height=15)
-multiplot(t_maj_plot  , a_maj_plot  ,
-          t_xcorr_plot, a_xcorr_plot, 
-          t_nmi_plot  , a_nmi_plot  , cols=2)
-#dev.off()
+  xlab( "Number of Templates" ) + ylab( "Difference in mean Kappa between voting method and multiatlas" )
 @
@@
 
+<<a2a_multiatlas_means, results="asis",cache=FALSE>>=
+latex(merge(a_ma_means, t_ma_means, by="num_atlases"), colheads=c("Atlases", "ANTS", "ANIMAL"), file="")
+@
 \begin{table}

### 2012-11-14 paper.tex
Moved a2a figures to using facets for more better prettier.
Also included ADNI and SZ demographic data.

@@
 \begin{document}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlfunctioncall{library}(ggplot2)
-\hlfunctioncall{library}(Hmisc)
-\hlfunctioncall{library}(reshape2)
-\end{alltt}
-\end{kframe}
-\end{knitrout}
 
@@
 
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+{\scriptsize\ctable[caption={Descriptive Statistics by Baseline Dx},label=tab,rotate]{lccccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{Numbers after percents are frequencies.}\tnote[]{\indent Tests used:}\tnote[]{\textsuperscript{\normalfont 1}Kruskal-Wallis test; \textsuperscript{\normalfont 2}Pearson test}}{\FL
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}&\multicolumn{1}{c}{Test Statistic}\NN
+&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}&\ML
+Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.20~}{75.50 }{\scriptsize 78.45} &{\scriptsize 71.00~}{77.10 }{\scriptsize 81.40} &{\scriptsize 71.70~}{77.80 }{\scriptsize 81.75} &{\scriptsize 71.50~}{76.60 }{\scriptsize 81.30} &$ F_{2,66}=0.15 ,~ P=0.862 ^{1} $\NN
+Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}&$ \chi^{2}_{2}=0 ,~ P=1 ^{2} $\NN
+Education&{\scriptsize 16.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 15.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 16.5} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &$ F_{2,66}=3.44 ,~ P=0.038 ^{1} $\NN
+Marital~status~at~baseline~:~Married&74\%~{\scriptsize~(17)}&87\%~{\scriptsize~(20)}&87\%~{\scriptsize~(20)}&83\%~{\scriptsize~(57)}&$ \chi^{2}_{4}=3.04 ,~ P=0.551 ^{2} $\NN
+~~~~Widowed&22\%~{\scriptsize~(~5)}&13\%~{\scriptsize~(~3)}&13\%~{\scriptsize~(~3)}&16\%~{\scriptsize~(11)}&\NN
+~~~~Divorced&~4\%~{\scriptsize~(~1)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~1\%~{\scriptsize~(~1)}&\NN
+~~~~Never~married&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&\NN
+~~~~Unknown&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&\NN
+Ethnicity~:~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&$^{2}$\NN
+~~~~Not~Hisp/Latino&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(69)}&\NN
+~~~~Hisp/Latino&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
+Race~:~Am~Indian/Alaskan&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&$ \chi^{2}_{2}=5.61 ,~ P=0.061 ^{2} $\NN
+~~~~Asian&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
+~~~~Hawaiian/Other~PI&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
+~~~~Black&~17\%~{\scriptsize~(~4)}&~~4\%~{\scriptsize~(~1)}&~~0\%~{\scriptsize~(~0)}&~~7\%~{\scriptsize~(~5)}&\NN
+~~~~White&~83\%~{\scriptsize~(19)}&~96\%~{\scriptsize~(22)}&100\%~{\scriptsize~(23)}&~93\%~{\scriptsize~(64)}&\NN
+~~~~More~than~one&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
+~~~~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
+CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} &$ F_{2,66}=294.2 ,~ P<0.001 ^{1} $\NN
+ADAS~13&{\scriptsize  4.665~}{ 5.670 }{\scriptsize 12.335} &{\scriptsize 14.335~}{16.000 }{\scriptsize 20.500} &{\scriptsize 23.830~}{29.000 }{\scriptsize 31.665} &{\scriptsize 10.000~}{16.000 }{\scriptsize 25.330} &$ F_{2,66}=109.8 ,~ P<0.001 ^{1} $\NN
+MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} &$ F_{2,66}=94.17 ,~ P<0.001 ^{1} $\LL
+}}
+
 \todo{For minctracc -- we registered to TAL}
@@
 Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/a2a_facet}
+
+}
+
+\end{knitrout}
 
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/a2a1}
-
-}
-
-{\centering \includegraphics[width=\maxwidth]{figure/a2a2}
+{\centering \includegraphics[width=\maxwidth]{figure/a2a_facet_mameans}
 
@@
 Multi-atlas means:
+
+% latex.default(merge(a_ma_means, t_ma_means, by = "num_atlases"),      colheads = c("Atlases", "ANTS", "ANIMAL"), file = "") 
+%
+\begin{table}[!tbp]
+\begin{center}
+\begin{tabular}{lrrr}
+\hline\hline
+\multicolumn{1}{l}{merge}&\multicolumn{1}{c}{Atlases}&\multicolumn{1}{c}{ANTS}&\multicolumn{1}{c}{ANIMAL}\tabularnewline
+\hline
+1&$3$&$0.805414417498623$&$0.764266832660852$\tabularnewline
+2&$4$&$0.793646605043623$&$0.752420649183594$\tabularnewline
+3&$5$&$0.823292141139203$&$0.785546924392891$\tabularnewline
+4&$6$&$0.815578947302174$&$0.778345910844141$\tabularnewline
+5&$7$&$0.831936342756014$&$0.797761650124922$\tabularnewline
+6&$8$&$0.826298462726884$&$0.792443538092187$\tabularnewline
+7&$9$&$0.836877632304638$&$0.804568318173359$\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2012-11-15 paper.Rnw
Updating layout of paper

@@
 
-The hippocampus is of particular interest to many researchers because it is implicated in forms of brain dysfunction such as Alzheimer's disease and schizophrenia, and has functional significance in cognitive processes such as learning and memory.
-For many research questions involving neuroimaging data, accurate identification of the hippocampus and its subregions in participant MR images is a necessary first step in order to analyse neuroanatomical differences and changes in  sample populations.
-\todo{expand}
+The hippocampus is of particular interest to many researchers because it is
+implicated in forms of brain dysfunction such as Alzheimer's disease and
+schizophrenia, and has functional significance in cognitive processes such as
+learning and memory.  For many research questions involving neuroimaging data,
+accurate identification of the hippocampus and its subregions in participant MR
+images is a necessary first step in order to analyse neuroanatomical
+differences and changes in  sample populations.  \todo{expand}
 
-Currently, the gold standard for neuroanatomical segmentation is manual labelling by an expert human rater.
-This is problematic for several reasons.
-Practically, manual segmentation takes a significant investment of time and expertise (Hammers2003) which may not be readily available to researchers or clinicians.
-<blah> estimates that expert segmentation can take anywhere between 45 minutes to several days per image depending on the image resolution and protocol.
-The amount of data produced in neuroimaging experiments increasingly exceeds the capacity for identification of specific neuroanatomical structures by an expert manual rater.
-It may be difficult for laboratories to find or train, someone with the expertise to do such segmentation.
+Currently, the gold standard for neuroanatomical segmentation is manual
+labelling by an expert human rater.  This is problematic for several reasons.
+Practically, manual segmentation takes a significant investment of time and
+expertise (Hammers2003) which may not be readily available to researchers or
+clinicians.  <blah> estimates that expert segmentation can take anywhere
+between 45 minutes to several days per image depending on the image resolution
+and protocol.  The amount of data produced in neuroimaging experiments
+increasingly exceeds the capacity for identification of specific
+neuroanatomical structures by an expert manual rater.  It may be difficult for
+laboratories to find or train, someone with the expertise to do such
+segmentation.
 
-Compounding this is that the hippocampus is a problematic structure for segmentation.
-Its very definition is disputed and varies widely in the literature (Geuze2004).
-This has led to a variety of competing segmentation protocols in the literature which researchers can adhere to, including necessary ad-hoc modifications to adapt the definition in order to answer a particular research question.
-Efforts are underway to create a unified protocol (Jack2011) for volumetry applications.
-But, even with a unified protocol, in order to investigate certain research questions it may be necessary to select from different hippocampal definitions.
-For example, (Poppenk2011) found that overall hippocampal volume change did not predict recollection memory performance, but dividing the hippocampus into anterior and posterior regions did.
+Compounding this is that the hippocampus is a problematic structure for
+segmentation. Its very definition is disputed and varies widely in the
+literature (Geuze2004).  This has led to a variety of competing segmentation
+protocols in the literature which researchers can adhere to, including
+necessary ad-hoc modifications to adapt the definition in order to answer a
+particular research question.  Efforts are underway to create a unified
+protocol (Jack2011) for volumetry applications. But, even with a unified
+protocol, in order to investigate certain research questions it may be
+necessary to select from different hippocampal definitions. For example,
+(Poppenk2011) found that overall hippocampal volume change did not predict
+recollection memory performance, but dividing the hippocampus into anterior and
+posterior regions did.
 
-\todo{this bit needs to be a little clearer - was it just the division of the HC that led to the improved understanding or was it that spatially localized differences in volumes are a better way to describe the difference).}
+\todo{this bit needs to be a little clearer - was it just the division of the
+HC that led to the improved understanding or was it that spatially localized
+differences in volumes are a better way to describe the difference).}
 
-\todo{- Also, other situations: developmental studies -- need different definitions depending on the age range of subjects
--- additionally, may want to tweak definition to test certain hypotheses (e.g. include subiculum: yes/no, more or less 
- conservative WM/GM boundaries, compare with definitions from a different study,, etc...).}
+\todo{- Also, other situations: developmental studies -- need different
+definitions depending on the age range of subjects -- additionally, may want to
+tweak definition to test certain hypotheses (e.g. include subiculum: yes/no,
+more or less conservative WM/GM boundaries, compare with definitions from a
+different study,, etc...).}
 
-One can also expect neuroanatomical variability in this structure as it loses volume through aging and in the time course of different diseases.
-\todo{This bit isn't really "joined to the previous statement"  may need a little glue here}
+One can also expect neuroanatomical variability in this structure as it loses
+volume through aging and in the time course of different diseases.  \todo{This
+bit isn't really "joined to the previous statement"  may need a little glue
+here}
 
-Automated segmentation techniques are available and can minimize several of the confounds often observed in studies that involve manual volumetry, such as the time and expertise required for segmentation and intra- and inter-rater reliability.
-A class of segmentation approaches, known as atlas-based segmentation (also, template-based segmentation), make use of expertly labeled neuroanatomical atlases.
-An atlas MR image is warped to fit the subject neuroanatomy using nonlinear registration techniques [4, 11], and then the atlas labels are propagated via the registration transformation from the atlas to match the unique neuroanatomy of a particular subject.
-Segmentation accuracy is affected by nonlinear registration and resampling errors, and by the extent to which the atlas neuroanatomy differs from the underlying neuroanatomical variability of the population under study (citation needed).
-To address these limitations, several atlases can be used (multi-atlas segmentation).
-Labels from each atlas are propagated to the subjects, by way of an nonlinear image registration, producing as many labellings for each subject as atlases.
-A label fusion technique is employed to merge the candidate labellings for a subject into a single segmentation using a decision rule (for instance, voxel-wise voting).
-Not all of the atlases need be used to segment each subject.
-Selecting only those atlases that are most similar to the subject image has been shown to significantly improve overall registration and segmentation accuracy (Aljabar 2008, Collins2010, Leung2010, Lotjonen2010).
+Automated segmentation techniques are available and can minimize several of the
+confounds often observed in studies that involve manual volumetry, such as the
+time and expertise required for segmentation and intra- and inter-rater
+reliability.  A class of segmentation approaches, known as atlas-based
+segmentation (also, template-based segmentation), make use of expertly labeled
+neuroanatomical atlases. An atlas MR image is warped to fit the subject
+neuroanatomy using nonlinear registration techniques [4, 11], and then the
+atlas labels are propagated via the registration transformation from the atlas
+to match the unique neuroanatomy of a particular subject.  Segmentation
+accuracy is affected by nonlinear registration and resampling errors, and by
+the extent to which the atlas neuroanatomy differs from the underlying
+neuroanatomical variability of the population under study (citation needed).
+To address these limitations, several atlases can be used (multi-atlas
+segmentation). Labels from each atlas are propagated to the subjects, by way of
+an nonlinear image registration, producing as many labellings for each subject
+as atlases. A label fusion technique is employed to merge the candidate
+labellings for a subject into a single segmentation using a decision rule (for
+instance, voxel-wise voting). Not all of the atlases need be used to segment
+each subject. Selecting only those atlases that are most similar to the subject
+image has been shown to significantly improve overall registration and
+segmentation accuracy (Aljabar 2008, Collins2010, Leung2010, Lotjonen2010).
 
-\todo{briefly discuss other automated segmentation methods, and justify our focus on atlas-based methods}
+\todo{briefly discuss other automated segmentation methods, and justify our
+focus on atlas-based methods}
 
-In summary, whilst manual segmentation is the gold standard for neuroanatomical segmentation, it is time-consuming and requires an investment in expertise and training, and may be inflexible in the face of differing hippocampal definitions or changes.
-Atlas-based automated segmentation dramatically reduces the time needed for segmentation, and can be completed without any expert rater as long as an appropriate atlas library is available <be more specific (e.g.
-shares protocol, demographics, phenotypes of subject population)>.
-In the situation where such an atlas library does not exist, then an expert rater is needed.
-Many automated approaches rely on an atlas library containing between 30 and 80 expertly labeled brains (Heckemann2006, Collins2010) which, even though producing such a library may require less manual work than segmenting the entire set of subjects, it still poses a substantial time commitment.
+In summary, whilst manual segmentation is the gold standard for neuroanatomical
+segmentation, it is time-consuming and requires an investment in expertise and
+training, and may be inflexible in the face of differing hippocampal
+definitions or changes. Atlas-based automated segmentation dramatically reduces
+the time needed for segmentation, and can be completed without any expert rater
+as long as an appropriate atlas library is available <be more specific (e.g.
+shares protocol, demographics, phenotypes of subject population)>.  In the
+situation where such an atlas library does not exist, then an expert rater is
+needed. Many automated approaches rely on an atlas library containing between
+30 and 80 expertly labeled brains (Heckemann2006, Collins2010) which, even
+though producing such a library may require less manual work than segmenting
+the entire set of subjects, it still poses a substantial time commitment.
 
-In this paper we extend our automated method for subcortical segmentation (striatum, globus pallidus, and thalamus; Chakravarty2012) to hippocampal segmentation.
-This method, known as MAGeT brain, is suited for situations where manual effort is at a premium, and flexibility is needed in atlas definitions.
-Our strategy is to bootstrap the familiar multi-atlas-based segmentation process from only a very small number of atlases.
-In our previous work we used only a single atlas, based on serial histological slices, and validated the algorithm on 20 adolescent MR images.
-In this work, we extensively investigate MAGeT used with differently sized atlas and template libraries, registration algorithms, label fusion strategies, and hippocampal definitions.
-We evaluate our algorithm on a subset of images from the ADNI dataset, and compare our segmentations to the manual segmentations provided by ADNI.
-Additionally, we evaluate the segmentations produced with MAGeT using hippocampal definitions from five high-resolution T1 in-vivo atlases (Winterburn2012).
-<conclude with summary that this is detailed validation the algorithm in order to pick the best operating parameters, whereas our previous work showed the feasibility of the approach>
+In this paper we extend our automated method for subcortical segmentation
+(striatum, globus pallidus, and thalamus; Chakravarty2012) to hippocampal
+segmentation. This method, known as MAGeT brain, is suited for situations where
+manual effort is at a premium, and flexibility is needed in atlas definitions.
+Our strategy is to bootstrap the familiar multi-atlas-based segmentation
+process from only a very small number of atlases.  In our previous work we used
+only a single atlas, based on serial histological slices, and validated the
+algorithm on 20 adolescent MR images.  In this work, we extensively investigate
+MAGeT used with differently sized atlas and template libraries, registration
+algorithms, label fusion strategies, and hippocampal definitions.  We evaluate
+our algorithm on a subset of images from the ADNI dataset, and compare our
+segmentations to the manual segmentations provided by ADNI.  Additionally, we
+evaluate the segmentations produced with MAGeT using hippocampal definitions
+from five high-resolution T1 in-vivo atlases (Winterburn2012). <conclude with
+summary that this is detailed validation the algorithm in order to pick the
+best operating parameters, whereas our previous work showed the feasibility of
+the approach>
 
-The essential insight of MAGeT brain, generating a template library, is not new.
-Heckemann2006 compared a multi-atlas-based segmentation method to that of generating a template library from a single atlas and found poor performance so did not investigate this approach any further.
-The LEAP algorithm (Wolz2010) begins with 30 atlases and labels the most similar unlabeled images.
-The labeled images from this step are included in the atlas set for the following step, and this proceeds until all images have been labeled.
-<expand; include other multi-atlas methods too?
->
+The essential insight of MAGeT brain, generating a template library, is not
+new. Heckemann2006 compared a multi-atlas-based segmentation method to that of
+generating a template library from a single atlas and found poor performance so
+did not investigate this approach any further. The LEAP algorithm (Wolz2010)
+begins with 30 atlases and labels the most similar unlabeled images. The
+labeled images from this step are included in the atlas set for the following
+step, and this proceeds until all images have been labeled. <expand; include
+other multi-atlas methods too? >
 
-In this work we conduct a thorough validation of the MAGeT brain approach.
-Our aim is not to improve on segmentation accuracy beyond existing methods, but instead to provide a method that trades off manual segmentation expertise for computational processing time whilst providing useful accuracy for the research and clinical setting.
-<blah blah..
-this sucks>
+In this work we conduct a thorough validation of the MAGeT brain approach. Our
+aim is not to improve on segmentation accuracy beyond existing methods, but
+instead to provide a method that trades off manual segmentation expertise for
+computational processing time whilst providing useful accuracy for the research
+and clinical setting. <blah blah.. this sucks>
 
@@
 \subsection{Segmentation Algorithm}
-The segmentation approach we follow is an extension of traditional multi-atlas segmentation.
-In multi-atlas segmentation, the inputs are a template library consisting of a set of manually labelled MR images (atlases), and a set of unlabelled MR images to be labelled.
-Each unlabelled MR image is segmented by first propagating to it the labels from the template library images via image registration, and then these candidate labels are combined by a label fusion method (e.g.
-voxel-wise majority voting).
-The label fusion method used may determine which templates are used to segment a particular subject (for instance, voxel-wise majority vote takes labels from all templates, whereas a weighted majority vote selects only those templates most similar to the subject).
+The segmentation approach we follow is an extension of traditional multi-atlas
+segmentation. In multi-atlas segmentation, the inputs are a template library
+consisting of a set of manually labelled MR images (atlases), and a set of
+unlabelled MR images to be labelled.  Each unlabelled MR image is segmented by
+first propagating to it the labels from the template library images via image
+registration, and then these candidate labels are combined by a label fusion
+method (e.g. voxel-wise majority voting). The label fusion method used may
+determine which templates are used to segment a particular subject (for
+instance, voxel-wise majority vote takes labels from all templates, whereas a
+weighted majority vote selects only those templates most similar to the
+subject).
 
-Our extension to this approach is to add a preliminary stage in which we construct the template library using a much smaller set of atlases and a subset of the given unlabelled images.
-To begin, we (possibly arbitrarily) select from the unlabelled images to form the template library.
-Multiple labellings are given to each image in the template library via label propagation from each of the atlases (label fusion is not conducted).
-Traditional multi-atlas segmentation is then used to produce segmentations for the entire set of unlabelled images (including those used in the template library).
-Our algorithm is illustrated in pseudocode in Algorithm 1, and diagrammatically in Figure 1.
+Our extension to this approach is to add a preliminary stage in which we
+construct the template library using a much smaller set of atlases and a subset
+of the given unlabelled images.  To begin, we (possibly arbitrarily) select
+from the unlabelled images to form the template library.  Multiple labellings
+are given to each image in the template library via label propagation from each
+of the atlases (label fusion is not conducted).  Traditional multi-atlas
+segmentation is then used to produce segmentations for the entire set of
+unlabelled images (including those used in the template library).  Our
+algorithm is illustrated in pseudocode in Algorithm 1, and diagrammatically in
+Figure 1.
 
-- Essentially, we add an extra layer of label propagation between the set of atlases 
-- This allows the template library to better reflect the constitution of the images to be labelled.
+- Essentially, we add an extra layer of label propagation between the set of
+  atlases 
+- This allows the template library to better reflect the constitution of the
+  images to be labelled.
 
@@
 
-Before images were registered, the N3 algorithm [16] is first used to minimize the intensity nonuniformity in each of the atlases and unlabeled subject images.
-In this work we investigated the performance of two registration methods.
+Before images were registered, the N3 algorithm [16] is first used to minimize
+the intensity nonuniformity in each of the atlases and unlabeled subject
+images. In this work we investigated the performance of two registration
+methods.
 
@@
 
-The ANIMAL algorithm carries out Image registration is in two phases.
-In the first, a 12-parameter linear transformation (3 translations, rotations, scales, shears) is estimated between images using an algorithm that maximizes the correlation between blurred MR intensities and gradient magnitude over the whole brain [3].
-In the second phase, nonlinear registration is completed using the ANIMAL algorithm [4]: an iterative procedure that estimates a 3D deformation field between two MR images.
-At first, large deformations are estimated using blurred version of the input data.
-These larger deformations are then input to subsequent steps where the fit is refined by estimating smaller deformations on data blurred with a Gaussian kernel with a smaller FWHM.
-The final transformation is a set of local translations defined on a bed of equally spaced nodes that were estimated through the optimization of the correlation coefficient.
+The ANIMAL algorithm carries out Image registration is in two phases. In the
+first, a 12-parameter linear transformation (3 translations, rotations, scales,
+shears) is estimated between images using an algorithm that maximizes the
+correlation between blurred MR intensities and gradient magnitude over the
+whole brain [3]. In the second phase, nonlinear registration is completed using
+the ANIMAL algorithm [4]: an iterative procedure that estimates a 3D
+deformation field between two MR images. At first, large deformations are
+estimated using blurred version of the input data. These larger deformations
+are then input to subsequent steps where the fit is refined by estimating
+smaller deformations on data blurred with a Gaussian kernel with a smaller
+FWHM. The final transformation is a set of local translations defined on a bed
+of equally spaced nodes that were estimated through the optimization of the
+correlation coefficient. 
 \todo{should add the MNI website here; also we should add the parameters we used somewhere}
@@
 
-ANTs is a diffeomorphic registration algorithm which provides great flexibility over the choice of transformation model, objective function, and the consistency of the final transformation.
-The transformation is estimated in a hierarchical fashion where the MRI data is subsampled, allowing large deformations to be estimated and successively refined at later hierarchical stages (where the data is subsampled to a finer grid).
-The deformation field and the objective function are regularized with a Gaussian kernel at each level of the hierarchy.
-The ANTs algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}.
-We used an implementation of the ANTS algorithm compatible with the MINC data format, mincANTS \url{https://github.com/vfonov/mincANTS} \todo{(same here we should add the parameters we used).}
+ANTs is a diffeomorphic registration algorithm which provides great flexibility
+over the choice of transformation model, objective function, and the
+consistency of the final transformation. The transformation is estimated in a
+hierarchical fashion where the MRI data is subsampled, allowing large
+deformations to be estimated and successively refined at later hierarchical
+stages (where the data is subsampled to a finer grid). The deformation field
+and the objective function are regularized with a Gaussian kernel at each level
+of the hierarchy. The ANTs algorithm is freely available
+\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the ANTS
+algorithm compatible with the MINC data format, mincANTS
+\url{https://github.com/vfonov/mincANTS} \todo{(same here we should add the
+parameters we used).}
 
@@
 
-Label fusion is term given to the process of combining the information from several candidate labellings for an MR image into a single labelling.
-In this paper we explore the benefits of three different fusion methods.
+Label fusion is term given to the process of combining the information from
+several candidate labellings for an MR image into a single labelling.  In this
+paper we explore the benefits of three different fusion methods.
 
@@
 
-Labels are propagated from all template library images to a subject.
-Each output voxel is given the most frequent label at that voxel location amongst all candidate labellings.
-Ties are broken arbitrarily.
+Labels are propagated from all template library images to a subject.  Each
+output voxel is given the most frequent label at that voxel location amongst
+all candidate labellings.  Ties are broken arbitrarily.
 
@@
 
-Labels are propagated from only those template library images which most similar to the subject.  
-A voxel-wise majority vote is carried out from the labels from the top n ranked template library 
-images, where n is parameter set by the user.
+Labels are propagated from only those template library images which most
+similar to the subject.  A voxel-wise majority vote is carried out from the
+labels from the top n ranked template library images, where n is parameter set
+by the user.
 
-Similarity between a subject and a template image is measured using normalized cross-correlation of intensity values, over a region of interest defined on each template, after the subject has been linearly registered to the template.
-The heuristic used for defining a meaningful region of interest for each template is to use the extent of all the propagated labellings from each atlas (after linear registration only) and then dilating this region by three voxels.
+Similarity between a subject and a template image is measured using normalized
+cross-correlation of intensity values, over a region of interest defined on
+each template, after the subject has been linearly registered to the template.
+The heuristic used for defining a meaningful region of interest for each
+template is to use the extent of all the propagated labellings from each atlas
+(after linear registration only) and then dilating this region by three voxels.
 
-The utility to calculate the cross-correlation similarity measure is implemented as part of the ANIMAL toolkit. \todo{(ANIMAL or MINC)}
+The utility to calculate the cross-correlation similarity measure is
+implemented as part of the ANIMAL toolkit. \todo{(ANIMAL or MINC)}
 
@@
 
-This label fusion method is identical to the above process except a normalised mutual information score is used over the region of interest between a template and subject instead of cross-correlation.
+This label fusion method is identical to the above process except a normalised
+mutual information score is used over the region of interest between a template
+and subject instead of cross-correlation.
 
-The utility to calculate the normalised mutual information measure is implemented as part of the EZMinc package (based on ITK NMI routine).
+The utility to calculate the normalised mutual information measure is
+implemented as part of the EZMinc package (based on ITK NMI routine).
 
@@
 
-Each segmentation was evaluated against the 'gold-standard' manual segmentation from the dataset using the Dice Kappa ($\kappa$) overlap metric:
+Each segmentation was evaluated against the 'gold-standard' manual segmentation
+from the dataset using the Dice Kappa ($\kappa$) overlap metric:
 
@@
 
-where $a$ is the number of voxels common to the candidate segmentation and the gold standard and $b+c$ is the sum of the voxels uniquely identified by either the automatically generated candidate segmentation or the gold-standard.
+where $a$ is the number of voxels common to the candidate segmentation and the
+gold standard and $b+c$ is the sum of the voxels uniquely identified by either
+the automatically generated candidate segmentation or the gold-standard.
 \todo{I would do this after you define the experiments}
@@
 
-Three different experiments were conducted to evaluate the influence of registration method, atlas set size and resolution, template set size, and label fusion method and parameters.
+Three different experiments were conducted to evaluate the influence of
+registration method, atlas set size and resolution, template set size, and
+label fusion method and parameters.
 
@@
 
-In this experiment we performed repeated random sub-sampling cross-validation of the MAGeT algorithm with a pool of 69 images and labels from the ADNI dataset.
-The images are T1-weighted \todo{<insert description of scan/machine types>}.
-The hippocampal labels are provided as part of the ADNI dataset, and are produced \todo{<insert description of SNT labellings>}.
-The parameters we varied, and the ranges over which we varied are listed in \todo{table}.
-We performed 10 validations per parameter combination.
-In each validation trial, a random sample of subjects were chosen from the pool without replacement as the atlas set, whereas the template set was chosen randomly with replacement.
+In this experiment we performed repeated random sub-sampling cross-validation
+of the MAGeT algorithm with a pool of 69 images and labels from the ADNI
+dataset.  The images are T1-weighted \todo{<insert description of scan/machine
+types>}.  The hippocampal labels are provided as part of the ADNI dataset, and
+are produced \todo{<insert description of SNT labellings>}.  The parameters we
+varied, and the ranges over which we varied are listed in \todo{table}.  We
+performed 10 validations per parameter combination.  In each validation trial,
+a random sample of subjects were chosen from the pool without replacement as
+the atlas set, whereas the template set was chosen randomly with replacement.
 
@@
  \end{verbatim}
-These settings were adapted from the "reasonable starting point" given in the ANTS manual. \todo{<rationale?>}
+These settings were adapted from the "reasonable starting point" given in the
+ANTS manual. \todo{<rationale?>}
 
-For the purposes of this work we used the regularization parameters optimized in Robbins et al.[15]. \todo{right? (but added a second layer).}
+For the purposes of this work we used the regularization parameters optimized
+in Robbins et al.[15]. \todo{right? (but added a second layer).}
 
@@
 
-In this experiment we explored using the MAGeT brain algorithm with the Winterburn high-resolution atlases to segment 21 randomly chosen MR images from the ADNI dataset (seven each of healthy, MCI and AD subjects).
-The Winterburn atlases are digital segmentations of the hippocampus in five  in-vivo 300u isotropic T1-weighted MR scans, and include subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.
-Subjects in the Winterburn atlases range in age from 29-57 years (mean age of 37), and include two males and three females.
+In this experiment we explored using the MAGeT brain algorithm with the
+Winterburn high-resolution atlases to segment 21 randomly chosen MR images from
+the ADNI dataset (seven each of healthy, MCI and AD subjects).  The Winterburn
+atlases are digital segmentations of the hippocampus in five  in-vivo 300u
+isotropic T1-weighted MR scans, and include subfield segmentations for the
+cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.
+Subjects in the Winterburn atlases range in age from 29-57 years (mean age of
+37), and include two males and three females.
 
-Since hippocampal segmentation protocols differ between the ADNI labels and Winterburn atlases, this poses a problem for direct similarity comparisons between labels produced by MAGeT brain and the ADNI labels.
-\todo{explain why we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate the performance of MAGeT brain, we compared classification accuracy of subjects by diagnosis based on hippocampal volume using both the SMT labels and our produced labels.
-\todo{description of validation -- rms {\tt validate} or t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}
+Since hippocampal segmentation protocols differ between the ADNI labels and
+Winterburn atlases, this poses a problem for direct similarity comparisons
+between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
+we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate
+the performance of MAGeT brain, we compared classification accuracy of subjects
+by diagnosis based on hippocampal volume using both the SMT labels and our
+produced labels.  \todo{description of validation -- rms {\tt validate} or
+t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}
 
@@
 - consistently performs better than average naive performance by XXX
-- using ANTS, with a large enough template library (>12) MAGeT brain performs better than the average multi-atlas approach with the same number of atlases.
+- using ANTS, with a large enough template library (>12) MAGeT brain performs
+  better than the average multi-atlas approach with the same number of atlases.
   using ANIMAL, 5 or more atlases needed before boost seen. 
@@
 
-- A2A shows that if atlas population strongly(?) represents subject set variability, then free choice from atlas population will produce improvements (we know this b/c of extensive validation trials).
-- what about in the case where atlas population doesn???t strongly represent subject set variability (e.g. a priori atlas set)?  then, we can use atlas selection to refine atlas set?
+- A2A shows that if atlas population strongly(?) represents subject set
+  variability, then free choice from atlas population will produce improvements
+(we know this b/c of extensive validation trials). 
+- what about in the case where atlas population doesn???t strongly represent
+  subject set variability (e.g. a priori atlas set)?  then, we can use atlas
+selection to refine atlas set?
 

### 2012-11-21 paper.Rnw
Updated figures with sensible labels in preparation for Mallar's talk

@@
-\documentclass{article}
-\usepackage[margin=1.5in]{geometry}
-\usepackage{algorithm, color,algpseudocode, amsmath, url, geometry, ctable }
-
-\newcommand{\todo}[1]{
-    \addcontentsline{tdo}{todo}{\protect{#1}}
-    \textcolor{red}{#1}
-}
-
-\begin{document}
 \SweaveOpts{echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}
-
 <<loadpackages, results="hide", echo=FALSE, cache=FALSE>>=
@@
 
+\documentclass{article}
+\usepackage[margin=1.5in]{geometry}
+\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
+\usepackage[square,numbers]{natbib}
+\setlength{\marginparwidth}{1.2in}
+\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
+\newcommand{\todo}[1]{\marginnote{\textcolor{red}{TODO #1}}}
+\newcommand{\mc}[1]{\todo{(MC) #1}}
+
+%{\marginpar{ \textcolor{red}{\\ TODO: #1 \\} }
+\begin{document}
+
 \title{Hippocampal segmentation with MAGeT}
@@
 \maketitle
+
+\begin{abstract}
+Neuroimaging research often relies on automated anatomical segmentations of MR
+images of the brain. Current multi-atlas based approaches provide accurate
+segmentations of brain images by propagating manually derived segmentations of
+specific neuroanatomical structures to unlabelled data. These approaches often
+rely on a large number of such manually segmented atlases that take significant
+time and expertise to produce. We present an algorithm for the automatic
+segmentation of the hippocampus that minimizes the number of atlases needed
+while still achieving similar accuracy to multi-atlas approaches.
+\mc{Finish this... }
+\end{abstract}
 
@@
 schizophrenia, and has functional significance in cognitive processes such as
-learning and memory.  For many research questions involving neuroimaging data,
-accurate identification of the hippocampus and its subregions in participant MR
-images is a necessary first step in order to analyse neuroanatomical
-differences and changes in  sample populations.  \todo{expand}
+learning and memory.  For many research questions involving magnetic resonance
+imaging (MRI) data accurate identification of the hippocampus and its
+subregions is a necessary first step to better understand the individual
+neuroanatomy of subjects.
 
 Currently, the gold standard for neuroanatomical segmentation is manual
-labelling by an expert human rater.  This is problematic for several reasons.
-Practically, manual segmentation takes a significant investment of time and
-expertise (Hammers2003) which may not be readily available to researchers or
-clinicians.  <blah> estimates that expert segmentation can take anywhere
-between 45 minutes to several days per image depending on the image resolution
-and protocol.  The amount of data produced in neuroimaging experiments
-increasingly exceeds the capacity for identification of specific
-neuroanatomical structures by an expert manual rater.  It may be difficult for
-laboratories to find or train, someone with the expertise to do such
-segmentation.
+delineation by an expert human rater.  This is problematic for segmentation of
+the hippocampus for several reasons.  First, manual segmentation takes a
+significant investment of time and expertise \citep{Hammers2003} which may not
+be readily available to researchers or clinicians.  Second, the amount of data
+produced in neuroimaging experiments increasingly exceeds the capacity for
+identification of specific neuroanatomical structures by an expert manual
+rater.  Third, the true definition of hippocampal anatomy in MR images is
+disputed \citep{Geuze2004}, as evidenced by efforts to create an unified
+segmentation protocol \citep{Jack2011}.
 
-Compounding this is that the hippocampus is a problematic structure for
-segmentation. Its very definition is disputed and varies widely in the
-literature (Geuze2004).  This has led to a variety of competing segmentation
-protocols in the literature which researchers can adhere to, including
-necessary ad-hoc modifications to adapt the definition in order to answer a
-particular research question.  Efforts are underway to create a unified
-protocol (Jack2011) for volumetry applications. But, even with a unified
-protocol, in order to investigate certain research questions it may be
-necessary to select from different hippocampal definitions. For example,
-(Poppenk2011) found that overall hippocampal volume change did not predict
-recollection memory performance, but dividing the hippocampus into anterior and
-posterior regions did.
+Compounding each of these problems is the significant neuroanatomical
+variability in the hippocampus throughout the course of aging, development, and
+neuropsychiatric disorders \citep{Mouiha2011}.  Additionally, it may be
+necessary to use several different hippocampal definitions or, in fact, make
+specific modifications in the course of research. For example, Poppenk et al.
+\citep{Poppenk2011} found that subdividing the hippocampus into anterior and
+posterior regions resulted in a predictive relationship between volume
+difference of those regions and recollection memory performance.  Thus, while
+manual segmentation of the hippocampus is a necessary technique, to researchers
+or clinicians who do not have access to the needed human expertise its use may
+be infeasible.
 
-\todo{this bit needs to be a little clearer - was it just the division of the
-HC that led to the improved understanding or was it that spatially localized
-differences in volumes are a better way to describe the difference).}
+Automated segmentation techniques overcome the need for human expertise by
+performing segmentations computationally.  A popular class of automated
+methods, {\it multi-atlas-based segmentation}, rely on a set of expertly
+labeled neuroanatomical atlases. Each atlas is warped to fit a subject's
+neuroanatomy using nonlinear registration
+techniques \citep{Collins1995,Klein2009}.  Atlas labels are then transformed by
+this warping and a {\it label fusion} technique, such as voxel-wise voting, is
+used to merge the competing label definitions from each atlas into a final
+segmentation for a subject.
 
-\todo{- Also, other situations: developmental studies -- need different
-definitions depending on the age range of subjects -- additionally, may want to
-tweak definition to test certain hypotheses (e.g. include subiculum: yes/no,
-more or less conservative WM/GM boundaries, compare with definitions from a
-different study,, etc...).}
+Many descriptions of multi-atlas-based segmentation algorithms report relying
+on an atlas library containing anywhere between 30 and 80 expertly labeled
+brains \citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.  As
+noted, the production of an atlas library requires significant manual effort,
+and is limited since the choice of atlases or segmentation protocol may not
+reflect the underlying neuroanatomical variability of the population under
+study or be suited to answer the research questions at hand.
 
-One can also expect neuroanatomical variability in this structure as it loses
-volume through aging and in the time course of different diseases.  \todo{This
-bit isn't really "joined to the previous statement"  may need a little glue
-here}
+In this paper we propose an automated segmentation method to address the above
+issues of existing multi-atlas-based methods. Principally, our method aims to
+dramatically reduce the number of manually labelled atlases needed (under 10).
+This is achieved by using the small atlas library to boot-strap a much larger
+"template library", which is then used to segment the subjects in a similar
+fashion to basic multi-atlas segmentation. This approach has the additional
+advantage of using the unique subject population on hand to initialize the
+segmentation process and improve accuracy.
 
-Automated segmentation techniques are available and can minimize several of the
-confounds often observed in studies that involve manual volumetry, such as the
-time and expertise required for segmentation and intra- and inter-rater
-reliability.  A class of segmentation approaches, known as atlas-based
-segmentation (also, template-based segmentation), make use of expertly labeled
-neuroanatomical atlases. An atlas MR image is warped to fit the subject
-neuroanatomy using nonlinear registration techniques [4, 11], and then the
-atlas labels are propagated via the registration transformation from the atlas
-to match the unique neuroanatomy of a particular subject.  Segmentation
-accuracy is affected by nonlinear registration and resampling errors, and by
-the extent to which the atlas neuroanatomy differs from the underlying
-neuroanatomical variability of the population under study (citation needed).
-To address these limitations, several atlases can be used (multi-atlas
-segmentation). Labels from each atlas are propagated to the subjects, by way of
-an nonlinear image registration, producing as many labellings for each subject
-as atlases. A label fusion technique is employed to merge the candidate
-labellings for a subject into a single segmentation using a decision rule (for
-instance, voxel-wise voting). Not all of the atlases need be used to segment
-each subject. Selecting only those atlases that are most similar to the subject
-image has been shown to significantly improve overall registration and
-segmentation accuracy (Aljabar 2008, Collins2010, Leung2010, Lotjonen2010).
+The essential insight of generating a template library is not new.  Heckemann
+\citep{Heckemann2006} compared generating a template library from a single atlas
+to standard multi-atlas segmentation and found poor performance and so deemed
+the approach as inviable.  The LEAP algorithm \citep{Wolz2010} proceeds by
+iteratively segmenting the unlabelled image most similar to the atlas library
+images and then incorporating the now-labelled image into the atlas library,
+but requires 30 starting atlases.  The novelty of our method is to demonstrate
+the possibility of producing comparable segmentation accuracy to these and
+other multi-atlas-based methods while using significantly fewer manually
+created atlases.
 
-\todo{briefly discuss other automated segmentation methods, and justify our
-focus on atlas-based methods}
-
-In summary, whilst manual segmentation is the gold standard for neuroanatomical
-segmentation, it is time-consuming and requires an investment in expertise and
-training, and may be inflexible in the face of differing hippocampal
-definitions or changes. Atlas-based automated segmentation dramatically reduces
-the time needed for segmentation, and can be completed without any expert rater
-as long as an appropriate atlas library is available <be more specific (e.g.
-shares protocol, demographics, phenotypes of subject population)>.  In the
-situation where such an atlas library does not exist, then an expert rater is
-needed. Many automated approaches rely on an atlas library containing between
-30 and 80 expertly labeled brains (Heckemann2006, Collins2010) which, even
-though producing such a library may require less manual work than segmenting
-the entire set of subjects, it still poses a substantial time commitment.
-
-In this paper we extend our automated method for subcortical segmentation
-(striatum, globus pallidus, and thalamus; Chakravarty2012) to hippocampal
-segmentation. This method, known as MAGeT brain, is suited for situations where
-manual effort is at a premium, and flexibility is needed in atlas definitions.
-Our strategy is to bootstrap the familiar multi-atlas-based segmentation
-process from only a very small number of atlases.  In our previous work we used
-only a single atlas, based on serial histological slices, and validated the
-algorithm on 20 adolescent MR images.  In this work, we extensively investigate
-MAGeT used with differently sized atlas and template libraries, registration
-algorithms, label fusion strategies, and hippocampal definitions.  We evaluate
-our algorithm on a subset of images from the ADNI dataset, and compare our
-segmentations to the manual segmentations provided by ADNI.  Additionally, we
-evaluate the segmentations produced with MAGeT using hippocampal definitions
-from five high-resolution T1 in-vivo atlases (Winterburn2012). <conclude with
-summary that this is detailed validation the algorithm in order to pick the
-best operating parameters, whereas our previous work showed the feasibility of
-the approach>
-
-The essential insight of MAGeT brain, generating a template library, is not
-new. Heckemann2006 compared a multi-atlas-based segmentation method to that of
-generating a template library from a single atlas and found poor performance so
-did not investigate this approach any further. The LEAP algorithm (Wolz2010)
-begins with 30 atlases and labels the most similar unlabeled images. The
-labeled images from this step are included in the atlas set for the following
-step, and this proceeds until all images have been labeled. <expand; include
-other multi-atlas methods too? >
-
-In this work we conduct a thorough validation of the MAGeT brain approach. Our
-aim is not to improve on segmentation accuracy beyond existing methods, but
-instead to provide a method that trades off manual segmentation expertise for
-computational processing time whilst providing useful accuracy for the research
-and clinical setting. <blah blah.. this sucks>
+In our previous work \citep{Chakravarty2012}, we applied MAGeT brain to
+segmentation of the human striatum, globus pallidus, and thalamus using a
+single histologically-derived atlas. The main contribution of this paper is to
+extend our approach to the human hippocampus and perform a thorough validation
+over a range of atlas and template library sizes, which was not done in our
+previous work.  Due to the small number of atlases required, our method can
+easily accommodate different hippocampal definitions. Our aim is not to improve
+on segmentation accuracy beyond existing methods, but instead to provide a
+method that trades off manual segmentation expertise for computational
+processing time while providing sufficient accuracy for clinical and research
+applications.
 
 \section{Materials and Methods}
+
 explain two registration methods, definitions, fusion methods. 
@@
 \subsection{Segmentation Algorithm}
-The segmentation approach we follow is an extension of traditional multi-atlas
-segmentation. In multi-atlas segmentation, the inputs are a template library
-consisting of a set of manually labelled MR images (atlases), and a set of
-unlabelled MR images to be labelled.  Each unlabelled MR image is segmented by
-first propagating to it the labels from the template library images via image
-registration, and then these candidate labels are combined by a label fusion
-method (e.g. voxel-wise majority voting). The label fusion method used may
-determine which templates are used to segment a particular subject (for
-instance, voxel-wise majority vote takes labels from all templates, whereas a
-weighted majority vote selects only those templates most similar to the
-subject).
+\todo{Make sure we explain what MAGeT stands for}
+In this paper, we use the term {\it atlas} to mean any manually segmented MR
+image, and the term {\it atlas library} to mean a set of such images.  We use
+the term {\it template} to refer to any MR image, and associated labelling,
+used to segment another image, and the term {\it template library} to refer to
+a set of such images.  An atlas library may be used as a template library but,
+as we will discuss, a template library may also be composed of images with
+computer generated labellings.
 
-Our extension to this approach is to add a preliminary stage in which we
-construct the template library using a much smaller set of atlases and a subset
-of the given unlabelled images.  To begin, we (possibly arbitrarily) select
-from the unlabelled images to form the template library.  Multiple labellings
-are given to each image in the template library via label propagation from each
-of the atlases (label fusion is not conducted).  Traditional multi-atlas
-segmentation is then used to produce segmentations for the entire set of
-unlabelled images (including those used in the template library).  Our
-algorithm is illustrated in pseudocode in Algorithm 1, and diagrammatically in
-Figure 1.
+The segmentation approach we propose is best understood as an extension of
+basic multi-atlas segmentation \citep{Collins2010}.  In multi-atlas segmentation,
+an atlas library and unlabelled MR images are given as input.  Every atlas
+image is nonlinearly registered to each unlabelled image, and then each atlas'
+labels are propagated via the resulting transformations.  These labels are then
+fused to produce a single, definitive segmentation by some label fusion method
+(e.g.  voxel-wise majority vote).
 
-- Essentially, we add an extra layer of label propagation between the set of
-  atlases 
-- This allows the template library to better reflect the constitution of the
-  images to be labelled.
+Our extension adds a preliminary stage in which a template library is
+constructed from input images, and used in place of an atlas library in the
+standard multi-atlas-based method.  To create the template library, labels from
+each atlas image are propagated to each template library image via the
+transformation resulting from a non-linear registration between pair of images.
+As a result, each template library image has a label from each atlas.
+Basic multi-atlas segmentation is then used to produce segmentations for the
+entire set of unlabelled images (including those images used in the template
+library).
+
+Label fusion is performed by cross-correlation weighted voting, a strategy
+weighted towards an optimal combination of subjects from the template library
+which has been previously shown to improve segmentation
+accuracy \citep{Aljabar2009,Collins2010}. In this method, each template library
+image is ranked in similarity to each unlabelled image by the normalized
+cross-correlation of image intensities after linear registration in a region of
+interest (ROI) generously encompassing the hippocampus.  Only the top ranked
+template library image labels are used in a voxel-wise majority vote. The ROI
+is heuristically defined as the extent of all atlas labels after linear
+registration to the template, dilated by three voxels \citep{Chakravarty2012}.
 
@@
 correlation between blurred MR intensities and gradient magnitude over the
-whole brain [3]. In the second phase, nonlinear registration is completed using
-the ANIMAL algorithm [4]: an iterative procedure that estimates a 3D
+whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
+the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
 deformation field between two MR images. At first, large deformations are
@@
 correlation coefficient.
+
+For the purposes of this work we used the
+regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
+should be noted that the MAGeT brain algorithm is not dependent on this, or
+any, particular choice of registration method \citep{Chakravarty2012}.
+
 \todo{should add the MNI website here; also we should add the parameters we used somewhere}
@@
 <<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),colour=variable,geom="boxplot")
+qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),
+      measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),
+      colour=variable,geom="boxplot") + 
+        xlab("Diagnosis") + ylab("Total Hippocampal Volume (cm3)")
 @
@@
 <<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
-qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),measure.vars=c("FSL","MAPER","MAGeT")),colour=variable, geom="smooth", method="lm")
+qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),
+      measure.vars=c("FSL","MAPER","MAGeT")),
+      colour=variable, geom="smooth", method="lm") + 
+    geom_point() +
+    xlab("Total Hippocampal Volume (Manual Segmentation)") + 
+    ylab("Total Hippocampal Volume (Automatic Segmentation)")
 @
-  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
   \label{ADNI-baseline-volumes-plot}
@@
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
-qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm")
+qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm") + 
+    xlab("Total Hippocampal Volume (Manual Segmentation)") + 
+    ylab("Total Hippocampal Volume (MAGeT Segmentation)")
+
 @
-  \caption{Comparison of total HC volumes for MAGeT against manually rated volumes}
+  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total HC volumes for MAGeT against manually rated volumes of}
   \label{SZ_volumes}

### 2012-11-21 paper.tex
Updated figures with sensible labels in preparation for Mallar's talk

@@
 \usepackage[margin=1.5in]{geometry}
-\usepackage{algorithm, color,algpseudocode, amsmath, url, geometry, ctable }
+\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
+\usepackage[square,numbers]{natbib}
+\setlength{\marginparwidth}{1.2in}
+\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
+\newcommand{\todo}[1]{\marginnote{\textcolor{red}{TODO #1}}}
+\newcommand{\mc}[1]{\todo{(MC) #1}}
 
-\newcommand{\todo}[1]{
-    \addcontentsline{tdo}{todo}{\protect{#1}}
-    \textcolor{red}{#1}
-}
-
+%{\marginpar{ \textcolor{red}{\\ TODO: #1 \\} }
 \begin{document}
@@
 
+\begin{abstract}
+Neuroimaging research often relies on automated anatomical segmentations of MR
+images of the brain. Current multi-atlas based approaches provide accurate
+segmentations of brain images by propagating manually derived segmentations of
+specific neuroanatomical structures to unlabelled data. These approaches often
+rely on a large number of such manually segmented atlases that take significant
+time and expertise to produce. We present an algorithm for the automatic
+segmentation of the hippocampus that minimizes the number of atlases needed
+while still achieving similar accuracy to multi-atlas approaches.
+\mc{Finish this... }
+\end{abstract}
+
 \section{Introduction}
 
-The hippocampus is of particular interest to many researchers because it is implicated in forms of brain dysfunction such as Alzheimer's disease and schizophrenia, and has functional significance in cognitive processes such as learning and memory.
-For many research questions involving neuroimaging data, accurate identification of the hippocampus and its subregions in participant MR images is a necessary first step in order to analyse neuroanatomical differences and changes in  sample populations.
-\todo{expand}
+The hippocampus is of particular interest to many researchers because it is
+implicated in forms of brain dysfunction such as Alzheimer's disease and
+schizophrenia, and has functional significance in cognitive processes such as
+learning and memory.  For many research questions involving magnetic resonance
+imaging (MRI) data accurate identification of the hippocampus and its
+subregions is a necessary first step to better understand the individual
+neuroanatomy of subjects.
 
-Currently, the gold standard for neuroanatomical segmentation is manual labelling by an expert human rater.
-This is problematic for several reasons.
-Practically, manual segmentation takes a significant investment of time and expertise (Hammers2003) which may not be readily available to researchers or clinicians.
-<blah> estimates that expert segmentation can take anywhere between 45 minutes to several days per image depending on the image resolution and protocol.
-The amount of data produced in neuroimaging experiments increasingly exceeds the capacity for identification of specific neuroanatomical structures by an expert manual rater.
-It may be difficult for laboratories to find or train, someone with the expertise to do such segmentation.
+Currently, the gold standard for neuroanatomical segmentation is manual
+delineation by an expert human rater.  This is problematic for segmentation of
+the hippocampus for several reasons.  First, manual segmentation takes a
+significant investment of time and expertise \citep{Hammers2003} which may not
+be readily available to researchers or clinicians.  Second, the amount of data
+produced in neuroimaging experiments increasingly exceeds the capacity for
+identification of specific neuroanatomical structures by an expert manual
+rater.  Third, the true definition of hippocampal anatomy in MR images is
+disputed \citep{Geuze2004}, as evidenced by efforts to create an unified
+segmentation protocol \citep{Jack2011}.
 
-Compounding this is that the hippocampus is a problematic structure for segmentation.
-Its very definition is disputed and varies widely in the literature (Geuze2004).
-This has led to a variety of competing segmentation protocols in the literature which researchers can adhere to, including necessary ad-hoc modifications to adapt the definition in order to answer a particular research question.
-Efforts are underway to create a unified protocol (Jack2011) for volumetry applications.
-But, even with a unified protocol, in order to investigate certain research questions it may be necessary to select from different hippocampal definitions.
-For example, (Poppenk2011) found that overall hippocampal volume change did not predict recollection memory performance, but dividing the hippocampus into anterior and posterior regions did.
+Compounding each of these problems is the significant neuroanatomical
+variability in the hippocampus throughout the course of aging, development, and
+neuropsychiatric disorders \citep{Mouiha2011}.  Additionally, it may be
+necessary to use several different hippocampal definitions or, in fact, make
+specific modifications in the course of research. For example, Poppenk et al.
+\citep{Poppenk2011} found that subdividing the hippocampus into anterior and
+posterior regions resulted in a predictive relationship between volume
+difference of those regions and recollection memory performance.  Thus, while
+manual segmentation of the hippocampus is a necessary technique, to researchers
+or clinicians who do not have access to the needed human expertise its use may
+be infeasible.
 
-\todo{this bit needs to be a little clearer - was it just the division of the HC that led to the improved understanding or was it that spatially localized differences in volumes are a better way to describe the difference).}
+Automated segmentation techniques overcome the need for human expertise by
+performing segmentations computationally.  A popular class of automated
+methods, {\it multi-atlas-based segmentation}, rely on a set of expertly
+labeled neuroanatomical atlases. Each atlas is warped to fit a subject's
+neuroanatomy using nonlinear registration
+techniques \citep{Collins1995,Klein2009}.  Atlas labels are then transformed by
+this warping and a {\it label fusion} technique, such as voxel-wise voting, is
+used to merge the competing label definitions from each atlas into a final
+segmentation for a subject.
 
-\todo{- Also, other situations: developmental studies -- need different definitions depending on the age range of subjects
--- additionally, may want to tweak definition to test certain hypotheses (e.g. include subiculum: yes/no, more or less 
- conservative WM/GM boundaries, compare with definitions from a different study,, etc...).}
+Many descriptions of multi-atlas-based segmentation algorithms report relying
+on an atlas library containing anywhere between 30 and 80 expertly labeled
+brains \citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.  As
+noted, the production of an atlas library requires significant manual effort,
+and is limited since the choice of atlases or segmentation protocol may not
+reflect the underlying neuroanatomical variability of the population under
+study or be suited to answer the research questions at hand.
 
-One can also expect neuroanatomical variability in this structure as it loses volume through aging and in the time course of different diseases.
-\todo{This bit isn't really "joined to the previous statement"  may need a little glue here}
+In this paper we propose an automated segmentation method to address the above
+issues of existing multi-atlas-based methods. Principally, our method aims to
+dramatically reduce the number of manually labelled atlases needed (under 10).
+This is achieved by using the small atlas library to boot-strap a much larger
+"template library", which is then used to segment the subjects in a similar
+fashion to basic multi-atlas segmentation. This approach has the additional
+advantage of using the unique subject population on hand to initialize the
+segmentation process and improve accuracy.
 
-Automated segmentation techniques are available and can minimize several of the confounds often observed in studies that involve manual volumetry, such as the time and expertise required for segmentation and intra- and inter-rater reliability.
-A class of segmentation approaches, known as atlas-based segmentation (also, template-based segmentation), make use of expertly labeled neuroanatomical atlases.
-An atlas MR image is warped to fit the subject neuroanatomy using nonlinear registration techniques [4, 11], and then the atlas labels are propagated via the registration transformation from the atlas to match the unique neuroanatomy of a particular subject.
-Segmentation accuracy is affected by nonlinear registration and resampling errors, and by the extent to which the atlas neuroanatomy differs from the underlying neuroanatomical variability of the population under study (citation needed).
-To address these limitations, several atlases can be used (multi-atlas segmentation).
-Labels from each atlas are propagated to the subjects, by way of an nonlinear image registration, producing as many labellings for each subject as atlases.
-A label fusion technique is employed to merge the candidate labellings for a subject into a single segmentation using a decision rule (for instance, voxel-wise voting).
-Not all of the atlases need be used to segment each subject.
-Selecting only those atlases that are most similar to the subject image has been shown to significantly improve overall registration and segmentation accuracy (Aljabar 2008, Collins2010, Leung2010, Lotjonen2010).
+The essential insight of generating a template library is not new.  Heckemann
+\citep{Heckemann2006} compared generating a template library from a single atlas
+to standard multi-atlas segmentation and found poor performance and so deemed
+the approach as inviable.  The LEAP algorithm \citep{Wolz2010} proceeds by
+iteratively segmenting the unlabelled image most similar to the atlas library
+images and then incorporating the now-labelled image into the atlas library,
+but requires 30 starting atlases.  The novelty of our method is to demonstrate
+the possibility of producing comparable segmentation accuracy to these and
+other multi-atlas-based methods while using significantly fewer manually
+created atlases.
 
-\todo{briefly discuss other automated segmentation methods, and justify our focus on atlas-based methods}
-
-In summary, whilst manual segmentation is the gold standard for neuroanatomical segmentation, it is time-consuming and requires an investment in expertise and training, and may be inflexible in the face of differing hippocampal definitions or changes.
-Atlas-based automated segmentation dramatically reduces the time needed for segmentation, and can be completed without any expert rater as long as an appropriate atlas library is available <be more specific (e.g.
-shares protocol, demographics, phenotypes of subject population)>.
-In the situation where such an atlas library does not exist, then an expert rater is needed.
-Many automated approaches rely on an atlas library containing between 30 and 80 expertly labeled brains (Heckemann2006, Collins2010) which, even though producing such a library may require less manual work than segmenting the entire set of subjects, it still poses a substantial time commitment.
-
-In this paper we extend our automated method for subcortical segmentation (striatum, globus pallidus, and thalamus; Chakravarty2012) to hippocampal segmentation.
-This method, known as MAGeT brain, is suited for situations where manual effort is at a premium, and flexibility is needed in atlas definitions.
-Our strategy is to bootstrap the familiar multi-atlas-based segmentation process from only a very small number of atlases.
-In our previous work we used only a single atlas, based on serial histological slices, and validated the algorithm on 20 adolescent MR images.
-In this work, we extensively investigate MAGeT used with differently sized atlas and template libraries, registration algorithms, label fusion strategies, and hippocampal definitions.
-We evaluate our algorithm on a subset of images from the ADNI dataset, and compare our segmentations to the manual segmentations provided by ADNI.
-Additionally, we evaluate the segmentations produced with MAGeT using hippocampal definitions from five high-resolution T1 in-vivo atlases (Winterburn2012).
-<conclude with summary that this is detailed validation the algorithm in order to pick the best operating parameters, whereas our previous work showed the feasibility of the approach>
-
-The essential insight of MAGeT brain, generating a template library, is not new.
-Heckemann2006 compared a multi-atlas-based segmentation method to that of generating a template library from a single atlas and found poor performance so did not investigate this approach any further.
-The LEAP algorithm (Wolz2010) begins with 30 atlases and labels the most similar unlabeled images.
-The labeled images from this step are included in the atlas set for the following step, and this proceeds until all images have been labeled.
-<expand; include other multi-atlas methods too?
->
-
-In this work we conduct a thorough validation of the MAGeT brain approach.
-Our aim is not to improve on segmentation accuracy beyond existing methods, but instead to provide a method that trades off manual segmentation expertise for computational processing time whilst providing useful accuracy for the research and clinical setting.
-<blah blah..
-this sucks>
+In our previous work \citep{Chakravarty2012}, we applied MAGeT brain to
+segmentation of the human striatum, globus pallidus, and thalamus using a
+single histologically-derived atlas. The main contribution of this paper is to
+extend our approach to the human hippocampus and perform a thorough validation
+over a range of atlas and template library sizes, which was not done in our
+previous work.  Due to the small number of atlases required, our method can
+easily accommodate different hippocampal definitions. Our aim is not to improve
+on segmentation accuracy beyond existing methods, but instead to provide a
+method that trades off manual segmentation expertise for computational
+processing time while providing sufficient accuracy for clinical and research
+applications.
 
 \section{Materials and Methods}
+
 explain two registration methods, definitions, fusion methods. 
@@
 \subsection{Segmentation Algorithm}
-The segmentation approach we follow is an extension of traditional multi-atlas segmentation.
-In multi-atlas segmentation, the inputs are a template library consisting of a set of manually labelled MR images (atlases), and a set of unlabelled MR images to be labelled.
-Each unlabelled MR image is segmented by first propagating to it the labels from the template library images via image registration, and then these candidate labels are combined by a label fusion method (e.g.
-voxel-wise majority voting).
-The label fusion method used may determine which templates are used to segment a particular subject (for instance, voxel-wise majority vote takes labels from all templates, whereas a weighted majority vote selects only those templates most similar to the subject).
+\todo{Make sure we explain what MAGeT stands for}
+In this paper, we use the term {\it atlas} to mean any manually segmented MR
+image, and the term {\it atlas library} to mean a set of such images.  We use
+the term {\it template} to refer to any MR image, and associated labelling,
+used to segment another image, and the term {\it template library} to refer to
+a set of such images.  An atlas library may be used as a template library but,
+as we will discuss, a template library may also be composed of images with
+computer generated labellings.
 
-Our extension to this approach is to add a preliminary stage in which we construct the template library using a much smaller set of atlases and a subset of the given unlabelled images.
-To begin, we (possibly arbitrarily) select from the unlabelled images to form the template library.
-Multiple labellings are given to each image in the template library via label propagation from each of the atlases (label fusion is not conducted).
-Traditional multi-atlas segmentation is then used to produce segmentations for the entire set of unlabelled images (including those used in the template library).
-Our algorithm is illustrated in pseudocode in Algorithm 1, and diagrammatically in Figure 1.
+The segmentation approach we propose is best understood as an extension of
+basic multi-atlas segmentation \citep{Collins2010}.  In multi-atlas segmentation,
+an atlas library and unlabelled MR images are given as input.  Every atlas
+image is nonlinearly registered to each unlabelled image, and then each atlas'
+labels are propagated via the resulting transformations.  These labels are then
+fused to produce a single, definitive segmentation by some label fusion method
+(e.g.  voxel-wise majority vote).
 
-- Essentially, we add an extra layer of label propagation between the set of atlases 
-- This allows the template library to better reflect the constitution of the images to be labelled.
+Our extension adds a preliminary stage in which a template library is
+constructed from input images, and used in place of an atlas library in the
+standard multi-atlas-based method.  To create the template library, labels from
+each atlas image are propagated to each template library image via the
+transformation resulting from a non-linear registration between pair of images.
+As a result, each template library image has a label from each atlas.
+Basic multi-atlas segmentation is then used to produce segmentations for the
+entire set of unlabelled images (including those images used in the template
+library).
+
+Label fusion is performed by cross-correlation weighted voting, a strategy
+weighted towards an optimal combination of subjects from the template library
+which has been previously shown to improve segmentation
+accuracy \citep{Aljabar2009,Collins2010}. In this method, each template library
+image is ranked in similarity to each unlabelled image by the normalized
+cross-correlation of image intensities after linear registration in a region of
+interest (ROI) generously encompassing the hippocampus.  Only the top ranked
+template library image labels are used in a voxel-wise majority vote. The ROI
+is heuristically defined as the extent of all atlas labels after linear
+registration to the template, dilated by three voxels \citep{Chakravarty2012}.
 
@@
 
-The ANIMAL algorithm carries out Image registration is in two phases.
-In the first, a 12-parameter linear transformation (3 translations, rotations, scales, shears) is estimated between images using an algorithm that maximizes the correlation between blurred MR intensities and gradient magnitude over the whole brain [3].
-In the second phase, nonlinear registration is completed using the ANIMAL algorithm [4]: an iterative procedure that estimates a 3D deformation field between two MR images.
-At first, large deformations are estimated using blurred version of the input data.
-These larger deformations are then input to subsequent steps where the fit is refined by estimating smaller deformations on data blurred with a Gaussian kernel with a smaller FWHM.
-The final transformation is a set of local translations defined on a bed of equally spaced nodes that were estimated through the optimization of the correlation coefficient.
+The ANIMAL algorithm carries out Image registration is in two phases. In the
+first, a 12-parameter linear transformation (3 translations, rotations, scales,
+shears) is estimated between images using an algorithm that maximizes the
+correlation between blurred MR intensities and gradient magnitude over the
+whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
+the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
+deformation field between two MR images. At first, large deformations are
+estimated using blurred version of the input data. These larger deformations
+are then input to subsequent steps where the fit is refined by estimating
+smaller deformations on data blurred with a Gaussian kernel with a smaller
+FWHM. The final transformation is a set of local translations defined on a bed
+of equally spaced nodes that were estimated through the optimization of the
+correlation coefficient.
+
+For the purposes of this work we used the
+regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
+should be noted that the MAGeT brain algorithm is not dependent on this, or
+any, particular choice of registration method \citep{Chakravarty2012}.
+
 \todo{should add the MNI website here; also we should add the parameters we used somewhere}
@@
 
-  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \caption{{\em ADNI Baseline cohort} Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
   \label{ADNI-baseline-volumes-plot}
@@
 
-  \caption{Comparison of total HC volumes for MAGeT against manually rated volumes}
+  \caption{{\em First Episode Schizophrenic Patients} Comparison of total HC volumes for MAGeT against manually rated volumes of}
   \label{SZ_volumes}

### 2012-11-22 paper.Rnw
Updated figures for Mallar's presentation

@@
       colour=variable,geom="boxplot") + 
-        xlab("Diagnosis") + ylab("Total Hippocampal Volume (cm3)")
+        xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
 @

### 2012-11-22 paper.tex
Updated figures for Mallar's presentation

@@
 
-  \caption{{\em ADNI Baseline cohort} Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
   \label{ADNI-baseline-volumes-plot}
@@
 
-  \caption{{\em First Episode Schizophrenic Patients} Comparison of total HC volumes for MAGeT against manually rated volumes of}
+  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total HC volumes for MAGeT against manually rated volumes of}
   \label{SZ_volumes}

### 2012-11-22 paper.Rnw
Calculating r/p-values for ADNI/SZFEP plots, but display code currently commented out due to ugliness

@@
 library(ADNIMERGE)
+options(digits=3)
 @
@@
 <<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),
-      measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),
+qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),
       colour=variable,geom="boxplot") + 
         xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
+
 @
@@
 <<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
-qplot(SNT,value,data=melt(totals,id.vars=c("ID","SNT"),
-      measure.vars=c("FSL","MAPER","MAGeT")),
+by_snt_volumes = melt(totals,id.vars=c("ID","SNT"), measure.vars=c("FSL","MAPER","MAGeT"))
+fsl=subset(by_snt_volumes, variable=="FSL")
+maper=subset(by_snt_volumes, variable=="MAPER")
+maget=subset(by_snt_volumes, variable=="MAGeT")
+
+qplot(SNT,value,data=by_snt_volumes,
       colour=variable, geom="smooth", method="lm") + 
     geom_point() +
+#     geom_text(aes(
+#       label=paste("FSL       R=", round(cor(fsl$SNT,fsl$value, use="complete.obs"), 2), 
+#                   ", p<", round(as.double(t.test(fsl$SNT,fsl$value, paired=T)["p.value"]), 2)),
+#       x=1800,y=10000, hjust=0,size=2)) + 
+#     geom_text(aes(
+#       label=paste("MAPER R=", round(cor(maper$SNT,maper$value, use="complete.obs"), 2), 
+#                   ", p<", t.test(maper$SNT,maper$value, paired=T)["p.value"]), 
+#       x=1800,y= 9500, hjust=0, size=2)) +
+#     geom_text(aes(
+#       label=paste("MAGeT  R=", round(cor(maget$SNT,maget$value, use="complete.obs"),2), 
+#                   ", p<", t.test(maget$SNT,maget$value, paired=T)["p.value"]), 
+#       x=1800,y= 9000, hjust=0, size=2)) +
     xlab("Total Hippocampal Volume (Manual Segmentation)") + 
@@
 qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm") + 
+#   geom_text(aes(
+#       label=paste("R=", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
+#                   ", p<", t.test(Manual, MAGeT, paired=T)["p.value"]),
+#       x=6500,y=10000, hjust=0,size=2)) +
     xlab("Total Hippocampal Volume (Manual Segmentation)") + 

### 2012-11-30 paper.Rnw
Update with full ADNI baseline info; Outlined rational for experiments

@@
 library(ADNIMERGE)
+library(plyr)
 options(digits=3)
+theme_set(theme_bw())
 @
@@
 while still achieving similar accuracy to multi-atlas approaches.
-\mc{Finish this... }
+\todo{Finish this... }
 \end{abstract}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                         Experiment Rationale 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\subsubsection{Study Method Rationale - An Outline}
+\begin{description}
+\item[ADNI Repeated Random Subsampling Validation] \hfill \\ 
+{\bf Purpose:} to show method leads to improved accuracy (over multi-atlas method), and is robust
+w.r.t.
+registration method, atlas and template library composition and size, and various fusion methods.
+Additionally, to choose the best configuration of parameters (under the assumption that performance generalises).
+\\ \\ 
+{\bf Method:} using a set of 69 ADNI subjects, 13 from each disease category, perform 
+10 rounds of repeated random subsampling to pick atlas library and template library 
+for each parameter setting and then run MAGeT brain to segment non-atlas subjects.
+Also perform multi-atlas and single-atlas segmentation using the same parameter settings
+and atlas library size.
+\\ \\
+{\bf Results:} find significant improvement over multi-atlas performed with the same
+parameters. Also, find smoothed performance is monotonically increasing but asymptotic
+in size of both template and atlas library, with peak performance reached after 15
+templates. \todo{Can we statistically capture "peak" performance?  The point at which 
+gains become statistically insignificant?}
+
+\item[Winterburn-Atlas ADNI Baseline Segmentation] \hfill \\
+{\bf Purpose:} given that MAGeT is robust and accurate, show MAGeT performance on 
+a real-world task in which atlas set may not be representative of subjects.
+\\ \\
+{\bf Method:} We use best settings from first experiment (ANTS, majority vote, 
+template library of 20), and use Winterburn hi-res atlases to segment all ADNI Baseline 
+images.  Since our segmentation protocol differs from what can be segmented at 
+ADNI resolution a direct comparison is not appropriate.  Instead, we validate by 
+showing how well our segmentation volumes correlate with existing automated/manual 
+segmentation volumes. \todo{Also show disease prediction?}
+\\ \\
+{\bf Results: }
+
+\item[First Episode SZ Patients Segmentation] \hfill \\ 
+{\bf Purpose:} validate that MAGeT's performance generalises to other diseases.
+\\ \\
+{\bf Method:} Measure performance using best parameter settings on a different disease dataset.
+We use FEP SZ patients and controls, using two different atlas sets: a manual hippocampal segmentation of patients, and Winterburn atlas set.
+We validate the native-atlas set segmentations using Dice's Kappa, and the Winterburn-atlas-set segmentations by correlating volumes.
+\\ \\
+{\bf Results:} High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
+
+\item[Winterburn Atlas Reliability] \hfill \\
+{\bf Purpose:} validation of performance on hippocampal subregion segmentation
+\\ \\
+{\bf Method:} we use the Winterburn atlas set, using each high-res atlas in turn 
+to segment a standard-res MR of the same subject and compare the automatic 
+segmentation to the atlas segmentation.  We use a template library consisting 
+of standard-res MRs from the remaining Winterburn atlas subjects plus 11 other 
+standard-res healthy control subjects. 
+\\ \\
+{\bf Results:} TBD
+
+\end{description}
+\end{kframe}
+
 \subsubsection{ADNI Validation Experiment}
@@
 <<ADNI-baseline-volumes-prep,echo=F,cache=T>>=
-data = read.csv('data/ADNI_baseline_volumes.csv')
-unscaled=subset(data,Scaled==0 & Baseline==1 & DX != '')
-attach(unscaled)
-totals = data.frame(LONIUID,L+R,UCSFFSL_L+UCSFSL_R,MAPER_L+MAPER_R,UCSFSNT_L + UCSFSNT_R,SNT_L + SNT_R,DX)
-colnames(totals) <- c("ID","MAGeT","FSL","MAPER","UCSFSNT","SNT","DX")
+yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
+maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
+snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
+ucsd  = read.csv("data/ADNI_baseline_volumes/UCSDVOL.csv")
+fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
+fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv")
+
+# recode VISCODE
+
+viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
+yr1$VISCODE <- as.factor(yr1$Visit)
+levels(yr1$VISCODE) <-viscode_levels
+
+yr1$Scaled_2 <- grepl("Scaled_2", yr1$Description)
+
+subjects = subset(yr1, VISCODE == 'bl' & Scaled_2 == FALSE)
+
+# fetch just the columns we need, and do a little renaming
+mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
+maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
+snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
+ucsd_pruned   = data.frame(RID=ucsd$RID, VISCODE=ucsd$VISCODE, UCSD_L=ucsd$LHIPPOC, UCSD_R=ucsd$RHIPPOC)
+fs_pruned     = subset(fs, TEMPQC != "Fail")
+fs_pruned     = data.frame(FS_L=fs_pruned$ST29SV, FS_R=fs_pruned$ST88SV, Source = fs_pruned$IMAGEUID)
+fsl_pruned    = data.frame(FSL_L=fsl$X17, FSL_R=fsl$X53, Source=fsl$Source)
+
+# diagnoses
+dx = adnimerge[c("DX.bl", "RID", "VISCODE")]
+
+# Now create one data.frame with all the yr1 volume data we have from each datasource
+combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
+combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
+combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
+combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
+combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
+combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
+combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)
+
+attach(combined)
+vols = data.frame(RID = RID, VISCODE = VISCODE, DX = DX.bl,
+                  MAGeT = MAGeT_L + MAGeT_R, 
+                  MAPER = MAPER_L + MAPER_R, 
+                  SNT   = SNT_L + SNT_R, 
+                  UCSD  = UCSD_L + UCSD_R,
+                  FS    = FS_L + FS_R, 
+                  FSL   = FSL_L + FSL_R)
+write.csv(vols, "data/ADNI_baseline_volumes_combined_volumes.csv")
 @
@@
 <<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-qplot(DX,value,data=melt(totals,id.vars=c("ID","DX"),measure.vars=c("FSL","MAPER","MAGeT","SNT"),na.rm=T),
+totals = read.csv("data/ADNI_baseline_volumes_combined_volumes.csv")
+qplot(DX,value,data=melt(totals,measure.vars=c("FS", "FSL","UCSD", "MAPER","MAGeT","SNT"),na.rm=T),
       colour=variable,geom="boxplot") + 
@@
 \begin{figure}[h]
-<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
-by_snt_volumes = melt(totals,id.vars=c("ID","SNT"), measure.vars=c("FSL","MAPER","MAGeT"))
+<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
+by_snt_volumes = melt(totals, measure.vars=c("FSL","MAPER","MAGeT", "FS", "UCSD"))
 fsl=subset(by_snt_volumes, variable=="FSL")
@@
 
-\begin{figure}[h]
-<<SZ_volumes,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes',cache=T>>=
+\begin{figure}
+<<SZ_volumes,echo=F,fig=T, fig.align='center',cache=T>>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
@@
 \end{figure}
+
+Some other text here... 
 \end{document}

### 2012-11-30 paper.tex
Calculating r/p-values for ADNI/SZFEP plots, but display code currently commented out due to ugliness
Update with full ADNI baseline info; Outlined rational for experiments

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                         Experiment Rationale 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\subsubsection{Study Method Rationale - An Outline}
+\begin{description}
+\item[ADNI Repeated Random Subsampling Validation] \hfill \\ 
+{\bf Purpose:} to validate robustness of MAGeT accuracy when atlas set is representative 
+of subject set w.r.t. registration method, atlas and template library composition and size, and 
+various fusion methods. Additionally, to choose the best configuration of parameters 
+(under the assumption that performance generalises). 
+\\ \\ 
+{\bf Method:} using a set of 69 ADNI subjects, 13 from each disease category, perform 
+10 rounds of repeated random subsampling to pick atlas library and template library 
+for each parameter setting and then run MAGeT brain to segment non-atlas subjects.
+Also perform multi-atlas and single-atlas segmentation using the same parameter settings
+and atlas library size.
+\\ \\
+{\bf Results:} find significant improvement over multi-atlas performed with the same
+parameters. Also, find smoothed performance is monotonically increasing but asymptotic
+in size of both template and atlas library, with peak performance reached after 15
+templates. \todo{Can we statistically capture "peak" performance?  The point at which 
+gains become statistically insignificant?}
+
+\item[Winterburn Atlas Reliability] \hfill \\
+{\bf Purpose:} validation of reliability of whole hippocampus and hippocampal 
+subregion segmentations.
+\\ \\
+{\bf Method:} we use the Winterburn atlas set, using each high-res atlas in turn 
+to segment a standard-res MR of the same subject and compare the automatic 
+segmentation to the atlas segmentation.  We use a template library consisting 
+of standard-res MRs from the remaining Winterburn atlas subjects plus 11 other 
+standard-res healthy control subjects. 
+\\ \\
+{\bf Results:} TBD
+
+\item[Winterburn-Atlas ADNI Baseline Segmentation] \hfill \\
+{\bf Purpose:} given that MAGeT is robust and accurate, show MAGeT performance on a real-world task in which atlas set may not be representative of subjects.
+\\ \\
+{\bf Method:} We use best settings from first experiment (ANTS, majority vote, 
+template library of 20), and use Winterburn hi-res atlases to segment all ADNI Baseline 
+images.  Since our segmentation protocol differs from what can be segmented at 
+ADNI resolution a direct comparison is not appropriate.  Instead, we validate by 
+showing how well our segmentation volumes correlate with existing automated/manual 
+segmentation volumes. \todo{Also show disease prediction?}
+\\ \\
+{\bf Results: }
+
+\item[First Episode SZ Patients Segmentation] \hfill \\ 
+{\bf Purpose:} validate MAGeT's general performance.
+\\ \\
+{\bf Method:} Measure performance on a different disease dataset. We use FEP SZ patients and controls, 
+using two different atlas sets: a manual hippocampal segmentation of patients, and Winterburn 
+atlas set.  We validate the native-atlas set segmentations using Dice's Kappa, 
+and the Winterburn-atlas-set segmentations by correlating volumes. 
+\\ \\
+{\bf Results:} High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
+
+\end{description}
+\end{kframe}
+
 \subsubsection{ADNI Validation Experiment}
@@
 &\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}&\ML
-Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.20~}{75.50 }{\scriptsize 78.45} &{\scriptsize 71.00~}{77.10 }{\scriptsize 81.40} &{\scriptsize 71.70~}{77.80 }{\scriptsize 81.75} &{\scriptsize 71.50~}{76.60 }{\scriptsize 81.30} &$ F_{2,66}=0.15 ,~ P=0.862 ^{1} $\NN
+Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.2~}{75.5 }{\scriptsize 78.5} &{\scriptsize 71.0~}{77.1 }{\scriptsize 81.4} &{\scriptsize 71.7~}{77.8 }{\scriptsize 81.8} &{\scriptsize 71.5~}{76.6 }{\scriptsize 81.3} &$ F_{2,66}=0.15 ,~ P=0.862 ^{1} $\NN
 Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}&$ \chi^{2}_{2}=0 ,~ P=1 ^{2} $\NN
@@
 ~~~~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
-CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} &$ F_{2,66}=294.2 ,~ P<0.001 ^{1} $\NN
-ADAS~13&{\scriptsize  4.665~}{ 5.670 }{\scriptsize 12.335} &{\scriptsize 14.335~}{16.000 }{\scriptsize 20.500} &{\scriptsize 23.830~}{29.000 }{\scriptsize 31.665} &{\scriptsize 10.000~}{16.000 }{\scriptsize 25.330} &$ F_{2,66}=109.8 ,~ P<0.001 ^{1} $\NN
-MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} &$ F_{2,66}=94.17 ,~ P<0.001 ^{1} $\LL
+CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} &$ F_{2,66}=294 ,~ P<0.001 ^{1} $\NN
+ADAS~13&{\scriptsize  4.67~}{ 5.67 }{\scriptsize 12.34} &{\scriptsize 14.34~}{16.00 }{\scriptsize 20.50} &{\scriptsize 23.83~}{29.00 }{\scriptsize 31.66} &{\scriptsize 10.00~}{16.00 }{\scriptsize 25.33} &$ F_{2,66}=110 ,~ P<0.001 ^{1} $\NN
+MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} &$ F_{2,66}=94.2 ,~ P<0.001 ^{1} $\LL
 }}
@@
 
-\begin{figure}[h]
+\begin{figure}
 \begin{knitrout}

### 2012-12-05 paper.Rnw
Fleshed out methods a bit more.. worked in the rationale

@@
 \documentclass{article}
-\usepackage[margin=1.5in]{geometry}
+\usepackage[margin=1in]{geometry}
 \usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
@@
 \newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
-\newcommand{\todo}[1]{\marginnote{\textcolor{red}{TODO #1}}}
+\newcommand{\todo}[1]{\begin{kframe}{\textcolor{red}{TODO #1}}\end{kframe}}
 \newcommand{\mc}[1]{\todo{(MC) #1}}
@@
 
-\section{Materials and Methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%                               METHODS
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}
 
-explain two registration methods, definitions, fusion methods. 
-compare to each other, and then compare atlas-based, true multi-atlas, naive methods.
-trying to make choices as to the best reg/fusion/
+%%%%%                     MAGeT Brain Algorithm                            %%%%%
 
-evaluations
-described previously in Chakravarty2012.
-
-\subsection{Segmentation Algorithm}
+\subsection{MAGeT Brain Algorithm}
 \todo{Make sure we explain what MAGeT stands for}
+\todo{reference Chakravarty2012}
 In this paper, we use the term {\it atlas} to mean any manually segmented MR
@@
 
-Label fusion is performed by cross-correlation weighted voting, a strategy
-weighted towards an optimal combination of subjects from the template library
-which has been previously shown to improve segmentation
-accuracy \citep{Aljabar2009,Collins2010}. In this method, each template library
-image is ranked in similarity to each unlabelled image by the normalized
-cross-correlation of image intensities after linear registration in a region of
-interest (ROI) generously encompassing the hippocampus.  Only the top ranked
-template library image labels are used in a voxel-wise majority vote. The ROI
-is heuristically defined as the extent of all atlas labels after linear
-registration to the template, dilated by three voxels \citep{Chakravarty2012}.
-
 Source code can be found at \url{http://github.com/pipitone/MAGeTbrain}.
@@
 \begin{algorithm}
+  \scriptsize
   \caption{Pseudocode for the MAGeT Brain algorithm}
@@
 
-\subsection{Image Processing and Registration Methods}
+\todo{Note that algorithm does not specify a registration algorithm or fusion 
+method\citep{Chakravarty2012}.  Performance on these is test in experiments}
+
+%%%%%                             Subjects                                 %%%%%
+
+%  Describe subject datasets in general 
+\subsection{Subjects}
+
+\subsubsection{ADNI-1 1.5T Screening}
+\todo{- image characteristics}
+\todo{- describe SNT manual segmentations}
+\todo{- baseline/cross-validation dataset demographics?}
+
+<<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 + MMSE, 
+               data = adnimerge, subset = VISCODE == 'bl',   
+               method = "reverse", test=FALSE, overall=TRUE)
+
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
+      caption="ADNI-1 1.5T Screening demographics")
+@
+\todo{Check table only includes 1.5T from the Screening}
+
+\subsubsection{SZ First Episode Patients}
+\todo{image characteristics, demographics}
+<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
+fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
+fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
+tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
+               data=fep_sz_demographics, method = "reverse", test = FALSE)
+latex(tab, file="", size="scriptsize", 
+      caption="Schizophrenia First Episode Patient Demographics")
+@
+
+%%%%%                       Image pre-processing                           %%%%%
+
+\subsection{Image pre-processing}
 
@@
 the intensity nonuniformity in each of the atlases and unlabeled subject
-images. In this work we investigated the performance of two registration
-methods.
+images.
+
+%%%%%                       Registration Methods                           %%%%%
+
+\subsection{Registration}
 
 \subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}
-
 The ANIMAL algorithm carries out Image registration is in two phases. In the
@@
 
-For the purposes of this work we used the
-regularization parameters optimized in Robbins et al. \citep{Robbins2004}. It
-should be noted that the MAGeT brain algorithm is not dependent on this, or
-any, particular choice of registration method \citep{Chakravarty2012}.
+For the purposes of this work we used the regularization parameters optimized 
+in Robbins et al. \citep{Robbins2004}.
 
-\todo{should add the MNI website here; also we should add the parameters we used somewhere}
+\todo{link to MNI website here}
+\todo{command line}
 
@@
 algorithm compatible with the MINC data format, mincANTS
-\url{https://github.com/vfonov/mincANTS} \todo{(same here we should add the
-parameters we used).}
+\url{https://github.com/vfonov/mincANTS}.
 
-\subsection{Label Fusion Methods}
+We used the following command line when running the ANTS command, 
+\begin{verbatim}
+  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
+   --number-of-affine-iterations 10000x10000x10000x10000x10000 
+   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
+   --use-Histogram-Matching --MI-option 32x16000
+   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
+   -o transformation.xfm
+ \end{verbatim}
+These settings were adapted from the "reasonable starting point" given in the
+ANTS manual. \todo{reference manual?}
+
+%%%%%                           Label Fusion                               %%%%%
+
+\subsection{Label Fusion}
 
@@
 
-Labels are propagated from only those template library images which most
-similar to the subject.  A voxel-wise majority vote is carried out from the
-labels from the top n ranked template library images, where n is parameter set
-by the user.
+An optimal combination of subjects from the template library has previously been
+shown to improve segmentation accuracy \citep{Aljabar2009,Collins2010}.  In this
+method, each template library image is ranked in similarity to each unlabelled 
+image by the normalized cross-correlation (CC) of image intensities after linear
+registration, over a region of interest (ROI) generously encompassing the 
+hippocampus.  Only the top ranked
+template library image labels are used in a voxel-wise majority vote. The ROI
+is heuristically defined as the extent of all atlas labels after linear
+registration to the template, dilated by three voxels \citep{Chakravarty2012}.  
+The number of top ranked template library image labels is a configurable parameter.
 
-Similarity between a subject and a template image is measured using normalized
-cross-correlation of intensity values, over a region of interest defined on
-each template, after the subject has been linearly registered to the template.
-The heuristic used for defining a meaningful region of interest for each
-template is to use the extent of all the propagated labellings from each atlas
-(after linear registration only) and then dilating this region by three voxels.
-
-The utility to calculate the cross-correlation similarity measure is
-implemented as part of the ANIMAL toolkit. \todo{(ANIMAL or MINC)}
+The {\tt xcorr\_vol utility} from the ANIMAL toolkit is used to calculate the 
+cross-correlation similarity measure.
 
@@
 
-This label fusion method is identical to the above process except a normalised
-mutual information score is used over the region of interest between a template
-and subject instead of cross-correlation.
+This method is similar to cross-correlation weighted voting 
+except that image similarity is calculated by the normalised 
+mutual information score over the region of interest. \todo{reference?}
 
-The utility to calculate the normalised mutual information measure is
-implemented as part of the EZMinc package (based on ITK NMI routine).
+The {\tt itk\_similarity} utility from the EZMinc toolkit is used to calculate 
+the normalised mutual information measure.
+
+%%%%%                        Goodness-of-fit                               %%%%%
 
@@
 
-\subsection{Data and Experiments}
+%%%%%                            Experiments                               %%%%%
 
-Three different experiments were conducted to evaluate the influence of
-registration method, atlas set size and resolution, template set size, and
-label fusion method and parameters.
+\subsection{Experiments}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                         Experiment Rationale 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\subsubsection{Study Method Rationale - An Outline}
-\begin{description}
-\item[ADNI Repeated Random Subsampling Validation] \hfill \\ 
-{\bf Purpose:} to show method leads to improved accuracy (over multi-atlas method), and is robust
-w.r.t.
-registration method, atlas and template library composition and size, and various fusion methods.
-Additionally, to choose the best configuration of parameters (under the assumption that performance generalises).
-\\ \\ 
-{\bf Method:} using a set of 69 ADNI subjects, 13 from each disease category, perform 
-10 rounds of repeated random subsampling to pick atlas library and template library 
-for each parameter setting and then run MAGeT brain to segment non-atlas subjects.
-Also perform multi-atlas and single-atlas segmentation using the same parameter settings
-and atlas library size.
-\\ \\
-{\bf Results:} find significant improvement over multi-atlas performed with the same
-parameters. Also, find smoothed performance is monotonically increasing but asymptotic
-in size of both template and atlas library, with peak performance reached after 15
-templates. \todo{Can we statistically capture "peak" performance?  The point at which 
-gains become statistically insignificant?}
+Experiments were performed to assess the performance of MAGeT brain with various 
+parameter settings as well as on diverse datasets.  In each experiment we contrast 
+the performance of MAGeT brain with that standard single- and multi-atlas 
+segmentations derived from the same atlas library.
 
-\item[Winterburn-Atlas ADNI Baseline Segmentation] \hfill \\
-{\bf Purpose:} given that MAGeT is robust and accurate, show MAGeT performance on 
-a real-world task in which atlas set may not be representative of subjects.
-\\ \\
-{\bf Method:} We use best settings from first experiment (ANTS, majority vote, 
-template library of 20), and use Winterburn hi-res atlases to segment all ADNI Baseline 
-images.  Since our segmentation protocol differs from what can be segmented at 
-ADNI resolution a direct comparison is not appropriate.  Instead, we validate by 
-showing how well our segmentation volumes correlate with existing automated/manual 
-segmentation volumes. \todo{Also show disease prediction?}
-\\ \\
-{\bf Results: }
+%%%%%
+%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
+%%%%%
+\subsubsection{ADNI-1 cross-validation}
 
-\item[First Episode SZ Patients Segmentation] \hfill \\ 
-{\bf Purpose:} validate that MAGeT's performance generalises to other diseases.
-\\ \\
-{\bf Method:} Measure performance using best parameter settings on a different disease dataset.
-We use FEP SZ patients and controls, using two different atlas sets: a manual hippocampal segmentation of patients, and Winterburn atlas set.
-We validate the native-atlas set segmentations using Dice's Kappa, and the Winterburn-atlas-set segmentations by correlating volumes.
-\\ \\
-{\bf Results:} High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
+To test the accuracy of the MAGeT brain algorithm with different parameter 
+settings, repeated random sub-sampling cross-validation (RRSCV) was performed
+on a subset of the ADNI-1 dataset.
 
-\item[Winterburn Atlas Reliability] \hfill \\
-{\bf Purpose:} validation of performance on hippocampal subregion segmentation
-\\ \\
-{\bf Method:} we use the Winterburn atlas set, using each high-res atlas in turn 
-to segment a standard-res MR of the same subject and compare the automatic 
-segmentation to the atlas segmentation.  We use a template library consisting 
-of standard-res MRs from the remaining Winterburn atlas subjects plus 11 other 
-standard-res healthy control subjects. 
-\\ \\
-{\bf Results:} TBD
-
-\end{description}
-\end{kframe}
-
-\subsubsection{ADNI Validation Experiment}
-
-In this experiment we performed repeated random sub-sampling cross-validation
-of the MAGeT algorithm with a pool of 69 images and labels from the ADNI
-dataset.  The images are T1-weighted \todo{<insert description of scan/machine
-types>}.  The hippocampal labels are provided as part of the ADNI dataset, and
-are produced \todo{<insert description of SNT labellings>}.  The parameters we
-varied, and the ranges over which we varied are listed in \todo{table}.  We
-performed 10 validations per parameter combination.  In each validation trial,
-a random sample of subjects were chosen from the pool without replacement as
-the atlas set, whereas the template set was chosen randomly with replacement.
-
-\begin{table}
-    \begin{tabular}{l|l}
-        Parameter           & Values tested                                                      \\ \hline
-        Number of Templates & 3 to 20                                                            \\ 
-        Number of Atlases   & 3 to 9                                                             \\ 
-        Registration Method & ANTS or ANIMAL                                                     \\ 
-        Label Fusion Method & majority vote, cross-correlation weighted vote, NMI weighted vote. 
-    \end{tabular}
-\end{table}
-
-We used the following parameters for ANTS, during registration: 
-\begin{verbatim}
-  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
-   --number-of-affine-iterations 10000x10000x10000x10000x10000 
-   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
-   --use-Histogram-Matching --MI-option 32x16000
-   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
-   -o transformation.xfm
- \end{verbatim}
-These settings were adapted from the "reasonable starting point" given in the
-ANTS manual. \todo{<rationale?>}
-
-For the purposes of this work we used the regularization parameters optimized
-in Robbins et al.[15]. \todo{right? (but added a second layer).}
-
-\todo{multi-atlas comparison, "naive" comparison}
-
-\subsubsection{Winterburn High-resolution Hippocampal Atlas - ADNI Validation}
-
-In this experiment we explored using the MAGeT brain algorithm with the
-Winterburn high-resolution atlases to segment 21 randomly chosen MR images from
-the ADNI dataset (seven each of healthy, MCI and AD subjects).  The Winterburn
-atlases are digital segmentations of the hippocampus in five  in-vivo 300u
-isotropic T1-weighted MR scans, and include subfield segmentations for the
-cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.
-Subjects in the Winterburn atlases range in age from 29-57 years (mean age of
-37), and include two males and three females.
-
-Since hippocampal segmentation protocols differ between the ADNI labels and
-Winterburn atlases, this poses a problem for direct similarity comparisons
-between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
-we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate
-the performance of MAGeT brain, we compared classification accuracy of subjects
-by diagnosis based on hippocampal volume using both the SMT labels and our
-produced labels.  \todo{description of validation -- rms {\tt validate} or
-t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}
+{\bf Dataset evaluated.} 69 1.5T images were randomly selected from the 
+{\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
+are shown in Table X\todo{FIXME: table reference}.
 
@@
                  "382","431","1202")
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTMARRY + PTETHCAT + PTRACCAT + CDRSB + ADAS13 + MMSE, 
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 + MMSE, 
                data = adnimerge, subset = VISCODE == 'bl' & RID %in% adni_69_RIDs, 
-               method = "reverse", test=TRUE, overall=TRUE)
+               method = "reverse", test=FALSE, overall=TRUE)
 
-latex(tab, file="", size="scriptsize", where="h",landscape=TRUE, ctable=TRUE)
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
+      caption="ADNI-1 cross-validation subset demographics")
 @
 
-\todo{For minctracc -- we registered to TAL}
+{\bf Atlas and template libary.}  Atlases consisted of images taken from the 
+dataset, with corresponding manual labels provided by SNT.  Atlas library size
+was varied from 3 to 9 images.  The remaining images were segmented, with the 
+template library size varying from 3 to 20 images.  Template library images were
+selected randomly from the images to be segmented.
 
+{\bf Registration method.}  Both the ANTS and ANIMAL registration methods were 
+used.
+\todo{In ANIMAL, images aligned to TAL space}
+
+{\bf Label fusion.} Majority vote, Cross-correlation weighted majority vote, and
+Normalized Mutual Information weighted majority vote are used.  With the 
+weighted majority vote fusion methods, the number of top labels used in the 
+fusion was varied from 3 to 20 images.
+
+{\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV) 
+consists of repeated trials in which items from the dataset are randomly assigned to a training set or validation set.
+In each trial, performance on the validation set is measured, and then averaged across all trials.
+
+We performed RRSCV on each combination of parameters listed above: atlas library size, template library size, registration method, and label fusion method.
+We performed 10 trials per parameter combination.
+In each validation trial, the training set consisted of the images used as atlases, and the validation set consisted of the images to be segmented.
+The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were applied to segment the images in the validation set.
+Additionally, in each trial, the single-atlas segmentation was obtained for each atlas-template.
+
+The gold-standard for the segmentation accuracy of images in the validation set was the SNT manual labels.
+
+\todo{ number of registrations/comparisions }
+
+%%%%%
+%%%%%               Experiments: ADNI-1 Screening Validation               %%%%%
+%%%%%
+\subsubsection{ADNI-1 Screening Validation}
+
+To test the accuracy of MAGeT brain on a real-world task we segment the entire
+ADNI-1 dataset using an atlas set that is not representative of the subject set.
+
+{\bf Dataset evaluated.} All images from the {\em ADNI1:Screening 1.5T} 
+standardized dataset.
+
+{\bf Atlas and template libary.} The atlas library consisted of the entire 
+Winterburn atlas set.  The Winterburn atlases are digital segmentations of the
+hippocampus in five in-vivo 300u isotropic T1-weighted MR scans, and include 
+subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, 
+subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases range in 
+age from 29-57 years (mean age of 37), and include two males and three females.
+
+The template library consisted of 21 randomly selected images from the ADNI1 
+data dataset (7 healthy, MCI and AD subjects).
+
+{\bf Registration method.}  - ANTS as it performed best in the cross-validation
+experiment. \todo{how to reference results whilst still in Methods?}
+
+{\bf Label fusion.} Majority vote, as it is simpliest to run and performed 
+equally well in cross-validation experiment.
+
+{\bf Evaluation.}
+
+Since hippocampal segmentation protocols differ between the ADNI labels and
+Winterburn atlases, this poses a problem for direct similarity comparisons
+between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
+we did(n't) resegment the ADNI images with a the low-res protocol.}
+
+To evaluate the performance of MAGeT brain, we correlate our segmentation
+volumes with manual segmentation volumes, as well as with hippocampal volumes
+of established automated segmentation methods.
+
+Additionally, we compared classification accuracy of subjects
+by diagnosis based on hippocampal volume using both the SMT labels and our
+produced labels.  \todo{description of validation -- rms {\tt validate} or
+t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}
+
+%%%%%
+%%%%%          Experiments: SZ First Episode Patient Validation           %%%%%
+%%%%%
+\subsubsection{SZ First Episode Patient Validation}
+ To validate that MAGeT's performance generalises to other diseases.
+Measure performance using best parameter settings on a different disease dataset.
+
+{\bf Dataset evaluated.} SZ First Episode Patients
+
+{\bf Atlas and template libary.} two different atlas sets: a manual hippocampal
+segmentation of patients, and Winterburn atlas set.
+
+{\bf Registration method.} ANTS.
+
+{\bf Label fusion.} Majority vote.
+
+{\bf Evaluation.} We validate the FEP-atlas segmentations using Dice's Kappa,
+and the Winterburn-atlas segmentations by correlating volumes.
+
+%%%%%
+%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
+%%%%%
+\subsubsection{Winterburn Atlases Validation}
+{\bf Dataset evaluated.} T1 BRAVO scans of the same subjects included in the 
+Winterburn atlas set.  These scans are taken within .... weeks of the scans for 
+the Winterburn atlases.
+
+{\bf Atlas and template libary.} Atlas library is Winterburn T1 atlases.  
+Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.
+
+{\bf Registration method.} ANTS
+
+{\bf Label fusion.} Majority vote.
+
+{\bf Evaluation.}  Leave one out cross-validation (LOOCV) in which all five
+subjects are segmented in separate runs of MAGeT brain.  In each run, the subject
+to be segmented is excluded from the Atlas library (so only four atlases are used).
+
+Segmentation accuracy is judged by difference in hippocampal volume.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%                               Results
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Results}
-\subsubsection{A2A}
 
-Kappa vs. Number of templates: 
-Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))
-<<a2a_facet, cache=TRUE, fig=TRUE>>=
+%%%%%
+%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
+%%%%%
+\subsubsection{ADNI-1 Cross-Validation}
+
+ - find significant improvement over multi-atlas performed with the same
+  parameters. Also, find smoothed performance is monotonically increasing but asymptotic
+  in size of both template and atlas library, with peak performance reached after 15
+  templates. \todo{Can we statistically capture "peak" performance?  The point at which 
+  gains become statistically insignificant?}
+
+\begin{figure}
+<<ADNI1_cross_validation_means, cache=TRUE, fig=TRUE>>=
 t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
@@
 @
+  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))}
+  \label{ADNI1_cross_validation_means}
+\end{figure}
 
-<<a2a_facet_mameans, cache=TRUE, fig=TRUE>>=
+\begin{figure}
+<<ADN1-cross-validation-MA-difference, cache=TRUE, fig=TRUE>>=
 ggplot(all_data, aes(x=num_templates, y=k_minus_ma, colour=as.factor(num_atlases))) + 
@@
 @
-
-Kappa minus multi-atlas mean vs. Number of Templates
+  \caption{Difference in mean Kappa between MAGeT brain and multi-atlas}
+  \label{ADNI1_cross_validation_MA_difference}
+\end{figure}
 
@@
 <<a2a_multiatlas_means, results="asis",cache=FALSE>>=
-latex(merge(a_ma_means, t_ma_means, by="num_atlases"), colheads=c("Atlases", "ANTS", "ANIMAL"), file="")
+latex(merge(a_ma_means, t_ma_means, by="num_atlases"), colheads=c("Atlases", "ANTS", "ANIMAL"), file="", dec=2)
 @
@@
 \subsubsection{First Episode Schizophrenic patients}
-
-<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
-fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
-fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
-tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ, data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", size="scriptsize", where="h")
-@
+High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
 

### 2012-12-05 paper.tex
Fleshed out methods a bit more.. worked in the rationale

--- 2012-11-22 paper.tex
@@
 \usepackage{alltt}
-\usepackage[margin=1.5in]{geometry}
+\usepackage[margin=1in]{geometry}
 \usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
@@
 
-\subsection{Image Processing and Registration Methods}
+\todo{Note that algorithm does not specify a registration algorithm or fusion 
+method\citep{Chakravarty2012}.  Performance on these is test in experiments}
+
+%%%%%                             Subjects                                 %%%%%
+
+%  Describe subject datasets in general 
+\subsection{Subjects}
+
+\subsubsection{ADNI-1 1.5T Screening}
+\todo{- image characteristics}
+\todo{- describe SNT manual segmentations}
+\todo{- baseline/cross-validation dataset demographics?}
+
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+{\scriptsize\ctable[caption={ADNI-1 1.5T Screening demographics},label=tab,pos=!tbp,]{lrccccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{$N$\ is the number of non--missing values.}\tnote[]{Numbers after percents are frequencies.}}{\FL
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{EMCI}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}\NN
+&&\multicolumn{1}{c}{{\scriptsize $N=414$}}&\multicolumn{1}{c}{{\scriptsize $N=279$}}&\multicolumn{1}{c}{{\scriptsize $N=560$}}&\multicolumn{1}{c}{{\scriptsize $N=278$}}&\multicolumn{1}{c}{{\scriptsize $N=1531$}}\ML
+Age~at~baseline~\hfill\tiny{Years}&1531&{\scriptsize 70.9~}{74.0 }{\scriptsize 78.4} &{\scriptsize 66.0~}{71.0 }{\scriptsize 76.2} &{\scriptsize 69.4~}{74.3 }{\scriptsize 79.4} &{\scriptsize 70.9~}{75.9 }{\scriptsize 80.9} &{\scriptsize 69.6~}{74.0 }{\scriptsize 79.2} \NN
+Sex~:~Female&1531&50\%~{\scriptsize~(206)}&44\%~{\scriptsize~(123)}&38\%~{\scriptsize~(215)}&45\%~{\scriptsize~(125)}&44\%~{\scriptsize~(669)}\NN
+Education&1531&{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 17.8} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} \NN
+Ethnicity~:~Unknown&1531&~1\%~{\scriptsize~(~~~4)}&~1\%~{\scriptsize~(~~~2)}&~1\%~{\scriptsize~(~~~4)}&~1\%~{\scriptsize~(~~~3)}&~1\%~{\scriptsize~(~~13)}\NN
+~~~~Not~Hisp/Latino&&96\%~{\scriptsize~(~398)}&95\%~{\scriptsize~(~265)}&96\%~{\scriptsize~(~540)}&96\%~{\scriptsize~(~267)}&96\%~{\scriptsize~(1470)}\NN
+~~~~Hisp/Latino&&~3\%~{\scriptsize~(~~12)}&~4\%~{\scriptsize~(~~12)}&~3\%~{\scriptsize~(~~16)}&~3\%~{\scriptsize~(~~~8)}&~3\%~{\scriptsize~(~~48)}\NN
+CDR-SB&1531&{\scriptsize 0.0~}{0.0 }{\scriptsize 0.0} &{\scriptsize 0.5~}{1.0 }{\scriptsize 1.5} &{\scriptsize 1.0~}{1.5 }{\scriptsize 2.0} &{\scriptsize 3.5~}{4.5 }{\scriptsize 5.0} &{\scriptsize 0.0~}{1.0 }{\scriptsize 2.5} \NN
+ADAS~13&1511&{\scriptsize  6.0~}{ 9.0 }{\scriptsize 12.0} &{\scriptsize  8.0~}{12.0 }{\scriptsize 16.0} &{\scriptsize 14.3~}{18.7 }{\scriptsize 23.0} &{\scriptsize 24.0~}{29.0 }{\scriptsize 34.0} &{\scriptsize 10.0~}{15.3 }{\scriptsize 23.0} \NN
+MMSE&1531&{\scriptsize 29.0~}{29.0 }{\scriptsize 30.0} &{\scriptsize 28.0~}{29.0 }{\scriptsize 30.0} &{\scriptsize 26.0~}{27.0 }{\scriptsize 29.0} &{\scriptsize 21.2~}{23.0 }{\scriptsize 25.0} &{\scriptsize 26.0~}{28.0 }{\scriptsize 29.0} \LL
+}}
+
+\todo{Check table only includes 1.5T from the Screening}
+
+\subsubsection{SZ First Episode Patients}
+\todo{image characteristics, demographics}
+
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+\begin{table}[!tbp]
+\scriptsize
+\caption{Schizophrenia First Episode Patient Demographics\label{tab}} 
+\begin{center}
+\begin{tabular}{lrc}
+\hline\hline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{FEP}\tabularnewline
+&&\multicolumn{1}{c}{{\scriptsize $N=81$}}\tabularnewline
+\hline
+Age&80&{\scriptsize 21~}{23 }{\scriptsize 26} \tabularnewline
+Gender~:~M&81&63\%~{\scriptsize~(51)}\tabularnewline
+Handedness~:~ambi&81&~6\%~{\scriptsize~(~5)}\tabularnewline
+~~~~left&&~5\%~{\scriptsize~(~4)}\tabularnewline
+~~~~right&&89\%~{\scriptsize~(72)}\tabularnewline
+Education&81&{\scriptsize 11~}{13 }{\scriptsize 15} \tabularnewline
+SES~:~lower&81&31\%~{\scriptsize~(25)}\tabularnewline
+~~~~middle&&54\%~{\scriptsize~(44)}\tabularnewline
+~~~~upper&&15\%~{\scriptsize~(12)}\tabularnewline
+FSIQ&79&{\scriptsize  88~}{102 }{\scriptsize 109} \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table}
+
+%%%%%                       Image pre-processing                           %%%%%
+
+\subsection{Image pre-processing}
 
@@
 
-\subsection{Data and Experiments}
+%%%%%                            Experiments                               %%%%%
 
-Three different experiments were conducted to evaluate the influence of
-registration method, atlas set size and resolution, template set size, and
-label fusion method and parameters.
+\subsection{Experiments}
 
-\subsubsection{ADNI Validation Experiment}
+Experiments were performed to assess the performance of MAGeT brain with various 
+parameter settings as well as on diverse datasets.  In each experiment we contrast 
+the performance of MAGeT brain with that standard single- and multi-atlas 
+segmentations derived from the same atlas library.
 
-In this experiment we performed repeated random sub-sampling cross-validation
-of the MAGeT algorithm with a pool of 69 images and labels from the ADNI
-dataset.  The images are T1-weighted \todo{<insert description of scan/machine
-types>}.  The hippocampal labels are provided as part of the ADNI dataset, and
-are produced \todo{<insert description of SNT labellings>}.  The parameters we
-varied, and the ranges over which we varied are listed in \todo{table}.  We
-performed 10 validations per parameter combination.  In each validation trial,
-a random sample of subjects were chosen from the pool without replacement as
-the atlas set, whereas the template set was chosen randomly with replacement.
+%%%%%
+%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
+%%%%%
+\subsubsection{ADNI-1 cross-validation}
 
-\begin{table}
-    \begin{tabular}{l|l}
-        Parameter           & Values tested                                                      \\ \hline
-        Number of Templates & 3 to 20                                                            \\ 
-        Number of Atlases   & 3 to 9                                                             \\ 
-        Registration Method & ANTS or ANIMAL                                                     \\ 
-        Label Fusion Method & majority vote, cross-correlation weighted vote, NMI weighted vote. 
-    \end{tabular}
-\end{table}
+To test the accuracy of the MAGeT brain algorithm with different parameter 
+settings, repeated random sub-sampling cross-validation (RRSCV) was performed
+on a subset of the ADNI-1 dataset.
 
-We used the following parameters for ANTS, during registration: 
-\begin{verbatim}
-  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
-   --number-of-affine-iterations 10000x10000x10000x10000x10000 
-   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
-   --use-Histogram-Matching --MI-option 32x16000
-   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
-   -o transformation.xfm
- \end{verbatim}
-These settings were adapted from the "reasonable starting point" given in the
-ANTS manual. \todo{<rationale?>}
+{\bf Dataset evaluated.} 69 1.5T images were randomly selected from the 
+{\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
+are shown in Table X\todo{FIXME: table reference}.
 
-For the purposes of this work we used the regularization parameters optimized
-in Robbins et al.[15]. \todo{right? (but added a second layer).}
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+{\scriptsize\ctable[caption={ADNI-1 cross-validation subset demographics},label=tab,pos=!tbp,]{lcccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{Numbers after percents are frequencies.}}{\FL
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}\NN
+&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}\ML
+Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.2~}{75.5 }{\scriptsize 78.5} &{\scriptsize 71.0~}{77.1 }{\scriptsize 81.4} &{\scriptsize 71.7~}{77.8 }{\scriptsize 81.8} &{\scriptsize 71.5~}{76.6 }{\scriptsize 81.3} \NN
+Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}\NN
+Education&{\scriptsize 16.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 15.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 16.5} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} \NN
+Ethnicity~:~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}\NN
+~~~~Not~Hisp/Latino&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(69)}\NN
+~~~~Hisp/Latino&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}\NN
+CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} \NN
+ADAS~13&{\scriptsize  4.67~}{ 5.67 }{\scriptsize 12.34} &{\scriptsize 14.34~}{16.00 }{\scriptsize 20.50} &{\scriptsize 23.83~}{29.00 }{\scriptsize 31.66} &{\scriptsize 10.00~}{16.00 }{\scriptsize 25.33} \NN
+MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} \LL
+}}
 
-\todo{multi-atlas comparison, "naive" comparison}
+{\bf Atlas and template libary.}  Atlases consisted of images taken from the 
+dataset, with corresponding manual labels provided by SNT.  Atlas library size
+was varied from 3 to 9 images.  The remaining images were segmented, with the 
+template library size varying from 3 to 20 images.  Template library images were
+selected randomly from the images to be segmented.
 
-\subsubsection{Winterburn High-resolution Hippocampal Atlas - ADNI Validation}
+{\bf Registration method.}  Both the ANTS and ANIMAL registration methods were 
+used.
+\todo{In ANIMAL, images aligned to TAL space}
 
-In this experiment we explored using the MAGeT brain algorithm with the
-Winterburn high-resolution atlases to segment 21 randomly chosen MR images from
-the ADNI dataset (seven each of healthy, MCI and AD subjects).  The Winterburn
-atlases are digital segmentations of the hippocampus in five  in-vivo 300u
-isotropic T1-weighted MR scans, and include subfield segmentations for the
-cornus ammonis (CA) 1, CA4, dentate gyrus, subiculum, and CA 2 and 3 combined.
-Subjects in the Winterburn atlases range in age from 29-57 years (mean age of
-37), and include two males and three females.
+{\bf Label fusion.} Majority vote, Cross-correlation weighted majority vote, and
+Normalized Mutual Information weighted majority vote are used.  With the 
+weighted majority vote fusion methods, the number of top labels used in the 
+fusion was varied from 3 to 20 images.
+
+{\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV) 
+consists of repeated trials in which items from the dataset are randomly assigned to a training set or validation set.
+In each trial, performance on the validation set is measured, and then averaged across all trials.
+
+We performed RRSCV on each combination of parameters listed above: atlas library size, template library size, registration method, and label fusion method.
+We performed 10 trials per parameter combination.
+In each validation trial, the training set consisted of the images used as atlases, and the validation set consisted of the images to be segmented.
+The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were applied to segment the images in the validation set.
+Additionally, in each trial, the single-atlas segmentation was obtained for each atlas-template.
+
+The gold-standard for the segmentation accuracy of images in the validation set was the SNT manual labels.
+
+\todo{ number of registrations/comparisions }
+
+%%%%%
+%%%%%               Experiments: ADNI-1 Screening Validation               %%%%%
+%%%%%
+\subsubsection{ADNI-1 Screening Validation}
+
+To test the accuracy of MAGeT brain on a real-world task we segment the entire
+ADNI-1 dataset using an atlas set that is not representative of the subject set.
+
+{\bf Dataset evaluated.} All images from the {\em ADNI1:Screening 1.5T} 
+standardized dataset.
+
+{\bf Atlas and template libary.} The atlas library consisted of the entire 
+Winterburn atlas set.  The Winterburn atlases are digital segmentations of the
+hippocampus in five in-vivo 300u isotropic T1-weighted MR scans, and include 
+subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, 
+subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases range in 
+age from 29-57 years (mean age of 37), and include two males and three females.
+
+The template library consisted of 21 randomly selected images from the ADNI1 
+data dataset (7 healthy, MCI and AD subjects).
+
+{\bf Registration method.}  - ANTS as it performed best in the cross-validation
+experiment. \todo{how to reference results whilst still in Methods?}
+
+{\bf Label fusion.} Majority vote, as it is simpliest to run and performed 
+equally well in cross-validation experiment.
+
+{\bf Evaluation.}
 
@@
 between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
-we did(n't) resegment the ADNI images with a the low-res protocol.} To evaluate
-the performance of MAGeT brain, we compared classification accuracy of subjects
+we did(n't) resegment the ADNI images with a the low-res protocol.}
+
+To evaluate the performance of MAGeT brain, we correlate our segmentation
+volumes with manual segmentation volumes, as well as with hippocampal volumes
+of established automated segmentation methods.
+
+Additionally, we compared classification accuracy of subjects
 by diagnosis based on hippocampal volume using both the SMT labels and our
@@
 
-% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%%%%%
+%%%%%          Experiments: SZ First Episode Patient Validation           %%%%%
+%%%%%
+\subsubsection{SZ First Episode Patient Validation}
+ To validate that MAGeT's performance generalises to other diseases.
+Measure performance using best parameter settings on a different disease dataset.
+
+{\bf Dataset evaluated.} SZ First Episode Patients
+
+{\bf Atlas and template libary.} two different atlas sets: a manual hippocampal
+segmentation of patients, and Winterburn atlas set.
+
+{\bf Registration method.} ANTS.
+
+{\bf Label fusion.} Majority vote.
+
+{\bf Evaluation.} We validate the FEP-atlas segmentations using Dice's Kappa,
+and the Winterburn-atlas segmentations by correlating volumes.
+
+%%%%%
+%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
+%%%%%
+\subsubsection{Winterburn Atlases Validation}
+{\bf Dataset evaluated.} T1 BRAVO scans of the same subjects included in the 
+Winterburn atlas set.  These scans are taken within .... weeks of the scans for 
+the Winterburn atlases.
+
+{\bf Atlas and template libary.} Atlas library is Winterburn T1 atlases.  
+Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.
+
+{\bf Registration method.} ANTS
+
+{\bf Label fusion.} Majority vote.
+
+{\bf Evaluation.}  Leave one out cross-validation (LOOCV) in which all five
+subjects are segmented in separate runs of MAGeT brain.  In each run, the subject
+to be segmented is excluded from the Atlas library (so only four atlases are used).
+
+Segmentation accuracy is judged by difference in hippocampal volume.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-{\scriptsize\ctable[caption={Descriptive Statistics by Baseline Dx},label=tab,rotate]{lccccc}{\tnote[]{\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.}\tnote[]{Numbers after percents are frequencies.}\tnote[]{\indent Tests used:}\tnote[]{\textsuperscript{\normalfont 1}Kruskal-Wallis test; \textsuperscript{\normalfont 2}Pearson test}}{\FL
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}&\multicolumn{1}{c}{Test Statistic}\NN
-&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}&\ML
-Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.20~}{75.50 }{\scriptsize 78.45} &{\scriptsize 71.00~}{77.10 }{\scriptsize 81.40} &{\scriptsize 71.70~}{77.80 }{\scriptsize 81.75} &{\scriptsize 71.50~}{76.60 }{\scriptsize 81.30} &$ F_{2,66}=0.15 ,~ P=0.862 ^{1} $\NN
-Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}&$ \chi^{2}_{2}=0 ,~ P=1 ^{2} $\NN
-Education&{\scriptsize 16.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 15.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 16.5} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} &$ F_{2,66}=3.44 ,~ P=0.038 ^{1} $\NN
-Marital~status~at~baseline~:~Married&74\%~{\scriptsize~(17)}&87\%~{\scriptsize~(20)}&87\%~{\scriptsize~(20)}&83\%~{\scriptsize~(57)}&$ \chi^{2}_{4}=3.04 ,~ P=0.551 ^{2} $\NN
-~~~~Widowed&22\%~{\scriptsize~(~5)}&13\%~{\scriptsize~(~3)}&13\%~{\scriptsize~(~3)}&16\%~{\scriptsize~(11)}&\NN
-~~~~Divorced&~4\%~{\scriptsize~(~1)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~1\%~{\scriptsize~(~1)}&\NN
-~~~~Never~married&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&\NN
-~~~~Unknown&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&~0\%~{\scriptsize~(~0)}&\NN
-Ethnicity~:~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&$^{2}$\NN
-~~~~Not~Hisp/Latino&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(23)}&100\%~{\scriptsize~(69)}&\NN
-~~~~Hisp/Latino&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
-Race~:~Am~Indian/Alaskan&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&$ \chi^{2}_{2}=5.61 ,~ P=0.061 ^{2} $\NN
-~~~~Asian&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
-~~~~Hawaiian/Other~PI&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
-~~~~Black&~17\%~{\scriptsize~(~4)}&~~4\%~{\scriptsize~(~1)}&~~0\%~{\scriptsize~(~0)}&~~7\%~{\scriptsize~(~5)}&\NN
-~~~~White&~83\%~{\scriptsize~(19)}&~96\%~{\scriptsize~(22)}&100\%~{\scriptsize~(23)}&~93\%~{\scriptsize~(64)}&\NN
-~~~~More~than~one&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
-~~~~Unknown&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&~~0\%~{\scriptsize~(~0)}&\NN
-CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} &$ F_{2,66}=294.2 ,~ P<0.001 ^{1} $\NN
-ADAS~13&{\scriptsize  4.665~}{ 5.670 }{\scriptsize 12.335} &{\scriptsize 14.335~}{16.000 }{\scriptsize 20.500} &{\scriptsize 23.830~}{29.000 }{\scriptsize 31.665} &{\scriptsize 10.000~}{16.000 }{\scriptsize 25.330} &$ F_{2,66}=109.8 ,~ P<0.001 ^{1} $\NN
-MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} &$ F_{2,66}=94.17 ,~ P<0.001 ^{1} $\LL
-}}
+%                               Results
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Results}
 
-\todo{For minctracc -- we registered to TAL}
+%%%%%
+%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
+%%%%%
+\subsubsection{ADNI-1 Cross-Validation}
 
-\subsection{Results}
-\subsubsection{A2A}
+ - find significant improvement over multi-atlas performed with the same
+  parameters. Also, find smoothed performance is monotonically increasing but asymptotic
+  in size of both template and atlas library, with peak performance reached after 15
+  templates. \todo{Can we statistically capture "peak" performance?  The point at which 
+  gains become statistically insignificant?}
 
-Kappa vs. Number of templates: 
-Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))
+\begin{figure}
 \begin{knitrout}
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/a2a_facet}
+{\centering \includegraphics[width=\maxwidth]{figure/ADNI1_cross_validation_means}
 
@@
 
+  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))}
+  \label{ADNI1_cross_validation_means}
+\end{figure}
+
+\begin{figure}
 \begin{knitrout}
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/a2a_facet_mameans}
+{\centering \includegraphics[width=\maxwidth]{figure/ADN1-cross-validation-MA-difference}
 
@@
 
-Kappa minus multi-atlas mean vs. Number of Templates
+  \caption{Difference in mean Kappa between MAGeT brain and multi-atlas}
+  \label{ADNI1_cross_validation_MA_difference}
+\end{figure}
 
@@
 
-% latex.default(merge(a_ma_means, t_ma_means, by = "num_atlases"),      colheads = c("Atlases", "ANTS", "ANIMAL"), file = "") 
+% latex.default(merge(a_ma_means, t_ma_means, by = "num_atlases"),      colheads = c("Atlases", "ANTS", "ANIMAL"), file = "", dec = 2) 
 %
@@
 \hline
-1&$3$&$0.805414417498623$&$0.764266832660852$\tabularnewline
-2&$4$&$0.793646605043623$&$0.752420649183594$\tabularnewline
-3&$5$&$0.823292141139203$&$0.785546924392891$\tabularnewline
-4&$6$&$0.815578947302174$&$0.778345910844141$\tabularnewline
-5&$7$&$0.831936342756014$&$0.797761650124922$\tabularnewline
-6&$8$&$0.826298462726884$&$0.792443538092187$\tabularnewline
-7&$9$&$0.836877632304638$&$0.804568318173359$\tabularnewline
+1&$3.00$&$0.81$&$0.76$\tabularnewline
+2&$4.00$&$0.79$&$0.75$\tabularnewline
+3&$5.00$&$0.82$&$0.79$\tabularnewline
+4&$6.00$&$0.82$&$0.78$\tabularnewline
+5&$7.00$&$0.83$&$0.80$\tabularnewline
+6&$8.00$&$0.83$&$0.79$\tabularnewline
+7&$9.00$&$0.84$&$0.80$\tabularnewline
 \hline
@@
 \subsubsection{First Episode Schizophrenic patients}
+High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
 
-% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
-%
-\begin{table}[h]
-\scriptsize
-\caption{Descriptive Statistics by DX\label{tab}} 
-\begin{center}
-\begin{tabular}{lrc}
-\hline\hline
-\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{FEP}\tabularnewline
-&&\multicolumn{1}{c}{{\scriptsize $N=81$}}\tabularnewline
-\hline
-Age&80&{\scriptsize 21~}{23 }{\scriptsize 26} \tabularnewline
-Gender~:~M&81&63\%~{\scriptsize~(51)}\tabularnewline
-Handedness~:~ambi&81&~6\%~{\scriptsize~(~5)}\tabularnewline
-~~~~left&&~5\%~{\scriptsize~(~4)}\tabularnewline
-~~~~right&&89\%~{\scriptsize~(72)}\tabularnewline
-Education&81&{\scriptsize 11~}{13 }{\scriptsize 15} \tabularnewline
-SES~:~lower&81&31\%~{\scriptsize~(25)}\tabularnewline
-~~~~middle&&54\%~{\scriptsize~(44)}\tabularnewline
-~~~~upper&&15\%~{\scriptsize~(12)}\tabularnewline
-FSIQ&79&{\scriptsize  88~}{102 }{\scriptsize 109} \tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table}
-
-\begin{figure}[h]
+\begin{figure}
 \begin{knitrout}

### 2012-12-07 paper.Rnw
Further work on methods section

@@
 \SweaveOpts{echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}
+
 <<loadpackages, results="hide", echo=FALSE, cache=FALSE>>=
@@
 
-\documentclass{article}
+\documentclass[twocolumn]{article}
 \usepackage[margin=1in]{geometry}
@@
 \newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
-\newcommand{\todo}[1]{\begin{kframe}{\textcolor{red}{TODO #1}}\end{kframe}}
-\newcommand{\mc}[1]{\todo{(MC) #1}}
-
-%{\marginpar{ \textcolor{red}{\\ TODO: #1 \\} }
+\newcommand{\comment}[1]{\begin{kframe}{\textcolor{red}{#1}}\end{kframe}}
+\newcommand{\todo}[1]{\comment{TODO #1}}
+\newcommand{\mc}[1]{\comment{MC: #1}}
 \begin{document}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                               INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 %                               METHODS
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\todo{Note that algorithm does not specify a registration algorithm or fusion 
-method\citep{Chakravarty2012}.  Performance on these is test in experiments}
+\todo{Note that our algorithm does not specify a registration algorithm or
+fusion method\citep{Chakravarty2012}.  Performance on options is tested in our
+experiments}
 
@@
 \subsubsection{ADNI-1 1.5T Screening}
-\todo{- image characteristics}
-\todo{- describe SNT manual segmentations}
-\todo{- baseline/cross-validation dataset demographics?}
+\todo{include image characteristics}
+\todo{describe SNT manual segmentations}
 
+\begin{figure*}
 <<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 @
-\todo{Check table only includes 1.5T from the Screening}
+\end{figure*}
+
+\todo{Check demographics table only includes 1.5T from the Screening}
 
 \subsubsection{SZ First Episode Patients}
-\todo{image characteristics, demographics}
+\todo{include image characteristics}
+\begin{figure*}
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 @
+\end{figure*}
 
@@
 \todo{link to MNI website here}
-\todo{command line}
+\todo{show command line used}
 
@@
 These settings were adapted from the "reasonable starting point" given in the
-ANTS manual. \todo{reference manual?}
+ANTS manual. \todo{include a reference to manual?}
 
@@
 except that image similarity is calculated by the normalised 
-mutual information score over the region of interest. \todo{reference?}
+mutual information score over the region of interest. 
+\todo{is there a reference for NMI voting?}
 
@@
 
-Each segmentation was evaluated against the 'gold-standard' manual segmentation
-from the dataset using the Dice Kappa ($\kappa$) overlap metric:
+Two segmentation can be compared using the Dice Kappa ($\kappa$) overlap metric:
 
@@
 
-where $a$ is the number of voxels common to the candidate segmentation and the
-gold standard and $b+c$ is the sum of the voxels uniquely identified by either
-the automatically generated candidate segmentation or the gold-standard.
-\todo{I would do this after you define the experiments}
+where $a$ is the number of voxels common to both segmentations and $b+c$ is the
+sum of the voxels uniquely identified by either segmentation.  Using this metric
+we can compare an automatically generated segmentation to a gold-standard
+segmentation.
 
+%%%%%
 %%%%%                            Experiments                               %%%%%
-
+%%%%%
 \subsection{Experiments}
@@
 {\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
-are shown in Table X\todo{FIXME: table reference}.
+are shown in Table X
+\todo{FIXME: table reference}.
 
+\begin{table*}
 <<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 @
+\end{table*}
 
@@
 used.
+
 \todo{In ANIMAL, images aligned to TAL space}
@@
 {\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV)
-consists of repeated trials in which items from the dataset are randomly assigned to a training set or validation set.
-In each trial, performance on the validation set is measured, and then averaged across all trials.
+consists of repeated trials in which items from the dataset are randomly
+assigned to a training set or validation set. In each trial, performance on the
+validation set is measured, and then averaged across all trials.
 
-We performed RRSCV on each combination of parameters listed above: atlas library size, template library size, registration method, and label fusion method.
-We performed 10 trials per parameter combination.
-In each validation trial, the training set consisted of the images used as atlases, and the validation set consisted of the images to be segmented.
-The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were applied to segment the images in the validation set.
-Additionally, in each trial, the single-atlas segmentation was obtained for each atlas-template.
+We performed RRSCV on each combination of parameters listed above: atlas library
+size, template library size, registration method, and label fusion method.  We
+performed 10 trials per parameter combination.  In each validation trial, the
+training set consisted of the images used as atlases, and the validation set
+consisted of the images to be segmented. The MAGeT brain algorithm and the basic
+multi-atlas segmentation procedure were applied to segment the images in the
+validation set.  Additionally, in each trial, the single-atlas segmentation was
+obtained for each atlas-template.
 
-The gold-standard for the segmentation accuracy of images in the validation set was the SNT manual labels.
+The gold-standard for the segmentation accuracy of images in the validation set
+was the SNT manual labels.
 
-\todo{ number of registrations/comparisions }
+\todo{ show off the number of registrations/comparisions we did}
 
@@
 
-{\bf Registration method.}  - ANTS as it performed best in the cross-validation
-experiment. \todo{how to reference results whilst still in Methods?}
+{\bf Registration method.} ANTS, as it performed best in the cross-validation
+experiment.
+
+\todo{Is it okay to reference results whilst still in Methods?}
 
@@
 Winterburn atlases, this poses a problem for direct similarity comparisons
-between labels produced by MAGeT brain and the ADNI labels.  \todo{explain why
-we did(n't) resegment the ADNI images with a the low-res protocol.}
+between labels produced by MAGeT brain and the ADNI labels.
+
+\todo{explain why we did(n't) resegment the ADNI images with the low-res
+protocol and compare directly like that.}
 
@@
 by diagnosis based on hippocampal volume using both the SMT labels and our
-produced labels.  \todo{description of validation -- rms {\tt validate} or
-t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV}
+produced labels.
+
+\todo{Include a description of validation process.. Note that we could use rms
+{\tt validate} or t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV,
+with adni/our segmentations}
 
@@
 \subsubsection{SZ First Episode Patient Validation}
- To validate that MAGeT's performance generalises to other diseases.
-Measure performance using best parameter settings on a different disease dataset.
 
-{\bf Dataset evaluated.} SZ First Episode Patients
+{\bf Dataset evaluated.} To validate that MAGeT performance generalises to other
+diseases, we measure the performance using the best parameter settings previous
+found, on a dataset consiting of first episode schizophrenia patients.
+\todo{Refer to the FEP description above... or just include it here}.
 
-{\bf Atlas and template libary.} two different atlas sets: a manual hippocampal
-segmentation of patients, and Winterburn atlas set.
+{\bf Atlas and template libary.} - two different atlas sets: a manual
+hippocampal segmentation of patients, and Winterburn atlas set.
 
@@
 \subsubsection{Winterburn Atlases Validation}
-{\bf Dataset evaluated.} T1 BRAVO scans of the same subjects included in the 
+{\bf Dataset evaluated.} - T1 BRAVO scans of the same subjects included in the 
 Winterburn atlas set.  These scans are taken within .... weeks of the scans for 
@@
 
-{\bf Atlas and template libary.} Atlas library is Winterburn T1 atlases.  
+{\bf Atlas and template libary.} - Atlas library is Winterburn T1 atlases.  
 Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.
@@
 
-{\bf Evaluation.}  Leave one out cross-validation (LOOCV) in which all five
+{\bf Evaluation.}  - Leave one out cross-validation (LOOCV) in which all five
 subjects are segmented in separate runs of MAGeT brain.  In each run, the subject
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
 %                               Results
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Results}
+\section{Results}
 
@@
 %%%%%
-\subsubsection{ADNI-1 Cross-Validation}
+\subsection{ADNI-1 Cross-Validation}
 
 - find significant improvement over multi-atlas performed with the same
-  parameters. Also, find smoothed performance is monotonically increasing but asymptotic
-  in size of both template and atlas library, with peak performance reached after 15
-  templates. \todo{Can we statistically capture "peak" performance?  The point at which 
-  gains become statistically insignificant?}
+  parameters. Also, find smoothed performance is monotonically increasing but
+  asymptotic in size of both template and atlas library, with peak performance
+  reached after 15 templates.
+
+\todo{Can we statistically capture what "peak" performance is?  Something like,
+the point at which gains become statistically insignificant?}
 
@@
 @
-  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line fitted using GAM (generalised additive model) from R with defaults from ggplot2 (formula: y ~ s(x, bs = "cs"))}
+  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line
+  fitted using GAM (generalised additive model) from R with defaults from
+  ggplot2 (formula: y ~ s(x, bs = "cs"))}
   \label{ADNI1_cross_validation_means}
@@
   scale_colour_hue(name="Atlases") +
-  xlab( "Number of Templates" ) + ylab( "Difference in mean Kappa between voting method and multiatlas" )
+  xlab( "Number of Templates" ) + 
+  ylab( "Difference in mean Kappa between voting method and multiatlas" )
 @
@@
 
-Multi-atlas means:
+<<a2a_multiatlas_means, results="asis",cache=FALSE>>=
+latex(merge(a_ma_means, t_ma_means, by="num_atlases"),
+colheads=c("Atlases", "ANTS", "ANIMAL"), file="", dec=2,
+	caption="Multi-atlas means")
+@
 
-<<a2a_multiatlas_means, results="asis",cache=FALSE>>=
-latex(merge(a_ma_means, t_ma_means, by="num_atlases"), colheads=c("Atlases", "ANTS", "ANIMAL"), file="", dec=2)
-@
-\begin{table}
-    \begin{tabular}{c|c|c}
-        Number of Atlases & ANIMAL Kappa (Jaccard) & ANTS Kappa (Jaccard) \\ \hline
-        3                 & 0.76 (0.63)             & 0.80 (0.69)          \\ 
-        4                 & 0.75 (0.62)             & 0.79 (0.67)          \\ 
-        5                 & 0.79 (0.66)             & 0.82 (0.71)          \\ 
-        6                 & 0.78 (0.65)             & 0.82 (0.70)          \\ 
-        7                 & 0.80 (0.67)             & 0.83 (0.72)          \\ 
-        8                 & 0.79 (0.67)             & 0.83 (0.72)          \\ 
-        9                 & 0.80 (0.68)             & 0.84 (0.73)          \\
-    \end{tabular}
-\end{table}
+%\begin{table}
+%    \begin{tabular}{c|c|c}
+%        Number of Atlases & ANIMAL Kappa (Jaccard) & ANTS Kappa (Jaccard) \\ \hline
+%        3                 & 0.76 (0.63)             & 0.80 (0.69)          \\ 
+%        4                 & 0.75 (0.62)             & 0.79 (0.67)          \\ 
+%        5                 & 0.79 (0.66)             & 0.82 (0.71)          \\ 
+%        6                 & 0.78 (0.65)             & 0.82 (0.70)          \\ 
+%        7                 & 0.80 (0.67)             & 0.83 (0.72)          \\ 
+%        8                 & 0.79 (0.67)             & 0.83 (0.72)          \\ 
+%        9                 & 0.80 (0.68)             & 0.84 (0.73)          \\
+%    \end{tabular}
+%\end{table}
 
-- more atlases -> better performance
-- larger template library -> better performance, but tails off around 10-15 templates
-- no significant difference between majority or weighted vote methods (haven???t tested this statistically though). 
-- consistently performs better than average naive performance by XXX
-- using ANTS, with a large enough template library (>12) MAGeT brain performs
+Ideas:
+\begin{itemize}
+\item more atlases -> better performance
+\item  larger template library -> better performance, but tails off around 10-15 templates
+\item  no significant difference between majority or weighted vote methods (haven???t tested this statistically though). 
+\item  consistently performs better than average naive performance by XXX
+\item  using ANTS, with a large enough template library (>12) MAGeT brain performs
   better than the average multi-atlas approach with the same number of atlases.
   using ANIMAL, 5 or more atlases needed before boost seen. 
-- more atlases -> smaller template library required to improve on average multi-atlas performance
-- discuss variance?  best/worst case? --how often do we expect random template library selection to work decently
+\item  more atlases -> smaller template library required to improve on average multi-atlas performance
+\item  discuss variance?  best/worst case? --how often do we expect random template library selection to work decently
+\end{itemize}
 
-\todo{cost (in registrations) / benefit trade off graph:  show number of registrations per Kappa?  or hours of manual labour per Kappa?)}
+\todo{show cost (in registrations) / benefit tradec off graph:  show number of
+registrations per Kappa?  or hours of manual labour per Kappa?)}
 
-\subsubsection{Winterburn Atlas Segmentation of ADNI Baseline Images}
+%%%%%
+%%%%%                  Results: ADNI-1 Screening Validation               %%%%%
+%%%%%
+\subsection{ADNI-1 Screen Validation}
 
-- A2A shows that if atlas population strongly(?) represents subject set
+Ideas: 
+\begin{itemize}
+\item  A2A shows that if atlas population strongly(?) represents subject set
 variability, then free choice from atlas population will produce improvements
 (we know this b/c of extensive validation trials). 
-- what about in the case where atlas population doesn???t strongly represent
-  subject set variability (e.g. a priori atlas set)?  then, we can use atlas
-selection to refine atlas set?
+\item  what about in the case where atlas population doesn???t strongly
+represent subject set variability (e.g. a priori atlas set)?  then, we can use
+atlas selection to refine atlas set? 
+\end{itemize}
 
-\todo{Kappa against our manual rater is low}
+\todo{Kappa against our manual rater is low... explain that}
 
@@
 # recode VISCODE
-
 viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
@@
                   FSL   = FSL_L + FSL_R)
-write.csv(vols, "data/ADNI_baseline_volumes_combined_volumes.csv")
 @
@@
 <<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-totals = read.csv("data/ADNI_baseline_volumes_combined_volumes.csv")
+totals = vols
 qplot(DX,value,data=melt(totals,measure.vars=c("FS", "FSL","UCSD", "MAPER","MAGeT","SNT"),na.rm=T),
@@
 @
-  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
+  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
   \label{ADNI-baseline-volumes-plot}
@@
 
-\subsubsection{First Episode Schizophrenic patients}
-High volume correlation between Winterburn segmentation volumes and ground truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
+%%%%%
+%%%%%             Results: First Episode Patient Validation          %%%%%
+%%%%%
+\subsection{First Episode Schizophrenic Patients}
+High volume correlation between Winterburn segmentation volumes and ground
+truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
 
@@
 @
-  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total HC volumes for MAGeT against manually rated volumes of}
+  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
+  HC volumes for MAGeT against manually rated Hippocampal volumes}
   \label{SZ_volumes}
@@
 
-Some other text here... 
+%%%%%
+%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
+%%%%%
+\subsection{Winterburn Atlases Validation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                Discussion
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                               Conclusion 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+
+\bibliographystyle{plain}
+\bibliography{references}
 \end{document}

### 2013-01-16 paper.Rnw
Updated the intro organisation

@@
 \usepackage[square,numbers]{natbib}
+
+%draft mode
+\usepackage[colorlinks]{hyperref}
+\usepackage[colorinlistoftodos, textwidth=4cm, shadow]{todonotes}
+\usepackage[displaymath, tightpage]{preview}
+
 \setlength{\marginparwidth}{1.2in}
-\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
-\newcommand{\comment}[1]{\begin{kframe}{\textcolor{red}{#1}}\end{kframe}}
-\newcommand{\todo}[1]{\comment{TODO #1}}
-\newcommand{\mc}[1]{\comment{MC: #1}}
+
+%\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
+%\newcommand{\comment}[1]{\begin{kframe}{\textcolor{red}{#1}}\end{kframe}}
+%\newcommand{\todo}[1]{\comment{TODO #1}}
+%\newcommand{\mc}[1]{\comment{MC: #1}}
 \begin{document}
@@
 while still achieving similar accuracy to multi-atlas approaches.
-\todo{Finish this... }
+\todo[inline]{Finish this... }
 \end{abstract}
@@
 \todo{reference Chakravarty2012}
+
 In this paper, we use the term {\it atlas} to mean any manually segmented MR
@@
 
+The simplest form of multi-atlas segmentation combines labellings derived from
+several atlases by way of label fusion\citep{Collins2010}.  We will refer to
+this as {\em basic multi-atlas segmentation}.  The primary steps are as
+follows.  An atlas library and unlabelled MR images are given as input.  Each
+atlas image is nonlinearly registered to each unlabelled image, and then each
+atlas' labels are propagated via the resulting transformations.  The resulting
+labels are fused to produce a single, definitive segmentation by some label
+fusion method (for exampl, by voxel-wise majority voting).
+
 The segmentation approach we propose is best understood as an extension of
-basic multi-atlas segmentation \citep{Collins2010}.  In multi-atlas segmentation,
-an atlas library and unlabelled MR images are given as input.  Every atlas
-image is nonlinearly registered to each unlabelled image, and then each atlas'
-labels are propagated via the resulting transformations.  These labels are then
-fused to produce a single, definitive segmentation by some label fusion method
-(e.g.  voxel-wise majority vote).
+MAGeT brain adds a preliminary stage to this process in which the template
+library is constructed rather than given as input.  As before, MAGeT brain
+accepts an atlas library and unlabelled MR images as input.  Images for the
+template library are selected from
 
-Our extension adds a preliminary stage in which a template library is
-constructed from input images, and used in place of an atlas library in the
-standard multi-atlas-based method.  To create the template library, labels from
-each atlas image are propagated to each template library image via the
-transformation resulting from a non-linear registration between pair of images.
-As a result, each template library image has a label from each atlas.
-Basic multi-atlas segmentation is then used to produce segmentations for the
-entire set of unlabelled images (including those images used in the template
-library).
+To create the template library, labels from each atlas image are propagated to each template
+library image via the transformation resulting from a non-linear registration
+between pair of images.  As a result, each template library image has a label
+from each atlas.  Basic multi-atlas segmentation is then used to produce
+segmentations for the entire set of unlabelled images (including those images
+used in the template library).
+
+\todo{Note that our algorithm does not specify a registration algorithm or
+fusion method\citep{Chakravarty2012}.  Performance on options is tested in our
+experiments}
 
@@
   \begin{algorithmic}
-    \Function{MultiAtlas}{Templates, Subjects}
+    \Function{BasicMultiAtlasSegmentation}{Templates, Subjects}
       \ForAll{$subject$}
@@
 
-\todo{Note that our algorithm does not specify a registration algorithm or
-fusion method\citep{Chakravarty2012}.  Performance on options is tested in our
-experiments}
-
 %%%%%                             Subjects                                 %%%%%
@@
 
-\begin{figure*}
 <<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
-      caption="ADNI-1 1.5T Screening demographics")
+latex(tab, file="", size="scriptsize",  caption="ADNI-1 1.5T Screening demographics")
 @
-\end{figure*}
 
@@
 \todo{include image characteristics}
-\begin{figure*}
+
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
@@
                data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", size="scriptsize", 
+latex(tab, file="", size="scriptsize", ctable=FALSE,
       caption="Schizophrenia First Episode Patient Demographics")
 @
-\end{figure*}
 
@@
 We used the following command line when running the ANTS command, 
+{\small
 \begin{verbatim}
@@
  \end{verbatim}
+}
 These settings were adapted from the "reasonable starting point" given in the
@@
 
-\begin{table*}
 <<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
       caption="ADNI-1 cross-validation subset demographics")
 @
-\end{table*}
 

### 2013-01-17 paper.Rnw
Added short summary of competing methods

@@
 
-\documentclass[twocolumn]{article}
+\documentclass{article}
 \usepackage[margin=1in]{geometry}
@@
 variability in the hippocampus throughout the course of aging, development, and
-neuropsychiatric disorders \citep{Mouiha2011}.  Additionally, it may be
+neuropsychiatric disorders \citep{Mouiha2011b}.  Additionally, it may be
 necessary to use several different hippocampal definitions or, in fact, make
@@
 
+%  Automated methods 
+%
+%  Heckemann2006 - 
+%  Heckemann2011 -  Automatic morphometry in AD and MCI 
+%    - (MAPER). multi-atlas. 30 atlases, entire ADNI screening  
+%  Collins2010 - .. seg of HC and AM by augmenting ANIMAL
+%    - multi-atlas 80 atlases, atlas selection (NMI)
+%  Aljabar2009 - atlas selection and effect on accuracy
+%    - 
+%  Leung2010 - X-sectional and longitudinal HC vol in MCI, AD
+%    - (MAPS), multi-atlas + STAPLE. 55 atlases. baseline+1yr of ADNI seg.
+%    - atrophy rates
+%  Lotjonen2010 - Fast and robust
+%    - IBSR/ADNI.  
+%
+%
+%  Hippocampal Volume
+%
+%  Mouiha2011b - Hippocampal atrophy rates in AD
+%    - compare SNT vs FreeSurfer on ADNI
+%
 Automated segmentation techniques overcome the need for human expertise by
@@
 dramatically reduce the number of manually labelled atlases needed (under 10).
-This is achieved by using the small atlas library to boot-strap a much larger
+This is achieved by using the small atlas library to bootstrap a much larger
 "template library", which is then used to segment the subjects in a similar
@@
 
-\todo{Is it okay to reference results whilst still in Methods?}
+\todo{okay to reference results whilst still in Methods?}
 
@@
 with adni/our segmentations}
+
+\todo{Also consider self-consistency (i.e. Kappa of same subject at different
+field strengths) a la Heckemann2011}
+
+\todo{Also: group differences, expecting separation}
 

### 2013-01-22 paper.Rnw
Updated introduction to include a clearer argument for MAGeT brain as addressing segmentation with small atlas sets

@@
 variability in the hippocampus throughout the course of aging, development, and
-neuropsychiatric disorders \citep{Mouiha2011b}.  Additionally, it may be
-necessary to use several different hippocampal definitions or, in fact, make
-specific modifications in the course of research. For example, Poppenk et al.
+neuropsychiatric disorders \citep{Mouiha2011b}.  The result is that existing
+hippocampal atlases available to a researcher may not accurately represent
+neuroanatomy of a specific population under study.  Additionally, in the course
+of a research or clinical study, it may be necessary to make adjustments to
+hippocampal definition as a means of hypothesis testing.  For example, Poppenk
 \citep{Poppenk2011} found that subdividing the hippocampus into anterior and
 posterior regions resulted in a predictive relationship between volume
-difference of those regions and recollection memory performance.  Thus, while
-manual segmentation of the hippocampus is a necessary technique, to researchers
-or clinicians who do not have access to the needed human expertise its use may
-be infeasible.
+difference of those regions and recollection memory performance.  Making
+modifications to an set of atlas library segmentations may require the same
+level of expertise and manual effort needed to create the segmentations from
+scratch.
 
@@
 %    - IBSR/ADNI.  
-%
-%
-%  Hippocampal Volume
-%
 %  Mouiha2011b - Hippocampal atrophy rates in AD
@@
 performing segmentations computationally.  A popular class of automated
-methods, {\it multi-atlas-based segmentation}, rely on a set of expertly
-labeled neuroanatomical atlases. Each atlas is warped to fit a subject's
-neuroanatomy using nonlinear registration
-techniques \citep{Collins1995,Klein2009}.  Atlas labels are then transformed by
-this warping and a {\it label fusion} technique, such as voxel-wise voting, is
-used to merge the competing label definitions from each atlas into a final
-segmentation for a subject.
+methods, {\it multi-atlas-based segmentation}, use a series of expertly
+segmented MRIs (an atlas library) to act as templates against which a target
+image is labelled
+\cite{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
+In the basic method, each atlas is warped to fit a subject's neuroanatomy using
+nonlinear registration techniques \citep{Collins1995,Klein2009}.  Atlas labels
+are then transformed by this warping and a {\it label fusion} technique, such
+as voxel-wise voting, is used to merge the competing label definitions from
+each atlas into a final segmentation for a subject.
 
-Many descriptions of multi-atlas-based segmentation algorithms report relying
-on an atlas library containing anywhere between 30 and 80 expertly labeled
-brains \citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}.  As
-noted, the production of an atlas library requires significant manual effort,
-and is limited since the choice of atlases or segmentation protocol may not
-reflect the underlying neuroanatomical variability of the population under
-study or be suited to answer the research questions at hand.
+Many descriptions of multi-atlas-based hippocampal segmentation algorithms
+report relying on an atlas library containing anywhere between 30 and 80
+expertly labeled images
+\citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010} in order
+to achieve performance approach that of a manual rater.  As already noted, the
+production of an atlas library requires significant manual effort and
+expertise, as does the modification of an existing atlas library in order to
+better suit a target population or investigate a specific research hypothesis.
+Therefore, in this paper we tackle the problem of producing accurate
+segmentations using small numbers of manually segmented atlases.
 
-In this paper we propose an automated segmentation method to address the above
-issues of existing multi-atlas-based methods. Principally, our method aims to
-dramatically reduce the number of manually labelled atlases needed (under 10).
-This is achieved by using the small atlas library to bootstrap a much larger
-"template library", which is then used to segment the subjects in a similar
-fashion to basic multi-atlas segmentation. This approach has the additional
-advantage of using the unique subject population on hand to initialize the
-segmentation process and improve accuracy.
+In this paper we propose an extension to the basic multi-atlas-based
+segmentation scheme to address this issue.  Principally, we explore the
+possibility of using a small atlas library to bootstrap a much larger
+\emph{template library} composed of images taken from the target population.
+The template library is then used to segment the subjects in a similar fashion
+to basic multi-atlas segmentation: by label propagation and label fusion.  This
+intuition driving this approach is to leverage the unique neuroanatomy of
+subject population on hand to initialize the segmentation process and improve
+accuracy over direct propagation from the atlas library to unlabelled subjects.
 
 The essential insight of generating a template library is not new.  Heckemann
-\citep{Heckemann2006} compared generating a template library from a single atlas
-to standard multi-atlas segmentation and found poor performance and so deemed
-the approach as inviable.  The LEAP algorithm \citep{Wolz2010} proceeds by
-iteratively segmenting the unlabelled image most similar to the atlas library
-images and then incorporating the now-labelled image into the atlas library,
-but requires 30 starting atlases.  The novelty of our method is to demonstrate
-the possibility of producing comparable segmentation accuracy to these and
-other multi-atlas-based methods while using significantly fewer manually
-created atlases.
+\citep{Heckemann2006} briefly explored "indirect" segmentation by taking a
+single atlas and propagating those labels to intermediate subjects before
+fusing them in a target image space.  Heckemann compared indirect single-atlas
+segmentation to direct multi-atlas segmentation found the indirect approach
+performed worse.  In this paper we continue the same line of investigation but
+explore the performance when using multiple atlases, as well as several
+different registration and fusion methods.
 
-In our previous work \citep{Chakravarty2012}, we applied MAGeT brain to
+The LEAP algorithm \citep{Wolz2010} is another example of indirect segmentation
+previously explored.  LEAP proceeds by iteratively segmenting unlabelled images
+most similar to the atlas library images and then incorporating the
+now-labelled images into the atlas library for future iterations.  30 manually
+labelled atlases are required to start the process.  The novelty explored in
+our current work is demonstrate the viability of acheiving comparable
+segmentation accuracy while using significantly fewer manually created atlases.
+
+In previous work \citep{Chakravarty2012}, we applied MAGeT brain to
 segmentation of the human striatum, globus pallidus, and thalamus using a
 single histologically-derived atlas. The main contribution of this paper is to
-extend our approach to the human hippocampus and perform a thorough validation
-over a range of atlas and template library sizes, which was not done in our
-previous work.  Due to the small number of atlases required, our method can
-easily accommodate different hippocampal definitions. Our aim is not to improve
-on segmentation accuracy beyond existing methods, but instead to provide a
-method that trades off manual segmentation expertise for computational
-processing time while providing sufficient accuracy for clinical and research
-applications.
+extend our approach to the human hippocampus - a much more difficult structure
+to segment -- and perform a thorough validation over a range of atlas and
+template library sizes, registration methods, and fusion methods.  Our aim in
+this work is not necessarily to improve on segmentation accuracy beyond
+existing methods, but instead to provide a method that trades off manual
+segmentation expertise and time for computational processing time while still
+providing sufficient accuracy for clinical and research applications.
 

### 2013-01-24 paper.Rnw
Incorporated changes from some of Mallar's comments

@@
 \title{Hippocampal segmentation with MAGeT}
-\author{Jon Pipitone}
+\author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., Voineskos A., Chakravarty M.M., and the Alzheimer's Disease Neuroimaging Initiative}
 \maketitle
@@
 while still achieving similar accuracy to multi-atlas approaches.
-\todo[inline]{Finish this... }
+\todo[inline]{finish}
 \end{abstract}
@@
 The hippocampus is of particular interest to many researchers because it is
-implicated in forms of brain dysfunction such as Alzheimer's disease and
-schizophrenia, and has functional significance in cognitive processes such as
-learning and memory.  For many research questions involving magnetic resonance
-imaging (MRI) data accurate identification of the hippocampus and its
-subregions is a necessary first step to better understand the individual
-neuroanatomy of subjects.
+implicated in forms of brain dysfunction such as Alzheimer's
+disease\citep{Sabuncu2011} and schizophrenia\citep{Narr2004,Karnik-Henry2012},
+and has functional significance in cognitive processes such as learning and
+memory\citep{DenHeijer2012,Scoville2000}.  For many research questions
+involving magnetic resonance imaging (MRI) data accurate identification of the
+hippocampus and its subregions is a necessary first step to better understand
+the individual neuroanatomy of subjects.
 
 Currently, the gold standard for neuroanatomical segmentation is manual
-delineation by an expert human rater.  This is problematic for segmentation of
-the hippocampus for several reasons.  First, manual segmentation takes a
+delineation by an expert human rater.  This is problematic for hippocampal
+segmentation for several reasons.  First, manual segmentation takes a
 significant investment of time and expertise \citep{Hammers2003} which may not
@@
 variability in the hippocampus throughout the course of aging, development, and
-neuropsychiatric disorders \citep{Mouiha2011b}.  The result is that existing
-hippocampal atlases available to a researcher may not accurately represent
-neuroanatomy of a specific population under study.  Additionally, in the course
-of a research or clinical study, it may be necessary to make adjustments to
-hippocampal definition as a means of hypothesis testing.  For example, Poppenk
-\citep{Poppenk2011} found that subdividing the hippocampus into anterior and
-posterior regions resulted in a predictive relationship between volume
-difference of those regions and recollection memory performance.  Making
-modifications to an set of atlas library segmentations may require the same
-level of expertise and manual effort needed to create the segmentations from
-scratch.
+neuropsychiatric disorders \citep{Mouiha2011b,Sabuncu2011}.  The result is that
+existing hippocampal atlases available to a researcher may not accurately
+represent neuroanatomy of a specific population under study.  Additionally, in
+the course of a research or clinical study, it may be necessary to make
+adjustments to hippocampal definition as a means of hypothesis testing.  For
+example, Poppenk \citep{Poppenk2011} found that subdividing the hippocampus
+into anterior and posterior regions resulted in a predictive relationship
+between volume difference of those regions and recollection memory performance.
+Making modifications to an set of atlas library segmentations may require the
+same level of expertise and manual effort needed to create the segmentations
+from scratch.
 
@@
 %
-Automated segmentation techniques overcome the need for human expertise by
-performing segmentations computationally.  A popular class of automated
-methods, {\it multi-atlas-based segmentation}, use a series of expertly
-segmented MRIs (an atlas library) to act as templates against which a target
-image is labelled
+Fully-automated segmentation techniques overcome the need for human expertise.
+In particular, {\it atlas-based} approaches make use of one or more expertly
+segmented MRIs (the atlas library) as templates to guide the segmentation of a
+query subject's MRI
 \cite{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
-In the basic method, each atlas is warped to fit a subject's neuroanatomy using
-nonlinear registration techniques \citep{Collins1995,Klein2009}.  Atlas labels
-are then transformed by this warping and a {\it label fusion} technique, such
-as voxel-wise voting, is used to merge the competing label definitions from
-each atlas into a final segmentation for a subject.
+The basic strategy is as follows. Each atlas is warped to fit a subject's
+neuroanatomy using nonlinear registration techniques
+\citep{Collins1995,Klein2009}.  Atlas labels are then transformed by this
+warping to the subject's space, and a {\it label fusion} technique, such as
+voxel-wise voting, is used to merge the competing label definitions from each
+atlas into a final segmentation for the subject.
 
-Many descriptions of multi-atlas-based hippocampal segmentation algorithms
-report relying on an atlas library containing anywhere between 30 and 80
-expertly labeled images
-\citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010} in order
-to achieve performance approach that of a manual rater.  As already noted, the
-production of an atlas library requires significant manual effort and
-expertise, as does the modification of an existing atlas library in order to
-better suit a target population or investigate a specific research hypothesis.
-Therefore, in this paper we tackle the problem of producing accurate
-segmentations using small numbers of manually segmented atlases.
+Multi-atlas hippocampal segmentation algorithms achieve some of the best
+automated segmentation accuracies to-date, and approach the performance of a
+manual rater
+\citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}, but rely
+on an atlas library containing anywhere between 30 and 80 expertly labeled
+images to achieve this accuracy.  As already noted, the production of an atlas
+library requires significant manual effort and expertise, as does the
+modification of an existing atlas library in order to better suit a target
+population or investigate a specific research hypothesis.  Therefore, in this
+paper we tackle the problem of producing accurate segmentations using small
+numbers of manually segmented atlases.
 

### 2013-01-24 paper.Rnw
Full incorporated comments from Mallar into the introduction

@@
 \title{Hippocampal segmentation with MAGeT}
-\author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., Voineskos A., Chakravarty M.M., and the Alzheimer's Disease Neuroimaging Initiative}
+\author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., Voineskos
+A., Chakravarty M.M., and the Alzheimer's Disease Neuroimaging Initiative}
 \maketitle
@@
 The template library is then used to segment the subjects in a similar fashion
-to basic multi-atlas segmentation: by label propagation and label fusion.  This
-intuition driving this approach is to leverage the unique neuroanatomy of
-subject population on hand to initialize the segmentation process and improve
-accuracy over direct propagation from the atlas library to unlabelled subjects.
+to basic multi-atlas segmentation: by label propagation and label fusion.  We
+call our approach MAGeT brain: {\em m}ultiple {\em a}utomatically {\em
+ge}nerated {\em t}emplate {\em brain} segmentation. The intuition driving this
+approach is that by generating a template library we leverage the unique
+neuroanatomy of subject population on hand to initialize the segmentation
+process and improve accuracy over direct propagation from the atlas library to
+unlabelled subjects while also using fewer manually segmented atlases.
 
-The essential insight of generating a template library is not new.  Heckemann
+The insight of generating a template library is not new.  Heckemann
 \citep{Heckemann2006} briefly explored "indirect" segmentation by taking a
-single atlas and propagating those labels to intermediate subjects before
-fusing them in a target image space.  Heckemann compared indirect single-atlas
+single atlas and propagating those labels to intermediate subjects before fusing
+them in a target image space.  Heckemann compared indirect single-atlas
 segmentation to direct multi-atlas segmentation found the indirect approach
@@
 
-In previous work \citep{Chakravarty2012}, we applied MAGeT brain to
-segmentation of the human striatum, globus pallidus, and thalamus using a
-single histologically-derived atlas. The main contribution of this paper is to
-extend our approach to the human hippocampus - a much more difficult structure
-to segment -- and perform a thorough validation over a range of atlas and
-template library sizes, registration methods, and fusion methods.  Our aim in
-this work is not necessarily to improve on segmentation accuracy beyond
-existing methods, but instead to provide a method that trades off manual
-segmentation expertise and time for computational processing time while still
-providing sufficient accuracy for clinical and research applications.
+In previous work \citep{Chakravarty2012}, we applied MAGeT brain to segmentation
+of the human striatum, globus pallidus, and thalamus using a single
+histologically-derived atlas. The contribution of the present work is to extend
+our approach to the human hippocampus and perform a rigorous validation. We
+first conduct an intensive cross-validation of MAGeT on a pool of 69 MRIs from
+the ADNI database, comparing the performance of MAGeT against basic multi-atlas
+segmentation using a range of atlas and template library sizes, registration and
+fusion methods.  We then compare MAGeT segmentations of the entire ADNI
+screening dataset with those of established automated and manual segmentation
+techniques.  Finally, we compare MAGeT segmentations to those of manual
+segmentations from a dataset of first episode schizophrenic patient MRIs.
 

### 2013-01-30 paper.Rnw
Some work on improving the literature review in the introduction
Starting to pull apart the intro

@@
 
-%  Automated methods 
+% atlas-based methods -> a prior feature based methods?
 %
-%  Heckemann2006 - 
-%  Heckemann2011 -  Automatic morphometry in AD and MCI 
-%    - (MAPER). multi-atlas. 30 atlases, entire ADNI screening  
+% - basic multi-atlas method, 
 %  Collins2010 - .. seg of HC and AM by augmenting ANIMAL
@@
 %  Aljabar2009 - atlas selection and effect on accuracy
-%    - 
+% - extensions and practical applications
 %  Leung2010 - X-sectional and longitudinal HC vol in MCI, AD
@@
 %    - IBSR/ADNI.  
+% - Coutin's use of shape-based priors
+%
+% - multi-atlas with specialised/small atlas libraries
+%   - Fischl - freesurfer, high-resolution atlases + probability priors
+%   - Yushkevitch - invivo focal HF-MRI 
+%  
+%  Heckemann2011 -  Automatic morphometry in AD and MCI 
+%    - (MAPER). multi-atlas. 30 atlases, entire ADNI screening  
 %  Mouiha2011b - Hippocampal atrophy rates in AD
@@
 %
-Fully-automated segmentation techniques overcome the need for human expertise.
-In particular, {\it atlas-based} approaches make use of one or more expertly
-segmented MRIs (the atlas library) as templates to guide the segmentation of a
-query subject's MRI
-\cite{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
+%
+Fully-automated segmentation techniques require no human expertise beyond that
+of an initial configuration.  Multi-atlas automated segmentation algorithms make
+use of a series of expertly segmented MRIs (the atlas library) as models to
+guide the segmentation of a query subject's MRI, and achieve some of the best
+automated hippocampal segmentation accuracies to-date
+
+\citep{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
 The basic strategy is as follows. Each atlas is warped to fit a subject's
@@
 
-Multi-atlas hippocampal segmentation algorithms achieve some of the best
-automated segmentation accuracies to-date, and approach the performance of a
-manual rater
-\citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}, but rely
+but rely
 on an atlas library containing anywhere between 30 and 80 expertly labeled

### 2013-02-03 paper.Rnw
Outline of the intro

@@
 
-% atlas-based methods -> a prior feature based methods?
-%
-% - basic multi-atlas method, 
-%  Collins2010 - .. seg of HC and AM by augmenting ANIMAL
-%    - multi-atlas 80 atlases, atlas selection (NMI)
-%  Aljabar2009 - atlas selection and effect on accuracy
-% - extensions and practical applications
-%  Leung2010 - X-sectional and longitudinal HC vol in MCI, AD
-%    - (MAPS), multi-atlas + STAPLE. 55 atlases. baseline+1yr of ADNI seg.
-%    - atrophy rates
-%  Lotjonen2010 - Fast and robust
-%    - IBSR/ADNI.  
-% - Coutin's use of shape-based priors
-%
-% - multi-atlas with specialised/small atlas libraries
-%   - Fischl - freesurfer, high-resolution atlases + probability priors
-%   - Yushkevitch - invivo focal HF-MRI 
-%  
-%  Heckemann2011 -  Automatic morphometry in AD and MCI 
-%    - (MAPER). multi-atlas. 30 atlases, entire ADNI screening  
-%  Mouiha2011b - Hippocampal atrophy rates in AD
-%    - compare SNT vs FreeSurfer on ADNI
-%
-%
 Fully-automated segmentation techniques require no human expertise beyond that
@@
 
-\citep{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
 The basic strategy is as follows. Each atlas is warped to fit a subject's
@@
 atlas into a final segmentation for the subject.
+
+\citep{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
+
+## automated segmentation and the multi-atlas approach
+- automated segmentation techniques require no human intervention
+- model-based automatic segmentation approaches employ models of feature
+  variations as constraints during segmentation. 
+- multi-atlas segmentation techniques use many atlases and perform well
+- basic multi-atlas methods were pioneered by Heckmann et al in the MAPER
+  algorithm Heckemann 2006 and 2010. The strategy is as follows: 
+- individual atlases are treated as classifiers, registration is used to map the
+  classifiers to a subject space, and decision fusion is used to combine
+  expertise from classisfiers. define: atlas library
+- In Heckemann2006 a library of whole brain atlases are used, but Aljabar2009
+  shows that selecting a subset of the atlases most similar (using
+  cross-correlation of image intensities) to a subject and fusing only those
+  labels leads to better performance.  define: atlas selection.
+- Warfield2004: STAPLE as an alternative decision fusion technique?
+
+## application to  hippocampus
+- There are many applications of multi-atlas techniques to hippocampal
+  segmentation
+- collins2010 apply the multi-atlas approach to hippocampal and amygdala
+  segmentaiton using an atlas library of 80 atlases segmented using a custom
+  protocol (images selected from the ICBM152 dataset), the ANIMAL nonlinear
+  registration algorithm, normalised mutual information for atlas selection, and
+  majority vote for decision fusion, achieves a Dice Kappa similarity of 0.886.
+- Leung2010 (MAPS) uses an atlas library of 55 atlases (segmented using a
+  custom protocol) and STAPLE decision fusion to segment ADNI1 baseline and 12
+  month dataset. After selecting parameters for the best performance on the ADNI
+  dataset, they were able to achieve a Dice Kappa score of 0.89 compared to the
+  manual segmentations supplied with the ADNI data (see Methods for more
+  information on the ADNI dataset). 
+- Heckmann2011 used the MAPER algorithm (described earlier) with 30 atlases  to
+  segment ADNI1 1.5 and 3T baseline images, achieving a mean Dice Kappa
+  similarity score of 0.889. 
+- Lotjonen2012 explored various atlas selection methods, and used a custom
+  non-linear registration method based on intensity differences, with
+  post-processing of fused segmentations taking into account intensity
+  differences to tidy up the segmentations.  They applied this to 60 subjects
+  taken from the ADNI1 baseline dataset, using an atlas library of 30 subjects
+  (also taken from the ADNI1 baseline dataset) and achieved an mean Dice Kappa
+  similarity score of 0.885
+- Wolz2010 - differ significantly from the basic multi-atlas strategy in that
+  the atlas library is grown, starting with a set of manually labelled atlases
+  and successively incorporating unlabelled target images after labelling.
+  The sequence in which target images are labelled is such that the similarity
+  between the atlas images and the target images is minimised at each step,
+  effectively allowing for deformations between very dissimilar images to be
+  broken up into sequences of smaller deformations.  With an atlas library of
+  30 MR images, LEAP was used to segment the ADNI1 baseline dataset, achieving a
+  mean Dice score of 0.85 with manual segmentations.
+
+## specialised automated segmentation techiques of note
+- Fischl 2002 - describe a method which uses a probabilistic atlas of anatomical
+  and tissue classes along with spatial constraints for class labels encoded
+  using a Markov random field model. This forms the basis of FreeSurfer's whole
+  brain segmentation algorithm.
+- Fischl 2009 use a probabilistic atlas of anatomical classes as well as a
+  likelihood model of how those classes translate into MR image intensities.
+  Trained on manual segmentations of high resolution MR images, Bayesian
+  inference is used to classify voxels in an unseen image.  This forms the basis
+  of the FreeSurfer hippocampal subfield segmentation algorithm. 
+- Yushkevitch2010 describe a semi-automated method for hippocampal subfield
+  segmentation of focal T2 images.  The unlabelled MR image must be manually
+  partitioned into 'head', 'body' and 'tail', and then multi-atlas methods are
+  used to segment the image (ANTS registration, and similarity-weighted vote).
+  Finally, an AdaBoost-based bias correction classifier trained on texture,
+  spatial location and segmentation is applied to fix mislabelled voxels.
+
+## with all of these different methods, how much does the atlas library matter?
+- all of these methods require a sizeable atlas library which, as we've
+  discussed, requires significant effort to create. how much does an atlas
+  library matter for the performance of an automated method?
+- As noted, Aljabar2009's work on atlas selection shows clearly that the
+  composition of the atlas library has a significant effect on accuracy of
+  automated segmentation.
+- what anatomical features play a role here? How does protocol impact?  
+- Carmichael2005 - explores which aspects of the single-atlas strategy matter
+  most: registration, similarity between atlas and gold-standard delineation
+  protocol, and neuroanatomical similarity between the atlas image and subject
+  image. They found that most flexible registration (non-linear) method, atlas
+  segmentation that follows the same protocol as the gold-standard and,
+  importantly for us, an atlas with neuroanatomical features that match the
+  subject image lead to the highest overlap of labels. 
+- Nestor2012 explored the effect of segmentation protocol on automated
+  multi-atlas segmentation performance of the hippocampus in the ADNI1 1 Year
+  dataset, and found that protocols that more inclusive protocols which include
+  dorsal white-matter and posterior anatomy tended to produce higher overlap and
+  better accuracy at distinguishing disease classes.
+
+## the problem of needing a large atlas library, and the benefit of having a
+   small one
+- time expensive: if you have one atlas from a difficult to obtain source?
+  e.g. histological atlas
+- customising definitions is tough, but necessary to ask research questions
+
+%  Chupin's use of shape-based priors 
+%   - Mouiha2011b - Hippocampal atrophy rates in AD: compare SNT vs FreeSurfer
+%     on ADNI
+%
+
+## introducing MAGeT brain for bootstrapping automated hippocampal segmentation
+- MAGeT bootstrapping: tune the template library to your dataset, reduce the
+  number of input atlases you need, with the upshot that with fewer atlases it's
+  that much easier to change definitions in order to test hypothesis
 

### 2013-02-05 paper.Rnw
Complete introduction with rewrites to include even more literature

--- 2013-01-24 paper.Rnw
@@
 
-%  Automated methods 
+Fully-automated segmentation techniques require no human expertise beyond that
+of initial setup or training. One broad class of automated approach is that of
+the {\em model-based} approaches which employ models of anatomical feature
+variation to constrain segmentations. In this paper we focus on a subclass known
+as {\em multi-atlas} approaches which make use of a series of expertly segmented
+MRIs (the atlas library) as models to guide the segmentation of a query
+subject's MRI, and achieve some of the best automated hippocampal segmentation
+accuracies to-date.
+
+The multi-atlas segmentation technique was pioneered by Heckemann et al. with
+the MAPER algorithm \citep{Heckemann2006,Heckemann2011}. The basic strategy is as
+follows. Each atlas is warped to fit a subject's neuroanatomy using nonlinear
+registration techniques \citep{Collins1995,Klein2009}. Atlas labels are then
+transformed by this warping to the subject space and a {\it label fusion}
+technique, such as voxel-wise voting, is used to merge the label definitions
+from each atlas into a definitive segmentation. Typically, an atlas selection
+approach is used to improve accuracy by only using labels from atlases that are
+most similar to a subject image (cross-correlation of image intensities is a
+common similarity metric) \citep{Aljabar2009}.
+
+% - Warfield2004: STAPLE as an alternative decision fusion technique?
+
+Multi-atlas techniques have been applied very successfully to hippocampal
+segmentation. Collins et al. found that with an atlas library of 80 atlas
+images, the use of the ANIMAL nonlinear registration algorithm, the normalised
+mutual information as similarity metric for atlas selection, and majority vote
+for decision fusion they were able to achieve a mean Dice Kappa similarity score
+(DSC) to manual labels of 0.886\citep{Collins2010}. The Alzheimer's Disease
+Neuroimaging Initiative (ADNI) is a commonly used benchmarking dataset of MR
+images of controls and patients with MCI or Alzheimer's (see Methods for more
+information on the ADNI dataset).  Leung et al. tuned a multi-atlas  approach to
+the segmentation of ADNI1 1 year dataset of images and using 55 atlases and the
+STAPLE decision fusion method \citep{Warfield2004} achieved an mean DSC of 0.89
+to the manual segmentations supplied with the ADNI data  \citep{Leung2010}. The
+MAPER whole brain segmentation algorithm (described above), using 30 atlases,
+achieved a mean DSC score of 0.889 on all ADNI1 baseline images
+\citep{Heckemann2011}. In \citep{Lotjonen2012}, the authors use a proprietary
+non-linear registration method based on intensity differences, and
+post-processing step taking into account these intensity differences to tidy up
+fused segmentations. Using images from the ADNI1 baseline dataset, they achieved
+a mean DSC of 0.885 with 30 atlases. With the LEAP algorithm, Wolz et al.
+explored an elegant modification to the basic strategy: the atlas library is
+grown, beginning with a set of manually labelled atlases and successively
+incorporating unlabelled subject images after being labelling using multi-atlas
+techniques\citep{Wolz2010}.  The sequence in which subject images are labelled
+is chosen so that the similarity between the atlas images and the target images
+is minimised at each step, effectively allowing for deformations between very
+dissimilar images to be broken up into sequences of smaller deformations.  With
+an atlas library of 30 MR images, LEAP was used to segment the ADNI1 baseline
+dataset, achieving a mean Dice score of 0.85 with manual segmentations.
+
+While not purely nulti-atlas techniques, there are several important algorithms
+for hippocampal segmentation that inform our approach. The popular FreeSurfer
+application's whole brain segmentation algoritm uses a probabilistic atlas of
+anatomical and tissue classes along with spatial constraints for class labels
+encoded using a Markov random field model \citep{Fischl2002}. When segmenting
+hippocampal subfields, FreeSurfer employs a Bayesian inference algorithm using a
+probabilistic atlas of anatomical classes as a prior, and a likelihood model of
+how those classes translate into MR image intensities, both trained on manual
+segmentations of high resolution MR images \citep{Fischl2009}. Yushkevitch et
+al. describe a semi-automated method for hippocampal subfield segmentation of
+focal T2 images\citep{Yushkevitch2010}.  The unlabelled MR image must be
+manually partitioned into 'head', 'body' and 'tail', and then multi-atlas
+methods are used to segment the image.  Finally, an AdaBoost-based bias
+correction classifier is trained on texture, spatial location, and intensities
+of manual segmentations and is applied to fix mislabelled voxels.  
+%  Chupin's use of shape-based priors 
+%   - Mouiha2011b - Hippocampal atrophy rates in AD: compare SNT vs FreeSurfer
+%     on ADNI
 %
-%  Heckemann2006 - 
-%  Heckemann2011 -  Automatic morphometry in AD and MCI 
-%    - (MAPER). multi-atlas. 30 atlases, entire ADNI screening  
-%  Collins2010 - .. seg of HC and AM by augmenting ANIMAL
-%    - multi-atlas 80 atlases, atlas selection (NMI)
-%  Aljabar2009 - atlas selection and effect on accuracy
-%    - 
-%  Leung2010 - X-sectional and longitudinal HC vol in MCI, AD
-%    - (MAPS), multi-atlas + STAPLE. 55 atlases. baseline+1yr of ADNI seg.
-%    - atrophy rates
-%  Lotjonen2010 - Fast and robust
-%    - IBSR/ADNI.  
-%  Mouiha2011b - Hippocampal atrophy rates in AD
-%    - compare SNT vs FreeSurfer on ADNI
-%
-Fully-automated segmentation techniques overcome the need for human expertise.
-In particular, {\it atlas-based} approaches make use of one or more expertly
-segmented MRIs (the atlas library) as templates to guide the segmentation of a
-query subject's MRI
-\cite{Aljabar2009,Collins2010,Heckemann2011,Leung2010,Lotjonen2010,Leung2010,Wolz2010}.
-The basic strategy is as follows. Each atlas is warped to fit a subject's
-neuroanatomy using nonlinear registration techniques
-\citep{Collins1995,Klein2009}.  Atlas labels are then transformed by this
-warping to the subject's space, and a {\it label fusion} technique, such as
-voxel-wise voting, is used to merge the competing label definitions from each
-atlas into a final segmentation for the subject.
 
-Multi-atlas hippocampal segmentation algorithms achieve some of the best
-automated segmentation accuracies to-date, and approach the performance of a
-manual rater
-\citep{Heckemann2011,Collins2010,Aljabar2009,Leung2010,Lotjonen2010}, but rely
-on an atlas library containing anywhere between 30 and 80 expertly labeled
-images to achieve this accuracy.  As already noted, the production of an atlas
-library requires significant manual effort and expertise, as does the
-modification of an existing atlas library in order to better suit a target
-population or investigate a specific research hypothesis.  Therefore, in this
-paper we tackle the problem of producing accurate segmentations using small
-numbers of manually segmented atlases.
+Aside from the algorithmic choices used in multi-atlas segmentation, it is
+natural to ask about how the features of the atlases themselves impact the
+resulting segmentations. As noted, by choosing atlases ranked most similar to a
+subject image by voxel intensity profile segmentation accuracy is improved,
+suggesting that neuroanatomical similarity plays a role\citep{Aljabar2009}.
+Carmichael et al. explored this directly and found that when using only one
+atlas the important factors leading to improved accuracy are that the atlas have
+neuroanatomical features that match the subject, and that the atlas segmentation
+use the same protocol as the gold-standard \citep{Carmichael2005}. Nestor et al.
+found that hippocampal segmentation protocols that include more dorsal
+white-matter and posterior anatomy tended to produce higher overlap and better
+accuracy at distinguishing disease classes in the ADNI1 1 year dataset
+\citep{Nestor2012}. These results suggest both atlas library neuroanatomy and
+delineation protocol play a significant role in the resulting segmentation.
 
-In this paper we propose an extension to the basic multi-atlas-based
-segmentation scheme to address this issue.  Principally, we explore the
+Considered along with our earlier discussion on the difficulty of producing
+manual segmentations of MR images and the need for adaptable segmentation
+definitions in order to conduct research, this presents a real problem of labour
+and expertise when using existing multi-atlas segmentation methods which rely on
+relatively large atlas libraries (typically between 30 and 80 atlases). Indeed,
+it may be especially prohibitive to use these methods in situations where
+producing a single atlas is challenging (e.g. histology-based atlases, or
+atlases from very high resolution images). In this paper we address the problem
+of producing accurate segmentations using small numbers of manually segmented
+atlases.
+
+% MAGeT bootstrapping: tune the template library to your dataset, reduce the
+% number of input atlases you need, with the upshot that with fewer atlases it's
+% that much easier to change definitions in order to test hypothesis
+
+Our algorith, called MAGeT brain ({\em M}ultiple {\em A}utomatically {\em
+Ge}nerated {\em T}emplates), is an extension to the basic multi-atlas-based
+segmentation schema\citep{Chakravarty2012}. Principally, we explore the
 possibility of using a small atlas library to bootstrap a much larger
@@
 The template library is then used to segment the subjects in a similar fashion
-to basic multi-atlas segmentation: by label propagation and label fusion.  We
-call our approach MAGeT brain: {\em m}ultiple {\em a}utomatically {\em
-ge}nerated {\em t}emplate {\em brain} segmentation. The intuition driving this
-approach is that by generating a template library we leverage the unique
-neuroanatomy of subject population on hand to initialize the segmentation
-process and improve accuracy over direct propagation from the atlas library to
-unlabelled subjects while also using fewer manually segmented atlases.
+to basic multi-atlas segmentation: by label propagation and label fusion. The
+intuition driving this approach is that by generating a template library we
+leverage the unique neuroanatomy of subject population on hand to initialize the
+segmentation process and improve accuracy over direct propagation from the atlas
+library to unlabelled subjects while also using fewer manually segmented
+atlases.
 
-The insight of generating a template library is not new.  Heckemann
-\citep{Heckemann2006} briefly explored "indirect" segmentation by taking a
-single atlas and propagating those labels to intermediate subjects before fusing
-them in a target image space.  Heckemann compared indirect single-atlas
-segmentation to direct multi-atlas segmentation found the indirect approach
-performed worse.  In this paper we continue the same line of investigation but
-explore the performance when using multiple atlases, as well as several
+The insight of generating a template library is not new.  Heckemann et al.
+compared "indirect" segmentation -- taking a single atlas and propagating the
+labels to intermediate subjects before fusing them in a target image space -- to
+multi-atlas segmentation and found that the indirect approach performed worse
+\citep{Heckemann2006}. In this paper we continue the same line of investigation
+but explore the performance when using multiple atlases as well as the effect of
 different registration and fusion methods.
 
-The LEAP algorithm \citep{Wolz2010} is another example of indirect segmentation
-previously explored.  LEAP proceeds by iteratively segmenting unlabelled images
-most similar to the atlas library images and then incorporating the
-now-labelled images into the atlas library for future iterations.  30 manually
-labelled atlases are required to start the process.  The novelty explored in
-our current work is demonstrate the viability of acheiving comparable
-segmentation accuracy while using significantly fewer manually created atlases.
+The LEAP algorithm \citep{Wolz2010}, described above, is another example of
+indirect segmentation previously explored.  LEAP proceeds by iteratively
+segmenting unlabelled images most similar to the atlas library images and then
+incorporating the labelled images into the atlas library for future iterations.
+The novelty explored in our current work is demonstrate the viability of
+achieving comparable segmentation accuracy while using significantly fewer
+manually created atlases.
 
@@
 labels are fused to produce a single, definitive segmentation by some label
-fusion method (for exampl, by voxel-wise majority voting).
+fusion method (for example, by voxel-wise majority voting).
 

### 2013-02-05 paper.Rnw
Updated methods to clarify MAGeT, and add canned description of ADNI1

@@
 The novelty explored in our current work is demonstrate the viability of
-achieving comparable segmentation accuracy while using significantly fewer
-manually created atlases.
+achieving comparable segmentation accuracy using the basic multi-atlas schema
+and using significantly fewer manually created atlases.
 
@@
 The simplest form of multi-atlas segmentation combines labellings derived from
-several atlases by way of label fusion\citep{Collins2010}.  We will refer to
+several atlases by way of label fusion \citep{Heckemann2006}.  We will refer to
 this as {\em basic multi-atlas segmentation}.  The primary steps are as
-follows.  An atlas library and unlabelled MR images are given as input.  Each
-atlas image is nonlinearly registered to each unlabelled image, and then each
-atlas' labels are propagated via the resulting transformations.  The resulting
-labels are fused to produce a single, definitive segmentation by some label
-fusion method (for example, by voxel-wise majority voting).
+follow are: (1) an atlas library and unlabelled MR images are given as input,
+(2) each atlas image is nonlinearly registered to each unlabelled image, (4)
+each atlas' labels are propagated via the resulting transformations to the
+unlabelled image space, and (5) the resulting labels are fused to produce a
+single, definitive segmentation. The particular registration and voting method
+used are left unspecified.
 
-The segmentation approach we propose is best understood as an extension of
-MAGeT brain adds a preliminary stage to this process in which the template
-library is constructed rather than given as input.  As before, MAGeT brain
-accepts an atlas library and unlabelled MR images as input.  Images for the
-template library are selected from
+The segmentation approach we propose is best understood as an extension of MAGeT
+brain adds a preliminary stage to this process in which the template library is
+constructed rather than given as input.  As before, MAGeT brain accepts an atlas
+library and unlabelled MR subject images as input.  Template library images are
+selected from subject images.  The choice of subjects used in the template
+library can be made to reflect the neuroanatomy or demographics of the subject
+set as a whole (for instance, by sampling equally from cases and controls).
 
-To create the template library, labels from each atlas image are propagated to each template
-library image via the transformation resulting from a non-linear registration
-between pair of images.  As a result, each template library image has a label
-from each atlas.  Basic multi-atlas segmentation is then used to produce
-segmentations for the entire set of unlabelled images (including those images
-used in the template library).
+To create the template library, labels from each atlas image are propagated to
+each template library image via the transformation resulting from a non-linear
+registration between pair of images.  As a result, each template library image
+has a label from each atlas.  Basic multi-atlas segmentation is then used to
+produce segmentations for the entire set of unlabelled images (including those
+images used in the template library). Figure X describes the MAGeT brain
+algorithm in pseudocode.
 
-\todo{Note that our algorithm does not specify a registration algorithm or
-fusion method\citep{Chakravarty2012}.  Performance on options is tested in our
-experiments}
-
-Source code can be found at \url{http://github.com/pipitone/MAGeTbrain}.
+\footnote{Source code can be found at
+\url{http://github.com/pipitone/MAGeTbrain}.}
 
@@
 
+Clinical, demographic and pre-processed T1-weighted MRI used in this paper were
+downloaded by the authors from the ADNI1 database (adni.loni.ucla.edu) from
+between March 2012 and August 2012.
+
+All images were acquired using 1.5T scanners (General Electric Healthcare,
+Philips Medical Systems or Siemens Medical Solutions) at multiple sites from the
+ADNI website.  "Representative imaging parameters were TR = 2400ms, TI = 1000ms,
+TE = 3.5ms, flip angle = 8°, field of view = 240 × 240mm and 160 sagittal
+1.2mm-thick-slices and a 192 × 192 matrix yielding a voxel resolution of 1.25 ×
+1.25 × 1.2 mm3, or 180 sagittal 1.2mm-thick-slices with a 256 × 256 matrix
+resolution of 1.25 × 1.25 × 1.2 mm3, or 180 sagittal 1.2mm-thick-slices with a
+256 × 256 matrix resolution of 1.25 × 1.25 × 1.2 mm3, or 180 sagittal
+1.2mm-thick-slices with a 256 × 256 matrix yielding a voxel resolution of 0.94 ×
+0.94 × 1.2 mm3." The details of the ADNI MR imaging protocol are described in
+\citep{Jack2008a}. The T1-weighted scans that were designated to be the “best”
+after quality control were processed using the standard ADNI image processing
+pipeline, which included post-acquisition correction of gradient warping
+(Jovicich et al., 2006), B1 non-uniformity correction (Narayana et al., 1998)
+depending on the scanner and coil type, intensity non-uniformity correction
+(Sled et al., 1998) and phantom based scaling correction (Gunter et al., 2006) -
+the geometric phantom scan having been acquired with each patient scan.
+
+%We downloaded pre-processed baseline and 12-month repeat volumetric T1-weighted
+%MR scans acquired using 1.5T scanners (General Electric Healthcare, Philips
+%Medical Systems or Siemens Medical Solutions) at multiple sites from the ADNI
+%website.  Representative imaging parameters were TR = 2400ms, TI = 1000ms, TE =
+%3.5ms, flip angle = 8°, field of view = 240 × 240mm and 160 sagittal
+%1.2mm-thick-slices and a 192 × 192 matrix yielding a voxel resolution of 1.25 ×
+%1.25 × 1.2 mm3, or 180 sagittal 1.2mm-thick-slices with a 256 × 256 matrix
+%resolution of 1.25 × 1.25 × 1.2 mm3, or 180 sagittal 1.2mm-thick-slices with a
+%256 × 256 matrix resolution of 1.25 × 1.25 × 1.2 mm3, or 180 sagittal
+%1.2mm-thick-slices with a 256 × 256 matrix yielding a voxel resolution of 0.94 ×
+%0.94 × 1.2 mm3. The details of the ADNI MR imaging protocol are described in
+%(Jack, Jr. et al., 2008a), and listed on the ADNI website
+%(http://www.loni.ucla.edu/ADNI/Research/Cores/). The T1-weighted volumetric
+%scans that were designated to be the “best” after quality control were processed
+%using the standard ADNI image processing pipeline, which included
+%post-acquisition correction of gradient warping (Jovicich et al., 2006), B1
+%non-uniformity correction (Narayana et al., 1998) depending on the scanner and
+%coil type, intensity non-uniformity correction (Sled et al., 1998) and phantom
+%based scaling correction (Gunter et al., 2006) - the geometric phantom scan
+%having been acquired with each patient scan.
+%
+%Clinical and demographic data are shown in Tables 1-4. Table 1 shows the
+%demographic data of the 15 randomly-selected subjects (5 control, 5 MCI and 5
+%AD) used for method optimisation.  Table 2 shows the demographic data of the 30
+%randomly-selected subjects (10 control, 10 MCI and 10 AD) used for method
+%validation (note that this subset of 30 subjects was separate from the subset of
+%15 subjects used for method optimisation). Table 3 shows the demographic data of
+%682 subjects (200 control, 335 MCI and 147 AD) in our full dataset. We also
+%sub-divided the MCI subjects into three subgroups (8 reverters, 204 stable and
+%123 converters) based on their follow-up clinical diagnoses determined up to 36
+%months after baseline.  Table 4 shows the demographic data of the MCI subgroups.
+%
 <<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=

### 2013-02-06 paper.Rnw
Updated methods with ADNI1 description and minctracc parameters

--- 2013-02-05 paper.Rnw
@@
   \includegraphics[width=0.5\textwidth]{figure/MAGeT-figure.png}
-  \caption{Diagram of the MAGeT Brain algorithm}
+  \caption{Diagram of the MAGeT Brain algorithm \label{fig:MAGeT-diagram}}
 \end{figure}
@@
 
-\subsubsection{ADNI-1 1.5T Screening}
-\todo{include image characteristics}
+\subsubsection{ADNI1 1.5T Screening Dataset}
+
+Clinical, demographic and pre-processed T1-weighted MRI used in this paper were
+downloaded by the authors from the ADNI1 database (adni.loni.ucla.edu) between
+March 2012 and August 2012. The image dataset download was the "ADNI1:Screening
+1.5T" standardized dataset available from ADNI
+(http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/)\citep{Wyman2012}.
+This image collection contains uniformly preprocessed images which have been
+designated to be the "best" after quality control.  All images were acquired
+using 1.5T scanners (General Electric Healthcare, Philips Medical Systems or
+Siemens Medical Solutions) at multiple sites using the protocol described in
+\citep{Jack2008a}.  Representative 1.5T imaging parameters were TR = 2400ms, TI
+= 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240 x 240mm, a
+192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel resolution of
+1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown in table
+\ref{tab:ADNI1_1.5T_Screening_Dataset_Demographics}.
+
+%Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
+%flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
+%(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
+%$mm^3$.
+
 \todo{describe SNT manual segmentations}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1 Screening Dataset Demographics Table                                  %
+%                                                                             %
 <<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 + MMSE, 
-               data = adnimerge, subset = VISCODE == 'bl',   
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+               MMSE, data = adnimerge, subset = VISCODE == 'bl',   
                method = "reverse", test=FALSE, overall=TRUE)
 
-latex(tab, file="", size="scriptsize",  caption="ADNI-1 1.5T Screening demographics")
+latex(tab, file="", size="scriptsize", 
+    caption="ADNI1 1.5T Screening dataset demographics",
+    title="tab:ADN1_1.5T_Screening_Dataset_Demographics")
 @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsubsection{SZ First Episode Patients}
+
 \todo{include image characteristics}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% First Episode SZ Demographics Table                                         %
+%                                                                             %
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 latex(tab, file="", size="scriptsize", ctable=FALSE,
-      caption="Schizophrenia First Episode Patient Demographics")
+      caption="Schizophrenia First Episode Patient Demographics",
+      title="tab:SZFEP_Demographics")
 @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsubsection{Automatic Normalization and Image Matching and Anatomical Labeling (ANIMAL)}
-The ANIMAL algorithm carries out Image registration is in two phases. In the
+\subsubsection{Automatic Normalization and Image Matching and Anatomical
+Labeling (ANIMAL)}
+
+The ANIMAL algorithm carries out image registration is in two phases. In the
 first, a 12-parameter linear transformation (3 translations, rotations, scales,
 shears) is estimated between images using an algorithm that maximizes the
-correlation between blurred MR intensities and gradient magnitude over the
-whole brain \citep{Collins}. In the second phase, nonlinear registration is completed using
-the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that estimates a 3D
-deformation field between two MR images. At first, large deformations are
-estimated using blurred version of the input data. These larger deformations
-are then input to subsequent steps where the fit is refined by estimating
-smaller deformations on data blurred with a Gaussian kernel with a smaller
-FWHM. The final transformation is a set of local translations defined on a bed
-of equally spaced nodes that were estimated through the optimization of the
-correlation coefficient.
+correlation between blurred MR intensities and gradient magnitude over the whole
+brain \citep{Collins}. In the second phase, nonlinear registration is completed
+using the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that
+estimates a 3D deformation field between two MR images. At first, large
+deformations are estimated using blurred version of the input data. These larger
+deformations are then input to subsequent steps where the fit is refined by
+estimating smaller deformations on data blurred with a Gaussian kernel with a
+smaller FWHM. The final transformation is a set of local translations defined on
+a bed of equally spaced nodes that were estimated through the optimization of
+the correlation coefficient.  For the purposes of this work we used the
+regularization parameters optimized in Robbins et al. \citep{Robbins2004},
+displayed in table \ref{tab:ANIMAL_params}.
 
-For the purposes of this work we used the regularization parameters optimized 
-in Robbins et al. \citep{Robbins2004}.
-
-\todo{link to MNI website here}
-\todo{show command line used}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ANIMAL parameters                                                           %
+%                                                                             %
+\begin{table*}[!tbp]
+\scriptsize
+\caption{ANIMAL Registration Parameters\label{tab:ANIMAL_params}} 
+\begin{center}
+\begin{tabular}{l | c c c}
+\hline 
+Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
+\hline 
+Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
+Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
+Iterations        &   30    &    30   &   10     \tabularnewline
+Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
+Sub-Lattice       &   6     &    6    &   6      \tabularnewline
+Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 {\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
-are shown in Table X
-\todo{FIXME: table reference}.
+are shown in Table \ref{tab:ADNI1_xvalidation_demographics}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1 X-Validation Dataset Demographics Table
+%
 <<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
@@
                  "382","431","1202")
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 + MMSE, 
-               data = adnimerge, subset = VISCODE == 'bl' & RID %in% adni_69_RIDs, 
-               method = "reverse", test=FALSE, overall=TRUE)
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
+               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
 
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption="ADNI-1 cross-validation subset demographics")
+      caption="ADNI-1 cross-validation subset demographics", 
+      title="tab:ADNI1_xvalidation_demographics")
 @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2013-02-06 paper.Rnw
Updates to the method section

@@
-\SweaveOpts{echo = FALSE, message = FALSE, warning = FALSE, fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE}
-
-<<loadpackages, results="hide", echo=FALSE, cache=FALSE>>=
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Sweave/Knitr init
+<<setup, include=FALSE, cache=FALSE>>=
+opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
+    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
 library(ggplot2)
@@
 @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \usepackage[margin=1in]{geometry}
-\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable }
-\usepackage[square,numbers]{natbib}
+\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
+\usepackage[round,authoryear]{natbib}
 
@@
 \usepackage[displaymath, tightpage]{preview}
-
 \setlength{\marginparwidth}{1.2in}
 
+%notes and TODO formatting
 %\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
@@
 
-\title{Hippocampal segmentation with MAGeT}
-\author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., Voineskos
-A., Chakravarty M.M., and the Alzheimer's Disease Neuroimaging Initiative}
+\title{Bootstrapping Multi-atlas Hippocampal Segmentation with MAGeT}
+\author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., \\ 
+Voineskos A., Chakravarty M.M., and \\ 
+the Alzheimer's Disease Neuroimaging Initiative}
 \maketitle
@@
 
-The segmentation approach we propose is best understood as an extension of MAGeT
-brain adds a preliminary stage to this process in which the template library is
-constructed rather than given as input.  As before, MAGeT brain accepts an atlas
-library and unlabelled MR subject images as input.  Template library images are
-selected from subject images.  The choice of subjects used in the template
-library can be made to reflect the neuroanatomy or demographics of the subject
-set as a whole (for instance, by sampling equally from cases and controls).
+The segmentation approach we propose is best understood as an extension of
+multi-atlas segmentation.  MAGeT brain adds a preliminary stage to this process
+in which the template library is constructed rather than given as input.  As
+before, MAGeT brain accepts an atlas library and unlabelled MR subject images as
+input.  Template library images are selected from subject images.  The choice of
+subjects used in the template library can be made to reflect the neuroanatomy or
+demographics of the subject set as a whole (for instance, by sampling equally
+from cases and controls).
 
@@
 produce segmentations for the entire set of unlabelled images (including those
-images used in the template library). Figure X describes the MAGeT brain
-algorithm in pseudocode.
-
-\footnote{Source code can be found at
-\url{http://github.com/pipitone/MAGeTbrain}.}
+images used in the template library). Figure \ref{alg:MAGeT} describes the MAGeT
+brain algorithm in pseudocode.  Source code can be found at
+\url{http://github.com/pipitone/MAGeTbrain}.
 
@@
   \caption{Pseudocode for the MAGeT Brain algorithm}
-  \label{pseudocodesdf}
+  \label{alg:MAGeT}
   \begin{algorithmic}
@@
 
-%  Describe subject datasets in general 
 \subsection{Subjects}
@@
 \subsubsection{ADNI1 1.5T Screening Dataset}
-
 Clinical, demographic and pre-processed T1-weighted MRI used in this paper were
@@
 1.5T" standardized dataset available from ADNI
-(http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/)\citep{Wyman2012}.
-This image collection contains uniformly preprocessed images which have been
-designated to be the "best" after quality control.  All images were acquired
-using 1.5T scanners (General Electric Healthcare, Philips Medical Systems or
-Siemens Medical Solutions) at multiple sites using the protocol described in
-\citep{Jack2008a}.  Representative 1.5T imaging parameters were TR = 2400ms, TI
-= 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240 x 240mm, a
-192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel resolution of
-1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown in table
-\ref{tab:ADNI1_1.5T_Screening_Dataset_Demographics}.
+\footnote{
+ \url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly preprocessed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
+= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
+x 240mm, a 192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel
+resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
+in table \ref{tab:ADNI1_1.5T_Screening_Dataset_Demographics}.
+
+For a subset of ADNI1 images, labels of the left and right hippocampi are
+available (herein refered to as SNT labels).  Semi-automated hippocampal
+volumetry was carried out using a commercially available high dimensional brain
+mapping tool (Medtronic Surgical Navigation Technologies, Louisville, CO), that
+has previously been validated and compared to manual tracing of the hippocampus
+\citep{Hsu2002}. Measurement of hippocampal volume is achieved first by placing
+manually 22 control points as local landmarks for the hippocampus on the
+individual brain MRI data: one landmark at the hippocampal head, one at the
+tail, and four per image (i.e., at the superior, inferior, medial and lateral
+boundaries) on five equally spaced images perpendicular to the long axis of the
+hippocampus. Second, fluid image transformation is used to match the individual
+brains to a template brain \citep{Christensen1997}. The pixels corresponding to
+the hippocampus are then labeled and counted to obtain volumes. This method of
+hippocampal voluming has a documented reliability of an intraclass coefficient
+better than .94 \citep{Hsu2002}.
 
@@
 
-\todo{describe SNT manual segmentations}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %                                                                             %
-<<ADNI-1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
+<<ADNI1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
+# RID < 2000 ensures only ADNI1 patients
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, subset = VISCODE == 'bl',   
+               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID < 2000,   
                method = "reverse", test=FALSE, overall=TRUE)
-
 latex(tab, file="", size="scriptsize", 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\todo{Check demographics table only includes 1.5T from the Screening}
 
@@
 
-Before images were registered, the N3 algorithm [16] is first used to minimize
-the intensity nonuniformity in each of the atlases and unlabeled subject
-images.
+Before images were registered, the N3 algorithm \citep{Sled1998} is first used
+to minimize the intensity nonuniformity in each of the atlases and unlabeled
+subject images.
 
@@
 ANTs is a diffeomorphic registration algorithm which provides great flexibility
-over the choice of transformation model, objective function, and the
-consistency of the final transformation. The transformation is estimated in a
-hierarchical fashion where the MRI data is subsampled, allowing large
-deformations to be estimated and successively refined at later hierarchical
-stages (where the data is subsampled to a finer grid). The deformation field
-and the objective function are regularized with a Gaussian kernel at each level
-of the hierarchy. The ANTs algorithm is freely available
-\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the ANTS
-algorithm compatible with the MINC data format, mincANTS
-\url{https://github.com/vfonov/mincANTS}.
+over the choice of transformation model, objective function, and the consistency
+of the final transformation. The transformation is estimated in a hierarchical
+fashion where the MRI data is subsampled, allowing large deformations to be
+estimated and successively refined at later hierarchical stages (where the data
+is subsampled to a finer grid). The deformation field and the objective function
+are regularized with a Gaussian kernel at each level of the hierarchy. The ANTs
+algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an
+implementation of the ANTS algorithm compatible with the MINC data format,
+mincANTS \url{https://github.com/vfonov/mincANTS}.
 
@@
 These settings were adapted from the "reasonable starting point" given in the
-ANTS manual. \todo{include a reference to manual?}
+ANTS manual 
+\footnote{\url{https://sourceforge.net/projects/advants/files/Documentation/}}.
 
@@
 
-Label fusion is term given to the process of combining the information from
+Label fusion is a term given to the process of combining the information from
 several candidate labellings for an MR image into a single labelling.  In this
@@
 registration, over a region of interest (ROI) generously encompassing the 
-hippocampus.  Only the top ranked
-template library image labels are used in a voxel-wise majority vote. The ROI
-is heuristically defined as the extent of all atlas labels after linear
-registration to the template, dilated by three voxels \citep{Chakravarty2012}.  
-The number of top ranked template library image labels is a configurable parameter.
+hippocampus.  Only the top ranked template library image labels are used in a
+voxel-wise majority vote. The ROI is heuristically defined as the extent of all
+atlas labels after linear registration to the template, dilated by three voxels
+\citep{Chakravarty2012}.  The number of top ranked template library image labels
+is a configurable parameter.
 
-The {\tt xcorr\_vol utility} from the ANIMAL toolkit is used to calculate the 
+The {\tt xcorr\_vol} utility from the ANIMAL toolkit is used to calculate the
 cross-correlation similarity measure.
@@
 
-This method is similar to cross-correlation weighted voting 
-except that image similarity is calculated by the normalised 
-mutual information score over the region of interest. 
-\todo{is there a reference for NMI voting?}
+This method is similar to cross-correlation weighted voting except that image
+similarity is calculated by the normalised mutual information score over the
+region of interest\citep{Studholme2001}.  The {\tt itk\_similarity} utility from
+the EZMinc toolkit\footnote{\url{https://github.com/vfonov/EZminc}} is used to
+calculate the normalised mutual information measure between to images.
 
-The {\tt itk\_similarity} utility from the EZMinc toolkit is used to calculate 
-the normalised mutual information measure.
+%%%%%                        Overlap Measure                               %%%%%
 
-%%%%%                        Goodness-of-fit                               %%%%%
+\subsection{Overlap Measure}
 
-\subsection{Goodness-of-fit}
-
-Two segmentation can be compared using the Dice Kappa ($\kappa$) overlap metric:
+The agreement between two segmentations can assessed using the Dice Similarity
+Coefficient (DSC):
 
 \begin{equation*} 
-\kappa=\frac{2a}{2a+b+c}
+DSC=\frac{2|A \cap B|}{|A| + |B|}
 \end{equation*}
 
-where $a$ is the number of voxels common to both segmentations and $b+c$ is the
-sum of the voxels uniquely identified by either segmentation.  Using this metric
-we can compare an automatically generated segmentation to a gold-standard
-segmentation.
+where $A$ and $B$ are the regions being compared, and the cardinality is the
+volume measured in voxels. Using this measure we can compare an automatically
+generated segmentation to a gold-standard segmentation.
 
@@
 
-\todo{In ANIMAL, images aligned to TAL space}
-
-{\bf Label fusion.} Majority vote, Cross-correlation weighted majority vote, and
+{\bf Label fusion.} Majority vote, cross-correlation weighted majority vote, and
 Normalized Mutual Information weighted majority vote are used.  With the
@@
 \subsubsection{ADNI-1 Screening Validation}
-
+\todo{comparison across field strengths, a la Heckemann2011?}
 To test the accuracy of MAGeT brain on a real-world task we segment the entire
@@
 
+#TODO: include Scaled_2? 
 subjects = subset(yr1, VISCODE == 'bl' & Scaled_2 == FALSE)
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{references}

### 2013-02-07 paper.Rnw

@@
 Compounding each of these problems is the significant neuroanatomical
-variability in the hippocampus throughout the course of aging, development, and
-neuropsychiatric disorders \citep{Mouiha2011b,Sabuncu2011}.  The result is that
+variability in the hippocampus throughout the course of aging, maturation, and
+neuropsychiatric disorders
+\citep{Mouiha2011b,Sabuncu2011,Giedd1998,Gogtay2006,Narr2002}.  The result is that
 existing hippocampal atlases available to a researcher may not accurately
@@
 between volume difference of those regions and recollection memory performance.
-Making modifications to an set of atlas library segmentations may require the
-same level of expertise and manual effort needed to create the segmentations
-from scratch.
+Making such modifications to a set of MRI data segmentations requires additional
+manual effort.
 
@@
 the MAPER algorithm \citep{Heckemann2006,Heckemann2011}. The basic strategy is as
-follows. Each atlas is warped to fit a subject's neuroanatomy using nonlinear
-registration techniques \citep{Collins1995,Klein2009}. Atlas labels are then
-transformed by this warping to the subject space and a {\it label fusion}
-technique, such as voxel-wise voting, is used to merge the label definitions
-from each atlas into a definitive segmentation. Typically, an atlas selection
-approach is used to improve accuracy by only using labels from atlases that are
-most similar to a subject image (cross-correlation of image intensities is a
-common similarity metric) \citep{Aljabar2009}.
+follows. Each atlas image is fit a subject's neuroanatomy using nonlinear
+registration techniques \citep{Collins1995,Klein2009}. This resulting nonlinear
+transformation is then applied to the atlas labels to bring them into the
+subject space.  Finally, a {\it label fusion} technique, such as voxel-wise
+voting, is used to merge these labels into a definitive segmentation for the
+subject. Typically, an atlas selection approach is used to improve accuracy by
+only using labels from atlases that are most similar to a subject image
+(cross-correlation of image intensities is a common similarity metric)
+\citep{Aljabar2009}.
 
@@
 The insight of generating a template library is not new.  Heckemann et al.
-compared "indirect" segmentation -- taking a single atlas and propagating the
+compared ``indirect'' segmentation -- taking a single atlas and propagating the
 labels to intermediate subjects before fusing them in a target image space -- to
@@
 incorporating the labelled images into the atlas library for future iterations.
-The novelty explored in our current work is demonstrate the viability of
+The novelty explored in our current work is to demonstrate the viability of
 achieving comparable segmentation accuracy using the basic multi-atlas schema

### 2013-02-11 paper.Rnw
Some of the methods rewritten for clarity

@@
 
-Fully-automated segmentation techniques require no human expertise beyond that
-of initial setup or training. One broad class of automated approach is that of
-the {\em model-based} approaches which employ models of anatomical feature
-variation to constrain segmentations. In this paper we focus on a subclass known
-as {\em multi-atlas} approaches which make use of a series of expertly segmented
-MRIs (the atlas library) as models to guide the segmentation of a query
-subject's MRI, and achieve some of the best automated hippocampal segmentation
-accuracies to-date.
+Automated segmentation techniques do not require human intervention but do
+require {\it a priori} anatomical information to guide segmentations.  In this
+paper we focus on methods that use manually segmented atlases as anatomical
+priors, as these methods achieve some of the best automated hippocampal
+segmentation accuracies to-date. This technique was pioneered by []\citep{} with
+single-atlas (or model-based) segmentation.  Volumetric image registration is
+used to estimate a fit between the neuroanatomy of an atlas and subject images.
+Labelling of the subject image is achieved by applying the resulting
+transformation to the atlas labels to bring them into the subject image space
+(\{em label propagation}). This method is limited in accuracy by the
+introduction of {methodological} errors in registration and partial volume
+effects in label resampling, and {conceptual} errors introduced when the anatomy
+of the atlas is not representative of the subject.
 
-The multi-atlas segmentation technique was pioneered by Heckemann et al. with
-the MAPER algorithm \citep{Heckemann2006,Heckemann2011}. The basic strategy is as
-follows. Each atlas image is fit a subject's neuroanatomy using nonlinear
-registration techniques \citep{Collins1995,Klein2009}. This resulting nonlinear
-transformation is then applied to the atlas labels to bring them into the
-subject space.  Finally, a {\it label fusion} technique, such as voxel-wise
-voting, is used to merge these labels into a definitive segmentation for the
-subject. Typically, an atlas selection approach is used to improve accuracy by
-only using labels from atlases that are most similar to a subject image
-(cross-correlation of image intensities is a common similarity metric)
-\citep{Aljabar2009}.
+Multi-atlas segmentation techniques address these limitations by combining
+segmentation information from a series of expertly segmentated atlases
+\citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2012,Aljabar2009,Leung2010,Wolz2010}.
+Each atlas image is registered to a subject image, and label propagation is
+performed to produce several labellings of the subject image (one from each
+atlas). A {\it label fusion} technique, such as voxel-wise voting, is used merge
+these labels into a definitive segmentation for the subject.
 
-% - Warfield2004: STAPLE as an alternative decision fusion technique?
+{\em Atlas selection} techniques are often used to filter out atlases dissimilar
+to a subject image in order to reduce {conceptual} errors \citep{Aljabar2009}.
 
-Multi-atlas techniques have been applied very successfully to hippocampal
-segmentation. Collins et al. found that with an atlas library of 80 atlas
-images, the use of the ANIMAL nonlinear registration algorithm, the normalised
-mutual information as similarity metric for atlas selection, and majority vote
-for decision fusion they were able to achieve a mean Dice Kappa similarity score
-(DSC) to manual labels of 0.886\citep{Collins2010}. The Alzheimer's Disease
-Neuroimaging Initiative (ADNI) is a commonly used benchmarking dataset of MR
-images of controls and patients with MCI or Alzheimer's (see Methods for more
-information on the ADNI dataset).  Leung et al. tuned a multi-atlas  approach to
-the segmentation of ADNI1 1 year dataset of images and using 55 atlases and the
-STAPLE decision fusion method \citep{Warfield2004} achieved an mean DSC of 0.89
-to the manual segmentations supplied with the ADNI data  \citep{Leung2010}. The
-MAPER whole brain segmentation algorithm (described above), using 30 atlases,
-achieved a mean DSC score of 0.889 on all ADNI1 baseline images
-\citep{Heckemann2011}. In \citep{Lotjonen2012}, the authors use a proprietary
-non-linear registration method based on intensity differences, and
-post-processing step taking into account these intensity differences to tidy up
+Multi-atlas methods have been very successfully applied to hippocampal
+segmentation. Collins et al. found near-manual segmentation performance using an
+atlas library of 80 T1-weighted atlas images from the ICBM152 dataset, the
+ANIMAL nonlinear registration algorithm, normalised mutual information as a
+similarity metric for atlas selection, and majority vote for label fusion
+\citep{Collins2010}.
+
+The Alzheimer's Disease Neuroimaging Initiative (ADNI) is a
+commonly used benchmarking dataset of MR images of controls and patients with
+MCI or Alzheimer's (see Methods for more information on the ADNI dataset).
+
+Leung et al. tuned parameters for registration and label fusion to the
+segmentation of ADNI1 1 year dataset of images with an atlas library of 55
+images \citep{Leung2010}.
+
+The MAPER whole brain segmentation algorithm \citep{Heckemann2006,Heckemann2011},
+using 30 atlases, on all ADNI1 baseline images \citep{Heckemann2011}.
+
+In \citep{Lotjonen2012}, the authors use a proprietary non-linear registration
+method based on intensity differences, and post-processing step taking into
+account these intensity differences to 
+- TODO: explain "tidy up"
+tidy up
 fused segmentations. Using images from the ADNI1 baseline dataset, they achieved
-a mean DSC of 0.885 with 30 atlases. With the LEAP algorithm, Wolz et al.
-explored an elegant modification to the basic strategy: the atlas library is
-grown, beginning with a set of manually labelled atlases and successively
-incorporating unlabelled subject images after being labelling using multi-atlas
-techniques\citep{Wolz2010}.  The sequence in which subject images are labelled
-is chosen so that the similarity between the atlas images and the target images
-is minimised at each step, effectively allowing for deformations between very
-dissimilar images to be broken up into sequences of smaller deformations.  With
-an atlas library of 30 MR images, LEAP was used to segment the ADNI1 baseline
-dataset, achieving a mean Dice score of 0.85 with manual segmentations.
+a mean DSC of 0.885 with 30 atlases.
+
+The LEAP algorithm, Wolz et al.  is an elegant modification to the basic
+multi-atlas strategy: the atlas library is grown, beginning with a set of
+manually labelled atlases and successively incorporating unlabelled subject
+images after being labelling using multi-atlas techniques\citep{Wolz2010}.  The
+sequence in which subject images are labelled is chosen so that the similarity
+between the atlas images and the target images is minimised at each step,
+effectively allowing for deformations between very dissimilar images to be
+broken up into sequences of smaller deformations.  With an atlas library of 30
+MR images, LEAP was used to segment the ADNI1 baseline dataset, achieving a mean
+Dice score of 0.85 with manual segmentations.
+
+- TODO: "probablistic methods?"
 
@@
 \subsection{MAGeT Brain Algorithm}
-\todo{Make sure we explain what MAGeT stands for}
-\todo{reference Chakravarty2012}
-
-In this paper, we use the term {\it atlas} to mean any manually segmented MR
-image, and the term {\it atlas library} to mean a set of such images.  We use
-the term {\it template} to refer to any MR image, and associated labelling,
-used to segment another image, and the term {\it template library} to refer to
-a set of such images.  An atlas library may be used as a template library but,
-as we will discuss, a template library may also be composed of images with
-computer generated labellings.
+We will use the term {\it atlas} to refer to two image volumes: an intensity
+image ({\it atlas image}) and a corresponding manual segmentation image ({\it
+atlas labels}). A {\it subject} image is an intensity image for which we would
+like an segmentation.  The term {\it template} will be used to refer more
+generically to any image and corresponding labelling, manual or computed, when
+it is {\it used as} as a model in the segmentation of another image. In other
+publications, the terms atlas and template are used interchangeably, but in this
+paper we force the distinction to make it clear when an image has a manual
+segmentation.  The terms {\it atlas library} and {\it template library} mean a
+set of such images.
 
 The simplest form of multi-atlas segmentation combines labellings derived from
-several atlases by way of label fusion \citep{Heckemann2006}.  We will refer to
-this as {\em basic multi-atlas segmentation}.  The primary steps are as
-follow are: (1) an atlas library and unlabelled MR images are given as input,
-(2) each atlas image is nonlinearly registered to each unlabelled image, (4)
-each atlas' labels are propagated via the resulting transformations to the
-unlabelled image space, and (5) the resulting labels are fused to produce a
-single, definitive segmentation. The particular registration and voting method
-used are left unspecified.
+several atlases by way of label fusion \citep{Heckemann2006,Heckemann2011}. We
+will refer to this as {\em basic multi-atlas segmentation}, and specific the
+basic schema as:
 
-The segmentation approach we propose is best understood as an extension of
-multi-atlas segmentation.  MAGeT brain adds a preliminary stage to this process
-in which the template library is constructed rather than given as input.  As
-before, MAGeT brain accepts an atlas library and unlabelled MR subject images as
-input.  Template library images are selected from subject images.  The choice of
-subjects used in the template library can be made to reflect the neuroanatomy or
-demographics of the subject set as a whole (for instance, by sampling equally
-from cases and controls).
+    1. an atlas library and set of subject images are given as
+       input, and the atlas library is used as a template library in the
+       following steps; 
+    2. each atlas intensity image is nonlinearly registered to each subject
+       intensity image; 
+    3. label images from each atlas are propagated via the resulting
+       transformations to the subject image space; and 
+    4. the resulting labels are fused to produce a single, definitive
+       segmentation.
 
-To create the template library, labels from each atlas image are propagated to
+The particular registration and voting method used are left unspecified.
+
+MAGeT brain is best understood as an extension of
+the basic multi-atlas segmentation schema. An additional stage is included in
+which the template library is constructed by labelling a subset of subject
+images, rather than by using the atlas library. The choice of subjects used in
+the template library can be made to reflect the neuroanatomy or demographics of
+the subject set as a whole (for instance, by sampling equally from cases and
+controls).
+
+Labels from each atlas image are propagated to
 each template library image via the transformation resulting from a non-linear

### 2013-02-13 paper.Rnw
Updated intro and methods

@@
 require {\it a priori} anatomical information to guide segmentations.  In this
-paper we focus on methods that use manually segmented atlases as anatomical
+paper we focus on methods that use manually segmented MRI atlases as anatomical
 priors, as these methods achieve some of the best automated hippocampal
-segmentation accuracies to-date. This technique was pioneered by []\citep{} with
-single-atlas (or model-based) segmentation.  Volumetric image registration is
-used to estimate a fit between the neuroanatomy of an atlas and subject images.
-Labelling of the subject image is achieved by applying the resulting
-transformation to the atlas labels to bring them into the subject image space
+segmentation accuracies to-date. This technique was first developed using a
+single atlas prior (known as single-atlas, or model-based,
+segmentation)\citep{Haller1997,Csernansky1998}.  Volumetric image registration
+is used to estimate a fit between the neuroanatomy of an atlas and target
+images.  Labelling of the target image is achieved by applying the resulting
+transformation to the atlas labels to bring them into the target image space
 (\{em label propagation}). This method is limited in accuracy by the
-introduction of {methodological} errors in registration and partial volume
-effects in label resampling, and {conceptual} errors introduced when the anatomy
-of the atlas is not representative of the subject.
+introduction of estimation errors in registration and partial volume effects in
+label resampling, and errors introduced when the anatomy of the atlas is
+unrepresentative of the target anatomy.
 
 Multi-atlas segmentation techniques address these limitations by combining
-segmentation information from a series of expertly segmentated atlases
+segmentation information from a series of expertly segmented atlases
 \citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2012,Aljabar2009,Leung2010,Wolz2010}.
-Each atlas image is registered to a subject image, and label propagation is
-performed to produce several labellings of the subject image (one from each
+Each atlas image is registered to a target image, and label propagation is
+performed to produce several labellings of the target image (one from each
 atlas). A {\it label fusion} technique, such as voxel-wise voting, is used merge
-these labels into a definitive segmentation for the subject.
-
-{\em Atlas selection} techniques are often used to filter out atlases dissimilar
-to a subject image in order to reduce {conceptual} errors \citep{Aljabar2009}.
+these labels into a definitive segmentation for the target.
 
@@
 In \citep{Lotjonen2012}, the authors use a proprietary non-linear registration
-method based on intensity differences, and post-processing step taking into
-account these intensity differences to 
-- TODO: explain "tidy up"
-tidy up
-fused segmentations. Using images from the ADNI1 baseline dataset, they achieved
-a mean DSC of 0.885 with 30 atlases.
+method based on intensity differences, and post-processing step using a graph
+cuts or EM approach with intensity spatial prior models to optimise the fused
+segmentations.  Using images from the ADNI1 baseline dataset and 30 atlases.
+
+{\em Atlas selection} techniques are often used to filter out atlases dissimilar
+to a target image in order to reduce {conceptual} errors \citep{Aljabar2009}.
 
@@
 multi-atlas strategy: the atlas library is grown, beginning with a set of
-manually labelled atlases and successively incorporating unlabelled subject
+manually labelled atlases and successively incorporating unlabelled target
 images after being labelling using multi-atlas techniques\citep{Wolz2010}.  The
-sequence in which subject images are labelled is chosen so that the similarity
+sequence in which target images are labelled is chosen so that the similarity
 between the atlas images and the target images is minimised at each step,
@@
 
-- TODO: "probablistic methods?"
+%- TODO: "probablistic methods?"
 
@@
 resulting segmentations. As noted, by choosing atlases ranked most similar to a
-subject image by voxel intensity profile segmentation accuracy is improved,
+target image by voxel intensity profile segmentation accuracy is improved,
 suggesting that neuroanatomical similarity plays a role\citep{Aljabar2009}.
@@
 atlas the important factors leading to improved accuracy are that the atlas have
-neuroanatomical features that match the subject, and that the atlas segmentation
+neuroanatomical features that match the target, and that the atlas segmentation
 use the same protocol as the gold-standard \citep{Carmichael2005}. Nestor et al.
@@
 \emph{template library} composed of images taken from the target population.
-The template library is then used to segment the subjects in a similar fashion
+The template library is then used to segment the targets in a similar fashion
 to basic multi-atlas segmentation: by label propagation and label fusion. The
 intuition driving this approach is that by generating a template library we
-leverage the unique neuroanatomy of subject population on hand to initialize the
+leverage the unique neuroanatomy of target population on hand to initialize the
 segmentation process and improve accuracy over direct propagation from the atlas
-library to unlabelled subjects while also using fewer manually segmented
+library to unlabelled targets while also using fewer manually segmented
 atlases.
@@
 compared ``indirect'' segmentation -- taking a single atlas and propagating the
-labels to intermediate subjects before fusing them in a target image space -- to
+labels to intermediate targets before fusing them in a target image space -- to
 multi-atlas segmentation and found that the indirect approach performed worse
@@
 histologically-derived atlas. The contribution of the present work is to extend
-our approach to the human hippocampus and perform a rigorous validation. We
-first conduct an intensive cross-validation of MAGeT on a pool of 69 MRIs from
-the ADNI database, comparing the performance of MAGeT against basic multi-atlas
-segmentation using a range of atlas and template library sizes, registration and
-fusion methods.  We then compare MAGeT segmentations of the entire ADNI
-screening dataset with those of established automated and manual segmentation
-techniques.  Finally, we compare MAGeT segmentations to those of manual
-segmentations from a dataset of first episode schizophrenic patient MRIs.
+our approach to the human hippocampus and perform a series of experiments to
+rigorous validate the method. First, we conduct an extensive cross-validation of
+MAGeT and basic multi-atlas segmentation on a subset of the ADNI1 dataset to
+assess the accuracy of MAGeT brain under various parameter settings (number of
+atlases and templates, registration and fusion methods). With the best
+performing parameter configuration discovered above, we estimate MAGeT
+intra-rater reliability by segmenting separately acquired T1 images of the
+atlas subjects. For this experiment, we use the Winterburn atlases: digital
+hippocampal subfield segmentations of five in-vivo high-resolution (300u
+isotropic) T1-weighted MR scans\citep{Winterburn2013}. To validate MAGeT in a
+real world situation, we segment the entire ADNI1 screening dataset and compare
+our segmentations to established automated and manual segmentations.
+Additionally, to ensure MAGeT brain accuracy across disease categories, we also
+compare MAGeT segmentations to manual segmentations of 139 first episode
+schizophrenic patients.
 
@@
 \subsection{MAGeT Brain Algorithm}
-We will use the term {\it atlas} to refer to two image volumes: an intensity
-image ({\it atlas image}) and a corresponding manual segmentation image ({\it
-atlas labels}). A {\it subject} image is an intensity image for which we would
-like an segmentation.  The term {\it template} will be used to refer more
-generically to any image and corresponding labelling, manual or computed, when
-it is {\it used as} as a model in the segmentation of another image. In other
-publications, the terms atlas and template are used interchangeably, but in this
-paper we force the distinction to make it clear when an image has a manual
-segmentation.  The terms {\it atlas library} and {\it template library} mean a
-set of such images.
+In this paper we will be making a distinction between an atlas and a template --
+typically these terms are used roughly interchangeably. The term {\it atlas} is
+taken to refer to two image volumes: an intensity image ({\it atlas image}) and
+a corresponding manual segmentation image ({\it atlas labels}). {\it Template}
+refers more generically to any image and corresponding labelling, manual or
+computed, when it is {\it used as} as a model in the segmentation of another
+image.  The terms {\it atlas library} and {\it template library} mean a set of
+such images. Additionally, we will use the terms {\it target} to refer to an
+intensity image for which we would like an segmentation.
 
@@
 several atlases by way of label fusion \citep{Heckemann2006,Heckemann2011}. We
-will refer to this as {\em basic multi-atlas segmentation}, and specific the
-basic schema as:
-
-    1. an atlas library and set of subject images are given as
-       input, and the atlas library is used as a template library in the
-       following steps; 
-    2. each atlas intensity image is nonlinearly registered to each subject
+will refer to this as {\em basic multi-atlas segmentation}. The basic schema as
+for this method is as follows: 
+\begin{itemize}
+\item   An atlas library and set of target images are given as input. The atlas
+        library is used as a template library in the following steps; 
+\item   Each atlas intensity image is nonlinearly registered to each target 
         intensity image; 
-    3. label images from each atlas are propagated via the resulting
-       transformations to the subject image space; and 
-    4. the resulting labels are fused to produce a single, definitive
+\item   label images from each atlas are propagated via the resulting
+        transformations to the target image space; and 
+\item   the resulting labels are fused to produce a single, definitive
         segmentation. 
-
+\end{itemize}       
 The particular registration and voting method used are left unspecified.
 
-MAGeT brain is best understood as an extension of
-the basic multi-atlas segmentation schema. An additional stage is included in
-which the template library is constructed by labelling a subset of subject
-images, rather than by using the atlas library. The choice of subjects used in
-the template library can be made to reflect the neuroanatomy or demographics of
-the subject set as a whole (for instance, by sampling equally from cases and
-controls).
+MAGeT brain is best understood as an extension of the basic multi-atlas
+segmentation schema. Additional stages construct the template library by
+labelling a subset of target images, rather than by using the atlas library.
+The choice of targets used in the template library can be made to reflect the
+neuroanatomy or demographics of the target set as a whole (for instance, by
+sampling equally from cases and controls). A truncated version of basic
+multi-atlas segmentation is used to label the template library images; label
+fusion is not performed. Instead, each template image gets a labelling from each
+atlas image. A second round of basic multi-atlas segmentation uses the template
+library to segment the entire set of target images (including those images used
+in the template library). Label fusion in this final step fuses all labels from
+all templates.  Figure \ref{alg:MAGeT} describes the MAGeT brain algorithm in
+pseudocode.
 
-Labels from each atlas image are propagated to
-each template library image via the transformation resulting from a non-linear
-registration between pair of images. As a result, each template library image
-has a label from each atlas. Basic multi-atlas segmentation is then used to
-produce segmentations for the entire set of unlabelled images (including those
-images used in the template library). Figure \ref{alg:MAGeT} describes the MAGeT
-brain algorithm in pseudocode.  Source code can be found at
+Source code for MAGeT brain can be found at
 \url{http://github.com/pipitone/MAGeTbrain}.
@@
     \Function{BasicMultiAtlasSegmentation}{Templates, Subjects}
-      \ForAll{$subject$}
+      \ForAll{$target$}
         \ForAll{$template$}
-          \State propagate all labels for template to subject space
-          \State store subject labels
+          \State propagate all labels for template to target space
+          \State store target labels
         \EndFor
-        \State fuse subject labels
+        \State fuse target labels
       \EndFor
@@
       \For{$i = 1 \to n$}
-        \State choose a subject to be used as a template
+        \State choose a target to be used as a template
         \State propagate labels from each atlas to template space
@@
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figure/MAGeT-figure.png}
-  \caption{Diagram of the MAGeT Brain algorithm \label{fig:MAGeT-diagram}}
-\end{figure}
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=0.5\textwidth]{figure/MAGeT-figure.png}
+%  \caption{Diagram of the MAGeT Brain algorithm \label{fig:MAGeT-diagram}}
+%\end{figure}
 
@@
 
-%%%%%                       Image pre-processing                           %%%%%
-
-\subsection{Image pre-processing}
-
-Before images were registered, the N3 algorithm \citep{Sled1998} is first used
-to minimize the intensity nonuniformity in each of the atlases and unlabeled
-subject images.
-
+c
 %%%%%                       Registration Methods                           %%%%%
 
-\subsection{Registration}
+\subsection{Registration Methods}
+
+Before registration, all images undergo preprocessing with the N3 algorithm
+\citep{Sled1998} to minimize intensity nonuniformity. In our experiments we use
+one of two non-linear image registration methods.
 
@@
 Label fusion is a term given to the process of combining the information from
-several candidate labellings for an MR image into a single labelling.  In this
-paper we explore the benefits of three different fusion methods.
+several candidate labellings for an intensity image into a single labelling.  In
+this paper we explore three different fusion methods.
 
@@
 
-Labels are propagated from all template library images to a subject.  Each
+Labels are propagated from all template library images to a target.  Each
 output voxel is given the most frequent label at that voxel location amongst
@@
 
-An optimal combination of subjects from the template library has previously been
+An optimal combination of targets from the template library has previously been
 shown to improve segmentation accuracy \citep{Aljabar2009,Collins2010}.  In this
@@
 
-The agreement between two segmentations can assessed using the Dice Similarity
-Coefficient (DSC):
+The Dice similarity index (SI) assesses the agreement between two segmentations.
+It is one of the most widely used measures of segmentation performance, and we
+use it as the basis of comparison throughout this paper.  Additionally, we
+report other commonly used measures:
 
-\begin{equation*} 
-DSC=\frac{2|A \cap B|}{|A| + |B|}
-\end{equation*}
+\begin{itemize}
+ \item \begin{equation*}Similarity index (SI) =\frac{2|A \cap B|}{|A| + |B|}\end{equation*}
+ \item \begin{equation*}Jaccard (J) =\frac{2|A \cap B|}{|A| + |B|}\end{equation*}
 
 where $A$ and $B$ are the regions being compared, and the cardinality is the
-volume measured in voxels. Using this measure we can compare an automatically
-generated segmentation to a gold-standard segmentation.
+volume measured in voxels. .
 
@@
 To test the accuracy of MAGeT brain on a real-world task we segment the entire
-ADNI-1 dataset using an atlas set that is not representative of the subject set.
+ADNI-1 dataset using an atlas set that is not representative of the target set.
 
@@
 
+- True for all parameter settings! (so works generally..)
 \todo{Can we statistically capture what "peak" performance is?  Something like,

### 2013-02-14 paper.Rnw
Mucho edits to the methods

@@
 transformation to the atlas labels to bring them into the target image space
-(\{em label propagation}). This method is limited in accuracy by the
+({\em label propagation}). This method is limited in accuracy by the
 introduction of estimation errors in registration and partial volume effects in
@@
 several atlases by way of label fusion \citep{Heckemann2006,Heckemann2011}. We
-will refer to this as {\em basic multi-atlas segmentation}. The basic schema as
-for this method is as follows: 
-\begin{itemize}
+will refer to this as {\em basic multi-atlas segmentation}. The schema as for
+this method is as follows: 
+\begin{enumerate}
 \item   An atlas library and set of target images are given as input. The atlas
@@
         intensity image; 
-\item   label images from each atlas are propagated via the resulting
+\item   Label images from each atlas are propagated via the resulting
         transformations to the target image space; and 
@@
         segmentation. 
-\end{itemize}       
+\end{enumerate}       
 The particular registration and voting method used are left unspecified.
@@
 MAGeT brain is best understood as an extension of the basic multi-atlas
-segmentation schema. Additional stages construct the template library by
-labelling a subset of target images, rather than by using the atlas library.
-The choice of targets used in the template library can be made to reflect the
-neuroanatomy or demographics of the target set as a whole (for instance, by
-sampling equally from cases and controls). A truncated version of basic
-multi-atlas segmentation is used to label the template library images; label
-fusion is not performed. Instead, each template image gets a labelling from each
-atlas image. A second round of basic multi-atlas segmentation uses the template
-library to segment the entire set of target images (including those images used
-in the template library). Label fusion in this final step fuses all labels from
-all templates.  Figure \ref{alg:MAGeT} describes the MAGeT brain algorithm in
-pseudocode.
+segmentation schema. Instead of using the atlas library to directly label the
+target images, a subset of the input images are selected as template images and
+then labelled.  The choice of targets used in the template library can be made
+to reflect the neuroanatomy or demographics of the target set as a whole (for
+instance, by sampling equally from cases and controls). Once the template
+library images have been chosen, a truncated version of basic multi-atlas
+segmentation is used to label the template library images without performing
+label fusion. Instead, each template image receives multiple labellings: one
+from each atlas image. A second round of basic multi-atlas segmentation uses the
+template library to segment the entire set of target images (including those
+images used in the template library). Label fusion in this final step fuses all
+labels from all templates. To summarize, figure \ref{alg:MAGeT} describes the
+MAGeT brain algorithm in pseudocode.
 
@@
 
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=0.5\textwidth]{figure/MAGeT-figure.png}
-%  \caption{Diagram of the MAGeT Brain algorithm \label{fig:MAGeT-diagram}}
-%\end{figure}
-
 %%%%%                             Subjects                                 %%%%%
@@
 \subsection{Subjects}
+Our experiments use three distinct subject datasets.
 
 \subsubsection{ADNI1 1.5T Screening Dataset}
-Clinical, demographic and pre-processed T1-weighted MRI used in this paper were
-downloaded by the authors from the ADNI1 database (adni.loni.ucla.edu) between
-March 2012 and August 2012. The image dataset download was the "ADNI1:Screening
-1.5T" standardized dataset available from ADNI
-\footnote{
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset download was the "ADNI1:Screening 1.5T"
+standardized dataset available from ADNI \footnote{
 \url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
@@
 \subsubsection{SZ First Episode Patients}
+Nomenclature: SZFEP Dataset
+SZFEP dataset demographics are shown in table \ref{tab:SZFEP_Demographics}.
 
@@
 
-c
+\subsubsection{Winterburn Atlases}
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal segmentations
+of of five in-vivo 300u isotropic T1-weighted MR images. The segmentations
+include subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate
+gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
+
+In addition to the high-resolution scans distributed as part of the Winterburn
+atlases, we also obtained additional T1 BRAVO scans of four of the five
+subjects.  \todo{demographics?}\todo{explain scan parameters in more detail..}
+
 %%%%%                       Registration Methods                           %%%%%
@@
 
-Before registration, all images undergo preprocessing with the N3 algorithm
+Before registration, all images underwent preprocessing with the N3 algorithm
 \citep{Sled1998} to minimize intensity nonuniformity. In our experiments we use
@@
 
-We used the following command line when running the ANTS command, 
+We used the following command line when running ANTS:
 {\small
@@
 several candidate labellings for an intensity image into a single labelling.  In
-this paper we explore three different fusion methods.
+this paper we explore three fusion methods.
 
@@
 
-%%%%%                        Overlap Measure                               %%%%%
+%%%%%                        Evaluation Measure                           %%%%%
 
-\subsection{Overlap Measure}
+\subsection{Assessing Segmentation Agreement}
 
-The Dice similarity index (SI) assesses the agreement between two segmentations.
-It is one of the most widely used measures of segmentation performance, and we
-use it as the basis of comparison throughout this paper.  Additionally, we
-report other commonly used measures:
+The Dice similarity coefficient (DSC) assesses the agreement between two
+segmentations.  It is one of the most widely used measures of segmentation
+performance, and we use it as the basis of comparison throughout this paper.
+Additionally, we report the Jaccard index, another commonly used similarity
+measure:
 
-\begin{itemize}
- \item \begin{equation*}Similarity index (SI) =\frac{2|A \cap B|}{|A| + |B|}\end{equation*}
- \item \begin{equation*}Jaccard (J) =\frac{2|A \cap B|}{|A| + |B|}\end{equation*}
+ \[\text{Dice's coefficient (DSC)} = \frac{2|A \cap B|}{|A| + |B|}\]
+
+ \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]
 
 where $A$ and $B$ are the regions being compared, and the cardinality is the
-volume measured in voxels. .
+volume measured in voxels.
 
@@
 
-Experiments were performed to assess the performance of MAGeT brain with various 
-parameter settings as well as on diverse datasets.  In each experiment we contrast 
-the performance of MAGeT brain with that standard single- and multi-atlas 
-segmentations derived from the same atlas library.
+The following experiments were performed to assess the performance of MAGeT
+brain with various parameter settings as well as on diverse datasets.  In each
+experiment we contrast the performance of MAGeT brain with that standard single-
+and multi-atlas segmentations derived from the same atlas library.
 
@@
 %%%%%
-\subsubsection{ADNI-1 cross-validation}
+\subsubsection{ADNI1 cross-validation}
+To test the effect of parameter settings on the MAGeT brain algorithm, repeated
+random sub-sampling cross-validation (RRSCV) was performed using a subset of the
+ADNI1 subjects for validation.  This form of cross-validation (described in
+detail below) allows us to validate MAGeT with various atlas and template
+library sizes, registration and label fusion methods.
 
-To test the accuracy of the MAGeT brain algorithm with different parameter 
-settings, repeated random sub-sampling cross-validation (RRSCV) was performed
-on a subset of the ADNI-1 dataset.
-
-{\bf Dataset evaluated.} 69 1.5T images were randomly selected from the 
-{\em ADNI1:Screening 1.5T} standardized dataset.  Demographics for this subset 
-are shown in Table \ref{tab:ADNI1_xvalidation_demographics}.
+{\bf Dataset evaluated.} 69 1.5T images were arbitrarily selected from the {\em
+ADNI1:Screening 1.5T} standardized dataset. 23 subjects were chosen from each
+disease category (cognitively normal (CN), MCI and AD). Demographics for this
+subset are shown in Table \ref{tab:ADNI1_xvalidation_demographics}.
 
@@
 
-{\bf Atlas and template libary.}  Atlases consisted of images taken from the
+{\bf Atlas and template libary.}  Atlases consist of images taken from the
 dataset, with corresponding manual labels provided by SNT.  Atlas library size
-was varied from 3 to 9 images.  The remaining images were segmented, with the
-template library size varying from 3 to 20 images.  Template library images were
-selected randomly from the images to be segmented.
+varied from 3 to 9 images.  The remaining 60-66 target images were segmented
+with the template library size varying from 3 to 20 images for each setting of
+atlas library size. Template library images were selected at random from the
+targets.
 
@@
 {\bf Label fusion.} Majority vote, cross-correlation weighted majority vote, and
-Normalized Mutual Information weighted majority vote are used.  With the
+Normalized Mutual Information weighted majority vote were used. With the
 weighted majority vote fusion methods, the number of top labels used in the
@@
 {\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV)
-consists of repeated trials in which items from the dataset are randomly
-assigned to a training set or validation set. In each trial, performance on the
-validation set is measured, and then averaged across all trials.
+consists of repeated trials in which items from the dataset are randomly sampled
+and assigned to a training set or validation set. In each trial, performance on
+the validation set is measured, and then averaged across all trials. This form
+of cross-validation allows
 
-We performed RRSCV on each combination of parameters listed above: atlas library
-size, template library size, registration method, and label fusion method.  We
-performed 10 trials per parameter combination.  In each validation trial, the
+With respect to this We performed RRSCV on each combination of parameters listed above: atlas library
+size, template library size, registration method, and label fusion method.
+
+We performed 10 trials per parameter combination.  In each validation trial, the
 training set consisted of the images used as atlases, and the validation set
-consisted of the images to be segmented. The MAGeT brain algorithm and the basic
-multi-atlas segmentation procedure were applied to segment the images in the
-validation set.  Additionally, in each trial, the single-atlas segmentation was
-obtained for each atlas-template.
+consisted of the images to be segmented.
+
+The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were
+applied to segment the images in the validation set.  Additionally, in each
+trial, the single-atlas segmentation was obtained for each atlas-template.
 

### 2013-02-15 paper.Rnw
Updated methods

@@
 atlas). A {\it label fusion} technique, such as voxel-wise voting, is used merge
-these labels into a definitive segmentation for the target.
+these labels into a definitive segmentation for the target.  In addition, {\em
+atlas selection} techniques are often used to exclude atlases from label fusion
+that are dissimilar to a target image in order to reduce error from
+unrepresentative anatomy \citep{Aljabar2009}. Cross-correlation or normalised
+mutual information of image intensities are common measures of image similarity
+used in atlas selection.
 
@@
 atlas library of 80 T1-weighted atlas images from the ICBM152 dataset, the
-ANIMAL nonlinear registration algorithm, normalised mutual information as a
-similarity metric for atlas selection, and majority vote for label fusion
-\citep{Collins2010}.
-
-The Alzheimer's Disease Neuroimaging Initiative (ADNI) is a
-commonly used benchmarking dataset of MR images of controls and patients with
+ANIMAL nonlinear registration algorithm, normalised mutual information as an
+atlas selection similarity metric, and majority vote for label fusion
+\citep{Collins2010}.  The Alzheimer's Disease Neuroimaging Initiative (ADNI) is
+a commonly used benchmarking dataset of MR images of controls and patients with
 MCI or Alzheimer's (see Methods for more information on the ADNI dataset).
-
 Leung et al. tuned parameters for registration and label fusion to the
 segmentation of ADNI1 1 year dataset of images with an atlas library of 55
-images \citep{Leung2010}.
+images \citep{Leung2010}.  The MAPER whole brain segmentation algorithm
+\citep{Heckemann2006,Heckemann2011}, using 30 atlases, on all ADNI1 baseline
+images \citep{Heckemann2011}.  Lotjonen et al. use images from the ADNI1
+baseline dataset and 30 atlases with a proprietary non-linear registration
+method based on intensity differences, and post-processing step using a graph
+cuts algorithm to optimise fused segmentations against a spatial intensity prior
+\citep{Lotjonen2012}.  
+\todo{No story here}
 
-The MAPER whole brain segmentation algorithm \citep{Heckemann2006,Heckemann2011},
-using 30 atlases, on all ADNI1 baseline images \citep{Heckemann2011}.
-
-In \citep{Lotjonen2012}, the authors use a proprietary non-linear registration
-method based on intensity differences, and post-processing step using a graph
-cuts or EM approach with intensity spatial prior models to optimise the fused
-segmentations.  Using images from the ADNI1 baseline dataset and 30 atlases.
-
-{\em Atlas selection} techniques are often used to filter out atlases dissimilar
-to a target image in order to reduce {conceptual} errors \citep{Aljabar2009}.
-
-The LEAP algorithm, Wolz et al.  is an elegant modification to the basic
-multi-atlas strategy: the atlas library is grown, beginning with a set of
-manually labelled atlases and successively incorporating unlabelled target
-images after being labelling using multi-atlas techniques\citep{Wolz2010}.  The
-sequence in which target images are labelled is chosen so that the similarity
-between the atlas images and the target images is minimised at each step,
-effectively allowing for deformations between very dissimilar images to be
-broken up into sequences of smaller deformations.  With an atlas library of 30
-MR images, LEAP was used to segment the ADNI1 baseline dataset, achieving a mean
-Dice score of 0.85 with manual segmentations.
+The LEAP algorithm is an elegant modification to the basic multi-atlas strategy
+\citep{Wolz2010}. The atlas library is grown, beginning with a set of manually
+labelled atlases, and successively incorporats unlabelled target images after
+being labelling using multi-atlas techniques. The sequence in which target
+images are labelled is chosen so that the similarity between the atlas images
+and the target images is minimised at each step, effectively allowing for
+deformations between very dissimilar images to be broken up into sequences of
+smaller deformations.  With an atlas library of 30 MR images, LEAP was used to
+segment the ADNI1 baseline dataset, achieving a mean Dice score of 0.85 with
+manual segmentations.
 
@@
 of manual segmentations and is applied to fix mislabelled voxels.
+
 %  Chupin's use of shape-based priors 
@@
 resulting segmentations. As noted, by choosing atlases ranked most similar to a
-target image by voxel intensity profile segmentation accuracy is improved,
-suggesting that neuroanatomical similarity plays a role\citep{Aljabar2009}.
-Carmichael et al. explored this directly and found that when using only one
-atlas the important factors leading to improved accuracy are that the atlas have
-neuroanatomical features that match the target, and that the atlas segmentation
-use the same protocol as the gold-standard \citep{Carmichael2005}. Nestor et al.
-found that hippocampal segmentation protocols that include more dorsal
-white-matter and posterior anatomy tended to produce higher overlap and better
-accuracy at distinguishing disease classes in the ADNI1 1 year dataset
-\citep{Nestor2012}. These results suggest both atlas library neuroanatomy and
-delineation protocol play a significant role in the resulting segmentation.
+target image by voxel intensity profile, segmentation accuracy is improved,
+suggesting that neuroanatomical similarity plays a strong role
+\citep{Aljabar2009}.  Carmichael et al. explored this directly and found that
+when using only one atlas the important factors leading to improved accuracy are
+that the atlas have neuroanatomical features that match the target, and that the
+atlas segmentation use the same protocol as the gold-standard
+\citep{Carmichael2005}. Nestor et al.  found that hippocampal segmentation
+protocols that include more dorsal white-matter and posterior anatomy tended to
+produce higher overlap and better accuracy at distinguishing disease classes in
+the ADNI1 1 year dataset \citep{Nestor2012}. These results suggest both atlas
+library neuroanatomy and delineation protocol play a significant role in the
+resulting segmentation.
 
@@
 
-\subsubsection{SZ First Episode Patients}
-Nomenclature: SZFEP Dataset
-SZFEP dataset demographics are shown in table \ref{tab:SZFEP_Demographics}.
-
-\todo{include image characteristics}
+\subsubsection{Schizophrenic First Episode Patients}
+\todo{Nomenclature: SFEP Dataset}
+\todo{describe image characteristics}
+\todo{describe expert segmentations}
+SFEP dataset demographics are shown in table \ref{tab:SZFEP_Demographics}.
 
@@
 \subsubsection{ADNI1 cross-validation}
-To test the effect of parameter settings on the MAGeT brain algorithm, repeated
-random sub-sampling cross-validation (RRSCV) was performed using a subset of the
-ADNI1 subjects for validation.  This form of cross-validation (described in
+To test the effect of parameter settings on the MAGeT brain algorithm, Monte
+Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
+the ADNI1 subjects for validation.  This form of cross-validation (described in
 detail below) allows us to validate MAGeT with various atlas and template
@@
 
-{\bf Evaluation.}  Repeated random sub-sampling cross-validation (RRSCV)
-consists of repeated trials in which items from the dataset are randomly sampled
-and assigned to a training set or validation set. In each trial, performance on
-the validation set is measured, and then averaged across all trials. This form
-of cross-validation allows
+{\bf Evaluation.}  Monte Carlo Cross-Validation (MCCV), also known as repeated
+random sub-sampling cross-validation, consists of repeated trials in which items
+from the dataset are randomly sampled (without replacement) and assigned to a
+training set or validation set \cite{Shao1993}.
 
-With respect to this We performed RRSCV on each combination of parameters listed above: atlas library
-size, template library size, registration method, and label fusion method.
+In each trial, the training set consists of images and SNT labels used as
+atlases, and the validation set consists of images to be segmented along with
+the corresponding SNT labels used as gold-standard segmentations. Both MAGeT
+brain and basic multi-atlas segmentation were performed in each trial, and the
+accuracy of the resulting segmentations were measured against the SNT labels.
 
-We performed 10 trials per parameter combination.  In each validation trial, the
-training set consisted of the images used as atlases, and the validation set
-consisted of the images to be segmented.
-
-The MAGeT brain algorithm and the basic multi-atlas segmentation procedure were
-applied to segment the images in the validation set.  Additionally, in each
-trial, the single-atlas segmentation was obtained for each atlas-template.
-
-The gold-standard for the segmentation accuracy of images in the validation set
-was the SNT manual labels.
-
-\todo{ show off the number of registrations/comparisions we did}
+We performed 10 trials on each combination of parameters listed above: atlas
+library size (1-9), template library size (1-20), registration method (ANTS vs.
+ANIMAL), and label fusion method (MV, XC-WV, NMI-WV). A total of 1080 parameter
+settings were evaluated over a total of 10800 trials.
 
@@
 \subsubsection{ADNI-1 Screening Validation}
-\todo{comparison across field strengths, a la Heckemann2011?}
 To test the accuracy of MAGeT brain on a real-world task we segment the entire
@@
 
-{\bf Atlas and template libary.} The atlas library consisted of the entire 
-Winterburn atlas set.  The Winterburn atlases are digital segmentations of the
-hippocampus in five in-vivo 300u isotropic T1-weighted MR scans, and include 
-subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate gyrus, 
-subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases range in 
-age from 29-57 years (mean age of 37), and include two males and three females.
+{\bf Atlas and template libary.} The Winterburn atlases were used as the atlas
+library.  The template library consisted of 21 randomly selected images from the
+ADNI1 data dataset (7 healthy, MCI and AD subjects).
 
-The template library consisted of 21 randomly selected images from the ADNI1 
-data dataset (7 healthy, MCI and AD subjects).
+{\bf Registration method.} The ANTS non-linear registration algorithm was used,
+as it performed best in the ADNI1 cross-validation experiment (see Results).
 
-{\bf Registration method.} ANTS, as it performed best in the cross-validation
-experiment.
-
-\todo{okay to reference results whilst still in Methods?}
-
-{\bf Label fusion.} Majority vote, as it is simpliest to run and performed 
-equally well in cross-validation experiment.
+{\bf Label fusion.} Majority vote was used performed equally well in the ADNI1
+cross-validation experiment and is the easiest label fusion method to operate.
 
 {\bf Evaluation.}
+Since the hippocampal segmentation protocols differ between the ADNI labels and
+Winterburn atlases, this poses a problem for direct similarity comparisons
+between labels produced by MAGeT brain and the ADNI labels.  To evaluate the
+performance of MAGeT brain, we compare the correlation of MAGeT segmentation
+volumes with manual segmentation (SNT) volumes. Additionally, we correlate the
+hippocampal volumes of established automated segmentation methods to MAGeT
+segmentations.
 
-Since hippocampal segmentation protocols differ between the ADNI labels and
-Winterburn atlases, this poses a problem for direct similarity comparisons
-between labels produced by MAGeT brain and the ADNI labels.
+Additionally, we compared classification accuracy of subjects by diagnosis based
+on hippocampal volume using both the SNT labels and our produced labels.
 
@@
 
-To evaluate the performance of MAGeT brain, we correlate our segmentation
-volumes with manual segmentation volumes, as well as with hippocampal volumes
-of established automated segmentation methods.
-
-Additionally, we compared classification accuracy of subjects
-by diagnosis based on hippocampal volume using both the SMT labels and our
-produced labels.
-
 \todo{Include a description of validation process.. Note that we could use rms
@@
 with adni/our segmentations}
-
-\todo{Also consider self-consistency (i.e. Kappa of same subject at different
-field strengths) a la Heckemann2011}
 
@@
 
-{\bf Dataset evaluated.} To validate that MAGeT performance generalises to other
-diseases, we measure the performance using the best parameter settings previous
-found, on a dataset consiting of first episode schizophrenia patients.
-\todo{Refer to the FEP description above... or just include it here}.
+{\bf Dataset evaluated.} To validate that MAGeT performance generalises to
+across disease categories, we measure the performance of MAGeT on the
+Schizophrenic First Episode Patient (SFEP) dataset.
 
@@
 hippocampal segmentation of patients, and Winterburn atlas set.  
+\todo{complete}
 
 {\bf Registration method.} ANTS.
+\todo{complete}
 
 {\bf Label fusion.} Majority vote.
+\todo{complete}
 
@@
 and the Winterburn-atlas segmentations by correlating volumes. 
+\todo{complete}
 
@@
 the Winterburn atlases.
+\todo{complete}
 
@@
 Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.
+\todo{complete}
 
 {\bf Registration method.} ANTS
+\todo{complete}
 
 {\bf Label fusion.} Majority vote.
+\todo{complete}
 
@@
 Segmentation accuracy is judged by difference in hippocampal volume.
+\todo{complete}
 

### 2013-02-26 paper.Rnw
Added SFEP dataset method details

@@
 \subsubsection{Schizophrenic First Episode Patients}
-\todo{Nomenclature: SFEP Dataset}
-\todo{describe image characteristics}
-\todo{describe expert segmentations}
-SFEP dataset demographics are shown in table \ref{tab:SZFEP_Demographics}.
+All patients were recruited and treated through the Prevention and Early
+Intervention Program for Psychoses (PEPP-Montreal), a specialized early
+intervention service at the Douglas Mental Health University Institute in
+Montreal, Canada. People aged 15–30 years from the local catchment area
+suffering from either affective or non-affective psychosis who had not taken
+antipsychotic medication for more than one month with an IQ above 70 were
+consecutively admitted as either in- or out-patients. Of those treated at PEPP,
+only patients aged 18 to 30 years with no previous history of neurological
+disease or head trauma causing loss of consciousness were eligible for the
+neuroimaging study; only those suffering from schizophrenia spectrum disorders
+were considered for this analysis.  For complete program details see Malla et
+al. \citep{Malla2003}.
 
+Scanning was carried out at the Montreal Neurological Institute on a 1.5-T
+Siemens whole body MRI system.  Structural T1 volumes were acquired for each
+participant using a three-dimensional (3D) gradient echo pulse sequence with
+sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
+angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular field-of-view
+for the images was 256mm (SI)×204mm (AP).  Subject demographics are shown in
+table \ref{tab:SZFEP_Demographics}.
+
+The hippocampus were traced following a validated protocol developed by Dr Jens
+Pruessner (Pruessner et al., 2000). A recent update to this protocol by Dr J
+Pruessner in 2006 allows to accurately and consistently subdivide the
+hippocampus into three different subregions: head, body, and tail.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\todo{Include a description of validation process.. Note that we could use rms
-{\tt validate} or t-test: contrast with QDA or LDA (Coupe 2011) used in LOOCV,
-with adni/our segmentations}
-
 \todo{Also: group differences, expecting separation}
@@
 \begin{figure}
-<<ADNI1_cross_validation_means, cache=TRUE, fig=TRUE>>=
+<<ADNI1_cross_validation_means, cache=TRUE>>=
 t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
@@
 \begin{figure}
-<<ADN1-cross-validation-MA-difference, cache=TRUE, fig=TRUE>>=
+<<ADN1-cross-validation-MA-difference, cache=TRUE>>=
 ggplot(all_data, aes(x=num_templates, y=k_minus_ma, colour=as.factor(num_atlases))) + 

### 2013-02-27 paper.Rnw
Added figures for a2a section

@@
 
-\begin{figure}
-<<ADNI1_cross_validation_means, cache=TRUE>>=
-t_data       = read.csv(file='data/a2a_tracc_results_2012_08_30.csv')
-a_data       = read.csv(file='data/a2a_ants_results_2012_08_30.csv')
-
-t_multiatlas = read.csv(file='data/a2a_tracc_multiatlas_2012_08_14.csv')
-a_multiatlas = read.csv(file='data/a2a_ants_multiatlas_2012_08_14.csv')
-t_ma_means   = aggregate( k ~ num_atlases , data=t_multiatlas, mean)
-a_ma_means   = aggregate( k ~ num_atlases , data=a_multiatlas, mean)
-
+<<ADN1_XVal_prep, include=FALSE>>=
+
+# all data from the validation trials
+t_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
+t_data       = subset(t_data, batch=="batch_1" & atlases %in% c(1,3,5,7,9))
+a_data       = t_data[,] #read.csv(gzfile('data/a2a_tracc/results_2013_01_03.csv.gz'))
+diagnoses    = read.csv('data/a2a_diagnoses.csv')
+t_data$se    = NULL
+t_data$sn    = NULL
+t_data$j     = NULL
+a_data$se    = NULL
+a_data$sn    = NULL
+a_data$j     = NULL       
 # add registration method to data frames (before joining)
@@
 
-# add column with kappa - multiatlas mean
-# num_atlases - 2 is a cheap trick to perform the necessary lookup into the ma_means tables
-t_data$k_minus_ma = t_data$k - t_ma_means[ t_data$num_atlases - 2, 2 ]
-a_data$k_minus_ma = a_data$k - a_ma_means[ a_data$num_atlases - 2, 2 ]
-
-all_data     = rbind(t_data,a_data)
-
-# use top_n as the number of templates for weighted voting schemes
-all_data[ all_data$method != "majvote", ]$num_templates = all_data[ all_data$method != "majvote", ]$top_n
-
-t_naive      = read.csv(file='data/a2a_tracc_naive.csv')
-a_naive      = read.csv(file='data/a2a_ants_naive_2012_08_30.csv')
-t_mean_naive = mean(t_naive$k)
-a_mean_naive = mean(a_naive$k)
-df_naive     = data.frame(
-                value = c(rep(t_mean_naive,3),rep(a_mean_naive,3)),
-                method = rep(c("majvote", "nmi", "xcorr"),2), 
-                reg_method = c(rep("ANIMAL",3),rep("ANTS",3)))
-
-method_labels <- list("Majority Vote" = "majvote", "NMI Vote" = "nmi", "Cross-correlation Vote" = "xcorr")
-levels(all_data$method) <- method_labels
-levels(df_naive$method) <- method_labels
-
-ggplot(all_data, aes(x=num_templates, y=k, colour=as.factor(num_atlases))) + 
-  facet_grid(method ~ reg_method) + 
-  geom_smooth() + 
-  geom_hline(data = df_naive, aes(yintercept = value, linetype="dashed")) + 
-  scale_colour_hue(name="Atlases") +
-  xlab( "Number of Templates" ) + ylab( "Mean Kappa" ) 
-@
-  \caption{Comparison of MAGeT performance on ADNI-1 subset.  Smoothing line
-  fitted using GAM (generalised additive model) from R with defaults from
-  ggplot2 (formula: y ~ s(x, bs = "cs"))}
-  \label{ADNI1_cross_validation_means}
-\end{figure}
+# baseline is multi-atlas majority vote (even though we have MA with other methods)
+t_ma = subset(t_data, approach=="multiatlas" & method=="majvote")
+a_ma = subset(a_data, approach=="multiatlas" & method=="majvote")
+t_mb = subset(t_data, approach=="mb")
+a_mb = subset(a_data, approach=="mb")
+
+ma = rbind(t_ma,a_ma)
+mb = rbind(t_mb,a_mb)
+ma$templates =  ma$atlases  # since there is no template library, these are equivalent
+mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n  # top_n is number of templates used
+
+all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"))
+all_data$k_diff = all_data$k.x - all_data$k.y
+method_labels <- list("Majority Vote" = "majvote", 
+                      "NMI Vote" = "nmi", 
+                      "Cross-correlation Vote" = "xcorr")
+levels(all_data$method.y) <- method_labels
+all_data       = merge(all_data, diagnoses)
+@
 
 \begin{figure}
-<<ADN1-cross-validation-MA-difference, cache=TRUE>>=
-ggplot(all_data, aes(x=num_templates, y=k_minus_ma, colour=as.factor(num_atlases))) + 
-  facet_grid(method ~ reg_method) + 
-  geom_smooth() + 
+<<ADN1-Xval-k_diff, cache=TRUE>>=
+ggplot(all_data, aes(x=templates.y, y=k_diff)) + 
+  facet_grid(reg_method~method.y) + 
+  stat_summary(fun.y=mean,geom='bar',position="dodge",
+               aes(y=k_diff,fill=as.factor(atlases))) + 
   scale_colour_hue(name="Atlases") +
   xlab( "Number of Templates" ) + 
-  ylab( "Difference in mean Kappa between voting method and multiatlas" )
+  ylab( "Difference in mean Kappa between voting method and multiatlas" ) +
+  coord_cartesian(ylim=c(0,0.07), xlim=c(0.5,10.5))
 @
@@
 
-<<a2a_multiatlas_means, results="asis",cache=FALSE>>=
-latex(merge(a_ma_means, t_ma_means, by="num_atlases"),
-colheads=c("Atlases", "ANTS", "ANIMAL"), file="", dec=2,
-	caption="Multi-atlas means")
-@
-
-%\begin{table}
-%    \begin{tabular}{c|c|c}
-%        Number of Atlases & ANIMAL Kappa (Jaccard) & ANTS Kappa (Jaccard) \\ \hline
-%        3                 & 0.76 (0.63)             & 0.80 (0.69)          \\ 
-%        4                 & 0.75 (0.62)             & 0.79 (0.67)          \\ 
-%        5                 & 0.79 (0.66)             & 0.82 (0.71)          \\ 
-%        6                 & 0.78 (0.65)             & 0.82 (0.70)          \\ 
-%        7                 & 0.80 (0.67)             & 0.83 (0.72)          \\ 
-%        8                 & 0.79 (0.67)             & 0.83 (0.72)          \\ 
-%        9                 & 0.80 (0.68)             & 0.84 (0.73)          \\
-%    \end{tabular}
-%\end{table}
+\begin{figure}
+<<ADN1-Xval-k_diff-ANTS-20_tmpls, cache=TRUE>>=
+a_20_tmpls = subset(all_data, 
+                      reg_method  == "ANTS" &
+                      method.y    == "Majority Vote" &
+                      templates.y == 20)
+ggplot(a_20_tmpls, aes(x=factor(atlases))) +
+  stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.x)) +
+  stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.x,group=1)) +
+  stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.y)) +
+  stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.y,group=1)) +
+  labs(x="Number of Atlases", y = "Kappa") +
+  scale_colour_hue(name="Method") + 
+  coord_cartesian(ylim=c(0.7,0.9))
+  #theme(legend.justification=c(1,0), legend.position=c(1,0))
+@
+  \caption{Difference in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
+  \label{}
+\end{figure}
+
+\begin{figure}
+<<ADN1-Xval-k_diff-ANTS-20_tmpls, cache=TRUE>>=
+a_20_tmpls = subset(all_data, 
+                      reg_method  == "ANTS" &
+                      method.y    == "Majority Vote" &
+                      templates.y == 20 &
+                      atlases %in% c(1,9))
+template_library = read.csv('data/a2a_ants/template_library_composition.csv')
+template_library$subject = NULL
+a_20_tmpls = merge(a_20_tmpls, template_library,
+                   by.x=c('timestamp', 'templates.y'), 
+                   by.y=c('timestamp', 'templates'))
+ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
+  facet_grid( . ~ atlases) + 
+  geom_smooth(method='lm',fullrange=F)
+
+@
+  \caption{Mean kappa vs. proportion of disease category in template library by disease category}
+  \label{}
+\end{figure}
 

### 2013-03-01 paper.Rnw
Added changes to the paper to reflect the new figures

@@
 %                                                                             %
-<<ADNI1-screening-demographics, echo=F, results="asis",cache=FALSE>>=
+<<ADNI1-scr-demographics, echo=F, results="asis",cache=FALSE>>=
 # RID < 2000 ensures only ADNI1 patients
@@
 %
-<<ADNI-69-demographics, echo=F, results="asis",cache=FALSE>>=
+<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
 adni_69_RIDs = c("295","413","619","685","729","782","938","954",
@@
 
-<<ADN1_XVal_prep, include=FALSE>>=
+<<ADN1-Xval_prep, include=FALSE>>=
 
@@
 t_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
-t_data       = subset(t_data, batch=="batch_1" & atlases %in% c(1,3,5,7,9))
+t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
 a_data       = t_data[,] #read.csv(gzfile('data/a2a_tracc/results_2013_01_03.csv.gz'))
@@
 all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"))
-all_data$k_diff = all_data$k.x - all_data$k.y
+all_data$k_diff = all_data$k.y - all_data$k.x 
 method_labels <- list("Majority Vote" = "majvote", 
@@
 all_data       = merge(all_data, diagnoses)
-@
+remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
+@
+
+\begin{figure}
+<<ADN1-Xval-k_mean, cache=TRUE>>=
+ggplot(all_data, aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
+  facet_grid(reg_method~method.y) + 
+  #stat_summary(fun.y=mean,geom='line',
+  #             aes(y=k.y)) + 
+  stat_smooth(method='lm', formula=y~log(x)) + 
+  scale_colour_hue(name="Atlases") +
+  xlab( "Number of Templates" ) + 
+  ylab( "mean Kappa" )
+@
+  \caption{MAGeT brain mean kappa by atlas and template library size}
+  \label{ADNI1_xval}
+\end{figure}
 
@@
 <<ADN1-Xval-k_diff, cache=TRUE>>=
-ggplot(all_data, aes(x=templates.y, y=k_diff)) + 
-  facet_grid(reg_method~method.y) + 
-  stat_summary(fun.y=mean,geom='bar',position="dodge",
-               aes(y=k_diff,fill=as.factor(atlases))) + 
+
+ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote"), 
+       aes(x=templates.y, y=k_diff, colour=as.factor(atlases))) + 
+  #facet_grid(reg_method~method.y) + 
+  #stat_summary(fun.y=mean,geom='line',position="dodge",
+  #             aes(y=k_diff,colour=as.factor(atlases))) + 
+  stat_smooth(method=lm, formula=y~log(x)) + 
   scale_colour_hue(name="Atlases") +
   xlab( "Number of Templates" ) + 
-  ylab( "Difference in mean Kappa between voting method and multiatlas" ) +
-  coord_cartesian(ylim=c(0,0.07), xlim=c(0.5,10.5))
-@
-  \caption{Difference in mean Kappa between MAGeT brain and multi-atlas}
+  ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
+  #coord_cartesian(ylim=c(0,0.07), xlim=c(0.5,10.5))
+@
+  \caption{Increase in mean Kappa between MAGeT brain and multi-atlas}
   \label{ADNI1_cross_validation_MA_difference}
@@
 \begin{figure}
-<<ADN1-Xval-k_diff-ANTS-20_tmpls, cache=TRUE>>=
+<<ADN1-Xval-k_diff-ANTS-20tmpls>>=
 a_20_tmpls = subset(all_data, 
@@
   stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.y,group=1)) +
-  labs(x="Number of Atlases", y = "Kappa") +
+  labs(x="Number of Atlases", y = "Kappa", 
+       title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
   scale_colour_hue(name="Method") + 
@@
 @
-  \caption{Difference in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
+  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
   \label{}
@@
 \begin{figure}
-<<ADN1-Xval-k_diff-ANTS-20_tmpls, cache=TRUE>>=
+<<ADN1-Xval-k-by-disease, cache=TRUE>>=
 a_20_tmpls = subset(all_data, 
@@
 
-<<ADNI-baseline-volumes-prep,echo=F,cache=T>>=
+<<ADNI-src-volumes-prep,echo=F,cache=T>>=
 yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
@@
 fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
-fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv")
+fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
 
@@
 # fetch just the columns we need, and do a little renaming
-mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
-maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
-snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
-ucsd_pruned   = data.frame(RID=ucsd$RID, VISCODE=ucsd$VISCODE, UCSD_L=ucsd$LHIPPOC, UCSD_R=ucsd$RHIPPOC)
+mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, 
+                           MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
+maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, 
+                           MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
+snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, 
+                           SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
+ucsd_pruned   = data.frame(RID=ucsd$RID, VISCODE=ucsd$VISCODE, 
+                           UCSD_L=ucsd$LHIPPOC, UCSD_R=ucsd$RHIPPOC)
 fs_pruned     = subset(fs, TEMPQC != "Fail")
-fs_pruned     = data.frame(FS_L=fs_pruned$ST29SV, FS_R=fs_pruned$ST88SV, Source = fs_pruned$IMAGEUID)
-fsl_pruned    = data.frame(FSL_L=fsl$X17, FSL_R=fsl$X53, Source=fsl$Source)
+fs_pruned     = data.frame(FS_L=fs_pruned$ST29SV, 
+                           FS_R=fs_pruned$ST88SV, Source = fs_pruned$IMAGEUID)
+fsl_pruned    = data.frame(FSL_L=fsl$X17, 
+                           FSL_R=fsl$X53, Source=fsl$Source)
 
@@
 \begin{figure}[h]
-<<ADNI-baseline-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
+<<ADNI-src-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
 totals = vols
@@
 \begin{figure}[h]
-<<ADNI-baseline-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
+<<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
 by_snt_volumes = melt(totals, measure.vars=c("FSL","MAPER","MAGeT", "FS", "UCSD"))
@@
 maget=subset(by_snt_volumes, variable=="MAGeT")
-
+with(by_snt_volumes,
 qplot(SNT,value,data=by_snt_volumes,
@@
     geom_point() +
-#     geom_text(aes(
-#       label=paste("FSL       R=", round(cor(fsl$SNT,fsl$value, use="complete.obs"), 2), 
-#                   ", p<", round(as.double(t.test(fsl$SNT,fsl$value, paired=T)["p.value"]), 2)),
-#       x=1800,y=10000, hjust=0,size=2)) + 
+     geom_text(aes(
+       label=paste("FSL       R=", round(cor(fsl$SNT,fsl$value, use="complete.obs"), 2), 
+                   ", p<", round(as.double(t.test(fsl$SNT,fsl$value, paired=T)["p.value"]), 2)),
+       x=1800,y=10000, hjust=0,size=2), data=data.frame()) + 
 #     geom_text(aes(
@@
     xlab("Total Hippocampal Volume (Manual Segmentation)") + 
-    ylab("Total Hippocampal Volume (Automatic Segmentation)")
+    ylab("Total Hippocampal Volume (Automatic Segmentation)"))
 @
@@
 
+<<ADNI-scr-summary-table, echo=F, results="asis",cache=FALSE>>=
+melted = melt(totals, measure.vars=c("FSL","MAPER","SNT", "FS", "UCSD"))
+cor_by_dx    = by(melted[c("MAGeT", "value")], c(melted['variable'], melted['DX']), 
+                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
+cor_overall  = by(melted[c("MAGeT", "value")], melted['variable'], 
+                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
+cor_by_dx = t(rbind(cor_by_dx))
+cor_overall = rbind(cor_overall)
+
+vol_by_dx = t(rbind(by(totals$MAGeT, totals$DX, mean)))
+colnames(vol_by_dx) <- "Mean Volume"
+vol_overall = matrix(mean(totals$MAGeT),dimnames=list("Overall"))
+
+dx      = cbind(vol_by_dx, cor_by_dx)
+overall <- cbind(vol_overall, cor_overall)
+
+latex(format.df(rbind(dx, overall), dec=2),
+      n.cgroup=c(1,5),cgroup=c("","Volume Correlation"),
+      n.rgroup=c(4,1), file="", size="scriptsize", )
+
+@
+
 %%%%%
@@
 \begin{figure}
-<<SZ_volumes,echo=F,fig=T, fig.align='center',cache=T>>=
+<<FEP_volumes,echo=F,fig=T, fig.align='center',cache=T>>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
+with(SZ_volumes,
 qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm") + 
-#   geom_text(aes(
-#       label=paste("R=", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
-#                   ", p<", t.test(Manual, MAGeT, paired=T)["p.value"]),
-#       x=6500,y=10000, hjust=0,size=2)) +
+   geom_text(aes(
+       label=paste("r = ", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
+                   ", p <", round(as.numeric(t.test(Manual, MAGeT, paired=T)["p.value"]), 6))),
+       x=7000,y=10000, hjust=0,vjust=-0.5,size=5,data=data.frame()) +
     xlab("Total Hippocampal Volume (Manual Segmentation)") + 
     ylab("Total Hippocampal Volume (MAGeT Segmentation)")
-
+)
 @

### 2013-03-01 r/a2a.r

@@ -1,192 +1,0 @@
-library(ggplot2)
-# Authors jon julie jason martin jens aristotle mallar
-#X11 = function (...) grDevices::x11(...,type='cairo')
-#X11 = CairoX11()
-
-multiplot <- function(..., plotlist=NULL, cols) {
-    require(grid)
-
-    # Make a list from the ... arguments and plotlist
-    plots <- c(list(...), plotlist)
-
-    numPlots = length(plots)
-
-    # Make the panel
-    plotCols = cols                          # Number of columns of plots
-    plotRows = ceiling(numPlots/plotCols) # Number of rows needed, calculated from # of cols
-
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(plotRows, plotCols)))
-    vplayout <- function(x, y)
-        viewport(layout.pos.row = x, layout.pos.col = y)
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-        curRow = ceiling(i/plotCols)
-        curCol = (i-1) %% plotCols + 1
-        print(plots[[i]], vp = vplayout(curRow, curCol ))
-    }
-
-}
-
-#####################################################
-# TRACC
-t_naive      = read.csv(file='a2a_tracc_naive.csv')
-t_data       = read.csv(file='a2a_tracc_results_2012_08_30.csv')
-t_maj        = subset(t_data, method == 'majvote')
-t_xcorr      = subset(t_data, method == 'xcorr')
-t_nmi        = subset(t_data, method == 'nmi')
-t_mean_naive = mean(t_naive$k)
-t_ylimits    = c(0.7,0.85)
-
-t_multiatlas = read.csv(file='a2a_tracc_multiatlas_2012_08_14.csv')
-t_ma_means   = aggregate( k ~ num_atlases , data=t_multiatlas, mean)
-t_maj$k_minus_ma_mean   <- t_maj$k   - t_ma_means[ t_maj$num_atlases - 3, 2 ]
-t_xcorr$k_minus_ma_mean <- t_xcorr$k - t_ma_means[ t_xcorr$num_atlases - 3, 2 ]
-t_nmi$k_minus_ma_mean   <- t_nmi$k   - t_ma_means[ t_nmi$num_atlases - 3, 2 ]
-
-# Majority Vote
[…]

### 2013-03-01 r/adni_classification.r

@@ -1,9 +1,0 @@
-# Classify the 69 ADNI subjects by hippocampal volume
-data = read.csv('adni_data.csv', sep=";")
-
-# defaults showing t-values for AD to NX 
-summary(lm((HC_left_vol + HC_right_vol) ~ DX + Sex + Age, data = data))
-
-# NX to AD/NX
-data$DX<-relevel(x=data$DX, ref="NX")
-summary(lm((HC_left_vol + HC_right_vol) ~ DX + Sex + Age, data = data))

### 2013-03-01 r/dan-model-validation.r

@@ -1,35 +1,0 @@
-library(rms)
-data = read.csv('adni_data.csv', sep=';')
-summary(data)
-dd <- datadist(data)
-options(datadist="dd")
-describe(data)
-mod <- lrm(data=data, DX ~ hc_left_vol + Age + Sex)
-mod
-modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex)
-modright <- lrm(data=data, DX ~ hc_right_vol + Age + Sex)
-modright <- lrm(data=data, DX ~ hc_right_vol + Age + Sex, subset=DX!=MCI)
-modright <- lrm(data=data, DX ~ hc_right_vol + Age + Sex, subset=DX!="MCI")
-modright
-modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="MCI")
-modleft
-
-summary(modleft)
-summary(modright)
-modleft
-modleft2 <- lrm(data=data, DX ~ Age + Sex, subset=DX!="MCI")
-lrtest(modleft, modleft2)
-validate(modleft, method=.632)
-modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="MCI", x=TRUE, y=TRUE)
-validate(modleft, method=.632)
-validate(modleft, method=".632")
-validate(modleft, method=boot)
-validate(modleft, method="boot")
-calibrate(modleft, method="boot")
-modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="MCI", x=TRUE, y=TRUE)
-modleft
-summary(modleft)
-print(modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="AD", x=TRUE, y=TRUE))
-summary(modleft)
-print(modleft <- lrm(data=data, DX ~ hc_left_vol + Age + Sex, subset=DX!="NX", x=TRUE, y=TRUE))
-savehistory("dan-model-validation.R")

### 2013-03-01 paper.Rnw
Winterburn atlas figure

@@
 
+<<WAval-boxplot,cache=T>>=
+# dirty dirty
+data = read.csv('data/WAval.csv')
+orig = subset(data, version == "orig")
+voted = subset(data, version == "voted")
+origtotal = orig[,2:6] + orig[,7:11]
+votedtotal = voted[,2:6] + voted[,7:11]
+pchange = abs((origtotal - votedtotal)/origtotal)
+diff = cbind(orig[,1],pchange)
+
+melted = melt(diff, id.vars="orig[, 1]")
+names(melted) <- c("subject","variable","value")
+ggplot(melted, aes(y=value,x=variable,colour=subject,group=subject)) + 
+  geom_point(size=3) + 
+  geom_lin() + 
+  labs(title="Percent difference in MAGeT volumes from original",
+       y="Percent difference in volume") 
+@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-03-07 paper.Rnw
Move movement on the results

--- 2013-03-01 paper.Rnw
@@
 <<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
-adni_69_RIDs = c("295","413","619","685","729","782","938","954",
-                 "1018","1155","907","981","222","324","448","546",
-                 "553","572","602","610","814","1341","675","681",
-                 "731","1130","41","68","70","101","249","293","316",
-                 "414","698","1206","1222","1339","751","1030","3",
-                 "5","10","16","22","53","168","183","241","1205",
-                 "328","1095","991","213","159","337","343","642",
-                 "753","1109","1217","29","56","1188","1293","149",
-                 "382","431","1202")
+adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
+                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
+                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
+                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
+                 1217,29,56,1188,1293,149,382,431,1202)
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
@@
 library.  The template library consisted of 21 randomly selected images from the
-ADNI1 data dataset (7 healthy, MCI and AD subjects).
+ADNI1 data dataset (7 healthy, MCI and AD subjects). Clinical and demographic
+data for the template library subjects are shown in Table
+\ref{tab:ADNI1_scr_tmpllib_demographics}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1 X-Validation Dataset Demographics Table
+%
+<<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
+scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
+                 0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
+                 0159)
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
+               scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+
+latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
+      caption="ADNI1 Screening Template Library demographics", 
+      title="tab:ADNI1_scr_tmpllib_demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%
@@
 
-- find significant improvement over multi-atlas performed with the same
-  parameters. Also, find smoothed performance is monotonically increasing but
-  asymptotic in size of both template and atlas library, with peak performance
-  reached after 15 templates.
-
-- True for all parameter settings! (so works generally..)
-\todo{Can we statistically capture what "peak" performance is?  Something like,
-the point at which gains become statistically insignificant?}
-
-<<ADN1-Xval_prep, include=FALSE>>=
-
+%                                                                              %
+% ADNI1 Screening Validation - results preparation                             %
+%                                                                              %
+<<ADN1-Xval-prep, include=FALSE, cache=TRUE>>=
 # all data from the validation trials
@@
 diagnoses    = read.csv('data/a2a_diagnoses.csv')
+
+# simplifies the table
 t_data$se    = NULL
@@
 a_data$j     = NULL
-# add registration method to data frames (before joining)
+
+# add registration method column 
 t_data       = cbind(t_data, reg_method = "ANIMAL")
@@
 
-# baseline is multi-atlas majority vote (even though we have MA with other methods)
+# separate MA majvote from MB so we can compute differences
 t_ma = subset(t_data, approach=="multiatlas" & method=="majvote")
@@
 mb = rbind(t_mb,a_mb)
-ma$templates =  ma$atlases  # since there is no template library, these are equivalent
-mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n  # top_n is number of templates used
+
+# set up a few equivalences for multi-atlas, to make nomenclature simpler
+ma$templates =  ma$atlases  # because no template library
+mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n
 
@@
 @
+%                                                                              %
+%                                                                              %
+
+The
 
 \begin{figure}
-<<ADN1-Xval-k_mean, cache=TRUE>>=
+<<ADN1-Xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
 ggplot(all_data, aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
@@
   xlab( "Number of Templates" ) + 
-  ylab( "mean Kappa" )
-@
-  \caption{MAGeT brain mean kappa by atlas and template library size}
-  \label{ADNI1_xval}
+  ylab( "Mean DSC" )
+@
+  \caption{MAGeT brain mean DSC by atlas and template library size}
+  \label{fig:ADNI1_xval-k_mean}
 \end{figure}
@@
 \begin{figure}
-<<ADN1-Xval-k_diff, cache=TRUE>>=
-
+<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
 ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote"), 
@@
 \begin{figure}
-<<ADN1-Xval-k_diff-ANTS-20tmpls>>=
+<<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
 a_20_tmpls = subset(all_data, 
@@
 \begin{figure}
-<<ADN1-Xval-k-by-disease, cache=TRUE>>=
+<<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
 a_20_tmpls = subset(all_data, 
@@
 
+- find significant improvement over multi-atlas performed with the same
+  parameters. Also, find smoothed performance is monotonically increasing but
+  asymptotic in size of both template and atlas library, with peak performance
+  reached after 15 templates.
+
+- True for all parameter settings! (so works generally..)
+\todo{Can we statistically capture what "peak" performance is?  Something like,
+the point at which gains become statistically insignificant?}
+
 Ideas:
@@
 <<ADNI-src-volumes-prep,echo=F,cache=T>>=
-yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+bl1.5T= read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
+bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
 mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
@@
 # recode VISCODE
-viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
-yr1$VISCODE <- as.factor(yr1$Visit)
-levels(yr1$VISCODE) <-viscode_levels
-
-yr1$Scaled_2 <- grepl("Scaled_2", yr1$Description)
+#viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
+#yr1$VISCODE <- as.factor(yr1$Visit)
+#levels(yr1$VISCODE) <-viscode_levels
+bl1.5T$VISCODE <- 'bl'
+bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)
 
 #TODO: include Scaled_2? 
-subjects = subset(yr1, VISCODE == 'bl' & Scaled_2 == FALSE)
+subjects = bl1.5T
 
@@
 attach(combined)
-vols = data.frame(RID = RID, VISCODE = VISCODE, DX = DX.bl,
+totals = data.frame(RID = RID, VISCODE = VISCODE, DX = DX.bl,
                   MAGeT = MAGeT_L + MAGeT_R, 
@@
                   FSL   = FSL_L + FSL_R)
+detach(combined)
+complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
 @
@@
 \begin{figure}[h]
-<<ADNI-src-volumes-boxplot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-prep',cache=T>>=
-totals = vols
-qplot(DX,value,data=melt(totals,measure.vars=c("FS", "FSL","UCSD", "MAPER","MAGeT","SNT"),na.rm=T),
+<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T>>=
+qplot(DX,value,data=melt(complete_totals,measure.vars=c("FS", "FSL","UCSD", "MAPER","MAGeT","SNT")),
       colour=variable,geom="boxplot") + 
       xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
-
 @
   \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
-  \label{ADNI-baseline-volumes-boxplot}
+  \label{ADNI-scr-volumes-boxplot}
+\end{figure}
+
+\begin{figure}[h]
+<<ADNI-scr-histogram,dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
+melted = melt(complete_totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
+ggplot(data=melted, aes(x=value)) + 
+       geom_histogram(binwidth=250) +
+       facet_grid( variable ~ DX) 
+@
+  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
+  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \label{ADNI-scr-histogram}
 \end{figure}
@@
 <<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
-by_snt_volumes = melt(totals, measure.vars=c("FSL","MAPER","MAGeT", "FS", "UCSD"))
-fsl=subset(by_snt_volumes, variable=="FSL")
-maper=subset(by_snt_volumes, variable=="MAPER")
-maget=subset(by_snt_volumes, variable=="MAGeT")
+by_snt_volumes = melt(totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
+
 with(by_snt_volumes,
-qplot(SNT,value,data=by_snt_volumes,
-      colour=variable, geom="smooth", method="lm") + 
-    geom_point() +
-     geom_text(aes(
-       label=paste("FSL       R=", round(cor(fsl$SNT,fsl$value, use="complete.obs"), 2), 
-                   ", p<", round(as.double(t.test(fsl$SNT,fsl$value, paired=T)["p.value"]), 2)),
-       x=1800,y=10000, hjust=0,size=2), data=data.frame()) + 
+ggplot(data=by_snt_volumes, aes(x=value)) + geom_histogram(binwidth=250) +
+    facet_grid( variable ~ DX) 
+     )
 #     geom_text(aes(
@@
 #       x=1800,y= 9000, hjust=0, size=2)) +
-    xlab("Total Hippocampal Volume (Manual Segmentation)") + 
-    ylab("Total Hippocampal Volume (Automatic Segmentation)"))
+#   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
+#   ylab("Total Hippocampal Volume (Automatic Segmentation)")) 
+@
+  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
+  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \label{ADNI-baseline-volumes-plot}
+\end{figure}
+
+\begin{figure}[h]
+<<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
+by_snt_volumes = melt(totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
+
+with(by_snt_volumes,
+ggplot(data=by_snt_volumes, aes(x=value)) + geom_histogram(binwidth=250) +
+    facet_grid( variable ~ DX) 
+     )
+#     geom_text(aes(
+#       label=paste("MAPER R=", round(cor(maper$SNT,maper$value, use="complete.obs"), 2), 
+#                   ", p<", t.test(maper$SNT,maper$value, paired=T)["p.value"]), 
+#       x=1800,y= 9500, hjust=0, size=2)) +
+#     geom_text(aes(
+#       label=paste("MAGeT  R=", round(cor(maget$SNT,maget$value, use="complete.obs"),2), 
+#                   ", p<", t.test(maget$SNT,maget$value, paired=T)["p.value"]), 
+#       x=1800,y= 9000, hjust=0, size=2)) +
+#   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
+#   ylab("Total Hippocampal Volume (Automatic Segmentation)")) 
 @
@@
       n.cgroup=c(1,5),cgroup=c("","Volume Correlation"),
-      n.rgroup=c(4,1), file="", size="scriptsize", )
+      n.rgroup=c(4,1), file="", size="scriptsize")
 
@@
 \begin{figure}
-<<FEP_volumes,echo=F,fig=T, fig.align='center',cache=T>>=
+<<FEP-volumes,cache=T>>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
@@
 qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm") + 
-   geom_text(aes(
-       label=paste("r = ", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
-                   ", p <", round(as.numeric(t.test(Manual, MAGeT, paired=T)["p.value"]), 6))),
-       x=7000,y=10000, hjust=0,vjust=-0.5,size=5,data=data.frame()) +
+#   geom_text(aes(
+#       label=paste("r = ", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
+#                   ", p <", round(as.numeric(t.test(Manual, MAGeT, paired=T)["p.value"]), 6))),
+#       x=7000,y=10000, hjust=0,vjust=-0.5,size=5,data=data.frame()) +
    xlab("Total Hippocampal Volume (Manual Segmentation)") + 
@@
 
+<<WAval-vol-boxplot,cache=T>>=
+# dirty dirty
+data = subset(read.csv('data/WAval.csv'), file != 'anusha')
+gold_0.3mm = subset(data, version == "gold_0.3mm")
+gold_0.6mm = subset(data, version == "gold_0.6mm")
+mb_bravo   = subset(data, version == "mb_bravo")
+mb_0.6mm   = subset(data, version == "mb_0.6mm")
+
+gold_0.3mm_total = gold_0.3mm[,2:6] + gold_0.3mm[,7:11]
+gold_0.6mm_total = gold_0.6mm[,2:6] + gold_0.6mm[,7:11]
+mb_bravo_total   = mb_bravo[,2:6] + mb_bravo[,7:11]
+mb_0.6mm_total   = mb_0.6mm[,2:6] + mb_0.6mm[,7:11]
+
+gold_diff     = (gold_0.3mm_total - gold_0.6mm_total)/gold_0.3mm_total
+mb_bravo_diff = (gold_0.3mm_total - mb_bravo_total)/gold_0.3mm_total
+mb_0.6mm_diff = (gold_0.3mm_total - mb_0.6mm_total)/gold_0.3mm_total
+
+subjects = gold_0.6mm[,1]
+mb0.6mm_diff_over_gold_diff = cbind(subjects, mb_0.6mm_diff - mean(gold_diff) , res="diffdff")
+gold_diff     = cbind(subjects,gold_diff, res="0.3mm gold & 0.6mm gold")
+mb_bravo_diff = cbind(subjects,mb_bravo_diff, res="0.3mm gold & MAGeT-on-BRAVO")
+mb_0.6mm_diff = cbind(subjects,mb_0.6mm_diff, res="0.3mm gold & MAGeT-at-0.6mm")
+
+diff = rbind(mb_bravo_diff,mb_0.6mm_diff,gold_diff, mb0.6mm_diff_over_gold_diff)
+names(diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
+
+melted = melt(diff, id.vars=c("subject", "res"))
+names(melted) <- c("subject","res", "region","voldiff")
+ggplot(melted, aes(y=voldiff,x=region)) +
+  geom_boxplot() + facet_grid( . ~ res)  + 
+  labs(title="Percent difference in MAGeT volumes from original",
+       y="Percent difference in volume", 
+       x="Region") 
+@
+
+<<WAval-kappa,cache=T>>=
+data = read.csv('data/WAval-kappa.csv')
+ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)
+@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-03-12 paper.Rnw
Added normals and normality tests to the ADNI1 Screening histogram as per Tris's comments

@@
 melted   = melt(complete_totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
+binwidth = 250
+
+# prepare the normals
+grid <- with(melted, seq(min(value), max(value), length = 100))
+normaldf <- ddply(melted, c("variable", "DX"), function(df) {
+  data.frame( 
+    value = grid,
+    density = dnorm(grid, mean(df$value), sd(df$value)),
+    freq = length(df$value) * dnorm(grid, mean(df$value), sd(df$value)) * binwidth,
+    length = length(df$value)
+  )
+})
+
+testsdf <- ddply(melted, c("variable", "DX"), function(df) {
+  shapiro = shapiro.test(df$value)
+  data.frame(
+    shapiro.w = shapiro$statistic,
+    shapiro.p = shapiro$p.value,
+    mean = mean(df$value),
+    sd = sd(df$value)
+  )
+})
+
+label_x_pos = 10000
 ggplot(data=melted, aes(x=value)) + 
-       geom_histogram(binwidth=250) +
-       facet_grid( variable ~ DX) 
-@
-  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
-  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  geom_histogram(binwidth=binwidth, alpha=0.8) + 
+  geom_line(aes(y=freq, colour='red'), data=normaldf) +
+  geom_text(aes(y=29,x=label_x_pos,hjust=1,size=1, 
+                label=paste("list(W ==",signif(shapiro.w, digits=3), ", p == ", 
+                            signif(shapiro.p,digits=3), ")")
+                ), parse=TRUE,
+            data=testsdf) + 
+  geom_text(aes(y=25,x=label_x_pos,hjust=1,size=1, 
+                label=paste("sigma ==",signif(sd, digits=3))), parse=TRUE,
+            data=testsdf) + 
+  geom_text(aes(y=21,x=label_x_pos,hjust=1,size=1, 
+                label=paste("mu ==",signif(mean, digits=3))), parse=TRUE,
+            data=testsdf) + 
+  xlim(0, 10000) + ylim(0,30) + 
+  facet_grid(variable ~ DX) + 
+  scale_colour_discrete(legend=FALSE) + 
+  scale_size_continuous(legend=FALSE) + 
+  labs(title="Total hippocampus volume by disease and measurement method", 
+       x=expression(paste("Total hippocampal volume (", mm^2, ")")),
+       y="Frequency") + 
+  opts(axis.title.x = theme_text(vjust=-0.1))
+@
+  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes as measured by SNT, FSL, FreeSurfer (FS), MAPER, and MAGeT brain (MAGeT).
+A fitted normal curve is shown in red.
+$W$ is the Shapiro-Wilkes test statistic measuring normality of the data (low p-values indicate non-normality).}
   \label{ADNI-scr-histogram}

### 2013-03-19 paper.Rnw
Bland-Altman plot, plus stddev test of X-validation with poor results... booo

--- 2013-03-07 paper.Rnw
@@
 theme_set(theme_bw())
+# bland_altman_plot <- function(dat){
+#   dat$diff <- dat$pre - dat$post;
+#   
+#   return ggplot(dat, aes(pre, diff)) + 
+#     geom_hline(yintercept = mean(dat$diff) + c(-2, 0, 2) * sd(dat$diff), 
+#                linetype=2, color='brown') + 
+#     geom_point() 
+# }
 @
@@
 %                                                                              %
-
-The
+ % Bland-Altman plot for MA and MAGeT
 
@@
   #facet_grid(reg_method~method.y) + 
-  #stat_summary(fun.y=mean,geom='line',position="dodge",
-  #             aes(y=k_diff,colour=as.factor(atlases))) + 
-  stat_smooth(method=lm, formula=y~log(x)) + 
+  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
+               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
+  stat_summary(fun.y=mean, geom="point", 
+               aes(y=k_diff,colour=as.factor(atlases))) + 
+  #stat_smooth(method=lm, formula=y~log(x)) + 
+  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
+  coord_cartesian(ylim=c(-0.1,0.05)) + 
+  scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
   scale_colour_hue(name="Atlases") +
+  scale_alpha_continuous(legend=FALSE) + 
+  scale_linetype_discrete(legend=FALSE) + 
   xlab( "Number of Templates" ) + 
   ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
-  #coord_cartesian(ylim=c(0,0.07), xlim=c(0.5,10.5))
 @
@@
 \end{figure}
+
+\begin{figure}
+<<ADN1-Xval-k-sd-t-test-p-plot, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+    ANTSmajvote = subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote")
+
+stddev = ddply(ANTSmajvote, c("subject", "label","atlases","templates.y"), function (df) {
+  data.frame(
+    MA = sd(df$k.x),
+    MB = sd(df$k.y) 
+  )
+})
+
+stddev.t.test = ddply(stddev, c("atlases","templates.y"), function (df) {
+  t = t.test(df$MA, df$MB)
+  data.frame(
+    statistic = t$statistic, 
+    p.value = t$p.value
+  )
+})
+
+ggplot(stddev.t.test, 
+       aes(x=templates.y, y=p.value, colour=as.factor(atlases))) + 
+  geom_line() + 
+  geom_hline(aes(yintercept=0.05), linetype='dashed') + 
+  scale_colour_hue(name="Atlases") +
+  scale_alpha_continuous(guide='none') + 
+  scale_linetype_discrete(guide='none') + 
+  scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
+  scale_y_log10() + 
+  xlab( "Number of Templates" ) + 
+  ylab( "p-value" ) 
+@
+\end{figure}
+
+%- show variance in MAGeT performance by showing nine plots, each with mean multi-atlas performance plotted.
 
@@
 \begin{figure}
+<<ADN1-Xval-ANTS-20tmpls-t-test-sd, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+a_20_tmpls = subset(all_data, 
+                      reg_method  == "ANTS" &
+                      method.y    == "Majority Vote" &
+                      templates.y == 20 &
+                      atlases     == 9 & 
+                      volume.x < 10000)
+stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
+  data.frame(
+    MA = sd(df$k.x),  #MA
+    MB = sd(df$k.y)  #MB
+  )
+})
+
+# TODO: against SNT
+melted = melt(stddev, id.vars=c("subject", "label"))
+ggplot(melted, aes(x=variable, y=value)) + geom_boxplot()
+t.test(stddev$MA, stddev$MB)
+@
+  \caption{}
+  \label{}
+\end{figure}
+
+\begin{figure}
+<<ADN1-Xval-ANTS-20tmpls-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+a_20_tmpls = subset(all_data, 
+                      reg_method  == "ANTS" &
+                      method.y    == "Majority Vote" &
+                      templates.y == 20 &
+                      atlases     == 9 &
+                      volume.x < 10000)
+avg = ddply(a_20_tmpls, c("subject", "label"), function (df) {
+  data.frame(
+    pre = mean(df$volume.x),  #MA
+    post = mean(df$volume.y)  #MB
+  )
+})
+# TODO: against SNT
+
+#df = data.frame(pre=a_20_tmpls$volume.x, post=a_20_tmpls$volume.y)
+bland_altman_plot(avg) + 
+  xlab('Mean of MA and MAGeT HC Left/Right Volumes') + 
+  ylab('MA - MAGeT Volume') + 
+  opts(title = 'Bland-Altman Plot')
+@
+  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
+  \label{}
+\end{figure}
+
+\begin{figure}
 <<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
@@
   geom_smooth(method='lm',fullrange=F) 
-
+ # add 
 @
@@
 registrations per Kappa?  or hours of manual labour per Kappa?)}
+% discuss run-time for each of MAGet and multi-atlas
 
@@
 combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
+#combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
 combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
@@
 attach(combined)
-totals = data.frame(RID = RID, VISCODE = VISCODE, DX = DX.bl,
+totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                   MAGeT = MAGeT_L + MAGeT_R, 
@@
                   SNT   = SNT_L + SNT_R, 
-                  UCSD  = UCSD_L + UCSD_R,
+                  #UCSD  = UCSD_L + UCSD_R,
                   FS    = FS_L + FS_R, 
                   FSL   = FSL_L + FSL_R)
+
 detach(combined)
-complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
+#complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
+totals.complete = na.omit(totals)
 @
@@
 <<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T>>=
-qplot(DX,value,data=melt(complete_totals,measure.vars=c("FS", "FSL","UCSD", "MAPER","MAGeT","SNT")),
+qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
       colour=variable,geom="boxplot") + 
@@
 
+% paired t-test against SNT to show similarity (should be low) 
+% Kruskal-Willis against  (should be high?)
+% post-hoc power analysis to say we can detect this change
 \begin{figure}[h]
 <<ADNI-scr-histogram,dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
-melted = melt(complete_totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
+melted   = melt(totals.complete, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
+binwidth = 250
+
+# prepare the normals
+grid <- with(melted, seq(min(value), max(value), length = 100))
+normaldf <- ddply(melted, c("variable", "DX"), function(df) {
+  data.frame( 
+    value = grid,
+    density = dnorm(grid, mean(df$value), sd(df$value)),
+    freq = length(df$value) * dnorm(grid, mean(df$value), sd(df$value)) * binwidth,
+    length = length(df$value)
+  )
+})
+
+testsdf <- ddply(melted, c("variable", "DX"), function(df) {
+  shapiro = shapiro.test(df$value)
+  data.frame(
+    shapiro.w = shapiro$statistic,
+    shapiro.p = shapiro$p.value,
+    mean = mean(df$value),
+    sd = sd(df$value)
+  )
+})
+
+melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
+snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
+  ks = ks.test(df$SNT, df$value)
+  data.frame( 
+    ks.D = ks$statistic,
+    ks.p = ks$p.value
+  )
+})
+
+label_x_pos = 10000
 ggplot(data=melted, aes(x=value)) + 
-       geom_histogram(binwidth=250) +
-       facet_grid( variable ~ DX) 
-@
-  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
-  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  geom_histogram(binwidth=binwidth, alpha=0.8) + 
+  geom_line(aes(y=freq, colour='red'), data=normaldf) +
+  geom_text(aes(y=29,x=label_x_pos,hjust=1,size=1, 
+                label=paste("list(W ==",signif(shapiro.w, digits=3), ", p == ", 
+                            signif(shapiro.p,digits=3), ")")), 
+            parse=TRUE, data=testsdf) +
+  geom_text(aes(y=25,x=label_x_pos,hjust=1,size=1, 
+                label=paste("list(D ==",signif(ks.D, digits=3), ", p == ", 
+                            signif(ks.p,digits=3), ")")), 
+            parse=TRUE, data=snt.tests) + 
+  geom_text(aes(y=21,x=label_x_pos,hjust=1,size=1, 
+                label=paste("sigma ==",signif(sd, digits=3))), 
+            parse=TRUE, data=testsdf) + 
+  geom_text(aes(y=17,x=label_x_pos,hjust=1,size=1, 
+                label=paste("mu ==",signif(mean, digits=3))), 
+            parse=TRUE,data=testsdf) + 
+  xlim(0, 10000) + ylim(0,30) + 
+  facet_grid(variable ~ DX) + 
+  scale_colour_discrete(legend=FALSE) + 
+  scale_size_continuous(legend=FALSE) + 
+  labs(title="Total hippocampus volume by disease and measurement method", 
+       x=expression(paste("Total hippocampal volume (", mm^2, ")")),
+       y="Frequency") + 
+  opts(axis.title.x = theme_text(vjust=-0.1))
+@
+  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes as measured by SNT, FSL, FreeSurfer (FS), MAPER, and MAGeT brain (MAGeT).
+A fitted normal curve is shown in red.
+$W$ is the Shapiro-Wilkes test statistic measuring normality of the data (signficance indicates non-normality).
+$D$ is the Kolmogorov-Smirnov test statistic measuring the goodness-of-fit between the distribution of measured volumes and SNT volumes (significance indicates a difference).}
   \label{ADNI-scr-histogram}
@@
 
-\begin{figure}[h]
-<<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
-by_snt_volumes = melt(totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
-
-with(by_snt_volumes,
-ggplot(data=by_snt_volumes, aes(x=value)) + geom_histogram(binwidth=250) +
-    facet_grid( variable ~ DX) 
-     )
-#     geom_text(aes(
-#       label=paste("MAPER R=", round(cor(maper$SNT,maper$value, use="complete.obs"), 2), 
-#                   ", p<", t.test(maper$SNT,maper$value, paired=T)["p.value"]), 
-#       x=1800,y= 9500, hjust=0, size=2)) +
-#     geom_text(aes(
-#       label=paste("MAGeT  R=", round(cor(maget$SNT,maget$value, use="complete.obs"),2), 
-#                   ", p<", t.test(maget$SNT,maget$value, paired=T)["p.value"]), 
-#       x=1800,y= 9000, hjust=0, size=2)) +
-#   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
-#   ylab("Total Hippocampal Volume (Automatic Segmentation)")) 
-@
-  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
-  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
-  \label{ADNI-baseline-volumes-plot}
-\end{figure}
-
-\begin{figure}[h]
-<<ADNI-src-volumes-plot,echo=F,fig=T, fig.align='center',dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
-by_snt_volumes = melt(totals, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
-
-with(by_snt_volumes,
-ggplot(data=by_snt_volumes, aes(x=value)) + geom_histogram(binwidth=250) +
-    facet_grid( variable ~ DX) 
-     )
-#     geom_text(aes(
-#       label=paste("MAPER R=", round(cor(maper$SNT,maper$value, use="complete.obs"), 2), 
-#                   ", p<", t.test(maper$SNT,maper$value, paired=T)["p.value"]), 
-#       x=1800,y= 9500, hjust=0, size=2)) +
-#     geom_text(aes(
-#       label=paste("MAGeT  R=", round(cor(maget$SNT,maget$value, use="complete.obs"),2), 
-#                   ", p<", t.test(maget$SNT,maget$value, paired=T)["p.value"]), 
-#       x=1800,y= 9000, hjust=0, size=2)) +
-#   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
-#   ylab("Total Hippocampal Volume (Automatic Segmentation)")) 
-@
-  \caption{{\bf ADNI Baseline cohort.} Comparison of HC volumes by FreeSurfer
-  (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
-  \label{ADNI-baseline-volumes-plot}
-\end{figure}
-
+%
+
+% show as a plot
 <<ADNI-scr-summary-table, echo=F, results="asis",cache=FALSE>>=
-melted = melt(totals, measure.vars=c("FSL","MAPER","SNT", "FS", "UCSD"))
+# a rather dirty way of computing the pairwise correlations and then forming
+# a printable table out of them.  Sorry.
+melted = melt(totals.complete, measure.vars=c("SNT","FSL", "FS", "MAPER"))
 cor_by_dx    = by(melted[c("MAGeT", "value")], c(melted['variable'], melted['DX']), 
@@
 
-vol_by_dx = t(rbind(by(totals$MAGeT, totals$DX, mean)))
+vol_by_dx = t(rbind(by(totals.complete$MAGeT, totals.complete$DX, mean)))
 colnames(vol_by_dx) <- "Mean Volume"
-vol_overall = matrix(mean(totals$MAGeT),dimnames=list("Overall"))
+vol_overall = matrix(mean(totals.complete$MAGeT),dimnames=list("Overall"))
 
 dx      = cbind(vol_by_dx, cor_by_dx)
-overall <- cbind(vol_overall, cor_overall)
-
-latex(format.df(rbind(dx, overall), dec=2),
-      n.cgroup=c(1,5),cgroup=c("","Volume Correlation"),
-      n.rgroup=c(4,1), file="", size="scriptsize")
-
+overall = cbind(vol_overall, cor_overall)
+df      = format.df(rbind(dx, overall), dec=2)
+
+latex(df, title="",
+      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
+      )#file="", size="scriptsize")
 @
@@
 
-gold_diff     = (gold_0.3mm_total - gold_0.6mm_total)/gold_0.3mm_total
-mb_bravo_diff = (gold_0.3mm_total - mb_bravo_total)/gold_0.3mm_total
-mb_0.6mm_diff = (gold_0.3mm_total - mb_0.6mm_total)/gold_0.3mm_total
+# calculate percent error from the gold standard measured at 0.3mm resolution
+pe_gold0.6mm = (gold_0.6mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
+pe_mbbravo   = (mb_bravo_total - gold_0.3mm_total)/gold_0.3mm_total * 100
+pe_mb0.6mm   = (mb_0.6mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
 
 subjects = gold_0.6mm[,1]
-mb0.6mm_diff_over_gold_diff = cbind(subjects, mb_0.6mm_diff - mean(gold_diff) , res="diffdff")
-gold_diff     = cbind(subjects,gold_diff, res="0.3mm gold & 0.6mm gold")
-mb_bravo_diff = cbind(subjects,mb_bravo_diff, res="0.3mm gold & MAGeT-on-BRAVO")
-mb_0.6mm_diff = cbind(subjects,mb_0.6mm_diff, res="0.3mm gold & MAGeT-at-0.6mm")
-
-diff = rbind(mb_bravo_diff,mb_0.6mm_diff,gold_diff, mb0.6mm_diff_over_gold_diff)
-names(diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
-
-melted = melt(diff, id.vars=c("subject", "res"))
+pe_gold0.6mm = cbind(subjects,pe_gold0.6mm, res="Resampling to 0.6mm")
+pe_mbbravo   = cbind(subjects,pe_mbbravo,   res="MAGeT-on-BRAVO")
+pe_mb0.6mm   = cbind(subjects,pe_mb0.6mm,   res="MAGeT-at-0.6mm")
+
+pe = rbind(pe_mbbravo,pe_mb0.6mm, pe_gold0.6mm)
+names(pe) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
+
+melted = melt(pe, id.vars=c("subject", "res"))
+names(melted) <- c("subject","res", "region","percenterr")
+ggplot(melted, aes(y=percenterr,x=region)) +
+  geom_boxplot(aes(colour=res)) +
+  labs(title="Percent error in volumes under different measurement regimes",
+       y="Percent error in volume", 
+       x="Region") 
+@
+
+<<WAval-error,cach=T>>=
+# compare % error between MAGeT and resampling
+mb0.6mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.6mm_diff), res="diffdff")
+names(mb0.6mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
+melted = melt(mb0.6mm_diff_over_gold_diff, id.vars=c("subject", "res"))
 names(melted) <- c("subject","res", "region","voldiff")
 ggplot(melted, aes(y=voldiff,x=region)) +
-  geom_boxplot() + facet_grid( . ~ res)  + 
-  labs(title="Percent difference in MAGeT volumes from original",
-       y="Percent difference in volume", 
+  geom_boxplot() +
+  labs(title="Absolute difference in % difference between MAGeT and resampling",
+       y="Absolute difference in % difference in volume", 
        x="Region")
 @
-
 <<WAval-kappa,cache=T>>=

### 2013-03-20 paper.Rnw
Updated figures after presentation to the lab

@@
 <<ADN1-Xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ggplot(all_data, aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
+ggplot(subset(all_data, reg_method=="ANTS"), aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
   facet_grid(reg_method~method.y) + 
-  #stat_summary(fun.y=mean,geom='line',
-  #             aes(y=k.y)) + 
-  stat_smooth(method='lm', formula=y~log(x)) + 
+  stat_summary(fun.y=mean,geom='line',
+               aes(y=k.y, weight=1)) + 
+  #stat_smooth(method='lm', formula=y~log(x)) + 
   scale_colour_hue(name="Atlases") +
@@
 <<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote"), 
+ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote" &
+                templates.y %in% seq(1,20,2)), 
        aes(x=templates.y, y=k_diff, colour=as.factor(atlases))) + 
@@
                aes(y=k_diff,colour=as.factor(atlases))) + 
+  stat_summary(fun.y=mean, geom="line", 
+               aes(y=k_diff,colour=as.factor(atlases))) + 
   #stat_smooth(method=lm, formula=y~log(x)) + 
@@
   coord_cartesian(ylim=c(-0.1,0.05)) + 
-  scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
+  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
   scale_colour_hue(name="Atlases") +
@@
 
-stddev = ddply(ANTSmajvote, c("subject", "label","atlases","templates.y"), function (df) {
+data.stats = ddply(ANTSmajvote, c("subject", "label","atlases","templates.y"), function (df) {
   data.frame(
-    MA = sd(df$k.x),
-    MB = sd(df$k.y) 
+    MA.sd  =  sd(df$k.x),
+    MB.sd  =  sd(df$k.y),
+    MA.var = var(df$k.x),
+    MB.var = var(df$k.y)    
   )
@@
 
-stddev.t.test = ddply(stddev, c("atlases","templates.y"), function (df) {
-  t = t.test(df$MA, df$MB)
+stats.tests = ddply(data.stats, c("atlases","templates.y"), function (df) {
+  t.var = t.test(df$MA.var, df$MB.var)
+  t.sd  = t.test(df$MA.sd , df$MB.sd)
   data.frame(
-    statistic = t$statistic, 
-    p.value = t$p.value
+    stat      = c("var", "sd"),
+    statistic = c(t.var$statistic, t.sd$statistic),
+    p.value   = c(t.var$p.value, t.sd$p.value)
   )
@@
 
-ggplot(stddev.t.test, 
+ggplot(stats.tests, 
        aes(x=templates.y, y=p.value, colour=as.factor(atlases))) + 
+  facet_grid( . ~ stat) + 
   geom_line() + 
@@
   scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
-  scale_y_log10() + 
   xlab( "Number of Templates" ) + 
   ylab( "p-value" )
+
 @
 \end{figure}
+
+<<ADN1-Xval-sd-by-subject, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+SDs = data.stats[c('subject','label','atlases','templates.y','MA.sd','MB.sd')]
+SDs = subset(SDs, templates.y %in% c(5,10,15,20) & atlases == 5)
+
+melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
+ggplot(melted, aes(x=templates.y, y=value, 
+                   group=interaction(templates.y, variable), colour=as.factor(variable))) + 
+  scale_x_continuous(breaks=c(5,10,15,20)) + 
+  geom_violin() 
+@
 

### 2013-03-28 paper.Rnw
Updates to the results section...

--- 2013-03-19 paper.Rnw
@@
 <<setup, include=FALSE, cache=FALSE>>=
-opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
-    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
+#opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
+#    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
 library(ggplot2)
@@
 %                                                                              %
- % Bland-Altman plot for MA and MAGeT
+
+% TODO: Bland-Altman plot for MA and MAGeT
+
+Hippocampal MAGeT-based segmentations using both ANIMAL and ANTs registration
+algorithm demonstrate good overlap with manually derived gold-standards ( Figure
+\ref{fig:ADNI1-xval-k-mean}, \todo{table?}) using each label fusion method.
+Qualitatively, both ANIMAL and ANTs-based segmentations demonstrate trend
+overlap accuracy that increases with the size of atlas library and template
+library, with at least 12 templates needed for optimal results
+\todo{quantify?}).  No significant difference in segmentation accuracy is seen
+when either ANIMAL or ANTs registration is used with different label fusion
+techniques, at any atlas or template library sizes (\todo{lm of interaction?).
+In every parameter configuration, the use of MAGeT Brain with ANTs registration
+shows increased segmentation accuracy over MAGeT Brain with ANIMAL registration
+(\todo{interaction?}).
+
+It is interesting to note that with an even number of templates, MAGeT brain
+shows a small decrease in performance relative to when one fewer template image
+is used (\todo{show significance of this}). See section (\todo{ref}) for a
+discussion of this behaviour. In the remainder of this section, only results
+from odd-sized template libraries will be shown.
 
 \begin{figure}
-<<ADN1-Xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ggplot(all_data, aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
+<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+
+ggplot(subset(all_data, method == 'majvote'), 
+       aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
   facet_grid(reg_method~method.y) + 
@@
 \end{figure}
+
+With an increasing number of templates, MAGeT-based segmentating using ANTs
+registration and majority vote label fusion shows improvement in overlap
+accuracy over multi-atlas-based segmentation, using the same number of atlases
+and voting method Figure \todo{ref}. The magnitude of improvement over
+multi-atlas-based segmentation decreases with an increasing number of atlases,
+with accuracy converging with 7 atlases.  Peak improvement in MAGeT Brain
+accuracy (0.02 DSC) is found when one atlas is used with a template library of
+20 images. \todo{stats?}
 
@@
 
+In addition to an improvement in accuracy over multi-atlas-based segmentation,
+MAGeT Brain also shows a decrease in the variability of segmentation accuracy
+with different input atlases (Figure \todo{ref}).  Additionally, variability
+decreases with the size of template library, and is significantly decreased
+(p<0.05) when using 3-7 atlases and 17 or more templates.
+
 \begin{figure}
@@
   scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
-  scale_y_log10() + 
   xlab( "Number of Templates" ) + 
@@
 @
-  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
+  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS,
+  20 templates}
   \label{}
@@
 
-\begin{figure}
-<<ADN1-Xval-ANTS-20tmpls-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-a_20_tmpls = subset(all_data, 
-                      reg_method  == "ANTS" &
-                      method.y    == "Majority Vote" &
-                      templates.y == 20 &
-                      atlases     == 9 &
-                      volume.x < 10000)
-avg = ddply(a_20_tmpls, c("subject", "label"), function (df) {
-  data.frame(
-    pre = mean(df$volume.x),  #MA
-    post = mean(df$volume.y)  #MB
-  )
-})
-# TODO: against SNT
-
-#df = data.frame(pre=a_20_tmpls$volume.x, post=a_20_tmpls$volume.y)
-bland_altman_plot(avg) + 
-  xlab('Mean of MA and MAGeT HC Left/Right Volumes') + 
-  ylab('MA - MAGeT Volume') + 
-  opts(title = 'Bland-Altman Plot')
-@
-  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
-  \label{}
-\end{figure}
-
-\begin{figure}
-<<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-a_20_tmpls = subset(all_data, 
-                      reg_method  == "ANTS" &
-                      method.y    == "Majority Vote" &
-                      templates.y == 20 &
-                      atlases %in% c(1,9))
-template_library = read.csv('data/a2a_ants/template_library_composition.csv')
-template_library$subject = NULL
-a_20_tmpls = merge(a_20_tmpls, template_library,
-                   by.x=c('timestamp', 'templates.y'), 
-                   by.y=c('timestamp', 'templates'))
-ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
-  facet_grid( . ~ atlases) + 
-  geom_smooth(method='lm',fullrange=F) 
- # add 
-@
-  \caption{Mean kappa vs. proportion of disease category in template library by disease category}
-  \label{}
-\end{figure}
-
-- find significant improvement over multi-atlas performed with the same
-  parameters. Also, find smoothed performance is monotonically increasing but
-  asymptotic in size of both template and atlas library, with peak performance
-  reached after 15 templates.
-
-- True for all parameter settings! (so works generally..)
-\todo{Can we statistically capture what "peak" performance is?  Something like,
-the point at which gains become statistically insignificant?}
-
-Ideas:
-\begin{itemize}
-\item more atlases -> better performance
-\item  larger template library -> better performance, but tails off around 10-15 templates
-\item  no significant difference between majority or weighted vote methods (haven???t tested this statistically though). 
-\item  consistently performs better than average naive performance by XXX
-\item  using ANTS, with a large enough template library (>12) MAGeT brain performs
-  better than the average multi-atlas approach with the same number of atlases.
-  using ANIMAL, 5 or more atlases needed before boost seen. 
-\item  more atlases -> smaller template library required to improve on average multi-atlas performance
-\item  discuss variance?  best/worst case? --how often do we expect random template library selection to work decently
-\end{itemize}
-
-\todo{show cost (in registrations) / benefit tradec off graph:  show number of
-registrations per Kappa?  or hours of manual labour per Kappa?)}
+% \begin{figure}
+% <<ADN1-Xval-ANTS-20tmpls-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+% a_20_tmpls = subset(all_data, 
+%                       reg_method  == "ANTS" &
+%                       method.y    == "Majority Vote" &
+%                       templates.y == 20 &
+%                       atlases     == 9 &
+%                       volume.x < 10000)
+% avg = ddply(a_20_tmpls, c("subject", "label"), function (df) {
+%   data.frame(
+%     pre = mean(df$volume.x),  #MA
+%     post = mean(df$volume.y)  #MB
+%   )
+% })
+% # TODO: against SNT
+% 
+% #df = data.frame(pre=a_20_tmpls$volume.x, post=a_20_tmpls$volume.y)
+% bland_altman_plot(avg) + 
+%   xlab('Mean of MA and MAGeT HC Left/Right Volumes') + 
+%   ylab('MA - MAGeT Volume') + 
+%   opts(title = 'Bland-Altman Plot')
+% @
+%   \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
+%   \label{}
+% \end{figure}
+% 
+% 
+% \begin{figure}
+% <<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+% a_20_tmpls = subset(all_data, 
+%                       reg_method  == "ANTS" &
+%                       method.y    == "Majority Vote" &
+%                       templates.y == 20 &
+%                       atlases %in% c(1,9))
+% template_library = read.csv('data/a2a_ants/template_library_composition.csv')
+% template_library$subject = NULL
+% a_20_tmpls = merge(a_20_tmpls, template_library,
+%                    by.x=c('timestamp', 'templates.y'), 
+%                    by.y=c('timestamp', 'templates'))
+% ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
+%   facet_grid( . ~ atlases) + 
+%   geom_smooth(method='lm',fullrange=F) 
+%  # add 
+% @
+%   \caption{Mean kappa vs. proportion of disease category in template library by disease category}
+%   \label{}
+% \end{figure}
+% 
+% 
+% \todo{show cost (in registrations) / benefit tradec off graph:  show number of
+% registrations per Kappa?  or hours of manual labour per Kappa?)}
 % discuss run-time for each of MAGet and multi-atlas
@@
 %%%%%
-\subsection{ADNI-1 Screen Validation}
-
-Ideas: 
-\begin{itemize}
-\item  A2A shows that if atlas population strongly(?) represents subject set
-variability, then free choice from atlas population will produce improvements
-(we know this b/c of extensive validation trials). 
-\item  what about in the case where atlas population doesn???t strongly
-represent subject set variability (e.g. a priori atlas set)?  then, we can use
-atlas selection to refine atlas set? 
-\end{itemize}
-
-\todo{Kappa against our manual rater is low... explain that}
+\subsection{ADNI-1 Screening Validation}
+
+Based on the results from the ADNI1 Cross-Validation experiment, MAGeT Brain was
+configured with a template library of 20 subject images and majority vote label
+fusion.
+
+% \begin{itemize}
+% \item  A2A shows that if atlas population strongly(?) represents subject set
+%        variability, then free choice from atlas population will produce
+%        improvements (we know this b/c of extensive validation trials). 
+% \item  what about in the case where atlas population doesn???t strongly
+%        represent subject set variability (e.g. a priori atlas set)?  then, we
+%        can use atlas selection to refine atlas set? 
+% \end{itemize}
+% 
+% \todo{Kappa against our manual rater is low... explain that}
 
@@
 \begin{figure}[h]
-<<ADNI-scr-histogram,dependson='ADNI-baseline-volumes-boxplot',cache=T>>=
+<<ADNI-scr-histogram,cache=T>>=
 melted   = melt(totals.complete, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
@@
   facet_grid(variable ~ DX) + 
-  scale_colour_discrete(legend=FALSE) + 
-  scale_size_continuous(legend=FALSE) + 
+  scale_colour_discrete(guide="none") + 
+  scale_size_continuous(guide="none") + 
   labs(title="Total hippocampus volume by disease and measurement method", 
@@
        y="Frequency") + 
-  opts(axis.title.x = theme_text(vjust=-0.1))
-@
-  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes as measured by SNT, FSL, FreeSurfer (FS), MAPER, and MAGeT brain (MAGeT).
-A fitted normal curve is shown in red.
-$W$ is the Shapiro-Wilkes test statistic measuring normality of the data (signficance indicates non-normality).
-$D$ is the Kolmogorov-Smirnov test statistic measuring the goodness-of-fit between the distribution of measured volumes and SNT volumes (significance indicates a difference).}
+  theme(axis.title.x = element_text(vjust=-0.1))
+@
+  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes
+  as measured by SNT, FSL, FreeSurfer (FS), MAPER, and MAGeT brain (MAGeT). A
+  fitted normal curve is shown in red. $W$ is the Shapiro-Wilkes test statistic
+  measuring normality of the data (signficance indicates non-normality). $D$ is
+  the Kolmogorov-Smirnov test statistic measuring the goodness-of-fit between
+  the distribution of measured volumes and SNT volumes (significance indicates a
+  difference).}
   \label{ADNI-scr-histogram}
@@
 # a printable table out of them.  Sorry.
-melted = melt(totals.complete, measure.vars=c("SNT","FSL", "FS", "MAPER"))
-cor_by_dx    = by(melted[c("MAGeT", "value")], c(melted['variable'], melted['DX']), 
-                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
-cor_overall  = by(melted[c("MAGeT", "value")], melted['variable'], 
-                  function(x) cor(x$MAGeT,x$value, use="pairwise.complete.obs"))
+melted = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
+cor_by_dx    = by(melted[c("SNT", "value")], c(melted['variable'], melted['DX']), 
+                  function(x) cor(x$SNT,x$value, use="pairwise.complete.obs"))
+cor_overall  = by(melted[c("SNT", "value")], melted['variable'], 
+                  function(x) cor(x$SNT,x$value, use="pairwise.complete.obs"))
 cor_by_dx = t(rbind(cor_by_dx))
@@
       n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
-      )#file="", size="scriptsize")
+      file="", size="scriptsize")
 @
@@
 <<WAval-kappa,cache=T>>=
-data = read.csv('data/WAval-kappa.csv')
+#data = read.csv('data/WAval-kappa.csv')
 ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)

### 2013-03-28 paper.Rnw

@@
   facet_grid(reg_method~method.y) + 
-  #stat_summary(fun.y=mean,geom='line',
-  #             aes(y=k.y)) + 
-  stat_smooth(method='lm', formula=y~log(x)) + 
+  stat_summary(fun.y=mean,geom='line',
+               aes(y=k.y, weight=1)) + 
+  #stat_smooth(method='lm', formula=y~log(x)) + 
   scale_colour_hue(name="Atlases") +
@@
   ylab( "p-value" )
+
 @
 \end{figure}
+
+<<ADN1-Xval-sd-by-subject, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+SDs = data.stats[c('subject','label','atlases','templates.y','MA.sd','MB.sd')]
+SDs = subset(SDs, templates.y %in% c(5,10,15,20) & atlases == 5)
+
+melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
+ggplot(melted, aes(x=templates.y, y=value, 
+                   group=interaction(templates.y, variable), colour=as.factor(variable))) + 
+  scale_x_continuous(breaks=c(5,10,15,20)) + 
+  geom_violin() 
+@
 

### 2013-04-03 paper.Rnw
Small updates to the methods wording, Park QC added, bad a2a_tracc data added

--- 2013-03-28 paper.Rnw
@@
 
+\newcommand{\mb}{MAGeT Brain}
+
 %notes and TODO formatting
@@
 
-\title{Bootstrapping Multi-atlas Hippocampal Segmentation with MAGeT}
+\title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
 \author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., \\ 
@@
 
-Our algorith, called MAGeT brain ({\em M}ultiple {\em A}utomatically {\em
+Our algorith, called \mb ({\em M}ultiple {\em A}utomatically {\em
 Ge}nerated {\em T}emplates), is an extension to the basic multi-atlas-based
@@
 
-In previous work \citep{Chakravarty2012}, we applied MAGeT brain to segmentation
+In previous work \citep{Chakravarty2012}, we applied \mb to segmentation
 of the human striatum, globus pallidus, and thalamus using a single
@@
 rigorous validate the method. First, we conduct an extensive cross-validation of
-MAGeT and basic multi-atlas segmentation on a subset of the ADNI1 dataset to
-assess the accuracy of MAGeT brain under various parameter settings (number of
+\mb and basic multi-atlas segmentation on a subset of the ADNI1 dataset to
+assess the accuracy of \mb under various parameter settings (number of
 atlases and templates, registration and fusion methods). With the best
-performing parameter configuration discovered above, we estimate MAGeT
+performing parameter configuration discovered above, we estimate \mb
 intra-rater reliability by segmenting separately acquired T1 images of the
@@
 hippocampal subfield segmentations of five in-vivo high-resolution (300u
-isotropic) T1-weighted MR scans\citep{Winterburn2013}. To validate MAGeT in a
+isotropic) T1-weighted MR scans\citep{Winterburn2013}. To validate \mb in a
 real world situation, we segment the entire ADNI1 screening dataset and compare
 our segmentations to established automated and manual segmentations.
-Additionally, to ensure MAGeT brain accuracy across disease categories, we also
-compare MAGeT segmentations to manual segmentations of 139 first episode
+Additionally, to ensure \mb accuracy across disease categories, we also
+compare \mb segmentations to manual segmentations of 139 first episode
 schizophrenic patients.
@@
 
-\subsection{MAGeT Brain Algorithm}
+\subsection{\mb Algorithm}
 In this paper we will be making a distinction between an atlas and a template --
@@
 
-MAGeT brain is best understood as an extension of the basic multi-atlas
+\mb is best understood as an extension of the basic multi-atlas
 segmentation schema. Instead of using the atlas library to directly label the
@@
 labels from all templates. To summarize, figure \ref{alg:MAGeT} describes the
-MAGeT brain algorithm in pseudocode.
-
-Source code for MAGeT brain can be found at
+\mb algorithm in pseudocode.
+
+Source code for \mb can be found at
 \url{http://github.com/pipitone/MAGeTbrain}.
@@
   \scriptsize
-  \caption{Pseudocode for the MAGeT Brain algorithm}
+  \caption{Pseudocode for the \mb algorithm}
   \label{alg:MAGeT}
@@
 
-The following experiments were performed to assess the performance of MAGeT
-brain with various parameter settings as well as on diverse datasets.  In each
-experiment we contrast the performance of MAGeT brain with that standard single-
-and multi-atlas segmentations derived from the same atlas library.
+The following experiments were performed to assess the performance of \mb with
+various parameter settings as well as on diverse datasets.  In each experiment
+we contrast the performance of \mb with that standard single- and multi-atlas
+segmentations derived from the same atlas library.
 
@@
 \subsubsection{ADNI1 cross-validation}
-To test the effect of parameter settings on the MAGeT brain algorithm, Monte
+To test the effect of parameter settings on the \mb algorithm, Monte
 Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
 the ADNI1 subjects for validation.  This form of cross-validation (described in
-detail below) allows us to validate MAGeT with various atlas and template
+detail below) allows us to validate \mb with various atlas and template
 library sizes, registration and label fusion methods.
@@
 
-{\bf Atlas and template libary.}  Atlases consist of images taken from the
-dataset, with corresponding manual labels provided by SNT.  Atlas library size
-varied from 3 to 9 images.  The remaining 60-66 target images were segmented
-with the template library size varying from 3 to 20 images for each setting of
-atlas library size. Template library images were selected at random from the
-targets.
+{\bf Atlas and template libary.}  The atlas and template libraries are selected
+from the dataset evaluated.  Atlases consist of images taken from the
+dataset, with corresponding manual labels provided by SNT. Template library 
+images were selected at random from remaining images not used as atlases. The
+atlas library size ranged from 1 to 9 images. Template library size ranged from
+1 to 20 images.
 
@@
 weighted majority vote fusion methods, the number of top labels used in the
-fusion was varied from 3 to 20 images.
+fusion was varied from 1 to 20 images.
 
 {\bf Evaluation.}  Monte Carlo Cross-Validation (MCCV), also known as repeated
-random sub-sampling cross-validation, consists of repeated trials in which items
-from the dataset are randomly sampled (without replacement) and assigned to a
-training set or validation set \cite{Shao1993}.
-
-In each trial, the training set consists of images and SNT labels used as
-atlases, and the validation set consists of images to be segmented along with
-the corresponding SNT labels used as gold-standard segmentations. Both MAGeT
-brain and basic multi-atlas segmentation were performed in each trial, and the
-accuracy of the resulting segmentations were measured against the SNT labels.
-
-We performed 10 trials on each combination of parameters listed above: atlas
-library size (1-9), template library size (1-20), registration method (ANTS vs.
-ANIMAL), and label fusion method (MV, XC-WV, NMI-WV). A total of 1080 parameter
-settings were evaluated over a total of 10800 trials.
+random sub-sampling cross-validation, consists of repeated rounds of validation 
+in which items from the dataset are randomly sampled (without replacement) and 
+assigned to a training set or validation set \cite{Shao1993}.
+
+In this experiment, each validation round consists of the following steps. First, 
+an atlas library is selected (the training set) and, from the remaining images, 
+a template library and subject to be segmented is chosen (the validation set). 
+Second, \mb and basic multi-atlas segmentation are performed on the subject using the 
+selected atlas and template libraries. Third, the accuracy of the resulting 
+segmentations are measured against the SNT labels.
+
+A total of 10 validation rounds are perfomed on each subject in the dataset, for each 
+combination of parameter settings: atlas library size (1-9), template library 
+size (1-20), registration method (ANTS vs. ANIMAL), and label fusion method 
+(MV, XC-WV, NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were
+conducted, resulting in a total of 1'490'400 segmentations analysed.
 
@@
 \subsubsection{ADNI-1 Screening Validation}
-To test the accuracy of MAGeT brain on a real-world task we segment the entire
+To test the accuracy of \mb on a real-world task we segment the entire
 ADNI-1 dataset using an atlas set that is not representative of the target set.
@@
 Winterburn atlases, this poses a problem for direct similarity comparisons
-between labels produced by MAGeT brain and the ADNI labels.  To evaluate the
-performance of MAGeT brain, we compare the correlation of MAGeT segmentation
+between labels produced by \mb and the ADNI labels.  To evaluate the
+performance of \mb, we compare the correlation of \mb segmentation
 volumes with manual segmentation (SNT) volumes. Additionally, we correlate the
-hippocampal volumes of established automated segmentation methods to MAGeT
+hippocampal volumes of established automated segmentation methods to \mb
 segmentations.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 X-Validation Dataset Demographics Table
+% ADNI1 Screening Template Library Demographics Table
 %
@@
 
-{\bf Dataset evaluated.} To validate that MAGeT performance generalises to
-across disease categories, we measure the performance of MAGeT on the
+{\bf Dataset evaluated.} To validate that \mb performance generalises to
+across disease categories, we measure the performance of \mb on the
 Schizophrenic First Episode Patient (SFEP) dataset.
@@
 {\bf Evaluation.}  - Leave one out cross-validation (LOOCV) in which all five
-subjects are segmented in separate runs of MAGeT brain.  In each run, the subject
+subjects are segmented in separate runs of \mb.  In each run, the subject
 to be segmented is excluded from the Atlas library (so only four atlases are used).
@@
 %                                                                              %
-<<ADN1-Xval-prep, include=FALSE, cache=TRUE>>=
+<<ADN1-xval-prep, include=FALSE, cache=TRUE>>=
 # all data from the validation trials
-t_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
+a_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
+#t_data       = read.csv(gzfile('data/a2a_tracc/xfmtrials_results_2013-04-01.csv.gz'))
+t_data       = read.csv(gzfile('data/a2a_tracc/results_2012_12_31.csv.gz'))
+diagnoses    = read.csv('data/a2a_diagnoses.csv')
+
+# simplifies the table
 t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
-a_data       = t_data[,] #read.csv(gzfile('data/a2a_tracc/results_2013_01_03.csv.gz'))
-diagnoses    = read.csv('data/a2a_diagnoses.csv')
-
-# simplifies the table
+a_data       = subset(a_data, atlases %in% c(1,3,5,7,9))
 t_data$se    = NULL
@@
 # separate MA majvote from MB so we can compute differences
+# we only compare multiatlas using majority vote
 t_ma = subset(t_data, approach=="multiatlas" & method=="majvote")
@@
 
-all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"))
-all_data$k_diff = all_data$k.y - all_data$k.x 
-method_labels <- list("Majority Vote" = "majvote", 
+# aggregate over batches
+ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases + templates + top_n + label, data = ma, mean)
+mb_mean = aggregate(k ~ subject + reg_method + approach + method + atlases + templates + top_n + label, data = mb, mean)
+
+all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"), 
+                 suffixes=c(".ma", ".mb"))
+
+all_data_mean = merge(ma_mean, mb_mean, by=c("reg_method", "label", "atlases", "subject"), 
+                 suffixes=c(".ma", ".mb"))
+
+all_data$k_diff = all_data$k.mb - all_data$k.ma
+all_data_mean$k_diff = all_data_mean$k.mb - all_data_mean$k.ma
+
+vote_levels <- list("Majority Vote" = "majvote", 
                       "NMI Vote" = "nmi", 
                       "Cross-correlation Vote" = "xcorr")
-levels(all_data$method.y) <- method_labels
+levels(all_data$method.mb) <- vote_levels
+levels(all_data_mean$method.mb) <- vote_levels
+
 all_data       = merge(all_data, diagnoses)
-remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
+all_data_mean  = merge(all_data_mean, diagnoses)
+#remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
 @
@@
 
-% TODO: Bland-Altman plot for MA and MAGeT
-
-Hippocampal MAGeT-based segmentations using both ANIMAL and ANTs registration
-algorithm demonstrate good overlap with manually derived gold-standards ( Figure
-\ref{fig:ADNI1-xval-k-mean}, \todo{table?}) using each label fusion method.
-Qualitatively, both ANIMAL and ANTs-based segmentations demonstrate trend
-overlap accuracy that increases with the size of atlas library and template
-library, with at least 12 templates needed for optimal results
-\todo{quantify?}).  No significant difference in segmentation accuracy is seen
-when either ANIMAL or ANTs registration is used with different label fusion
-techniques, at any atlas or template library sizes (\todo{lm of interaction?).
-In every parameter configuration, the use of MAGeT Brain with ANTs registration
-shows increased segmentation accuracy over MAGeT Brain with ANIMAL registration
-(\todo{interaction?}).
-
-It is interesting to note that with an even number of templates, MAGeT brain
-shows a small decrease in performance relative to when one fewer template image
-is used (\todo{show significance of this}). See section (\todo{ref}) for a
-discussion of this behaviour. In the remainder of this section, only results
-from odd-sized template libraries will be shown.
+%
+% synonyms for significant: 
+% powerful, suggestive, apparent, definite, evident, large, marked, material, 
+% noticeable, plain, pronounced, substantial, substantive, 
+%
+% antonyms: 
+% negligible, minor, no appreciable, scant
+
+In this experiment we conducted 10 rounds of \mb and multi-atlas segmentation of 
+each of 69 subjects, over a range of atlas and template library sizes, with ANTS 
+and ANIMAL registration algorithms, and three kinds of label fusion techniques. 
+Hippocampal \mb-based segmentations using both ANIMAL and ANTs registration
+algorithm demonstrate good overlap with manually derived gold-standards (Figure
+\ref{fig:ADNI1-xval-k-mean}). Qualitatively, both ANIMAL and ANTs-based 
+segmentations demonstrate trend overlap accuracy that increases with the size of 
+atlas library and template library. With 12 or more templates \mb library of at 
+least 12 templates
+
+No marked difference in segmentation accuracy is seen when either ANIMAL or ANTs
+registration is used with different label fusion techniques, at any atlas or
+template library sizes.  In every parameter configuration, the use of \mb with 
+ANTs registration shows increased segmentation accuracy over \mb with ANIMAL registration (\todo{interaction?}).
+
+It is interesting to note that with an even number of templates, \mb shows a
+small decrease in performance relative to when one fewer template image is used
+(\todo{show significance of this}). See section (\todo{ref}) for a discussion of
+this behaviour. In the remainder of this section, only results from odd-sized
+template libraries will be shown.
 
@@
 
-ggplot(subset(all_data, method == 'majvote'), 
-       aes(x=templates.y,y=k.y, colour=as.factor(atlases))) + 
-  facet_grid(reg_method~method.y) + 
-  #stat_summary(fun.y=mean,geom='line',
-  #             aes(y=k.y)) + 
-  stat_smooth(method='lm', formula=y~log(x)) + 
+ggplot(all_data_mean, 
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
+  facet_grid(reg_method~method.mb) + 
+  stat_summary(fun.y=mean,geom='line',
+               aes(y=k.mb, weight=1)) + 
+  #stat_smooth(method='lm', formula=y~log(x)) + 
   scale_colour_hue(name="Atlases") +
@@
 @
-  \caption{MAGeT brain mean DSC by atlas and template library size}
+  \caption{\mb mean DSC by atlas and template library size}
   \label{fig:ADNI1_xval-k_mean}
@@
 
-With an increasing number of templates, MAGeT-based segmentating using ANTs
+With an increasing number of templates, \mb-based segmentating using ANTs
 registration and majority vote label fusion shows improvement in overlap
@@
 multi-atlas-based segmentation decreases with an increasing number of atlases,
-with accuracy converging with 7 atlases.  Peak improvement in MAGeT Brain
-accuracy (0.02 DSC) is found when one atlas is used with a template library of
-20 images. \todo{stats?}
+with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (0.02
+DSC) is found when one atlas is used with a template library of 20 images.
+\todo{stats?}
 
@@
 <<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ggplot(subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote"), 
-       aes(x=templates.y, y=k_diff, colour=as.factor(atlases))) + 
+ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
+                templates.mb %in% seq(1,20,2)), 
+       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
   #facet_grid(reg_method~method.y) + 
@@
 @
-  \caption{Increase in mean Kappa between MAGeT brain and multi-atlas}
+  \caption{Increase in mean Kappa between \mb and multi-atlas}
   \label{ADNI1_cross_validation_MA_difference}
@@
 In addition to an improvement in accuracy over multi-atlas-based segmentation,
-MAGeT Brain also shows a decrease in the variability of segmentation accuracy
+\mb also shows a decrease in the variability of segmentation accuracy
 with different input atlases (Figure \todo{ref}).  Additionally, variability
@@
 <<ADN1-Xval-k-sd-t-test-p-plot, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-    ANTSmajvote = subset(all_data, reg_method=="ANTS" & method.y=="Majority Vote")
-
-stddev = ddply(ANTSmajvote, c("subject", "label","atlases","templates.y"), function (df) {
+ANTSmajvote = subset(all_data, reg_method=="ANTS" & method.mb=="Majority Vote")
+
+data.stats = ddply(ANTSmajvote, c("subject", "label","atlases","templates.mb"), function (df) {
   data.frame(
-    MA = sd(df$k.x),
-    MB = sd(df$k.y) 
+    MA.sd  =  sd(df$k.ma),
+    MB.sd  =  sd(df$k.mb),
+    MA.var = var(df$k.ma),
+    MB.var = var(df$k.mb)
   )
@@
 
-stddev.t.test = ddply(stddev, c("atlases","templates.y"), function (df) {
-  t = t.test(df$MA, df$MB)
+stats.tests = ddply(data.stats, c("atlases","templates.mb"), function (df) {
+  t.var = t.test(df$MA.var, df$MB.var)
+  t.sd  = t.test(df$MA.sd , df$MB.sd)
   data.frame(
-    statistic = t$statistic, 
-    p.value = t$p.value
+    stat      = c("var", "sd"),
+    statistic = c(t.var$statistic, t.sd$statistic),
+    p.value   = c(t.var$p.value, t.sd$p.value)
   )
@@
 
-ggplot(stddev.t.test, 
-       aes(x=templates.y, y=p.value, colour=as.factor(atlases))) + 
+ggplot(stats.tests, 
+       aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
+  facet_grid( . ~ stat) + 
   geom_line() + 
@@
   scale_linetype_discrete(guide='none') + 
-  scale_x_continuous(breaks=1:20) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
+  scale_x_continuous(breaks=seq(1,20,by=2)) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
   xlab( "Number of Templates" ) + 
   ylab( "p-value" )
+
 @
 \end{figure}
+
+<<ADNI1-Xval-sd-by-subject, cache=TRUE>>=
+SDs = data.stats[c('subject','label','atlases','templates.mb','MA.sd','MB.sd')]
+SDs = subset(SDs, templates.mb %in% c(5,10,15,20) & atlases == 5)
+
+melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
+ggplot(melted, aes(x=templates.mb, y=value, 
+                   group=interaction(templates.mb, variable), colour=as.factor(variable))) + 
+  scale_x_continuous(breaks=c(5,10,15,20)) + 
+  geom_violin() 
+@
 
@@
 <<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-a_20_tmpls = subset(all_data, 
+a_20_tmpls = subset(all_data_mean, 
                       reg_method  == "ANTS" &
-                      method.y    == "Majority Vote" &
-                      templates.y == 20)
+                      method.mb   == "Majority Vote" &
+                      templates.mb== 20)
 ggplot(a_20_tmpls, aes(x=factor(atlases))) +
-  stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.x)) +
-  stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.x,group=1)) +
-  stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.y)) +
-  stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.y,group=1)) +
+  #geom_point(aes(y=k.mb,group=1,colour="MAGeT"),position="jitter") + 
+  #geom_point(aes(y=k.ma,group=2,colour="MA"),position="jitter") + 
+  geom_violin(aes(y=k.mb,group=factor(atlases),colour="MAGeT")) + 
+  geom_violin(aes(y=k.ma,group=factor(atlases),colour="MA")) + 
+  geom_smooth(aes(y=k.mb,group=1,colour="MAGeT")) +
+  geom_smooth(aes(y=k.ma,group=2,colour="MA")) +
+
+  #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
+  #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
+  #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
+  #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
   labs(x="Number of Atlases", y = "Kappa", 
@@
 @
-  \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS,
+  \caption{Change in mean Kappa between multi-atlas and \mb using ANTS,
   20 templates}
@@
                       reg_method  == "ANTS" &
-                      method.y    == "Majority Vote" &
-                      templates.y == 20 &
+                      method.mb    == "Majority Vote" &
+                      templates.mb == 20 &
                       atlases     == 9 & 
-                      volume.x < 10000)
+                      volume.ma < 10000)
 stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
   data.frame(
-    MA = sd(df$k.x),  #MA
-    MB = sd(df$k.y)  #MB
+    MA = sd(df$k.ma),  #MA
+    MB = sd(df$k.mb)  #MB
   )
@@
 % @
-%   \caption{Change in mean Kappa between multi-atlas and MAGeT brain using ANTS, 20 templates}
+%   \caption{Change in mean Kappa between multi-atlas and \mb using ANTS, 20 templates}
 %   \label{}
@@
 
-Based on the results from the ADNI1 Cross-Validation experiment, MAGeT Brain was
-configured with a template library of 20 subject images and majority vote label
-fusion.
+Based on the results from the ADNI1 Cross-Validation experiment, in this
+experiment \mb was configured with a template library of 21 randomly
+chosen subject images (7 from each disease class) and used majority vote label
+fusion. Overall co
 
@@
 <<ADNI-src-volumes-prep,echo=F,cache=T>>=
-bl1.5T= read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
-bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
+#bl1.5T= read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
+#bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
+yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
@@
 fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
-
+qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")
 # recode VISCODE
@@
 #levels(yr1$VISCODE) <-viscode_levels
-bl1.5T$VISCODE <- 'bl'
-bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)
+#bl1.5T$VISCODE <- 'bl'
+#bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)
 
 #TODO: include Scaled_2? 
-subjects = bl1.5T
+subjects = subset(yr1,Visit ==1)
+subjects$VISCODE = 'bl'
 
@@
                            FSL_R=fsl$X53, Source=fsl$Source)
+qc_pruned     = qc[,c("Image.Data.ID","MAGeT_QC","FSL_QC")]
+qc_pruned     = subset(qc_pruned, MAGeT_QC == 1 & FSL_QC == 1)
 
@@
 combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)
+combined = merge(combined, qc_pruned   , by=c("Image.Data.ID"))
 
@@
 @
-  \caption{Comparison of HC volumes by FreeSurfer (FSF), MAGeT brain (MAGeT), MAPER, and manual (SNT).}
+  \caption{Comparison of HC volumes by FreeSurfer (FSF), \mb (MAGeT), MAPER, and manual (SNT).}
   \label{ADNI-scr-volumes-boxplot}
@@
   \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes
-  as measured by SNT, FSL, FreeSurfer (FS), MAPER, and MAGeT brain (MAGeT). A
+  as measured by SNT, FSL, FreeSurfer (FS), MAPER, and \mb (MAGeT). A
   fitted normal curve is shown in red. $W$ is the Shapiro-Wilkes test statistic
@@
   \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
-  HC volumes for MAGeT against manually rated Hippocampal volumes}
+  HC volumes for \mb against manually rated Hippocampal volumes}
   \label{SZ_volumes}
@@
 @
-<<WAval-kappa,cache=T>>=
-#data = read.csv('data/WAval-kappa.csv')
-ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)
-@
+
+%<<WAval-kappa,cache=T>>=
+%#data = read.csv('data/WAval-kappa.csv')
+%ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)
+%@
 

### 2013-04-04 paper.Rnw
Fixed QC results, and added a table of segmentation totals for ADNI1 Screening Dataset

@@
 \newcommand{\mb}{MAGeT Brain}
+\newcommand{\ants}{ANTS}
+\newcommand{\animal}{ANIMAL}
 
@@
 atlas library of 80 T1-weighted atlas images from the ICBM152 dataset, the
-ANIMAL nonlinear registration algorithm, normalised mutual information as an
+\animal nonlinear registration algorithm, normalised mutual information as an
 atlas selection similarity metric, and majority vote for label fusion
@@
 \subsubsection{Automatic Normalization and Image Matching and Anatomical
-Labeling (ANIMAL)}
-
-The ANIMAL algorithm carries out image registration is in two phases. In the
+Labeling (\animal)}
+
+The \animal algorithm carries out image registration is in two phases. In the
 first, a 12-parameter linear transformation (3 translations, rotations, scales,
@@
 brain \citep{Collins}. In the second phase, nonlinear registration is completed
-using the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that
+using the \animal algorithm \citep{Collins1995}: an iterative procedure that
 estimates a 3D deformation field between two MR images. At first, large
@@
 \scriptsize
-\caption{ANIMAL Registration Parameters\label{tab:ANIMAL_params}} 
+\caption{\animal Registration Parameters\label{tab:ANIMAL_params}} 
 \begin{center}
@@
 
-\subsubsection{Automatic Normalization Tools (ANTS)}
-
-ANTs is a diffeomorphic registration algorithm which provides great flexibility
+\subsubsection{Automatic Normalization Tools (\ants)}
+
+\ants is a diffeomorphic registration algorithm which provides great flexibility
 over the choice of transformation model, objective function, and the consistency
@@
 is subsampled to a finer grid). The deformation field and the objective function
-are regularized with a Gaussian kernel at each level of the hierarchy. The ANTs
+are regularized with a Gaussian kernel at each level of the hierarchy. The \ants
 algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an
-implementation of the ANTS algorithm compatible with the MINC data format,
+implementation of the \ants algorithm compatible with the MINC data format,
 mincANTS \url{https://github.com/vfonov/mincANTS}.
 
-We used the following command line when running ANTS:
+We used the following command line when running \ants:
 {\small
@@
 These settings were adapted from the "reasonable starting point" given in the
-ANTS manual 
+\ants manual 
 \footnote{\url{https://sourceforge.net/projects/advants/files/Documentation/}}.
@@
 
-The {\tt xcorr\_vol} utility from the ANIMAL toolkit is used to calculate the
+The {\tt xcorr\_vol} utility from the \animal toolkit is used to calculate the
 cross-correlation similarity measure.
@@
 Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
-the ADNI1 subjects for validation.  This form of cross-validation (described in
+the ADNI1 dataset for validation.  This form of cross-validation (described in
 detail below) allows us to validate \mb with various atlas and template
@@
 
-{\bf Registration method.} Both the ANTS and ANIMAL registration methods were
+{\bf Registration method.} Both the \ants and \animal registration methods were
 used.
@@
 
-A total of 10 validation rounds are perfomed on each subject in the dataset, for each 
-combination of parameter settings: atlas library size (1-9), template library 
-size (1-20), registration method (ANTS vs. ANIMAL), and label fusion method 
-(MV, XC-WV, NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were
-conducted, resulting in a total of 1'490'400 segmentations analysed.
+A total of ten validation rounds are perfomed on each subject in the dataset,
+for each combination of parameter settings. Reported subject segmentation
+accuracy is the averaged over the ten validation rounds. The parameter settings
+we explore are: atlas library size (1-9), template library size (1-20),
+registration method (\ants vs.  \animal), and label fusion method (MV, XC-WV,
+NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were conducted,
+resulting in a total of 1'490'400 segmentations analysed.
 
@@
 
-{\bf Registration method.} The ANTS non-linear registration algorithm was used,
+{\bf Registration method.} The \ants non-linear registration algorithm was used,
 as it performed best in the ADNI1 cross-validation experiment (see Results).
@@
 Since the hippocampal segmentation protocols differ between the ADNI labels and
-Winterburn atlases, this poses a problem for direct similarity comparisons
-between labels produced by \mb and the ADNI labels.  To evaluate the
-performance of \mb, we compare the correlation of \mb segmentation
+Winterburn atlases, this poses a problem for direct evaluation between labels
+produced by \mb and the ADNI labels in terms of overlap; we would not expect
+different segmentation protocols to have a high degree of overlap.  Instead, to
+evaluate the performance of \mb we compare the correlation of \mb segmentation
 volumes with manual segmentation (SNT) volumes.  Additionally, we correlate the
@@
 segmentations.
-
-Additionally, we compared classification accuracy of subjects by diagnosis based
-on hippocampal volume using both the SNT labels and our produced labels.
-
-\todo{explain why we did(n't) resegment the ADNI images with the low-res
-protocol and compare directly like that.}
-
-\todo{Also: group differences, expecting separation}
 
@@
 
-{\bf Registration method.} ANTS.
+{\bf Registration method.} \ants.
 \todo{complete}
@@
 
-{\bf Registration method.} ANTS
+{\bf Registration method.} \ants
 \todo{complete}
@@
 %%%%%
-\subsection{ADNI-1 Cross-Validation}
+\subsection{ADNI1 Cross-Validation}
 
@@
 In this experiment we conducted 10 rounds of \mb and multi-atlas segmentation of
-each of 69 subjects, over a range of atlas and template library sizes, with ANTS 
-and ANIMAL registration algorithms, and three kinds of label fusion techniques. 
-Hippocampal \mb-based segmentations using both ANIMAL and ANTs registration
-algorithm demonstrate good overlap with manually derived gold-standards (Figure
-\ref{fig:ADNI1-xval-k-mean}). Qualitatively, both ANIMAL and ANTs-based 
+each of 69 subjects at a range of atlas and template library sizes, registration
+algorithm (\ants or \animal), and three label fusion techniques.  Hippocampal
+\mb-based segmentations using both \animal and \ants registration algorithm
+demonstrate good overlap with manually derived gold-standards (Figure
+\ref{fig:ADNI1-xval-k-mean}). Qualitatively, both \animal and \ants-based
 segmentations demonstrate trend overlap accuracy that increases with the size of
@@
 
-No marked difference in segmentation accuracy is seen when either ANIMAL or ANTs
+No marked difference in segmentation accuracy is seen when either \animal or \ants
 registration is used with different label fusion techniques, at any atlas or
 template library sizes.  In every parameter configuration, the use of \mb with
-ANTs registration shows increased segmentation accuracy over \mb with ANIMAL registration (\todo{interaction?}).
+\ants registration shows increased segmentation accuracy over \mb with \animal
+registration. In the remainder of this section, only results using the \ants
+registration algorithm will be shown.
 
 It is interesting to note that with an even number of templates, \mb shows a
-small decrease in performance relative to when one fewer template image is used
-(\todo{show significance of this}). See section (\todo{ref}) for a discussion of
-this behaviour. In the remainder of this section, only results from odd-sized
-template libraries will be shown.
+small decrease in performance relative to when one fewer template image is used.
+See section (\todo{ref}) for a discussion of this behaviour. In the remainder of
+this section, only results from odd-sized template libraries will be shown.
 
@@
 @
-  \caption{\mb mean DSC by atlas and template library size}
-  \label{fig:ADNI1_xval-k_mean}
+  \label{fig:ADNI1-xval-k_mean}
+  \caption{
+  {\em \mb mean DSC by atlas and template library size, registration algorithm,
+  and label fusion method. Mean DSC is calculated over each subject from the
+  ADNI1 Cross-Validation } 
+  }
 \end{figure}
 
-With an increasing number of templates, \mb-based segmentating using ANTs
+With an increasing number of templates, \mb-based segmentating using \ants
 registration and majority vote label fusion shows improvement in overlap
 accuracy over multi-atlas-based segmentation, using the same number of atlases
-and voting method Figure \todo{ref}. The magnitude of improvement over
+and voting method (Figure \todo{fig:ADNI1-xva-k-diff}. The magnitude of improvement over
 multi-atlas-based segmentation decreases with an increasing number of atlases,
-with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (0.02
+with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (~0.02
 DSC) is found when one atlas is used with a template library of 20 images.
 \todo{stats?}
+
+\todo{mention failures}
 
@@
   \caption{Increase in mean Kappa between \mb and multi-atlas}
-  \label{ADNI1_cross_validation_MA_difference}
+  \label{ADNI1-xval-k-diff}
 \end{figure}
@@
 @
-  \caption{Change in mean Kappa between multi-atlas and \mb using ANTS,
+  \caption{Change in mean Kappa between multi-atlas and \mb using \ants,
   20 templates}
@@
 %%%%%
-\subsection{ADNI-1 Screening Validation}
+\subsection{ADNI1 Screening Validation}
 
@@
 <<ADNI-src-volumes-prep,echo=F,cache=T>>=
-#bl1.5T= read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
+bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
 #bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
-yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+#yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
@@
 snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
-ucsd  = read.csv("data/ADNI_baseline_volumes/UCSDVOL.csv")
 fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
@@
 #levels(yr1$VISCODE) <-viscode_levels
-#bl1.5T$VISCODE <- 'bl'
 #bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)
 
-#TODO: include Scaled_2? 
-subjects = subset(yr1,Visit ==1)
+#subjects = subset(yr1,Visit ==1)
+subjects = bl
 subjects$VISCODE = 'bl'
@@
                            SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
-ucsd_pruned   = data.frame(RID=ucsd$RID, VISCODE=ucsd$VISCODE, 
-                           UCSD_L=ucsd$LHIPPOC, UCSD_R=ucsd$RHIPPOC)
-fs_pruned     = subset(fs, TEMPQC != "Fail")
-fs_pruned     = data.frame(FS_L=fs_pruned$ST29SV, 
-                           FS_R=fs_pruned$ST88SV, Source = fs_pruned$IMAGEUID)
+fs_pruned     = data.frame(TEMPQC =fs$TEMPQC, 
+                           FS_L=fs$ST29SV, 
+                           FS_R=fs$ST88SV, 
+                           Source = fs$IMAGEUID)
 fsl_pruned    = data.frame(FSL_L=fsl$X17, 
                            FSL_R=fsl$X53, Source=fsl$Source)
-qc_pruned     = qc[,c("Image.Data.ID","MAGeT_QC","FSL_QC")]
-qc_pruned     = subset(qc_pruned, MAGeT_QC == 1 & FSL_QC == 1)
 
 # diagnoses
-dx = adnimerge[c("DX.bl", "RID", "VISCODE")]
+dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
 
@@
 combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-#combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
 combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
@@
 combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)
-combined = merge(combined, qc_pruned   , by=c("Image.Data.ID"))
-
-attach(combined)
+combined = merge(combined, qc          , by=c("Image.Data.ID"), all.x=TRUE)
+
+# A table showing the number of images (and failures) from each package
+#           Total  MAGeT  FSL  FS   MAPER
+# Images    
+# Failures
+# 
+# Total number complete images =
+
+# Step 2: Of those, get counts for the numbers that fail QC
+package_totals = data.frame(
+  Screening  = c(length(subjects$VISCODE), 'N/A'),
+  SNT        = c(length(which(!is.na(combined$SNT_L)))    , "--"),
+  MAGeT      = c(length(which(!is.na(combined$MAGeT_L)))  , length(which(combined$MAGeT_QC==0))), 
+  MAPER      = c(length(which(!is.na(combined$MAPER_L)))  , "--"), 
+  FSL        = c(length(which(!is.na(combined$FSL_L)))    , length(which(combined$FSL_QC==0))), 
+  FS         = c(length(which(!is.na(combined$FS_L)))     , length(which(combined$TEMPQC == "Fail"))),
+  row.names  = c("Images", "Failures")
+)
+
+all_qc = subset(all_qc, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )
+
+attach(all_qc)
 totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
@@
                   FSL   = FSL_L + FSL_R)
-
-detach(combined)
-#complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
+detach(all_qc)
 totals.complete = na.omit(totals)
 @
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1 Screening Dataset Package Totals & Failures                           %
+%                                                                             %
+<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
+latex(package_totals, file="", size="scriptsize", 
+    caption="ADNI1 1.5T Screening Dataset Segmentations",
+    title="",
+    label="tab:ADN1-1.5T-Screening-Dataset-Segmentations")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2013-04-05 paper.Rnw
Lots of updates: a2a_tracc results updated, mucho updates to analysis

@@
 theme_set(theme_bw())
-# bland_altman_plot <- function(dat){
-#   dat$diff <- dat$pre - dat$post;
-#   
-#   return ggplot(dat, aes(pre, diff)) + 
-#     geom_hline(yintercept = mean(dat$diff) + c(-2, 0, 2) * sd(dat$diff), 
-#                linetype=2, color='brown') + 
-#     geom_point() 
-# }
+
+bland_altman_plot <- function(x, y) {
+  data = data.frame(x=x,y=y,diff=x-y)
+  return(ggplot(data, aes(x= x, y = diff)) + 
+    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
+               linetype=2, color='brown') + 
+    geom_point())
+}
 @
@@
 isotropic) T1-weighted MR scans\citep{Winterburn2013}. To validate \mb in a
-real world situation, we segment the entire ADNI1 screening dataset and compare
+real world situation, we segment the entire ADNI1 Complete 1Yr dataset and compare
 our segmentations to established automated and manual segmentations.
@@
 
-\subsubsection{ADNI1 1.5T Screening Dataset}
+\subsubsection{ADNI1 1.5T Complete 1Yr Dataset}
 Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
 authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset download was the "ADNI1:Screening 1.5T"
+August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
 standardized dataset available from ADNI \footnote{
@@
 resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
-in table \ref{tab:ADNI1_1.5T_Screening_Dataset_Demographics}.
+in table \ref{tab:ADNI1_1.5T_Complete 1Yr_Dataset_Demographics}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 Screening Dataset Demographics Table                                  %
+% ADNI1 Complete 1Yr Dataset Demographics Table                                  %
 %                                                                             %
@@
 # RID < 2000 ensures only ADNI1 patients
+yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID < 2000,   
+               MMSE, data = adnimerge, 
+               subset =  RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
                method = "reverse", test=FALSE, overall=TRUE)
+
 latex(tab, file="", size="scriptsize", 
-    caption="ADNI1 1.5T Screening dataset demographics",
-    title="tab:ADN1_1.5T_Screening_Dataset_Demographics")
+    caption="ADNI1 1.5T_Complete_1Yr dataset demographics",
+    title="tab:ADN1_1.5T_Complete_1Yr_Dataset_Demographics")
 @
@@
 % First Episode SZ Demographics Table                                         %
-%                                                                             %
+%
+
+\begin{figure}
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
@@
 @
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \caption{
+  }
+\end{figure}
 
@@
 {\bf Dataset evaluated.} 69 1.5T images were arbitrarily selected from the {\em
-ADNI1:Screening 1.5T} standardized dataset. 23 subjects were chosen from each
+ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
 disease category (cognitively normal (CN), MCI and AD). Demographics for this
@@
 %%%%%
-%%%%%               Experiments: ADNI-1 Screening Validation               %%%%%
-%%%%%
-\subsubsection{ADNI-1 Screening Validation}
+%%%%%               Experiments: ADNI-1 Complete 1Yr Validation               %%%%%
+%%%%%
+\subsubsection{ADNI-1 Complete 1Yr Validation}
 To test the accuracy of \mb on a real-world task we segment the entire
@@
 
-{\bf Dataset evaluated.} All images from the {\em ADNI1:Screening 1.5T}
+{\bf Dataset evaluated.} All images from the {\em ADNI1:Complete 1Yr 1.5T}
 standardized dataset.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 Screening Template Library Demographics Table
+% ADNI1 Complete 1Yr Template Library Demographics Table
 %
@@
 latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
-      caption="ADNI1 Screening Template Library demographics", 
+      caption="ADNI1 Complete 1Yr Template Library demographics", 
       title="tab:ADNI1_scr_tmpllib_demographics")
@@
 %                                                                              %
-% ADNI1 Screening Validation - results preparation                             %
+% ADNI1 Complete 1Yr Validation - results preparation                             %
 %                                                                              %
-<<ADN1-xval-prep, include=FALSE, cache=TRUE>>=
+<<ADN1-xval-load, include=FALSE, cache=TRUE>>=
 # all data from the validation trials
@@
 #t_data       = read.csv(gzfile('data/a2a_tracc/xfmtrials_results_2013-04-01.csv.gz'))
-t_data       = read.csv(gzfile('data/a2a_tracc/results_2012_12_31.csv.gz'))
+t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz'))
 diagnoses    = read.csv('data/a2a_diagnoses.csv')
-
-# simplifies the table
+@
+
+<<ADN1-xval-prep, include=FALSE, cache=TRUE>>=
+# simplify the data
 t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
@@
 # we only compare multiatlas using majority vote
-t_ma = subset(t_data, approach=="multiatlas" & method=="majvote")
+t_ma = subset(t_data, approach=="ma" & method=="majvote")
+t_ma$approach="multiatlas"
 a_ma = subset(a_data, approach=="multiatlas" & method=="majvote")
@@
 segmentations demonstrate trend overlap accuracy that increases with the size of
-atlas library and template library. With 12 or more templates \mb library of at
-least 12 templates
+atlas library and template library. Improvement in accuracy diminishes noticeably
+with template libraries larger than ten images.
 
@@
 template library sizes.  In every parameter configuration, the use of \mb with
-\ants registration shows increased segmentation accuracy over \mb with \animal
+\ants registration shows a pronounced increase in segmentation accuracy over \mb with \animal
 registration. In the remainder of this section, only results using the \ants
@@
   \caption{
-  {\em \mb mean DSC by atlas and template library size, registration algorithm,
-  and label fusion method. Mean DSC is calculated over each subject from the
-  ADNI1 Cross-Validation } 
+  {\em Mean DSC by atlas and template library size, registration algorithm,
+  and label fusion method. The mean DSC score is calculated over all subjects 
+  from the ADNI1 Cross-Validation dataset for each parameter setting.} 
   }
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
-@
-  \caption{Increase in mean Kappa between \mb and multi-atlas}
+  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) 
+@
+  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
+  a range of parameter settings.}
   \label{ADNI1-xval-k-diff}
@@
 \mb also shows a decrease in the variability of segmentation accuracy 
-with different input atlases (Figure \todo{ref}).  Additionally, variability
-decreases with the size of template library, and is significantly decreased
-(p<0.05) when using 3-7 atlases and 17 or more templates.
+(Figure \todo{ADNI1-xval-variability}).  The size of template library 
+necessary to reach a significant (p<0.5) decrease in variance and standard 
+deviation grows with the size of atlas library used.  A template library of 19
+images is sufficient to show significant decrease in variance and standard 
+deviation for 3-7 atlases.
 
 \begin{figure}
-<<ADN1-Xval-k-sd-t-test-p-plot, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ANTSmajvote = subset(all_data, reg_method=="ANTS" & method.mb=="Majority Vote")
+<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+ANTSmajvote = subset(all_data, 
+                     reg_method=="ANTS" & 
+                     method.mb=="Majority Vote" &
+                    templates.mb %in% seq(1,20,2))
 
@@
   data.frame(
-    stat      = c("var", "sd"),
+    stat      = c("Variance", "Standard Deviation"),
     statistic = c(t.var$statistic, t.sd$statistic),
-    p.value   = c(t.var$p.value, t.sd$p.value)
+    p.value   = c(t.var$p.value, t.sd$p.value),
+    direction = c(t.var$statistic < 0, t.sd$statistic < 0)
   )
@@
 ggplot(stats.tests, 
-       aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
+  aes(x=templates.mb, y=p.value, colour=as.factor(atlases), 
+      size = abs(statistic))) + 
   facet_grid( . ~ stat) + 
-  geom_line() + 
-  geom_hline(aes(yintercept=0.05), linetype='dashed') + 
+  geom_line(aes(size=0.1)) + 
+  geom_point(aes(shape = factor(direction))) + 
+  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
+  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
   scale_colour_hue(name="Atlases") +
@@
   scale_linetype_discrete(guide='none') + 
-  scale_x_continuous(breaks=seq(1,20,by=2)) + scale_y_continuous(breaks=seq(0,1,by=0.05)) + 
+  #scale_size_continuous(guide='none') + 
+  scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
+  scale_size_area(name="t-statistic magnitude") + 
+  scale_x_continuous(breaks=seq(1,20,by=2)) + 
+  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
   xlab( "Number of Templates" ) + 
   ylab( "p-value" ) 
-
-@
+@
+  \caption{Difference in variability of \mb vs. multi-atlas segmentation accuracy. 
+  In these plots we consider the segmentation accuracy within each subject across
+  the 10 rounds of validation per parameter setting. Standard deviation and 
+  variance is computed per subject for \mb and multi-atlas, and the distribution
+  of these test statistics is compared with a t-test. The p-value of this test, 
+  is shown on the y-axis (scaled logarithmically).}
+  \label{ADNI1-xval-variability}
 \end{figure}
 
-<<ADNI1-Xval-sd-by-subject, cache=TRUE>>=
-SDs = data.stats[c('subject','label','atlases','templates.mb','MA.sd','MB.sd')]
-SDs = subset(SDs, templates.mb %in% c(5,10,15,20) & atlases == 5)
-
-melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
-ggplot(melted, aes(x=templates.mb, y=value, 
-                   group=interaction(templates.mb, variable), colour=as.factor(variable))) + 
-  scale_x_continuous(breaks=c(5,10,15,20)) + 
-  geom_violin() 
-@
-
-%- show variance in MAGeT performance by showing nine plots, each with mean multi-atlas performance plotted.
+% <<ADNI1-Xval-sd-by-subject, cache=TRUE>>=
+% SDs = data.stats[c('subject','label','atlases','templates.mb','MA.sd','MB.sd')]
+% SDs = subset(SDs, templates.mb %in% c(5,10,15,20) & atlases == 5)
+% 
+% melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
+% ggplot(melted, aes(x=templates.mb, y=value, 
+%                    group=interaction(templates.mb, variable), colour=as.factor(variable))) + 
+%   scale_x_continuous(breaks=c(5,10,15,20)) + 
+%   geom_violin() 
+% @
+% \begin{figure}
+% <<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+% a_20_tmpls = subset(all_data_mean, 
+%                       reg_method  == "ANTS" &
+%                       method.mb   == "Majority Vote" &
+%                       templates.mb== 20)
+% ggplot(a_20_tmpls, aes(x=factor(atlases))) +
+%   #geom_point(aes(y=k.mb,group=1,colour="MAGeT"),position="jitter") + 
+%   #geom_point(aes(y=k.ma,group=2,colour="MA"),position="jitter") + 
+%   geom_violin(aes(y=k.mb,group=factor(atlases),colour="MAGeT")) + 
+%   geom_violin(aes(y=k.ma,group=factor(atlases),colour="MA")) + 
+%   geom_smooth(aes(y=k.mb,group=1,colour="MAGeT")) +
+%   geom_smooth(aes(y=k.ma,group=2,colour="MA")) +
+%   
+%   #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
+%   #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
+%   #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
+%   #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
+%   labs(x="Number of Atlases", y = "Kappa", 
+%        title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
+%   scale_colour_hue(name="Method") + 
+%   coord_cartesian(ylim=c(0.7,0.9))
+%   #theme(legend.justification=c(1,0), legend.position=c(1,0))
+% @
+%   \caption{Change in mean Kappa between multi-atlas and \mb using \ants,
+%   20 templates}
+%   \label{}
+% \end{figure}
+% 
+% \begin{figure}
+% <<ADN1-Xval-ANTS-20tmpls-t-test-sd, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+% a_20_tmpls = subset(all_data, 
+%                       reg_method  == "ANTS" &
+%                       method.mb    == "Majority Vote" &
+%                       templates.mb == 20 &
+%                       atlases     == 9 & 
+%                       volume.ma < 10000)
+% stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
+%   data.frame(
+%     MA = sd(df$k.ma),  #MA
+%     MB = sd(df$k.mb)  #MB
+%   )
+% })
+% 
+% 
+% # TODO: against SNT
+% melted = melt(stddev, id.vars=c("subject", "label"))
+% ggplot(melted, aes(x=variable, y=value)) + geom_boxplot()
+% t.test(stddev$MA, stddev$MB)
+% @
+%   \caption{}
+%   \label{}
+% \end{figure}
+% 
+%
+
+Finally, \mb segmentation volumes show differential agreement with manual (SNT) 
+volumes throughout the range of hippocampal volumes (Figure 
+\ref{ADNI1-xval-Bland-Altman}). \mb shows a fixed bias towards producing smaller
+segmentations, and shows a negligble proportional bias towards smaller 
+segmentations inversely proportional to volume size. Nevertheless, when using 
+3 or more atlases, \mb volumes are within the 95\% limits of agreement with 
+manual labels.
 
 \begin{figure}
-<<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-a_20_tmpls = subset(all_data_mean, 
-                      reg_method  == "ANTS" &
-                      method.mb   == "Majority Vote" &
-                      templates.mb== 20)
-ggplot(a_20_tmpls, aes(x=factor(atlases))) +
-  #geom_point(aes(y=k.mb,group=1,colour="MAGeT"),position="jitter") + 
-  #geom_point(aes(y=k.ma,group=2,colour="MA"),position="jitter") + 
-  geom_violin(aes(y=k.mb,group=factor(atlases),colour="MAGeT")) + 
-  geom_violin(aes(y=k.ma,group=factor(atlases),colour="MA")) + 
-  geom_smooth(aes(y=k.mb,group=1,colour="MAGeT")) +
-  geom_smooth(aes(y=k.ma,group=2,colour="MA")) +
-
-  #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
-  #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
-  #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
-  #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
-  labs(x="Number of Atlases", y = "Kappa", 
-       title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
-  scale_colour_hue(name="Method") + 
-  coord_cartesian(ylim=c(0.7,0.9))
-  #theme(legend.justification=c(1,0), legend.position=c(1,0))
-@
-  \caption{Change in mean Kappa between multi-atlas and \mb using \ants,
-  20 templates}
-  \label{}
-\end{figure}
-
-\begin{figure}
-<<ADN1-Xval-ANTS-20tmpls-t-test-sd, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+
+bland_altman_plot <- function(data, x, y) {
+  data$diff = data[x] - data[y] 
+  return(ggplot(data, aes(x= x, y = diff)) + 
+    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
+               linetype=2, color='brown') + 
+    geom_point())
+}
 a_20_tmpls = subset(all_data, 
@@
                       templates.mb == 20 &
-                      atlases     == 9 & 
+                      atlases      %in% seq(1,9,by =2) &
                       volume.ma     < 10000)
-stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
+total_HC_volume = ddply(a_20_tmpls, c("subject", "label", "atlases"), function (df) {
   data.frame(
-    MA = sd(df$k.ma),  #MA
-    MB = sd(df$k.mb)  #MB
+    ma = sum(df$volume.ma),
+    mb = sum(df$volume.mb)
   )
 })
-
-# TODO: against SNT
-melted = melt(stddev, id.vars=c("subject", "label"))
-ggplot(melted, aes(x=variable, y=value)) + geom_boxplot()
-t.test(stddev$MA, stddev$MB)
-@
-  \caption{}
-  \label{}
+total_HC_volume$diff = total_HC_volume$ma - total_HC_volume$mb
+
+ggplot(total_HC_volume, aes(x=ma, y = diff, group=atlases)) + 
+    geom_hline(yintercept = mean(total_HC_volume$diff) + c(-2, 0, 2) * sd(total_HC_volume$diff), 
+               linetype=2, color='brown') + 
+    geom_point() + 
+  facet_grid( . ~ atlases) + 
+  xlab('Mean multi-atlas and MAGeT Brain total Hippocampus volume') + 
+  ylab('Multi-atlas - MAGeT Brain volume') 
+@
+  \caption{Bland-Altman plot comparing multi-atlas and \mb hippocampal volumes 
+  when using 9 atlases, and a 20 image template library.}
+  \label{ADNI1-xval-Bland-Altman}
 \end{figure}
-
-% \begin{figure}
-% <<ADN1-Xval-ANTS-20tmpls-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-% a_20_tmpls = subset(all_data, 
-%                       reg_method  == "ANTS" &
-%                       method.y    == "Majority Vote" &
-%                       templates.y == 20 &
-%                       atlases     == 9 &
-%                       volume.x < 10000)
-% avg = ddply(a_20_tmpls, c("subject", "label"), function (df) {
-%   data.frame(
-%     pre = mean(df$volume.x),  #MA
-%     post = mean(df$volume.y)  #MB
-%   )
-% })
-% # TODO: against SNT
-% 
-% #df = data.frame(pre=a_20_tmpls$volume.x, post=a_20_tmpls$volume.y)
-% bland_altman_plot(avg) + 
-%   xlab('Mean of MA and MAGeT HC Left/Right Volumes') + 
-%   ylab('MA - MAGeT Volume') + 
-%   opts(title = 'Bland-Altman Plot')
-% @
-%   \caption{Change in mean Kappa between multi-atlas and \mb using ANTS, 20 templates}
-%   \label{}
-% \end{figure}
 % 
@@
 %%%%%
-%%%%%                  Results: ADNI-1 Screening Validation               %%%%%
-%%%%%
-\subsection{ADNI1 Screening Validation}
-
-Based on the results from the ADNI1 Cross-Validation experiment, in this
-experiment \mb was configured with a template library of 21 randomly
-chosen subject images (7 from each disease class) and used majority vote label
-fusion. Overall co
+%%%%%                  Results: ADNI-1 Complete 1Yr Validation               %%%%%
+%%%%%
+\subsection{ADNI1 Complete 1Yr Validation}
 
@@
 <<ADNI-src-volumes-prep,echo=F,cache=T>>=
-bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
+#bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
 #bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
 #yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
@@
 package_totals = data.frame(
-  Screening  = c(length(subjects$VISCODE), 'N/A'),
+  Total     = c(length(subjects$VISCODE), 'N/A'),
   SNT        = c(length(which(!is.na(combined$SNT_L)))    , "--"),
@@
 
+Based on the results from the ADNI1 Cross-Validation experiment, in this
+experiment \mb was configured with a template library of 21 randomly
+chosen subject images (7 from each disease class) and used majority vote label
+fusion. The entire ADNI1 Complete 1Yr 1.5T dataset was segmented by \mb and we 
+now compare the resulting volumes with those obtained by manual segmentation (SNT), 
+and other automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
+\ref{tab:ADN1-1.5T-Complete-1yr-Dataset-Segmentations} shows the total count of 
+segmentations available, including a count of those which have failed a quality 
+control inspection. Only those images which had segmentations from every method
+are included in the following analysis (a total of 
+\Sexpr{length(totals.complete$RID)} images).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % ADNI1 Screening Dataset Package Totals & Failures                           %
-%                                                                             %
+%                        
+%
+\begin{figure}
 <<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
 latex(package_totals, file="", size="scriptsize", 
-    caption="ADNI1 1.5T Screening Dataset Segmentations",
+    caption="ADNI1 1.5T Complete 1yr Dataset Segmentations",
     title="",
-    label="tab:ADN1-1.5T-Screening-Dataset-Segmentations")
-@
+    label="tab:ADN1-1.5T-Complete-1yr-Dataset-Segmentations")
+@
+  \caption{Pass/fail quality control indicators were supplied with the FreeSurfer
+  volumes downloaded from the ADNI website (we used the temporal lobe quality
+  control indicator, TEMPQC). One of the authors (MP) performed visual 
+  quality inspection for MAGeT and FSL segmentations.  
+  }
+\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 @
-  \caption{Comparison of HC volumes by FreeSurfer (FSF), \mb (MAGeT), MAPER, and manual (SNT).}
+  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), , FSL, 
+  MAPER, \mb (MAGeT) and manual (SNT).}
   \label{ADNI-scr-volumes-boxplot}
@@
 <<FEP-volumes,cache=T>>=
-SZ_volumes=read.csv("data/SZ_volumes_by_method.csv")
-with(SZ_volumes,
-qplot(Manual,MAGeT,data=SZ_volumes) + geom_smooth(method="lm") + 
+SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
+casted = dcast(SZ_volumes, Subject + Label ~ Method)
+ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
+  geom_smooth(method="lm") + 
+  geom_point() + 
 #   geom_text(aes(
@@
 
+<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes'>>=
+bland_altman_plot(casted$Manual, casted$MAGeT) + 
+  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
+  ylab('Manual - MAGeT Brain volume') 
+@
+  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
+  when using five Winterburn atlases, and a 21 image template library.}
+  \label{}
+\end{figure}
+
 %%%%%
@@
 gold_0.3mm = subset(data, version == "gold_0.3mm")
-gold_0.6mm = subset(data, version == "gold_0.6mm")
+gold_0.9mm = subset(data, version == "gold_0.9mm")
 mb_bravo   = subset(data, version == "mb_bravo")
-mb_0.6mm   = subset(data, version == "mb_0.6mm")
+mb_0.9mm   = subset(data, version == "mb_0.9mm")
 
 gold_0.3mm_total = gold_0.3mm[,2:6] + gold_0.3mm[,7:11]
-gold_0.6mm_total = gold_0.6mm[,2:6] + gold_0.6mm[,7:11]
+gold_0.9mm_total = gold_0.9mm[,2:6] + gold_0.9mm[,7:11]
 mb_bravo_total   = mb_bravo[,2:6] + mb_bravo[,7:11]
-mb_0.6mm_total   = mb_0.6mm[,2:6] + mb_0.6mm[,7:11]
+mb_0.9mm_total   = mb_0.6mm[,2:6] + mb_0.9mm[,7:11]
 
 # calculate percent error from the gold standard measured at 0.3mm resolution
-pe_gold0.6mm = (gold_0.6mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
+pe_gold0.9mm = (gold_0.9mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
 pe_mbbravo   = (mb_bravo_total - gold_0.3mm_total)/gold_0.3mm_total * 100
-pe_mb0.6mm   = (mb_0.6mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
-
-subjects = gold_0.6mm[,1]
-pe_gold0.6mm = cbind(subjects,pe_gold0.6mm, res="Resampling to 0.6mm")
+pe_mb0.9mm   = (mb_0.9mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
+
+subjects = gold_0.9mm[,1]
+pe_gold0.9mm = cbind(subjects,pe_gold0.9mm, res="Resampling to 0.9mm")
 pe_mbbravo   = cbind(subjects,pe_mbbravo,   res="MAGeT-on-BRAVO")
-pe_mb0.6mm   = cbind(subjects,pe_mb0.6mm,   res="MAGeT-at-0.6mm")
+pe_mb0.6mm   = cbind(subjects,pe_mb0.9mm,   res="MAGeT-at-0.9mm")
 
@@
 # compare % error between MAGeT and resampling
-mb0.6mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.6mm_diff), res="diffdff")
-names(mb0.6mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
-melted = melt(mb0.6mm_diff_over_gold_diff, id.vars=c("subject", "res"))
+mb0.9mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.9mm_diff), res="diffdff")
+names(mb0.9mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
+melted = melt(mb0.9mm_diff_over_gold_diff, id.vars=c("subject", "res"))
 names(melted) <- c("subject","res", "region","voldiff")

### 2013-04-05 paper.Rnw
Updates to figures with captions and sensible sizes, fixed latex references to figures

@@
 <<setup, include=FALSE, cache=FALSE>>=
-#opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
-#    fig.align = 'center', tidy = FALSE, comment = NA, cache = TRUE)
+library(knitr)
+opts_chunk$set(echo = FALSE, 
+               message = FALSE, 
+               warning = FALSE, 
+               fig.align = 'center', 
+               tidy = FALSE, 
+               comment = NA, 
+               cache = TRUE,
+               fig.width = 6, 
+               fig.height = 6)
 library(ggplot2)
@@
 
-\newcommand{\mb}{MAGeT Brain}
-\newcommand{\ants}{ANTS}
-\newcommand{\animal}{ANIMAL}
+\newcommand{\mb}{MAGeT-Brain }
+\newcommand{\ants}{ANTS }
+\newcommand{\animal}{ANIMAL }
 
@@
 resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
-in table \ref{tab:ADNI1_1.5T_Complete 1Yr_Dataset_Demographics}.
+in table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.
 
@@
 latex(tab, file="", size="scriptsize", 
-    caption="ADNI1 1.5T_Complete_1Yr dataset demographics",
-    title="tab:ADN1_1.5T_Complete_1Yr_Dataset_Demographics")
+    caption="ADNI1 1.5T Complete 1Yr dataset demographics",
+    label="tab:ADN1-1.5T-Complete-1Yr-Dataset-Demographics")
 @
@@
 for the images was 256mm (SI)×204mm (AP).  Subject demographics are shown in
-table \ref{tab:SZFEP_Demographics}.
+table \ref{tab:SZFEP-Demographics}.
 
@@
 
-\begin{figure}
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
@@
                data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", size="scriptsize", ctable=FALSE,
+latex(tab, file="", ctable=FALSE,
       caption="Schizophrenia First Episode Patient Demographics",
-      title="tab:SZFEP_Demographics")
+      title="tab:SZFEP-Demographics")
 @
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \caption{
-  }
-\end{figure}
 
@@
 regularization parameters optimized in Robbins et al. \citep{Robbins2004},
-displayed in table \ref{tab:ANIMAL_params}.
+displayed in table \ref{tab:ANIMAL-params}.
 
@@
 %                                                                             %
-\begin{table*}[!tbp]
-\scriptsize
-\caption{\animal Registration Parameters\label{tab:ANIMAL_params}} 
+\begin{table}[!tbp]
 \begin{center}
+\caption{\animal  registration parameters}
 \begin{tabular}{l | c c c}
@@
 \end{center}
-\end{table*}
+\label{tab:ANIMAL-params}
+\end{table}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 disease category (cognitively normal (CN), MCI and AD). Demographics for this
-subset are shown in Table \ref{tab:ADNI1_xvalidation_demographics}.
+subset are shown in Table \ref{tab:ADNI1-xvalidation-demographics}.
 
@@
       caption="ADNI-1 cross-validation subset demographics", 
-      title="tab:ADNI1_xvalidation_demographics")
+      title="tab:ADNI1-xvalidation-demographics")
 @
@@
 data for the template library subjects are shown in Table
-\ref{tab:ADNI1_scr_tmpllib_demographics}.
+\ref{tab:ADNI1-scr-tmpllib-demographics}.
 
@@
       caption="ADNI1 Complete 1Yr Template Library demographics", 
-      title="tab:ADNI1_scr_tmpllib_demographics")
+      title="tab:ADNI1-scr-tmpllib-demographics")
 @
@@
 <<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-
 ggplot(all_data_mean, 
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Mean DSC" )
+  ylab( "Mean DSC" ) + 
+  opts(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   \caption{
-  {\em Mean DSC by atlas and template library size, registration algorithm,
-  and label fusion method. The mean DSC score is calculated over all subjects 
+  {\bf Mean DSC by atlas and template library size, registration algorithm,
+  and label fusion method.} The mean DSC score is calculated over all subjects 
   from the ADNI1 Cross-Validation dataset for each parameter setting.}
-  }
 \end{figure}
@@
   geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
-  coord_cartesian(ylim=c(-0.1,0.05)) + 
   scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) 
+  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) +
+  opts(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   facet_grid( . ~ stat) + 
-  geom_line(aes(size=0.1)) + 
+  geom_line(aes(size=0.1,area=0.1)) + 
   geom_point(aes(shape = factor(direction))) + 
@@
   geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
-  scale_colour_hue(name="Atlases") +
-  scale_alpha_continuous(guide='none') + 
-  scale_linetype_discrete(guide='none') + 
-  #scale_size_continuous(guide='none') + 
   scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
@@
   xlab( "Number of Templates" ) + 
-  ylab( "p-value" ) 
-@
-  \caption{Difference in variability of \mb vs. multi-atlas segmentation accuracy. 
+  ylab( "p-value" ) +
+  scale_colour_hue(name="Atlases") +
+  scale_alpha_continuous(guide='none') + 
+  scale_linetype_discrete(guide='none') + 
+  scale_size_continuous(guide='none') + 
+  opts(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
+  accuracy.}
   In these plots we consider the segmentation accuracy within each subject across
@@
 \begin{figure}
-<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-
-bland_altman_plot <- function(data, x, y) {
-  data$diff = data[x] - data[y] 
-  return(ggplot(data, aes(x= x, y = diff)) + 
-    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + 
-    geom_point())
-}
+<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=4>>=
 a_20_tmpls = subset(all_data, 
@@
 
-% \begin{itemize}
-% \item  A2A shows that if atlas population strongly(?) represents subject set
-%        variability, then free choice from atlas population will produce
-%        improvements (we know this b/c of extensive validation trials). 
-% \item  what about in the case where atlas population doesn???t strongly
-%        represent subject set variability (e.g. a priori atlas set)?  then, we
-%        can use atlas selection to refine atlas set? 
-% \end{itemize}
-% 
-% \todo{Kappa against our manual rater is low... explain that}
-
 <<ADNI-src-volumes-prep,echo=F,cache=T>>=
-#bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
-#bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
-#yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
@@
 qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")
-# recode VISCODE
-#viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
-#yr1$VISCODE <- as.factor(yr1$Visit)
-#levels(yr1$VISCODE) <-viscode_levels
-#bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)
-
-#subjects = subset(yr1,Visit ==1)
+
 subjects = bl
-subjects$VISCODE = 'bl'
+subjects$VISCODE = 'bl' # to match up with adnimerge
 
@@
 
-# A table showing the number of images (and failures) from each package
-#           Total  MAGeT  FSL  FS   MAPER
-# Images    
-# Failures
-# 
-# Total number complete images =
-
 # Step 2: Of those, get counts for the numbers that fail QC
@@
                   SNT   = SNT_L + SNT_R, 
-                  #UCSD  = UCSD_L + UCSD_R,
                   FS    = FS_L + FS_R, 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 Screening Dataset Package Totals & Failures                           %
-%                        
-%
-\begin{figure}
+% ADNI1 1yr Complete Dataset Package Totals & Failures                        %
+%                        
+%
 <<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
+caption = paste(
+  "Pass/fail quality control indicators were supplied with the FreeSurfer",
+  "volumes downloaded from the ADNI website (we used the temporal lobe quality",
+  "control indicator, TEMPQC). One of the authors (MP) performed visual",
+  "quality inspection for MAGeT and FSL segmentations.", sep=" ")
 latex(package_totals, file="", size="scriptsize", 
-    caption="ADNI1 1.5T Complete 1yr Dataset Segmentations",
-    title="",
-    label="tab:ADN1-1.5T-Complete-1yr-Dataset-Segmentations")
-@
-  \caption{Pass/fail quality control indicators were supplied with the FreeSurfer
-  volumes downloaded from the ADNI website (we used the temporal lobe quality
-  control indicator, TEMPQC). One of the authors (MP) performed visual 
-  quality inspection for MAGeT and FSL segmentations.  
-  }
-\end{figure}
+    caption = caption,
+    title = "",
+    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
+@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+Comparing total hippocampal volume across segmentation methods, we find close
+agreement between manual volumes and MAPER and MAGeT volumes across disease 
+categories (Figure \ref{ADNI-scr-volumes-boxplot}). FSL and FreeSurfer both 
+produce volumes which are demonstratively larger than manually segmented volumes.
+
 \begin{figure}[h]
-<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T>>=
+<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
+snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
+  data.frame( c = cor(df$SNT, df$value) ) 
+})
+
 qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
@@
 @
-  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), , FSL, 
+  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
   MAPER, \mb (MAGeT) and manual (SNT).}
@@
 
-latex(df, title="",
-      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
-      file="", size="scriptsize")
+#latex(df, title="",
+#      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
+#      file="", size="scriptsize")
 @
@@
    ylab("Total Hippocampal Volume (MAGeT Segmentation)")
-)
 @
@@
 
+\begin{figure}
 <<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes'>>=
@@
 
+\begin{figure}
 <<WAval-vol-boxplot,cache=T>>=
@@
 subjects = gold_0.9mm[,1]
-pe_gold0.9mm = cbind(subjects,pe_gold0.9mm, res="Resampling to 0.9mm")
-pe_mbbravo   = cbind(subjects,pe_mbbravo,   res="MAGeT-on-BRAVO")
-pe_mb0.6mm   = cbind(subjects,pe_mb0.9mm,   res="MAGeT-at-0.9mm")
+pe_gold0.9mm = cbind(subjects,pe_gold0.9mm, res="Resampled")
+pe_mbbravo   = cbind(subjects,pe_mbbravo,   res="BRAVO")
+pe_mb0.6mm   = cbind(subjects,pe_mb0.9mm,   res="0.9mm voxels")
 
@@
 names(pe) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
-
 melted = melt(pe, id.vars=c("subject", "res"))
 names(melted) <- c("subject","res", "region","percenterr")
+
 ggplot(melted, aes(y=percenterr,x=region)) +
   geom_boxplot(aes(colour=res)) +
-  labs(title="Percent error in volumes under different measurement regimes",
-       y="Percent error in volume", 
-       x="Region") 
-@
-
-<<WAval-error,cach=T>>=
-# compare % error between MAGeT and resampling
-mb0.9mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.9mm_diff), res="diffdff")
-names(mb0.9mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
-melted = melt(mb0.9mm_diff_over_gold_diff, id.vars=c("subject", "res"))
-names(melted) <- c("subject","res", "region","voldiff")
-ggplot(melted, aes(y=voldiff,x=region)) +
-  geom_boxplot() +
-  labs(title="Absolute difference in % difference between MAGeT and resampling",
-       y="Absolute difference in % difference in volume", 
-       x="Region")
-@
+  labs(y="Percent error in volume", x="Region") +
+  opts(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Percent error in segmentation volume by Hippocampus subregion.}}
+  \label{WAval-vol-boxplot}
+\end{figure}
+% 
+% <<WAval-error,cach=T>>=
+% # compare % error between MAGeT and resampling
+% mb0.9mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.9mm_diff), res="diffdff")
+% names(mb0.9mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
+% melted = melt(mb0.9mm_diff_over_gold_diff, id.vars=c("subject", "res"))
+% names(melted) <- c("subject","res", "region","voldiff")
+% ggplot(melted, aes(y=voldiff,x=region)) +
+%   geom_boxplot() +
+%   labs(title="Absolute difference in % difference between MAGeT and resampling",
+%        y="Absolute difference in % difference in volume", 
+%        x="Region")
+% @
 

### 2013-04-09 paper.Rnw
Results section DONE! Methods section DONE!  First draft anyway

@@
     geom_point())
+}
+
+lm_eqn = function(m) {
+  l <- list(a = format(coef(m)[1], digits = 2),
+      b = format(abs(coef(m)[2]), digits = 2),
+      r2 = format(summary(m)$r.squared, digits = 3));
+  if (coef(m)[2] >= 0)  {
+    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
+  } else {
+    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
+  }
+
+  as.character(as.expression(eq))
 }
@@
 yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
+yr1 = subset(yr1, !grepl("Scaled_2", Description))
+yr1 = yr1[,c("RID","VISCODE")]
+
+adnimerge_yr1 = merge(adnimerge, yr1) 
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
                MMSE, data = adnimerge, 
-               subset =  RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
+               subset =  adnimerge_yr1, #RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
                method = "reverse", test=FALSE, overall=TRUE)
@@
 
-\subsubsection{Schizophrenic First Episode Patients}
+\subsubsection{Schizophrenia First Episode Patients}
+\label{sec:SZFEP-dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 \subsubsection{SZ First Episode Patient Validation}
-
-{\bf Dataset evaluated.} To validate that \mb performance generalises to
-across disease categories, we measure the performance of \mb on the
-Schizophrenic First Episode Patient (SFEP) dataset.
-
-{\bf Atlas and template libary.} - two different atlas sets: a manual
-hippocampal segmentation of patients, and Winterburn atlas set.  
-\todo{complete}
-
-{\bf Registration method.} \ants.
-\todo{complete}
-
-{\bf Label fusion.} Majority vote.
-\todo{complete}
-
-{\bf Evaluation.} We validate the FEP-atlas segmentations using Dice's Kappa,
-and the Winterburn-atlas segmentations by correlating volumes. 
-\todo{complete}
+The previous experiments explore \mb on an Alzheimer's disease dataset (ADNI).  To validate
+that \mb algorithm works effectively with other diseased brain images, 
+the performance of \mb is measured on a Schizophrenia dataset.
+
+{\bf Dataset evaluated.} The entire Schizophrenia First Episode Patient (SZFEP) 
+dataset (see section \ref{sec:SZFEP-dataset}).
+
+{\bf Atlas and template libary.} The atlas library is composed of the Winterburn 
+T1 atlas images, with whole hippocampus segmentations (no subregions). The 
+template library is composed of 21 images selected at random from the SZFEP dataset.
+
+{\bf Registration method.} The \ants registration algorithm is used for image
+registration.
+
+{\bf Label fusion.} The majority-vote method is used for label fusion.
+
+{\bf Evaluation.} The manual segmentation protocol used to segment the Winterburn
+atlases is similar to, but different from, the protocol used to segment the SZFEP
+dataset. Therefore, rather than use an overlap metric, \mb hippocampal volumes
+are compared to the corresponding manual segmentation volumes.
 
@@
 %%%%%
-\subsubsection{Winterburn Atlases Validation}
-{\bf Dataset evaluated.} - T1 BRAVO scans of the same subjects included in the 
-Winterburn atlas set.  These scans are taken within .... weeks of the scans for 
-the Winterburn atlases.
-\todo{complete}
-
-{\bf Atlas and template libary.} - Atlas library is Winterburn T1 atlases.  
-Template library consists of all five T1 BRAVOs, plus 15 T1 healthy control images.
-\todo{complete}
-
-{\bf Registration method.} \ants
-\todo{complete}
-
-{\bf Label fusion.} Majority vote.
-\todo{complete}
-
-{\bf Evaluation.}  - Leave one out cross-validation (LOOCV) in which all five
-subjects are segmented in separate runs of \mb.  In each run, the subject
-to be segmented is excluded from the Atlas library (so only four atlases are used).
-
-Segmentation accuracy is judged by difference in hippocampal volume.
-\todo{complete}
+\subsubsection{Winterburn Atlases Cross-Validation}
+In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
+segmentation is tested using a leave-one-out cross-validation (LOOCV) design.
+
+{\bf Dataset evaluated.} \mb segmentations are evaluated on two datasets. The 
+WA-BRAVO dataset consists of separately acquired 3T T1 BRAVO images (0.9mm-isotropic voxel)
+of four of the five Winterburn atlas subjects. The WA-Resampled dataset consists
+of the Winterburn atlas images and segmentations downsampled to 0.9mm-isotropic 
+voxel resolution.
+
+{\bf Atlas and template libary.} The atlas library is composed of the Winterburn 
+T1 atlas images and hippocampal subregion segmentations. The template library 
+consists of 15 3T T1 images (0.9mm-isotropic voxels) of healthy subjects as well
+as all of the images from the dataset being evaluated.
+
+\todo{reference, demographics for health control images}.
+
+{\bf Registration method.} The \ants image registration method is used.
+
+{\bf Label fusion.} The majority vote label fusion method is used.
+
+{\bf Evaluation.}  Leave-one-out-cross-validation (LOOCV) is performed separately 
+for each dataset evaluated (WA-BRAVO, and WA-Resampled) as follows. Each subject
+in the dataset is segmented by \mb using an atlas library in which that subject's 
+image is excluded. The segmentation volumes of each hippocampal subregion are
+compared to the expert manual segmentations of the unmodified Winterburn atlases. 
+As a point of comparison, we also calculate the subregion volumes from the Winterburn 
+atlas segmentations after downsampling to 0.9mm-isotropic voxels.
 
@@
 small decrease in performance relative to when one fewer template image is used.
-See section (\todo{ref}) for a discussion of this behaviour. In the remainder of
+See section \todo{ref} for a discussion of this behaviour. In the remainder of
 this section, only results from odd-sized template libraries will be shown.
@@
 accuracy over multi-atlas-based segmentation, using the same number of atlases
-and voting method (Figure \todo{fig:ADNI1-xva-k-diff}. The magnitude of improvement over
+and voting method (Figure \ref{fig:ADNI1-xva-k-diff}). The magnitude of improvement over
 multi-atlas-based segmentation decreases with an increasing number of atlases,
@@
 DSC) is found when one atlas is used with a template library of 20 images.
-\todo{stats?}
 
@@
   accuracy.}
-  In these plots we consider the segmentation accuracy within each subject across
-  the 10 rounds of validation per parameter setting. Standard deviation and 
-  variance is computed per subject for \mb and multi-atlas, and the distribution
-  of these test statistics is compared with a t-test. The p-value of this test, 
-  is shown on the y-axis (scaled logarithmically).}
+  Variability of segmentation accuracy within 10 rounds of validation per subject. 
+  Standard deviation and variance is computed per subject for both \mb and 
+  multi-atlas, and the distribution of these test statistics is compared with a 
+  t-test. The p-value of this test, is shown on the y-axis (scaled logarithmically).}
   \label{ADNI1-xval-variability}
@@
 volumes throughout the range of hippocampal volumes (Figure 
-\ref{ADNI1-xval-Bland-Altman}). \mb shows a fixed bias towards producing smaller
-segmentations, and shows a negligble proportional bias towards smaller 
-segmentations inversely proportional to volume size. Nevertheless, when using 
-3 or more atlases, \mb volumes are within the 95\% limits of agreement with 
-manual labels.
+\ref{ADNI1-xval-Bland-Altman}). \mb shows a negligble fixed bias towards 
+producing smaller segmentations, and shows a slight proportional bias towards 
+smaller segmentations with larger hippocampi.
+\todo{Note limits of agreement?}
 
 \begin{figure}
-<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=4>>=
-a_20_tmpls = subset(all_data, 
+<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=8>>=
+df_subset = subset(all_data, 
                       reg_method   == "ANTS" &
                       method.mb    == "Majority Vote" &
-                      templates.mb == 20 &
-                      atlases      %in% seq(1,9,by =2) &
+                      templates.mb %in% c(1,11,19) &
+                      atlases      %in% c(1,5,9) &
                       volume.ma     < 10000)
-total_HC_volume = ddply(a_20_tmpls, c("subject", "label", "atlases"), function (df) {
+
+# get SNT data formatted like 
+snt   = read.csv("data/a2a_snt_volumes.csv")
+df_merged = merge(df_subset, snt, all.x = TRUE)
+
+hc_vol = ddply(df_merged, c("subject", "label", "atlases", "templates.mb"), function (df) {
   data.frame(
+    snt = sum(df$volume.snt),
     ma = sum(df$volume.ma),
@@
 })
-total_HC_volume$diff = total_HC_volume$ma - total_HC_volume$mb
-
-ggplot(total_HC_volume, aes(x=ma, y = diff, group=atlases)) + 
-    geom_hline(yintercept = mean(total_HC_volume$diff) + c(-2, 0, 2) * sd(total_HC_volume$diff), 
-               linetype=2, color='brown') + 
+hc_vol$diff = hc_vol$snt - hc_vol$mb
+
+group.stats = ddply(hc_vol, c("atlases", "templates.mb"), function (df) {
+  data.frame(
+    mean = mean(df$diff),
+    sd   = sd(df$diff)
+  )
+})
+hc_vol = merge(hc_vol, group.stats)
+hc_vol$Atlases = hc_vol$atlases
+hc_vol$Templates = hc_vol$templates.mb
+
+ggplot(hc_vol, aes(x=ma, y = diff, group=Atlases)) + 
+    geom_hline(aes(yintercept = mean + c(-2, 0, 2) * sd, colour='brown'), 
+                linetype=2) + 
     geom_point() + 
-  facet_grid( . ~ atlases) + 
-  xlab('Mean multi-atlas and MAGeT Brain total Hippocampus volume') + 
-  ylab('Multi-atlas - MAGeT Brain volume') 
-@
-  \caption{Bland-Altman plot comparing multi-atlas and \mb hippocampal volumes 
-  when using 9 atlases, and a 20 image template library.}
+    geom_smooth(method="lm") + 
+  facet_grid(Atlases~Templates, labeller = label_both) + 
+  xlab(expression('Mean manual and MAGeT-Brain total hippocampus volume ' (mm^3))) + 
+  ylab(expression('Manual - MAGeT-Brain total hippocampus volume ' (mm^3))) 
+@
+  \caption{Bland-Altman plots comparing manual and \mb hippocampal volumes 
+  when using a varying number of atlases and templates.}
   \label{ADNI1-xval-Bland-Altman}
@@
 <<ADNI-src-volumes-prep,echo=F,cache=T>>=
-bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+#bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
@@
 
-subjects = bl
-subjects$VISCODE = 'bl' # to match up with adnimerge
+# Get subjects, and diagnoses
+yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+yr1 = subset(yr1, !grepl("Scaled_2", Description)) #TODO: remove Scaled_2? 
+yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
+subjects = yr1[,c("Image.Data.ID", "RID","VISCODE")]  # we only need these fields
+dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
+subjects = merge(subjects, dx)
 
@@
 
-# diagnoses
-dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
-
 # Now create one data.frame with all the yr1 volume data we have from each datasource
@@
 combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)
 combined = merge(combined, qc          , by=c("Image.Data.ID"), all.x=TRUE)
@@
 are included in the following analysis (a total of
-\Sexpr{length(totals.complete$RID)} images).
+\Sexpr{length(totals.complete$RID)} images; Table 
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals}).
 
@@
 
-Comparing total hippocampal volume across segmentation methods, we find close
-agreement between manual volumes and MAPER and MAGeT volumes across disease 
-categories (Figure \ref{ADNI-scr-volumes-boxplot}). FSL and FreeSurfer both 
-produce volumes which are demonstratively larger than manually segmented volumes.
-
 \begin{figure}[h]
@@
   MAPER, \mb (MAGeT) and manual (SNT).}
-  \label{ADNI-scr-volumes-boxplot}
+  \label{fig:ADNI-scr-volumes-boxplot}
 \end{figure}
@@
 \begin{figure}[h]
-<<ADNI-scr-histogram,cache=T>>=
+<<ADNI-scr-histogram-prep,cache=T>>=
 melted   = melt(totals.complete, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
@@
 }) 
-
-label_x_pos = 10000
+@
+
+<<ADNI1-scr-histogram,dependson='ADNI-scr-histogram-prep',cache=T>>=
+
 ggplot(data=melted, aes(x=value)) + 
@@
   geom_line(aes(y=freq, colour='red'), data=normaldf) +
-  geom_text(aes(y=29,x=label_x_pos,hjust=1,size=1, 
+  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=1.5,size=1, 
                 label=paste("list(W ==",signif(shapiro.w, digits=3), ", p == ", 
-                            signif(shapiro.p,digits=3), ")")), 
+                            round(shapiro.p,digits=3), ")")), 
             parse=TRUE, data=testsdf) +
-  geom_text(aes(y=25,x=label_x_pos,hjust=1,size=1, 
+  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=3,size=1, 
                 label=paste("list(D ==",signif(ks.D, digits=3), ", p == ", 
-                            signif(ks.p,digits=3), ")")), 
+                            round(ks.p,digits=3), ")")), 
             parse=TRUE, data=snt.tests) + 
-  geom_text(aes(y=21,x=label_x_pos,hjust=1,size=1, 
+  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=5.5,size=1, 
                 label=paste("sigma ==",signif(sd, digits=3))), 
             parse=TRUE, data=testsdf) + 
-  geom_text(aes(y=17,x=label_x_pos,hjust=1,size=1, 
+  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=6,size=1, 
                 label=paste("mu ==",signif(mean, digits=3))), 
             parse=TRUE,data=testsdf) + 
-  xlim(0, 10000) + ylim(0,30) + 
+  #xlim(0, 10000) + 
+  ylim(0,40) + 
   facet_grid(variable ~ DX) + 
@@
   scale_size_continuous(guide="none") + 
-  labs(title="Total hippocampus volume by disease and measurement method", 
-       x=expression(paste("Total hippocampal volume (", mm^2, ")")),
+  labs(x=expression(paste("Total hippocampal volume (", mm^3, ")")),
        y="Frequency") + 
@@
   difference).}
-  \label{ADNI-scr-histogram}
+  \label{fig:ADNI-scr-histogram}
 \end{figure}
 
-%
+Comparing total hippocampal volume, we find close
+agreement between manual volumes, MAPER and MAGeT volumes across disease 
+categories (Figure \ref{fig:ADNI-scr-volumes-boxplot}, Figure 
+\ref{fig:ADNI-scr-histogram}). FSL and FreeSurfer both produce volumes in close 
+agreement with each other, but which are consistently larger than manually 
+segmented volumes in each disease category.
 
@@
 \subsection{First Episode Schizophrenic Patients}
-High volume correlation between Winterburn segmentation volumes and ground
-truth.  (High-ish?) Kappa when using manual segmentations as Atlases.
+In this experiment we move from an Alzheimer's disease dataset to a dataset of 
+first episode schizophrenia parients.  This dataset is segmented using \mb with the
+Winterburn atlases, and a template library of 21 subject images selected at random. 
+Expert manual whole hippocampal segmentations are used as gold standards.
+
+\mb produces hippocampus segmentation volumes that are highly correlated with 
+manual segmentation volumes (Figure \ref{fig:SZ-volumes}).  Additionally, as we 
+saw in the previous experiments, \mb volumes show a fixed bias towards smaller 
+volumes (although the bias is negligble), and a proportional bias towards 
+producing smaller segmentations for larger hippocampi (Figure \ref{fig:FEP-SZ-Bland-Altman}.
 
 \begin{figure}
-<<FEP-volumes,cache=T>>=
+<<FEP-volumes,cache=FALSE,dependson="setup">>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
 casted = dcast(SZ_volumes, Subject + Label ~ Method)
+lm = lm(MAGeT ~ Manual, casted)
 ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
-  geom_smooth(method="lm") + 
+  geom_smooth(method="lm", formula=y~x) + 
   geom_point() + 
-#   geom_text(aes(
-#       label=paste("r = ", round(cor(Manual,MAGeT, use="complete.obs"), 2), 
-#                   ", p <", round(as.numeric(t.test(Manual, MAGeT, paired=T)["p.value"]), 6))),
-#       x=7000,y=10000, hjust=0,vjust=-0.5,size=5,data=data.frame()) +
-   xlab("Total Hippocampal Volume (Manual Segmentation)") + 
-   ylab("Total Hippocampal Volume (MAGeT Segmentation)")
+  geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
+  xlab(expression("Manual left / right hippocampus volume " (mm^3))) + 
+  ylab(expression("MAGeT left / right hippocampus volume "  (mm^3))) + 
+  scale_size_continuous(guide="none") 
 @
@@
   HC volumes for \mb against manually rated Hippocampal volumes}
-  \label{SZ_volumes}
+  \label{fig:SZ-volumes}
 \end{figure}
@@
 \begin{figure}
-<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes'>>=
+<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes', fig.width=4, fig.height=4>>=
+bland_altman_plot <- function(x, y) {
+  data = data.frame(x=x,y=y,diff=x-y)
+  upper = mean(data$diff) + 2 * sd(data$diff)
+  lower = mean(data$diff) - 2 * sd(data$diff)
+  return(ggplot(data, aes(x= x, y = diff)) + 
+    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
+               linetype=2, color='brown') + 
+    annotate("text", y = upper, x=Inf, 
+              label=paste("std =",round(upper,0)), 
+              hjust=1.1,vjust=1.5) + 
+    annotate("text", y = lower, x=Inf, 
+              label=paste("-std =",round(lower,0)), 
+              hjust=1.1,vjust=1.5) + 
+    geom_point())
+}
+
 bland_altman_plot(casted$Manual, casted$MAGeT) + 
   xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
-  ylab('Manual - MAGeT Brain volume') 
+  ylab('Manual - MAGeT Brain volume') + 
+  geom_smooth(method="lm", formula = y~x)
 @
@@
   when using five Winterburn atlases, and a 21 image template library.}
-  \label{}
+  \label{fig:FEP-SZ-Bland-Altman}
 \end{figure}
@@
 %%%%%
-\subsection{Winterburn Atlases Validation}
+\subsection{Winterburn Atlases Cross-Validation}
+
+The final experiment explores \mb segmentations of hippocampal subfields. To
+achieve this, a leave-one-out validation is conducted in which lower-resolution
+images ($0.9mm^3$ voxels) of each Winterburn atlas subject is segmented using 
+the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
+atlases when downsampled to $0.9mm^3$ voxels are also computed.
+
+In general, across hippocampal subregions the percent error in volume between 
+\mb segmentations and the manual Winterburn atlas segmentations compares 
+favourably to error when resampling the atlas segmentations (Figure 
+\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate subregions 
+all show near or smaller percent errors. The Sibiculum and CA2/CA3 subregions 
+show distinctly larger than resampling error.
 
@@
 # dirty dirty
-data = subset(read.csv('data/WAval.csv'), file != 'anusha')
-gold_0.3mm = subset(data, version == "gold_0.3mm")
-gold_0.9mm = subset(data, version == "gold_0.9mm")
-mb_bravo   = subset(data, version == "mb_bravo")
-mb_0.9mm   = subset(data, version == "mb_0.9mm")
-
-gold_0.3mm_total = gold_0.3mm[,2:6] + gold_0.3mm[,7:11]
-gold_0.9mm_total = gold_0.9mm[,2:6] + gold_0.9mm[,7:11]
-mb_bravo_total   = mb_bravo[,2:6] + mb_bravo[,7:11]
-mb_0.9mm_total   = mb_0.6mm[,2:6] + mb_0.9mm[,7:11]
-
-# calculate percent error from the gold standard measured at 0.3mm resolution
-pe_gold0.9mm = (gold_0.9mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
-pe_mbbravo   = (mb_bravo_total - gold_0.3mm_total)/gold_0.3mm_total * 100
-pe_mb0.9mm   = (mb_0.9mm_total - gold_0.3mm_total)/gold_0.3mm_total * 100
-
-subjects = gold_0.9mm[,1]
-pe_gold0.9mm = cbind(subjects,pe_gold0.9mm, res="Resampled")
-pe_mbbravo   = cbind(subjects,pe_mbbravo,   res="BRAVO")
-pe_mb0.6mm   = cbind(subjects,pe_mb0.9mm,   res="0.9mm voxels")
-
-pe = rbind(pe_mbbravo,pe_mb0.6mm, pe_gold0.6mm)
-names(pe) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
-melted = melt(pe, id.vars=c("subject", "res"))
-names(melted) <- c("subject","res", "region","percenterr")
-
-ggplot(melted, aes(y=percenterr,x=region)) +
-  geom_boxplot(aes(colour=res)) +
+data = subset(read.csv('data/WAval.csv'))
+casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
+casted = rename(casted, c("variable"="region"))
+
+error = ddply(casted, c("file", "region"), function(df) {
+  with(df, 
+  data.frame(
+    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
+    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
+    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
+
+levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
+                            "Subiculum"="X2","Subiculum"="X102", 
+                            "CA4"="X4","CA4"="X104", 
+                            "CA2/CA3"="X5","CA2/CA3"="X105", 
+                            "Dentate"="X6","Dentate"="X106")
+melted = melt(error,id.vars=c("file", "region"))
+melted = rename(melted, c("variable"="measure", "value"="percenterr"))
+levels(melted$measure) <- list("Resampled WA"="resampled",
+                               "MAGeT-Brain(Resampled WA)" = "mb_0.9mm",
+                               "MAGeT-Brain(BRAVO)"="mb_bravo")
+
+ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
+  geom_boxplot() +
   labs(y="Percent error in volume", x="Region") +
-  opts(legend.direction = "horizontal", legend.position = "bottom")
-@
-  \caption{{\bf Percent error in segmentation volume by Hippocampus subregion.}}
-  \label{WAval-vol-boxplot}
+  scale_colour_discrete("") + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Percent error in segmentation volume by hippocampus subregion.} 
+  {\em Resampled WA} are volumes of the manual segmentations of the Winterburn atlases 
+  after resampling to $0.9mm^3$.  {\em \mb(Resampled WA)} volumes are
+  \mb segmentations of the Winterburn atlas images after resampling to $0.9mm^3$ voxels.
+  {\em \mb(BRAVO) \mb} volumes are \mb segmentations of T1 BRAVO images ($0.9mm^3$ voxels)      
+  acquired separately 
+  of four of the five Winterburn atlas subjects. Percent error is measured against the volumes of
+  of the unmodified Winterburn atlas segmentations.}
+  \label{fig:WAval-vol-boxplot}
 \end{figure}

### 2013-04-11 paper.Rnw
Step 1 of a significant re-org of the paper structure -- squashing methods

@@
 \usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
+\usepackage{titlesec}
 \usepackage[round,authoryear]{natbib}
@@
 
+% add \subsubsubsection
+\setcounter{secnumdepth}{4}  
+\titleformat{\paragraph}
+{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+\titlespacing*{\paragraph}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+\newcommand{\subsubsubsection}[1]{\paragraph{#1}}
+
+% shortcuts 
 \newcommand{\mb}{MAGeT-Brain }
@@
 atlas subjects. For this experiment, we use the Winterburn atlases: digital
-hippocampal subfield segmentations of five in-vivo high-resolution (300u
+hippocampal subfield segmentations of five {\em in vivo} high-resolution (300$\mu$
 isotropic) T1-weighted MR scans\citep{Winterburn2013}. To validate \mb in a
@@
 
-\subsection{\mb Algorithm}
+\subsection{The \mb Algorithm}
 In this paper we will be making a distinction between an atlas and a template --
@@
 
-%%%%%                             Subjects                                 %%%%%
-
-\subsection{Subjects}
-Our experiments use three distinct subject datasets.
-
-\subsubsection{ADNI1 1.5T Complete 1Yr Dataset}
-Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
-standardized dataset available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly preprocessed images
-which have been designated to be the "best" after quality control.  All images
-were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
-Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
-= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
-x 240mm, a 192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel
-resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
-in table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.
-
-For a subset of ADNI1 images, labels of the left and right hippocampi are
-available (herein refered to as SNT labels).  Semi-automated hippocampal
-volumetry was carried out using a commercially available high dimensional brain
-mapping tool (Medtronic Surgical Navigation Technologies, Louisville, CO), that
-has previously been validated and compared to manual tracing of the hippocampus
-\citep{Hsu2002}. Measurement of hippocampal volume is achieved first by placing
-manually 22 control points as local landmarks for the hippocampus on the
-individual brain MRI data: one landmark at the hippocampal head, one at the
-tail, and four per image (i.e., at the superior, inferior, medial and lateral
-boundaries) on five equally spaced images perpendicular to the long axis of the
-hippocampus. Second, fluid image transformation is used to match the individual
-brains to a template brain \citep{Christensen1997}. The pixels corresponding to
-the hippocampus are then labeled and counted to obtain volumes. This method of
-hippocampal voluming has a documented reliability of an intraclass coefficient
-better than .94 \citep{Hsu2002}.
-
-%Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
-%flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
-%(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
-%$mm^3$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 Complete 1Yr Dataset Demographics Table                                  %
-%                                                                             %
-<<ADNI1-scr-demographics, echo=F, results="asis",cache=FALSE>>=
-# RID < 2000 ensures only ADNI1 patients
-yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
-yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
-yr1 = subset(yr1, !grepl("Scaled_2", Description))
-yr1 = yr1[,c("RID","VISCODE")]
-
-adnimerge_yr1 = merge(adnimerge, yr1) 
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, 
-               subset =  adnimerge_yr1, #RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
-               method = "reverse", test=FALSE, overall=TRUE)
-
-latex(tab, file="", size="scriptsize", 
-    caption="ADNI1 1.5T Complete 1Yr dataset demographics",
-    label="tab:ADN1-1.5T-Complete-1Yr-Dataset-Demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Schizophrenia First Episode Patients}
-\label{sec:SZFEP-dataset}
-All patients were recruited and treated through the Prevention and Early
-Intervention Program for Psychoses (PEPP-Montreal), a specialized early
-intervention service at the Douglas Mental Health University Institute in
-Montreal, Canada. People aged 15–30 years from the local catchment area
-suffering from either affective or non-affective psychosis who had not taken
-antipsychotic medication for more than one month with an IQ above 70 were
-consecutively admitted as either in- or out-patients. Of those treated at PEPP,
-only patients aged 18 to 30 years with no previous history of neurological
-disease or head trauma causing loss of consciousness were eligible for the
-neuroimaging study; only those suffering from schizophrenia spectrum disorders
-were considered for this analysis.  For complete program details see Malla et
-al. \citep{Malla2003}.
-
-Scanning was carried out at the Montreal Neurological Institute on a 1.5-T
-Siemens whole body MRI system.  Structural T1 volumes were acquired for each
-participant using a three-dimensional (3D) gradient echo pulse sequence with
-sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
-angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular field-of-view
-for the images was 256mm (SI)×204mm (AP).  Subject demographics are shown in
-table \ref{tab:SZFEP-Demographics}.
-
-The hippocampus were traced following a validated protocol developed by Dr Jens
-Pruessner (Pruessner et al., 2000). A recent update to this protocol by Dr J
-Pruessner in 2006 allows to accurately and consistently subdivide the
-hippocampus into three different subregions: head, body, and tail.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% First Episode SZ Demographics Table                                         %
-%
-
-<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
-fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
-fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
-tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
-               data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", ctable=FALSE,
-      caption="Schizophrenia First Episode Patient Demographics",
-      title="tab:SZFEP-Demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Winterburn Atlases}
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal segmentations
-of of five in-vivo 300u isotropic T1-weighted MR images. The segmentations
-include subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate
-gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases
-range in age from 29-57 years (mean age of 37), and include two males and three
-females.
-
-In addition to the high-resolution scans distributed as part of the Winterburn
-atlases, we also obtained additional T1 BRAVO scans of four of the five
-subjects.  \todo{demographics?}\todo{explain scan parameters in more detail..}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%                       Registration Methods                          %%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Registration Methods}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ANIMAL parameters                                                           %
-%                                                                             %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ANIMAL parameters                    
+%                                     
 \begin{table}[!tbp]
@@
 \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%                           Label Fusion                              %%%%%
-
-\subsection{Label Fusion}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Label fusion methods}
 Label fusion is a term given to the process of combining the information from
@@
 
-\subsubsection{Voxel-wise Majority Vote}
-
+\begin{description}
+  \item[Voxel-wise Majority Vote] \hfill \\
   Labels are propagated from all template library images to a target.  Each
@@
 
-\subsubsection{Cross-correlation Weighted Majority Vote}
-
+  \item[Cross-correlation Weighted Majority Vote] \hfill \\
   An optimal combination of targets from the template library has previously been
@@
 
-\subsubsection{Normalised Mutual Information Weighted Majority Vote}
-
+  \item[Normalised Mutual Information Weighted Majority Vote] \hfill \\
   This method is similar to cross-correlation weighted voting except that image
   similarity is calculated by the normalised mutual information score over the
-region of interest\citep{Studholme2001}.  The {\tt itk\_similarity} utility from
-the EZMinc toolkit\footnote{\url{https://github.com/vfonov/EZminc}} is used to
-calculate the normalised mutual information measure between to images.
-
-%%%%%                        Evaluation Measure                           %%%%%
-
-\subsection{Assessing Segmentation Agreement}
-
-The Dice similarity coefficient (DSC) assesses the agreement between two
-segmentations.  It is one of the most widely used measures of segmentation
-performance, and we use it as the basis of comparison throughout this paper.
-Additionally, we report the Jaccard index, another commonly used similarity
-measure:
-
- \[\text{Dice's coefficient (DSC)} = \frac{2|A \cap B|}{|A| + |B|}\]
-
- \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]
-
-where $A$ and $B$ are the regions being compared, and the cardinality is the
-volume measured in voxels.
-
-%%%%%
+  region of interest\citep{Studholme2001}.  The {\tt itk\_similarity} utility
+  from the EZMinc toolkit\footnote{\url{https://github.com/vfonov/EZminc}} is
+  used to calculate the normalised mutual information measure between to images.
+\end{description}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%                            Experiments                              %%%%%
-%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Experiments}
 
-The following experiments were performed to assess the performance of \mb with
-various parameter settings as well as on diverse datasets.  In each experiment
-we contrast the performance of \mb with that standard single- and multi-atlas
-segmentations derived from the same atlas library.
-
-%%%%%
-%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
-%%%%%
-\subsubsection{ADNI1 cross-validation}
+The following section describes the experiments we conducted to assess the
+segmentation quality of the \mb algorithm. The first two experiments assess the
+validity of \mb using a cross-validation design. Experiment 1 investigates the
+accuracy of \mb whole hippocampus segmentation over a wide range of parameter
+settings. This enables us to choose the parameter settings offering the best
+performance for use in subsequent experiments. Experiment 2 tests hippocampal
+subfield segmentation quality. The last two experiments assess the validity of
+the \mb algorithm when applied to different diseases: Alzheimer's disease
+(Experiment 3) and first episode schizophrenia patients (Experiment 4).
+
+%The ADNI experiment includes comparison of volumes with other automated
+%segmentation methods. Evaluation is done by correlating volumes b/c overlap
+%metrics assume identical segmentation protocol.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%         Experiment 1: Whole Hippocampus Cross-Validation            %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection*{Experiment 1: Whole Hippocampus Cross-Validation}
+
 To test the effect of parameter settings on the \mb algorithm, Monte
 Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
-the ADNI1 dataset for validation.  This form of cross-validation (described in
-detail below) allows us to validate \mb with various atlas and template
-library sizes, registration and label fusion methods.
-
-{\bf Dataset evaluated.} 69 1.5T images were arbitrarily selected from the {\em
-ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
-disease category (cognitively normal (CN), MCI and AD). Demographics for this
-subset are shown in Table \ref{tab:ADNI1-xvalidation-demographics}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+the ADNI1 dataset for validation (see below).  This form of cross-validation
+allows us to validate \mb with various atlas and template library sizes,
+registration and label fusion methods.
+
+Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
+cross-validation, consists of repeated rounds of validation in which items from
+the dataset are randomly sampled (without replacement) and assigned to a
+training set or validation set \cite{Shao1993}.
+
+In this experiment, each validation round consists of the following steps. First, 
+an atlas library is selected (the training set) and, from the remaining images, 
+a template library and subject to be segmented is chosen (the validation set). 
+Second, \mb and basic multi-atlas segmentation are performed on the subject using the 
+selected atlas and template libraries. Third, the accuracy of the resulting 
+segmentations are measured against the SNT labels.
+
+A total of ten validation rounds are perfomed on each subject in the dataset,
+for each combination of parameter settings. Reported subject segmentation
+accuracy is the averaged over the ten validation rounds. The parameter settings
+we explore are: atlas library size (1-9), template library size (1-20),
+registration method (\ants vs.  \animal), and label fusion method (MV, XC-WV,
+NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were conducted,
+resulting in a total of 1'490'400 segmentations analysed.
+
+\subsubsubsection{Dataset evaluated.}
+
+69 1.5T images were arbitrarily selected from the {\em ADNI1:Complete 1Yr 1.5T}
+standardized dataset. 23 subjects were chosen from each disease category
+(cognitively normal (CN), MCI and AD). Demographics for this subset are shown in
+Table \ref{tab:ADNI1-xvalidation-demographics}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 % ADNI1 X-Validation Dataset Demographics Table
@@
 @
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-{\bf Atlas and template libary.}  The atlas and template libraries are selected
-from the dataset evaluated.  Atlases consist of images taken from the
-dataset, with corresponding manual labels provided by SNT. Template library 
-images were selected at random from remaining images not used as atlases. The
-atlas library size ranged from 1 to 9 images. Template library size ranged from
-1 to 20 images.
-
-{\bf Registration method.} Both the \ants and \animal registration methods were
-used.
-
-{\bf Label fusion.} Majority vote, cross-correlation weighted majority vote, and
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The atlas and template libraries are selected from the dataset evaluated.
+Atlases consist of images taken from the dataset, with corresponding manual
+labels provided by SNT. Template library images were selected at random from
+remaining images not used as atlases. The atlas library size ranged from 1 to 9
+images. Template library size ranged from 1 to 20 images.
+
+Both the \ants and \animal registration methods were used.
+
+Majority vote, cross-correlation weighted majority vote, and
 Normalized Mutual Information weighted majority vote were used. With the
@@
 
-{\bf Evaluation.}  Monte Carlo Cross-Validation (MCCV), also known as repeated
-random sub-sampling cross-validation, consists of repeated rounds of validation 
-in which items from the dataset are randomly sampled (without replacement) and 
-assigned to a training set or validation set \cite{Shao1993}.
-
-In this experiment, each validation round consists of the following steps. First, 
-an atlas library is selected (the training set) and, from the remaining images, 
-a template library and subject to be segmented is chosen (the validation set). 
-Second, \mb and basic multi-atlas segmentation are performed on the subject using the 
-selected atlas and template libraries. Third, the accuracy of the resulting 
-segmentations are measured against the SNT labels.
-
-A total of ten validation rounds are perfomed on each subject in the dataset,
-for each combination of parameter settings. Reported subject segmentation
-accuracy is the averaged over the ten validation rounds. The parameter settings
-we explore are: atlas library size (1-9), template library size (1-20),
-registration method (\ants vs.  \animal), and label fusion method (MV, XC-WV,
-NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were conducted,
-resulting in a total of 1'490'400 segmentations analysed.
-
-%%%%%
-%%%%%               Experiments: ADNI-1 Complete 1Yr Validation               %%%%%
-%%%%%
-\subsubsection{ADNI-1 Complete 1Yr Validation}
+\subsubsubsection{Evaluation method}
+
+The Dice similarity coefficient (DSC) assesses the agreement between two
+segmentations.  It is one of the most widely used measures of segmentation
+performance, and we use it as the basis of comparison throughout this paper.
+Additionally, we report the Jaccard index, another commonly used similarity
+measure:
+
+ \[\text{Dice's coefficient (DSC)} = \frac{2|A \cap B|}{|A| + |B|}\]
+
+ \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]
+
+where $A$ and $B$ are the regions being compared, and the cardinality is the
+volume measured in voxels.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%        Experiment 2: Hippocampal Subfield Cross-Validation          %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Winterburn Atlases Cross-Validation}
+In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
+segmentation is tested using a leave-one-out cross-validation (LOOCV) design.
+Leave-one-out-cross-validation (LOOCV) is performed separately for each dataset
+evaluated (WA-BRAVO, and WA-Resampled) as follows. Each subject in the dataset
+is segmented by \mb using an atlas library in which that subject's image is
+excluded.
+
+The atlas library is composed of the Winterburn T1 atlas images and hippocampal
+subregion segmentations. The template library consists of 15 3T T1 images
+(0.9mm-isotropic voxels) of healthy subjects as well as all of the images from
+the dataset being evaluated.
+
+- note registration and fusion method were chosen 
+\todo{reference, demographics for health control images}.
+
+\subsubsubsection{Dataset evaluated} 
+\mb segmentations are evaluated on two datasets. The WA-BRAVO dataset consists
+of separately acquired 3T T1 BRAVO images (0.9mm-isotropic voxel) of four of the
+five Winterburn atlas subjects.  The WA-Resampled dataset consists of the
+Winterburn atlas images and segmentations downsampled to 0.9mm-isotropic voxel
+resolution.
+
+\subsubsubsection{Evaluation method}  
+The segmentation volumes of each hippocampal subregion are compared to
+the expert manual segmentations of the unmodified Winterburn atlases.  As a
+point of comparison, we also calculate the subregion volumes from the Winterburn
+atlas segmentations after downsampling to 0.9mm-isotropic voxels.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%         Experiment 3: Application of \mb to the segmentation        %%%%%
+%%%%%                       of schizophrenia patients                     %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{SZ First Episode Patient Validation}
+To validate that \mb algorithm works effectively with diseased brain images the
+performance of \mb is measured on a Schizophrenia dataset.
+
+The atlas library is composed of the Winterburn T1 atlas images, with whole
+hippocampus segmentations (no subregions). The template library is composed of
+21 images selected at random from the SZFEP dataset.
+
+- note registration and fusion method were chosen 
+\subsubsubsection{Dataset Evaluated}
+
+\subsubsubsection{Evaluation method}
+The manual segmentation protocol used to segment the Winterburn atlases is
+similar to, but different from, the protocol used to segment the SZFEP dataset.
+Therefore, rather than use an overlap metric, \mb hippocampal volumes are
+compared to the corresponding manual segmentation volumes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%         EXPERIMENT 3: Application of \mb to the segmentation        %%%%%
+%%%%%                       of ADNI subjects                              %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Experiment 4: Segmentation of Alzheimer's Disease Patients}
+
 To test the accuracy of \mb on a real-world task we segment the entire
@@
 
-{\bf Dataset evaluated.} All images from the {\em ADNI1:Complete 1Yr 1.5T}
+The Winterburn atlases were used as the atlas
+library.  The template library consisted of 21 randomly selected images from the
+ADNI1 data dataset (7 healthy, MCI and AD subjects).
+
+- note registration and fusion method were chosen
+
+\ref{tab:ADNI1-scr-tmpllib-demographics}.
+
+\subsubsubsection{Dataset evaluated} 
+All images from the {\em ADNI1:Complete 1Yr 1.5T}
 standardized dataset.
 
-{\bf Atlas and template libary.} The Winterburn atlases were used as the atlas
-library.  The template library consisted of 21 randomly selected images from the
-ADNI1 data dataset (7 healthy, MCI and AD subjects). Clinical and demographic
-data for the template library subjects are shown in Table
-\ref{tab:ADNI1-scr-tmpllib-demographics}.
-
-{\bf Registration method.} The \ants non-linear registration algorithm was used,
-as it performed best in the ADNI1 cross-validation experiment (see Results).
-
-{\bf Label fusion.} Majority vote was used performed equally well in the ADNI1
-cross-validation experiment and is the easiest label fusion method to operate.
-
-{\bf Evaluation.}
+\subsubsubsection{Evaluation method}
 Since the hippocampal segmentation protocols differ between the ADNI labels and
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 % ADNI1 Complete 1Yr Template Library Demographics Table
 %
-<<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
-scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
-                 0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
-                 0159)
+%
+% <<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
+% scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
+%                  0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
+%                  0159)
+% tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+%                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
+%                scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+% 
+% latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
+%       caption="ADNI1 Complete 1Yr Template Library demographics", 
+%       title="tab:ADNI1-scr-tmpllib-demographics")
+% @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%                             Subjects                                 %%%%%
+\subsection{Subjects}
+
+\subsubsection{Winterburn Atlases}
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal segmentations
+of of five in-vivo 300u isotropic T1-weighted MR images. The segmentations
+include subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate
+gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
+
+In addition to the high-resolution scans distributed as part of the Winterburn
+atlases, we also obtained additional T1 BRAVO scans of four of the five
+subjects.  \todo{demographics?}\todo{explain scan parameters in more detail..}
+
+\subsubsubsection{FEP dataset}
+All patients were recruited and treated through the Prevention and Early
+Intervention Program for Psychoses (PEPP-Montreal), a specialized early
+intervention service at the Douglas Mental Health University Institute in
+Montreal, Canada. People aged 15–30 years from the local catchment area
+suffering from either affective or non-affective psychosis who had not taken
+antipsychotic medication for more than one month with an IQ above 70 were
+consecutively admitted as either in- or out-patients. Of those treated at PEPP,
+only patients aged 18 to 30 years with no previous history of neurological
+disease or head trauma causing loss of consciousness were eligible for the
+neuroimaging study; only those suffering from schizophrenia spectrum disorders
+were considered for this analysis.  For complete program details see Malla et
+al. \citep{Malla2003}.
+
+Scanning was carried out at the Montreal Neurological Institute on a 1.5-T
+Siemens whole body MRI system.  Structural T1 volumes were acquired for each
+participant using a three-dimensional (3D) gradient echo pulse sequence with
+sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
+angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular
+field-of-view for the images was 256mm (SI)×204mm (AP).  Subject demographics
+are shown in table \ref{tab:SZFEP-Demographics}.
+
+The hippocampus were traced following a validated protocol developed by Dr Jens
+Pruessner \todo{cite: (Pruessner et al., 2000)}. A recent update to this
+protocol by Dr J Pruessner in 2006 allows to accurately and consistently
+subdivide the hippocampus into three different subregions: head, body, and tail.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% First Episode SZ Demographics Table
+%
+<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
+fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
+fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
+tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
+               data=fep_sz_demographics, method = "reverse", test = FALSE)
+latex(tab, file="", ctable=FALSE,
+      caption="Schizophrenia First Episode Patient Demographics",
+      title="tab:SZFEP-Demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{ADNI1 1.5T Complete 1Yr Dataset}
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
+standardized dataset available from ADNI \footnote{
+\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly preprocessed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
+= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
+x 240mm, a 192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel
+resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
+in table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.
+
+% The following blurb is taken (except for the first sentence) verbatim 
+For a subset of ADNI1 images, labels of the left and right hippocampi are
+available (herein refered to as SNT labels).  Semi-automated hippocampal
+volumetry was carried out using a commercially available high dimensional brain
+mapping tool (Medtronic Surgical Navigation Technologies, Louisville, CO), that
+has previously been validated and compared to manual tracing of the hippocampus
+\citep{Hsu2002}. Measurement of hippocampal volume is achieved first by placing
+manually 22 control points as local landmarks for the hippocampus on the
+individual brain MRI data: one landmark at the hippocampal head, one at the
+tail, and four per image (i.e., at the superior, inferior, medial and lateral
+boundaries) on five equally spaced images perpendicular to the long axis of the
+hippocampus. Second, fluid image transformation is used to match the individual
+brains to a template brain \citep{Christensen1997}. The pixels corresponding to
+the hippocampus are then labeled and counted to obtain volumes. This method of
+hippocampal voluming has a documented reliability of an intraclass coefficient
+better than .94 \citep{Hsu2002}.
+
+%Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
+%flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
+%(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
+%$mm^3$.
+
+%                                                                             
+%                                                                             
+% ADNI1 Complete 1Yr Dataset Demographics Table 
+%                                                                             
+%                                                                             
+<<ADNI1-scr-demographics, echo=F, results="asis",cache=FALSE>>=
+# RID < 2000 ensures only ADNI1 patients
+yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
+yr1 = subset(yr1, !grepl("Scaled_2", Description))
+yr1 = yr1[,c("RID","VISCODE")]
+
+adnimerge_yr1 = merge(adnimerge, yr1) 
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-               scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-
-latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
-      caption="ADNI1 Complete 1Yr Template Library demographics", 
-      title="tab:ADNI1-scr-tmpllib-demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%
-%%%%%          Experiments: SZ First Episode Patient Validation           %%%%%
-%%%%%
-\subsubsection{SZ First Episode Patient Validation}
-The previous experiments explore \mb on an Alzheimer's disease dataset (ADNI).  To validate
-that \mb algorithm works effectively with other diseased brain images, 
-the performance of \mb is measured on a Schizophrenia dataset.
-
-{\bf Dataset evaluated.} The entire Schizophrenia First Episode Patient (SZFEP) 
-dataset (see section \ref{sec:SZFEP-dataset}).
-
-{\bf Atlas and template libary.} The atlas library is composed of the Winterburn 
-T1 atlas images, with whole hippocampus segmentations (no subregions). The 
-template library is composed of 21 images selected at random from the SZFEP dataset.
-
-{\bf Registration method.} The \ants registration algorithm is used for image
-registration.
-
-{\bf Label fusion.} The majority-vote method is used for label fusion.
-
-{\bf Evaluation.} The manual segmentation protocol used to segment the Winterburn
-atlases is similar to, but different from, the protocol used to segment the SZFEP
-dataset. Therefore, rather than use an overlap metric, \mb hippocampal volumes
-are compared to the corresponding manual segmentation volumes.
-
-%%%%%
-%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
-%%%%%
-\subsubsection{Winterburn Atlases Cross-Validation}
-In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
-segmentation is tested using a leave-one-out cross-validation (LOOCV) design.
-
-{\bf Dataset evaluated.} \mb segmentations are evaluated on two datasets. The 
-WA-BRAVO dataset consists of separately acquired 3T T1 BRAVO images (0.9mm-isotropic voxel)
-of four of the five Winterburn atlas subjects. The WA-Resampled dataset consists
-of the Winterburn atlas images and segmentations downsampled to 0.9mm-isotropic 
-voxel resolution.
-
-{\bf Atlas and template libary.} The atlas library is composed of the Winterburn 
-T1 atlas images and hippocampal subregion segmentations. The template library 
-consists of 15 3T T1 images (0.9mm-isotropic voxels) of healthy subjects as well
-as all of the images from the dataset being evaluated.
-
-\todo{reference, demographics for health control images}.
-
-{\bf Registration method.} The \ants image registration method is used.
-
-{\bf Label fusion.} The majority vote label fusion method is used.
-
-{\bf Evaluation.}  Leave-one-out-cross-validation (LOOCV) is performed separately 
-for each dataset evaluated (WA-BRAVO, and WA-Resampled) as follows. Each subject
-in the dataset is segmented by \mb using an atlas library in which that subject's 
-image is excluded. The segmentation volumes of each hippocampal subregion are
-compared to the expert manual segmentations of the unmodified Winterburn atlases. 
-As a point of comparison, we also calculate the subregion volumes from the Winterburn 
-atlas segmentations after downsampling to 0.9mm-isotropic voxels.
+               MMSE, data = adnimerge, 
+               subset =  adnimerge_yr1, #RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
+               method = "reverse", test=FALSE, overall=TRUE)
+
+latex(tab, file="", size="scriptsize", 
+    caption="ADNI1 1.5T Complete 1Yr dataset demographics",
+    label="tab:ADN1-1.5T-Complete-1Yr-Dataset-Demographics")
+@ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %                                                                              %
-<<ADN1-xval-load, include=FALSE, cache=TRUE>>=
+<<ADNI1-xval-load, include=FALSE, cache=TRUE>>=
 # all data from the validation trials
@@
 
-<<ADN1-xval-prep, include=FALSE, cache=TRUE>>=
+<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
 # simplify the data
@@
 all_data_mean  = merge(all_data_mean, diagnoses)
-#remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
+remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
 @
@@
 
+\todo{Include multi-atlas 9-atlas in plot as a comparison}
+
 \begin{figure}
@@
 combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, qc          , by=c("Image.Data.ID"), all.x=TRUE)
+combined = merge(combined, qc          , by=c("Image.Data.ID", "RID"), all.x=TRUE)
 
@@
 
-all_qc = subset(all_qc, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )
+all_qc = subset(combined, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )
 
@@
 
+\begin{figure}
+  \includegraphics[width=6in,height=3in]{figure/ADNI1_MAGeT_montage/figure.pdf}
+  \caption{A caption}
+  \label{fig:ADNI1-segmentation-montage}
+\end{figure}
+
 % show as a plot

### 2013-04-11 paper.Rnw
Major reorg of the methods is complete

@@
 \usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
-\usepackage{titlesec}
+\usepackage{titlesec, siunitx}
 \usepackage[round,authoryear]{natbib}
@@
 % add \subsubsubsection
-\setcounter{secnumdepth}{4}  
-\titleformat{\paragraph}
-{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-\titlespacing*{\paragraph}
-{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+% \setcounter{secnumdepth}{4}  
+% \titleformat{\paragraph}
+% {\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+% \titlespacing*{\paragraph}
+% {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 \newcommand{\subsubsubsection}[1]{\paragraph{#1}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Experiment 1: Whole Hippocampus Cross-Validation}
-
-To test the effect of parameter settings on the \mb algorithm, Monte
-Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
-the ADNI1 dataset for validation (see below).  This form of cross-validation
-allows us to validate \mb with various atlas and template library sizes,
-registration and label fusion methods.
-
-Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
-cross-validation, consists of repeated rounds of validation in which items from
-the dataset are randomly sampled (without replacement) and assigned to a
-training set or validation set \cite{Shao1993}.
-
-In this experiment, each validation round consists of the following steps. First, 
-an atlas library is selected (the training set) and, from the remaining images, 
-a template library and subject to be segmented is chosen (the validation set). 
-Second, \mb and basic multi-atlas segmentation are performed on the subject using the 
-selected atlas and template libraries. Third, the accuracy of the resulting 
-segmentations are measured against the SNT labels.
-
-A total of ten validation rounds are perfomed on each subject in the dataset,
-for each combination of parameter settings. Reported subject segmentation
-accuracy is the averaged over the ten validation rounds. The parameter settings
-we explore are: atlas library size (1-9), template library size (1-20),
-registration method (\ants vs.  \animal), and label fusion method (MV, XC-WV,
-NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were conducted,
-resulting in a total of 1'490'400 segmentations analysed.
-
-\subsubsubsection{Dataset evaluated.}
-
-69 1.5T images were arbitrarily selected from the {\em ADNI1:Complete 1Yr 1.5T}
-standardized dataset. 23 subjects were chosen from each disease category
-(cognitively normal (CN), MCI and AD). Demographics for this subset are shown in
-Table \ref{tab:ADNI1-xvalidation-demographics}.
+\subsubsection{Experiment 1: Whole Hippocampus Cross-Validation}
+
+Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
+of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
+cross-validation allows us to rigorously validate a large number of parameter
+settings of \mb.
+
+\paragraph{ADNI1 dataset}
+
+69 1.5T images were arbitrarily selected from the baseline scans in the {\em
+ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
+disease category: cognitively normal (CN), mild cognitive impairment (MCI) and
+Alzheimer's disease (AD). Demographics for this subset are shown in Table
+\ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
+right whole hippocampi are available. These labels have been generated using a
+semiautomatic tool (SNT, Medtronic Surgical Navigation Technologies, Louisville,
+CO; see Supplementary Materials).
+
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
+standardized dataset available from ADNI \footnote{
+\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly preprocessed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
+= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
+x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.
 
@@
 
-The atlas and template libraries are selected from the dataset evaluated.
-Atlases consist of images taken from the dataset, with corresponding manual
-labels provided by SNT. Template library images were selected at random from
-remaining images not used as atlases. The atlas library size ranged from 1 to 9
-images. Template library size ranged from 1 to 20 images.
-
-Both the \ants and \animal registration methods were used.
-
-Majority vote, cross-correlation weighted majority vote, and
-Normalized Mutual Information weighted majority vote were used. With the
-weighted majority vote fusion methods, the number of top labels used in the
-fusion was varied from 1 to 20 images.
-
-\subsubsubsection{Evaluation method}
+\paragraph{Experiment details}
+
+Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
+cross-validation, consists of repeated rounds of validation conducted on a fixed
+dataset \cite{Shao1993}. In each round, the dataset is randomly partitioned into
+a training set and a validation set. The method to be validated is then given
+the training data, and its output is compared with the validation set.
+
+In this experiment, our dataset consists of 69 1.5T images and corresponding
+manual segmentations taken from the ADNI1 project (see below for details). In each
+validation round, the training set is used as an atlas library for both \mb and
+basic multi-atlas. The remaining images are segmented by both methods, and the
+resulting segmentations are compared to the manual segmentations for the images.
+
+A total of ten validation rounds are perfomed on each subject in the dataset,
+over each combination of parameter settings. The parameter settings
+we explore are: atlas library size (1-9), template library size (1-20),
+registration method (\ants or \animal), and label fusion method (majority vote,
+cross-correlation weighted majority vote, and normalized mutual information
+weighted majority vote).  A total of $10 \times 69 \times 9 \times 20 \times 2
+\times 3 = \num{7452000}$ validation rounds were conducted, resulting in a total
+of $\num{1490400}$ segmentations analysed.
+
+\paragraph{Evaluation method}
 
@@
 segmentations. It is one of the most widely used measures of segmentation
-performance, and we use it as the basis of comparison throughout this paper.
+performance, and we use it as the basis of comparison in this experiment.
 Additionally, we report the Jaccard index, another commonly used similarity
@@
 volume measured in voxels.
+
+Reported subject segmentation accuracy is averaged over the ten validation
+rounds for each parameter setting.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Winterburn Atlases Cross-Validation}
+\subsubsection{Experiment 2: Winterburn Atlases Cross-Validation}
 In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
 segmentation is tested using a leave-one-out cross-validation (LOOCV) design.
-Leave-one-out-cross-validation (LOOCV) is performed separately for each dataset
-evaluated (WA-BRAVO, and WA-Resampled) as follows. Each subject in the dataset
-is segmented by \mb using an atlas library in which that subject's image is
-excluded.
-
-The atlas library is composed of the Winterburn T1 atlas images and hippocampal
-subregion segmentations. The template library consists of 15 3T T1 images
-(0.9mm-isotropic voxels) of healthy subjects as well as all of the images from
-the dataset being evaluated.
-
-- note registration and fusion method were chosen 
-\todo{reference, demographics for health control images}.
-
-\subsubsubsection{Dataset evaluated} 
-\mb segmentations are evaluated on two datasets. The WA-BRAVO dataset consists
-of separately acquired 3T T1 BRAVO images (0.9mm-isotropic voxel) of four of the
-five Winterburn atlas subjects.  The WA-Resampled dataset consists of the
-Winterburn atlas images and segmentations downsampled to 0.9mm-isotropic voxel
-resolution.
-
-\subsubsubsection{Evaluation method}  
-The segmentation volumes of each hippocampal subregion are compared to
-the expert manual segmentations of the unmodified Winterburn atlases.  As a
-point of comparison, we also calculate the subregion volumes from the Winterburn
-atlas segmentations after downsampling to 0.9mm-isotropic voxels.
+The optimal parameter settings found in Experiment 1 are used.
+
+\paragraph{Winterburn Atlases dataset} 
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+segmentations of five in-vivo $300\mu$ isotropic T1-weighted MR images. The
+segmentations include subfield segmentations for the cornus ammonis (CA) 1, CA4,
+dentate gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn
+atlases range in age from 29-57 years (mean age of 37), and include two males
+and three females.
+
+In addition to the high-resolution scans distributed as part of the Winterburn
+atlases, we also obtained additional 3T T1 BRAVO images (0.9mm-isotropic voxels)
+of four of the five Winterburn atlas subjects.
+
+\paragraph{Experiment details} 
+Leave-one-out cross-validation (LOOCV) is an approach in which the method to be
+validated is given all but one item in a dataset as training data, and the
+output is compared with the left-out item. This is done, in turn, for each item
+in the dataset.
+
+In this experiment, the five $300\mu$-isotropic voxel Winterburn atlases are
+used as the atlas library for \mb segmentation. LOOCV is conducted separately
+for two different input datasets: the four Winterburn atlas subject BRAVO MR
+images (referred to as the {\em Winterburn BRAVO} dataset), and the five
+Winterburn atlas subject images resampled to 0.9mm-isotropic voxel resolution
+(referred to as the {\em Winterburn Resampled} dataset. Each subject in the
+dataset is segmented by \mb with that subject's image excluded from the atlas
+library. The template library consists of 3T T1 images (0.9mm-isotropic
+voxels) of healthy subjects in addition to the images from the dataset being
+evaluated.
+
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used.
+
+\paragraph{Evaluation method}  
+The segmentation volumes of each hippocampal subregion are compared to the
+expert manual segmentations of the unmodified Winterburn atlases.  As a point of
+comparison, we also calculate the subregion volumes from the Winterburn atlas
+segmentations after downsampling to 0.9mm-isotropic voxels.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{SZ First Episode Patient Validation}
-To validate that \mb algorithm works effectively with diseased brain images the
-performance of \mb is measured on a Schizophrenia dataset.
-
-The atlas library is composed of the Winterburn T1 atlas images, with whole
-hippocampus segmentations (no subregions). The template library is composed of
-21 images selected at random from the SZFEP dataset.
-
-- note registration and fusion method were chosen 
-\subsubsubsection{Dataset Evaluated}
-
-\subsubsubsection{Evaluation method}
-The manual segmentation protocol used to segment the Winterburn atlases is
-similar to, but different from, the protocol used to segment the SZFEP dataset.
-Therefore, rather than use an overlap metric, \mb hippocampal volumes are
-compared to the corresponding manual segmentation volumes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%         EXPERIMENT 3: Application of \mb to the segmentation        %%%%%
-%%%%%                       of ADNI subjects                              %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 4: Segmentation of Alzheimer's Disease Patients}
-
-To test the accuracy of \mb on a real-world task we segment the entire
-ADNI-1 dataset using an atlas set that is not representative of the target set.
-
-The Winterburn atlases were used as the atlas
-library.  The template library consisted of 21 randomly selected images from the
-ADNI1 data dataset (7 healthy, MCI and AD subjects).
-
-- note registration and fusion method were chosen
-
-\ref{tab:ADNI1-scr-tmpllib-demographics}.
-
-\subsubsubsection{Dataset evaluated} 
-All images from the {\em ADNI1:Complete 1Yr 1.5T}
-standardized dataset.
-
-\subsubsubsection{Evaluation method}
-Since the hippocampal segmentation protocols differ between the ADNI labels and
-Winterburn atlases, this poses a problem for direct evaluation between labels
-produced by \mb and the ADNI labels in terms of overlap; we would not expect
-different segmentation protocols to have a high degree of overlap.  Instead, to
-evaluate the performance of \mb we compare the correlation of \mb segmentation
-volumes with manual segmentation (SNT) volumes.  Additionally, we correlate the
-hippocampal volumes of established automated segmentation methods to \mb
+\subsubsection{Experiment 3: Application to the segmentation first episode
+schizophrenia patients}
+
+To validate that \mb algorithm works effectively with diseased brain images \mb
+is applied to a dataset of Schizophrenia patient MR images. The resulting
+segmentations are assessed for quality by comparison with expert manual
 segmentations.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% ADNI1 Complete 1Yr Template Library Demographics Table
-%
-%
-% <<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
-% scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
-%                  0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
-%                  0159)
-% tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-%                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-%                scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-% 
-% latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
-%       caption="ADNI1 Complete 1Yr Template Library demographics", 
-%       title="tab:ADNI1-scr-tmpllib-demographics")
-% @
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%                             Subjects                                 %%%%%
-\subsection{Subjects}
-
-\subsubsection{Winterburn Atlases}
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal segmentations
-of of five in-vivo 300u isotropic T1-weighted MR images. The segmentations
-include subfield segmentations for the cornus ammonis (CA) 1, CA4, dentate
-gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn atlases
-range in age from 29-57 years (mean age of 37), and include two males and three
-females.
-
-In addition to the high-resolution scans distributed as part of the Winterburn
-atlases, we also obtained additional T1 BRAVO scans of four of the five
-subjects.  \todo{demographics?}\todo{explain scan parameters in more detail..}
-
-\subsubsubsection{FEP dataset}
+\paragraph{SZ-FEP dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 
-Scanning was carried out at the Montreal Neurological Institute on a 1.5-T
-Siemens whole body MRI system.  Structural T1 volumes were acquired for each
-participant using a three-dimensional (3D) gradient echo pulse sequence with
-sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
+Scanning of 81 subjects was carried out at the Montreal Neurological Institute
+on a 1.5-T Siemens whole body MRI system.  Structural T1 volumes were acquired
+for each participant using a three-dimensional (3D) gradient echo pulse sequence
+with sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
 angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular
-field-of-view for the images was 256mm (SI)×204mm (AP).  Subject demographics
-are shown in table \ref{tab:SZFEP-Demographics}.
+field-of-view for the images was 256mm (SI)$\times$204mm (AP).  Subject
+demographics are shown in table \ref{tab:SZFEP-Demographics}.
 
@@
 subdivide the hippocampus into three different subregions: head, body, and tail.
+
+\paragraph{Experiment details} 
+\mb is configured with an atlas library composed of the Winterburn T1 atlases
+(see Experiment 2).  All images from the SZ-FEP dataset are segmented.  The optimal
+size of template library, registration method, and label fusion method found in
+Experiment 1 are used.
+
+\paragraph{Evaluation method}
+The manual segmentation protocol used to segment the Winterburn atlases is
+similar to, but different from, the protocol used to segment the SZFEP dataset.
+Therefore, rather than use an overlap metric, \mb hippocampal volumes are
+compared to the corresponding manual segmentation volumes.
+
+\todo{discuss Bland-Altman?  Or bias analysis}
 
@@
 
-\subsubsection{ADNI1 1.5T Complete 1Yr Dataset}
-Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
-standardized dataset available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly preprocessed images
-which have been designated to be the "best" after quality control.  All images
-were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
-Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
-= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
-x 240mm, a 192 x 192 x 166 matrix (x, y, and z directions) yielding a voxel
-resolution of 1.25 x 1.25 x 1.2 $mm^3$. Clinical and demographic data are shown
-in table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.
-
-% The following blurb is taken (except for the first sentence) verbatim 
-For a subset of ADNI1 images, labels of the left and right hippocampi are
-available (herein refered to as SNT labels).  Semi-automated hippocampal
-volumetry was carried out using a commercially available high dimensional brain
-mapping tool (Medtronic Surgical Navigation Technologies, Louisville, CO), that
-has previously been validated and compared to manual tracing of the hippocampus
-\citep{Hsu2002}. Measurement of hippocampal volume is achieved first by placing
-manually 22 control points as local landmarks for the hippocampus on the
-individual brain MRI data: one landmark at the hippocampal head, one at the
-tail, and four per image (i.e., at the superior, inferior, medial and lateral
-boundaries) on five equally spaced images perpendicular to the long axis of the
-hippocampus. Second, fluid image transformation is used to match the individual
-brains to a template brain \citep{Christensen1997}. The pixels corresponding to
-the hippocampus are then labeled and counted to obtain volumes. This method of
-hippocampal voluming has a documented reliability of an intraclass coefficient
-better than .94 \citep{Hsu2002}.
-
-%Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
-%flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
-%(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
-%$mm^3$.
-
-%                                                                             
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%         EXPERIMENT 3: Application of \mb to the segmentation        %%%%%
+%%%%%                       of ADNI subjects                              %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Experiment 4: Application to the segmentation of Alzheimer's
+disease patients}
+
+To validate that \mb algorithm works as well as established automated methods,
+\mb is applied to the ADNI1 dataset and the resulting segmentations are
+compared to those produced by FreeSurfer, FSL, MAPER, and by expert manual
+segmentation.
+
+\paragraph{ADNI1 dataset revisited} 
+All images from the {\em ADNI1:Complete 1Yr 1.5T} standardized dataset described
+in Experiment 1 are used.  Clinical and demographic data are shown in table
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.
+
+\paragraph{Experiment details} 
+\mb is configured with an atlas library composed of the Winterburn T1 atlases
+(see Experiment 2). All images from the ADN11 Complete 1Yr 1.5T dataset are
+segmented. The optimal size of template library, registration method, and label
+fusion method found in Experiment 1 are used. The template library is composed
+of equal numbers of images from each disease class (AD, MCI, and cognitively
+normal controls).
+
+\paragraph{Evaluation method}
+Since the hippocampal segmentation protocols differ between the ADNI labels and
+Winterburn atlases, this poses a problem for direct evaluation between labels
+produced by \mb and the ADNI labels in terms of overlap; we would not expect
+different segmentation protocols to have a high degree of overlap.  Instead, to
+evaluate the performance of \mb we compare the correlation of \mb segmentation
+volumes with manual segmentation (SNT) volumes. Additionally, we correlate the
+hippocampal volumes of established automated segmentation methods to \mb
+segmentations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                                                                             
 % ADNI1 Complete 1Yr Dataset Demographics Table 
-%                                                                             
 %                                                                             
@@
 
-adnimerge_yr1 = merge(adnimerge, yr1) 
+adnimerge.yr1 = merge(adnimerge, yr1) 
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, 
-               subset =  adnimerge_yr1, #RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
+               MMSE, data = adnimerge.yr1, 
+               #subset = RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
                method = "reverse", test=FALSE, overall=TRUE)
@@
     label="tab:ADN1-1.5T-Complete-1Yr-Dataset-Demographics")
-@ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\section{Supplementary Materials}
+
+\subsection{ADNI Manual Labels}
+% The following blurb is taken (except for the first sentence) verbatim 
+Semi-automated hippocampal volumetry was carried out using a commercially
+available high dimensional brain mapping tool (Medtronic Surgical Navigation
+Technologies, Louisville, CO), that has previously been validated and compared
+to manual tracing of the hippocampus \citep{Hsu2002}. Measurement of hippocampal
+volume is achieved first by placing manually 22 control points as local
+landmarks for the hippocampus on the individual brain MRI data: one landmark at
+the hippocampal head, one at the tail, and four per image (i.e., at the
+superior, inferior, medial and lateral boundaries) on five equally spaced images
+perpendicular to the long axis of the hippocampus. Second, fluid image
+transformation is used to match the individual brains to a template brain
+\citep{Christensen1997}. The pixels corresponding to the hippocampus are then
+labeled and counted to obtain volumes. This method of hippocampal voluming has a
+documented reliability of an intraclass coefficient better than .94
+\citep{Hsu2002}.
+
+%Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
+%flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
+%(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
+%$mm^3$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% ADNI1 Complete 1Yr Template Library Demographics Table
+%
+%
+% <<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
+% scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
+%                  0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
+%                  0159)
+% tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+%                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
+%                scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+% 
+% latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
+%       caption="ADNI1 Complete 1Yr Template Library demographics", 
+%       title="tab:ADNI1-scr-tmpllib-demographics")
+% @
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{plainnat}

### 2013-04-13 paper.Rnw
Fixed results section order, and exported some analysis code to files

@@
 \usepackage[round,authoryear]{natbib}
+\usepackage[section]{placeins}  % keep floats in their place.
 
@@
 for two different input datasets: the four Winterburn atlas subject BRAVO MR
-images (referred to as the {\em Winterburn BRAVO} dataset), and the five
+images (referred to as the {\em WA BRAVO} dataset), and the five
 Winterburn atlas subject images resampled to 0.9mm-isotropic voxel resolution
-(referred to as the {\em Winterburn Resampled} dataset. Each subject in the
+(referred to as the {\em WA Resampled} dataset. Each subject in the
 dataset is segmented by \mb with that subject's image excluded from the atlas
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                               Results
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                               RESULTS                                        %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%
-%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
-%%%%%
-\subsection{ADNI1 Cross-Validation}
-
-%                                                                              %
-% ADNI1 Complete 1Yr Validation - results preparation                             %
-%                                                                              %
-<<ADNI1-xval-load, include=FALSE, cache=TRUE>>=
-# all data from the validation trials
-a_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
-#t_data       = read.csv(gzfile('data/a2a_tracc/xfmtrials_results_2013-04-01.csv.gz'))
-t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz'))
-diagnoses    = read.csv('data/a2a_diagnoses.csv')
-@
-
-<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-# simplify the data
-t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
-a_data       = subset(a_data, atlases %in% c(1,3,5,7,9))
-t_data$se    = NULL
-t_data$sn    = NULL
-t_data$j     = NULL
-a_data$se    = NULL
-a_data$sn    = NULL
-a_data$j     = NULL
-
-# add registration method column 
-t_data       = cbind(t_data, reg_method = "ANIMAL")
-a_data       = cbind(a_data, reg_method = "ANTS")
-
-# separate MA majvote from MB so we can compute differences
-# we only compare multiatlas using majority vote
-t_ma = subset(t_data, approach=="ma" & method=="majvote")
-t_ma$approach="multiatlas"
-a_ma = subset(a_data, approach=="multiatlas" & method=="majvote")
-t_mb = subset(t_data, approach=="mb")
-a_mb = subset(a_data, approach=="mb")
-
-ma = rbind(t_ma,a_ma)
-mb = rbind(t_mb,a_mb)
-
-# set up a few equivalences for multi-atlas, to make nomenclature simpler
-ma$templates =  ma$atlases  # because no template library
-mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n
-
-# aggregate over batches
-ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases + templates + top_n + label, data = ma, mean)
-mb_mean = aggregate(k ~ subject + reg_method + approach + method + atlases + templates + top_n + label, data = mb, mean)
-
-all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"), 
-                 suffixes=c(".ma", ".mb"))
-
-all_data_mean = merge(ma_mean, mb_mean, by=c("reg_method", "label", "atlases", "subject"), 
-                 suffixes=c(".ma", ".mb"))
-
-all_data$k_diff = all_data$k.mb - all_data$k.ma
-all_data_mean$k_diff = all_data_mean$k.mb - all_data_mean$k.ma
-
-vote_levels <- list("Majority Vote" = "majvote", 
-                      "NMI Vote" = "nmi", 
-                      "Cross-correlation Vote" = "xcorr")
-levels(all_data$method.mb) <- vote_levels
-levels(all_data_mean$method.mb) <- vote_levels
-
-all_data       = merge(all_data, diagnoses)
-all_data_mean  = merge(all_data_mean, diagnoses)
-remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
-@
-%                                                                              %
-%                                                                              %
-
-%
-% synonyms for significant: 
-% powerful, suggestive, apparent, definite, evident, large, marked, material, 
-% noticeable, plain, pronounced, substantial, substantive, 
-%
-% antonyms: 
-% negligible, minor, no appreciable, scant
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%     RESULTS: Experiment 1: Whole Hippocampus Cross-Validation       %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment 1 Results: Whole Hippocampus Cross-Validation}
 
@@
 
+With an increasing number of templates, \mb-based segmentating using \ants
+registration and majority vote label fusion shows improvement in overlap
+accuracy over multi-atlas-based segmentation, using the same number of atlases
+and voting method (Figure \ref{fig:ADNI1-xva-k-diff}). The magnitude of improvement over
+multi-atlas-based segmentation decreases with an increasing number of atlases,
+with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (~0.02
+DSC) is found when one atlas is used with a template library of 20 images.
+
+\todo{mention failures}
+
+In addition to an improvement in accuracy over multi-atlas-based segmentation,
+\mb also shows a decrease in the variability of segmentation accuracy 
+(Figure \todo{ADNI1-xval-variability}).  The size of template library 
+necessary to reach a significant (p<0.5) decrease in variance and standard 
+deviation grows with the size of atlas library used.  A template library of 19
+images is sufficient to show significant decrease in variance and standard 
+deviation for 3-7 atlases.
+
+Finally, \mb segmentation volumes show differential agreement with manual (SNT) 
+volumes throughout the range of hippocampal volumes (Figure 
+\ref{ADNI1-xval-Bland-Altman}). \mb shows a negligble fixed bias towards 
+producing smaller segmentations, and shows a slight proportional bias towards 
+smaller segmentations with larger hippocampi.
+\todo{Note limits of agreement?}
+
+\todo{Include multi-atlas 9-atlas in plot as a comparison}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: prep
+% Prepares the data in a the form needed for plotting.
+<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
+all_data  <- read.csv('data/cache/ADNI-XVAL:all_data.csv')
+all_data_mean  <- read.csv('data/cache/ADNI-XVAL:all_data_mean.csv')
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
 \begin{figure}
@@
   \caption{
-  {\bf Mean DSC by atlas and template library size, registration algorithm,
-  and label fusion method.} The mean DSC score is calculated over all subjects 
-  from the ADNI1 Cross-Validation dataset for each parameter setting.}
+  {\bf Mean DSC of \mb segmentations across 69 ADNI1 subjects, by atlas and
+  template library size, registration algorithm, and label fusion method.} 
+  }
 \end{figure}
-
-With an increasing number of templates, \mb-based segmentating using \ants
-registration and majority vote label fusion shows improvement in overlap
-accuracy over multi-atlas-based segmentation, using the same number of atlases
-and voting method (Figure \ref{fig:ADNI1-xva-k-diff}). The magnitude of improvement over
-multi-atlas-based segmentation decreases with an increasing number of atlases,
-with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (~0.02
-DSC) is found when one atlas is used with a template library of 20 images.
-
-\todo{mention failures}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: "increase" in mean kappa over multi-atlas
 \begin{figure}
-<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=6,fig.height=4>>=
 ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) +
+  ylab( "Difference in mean DSC" ) +
   opts(legend.direction = "horizontal", legend.position = "bottom")
@@
 \end{figure}
-
-In addition to an improvement in accuracy over multi-atlas-based segmentation,
-\mb also shows a decrease in the variability of segmentation accuracy 
-(Figure \todo{ADNI1-xval-variability}).  The size of template library 
-necessary to reach a significant (p<0.5) decrease in variance and standard 
-deviation grows with the size of atlas library used.  A template library of 19
-images is sufficient to show significant decrease in variance and standard 
-deviation for 3-7 atlases.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: show variability of MAGeT over parameters
 \begin{figure}
@@
 \end{figure}
-
-% <<ADNI1-Xval-sd-by-subject, cache=TRUE>>=
-% SDs = data.stats[c('subject','label','atlases','templates.mb','MA.sd','MB.sd')]
-% SDs = subset(SDs, templates.mb %in% c(5,10,15,20) & atlases == 5)
-% 
-% melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
-% ggplot(melted, aes(x=templates.mb, y=value, 
-%                    group=interaction(templates.mb, variable), colour=as.factor(variable))) + 
-%   scale_x_continuous(breaks=c(5,10,15,20)) + 
-%   geom_violin() 
-% @
-% \begin{figure}
-% <<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-% a_20_tmpls = subset(all_data_mean, 
-%                       reg_method  == "ANTS" &
-%                       method.mb   == "Majority Vote" &
-%                       templates.mb== 20)
-% ggplot(a_20_tmpls, aes(x=factor(atlases))) +
-%   #geom_point(aes(y=k.mb,group=1,colour="MAGeT"),position="jitter") + 
-%   #geom_point(aes(y=k.ma,group=2,colour="MA"),position="jitter") + 
-%   geom_violin(aes(y=k.mb,group=factor(atlases),colour="MAGeT")) + 
-%   geom_violin(aes(y=k.ma,group=factor(atlases),colour="MA")) + 
-%   geom_smooth(aes(y=k.mb,group=1,colour="MAGeT")) +
-%   geom_smooth(aes(y=k.ma,group=2,colour="MA")) +
-%   
-%   #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
-%   #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
-%   #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
-%   #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
-%   labs(x="Number of Atlases", y = "Kappa", 
-%        title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
-%   scale_colour_hue(name="Method") + 
-%   coord_cartesian(ylim=c(0.7,0.9))
-%   #theme(legend.justification=c(1,0), legend.position=c(1,0))
-% @
-%   \caption{Change in mean Kappa between multi-atlas and \mb using \ants,
-%   20 templates}
-%   \label{}
-% \end{figure}
-% 
-% \begin{figure}
-% <<ADN1-Xval-ANTS-20tmpls-t-test-sd, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-% a_20_tmpls = subset(all_data, 
-%                       reg_method  == "ANTS" &
-%                       method.mb    == "Majority Vote" &
-%                       templates.mb == 20 &
-%                       atlases     == 9 & 
-%                       volume.ma < 10000)
-% stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
-%   data.frame(
-%     MA = sd(df$k.ma),  #MA
-%     MB = sd(df$k.mb)  #MB
-%   )
-% })
-% 
-% 
-% # TODO: against SNT
-% melted = melt(stddev, id.vars=c("subject", "label"))
-% ggplot(melted, aes(x=variable, y=value)) + geom_boxplot()
-% t.test(stddev$MA, stddev$MB)
-% @
-%   \caption{}
-%   \label{}
-% \end{figure}
-% 
-%
-
-Finally, \mb segmentation volumes show differential agreement with manual (SNT) 
-volumes throughout the range of hippocampal volumes (Figure 
-\ref{ADNI1-xval-Bland-Altman}). \mb shows a negligble fixed bias towards 
-producing smaller segmentations, and shows a slight proportional bias towards 
-smaller segmentations with larger hippocampi.
-\todo{Note limits of agreement?}
-
-\todo{Include multi-atlas 9-atlas in plot as a comparison}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: Bland-Altman plots
 \begin{figure}
@@
 \end{figure}
-% 
-% 
-% \begin{figure}
-% <<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-% a_20_tmpls = subset(all_data, 
-%                       reg_method  == "ANTS" &
-%                       method.y    == "Majority Vote" &
-%                       templates.y == 20 &
-%                       atlases %in% c(1,9))
-% template_library = read.csv('data/a2a_ants/template_library_composition.csv')
-% template_library$subject = NULL
-% a_20_tmpls = merge(a_20_tmpls, template_library,
-%                    by.x=c('timestamp', 'templates.y'), 
-%                    by.y=c('timestamp', 'templates'))
-% ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
-%   facet_grid( . ~ atlases) + 
-%   geom_smooth(method='lm',fullrange=F) 
-%  # add 
-% @
-%   \caption{Mean kappa vs. proportion of disease category in template library by disease category}
-%   \label{}
-% \end{figure}
-% 
-% 
-% \todo{show cost (in registrations) / benefit tradec off graph:  show number of
-% registrations per Kappa?  or hours of manual labour per Kappa?)}
-% discuss run-time for each of MAGet and multi-atlas
-
-%%%%%
-%%%%%                  Results: ADNI-1 Complete 1Yr Validation               %%%%%
-%%%%%
-\subsection{ADNI1 Complete 1Yr Validation}
-
-<<ADNI-src-volumes-prep,echo=F,cache=T>>=
-#bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
-mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
-maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
-snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
-fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
-fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
-qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")
-
-# Get subjects, and diagnoses
-yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
-yr1 = subset(yr1, !grepl("Scaled_2", Description)) #TODO: remove Scaled_2? 
-yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
-subjects = yr1[,c("Image.Data.ID", "RID","VISCODE")]  # we only need these fields
-dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
-subjects = merge(subjects, dx)
-
-# fetch just the columns we need, and do a little renaming
-mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, 
-                           MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
-maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, 
-                           MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
-snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, 
-                           SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
-fs_pruned     = data.frame(TEMPQC =fs$TEMPQC, 
-                           FS_L=fs$ST29SV, 
-                           FS_R=fs$ST88SV, 
-                           Source = fs$IMAGEUID)
-fsl_pruned    = data.frame(FSL_L=fsl$X17, 
-                           FSL_R=fsl$X53, Source=fsl$Source)
-
-# Now create one data.frame with all the yr1 volume data we have from each datasource
-combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
-combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
-combined = merge(combined, qc          , by=c("Image.Data.ID", "RID"), all.x=TRUE)
-
-# Step 2: Of those, get counts for the numbers that fail QC
-package_totals = data.frame(
-  Total     = c(length(subjects$VISCODE), 'N/A'),
-  SNT        = c(length(which(!is.na(combined$SNT_L)))    , "--"),
-  MAGeT      = c(length(which(!is.na(combined$MAGeT_L)))  , length(which(combined$MAGeT_QC==0))), 
-  MAPER      = c(length(which(!is.na(combined$MAPER_L)))  , "--"), 
-  FSL        = c(length(which(!is.na(combined$FSL_L)))    , length(which(combined$FSL_QC==0))), 
-  FS         = c(length(which(!is.na(combined$FS_L)))     , length(which(combined$TEMPQC == "Fail"))),
-  row.names  = c("Images", "Failures")
-)
-
-all_qc = subset(combined, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )
-
-attach(all_qc)
-totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
-                  MAGeT = MAGeT_L + MAGeT_R, 
-                  MAPER = MAPER_L + MAPER_R, 
-                  SNT   = SNT_L + SNT_R, 
-                  FS    = FS_L + FS_R, 
-                  FSL   = FSL_L + FSL_R)
-detach(all_qc)
-totals.complete = na.omit(totals)
-@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\FloatBarrier
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
+This experiment explores \mb segmentations of hippocampal subfields. To
+achieve this, a leave-one-out validation is conducted in which lower-resolution
+images ($0.9mm^3$ voxels) of each Winterburn atlas subject is segmented using 
+the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
+atlases when downsampled to $0.9mm^3$ voxels are also computed.
+
+In general, across hippocampal subregions the percent error in volume between 
+\mb segmentations and the manual Winterburn atlas segmentations compares 
+favourably to error when resampling the atlas segmentations (Figure 
+\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate subregions 
+all show near or smaller percent errors. The Sibiculum and CA2/CA3 subregions 
+show distinctly larger than resampling error.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Winterburn-XVAL: compare resampling
+\begin{figure}[h!]
+<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
+data = subset(read.csv('data/WAval.csv'))
+casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
+casted = rename(casted, c("variable"="region"))
+
+error = ddply(casted, c("file", "region"), function(df) {
+  with(df, 
+  data.frame(
+    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
+    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
+    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
+
+levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
+                            "Subiculum"="X2","Subiculum"="X102", 
+                            "CA4"="X4","CA4"="X104", 
+                            "CA2/CA3"="X5","CA2/CA3"="X105", 
+                            "Dentate"="X6","Dentate"="X106")
+melted = melt(error,id.vars=c("file", "region"))
+melted = rename(melted, c("variable"="measure", "value"="percenterr"))
+levels(melted$measure) <- list("WA Resampled"="resampled",
+                               "MAGeT + WA Resampled" = "mb_0.9mm",
+                               "MAGeT + WA BRAVO"="mb_bravo")
+
+ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
+  geom_boxplot() +
+  labs(y="Percent error in volume", x="Region") +
+  scale_colour_discrete("") + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Percent error in segmentation volume by hippocampus subregion.} 
+  Percent error is measured against the volumes of of the unmodified Winterburn
+  atlas segmentations.
+  {\bf WA Resampled} are volumes of the manual segmentations of the Winterburn
+  atlases after resampling to $0.9mm^3$.  
+  {\bf MAGeT + WA Resampled} volumes are \mb segmentations of the Winterburn atlas
+  images after resampling to $0.9mm^3$ voxels.
+  {\bf MAGeT + WA BRAVO} volumes are \mb segmentations of T1 BRAVO images
+  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
+  subjects. 
+  } 
+  \label{fig:WAval-vol-boxplot}
+\end{figure}
+
+\FloatBarrier
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%    RESULTS: Experiment 3, Application of \mb to the segmentation    %%%%%
+%%%%%                       of schizophrenia patients                     %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment 3 Results: Application to the segmentation first episode
+\\ \mbox{schizophrenia patients}}
+
+In this experiment \mb is applied to a dataset of images of first episode
+schizophrenia parients, using the Winterburn atlases and a template library of
+21 subject images selected at random.  Expert manual whole hippocampal
+segmentations are used as gold standards.
+
+\mb produces hippocampus segmentation volumes that are highly correlated with 
+manual segmentation volumes (Figure \ref{fig:SZ-volumes}).  Additionally, as we 
+saw in the previous experiments, \mb volumes show a fixed bias towards smaller 
+volumes (although the bias is negligble), and a proportional bias towards 
+producing smaller segmentations for larger hippocampi (Figure \ref{fig:FEP-SZ-Bland-Altman}.
+
+\begin{figure}
+<<FEP-volumes,cache=FALSE,dependson="setup">>=
+SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
+casted = dcast(SZ_volumes, Subject + Label ~ Method)
+lm = lm(MAGeT ~ Manual, casted)
+ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
+  geom_smooth(method="lm", formula=y~x) + 
+  geom_point() + 
+  geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
+  xlab(expression("Manual left / right hippocampus volume " (mm^3))) + 
+  ylab(expression("MAGeT left / right hippocampus volume "  (mm^3))) + 
+  scale_size_continuous(guide="none") 
+@
+  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
+  HC volumes for \mb against manually rated Hippocampal volumes}
+  \label{fig:SZ-volumes}
+\end{figure}
+
+\begin{figure}
+<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes', fig.width=4, fig.height=4>>=
+bland_altman_plot <- function(x, y) {
+  data = data.frame(x=x,y=y,diff=x-y)
+  upper = mean(data$diff) + 2 * sd(data$diff)
+  lower = mean(data$diff) - 2 * sd(data$diff)
+  return(ggplot(data, aes(x= x, y = diff)) + 
+    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
+               linetype=2, color='brown') + 
+    annotate("text", y = upper, x=Inf, 
+              label=paste("std =",round(upper,0)), 
+              hjust=1.1,vjust=1.5) + 
+    annotate("text", y = lower, x=Inf, 
+              label=paste("-std =",round(lower,0)), 
+              hjust=1.1,vjust=1.5) + 
+    geom_point())
+}
+
+bland_altman_plot(casted$Manual, casted$MAGeT) + 
+  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
+  ylab('Manual - MAGeT Brain volume') + 
+  geom_smooth(method="lm", formula = y~x)
+@
+  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
+  when using five Winterburn atlases, and a 21 image template library.}
+  \label{fig:FEP-SZ-Bland-Altman}
+\end{figure}
+
+\FloatBarrier
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%                           RESULTS                                   %%%%%
+%%%%%                                                                     %%%%%
+%%%%%         EXPERIMENT 3: Application of \mb to the segmentation        %%%%%
+%%%%%                       of ADNI subjects                              %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment 4 Results: Application to the segmentation of Alzheimer's
+disease patients}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 1yr Complete Dataset Package Totals & Failures                        %
-%                        
+Comparing total hippocampal volume, we find close agreement between manual
+volumes, MAPER and MAGeT volumes across disease categories (Figure
+\ref{fig:ADNI-scr-volumes-boxplot}, Figure \ref{fig:ADNI-scr-histogram}). FSL
+and FreeSurfer both produce volumes in close agreement with each other, but
+which are consistently larger than manually segmented volumes in each disease
+category.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-COMPLETE
+% Prepares the data in a the form needed for plotting.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+<<ADNI-src-volumes-prep,echo=F,cache=T>>=
+totals.complete  <- read.csv('data/cache/ADNI1:totals.complete.csv')
+totals           <- read.csv('data/cache/ADNI1:totals.csv')
+package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1 1yr Complete Dataset Package Totals & Failures
 %                        
@@
 @
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 <<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
-melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
-snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
-  data.frame( c = cor(df$SNT, df$value) ) 
+snt.tests <- ddply(totals.complete, c("DX"), function(df) {
+  data.frame( corr = cor(df$SNT, df$MAGeT),
+              value = -Inf) 
 })
@@
       colour=variable,geom="boxplot") + 
+      geom_text(aes(y=value,x=DX, vjust=-1.4, size=1,
+                    label=paste("r==", signif(corr, digits=3))), 
+            colour='black',parse=TRUE, data=snt.tests) +
       xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
@@
 
-Comparing total hippocampal volume, we find close
-agreement between manual volumes, MAPER and MAGeT volumes across disease 
-categories (Figure \ref{fig:ADNI-scr-volumes-boxplot}, Figure 
-\ref{fig:ADNI-scr-histogram}). FSL and FreeSurfer both produce volumes in close 
-agreement with each other, but which are consistently larger than manually 
-segmented volumes in each disease category.
-
 \begin{figure}
@@
 
-% show as a plot
-<<ADNI-scr-summary-table, echo=F, results="asis",cache=FALSE>>=
-# a rather dirty way of computing the pairwise correlations and then forming
-# a printable table out of them.  Sorry.
-melted = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
-cor_by_dx    = by(melted[c("SNT", "value")], c(melted['variable'], melted['DX']), 
-                  function(x) cor(x$SNT,x$value, use="pairwise.complete.obs"))
-cor_overall  = by(melted[c("SNT", "value")], melted['variable'], 
-                  function(x) cor(x$SNT,x$value, use="pairwise.complete.obs"))
-cor_by_dx = t(rbind(cor_by_dx))
-cor_overall = rbind(cor_overall)
-
-vol_by_dx = t(rbind(by(totals.complete$MAGeT, totals.complete$DX, mean)))
-colnames(vol_by_dx) <- "Mean Volume"
-vol_overall = matrix(mean(totals.complete$MAGeT),dimnames=list("Overall"))
-
-dx      = cbind(vol_by_dx, cor_by_dx)
-overall = cbind(vol_overall, cor_overall)
-df      = format.df(rbind(dx, overall), dec=2)
-
-#latex(df, title="",
-#      n.cgroup=c(1,ncol(df)-1),cgroup=c("","Volume Correlation"),
-#      file="", size="scriptsize")
-@
-
-%%%%%
-%%%%%             Results: First Episode Patient Validation          %%%%%
-%%%%%
-\subsection{First Episode Schizophrenic Patients}
-In this experiment we move from an Alzheimer's disease dataset to a dataset of 
-first episode schizophrenia parients.  This dataset is segmented using \mb with the
-Winterburn atlases, and a template library of 21 subject images selected at random. 
-Expert manual whole hippocampal segmentations are used as gold standards.
-
-\mb produces hippocampus segmentation volumes that are highly correlated with 
-manual segmentation volumes (Figure \ref{fig:SZ-volumes}).  Additionally, as we 
-saw in the previous experiments, \mb volumes show a fixed bias towards smaller 
-volumes (although the bias is negligble), and a proportional bias towards 
-producing smaller segmentations for larger hippocampi (Figure \ref{fig:FEP-SZ-Bland-Altman}.
-
-\begin{figure}
-<<FEP-volumes,cache=FALSE,dependson="setup">>=
-SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
-casted = dcast(SZ_volumes, Subject + Label ~ Method)
-lm = lm(MAGeT ~ Manual, casted)
-ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
-  geom_smooth(method="lm", formula=y~x) + 
-  geom_point() + 
-  geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
-  xlab(expression("Manual left / right hippocampus volume " (mm^3))) + 
-  ylab(expression("MAGeT left / right hippocampus volume "  (mm^3))) + 
-  scale_size_continuous(guide="none") 
-@
-  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
-  HC volumes for \mb against manually rated Hippocampal volumes}
-  \label{fig:SZ-volumes}
-\end{figure}
-
-\begin{figure}
-<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes', fig.width=4, fig.height=4>>=
-bland_altman_plot <- function(x, y) {
-  data = data.frame(x=x,y=y,diff=x-y)
-  upper = mean(data$diff) + 2 * sd(data$diff)
-  lower = mean(data$diff) - 2 * sd(data$diff)
-  return(ggplot(data, aes(x= x, y = diff)) + 
-    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + 
-    annotate("text", y = upper, x=Inf, 
-              label=paste("std =",round(upper,0)), 
-              hjust=1.1,vjust=1.5) + 
-    annotate("text", y = lower, x=Inf, 
-              label=paste("-std =",round(lower,0)), 
-              hjust=1.1,vjust=1.5) + 
-    geom_point())
-}
-
-bland_altman_plot(casted$Manual, casted$MAGeT) + 
-  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
-  ylab('Manual - MAGeT Brain volume') + 
-  geom_smooth(method="lm", formula = y~x)
-@
-  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
-  when using five Winterburn atlases, and a 21 image template library.}
-  \label{fig:FEP-SZ-Bland-Altman}
-\end{figure}
-
-%%%%%
-%%%%%              Experiments: Winterburn Atlases Validation              %%%%%
-%%%%%
-\subsection{Winterburn Atlases Cross-Validation}
-
-The final experiment explores \mb segmentations of hippocampal subfields. To
-achieve this, a leave-one-out validation is conducted in which lower-resolution
-images ($0.9mm^3$ voxels) of each Winterburn atlas subject is segmented using 
-the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
-atlases when downsampled to $0.9mm^3$ voxels are also computed.
-
-In general, across hippocampal subregions the percent error in volume between 
-\mb segmentations and the manual Winterburn atlas segmentations compares 
-favourably to error when resampling the atlas segmentations (Figure 
-\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate subregions 
-all show near or smaller percent errors. The Sibiculum and CA2/CA3 subregions 
-show distinctly larger than resampling error.
-
-\begin{figure}
-<<WAval-vol-boxplot,cache=T>>=
-# dirty dirty
-data = subset(read.csv('data/WAval.csv'))
-casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
-casted = rename(casted, c("variable"="region"))
-
-error = ddply(casted, c("file", "region"), function(df) {
-  with(df, 
-  data.frame(
-    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
-
-levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
-                            "Subiculum"="X2","Subiculum"="X102", 
-                            "CA4"="X4","CA4"="X104", 
-                            "CA2/CA3"="X5","CA2/CA3"="X105", 
-                            "Dentate"="X6","Dentate"="X106")
-melted = melt(error,id.vars=c("file", "region"))
-melted = rename(melted, c("variable"="measure", "value"="percenterr"))
-levels(melted$measure) <- list("Resampled WA"="resampled",
-                               "MAGeT-Brain(Resampled WA)" = "mb_0.9mm",
-                               "MAGeT-Brain(BRAVO)"="mb_bravo")
-
-ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
-  geom_boxplot() +
-  labs(y="Percent error in volume", x="Region") +
-  scale_colour_discrete("") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
-  \caption{{\bf Percent error in segmentation volume by hippocampus subregion.} 
-  {\em Resampled WA} are volumes of the manual segmentations of the Winterburn atlases 
-  after resampling to $0.9mm^3$.  {\em \mb(Resampled WA)} volumes are
-  \mb segmentations of the Winterburn atlas images after resampling to $0.9mm^3$ voxels.
-  {\em \mb(BRAVO) \mb} volumes are \mb segmentations of T1 BRAVO images ($0.9mm^3$ voxels)      
-  acquired separately 
-  of four of the five Winterburn atlas subjects. Percent error is measured against the volumes of
-  of the unmodified Winterburn atlas segmentations.}
-  \label{fig:WAval-vol-boxplot}
-\end{figure}
-% 
-% <<WAval-error,cach=T>>=
-% # compare % error between MAGeT and resampling
-% mb0.9mm_diff_over_gold_diff = cbind(subjects, abs(gold_diff) - abs(mb_0.9mm_diff), res="diffdff")
-% names(mb0.9mm_diff_over_gold_diff) <- c("subject", "CA1", "Subiculum", "CA4", "CA2/CA3", "Dentate", "res")
-% melted = melt(mb0.9mm_diff_over_gold_diff, id.vars=c("subject", "res"))
-% names(melted) <- c("subject","res", "region","voldiff")
-% ggplot(melted, aes(y=voldiff,x=region)) +
-%   geom_boxplot() +
-%   labs(title="Absolute difference in % difference between MAGeT and resampling",
-%        y="Absolute difference in % difference in volume", 
-%        x="Region")
-% @
-
-%<<WAval-kappa,cache=T>>=
-%#data = read.csv('data/WAval-kappa.csv')
-%ggplot(data, aes(x=label, y=kappa)) + geom_boxplot(y=kappa)
-%@
+\FloatBarrier
 
@@
 \section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Tabulation of segmentation accuracies of existing methods (ADNI)
+%
+<<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
+tab = read.csv("data/ADNI-existing-work.csv")
+tab$Notes <- NULL
+
+adni_lit = subset(tab, Dataset=="ADNI")
+adni_lit$Dataset <- NULL
+other_lit = subset(tab, Dataset!="ADNI")
+other_lit$Dataset <- NULL
+
+latex(adni_lit, file="", 
+      size="scriptsize",
+      caption="Caption", 
+      rowname=NULL,
+      col.just=c("p{1.5in}l","p{0.5in}","p{.5in}","p{1in}"),
+      title="")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
-%flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
-%(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
-%$mm^3$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% ADNI1 Complete 1Yr Template Library Demographics Table
-%
-%
-% <<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
-% scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
-%                  0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
-%                  0159)
-% tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-%                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-%                scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-% 
-% latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
-%       caption="ADNI1 Complete 1Yr Template Library demographics", 
-%       title="tab:ADNI1-scr-tmpllib-demographics")
-% @
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \bibliographystyle{plainnat}

### 2013-04-13 cache_analysis.R

@@ -1,0 +1,139 @@
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+# ADNI1-XVAL: prep
+# Prepares the data in a the form needed for plotting.
+#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+# all data from the validation trials
+a_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
+t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz'))
+diagnoses    = read.csv('data/a2a_diagnoses.csv')
+
+# simplify the data
+t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
+a_data       = subset(a_data, atlases %in% c(1,3,5,7,9))
+t_data$se    = NULL
+t_data$sn    = NULL
+t_data$j     = NULL
+a_data$se    = NULL
+a_data$sn    = NULL
+a_data$j     = NULL       
+
+# add registration method column 
+t_data       = cbind(t_data, reg_method = "ANIMAL")
+a_data       = cbind(a_data, reg_method = "ANTS")
+
+# separate MA majvote from MB so we can compute differences
+# we only compare multiatlas using majority vote
+t_ma = subset(t_data, approach=="ma" & method=="majvote")
+t_ma$approach="multiatlas"
+a_ma = subset(a_data, approach=="multiatlas" & method=="majvote")
+t_mb = subset(t_data, approach=="mb")
+a_mb = subset(a_data, approach=="mb")
+
+ma = rbind(t_ma,a_ma)
+mb = rbind(t_mb,a_mb)
+
+# set up a few equivalences for multi-atlas, to make nomenclature simpler
+ma$templates =  ma$atlases  # because no template library
+mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n 
+
+# aggregate over batches
+ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
+                    templates + top_n + label, data = ma, mean)
+mb_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
+                    templates + top_n + label, data = mb, mean)                    
+
[…]

### 2013-04-13 unused_code.R

@@ -1,0 +1,162 @@
+#Representative 3T image parameters were TR = 2300ms, TI = 900ms, TE = 3.0ms,
+#flip angle = $8{\circ}$, field of view = 256 x 240mm, a 256 x 256 x 166 matrix
+#(x, y, and z directions) yielding a voxel resolution of 1.00 x 1.00 x 1.2
+#$mm^3$.
+
+########################################
+#
+# ADNI1 Complete 1Yr Template Library Demographics Table
+#
+#
+# <<ADNI1-scr-tmpllib-demographics, echo=F, results="asis",cache=FALSE>>=
+# scr_tmpllib_RIDs = c(0295, 0413, 0619, 0685, 0729, 0782, 0938, 0954, 1018, 1155,
+#                  0907, 0981, 0448, 0553, 0814, 1130, 0698, 1339, 1095, 0991,
+#                  0159)
+# tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+#                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID #in#
+#                scr_tmpllib_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+# 
+# latex(tab, landscape=FALSE, ctable=FALSE, file="", size="scriptsize",
+#       caption="ADNI1 Complete 1Yr Template Library demographics", 
+#       title="tab:ADNI1-scr-tmpllib-demographics")
+# @
+########################################
+
+
+# <<ADNI1-Xval-sd-by-subject, cache=TRUE>>=
+# SDs = data.stats[c('subject','label','atlases','templates.mb','MA.sd','MB.sd')]
+# SDs = subset(SDs, templates.mb #in# c(5,10,15,20) & atlases == 5)
+# 
+# melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
+# ggplot(melted, aes(x=templates.mb, y=value, 
+#                    group=interaction(templates.mb, variable), colour=as.factor(variable))) + 
+#   scale_x_continuous(breaks=c(5,10,15,20)) + 
+#   geom_violin() 
+# @
+# \begin{figure}
+# <<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+# a_20_tmpls = subset(all_data_mean, 
+#                       reg_method  == "ANTS" &
+#                       method.mb   == "Majority Vote" &
[…]

### 2013-04-13 paper.Rnw
Fix up layout of ADNI automated methods survey table

@@
 <<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
+caption = paste(
+  "Survey of automated segmentation accuracy of the ADNI dataset",
+  sep=" ")
+
 tab = read.csv("data/ADNI-existing-work.csv")
@@
       size="scriptsize",
-      caption="Caption", 
+      caption=caption, 
       rowname=NULL,
-      col.just=c("p{1.5in}l","p{0.5in}","p{.5in}","p{1in}"),
+      col.just=c("p{1.5in}l","p{0.5in}","p{1.3in}","p{1.5in}"),
       title="")

### 2013-05-01 paper.Rnw
ADNI1 boxplot now has nicer labels; correct subfield labels; and a table of method/volume correlations

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-COMPLETE
+% Prepares the data in a the form needed for plotting.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+<<ADNI-src-volumes-prep,echo=F,cache=T>>=
+totals.complete  <- read.csv('data/cache/ADNI1:totals.complete.csv')
+totals           <- read.csv('data/cache/ADNI1:totals.csv')
+package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                                                                             
@@
 yr1 = yr1[,c("RID","VISCODE")]
-
+yr1 = 
 adnimerge.yr1 = merge(adnimerge, yr1) 
@@
                             "Subiculum"="X2","Subiculum"="X102", 
-                            "CA4"="X4","CA4"="X104", 
+                            "CA4/DG"="X4","CA4/DG"="X104", 
                             "CA2/CA3"="X5","CA2/CA3"="X105", 
-                            "Dentate"="X6","Dentate"="X106")
+                            "SR/SL/SM"="X6","SR/SL/SM"="X106")
 melted = melt(error,id.vars=c("file", "region"))
 melted = rename(melted, c("variable"="measure", "value"="percenterr"))
-levels(melted$measure) <- list("WA Resampled"="resampled",
-                               "MAGeT + WA Resampled" = "mb_0.9mm",
-                               "MAGeT + WA BRAVO"="mb_bravo")
+levels(melted$measure) <- list("Subsampled"="resampled",
+                               "MAGeT + Subsampled" = "mb_0.9mm",
+                               "MAGeT + BRAVO"="mb_bravo")
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-COMPLETE
-% Prepares the data in a the form needed for plotting.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<ADNI-src-volumes-prep,echo=F,cache=T>>=
-totals.complete  <- read.csv('data/cache/ADNI1:totals.complete.csv')
-totals           <- read.csv('data/cache/ADNI1:totals.csv')
-package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
-@
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % ADNI1 1yr Complete Dataset Package Totals & Failures
@@
 <<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
+      colour=variable,geom="boxplot") + 
+      #geom_text(aes(y=value,x=DX, vjust=-1.4, size=1,
+      #              label=paste("r==", signif(corr, digits=3))), 
+      #      colour='black',parse=TRUE, data=snt.tests) +
+      xlab("Diagnosis") + 
+      ylab(expression(paste("Total hippocampal volume (", mm^3, ")")))
+@
+  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
+  MAPER, \mb (MAGeT) and manual (SNT).}
+  \label{fig:ADNI-scr-volumes-boxplot}
+\end{figure}
+
+\begin{figure}[h]
+<<ADNI-scr-volumes-plot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+            variable.name = "method", value.name = "volume")
+
 snt.tests <- ddply(totals.complete, c("DX"), function(df) {
@@
 
-qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
-      colour=variable,geom="boxplot") + 
-      geom_text(aes(y=value,x=DX, vjust=-1.4, size=1,
-                    label=paste("r==", signif(corr, digits=3))), 
-            colour='black',parse=TRUE, data=snt.tests) +
-      xlab("Diagnosis") + ylab("Total Hippocampal Volume (mm3)")
+ggplot(data=melted, aes(x=SNT, y=volume, colour=method)) + 
+      geom_point() + 
+      geom_smooth(method="lm") + 
+      #geom_text(aes(y=value,x=DX, vjust=-1.4, size=1,
+      #              label=paste("r==", signif(corr, digits=3))), 
+      #      colour='black',parse=TRUE, data=snt.tests) +
+      xlab(expression(paste("Manual total hippocampal volume (", mm^3, ")"))) + 
+      ylab(expression(paste("Total hippocampal volume (", mm^3, ")")))
 @
@@
   MAPER, \mb (MAGeT) and manual (SNT).}
-  \label{fig:ADNI-scr-volumes-boxplot}
+  \label{fig:ADNI-scr-volumes-plot}
 \end{figure}
+
+<<ADNI-scr-volumes-table,dependson='ADNI-baseline-volumes-prep',cache=T,results='asis'>>=
+melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+            variable.name = "method", value.name = "volume")
+
+snt.tests <- ddply(totals.complete, c("DX"), function(df) {
+  round(data.frame( MAGeT = cor(df$SNT, df$MAGeT),
+              MAPER = cor(df$SNT, df$MAGeT),
+              FS    = cor(df$SNT, df$FS),
+              FSL   = cor(df$SNT, df$FSL)
+              ), 2)
+})
+
+latex(snt.tests, #file="", 
+      size="scriptsize",
+      caption="", row.names="",
+      title="")
+@
 

### 2013-05-02 paper.Rnw
Integrated Mallar's comments on the intro, and fixed references

@@
 %draft mode
-\usepackage[colorlinks]{hyperref}
 \usepackage[colorinlistoftodos, textwidth=4cm, shadow]{todonotes}
@@
 
-The hippocampus is of particular interest to many researchers because it is
-implicated in forms of brain dysfunction such as Alzheimer's
-disease\citep{Sabuncu2011} and schizophrenia\citep{Narr2004,Karnik-Henry2012},
-and has functional significance in cognitive processes such as learning and
-memory\citep{DenHeijer2012,Scoville2000}.  For many research questions
-involving magnetic resonance imaging (MRI) data accurate identification of the
-hippocampus and its subregions is a necessary first step to better understand
-the individual neuroanatomy of subjects.
-
-Currently, the gold standard for neuroanatomical segmentation is manual
-delineation by an expert human rater.  This is problematic for hippocampal
-segmentation for several reasons.  First, manual segmentation takes a
-significant investment of time and expertise \citep{Hammers2003} which may not
-be readily available to researchers or clinicians.  Second, the amount of data
-produced in neuroimaging experiments increasingly exceeds the capacity for
-identification of specific neuroanatomical structures by an expert manual
-rater.  Third, the true definition of hippocampal anatomy in MR images is
-disputed \citep{Geuze2004}, as evidenced by efforts to create an unified
-segmentation protocol \citep{Jack2011}.
-
-Compounding each of these problems is the significant neuroanatomical
-variability in the hippocampus throughout the course of aging, maturation, and
-neuropsychiatric disorders
-\citep{Mouiha2011b,Sabuncu2011,Giedd1998,Gogtay2006,Narr2002}.  The result is that
-existing hippocampal atlases available to a researcher may not accurately
-represent neuroanatomy of a specific population under study.  Additionally, in
-the course of a research or clinical study, it may be necessary to make
-adjustments to hippocampal definition as a means of hypothesis testing.  For
-example, Poppenk \citep{Poppenk2011} found that subdividing the hippocampus
-into anterior and posterior regions resulted in a predictive relationship
-between volume difference of those regions and recollection memory performance.
-Making such modifications to a set of MRI data segmentations requires additional
-manual effort.
-
-Automated segmentation techniques do not require human intervention but do
-require {\it a priori} anatomical information to guide segmentations.  In this
-paper we focus on methods that use manually segmented MRI atlases as anatomical
-priors, as these methods achieve some of the best automated hippocampal
-segmentation accuracies to-date. This technique was first developed using a
-single atlas prior (known as single-atlas, or model-based,
-segmentation)\citep{Haller1997,Csernansky1998}.  Volumetric image registration
-is used to estimate a fit between the neuroanatomy of an atlas and target
-images.  Labelling of the target image is achieved by applying the resulting
-transformation to the atlas labels to bring them into the target image space
-({\em label propagation}). This method is limited in accuracy by the
-introduction of estimation errors in registration and partial volume effects in
-label resampling, and errors introduced when the anatomy of the atlas is
-unrepresentative of the target anatomy.
-
-Multi-atlas segmentation techniques address these limitations by combining
-segmentation information from a series of expertly segmented atlases
-\citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2012,Aljabar2009,Leung2010,Wolz2010}.
+The hippocampus is a neuroanatomical structure situated in the medial temporal
+lobe of the brain, and has long been associated with learning and memory
+\citep{DenHeijer2012,Scoville2000}.  In addition to its known functional roles,
+the hippocampus is of interest to neuroscientists because it is implicated in
+several forms of brain disfunction such as Alzheimer's disease
+\citep{Sabuncu2011} and schizophrenia \citep{Narr2004,Karnik-Henry2012}.  In
+neuroimaging experiments, magnetic resonance images (MRI) are often used for the
+identification of the hippocampus.  As such, accurate segmentation of the
+hippocampus and its subregions in MRI is a necessary first step to better
+understand the unique neuroanatomy of subjects. Typically, the gold standard for
+neuroanatomical segmentation is manual delineation by an expert human. However
+the rapid increase in the availability of MRI data and the time and expertise
+required for manual segmentation is prohibitive
+\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007,Weiner2013}.
+Further, there is little agreement between researchers regarding how exactly the
+hippocampus should be identified in MRI images \citep{Geuze2004} and this has
+led to efforts to create an unified hippocampal segmentation protocol
+\citep{Jack2011}.
+
+Automated segmentation techniques that are reliable, objective, and reproducible
+are a necessary alternative to manual segmentation.  In the case of classical
+model-based segmentation methods \citep{Haller1997,Csernansky1998}, an MRI atlas
+that was previously manually labelled by an expert rater is matched to target
+images using nonlinear registration methods.  The resulting nonlinear
+transformation is applied to the manual labels (ie. {\em label propagation}) to
+apply them to the target image.  While this methodology has been used
+successfully in several contexts
+\citep{MallarMedIA200809,MallarHBM20082009,Collins1995,Haller1997}, it is limited
+in accuracy by the introduction of errors due inaccuracies in the nonlinear
+transformation itself, partial volume effects in label resampling, and
+irreconcilable differences between the neuroanatomy represented within the atlas
+and target images.
+
+One methodology that can be used to mitigate these sources of errors involves
+the use of multiple manually segmented atlases and probabilistic segmentation
+techniques, such as those found in the FreeSurfer package \citep{Fischl2002},
+FreeSurfer uses a probabilistic atlas of anatomical and tissue classes along
+with spatial constraints for class labels encoded using a Markov random field
+model to segment the entire brain.
+
+More recently, many groups have been using multiple atlases to improve overall
+segmentation accuracy (ie. multi-atlas segmentation) over model-based approaches
+\citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
 Each atlas image is registered to a target image, and label propagation is
 performed to produce several labellings of the target image (one from each
-atlas). A {\it label fusion} technique, such as voxel-wise voting, is used merge
-these labels into a definitive segmentation for the target.  In addition, {\em
-atlas selection} techniques are often used to exclude atlases from label fusion
-that are dissimilar to a target image in order to reduce error from
-unrepresentative anatomy \citep{Aljabar2009}. Cross-correlation or normalised
-mutual information of image intensities are common measures of image similarity
-used in atlas selection.
-
-Multi-atlas methods have been very successfully applied to hippocampal
-segmentation. Collins et al. found near-manual segmentation performance using an
-atlas library of 80 T1-weighted atlas images from the ICBM152 dataset, the
-\animal nonlinear registration algorithm, normalised mutual information as an
-atlas selection similarity metric, and majority vote for label fusion
-\citep{Collins2010}.  The Alzheimer's Disease Neuroimaging Initiative (ADNI) is
-a commonly used benchmarking dataset of MR images of controls and patients with
-MCI or Alzheimer's (see Methods for more information on the ADNI dataset).
-Leung et al. tuned parameters for registration and label fusion to the
-segmentation of ADNI1 1 year dataset of images with an atlas library of 55
-images \citep{Leung2010}.  The MAPER whole brain segmentation algorithm
-\citep{Heckemann2006,Heckemann2011}, using 30 atlases, on all ADNI1 baseline
-images \citep{Heckemann2011}.  Lotjonen et al. use images from the ADNI1
-baseline dataset and 30 atlases with a proprietary non-linear registration
-method based on intensity differences, and post-processing step using a graph
-cuts algorithm to optimise fused segmentations against a spatial intensity prior
-\citep{Lotjonen2012}.  
-\todo{No story here}
-
-The LEAP algorithm is an elegant modification to the basic multi-atlas strategy
-\citep{Wolz2010}. The atlas library is grown, beginning with a set of manually
-labelled atlases, and successively incorporats unlabelled target images after
-being labelling using multi-atlas techniques. The sequence in which target
-images are labelled is chosen so that the similarity between the atlas images
-and the target images is minimised at each step, effectively allowing for
-deformations between very dissimilar images to be broken up into sequences of
-smaller deformations.  With an atlas library of 30 MR images, LEAP was used to
-segment the ADNI1 baseline dataset, achieving a mean Dice score of 0.85 with
-manual segmentations.
-
-%- TODO: "probablistic methods?"
-
-While not purely nulti-atlas techniques, there are several important algorithms
-for hippocampal segmentation that inform our approach. The popular FreeSurfer
-application's whole brain segmentation algoritm uses a probabilistic atlas of
-anatomical and tissue classes along with spatial constraints for class labels
-encoded using a Markov random field model \citep{Fischl2002}. When segmenting
-hippocampal subfields, FreeSurfer employs a Bayesian inference algorithm using a
-probabilistic atlas of anatomical classes as a prior, and a likelihood model of
-how those classes translate into MR image intensities, both trained on manual
-segmentations of high resolution MR images \citep{Fischl2009}. Yushkevitch et
-al. describe a semi-automated method for hippocampal subfield segmentation of
-focal T2 images\citep{Yushkevitch2010}.  The unlabelled MR image must be
-manually partitioned into 'head', 'body' and 'tail', and then multi-atlas
-methods are used to segment the image.  Finally, an AdaBoost-based bias
-correction classifier is trained on texture, spatial location, and intensities
-of manual segmentations and is applied to fix mislabelled voxels.
-
-%  Chupin's use of shape-based priors 
-%   - Mouiha2011b - Hippocampal atrophy rates in AD: compare SNT vs FreeSurfer
-%     on ADNI
-%
-
-Aside from the algorithmic choices used in multi-atlas segmentation, it is
-natural to ask about how the features of the atlases themselves impact the
-resulting segmentations. As noted, by choosing atlases ranked most similar to a
-target image by voxel intensity profile, segmentation accuracy is improved,
-suggesting that neuroanatomical similarity plays a strong role
-\citep{Aljabar2009}.  Carmichael et al. explored this directly and found that
-when using only one atlas the important factors leading to improved accuracy are
-that the atlas have neuroanatomical features that match the target, and that the
-atlas segmentation use the same protocol as the gold-standard
-\citep{Carmichael2005}. Nestor et al.  found that hippocampal segmentation
-protocols that include more dorsal white-matter and posterior anatomy tended to
-produce higher overlap and better accuracy at distinguishing disease classes in
-the ADNI1 1 year dataset \citep{Nestor2012}. These results suggest both atlas
-library neuroanatomy and delineation protocol play a significant role in the
-resulting segmentation.
-
-Considered along with our earlier discussion on the difficulty of producing
-manual segmentations of MR images and the need for adaptable segmentation
-definitions in order to conduct research, this presents a real problem of labour
-and expertise when using existing multi-atlas segmentation methods which rely on
-relatively large atlas libraries (typically between 30 and 80 atlases). Indeed,
-it may be especially prohibitive to use these methods in situations where
-producing a single atlas is challenging (e.g. histology-based atlases, or
-atlases from very high resolution images). In this paper we address the problem
-of producing accurate segmentations using small numbers of manually segmented
-atlases.
-
-% MAGeT bootstrapping: tune the template library to your dataset, reduce the
-% number of input atlases you need, with the upshot that with fewer atlases it's
-% that much easier to change definitions in order to test hypothesis
-
-Our algorith, called \mb ({\em M}ultiple {\em A}utomatically {\em
-Ge}nerated {\em T}emplates), is an extension to the basic multi-atlas-based
-segmentation schema\citep{Chakravarty2012}. Principally, we explore the
-possibility of using a small atlas library to bootstrap a much larger
-\emph{template library} composed of images taken from the target population.
-The template library is then used to segment the targets in a similar fashion
-to basic multi-atlas segmentation: by label propagation and label fusion. The
-intuition driving this approach is that by generating a template library we
-leverage the unique neuroanatomy of target population on hand to initialize the
-segmentation process and improve accuracy over direct propagation from the atlas
-library to unlabelled targets while also using fewer manually segmented
-atlases.
-
-The insight of generating a template library is not new.  Heckemann et al.
-compared ``indirect'' segmentation -- taking a single atlas and propagating the
-labels to intermediate targets before fusing them in a target image space -- to
-multi-atlas segmentation and found that the indirect approach performed worse
-\citep{Heckemann2006}. In this paper we continue the same line of investigation
-but explore the performance when using multiple atlases as well as the effect of
-different registration and fusion methods.
-
-The LEAP algorithm \citep{Wolz2010}, described above, is another example of
-indirect segmentation previously explored.  LEAP proceeds by iteratively
-segmenting unlabelled images most similar to the atlas library images and then
-incorporating the labelled images into the atlas library for future iterations.
-The novelty explored in our current work is to demonstrate the viability of
-achieving comparable segmentation accuracy using the basic multi-atlas schema
-and using significantly fewer manually created atlases.
-
-In previous work \citep{Chakravarty2012}, we applied \mb to segmentation
-of the human striatum, globus pallidus, and thalamus using a single
-histologically-derived atlas. The contribution of the present work is to extend
-our approach to the human hippocampus and perform a series of experiments to
-rigorous validate the method. First, we conduct an extensive cross-validation of
-\mb and basic multi-atlas segmentation on a subset of the ADNI1 dataset to
-assess the accuracy of \mb under various parameter settings (number of
-atlases and templates, registration and fusion methods). With the best
-performing parameter configuration discovered above, we estimate \mb
-intra-rater reliability by segmenting separately acquired T1 images of the
-atlas subjects. For this experiment, we use the Winterburn atlases: digital
-hippocampal subfield segmentations of five {\em in vivo} high-resolution (300$\mu$
-isotropic) T1-weighted MR scans\citep{Winterburn2013}. To validate \mb in a
-real world situation, we segment the entire ADNI1 Complete 1Yr dataset and compare
-our segmentations to established automated and manual segmentations.
-Additionally, to ensure \mb accuracy across disease categories, we also
-compare \mb segmentations to manual segmentations of 139 first episode
-schizophrenic patients.
+atlas). A {\it label fusion} technique, such as voxel-wise voting, is used to
+merge these labels into a definitive segmentation for the target.  In addition,
+weighted voting procedures that use {\em atlas selection} techniques are often
+used to exclude atlases from label fusion that are dissimilar to a target image
+in order to reduce error from unrepresentative anatomy \citep{Aljabar2009}.
+This involves the selection of a subset of atlases using a similarity metric
+such a cross-correlation \citep{Aljabar2009} or normalized mutual information.
+Such selection has the added benefit of significantly reducing the number of
+nonlinear registrations.  For example Collins and Pruessner \citep{Collins2010}
+demonstrated that only 14 atlases, selected based on highest similarity between
+medial temporal lobe neuroanatomy as evaluated by normalized mutual information
+\citep{Studholme1999} from a library of 80 atlases, were required to achieve
+accurate segmentations of the hippocampus.  Several methods have been explored
+for label fusion including the STAPLE algorithm \citep{Warfield2004} that
+computes a probabilistic segmentation using an expectation maximization
+framework from an set of competing segmentations; or others where a subset of
+segmentations can be estimated using metrics such as the sum of squared
+differences in the regions of interest to be segmented \citep{Coupe2011}.
+
+However, many of these methods require significant investment of time and
+resources for the creation of the atlas library; ranging from atlas libraries
+that require between 30 \citep{Heckemann2006} and 80 \citep{Collins2010}
+manually segmented atlases.  This strategy has the main drawback of being
+inflexible as it does not easily accommodate varying the definition of the
+hippocampal anatomy (such as the commonly used heuristic of subdividing the
+hippocampus in to head, body, and tail \citep{Poppenk2011,Pruessner2000}).
+Furthermore, none of these methods have demonstrated sufficient flexibility to
+accommodate atlases that are somehow exceptional such as those derived from
+serial histological data \citep{Chakravarty06a,Yelnik2007} or high-resolution MRI
+data that enables robust identification of hippocampal subfields
+\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}.  Due to
+the recent availability of the latter, there has been increased interest in the
+use of probabilistic methods for the identification of the hippocampal subfields
+on standard T1-weighted images.  Our group recently demonstrated that through
+use of an intermediary automated segmentation stage, robust and accurate
+segmentation of the striatum, pallidum, and thalamus using a single atlas
+derived from serial histological data \citep{Chakravarty2013} is possible.  The
+novelty of this manuscript is the extension our multi-atlas methodology to the
+hippocampus using more than a single input atlas, while simultaneously limiting
+the number of possible inputs used during segmentation, and demonstrating that
+accurate identification of the hippocampal subfields is indeed possible using
+this methodology.
+
+There are few methods that have attempted to perform multi-atlas segmentation
+with a limited number of input atlases. The LEAP algorithm is an elegant
+modification to the basic multi-atlas strategy \citep{Wolz2010} in which the
+atlas library is grown, beginning with a set of manually labelled atlases, and
+successively incorporating unlabelled target images after themselves being
+labelling using multi-atlas techniques. The sequence in which target images are
+labelled is chosen so that the similarity between the atlas images and the
+target images is minimised at each step, effectively allowing for deformations
+between very dissimilar images to be broken up into sequences of smaller
+deformations.  Although Wolz et al. begin with an atlas library of 30 MR images,
+this method could theoretically work using a much smaller atlas library.  In
+their validation, LEAP was used to segment the whole hippocampus in the ADNI1
+baseline dataset, achieving a mean Dice score of 0.85 with manual segmentations.
+
+To the best of our knowledge there are two other segmentation methods that
+attempt to define the hippocampal subfields using standard T1-weighted data.
+The first is included with the FreeSurfer package \citep{VanLeemput2009}.  This
+work is limited as it omits the tail of the hippocampus and the segmentation
+protocol has yet to be fully validated.  Nonetheless, they demonstrate that the
+applicability of their work using data from 10 subjects.  In the second method,
+Yushkevich and colleagues \citep{Yushkevich2009} acquired and labelled
+hippocampal subfields on high-resolution MRI data from post-mortem medial
+temporal lobe samples an demonstrate it's applicability in the segmentation of
+other MRI-volumes using nonlinear registration guided using manually derived
+hippocampus masks and specific landmarks. In their work they demonstrate
+accurate parcellations of the subfields.
+
+Here we address the issue of limiting the number of input atlases by tuning our
+algorithm, for segmentation of the entire hippocampus, using a multi-fold
+experiment performed on a subset Alzheimer's Disease Neuroimaging Initiative
+(ADNI) 1 dataset.  The tuning includes \todo{complete}. Based on the parameters
+we find in our experiment, we validate our algorithm using all of the data
+available in the ADNI Complete 1Yr sample and compare our segmentations to the
+other segmentations that are available through the ADNI informatics portal.  To
+ensure that we have not over-fit our parameters to the aging or
+neurodegenerative brain, we also apply our segmentations to a dataset of normal
+controls and individuals suffering from first episode psychosis. Finally, we
+perform a leave-one-out validation experiment to determine if the subfields can
+be accurately identified using our multi-atlas framework.
 
@@
 \subsection{The \mb Algorithm}
-In this paper we will be making a distinction between an atlas and a template --
-typically these terms are used roughly interchangeably. The term {\it atlas} is
-taken to refer to two image volumes: an intensity image ({\it atlas image}) and
-a corresponding manual segmentation image ({\it atlas labels}). {\it Template}
-refers more generically to any image and corresponding labelling, manual or
-computed, when it is {\it used as} as a model in the segmentation of another
-image.  The terms {\it atlas library} and {\it template library} mean a set of
-such images. Additionally, we will use the terms {\it target} to refer to an
-intensity image for which we would like an segmentation.
+In this paper we will make a distinction between an atlas and a template image.
+The term {\it atlas} refers to two image volumes: an intensity image ({\it atlas
+image}) and a corresponding manual segmentation image ({\it atlas labels}). The
+term {\it template} refers to the {\it use} an any image and corresponding label
+image (whether the labels are manual or computed) as a model in the segmentation
+of another image.  The terms {\it atlas library} and {\it template library} mean
+a set of such images.  Additionally, we will use the terms {\it target} to refer
+to an intensity image for which we would like an segmentation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                       Registration Methods                          %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Registration Methods}
+%%%%%                            Experiments                              %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiments}
+
+The following section describes the experiments we conducted to assess the
+segmentation quality of the \mb algorithm. The first two experiments assess the
+validity of \mb using a cross-validation design. Experiment 1 investigates the
+accuracy of \mb whole hippocampus segmentation over a wide range of parameter
+settings. This enables us to choose the parameter settings offering the best
+performance for use in subsequent experiments. Experiment 2 tests hippocampal
+subfield segmentation quality. The last two experiments assess the validity of
+the \mb algorithm when applied to different diseases: Alzheimer's disease
+(Experiment 3) and first episode schizophrenia patients (Experiment 4).
+
+%The ADNI experiment includes comparison of volumes with other automated
+%segmentation methods. Evaluation is done by correlating volumes b/c overlap
+%metrics assume identical segmentation protocol.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%         Experiment 1: Whole Hippocampus Cross-Validation            %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Experiment 1: Whole Hippocampus Cross-Validation}
+
+Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
+of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
+cross-validation allows us to rigorously validate a large number of parameter
+settings of \mb.
+
+\paragraph{ADNI1 dataset}
+
+69 1.5T images were arbitrarily selected from the baseline scans in the {\em
+ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
+disease category: cognitively normal (CN), mild cognitive impairment (MCI) and
+Alzheimer's disease (AD). Demographics for this subset are shown in Table
+\ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
+right whole hippocampi are available. These labels have been generated using a
+semiautomatic tool (SNT, Medtronic Surgical Navigation Technologies, Louisville,
+CO; see Supplementary Materials).
+
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
+standardized dataset available from ADNI \footnote{
+\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly preprocessed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR
+= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
+x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% ADNI1 X-Validation Dataset Demographics Table
+%
+<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
+adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
+                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
+                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
+                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
+                 1217,29,56,1188,1293,149,382,431,1202)
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
+               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption="ADNI-1 cross-validation subset demographics", 
+      title="tab:ADNI1-xvalidation-demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Experiment details}
+
+Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
+cross-validation, consists of repeated rounds of validation conducted on a fixed
+dataset \cite{Shao1993}. In each round, the dataset is randomly partitioned into
+a training set and a validation set. The method to be validated is then given
+the training data, and its output is compared with the validation set.
+
+In this experiment, our dataset consists of 69 1.5T images and corresponding
+manual segmentations taken from the ADNI1 project (see below for details). In each
+validation round, the training set is used as an atlas library for both \mb and
+basic multi-atlas. The remaining images are segmented by both methods, and the
+resulting segmentations are compared to the manual segmentations for the images.
+
+A total of ten validation rounds are perfomed on each subject in the dataset,
+over each combination of parameter settings. The parameter settings
+we explore are: atlas library size (1-9), template library size (1-20),
+registration method (\ants or \animal), and label fusion method (majority vote,
+cross-correlation weighted majority vote, and normalized mutual information
+weighted majority vote).  A total of $10 \times 69 \times 9 \times 20 \times 2
+\times 3 = \num{7452000}$ validation rounds were conducted, resulting in a total
+of $\num{1490400}$ segmentations analysed.
 
 Before registration, all images underwent preprocessing with the N3 algorithm
-\citep{Sled1998} to minimize intensity nonuniformity. In our experiments we use
+\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we use
 one of two non-linear image registration methods.
 
-\subsubsection{Automatic Normalization and Image Matching and Anatomical
+\subparagraph{Automatic Normalization and Image Matching and Anatomical
 Labeling (\animal)}
@@
 
-\subsubsection{Automatic Normalization Tools (\ants)}
+\subparagraph{Automatic Normalization Tools (\ants)}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                           Label Fusion                              %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Label fusion methods}
+\subparagraph{Label fusion methods}
 Label fusion is a term given to the process of combining the information from
 several candidate labellings for an intensity image into a single labelling.  In
-this paper we explore three fusion methods.
+this experiment we explore three fusion methods.
 
@@
 \end{description}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                            Experiments                              %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiments}
-
-The following section describes the experiments we conducted to assess the
-segmentation quality of the \mb algorithm. The first two experiments assess the
-validity of \mb using a cross-validation design. Experiment 1 investigates the
-accuracy of \mb whole hippocampus segmentation over a wide range of parameter
-settings. This enables us to choose the parameter settings offering the best
-performance for use in subsequent experiments. Experiment 2 tests hippocampal
-subfield segmentation quality. The last two experiments assess the validity of
-the \mb algorithm when applied to different diseases: Alzheimer's disease
-(Experiment 3) and first episode schizophrenia patients (Experiment 4).
-
-%The ADNI experiment includes comparison of volumes with other automated
-%segmentation methods. Evaluation is done by correlating volumes b/c overlap
-%metrics assume identical segmentation protocol.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%         Experiment 1: Whole Hippocampus Cross-Validation            %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 1: Whole Hippocampus Cross-Validation}
-
-Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
-cross-validation allows us to rigorously validate a large number of parameter
-settings of \mb.
-
-\paragraph{ADNI1 dataset}
-
-69 1.5T images were arbitrarily selected from the baseline scans in the {\em
-ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
-disease category: cognitively normal (CN), mild cognitive impairment (MCI) and
-Alzheimer's disease (AD). Demographics for this subset are shown in Table
-\ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
-right whole hippocampi are available. These labels have been generated using a
-semiautomatic tool (SNT, Medtronic Surgical Navigation Technologies, Louisville,
-CO; see Supplementary Materials).
-
-Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
-standardized dataset available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly preprocessed images
-which have been designated to be the "best" after quality control.  All images
-were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
-Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008a}.  Representative 1.5T imaging parameters were TR
-= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
-x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
-yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% ADNI1 X-Validation Dataset Demographics Table
-%
-<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
-adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
-                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
-                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
-                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
-                 1217,29,56,1188,1293,149,382,431,1202)
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption="ADNI-1 cross-validation subset demographics", 
-      title="tab:ADNI1-xvalidation-demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{Experiment details}
-
-Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
-cross-validation, consists of repeated rounds of validation conducted on a fixed
-dataset \cite{Shao1993}. In each round, the dataset is randomly partitioned into
-a training set and a validation set. The method to be validated is then given
-the training data, and its output is compared with the validation set.
-
-In this experiment, our dataset consists of 69 1.5T images and corresponding
-manual segmentations taken from the ADNI1 project (see below for details). In each
-validation round, the training set is used as an atlas library for both \mb and
-basic multi-atlas. The remaining images are segmented by both methods, and the
-resulting segmentations are compared to the manual segmentations for the images.
-
-A total of ten validation rounds are perfomed on each subject in the dataset,
-over each combination of parameter settings. The parameter settings
-we explore are: atlas library size (1-9), template library size (1-20),
-registration method (\ants or \animal), and label fusion method (majority vote,
-cross-correlation weighted majority vote, and normalized mutual information
-weighted majority vote).  A total of $10 \times 69 \times 9 \times 20 \times 2
-\times 3 = \num{7452000}$ validation rounds were conducted, resulting in a total
-of $\num{1490400}$ segmentations analysed.
 
@@
 
-latex(snt.tests, #file="", 
+latex(snt.tests, file="", 
       size="scriptsize",

### 2013-05-03 paper.Rnw
Methods: adni xval experiment now contains all registration and label fusion details; the useless histogram chart is gone

@@
 cross-validation allows us to rigorously validate a large number of parameter
-settings of \mb.
+settings of \mb (atlas and template library sizes, registration algorithm, and 
+label fusion method) and select the best parameters to use in subsequent 
+experiments.
 
@@
 Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
+authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
 August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
@@
 In this experiment, our dataset consists of 69 1.5T images and corresponding
-manual segmentations taken from the ADNI1 project (see below for details). In each
-validation round, the training set is used as an atlas library for both \mb and
-basic multi-atlas. The remaining images are segmented by both methods, and the
-resulting segmentations are compared to the manual segmentations for the images.
+manual segmentations. In each validation round, the dataset is partitioned into 
+a training set consisting of images and their manual labels to be used as an atlas 
+library, and a validation set consisting of the remaining images segmented by both 
+\mb and multi-atlas. The resulting segmentations are compared to the manual 
+segmentations for the images.
 
@@
 \mb produces hippocampus segmentation volumes that are highly correlated with 
-manual segmentation volumes (Figure \ref{fig:SZ-volumes}).  Additionally, as we 
-saw in the previous experiments, \mb volumes show a fixed bias towards smaller 
-volumes (although the bias is negligble), and a proportional bias towards 
-producing smaller segmentations for larger hippocampi (Figure \ref{fig:FEP-SZ-Bland-Altman}.
+manual segmentation volumes (Figure \ref{fig:SZ-volumes}).
 
@@
 
-\begin{figure}
-<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes', fig.width=4, fig.height=4>>=
-bland_altman_plot <- function(x, y) {
-  data = data.frame(x=x,y=y,diff=x-y)
-  upper = mean(data$diff) + 2 * sd(data$diff)
-  lower = mean(data$diff) - 2 * sd(data$diff)
-  return(ggplot(data, aes(x= x, y = diff)) + 
-    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + 
-    annotate("text", y = upper, x=Inf, 
-              label=paste("std =",round(upper,0)), 
-              hjust=1.1,vjust=1.5) + 
-    annotate("text", y = lower, x=Inf, 
-              label=paste("-std =",round(lower,0)), 
-              hjust=1.1,vjust=1.5) + 
-    geom_point())
-}
-
-bland_altman_plot(casted$Manual, casted$MAGeT) + 
-  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
-  ylab('Manual - MAGeT Brain volume') + 
-  geom_smooth(method="lm", formula = y~x)
-@
-  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
-  when using five Winterburn atlases, and a 21 image template library.}
-  \label{fig:FEP-SZ-Bland-Altman}
-\end{figure}
-
 \FloatBarrier
@@
 @
-
-% paired t-test against SNT to show similarity (should be low) 
-% Kruskal-Willis against  (should be high?)
-% post-hoc power analysis to say we can detect this change
-\begin{figure}[h]
-<<ADNI-scr-histogram-prep,cache=T>>=
-melted   = melt(totals.complete, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
-binwidth = 250
-
-# prepare the normals
-grid <- with(melted, seq(min(value), max(value), length = 100))
-normaldf <- ddply(melted, c("variable", "DX"), function(df) {
-  data.frame( 
-    value = grid,
-    density = dnorm(grid, mean(df$value), sd(df$value)),
-    freq = length(df$value) * dnorm(grid, mean(df$value), sd(df$value)) * binwidth,
-    length = length(df$value)
-  )
-})
-
-testsdf <- ddply(melted, c("variable", "DX"), function(df) {
-  shapiro = shapiro.test(df$value)
-  data.frame(
-    shapiro.w = shapiro$statistic,
-    shapiro.p = shapiro$p.value,
-    mean = mean(df$value),
-    sd = sd(df$value)
-  )
-})
-
-melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
-snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
-  ks = ks.test(df$SNT, df$value)
-  data.frame( 
-    ks.D = ks$statistic,
-    ks.p = ks$p.value
-  )
-}) 
-@
-
-<<ADNI1-scr-histogram,dependson='ADNI-scr-histogram-prep',cache=T>>=
-
-ggplot(data=melted, aes(x=value)) + 
-  geom_histogram(binwidth=binwidth, alpha=0.8) + 
-  geom_line(aes(y=freq, colour='red'), data=normaldf) +
-  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=1.5,size=1, 
-                label=paste("list(W ==",signif(shapiro.w, digits=3), ", p == ", 
-                            round(shapiro.p,digits=3), ")")), 
-            parse=TRUE, data=testsdf) +
-  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=3,size=1, 
-                label=paste("list(D ==",signif(ks.D, digits=3), ", p == ", 
-                            round(ks.p,digits=3), ")")), 
-            parse=TRUE, data=snt.tests) + 
-  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=5.5,size=1, 
-                label=paste("sigma ==",signif(sd, digits=3))), 
-            parse=TRUE, data=testsdf) + 
-  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=6,size=1, 
-                label=paste("mu ==",signif(mean, digits=3))), 
-            parse=TRUE,data=testsdf) + 
-  #xlim(0, 10000) + 
-  ylim(0,40) + 
-  facet_grid(variable ~ DX) + 
-  scale_colour_discrete(guide="none") + 
-  scale_size_continuous(guide="none") + 
-  labs(x=expression(paste("Total hippocampal volume (", mm^3, ")")),
-       y="Frequency") + 
-  theme(axis.title.x = element_text(vjust=-0.1))
-@
-  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes
-  as measured by SNT, FSL, FreeSurfer (FS), MAPER, and \mb (MAGeT). A
-  fitted normal curve is shown in red. $W$ is the Shapiro-Wilkes test statistic
-  measuring normality of the data (signficance indicates non-normality). $D$ is
-  the Kolmogorov-Smirnov test statistic measuring the goodness-of-fit between
-  the distribution of measured volumes and SNT volumes (significance indicates a
-  difference).}
-  \label{fig:ADNI-scr-histogram}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=6in,height=3in]{figure/ADNI1_MAGeT_montage/figure.pdf}
-  \caption{A caption}
-  \label{fig:ADNI1-segmentation-montage}
-\end{figure}
 

### 2013-05-17 paper.Rnw
Rewrote bits of the methods (MB description, etc.) for clarity and completeness.
Also added in a bland-altman plot for ADNI1-1yr dataset

@@
 \begin{document}
+%\SweaveOpts{concordance=TRUE}
 
 \title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
-\author{Pipitone J., Winterburn J., Lerch J., Pruessner J., Lepage M., \\ 
+\author{Pipitone J., Winterburn J., Lett, T., Lerch J., Pruessner J., Lepage M., \\ 
 Voineskos A., Chakravarty M.M., and \\ 
@@
 \subsection{The \mb Algorithm}
-In this paper we will make a distinction between an atlas and a template image.
-The term {\it atlas} refers to two image volumes: an intensity image ({\it atlas
-image}) and a corresponding manual segmentation image ({\it atlas labels}). The
-term {\it template} refers to the {\it use} an any image and corresponding label
-image (whether the labels are manual or computed) as a model in the segmentation
-of another image.  The terms {\it atlas library} and {\it template library} mean
-a set of such images.  Additionally, we will use the terms {\it target} to refer
-to an intensity image for which we would like an segmentation.
-
-The simplest form of multi-atlas segmentation combines labellings derived from
-several atlases by way of label fusion \citep{Heckemann2006,Heckemann2011}. We
-will refer to this as {\em basic multi-atlas segmentation}. The schema as for
-this method is as follows: 
-\begin{enumerate}
-\item   An atlas library and set of target images are given as input. The atlas
-        library is used as a template library in the following steps; 
-\item   Each atlas intensity image is nonlinearly registered to each target 
-        intensity image; 
-\item   Label images from each atlas are propagated via the resulting
-        transformations to the target image space; and 
-\item   the resulting labels are fused to produce a single, definitive
-        segmentation. 
-\end{enumerate}       
-The particular registration and voting method used are left unspecified.
-
-\mb is best understood as an extension of the basic multi-atlas
-segmentation schema. Instead of using the atlas library to directly label the
-target images, a subset of the input images are selected as template images and
-then labelled.  The choice of targets used in the template library can be made
-to reflect the neuroanatomy or demographics of the target set as a whole (for
-instance, by sampling equally from cases and controls). Once the template
-library images have been chosen, a truncated version of basic multi-atlas
-segmentation is used to label the template library images without performing
-label fusion. Instead, each template image receives multiple labellings: one
-from each atlas image. A second round of basic multi-atlas segmentation uses the
-template library to segment the entire set of target images (including those
-images used in the template library). Label fusion in this final step fuses all
-labels from all templates. To summarize, figure \ref{alg:MAGeT} describes the
-\mb algorithm in pseudocode.
-
-Source code for \mb can be found at
-\url{http://github.com/pipitone/MAGeTbrain}.
+In this paper, we will use the term {\em label} to mean any segmentation (manual or 
+derived) of an MR image.  {\em Label propagation}, or {\em labelling}, is the 
+process by which two images are registered and the resulting transformation 
+is applied to the labels from one image to bring them into alignment with the 
+other image. We will use the term {\em atlas} to mean a manually segmented image,
+and the term {\em template} to mean an automatically segmented image (i.e. 
+via label propagation). The terms {\em atlas library} and {\em template library} 
+describe any set of such images.Additionally, we will use the term {\em target} 
+to refer to an unlabelled image that is undergoing segmentation.
+
+The simplest form of multi-atlas segmentation, (which we will call {\em basic 
+multi-atlas segmentation}), involves three steps.  First, each labelled 
+input image (i.e. atlas or template) is registered to an unlabelled target image.
+Second, the labels from each image are propogated to the target image space. 
+Third, the labels are combined into a single labelling by way of a label fusion method 
+\citep{Heckemann2006, Heckemann2011}. This method is described in detail in other
+publications \citep{Collins2010,Heckemann2011,Aljabar2009}.
+
+\mb bootstraps the creation of a large template library given a limited input atlas 
+library, and then uses the template library in basic multi-atlas segmentation. 
+Images for the template library are selected from a set of input target images, 
+either arbitrarily or so as to reflect the neuroanatomy or demographics of the 
+target set as a whole (for instance, by sampling equally from cases and controls). 
+The template library images are then labelled by each of the atlases. Basic 
+multi-atlas segmentation is then conducted using the template library
+to segment the entire set of target images (including the targets whose images
+are used in the construction of the template library).  Since each template
+library image has multiple labels (one from each atlas), the final number of 
+labels to be fused for each target may be quite large (i.e. \# of atlas $\times$ 
+\# of templates).
+
+Figure \ref{alg:MAGeT} describes the \mb algorithm in pseudocode. Source code 
+for \mb can be found at \url{http://github.com/pipitone/MAGeTbrain}.
 
@@
 
-\subparagraph{Label fusion methods}
+\paragraph{Label fusion methods}
 Label fusion is a term given to the process of combining the information from
@@
 
-\begin{description}
-  \item[Voxel-wise Majority Vote] \hfill \\
+\subparagraph{Voxel-wise Majority Vote}
   Labels are propagated from all template library images to a target.  Each
@@
 
-  \item[Cross-correlation Weighted Majority Vote] \hfill \\
+\subparagraph{Cross-correlation Weighted Majority Vote}
   An optimal combination of targets from the template library has previously been
@@
 
-  \item[Normalised Mutual Information Weighted Majority Vote] \hfill \\
+\subparagraph{Normalised Mutual Information Weighted Majority Vote}
   This method is similar to cross-correlation weighted voting except that image
@@
   used to calculate the normalised mutual information measure between to images.
-\end{description}
 
@@
 
-Reported subject segmentation accuracy is averaged over the ten validation
+The manual segmentations (SNT) provided as part of the ADNI dataset are used as 
+the gold standard to compare with. We compare both \mb labels and multi-atlas 
+labels.  The segmentation accuracy reported is averaged over the ten validation
 rounds for each parameter setting.
+
+In order to investigate the performance of \mb in a real world setting in which 
+only one set of atlas and template images are used, we explore the variability 
+in label agreement at fixed parameter settings when the choice for atlas and 
+template images is varied. This is achieved by first computing the standard 
+deviation and variance of DSC scores in each block 
+of ten validation rounds per subject. The distribution of these statistics across 
+all subjects is then compared between \mb and multi-atlas using a Student's 
+t-test. A significant difference between distributions is taken to show either
+a larger or smaller level of variability between methods.
 
@@
 used as the atlas library for \mb segmentation. LOOCV is conducted separately
-for two different input datasets: the four Winterburn atlas subject BRAVO MR
-images (referred to as the {\em WA BRAVO} dataset), and the five
+for two different input datasets: the four Winterburn atlas subject T1 BRAVO
+images (referred to as the {\em BRAVO} dataset), and the five
 Winterburn atlas subject images resampled to 0.9mm-isotropic voxel resolution
-(referred to as the {\em WA Resampled} dataset. Each subject in the
+(referred to as the {\em Resampled} dataset. Each subject in the
 dataset is segmented by \mb with that subject's image excluded from the atlas
@@
 \paragraph{Evaluation method}  
-The segmentation volumes of each hippocampal subregion are compared to the
-expert manual segmentations of the unmodified Winterburn atlases.  As a point of
-comparison, we also calculate the subregion volumes from the Winterburn atlas
-segmentations after downsampling to 0.9mm-isotropic voxels.
+\mb performance is measured with respect to straight-forward nearest-neighbour
+resampling of the Winterburn atlas labels to 0.9mm-isotropic voxels (i.e. no 
+label propagation). The volumes of each hippocampal subregion in the downsampled
+Winterburn atlas labels are compared to the volumes obtained from the full 
+resolution Winterburn atlas labels, the \mb labels on the downsampled Winterburn 
+atlas images, and the \mb labels on the T1 BRAVO acquisitions.
 
@@
 similar to, but different from, the protocol used to segment the SZFEP dataset.
-Therefore, rather than use an overlap metric, \mb hippocampal volumes are
-compared to the corresponding manual segmentation volumes.
-
-\todo{discuss Bland-Altman?  Or bias analysis}
+Therefore, rather than evaluate using an overlap metric, such as DSC, which 
+assumes a fixed segmentation protocol, we instead correlate whole hippocampus 
+volume between \mb and manual segmentations.
 
@@
 deviation for 3-7 atlases.
-
-Finally, \mb segmentation volumes show differential agreement with manual (SNT) 
-volumes throughout the range of hippocampal volumes (Figure 
-\ref{ADNI1-xval-Bland-Altman}). \mb shows a negligble fixed bias towards 
-producing smaller segmentations, and shows a slight proportional bias towards 
-smaller segmentations with larger hippocampi.
-\todo{Note limits of agreement?}
-
-\todo{Include multi-atlas 9-atlas in plot as a comparison}
 
@@
   t.sd  = t.test(df$MA.sd , df$MB.sd)
+  diff_mean_var = mean(df$MA.var - df$MB.var)
+  diff_mean_sd  = mean(df$MA.sd -  df$MB.sd)
+
+  # test statistic is positive if MA mean is larger than MB
+  # i.e. that MA has a greater mean of variances/SDs
+  # i.e. that MB has a smaller mean, and so is doing "better"
   data.frame(
@@
     p.value   = c(t.var$p.value, t.sd$p.value),
-    direction = c(t.var$statistic < 0, t.sd$statistic < 0)
+    direction = c(diff_mean_var, diff_mean_sd)
   )
@@
 
-ggplot(stats.tests, 
-  aes(x=templates.mb, y=p.value, colour=as.factor(atlases), 
-      size = abs(statistic))) + 
+ggplot(subset(stats.tests, direction > 0), 
+  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
   facet_grid( . ~ stat) + 
-  geom_line(aes(size=0.1,area=0.1)) + 
-  geom_point(aes(shape = factor(direction))) + 
+  geom_line(size=1) +
+  geom_point(size=3) + 
   geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
   geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
-  scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
-  scale_size_area(name="t-statistic magnitude") + 
   scale_x_continuous(breaks=seq(1,20,by=2)) + 
@@
   scale_size_continuous(guide='none') + 
-  opts(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   multi-atlas, and the distribution of these test statistics is compared with a 
-  t-test. The p-value of this test, is shown on the y-axis (scaled logarithmically).}
+  t-test. The p-value of this test, is shown on the y-axis (scaled logarithmically).
+  Only points where \mb mean variability is lower multi-atlas are shown.}
   \label{ADNI1-xval-variability}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: Bland-Altman plots
-\begin{figure}
-<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=8>>=
-df_subset = subset(all_data, 
-                      reg_method   == "ANTS" &
-                      method.mb    == "Majority Vote" &
-                      templates.mb %in% c(1,11,19) &
-                      atlases      %in% c(1,5,9) &
-                      volume.ma     < 10000)
-
-# get SNT data formatted like 
-snt   = read.csv("data/a2a_snt_volumes.csv")
-df_merged = merge(df_subset, snt, all.x = TRUE)
-
-hc_vol = ddply(df_merged, c("subject", "label", "atlases", "templates.mb"), function (df) {
-  data.frame(
-    snt = sum(df$volume.snt),
-    ma = sum(df$volume.ma),
-    mb = sum(df$volume.mb)
-  )
-})
-hc_vol$diff = hc_vol$snt - hc_vol$mb
-
-group.stats = ddply(hc_vol, c("atlases", "templates.mb"), function (df) {
-  data.frame(
-    mean = mean(df$diff),
-    sd   = sd(df$diff)
-  )
-})
-hc_vol = merge(hc_vol, group.stats)
-hc_vol$Atlases = hc_vol$atlases
-hc_vol$Templates = hc_vol$templates.mb
-
-ggplot(hc_vol, aes(x=ma, y = diff, group=Atlases)) + 
-    geom_hline(aes(yintercept = mean + c(-2, 0, 2) * sd, colour='brown'), 
-                linetype=2) + 
-    geom_point() + 
-    geom_smooth(method="lm") + 
-  facet_grid(Atlases~Templates, labeller = label_both) + 
-  xlab(expression('Mean manual and MAGeT-Brain total hippocampus volume ' (mm^3))) + 
-  ylab(expression('Manual - MAGeT-Brain total hippocampus volume ' (mm^3))) 
-@
-  \caption{Bland-Altman plots comparing manual and \mb hippocampal volumes 
-  when using a varying number of atlases and templates.}
-  \label{ADNI1-xval-Bland-Altman}
 \end{figure}
@@
 
+Additionally \mb volumes show a fixed bias towards larger volumes by an 
+average $-205.2mm^3$, and a proportional bias towards producing smaller 
+segmentations for larger hippocampi (Figure \ref{fig:ADNI1-src-Bland-Altman}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 })
-
+caption = paste(
+  "Pearson correlation between manual (SNT) whole hippocampus volume", 
+  "and whole hippocampus volume by method.", sep=" ")
 latex(snt.tests, file="", 
       size="scriptsize",
-      caption="", row.names="",
+      caption=caption, row.names="",
       title="")
 @
+
+\begin{figure}
+<<ADN1-src-Bland-Atlman, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=4, fig.height=4>>=
+melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+            variable.name = "method", value.name = "volume")
+melted=subset(melted, method=="MAGeT")
+
+bland_altman_plot <- function(x, y) {
+  data = data.frame(x=x,y=y,diff=x-y)
+  upper = mean(data$diff) + 2 * sd(data$diff)
+  lower = mean(data$diff) - 2 * sd(data$diff)
+  return(ggplot(data, aes(x= x, y = diff)) + 
+    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
+               linetype=2, color='brown') + 
+    annotate("text", y = upper, x=Inf, 
+              label=paste("std =",round(upper,0)), 
+              hjust=1.1,vjust=1.5) + 
+    annotate("text", y = lower, x=Inf, 
+              label=paste("-std =",round(lower,0)), 
+              hjust=1.1,vjust=1.5) + 
+    geom_point())
+}
+
+melted$diff = melted$SNT - melted$volume
+limits = mean(melted$diff) + c(-2, 0, 2) * sd(melted$diff)
+
+ggplot(melted, aes(x= SNT, y = diff)) + 
+  geom_hline(yintercept = limits, linetype=2, color='brown') + 
+ annotate("text", y = limits, x=Inf, label=round(limits,1), 
+            hjust=1.1,vjust=-0.8, size=3) + 
+  geom_point() + 
+  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
+  ylab('Manual - MAGeT Brain volume') + 
+  geom_smooth(method="lm", formula = y~x)
+@
+  \caption{Bland-Altman plot comparing manual (SNT) and \mb hippocampal volumes 
+  in the ADNI1 1.5T 1Yr Complete dataset.}
+  \label{fig:ADNI-scr-Bland-Altman}
+\end{figure}
 

### 2013-05-21 paper.Rnw
Correlations on the ADNI1 volumes figure, subregions figure

@@
 \newcommand{\animal}{ANIMAL }
-
+\newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T }
 %notes and TODO formatting
@@
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
-ADNI1:Complete 1Yr 1.5T} standardized dataset. 23 subjects were chosen from each
+\adnidataset} standardized dataset. 23 subjects were chosen from each
 disease category: cognitively normal (CN), mild cognitive impairment (MCI) and
@@
 authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
+August 2012. The image dataset download was the "\adnidataset"
 standardized dataset available from ADNI \footnote{
@@
 %%%%%                                                                     %%%%%
-%%%%%         EXPERIMENT 3: Application of \mb to the segmentation        %%%%%
+%%%%%         EXPERIMENT 4: Application of \mb to the segmentation        %%%%%
 %%%%%                       of ADNI subjects                              %%%%%
@@
 \paragraph{ADNI1 dataset revisited} 
-All images from the {\em ADNI1:Complete 1Yr 1.5T} standardized dataset described
+All images from the {\em \adnidataset} standardized dataset described
 in Experiment 1 are used.  Clinical and demographic data are shown in table
@@
 \mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2). All images from the ADN11 Complete 1Yr 1.5T dataset are
+(see Experiment 2). All images from the \adnidataset dataset are
 segmented. The optimal size of template library, registration method, and label
@@
   geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
-  xlab(expression("Manual left / right hippocampus volume " (mm^3))) + 
-  ylab(expression("MAGeT left / right hippocampus volume "  (mm^3))) + 
+  xlab(expression("Manual unilateral whole hippocampus volume " (mm^3))) + 
+  ylab(expression("MAGeT unilateral whole hippocampus volume "  (mm^3))) + 
   scale_size_continuous(guide="none") 
@@
 chosen subject images (7 from each disease class) and used majority vote label
-fusion. The entire ADNI1 Complete 1Yr 1.5T dataset was segmented by \mb and we 
+fusion. The entire \adnidataset dataset was segmented by \mb and we 
 now compare the resulting volumes with those obtained by manual segmentation (SNT), 
@@
       xlab("Diagnosis") + 
-      ylab(expression(paste("Total hippocampal volume (", mm^3, ")")))
+      ylab(expression(paste("Bilateral whole hippocampal volume (", mm^3, ")")))
 @
@@
 melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
-            variable.name = "method", value.name = "volume")
-
-snt.tests <- ddply(totals.complete, c("DX"), function(df) {
-  data.frame( corr = cor(df$SNT, df$MAGeT),
-              value = -Inf) 
-})
-
-ggplot(data=melted, aes(x=SNT, y=volume, colour=method)) + 
+            variable.name = "Method", value.name = "volume")
+
+df = totals.complete
+levels(melted) = c("FS" = ),
+                   "FSL"= paste("FSL", round(cor(df$SNT, df$FSL),2)),
+                   "MAPER"= paste("MAPER", round(cor(df$SNT, df$MAPER),2)),
+                   "MAGeT"= paste("MAGeT", round(cor(df$SNT, df$FS),2)))
+
+ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
     geom_point() + 
     geom_smooth(method="lm") + 
-      #geom_text(aes(y=value,x=DX, vjust=-1.4, size=1,
-      #              label=paste("r==", signif(corr, digits=3))), 
-      #      colour='black',parse=TRUE, data=snt.tests) +
-      xlab(expression(paste("Manual total hippocampal volume (", mm^3, ")"))) + 
-      ylab(expression(paste("Total hippocampal volume (", mm^3, ")")))
+    xlab(expression("Manual bilateral whole hippocampal volume " (mm^3))) + 
+    ylab(expression("Automated bilateral whole hippocampal volume " (mm^3))) +
+    annotate("text", y = c(3000,2500,2000,1500), x=Inf-100, size=3,
+              label = c(paste("FS r=", round(cor(df$SNT, df$FS),2)),
+                        paste("FSL r=", round(cor(df$SNT, df$FSL),2)),
+                        paste("MAPER r=", round(cor(df$SNT, df$MAPER),2)),
+                        paste("MAGeT r=", round(cor(df$SNT, df$MAGeT),2))),
+              hjust=1.1,vjust=1.5) +
+    theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
 \end{figure}
-
-<<ADNI-scr-volumes-table,dependson='ADNI-baseline-volumes-prep',cache=T,results='asis'>>=
-melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
-            variable.name = "method", value.name = "volume")
-
-snt.tests <- ddply(totals.complete, c("DX"), function(df) {
-  round(data.frame( MAGeT = cor(df$SNT, df$MAGeT),
-              MAPER = cor(df$SNT, df$MAGeT),
-              FS    = cor(df$SNT, df$FS),
-              FSL   = cor(df$SNT, df$FSL)
-              ), 2)
-})
-caption = paste(
-  "Pearson correlation between manual (SNT) whole hippocampus volume", 
-  "and whole hippocampus volume by method.", sep=" ")
-latex(snt.tests, file="", 
-      size="scriptsize",
-      caption=caption, row.names="",
-      title="")
-@
 
@@
   geom_point() + 
-  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
-  ylab('Manual - MAGeT Brain volume') + 
+  xlab('Mean of manual and MAGeT-Brain hippocampal volume') + 
+  ylab('Difference between Manual and MAGeT-Brain hippocampal volume') + 
   geom_smooth(method="lm", formula = y~x)
@@
   \caption{Bland-Altman plot comparing manual (SNT) and \mb hippocampal volumes 
-  in the ADNI1 1.5T 1Yr Complete dataset.}
+  in the \adnidataset dataset.}
   \label{fig:ADNI-scr-Bland-Altman}
@@
 
+\begin{figure}
+<<ADN1-src-subfield, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=4, fig.height=4>>=
+adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
+                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
+                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
+                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
+                 1217,29,56,1188,1293,149,382,431,1202)
+mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
+
+# Get subjects, and diagnoses
+base = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
+base = subset(base, !grepl("Scaled_2", Description)) 
+base$VISCODE <- factor(base$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
+subjects = yr1[,c("Image.Data.ID", "RID","VISCODE")]  # we only need these fields
+dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
+subjects = merge(subjects, dx)
+
+# Now create one data.frame with all the yr1 volume 
+# data we have from each datasource
+subregions = merge(subjects, mb, by.x="Image.Data.ID", by.y="SourceImageID", 
+                   all.x=TRUE)
+subregions = subset(subregions, VISCODE == "bl" & RID %in% adni_69_RIDs)
+attach(subregions)
+subregions$X1 = mean(X1, X101)
+subregions$X2 = mean(X2, X102)
+subregions$X4 = mean(X4, X104)
+subregions$X5 = mean(X5, X105)
+subregions$X6 = mean(X6, X106)
+detach(subregions)
+subregions = subregions[,!names(subregions) %in% c("Left_Hippo","Right_Hippo",
+                                                   "X101","X102","X103","X105",
+                                                   "X106")]
+subregions.melted = melt(subregions, measure.vars=c("X1","X2","X4","X5","X6"),
+                         variable.name="Subregion", value.name="Volume")
+levels(subregions.melted$Subregion) <-list("CA1"="X1", "CA2/CA3"="X5",
+                                           "CA4/DG"="X4", 
+                                           "SR/SL/SM"="X6",
+                                           "Subiculum"="X2")
+ggplot(subregions.melted, aes(x=Subregion, y=Volume)) +
+  geom_boxplot()
+@
+  \caption{}
+  \label{fig:ADNI-scr-subfields}
+\end{figure}
 \FloatBarrier

### 2013-05-21 paper.Rnw
Mallar's suggestions plus fixes

@@
 \todo[inline]{finish}
+
 \end{abstract}
@@
 experiment performed on a subset Alzheimer's Disease Neuroimaging Initiative
-(ADNI) 1 dataset.  The tuning includes \todo{complete}. Based on the parameters
-we find in our experiment, we validate our algorithm using all of the data
-available in the ADNI Complete 1Yr sample and compare our segmentations to the
-other segmentations that are available through the ADNI informatics portal.  To
-ensure that we have not over-fit our parameters to the aging or
-neurodegenerative brain, we also apply our segmentations to a dataset of normal
-controls and individuals suffering from first episode psychosis. Finally, we
-perform a leave-one-out validation experiment to determine if the subfields can
-be accurately identified using our multi-atlas framework.
+(ADNI) 1 dataset. Based on the parameters we find in our experiment, we validate
+our algorithm using all of the data available in the ADNI Complete 1Yr sample
+and compare our segmentations to the other segmentations that are available
+through the ADNI informatics portal.  To ensure that we have not over-fit our
+parameters to the aging or neurodegenerative brain, we also apply our
+segmentations to a dataset of normal controls and individuals suffering from
+first episode psychosis. Finally, we perform a leave-one-out validation
+experiment to determine if the subfields can be accurately identified using our
+multi-atlas framework.
 
@@
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
-\adnidataset} standardized dataset. 23 subjects were chosen from each
-disease category: cognitively normal (CN), mild cognitive impairment (MCI) and
+\adnidataset} standardized dataset. 23 subjects were chosen from each disease
+category: cognitively normal (CN), mild cognitive impairment (MCI) and
 Alzheimer's disease (AD). Demographics for this subset are shown in Table
 \ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
-right whole hippocampi are available. These labels have been generated using a
-semiautomatic tool (SNT, Medtronic Surgical Navigation Technologies, Louisville,
-CO; see Supplementary Materials).
+right whole hippocampi are available. These labels have been generated using the
+SNT tool from Medtronic Surgical Navigation Technologies, Louisville, CO (see
+Supplementary Materials for detailed discussion of the manual segmentation
+process used).
 
@@
 this experiment we explore three fusion methods. 
-
-\subparagraph{Voxel-wise Majority Vote}
+\begin{description}
+  \item[Voxel-wise Majority Vote]
   Labels are propagated from all template library images to a target.  Each
@@
 
-\subparagraph{Cross-correlation Weighted Majority Vote}
+  \item[Cross-correlation Weighted Majority Vote]
   An optimal combination of targets from the template library has previously been
@@
 
-\subparagraph{Normalised Mutual Information Weighted Majority Vote}
+  \item[Normalised Mutual Information Weighted Majority Vote]
   This method is similar to cross-correlation weighted voting except that image
@@
   used to calculate the normalised mutual information measure between to images.
-
+\end{description}
 \paragraph{Evaluation method}
@@
 The manual segmentations (SNT) provided as part of the ADNI dataset are used as 
-the gold standard to compare with. We compare both \mb labels and multi-atlas 
-labels.  The segmentation accuracy reported is averaged over the ten validation
+the gold standard to compare with.  The segmentation accuracy reported is averaged over the ten validation
 rounds for each parameter setting.
@@
 images (referred to as the {\em BRAVO} dataset), and the five
-Winterburn atlas subject images resampled to 0.9mm-isotropic voxel resolution
-(referred to as the {\em Resampled} dataset. Each subject in the
+Winterburn atlas subject images subsampled to 0.9mm-isotropic voxel resolution
+(referred to as the {\em Subsampled} dataset. Each subject in the
 dataset is segmented by \mb with that subject's image excluded from the atlas
@@
 resampling of the Winterburn atlas labels to 0.9mm-isotropic voxels (i.e. no 
-label propagation). The volumes of each hippocampal subregion in the downsampled
+label propagation). The volumes of each hippocampal subregion in the subsampled
 Winterburn atlas labels are compared to the volumes obtained from the full 
-resolution Winterburn atlas labels, the \mb labels on the downsampled Winterburn 
+resolution Winterburn atlas labels, the \mb labels on the subsampled Winterburn 
 atlas images, and the \mb labels on the T1 BRAVO acquisitions.
@@
 
-To validate that \mb algorithm works effectively with diseased brain images \mb
-is applied to a dataset of Schizophrenia patient MR images. The resulting
-segmentations are assessed for quality by comparison with expert manual
-segmentations.
+To validate that \mb algorithm works effectively in the context of other
+neuropsychiatric disorders, we use the Winterburn atlases with \mb to predict
+the hippocampal segmentation of dataset of Schizophrenia patient MR images. The
+resulting segmentations are assessed for quality by comparison with expert
+manual segmentations.
 
@@
 subdivide the hippocampus into three different subregions: head, body, and tail.
+We use these segmentations as to validate our implementation of \mb.
 
@@
 \paragraph{Evaluation method}
-The manual segmentation protocol used to segment the Winterburn atlases is
-similar to, but different from, the protocol used to segment the SZFEP dataset.
-Therefore, rather than evaluate using an overlap metric, such as DSC, which 
-assumes a fixed segmentation protocol, we instead correlate whole hippocampus 
-volume between \mb and manual segmentations.
+The manual segmentation protocol used to segment the Winterburn atlases has some
+neuroanatomical differences in comparison to the protocol used to segment the
+SZFEP dataset.  Therefore, rather than evaluate using an overlap metric, such as
+DSC, which assumes a fixed segmentation protocol, we instead correlate whole
+hippocampus volume between \mb and manual segmentations.
+
+\todo{done elsewhere?}
 
@@
 \mb-based segmentations using both \animal and \ants registration algorithm
-demonstrate good overlap with manually derived gold-standards (Figure
+demonstrate good overlap with SNT derived gold-standards (Figure
 \ref{fig:ADNI1-xval-k-mean}). Qualitatively, both \animal and \ants-based
@@
 \mb also shows a decrease in the variability of segmentation accuracy 
-(Figure \todo{ADNI1-xval-variability}).  The size of template library 
+(Figure \ref{ADNI1-xval-variability}).  The size of template library 
 necessary to reach a significant (p<0.5) decrease in variance and standard 
@@
   Only points where \mb mean variability is lower multi-atlas are shown.}
+  \todo{note log scale}
   \label{ADNI1-xval-variability}
@@
 the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
-atlases when downsampled to $0.9mm^3$ voxels are also computed.
+atlases when subsampled to $0.9mm^3$ voxels are also computed.
 
@@
 favourably to error when resampling the atlas segmentations (Figure
-\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate subregions 
-all show near or smaller percent errors. The Sibiculum and CA2/CA3 subregions 
-show distinctly larger than resampling error.
+\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
+subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
+subregions show distinctly larger \todo{what?} than resampling error. 
+\todo{note direction of change}
 
@@
   atlas segmentations.
-  {\bf WA Resampled} are volumes of the manual segmentations of the Winterburn
+  {\bf Subsampled} are volumes of the manual segmentations of the Winterburn
   atlases after resampling to $0.9mm^3$.  
-  {\bf MAGeT + WA Resampled} volumes are \mb segmentations of the Winterburn atlas
+  {\bf MAGeT + WA Subsampled} volumes are \mb segmentations of the Winterburn atlas
   images after resampling to $0.9mm^3$ voxels.
@@
   ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
-  subjects. 
-  } 
+  subjects.} 
   \label{fig:WAval-vol-boxplot}
@@
 manual segmentation volumes (Figure \ref{fig:SZ-volumes}). 
+\todo{p-value}
 
@@
 and FreeSurfer both produce volumes in close agreement with each other, but
-which are consistently larger than manually segmented volumes in each disease
-category.
+which are consistently larger \todo{ADD AN ESTIMATE OF THE DIFFERENCE HERE} than
+manually segmented volumes in each disease category.
 
@@
 average $-205.2mm^3$, and a proportional bias towards producing smaller 
-segmentations for larger hippocampi (Figure \ref{fig:ADNI1-src-Bland-Altman}.
+segmentations for larger hippocampi \todo{WHEN COMPARED TO SNT SEGMENTATIONS} (Figure \ref{fig:ADNI1-src-Bland-Altman}.
+
+Figure \ref{fig:ADNI-scr-subfields} shows the subfield volumes for the 69 ADNI1
+subjects used in Experiment 1.
 
@@
 df = totals.complete
-levels(melted) = c("FS" = ),
+levels(melted) = c("FS" = paste("FSL", round(cor(df$SNT, df$FS),2)),
                    "FSL"= paste("FSL", round(cor(df$SNT, df$FSL),2)),
                    "MAPER"= paste("MAPER", round(cor(df$SNT, df$MAPER),2)),
-                   "MAGeT"= paste("MAGeT", round(cor(df$SNT, df$FS),2)))
+                   "MAGeT"= paste("MAGeT", round(cor(df$SNT, df$MAGeT),2)))
 

### 2013-05-22 figure/ADNI1_SNT_MB_montage/montage_comparison.sh
Adding MB/SNT montage

@@ -1,0 +1,32 @@
+#!/bin/bash
+
+tmp=$(mktemp -d)
+ana=$1
+lbl=$2
+cmp=$3
+slice=$4
+out=$5
+step=0.9
+
+reorder_labels.sh $lbl $tmp/_ordered_lbl.mnc
+reorder_labels.sh $cmp $tmp/_ordered_cmp.mnc
+autocrop+ -byte -keep_real_range -nearest_neighbour -isostep $step $tmp/_ordered_lbl.mnc $tmp/isostep_lbl.mnc 
+autocrop+ -byte -keep_real_range -nearest_neighbour -isostep $step $tmp/_ordered_cmp.mnc $tmp/isostep_cmp.mnc 
+
+autocrop -isostep $step $ana $tmp/isostep_ana.mnc 
+
+mincreshape $tmp/isostep_lbl.mnc $tmp/cropped_lbl.mnc -start 16,64,52 -count 98,98,98 
+mincreshape $tmp/isostep_cmp.mnc $tmp/cropped_cmp.mnc -start 16,64,52 -count 98,98,98 
+mincreshape $tmp/isostep_ana.mnc $tmp/cropped_ana.mnc  -start 16,64,52 -count 98,98,98 
+
+mincpik -sagittal -auto_range -slice ${slice} $tmp/cropped_ana.mnc $tmp/ana_${slice}_sagittal.png 
+mincpik -sagittal -lookup -spectral -image_range 0 20 -slice ${slice} $tmp/cropped_lbl.mnc $tmp/lbl_${slice}_sagittal.png 
+mincpik -sagittal -lookup -spectral -image_range 0 20 -slice ${slice} $tmp/cropped_cmp.mnc $tmp/cmp_${slice}_sagittal.png 
+
+composite -dissolve 55 -tile -gravity south $tmp/ana_${slice}_sagittal.png $tmp/lbl_${slice}_sagittal.png $tmp/lbl_overlay_${slice}_sagittal.png 
+composite -dissolve 55 -tile -gravity south $tmp/ana_${slice}_sagittal.png $tmp/cmp_${slice}_sagittal.png $tmp/cmp_overlay_${slice}_sagittal.png 
+
+montage -geometry 200x200 -tile 3x1 -background black $tmp/ana_${slice}_sagittal.png \
+                                                      $tmp/lbl_overlay_${slice}_sagittal.png \
+                                                      $tmp/cmp_overlay_${slice}_sagittal.png  $out
+echo $tmp

### 2013-05-22 figure/ADNI1_SNT_MB_montage/montage_comparison.sh
Updated montage to zoom in on the hippocampus

@@ -15,9 +15,11 @@ autocrop+ -byte -keep_real_range -nearest_neighbour -isostep $step $tmp/_ordered
 
 autocrop -isostep $step $ana $tmp/isostep_ana.mnc 
 
-mincreshape $tmp/isostep_lbl.mnc $tmp/cropped_lbl.mnc -start 16,64,52 -count 98,98,98 
-mincreshape $tmp/isostep_cmp.mnc $tmp/cropped_cmp.mnc -start 16,64,52 -count 98,98,98 
-mincreshape $tmp/isostep_ana.mnc $tmp/cropped_ana.mnc  -start 16,64,52 -count 98,98,98 
+window="-start 16,64,52 -count 98,98,98"
+window="-start 40,90,52 -count 50,50,98"
+mincreshape $tmp/isostep_lbl.mnc $tmp/cropped_lbl.mnc $window
+mincreshape $tmp/isostep_cmp.mnc $tmp/cropped_cmp.mnc $window
+mincreshape $tmp/isostep_ana.mnc $tmp/cropped_ana.mnc $window
 
 mincpik -sagittal -auto_range -slice ${slice} $tmp/cropped_ana.mnc $tmp/ana_${slice}_sagittal.png 
 mincpik -sagittal -lookup -spectral -image_range 0 20 -slice ${slice} $tmp/cropped_lbl.mnc $tmp/lbl_${slice}_sagittal.png 
@@ -26,7 +28,7 @@ mincpik -sagittal -lookup -spectral -image_range 0 20 -slice ${slice} $tmp/cropp
 composite -dissolve 55 -tile -gravity south $tmp/ana_${slice}_sagittal.png $tmp/lbl_${slice}_sagittal.png $tmp/lbl_overlay_${slice}_sagittal.png 
 composite -dissolve 55 -tile -gravity south $tmp/ana_${slice}_sagittal.png $tmp/cmp_${slice}_sagittal.png $tmp/cmp_overlay_${slice}_sagittal.png 
 
-montage -geometry 200x200 -tile 3x1 -background black $tmp/ana_${slice}_sagittal.png \
-                                                      $tmp/lbl_overlay_${slice}_sagittal.png \
-                                                      $tmp/cmp_overlay_${slice}_sagittal.png  $out
+montage -tile x1 -geometry +2+2 -background black $tmp/ana_${slice}_sagittal.png \
+                                   $tmp/lbl_overlay_${slice}_sagittal.png \
[…]

### 2013-05-22 figure/ADNI1_SNT_MB_montage/montage_comparison.sh
updated montage with markers pointing out idiosyncratic segmentations

@@ -22,13 +22,13 @@ mincreshape $tmp/isostep_cmp.mnc $tmp/cropped_cmp.mnc $window
 mincreshape $tmp/isostep_ana.mnc $tmp/cropped_ana.mnc $window
 
 mincpik -sagittal -auto_range -slice ${slice} $tmp/cropped_ana.mnc $tmp/ana_${slice}_sagittal.png 
-mincpik -sagittal -lookup -spectral -image_range 0 20 -slice ${slice} $tmp/cropped_lbl.mnc $tmp/lbl_${slice}_sagittal.png 
-mincpik -sagittal -lookup -spectral -image_range 0 20 -slice ${slice} $tmp/cropped_cmp.mnc $tmp/cmp_${slice}_sagittal.png 
+mincpik -sagittal -lookup -spectral -image_range 0 2 -slice ${slice} $tmp/cropped_lbl.mnc $tmp/lbl_${slice}_sagittal.png 
+mincpik -sagittal -lookup -spectral -image_range 0 4 -slice ${slice} $tmp/cropped_cmp.mnc $tmp/cmp_${slice}_sagittal.png 
 
 composite -dissolve 55 -tile -gravity south $tmp/ana_${slice}_sagittal.png $tmp/lbl_${slice}_sagittal.png $tmp/lbl_overlay_${slice}_sagittal.png 
 composite -dissolve 55 -tile -gravity south $tmp/ana_${slice}_sagittal.png $tmp/cmp_${slice}_sagittal.png $tmp/cmp_overlay_${slice}_sagittal.png 
 
-montage -tile x1 -geometry +2+2 -background black $tmp/ana_${slice}_sagittal.png \
+montage -tile x1 -geometry +0+0 -background black $tmp/ana_${slice}_sagittal.png \
                                    $tmp/lbl_overlay_${slice}_sagittal.png \
                                    $tmp/cmp_overlay_${slice}_sagittal.png  $out
-echo $tmp
+rm -rf $tmp

### 2013-05-30 paper.Rnw
Added discussion notes and removed even numbers from the A2A plots

@@
   \citep{Chakravarty2012}.  The number of top ranked template library image labels
-  is a configurable parameter.
+  is a configurable parameter and displayed as the size of the template library
+  in the rest of the paper.
 
@@
 <<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ggplot(all_data_mean, 
-       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
+ggplot(subset(all_data_mean, (templates.mb*atlases) %% 2 == 1),
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
   facet_grid(reg_method~method.mb) + 
   stat_summary(fun.y=mean,geom='line',
-               aes(y=k.mb, weight=1)) + 
+               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+    stat_summary(fun.y=mean,geom='point',
+               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
   #stat_smooth(method='lm', formula=y~log(x)) + 
@@
   ylab( "Mean DSC" ) + 
-  opts(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
 <<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=6,fig.height=4>>=
-ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
-                templates.mb %in% seq(1,20,2)), 
+ggplot(subset(all_data_mean, reg_method=="ANTS" & 
+                method.mb=="Majority Vote" & (templates.mb*atlases) %% 2 == 1), 
        aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
@@
   scale_colour_hue(name="Atlases") +
-  scale_alpha_continuous(legend=FALSE) + 
-  scale_linetype_discrete(legend=FALSE) + 
+  scale_alpha_continuous(guide = "none") + 
+  scale_linetype_discrete(guide = "none") + 
   xlab( "Number of Templates" ) + 
   ylab( "Difference in mean DSC" ) +
-  opts(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   Only points where \mb mean variability is lower multi-atlas are shown.}
-  \todo{note log scale}
   \label{ADNI1-xval-variability}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 3 Results: Application to the segmentation first episode
+\subsection{Experiment 3 Results: Application to the segmentation of first episode
 \\ \mbox{schizophrenia patients}}
@@
 
+\begin{figure}
+  \begin{centering}
+    \includegraphics[width=6in]{figure/FEP-montage.pdf}
+  \end{centering}
+  \caption{Sagittal slices from two subjects showing a comparison of the
+  original Winterburn atlas subfield segmentations (at 0.3mm3-isotropic voxel
+  resolution), the downsampled Winterburn segmentations (at 0.9mm3-isotropic
+  voxel resolution), and the MAGeT brain labels on the downsampled atlas image.}
+  \label{fig:FEP-montage}
+\end{figure}
+
 \FloatBarrier
@@
 <<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+melted = melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
+              variable.name="Method", value.name="Volume")
+correlations = ddply(melted, c("DX","Method"), function (df) {
+  data.frame(
+    pearson = cor(df$SNT, df$Volume)
+  )
+})
+
 qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
       colour=variable,geom="boxplot") + 
-      #geom_text(aes(y=value,x=DX, vjust=-1.4, size=1,
-      #              label=paste("r==", signif(corr, digits=3))), 
-      #      colour='black',parse=TRUE, data=snt.tests) +
+      #geom_text(aes(y=Inf,x=interaction(DX,Method), vjust=2, label=pearson, colour=Method), size=3, data=correlations) +
       xlab("Diagnosis") + 
@@
 ggplot(melted, aes(x= SNT, y = diff)) + 
+  facet_grid( DX ~ . ) + 
   geom_hline(yintercept = limits, linetype=2, color='brown') + 
@@
 \begin{figure}
-<<ADN1-src-subfield, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=4, fig.height=4>>=
-adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
-                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
-                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
-                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
-                 1217,29,56,1188,1293,149,382,431,1202)
-mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
-
-# Get subjects, and diagnoses
-base = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
-base = subset(base, !grepl("Scaled_2", Description)) 
-base$VISCODE <- factor(base$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
-subjects = yr1[,c("Image.Data.ID", "RID","VISCODE")]  # we only need these fields
-dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
-subjects = merge(subjects, dx)
-
-# Now create one data.frame with all the yr1 volume 
-# data we have from each datasource
-subregions = merge(subjects, mb, by.x="Image.Data.ID", by.y="SourceImageID", 
-                   all.x=TRUE)
-subregions = subset(subregions, VISCODE == "bl" & RID %in% adni_69_RIDs)
-attach(subregions)
-subregions$X1 = mean(X1, X101)
-subregions$X2 = mean(X2, X102)
-subregions$X4 = mean(X4, X104)
-subregions$X5 = mean(X5, X105)
-subregions$X6 = mean(X6, X106)
-detach(subregions)
-subregions = subregions[,!names(subregions) %in% c("Left_Hippo","Right_Hippo",
-                                                   "X101","X102","X103","X105",
-                                                   "X106")]
-subregions.melted = melt(subregions, measure.vars=c("X1","X2","X4","X5","X6"),
-                         variable.name="Subregion", value.name="Volume")
-levels(subregions.melted$Subregion) <-list("CA1"="X1", "CA2/CA3"="X5",
-                                           "CA4/DG"="X4", 
-                                           "SR/SL/SM"="X6",
-                                           "Subiculum"="X2")
-ggplot(subregions.melted, aes(x=Subregion, y=Volume)) +
-  geom_boxplot()
-@
-  \caption{}
+  \begin{centering}
+    \includegraphics[width=6in]{figure/ADNI1_SNT_MB_montage/montage.pdf}
+  \end{centering}
+  \caption{Manual and MAGeT segmentation results for 30 ADNI1 subjects (10
+  subjects randomly selected from each disease category in the subject pool used
+  in Experiment 1). Sagittal slices are shown for subject unlabelled T1-weighted
+  anatomical image, SNT manual label (green), and \mb label (blue). Noted are
+  examples of common segmentation idiosyncracies: 
+  {\em (a)} over-estimation of hippocampal head and  
+  {\em (b)} translated manual segmentation by SNT; 
+  {\em (c)} under-estimation of hippocampal body and
+  {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb.}
   \label{fig:ADNI-scr-subfields}
 \end{figure}
+
 \FloatBarrier
@@
 
+\begin{verbatim}
+
+Experiment 1: 
+- address the absence of weighted-voting effects. 
+  - hypothesis: the benefits of weighted voting only outweigh the resampling
+    error effects when choosing from a large library.  In this experiment, the 
+    template library (during weighted voting) consists of only 20 templates, so
+    the most we could hope for is a "squashing" upwards of the curve towards 
+    the 20-template limit (which is the same across all cases in this experiment).
+  - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
+    Perhaps in future experiments with larger template libraries, this could be
+    explored further.
+- Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
+  - couldn't get STAPLE to work with our image formats
+  - Idea: expect smoothness across range of templates (no even # dips, below)
+  - perhaps more sophisticated fusion methods would boost results over majority 
+    vote based techniques.
+
+- why does MAGeT brain show a dip in average Kappa when using an even number
+  templates. Does MA show this same pattern? (yes)
+  - Hypothesis: our voting method is biased when breaking ties to choose the 
+    label with the lowest numeric value.
+  - If we compute the total number of labels fused (atlases * templates) then 
+    both MB/MA perform worse when fusing an even number of labels.
+
+- we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
+  using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
+  this increase justify the extra effort involved in MB? 
+  - decrease in variability
+  - we are comparing "true" averages (see above)... does this make our criteria 
+    for a worthy improvement less strict?
+
+- how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
+- because of the extensive cross-validation in experiment 1, we are very likely
+  showing results that approach the true average of MB and MA on that dataset
+  (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
+  each).
+- even with this caveat, do we think we do well enough?  i.e. other than parameter 
+  tuning, to what extent does this experiment tell us about how MB would do in 
+  practice.
+
+- effects we are seeing is not only averaging effects (Chakravarty et al. 2012).
+
+- take aways: as in other studies (Aljabar, Heckemann) we find that performance
+  scales with the number of inputs (approx. log(n)), and that a large enough
+  template library can boost performance with a small number of atlases. The use 
+  of templates /can/ have a negative impact on performance (Figure 2), that is 
+  balanced by the improvement of growing the template library.  Likely the negative
+  impact on peformance is as a result of resampling/mislabelling error. In other 
+  words, tuning MAGeT brain involves balancing the tension between an improvement 
+  in peformance due to increase neuroanatomical capture and decreased peformance
+  due to resampling error.
+
+Experiment 2: 
+- describe resampling error during downsampling (essentially partial volume effect 
+  due to averaging/majority vote in nearest neighbour selection)
+- Why would some structures have greater downsampling error than others? (i.e. what
+  is it about a structure that would make it especially prone to resampling error)
+  - since error is discretised to whole voxels, smaller regions will show larger
+    percent error
+  - average unilateral volume (approx; mm3): 
+     region   volume   downsampling error
+     CA1      800            2%
+     CA4/DG   600          -30%
+     SR/SL/SM 700           30%
+     Sibculum 350            0%
+     CA2/3    200           15%
+
+- observation: when downsampling error is small (<10 percent), MB error is larger, 
+  and vice versa when downsampling error is large.  Could this a case where the 
+  inevitable MAGeT resampling error outweighs the small downsampling error?
+- observation: MAGeT produces similar volumes for the downsampled and BRAVO images
+  which demonstrates reliability. Additionally, MB's error is in the same *direction*
+  as the downsampling error except for SR/SL/SM
+- for MB to show less error than resampling means that voting across templates is 
+  in aggregate performing better than local nearest neighbour fitting. presumably 
+  we'd see the same improvement with basic multiatlas as well.
+
+- take away: MB produces subregion volumes that are comparable or better than
+  resampling error, except for the CA2/3 where error is near 25 percent (but 
+  even then resampling error is ~15%)
+
+Experiment 3: 
+- what is an "acceptable" r^2 value? 
+- take away: MB has proven to be robust with atlases derived from three different
+  segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
+  populations (older with AD progression; young and healthy; young and SZ).
+
+Experiment 4: 
+- address the smaller difference in mean volume across disease classes. smaller 
+  than every other method.
+
+- address the segmentation bias towards over-estimating smaller hippocampi, and 
+  underestimating larger HC
+
+- visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
+  is lower than the other methods (but we may be biased. :-).
+
+- take away: MB, with the Winterburn atlases, produces HC volumes more inline with
+  SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required.
+
+\end{verbatim}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-05-30 cache_analysis.R

@@ -8,8 +8,8 @@ t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz
 diagnoses    = read.csv('data/a2a_diagnoses.csv')
 
 # simplify the data
-t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
-a_data       = subset(a_data, atlases %in% c(1,3,5,7,9))
+t_data       = subset(t_data, atlases %in% seq(1,9,1))
+a_data       = subset(a_data, atlases %in% seq(1,9,1))
 t_data$se    = NULL
 t_data$sn    = NULL
 t_data$j     = NULL
@@ -31,6 +31,9 @@ a_mb = subset(a_data, approach=="mb")
 
 ma = rbind(t_ma,a_ma)
 mb = rbind(t_mb,a_mb)
+
+ma$num_labels = ma$atlases
+mb$num_labels = mb$atlases * mb$templates 
 
 # set up a few equivalences for multi-atlas, to make nomenclature simpler
 ma$templates =  ma$atlases  # because no template library

### 2013-05-30 paper.Rnw
bland-altman for ADNI-XVAL

@@
 
-bland_altman_plot <- function(x, y) {
-  data = data.frame(x=x,y=y,diff=x-y)
-  return(ggplot(data, aes(x= x, y = diff)) + 
+bland_altman_plot <- function(x, y, c) {
+  data = data.frame(x=x,y=y,diff=x-y, c=c)
+  return(ggplot(data, aes(x= x, y = diff, colour = as.factor(c))) + 
     geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + 
-    geom_point())
+               linetype=2, color='brown') + geom_point())
 }
@@
 
+\begin{figure}
+<<ADN1-bland-altman, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
+a2a.best = subset(all_data, atlases == 9 &
+                    templates.mb == 19 &
+                    reg_method=='ANTS' &
+                    method.mb =='Majority Vote')
+a2a.best = merge(a2a.best, snt.unilateral.vols)
+bland_altman_plot(a2a.best$volume.snt, a2a.best$volume.ma, a2a.best$DX) + geom_smooth(method="lm")
+@
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-ggplot(melted, aes(x= SNT, y = diff)) + 
-  facet_grid( DX ~ . ) + 
+ggplot(melted, aes(x= SNT, y = diff, colour=DX)) + 
   geom_hline(yintercept = limits, linetype=2, color='brown') + 

### 2013-06-04 paper.Rnw
New figures added to results section
Two montages, and an updated ADNI1-scr Bland-Altman figure.

--- 2013-05-30 paper.Rnw
@@
 
-bland_altman_plot <- function(x, y) {
-  data = data.frame(x=x,y=y,diff=x-y)
-  return(ggplot(data, aes(x= x, y = diff)) + 
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  require(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
+bland_altman_plot <- function(x, y, c) {
+  data = data.frame(x=x,y=y,diff=x-y, m = (x+y)/2, c=c)
+  return(ggplot(data, aes(x= m, y = diff, colour = as.factor(c))) + 
     geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + 
-    geom_point())
+               linetype=2, color='brown') + geom_point())
 }
@@
 The manual segmentations (SNT) provided as part of the ADNI dataset are used as 
-the gold standard to compare with.  The segmentation accuracy reported is averaged over the ten validation
-rounds for each parameter setting.
+the gold standard to compare with.  The segmentation accuracy reported is averaged 
+over the ten validation rounds for each parameter setting.
 
@@
 atlas library and template library. Improvement in accuracy diminishes noticeably
-with template libraries larger than ten images.
+with template libraries larger than roughly ten images.
 
 No marked difference in segmentation accuracy is seen when either \animal or \ants
-registration is used with different label fusion techniques, at any atlas or
-template library sizes.  In every parameter configuration, the use of \mb with
-\ants registration shows a pronounced increase in segmentation accuracy over \mb with \animal
-registration. In the remainder of this section, only results using the \ants
-registration algorithm will be shown.
-
-It is interesting to note that with an even number of templates, \mb shows a
-small decrease in performance relative to when one fewer template image is used.
-See section \todo{ref} for a discussion of this behaviour. In the remainder of
-this section, only results from odd-sized template libraries will be shown.
-
-With an increasing number of templates, \mb-based segmentating using \ants
-registration and majority vote label fusion shows improvement in overlap
-accuracy over multi-atlas-based segmentation, using the same number of atlases
+registration is used with any number of atlases or templates.  In every parameter 
+configuration, the use of \mb with \ants registration shows a pronounced increase
+in segmentation accuracy over \mb with \animal registration.  Surprisingly, the 
+label fusion method used does not significantly improve label accuracy, contrary 
+to the findings of Aljabar et al. \cite{Aljabar2009} when using weighted voting
+on much larger altas/template libraries. In the remainder of  this section, only 
+results using the \ants registration algorithm and majority vote fusion will 
+be shown.
+
+With an increasing number of templates, \mb shows improvement in overlap
+accuracy over multi-atlas-based segmentation when using the same number of atlases
 and voting method (Figure \ref{fig:ADNI1-xva-k-diff}). The magnitude of improvement over
@@
 deviation for 3-7 atlases. 
+\todo{Tris: only show variance OR standard deviation because otherwise the story gets complicated... why compare measures?}
+
+It is interesting to note that with an even number of templates, \mb shows a
+small decrease in performance relative to when one fewer template image is used.
+See section \todo{ref} for a discussion of this behaviour. In the remainder of
+this section, only results from odd-sized template libraries will be shown.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+<<ADN1-library-composition, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+# These composition datasheets describe the subjects and subject DX in the 
+# template and atlas libraries for each voting run
+atlas_comp = read.csv('data/a2a_ants/atlas_library_composition.csv')
+templ_comp = read.csv('data/a2a_ants/template_library_composition.csv')
+a2a.best = subset(all_data, atlases == 9 &
+                    templates.mb == 19 &
+                    reg_method=='ANTS' &
+                    method.mb =='Majority Vote')
+
+cumulative_dist = 
+  ggplot(a2a.best, aes(x=k.mb)) + stat_ecdf() + scale_y_continuous(breaks=seq(0,1,0.1))
+
+a2a.best = subset(a2a.best, k.mb > 0.75) # drop the guys that suck
+
+a2a.best = merge(a2a.best, atlas_comp, by=c("timestamp","atlases"))
+a2a.best = merge(a2a.best, templ_comp, by.x=c("timestamp","templates.mb"), by.y=c("timestamp","templates"))
+a2a.best$templates_AD_percent = (a2a.best$templates_AD / 19) * 100
+a2a.best$templates_MCI_percent = (a2a.best$templates_MCI / 19) * 100
+a2a.best.melted = melt(a2a.best, measure.vars=c("templates_AD_percent", "templates_MCI_percent"))
+
+k_by_composition = 
+  ggplot(a2a.best.melted, aes(x=value, y=k.mb, colour=DX)) + facet_grid(variable ~ DX) + 
+  geom_boxplot(aes(group=as.factor(value), colour=DX)) +  geom_smooth(method="lm")
+
+snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
+a2a.best = merge(a2a.best, snt.unilateral.vols)
+
+# bland-altman plots
+bland_mb_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.mb, DX) + 
+  geom_smooth(method="lm") + 
+  scale_y_continuous(breaks=seq(-500, 500, 50)))
+
+bland_ma_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.ma, DX) + 
+  geom_smooth(method="lm") + 
+  scale_y_continuous(breaks=seq(-500, 500, 50)))
+
+# just the trials with a high percentage of MCI in the template library
+a2a.best.highMCI = subset(a2a.best, templates_MCI_percent > 50)
+
+bland_mb_highMCI = with(a2a.best.highMCI, bland_altman_plot(volume.snt, volume.ma, DX) + 
+  geom_smooth(method="lm") + 
+  scale_y_continuous(breaks=seq(-500, 500, 50)))
+
+@
+\end{figure}
 
@@
 
-ggplot(subset(stats.tests, direction > 0), 
+ggplot(subset(stats.tests, direction > 0, stat == "Variance"), 
   aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
-  facet_grid( . ~ stat) + 
+  #facet_grid( . ~ stat) + 
   geom_line(size=1) +
@@
   xlab( "Number of Templates" ) + 
-  ylab( "p-value" ) +
+  ylab( "Difference of variance during validation (p-value)" ) +
   scale_colour_hue(name="Atlases") +
@@
   accuracy.}
-  Variability of segmentation accuracy within 10 rounds of validation per subject. 
-  Standard deviation and variance is computed per subject for both \mb and 
-  multi-atlas, and the distribution of these test statistics is compared with a 
-  t-test. The p-value of this test, is shown on the y-axis (scaled logarithmically).
-  Only points where \mb mean variability is lower multi-atlas are shown.}
+  Variance of segmentation accuracy between \mb and multi-atlas segmentation 
+  is computed for each subject across all ten rounds of validation. Shown on the 
+  y-axis (scaled logarithmically) is the  p-value resulting from a t-test 
+  comparing the distribution of variances at each parameter setting 
+  (atlas/template library size). Only points where \mb mean variability is lower 
+  than multi-atlas are shown.}
   \label{ADNI1-xval-variability}
@@
 subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
-subregions show distinctly larger \todo{what?} than resampling error. 
-\todo{note direction of change}
+subregions show distinctly larger volumes than resampling error. 
+\todo{note direction of change}.
+
+Figure \ref{fig:subfield-montage} shows a qualitative comparision of \mb subfield
+segmentation.
 
@@
 
+\begin{figure}
+  \begin{centering}
+    \includegraphics[width=6in]{figure/subfield-montage.pdf}
+  \end{centering}
+  \caption{Sagittal slices from two subjects showing a comparison of the
+  original Winterburn atlas subfield segmentations (at 0.3mm3-isotropic voxel
+  resolution), the subsampled Winterburn segmentations (at 0.9mm3-isotropic
+  voxel resolution), and the MAGeT brain labels on the subsampled atlas image.}
+  \label{fig:subfield-montage}
+\end{figure}
+
 \FloatBarrier
@@
 
-\begin{figure}
-  \begin{centering}
-    \includegraphics[width=6in]{figure/FEP-montage.pdf}
-  \end{centering}
-  \caption{Sagittal slices from two subjects showing a comparison of the
-  original Winterburn atlas subfield segmentations (at 0.3mm3-isotropic voxel
-  resolution), the downsampled Winterburn segmentations (at 0.9mm3-isotropic
-  voxel resolution), and the MAGeT brain labels on the downsampled atlas image.}
-  \label{fig:FEP-montage}
-\end{figure}
-
 \FloatBarrier
@@
 
-Comparing total hippocampal volume, we find close agreement between manual
-volumes, MAPER and MAGeT volumes across disease categories (Figure
-\ref{fig:ADNI-scr-volumes-boxplot}, Figure \ref{fig:ADNI-scr-histogram}). FSL
-and FreeSurfer both produce volumes in close agreement with each other, but
-which are consistently larger \todo{ADD AN ESTIMATE OF THE DIFFERENCE HERE} than
-manually segmented volumes in each disease category.
-
-Additionally \mb volumes show a fixed bias towards larger volumes by an 
-average $-205.2mm^3$, and a proportional bias towards producing smaller 
-segmentations for larger hippocampi \todo{WHEN COMPARED TO SNT SEGMENTATIONS} (Figure \ref{fig:ADNI1-src-Bland-Altman}.
-
-Figure \ref{fig:ADNI-scr-subfields} shows the subfield volumes for the 69 ADNI1
-subjects used in Experiment 1.
+We find a close relationship in total bilateral hippocampal volume between
+all methods and manually segmented volumes (Figure \ref{fig:ADNI-scr-volumes-plot}). 
+Volumes are correlated with Pearson-r > 0.78 for all methods across disease categories.
+Within disease categories (Figure \ref{fig:ADNI-scr-volumes-boxplot}),
+\mb is consistently well correlated to manual volumes (Pearson-r > 0.85), but appears to 
+slightly over-estimate the volume of the AD hippocampus.
+
+To investigate the level of agreement with manually segmented hippocampal volumes, 
+we constructed Bland-Altman plots for each method (Figure \ref{fig:ADNI1-src-Bland-Altman}).  
+As Bland \& Altman, 1985 \cite{Bland1986} noted, high 
+correlation amongst measures of the same quantity does not necessarily 
+imply agreement (as correlation can be driven by a large range in true values, 
+for instance).  What is most striking in Figure \ref{fig:ADNI1-src-Bland-Altman}
+is that all methods show an obvious proportional bias: FreeSurfer and FSL 
+markedly under-estimate smaller hippocampi and over-estimate large hippocampi, 
+whereas MAPER and \mb more conservatively show the reverse bias.  Additionally, 
+all methods show a fixed bias, with FreeSurfer and FSL most dramatically 
+over-estimating hippocampal volume by $2600mm^3$ and $2800mm^3$ on average, respectively, 
+and MAPER and \mb within $250mm^3$ on average.
+
+Figure \ref{fig:ADNI-scr-subfields} shows a qualitative comparison of \mb and 
+manual (SNT) hippocampal segmentations for 10 randomly selected subjects in 
+each disease category, and illustrates some of the common errors found during 
+visual inspection. Mostly frequently, we find \mb improperly includes the vestigial
+hippocampal sulcus and, although not anatomically incorrect, \mb under-estimates 
+the hippocampal body in comparison to the manual (SNT) segmentation.
 
@@
 \begin{figure}[h]
-<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
 melted = melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
@@
       colour=variable,geom="boxplot") + 
-      #geom_text(aes(y=Inf,x=interaction(DX,Method), vjust=2, label=pearson, colour=Method), size=3, data=correlations) +
+      geom_text(aes(y=-Inf,x=c(.7,.85,1,1.15, 1.7,1.85,2,2.15, 2.7,2.85,3,3.15), vjust=-5, label=round(pearson,2)), 
+                colour = 'black', size=3, data=correlations) +
       xlab("Diagnosis") + 
-      ylab(expression(paste("Bilateral whole hippocampal volume (", mm^3, ")")))
+      ylab(expression(paste("Bilateral whole hippocampal volume (", mm^3, ")"))) + 
+      theme(legend.direction = "horizontal", legend.position = "bottom")
 @
   \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
-  MAPER, \mb (MAGeT) and manual (SNT).}
+  MAPER, \mb (MAGeT) and manual (SNT) by disease category.  Pearson correlation 
+  with manual volumes are shown below each box-and-whisker.}
   \label{fig:ADNI-scr-volumes-boxplot}
@@
 df = totals.complete
-levels(melted) = c("FS" = paste("FSL", round(cor(df$SNT, df$FS),2)),
-                   "FSL"= paste("FSL", round(cor(df$SNT, df$FSL),2)),
-                   "MAPER"= paste("MAPER", round(cor(df$SNT, df$MAPER),2)),
-                   "MAGeT"= paste("MAGeT", round(cor(df$SNT, df$MAGeT),2)))
 
@@
   \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
-  MAPER, \mb (MAGeT) and manual (SNT).}
+  MAPER, \mb (MAGeT) and manual (SNT). Pearson correlation with manual volumes
+  are shown for each method.}
   \label{fig:ADNI-scr-volumes-plot}
@@
 \begin{figure}
-<<ADN1-src-Bland-Atlman, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=4, fig.height=4>>=
+<<ADN1-src-Bland-Atlman, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=6, fig.height=6>>=
 melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
             variable.name = "method", value.name = "volume")
-melted=subset(melted, method=="MAGeT")
-
-bland_altman_plot <- function(x, y) {
-  data = data.frame(x=x,y=y,diff=x-y)
-  upper = mean(data$diff) + 2 * sd(data$diff)
-  lower = mean(data$diff) - 2 * sd(data$diff)
-  return(ggplot(data, aes(x= x, y = diff)) + 
-    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + 
-    annotate("text", y = upper, x=Inf, 
-              label=paste("std =",round(upper,0)), 
-              hjust=1.1,vjust=1.5) + 
-    annotate("text", y = lower, x=Inf, 
-              label=paste("-std =",round(lower,0)), 
-              hjust=1.1,vjust=1.5) + 
-    geom_point())
-}
-
 melted$diff = melted$SNT - melted$volume
-limits = mean(melted$diff) + c(-2, 0, 2) * sd(melted$diff)
-
-ggplot(melted, aes(x= SNT, y = diff)) + 
-  facet_grid( DX ~ . ) + 
-  geom_hline(yintercept = limits, linetype=2, color='brown') + 
-  annotate("text", y = limits, x=Inf, label=round(limits,1), 
-            hjust=1.1,vjust=-0.8, size=3) + 
+melted$mean = ( melted$SNT + melted$volume ) / 2
+limits = ddply(melted, c("method"), function (df) { 
+  data.frame(
+    y = mean(df$diff) + c(-1.96,0,2) * sd(df$diff) 
+)})
+
+ggplot(melted, aes(x= mean, y = diff, colour=DX)) + 
+  facet_wrap( ~ method, nrow=2, scales="free") + 
   geom_point() +  
-  xlab('Mean of manual and MAGeT-Brain hippocampal volume') + 
-  ylab('Difference between Manual and MAGeT-Brain hippocampal volume') + 
-  geom_smooth(method="lm", formula = y~x)
-@
-  \caption{Bland-Altman plot comparing manual (SNT) and \mb hippocampal volumes 
-  in the \adnidataset dataset.}
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown", data=limits) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits, hjust=1.1,vjust=-0.8, size=3) + 
+  xlab('Mean of manual and automated total hippocampal volume') + 
+  ylab('Difference between manual and automated total hippocampal volume') + 
+  scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
+  scale_x_continuous(breaks=seq(0,10000, 1000)) + 
+  geom_smooth(method="lm", formula = y~x) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{Bland-Altman plot comparing manual (SNT) and automated total hippocampal volumes 
+  in the \adnidataset dataset.  The overall mean difference, and limits of 
+  agreement ($\pm 1.96SD$) are shown by dashed horizontal lines.}
   \label{fig:ADNI-scr-Bland-Altman}

### 2013-06-04 cache_analysis.R

@@ -41,9 +41,9 @@ mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n
 
 # aggregate over batches
 ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
-                    templates + top_n + label, data = ma, mean)
+                    templates + top_n + label + volume, data = ma, mean)
 mb_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
-                    templates + top_n + label, data = mb, mean)                    
+                    templates + top_n + label + volume, data = mb, mean)                    
 
 all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label",
                               "reg_method", "subject"), 

### 2013-07-16 paper.Rnw
Rest of Mallar\'s comments for methods/results

@@
 \title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
-\author{Pipitone J., Winterburn J., Lett, T., Lerch J., Pruessner J., Lepage M., \\ 
+\author{Pipitone J., Park M., Winterburn J., Lett, T., Lerch J., Pruessner J., Lepage M., \\ 
 Voineskos A., Chakravarty M.M., and \\ 
 the Alzheimer's Disease Neuroimaging Initiative}
+\todo{Full names}
+\todo{Associations}
 \maketitle
@@
 understand the unique neuroanatomy of subjects. Typically, the gold standard for
-neuroanatomical segmentation is manual delineation by an expert human. However
+neuroanatomical segmentation is manual delineation by an expert human rater. However
 the rapid increase in the availability of MRI data and the time and expertise
@@
 the use of multiple manually segmented atlases and probabilistic segmentation
-techniques, such as those found in the FreeSurfer package \citep{Fischl2002},
+techniques, such as those found in the FreeSurfer package \citep{Fischl2002}.
 FreeSurfer uses a probabilistic atlas of anatomical and tissue classes along
@@
 hippocampal subfields on high-resolution MRI data from post-mortem medial
-temporal lobe samples an demonstrate it's applicability in the segmentation of
+temporal lobe samples and demonstrate accurate parcellations of hippocampal subfields in
 other MRI-volumes using nonlinear registration guided using manually derived
-hippocampus masks and specific landmarks. In their work they demonstrate
-accurate parcellations of the subfields.
+hippocampus masks and specific landmarks. \todo{flow is terrible in this sentence...}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 2: Winterburn Atlases Cross-Validation}
+\subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
 In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
 segmentation is tested using a leave-one-out cross-validation (LOOCV) design. 
-The optimal parameter settings found in Experiment 1 are used.
+The optimal parameter settings for \mb found in Experiment 1 are used in this 
+experiment.
 
@@
 
-In this experiment, the five $300\mu$-isotropic voxel Winterburn atlases are
-used as the atlas library for \mb segmentation. LOOCV is conducted separately
-for two different input datasets: the four Winterburn atlas subject T1 BRAVO
-images (referred to as the {\em BRAVO} dataset), and the five
-Winterburn atlas subject images subsampled to 0.9mm-isotropic voxel resolution
-(referred to as the {\em Subsampled} dataset. Each subject in the
-dataset is segmented by \mb with that subject's image excluded from the atlas
-library. The template library consists of 3T T1 images (0.9mm-isotropic
-voxels) of healthy subjects in addition to the images from the dataset being
-evaluated.
-
-The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used.
+In this experiment the Winterburn atlases are used as the cross-validation dataset. 
+The five $300\mu$-isotropic voxel images and labels are used as atlases, and 
+LOOCV is conducted using the five Winterburn atlas subject images 
+subsampled (using trilinear interpolation) to 0.9mm-isotropic voxel 
+resolution (referred to as the {\em Subsampled} dataset) as input subjects. Subsampling
+the subject images allows us to assess \mb in a typical segmentation scenario (high-resolution atlases
+and lower-resolution subjects).  The template library is composed of the subject
+images, plus an additional set of 3T T1 images (0.9mm-isotropic voxels) of healthy
+subjects acquired separately (Table \ref{table:WAval-healthy-demographics}). The 
+optimal size of template library, registration method, and label fusion method 
+found in Experiment 1 are used. Each resampled Winterburn atlas subject image is 
+segmented by \mb with that subject's image excluded from the atlas library.
+
+We reproduce this experiment in a separate LOOCV run in which the resampled 
+Winterburn atlas subject images are substituted for separately acquired T1 
+BRAVO images of four of the subjects (referred to as the {\em BRAVO} dataset).
 
 \paragraph{Evaluation method}  
-\mb performance is measured with respect to straight-forward nearest-neighbour
-resampling of the Winterburn atlas labels to 0.9mm-isotropic voxels (i.e. no 
-label propagation). The volumes of each hippocampal subregion in the subsampled
-Winterburn atlas labels are compared to the volumes obtained from the full 
-resolution Winterburn atlas labels, the \mb labels on the subsampled Winterburn 
-atlas images, and the \mb labels on the T1 BRAVO acquisitions.
+To assess the \mb LOOCV segmentations we compute the relative percent error in 
+hippocampal volume with the full resolution Winterburn atlas 
+segmentations. In addition, by computing the relative error in volume of
+the Winterburn atlas labels resampled (with nearest-neighbour interpolation) to 0.9mm-isotropic voxels, we obtain a baseline error to assess against.
 
@@
 
-The hippocampus were traced following a validated protocol developed by Dr Jens
-Pruessner \todo{cite: (Pruessner et al., 2000)}. A recent update to this
-protocol by Dr J Pruessner in 2006 allows to accurately and consistently
-subdivide the hippocampus into three different subregions: head, body, and tail.
-We use these segmentations as to validate our implementation of \mb.
+Each subject hippocampus is traced following a validated segmentation protocol 
+\citep{Pruessner2000}.
 
@@
 \mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2).  All images from the SZ-FEP dataset are segmented.  The optimal
-size of template library, registration method, and label fusion method found in
-Experiment 1 are used.
+(see Experiment 2).  All images from the SZ-FEP dataset are segmented by \mb.  The 
+optimal size of template library, registration method, and label fusion method 
+found in Experiment 1 are used.
 
@@
 
-\todo{done elsewhere?}
+\todo{take description from below}
 
@@
 Since the hippocampal segmentation protocols differ between the ADNI labels and
-Winterburn atlases, this poses a problem for direct evaluation between labels
-produced by \mb and the ADNI labels in terms of overlap; we would not expect
-different segmentation protocols to have a high degree of overlap.  Instead, to
-evaluate the performance of \mb we compare the correlation of \mb segmentation
-volumes with manual segmentation (SNT) volumes. Additionally, we correlate the
-hippocampal volumes of established automated segmentation methods to \mb
-segmentations.
+Winterburn atlases, this poses a problem for evaluation of by measuring overlap. 
+That is, different segmentation protocols will necessarily produce segmentations
+that do not perfectly overlap, and so the degree of overlap cannot be used soley 
+to assess two segmentaiton methods that each use different protocols.
+
+Instead, we assess the correlation in hippocampal volume between \mb and manual segmentation
+(SNT). For reference, we also compute the correlation in hippocampal volume between
+the existing, established automated segmentation methods: FSL, FreeSurfer, and 
+MAPER.  Additionally, we evaluate the volume-related fixed and proportional 
+biases in all segmentation methods using Bland-Altman plots \todo{Bland-Altman ref}.
+
+\todo{references} with manual segmentations. 
+\todo{"hippocampal volume" = mean hemispheric volume?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+
+<<ADN1-xval-k-mean-mb.ma, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+ss = subset(all_data_mean, reg_method == 'ANTS' & method.mb == 'Majority Vote')
+ss = subset(ss, (templates.mb*atlases) %% 2 == 1)
+ma_lines = ddply(ss, c("atlases"), function (df) {data.frame(yint = mean(df$k.ma))})
+
+ggplot(ss,
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
+  facet_grid(reg_method~method.mb) + 
+  stat_summary(fun.y=mean,geom='line',
+               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+  stat_summary(fun.y=mean,geom='point',
+               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+  geom_hline(yintercept=ma_lines$yint, colour = c("red", "orange", "green", "cyan", "purple"), linetype="dashed") + 
+  #stat_smooth(method='lm', formula=y~log(x)) + 
+  scale_colour_hue(name="Atlases") +
+  xlab( "Number of Templates" ) + 
+  ylab( "Mean DSC" ) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \label{fig:ADNI1-xval-k_mean}
+  \caption{
+  {\bf Mean DSC of \mb segmentations across 69 ADNI1 subjects, by atlas and
+  template library size, registration algorithm, and label fusion method.} 
+  }
+\end{figure}
 
@@
 <<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ANTSmajvote = subset(all_data, 
+ants = subset(all_data_mean, 
                      reg_method=="ANTS" & 
@@
 
-data.stats = ddply(ANTSmajvote, c("subject", "label","atlases","templates.mb"), function (df) {
+ants.stats = ddply(ants, c("subject", "atlases","templates.mb"), 
+                   function (df) {
                      data.frame(
@@
 
-ggplot(subset(stats.tests, direction > 0, stat == "Variance"), 
+ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
   aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-This experiment explores \mb segmentations of hippocampal subfields. To
+This experiment explores \mb segmentation of hippocampal subfields. To
 achieve this, a leave-one-out validation is conducted in which lower-resolution
-images ($0.9mm^3$ voxels) of each Winterburn atlas subject is segmented using 
-the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
-atlases when subsampled to $0.9mm^3$ voxels are also computed.
-
-In general, across hippocampal subregions the percent error in volume between
-\mb segmentations and the manual Winterburn atlas segmentations compares
-favourably to error when resampling the atlas segmentations (Figure
+images ($0.9mm^3$ voxels) of each Winterburn atlas subject are segmented using 
+the remaining Winterburn atlases.
+
+In general, across hippocampal subregions the percent error in volume of
+\mb segmentations compares
+favourably to the error resulting from image resampling (Figure
 \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
 subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
-subregions show distinctly larger volumes than resampling error. 
+subregions show distinctly larger error than is found through resampling.  
 \todo{note direction of change}.
@@
 ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
+  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
   geom_boxplot() +
-  labs(y="Percent error in volume", x="Region") +
+  labs(y=expression(paste("Percent error in volume relative to 0.3", mm^3, " Winterburn Atlases")), x="Region") +
   scale_colour_discrete("") + 
@@
   \label{fig:subfield-montage}
+  \todo{MAGeT on BRAVO.  Also, show coronal views}
 \end{figure}
@@
 manual segmentation volumes (Figure \ref{fig:SZ-volumes}). 
-\todo{p-value}
+\todo{r and p-values}
 
@@
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
-casted = dcast(SZ_volumes, Subject + Label ~ Method)
+mean_sz_vols = ddply(SZ_volumes, c('Subject', 'Method'), function (df) {
+  data.frame(Volume = mean(df$Volume))})
+
+casted = dcast(mean_sz_vols, Subject + Label ~ Method)
 lm = lm(MAGeT ~ Manual, casted)
@@
   geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
-  xlab(expression("Manual unilateral whole hippocampus volume " (mm^3))) + 
-  ylab(expression("MAGeT unilateral whole hippocampus volume "  (mm^3))) + 
+  xlab(expression("Mean Manual hippocampus volume " (mm^3))) + 
+  ylab(expression("Mean MAGeT hippocampus volume "  (mm^3))) + 
   scale_size_continuous(guide="none") 
@@
   \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
-  HC volumes for \mb against manually rated Hippocampal volumes}
+  HC volume for \mb against manually rated Hippocampal volumes}
   \label{fig:SZ-volumes}
@@
 <<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
-melted = melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
+melted = melt(totals.complete,measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"),
               variable.name="Method", value.name="Volume")
@@
 })
-
-qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
-      colour=variable,geom="boxplot") + 
-      geom_text(aes(y=-Inf,x=c(.7,.85,1,1.15, 1.7,1.85,2,2.15, 2.7,2.85,3,3.15), vjust=-5, label=round(pearson,2)), 
-                colour = 'black', size=3, data=correlations) +
+m2 = melt(totals.complete,measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
+levels(m2$Method) = c("FreeSurfer"= "FreeSurfer", "FSL" = "FSL", "MAPER" = "MAPER", "MAGeT" = "MAGeT", "SNT" = "Manual")
+m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
+
+qplot(DX,value,data=m2,
+      colour=Method,geom="boxplot") + 
+      #geom_text(aes(y=-Inf,x=c(.7,.85,1,1.15, 1.7,1.85,2,2.15, 2.7,2.85,3,3.15), vjust=-5, label=round(pearson,2)), 
+      #          colour = 'black', size=3, data=correlations) +
       xlab("Diagnosis") + 
-      ylab(expression(paste("Bilateral whole hippocampal volume (", mm^3, ")"))) + 
-      theme(legend.direction = "horizontal", legend.position = "bottom")
+      ylab(expression(paste("Combined whole hippocampal volume (", mm^3, ")"))) + 
+      theme(legend.direction = "horizontal", legend.position = "bottom", text=element_text(size=18))
 @
@@
 <<ADN1-src-Bland-Atlman, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=6, fig.height=6>>=
-melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+names(totals.complete)[8] = "FreeSurfer"
+melted=melt(totals.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
             variable.name = "method", value.name = "volume")
+
 melted$diff = melted$SNT - melted$volume
+melted = subset(melted, (method == "FreeSurfer") | 
+                        (method == "FSL"   & diff > -5000 & diff < 900) | 
+                        (method == "MAPER" & diff > -500 & diff < 1000) | 
+                        (method == "MAGeT" & diff > -1100))
+#melted = subset(melted, method == "MAGeT")
+melted$DX = factor(melted$DX, levels(melted$DX)[c(2,3,1)])
 melted$mean = ( melted$SNT + melted$volume ) / 2
@@
 
-ggplot(melted, aes(x= mean, y = diff, colour=DX)) + 
+names(melted)[4] = "Diagnosis"
+ggplot(melted, aes(x= mean, y = diff, colour=Diagnosis)) + 
   facet_wrap( ~ method, nrow=2, scales="free") + 
@@
   geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits, hjust=1.1,vjust=-0.8, size=3) + 
-  xlab('Mean of manual and automated total hippocampal volume') + 
-  ylab('Difference between manual and automated total hippocampal volume') + 
+  xlab(expression('Mean of manual and automated volume' (mm^3))) + 
+  ylab(expression('Difference in manual and automated volumes' (mm^3))) + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
@@
   geom_smooth(method="lm", formula = y~x) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom", text=element_text(size=20))
 @
@@
 \begin{verbatim}
+
+In this manuscript we have presented the implementation and validation of the MAGeT Brain framework; a methodology that requires very few input atlases in order to provide accurate and reliable segmentations.
+We have demonstrated that our methodology robustly provides accurate and consistent segmentations in populations with different ageing and neuropsychiatric characteristics.
+Further, we have demonstrated that algorithmic performance is not dependent on a single definition of the hippocampus and can easily differing hippocampal definitions \cite{Winterburn,Pruessner,and SNT}.
+Finally, we also demonstrate that MAGeT Brain provides accurate automatic identification of the hippocampal subfields despite contrast and resolution limitations in standard T1-weighted image volumes.
+
+Throughout the cross validation in Experiment #1 (10-fold cross validation in the ADNI sample) we demonstrate that two parameter choices increase segmentation accuracy; namely increasing the number of atlases and the number of templates.
+However after setting the parameters to 5 atlases and 15 templates that there are diminishing returns with respect to this improvement.
+We were somewhat surprised to find that cross-correlation and normalized mutual information based weighted voting did little to improve segmentation accuracy.
+This suggests that using an intermediate template library generated by accurate nonlinear registration methods such as ANTs \cite{Avants|} sufficiently filters sources of error such as nonlinear registration error and partial volume effects that arise through the use of nearest neighbour resampling.
+From our previous work on the MAGeT Brain algorithm we have shown that these increases in error are not simply a smoothing or averaging effect \cite{Chakravarty}.
+Although, the goal of this manuscript was to not exhaustively test or validate multiple different voting strategies in the context of our segmentation algorithm, it is important to note that other strategies for voting are available.
+For example, other groups have used the STAPLE algorithm \cite{Warfield} (or variants of the STAPLE algorithm \cite{stuff from the Duchesne group here}).
+Can then be weighted each segmentation based upon its estimated performance level with respect to the other available candidate segmentations.
+Further, the sensitivity and specificity parameters could also be tuned to potentially improve segmentation accuracy and reliability.
+It is rather likely that using more sophisticated voting methods would have a positive effect on the overall segmentation performance.
+To this end, more work is required in order to determine the source of the slight decrease in segmentation performance when the number of templates are set to an even number.
+Our initial concern was that this dip in performance was a by-product of the MAGeT Brain algorithm itself;  however we determined that this pattern was also true in the analysis of multi-atlas segmentations we used in our experiments.
+We believe that our voting methodology is biased towards labels with the lowest numeric values when breaking ties, thus causing the slight bias observed when using an even number of templates.
+This is another area where the voting scheme could be used to improve performance.
+However, it is worth noting that this limitation was previously identified by Heckemann et al.
+\cite{Heckemann2006a} and, subsequently, other groups have not even considered the case where an even number of candidate labels are considered (eg: \cite{Leung2010}).
+
+Another concern is the moderate-to-small improvement observed in MAGeT Brain in comparison to multi-atlas segmentation when using the same number of atlases.
+The actual benefit in using MAGeT Brain is consistency of the labelling regardless of atlas or template choice.
+This is an important consideration that few have touched on previously.
+The 10-fold Monte Carlo cross validation that we present in Experiment 1 is amongst one of the most stringent performed in the multi-atlas/segmentation literature.
+To the best of our knowledge other groups have only ever performed a 3-fold validation \cite{????}.
+This suggests that our results are reflective of a true average over the choice of parameter settings and are independent of atlas or template choice (provided the input atlases are properly segmented).
+
+In comparison to other methodologies in the field, it is important to note that MAGeT Brain performs quite well.
+Table 6 gives a survey of some of the most recent multi-atlas implementations in the fields and the kappa values that are reported that validate their results using the ADNI dataset, many of which use the same SNT labels for populating their template library and for gold standards for evaluation.
+While it is difficult to compare segmentation results across studies, gold-standards, evaluation metrics, and algorithms it is worth noting that the methods summarized in Table 6 require more atlases (between 16-55) than our MAGeT Brain implementation with the Winterburn atlases \cite{Winterburn}.
+Amongst these methods only 4 yields mean Kappas that are higher than the ones that are reported here.
+There are some important differences between our method and these specific methods.
+Others have reported the difficulty with misregistrations in candidate segmentation (ie: segmentations generated that are then input in the voxel-voting procedure \cite{Collins}.
+The work of Leung \cite{Leung2010} (work also reports some of the highest kappas for the segmentation of ADNI data) tackles this problems by using an intensity threshold that is estimated heuristically at the time of segmentation.
+While this works for the ADNI dataset (which is partially homogenized with respect to image acquisition and pre-processing), it is unclear if this type of heuristic is applicable to other datasets.
+In all cases, these methods require more atlases than our implementation with the Winterburn atlases.
+Lotjonen et al.
+\cite{Lotjonen} achieve highly accurate segmentation but correct their segmentations using classifications derived using an expectation maximization framework.
+In their initial work, Chupin and colleagues develop their probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy patients.
+In subsequent methods, they retuned their methodology using the ADNI sample with a hierarchical experimentation protocol that involved.
+These methods suggest that some post-processing of the final segmentations would improve accuracy of the segmentation.
+While that may be true, there is little consensus regarding how to achieve this.
+Further, to the best of our knowledge, no other groups have validated their work using multiple atlas labels, different acquisitions, and disease populations in order to demonstrate the robustness of their technique.
+This is one of the clear strengths of this work.
+Further, unlike some of the algorithms mentioned our implementation does not require retuning for new populations or datasets as it inherently models the variability of the dataset through the template library.
+However it should be noted that the increased accuracy that follows increasing the number of atlases and templates comes at an increased computational cost (O(log(n), as previously mentioned in other work \cite{Heckemann}).
+
+In comparison to the methods that we compared and were available through the ADNI database (FreeSufer, MAPER) and the method that initialized ourselves (FSL-FIRST) we find extremely variable performance of all methods.
+With the exception of FIRST all methods correlate well with the SNT volumes provided in the ADNI database.
+However, FreeSufer and FIRST provide radically different definitions of the size of the hippocampus in comparison to the other methods.
+Further, when estimating bias of these methods relative to SNT hippocampal volumes we see that large hippocampi are over estimated while small hippocampi are under estimated.
+By comparison, our method and MAPER are far more conservative suggesting that these methods may be better suited for estimating true-positives.
+While this comparison with these methods using only volume, more work is needed to better understand the differences between our method and other methods.
+
+Finally, we have also demonstrated that our algorithmic framework is appropriate for the segmentation of hippocampal subfields in standard T1-weighted data.
+This has started to be come a burgeoning topic in the segmentation literature, although very few methods are available for the segmentation of 3T data \cite{Yushkevich and van Leemput}.
+However, the FreeSufer implementation of subfield segmentation.
+While recent work demonstrates that subfield segmentations can be used for classification of AD, MCI, and NC, there has been no explicit validation of the methodology based on accruacy or precision.
+Although the initial hippocampal subfield segmentations from the Yushkevich group have been demonstrated to work on the ADNI population, there has also been no validation of their work.
+In addition, their work requires some manual initialization to properly function.
+Our work demonstrates that we can reliably identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
+The fact that CA2/CA3 and molecular layers cannot be reasonably identified should not be surprising.
+These are extremely thin and spatially convoluted regions that originally required high-resolution MRI for identification.
+It is likely that they would be that extents of these regions are well below the resolution offered by standard T1-weighted images.
+In fact many manual segmentation methodologies do not attempt to parse these regions either \cite{Wisse, Mueller}
+
+In conclusion, we have presented a flexible multi-atlas framework.
+It has considerable advantages over other methods as only a small set of atlases is required to initialize the algorithm.
+We demonstrate that our method works robustly over hippocampal definitions, different populations, and different acquisition types.
+Finally, we also demonstrate that using this method that accurate identification of the hippocampal subfields in also possible.
 

### 2013-07-17 paper.Rnw
'nother day, 'nother dollar, Donegan

@@
 over the choice of transformation model, objective function, and the consistency
-of the final transformation. The transformation is estimated in a hierarchical
-fashion where the MRI data is subsampled, allowing large deformations to be
-estimated and successively refined at later hierarchical stages (where the data
-is subsampled to a finer grid). The deformation field and the objective function
-are regularized with a Gaussian kernel at each level of the hierarchy. The \ants
-algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an
-implementation of the \ants algorithm compatible with the MINC data format,
-mincANTS \url{https://github.com/vfonov/mincANTS}.
+of the final transformation \citep{Avants}. The transformation is estimated in a
+hierarchical fashion where the MRI data is subsampled, allowing large
+deformations to be estimated and successively refined at later hierarchical
+stages (where the data is subsampled to a finer grid). The deformation field and
+the objective function are regularized with a Gaussian kernel at each level of
+the hierarchy. The \ants algorithm is freely available
+\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the \ants
+algorithm compatible with the MINC data format, mincANTS
+\url{https://github.com/vfonov/mincANTS}.
 
@@
 subregions show distinctly larger error than is found through resampling.  
-\todo{note direction of change}.
+\todo{note direction of change.}
 
@@
   resolution), the subsampled Winterburn segmentations (at 0.9mm3-isotropic
-  voxel resolution), and the MAGeT brain labels on the subsampled atlas image.}
+  voxel resolution), and the \mb labels on the subsampled atlas image.}
   \label{fig:subfield-montage}
@@
 
+In this manuscript we have presented the implementation and validation of the
+\mb framework -- a methodology that requires very few input atlases in
+order to provide accurate and reliable segmentations.  We have demonstrated that
+our methodology robustly provides accurate and consistent segmentations in
+populations with different ageing and neuropsychiatric characteristics.
+Further, we have demonstrated that algorithmic performance is not dependent on a
+single definition of the hippocampus but is effective with differing hippocampal
+definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}.  Finally, we also demonstrate
+that \mb provides accurate automatic identification of the hippocampal
+subfields despite contrast and resolution limitations in standard T1-weighted
+image volumes.
+
+Throughout the cross validation in Experiment 1 (10-fold cross validation in the
+\adnidataset dataset subsample) we find that two parameter choices improve 
+segmentation accuracy: increasing the number of atlases, and the number of
+templates. However, after setting the parameters to 5 atlases and 15 templates
+there are diminishing returns with respect to this improvement. We were
+somewhat surprised to find that cross-correlation and normalized mutual
+information based weighted voting did little to improve segmentation accuracy.
+This suggests that using an intermediate template library generated by accurate
+nonlinear registration methods such as \ants \cite{Avants} sufficiently filters
+sources of error such as nonlinear registration error and partial volume effects
+that arise through the use of nearest neighbour resampling. From our previous
+work on the \mb algorithm we have shown that these increases in error
+are not simply a smoothing or averaging effect \cite{Chakravarty2013}.
+
+Although, the goal of this manuscript was to not exhaustively test or validate
+multiple different voting strategies in the context of our segmentation
+algorithm, it is important to note that other strategies for voting are
+available.  For example, other groups have used the STAPLE algorithm
+\cite{Warfield2004} (or variants of the STAPLE algorithm \cite{stuff from the
+Duchesne group here} which weights each segmentation based upon its estimated
+performance level with respect to the other available candidate segmentations.
+Further, the sensitivity and specificity parameters could also be tuned to
+potentially improve segmentation accuracy and reliability.  It is rather likely
+that using more sophisticated voting methods would have a positive effect on the
+overall segmentation performance.
+
+To this end, more work is required in order to determine the source of the
+slight decrease in segmentation performance when the number of templates are set
+to an even number.  Our initial concern was that this dip in performance was a
+by-product of the \mb algorithm itself.  However, we determined that
+this pattern was also true in the analysis of multi-atlas segmentations we used
+in our experiments.  We believe that our majority voting methodology is biased
+towards labels with the lowest numeric values when breaking ties, thus causing
+the slight bias observed when using an even number of templates. This is another
+area where the voting scheme could be used to improve performance. However, it
+is worth noting that this limitation was previously identified by Heckemann et
+al. \cite{Heckemann2006a} and, subsequently, other groups have not even
+considered the potential pitfalls of an even number of candidate labels (eg.
+\cite{Leung2010}).
+
+Another concern is the moderate-to-small improvement observed in \mb in
+comparison to multi-atlas segmentation when using the same number of atlases.
+The actual benefit in using \mb is consistency of the labelling regardless of
+atlas or template choice. This is an important consideration that few have
+touched on previously. The 10-fold Monte Carlo cross validation that we present
+in Experiment 1 is amongst one of the most stringent performed in the
+multi-atlas/segmentation literature.  To the best of our knowledge other groups
+have only ever performed a 3-fold validation \cite{????}.  This suggests that
+our results are reflective of a true average over the choice of parameter
+settings and are independent of atlas or template choice (provided the input
+atlases are properly segmented).
+
+\todo{additionally, we could be seeing a ceiling effect: reaching the limit of
+what is achievable with a multi-atlas-based segmentation method}
+
+In comparison to other methodologies in the field, it is important to note that
+\mb performs quite well.  Table 6 gives a survey of some of the most recent
+multi-atlas implementations with the reported kappa values used to validate the
+algorithm on the ADNI dataset (many of which use the same SNT labels for
+populating their template library and as gold standards for evaluation).  While
+it is difficult to compare segmentation results across studies, gold-standards,
+evaluation metrics, and algorithms it is worth noting that the methods
+summarized in Table 6 require more atlases (between 16-55) than our \mb
+implementation with the Winterburn atlases \cite{Winterburn2013}.  Amongst these
+methods only method 4 yields mean Kappas that are higher than the ones that are
+reported here.
+
+There are some important differences between our method
+and these specific methods.  Others have reported the difficulty with
+misregistrations in candidate segmentation (i.e. segmentations generated that are
+then input in the voxel-voting procedure \cite{Collins2010}). The work of Leung
+\cite{Leung2010} (work also reports some of the highest kappas for the
+segmentation of ADNI data) tackles this problems by using an intensity threshold
+that is estimated heuristically at the time of segmentation.  While this works
+for the ADNI dataset (which is partially homogenized with respect to image
+acquisition and pre-processing), it is unclear if this type of heuristic is
+applicable to other datasets.  In all cases, these methods require more atlases
+than our implementation with the Winterburn atlases.  Lotjonen et al.
+\cite{Lotjonen2010} achieve highly accurate segmentation but correct their
+segmentations using classifications derived using an expectation maximization
+framework.  In their initial work, Chupin and colleagues develop their
+probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy
+patients.  In subsequent methods, they retuned their methodology using the ADNI
+sample with a hierarchical experimentation protocol that involved.  These
+methods suggest that some post-processing of the final segmentations would
+improve accuracy of the segmentation.  While that may be true, there is little
+consensus regarding how to achieve this.
+
+To the best of our knowledge, no other groups have validated their work using
+multiple atlas segmentation protocols, different acquisitions, and disease
+populations in order to demonstrate the robustness of their technique.  This is
+one of the clear strengths of this work.  Furthermore, unlike some of the
+algorithms mentioned, our implementation does not require retuning for new
+populations or datasets as it inherently models the variability of the dataset
+through the template library.  However it should be noted that the increased
+accuracy that follows increasing the number of atlases and templates comes at an
+increased computational cost ($\Omicron(log(n))$), as previously mentioned in
+other work \cite{Heckemann}).
+
+In comparison to the methods that we compared and were available through the
+ADNI database (FreeSufer, MAPER) and the method that we initialized ourselves
+(FSL-FIRST) we find extremely variable performance of all methods.  With the
+exception of FSL all methods correlate well with the SNT volumes provided in
+the ADNI database.  However, FreeSufer and FIRST provide radically different
+definitions of the size of the hippocampus in comparison to the other methods.
+Further, when estimating bias of these methods relative to SNT hippocampal
+volumes we see that large hippocampi are over estimated while small hippocampi
+are under estimated.  By comparison, our method and MAPER are far more
+conservative suggesting that these methods may be better suited for estimating
+true-positives.  While this comparison with these methods using only volume,
+more work is needed to better understand the differences between our method and
+other methods.
+
+Finally, we have also demonstrated that our algorithmic framework is appropriate
+for the segmentation of hippocampal subfields in standard T1-weighted data.
+This has started to become a burgeoning topic in the segmentation literature,
+although very few methods are available for the segmentation of 3T data
+\cite{Yushkevich2009, and VanLeemput2009}.  However, the FreeSufer implementation of
+subfield segmentation.  While recent work demonstrates that subfield
+segmentations can be used for classification of AD, MCI, and NC, there has been
+no explicit validation of the methodology based on accruacy or precision.
+Although the initial hippocampal subfield segmentations from the Yushkevich
+group have been demonstrated to work on the ADNI population, there has also been
+no validation of their work.  In addition, their work requires some manual
+initialization to properly function.  Our work demonstrates that we can reliably
+identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
+The fact that CA2/CA3 and molecular layers cannot be reasonably identified
+should not be surprising.  These are extremely thin and spatially convoluted
+regions that originally required high-resolution MRI for identification.  It is
+likely that they would be that extents of these regions are well below the
+resolution offered by standard T1-weighted images.  In fact many manual
+segmentation methodologies do not attempt to parse these regions either
+\cite{Wisse2012, Mueller2009}
+
+In conclusion, we have presented a flexible multi-atlas framework.  It has
+considerable advantages over other methods as only a small set of atlases is
+required to initialize the algorithm.  We demonstrate that our method works
+robustly over hippocampal definitions, different populations, and different
+acquisition types.  Finally, we also demonstrate that using this method that
+accurate identification of the hippocampal subfields in also possible.
+
 \begin{verbatim}
-
-In this manuscript we have presented the implementation and validation of the MAGeT Brain framework; a methodology that requires very few input atlases in order to provide accurate and reliable segmentations.
-We have demonstrated that our methodology robustly provides accurate and consistent segmentations in populations with different ageing and neuropsychiatric characteristics.
-Further, we have demonstrated that algorithmic performance is not dependent on a single definition of the hippocampus and can easily differing hippocampal definitions \cite{Winterburn,Pruessner,and SNT}.
-Finally, we also demonstrate that MAGeT Brain provides accurate automatic identification of the hippocampal subfields despite contrast and resolution limitations in standard T1-weighted image volumes.
-
-Throughout the cross validation in Experiment #1 (10-fold cross validation in the ADNI sample) we demonstrate that two parameter choices increase segmentation accuracy; namely increasing the number of atlases and the number of templates.
-However after setting the parameters to 5 atlases and 15 templates that there are diminishing returns with respect to this improvement.
-We were somewhat surprised to find that cross-correlation and normalized mutual information based weighted voting did little to improve segmentation accuracy.
-This suggests that using an intermediate template library generated by accurate nonlinear registration methods such as ANTs \cite{Avants|} sufficiently filters sources of error such as nonlinear registration error and partial volume effects that arise through the use of nearest neighbour resampling.
-From our previous work on the MAGeT Brain algorithm we have shown that these increases in error are not simply a smoothing or averaging effect \cite{Chakravarty}.
-Although, the goal of this manuscript was to not exhaustively test or validate multiple different voting strategies in the context of our segmentation algorithm, it is important to note that other strategies for voting are available.
-For example, other groups have used the STAPLE algorithm \cite{Warfield} (or variants of the STAPLE algorithm \cite{stuff from the Duchesne group here}).
-Can then be weighted each segmentation based upon its estimated performance level with respect to the other available candidate segmentations.
-Further, the sensitivity and specificity parameters could also be tuned to potentially improve segmentation accuracy and reliability.
-It is rather likely that using more sophisticated voting methods would have a positive effect on the overall segmentation performance.
-To this end, more work is required in order to determine the source of the slight decrease in segmentation performance when the number of templates are set to an even number.
-Our initial concern was that this dip in performance was a by-product of the MAGeT Brain algorithm itself;  however we determined that this pattern was also true in the analysis of multi-atlas segmentations we used in our experiments.
-We believe that our voting methodology is biased towards labels with the lowest numeric values when breaking ties, thus causing the slight bias observed when using an even number of templates.
-This is another area where the voting scheme could be used to improve performance.
-However, it is worth noting that this limitation was previously identified by Heckemann et al.
-\cite{Heckemann2006a} and, subsequently, other groups have not even considered the case where an even number of candidate labels are considered (eg: \cite{Leung2010}).
-
-Another concern is the moderate-to-small improvement observed in MAGeT Brain in comparison to multi-atlas segmentation when using the same number of atlases.
-The actual benefit in using MAGeT Brain is consistency of the labelling regardless of atlas or template choice.
-This is an important consideration that few have touched on previously.
-The 10-fold Monte Carlo cross validation that we present in Experiment 1 is amongst one of the most stringent performed in the multi-atlas/segmentation literature.
-To the best of our knowledge other groups have only ever performed a 3-fold validation \cite{????}.
-This suggests that our results are reflective of a true average over the choice of parameter settings and are independent of atlas or template choice (provided the input atlases are properly segmented).
-
-In comparison to other methodologies in the field, it is important to note that MAGeT Brain performs quite well.
-Table 6 gives a survey of some of the most recent multi-atlas implementations in the fields and the kappa values that are reported that validate their results using the ADNI dataset, many of which use the same SNT labels for populating their template library and for gold standards for evaluation.
-While it is difficult to compare segmentation results across studies, gold-standards, evaluation metrics, and algorithms it is worth noting that the methods summarized in Table 6 require more atlases (between 16-55) than our MAGeT Brain implementation with the Winterburn atlases \cite{Winterburn}.
-Amongst these methods only 4 yields mean Kappas that are higher than the ones that are reported here.
-There are some important differences between our method and these specific methods.
-Others have reported the difficulty with misregistrations in candidate segmentation (ie: segmentations generated that are then input in the voxel-voting procedure \cite{Collins}.
-The work of Leung \cite{Leung2010} (work also reports some of the highest kappas for the segmentation of ADNI data) tackles this problems by using an intensity threshold that is estimated heuristically at the time of segmentation.
-While this works for the ADNI dataset (which is partially homogenized with respect to image acquisition and pre-processing), it is unclear if this type of heuristic is applicable to other datasets.
-In all cases, these methods require more atlases than our implementation with the Winterburn atlases.
-Lotjonen et al.
-\cite{Lotjonen} achieve highly accurate segmentation but correct their segmentations using classifications derived using an expectation maximization framework.
-In their initial work, Chupin and colleagues develop their probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy patients.
-In subsequent methods, they retuned their methodology using the ADNI sample with a hierarchical experimentation protocol that involved.
-These methods suggest that some post-processing of the final segmentations would improve accuracy of the segmentation.
-While that may be true, there is little consensus regarding how to achieve this.
-Further, to the best of our knowledge, no other groups have validated their work using multiple atlas labels, different acquisitions, and disease populations in order to demonstrate the robustness of their technique.
-This is one of the clear strengths of this work.
-Further, unlike some of the algorithms mentioned our implementation does not require retuning for new populations or datasets as it inherently models the variability of the dataset through the template library.
-However it should be noted that the increased accuracy that follows increasing the number of atlases and templates comes at an increased computational cost (O(log(n), as previously mentioned in other work \cite{Heckemann}).
-
-In comparison to the methods that we compared and were available through the ADNI database (FreeSufer, MAPER) and the method that initialized ourselves (FSL-FIRST) we find extremely variable performance of all methods.
-With the exception of FIRST all methods correlate well with the SNT volumes provided in the ADNI database.
-However, FreeSufer and FIRST provide radically different definitions of the size of the hippocampus in comparison to the other methods.
-Further, when estimating bias of these methods relative to SNT hippocampal volumes we see that large hippocampi are over estimated while small hippocampi are under estimated.
-By comparison, our method and MAPER are far more conservative suggesting that these methods may be better suited for estimating true-positives.
-While this comparison with these methods using only volume, more work is needed to better understand the differences between our method and other methods.
-
-Finally, we have also demonstrated that our algorithmic framework is appropriate for the segmentation of hippocampal subfields in standard T1-weighted data.
-This has started to be come a burgeoning topic in the segmentation literature, although very few methods are available for the segmentation of 3T data \cite{Yushkevich and van Leemput}.
-However, the FreeSufer implementation of subfield segmentation.
-While recent work demonstrates that subfield segmentations can be used for classification of AD, MCI, and NC, there has been no explicit validation of the methodology based on accruacy or precision.
-Although the initial hippocampal subfield segmentations from the Yushkevich group have been demonstrated to work on the ADNI population, there has also been no validation of their work.
-In addition, their work requires some manual initialization to properly function.
-Our work demonstrates that we can reliably identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
-The fact that CA2/CA3 and molecular layers cannot be reasonably identified should not be surprising.
-These are extremely thin and spatially convoluted regions that originally required high-resolution MRI for identification.
-It is likely that they would be that extents of these regions are well below the resolution offered by standard T1-weighted images.
-In fact many manual segmentation methodologies do not attempt to parse these regions either \cite{Wisse, Mueller}
-
-In conclusion, we have presented a flexible multi-atlas framework.
-It has considerable advantages over other methods as only a small set of atlases is required to initialize the algorithm.
-We demonstrate that our method works robustly over hippocampal definitions, different populations, and different acquisition types.
-Finally, we also demonstrate that using this method that accurate identification of the hippocampal subfields in also possible.
-
 Experiment 1: 

### 2013-07-19 paper.Rnw
References fixes, author affiliations, figure fixes

@@
 \usepackage[section]{placeins}  % keep floats in their place.
+\usepackage{authblk}    % Listing Author affiliations
 
@@
 
-% add \subsubsubsection
-% \setcounter{secnumdepth}{4}  
-% \titleformat{\paragraph}
-% {\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-% \titlespacing*{\paragraph}
-% {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+% styling
 \newcommand{\subsubsubsection}[1]{\paragraph{#1}}
@@
 \newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T }
-%notes and TODO formatting
-%\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
-%\newcommand{\comment}[1]{\begin{kframe}{\textcolor{red}{#1}}\end{kframe}}
-%\newcommand{\todo}[1]{\comment{TODO #1}}
-%\newcommand{\mc}[1]{\comment{MC: #1}}
+
+% title and authors
+\title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
+\author[1]{Jon Pipitone}
+\author[1]{Matt T. Park}
+\author[1]{Julie Winterburn}
+\author[1]{Tristram A. Lett}
+\author[2,3]{Jason P. Lerch}
+\author[4]{Jens C. Pruessner}
+\author[4]{Martin Lepage}
+\author[1,5]{Aristotle Voineskos}
+\author[1,5,6]{M. Mallar Chakravarty}
+\author{Alzheimer's Disease Neuroimaging Initiative}
+\affil[1]{Kimel Family Translational Imaging-Genetics Lab, Centre for Addiction and
+Mental Health, Toronto, ON, Canada}
+\affil[2]{Neurosciences and Mental Health, Hospital for Sick Children, Toronto,
+ON,Canada}
+\affil[3]{Department of Medical Biophysics, University of Toronto, Toronto, ON,
+Canada}
+\affil[4]{Brain Imaging Group, Douglas Hospital Research Centre, Verdun, QC, Canada}
+\affil[5]{Department of Psychiatry, University of Toronto, Toronto, ON, Canada}
+\affil[6]{Institute of Biomedical Engineering, University of Toronto, Toronto,
+ON, Canada}
+
+\renewcommand\Authands{ and }
+
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
-
-\title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
-\author{Pipitone J., Park M., Winterburn J., Lett, T., Lerch J., Pruessner J., Lepage M., \\ 
-Voineskos A., Chakravarty M.M., and \\ 
-the Alzheimer's Disease Neuroimaging Initiative}
-\todo{Full names}
-\todo{Associations}
 \maketitle
@@
 \begin{abstract}
-Neuroimaging research often relies on automated anatomical segmentations of MR
-images of the brain. Current multi-atlas based approaches provide accurate
-segmentations of brain images by propagating manually derived segmentations of
-specific neuroanatomical structures to unlabelled data. These approaches often
-rely on a large number of such manually segmented atlases that take significant
-time and expertise to produce. We present an algorithm for the automatic
-segmentation of the hippocampus that minimizes the number of atlases needed
-while still achieving similar accuracy to multi-atlas approaches.
-\todo[inline]{finish}
+%Neuroimaging research often relies on automated anatomical segmentations of MR
+%images of the brain. Current multi-atlas based approaches provide accurate
+%segmentations of brain images by propagating manually derived segmentations of
+%specific neuroanatomical structures to unlabelled data. These approaches often
+%rely on a large number of such manually segmented atlases that take significant
+%time and expertise to produce. We present an algorithm for the automatic
+%segmentation of the hippocampus that minimizes the number of atlases needed
+%while still achieving similar accuracy to multi-atlas approaches.
 
@@
 required for manual segmentation is prohibitive
-\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007,Weiner2013}.
+\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}.
 Further, there is little agreement between researchers regarding how exactly the
@@
 successfully in several contexts
-\citep{MallarMedIA200809,MallarHBM20082009,Collins1995,Haller1997}, it is limited
+\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
 in accuracy by the introduction of errors due inaccuracies in the nonlinear
@@
 accommodate atlases that are somehow exceptional such as those derived from
-serial histological data \citep{Chakravarty06a,Yelnik2007} or high-resolution MRI
+serial histological data \citep{Chakravarty2006,Yelnik2007} or high-resolution MRI
 data that enables robust identification of hippocampal subfields
@@
 segmentation of the striatum, pallidum, and thalamus using a single atlas
-derived from serial histological data \citep{Chakravarty2013} is possible.  The
+derived from serial histological data \citep{MallarChakravarty2012} is possible.  The
 novelty of this manuscript is the extension our multi-atlas methodology to the
@@
 over the choice of transformation model, objective function, and the consistency
-of the final transformation \citep{Avants}. The transformation is estimated in a
+of the final transformation \citep{Avants2008}. The transformation is estimated in a
 hierarchical fashion where the MRI data is subsampled, allowing large
@@
   atlas labels after linear registration to the template, dilated by three voxels
-  \citep{Chakravarty2012}.  The number of top ranked template library image labels
+  \citep{MallarChakravarty2012}.  The number of top ranked template library image labels
   is a configurable parameter and displayed as the size of the template library
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<ADNI-src-volumes-prep,echo=F,cache=T>>=
-totals.complete  <- read.csv('data/cache/ADNI1:totals.complete.csv')
-totals           <- read.csv('data/cache/ADNI1:totals.csv')
-package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
+<<ADNI-volumes-prep,echo=F,cache=T>>=
+means.complete   <- read.csv(gzfile('data/cache/ADNI1:means.complete.csv.gz'))
+package_totals   <- read.csv(gzfile('data/cache/ADNI1:package_totals.csv.gz'))
 @
@@
 %                                                                             
-<<ADNI1-scr-demographics, echo=F, results="asis",cache=FALSE>>=
+<<ADNI-demographics, echo=F, results="asis",cache=FALSE>>=
 # RID < 2000 ensures only ADNI1 patients
@@
 <<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-all_data  <- read.csv('data/cache/ADNI-XVAL:all_data.csv')
-all_data_mean  <- read.csv('data/cache/ADNI-XVAL:all_data_mean.csv')
+all_data  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data.csv.gz'))
+all_data_mean  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data_mean.csv.gz'))
 @
@@
   \label{fig:subfield-montage}
+\end{figure}
+
 \todo{MAGeT on BRAVO.  Also, show coronal views}
-\end{figure}
 
@@
 
-casted = dcast(mean_sz_vols, Subject + Label ~ Method)
+casted = dcast(mean_sz_vols, Subject ~ Method, value.var = "Volume")
 lm = lm(MAGeT ~ Manual, casted)
@@
 are included in the following analysis (a total of
-\Sexpr{length(totals.complete$RID)} images; Table 
+\Sexpr{length(means.complete$RID)} images; Table 
 \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals}).
@@
 We find a close relationship in total bilateral hippocampal volume between
-all methods and manually segmented volumes (Figure \ref{fig:ADNI-scr-volumes-plot}). 
+all methods and manually segmented volumes (Figure \ref{fig:ADNI-volumes-plot}). 
 Volumes are correlated with Pearson-r > 0.78 for all methods across disease categories.
-Within disease categories (Figure \ref{fig:ADNI-scr-volumes-boxplot}),
+Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}),
 \mb is consistently well correlated to manual volumes (Pearson-r > 0.85), but appears to 
@@
 To investigate the level of agreement with manually segmented hippocampal volumes, 
-we constructed Bland-Altman plots for each method (Figure \ref{fig:ADNI1-src-Bland-Altman}).  
+we constructed Bland-Altman plots for each method (Figure \ref{fig:ADNI1-Bland-Altman}).  
 As Bland \& Altman, 1985 \cite{Bland1986} noted, high 
@@
 imply agreement (as correlation can be driven by a large range in true values, 
-for instance).  What is most striking in Figure \ref{fig:ADNI1-src-Bland-Altman}
+for instance).  What is most striking in Figure \ref{fig:ADNI1-Bland-Altman}
 is that all methods show an obvious proportional bias: FreeSurfer and FSL 
@@
 
-Figure \ref{fig:ADNI-scr-subfields} shows a qualitative comparison of \mb and 
+Figure \ref{fig:ADNI-subfields} shows a qualitative comparison of \mb and 
 manual (SNT) hippocampal segmentations for 10 randomly selected subjects in 
@@
 %                        
-<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
+<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
 caption = paste(
@@
 \begin{figure}[h]
-<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
-melted = melt(totals.complete,measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"),
+<<ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
+melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
               variable.name="Method", value.name="Volume")
@@
 })
-m2 = melt(totals.complete,measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
-levels(m2$Method) = c("FreeSurfer"= "FreeSurfer", "FSL" = "FSL", "MAPER" = "MAPER", "MAGeT" = "MAGeT", "SNT" = "Manual")
+m2 = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
+levels(m2$Method) = c("FS"= "FreeSurfer", "FSL" = "FSL", "MAPER" = "MAPER", "MAGeT" = "MAGeT", "SNT" = "Manual")
 m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
@@
       xlab("Diagnosis") + 
-      ylab(expression(paste("Combined whole hippocampal volume (", mm^3, ")"))) + 
+      ylab(expression(paste("Mean hemispheric hippocampal volume (", mm^3, ")"))) + 
       theme(legend.direction = "horizontal", legend.position = "bottom", text=element_text(size=18))
@@
   with manual volumes are shown below each box-and-whisker.}
-  \label{fig:ADNI-scr-volumes-boxplot}
+  \label{fig:ADNI-volumes-boxplot}
 \end{figure}
@@
 \begin{figure}[h]
-<<ADNI-scr-volumes-plot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
-melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+<<ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
             variable.name = "Method", value.name = "volume")
 
-df = totals.complete
+df = means.complete
 
@@
     geom_smooth(method="lm") + 
-    xlab(expression("Manual bilateral whole hippocampal volume " (mm^3))) + 
-    ylab(expression("Automated bilateral whole hippocampal volume " (mm^3))) +
-    annotate("text", y = c(3000,2500,2000,1500), x=Inf-100, size=3,
-              label = c(paste("FS r=", round(cor(df$SNT, df$FS),2)),
-                        paste("FSL r=", round(cor(df$SNT, df$FSL),2)),
-                        paste("MAPER r=", round(cor(df$SNT, df$MAPER),2)),
-                        paste("MAGeT r=", round(cor(df$SNT, df$MAGeT),2))),
-              hjust=1.1,vjust=1.5) +
+    xlab(expression("Manual mean hemispheric hippocampal volume " (mm^3))) + 
+    ylab(expression("Automated mean hemispheric hippocampal volume " (mm^3))) +
+#    annotate("text", y = c(3000,2500,2000,1500), x=Inf-100, size=3,
+#              label = c(paste("FS r=", round(cor(df$SNT, df$FS),2)),
+#                        paste("FSL r=", round(cor(df$SNT, df$FSL),2)),
+#                        paste("MAPER r=", round(cor(df$SNT, df$MAPER),2)),
+#                        paste("MAGeT r=", round(cor(df$SNT, df$MAGeT),2))),
+#              hjust=1.1,vjust=1.5) +
     theme(legend.direction = "horizontal", legend.position = "bottom")
@@
   are shown for each method.}
-  \label{fig:ADNI-scr-volumes-plot}
+  \label{fig:ADNI-volumes-plot}
 \end{figure}
@@
 \begin{figure}
-<<ADN1-src-Bland-Atlman, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=6, fig.height=6>>=
-names(totals.complete)[8] = "FreeSurfer"
-melted=melt(totals.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
+<<ADN1-Bland-Atlman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=6, fig.height=6>>=
+names(means.complete)[8] = "FreeSurfer"
+melted=melt(means.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
             variable.name = "method", value.name = "volume")
@@
   geom_smooth(method="lm", formula = y~x) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", text=element_text(size=20))
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   agreement ($\pm 1.96SD$) are shown by dashed horizontal lines.}
-  \label{fig:ADNI-scr-Bland-Altman}
+  \label{fig:ADNI-Bland-Altman}
 \end{figure}
@@
   {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb.}
-  \label{fig:ADNI-scr-subfields}
+  \label{fig:ADNI-subfields}
 \end{figure}
@@
 This suggests that using an intermediate template library generated by accurate
-nonlinear registration methods such as \ants \cite{Avants} sufficiently filters
+nonlinear registration methods such as \ants \cite{Avants2008} sufficiently filters
 sources of error such as nonlinear registration error and partial volume effects
@@
 work on the \mb algorithm we have shown that these increases in error
-are not simply a smoothing or averaging effect \cite{Chakravarty2013}.
+are not simply a smoothing or averaging effect \cite{MallarChakravarty2012}.
 
@@
 available.  For example, other groups have used the STAPLE algorithm
-\cite{Warfield2004} (or variants of the STAPLE algorithm \cite{stuff from the
-Duchesne group here} which weights each segmentation based upon its estimated
+\cite{Warfield2004} (or variants of the STAPLE algorithm \cite{Duchesne2012})
+which weights each segmentation based upon its estimated
 performance level with respect to the other available candidate segmentations.
@@
 multi-atlas/segmentation literature.  To the best of our knowledge other groups
-have only ever performed a 3-fold validation \cite{????}.  This suggests that
+have only ever performed a 3-fold validation \cite{?}.  This suggests that
 our results are reflective of a true average over the choice of parameter
@@
 accuracy that follows increasing the number of atlases and templates comes at an
-increased computational cost ($\Omicron(log(n))$), as previously mentioned in
-other work \cite{Heckemann}).
+increased computational cost ($O(log(n))$), as previously mentioned in
+other work \cite{Heckemann2006}). \todo{check this: time complexity seems suspicious}
 
@@
 although very few methods are available for the segmentation of 3T data
-\cite{Yushkevich2009, and VanLeemput2009}.  However, the FreeSufer implementation of
+\cite{Yushkevich2009,VanLeemput2009}. However, the FreeSufer implementation of
 subfield segmentation. While recent work demonstrates that subfield
@@
 
-\begin{verbatim}
-Experiment 1: 
-- address the absence of weighted-voting effects. 
-  - hypothesis: the benefits of weighted voting only outweigh the resampling
-    error effects when choosing from a large library.  In this experiment, the 
-    template library (during weighted voting) consists of only 20 templates, so
-    the most we could hope for is a "squashing" upwards of the curve towards 
-    the 20-template limit (which is the same across all cases in this experiment).
-  - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
-    Perhaps in future experiments with larger template libraries, this could be
-    explored further.
-- Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
-  - couldn't get STAPLE to work with our image formats
-  - Idea: expect smoothness across range of templates (no even # dips, below)
-  - perhaps more sophisticated fusion methods would boost results over majority 
-    vote based techniques.
-
-- why does MAGeT brain show a dip in average Kappa when using an even number
-  templates. Does MA show this same pattern? (yes)
-  - Hypothesis: our voting method is biased when breaking ties to choose the 
-    label with the lowest numeric value.
-  - If we compute the total number of labels fused (atlases * templates) then 
-    both MB/MA perform worse when fusing an even number of labels.
-
-- we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
-  using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
-  this increase justify the extra effort involved in MB? 
-  - decrease in variability
-  - we are comparing "true" averages (see above)... does this make our criteria 
-    for a worthy improvement less strict?
-
-- how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
-- because of the extensive cross-validation in experiment 1, we are very likely
-  showing results that approach the true average of MB and MA on that dataset
-  (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
-  each).
-- even with this caveat, do we think we do well enough?  i.e. other than parameter 
-  tuning, to what extent does this experiment tell us about how MB would do in 
-  practice.
-
-- effects we are seeing is not only averaging effects (Chakravarty et al. 2012).
-
-- take aways: as in other studies (Aljabar, Heckemann) we find that performance
-  scales with the number of inputs (approx. log(n)), and that a large enough
-  template library can boost performance with a small number of atlases. The use 
-  of templates /can/ have a negative impact on performance (Figure 2), that is 
-  balanced by the improvement of growing the template library.  Likely the negative
-  impact on peformance is as a result of resampling/mislabelling error. In other 
-  words, tuning MAGeT brain involves balancing the tension between an improvement 
-  in peformance due to increase neuroanatomical capture and decreased peformance
-  due to resampling error.
-
-Experiment 2: 
-- describe resampling error during downsampling (essentially partial volume effect 
-  due to averaging/majority vote in nearest neighbour selection)
-- Why would some structures have greater downsampling error than others? (i.e. what
-  is it about a structure that would make it especially prone to resampling error)
-  - since error is discretised to whole voxels, smaller regions will show larger
-    percent error
-  - average unilateral volume (approx; mm3): 
-     region   volume   downsampling error
-     CA1      800            2%
-     CA4/DG   600          -30%
-     SR/SL/SM 700           30%
-     Sibculum 350            0%
-     CA2/3    200           15%
-
-- observation: when downsampling error is small (<10 percent), MB error is larger, 
-  and vice versa when downsampling error is large.  Could this a case where the 
-  inevitable MAGeT resampling error outweighs the small downsampling error?
-- observation: MAGeT produces similar volumes for the downsampled and BRAVO images
-  which demonstrates reliability. Additionally, MB's error is in the same *direction*
-  as the downsampling error except for SR/SL/SM
-- for MB to show less error than resampling means that voting across templates is 
-  in aggregate performing better than local nearest neighbour fitting. presumably 
-  we'd see the same improvement with basic multiatlas as well.
-
-- take away: MB produces subregion volumes that are comparable or better than
-  resampling error, except for the CA2/3 where error is near 25 percent (but 
-  even then resampling error is ~15%)
-
-Experiment 3: 
-- what is an "acceptable" r^2 value? 
-- take away: MB has proven to be robust with atlases derived from three different
-  segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
-  populations (older with AD progression; young and healthy; young and SZ).
-
-Experiment 4: 
-- address the smaller difference in mean volume across disease classes. smaller 
-  than every other method.
-
-- address the segmentation bias towards over-estimating smaller hippocampi, and 
-  underestimating larger HC
-
-- visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
-  is lower than the other methods (but we may be biased. :-).
-
-- take away: MB, with the Winterburn atlases, produces HC volumes more inline with
-  SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required.
-
-\end{verbatim}
+% \begin{verbatim}
+% Experiment 1: 
+% - address the absence of weighted-voting effects. 
+%   - hypothesis: the benefits of weighted voting only outweigh the resampling
+%     error effects when choosing from a large library.  In this experiment, the 
+%     template library (during weighted voting) consists of only 20 templates, so
+%     the most we could hope for is a "squashing" upwards of the curve towards 
+%     the 20-template limit (which is the same across all cases in this experiment).
+%   - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
+%     Perhaps in future experiments with larger template libraries, this could be
+%     explored further.
+% - Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
+%   - couldn't get STAPLE to work with our image formats
+%   - Idea: expect smoothness across range of templates (no even # dips, below)
+%   - perhaps more sophisticated fusion methods would boost results over majority 
+%     vote based techniques.
+% 
+% - why does MAGeT brain show a dip in average Kappa when using an even number
+%   templates. Does MA show this same pattern? (yes)
+%   - Hypothesis: our voting method is biased when breaking ties to choose the 
+%     label with the lowest numeric value.
+%   - If we compute the total number of labels fused (atlases * templates) then 
+%     both MB/MA perform worse when fusing an even number of labels. 
+% 
+% - we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
+%   using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
+%   this increase justify the extra effort involved in MB? 
+%   - decrease in variability
+%   - we are comparing "true" averages (see above)... does this make our criteria 
+%     for a worthy improvement less strict? 
+%     
+% 
+% - how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
+% - because of the extensive cross-validation in experiment 1, we are very likely
+%   showing results that approach the true average of MB and MA on that dataset
+%   (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
+%   each).
+% - even with this caveat, do we think we do well enough?  i.e. other than parameter 
+%   tuning, to what extent does this experiment tell us about how MB would do in 
+%   practice. 
+% 
+% - effects we are seeing is not only averaging effects (Chakravarty et al. 2012). 
+% 
+% - take aways: as in other studies (Aljabar, Heckemann) we find that performance
+%   scales with the number of inputs (approx. log(n)), and that a large enough
+%   template library can boost performance with a small number of atlases. The use 
+%   of templates /can/ have a negative impact on performance (Figure 2), that is 
+%   balanced by the improvement of growing the template library.  Likely the negative
+%   impact on peformance is as a result of resampling/mislabelling error. In other 
+%   words, tuning MAGeT brain involves balancing the tension between an improvement 
+%   in peformance due to increase neuroanatomical capture and decreased peformance
+%   due to resampling error. 
+% 
+% Experiment 2: 
+% - describe resampling error during downsampling (essentially partial volume effect 
+%   due to averaging/majority vote in nearest neighbour selection)
+% - Why would some structures have greater downsampling error than others? (i.e. what
+%   is it about a structure that would make it especially prone to resampling error)
+%   - since error is discretised to whole voxels, smaller regions will show larger
+%     percent error
+%   - average unilateral volume (approx; mm3): 
+%      region   volume   downsampling error
+%      CA1      800            2%
+%      CA4/DG   600          -30%
+%      SR/SL/SM 700           30%
+%      Sibculum 350            0%
+%      CA2/3    200           15%
+% 
+% - observation: when downsampling error is small (<10 percent), MB error is larger, 
+%   and vice versa when downsampling error is large.  Could this a case where the 
+%   inevitable MAGeT resampling error outweighs the small downsampling error?
+% - observation: MAGeT produces similar volumes for the downsampled and BRAVO images
+%   which demonstrates reliability. Additionally, MB's error is in the same *direction*
+%   as the downsampling error except for SR/SL/SM
+% - for MB to show less error than resampling means that voting across templates is 
+%   in aggregate performing better than local nearest neighbour fitting. presumably 
+%   we'd see the same improvement with basic multiatlas as well.
+% 
+% - take away: MB produces subregion volumes that are comparable or better than
+%   resampling error, except for the CA2/3 where error is near 25 percent (but 
+%   even then resampling error is ~15%)
+% 
+% Experiment 3: 
+% - what is an "acceptable" r^2 value? 
+% - take away: MB has proven to be robust with atlases derived from three different
+%   segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
+%   populations (older with AD progression; young and healthy; young and SZ). 
+%   
+% Experiment 4: 
+% - address the smaller difference in mean volume across disease classes. smaller 
+%   than every other method. 
+%   
+% - address the segmentation bias towards over-estimating smaller hippocampi, and 
+%   underestimating larger HC
+%   
+% - visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
+%   is lower than the other methods (but we may be biased. :-). 
+%   
+% - take away: MB, with the Winterburn atlases, produces HC volumes more inline with
+%   SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required. 
+% 
+% \end{verbatim}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-07-19 cache_analysis.R

@@ -1,3 +1,8 @@
+library(Hmisc)
+library(reshape2)
+library(ADNIMERGE)
+library(plyr)
+
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # ADNI1-XVAL: prep
 # Prepares the data in a the form needed for plotting.
@@ -8,8 +13,8 @@ t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz
 diagnoses    = read.csv('data/a2a_diagnoses.csv')
 
 # simplify the data
-t_data       = subset(t_data, atlases %in% seq(1,9,1))
-a_data       = subset(a_data, atlases %in% seq(1,9,1))
+t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
+a_data       = subset(a_data, atlases %in% c(1,3,5,7,9))
 t_data$se    = NULL
 t_data$sn    = NULL
 t_data$j     = NULL
@@ -32,18 +37,15 @@ a_mb = subset(a_data, approach=="mb")
 ma = rbind(t_ma,a_ma)
 mb = rbind(t_mb,a_mb)
 
-ma$num_labels = ma$atlases
-mb$num_labels = mb$atlases * mb$templates 
-
 # set up a few equivalences for multi-atlas, to make nomenclature simpler
 ma$templates =  ma$atlases  # because no template library
 mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n 
 
 # aggregate over batches
 ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
-                    templates + top_n + label + volume, data = ma, mean)
+                    templates + top_n + label, data = ma, mean)
 mb_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
-                    templates + top_n + label + volume, data = mb, mean)                    
+                    templates + top_n + label, data = mb, mean)                    
 
 all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label",
                               "reg_method", "subject"), 
@@ -128,15 +130,14 @@ package_totals = data.frame(
 all_qc = subset(combined, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )
 
 attach(all_qc)
[…]

### 2013-07-21 paper.Rnw
Figures formatted nicely, lots of other fixes.
In particular, I went through all of the TODOs and patched them up,
except for those that will require more thinking/work/or input.
For instance, the montages are still not complete, but I elaborated
the todo's with figure references so that other readers won't be
weirded out.

Also, I fixed a bunch of references, and put in citet references
where necessary.

Lastly, in addition to many wording changes and spelling/grammar
fixes, I added in errorbars on the ADNI-Xval plot, changed plot
text and point sizes, and included p/t-values for Pearson
corr

@@
 options(digits=3)
-theme_set(theme_bw())
+theme_set(theme_bw(base_size = 12))
 
@@
 \newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T }
+\newcommand{\FEPdataset}{SZ-FEP }
 
@@
 segmentation of the striatum, pallidum, and thalamus using a single atlas
-derived from serial histological data \citep{MallarChakravarty2012} is possible.  The
-novelty of this manuscript is the extension our multi-atlas methodology to the
-hippocampus using more than a single input atlas, while simultaneously limiting
-the number of possible inputs used during segmentation, and demonstrating that
-accurate identification of the hippocampal subfields is indeed possible using
-this methodology.
+derived from serial histological data is possible \citep{MallarChakravarty2012}.
+The novelty of this manuscript is the extension our multi-atlas methodology to
+the hippocampus using more than a single input atlas, while simultaneously
+limiting the number of possible inputs used during segmentation, and
+demonstrating that accurate identification of the hippocampal subfields is
+indeed possible using this methodology.
 
@@
 between very dissimilar images to be broken up into sequences of smaller
-deformations.  Although Wolz et al. begin with an atlas library of 30 MR images,
-this method could theoretically work using a much smaller atlas library.  In
-their validation, LEAP was used to segment the whole hippocampus in the ADNI1
+deformations.  Although \citet{Wolz2010} begin with an atlas library of 30 MR
+images, this method could theoretically work using a much smaller atlas library.
+In their validation, LEAP was used to segment the whole hippocampus in the ADNI1
 baseline dataset, achieving a mean Dice score of 0.85 with manual segmentations.
@@
 applicability of their work using data from 10 subjects.  In the second method,
-Yushkevich and colleagues \citep{Yushkevich2009} acquired and labelled
-hippocampal subfields on high-resolution MRI data from post-mortem medial
-temporal lobe samples and demonstrate accurate parcellations of hippocampal subfields in
-other MRI-volumes using nonlinear registration guided using manually derived
-hippocampus masks and specific landmarks. \todo{flow is terrible in this sentence...}
+\citet{Yushkevich2009} acquired and labelled hippocampal subfields on
+high-resolution MRI data from post-mortem medial temporal lobe samples. Using
+nonlinear registration guided by manually derived hippocampus masks and specific
+landmarks, they demonstrate accurate parcellation of hippocampal subfields in
+unlabelled MRI volumes.
 
@@
 input image (i.e. atlas or template) is registered to an unlabelled target image.
-Second, the labels from each image are propogated to the target image space. 
+Second, the labels from each image are propagated to the target image space. 
 Third, the labels are combined into a single labelling by way of a label fusion method 
@@
 \ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
-right whole hippocampi are available. These labels have been generated using the
-SNT tool from Medtronic Surgical Navigation Technologies, Louisville, CO (see
-Supplementary Materials for detailed discussion of the manual segmentation
-process used).
+right whole hippocampi are available \citep{Hsu2002}. These labels have been
+generated using the SNT tool from Medtronic Surgical Navigation Technologies,
+Louisville, CO (see Supplementary Materials for detailed discussion of the
+manual segmentation process used).
 
@@
 cross-validation, consists of repeated rounds of validation conducted on a fixed
-dataset \cite{Shao1993}. In each round, the dataset is randomly partitioned into
+dataset \citep{Shao1993}. In each round, the dataset is randomly partitioned into
 a training set and a validation set. The method to be validated is then given
@@
 
-A total of ten validation rounds are perfomed on each subject in the dataset,
+A total of ten validation rounds are performed on each subject in the dataset,
 over each combination of parameter settings. The parameter settings
@@
 the correlation coefficient.  For the purposes of this work we used the
-regularization parameters optimized in Robbins et al. \citep{Robbins2004},
-displayed in table \ref{tab:ANIMAL-params}.
+regularization parameters optimized in \citet{Robbins2004}, displayed in table
+\ref{tab:ANIMAL-params}.
 
@@
 The manual segmentations (SNT) provided as part of the ADNI dataset are used as
-the gold standard to compare with.  The segmentation accuracy reported is averaged 
-over the ten validation rounds for each parameter setting.
+the gold standard to compare with \citep{Hsu2002}.  The segmentation accuracy
+reported is averaged over the ten validation rounds for each parameter setting.
 
@@
 template images is varied. This is achieved by first computing the standard 
-deviation and variance of DSC scores in each block 
-of ten validation rounds per subject. The distribution of these statistics across 
-all subjects is then compared between \mb and multi-atlas using a Student's 
-t-test. A significant difference between distributions is taken to show either
-a larger or smaller level of variability between methods.
+deviation and variance of DSC scores in each block of ten validation rounds per
+subject. The distribution of these statistics across all subjects is then
+compared between \mb and multi-atlas using a Student's t-test. A significant
+difference between distributions is taken to show either a larger or smaller
+level of variability between methods.
 
@@
 
-In this experiment the Winterburn atlases are used as the cross-validation dataset. 
-The five $300\mu$-isotropic voxel images and labels are used as atlases, and 
-LOOCV is conducted using the five Winterburn atlas subject images 
-subsampled (using trilinear interpolation) to 0.9mm-isotropic voxel 
-resolution (referred to as the {\em Subsampled} dataset) as input subjects. Subsampling
-the subject images allows us to assess \mb in a typical segmentation scenario (high-resolution atlases
-and lower-resolution subjects).  The template library is composed of the subject
-images, plus an additional set of 3T T1 images (0.9mm-isotropic voxels) of healthy
-subjects acquired separately (Table \ref{table:WAval-healthy-demographics}). The 
-optimal size of template library, registration method, and label fusion method 
-found in Experiment 1 are used. Each resampled Winterburn atlas subject image is 
-segmented by \mb with that subject's image excluded from the atlas library.
+In this experiment the Winterburn atlases are used as the cross-validation
+dataset.  The five $300\mu$-isotropic voxel images and labels are used as
+atlases, and LOOCV is conducted using the five Winterburn atlas subject images
+subsampled (using trilinear interpolation) to 0.9mm-isotropic voxel resolution
+(referred to as the {\em Subsampled} dataset) as input subjects. Subsampling the
+subject images allows us to assess \mb in a typical segmentation scenario
+(high-resolution atlases and lower-resolution subjects).  The template library
+is composed of the subject images, plus an additional set of 3T T1 images
+(0.9mm-isotropic voxels) of healthy subjects acquired separately (Table
+\ref{table:WAval-healthy-demographics}). The optimal size of template library,
+registration method, and label fusion method found in Experiment 1 are used.
+Each resampled Winterburn atlas subject image is segmented by \mb with that
+subject's image excluded from the atlas library.
 
@@
 hippocampal volume with the full resolution Winterburn atlas 
-segmentations. In addition, by computing the relative error in volume of
-the Winterburn atlas labels resampled (with nearest-neighbour interpolation) to 0.9mm-isotropic voxels, we obtain a baseline error to assess against.
+segmentations. In addition, by computing the relative error in volume of the
+Winterburn atlas labels resampled (with nearest-neighbour interpolation) to
+0.9mm-isotropic voxels, we obtain a baseline error to assess against.
 
@@
 
-\paragraph{SZ-FEP dataset}
+\paragraph{\FEPdataset dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 neuroimaging study; only those suffering from schizophrenia spectrum disorders
-were considered for this analysis.  For complete program details see Malla et
-al. \citep{Malla2003}.
+were considered for this analysis.  For complete program details see
+\citet{Malla2003}.
 
@@
 \mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2).  All images from the SZ-FEP dataset are segmented by \mb.  The 
+(see Experiment 2).  All images from the \FEPdataset dataset are segmented by \mb.  The 
 optimal size of template library, registration method, and label fusion method 
@@
 \paragraph{Evaluation method}
-The manual segmentation protocol used to segment the Winterburn atlases has some
-neuroanatomical differences in comparison to the protocol used to segment the
-SZFEP dataset. Therefore, rather than evaluate using an overlap metric, such as
-DSC, which assumes a fixed segmentation protocol, we instead correlate whole
-hippocampus volume between \mb and manual segmentations.
-
-\todo{take description from below}
+The manual and Winterburn hippocampal segmentation protocols differ slightly in
+the neuroanatomical features that are delineated.  This poses a problem for
+evaluation by measuring overlap.  That is, since different protocols will
+necessarily produce segmentations that do not perfectly overlap, the degree of
+overlap cannot be solely used to compare segmentation methods using different
+protocols. In place of an overlap metric, we can assess the degree of correlation in
+average hippocampal volume of the subjects produced by each method.
+Specifically, Pearson correlation is used.
 
@@
 \paragraph{Evaluation method}
-Since the hippocampal segmentation protocols differ between the ADNI labels and
-Winterburn atlases, this poses a problem for evaluation of by measuring overlap. 
-That is, different segmentation protocols will necessarily produce segmentations
-that do not perfectly overlap, and so the degree of overlap cannot be used soley 
-to assess two segmentaiton methods that each use different protocols.
-
-Instead, we assess the correlation in hippocampal volume between \mb and manual segmentation
-(SNT). For reference, we also compute the correlation in hippocampal volume between
-the existing, established automated segmentation methods: FSL, FreeSurfer, and 
-MAPER.  Additionally, we evaluate the volume-related fixed and proportional 
-biases in all segmentation methods using Bland-Altman plots \todo{Bland-Altman ref}.
-
-\todo{references} with manual segmentations. 
-\todo{"hippocampal volume" = mean hemispheric volume?}
+As in Experiment 3, the manual (SNT) and Winterburn hippocampal segmentation
+protocols differ in the neuroanatomical features delineated, and so we must
+assess \mb by the degree of correlation of average hippocampal volume across all
+subjects produced by \mb and by manual segmentation.  Specifically, Pearson
+correlation is used.  For comparison, we also compute the correlation in
+hippocampal volume between the existing, established automated segmentation
+methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
+volume-related fixed and proportional biases in all segmentation methods using
+Bland-Altman plots \citep{Bland1986}.
 
@@
 % ADNI1-COMPLETE
-% Prepares the data in a the form needed for plotting.
+% Prepares the data in the form needed for plotting.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 label fusion method used does not significantly improve label accuracy, contrary 
-to the findings of Aljabar et al. \cite{Aljabar2009} when using weighted voting
-on much larger altas/template libraries. In the remainder of  this section, only 
-results using the \ants registration algorithm and majority vote fusion will 
-be shown.
+to the findings of \citet{Aljabar2009} when using weighted voting on much larger
+atlas/template libraries. In the remainder of  this section, only results using
+the \ants registration algorithm and majority vote fusion will be shown.
 
@@
 
-\todo{mention failures}
-
 In addition to an improvement in accuracy over multi-atlas-based segmentation,
@@
 (Figure \ref{ADNI1-xval-variability}).  The size of template library 
-necessary to reach a significant (p<0.5) decrease in variance and standard 
+necessary to reach a significant ($p < 0.05$) decrease in variance and standard 
 deviation grows with the size of atlas library used.  A template library of 19
@@
 deviation for 3-7 atlases.
-\todo{Tris: only show variance OR standard deviation because otherwise the story gets complicated... why compare measures?}
-
-It is interesting to note that with an even number of templates, \mb shows a
-small decrease in performance relative to when one fewer template image is used.
-See section \todo{ref} for a discussion of this behaviour. In the remainder of
-this section, only results from odd-sized template libraries will be shown.
+
+We have omitted results obtained when using an even number of atlases or
+templates since with this configuration we found significantly decreased
+performance. We believe this is as a result of an inherent bias in the majority
+vote fusion method used (see Discussion).
 
@@
 
-\begin{figure}
-<<ADN1-library-composition, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-# These composition datasheets describe the subjects and subject DX in the 
-# template and atlas libraries for each voting run
-atlas_comp = read.csv('data/a2a_ants/atlas_library_composition.csv')
-templ_comp = read.csv('data/a2a_ants/template_library_composition.csv')
-a2a.best = subset(all_data, atlases == 9 &
-                    templates.mb == 19 &
-                    reg_method=='ANTS' &
-                    method.mb =='Majority Vote')
-
-cumulative_dist = 
-  ggplot(a2a.best, aes(x=k.mb)) + stat_ecdf() + scale_y_continuous(breaks=seq(0,1,0.1))
-
-a2a.best = subset(a2a.best, k.mb > 0.75) # drop the guys that suck
-
-a2a.best = merge(a2a.best, atlas_comp, by=c("timestamp","atlases"))
-a2a.best = merge(a2a.best, templ_comp, by.x=c("timestamp","templates.mb"), by.y=c("timestamp","templates"))
-a2a.best$templates_AD_percent = (a2a.best$templates_AD / 19) * 100
-a2a.best$templates_MCI_percent = (a2a.best$templates_MCI / 19) * 100
-a2a.best.melted = melt(a2a.best, measure.vars=c("templates_AD_percent", "templates_MCI_percent"))
-
-k_by_composition = 
-  ggplot(a2a.best.melted, aes(x=value, y=k.mb, colour=DX)) + facet_grid(variable ~ DX) + 
-  geom_boxplot(aes(group=as.factor(value), colour=DX)) +  geom_smooth(method="lm")
-
-snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
-a2a.best = merge(a2a.best, snt.unilateral.vols)
-
-# bland-altman plots
-bland_mb_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.mb, DX) + 
-  geom_smooth(method="lm") + 
-  scale_y_continuous(breaks=seq(-500, 500, 50)))
-
-bland_ma_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.ma, DX) + 
-  geom_smooth(method="lm") + 
-  scale_y_continuous(breaks=seq(-500, 500, 50)))
-
-# just the trials with a high percentage of MCI in the template library
-a2a.best.highMCI = subset(a2a.best, templates_MCI_percent > 50)
-
-bland_mb_highMCI = with(a2a.best.highMCI, bland_altman_plot(volume.snt, volume.ma, DX) + 
-  geom_smooth(method="lm") + 
-  scale_y_continuous(breaks=seq(-500, 500, 50)))
-
-@
-\end{figure}
+% \begin{figure}
+% <<ADN1-library-composition, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+% # These composition datasheets describe the subjects and subject DX in the 
+% # template and atlas libraries for each voting run
+% atlas_comp = read.csv('data/a2a_ants/atlas_library_composition.csv')
+% templ_comp = read.csv('data/a2a_ants/template_library_composition.csv')
+% a2a.best = subset(all_data, atlases == 9 &
+%                     templates.mb == 19 &
+%                     reg_method=='ANTS' &
+%                     method.mb =='Majority Vote')
+% 
+% cumulative_dist = 
+%   ggplot(a2a.best, aes(x=k.mb)) + stat_ecdf() + scale_y_continuous(breaks=seq(0,1,0.1))
+% 
+% a2a.best = subset(a2a.best, k.mb > 0.75) # drop the guys that suck
+% 
+% a2a.best = merge(a2a.best, atlas_comp, by=c("timestamp","atlases"))
+% a2a.best = merge(a2a.best, templ_comp, by.x=c("timestamp","templates.mb"), by.y=c("timestamp","templates"))
+% a2a.best$templates_AD_percent = (a2a.best$templates_AD / 19) * 100
+% a2a.best$templates_MCI_percent = (a2a.best$templates_MCI / 19) * 100
+% a2a.best.melted = melt(a2a.best, measure.vars=c("templates_AD_percent", "templates_MCI_percent"))
+% 
+% k_by_composition = 
+%   ggplot(a2a.best.melted, aes(x=value, y=k.mb, colour=DX)) + facet_grid(variable ~ DX) + 
+%   geom_boxplot(aes(group=as.factor(value), colour=DX)) +  geom_smooth(method="lm") 
+% 
+% snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
+% a2a.best = merge(a2a.best, snt.unilateral.vols)
+% 
+% # bland-altman plots
+% bland_mb_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.mb, DX) + 
+%   geom_smooth(method="lm") + 
+%   scale_y_continuous(breaks=seq(-500, 500, 50)))
+% 
+% bland_ma_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.ma, DX) + 
+%   geom_smooth(method="lm") + 
+%   scale_y_continuous(breaks=seq(-500, 500, 50)))
+% 
+% # just the trials with a high percentage of MCI in the template library
+% a2a.best.highMCI = subset(a2a.best, templates_MCI_percent > 50)
+% 
+% bland_mb_highMCI = with(a2a.best.highMCI, bland_altman_plot(volume.snt, volume.ma, DX) + 
+%   geom_smooth(method="lm") + 
+%   scale_y_continuous(breaks=seq(-500, 500, 50)))
+% 
+%   
+% @
+% \end{figure}
 
@@
 \begin{figure}
-<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=7, fig.height=7>>=
+stderr <- function(x) sqrt(var(x)/length(x))
 ggplot(subset(all_data_mean, (templates.mb*atlases) %% 2 == 1),
@@
                aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
-  stat_summary(fun.y=mean,geom='point',
-               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+  stat_summary(fun.y=mean,
+               fun.ymin=function (x) mean(x) - stderr(x),
+               fun.ymax=function (x) mean(x) + stderr(x), 
+               geom='errorbar',
+               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
   #stat_smooth(method='lm', formula=y~log(x)) + 
-  scale_colour_hue(name="Atlases") +
+  scale_colour_hue(name="Number of Atlases") +
   xlab( "Number of Templates" ) + 
   ylab( "Mean DSC" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text = element_text(size = 8))
 @
   \label{fig:ADNI1-xval-k_mean}
-  \caption{
-  {\bf Mean DSC of \mb segmentations across 69 ADNI1 subjects, by atlas and
-  template library size, registration algorithm, and label fusion method.} 
-  }
+  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with manual
+  segmentations (SNT) for 69 ADNI1 subjects vs. atlas and template library size,
+  registration algorithm, and label fusion method. Points above zero indicate an
+  \mb parameter settings yielding a higher mean DSC score than multi-atlas
+  segmentation using the same number of atlases.}
 \end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}
-
-<<ADN1-xval-k-mean-mb.ma, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ss = subset(all_data_mean, reg_method == 'ANTS' & method.mb == 'Majority Vote')
-ss = subset(ss, (templates.mb*atlases) %% 2 == 1)
-ma_lines = ddply(ss, c("atlases"), function (df) {data.frame(yint = mean(df$k.ma))})
-
-ggplot(ss,
-       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
-  facet_grid(reg_method~method.mb) + 
-  stat_summary(fun.y=mean,geom='line',
-               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
-  stat_summary(fun.y=mean,geom='point',
-               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
-  geom_hline(yintercept=ma_lines$yint, colour = c("red", "orange", "green", "cyan", "purple"), linetype="dashed") + 
-  #stat_smooth(method='lm', formula=y~log(x)) + 
-  scale_colour_hue(name="Atlases") +
-  xlab( "Number of Templates" ) + 
-  ylab( "Mean DSC" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
-  \label{fig:ADNI1-xval-k_mean}
-  \caption{
-  {\bf Mean DSC of \mb segmentations across 69 ADNI1 subjects, by atlas and
-  template library size, registration algorithm, and label fusion method.} 
-  }
-\end{figure}
 
@@
 \begin{figure}
-<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=6,fig.height=4>>=
+<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep'>>=
 ggplot(subset(all_data_mean, reg_method=="ANTS" & 
@@
   scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
-  scale_colour_hue(name="Atlases") +
+  scale_colour_hue(name="Number of Atlases") +
   scale_alpha_continuous(guide = "none") + 
@@
 @
-  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
-  a range of parameter settings.}
+  \caption{Difference in mean Dice's Similarity Coefficient of \mb and
+  multi-atlas segmentations with manual (SNT) segmentations for a range of atlas
+  and template library sizes.}
   \label{ADNI1-xval-k-diff}
@@
 subregions show distinctly larger error than is found through resampling.
-\todo{note direction of change.}
-
-Figure \ref{fig:subfield-montage} shows a qualitative comparision of \mb subfield
+
+Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb subfield
 segmentation.
@@
   \caption{Sagittal slices from two subjects showing a comparison of the
-  original Winterburn atlas subfield segmentations (at 0.3mm3-isotropic voxel
-  resolution), the subsampled Winterburn segmentations (at 0.9mm3-isotropic
+  original Winterburn atlas subfield segmentations (at 0.3mm-isotropic voxel
+  resolution), the subsampled Winterburn segmentations (at 0.9mm-isotropic
   voxel resolution), and the \mb labels on the subsampled atlas image.}
@@
 
-\todo{MAGeT on BRAVO.  Also, show coronal views}
+\todo[inline]{Figure \ref{fig:subfield-montage}. Show MAGeT segmentations on
+BRAVO images.  Also, show coronal/transverse slices and MNI coordinates}
 
@@
 In this experiment \mb is applied to a dataset of images of first episode
-schizophrenia parients, using the Winterburn atlases and a template library of
+schizophrenia patients, using the Winterburn atlases and a template library of
 21 subject images selected at random.  Expert manual whole hippocampal
 segmentations are used as gold standards.
-
-\mb produces hippocampus segmentation volumes that are highly correlated with 
-manual segmentation volumes (Figure \ref{fig:SZ-volumes}). 
-\todo{r and p-values}
 
@@
 @
-  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
-  HC volume for \mb against manually rated Hippocampal volumes}
+  \caption{Subject mean hippocampal volumes by \mb vs. manually segmented
+   volumes from the First Episode Patients with Schizophrenia dataset. A linear fit
+   and standard error are shown.}
   \label{fig:SZ-volumes}
 \end{figure}
+
+<<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
+  fep.cor.test = cor.test(casted$Manual, casted$MAGeT)
+@
+\mb produces hippocampus segmentation volumes that are highly correlated with 
+manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
+$t = \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001,
+fep.cor.test$p.value)}$; Figure \ref{fig:SZ-volumes}).
 
@@
 Based on the results from the ADNI1 Cross-Validation experiment, in this
-experiment \mb was configured with a template library of 21 randomly
-chosen subject images (7 from each disease class) and used majority vote label
-fusion. The entire \adnidataset dataset was segmented by \mb and we 
-now compare the resulting volumes with those obtained by manual segmentation (SNT), 
-and other automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
-\ref{tab:ADN1-1.5T-Complete-1yr-Dataset-Segmentations} shows the total count of 
+experiment \mb was configured with a template library of 21 randomly chosen
+subject images (7 from each disease class) and used majority vote label fusion.
+The entire \adnidataset dataset was segmented by \mb and we now compare the
+resulting volumes with those obtained by manual segmentation (SNT), and other
+automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
 segmentations available, including a count of those which have failed a quality
@@
 are included in the following analysis (a total of
-\Sexpr{length(means.complete$RID)} images; Table 
-\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals}).
+\Sexpr{length(means.complete$RID)} images).
 
@@
 all methods and manually segmented volumes (Figure \ref{fig:ADNI-volumes-plot}). 
-Volumes are correlated with Pearson-r > 0.78 for all methods across disease categories.
-Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}),
-\mb is consistently well correlated to manual volumes (Pearson-r > 0.85), but appears to 
-slightly over-estimate the volume of the AD hippocampus.
-
-To investigate the level of agreement with manually segmented hippocampal volumes, 
-we constructed Bland-Altman plots for each method (Figure \ref{fig:ADNI1-Bland-Altman}).  
-As Bland \& Altman, 1985 \cite{Bland1986} noted, high 
-correlation amongst measures of the same quantity does not necessarily 
-imply agreement (as correlation can be driven by a large range in true values, 
-for instance).  What is most striking in Figure \ref{fig:ADNI1-Bland-Altman}
-is that all methods show an obvious proportional bias: FreeSurfer and FSL 
-markedly under-estimate smaller hippocampi and over-estimate large hippocampi, 
-whereas MAPER and \mb more conservatively show the reverse bias.  Additionally, 
-all methods show a fixed bias, with FreeSurfer and FSL most dramatically 
-over-estimating hippocampal volume by $2600mm^3$ and $2800mm^3$ on average, respectively, 
-and MAPER and \mb within $250mm^3$ on average.
+Volumes are correlated with Pearson $r > 0.78$ for all methods across disease
+categories.  Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}),
+\mb is consistently well correlated to manual volumes (Pearson $r > 0.85$), but
+appears to slightly over-estimate the volume of the AD hippocampus.
+
+To investigate the level of agreement with manually segmented hippocampal
+volumes, we constructed Bland-Altman plots for each method (Figure
+\ref{fig:ADNI-Bland-Altman}).  As \citet{Bland1986} noted, high correlation
+amongst measures of the same quantity does not necessarily imply agreement (as
+correlation can be driven by a large range in true values, for instance).  What
+is most striking in Figure \ref{fig:ADNI-Bland-Altman} is that all methods show
+an obvious proportional bias: FreeSurfer and FSL markedly under-estimate smaller
+hippocampi and over-estimate large hippocampi, whereas MAPER and \mb more
+conservatively show the reverse bias.  Additionally, all methods show a fixed
+bias, with FreeSurfer and FSL most dramatically over-estimating hippocampal
+volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and MAPER and
+\mb within $250 mm^3$ on average.
 
@@
       xlab("Diagnosis") + 
-      ylab(expression(paste("Mean hemispheric hippocampal volume (", mm^3, ")"))) + 
-      theme(legend.direction = "horizontal", legend.position = "bottom", text=element_text(size=18))
+      ylab(expression(paste("Hippocampal volume (", mm^3, ")"))) + 
+      theme(legend.direction = "horizontal", legend.position = "bottom")
 @
-  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
-  MAPER, \mb (MAGeT) and manual (SNT) by disease category.  Pearson correlation 
-  with manual volumes are shown below each box-and-whisker.}
+  \caption{Subject mean hippocampal volume as measured in the \adnidataset
+dataset by FreeSurfer, FSL, MAPER, \mb, and manual raters (SNT) vs. disease
+category.}
   \label{fig:ADNI-volumes-boxplot}
@@
 ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
-    geom_point() + 
+    geom_point(size=1) + 
     geom_smooth(method="lm") + 
-    xlab(expression("Manual mean hemispheric hippocampal volume " (mm^3))) + 
-    ylab(expression("Automated mean hemispheric hippocampal volume " (mm^3))) +
-#    annotate("text", y = c(3000,2500,2000,1500), x=Inf-100, size=3,
-#              label = c(paste("FS r=", round(cor(df$SNT, df$FS),2)),
-#                        paste("FSL r=", round(cor(df$SNT, df$FSL),2)),
-#                        paste("MAPER r=", round(cor(df$SNT, df$MAPER),2)),
-#                        paste("MAGeT r=", round(cor(df$SNT, df$MAGeT),2))),
-#              hjust=1.1,vjust=1.5) +
+    xlab(expression("Manual hippocampal volume " (mm^3))) + 
+    ylab(expression("Automated hippocampal volume " (mm^3))) +
+    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
+              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
+                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
+                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
+                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
+              hjust=1.1,vjust=1.5) +
     theme(legend.direction = "horizontal", legend.position = "bottom")
 @
-  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
-  MAPER, \mb (MAGeT) and manual (SNT). Pearson correlation with manual volumes
-  are shown for each method.}
+  \caption{Subject mean hippocampal volume as measured in the \adnidataset
+dataset by each of the four automated methods investigated (FreeSurfer (FS),
+FSL, MAPER, \mb) vs. manual rating (SNT). Linear fit lines and pearson
+correlations are shown for each method.}
   \label{fig:ADNI-volumes-plot}
@@
 \begin{figure}
-<<ADN1-Bland-Atlman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=6, fig.height=6>>=
+<<ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
 names(means.complete)[8] = "FreeSurfer"
@@
 melted = subset(melted, (method == "FreeSurfer") | 
-                        (method == "FSL"   & diff > -5000 & diff < 900) | 
-                        (method == "MAPER" & diff > -500 & diff < 1000) | 
-                        (method == "MAGeT" & diff > -1100))
+                        (method == "FSL"  ) | # & diff > -5000 & diff < 900) | 
+                        (method == "MAPER") | # & diff > -500 & diff < 1000) | 
+                        (method == "MAGeT"))  # & diff > -1100))
 #melted = subset(melted, method == "MAGeT")
@@
 ggplot(melted, aes(x= mean, y = diff, colour=Diagnosis)) + 
-  facet_wrap( ~ method, nrow=2, scales="free") + 
-  geom_point() +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown", data=limits) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits, hjust=1.1,vjust=-0.8, size=3) + 
-  xlab(expression('Mean of manual and automated volume' (mm^3))) + 
-  ylab(expression('Difference in manual and automated volumes' (mm^3))) + 
+  facet_wrap( ~ method, nrow=2, scales="fixed") + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=1) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits,
+            hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of manual and automated hippocampal volume ' (mm^3))) + 
+  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
   scale_x_continuous(breaks=seq(0,10000, 1000)) + 
-  geom_smooth(method="lm", formula = y~x) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
 @
-  \caption{Bland-Altman plot comparing manual (SNT) and automated total hippocampal volumes 
-  in the \adnidataset dataset.  The overall mean difference, and limits of 
-  agreement ($\pm 1.96SD$) are shown by dashed horizontal lines.}
+  \caption{Bland-Altman plots comparing subject mean hippocampal volume as
+measured in the \adnidataset dataset by manual raters (SNT) and each of the four
+automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall mean
+difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
+horizontal lines. Linear fit lines are shown for each diagnosis group. Note,
+points below the mean difference indicate overestimation of the volume with
+respect to the manual rating, and vice versa. }
+
   \label{fig:ADNI-Bland-Altman}
@@
   anatomical image, SNT manual label (green), and \mb label (blue). Noted are
-  examples of common segmentation idiosyncracies: 
+  examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
@@
 \end{figure}
+
+\todo[inline]{Figure \ref{fig:ADNI-subfields}. Show coronal/transverse slices
+and MNI coordinates}
 
@@
 This suggests that using an intermediate template library generated by accurate
-nonlinear registration methods such as \ants \cite{Avants2008} sufficiently filters
+nonlinear registration methods such as \ants \citep{Avants2008} sufficiently filters
 sources of error such as nonlinear registration error and partial volume effects
@@
 work on the \mb algorithm we have shown that these increases in error
-are not simply a smoothing or averaging effect \cite{MallarChakravarty2012}.
+are not simply a smoothing or averaging effect \citep{MallarChakravarty2012}.
 
@@
 available.  For example, other groups have used the STAPLE algorithm
-\cite{Warfield2004} (or variants of the STAPLE algorithm \cite{Duchesne2012})
+\citep{Warfield2004} (or variants of the STAPLE algorithm \citep{Duchesne2012})
 which weights each segmentation based upon its estimated
@@
 area where the voting scheme could be used to improve performance. However, it
-is worth noting that this limitation was previously identified by Heckemann et
-al. \cite{Heckemann2006a} and, subsequently, other groups have not even
-considered the potential pitfalls of an even number of candidate labels (eg.
-\cite{Leung2010}).
+is worth noting that this limitation was previously identified by
+\citet{Heckemann2006a} and, subsequently, other groups have not even considered
+the potential pitfalls of an even number of candidate labels (eg.
+\citet{Leung2010}).
 
@@
 multi-atlas/segmentation literature.  To the best of our knowledge other groups
-have only ever performed a 3-fold validation \cite{?}.  This suggests that
+have only ever performed a 3-fold validation \citep{?}.  This suggests that
 our results are reflective of a true average over the choice of parameter
@@
 
-\todo{additionally, we could be seeing a ceiling effect: reaching the limit of
-what is achievable with a multi-atlas-based segmentation method}
+\todo[inline]{could we be seeing a ceiling effect... reaching the limit of what
+is achievable with a multi-atlas-based segmentation method}
 
@@
 summarized in Table 6 require more atlases (between 16-55) than our \mb
-implementation with the Winterburn atlases \cite{Winterburn2013}.  Amongst these
+implementation with the Winterburn atlases \citep{Winterburn2013}.  Amongst these
 methods only method 4 yields mean Kappas that are higher than the ones that are
@@
 and these specific methods.  Others have reported the difficulty with
-misregistrations in candidate segmentation (i.e. segmentations generated that are
-then input in the voxel-voting procedure \cite{Collins2010}). The work of Leung
-\cite{Leung2010} (work also reports some of the highest kappas for the
-segmentation of ADNI data) tackles this problems by using an intensity threshold
-that is estimated heuristically at the time of segmentation.  While this works
-for the ADNI dataset (which is partially homogenized with respect to image
-acquisition and pre-processing), it is unclear if this type of heuristic is
-applicable to other datasets.  In all cases, these methods require more atlases
-than our implementation with the Winterburn atlases.  Lotjonen et al.
-\cite{Lotjonen2010} achieve highly accurate segmentation but correct their
+mis-registrations in candidate segmentation (i.e. segmentations generated that are
+then input in the voxel-voting procedure \citep{Collins2010}). The work of 
+\citet{Leung2010} tackles this problem by using an intensity threshold
+that is estimated heuristically at the time of segmentation (this work also
+reports some of the highest kappas for the segmentation of ADNI data).  While
+this works for the ADNI dataset (which is partially homogenized with respect to
+image acquisition and pre-processing), it is unclear if this type of heuristic
+is applicable to other datasets.  In all cases, these methods require more
+atlases than our implementation with the Winterburn atlases.
+\citet{Lotjonen2010} achieve highly accurate segmentation but correct their
 segmentations using classifications derived using an expectation maximization
-framework.  In their initial work, Chupin and colleagues develop their
+framework.  In their initial work, \citet{Chupin2009} develop their
 probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy
@@
 increased computational cost ($O(log(n))$), as previously mentioned in
-other work \cite{Heckemann2006}). \todo{check this: time complexity seems suspicious}
+other work \citep{Heckemann2006}. 
+\todo[inline]{time complexity seems suspicious}
 
 In comparison to the methods that we compared and were available through the
-ADNI database (FreeSufer, MAPER) and the method that we initialized ourselves
+ADNI database (FreeSurfer, MAPER) and the method that we initialized ourselves
 (FSL-FIRST) we find extremely variable performance of all methods.  With the
 exception of FSL all methods correlate well with the SNT volumes provided in
-the ADNI database.  However, FreeSufer and FIRST provide radically different
+the ADNI database.  However, FreeSurfer and FIRST provide radically different
 definitions of the size of the hippocampus in comparison to the other methods.
@@
 although very few methods are available for the segmentation of 3T data
-\cite{Yushkevich2009,VanLeemput2009}. However, the FreeSufer implementation of
+\citep{Yushkevich2009,VanLeemput2009}. However, the FreeSurfer implementation of
 subfield segmentation. While recent work demonstrates that subfield
 segmentations can be used for classification of AD, MCI, and NC, there has been
-no explicit validation of the methodology based on accruacy or precision.
+no explicit validation of the methodology based on accuracy or precision.
 Although the initial hippocampal subfield segmentations from the Yushkevich
@@
 segmentation methodologies do not attempt to parse these regions either
-\cite{Wisse2012, Mueller2009}
-
+\citep{Wisse2012, Mueller2009}
+
+\subsection{Conclusion}
 In conclusion, we have presented a flexible multi-atlas framework.  It has
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{abbrvnat}
 \bibliography{references}

### 2013-07-29 paper.Rnw
Added contact block

@@
 
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  require(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-
-bland_altman_plot <- function(x, y, c) {
-  data = data.frame(x=x,y=y,diff=x-y, m = (x+y)/2, c=c)
-  return(ggplot(data, aes(x= m, y = diff, colour = as.factor(c))) + 
-    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
-               linetype=2, color='brown') + geom_point())
-}
-
 lm_eqn = function(m) {
@@
 \newcommand{\subsubsubsection}[1]{\paragraph{#1}}
+\renewcommand\Affilfont{\itshape\small}  %authblk
 
@@
 \author[1,5,6]{M. Mallar Chakravarty}
-\author{Alzheimer's Disease Neuroimaging Initiative}
+\author[ ]{Alzheimer's Disease Neuroimaging Initiative}
 \affil[1]{Kimel Family Translational Imaging-Genetics Lab, Centre for Addiction and
@@
 
+\date{}
+
 \begin{document}
@@
 %while still achieving similar accuracy to multi-atlas approaches.
-
 \end{abstract}
+
+\parbox{4in}{
+\textbf{Contact:} \\
+Jon Pipitone and M. Mallar Chakravarty \\
+Kimel Family Translation Imaging-Genetics Research Laboratory \\
+Research Imaging Centre \\
+Centre for Addiction and Mental Health \\
+250 College St. \\
+Toronto, Canada   M5T 1R8 \\
+jon.pipitone@camh.ca; mallar.chakravarty@camh.ca}
 

### 2013-08-01 paper.Rnw
Figure/table references fixed up, fixed totals/QC table

@@
 \usepackage{authblk}            % Listing Author affiliations
+\usepackage[hypcap]{caption}
 
@@
 transformation is applied to the manual labels (ie. {\em label propagation}) to
-apply them to the target image.  While this methodology has been used
+move them into the target image space.  While this methodology has been used
 successfully in several contexts
@@
 accuracy of \mb whole hippocampus segmentation over a wide range of parameter
-settings. This enables us to choose the parameter settings offering the best
-performance for use in subsequent experiments. Experiment 2 tests hippocampal
-subfield segmentation quality. The last two experiments assess the validity of
-the \mb algorithm when applied to different diseases: Alzheimer's disease
-(Experiment 3) and first episode schizophrenia patients (Experiment 4).
-
-%The ADNI experiment includes comparison of volumes with other automated
-%segmentation methods. Evaluation is done by correlating volumes b/c overlap
-%metrics assume identical segmentation protocol.
+settings. The results of this experiment enable us to choose the parameter
+settings offering the best performance for use in subsequent experiments.
+Experiment 2 tests hippocampal subfield segmentation quality. The last two
+experiments assess the validity of the \mb algorithm when applied to different
+diseases: first episode schizophrenia (Experiment 3), and Alzheimer's disease
+(Experiment 4). Additionally, in experiment 4, we compare \mb segmentations with
+those of well-known automated and manual methods and assess segmentation bias.
 
@@
                  1217,29,56,1188,1293,149,382,431,1202)
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
                adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-
+caption = paste(
+  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
+  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
+  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption="ADNI-1 cross-validation subset demographics", 
+      caption=caption, caption.loc = c("top"),
       title="tab:ADNI1-xvalidation-demographics")
@@
 the correlation coefficient.  For the purposes of this work we used the
-regularization parameters optimized in \citet{Robbins2004}, displayed in table
+regularization parameters optimized in \citet{Robbins2004}, displayed in Table
 \ref{tab:ANIMAL-params}.
@@
 %                                     
-\begin{table}[!tbp]
+\begin{table*}[!tbp]
+\scriptsize
+\caption{\textbf{\animal registration parameters.} \label{tab:ANIMAL-params}}
 \begin{center}
-\caption{\animal  registration parameters}
 \begin{tabular}{l | c c c}
@@
 \end{center}
-\label{tab:ANIMAL-params}
-\end{table}
+\end{table*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 (0.9mm-isotropic voxels) of healthy subjects acquired separately (Table
-\ref{table:WAval-healthy-demographics}). The optimal size of template library,
+\ref{tab:WAval-healthy-demographics}). The optimal size of template library,
 registration method, and label fusion method found in Experiment 1 are used.
@@
 subject's image excluded from the atlas library.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Subfield Validation Template library demographics (healthy controls)
+%                                     
+<<WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
+master = read.csv('data/DT_study.csv')
+tab <- summary(~Age + Sex + Education + Handedness, 
+               data = master, method="reverse", na.rm = TRUE, test=FALSE)
+caption = paste(
+  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
+  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
+  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      title="tab:ADNI1-xvalidation-demographics")
+@
+\begin{table*}[!tbp]
+\scriptsize
+\caption{\textbf{Subfield validation template library demographics.}
+\label{tab:WAval-healthy-demographics}}
+\begin{center}
+\begin{tabular}{l | c c c}
+\hline 
+Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
+\hline 
+Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
+Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
+Iterations        &   30    &    30   &   10     \tabularnewline
+Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
+Sub-Lattice       &   6     &    6    &   6      \tabularnewline
+Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 field-of-view for the images was 256mm (SI)$\times$204mm (AP).  Subject
-demographics are shown in table \ref{tab:SZFEP-Demographics}.
+demographics are shown in Table \ref{tab:SZFEP-Demographics}.
 
@@
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
+caption = paste(
+  "\\textbf{Schizophrenia First Episode Patient Demographics.}",
+  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
+  "State Examination (MMSE).", sep=" ") 
 fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
@@
                data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", ctable=FALSE,
-      caption="Schizophrenia First Episode Patient Demographics",
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
       title="tab:SZFEP-Demographics")
@@
 All images from the {\em \adnidataset} standardized dataset described
-in Experiment 1 are used.  Clinical and demographic data are shown in table
-\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Demographics}.
+in Experiment 1 are used.  Clinical and demographic data are shown in Table
+\ref{tab:ADNI1-Dataset-Demographics}.
 
@@
 %                                                                             
-<<ADNI-demographics, echo=F, results="asis",cache=FALSE>>=
+<<ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
 # RID < 2000 ensures only ADNI1 patients
@@
 adnimerge.yr1 = merge(adnimerge, yr1) 
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + PTETHCAT + CDRSB + ADAS13 +
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
                MMSE, data = adnimerge.yr1, 
@@
                method = "reverse", test=FALSE, overall=TRUE)
-
-latex(tab, file="", size="scriptsize", 
-    caption="ADNI1 1.5T Complete 1Yr dataset demographics",
-    label="tab:ADN1-1.5T-Complete-1Yr-Dataset-Demographics")
+caption = paste(
+  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.}", 
+  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
+  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+    caption=caption, caption.loc = c("top"),
+    title="tab:ADNI1-Dataset-Demographics",
+    label="tab:ADNI1-Dataset-Demographics")
 @
@@
 accuracy over multi-atlas-based segmentation when using the same number of atlases
-and voting method (Figure \ref{fig:ADNI1-xva-k-diff}). The magnitude of improvement over
+and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The magnitude of improvement over
 multi-atlas-based segmentation decreases with an increasing number of atlases,
@@
 \begin{figure}
-<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=7, fig.height=7>>=
+<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-prep', fig.width=7, fig.height=7>>=
 stderr <- function(x) sqrt(var(x)/length(x))
@@
 @
-  \label{fig:ADNI1-xval-k_mean}
   \caption{Mean Dice's Similarity Coefficient of \mb segmentations with manual
@@
   \mb parameter settings yielding a higher mean DSC score than multi-atlas
-  segmentation using the same number of atlases.}
+  segmentation using the same number of atlases. \label{fig:ADNI1-xval-k-mean}}
 \end{figure}
@@
 \begin{figure}
-<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep'>>=
+<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-prep'>>=
 ggplot(subset(all_data_mean, reg_method=="ANTS" & 
@@
   multi-atlas segmentations with manual (SNT) segmentations for a range of atlas
-  and template library sizes.}
-  \label{ADNI1-xval-k-diff}
+  and template library sizes. \label{fig:ADNI1-xval-k-diff}}
 \end{figure}
@@
 segmentations available, including a count of those which have failed a quality
-control inspection. Only those images which had segmentations from every method
+control inspection. Quality control for FreeSurfer segmentations was provided along
+with the segmentations themselves from ADNI \cite{Hartig2010}. One author (MP) 
+inspected
+
+Only those images which had segmentations from every method
 are included in the following analysis (a total of
@@
 <<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
-caption = paste(
-  "Pass/fail quality control indicators were supplied with the FreeSurfer",
-  "volumes downloaded from the ADNI website (we used the temporal lobe quality",
-  "control indicator, TEMPQC). One of the authors (MP) performed visual",
-  "quality inspection for MAGeT and FSL segmentations.", sep=" ")
-latex(package_totals, file="", size="scriptsize", 
-    caption = caption,
-    title = "",
+package_totals$Total <- NULL
+latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
+    caption="\\adnidataset dataset segmentation totals and QC failures.", 
+    caption.loc = c("top"), rowname=NULL, 
     label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@@
 available.  For example, other groups have used the STAPLE algorithm
-\citep{Warfield2004} (or variants of the STAPLE algorithm \citep{Duchesne2012})
+\citep{Warfield2004} (or variants of the STAPLE algorithm \citep{Robitaille2012})
 which weights each segmentation based upon its estimated
@@
 atlas or template choice. This is an important consideration that few have
-touched on previously. The 10-fold Monte Carlo cross validation that we present
+touched on previously. The 10-fold Monte Carlo leave-one-out cross-validation that we present
 in Experiment 1 is amongst one of the most stringent performed in the
 multi-atlas/segmentation literature.  To the best of our knowledge other groups
-have only ever performed a 3-fold validation \citep{?}.  This suggests that
+have only ever performed cross-validation \citep{?}.  This suggests that
 our results are reflective of a true average over the choice of parameter
@@
 atlases are properly segmented).
+
+% Validation in other works
+% Chupin 2009: LOOCV, 1-fold validation
+% Heckemann 2011: QC of segmentations, comparison across field strengths, 
+%                 agreement between two different atlas sets
+% Coupe 2012: LOOCV
+% Lotjonen 2010: LOOCV, 1-fold validation
+% Leung 2010: 1-fold, QC, classification
+% Wolz 2010: 1-fold over different leap settings, classification
+% Collins 2010: 1-fold validation, QC
+%
 

### 2013-08-03 paper.Rnw
Existing work table updated, last few fixes.
In particular, I added a blurb about how MP did QC. Also, added in a short
comparison to other existing work (referencing the table) and referenced
the table when justifying our low overlap result

@@
 images using nonlinear registration methods.  The resulting nonlinear
-transformation is applied to the manual labels (ie. {\em label propagation}) to
+transformation is applied to the manual labels (i.e. {\em label propagation}) to
 move them into the target image space.  While this methodology has been used
@@
 More recently, many groups have been using multiple atlases to improve overall
-segmentation accuracy (ie. multi-atlas segmentation) over model-based approaches
+segmentation accuracy (i.e. multi-atlas segmentation) over model-based approaches
 \citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
@@
 Such selection has the added benefit of significantly reducing the number of
-nonlinear registrations.  For example Collins and Pruessner \citep{Collins2010}
+nonlinear registrations.  For example \citet{Collins2010}
 demonstrated that only 14 atlases, selected based on highest similarity between
@@
 images, this method could theoretically work using a much smaller atlas library.
-In their validation, LEAP was used to segment the whole hippocampus in the ADNI1
-baseline dataset, achieving a mean Dice score of 0.85 with manual segmentations.
+In their validation, LEAP was used to segment the whole hippocampus in the
+ADNI-1 baseline dataset, achieving a mean Dice score of 0.85 with manual
+segmentations.
 
@@
       caption=caption, caption.loc = c("top"),
-      title="tab:ADNI1-xvalidation-demographics")
+      label="tab:ADNI1-xvalidation-demographics")
 @
@@
 performance, and we use it as the basis of comparison in this experiment.
-Additionally, we report the Jaccard index, another commonly used similarity
-measure:
+%Additionally, we report the Jaccard index, another commonly used similarity
+%measure:
 
@@
 
- \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]
+% \[\text{Jaccard (J)} = \frac{|A \cap B|}{|A \cup B|} = \frac{DSC}{(2-DSC)}\]
 
@@
       caption=caption, caption.loc = c("top"),
-      title="tab:ADNI1-xvalidation-demographics")
-@
-\begin{table*}[!tbp]
-\scriptsize
-\caption{\textbf{Subfield validation template library demographics.}
-\label{tab:WAval-healthy-demographics}}
-\begin{center}
-\begin{tabular}{l | c c c}
-\hline 
-Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
-\hline 
-Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
-Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
-Iterations        &   30    &    30   &   10     \tabularnewline
-Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
-Sub-Lattice       &   6     &    6    &   6      \tabularnewline
-Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
+      title="tab:WAval-healthy-demographics")
+@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 <<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ants = subset(all_data_mean, 
+ants = subset(all_data, 
                      reg_method=="ANTS" & 
@@
 
-ants.stats = ddply(ants, c("subject", "atlases","templates.mb"), 
+ants.stats = ddply(ants, c("subject", "label", "atlases","templates.mb"), 
                    function (df) {
@@
 
-stats.tests = ddply(data.stats, c("atlases","templates.mb"), function (df) {
+stats.tests = ddply(ants.stats, c("atlases","templates.mb"), function (df) {
   t.var = t.test(df$MA.var, df$MB.var)
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Difference of variance during validation (p-value)" ) +
+  ylab( "Difference of variance during validation (p-value)") +
   scale_colour_hue(name="Atlases") +
@@
 control inspection. Quality control for FreeSurfer segmentations was provided along
-with the segmentations themselves from ADNI \cite{Hartig2010}. One author (MP) 
-inspected
-
-Only those images which had segmentations from every method
-are included in the following analysis (a total of
-\Sexpr{length(means.complete$RID)} images).
+with the segmentations from ADNI \cite{Hartig2010}. One of the authors (MP) 
+performed visual quality inspection for MAGeT and FSL segmentations using similar 
+quality control guidelines (if either hippocampus was under or over segmented
+by greater than 10mm in three or more slices then the segmentation did not pass).
+
+Only those images which had quality segmentations from each method are 
+included in the following analysis (a total of \Sexpr{length(means.complete$RID)}
+images).
 
@@
 latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
-    caption="\\adnidataset dataset segmentation totals and QC failures.", 
+    caption="\\textbf{Number of segmented images and quality control failures of 
+             \\adnidataset dataset by method.}", 
     caption.loc = c("top"), rowname=NULL, 
@@
 <<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
-caption = paste(
-  "Survey of automated segmentation accuracy of the ADNI dataset",
-  sep=" ")
-
-tab = read.csv("data/ADNI-existing-work.csv")
-tab$Notes <- NULL
-
-adni_lit = subset(tab, Dataset=="ADNI")
-adni_lit$Dataset <- NULL
-other_lit = subset(tab, Dataset!="ADNI")
-other_lit$Dataset <- NULL
-
-latex(adni_lit, file="", 
-      size="scriptsize",
-      caption=caption, 
-      rowname=NULL,
-      col.just=c("p{1.5in}l","p{0.5in}","p{1.3in}","p{1.5in}"),
-      title="")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+# caption = 
+#   "\\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
+#    ADNI dataset.} For each method, the number of manually labelled atlases used for 
+#    training, the best Dice's overlap measure, the disease classes measured, and 
+#    the validation procedure are shown. LOOCV = Leave-one-out cross-validation. 
+#    Some studies of automated segmentation of ADNI images are excluded because they 
+#    do not not provide overlap measures for the hippocampus \\cite{Heckemann2011, Chupin2009}.
+#   "
+# 
+# tab = read.csv("data/ADNI-existing-work.csv")
+# tab$Notes <- NULL
+# 
+# adni_lit = subset(tab, Dataset=="ADNI")
+# adni_lit$Dataset <- NULL
+# 
+# 
+# latex(adni_lit, #file="", 
+#       size="scriptsize",
+#       caption=caption, 
+#       rowname=NULL,
+#       col.just=c("p{2in}","p{1in}","p{1in}","p{1in}","p{1in}"),
+#       title="")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Tabulation of segmentation accuracies of existing methods (ADNI)
+%
+%
+<<mb-best-kappa, echo=F, cache=FALSE>>=
+ants.mean = subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &      templates.mb == 19 & atlases == 9)
+mb.best.kappa = mean(ants.mean$k.mb)
+@
+
+\begin{table}[!tbp]
+\scriptsize
+\caption{\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
+   ADNI dataset.} For each method, the number of manually labelled atlases used for 
+   training, the best Dice's overlap measure, the disease classes measured, and 
+   the validation procedure are shown. Unless specified, validation datasets are
+   composed equally of subjects diagnosed with Alzheimer's Disease (AD), Mild 
+   Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV = Leave-one-out 
+   cross-validation. 
+   Some studies of automated segmentation of ADNI images are excluded because they 
+   do not not provide overlap measures for the hippocampus \cite{Heckemann2011,Chupin2009}.
+  \label{tab:other-methods}} 
+\begin{center}
+\begin{tabular}{ l  c  p{1in}  l  p{2in}  }
+\hline\hline
+Method & Atlases & DSC & Reference & Validation \tabularnewline
+\hline
+\mb                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
+LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
+ACM (AdaBoost-based)    &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
+Patch-based label fusion&16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
+Multi-atlas             &30&0.885                 &\cite{Lotjonen2010}&Segmentation of 60 subjects\tabularnewline
+Multi-atlas + weighted fusion&20&0.898 (CN) \newline 0.798 (left HC, MCI) &\cite{Wang2011}  &10 rounds of Monte Carlo CV on 20 subjects, pool of 139 (CN/MCI)\tabularnewline
+Multi-atlas (MAPS)    &55&0.890                   &\cite{Leung2010}&Segmentation of 30 subjects (10 AD, MCI, and CN)\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-Throughout the cross validation in Experiment 1 (10-fold cross validation in the
+Throughout the cross validation in Experiment 1 (10-fold Monte Carlo cross 
+validation in the
 \adnidataset dataset subsample) we find that two parameter choices improve 
@@
 in Experiment 1 is amongst one of the most stringent performed in the
-multi-atlas/segmentation literature.  To the best of our knowledge other groups
-have only ever performed cross-validation \citep{?}.  This suggests that
+multi-atlas/segmentation literature.  To the best of our knowledge, other groups
+using ADNI data for validation do at most a single round of leave-one-out-validation
+(Table \ref{tab:other-methods}), aside from \citep{Wang2011} who perform a 
+similar validation as us. The thoroughness of our validation suggests that
 our results are reflective of a true average over the choice of parameter
@@
 
-% Validation in other works
+% Validation of ADNI in other works
 % Chupin 2009: LOOCV, 1-fold validation
@@
 %                 agreement between two different atlas sets
-% Coupe 2012: LOOCV
+% Coupe 2011e: LOOCV
 % Lotjonen 2010: LOOCV, 1-fold validation
@@
 % Collins 2010: 1-fold validation, QC
-%
+% Morra 2008: LOOCV with 21 atlases, 2 manual raters
+% Wang 2011: 10 rounds of MCCV with 20 atlases and 20 subjects, chosen from 
+%            a pool of 139 images
 

### 2013-08-07 comments/paper_MMC_discussion.Rnw
some of aristotle's comments addressed

--- 2013-07-29 paper.Rnw
@@
 options(digits=3)
-theme_set(theme_bw(base_size = 12))
+theme_set(theme_bw())
+
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  require(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
+bland_altman_plot <- function(x, y, c) {
+  data = data.frame(x=x,y=y,diff=x-y, m = (x+y)/2, c=c)
+  return(ggplot(data, aes(x= m, y = diff, colour = as.factor(c))) + 
+    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
+               linetype=2, color='brown') + geom_point())
+}
 
@@
 \usepackage[section]{placeins}  % keep floats in their place.
-\usepackage{authblk}    % Listing Author affiliations
 
@@
 
-% styling
+% add \subsubsubsection
+% \setcounter{secnumdepth}{4}  
+% \titleformat{\paragraph}
+% {\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+% \titlespacing*{\paragraph}
+% {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 \newcommand{\subsubsubsection}[1]{\paragraph{#1}}
-\renewcommand\Affilfont{\itshape\small}  %authblk
 
@@
 \newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T }
-\newcommand{\FEPdataset}{SZ-FEP }
-
-% title and authors
+%notes and TODO formatting
+%\newcommand{\marginnote}[1]{\-\marginpar[\raggedleft\footnotesize #1]{\raggedright\footnotesize #1}}
+%\newcommand{\comment}[1]{\begin{kframe}{\textcolor{red}{#1}}\end{kframe}}
+%\newcommand{\todo}[1]{\comment{TODO #1}}
+%\newcommand{\mc}[1]{\comment{MC: #1}}
+\begin{document}
+%\SweaveOpts{concordance=TRUE}
+
 \title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
-\author[1]{Jon Pipitone}
-\author[1]{Matt T. Park}
-\author[1]{Julie Winterburn}
-\author[1]{Tristram A. Lett}
-\author[2,3]{Jason P. Lerch}
-\author[4]{Jens C. Pruessner}
-\author[4]{Martin Lepage}
-\author[1,5]{Aristotle Voineskos}
-\author[1,5,6]{M. Mallar Chakravarty}
-\author[ ]{Alzheimer's Disease Neuroimaging Initiative}
-\affil[1]{Kimel Family Translational Imaging-Genetics Lab, Centre for Addiction and
-Mental Health, Toronto, ON, Canada}
-\affil[2]{Neurosciences and Mental Health, Hospital for Sick Children, Toronto,
-ON,Canada}
-\affil[3]{Department of Medical Biophysics, University of Toronto, Toronto, ON,
-Canada}
-\affil[4]{Brain Imaging Group, Douglas Hospital Research Centre, Verdun, QC, Canada}
-\affil[5]{Department of Psychiatry, University of Toronto, Toronto, ON, Canada}
-\affil[6]{Institute of Biomedical Engineering, University of Toronto, Toronto,
-ON, Canada}
-
-\renewcommand\Authands{ and }
-
-\date{}
-
-\begin{document}
+\author{Pipitone J., Winterburn J., Lett, T., Lerch J., Pruessner J., Lepage M., \\ 
+Voineskos A., Chakravarty M.M., and \\ 
+the Alzheimer's Disease Neuroimaging Initiative}
 \maketitle
@@
 \begin{abstract}
-%Neuroimaging research often relies on automated anatomical segmentations of MR
-%images of the brain. Current multi-atlas based approaches provide accurate
-%segmentations of brain images by propagating manually derived segmentations of
-%specific neuroanatomical structures to unlabelled data. These approaches often
-%rely on a large number of such manually segmented atlases that take significant
-%time and expertise to produce. We present an algorithm for the automatic
-%segmentation of the hippocampus that minimizes the number of atlases needed
-%while still achieving similar accuracy to multi-atlas approaches.
+Neuroimaging research often relies on automated anatomical segmentations of MR
+images of the brain. Current multi-atlas based approaches provide accurate
+segmentations of brain images by propagating manually derived segmentations of
+specific neuroanatomical structures to unlabelled data. These approaches often
+rely on a large number of such manually segmented atlases that take significant
+time and expertise to produce. We present an algorithm for the automatic
+segmentation of the hippocampus that minimizes the number of atlases needed
+while still achieving similar accuracy to multi-atlas approaches.
+\todo[inline]{finish}
+
 \end{abstract}
-
-\parbox{4in}{
-\textbf{Contact:} \\
-Jon Pipitone and M. Mallar Chakravarty \\
-Kimel Family Translation Imaging-Genetics Research Laboratory \\
-Research Imaging Centre \\
-Centre for Addiction and Mental Health \\
-250 College St. \\
-Toronto, Canada   M5T 1R8 \\
-jon.pipitone@camh.ca; mallar.chakravarty@camh.ca}
 
@@
 understand the unique neuroanatomy of subjects. Typically, the gold standard for
-neuroanatomical segmentation is manual delineation by an expert human rater. However
+neuroanatomical segmentation is manual delineation by an expert human. However
 the rapid increase in the availability of MRI data and the time and expertise
 required for manual segmentation is prohibitive
-\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}.
+\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007,Weiner2013}.
 Further, there is little agreement between researchers regarding how exactly the
@@
 successfully in several contexts
-\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
+\citep{MallarMedIA200809,MallarHBM20082009,Collins1995,Haller1997}, it is limited
 in accuracy by the introduction of errors due inaccuracies in the nonlinear
@@
 the use of multiple manually segmented atlases and probabilistic segmentation
-techniques, such as those found in the FreeSurfer package \citep{Fischl2002}.
+techniques, such as those found in the FreeSurfer package \citep{Fischl2002},
 FreeSurfer uses a probabilistic atlas of anatomical and tissue classes along
@@
 accommodate atlases that are somehow exceptional such as those derived from
-serial histological data \citep{Chakravarty2006,Yelnik2007} or high-resolution MRI
+serial histological data \citep{Chakravarty06a,Yelnik2007} or high-resolution MRI
 data that enables robust identification of hippocampal subfields
@@
 segmentation of the striatum, pallidum, and thalamus using a single atlas
-derived from serial histological data is possible \citep{MallarChakravarty2012}.
-The novelty of this manuscript is the extension our multi-atlas methodology to
-the hippocampus using more than a single input atlas, while simultaneously
-limiting the number of possible inputs used during segmentation, and
-demonstrating that accurate identification of the hippocampal subfields is
-indeed possible using this methodology.
+derived from serial histological data \citep{Chakravarty2013} is possible.  The
+novelty of this manuscript is the extension our multi-atlas methodology to the
+hippocampus using more than a single input atlas, while simultaneously limiting
+the number of possible inputs used during segmentation, and demonstrating that
+accurate identification of the hippocampal subfields is indeed possible using
+this methodology.
 
@@
 between very dissimilar images to be broken up into sequences of smaller
-deformations.  Although \citet{Wolz2010} begin with an atlas library of 30 MR
-images, this method could theoretically work using a much smaller atlas library.
-In their validation, LEAP was used to segment the whole hippocampus in the ADNI1
+deformations.  Although Wolz et al. begin with an atlas library of 30 MR images,
+this method could theoretically work using a much smaller atlas library.  In
+their validation, LEAP was used to segment the whole hippocampus in the ADNI1
 baseline dataset, achieving a mean Dice score of 0.85 with manual segmentations.
@@
 applicability of their work using data from 10 subjects.  In the second method,
-\citet{Yushkevich2009} acquired and labelled hippocampal subfields on
-high-resolution MRI data from post-mortem medial temporal lobe samples. Using
-nonlinear registration guided by manually derived hippocampus masks and specific
-landmarks, they demonstrate accurate parcellation of hippocampal subfields in
-unlabelled MRI volumes.
+Yushkevich and colleagues \citep{Yushkevich2009} acquired and labelled
+hippocampal subfields on high-resolution MRI data from post-mortem medial
+temporal lobe samples an demonstrate it's applicability in the segmentation of
+other MRI-volumes using nonlinear registration guided using manually derived
+hippocampus masks and specific landmarks. In their work they demonstrate
+accurate parcellations of the subfields.
 
@@
 input image (i.e. atlas or template) is registered to an unlabelled target image.
-Second, the labels from each image are propagated to the target image space. 
+Second, the labels from each image are propogated to the target image space. 
 Third, the labels are combined into a single labelling by way of a label fusion method 
@@
 \ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
-right whole hippocampi are available \citep{Hsu2002}. These labels have been
-generated using the SNT tool from Medtronic Surgical Navigation Technologies,
-Louisville, CO (see Supplementary Materials for detailed discussion of the
-manual segmentation process used).
+right whole hippocampi are available. These labels have been generated using the
+SNT tool from Medtronic Surgical Navigation Technologies, Louisville, CO (see
+Supplementary Materials for detailed discussion of the manual segmentation
+process used).
 
@@
 cross-validation, consists of repeated rounds of validation conducted on a fixed
-dataset \citep{Shao1993}. In each round, the dataset is randomly partitioned into
+dataset \cite{Shao1993}. In each round, the dataset is randomly partitioned into
 a training set and a validation set. The method to be validated is then given
@@
 
-A total of ten validation rounds are performed on each subject in the dataset,
+A total of ten validation rounds are perfomed on each subject in the dataset,
 over each combination of parameter settings. The parameter settings
@@
 the correlation coefficient.  For the purposes of this work we used the
-regularization parameters optimized in \citet{Robbins2004}, displayed in table
-\ref{tab:ANIMAL-params}.
+regularization parameters optimized in Robbins et al. \citep{Robbins2004},
+displayed in table \ref{tab:ANIMAL-params}.
 
@@
 over the choice of transformation model, objective function, and the consistency
-of the final transformation \citep{Avants2008}. The transformation is estimated in a
-hierarchical fashion where the MRI data is subsampled, allowing large
-deformations to be estimated and successively refined at later hierarchical
-stages (where the data is subsampled to a finer grid). The deformation field and
-the objective function are regularized with a Gaussian kernel at each level of
-the hierarchy. The \ants algorithm is freely available
-\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the \ants
-algorithm compatible with the MINC data format, mincANTS
-\url{https://github.com/vfonov/mincANTS}.
+of the final transformation. The transformation is estimated in a hierarchical
+fashion where the MRI data is subsampled, allowing large deformations to be
+estimated and successively refined at later hierarchical stages (where the data
+is subsampled to a finer grid). The deformation field and the objective function
+are regularized with a Gaussian kernel at each level of the hierarchy. The \ants
+algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an
+implementation of the \ants algorithm compatible with the MINC data format,
+mincANTS \url{https://github.com/vfonov/mincANTS}.
 
@@
   atlas labels after linear registration to the template, dilated by three voxels
-  \citep{MallarChakravarty2012}.  The number of top ranked template library image labels
+  \citep{Chakravarty2012}.  The number of top ranked template library image labels
   is a configurable parameter and displayed as the size of the template library
@@
 The manual segmentations (SNT) provided as part of the ADNI dataset are used as 
-the gold standard to compare with \citep{Hsu2002}.  The segmentation accuracy
-reported is averaged over the ten validation rounds for each parameter setting.
+the gold standard to compare with.  The segmentation accuracy reported is averaged 
+over the ten validation rounds for each parameter setting.
 
@@
 template images is varied. This is achieved by first computing the standard 
-deviation and variance of DSC scores in each block of ten validation rounds per
-subject. The distribution of these statistics across all subjects is then
-compared between \mb and multi-atlas using a Student's t-test. A significant
-difference between distributions is taken to show either a larger or smaller
-level of variability between methods.
+deviation and variance of DSC scores in each block 
+of ten validation rounds per subject. The distribution of these statistics across 
+all subjects is then compared between \mb and multi-atlas using a Student's 
+t-test. A significant difference between distributions is taken to show either
+a larger or smaller level of variability between methods.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
+\subsubsection{Experiment 2: Winterburn Atlases Cross-Validation}
 In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
 segmentation is tested using a leave-one-out cross-validation (LOOCV) design.
-The optimal parameter settings for \mb found in Experiment 1 are used in this 
-experiment.
+The optimal parameter settings found in Experiment 1 are used.
 
@@
 
-In this experiment the Winterburn atlases are used as the cross-validation
-dataset.  The five $300\mu$-isotropic voxel images and labels are used as
-atlases, and LOOCV is conducted using the five Winterburn atlas subject images
-subsampled (using trilinear interpolation) to 0.9mm-isotropic voxel resolution
-(referred to as the {\em Subsampled} dataset) as input subjects. Subsampling the
-subject images allows us to assess \mb in a typical segmentation scenario
-(high-resolution atlases and lower-resolution subjects).  The template library
-is composed of the subject images, plus an additional set of 3T T1 images
-(0.9mm-isotropic voxels) of healthy subjects acquired separately (Table
-\ref{table:WAval-healthy-demographics}). The optimal size of template library,
-registration method, and label fusion method found in Experiment 1 are used.
-Each resampled Winterburn atlas subject image is segmented by \mb with that
-subject's image excluded from the atlas library.
-
-We reproduce this experiment in a separate LOOCV run in which the resampled 
-Winterburn atlas subject images are substituted for separately acquired T1 
-BRAVO images of four of the subjects (referred to as the {\em BRAVO} dataset).
+In this experiment, the five $300\mu$-isotropic voxel Winterburn atlases are
+used as the atlas library for \mb segmentation. LOOCV is conducted separately
+for two different input datasets: the four Winterburn atlas subject T1 BRAVO
+images (referred to as the {\em BRAVO} dataset), and the five
+Winterburn atlas subject images subsampled to 0.9mm-isotropic voxel resolution
+(referred to as the {\em Subsampled} dataset. Each subject in the
+dataset is segmented by \mb with that subject's image excluded from the atlas
+library. The template library consists of 3T T1 images (0.9mm-isotropic
+voxels) of healthy subjects in addition to the images from the dataset being
+evaluated.
+
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used.
 
 \paragraph{Evaluation method}  
-To assess the \mb LOOCV segmentations we compute the relative percent error in 
-hippocampal volume with the full resolution Winterburn atlas 
-segmentations. In addition, by computing the relative error in volume of the
-Winterburn atlas labels resampled (with nearest-neighbour interpolation) to
-0.9mm-isotropic voxels, we obtain a baseline error to assess against.
+\mb performance is measured with respect to straight-forward nearest-neighbour
+resampling of the Winterburn atlas labels to 0.9mm-isotropic voxels (i.e. no 
+label propagation). The volumes of each hippocampal subregion in the subsampled
+Winterburn atlas labels are compared to the volumes obtained from the full 
+resolution Winterburn atlas labels, the \mb labels on the subsampled Winterburn 
+atlas images, and the \mb labels on the T1 BRAVO acquisitions.
 
@@
 
-\paragraph{\FEPdataset dataset}
+\paragraph{SZ-FEP dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 neuroimaging study; only those suffering from schizophrenia spectrum disorders
-were considered for this analysis.  For complete program details see
-\citet{Malla2003}.
+were considered for this analysis.  For complete program details see Malla et
+al. \citep{Malla2003}.
 
@@
 
-Each subject hippocampus is traced following a validated segmentation protocol 
-\citep{Pruessner2000}.
+The hippocampus were traced following a validated protocol developed by Dr Jens
+Pruessner \todo{cite: (Pruessner et al., 2000)}. A recent update to this
+protocol by Dr J Pruessner in 2006 allows to accurately and consistently
+subdivide the hippocampus into three different subregions: head, body, and tail.
+We use these segmentations as to validate our implementation of \mb.
 
@@
 \mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2).  All images from the \FEPdataset dataset are segmented by \mb.  The 
-optimal size of template library, registration method, and label fusion method 
-found in Experiment 1 are used.
+(see Experiment 2).  All images from the SZ-FEP dataset are segmented.  The optimal
+size of template library, registration method, and label fusion method found in
+Experiment 1 are used.
 
 \paragraph{Evaluation method}
-The manual and Winterburn hippocampal segmentation protocols differ slightly in
-the neuroanatomical features that are delineated.  This poses a problem for
-evaluation by measuring overlap.  That is, since different protocols will
-necessarily produce segmentations that do not perfectly overlap, the degree of
-overlap cannot be solely used to compare segmentation methods using different
-protocols. In place of an overlap metric, we can assess the degree of correlation in
-average hippocampal volume of the subjects produced by each method.
-Specifically, Pearson correlation is used.
+The manual segmentation protocol used to segment the Winterburn atlases has some
+neuroanatomical differences in comparison to the protocol used to segment the
+SZFEP dataset.  Therefore, rather than evaluate using an overlap metric, such as
+DSC, which assumes a fixed segmentation protocol, we instead correlate whole
+hippocampus volume between \mb and manual segmentations.
+
+\todo{done elsewhere?}
 
@@
 \paragraph{Evaluation method}
-As in Experiment 3, the manual (SNT) and Winterburn hippocampal segmentation
-protocols differ in the neuroanatomical features delineated, and so we must
-assess \mb by the degree of correlation of average hippocampal volume across all
-subjects produced by \mb and by manual segmentation.  Specifically, Pearson
-correlation is used.  For comparison, we also compute the correlation in
-hippocampal volume between the existing, established automated segmentation
-methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
-volume-related fixed and proportional biases in all segmentation methods using
-Bland-Altman plots \citep{Bland1986}.
+Since the hippocampal segmentation protocols differ between the ADNI labels and
+Winterburn atlases, this poses a problem for direct evaluation between labels
+produced by \mb and the ADNI labels in terms of overlap; we would not expect
+different segmentation protocols to have a high degree of overlap.  Instead, to
+evaluate the performance of \mb we compare the correlation of \mb segmentation
+volumes with manual segmentation (SNT) volumes. Additionally, we correlate the
+hippocampal volumes of established automated segmentation methods to \mb
+segmentations.
 
@@
 % ADNI1-COMPLETE
-% Prepares the data in the form needed for plotting.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<ADNI-volumes-prep,echo=F,cache=T>>=
-means.complete   <- read.csv(gzfile('data/cache/ADNI1:means.complete.csv.gz'))
-package_totals   <- read.csv(gzfile('data/cache/ADNI1:package_totals.csv.gz'))
+% Prepares the data in a the form needed for plotting.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+<<ADNI-src-volumes-prep,echo=F,cache=T>>=
+totals.complete  <- read.csv('data/cache/ADNI1:totals.complete.csv')
+totals           <- read.csv('data/cache/ADNI1:totals.csv')
+package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
 @
@@
 %                                                                             
-<<ADNI-demographics, echo=F, results="asis",cache=FALSE>>=
+<<ADNI1-scr-demographics, echo=F, results="asis",cache=FALSE>>=
 # RID < 2000 ensures only ADNI1 patients
@@
 label fusion method used does not significantly improve label accuracy, contrary 
-to the findings of \citet{Aljabar2009} when using weighted voting on much larger
-atlas/template libraries. In the remainder of  this section, only results using
-the \ants registration algorithm and majority vote fusion will be shown.
+to the findings of Aljabar et al. \cite{Aljabar2009} when using weighted voting
+on much larger altas/template libraries. In the remainder of  this section, only 
+results using the \ants registration algorithm and majority vote fusion will 
+be shown.
 
@@
 
+\todo{mention failures}
+
 In addition to an improvement in accuracy over multi-atlas-based segmentation,
@@
 (Figure \ref{ADNI1-xval-variability}).  The size of template library 
-necessary to reach a significant ($p < 0.05$) decrease in variance and standard 
+necessary to reach a significant (p<0.5) decrease in variance and standard 
 deviation grows with the size of atlas library used.  A template library of 19
@@
 deviation for 3-7 atlases. 
-
-We have omitted results obtained when using an even number of atlases or
-templates since with this configuration we found significantly decreased
-performance. We believe this is as a result of an inherent bias in the majority
-vote fusion method used (see Discussion).
+\todo{Tris: only show variance OR standard deviation because otherwise the story gets complicated... why compare measures?}
+
+It is interesting to note that with an even number of templates, \mb shows a
+small decrease in performance relative to when one fewer template image is used.
+See section \todo{ref} for a discussion of this behaviour. In the remainder of
+this section, only results from odd-sized template libraries will be shown.
 
@@
 <<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-all_data  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data.csv.gz'))
-all_data_mean  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data_mean.csv.gz'))
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{figure}
-% <<ADN1-library-composition, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-% # These composition datasheets describe the subjects and subject DX in the 
-% # template and atlas libraries for each voting run
-% atlas_comp = read.csv('data/a2a_ants/atlas_library_composition.csv')
-% templ_comp = read.csv('data/a2a_ants/template_library_composition.csv')
-% a2a.best = subset(all_data, atlases == 9 &
-%                     templates.mb == 19 &
-%                     reg_method=='ANTS' &
-%                     method.mb =='Majority Vote')
-% 
-% cumulative_dist = 
-%   ggplot(a2a.best, aes(x=k.mb)) + stat_ecdf() + scale_y_continuous(breaks=seq(0,1,0.1))
-% 
-% a2a.best = subset(a2a.best, k.mb > 0.75) # drop the guys that suck
-% 
-% a2a.best = merge(a2a.best, atlas_comp, by=c("timestamp","atlases"))
-% a2a.best = merge(a2a.best, templ_comp, by.x=c("timestamp","templates.mb"), by.y=c("timestamp","templates"))
-% a2a.best$templates_AD_percent = (a2a.best$templates_AD / 19) * 100
-% a2a.best$templates_MCI_percent = (a2a.best$templates_MCI / 19) * 100
-% a2a.best.melted = melt(a2a.best, measure.vars=c("templates_AD_percent", "templates_MCI_percent"))
-% 
-% k_by_composition = 
-%   ggplot(a2a.best.melted, aes(x=value, y=k.mb, colour=DX)) + facet_grid(variable ~ DX) + 
-%   geom_boxplot(aes(group=as.factor(value), colour=DX)) +  geom_smooth(method="lm") 
-% 
-% snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
-% a2a.best = merge(a2a.best, snt.unilateral.vols)
-% 
-% # bland-altman plots
-% bland_mb_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.mb, DX) + 
-%   geom_smooth(method="lm") + 
-%   scale_y_continuous(breaks=seq(-500, 500, 50)))
-% 
-% bland_ma_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.ma, DX) + 
-%   geom_smooth(method="lm") + 
-%   scale_y_continuous(breaks=seq(-500, 500, 50)))
-% 
-% # just the trials with a high percentage of MCI in the template library
-% a2a.best.highMCI = subset(a2a.best, templates_MCI_percent > 50)
-% 
-% bland_mb_highMCI = with(a2a.best.highMCI, bland_altman_plot(volume.snt, volume.ma, DX) + 
-%   geom_smooth(method="lm") + 
-%   scale_y_continuous(breaks=seq(-500, 500, 50)))
-% 
-%   
-% @
-% \end{figure}
+all_data  <- read.csv('data/cache/ADNI-XVAL:all_data.csv')
+all_data_mean  <- read.csv('data/cache/ADNI-XVAL:all_data_mean.csv')
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+<<ADN1-library-composition, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+# These composition datasheets describe the subjects and subject DX in the 
+# template and atlas libraries for each voting run
+atlas_comp = read.csv('data/a2a_ants/atlas_library_composition.csv')
+templ_comp = read.csv('data/a2a_ants/template_library_composition.csv')
+a2a.best = subset(all_data, atlases == 9 &
+                    templates.mb == 19 &
+                    reg_method=='ANTS' &
+                    method.mb =='Majority Vote')
+
+cumulative_dist = 
+  ggplot(a2a.best, aes(x=k.mb)) + stat_ecdf() + scale_y_continuous(breaks=seq(0,1,0.1))
+
+a2a.best = subset(a2a.best, k.mb > 0.75) # drop the guys that suck
+
+a2a.best = merge(a2a.best, atlas_comp, by=c("timestamp","atlases"))
+a2a.best = merge(a2a.best, templ_comp, by.x=c("timestamp","templates.mb"), by.y=c("timestamp","templates"))
+a2a.best$templates_AD_percent = (a2a.best$templates_AD / 19) * 100
+a2a.best$templates_MCI_percent = (a2a.best$templates_MCI / 19) * 100
+a2a.best.melted = melt(a2a.best, measure.vars=c("templates_AD_percent", "templates_MCI_percent"))
+
+k_by_composition = 
+  ggplot(a2a.best.melted, aes(x=value, y=k.mb, colour=DX)) + facet_grid(variable ~ DX) + 
+  geom_boxplot(aes(group=as.factor(value), colour=DX)) +  geom_smooth(method="lm")
+
+snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
+a2a.best = merge(a2a.best, snt.unilateral.vols)
+
+# bland-altman plots
+bland_mb_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.mb, DX) + 
+  geom_smooth(method="lm") + 
+  scale_y_continuous(breaks=seq(-500, 500, 50)))
+
+bland_ma_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.ma, DX) + 
+  geom_smooth(method="lm") + 
+  scale_y_continuous(breaks=seq(-500, 500, 50)))
+
+# just the trials with a high percentage of MCI in the template library
+a2a.best.highMCI = subset(a2a.best, templates_MCI_percent > 50)
+
+bland_mb_highMCI = with(a2a.best.highMCI, bland_altman_plot(volume.snt, volume.ma, DX) + 
+  geom_smooth(method="lm") + 
+  scale_y_continuous(breaks=seq(-500, 500, 50)))
+
+@
+\end{figure}
 
@@
 \begin{figure}
-<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=7, fig.height=7>>=
-stderr <- function(x) sqrt(var(x)/length(x))
+<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
 ggplot(subset(all_data_mean, (templates.mb*atlases) %% 2 == 1),
@@
                aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
-  stat_summary(fun.y=mean,
-               fun.ymin=function (x) mean(x) - stderr(x),
-               fun.ymax=function (x) mean(x) + stderr(x), 
-               geom='errorbar',
-               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
+    stat_summary(fun.y=mean,geom='point',
+               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
   #stat_smooth(method='lm', formula=y~log(x)) + 
-  scale_colour_hue(name="Number of Atlases") +
+  scale_colour_hue(name="Atlases") +
   xlab( "Number of Templates" ) + 
   ylab( "Mean DSC" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text = element_text(size = 8))
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
   \label{fig:ADNI1-xval-k_mean}
-  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with manual
-  segmentations (SNT) for 69 ADNI1 subjects vs. atlas and template library size,
-  registration algorithm, and label fusion method. Points above zero indicate an
-  \mb parameter settings yielding a higher mean DSC score than multi-atlas
-  segmentation using the same number of atlases.}
+  \caption{
+  {\bf Mean DSC of \mb segmentations across 69 ADNI1 subjects, by atlas and
+  template library size, registration algorithm, and label fusion method.} 
+  }
 \end{figure}
@@
 \begin{figure}
-<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep'>>=
+<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=6,fig.height=4>>=
 ggplot(subset(all_data_mean, reg_method=="ANTS" & 
@@
   scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
-  scale_colour_hue(name="Number of Atlases") +
+  scale_colour_hue(name="Atlases") +
   scale_alpha_continuous(guide = "none") + 
@@
 @
-  \caption{Difference in mean Dice's Similarity Coefficient of \mb and
-  multi-atlas segmentations with manual (SNT) segmentations for a range of atlas
-  and template library sizes.}
+  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
+  a range of parameter settings.}
   \label{ADNI1-xval-k-diff}
@@
 <<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ants = subset(all_data_mean, 
+ANTSmajvote = subset(all_data, 
                      reg_method=="ANTS" & 
@@
 
-ants.stats = ddply(ants, c("subject", "atlases","templates.mb"), 
-                   function (df) {
+data.stats = ddply(ANTSmajvote, c("subject", "label","atlases","templates.mb"), function (df) {
   data.frame(
@@
 
-ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
+ggplot(subset(stats.tests, direction > 0, stat == "Variance"), 
   aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-This experiment explores \mb segmentation of hippocampal subfields. To
+This experiment explores \mb segmentations of hippocampal subfields. To
 achieve this, a leave-one-out validation is conducted in which lower-resolution
-images ($0.9mm^3$ voxels) of each Winterburn atlas subject are segmented using 
-the remaining Winterburn atlases.
-
-In general, across hippocampal subregions the percent error in volume of
-\mb segmentations compares
-favourably to the error resulting from image resampling (Figure
+images ($0.9mm^3$ voxels) of each Winterburn atlas subject is segmented using 
+the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
+atlases when subsampled to $0.9mm^3$ voxels are also computed.
+
+In general, across hippocampal subregions the percent error in volume between
+\mb segmentations and the manual Winterburn atlas segmentations compares
+favourably to error when resampling the atlas segmentations (Figure
 \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
 subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
-subregions show distinctly larger error than is found through resampling.
-
-Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb subfield
+subregions show distinctly larger volumes than resampling error. 
+\todo{note direction of change}.
+
+Figure \ref{fig:subfield-montage} shows a qualitative comparision of \mb subfield
 segmentation.
@@
 ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
-  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
   geom_boxplot() +
-  labs(y=expression(paste("Percent error in volume relative to 0.3", mm^3, " Winterburn Atlases")), x="Region") +
+  labs(y="Percent error in volume", x="Region") +
   scale_colour_discrete("") + 
@@
   \caption{Sagittal slices from two subjects showing a comparison of the
-  original Winterburn atlas subfield segmentations (at 0.3mm-isotropic voxel
-  resolution), the subsampled Winterburn segmentations (at 0.9mm-isotropic
-  voxel resolution), and the \mb labels on the subsampled atlas image.}
+  original Winterburn atlas subfield segmentations (at 0.3mm3-isotropic voxel
+  resolution), the subsampled Winterburn segmentations (at 0.9mm3-isotropic
+  voxel resolution), and the MAGeT brain labels on the subsampled atlas image.}
   \label{fig:subfield-montage}
@@
 
-\todo[inline]{Figure \ref{fig:subfield-montage}. Show MAGeT segmentations on
-BRAVO images.  Also, show coronal/transverse slices and MNI coordinates}
-
 \FloatBarrier
@@
 In this experiment \mb is applied to a dataset of images of first episode
-schizophrenia patients, using the Winterburn atlases and a template library of
+schizophrenia parients, using the Winterburn atlases and a template library of
 21 subject images selected at random.  Expert manual whole hippocampal
 segmentations are used as gold standards.
+
+\mb produces hippocampus segmentation volumes that are highly correlated with 
+manual segmentation volumes (Figure \ref{fig:SZ-volumes}). 
+\todo{p-value}
 
@@
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
-mean_sz_vols = ddply(SZ_volumes, c('Subject', 'Method'), function (df) {
-  data.frame(Volume = mean(df$Volume))})
-
-casted = dcast(mean_sz_vols, Subject ~ Method, value.var = "Volume")
+casted = dcast(SZ_volumes, Subject + Label ~ Method)
 lm = lm(MAGeT ~ Manual, casted)
@@
   geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
-  xlab(expression("Mean Manual hippocampus volume " (mm^3))) + 
-  ylab(expression("Mean MAGeT hippocampus volume "  (mm^3))) + 
+  xlab(expression("Manual unilateral whole hippocampus volume " (mm^3))) + 
+  ylab(expression("MAGeT unilateral whole hippocampus volume "  (mm^3))) + 
   scale_size_continuous(guide="none") 
 @
-  \caption{Subject mean hippocampal volumes by \mb vs. manually segmented
-   volumes from the First Episode Patients with Schizophrenia dataset. A linear fit
-   and standard error are shown.}
+  \caption{{\bf First Episode Schizophrenic Patients.} Comparison of total
+  HC volumes for \mb against manually rated Hippocampal volumes}
   \label{fig:SZ-volumes}
@@
 
-<<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
-  fep.cor.test = cor.test(casted$Manual, casted$MAGeT)
-@
-\mb produces hippocampus segmentation volumes that are highly correlated with 
-manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
-$t = \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001,
-fep.cor.test$p.value)}$; Figure \ref{fig:SZ-volumes}).
-
 \FloatBarrier
@@
 Based on the results from the ADNI1 Cross-Validation experiment, in this
-experiment \mb was configured with a template library of 21 randomly chosen
-subject images (7 from each disease class) and used majority vote label fusion.
-The entire \adnidataset dataset was segmented by \mb and we now compare the
-resulting volumes with those obtained by manual segmentation (SNT), and other
-automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
-\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
+experiment \mb was configured with a template library of 21 randomly
+chosen subject images (7 from each disease class) and used majority vote label
+fusion. The entire \adnidataset dataset was segmented by \mb and we 
+now compare the resulting volumes with those obtained by manual segmentation (SNT), 
+and other automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
+\ref{tab:ADN1-1.5T-Complete-1yr-Dataset-Segmentations} shows the total count of 
 segmentations available, including a count of those which have failed a quality 
@@
 are included in the following analysis (a total of
-\Sexpr{length(means.complete$RID)} images).
+\Sexpr{length(totals.complete$RID)} images; Table 
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals}).
 
 We find a close relationship in total bilateral hippocampal volume between
-all methods and manually segmented volumes (Figure \ref{fig:ADNI-volumes-plot}). 
-Volumes are correlated with Pearson $r > 0.78$ for all methods across disease
-categories.  Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}),
-\mb is consistently well correlated to manual volumes (Pearson $r > 0.85$), but
-appears to slightly over-estimate the volume of the AD hippocampus.
-
-To investigate the level of agreement with manually segmented hippocampal
-volumes, we constructed Bland-Altman plots for each method (Figure
-\ref{fig:ADNI-Bland-Altman}).  As \citet{Bland1986} noted, high correlation
-amongst measures of the same quantity does not necessarily imply agreement (as
-correlation can be driven by a large range in true values, for instance).  What
-is most striking in Figure \ref{fig:ADNI-Bland-Altman} is that all methods show
-an obvious proportional bias: FreeSurfer and FSL markedly under-estimate smaller
-hippocampi and over-estimate large hippocampi, whereas MAPER and \mb more
-conservatively show the reverse bias.  Additionally, all methods show a fixed
-bias, with FreeSurfer and FSL most dramatically over-estimating hippocampal
-volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and MAPER and
-\mb within $250 mm^3$ on average.
-
-Figure \ref{fig:ADNI-subfields} shows a qualitative comparison of \mb and 
+all methods and manually segmented volumes (Figure \ref{fig:ADNI-scr-volumes-plot}). 
+Volumes are correlated with Pearson-r > 0.78 for all methods across disease categories.
+Within disease categories (Figure \ref{fig:ADNI-scr-volumes-boxplot}),
+\mb is consistently well correlated to manual volumes (Pearson-r > 0.85), but appears to 
+slightly over-estimate the volume of the AD hippocampus.
+
+To investigate the level of agreement with manually segmented hippocampal volumes, 
+we constructed Bland-Altman plots for each method (Figure \ref{fig:ADNI1-src-Bland-Altman}).  
+As Bland \& Altman, 1985 \cite{Bland1986} noted, high 
+correlation amongst measures of the same quantity does not necessarily 
+imply agreement (as correlation can be driven by a large range in true values, 
+for instance).  What is most striking in Figure \ref{fig:ADNI1-src-Bland-Altman}
+is that all methods show an obvious proportional bias: FreeSurfer and FSL 
+markedly under-estimate smaller hippocampi and over-estimate large hippocampi, 
+whereas MAPER and \mb more conservatively show the reverse bias.  Additionally, 
+all methods show a fixed bias, with FreeSurfer and FSL most dramatically 
+over-estimating hippocampal volume by $2600mm^3$ and $2800mm^3$ on average, respectively, 
+and MAPER and \mb within $250mm^3$ on average.
+
+Figure \ref{fig:ADNI-scr-subfields} shows a qualitative comparison of \mb and 
 manual (SNT) hippocampal segmentations for 10 randomly selected subjects in 
@@
 %                        
-<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
+<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
 caption = paste(
@@
 \begin{figure}[h]
-<<ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
-melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
+<<ADNI-scr-volumes-boxplot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
+melted = melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
               variable.name="Method", value.name="Volume")
@@
 })
-m2 = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
-levels(m2$Method) = c("FS"= "FreeSurfer", "FSL" = "FSL", "MAPER" = "MAPER", "MAGeT" = "MAGeT", "SNT" = "Manual")
-m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
-
-qplot(DX,value,data=m2,
-      colour=Method,geom="boxplot") + 
-      #geom_text(aes(y=-Inf,x=c(.7,.85,1,1.15, 1.7,1.85,2,2.15, 2.7,2.85,3,3.15), vjust=-5, label=round(pearson,2)), 
-      #          colour = 'black', size=3, data=correlations) +
+
+qplot(DX,value,data=melt(totals.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT")),
+      colour=variable,geom="boxplot") + 
+      geom_text(aes(y=-Inf,x=c(.7,.85,1,1.15, 1.7,1.85,2,2.15, 2.7,2.85,3,3.15), vjust=-5, label=round(pearson,2)), 
+                colour = 'black', size=3, data=correlations) +
       xlab("Diagnosis") + 
-      ylab(expression(paste("Hippocampal volume (", mm^3, ")"))) + 
+      ylab(expression(paste("Bilateral whole hippocampal volume (", mm^3, ")"))) + 
       theme(legend.direction = "horizontal", legend.position = "bottom")
 @
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by FreeSurfer, FSL, MAPER, \mb, and manual raters (SNT) vs. disease
-category.}
-  \label{fig:ADNI-volumes-boxplot}
+  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
+  MAPER, \mb (MAGeT) and manual (SNT) by disease category.  Pearson correlation 
+  with manual volumes are shown below each box-and-whisker.}
+  \label{fig:ADNI-scr-volumes-boxplot}
 \end{figure}
@@
 \begin{figure}[h]
-<<ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
-melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+<<ADNI-scr-volumes-plot,dependson='ADNI-baseline-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
             variable.name = "Method", value.name = "volume")
 
-df = means.complete
+df = totals.complete
 
 ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
-    geom_point(size=1) + 
+    geom_point() + 
     geom_smooth(method="lm") + 
-    xlab(expression("Manual hippocampal volume " (mm^3))) + 
-    ylab(expression("Automated hippocampal volume " (mm^3))) +
-    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
-              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
-                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
-                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
-                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
+    xlab(expression("Manual bilateral whole hippocampal volume " (mm^3))) + 
+    ylab(expression("Automated bilateral whole hippocampal volume " (mm^3))) +
+    annotate("text", y = c(3000,2500,2000,1500), x=Inf-100, size=3,
+              label = c(paste("FS r=", round(cor(df$SNT, df$FS),2)),
+                        paste("FSL r=", round(cor(df$SNT, df$FSL),2)),
+                        paste("MAPER r=", round(cor(df$SNT, df$MAPER),2)),
+                        paste("MAGeT r=", round(cor(df$SNT, df$MAGeT),2))),
               hjust=1.1,vjust=1.5) +
@@
 @
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by each of the four automated methods investigated (FreeSurfer (FS),
-FSL, MAPER, \mb) vs. manual rating (SNT). Linear fit lines and pearson
-correlations are shown for each method.}
-  \label{fig:ADNI-volumes-plot}
+  \caption{Comparison of hippocampus volumes obtained by FreeSurfer (FS), FSL, 
+  MAPER, \mb (MAGeT) and manual (SNT). Pearson correlation with manual volumes
+  are shown for each method.}
+  \label{fig:ADNI-scr-volumes-plot}
 \end{figure}
@@
 \begin{figure}
-<<ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
-names(means.complete)[8] = "FreeSurfer"
-melted=melt(means.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
+<<ADN1-src-Bland-Atlman, cache=TRUE, dependson='ADNI-baseline-volumes-prep', fig.width=6, fig.height=6>>=
+melted=melt(totals.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
             variable.name = "method", value.name = "volume")
-
 melted$diff = melted$SNT - melted$volume
-melted = subset(melted, (method == "FreeSurfer") | 
-                        (method == "FSL"  ) | # & diff > -5000 & diff < 900) | 
-                        (method == "MAPER") | # & diff > -500 & diff < 1000) | 
-                        (method == "MAGeT"))  # & diff > -1100))
-#melted = subset(melted, method == "MAGeT")
-melted$DX = factor(melted$DX, levels(melted$DX)[c(2,3,1)])
 melted$mean = ( melted$SNT + melted$volume ) / 2
@@
 
-names(melted)[4] = "Diagnosis"
-ggplot(melted, aes(x= mean, y = diff, colour=Diagnosis)) + 
-  facet_wrap( ~ method, nrow=2, scales="fixed") + 
-  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=1) +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
-             data=limits, size=0.3, alpha=0.5) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits,
-            hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of manual and automated hippocampal volume ' (mm^3))) + 
-  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
+ggplot(melted, aes(x= mean, y = diff, colour=DX)) + 
+  facet_wrap( ~ method, nrow=2, scales="free") + 
+  geom_point() +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown", data=limits) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits, hjust=1.1,vjust=-0.8, size=3) + 
+  xlab('Mean of manual and automated total hippocampal volume') + 
+  ylab('Difference between manual and automated total hippocampal volume') + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
   scale_x_continuous(breaks=seq(0,10000, 1000)) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text.y = element_text(angle = 90, hjust=0.5), 
-        axis.text = element_text(size = 8))
-@
-  \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset dataset by manual raters (SNT) and each of the four
-automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall mean
-difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
-horizontal lines. Linear fit lines are shown for each diagnosis group. Note,
-points below the mean difference indicate overestimation of the volume with
-respect to the manual rating, and vice versa. }
-
-  \label{fig:ADNI-Bland-Altman}
+  geom_smooth(method="lm", formula = y~x) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{Bland-Altman plot comparing manual (SNT) and automated total hippocampal volumes 
+  in the \adnidataset dataset.  The overall mean difference, and limits of 
+  agreement ($\pm 1.96SD$) are shown by dashed horizontal lines.}
+  \label{fig:ADNI-scr-Bland-Altman}
 \end{figure}
@@
   anatomical image, SNT manual label (green), and \mb label (blue). Noted are
-  examples of common segmentation idiosyncrasies: 
+  examples of common segmentation idiosyncracies: 
   {\em (a)} over-estimation of hippocampal head and  
@@
   {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb.}
-  \label{fig:ADNI-subfields}
+  \label{fig:ADNI-scr-subfields}
 \end{figure}
-
-\todo[inline]{Figure \ref{fig:ADNI-subfields}. Show coronal/transverse slices
-and MNI coordinates}
 
@@
 
-In this manuscript we have presented the implementation and validation of the
-\mb framework -- a methodology that requires very few input atlases in
-order to provide accurate and reliable segmentations.  We have demonstrated that
-our methodology robustly provides accurate and consistent segmentations in
-populations with different ageing and neuropsychiatric characteristics.
-Further, we have demonstrated that algorithmic performance is not dependent on a
-single definition of the hippocampus but is effective with differing hippocampal
-definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}.  Finally, we also demonstrate
-that \mb provides accurate automatic identification of the hippocampal
-subfields despite contrast and resolution limitations in standard T1-weighted
-image volumes.
-
-Throughout the cross validation in Experiment 1 (10-fold cross validation in the
-\adnidataset dataset subsample) we find that two parameter choices improve 
-segmentation accuracy: increasing the number of atlases, and the number of
-templates. However, after setting the parameters to 5 atlases and 15 templates
-there are diminishing returns with respect to this improvement. We were
-somewhat surprised to find that cross-correlation and normalized mutual
-information based weighted voting did little to improve segmentation accuracy.
-This suggests that using an intermediate template library generated by accurate
-nonlinear registration methods such as \ants \citep{Avants2008} sufficiently filters
-sources of error such as nonlinear registration error and partial volume effects
-that arise through the use of nearest neighbour resampling. From our previous
-work on the \mb algorithm we have shown that these increases in error
-are not simply a smoothing or averaging effect \citep{MallarChakravarty2012}.
-
-Although, the goal of this manuscript was to not exhaustively test or validate
-multiple different voting strategies in the context of our segmentation
-algorithm, it is important to note that other strategies for voting are
-available.  For example, other groups have used the STAPLE algorithm
-\citep{Warfield2004} (or variants of the STAPLE algorithm \citep{Duchesne2012})
-which weights each segmentation based upon its estimated
-performance level with respect to the other available candidate segmentations.
-Further, the sensitivity and specificity parameters could also be tuned to
-potentially improve segmentation accuracy and reliability.  It is rather likely
-that using more sophisticated voting methods would have a positive effect on the
-overall segmentation performance.
-
-To this end, more work is required in order to determine the source of the
-slight decrease in segmentation performance when the number of templates are set
-to an even number.  Our initial concern was that this dip in performance was a
-by-product of the \mb algorithm itself.  However, we determined that
-this pattern was also true in the analysis of multi-atlas segmentations we used
-in our experiments.  We believe that our majority voting methodology is biased
-towards labels with the lowest numeric values when breaking ties, thus causing
-the slight bias observed when using an even number of templates. This is another
-area where the voting scheme could be used to improve performance. However, it
-is worth noting that this limitation was previously identified by
-\citet{Heckemann2006a} and, subsequently, other groups have not even considered
-the potential pitfalls of an even number of candidate labels (eg.
-\citet{Leung2010}).
-
-Another concern is the moderate-to-small improvement observed in \mb in
-comparison to multi-atlas segmentation when using the same number of atlases.
-The actual benefit in using \mb is consistency of the labelling regardless of
-atlas or template choice. This is an important consideration that few have
-touched on previously. The 10-fold Monte Carlo cross validation that we present
-in Experiment 1 is amongst one of the most stringent performed in the
-multi-atlas/segmentation literature.  To the best of our knowledge other groups
-have only ever performed a 3-fold validation \citep{?}.  This suggests that
-our results are reflective of a true average over the choice of parameter
-settings and are independent of atlas or template choice (provided the input
-atlases are properly segmented).
-
-\todo[inline]{could we be seeing a ceiling effect... reaching the limit of what
-is achievable with a multi-atlas-based segmentation method}
-
-In comparison to other methodologies in the field, it is important to note that
-\mb performs quite well.  Table 6 gives a survey of some of the most recent
-multi-atlas implementations with the reported kappa values used to validate the
-algorithm on the ADNI dataset (many of which use the same SNT labels for
-populating their template library and as gold standards for evaluation).  While
-it is difficult to compare segmentation results across studies, gold-standards,
-evaluation metrics, and algorithms it is worth noting that the methods
-summarized in Table 6 require more atlases (between 16-55) than our \mb
-implementation with the Winterburn atlases \citep{Winterburn2013}.  Amongst these
-methods only method 4 yields mean Kappas that are higher than the ones that are
-reported here.
-
-There are some important differences between our method
-and these specific methods.  Others have reported the difficulty with
-mis-registrations in candidate segmentation (i.e. segmentations generated that are
-then input in the voxel-voting procedure \citep{Collins2010}). The work of 
-\citet{Leung2010} tackles this problem by using an intensity threshold
-that is estimated heuristically at the time of segmentation (this work also
-reports some of the highest kappas for the segmentation of ADNI data).  While
-this works for the ADNI dataset (which is partially homogenized with respect to
-image acquisition and pre-processing), it is unclear if this type of heuristic
-is applicable to other datasets.  In all cases, these methods require more
-atlases than our implementation with the Winterburn atlases.
-\citet{Lotjonen2010} achieve highly accurate segmentation but correct their
-segmentations using classifications derived using an expectation maximization
-framework.  In their initial work, \citet{Chupin2009} develop their
-probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy
-patients.  In subsequent methods, they retuned their methodology using the ADNI
-sample with a hierarchical experimentation protocol that involved.  These
-methods suggest that some post-processing of the final segmentations would
-improve accuracy of the segmentation.  While that may be true, there is little
-consensus regarding how to achieve this.
-
-To the best of our knowledge, no other groups have validated their work using
-multiple atlas segmentation protocols, different acquisitions, and disease
-populations in order to demonstrate the robustness of their technique.  This is
-one of the clear strengths of this work.  Furthermore, unlike some of the
-algorithms mentioned, our implementation does not require retuning for new
-populations or datasets as it inherently models the variability of the dataset
-through the template library.  However it should be noted that the increased
-accuracy that follows increasing the number of atlases and templates comes at an
-increased computational cost ($O(log(n))$), as previously mentioned in
-other work \citep{Heckemann2006}. 
-\todo[inline]{time complexity seems suspicious}
-
-In comparison to the methods that we compared and were available through the
-ADNI database (FreeSurfer, MAPER) and the method that we initialized ourselves
-(FSL-FIRST) we find extremely variable performance of all methods.  With the
-exception of FSL all methods correlate well with the SNT volumes provided in
-the ADNI database.  However, FreeSurfer and FIRST provide radically different
-definitions of the size of the hippocampus in comparison to the other methods.
-Further, when estimating bias of these methods relative to SNT hippocampal
-volumes we see that large hippocampi are over estimated while small hippocampi
-are under estimated.  By comparison, our method and MAPER are far more
-conservative suggesting that these methods may be better suited for estimating
-true-positives.  While this comparison with these methods using only volume,
-more work is needed to better understand the differences between our method and
-other methods.
-
-Finally, we have also demonstrated that our algorithmic framework is appropriate
-for the segmentation of hippocampal subfields in standard T1-weighted data.
-This has started to become a burgeoning topic in the segmentation literature,
-although very few methods are available for the segmentation of 3T data
-\citep{Yushkevich2009,VanLeemput2009}. However, the FreeSurfer implementation of
-subfield segmentation. While recent work demonstrates that subfield
-segmentations can be used for classification of AD, MCI, and NC, there has been
-no explicit validation of the methodology based on accuracy or precision.
-Although the initial hippocampal subfield segmentations from the Yushkevich
-group have been demonstrated to work on the ADNI population, there has also been
-no validation of their work.  In addition, their work requires some manual
-initialization to properly function.  Our work demonstrates that we can reliably
-identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
-The fact that CA2/CA3 and molecular layers cannot be reasonably identified
-should not be surprising.  These are extremely thin and spatially convoluted
-regions that originally required high-resolution MRI for identification.  It is
-likely that they would be that extents of these regions are well below the
-resolution offered by standard T1-weighted images.  In fact many manual
-segmentation methodologies do not attempt to parse these regions either
-\citep{Wisse2012, Mueller2009}
-
-\subsection{Conclusion}
-In conclusion, we have presented a flexible multi-atlas framework.  It has
-considerable advantages over other methods as only a small set of atlases is
-required to initialize the algorithm.  We demonstrate that our method works
-robustly over hippocampal definitions, different populations, and different
-acquisition types.  Finally, we also demonstrate that using this method that
-accurate identification of the hippocampal subfields in also possible.
-
-% \begin{verbatim}
-% Experiment 1: 
-% - address the absence of weighted-voting effects. 
-%   - hypothesis: the benefits of weighted voting only outweigh the resampling
-%     error effects when choosing from a large library.  In this experiment, the 
-%     template library (during weighted voting) consists of only 20 templates, so
-%     the most we could hope for is a "squashing" upwards of the curve towards 
-%     the 20-template limit (which is the same across all cases in this experiment).
-%   - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
-%     Perhaps in future experiments with larger template libraries, this could be
-%     explored further.
-% - Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
-%   - couldn't get STAPLE to work with our image formats
-%   - Idea: expect smoothness across range of templates (no even # dips, below)
-%   - perhaps more sophisticated fusion methods would boost results over majority 
-%     vote based techniques.
-% 
-% - why does MAGeT brain show a dip in average Kappa when using an even number
-%   templates. Does MA show this same pattern? (yes)
-%   - Hypothesis: our voting method is biased when breaking ties to choose the 
-%     label with the lowest numeric value.
-%   - If we compute the total number of labels fused (atlases * templates) then 
-%     both MB/MA perform worse when fusing an even number of labels. 
-% 
-% - we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
-%   using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
-%   this increase justify the extra effort involved in MB? 
-%   - decrease in variability
-%   - we are comparing "true" averages (see above)... does this make our criteria 
-%     for a worthy improvement less strict? 
-%     
-% 
-% - how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
-% - because of the extensive cross-validation in experiment 1, we are very likely
-%   showing results that approach the true average of MB and MA on that dataset
-%   (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
-%   each).
-% - even with this caveat, do we think we do well enough?  i.e. other than parameter 
-%   tuning, to what extent does this experiment tell us about how MB would do in 
-%   practice. 
-% 
-% - effects we are seeing is not only averaging effects (Chakravarty et al. 2012). 
-% 
-% - take aways: as in other studies (Aljabar, Heckemann) we find that performance
-%   scales with the number of inputs (approx. log(n)), and that a large enough
-%   template library can boost performance with a small number of atlases. The use 
-%   of templates /can/ have a negative impact on performance (Figure 2), that is 
-%   balanced by the improvement of growing the template library.  Likely the negative
-%   impact on peformance is as a result of resampling/mislabelling error. In other 
-%   words, tuning MAGeT brain involves balancing the tension between an improvement 
-%   in peformance due to increase neuroanatomical capture and decreased peformance
-%   due to resampling error. 
-% 
-% Experiment 2: 
-% - describe resampling error during downsampling (essentially partial volume effect 
-%   due to averaging/majority vote in nearest neighbour selection)
-% - Why would some structures have greater downsampling error than others? (i.e. what
-%   is it about a structure that would make it especially prone to resampling error)
-%   - since error is discretised to whole voxels, smaller regions will show larger
-%     percent error
-%   - average unilateral volume (approx; mm3): 
-%      region   volume   downsampling error
-%      CA1      800            2%
-%      CA4/DG   600          -30%
-%      SR/SL/SM 700           30%
-%      Sibculum 350            0%
-%      CA2/3    200           15%
-% 
-% - observation: when downsampling error is small (<10 percent), MB error is larger, 
-%   and vice versa when downsampling error is large.  Could this a case where the 
-%   inevitable MAGeT resampling error outweighs the small downsampling error?
-% - observation: MAGeT produces similar volumes for the downsampled and BRAVO images
-%   which demonstrates reliability. Additionally, MB's error is in the same *direction*
-%   as the downsampling error except for SR/SL/SM
-% - for MB to show less error than resampling means that voting across templates is 
-%   in aggregate performing better than local nearest neighbour fitting. presumably 
-%   we'd see the same improvement with basic multiatlas as well.
-% 
-% - take away: MB produces subregion volumes that are comparable or better than
-%   resampling error, except for the CA2/3 where error is near 25 percent (but 
-%   even then resampling error is ~15%)
-% 
-% Experiment 3: 
-% - what is an "acceptable" r^2 value? 
-% - take away: MB has proven to be robust with atlases derived from three different
-%   segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
-%   populations (older with AD progression; young and healthy; young and SZ). 
-%   
-% Experiment 4: 
-% - address the smaller difference in mean volume across disease classes. smaller 
-%   than every other method. 
-%   
-% - address the segmentation bias towards over-estimating smaller hippocampi, and 
-%   underestimating larger HC
-%   
-% - visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
-%   is lower than the other methods (but we may be biased. :-). 
-%   
-% - take away: MB, with the Winterburn atlases, produces HC volumes more inline with
-%   SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required. 
-% 
-% \end{verbatim}
-
+%\begin{verbatim}
+
+In this manuscript we have presented the implementation and validation of the MAGeT Brain framework; a methodology that requires very few input atlases in order to provide accurate and reliable segmentations.
+We have demonstrated that our methodology robustly provides accurate and consistent segmentations in populations with different ageing and neuropsychiatric characteristics.
+Further, we have demonstrated that algorithmic performance is not dependent on a single definition of the hippocampus and can easily differing hippocampal definitions \cite{Winterburn,Pruessner,and SNT}.
+Finally, we also demonstrate that MAGeT Brain provides accurate automatic identification of the hippocampal subfields despite contrast and resolution limitations in standard T1-weighted image volumes.
+
+Throughout the cross validation in Experiment #1 (10-fold cross validation in the ADNI sample) we demonstrate that two parameter choices increase segmentation accuracy; namely increasing the number of atlases and the number of templates.
+However after setting the parameters to 5 atlases and 15 templates that there are diminishing returns with respect to this improvement.
+We were somewhat surprised to find that cross-correlation and normalized mutual information based weighted voting did little to improve segmentation accuracy.
+This suggests that using an intermediate template library generated by accurate nonlinear registration methods such as ANTs \cite{Avants|} sufficiently filters sources of error such as nonlinear registration error and partial volume effects that arise through the use of nearest neighbour resampling.
+From our previous work on the MAGeT Brain algorithm we have shown that these increases in error are not simply a smoothing or averaging effect \cite{Chakravarty}.
+Although, the goal of this manuscript was to not exhaustively test or validate multiple different voting strategies in the context of our segmentation algorithm, it is important to note that other strategies for voting are available.
+For example, other groups have used the STAPLE algorithm \cite{Warfield} (or variants of the STAPLE algorithm \cite{stuff from the Duchesne group here}).
+Can then be weighted each segmentation based upon its estimated performance level with respect to the other available candidate segmentations.
+Further, the sensitivity and specificity parameters could also be tuned to potentially improve segmentation accuracy and reliability.
+It is rather likely that using more sophisticated voting methods would have a positive effect on the overall segmentation performance.
+To this end, more work is required in order to determine the source of the slight decrease in segmentation performance when the number of templates are set to an even number.
+Our initial concern was that this dip in performance was a by-product of the MAGeT Brain algorithm itself;  however we determined that this pattern was also true in the analysis of multi-atlas segmentations we used in our experiments.
+We believe that our voting methodology is biased towards labels with the lowest numeric values when breaking ties, thus causing the slight bias observed when using an even number of templates.
+This is another area where the voting scheme could be used to improve performance.
+However, it is worth noting that this limitation was previously identified by Heckemann et al.
+\cite{Heckemann2006a} and, subsequently, other groups have not even considered the case where an even number of candidate labels are considered (eg: \cite{Leung2010}).
+
+Another concern is the moderate-to-small improvement observed in MAGeT Brain in comparison to multi-atlas segmentation when using the same number of atlases.
+The actual benefit in using MAGeT Brain is consistency of the labelling regardless of atlas or template choice.
+This is an important consideration that few have touched on previously.
+The 10-fold Monte Carlo cross validation that we present in Experiment 1 is amongst one of the most stringent performed in the multi-atlas/segmentation literature.
+To the best of our knowledge other groups have only ever performed a 3-fold validation \cite{????}.
+This suggests that our results are reflective of a true average over the choice of parameter settings and are independent of atlas or template choice (provided the input atlases are properly segmented).
+
+In comparison to other methodologies in the field, it is important to note that MAGeT Brain performs quite well.
+Table 6 gives a survey of some of the most recent multi-atlas implementations in the fields and the kappa values that are reported that validate their results using the ADNI dataset, many of which use the same SNT labels for populating their template library and for gold standards for evaluation.
+While it is difficult to compare segmentation results across studies, gold-standards, evaluation metrics, and algorithms it is worth noting that the methods summarized in Table 6 require more atlases (between 16-55) than our MAGeT Brain implementation with the Winterburn atlases \cite{Winterburn}.
+Amongst these methods only 4 yields mean Kappas that are higher than the ones that are reported here.
+There are some important differences between our method and these specific methods.
+Others have reported the difficulty with misregistrations in candidate segmentation (ie: segmentations generated that are then input in the voxel-voting procedure \cite{Collins}.
+The work of Leung \cite{Leung2010} (work also reports some of the highest kappas for the segmentation of ADNI data) tackles this problems by using an intensity threshold that is estimated heuristically at the time of segmentation.
+While this works for the ADNI dataset (which is partially homogenized with respect to image acquisition and pre-processing), it is unclear if this type of heuristic is applicable to other datasets.
+In all cases, these methods require more atlases than our implementation with the Winterburn atlases.
+Lotjonen et al.
+\cite{Lotjonen} achieve highly accurate segmentation but correct their segmentations using classifications derived using an expectation maximization framework.
+In their initial work, Chupin and colleagues develop their probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy patients.
+In subsequent methods, they retuned their methodology using the ADNI sample with a hierarchical experimentation protocol that involved.
+These methods suggest that some post-processing of the final segmentations would improve accuracy of the segmentation.
+While that may be true, there is little consensus regarding how to achieve this.
+Further, to the best of our knowledge, no other groups have validated their work using multiple atlas labels, different acquisitions, and disease populations in order to demonstrate the robustness of their technique.
+This is one of the clear strengths of this work.
+Further, unlike some of the algorithms mentioned our implementation does not require retuning for new populations or datasets as it inherently models the variability of the dataset through the template library.
+However it should be noted that the increased accuracy that follows increasing the number of atlases and templates comes at an increased computational cost (O(log(n), as previously mentioned in other work \cite{Heckemann}).
+
+In comparison to the methods that we compared and were available through the ADNI database (FreeSufer, MAPER) and the method that initialized ourselves (FSL-FIRST) we find extremely variable performance of all methods.
+With the exception of FIRST all methods correlate well with the SNT volumes provided in the ADNI database.
+However, FreeSufer and FIRST provide radically different definitions of the size of the hippocampus in comparison to the other methods.
+Further, when estimating bias of these methods relative to SNT hippocampal volumes we see that large hippocampi are over estimated while small hippocampi are under estimated.
+By comparison, our method and MAPER are far more conservative suggesting that these methods may be better suited for estimating true-positives.
+While this comparison with these methods using only volume, more work is needed to better understand the differences between our method and other methods.
+
+Finally, we have also demonstrated that our algorithmic framework is appropriate for the segmentation of hippocampal subfields in standard T1-weighted data.
+This has started to be come a burgeoning topic in the segmentation literature, although very few methods are available for the segmentation of 3T data \cite{Yushkevich and van Leemput}.
+However, the FreeSufer implementation of subfield segmentation.
+While recent work demonstrates that subfield segmentations can be used for classification of AD, MCI, and NC, there has been no explicit validation of the methodology based on accruacy or precision.
+Although the initial hippocampal subfield segmentations from the Yushkevich group have been demonstrated to work on the ADNI population, there has also been no validation of their work.
+In addition, their work requires some manual initialization to properly function.
+Our work demonstrates that we can reliably identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
+The fact that CA2/CA3 and molecular layers cannot be reasonably identified should not be surprising.
+These are extremely thin and spatially convoluted regions that originally required high-resolution MRI for identification.
+It is likely that they would be that extents of these regions are well below the resolution offered by standard T1-weighted images.
+In fact many manual segmentation methodologies do not attempt to parse these regions either \cite{Wisse, Mueller}
+
+In conclusion, we have presented a flexible multi-atlas framework.
+It has considerable advantages over other methods as only a small set of atlases is required to initialize the algorithm.
+We demonstrate that our method works robustly over hippocampal definitions, different populations, and different acquisition types.
+Finally, we also demonstrate that using this method that accurate identification of the hippocampal subfields in also possible.
+
+Experiment 1: 
+- address the absence of weighted-voting effects. 
+  - hypothesis: the benefits of weighted voting only outweigh the resampling
+    error effects when choosing from a large library.  In this experiment, the 
+    template library (during weighted voting) consists of only 20 templates, so
+    the most we could hope for is a "squashing" upwards of the curve towards 
+    the 20-template limit (which is the same across all cases in this experiment).
+  - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
+    Perhaps in future experiments with larger template libraries, this could be
+    explored further.
+- Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
+  - couldn't get STAPLE to work with our image formats
+  - Idea: expect smoothness across range of templates (no even # dips, below)
+  - perhaps more sophisticated fusion methods would boost results over majority 
+    vote based techniques.
+
+- why does MAGeT brain show a dip in average Kappa when using an even number
+  templates. Does MA show this same pattern? (yes)
+  - Hypothesis: our voting method is biased when breaking ties to choose the 
+    label with the lowest numeric value.
+  - If we compute the total number of labels fused (atlases * templates) then 
+    both MB/MA perform worse when fusing an even number of labels.
+
+- we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
+  using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
+  this increase justify the extra effort involved in MB? 
+  - decrease in variability
+  - we are comparing "true" averages (see above)... does this make our criteria 
+    for a worthy improvement less strict?
+
+- how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
+- because of the extensive cross-validation in experiment 1, we are very likely
+  showing results that approach the true average of MB and MA on that dataset
+  (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
+  each).
+- even with this caveat, do we think we do well enough?  i.e. other than parameter 
+  tuning, to what extent does this experiment tell us about how MB would do in 
+  practice.
+
+- effects we are seeing is not only averaging effects (Chakravarty et al. 2012).
+
+- take aways: as in other studies (Aljabar, Heckemann) we find that performance
+  scales with the number of inputs (approx. log(n)), and that a large enough
+  template library can boost performance with a small number of atlases. The use 
+  of templates /can/ have a negative impact on performance (Figure 2), that is 
+  balanced by the improvement of growing the template library.  Likely the negative
+  impact on peformance is as a result of resampling/mislabelling error. In other 
+  words, tuning MAGeT brain involves balancing the tension between an improvement 
+  in peformance due to increase neuroanatomical capture and decreased peformance
+  due to resampling error.
+
+Experiment 2: 
+- describe resampling error during downsampling (essentially partial volume effect 
+  due to averaging/majority vote in nearest neighbour selection)
+- Why would some structures have greater downsampling error than others? (i.e. what
+  is it about a structure that would make it especially prone to resampling error)
+  - since error is discretised to whole voxels, smaller regions will show larger
+    percent error
+  - average unilateral volume (approx; mm3): 
+     region   volume   downsampling error
+     CA1      800            2%
+     CA4/DG   600          -30%
+     SR/SL/SM 700           30%
+     Sibculum 350            0%
+     CA2/3    200           15%
+
+- observation: when downsampling error is small (<10 percent), MB error is larger, 
+  and vice versa when downsampling error is large.  Could this a case where the 
+  inevitable MAGeT resampling error outweighs the small downsampling error?
+- observation: MAGeT produces similar volumes for the downsampled and BRAVO images
+  which demonstrates reliability. Additionally, MB's error is in the same *direction*
+  as the downsampling error except for SR/SL/SM
+- for MB to show less error than resampling means that voting across templates is 
+  in aggregate performing better than local nearest neighbour fitting. presumably 
+  we'd see the same improvement with basic multiatlas as well.
+
+- take away: MB produces subregion volumes that are comparable or better than
+  resampling error, except for the CA2/3 where error is near 25 percent (but 
+  even then resampling error is ~15%)
+
+Experiment 3: 
+- what is an "acceptable" r^2 value? 
+- take away: MB has proven to be robust with atlases derived from three different
+  segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
+  populations (older with AD progression; young and healthy; young and SZ).
+
+Experiment 4: 
+- address the smaller difference in mean volume across disease classes. smaller 
+  than every other method.
+
+- address the segmentation bias towards over-estimating smaller hippocampi, and 
+  underestimating larger HC
+
+- visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
+  is lower than the other methods (but we may be biased. :-).
+
+- take away: MB, with the Winterburn atlases, produces HC volumes more inline with
+  SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required.
+
+\end{verbatim}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
 
@@
 
-\bibliographystyle{abbrvnat}
+\bibliographystyle{plainnat}
 \bibliography{references}

### 2013-08-07 paper.Rnw
some of aristotle's comments addressed

@@
 \author[1]{Julie Winterburn}
-\author[1]{Tristram A. Lett}
+\author[1,7]{Tristram A. Lett}
 \author[2,3]{Jason P. Lerch}
@@
 \author[4]{Martin Lepage}
-\author[1,5]{Aristotle Voineskos}
+\author[1,5,7]{Aristotle N. Voineskos}
 \author[1,5,6]{M. Mallar Chakravarty}
@@
 ON, Canada}
+\affil[7]{Institute of Medical Science, University of Toronto, Toronto, ON, Canada}
 
@@
 understand the unique neuroanatomy of subjects. Typically, the gold standard for
-neuroanatomical segmentation is manual delineation by an expert human rater. However
-the rapid increase in the availability of MRI data and the time and expertise
-required for manual segmentation is prohibitive
-\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}.
-Further, there is little agreement between researchers regarding how exactly the
-hippocampus should be identified in MRI images \citep{Geuze2004} and this has
-led to efforts to create an unified hippocampal segmentation protocol
-\citep{Jack2011}.
+neuroanatomical segmentation is manual delineation by an expert human rater.
+However with the availability of increasingly large MRI datasets, the time and
+expertise required for manual segmentation becomes prohibitive
+\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}.  Further, there is
+little agreement between researchers regarding how exactly the hippocampus
+should be identified in MRI images \citep{Geuze2004} and this has led to efforts
+to create an unified hippocampal segmentation protocol \citep{Jack2011}.
 
@@
 This involves the selection of a subset of atlases using a similarity metric
-such a cross-correlation \citep{Aljabar2009} or normalized mutual information.
+such as cross-correlation \citep{Aljabar2009} or normalized mutual information.
 Such selection has the added benefit of significantly reducing the number of
@@
 computes a probabilistic segmentation using an expectation maximization
-framework from an set of competing segmentations; or others where a subset of
+framework from an set of competing segmentations, or others where a subset of
 segmentations can be estimated using metrics such as the sum of squared
@@
 However, many of these methods require significant investment of time and
-resources for the creation of the atlas library; ranging from atlas libraries
-that require between 30 \citep{Heckemann2006} and 80 \citep{Collins2010}
-manually segmented atlases.  This strategy has the main drawback of being
-inflexible as it does not easily accommodate varying the definition of the
-hippocampal anatomy (such as the commonly used heuristic of subdividing the
-hippocampus in to head, body, and tail \citep{Poppenk2011,Pruessner2000}).
-Furthermore, none of these methods have demonstrated sufficient flexibility to
-accommodate atlases that are somehow exceptional such as those derived from
-serial histological data \citep{Chakravarty2006,Yelnik2007} or high-resolution MRI
-data that enables robust identification of hippocampal subfields
-\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}.  Due to
-the recent availability of the latter, there has been increased interest in the
-use of probabilistic methods for the identification of the hippocampal subfields
-on standard T1-weighted images.  Our group recently demonstrated that through
-use of an intermediary automated segmentation stage, robust and accurate
+resources for the creation of the atlas library ranging between 30
+\citep{Heckemann2006} and 80 \citep{Collins2010} manually segmented atlases.
+This strategy has the main drawback of being inflexible as it does not easily
+accommodate varying the definition of the hippocampal anatomy (such as the
+commonly used heuristic of subdividing the hippocampus into head, body, and tail
+\citep{Poppenk2011,Pruessner2000}).  Furthermore, none of these methods have
+demonstrated sufficient flexibility to accommodate atlases that are somehow
+exceptional such as those derived from serial histological data
+\citep{Chakravarty2006,Yelnik2007} or high-resolution MRI data that enables
+robust identification of hippocampal subfields
+\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}.  Due
+to the recent availability of the latter, there has been increased interest in
+the use of probabilistic methods for the identification of the hippocampal
+subfields on standard T1-weighted images.  Our group recently demonstrated that
+through use of an intermediary automated segmentation stage, robust and accurate
 segmentation of the striatum, pallidum, and thalamus using a single atlas
@@
 work is limited as it omits the tail of the hippocampus and the segmentation
-protocol has yet to be fully validated.  Nonetheless, they demonstrate that the
-applicability of their work using data from 10 subjects.  In the second method,
-\citet{Yushkevich2009} acquired and labelled hippocampal subfields on
-high-resolution MRI data from post-mortem medial temporal lobe samples. Using
-nonlinear registration guided by manually derived hippocampus masks and specific
-landmarks, they demonstrate accurate parcellation of hippocampal subfields in
-unlabelled MRI volumes.
-
-Here we address the issue of limiting the number of input atlases by tuning our
-algorithm, for segmentation of the entire hippocampus, using a multi-fold
-experiment performed on a subset Alzheimer's Disease Neuroimaging Initiative
-(ADNI) 1 dataset. Based on the parameters we find in our experiment, we validate
-our algorithm using all of the data available in the ADNI Complete 1Yr sample
-and compare our segmentations to the other segmentations that are available
-through the ADNI informatics portal.  To ensure that we have not over-fit our
-parameters to the aging or neurodegenerative brain, we also apply our
-segmentations to a dataset of normal controls and individuals suffering from
-first episode psychosis. Finally, we perform a leave-one-out validation
+protocol has yet to be fully validated.  Nonetheless, it  demonstrates that the
+applicability of hippocampal subfield segmentation using data from 10 subjects.
+In the second method, \citet{Yushkevich2009} acquired and labelled hippocampal
+subfields on high-resolution MRI data from post-mortem medial temporal lobe
+samples. Using nonlinear registration guided by manually derived hippocampus
+masks and specific landmarks, they demonstrate accurate parcellation of
+hippocampal subfields in unlabelled MRI volumes.
+
+In this paper we undertake a thorough validation of the \mb algorithm for
+segmentation of the hippocampus.  First, we address the feasibility and accuracy
+of whole hippocampus segmentation with a limited number of input atlases by
+performing a multi-fold experiment using a subset of the Alzheimer's Disease
+Neuroimaging Initiative (ADNI) 1 dataset. Based on the parameters we find in our
+experiment, we then validate our algorithm using all of the data available in
+the ADNI Complete 1Yr sample and compare our segmentations to the other
+segmentations that are available through the ADNI informatics portal. To ensure
+that we have not over-fit our parameters to the aging or neurodegenerative
+brain, we also apply our segmentations to a dataset of individuals suffering
+from first episode psychosis. Finally, we perform a leave-one-out validation
 experiment to determine if the subfields can be accurately identified using our
@@
 via label propagation). The terms {\em atlas library} and {\em template library} 
-describe any set of such images.Additionally, we will use the term {\em target} 
+describe any set of such images. Additionally, we will use the term {\em target} 
 to refer to an unlabelled image that is undergoing segmentation.
@@
                adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-caption = paste(
-  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
-  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
-  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
+caption = 
+  "\\textbf{ADNI-1 cross-validation subset demographics.} 
+  CN     - Cognitively Normal. 
+  LMCI   - Late-onset Mild Cognitive Impairment. 
+  AD     - Alzheimer's Disease.  
+  Hisp   - Hispanic. 
+  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
+  ADAS   - Alzheimer's Disease Assessment Scale.  
+  MMSE   - Mini-Mental State Examination."
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
@@
 estimating smaller deformations on data blurred with a Gaussian kernel with a
-smaller FWHM. The final transformation is a set of local translations defined on
-a bed of equally spaced nodes that were estimated through the optimization of
-the correlation coefficient.  For the purposes of this work we used the
-regularization parameters optimized in \citet{Robbins2004}, displayed in Table
-\ref{tab:ANIMAL-params}.
+smaller full width at half maximum (FWHM). The final transformation is a set of
+local translations defined on a bed of equally spaced nodes that were estimated
+through the optimization of the correlation coefficient.  For the purposes of
+this work we used the regularization parameters optimized in
+\citet{Robbins2004}, displayed in Table \ref{tab:ANIMAL-params}.
 
@@
 The manual segmentations (SNT) provided as part of the ADNI dataset are used as
-the gold standard to compare with \citep{Hsu2002}.  The segmentation accuracy
-reported is averaged over the ten validation rounds for each parameter setting.
+the gold standard with which to compare \citep{Hsu2002}. The segmentation
+accuracy reported is averaged over the ten validation rounds for each parameter
+setting.
 
 In order to investigate the performance of \mb in a real world setting in which 
-only one set of atlas and template images are used, we explore the variability 
-in label agreement at fixed parameter settings when the choice for atlas and 
-template images is varied. This is achieved by first computing the standard 
-deviation and variance of DSC scores in each block of ten validation rounds per
-subject. The distribution of these statistics across all subjects is then
-compared between \mb and multi-atlas using a Student's t-test. A significant
-difference between distributions is taken to show either a larger or smaller
-level of variability between methods.
+only one size of atlas and template libraries are used, we explore the
+variability in label agreement at fixed parameter settings under varying  
+atlas and template library compositions. This is achieved by first computing the
+standard deviation and variance of DSC scores in each block of ten validation
+rounds per subject. The distribution of these statistics across all subjects is
+then compared between \mb and multi-atlas using a Student's t-test. A
+significant difference between distributions is taken to show either a larger or
+smaller level of variability between methods.
 
@@
                data = master, method="reverse", na.rm = TRUE, test=FALSE)
-caption = paste(
-  "\\textbf{ADNI-1 cross-validation subset demographics.}", 
-  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
-  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
+caption = 
+  "\\textbf{Subfield validation healthy control demographics.}
+  CN     - Cognitively Normal. 
+  LMCI   - Late-onset Mild Cognitive Impairment. 
+  AD     - Alzheimer's Disease.  
+  Hisp   - Hispanic. 
+  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
+  ADAS   - Alzheimer's Disease Assessment Scale.  
+  MMSE   - Mini-Mental State Examination."
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 3: Application to the segmentation first episode
+\subsubsection{Experiment 3: Application to the segmentation of first episode
 schizophrenia patients}
@@
 intervention service at the Douglas Mental Health University Institute in
-Montreal, Canada. People aged 15–30 years from the local catchment area
+Montreal, Canada. People aged 15 to 30 years from the local catchment area
 suffering from either affective or non-affective psychosis who had not taken
@@
 The manual and Winterburn hippocampal segmentation protocols differ slightly in
-the neuroanatomical features that are delineated.  This poses a problem for
-evaluation by measuring overlap.  That is, since different protocols will
-necessarily produce segmentations that do not perfectly overlap, the degree of
-overlap cannot be solely used to compare segmentation methods using different
-protocols. In place of an overlap metric, we can assess the degree of correlation in
-average hippocampal volume of the subjects produced by each method.
-Specifically, Pearson correlation is used.
+the neuroanatomical features that are delineated. This difference poses a
+problem for evaluation by measuring overlap.  That is, since different protocols
+will necessarily produce segmentations that do not perfectly overlap, the degree
+of overlap cannot be solely used to compare segmentation methods using different
+protocols. In place of an overlap metric, we can assess the degree of
+correlation in average hippocampal volume of the subjects produced by each
+method. Specifically, a Pearson correlation is used.
 
@@
 <<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
-caption = paste(
-  "\\textbf{Schizophrenia First Episode Patient Demographics.}",
-  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
-  "State Examination (MMSE).", sep=" ") 
+caption = 
+  "\\textbf{Schizophrenia First Episode Patient Demographics.}
+  ambi - ambidextrous. 
+  SES  - Socioeconomic Status score. 
+  FSIQ - Full Scale IQ.
+"
 fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
@@
 \paragraph{ADNI1 dataset revisited} 
-All images from the {\em \adnidataset} standardized dataset described
-in Experiment 1 are used.  Clinical and demographic data are shown in Table
-\ref{tab:ADNI1-Dataset-Demographics}.
+All images $1919$ images from the {\em \adnidataset} standardized dataset
+described in Experiment 1 are used.  Clinical and demographic data for these
+subjects are shown in Table \ref{tab:ADNI1-Dataset-Demographics}.
+
+Based on the results from the ADNI1 Cross-Validation experiment, in this
+experiment \mb was configured with a template library of 21 randomly chosen
+subject images (7 from each disease class) and used majority vote label fusion.
+The entire \adnidataset dataset was segmented by \mb and we now compare the
+resulting volumes with those obtained by manual segmentation (SNT), and other
+automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
+segmentations available, including a count of those which have failed a quality
+control inspection. Quality control for FreeSurfer segmentations was provided along
+with the segmentations from ADNI \cite{Hartig2010}. One of the authors (MP) 
+performed visual quality inspection for MAGeT and FSL segmentations using similar 
+quality control guidelines (if either hippocampus was under or over segmented
+by greater than 10mm in three or more slices then the segmentation did not
+pass).  Only those images which had quality segmentations from each method are
+included in the following analysis (a total of
+\Sexpr{length(means.complete$RID)} images).
 
@@
 \mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2). All images from the \adnidataset dataset are
-segmented. The optimal size of template library, registration method, and label
-fusion method found in Experiment 1 are used. The template library is composed
-of equal numbers of images from each disease class (AD, MCI, and cognitively
-normal controls).
+(see Experiment 2). All images from the \adnidataset dataset are segmented. The
+optimal size of template library, registration method, and label fusion method
+found in Experiment 1 are used. The template library is composed of 21 randomly
+chosen subject images (7 from each disease class).
 
@@
                method = "reverse", test=FALSE, overall=TRUE)
-caption = paste(
-  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.}", 
-  "Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease",
-  "Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).", sep=" ") 
+caption = 
+  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} 
+  CN     - Cognitively Normal. 
+  LMCI   - Late-onset Mild Cognitive Impairment. 
+  AD     - Alzheimer's Disease.  
+  Hisp   - Hispanic. 
+  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
+  ADAS   - Alzheimer's Disease Assessment Scale.  
+  MMSE   - Mini-Mental State Examination."
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
@@
 quality control guidelines (if either hippocampus was under or over segmented
-by greater than 10mm in three or more slices then the segmentation did not pass).
-
-Only those images which had quality segmentations from each method are 
-included in the following analysis (a total of \Sexpr{length(means.complete$RID)}
-images).
+by greater than 10mm in three or more slices then the segmentation did not
+pass).  Only those images which had quality segmentations from each method are
+included in the following analysis (a total of
+\Sexpr{length(means.complete$RID)} images).
 

### 2013-08-18 paper.Rnw
Completed responses to Aristotle's comments

@@
 
-To validate that \mb algorithm works as well as established automated methods,
-\mb is applied to the ADNI1 dataset and the resulting segmentations are
-compared to those produced by FreeSurfer, FSL, MAPER, and by expert manual
-segmentation.
+To validate the \mb algorithm as respect to other established automated
+methods, \mb is applied to large dataset from the ADNI project, and the
+resulting segmentations are compared to those produced by FreeSurfer, FSL,
+MAPER, and by expert semi-automated segmentation (SNT).
 
 \paragraph{ADNI1 dataset revisited} 
-All images $1919$ images from the {\em \adnidataset} standardized dataset
-described in Experiment 1 are used.  Clinical and demographic data for these
-subjects are shown in Table \ref{tab:ADNI1-Dataset-Demographics}.
-
-Based on the results from the ADNI1 Cross-Validation experiment, in this
-experiment \mb was configured with a template library of 21 randomly chosen
-subject images (7 from each disease class) and used majority vote label fusion.
-The entire \adnidataset dataset was segmented by \mb and we now compare the
-resulting volumes with those obtained by manual segmentation (SNT), and other
-automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
-\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
-segmentations available, including a count of those which have failed a quality
-control inspection. Quality control for FreeSurfer segmentations was provided along
-with the segmentations from ADNI \cite{Hartig2010}. One of the authors (MP) 
-performed visual quality inspection for MAGeT and FSL segmentations using similar 
-quality control guidelines (if either hippocampus was under or over segmented
-by greater than 10mm in three or more slices then the segmentation did not
-pass).  Only those images which had quality segmentations from each method are
-included in the following analysis (a total of
-\Sexpr{length(means.complete$RID)} images).
+The $1919$ images from the {\em \adnidataset} standardized dataset, described
+in Experiment 1, were initially considered.  SNT and FreeSurfer hippocampal
+volumes were provided by ADNI, along with FreeSurfer quality control data
+(guidelines described in \citep{Hartig2010}). Clinical and demographic data for
+the entire \adnidataset dataset are shown in Table
+\ref{tab:ADNI1-Dataset-Demographics}.
 
 \paragraph{Experiment details} 
-\mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2). All images from the \adnidataset dataset are segmented. The
-optimal size of template library, registration method, and label fusion method
-found in Experiment 1 are used. The template library is composed of 21 randomly
-chosen subject images (7 from each disease class).
+\mb was configured with an atlas library composed of the five Winterburn T1
+atlases (described in Experiment 2). The optimal size of template library,
+registration method, and label fusion method found in Experiment 1 were used.
+The FSL FIRST segmentation method was used  via the \verb+run_first_all+ script
+according to the FIRST user guide
+\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
+from the \adnidataset dataset were segmented by both methods.
+
+One author (MP) performed visual quality inspection for MAGeT and FSL
+segmentations using similar quality control guidelines (if either hippocampus
+was under or over segmented by greater than 10mm in three or more slices then
+the segmentation did not pass).  Only images meeting the conditions of having
+segmentations from all methods (SNT, MAPER, FreeSurfer, FSL, and \mb) and also
+passing quality control inspection are included in the analysis.
 
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-This experiment explores \mb segmentation of hippocampal subfields. To
-achieve this, a leave-one-out validation is conducted in which lower-resolution
-images ($0.9mm^3$ voxels) of each Winterburn atlas subject are segmented using 
-the remaining Winterburn atlases.
-
-In general, across hippocampal subregions the percent error in volume of
-\mb segmentations compares
-favourably to the error resulting from image resampling (Figure
-\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
+This experiment explores \mb segmentation of hippocampal subfields. To achieve
+this, a leave-one-out validation was conducted in which lower-resolution images
+($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
+remaining Winterburn atlases.
+
+In general, across hippocampal subregions the percent error in volume of \mb
+segmentations compares favourably to the error resulting from image resampling
+(Figure \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
 subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
@@
 subject images (7 from each disease class) and used majority vote label fusion.
-The entire \adnidataset dataset was segmented by \mb and we now compare the
-resulting volumes with those obtained by manual segmentation (SNT), and other
-automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
-\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
-segmentations available, including a count of those which have failed a quality
-control inspection. Quality control for FreeSurfer segmentations was provided along
-with the segmentations from ADNI \cite{Hartig2010}. One of the authors (MP) 
-performed visual quality inspection for MAGeT and FSL segmentations using similar 
-quality control guidelines (if either hippocampus was under or over segmented
-by greater than 10mm in three or more slices then the segmentation did not
-pass).  Only those images which had quality segmentations from each method are
-included in the following analysis (a total of
-\Sexpr{length(means.complete$RID)} images).
-
-We find a close relationship in total bilateral hippocampal volume between
+The entire \adnidataset dataset was segmented by \mb, and the resulting volumes
+compared with those obtained by expert semi-automated segmentation (SNT), and
+three other automated segmentation techniques: MAPER, FreeSurfer, and FSL.
+Table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total
+count of segmentations available, including a count of those which have failed
+a quality control inspection.  A total of \Sexpr{length(means.complete$RID)}
+images are included in the following analysis, having met quality criteria and
+having segmentations from every method.
+
+We found a close relationship in total bilateral hippocampal volume between
 all methods and manually segmented volumes (Figure \ref{fig:ADNI-volumes-plot}). 
@@
 available.  For example, other groups have used the STAPLE algorithm
-\citep{Warfield2004} (or variants of the STAPLE algorithm \citep{Robitaille2012})
-which weights each segmentation based upon its estimated
-performance level with respect to the other available candidate segmentations.
-Further, the sensitivity and specificity parameters could also be tuned to
-potentially improve segmentation accuracy and reliability.  It is rather likely
-that using more sophisticated voting methods would have a positive effect on the
-overall segmentation performance.
+\citep{Warfield2004} (or variants of the STAPLE algorithm
+\citep{Robitaille2012}) which weights each segmentation based upon its
+estimated performance level with respect to the other available candidate
+segmentations.  Further, the sensitivity and specificity parameters could also
+be tuned to potentially improve segmentation accuracy and reliability.  It is
+likely that using more sophisticated voting methods would have a positive
+effect on the overall segmentation performance, as demonstrated by the STAPLE
+algorithm.
 
 To this end, more work is required in order to determine the source of the
-slight decrease in segmentation performance when the number of templates are set
-to an even number.  Our initial concern was that this dip in performance was a
-by-product of the \mb algorithm itself.  However, we determined that
-this pattern was also true in the analysis of multi-atlas segmentations we used
-in our experiments.  We believe that our majority voting methodology is biased
-towards labels with the lowest numeric values when breaking ties, thus causing
-the slight bias observed when using an even number of templates. This is another
-area where the voting scheme could be used to improve performance. However, it
-is worth noting that this limitation was previously identified by
+slight decrease in segmentation performance when the number of templates are
+set to an even number.  Our initial concern was that this dip in performance
+was a by-product of the \mb algorithm itself.  However, this this pattern is
+also found in the results of the multi-atlas segmentations we used in our
+experiments.  We believe that our majority voting methodology is biased towards
+labels with the lowest numeric values when breaking ties (by way of the
+implementation of the \verb+mode+ function used to determine majority), thus
+causing the slight bias observed when using an even number of templates. This
+is another area where the voting scheme could be used to improve performance.
+However, it is worth noting that this limitation was previously identified by
 \citet{Heckemann2006a} and, subsequently, other groups have not even considered
@@
 multi-atlas/segmentation literature. To the best of our knowledge, other groups
-using ADNI data for validation do at most a single round of leave-one-out-validation
-(Table \ref{tab:other-methods}), aside from \citep{Wang2011} who perform a 
-similar validation as us. The thoroughness of our validation suggests that
-our results are reflective of a true average over the choice of parameter
-settings and are independent of atlas or template choice (provided the input
-atlases are properly segmented).
+using ADNI data for validation do at most a single round of
+leave-one-out-validation (Table \ref{tab:other-methods}), aside from
+\citep{Wang2011} who perform a similar validation as us). The thoroughness of
+our validation suggests that our results are reflective of a true average over
+the choice of parameter settings and are independent of atlas or template
+choice (provided the input atlases are properly segmented.
 
@@
 probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy
-patients.  In subsequent methods, they retuned their methodology using the ADNI
-sample with a hierarchical experimentation protocol that involved.  These
-methods suggest that some post-processing of the final segmentations would
-improve accuracy of the segmentation.  While that may be true, there is little
-consensus regarding how to achieve this.
+patients, and then use this method to segment an ADNI sample, with a
+hierarchical experimentation protocol that involved \todo{complete this
+sentence}.  These methods suggest that some post-processing of the final
+segmentations would improve accuracy of the segmentation. While that may be
+true, there is little consensus regarding how to achieve this.
 
@@
 volumes we see that large hippocampi are over estimated while small hippocampi
-are under estimated.  By comparison, our method and MAPER are far more
-conservative suggesting that these methods may be better suited for estimating
-true-positives.  While this comparison with these methods using only volume,
-more work is needed to better understand the differences between our method and
-other methods.
-
-Finally, we have also demonstrated that our algorithmic framework is appropriate
-for the segmentation of hippocampal subfields in standard T1-weighted data.
-This has started to become a burgeoning topic in the segmentation literature,
-although very few methods are available for the segmentation of 3T data
-\citep{Yushkevich2009,VanLeemput2009}. However, the FreeSurfer implementation of
-subfield segmentation. While recent work demonstrates that subfield
-segmentations can be used for classification of AD, MCI, and NC, there has been
-no explicit validation of the methodology based on accuracy or precision.
-Although the initial hippocampal subfield segmentations from the Yushkevich
-group have been demonstrated to work on the ADNI population, there has also been
-no validation of their work.  In addition, their work requires some manual
-initialization to properly function.  Our work demonstrates that we can reliably
-identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
-The fact that CA2/CA3 and molecular layers cannot be reasonably identified
-should not be surprising.  These are extremely thin and spatially convoluted
-regions that originally required high-resolution MRI for identification.  It is
-likely that they would be that extents of these regions are well below the
-resolution offered by standard T1-weighted images.  In fact many manual
-segmentation methodologies do not attempt to parse these regions either
-\citep{Wisse2012, Mueller2009}
-
-\subsection{Conclusion}
+are under estimated. By comparison, \mb and MAPER are far more conservative in
+volume estimation, suggesting these methods may be better suited for estimating
+true-positives. However, in this analysis we have only compared methods by
+total hippocampal volume, and so more work is needed to understand the full
+extent to which these methods differ.
+
+Finally, we have also demonstrated that our algorithmic framework is
+appropriate for the segmentation of hippocampal subfields in standard
+T1-weighted data.  Subfield segmentation is a burgeoning topic in the
+literature although very few automated methods are available for the
+segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. However, the
+FreeSurfer implementation of subfield segmentation \todo{complete}. While
+recent work demonstrates that subfield segmentations can be used for
+classification of AD, MCI, and NC, there has been no explicit validation of the
+methodology based on accuracy or precision.  Although the initial hippocampal
+subfield segmentations from the Yushkevich group have been demonstrated to work
+on the ADNI population, there has also been no validation of their work.  In
+addition, their work requires some manual initialization to properly function.
+Our work demonstrates that we can reliably identify the CA1, subiculum, and CA4
+dentate with only modest amounts of error.  The fact that CA2/CA3 and molecular
+layers cannot be reasonably identified should not be surprising as these are
+extremely thin and spatially convoluted regions that originally required
+high-resolution MRI for identification.  It is likely that extents of these
+regions are well below the resolution offered by standard T1-weighted images
+and, in fact, many manual segmentation methodologies do not attempt to parse
+these regions either \citep{Wisse2012, Mueller2009}.
+
 In conclusion, we have presented a flexible multi-atlas framework.  It has

### 2013-08-20 paper.Rnw
Half of Matt's comments responded to

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% TODO: 
+%  - check tense: we are is have has 
+%  - update montages with TAL coordinates
+%  - add MAGeT brain diagram, and remove pseudocode
+%  - check that \gs is used instead of Manual 
 % Sweave/Knitr init
@@
 \usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
-\usepackage{titlesec, siunitx}
+\usepackage{titlesec, siunitx, graphicx}
 \usepackage[round,authoryear]{natbib}
@@
 \newcommand{\FEPdataset}{SZ-FEP }
+\newcommand{\gs}{Gold Standard }
 
@@
 successfully in several contexts
-\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
-in accuracy by the introduction of errors due inaccuracies in the nonlinear
-transformation itself, partial volume effects in label resampling, and
-irreconcilable differences between the neuroanatomy represented within the atlas
-and target images.
+\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited in
+accuracy due to inaccuracies in the estimated nonlinear transformation itself,
+partial volume effects in label resampling, and irreconcilable differences
+between the neuroanatomy represented within the atlas and target images.
 
@@
 atlas). A {\it label fusion} technique, such as voxel-wise voting, is used to
-merge these labels into a definitive segmentation for the target.  In addition,
+merge these labels into the definitive segmentation for the target. In addition,
 weighted voting procedures that use {\em atlas selection} techniques are often
@@
 Such selection has the added benefit of significantly reducing the number of
-nonlinear registrations.  For example \citet{Collins2010}
-demonstrated that only 14 atlases, selected based on highest similarity between
-medial temporal lobe neuroanatomy as evaluated by normalized mutual information
-\citep{Studholme1999} from a library of 80 atlases, were required to achieve
-accurate segmentations of the hippocampus.  Several methods have been explored
-for label fusion including the STAPLE algorithm \citep{Warfield2004} that
-computes a probabilistic segmentation using an expectation maximization
-framework from an set of competing segmentations, or others where a subset of
-segmentations can be estimated using metrics such as the sum of squared
-differences in the regions of interest to be segmented \citep{Coupe2011}.
+nonlinear registrations. For example \citet{Collins2010} demonstrated that only
+14 atlases, selected based on highest similarity between medial temporal lobe
+neuroanatomy as evaluated by normalized mutual information \citep{Studholme1999}
+from a library of 80 atlases, were required to achieve accurate segmentations of
+the hippocampus. Additionally, several methods have been explored for label
+fusion, including the STAPLE algorithm \citep{Warfield2004} that computes a
+probabilistic segmentation using an expectation maximization framework from an
+set of competing segmentations, or others where a subset of segmentations can be
+estimated using metrics such as the sum of squared differences in the regions of
+interest to be segmented \citep{Coupe2011}.
 
@@
 derived from serial histological data is possible \citep{MallarChakravarty2012}.
-The novelty of this manuscript is the extension our multi-atlas methodology to
-the hippocampus using more than a single input atlas, while simultaneously
+The novelty of this manuscript is the extension of our multi-atlas methodology
+to the hippocampus using more than a single input atlas, while simultaneously
 limiting the number of possible inputs used during segmentation, and
@@
 successively incorporating unlabelled target images after themselves being
-labelling using multi-atlas techniques. The sequence in which target images are
+labelled using multi-atlas techniques. The sequence in which target images are
 labelled is chosen so that the similarity between the atlas images and the
@@
 To the best of our knowledge there are two other segmentation methods that
-attempt to define the hippocampal subfields using standard T1-weighted data.
-The first is included with the FreeSurfer package \citep{VanLeemput2009}.  This
-work is limited as it omits the tail of the hippocampus and the segmentation
-protocol has yet to be fully validated.  Nonetheless, it  demonstrates that the
-applicability of hippocampal subfield segmentation using data from 10 subjects.
-In the second method, \citet{Yushkevich2009} acquired and labelled hippocampal
-subfields on high-resolution MRI data from post-mortem medial temporal lobe
-samples. Using nonlinear registration guided by manually derived hippocampus
-masks and specific landmarks, they demonstrate accurate parcellation of
-hippocampal subfields in unlabelled MRI volumes.
+attempt to define the hippocampus and its subfields using standard T1-weighted
+data. The first is included with the FreeSurfer package \citep{VanLeemput2009}.
+This work is limited as it omits the tail of the hippocampus and the
+segmentation protocol has yet to be fully validated. Nonetheless, it
+demonstrates that the applicability of hippocampal subfield segmentation using
+data from 10 subjects. In the second method, \citet{Yushkevich2009} acquired
+and labelled hippocampal subfields on high-resolution MRI data from post-mortem
+medial temporal lobe samples. Using nonlinear registration guided by manually
+derived hippocampus masks and specific landmarks, they demonstrate accurate
+parcellation of hippocampal subfields in unlabelled MRI volumes.
 
@@
 Second, the labels from each image are propagated to the target image space. 
-Third, the labels are combined into a single labelling by way of a label fusion method 
-\citep{Heckemann2006, Heckemann2011}. This method is described in detail in other
-publications \citep{Collins2010,Heckemann2011,Aljabar2009}.
+Third, the labels are combined into a single label by label fusion
+\citep{Heckemann2006, Heckemann2011}. This method is described in detail in
+other publications \citep{Collins2010,Heckemann2011,Aljabar2009}.
 
@@
 target set as a whole (for instance, by sampling equally from cases and controls). 
-The template library images are then labelled by each of the atlases. Basic 
-multi-atlas segmentation is then conducted using the template library
-to segment the entire set of target images (including the targets whose images
-are used in the construction of the template library).  Since each template
-library image has multiple labels (one from each atlas), the final number of 
-labels to be fused for each target may be quite large (i.e. \# of atlas $\times$ 
-\# of templates).
+The template library images are automatically labelled by each of the atlases
+via label propagation. Basic multi-atlas segmentation is then conducted using
+the template library to segment the entire set of target images (including the
+targets whose images are used in the construction of the template library).
+Since each template library image has multiple labels (one from each atlas), the
+final number of labels to be fused for each target may be quite large (i.e. \#
+of atlas $\times$ \# of templates).
 
@@
 
-The following section describes the experiments we conducted to assess the
+The following section describes for experiments conducted to assess the
 segmentation quality of the \mb algorithm. The first two experiments assess the
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
+of images and SNT \gs segmentations from the ADNI1 dataset. This form of
 cross-validation allows us to rigorously validate a large number of parameter
@@
 
-\paragraph{ADNI1 dataset}
-
+\paragraph{\adnidataset dataset} 
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
@@
 Alzheimer's disease (AD). Demographics for this subset are shown in Table
-\ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
-right whole hippocampi are available \citep{Hsu2002}. These labels have been
-generated using the SNT tool from Medtronic Surgical Navigation Technologies,
-Louisville, CO (see Supplementary Materials for detailed discussion of the
-manual segmentation process used).
+\ref{tab:ADNI1-xvalidation-demographics}.  Semi-automated segmentations of the
+left and right whole hippocampi are made available with a subset of ADNI images
+\citep{Hsu2002}. These labels have been generated using the SNT tool from
+Medtronic Surgical Navigation Technologies, Louisville, CO (see Supplementary
+Materials for detailed discussion of the segmentation process used). Throughout
+this paper, we will use these segmentations as the \gs for comparison.
 
@@
 authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset download was the "\adnidataset"
-standardized dataset available from ADNI \footnote{
+August 2012. The image dataset used was the "\adnidataset" standardized dataset
+available from ADNI \footnote{
 \url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
@@
 Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR
-= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
-x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
+2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240 x
+240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
 yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.
@@
 
-In this experiment, our dataset consists of 69 1.5T images and corresponding
-manual segmentations. In each validation round, the dataset is partitioned into 
-a training set consisting of images and their manual labels to be used as an atlas 
-library, and a validation set consisting of the remaining images segmented by both 
-\mb and multi-atlas. The resulting segmentations are compared to the manual 
-segmentations for the images.
+In this experiment, our dataset consists of 69 1.5T images and corresponding \gs
+segmentations. In each validation round, the dataset is partitioned into a
+training set consisting of images and their \gs segmentations to be used as an
+atlas library, and a validation set consisting of the remaining images segmented
+by both \mb and multi-atlas. The computed segmentations are compared to the
+\gs segmentations (see below).
 
@@
 Before registration, all images underwent preprocessing with the N3 algorithm
-\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we use
-one of two non-linear image registration methods.
+\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we
+compared two non-linear image registration methods:
 
@@
 
-The \animal algorithm carries out image registration is in two phases. In the
+The \animal algorithm carries out image registration in two phases. In the
 first, a 12-parameter linear transformation (3 translations, rotations, scales,
@@
   output voxel is given the most frequent label at that voxel location amongst
-  all candidate labellings.  Ties are broken arbitrarily.
+  all candidate labellings.
 
@@
 \end{description}
+
 \paragraph{Evaluation method}
 
-The Dice similarity coefficient (DSC) assesses the agreement between two
-segmentations. It is one of the most widely used measures of segmentation
-performance, and we use it as the basis of comparison in this experiment.
+The Dice similarity coefficient (DSC), also known as Dice's Kappa, assesses the
+agreement between two segmentations. It is one of the most widely used measures
+of segmentation accuracy, and we use it as the basis of comparison in this
+experiment.
 %Additionally, we report the Jaccard index, another commonly used similarity
@@
 
-The manual segmentations (SNT) provided as part of the ADNI dataset are used as
-the gold standard with which to compare \citep{Hsu2002}. The segmentation
-accuracy reported is averaged over the ten validation rounds for each parameter
-setting.
+The automatically produces labels are compared to the \gs SNT labels provided by
+ADNI. The segmentation accuracy reported for a parameter setting is the average
+over ten validation rounds.
 
@@
 In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
-segmentation is tested using a leave-one-out cross-validation (LOOCV) design. 
-The optimal parameter settings for \mb found in Experiment 1 are used in this 
-experiment.
+segmentation is tested using a leave-one-out cross-validation (LOOCV) design on
+two image sets. The optimal parameter settings for \mb found in Experiment 1 are
+used in this experiment.
 
@@
 In addition to the high-resolution scans distributed as part of the Winterburn
-atlases, we also obtained additional 3T T1 BRAVO images (0.9mm-isotropic voxels)
-of four of the five Winterburn atlas subjects.
+atlases, we also obtained separate 3T T1 BRAVO acquisitions (0.9mm-isotropic
+voxels) of four of the five Winterburn atlas subjects.
 
@@
 Leave-one-out cross-validation (LOOCV) is an approach in which the method to be
-validated is given all but one item in a dataset as training data, and the
-output is compared with the left-out item. This is done, in turn, for each item
-in the dataset.
-
-In this experiment the Winterburn atlases are used as the cross-validation
-dataset.  The five $300\mu$-isotropic voxel images and labels are used as
-atlases, and LOOCV is conducted using the five Winterburn atlas subject images
-subsampled (using trilinear interpolation) to 0.9mm-isotropic voxel resolution
-(referred to as the {\em Subsampled} dataset) as input subjects. Subsampling the
-subject images allows us to assess \mb in a typical segmentation scenario
-(high-resolution atlases and lower-resolution subjects).  The template library
-is composed of the subject images, plus an additional set of 3T T1 images
-(0.9mm-isotropic voxels) of healthy subjects acquired separately (Table
-\ref{tab:WAval-healthy-demographics}). The optimal size of template library,
-registration method, and label fusion method found in Experiment 1 are used.
-Each resampled Winterburn atlas subject image is segmented by \mb with that
-subject's image excluded from the atlas library.
+validated is given all but one item in a dataset as training data (in our case,
+atlas library images and labels), and the output accuracy is measured on the
+left-out item. This is done, in turn, for each item in the dataset.
+
+\begin{figure}
+  \begin{centering}
+    \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
+  \end{centering}
+  \caption{A schematic illustration of the experimental set up of the
+  leave-one-out cross-validation (LOOCV) subfield experiment.}
+  \label{fig:LOOCV-schematic}
+\end{figure}
+
+In part 2a of this experiment, we have modified the LOOCV experimental
+design in order to compare \mb segmentation accuracy with the accuracy of image
+resampling error. The five $300\mu$-isotropic voxel Winterburn atlas images and
+labels are used as the atlas library, and LOOCV is conducted using the five Winterburn
+atlas images subsampled (using trilinear interpolation) to
+0.9mm-isotropic voxel resolution (referred to as the {\em Subsampled} dataset)
+as input subjects. Subsampling the subject images allows us to assess \mb in a
+typical segmentation scenario (high-resolution atlases and lower-resolution
+subjects).  The template library is composed of all five subject images, plus an
+additional set of 3T T1 images (0.9mm-isotropic voxels) of healthy subjects
+acquired separately (Table \ref{tab:WAval-healthy-demographics}). The optimal
+size of template library, registration method, and label fusion method found in
+Experiment 1 are used.  As with traditional LOOCV, each subject image is
+segmented by \mb with that subject's image excluded from the atlas library.  In
+part 2b of this experiment, we carry out a similar validation in which the
+subject images are separately acquired T1 BRAVO images of four of the five
+Winterburn atlas subjects (referred to as the {\em BRAVO} dataset).  Figure
+\ref{fig:LOOCV-schematic} illustrates the experimental set up of both parts of
+this experiment.
 
@@
 
-We reproduce this experiment in a separate LOOCV run in which the resampled 
-Winterburn atlas subject images are substituted for separately acquired T1 
-BRAVO images of four of the subjects (referred to as the {\em BRAVO} dataset).
-
 \paragraph{Evaluation method}  
@@
 
-To validate that \mb algorithm works effectively in the context of other
-neuropsychiatric disorders, we use the Winterburn atlases with \mb to predict
-the hippocampal segmentation of dataset of Schizophrenia patient MR images. The
-resulting segmentations are assessed for quality by comparison with expert
-manual segmentations.
+To validate that \mb works effectively in the context of other neuropsychiatric
+disorders, in this experiment we use the Winterburn atlases to predict the
+hippocampal segmentations of patients with Schizophrenia. The resulting
+segmentations are assessed for quality by comparison with expert manual
+segmentations.
 
@@
 
-Each subject hippocampus is traced following a validated segmentation protocol 
-\citep{Pruessner2000}.
+Manual segmentation of each subject hippocampus is produced following a
+validated segmentation protocol \citep{Pruessner2000}.
 
@@
 The manual and Winterburn hippocampal segmentation protocols differ slightly in
-the neuroanatomical features that are delineated. This difference poses a
-problem for evaluation by measuring overlap.  That is, since different protocols
-will necessarily produce segmentations that do not perfectly overlap, the degree
-of overlap cannot be solely used to compare segmentation methods using different
-protocols. In place of an overlap metric, we can assess the degree of
-correlation in average hippocampal volume of the subjects produced by each
-method. Specifically, a Pearson correlation is used.
+the neuroanatomical features that are delineated \citep{Winterburn2013}. This
+difference poses a problem for evaluation by measuring overlap.  That is, since
+different protocols will necessarily produce segmentations that do not perfectly
+overlap, the degree of overlap cannot be solely used to compare segmentation
+methods using different protocols. In place of an overlap metric, we can assess
+the degree of correlation in average hippocampal volume of the subjects produced
+by each method. Specifically, a Pearson correlation is used.
 
@@
 
-To validate the \mb algorithm as respect to other established automated
-methods, \mb is applied to large dataset from the ADNI project, and the
+To validate the \mb algorithm with respect to other established automated
+methods, \mb is applied to large dataset from the ADNI project and the
 resulting segmentations are compared to those produced by FreeSurfer, FSL,
-MAPER, and by expert semi-automated segmentation (SNT).
+MAPER and by expert semi-automated manual segmentations (SNT).
 
@@
 \paragraph{Evaluation method}
-As in Experiment 3, the manual (SNT) and Winterburn hippocampal segmentation
+As in Experiment 3, the SNT \gs and Winterburn hippocampal segmentation
 protocols differ in the neuroanatomical features delineated, and so we must
 assess \mb by the degree of correlation of average hippocampal volume across all
-subjects produced by \mb and by manual segmentation.  Specifically, Pearson
-correlation is used.  For comparison, we also compute the correlation in
-hippocampal volume between the existing, established automated segmentation
-methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
+subjects. Specifically, Pearson correlation is used. We also compute the
+correlation in hippocampal volume between the existing, established automated
+segmentation methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
 volume-related fixed and proportional biases in all segmentation methods using
@@
 @
-  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with manual
-  segmentations (SNT) for 69 ADNI1 subjects vs. atlas and template library size,
-  registration algorithm, and label fusion method. Points above zero indicate an
-  \mb parameter settings yielding a higher mean DSC score than multi-atlas
-  segmentation using the same number of atlases. \label{fig:ADNI1-xval-k-mean}}
+  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with 
+  SNT \gs segmentations for 69 ADNI1 subjects vs. atlas and template library
+  size, registration algorithm, and label fusion method. Points above zero
+  indicate an \mb parameter settings yielding a higher mean DSC score than
+  multi-atlas segmentation using the same number of atlases.
+  \label{fig:ADNI1-xval-k-mean}}
 \end{figure}
@@
   \caption{Difference in mean Dice's Similarity Coefficient of \mb and
-  multi-atlas segmentations with manual (SNT) segmentations for a range of atlas
+  multi-atlas segmentations with \gs (SNT) segmentations for a range of atlas
   and template library sizes. \label{fig:ADNI1-xval-k-diff}}
@@
 21 subject images selected at random.  Expert manual whole hippocampal
-segmentations are used as gold standards.
+segmentations are used as a gold standard.
 
@@
 
-We found a close relationship in total bilateral hippocampal volume between
-all methods and manually segmented volumes (Figure \ref{fig:ADNI-volumes-plot}). 
-Volumes are correlated with Pearson $r > 0.78$ for all methods across disease
-categories.  Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}),
-\mb is consistently well correlated to manual volumes (Pearson $r > 0.85$), but
-appears to slightly over-estimate the volume of the AD hippocampus.
-
-To investigate the level of agreement with manually segmented hippocampal
+We found a close relationship in total bilateral hippocampal volume between all
+methods and SNT \gs volumes (Figure \ref{fig:ADNI-volumes-plot}).  Volumes are
+correlated with Pearson $r > 0.78$ for all methods across disease categories.
+Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}), \mb is
+consistently well correlated to manual volumes (Pearson $r > 0.85$), but appears
+to slightly over-estimate the volume of the AD hippocampus.
+
+To investigate the level of agreement with SNT \gs hippocampal
 volumes, we constructed Bland-Altman plots for each method (Figure
@@
 Figure \ref{fig:ADNI-subfields} shows a qualitative comparison of \mb and 
-manual (SNT) hippocampal segmentations for 10 randomly selected subjects in 
+SNT \gs hippocampal segmentations for 10 randomly selected subjects in 
 each disease category, and illustrates some of the common errors found during 
@@
 hippocampal sulcus and, although not anatomically incorrect, \mb under-estimates 
-the hippocampal body in comparison to the manual (SNT) segmentation.
+the hippocampal body in comparison to the SNT \gs segmentation.
 
@@
 m2 = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
-levels(m2$Method) = c("FS"= "FreeSurfer", "FSL" = "FSL", "MAPER" = "MAPER", "MAGeT" = "MAGeT", "SNT" = "Manual")
+levels(m2$Method) = c("FS"    = "FreeSurfer", 
+                      "FSL"   = "FSL", 
+                      "MAPER" = "MAPER", 
+                      "MAGeT" = "MAGeT", 
+                      "SNT"   = "SNT \\gs")
 m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
@@
   \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by FreeSurfer, FSL, MAPER, \mb, and manual raters (SNT) vs. disease
+dataset by FreeSurfer, FSL, MAPER, \mb, and SNT \gs vs. disease
 category.}
@@
     geom_smooth(method="lm") + 
-    xlab(expression("Manual hippocampal volume " (mm^3))) + 
+    xlab(expression("SNT \\gs hippocampal volume " (mm^3))) + 
     ylab(expression("Automated hippocampal volume " (mm^3))) +
@@
 dataset by each of the four automated methods investigated (FreeSurfer (FS),
-FSL, MAPER, \mb) vs. manual rating (SNT). Linear fit lines and pearson
+FSL, MAPER, \mb) vs. SNT \gs. Linear fit lines and pearson
 correlations are shown for each method.}
@@
             hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of manual and automated hippocampal volume ' (mm^3))) + 
+  xlab(expression('Mean of SNT \\gs and automated hippocampal volume ' (mm^3))) + 
   ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
@@
   \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset dataset by manual raters (SNT) and each of the four
-automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall mean
-difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
-horizontal lines. Linear fit lines are shown for each diagnosis group. Note,
-points below the mean difference indicate overestimation of the volume with
-respect to the manual rating, and vice versa. }
+measured in the \adnidataset dataset by SNT \gs segmentation and each of the
+four automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall
+mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
+dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
+Note, points below the mean difference indicate overestimation of the volume
+with respect to the \gs rating, and vice versa. }
 
@@
   \end{centering}
-  \caption{Manual and MAGeT segmentation results for 30 ADNI1 subjects (10
+  \caption{SNT \\gs and MAGeT segmentations for 30 ADNI1 subjects (10
   subjects randomly selected from each disease category in the subject pool used
   in Experiment 1). Sagittal slices are shown for subject unlabelled T1-weighted
-  anatomical image, SNT manual label (green), and \mb label (blue). Noted are
+  anatomical image, SNT \gs label (green), and \mb label (blue). Noted are
   examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
-  {\em (b)} translated manual segmentation by SNT; 
+  {\em (b)} translated SNT \gs segmentation; 
   {\em (c)} under-estimation of hippocampal body and
@@
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
-   ADNI dataset.} For each method, the number of manually labelled atlases used for 
-   training, the best Dice's overlap measure, the disease classes measured, and 
-   the validation procedure are shown. Unless specified, validation datasets are
-   composed equally of subjects diagnosed with Alzheimer's Disease (AD), Mild 
-   Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV = Leave-one-out 
-   cross-validation. 
+\caption{\textbf{Automated segmentation accuracy (overlap with SNT \gs labels)
+of the ADNI dataset.} For each method, the number of labelled atlases
+used for training, the best Dice's overlap measure, the disease classes
+measured, and the validation procedure are shown. Unless specified, validation
+datasets are composed equally of subjects diagnosed with Alzheimer's Disease
+(AD), Mild Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV =
+Leave-one-out cross-validation. 
    Some studies of automated segmentation of ADNI images are excluded because they 
@@
 
-\subsection{ADNI Manual Labels}
+\subsection{ADNI SNT Hippocampal Labels}
 % The following blurb is taken (except for the first sentence) verbatim 

### 2013-08-21 paper.Rnw
A few tense changes

@@
 atlas library, and a validation set consisting of the remaining images segmented
-by both \mb and multi-atlas. The computed segmentations are compared to the
+by both \mb and multi-atlas. The computed segmentations were compared to the
 \gs segmentations (see below).
 
-A total of ten validation rounds are performed on each subject in the dataset,
+A total of ten validation rounds were performed on each subject in the dataset,
 over each combination of parameter settings. The parameter settings
-we explore are: atlas library size (1-9), template library size (1-20),
+explored are: atlas library size (1-9), template library size (1-20),
 registration method (\ants or \animal), and label fusion method (majority vote,
@@
 
-To validate the \mb algorithm with respect to other established automated
-methods, \mb is applied to large dataset from the ADNI project and the
-resulting segmentations are compared to those produced by FreeSurfer, FSL,
-MAPER and by expert semi-automated manual segmentations (SNT).
+To validate \mb segmentation quality with respect to other established automated
+methods, \mb is applied to large dataset from the ADNI project and the resulting
+segmentations are compared to those produced by FreeSurfer, FSL, MAPER and by
+expert semi-automated manual segmentations (SNT).
 
 \paragraph{ADNI1 dataset revisited} 
-The $1919$ images from the {\em \adnidataset} standardized dataset, described
-in Experiment 1, were initially considered.  SNT and FreeSurfer hippocampal
-volumes were provided by ADNI, along with FreeSurfer quality control data
+The {\em \adnidataset} standardized dataset contains $1919$ images in total.
+SNT and FreeSurfer hippocampal volumes for a subset of images are provided by
+ADNI, along with quality control data for each FreeSurfer segmentation
 (guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
@@
 \mb was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 2). The optimal size of template library,
+atlases (described in Experiment 2), and the size of template library,
 registration method, and label fusion method found in Experiment 1 were used.
@@
 according to the FIRST user guide
-\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
-from the \adnidataset dataset were segmented by both methods.
+\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images in
+the \adnidataset dataset were segmented by both methods.
 

### 2013-08-22 paper.Rnw
A few small changes based on Matt's feedback

@@
 %  - add MAGeT brain diagram, and remove pseudocode
-%  - check that \gs is used instead of Manual 
 % Sweave/Knitr init
@@
 
-To validate \mb segmentation quality with respect to other established automated
-methods, \mb is applied to large dataset from the ADNI project and the resulting
-segmentations are compared to those produced by FreeSurfer, FSL, MAPER and by
-expert semi-automated manual segmentations (SNT).
+To validate \mb segmentation quality with respect to other established
+automated methods, \mb was applied to large dataset from the ADNI project and
+the resulting segmentations were compared to those produced by FreeSurfer, FSL,
+MAPER and the SNT \gs segmentations.
 
@@
 The {\em \adnidataset} standardized dataset contains $1919$ images in total.
-SNT and FreeSurfer hippocampal volumes for a subset of images are provided by
+SNT and FreeSurfer hippocampal volumes for a subset of images were provided by
 ADNI, along with quality control data for each FreeSurfer segmentation
 (guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
-the entire \adnidataset dataset are shown in Table
-\ref{tab:ADNI1-Dataset-Demographics}.
+the entire \adnidataset dataset are shown in Table \ref{tab:ADNI1-Dataset-Demographics}.
 
@@
 \mb was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 2), and the size of template library,
-registration method, and label fusion method found in Experiment 1 were used.
-The FSL FIRST segmentation method was used via the \verb+run_first_all+ script
-according to the FIRST user guide
-\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images in
-the \adnidataset dataset were segmented by both methods.
+atlases (described in Experiment 2), and size of template library,
+registration method, and label fusion method were determined by the optimal
+settings found in Experiment 1.  The FSL FIRST segmentation method was used via
+the \verb+run_first_all+ script according to the FIRST user guide
+\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
+in the \adnidataset dataset were segmented by both methods.
 
@@
 segmentations from all methods (SNT, MAPER, FreeSurfer, FSL, and \mb) and also
-passing quality control inspection are included in the analysis.
+passing quality control inspection were included in the analysis.
 
@@
 assess \mb by the degree of correlation of average hippocampal volume across all
-subjects. Specifically, Pearson correlation is used. We also compute the
+subjects. Specifically, a Pearson correlation is used. We also compute the
 correlation in hippocampal volume between the existing, established automated
@@
 atlas or template choice. This is an important consideration that few have
-touched on previously. The 10-fold Monte Carlo leave-one-out cross-validation that we present
-in Experiment 1 is amongst one of the most stringent performed in the
-multi-atlas/segmentation literature. To the best of our knowledge, other groups
-using ADNI data for validation do at most a single round of
+touched on previously. The 10-fold Monte Carlo leave-one-out cross-validation
+that we present in Experiment 1 is amongst one of the most stringent performed
+in the multi-atlas/segmentation literature. To the best of our knowledge, other
+groups using ADNI data for validation do at most a single round of
 leave-one-out-validation (Table \ref{tab:other-methods}), aside from

### 2013-08-23 paper.Rnw
Fully responded to Matt's comments

@@
 \newcommand{\gs}{Gold Standard }
+\newcommand{\fsl}{FSL }
+\newcommand{\freesurfer}{FreeSurfer }
+\newcommand{\maper}{MAPER }
 
@@
 \paragraph{Evaluation method}  
-
 The Dice similarity coefficient (DSC), also known as Dice's Kappa, assesses the
@@
 \paragraph{Evaluation method}
-The manual and Winterburn hippocampal segmentation protocols differ slightly in
+The Pruessner and Winterburn hippocampal segmentation protocols differ slightly in
 the neuroanatomical features that are delineated \citep{Winterburn2013}. This
@@
 As in Experiment 3, the SNT \gs and Winterburn hippocampal segmentation
-protocols differ in the neuroanatomical features delineated, and so we must
-assess \mb by the degree of correlation of average hippocampal volume across all
-subjects. Specifically, a Pearson correlation is used. We also compute the
-correlation in hippocampal volume between the existing, established automated
-segmentation methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
-volume-related fixed and proportional biases in all segmentation methods using
-Bland-Altman plots \citep{Bland1986}.
+protocols differ in the neuroanatomical features delineated, and so we 
+assessed \mb by the degree of (Pearson) correlation of average hippocampal volume
+across subjects. We also computed the correlation in hippocampal volume between
+the existing, established automated segmentation methods: FSL, FreeSurfer, and
+MAPER. Additionally, we evaluate the volume-related fixed and proportional
+biases in all segmentation methods using Bland-Altman plots \citep{Bland1986}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: prep
+% Prepares the data in a the form needed for plotting.
+<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
+all_data  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data.csv.gz'))
+all_data_mean  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data_mean.csv.gz'))
+
+mb.best.kappa = mean(subset(all_data_mean, reg_method   == "ANTS" & 
+                                  method.mb    == "Majority Vote" &
+                                  templates.mb == 19 & 
+                                  atlases      == 9)$k.mb)
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 In this experiment we conducted 10 rounds of \mb and multi-atlas segmentation of
-each of 69 subjects at a range of atlas and template library sizes, registration
-algorithm (\ants or \animal), and three label fusion techniques.  Hippocampal
-\mb-based segmentations using both \animal and \ants registration algorithm
-demonstrate good overlap with SNT derived gold-standards (Figure
-\ref{fig:ADNI1-xval-k-mean}). Qualitatively, both \animal and \ants-based
-segmentations demonstrate trend overlap accuracy that increases with the size of
-atlas library and template library. Improvement in accuracy diminishes noticeably
-with template libraries larger than roughly ten images.
-
-No marked difference in segmentation accuracy is seen when either \animal or \ants
-registration is used with any number of atlases or templates.  In every parameter 
-configuration, the use of \mb with \ants registration shows a pronounced increase
-in segmentation accuracy over \mb with \animal registration.  Surprisingly, the 
-label fusion method used does not significantly improve label accuracy, contrary 
-to the findings of \citet{Aljabar2009} when using weighted voting on much larger
-atlas/template libraries. In the remainder of  this section, only results using
-the \ants registration algorithm and majority vote fusion will be shown.
-
-With an increasing number of templates, \mb shows improvement in overlap
-accuracy over multi-atlas-based segmentation when using the same number of atlases
-and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The magnitude of improvement over
-multi-atlas-based segmentation decreases with an increasing number of atlases,
-with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (~0.02
-DSC) is found when one atlas is used with a template library of 20 images.
+of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
+images respectively, two registration algorithms (\ants or \animal), and three
+label fusion techniques (unweighted, cross-correlation, and normalised mutual
+information weighted majority vote).  Hippocampal \mb-based segmentations using
+both \animal and \ants registration algorithm demonstrate good overlap with SNT
+\gs segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=3)}
+when using 9 atlases, 19 templates, ANTS registration, and majority vote label
+fusion); Figure \ref{fig:ADNI1-xval-k-mean}).  Qualitatively, both \animal and
+\ants-based segmentations demonstrate trend overlap accuracy that increases with
+the size of atlas library and template library. Improvement in accuracy plateaus
+with template libraries larger than ten images.
+
+The use of \mb with \ants registration shows a pronounced increase
+in segmentation accuracy over \mb with \animal registration, across all
+other variable settings we tested. Additionally, by itself, using a weighted
+voting strategy did not significantly improve segmentation accuracy, contrary to
+the findings of \citet{Aljabar2009} in basic multi-atlas segmentation. Given
+these findings, in the remainder of this section only results using the \ants
+registration algorithm and majority vote fusion will be shown.
+
+With an increasing number of templates, \mb shows improvement over
+multi-atlas-based segmentation in overlap accuracy when using the same number of
+atlases and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The two methods
+converge in accuracy when using seven atlases.  Peak improvement in \mb accuracy
+(~0.02 DSC) is found when one atlas is used with a template library of 19
+images.
 
@@
 vote fusion method used (see Discussion).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: prep
-% Prepares the data in a the form needed for plotting.
-<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-all_data  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data.csv.gz'))
-all_data_mean  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data_mean.csv.gz'))
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   ylab( "Mean DSC" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
+  opts(legend.direction = "horizontal", legend.position = "bottom", 
         axis.text = element_text(size = 8))
@@
   size, registration algorithm, and label fusion method. Points above zero
-  indicate an \mb parameter settings yielding a higher mean DSC score than
+  indicate \mb parameter settings yielding a higher mean DSC score than
   multi-atlas segmentation using the same number of atlases.
@@
   ylab( "Difference in mean DSC" ) +
-  theme(legend.direction = "horizontal", legend.position = "bottom")
+  opts(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   scale_size_continuous(guide='none') + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
+  opts(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
   geom_boxplot() +
-  labs(y=expression(paste("Percent error in volume relative to 0.3", mm^3, " Winterburn Atlases")), x="Region") +
+  labs(y="Percent error in volume", x="Subfield") +
   scale_colour_discrete("") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
-  \caption{{\bf Percent error in segmentation volume by hippocampus subregion.} 
-  Percent error is measured against the volumes of of the unmodified Winterburn
-  atlas segmentations.
+  opts(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Average percent error in hemispheric subfield volume.} 
+  Percent error is measured against the volumes of the unmodified Winterburn
+  atlas subfield segmentations.
   {\bf Subsampled} are volumes of the manual segmentations of the Winterburn
   atlases after resampling to $0.9mm^3$.  
-  {\bf MAGeT + WA Subsampled} volumes are \mb segmentations of the Winterburn atlas
+  {\bf MAGeT + Subsampled} volumes are \mb segmentations of the Winterburn atlas
   images after resampling to $0.9mm^3$ voxels.
-  {\bf MAGeT + WA BRAVO} volumes are \mb segmentations of T1 BRAVO images
+  {\bf MAGeT + BRAVO} volumes are \mb segmentations of T1 BRAVO images
   ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
@@
 amongst measures of the same quantity does not necessarily imply agreement (as
-correlation can be driven by a large range in true values, for instance).  What
-is most striking in Figure \ref{fig:ADNI-Bland-Altman} is that all methods show
-an obvious proportional bias: FreeSurfer and FSL markedly under-estimate smaller
-hippocampi and over-estimate large hippocampi, whereas MAPER and \mb more
-conservatively show the reverse bias.  Additionally, all methods show a fixed
+correlation can be driven by a large range in true values, for instance).  All
+methods show an obvious proportional bias: FreeSurfer and FSL markedly
+under-estimate smaller hippocampi and over-estimate large hippocampi, whereas
+MAPER and \mb show a reverse, conservative bias (Figure
+\ref{fig:ADNI-Bland-Altman}).  Additionally, all methods show a fixed volume
 bias, with FreeSurfer and FSL most dramatically over-estimating hippocampal
@@
       ylab(expression(paste("Hippocampal volume (", mm^3, ")"))) + 
-      theme(legend.direction = "horizontal", legend.position = "bottom")
+      opts(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
               hjust=1.1,vjust=1.5) +
-    theme(legend.direction = "horizontal", legend.position = "bottom")
+    opts(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
 %
-<<mb-best-kappa, echo=F, cache=FALSE>>=
-ants.mean = subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &      templates.mb == 19 & atlases == 9)
-mb.best.kappa = mean(ants.mean$k.mb)
-@
-
 \begin{table}[!tbp]
@@
 Throughout the cross validation in Experiment 1 (10-fold Monte Carlo cross
-validation in the
-\adnidataset dataset subsample) we find that two parameter choices improve 
-segmentation accuracy: increasing the number of atlases, and the number of
-templates. However, after setting the parameters to 5 atlases and 15 templates
-there are diminishing returns with respect to this improvement. We were
-somewhat surprised to find that cross-correlation and normalized mutual
-information based weighted voting did little to improve segmentation accuracy.
-This suggests that using an intermediate template library generated by accurate
-nonlinear registration methods such as \ants \citep{Avants2008} sufficiently filters
-sources of error such as nonlinear registration error and partial volume effects
-that arise through the use of nearest neighbour resampling. From our previous
-work on the \mb algorithm we have shown that these increases in error
-are not simply a smoothing or averaging effect \citep{MallarChakravarty2012}.
+validation in the \adnidataset dataset subsample) we find that two parameter
+choices improve segmentation accuracy: increasing the number of atlases, and the
+number of templates. However, after setting the parameters to 5 atlases and 15
+templates there are diminishing returns with respect to this improvement.
+
+Previous work on multi-atlas segmentation methods
+\citep{Aljabar2009,Collins2010} has found that cross-correlation and normalized
+mutual information based weighted label fusion improves segmentation accuracy
+over simple majority vote label fusion. Selectively filtering out atlases with
+lower image similarity is believed to reduce sources of error from estimating
+deformations via nonlinear registration, partial volume effects from nearest
+neighbour image resampling, and neuroanatomical mismatch between atlases and
+subjects. That \mb does not see the same boost in performance from weighted
+voting may suggest that the neuroanatomical variability of a template library
+constructed from study subjects more closely matches any particular subject and
+thereby leaving less error to filter. From our previous work on the \mb
+algorithm we have shown that the reduction in error is not simply a smoothing or
+averaging effect \citep{MallarChakravarty2012}.
 
@@
 
-To this end, more work is required in order to determine the source of the
-slight decrease in segmentation performance when the number of templates are
-set to an even number.  Our initial concern was that this dip in performance
-was a by-product of the \mb algorithm itself.  However, this this pattern is
-also found in the results of the multi-atlas segmentations we used in our
+To this end, more work is required to determine the source of the slight
+decrease in segmentation performance when the number of templates are set to an
+even number.  Our initial concern was that this dip in performance was a
+by-product of the \mb algorithm itself.  However, this this pattern is also
+found in the results of the multi-atlas segmentations we used in our
 experiments.  We believe that our majority voting methodology is biased towards
@@
 implementation of the \verb+mode+ function used to determine majority), thus
-causing the slight bias observed when using an even number of templates. This
-is another area where the voting scheme could be used to improve performance.
+causing the slight bias observed when using an even number of templates. This is
+another area where the voting scheme could be used to improve performance.
 However, it is worth noting that this limitation was previously identified by
@@
 
-In comparison to other methodologies in the field, it is important to note that
-\mb performs quite well.  Table 6 gives a survey of some of the most recent
-multi-atlas implementations with the reported kappa values used to validate the
-algorithm on the ADNI dataset (many of which use the same SNT labels for
-populating their template library and as gold standards for evaluation).  While
-it is difficult to compare segmentation results across studies, gold-standards,
-evaluation metrics, and algorithms it is worth noting that the methods
-summarized in Table 6 require more atlases (between 16-55) than our \mb
-implementation with the Winterburn atlases \citep{Winterburn2013}.  Amongst these
-methods only method 4 yields mean Kappas that are higher than the ones that are
-reported here.
-
-There are some important differences between our method
-and these specific methods.  Others have reported the difficulty with
-mis-registrations in candidate segmentation (i.e. segmentations generated that are
-then input in the voxel-voting procedure \citep{Collins2010}). The work of 
-\citet{Leung2010} tackles this problem by using an intensity threshold
-that is estimated heuristically at the time of segmentation (this work also
-reports some of the highest kappas for the segmentation of ADNI data).  While
-this works for the ADNI dataset (which is partially homogenized with respect to
+In comparison to other methodologies in the field \mb performs favourably. Table
+\ref{tab:other-methods} surveys some of the most recent reported kappa values on
+the ADNI dataset using the same SNT labels for populating their template library
+and as gold standards for evaluation.  While it is difficult to compare
+segmentation results across studies, gold-standards, evaluation metrics, and
+algorithms it is worth noting that the methods summarized in Table
+\ref{tab:other-methods} require more atlases (between 16-55) than our \mb
+implementation with the Winterburn atlases \citep{Winterburn2013}.
+
+There are some important differences between our method and these specific
+methods.  Others have reported the difficulty with mis-registrations in
+candidate segmentation (i.e. segmentations generated that are then input in the
+voxel-voting procedure \citep{Collins2010}). The work of \citet{Leung2010}
+tackles this problem by using an intensity threshold that is estimated
+heuristically at the time of segmentation (this work also reports some of the
+highest kappas for the segmentation of ADNI data).  While this method is
+effective for the ADNI dataset (which is partially homogenized with respect to
 image acquisition and pre-processing), it is unclear if this type of heuristic
@@
 segmentations using classifications derived using an expectation maximization
-framework. In their initial work, \citet{Chupin2009} develop their
-probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy
-patients, and then use this method to segment an ADNI sample, with a
-hierarchical experimentation protocol that involved \todo{complete this
-sentence}.  These methods suggest that some post-processing of the final
-segmentations would improve accuracy of the segmentation. While that may be
-true, there is little consensus regarding how to achieve this.
+framework. In their initial work, \citet{Chupin2009} develop their probabilistic
+methodology using a cohort of 8 healthy controls and 15 epilepsy patients, and
+then use this method to segment an ADNI sample, with a hierarchical
+experimentation protocol that involved \todo{complete this sentence}.  These
+methods suggest that some post-processing of the final segmentations would
+improve accuracy of the segmentation. While that may be true, there is little
+consensus regarding how to achieve this.
 
@@
 In comparison to the methods that we compared and were available through the
-ADNI database (FreeSurfer, MAPER) and the method that we initialized ourselves
-(FSL-FIRST) we find extremely variable performance of all methods.  With the
-exception of FSL all methods correlate well with the SNT volumes provided in
-the ADNI database.  However, FreeSurfer and FIRST provide radically different
-definitions of the size of the hippocampus in comparison to the other methods.
-Further, when estimating bias of these methods relative to SNT hippocampal
-volumes we see that large hippocampi are over estimated while small hippocampi
-are under estimated. By comparison, \mb and MAPER are far more conservative in
-volume estimation, suggesting these methods may be better suited for estimating
-true-positives. However, in this analysis we have only compared methods by
-total hippocampal volume, and so more work is needed to understand the full
-extent to which these methods differ.
+ADNI database (\freesurfer, \maper) and \fsl-FIRST, we find extremely variable
+performance of all methods.  With the exception of \fsl all methods correlate
+well with the SNT volumes provided in the ADNI database.  However, \freesurfer
+and \fsl provide radically different definitions of the size of the hippocampus
+in comparison to the other methods.  Further, when estimating bias of these
+methods relative to SNT hippocampal volumes we see that large hippocampi are
+over estimated while small hippocampi are under estimated. By comparison, \mb
+and MAPER are far more conservative in volume estimation, suggesting these
+methods may be better suited for estimating true-positives. However, in this
+analysis we have only compared methods by total hippocampal volume, and so more
+work is needed to understand the full extent to which these methods differ.
 
@@
 
-In conclusion, we have presented a flexible multi-atlas framework.  It has
-considerable advantages over other methods as only a small set of atlases is
-required to initialize the algorithm.  We demonstrate that our method works
-robustly over hippocampal definitions, different populations, and different
-acquisition types.  Finally, we also demonstrate that using this method that
-accurate identification of the hippocampal subfields in also possible.
+In conclusion, we have presented a flexible multi-atlas framework that has
+considerable advantages over other methods since only a small number of atlases
+is required to initialize the algorithm.  We demonstrate that our method works
+robustly over hippocampal definitions, different disease populations, and
+different acquisition types. Finally, we also demonstrate that accurate
+identification of the hippocampal subfields is possible.
 
@@
 \bibliography{references}
+
+\section*{R Session Information}
+<<results = "asis">>=
+toLatex(sessionInfo(), locale=FALSE)
+@
 \end{document}

### 2013-08-26 comments/paper_7bae395_comments_jens.tex
Adding Jason and Jen's comments

--- 2013-08-21 paper.Rnw
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: 
-%  - check tense: we are is have has 
-%  - update montages with TAL coordinates
-%  - add MAGeT brain diagram, and remove pseudocode
-%  - check that \gs is used instead of Manual 
 % Sweave/Knitr init
-<<setup, include=FALSE, cache=FALSE>>=
-library(knitr)
-opts_chunk$set(echo = FALSE, 
-               message = FALSE, 
-               warning = FALSE, 
-               fig.align = 'center', 
-               tidy = FALSE, 
-               comment = NA, 
-               cache = TRUE,
-               fig.width = 6, 
-               fig.height = 6)
-library(ggplot2)
-library(Hmisc)
-library(reshape2)
-library(ADNIMERGE)
-library(plyr)
-options(digits=3)
-theme_set(theme_bw(base_size = 12))
-
-lm_eqn = function(m) {
-  l <- list(a = format(coef(m)[1], digits = 2),
-      b = format(abs(coef(m)[2]), digits = 2),
-      r2 = format(summary(m)$r.squared, digits = 3));
-  if (coef(m)[2] >= 0)  {
-    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
-  } else {
-    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass{article}\usepackage{graphicx, color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
 }
-
-  as.character(as.expression(eq))
-}
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass{article}
+\makeatother
+
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
+\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
+\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
+\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
+\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
+\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
+\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
+\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
+\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
+\newcommand{\hlslot}[1]{\textit{#1}}%
+\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage[margin=1in]{geometry}
 \usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
-\usepackage{titlesec, siunitx, graphicx}
+\usepackage{titlesec, siunitx}
 \usepackage[round,authoryear]{natbib}
@@
 \newcommand{\FEPdataset}{SZ-FEP }
-\newcommand{\gs}{Gold Standard }
 
@@
 \author[1]{Julie Winterburn}
-\author[1,7]{Tristram A. Lett}
+\author[1]{Tristram A. Lett}
 \author[2,3]{Jason P. Lerch}
@@
 \author[4]{Martin Lepage}
-\author[1,5,7]{Aristotle N. Voineskos}
+\author[1,5]{Aristotle Voineskos}
 \author[1,5,6]{M. Mallar Chakravarty}
@@
 ON, Canada}
-\affil[7]{Institute of Medical Science, University of Toronto, Toronto, ON, Canada}
 
@@
 \begin{abstract}
-%Neuroimaging research often relies on automated anatomical segmentations of MR
-%images of the brain. Current multi-atlas based approaches provide accurate
-%segmentations of brain images by propagating manually derived segmentations of
-%specific neuroanatomical structures to unlabelled data. These approaches often
-%rely on a large number of such manually segmented atlases that take significant
-%time and expertise to produce. We present an algorithm for the automatic
-%segmentation of the hippocampus that minimizes the number of atlases needed
-%while still achieving similar accuracy to multi-atlas approaches.
 \end{abstract}
@@
 
-The hippocampus is a neuroanatomical structure situated in the medial temporal
-lobe of the brain, and has long been associated with learning and memory
-\citep{DenHeijer2012,Scoville2000}. In addition to its known functional roles,
-the hippocampus is of interest to neuroscientists because it is implicated in
-several forms of brain disfunction such as Alzheimer's disease
+The hippocampus is a brain structure situated in the medial temporal
+lobe, and has long been associated with learning and memory
+\citep{DenHeijer2012,Scoville2000}.  The hippocampus is of interest to neuroscientists because it is implicated in
+many forms of brain disfunction, including Alzheimer's disease
 \citep{Sabuncu2011} and schizophrenia \citep{Narr2004,Karnik-Henry2012}.  In
-neuroimaging experiments, magnetic resonance images (MRI) are often used for the
-identification of the hippocampus. As such, accurate segmentation of the
+neuroimaging studies, structural magnetic resonance images (MRI) are often used for the volumetric assessment of the hippocampus.  As such, accurate segmentation of the
 hippocampus and its subregions in MRI is a necessary first step to better
-understand the unique neuroanatomy of subjects. Typically, the gold standard for
-neuroanatomical segmentation is manual delineation by an expert human rater.
-However with the availability of increasingly large MRI datasets, the time and
-expertise required for manual segmentation becomes prohibitive
-\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. Further, there is
-little agreement between researchers regarding how exactly the hippocampus
-should be identified in MRI images \citep{Geuze2004} and this has led to efforts
-to create an unified hippocampal segmentation protocol \citep{Jack2011}.
+understand interindividual variability of neuroanatomy of subjects. Typically, the gold standard for
+neuroanatomical segmentation is manual delineation by an expert human rater. However,
+the rapid increase in the number of available MRI data and the time and expertise
+required for manual segmentation makes this increasingly difficult
+\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}.
+Further, there is significant variation between segmentation protocols regarding specific anatomical boundaries of the 
+hippocampus \citep{Geuze2004} and this has
+led to efforts to create an unified hippocampal segmentation protocol
+\citep{Jack2011}.
 
 Automated segmentation techniques that are reliable, objective, and reproducible
-are a necessary alternative to manual segmentation. In the case of classical
+can be considered complementary to manual segmentation.  In the case of classical
 model-based segmentation methods \citep{Haller1997,Csernansky1998}, an MRI atlas
@@
 successfully in several contexts
-\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited in
-accuracy due to inaccuracies in the estimated nonlinear transformation itself,
-partial volume effects in label resampling, and irreconcilable differences
-between the neuroanatomy represented within the atlas and target images.
+\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
+in accuracy by the introduction of errors due inaccuracies in the nonlinear
+transformation itself, partial volume effects in label resampling, and
+irreconcilable differences between the neuroanatomy represented within the atlas
+and target images.
 
@@
 atlas). A {\it label fusion} technique, such as voxel-wise voting, is used to
-merge these labels into the definitive segmentation for the target. In addition,
+merge these labels into a definitive segmentation for the target.  In addition,
 weighted voting procedures that use {\em atlas selection} techniques are often
@@
 This involves the selection of a subset of atlases using a similarity metric
-such as cross-correlation \citep{Aljabar2009} or normalized mutual information.
+such a cross-correlation \citep{Aljabar2009} or normalized mutual information.
 Such selection has the added benefit of significantly reducing the number of
-nonlinear registrations. For example \citet{Collins2010} demonstrated that only
-14 atlases, selected based on highest similarity between medial temporal lobe
-neuroanatomy as evaluated by normalized mutual information \citep{Studholme1999}
-from a library of 80 atlases, were required to achieve accurate segmentations of
-the hippocampus. Additionally, several methods have been explored for label
-fusion, including the STAPLE algorithm \citep{Warfield2004} that computes a
-probabilistic segmentation using an expectation maximization framework from an
-set of competing segmentations, or others where a subset of segmentations can be
-estimated using metrics such as the sum of squared differences in the regions of
-interest to be segmented \citep{Coupe2011}.
+nonlinear registrations.  For example \citet{Collins2010}
+demonstrated that only 14 atlases, selected based on highest similarity between
+medial temporal lobe neuroanatomy as evaluated by normalized mutual information
+\citep{Studholme1999} from a library of 80 atlases, were required to achieve
+accurate segmentations of the hippocampus.  Several methods have been explored
+for label fusion including the STAPLE algorithm \citep{Warfield2004} that
+computes a probabilistic segmentation using an expectation maximization
+framework from an set of competing segmentations; or others where a subset of
+segmentations can be estimated using metrics such as the sum of squared
+differences in the regions of interest to be segmented \citep{Coupe2011}.
 
 However, many of these methods require significant investment of time and
-resources for the creation of the atlas library ranging between 30
-\citep{Heckemann2006} and 80 \citep{Collins2010} manually segmented atlases.
-This strategy has the main drawback of being inflexible as it does not easily
-accommodate varying the definition of the hippocampal anatomy (such as the
-commonly used heuristic of subdividing the hippocampus into head, body, and tail
-\citep{Poppenk2011,Pruessner2000}). Furthermore, none of these methods have
-demonstrated sufficient flexibility to accommodate atlases that are somehow
-exceptional such as those derived from serial histological data
-\citep{Chakravarty2006,Yelnik2007} or high-resolution MRI data that enables
-robust identification of hippocampal subfields
-\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}. Due
-to the recent availability of the latter, there has been increased interest in
-the use of probabilistic methods for the identification of the hippocampal
-subfields on standard T1-weighted images. Our group recently demonstrated that
-through use of an intermediary automated segmentation stage, robust and accurate
+resources for the creation of the atlas library; ranging from atlas libraries
+that require between 30 \citep{Heckemann2006} and 80 \citep{Collins2010}
+manually segmented atlases.  This strategy has the main drawback of being
+inflexible as it does not easily accommodate varying the definition of the
+hippocampal anatomy (such as the commonly used heuristic of subdividing the
+hippocampus in to head, body, and tail \citep{Poppenk2011,Pruessner2000}).
+Furthermore, none of these methods have demonstrated sufficient flexibility to
+accommodate atlases that are somehow exceptional such as those derived from
+serial histological data \citep{Chakravarty2006,Yelnik2007} or high-resolution MRI
+data that enables robust identification of hippocampal subfields
+\citep{Winterburn2013,Yushkevich2009,Mueller2009,VanLeemput2009,Wisse2012}.  Due to
+the recent availability of the latter, there has been increased interest in the
+use of probabilistic methods for the identification of the hippocampal subfields
+on standard T1-weighted images.  Our group recently demonstrated that through
+use of an intermediary automated segmentation stage, robust and accurate
 segmentation of the striatum, pallidum, and thalamus using a single atlas
 derived from serial histological data is possible \citep{MallarChakravarty2012}.
-The novelty of this manuscript is the extension of our multi-atlas methodology
-to the hippocampus using more than a single input atlas, while simultaneously
+The novelty of this manuscript is the extension our multi-atlas methodology to
+the hippocampus using more than a single input atlas, while simultaneously
 limiting the number of possible inputs used during segmentation, and
@@
 successively incorporating unlabelled target images after themselves being
-labelled using multi-atlas techniques. The sequence in which target images are
+labelling using multi-atlas techniques. The sequence in which target images are
 labelled is chosen so that the similarity between the atlas images and the
@@
 
-To the best of our knowledge there are two other segmentation methods that
-attempt to define the hippocampus and its subfields using standard T1-weighted
-data. The first is included with the FreeSurfer package \citep{VanLeemput2009}.
-This work is limited as it omits the tail of the hippocampus and the
-segmentation protocol has yet to be fully validated. Nonetheless, it
-demonstrates that the applicability of hippocampal subfield segmentation using
-data from 10 subjects. In the second method, \citet{Yushkevich2009} acquired
-and labelled hippocampal subfields on high-resolution MRI data from post-mortem
-medial temporal lobe samples. Using nonlinear registration guided by manually
-derived hippocampus masks and specific landmarks, they demonstrate accurate
-parcellation of hippocampal subfields in unlabelled MRI volumes.
-
-In this paper we undertake a thorough validation of the \mb algorithm for
-segmentation of the hippocampus. First, we address the feasibility and accuracy
-of whole hippocampus segmentation with a limited number of input atlases by
-performing a multi-fold experiment using a subset of the Alzheimer's Disease
-Neuroimaging Initiative (ADNI) 1 dataset. Based on the parameters we find in our
-experiment, we then validate our algorithm using all of the data available in
-the ADNI Complete 1Yr sample and compare our segmentations to the other
-segmentations that are available through the ADNI informatics portal. To ensure
-that we have not over-fit our parameters to the aging or neurodegenerative
-brain, we also apply our segmentations to a dataset of individuals suffering
-from first episode psychosis. Finally, we perform a leave-one-out validation
+To the best of our knowledge there are only two other segmentation methods that
+attempt to define the hippocampal subfields using standard T1-weighted data.
+The first is included with the FreeSurfer package \citep{VanLeemput2009}.  This
+work is limited as it omits the tail of the hippocampus and the segmentation
+protocol has yet to be fully validated.  Nonetheless, they demonstrate the
+applicability of their work using data from 10 subjects.  In the second method,
+\citet{Yushkevich2009} hippocampal subfields were acquired and labelled on
+high-resolution MRI data from post-mortem medial temporal lobe samples. Using
+nonlinear registration guided by manually derived hippocampus masks and specific
+landmarks, the authors demonstrated accurate parcellation of hippocampal subfields in
+unlabelled MRI volumes.
+
+Here we address the issue of limiting the number of input atlases by tuning our
+algorithm, for segmentation of the entire hippocampus, using a multi-fold
+experiment performed on a subset Alzheimer's Disease Neuroimaging Initiative
+(ADNI) 1 dataset. Based on the parameters we find in our experiment, we validate
+our algorithm using all of the data available in the ADNI Complete 1Yr sample
+and compare our segmentations to the other segmentations that are available
+through the ADNI informatics portal.  To ensure that we have not over-fit our
+parameters to the aging or neurodegenerative brain, we also apply our
+segmentations to a dataset of normal controls and individuals suffering from
+first episode psychosis. Finally, we perform a leave-one-out validation
 experiment to determine if the subfields can be accurately identified using our
@@
 via label propagation). The terms {\em atlas library} and {\em template library} 
-describe any set of such images. Additionally, we will use the term {\em target} 
+describe any set of such images.Additionally, we will use the term {\em target} 
 to refer to an unlabelled image that is undergoing segmentation.
@@
 Second, the labels from each image are propagated to the target image space. 
-Third, the labels are combined into a single label by label fusion
-\citep{Heckemann2006, Heckemann2011}. This method is described in detail in
-other publications \citep{Collins2010,Heckemann2011,Aljabar2009}.
+Third, the labels are combined into a single labelling by way of a label fusion method 
+\citep{Heckemann2006, Heckemann2011}. This method is described in detail in other
+publications \citep{Collins2010,Heckemann2011,Aljabar2009}.
 
@@
 target set as a whole (for instance, by sampling equally from cases and controls). 
-The template library images are automatically labelled by each of the atlases
-via label propagation. Basic multi-atlas segmentation is then conducted using
-the template library to segment the entire set of target images (including the
-targets whose images are used in the construction of the template library).
-Since each template library image has multiple labels (one from each atlas), the
-final number of labels to be fused for each target may be quite large (i.e. \#
-of atlas $\times$ \# of templates).
+The template library images are then labelled by each of the atlases. Basic 
+multi-atlas segmentation is then conducted using the template library
+to segment the entire set of target images (including the targets whose images
+are used in the construction of the template library).  Since each template
+library image has multiple labels (one from each atlas), the final number of 
+labels to be fused for each target may be quite large (i.e. \# of atlas $\times$ 
+\# of templates).
 
@@
 
-The following section describes for experiments conducted to assess the
+The following section describes the experiments we conducted to assess the
 segmentation quality of the \mb algorithm. The first two experiments assess the
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT \gs segmentations from the ADNI1 dataset. This form of
+of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
 cross-validation allows us to rigorously validate a large number of parameter
@@
 
-\paragraph{\adnidataset dataset} 
+\paragraph{ADNI1 dataset}
+
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
@@
 Alzheimer's disease (AD). Demographics for this subset are shown in Table
-\ref{tab:ADNI1-xvalidation-demographics}.  Semi-automated segmentations of the
-left and right whole hippocampi are made available with a subset of ADNI images
-\citep{Hsu2002}. These labels have been generated using the SNT tool from
-Medtronic Surgical Navigation Technologies, Louisville, CO (see Supplementary
-Materials for detailed discussion of the segmentation process used). Throughout
-this paper, we will use these segmentations as the \gs for comparison.
+\ref{tab:ADNI1-xvalidation-demographics}.  Manual segmentations of the left and
+right whole hippocampi are available \citep{Hsu2002}. These labels have been
+generated using the SNT tool from Medtronic Surgical Navigation Technologies,
+Louisville, CO (see Supplementary Materials for detailed discussion of the
+manual segmentation process used).
 
@@
 authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset used was the "\adnidataset" standardized dataset
-available from ADNI \footnote{
+August 2012. The image dataset download was the "\adnidataset"
+standardized dataset available from ADNI \footnote{
 \url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
@@
 Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
-2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240 x
-240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR
+= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
+x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
 yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.
@@
 %
-<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
-adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
-                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
-                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
-                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
-                 1217,29,56,1188,1293,149,382,431,1202)
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
-caption = 
-  "\\textbf{ADNI-1 cross-validation subset demographics.} 
-  CN     - Cognitively Normal. 
-  LMCI   - Late-onset Mild Cognitive Impairment. 
-  AD     - Alzheimer's Disease.  
-  Hisp   - Hispanic. 
-  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
-  ADAS   - Alzheimer's Disease Assessment Scale.  
-  MMSE   - Mini-Mental State Examination."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      label="tab:ADNI1-xvalidation-demographics")
-@
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+\begin{table*}[!tbp]
+\scriptsize
+\caption{\textbf{ADNI-1 cross-validation subset demographics.} Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).\label{tab:ADNI1-xvalidation-demographics}}
+\begin{center}
+\begin{tabular}{lcccc}
+\hline\hline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}\tabularnewline
+&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=23$}}&\multicolumn{1}{c}{{\scriptsize $N=69$}}\tabularnewline
+\hline
+Age~at~baseline~\hfill\tiny{Years}&{\scriptsize 72.2~}{75.5 }{\scriptsize 78.5} &{\scriptsize 71.0~}{77.1 }{\scriptsize 81.4} &{\scriptsize 71.7~}{77.8 }{\scriptsize 81.8} &{\scriptsize 71.5~}{76.6 }{\scriptsize 81.3} \tabularnewline
+Sex~:~Female&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(10)}&43\%~{\scriptsize~(30)}\tabularnewline
+Education&{\scriptsize 16.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 15.0~}{16.0 }{\scriptsize 18.0} &{\scriptsize 12.0~}{16.0 }{\scriptsize 16.5} &{\scriptsize 14.0~}{16.0 }{\scriptsize 18.0} \tabularnewline
+CDR-SB&{\scriptsize 0.00~}{0.00 }{\scriptsize 0.00} &{\scriptsize 0.75~}{1.50 }{\scriptsize 1.50} &{\scriptsize 4.00~}{4.50 }{\scriptsize 5.00} &{\scriptsize 0.00~}{1.50 }{\scriptsize 4.00} \tabularnewline
+ADAS~13&{\scriptsize  4.67~}{ 5.67 }{\scriptsize 12.34} &{\scriptsize 14.34~}{16.00 }{\scriptsize 20.50} &{\scriptsize 23.83~}{29.00 }{\scriptsize 31.66} &{\scriptsize 10.00~}{16.00 }{\scriptsize 25.33} \tabularnewline
+MMSE&{\scriptsize 28.5~}{29.0 }{\scriptsize 30.0} &{\scriptsize 25.0~}{27.0 }{\scriptsize 28.0} &{\scriptsize 21.0~}{23.0 }{\scriptsize 24.0} &{\scriptsize 24.0~}{27.0 }{\scriptsize 29.0} \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\Numbers after percents are frequencies.\end{table*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In this experiment, our dataset consists of 69 1.5T images and corresponding \gs
-segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and their \gs segmentations to be used as an
-atlas library, and a validation set consisting of the remaining images segmented
-by both \mb and multi-atlas. The computed segmentations were compared to the
-\gs segmentations (see below).
-
-A total of ten validation rounds were performed on each subject in the dataset,
+In this experiment, our dataset consists of 69 1.5T images and corresponding
+manual segmentations. In each validation round, the dataset is partitioned into 
+a training set consisting of images and their manual labels to be used as an atlas 
+library, and a validation set consisting of the remaining images segmented by both 
+\mb and multi-atlas. The resulting segmentations are compared to the manual 
+segmentations for the images.
+
+A total of ten validation rounds are performed on each subject in the dataset,
 over each combination of parameter settings. The parameter settings
-explored are: atlas library size (1-9), template library size (1-20),
+we explore are: atlas library size (1-9), template library size (1-20),
 registration method (\ants or \animal), and label fusion method (majority vote,
@@
 Before registration, all images underwent preprocessing with the N3 algorithm
-\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we
-compared two non-linear image registration methods:
+\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we use
+one of two non-linear image registration methods.
 
@@
 estimating smaller deformations on data blurred with a Gaussian kernel with a
-smaller full width at half maximum (FWHM). The final transformation is a set of
-local translations defined on a bed of equally spaced nodes that were estimated
-through the optimization of the correlation coefficient.  For the purposes of
-this work we used the regularization parameters optimized in
-\citet{Robbins2004}, displayed in Table \ref{tab:ANIMAL-params}.
+smaller FWHM. The final transformation is a set of local translations defined on
+a bed of equally spaced nodes that were estimated through the optimization of
+the correlation coefficient.  For the purposes of this work we used the
+regularization parameters optimized in \citet{Robbins2004}, displayed in Table
+\ref{tab:ANIMAL-params}.
 
@@
   output voxel is given the most frequent label at that voxel location amongst
-  all candidate labellings.
+  all candidate labellings.  Ties are broken arbitrarily.
 
@@
 \end{description}
-
 \paragraph{Evaluation method}
 
-The Dice similarity coefficient (DSC), also known as Dice's Kappa, assesses the
-agreement between two segmentations. It is one of the most widely used measures
-of segmentation accuracy, and we use it as the basis of comparison in this
-experiment.
+The Dice similarity coefficient (DSC) assesses the agreement between two
+segmentations. It is one of the most widely used measures of segmentation
+performance, and we use it as the basis of comparison in this experiment.
 %Additionally, we report the Jaccard index, another commonly used similarity
@@
 
-The automatically produces labels are compared to the \gs SNT labels provided by
-ADNI. The segmentation accuracy reported for a parameter setting is the average
-over ten validation rounds.
+The manual segmentations (SNT) provided as part of the ADNI dataset are used as
+the gold standard to compare with \citep{Hsu2002}.  The segmentation accuracy
+reported is averaged over the ten validation rounds for each parameter setting.
 
 In order to investigate the performance of \mb in a real world setting in which 
-only one size of atlas and template libraries are used, we explore the
-variability in label agreement at fixed parameter settings under varying  
-atlas and template library compositions. This is achieved by first computing the
-standard deviation and variance of DSC scores in each block of ten validation
-rounds per subject. The distribution of these statistics across all subjects is
-then compared between \mb and multi-atlas using a Student's t-test. A
-significant difference between distributions is taken to show either a larger or
-smaller level of variability between methods.
+only one set of atlas and template images are used, we explore the variability 
+in label agreement at fixed parameter settings when the choice for atlas and 
+template images is varied. This is achieved by first computing the standard 
+deviation and variance of DSC scores in each block of ten validation rounds per
+subject. The distribution of these statistics across all subjects is then
+compared between \mb and multi-atlas using a Student's t-test. A significant
+difference between distributions is taken to show either a larger or smaller
+level of variability between methods.
 
@@
 In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
-segmentation is tested using a leave-one-out cross-validation (LOOCV) design on
-two image sets. The optimal parameter settings for \mb found in Experiment 1 are
-used in this experiment.
+segmentation is tested using a leave-one-out cross-validation (LOOCV) design. 
+The optimal parameter settings for \mb found in Experiment 1 are used in this 
+experiment.
 
@@
 In addition to the high-resolution scans distributed as part of the Winterburn
-atlases, we also obtained separate 3T T1 BRAVO acquisitions (0.9mm-isotropic
-voxels) of four of the five Winterburn atlas subjects.
+atlases, we also obtained additional 3T T1 BRAVO images (0.9mm-isotropic voxels)
+of four of the five Winterburn atlas subjects.
 
@@
 Leave-one-out cross-validation (LOOCV) is an approach in which the method to be
-validated is given all but one item in a dataset as training data (in our case,
-atlas library images and labels), and the output accuracy is measured on the
-left-out item. This is done, in turn, for each item in the dataset.
-
-\begin{figure}
-  \begin{centering}
-    \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
-  \end{centering}
-  \caption{A schematic illustration of the experimental set up of the
-  leave-one-out cross-validation (LOOCV) subfield experiment.}
-  \label{fig:LOOCV-schematic}
-\end{figure}
-
-In part 2a of this experiment, we have modified the LOOCV experimental
-design in order to compare \mb segmentation accuracy with the accuracy of image
-resampling error. The five $300\mu$-isotropic voxel Winterburn atlas images and
-labels are used as the atlas library, and LOOCV is conducted using the five Winterburn
-atlas images subsampled (using trilinear interpolation) to
-0.9mm-isotropic voxel resolution (referred to as the {\em Subsampled} dataset)
-as input subjects. Subsampling the subject images allows us to assess \mb in a
-typical segmentation scenario (high-resolution atlases and lower-resolution
-subjects).  The template library is composed of all five subject images, plus an
-additional set of 3T T1 images (0.9mm-isotropic voxels) of healthy subjects
-acquired separately (Table \ref{tab:WAval-healthy-demographics}). The optimal
-size of template library, registration method, and label fusion method found in
-Experiment 1 are used.  As with traditional LOOCV, each subject image is
-segmented by \mb with that subject's image excluded from the atlas library.  In
-part 2b of this experiment, we carry out a similar validation in which the
-subject images are separately acquired T1 BRAVO images of four of the five
-Winterburn atlas subjects (referred to as the {\em BRAVO} dataset).  Figure
-\ref{fig:LOOCV-schematic} illustrates the experimental set up of both parts of
-this experiment.
+validated is given all but one item in a dataset as training data, and the
+output is compared with the left-out item. This is done, in turn, for each item
+in the dataset.
+
+In this experiment the Winterburn atlases are used as the cross-validation
+dataset.  The five $300\mu$-isotropic voxel images and labels are used as
+atlases, and LOOCV is conducted using the five Winterburn atlas subject images
+subsampled (using trilinear interpolation) to 0.9mm-isotropic voxel resolution
+(referred to as the {\em Subsampled} dataset) as input subjects. Subsampling the
+subject images allows us to assess \mb in a typical segmentation scenario
+(high-resolution atlases and lower-resolution subjects).  The template library
+is composed of the subject images, plus an additional set of 3T T1 images
+(0.9mm-isotropic voxels) of healthy subjects acquired separately (Table
+\ref{tab:WAval-healthy-demographics}). The optimal size of template library,
+registration method, and label fusion method found in Experiment 1 are used.
+Each resampled Winterburn atlas subject image is segmented by \mb with that
+subject's image excluded from the atlas library.
 
@@
 %                                     
-<<WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
-master = read.csv('data/DT_study.csv')
-tab <- summary(~Age + Sex + Education + Handedness, 
-               data = master, method="reverse", na.rm = TRUE, test=FALSE)
-caption = 
-  "\\textbf{Subfield validation healthy control demographics.}
-  CN     - Cognitively Normal. 
-  LMCI   - Late-onset Mild Cognitive Impairment. 
-  AD     - Alzheimer's Disease.  
-  Hisp   - Hispanic. 
-  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
-  ADAS   - Alzheimer's Disease Assessment Scale.  
-  MMSE   - Mini-Mental State Examination."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      title="tab:WAval-healthy-demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+\begin{table*}[!tbp]
+\scriptsize
+\caption{\textbf{ADNI-1 cross-validation subset demographics.} Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).\label{tab:WAval-healthy-demographics}}
+\begin{center}
+\begin{tabular}{lrc}
+\hline\hline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{ }\tabularnewline
+\hline
+Age&16&{\scriptsize 31.0~}{53.0 }{\scriptsize 63.8} \tabularnewline
+Sex~:~Male&16&38\%~{\scriptsize~(~6)}\tabularnewline
+Education~:~0.01&15&~7\%~{\scriptsize~(1)}\tabularnewline
+~~~~12&&13\%~{\scriptsize~(2)}\tabularnewline
+~~~~13&&13\%~{\scriptsize~(2)}\tabularnewline
+~~~~14&&20\%~{\scriptsize~(3)}\tabularnewline
+~~~~16&&13\%~{\scriptsize~(2)}\tabularnewline
+~~~~18&&33\%~{\scriptsize~(5)}\tabularnewline
+Handedness~:~R&16&94\%~{\scriptsize~(15)}\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We reproduce this experiment in a separate LOOCV run in which the resampled 
+Winterburn atlas subject images are substituted for separately acquired T1 
+BRAVO images of four of the subjects (referred to as the {\em BRAVO} dataset).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 3: Application to the segmentation of first episode
+\subsubsection{Experiment 3: Application to the segmentation first episode
 schizophrenia patients}
 
-To validate that \mb works effectively in the context of other neuropsychiatric
-disorders, in this experiment we use the Winterburn atlases to predict the
-hippocampal segmentations of patients with Schizophrenia. The resulting
-segmentations are assessed for quality by comparison with expert manual
-segmentations.
+To validate that \mb algorithm works effectively in the context of other
+neuropsychiatric disorders, we use the Winterburn atlases with \mb to predict
+the hippocampal segmentation of dataset of Schizophrenia patient MR images. The
+resulting segmentations are assessed for quality by comparison with expert
+manual segmentations.
 
@@
 intervention service at the Douglas Mental Health University Institute in
-Montreal, Canada. People aged 15 to 30 years from the local catchment area
+Montreal, Canada. People aged 15–30 years from the local catchment area
 suffering from either affective or non-affective psychosis who had not taken
@@
 
-Manual segmentation of each subject hippocampus is produced following a
-validated segmentation protocol \citep{Pruessner2000}.
+Each subject hippocampus is traced following a validated segmentation protocol 
+\citep{Pruessner2000}.
 
@@
 The manual and Winterburn hippocampal segmentation protocols differ slightly in
-the neuroanatomical features that are delineated \citep{Winterburn2013}. This
-difference poses a problem for evaluation by measuring overlap.  That is, since
-different protocols will necessarily produce segmentations that do not perfectly
-overlap, the degree of overlap cannot be solely used to compare segmentation
-methods using different protocols. In place of an overlap metric, we can assess
-the degree of correlation in average hippocampal volume of the subjects produced
-by each method. Specifically, a Pearson correlation is used.
+the neuroanatomical features that are delineated.  This poses a problem for
+evaluation by measuring overlap.  That is, since different protocols will
+necessarily produce segmentations that do not perfectly overlap, the degree of
+overlap cannot be solely used to compare segmentation methods using different
+protocols. In place of an overlap metric, we can assess the degree of correlation in
+average hippocampal volume of the subjects produced by each method.
+Specifically, Pearson correlation is used.
 
@@
 %
-<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
-caption = 
-  "\\textbf{Schizophrenia First Episode Patient Demographics.}
-  ambi - ambidextrous. 
-  SES  - Socioeconomic Status score. 
-  FSIQ - Full Scale IQ.
-"
-fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
-fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
-tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
-               data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      title="tab:SZFEP-Demographics")
-@
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+\begin{table*}[!tbp]
+\scriptsize
+\caption{\textbf{Schizophrenia First Episode Patient Demographics.} Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease State Examination (MMSE).\label{tab:SZFEP-Demographics}} 
+\begin{center}
+\begin{tabular}{lrc}
+\hline\hline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{FEP}\tabularnewline
+&&\multicolumn{1}{c}{{\scriptsize $N=81$}}\tabularnewline
+\hline
+Age&80&{\scriptsize 21~}{23 }{\scriptsize 26} \tabularnewline
+Gender~:~M&81&63\%~{\scriptsize~(51)}\tabularnewline
+Handedness~:~ambi&81&~6\%~{\scriptsize~(~5)}\tabularnewline
+~~~~left&&~5\%~{\scriptsize~(~4)}\tabularnewline
+~~~~right&&89\%~{\scriptsize~(72)}\tabularnewline
+Education&81&{\scriptsize 11~}{13 }{\scriptsize 15} \tabularnewline
+SES~:~lower&81&31\%~{\scriptsize~(25)}\tabularnewline
+~~~~middle&&54\%~{\scriptsize~(44)}\tabularnewline
+~~~~upper&&15\%~{\scriptsize~(12)}\tabularnewline
+FSIQ&79&{\scriptsize  88~}{102 }{\scriptsize 109} \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To validate \mb segmentation quality with respect to other established automated
-methods, \mb is applied to large dataset from the ADNI project and the resulting
-segmentations are compared to those produced by FreeSurfer, FSL, MAPER and by
-expert semi-automated manual segmentations (SNT).
+To validate that \mb algorithm works as well as established automated methods,
+\mb is applied to the ADNI1 dataset and the resulting segmentations are
+compared to those produced by FreeSurfer, FSL, MAPER, and by expert manual
+segmentation.
 
 \paragraph{ADNI1 dataset revisited} 
-The {\em \adnidataset} standardized dataset contains $1919$ images in total.
-SNT and FreeSurfer hippocampal volumes for a subset of images are provided by
-ADNI, along with quality control data for each FreeSurfer segmentation
-(guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
-the entire \adnidataset dataset are shown in Table
+All images from the {\em \adnidataset} standardized dataset described
+in Experiment 1 are used.  Clinical and demographic data are shown in Table
 \ref{tab:ADNI1-Dataset-Demographics}.
@@
 \paragraph{Experiment details} 
-\mb was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 2), and the size of template library,
-registration method, and label fusion method found in Experiment 1 were used.
-The FSL FIRST segmentation method was used via the \verb+run_first_all+ script
-according to the FIRST user guide
-\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images in
-the \adnidataset dataset were segmented by both methods.
-
-One author (MP) performed visual quality inspection for MAGeT and FSL
-segmentations using similar quality control guidelines (if either hippocampus
-was under or over segmented by greater than 10mm in three or more slices then
-the segmentation did not pass).  Only images meeting the conditions of having
-segmentations from all methods (SNT, MAPER, FreeSurfer, FSL, and \mb) and also
-passing quality control inspection are included in the analysis.
+\mb is configured with an atlas library composed of the Winterburn T1 atlases
+(see Experiment 2). All images from the \adnidataset dataset are
+segmented. The optimal size of template library, registration method, and label
+fusion method found in Experiment 1 are used. The template library is composed
+of equal numbers of images from each disease class (AD, MCI, and cognitively
+normal controls).
 
 \paragraph{Evaluation method}
-As in Experiment 3, the SNT \gs and Winterburn hippocampal segmentation
+As in Experiment 3, the manual (SNT) and Winterburn hippocampal segmentation
 protocols differ in the neuroanatomical features delineated, and so we must
 assess \mb by the degree of correlation of average hippocampal volume across all
-subjects. Specifically, Pearson correlation is used. We also compute the
-correlation in hippocampal volume between the existing, established automated
-segmentation methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
+subjects produced by \mb and by manual segmentation.  Specifically, Pearson
+correlation is used.  For comparison, we also compute the correlation in
+hippocampal volume between the existing, established automated segmentation
+methods: FSL, FreeSurfer, and MAPER.  Additionally, we evaluate the
 volume-related fixed and proportional biases in all segmentation methods using
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<ADNI-volumes-prep,echo=F,cache=T>>=
-means.complete   <- read.csv(gzfile('data/cache/ADNI1:means.complete.csv.gz'))
-package_totals   <- read.csv(gzfile('data/cache/ADNI1:package_totals.csv.gz'))
-@
 
@@
 %                                                                             
-<<ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
-# RID < 2000 ensures only ADNI1 patients
-yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
-yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
-yr1 = subset(yr1, !grepl("Scaled_2", Description))
-yr1 = yr1[,c("RID","VISCODE")]
-yr1 = 
-adnimerge.yr1 = merge(adnimerge, yr1) 
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge.yr1, 
-               #subset = RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
-               method = "reverse", test=FALSE, overall=TRUE)
-caption = 
-  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} 
-  CN     - Cognitively Normal. 
-  LMCI   - Late-onset Mild Cognitive Impairment. 
-  AD     - Alzheimer's Disease.  
-  Hisp   - Hispanic. 
-  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
-  ADAS   - Alzheimer's Disease Assessment Scale.  
-  MMSE   - Mini-Mental State Examination."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-    caption=caption, caption.loc = c("top"),
-    title="tab:ADNI1-Dataset-Demographics",
-    label="tab:ADNI1-Dataset-Demographics")
-@
+% latex.default(cstats, title = title, caption = caption, rowlabel = rowlabel,      col.just = col.just, numeric.dollar = FALSE, insert.bottom = legend,      rowname = lab, dcolumn = dcolumn, extracolheads = extracolheads,      extracolsize = Nsize, ...)
+%
+\begin{table*}[!tbp]
+\scriptsize
+\caption{\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} Clinical Dementia Rating-Sum of Boxes (CDR-SB), Alzheimer's Disease Assessment Scale (ADAS), Mini–Mental State Examination (MMSE).\label{tab:ADNI1-Dataset-Demographics}}
+\begin{center}
+\begin{tabular}{lrcccc}
+\hline\hline
+\multicolumn{1}{l}{}&\multicolumn{1}{c}{N}&\multicolumn{1}{c}{CN}&\multicolumn{1}{c}{LMCI}&\multicolumn{1}{c}{AD}&\multicolumn{1}{c}{Combined}\tabularnewline
+&&\multicolumn{1}{c}{{\scriptsize $N=584$}}&\multicolumn{1}{c}{{\scriptsize $N=931$}}&\multicolumn{1}{c}{{\scriptsize $N=404$}}&\multicolumn{1}{c}{{\scriptsize $N=1919$}}\tabularnewline
+\hline
+Age~at~baseline~\hfill\tiny{Years}&1919&{\scriptsize 72.4~}{75.8 }{\scriptsize 78.5} &{\scriptsize 70.5~}{75.1 }{\scriptsize 80.4} &{\scriptsize 70.1~}{75.3 }{\scriptsize 80.2} &{\scriptsize 71.1~}{75.3 }{\scriptsize 79.8} \tabularnewline
+Sex~:~Female&1919&48\%~{\scriptsize~(~278)}&35\%~{\scriptsize~(~327)}&49\%~{\scriptsize~(~198)}&42\%~{\scriptsize~(~803)}\tabularnewline
+Education&1919&{\scriptsize 14~}{16 }{\scriptsize 18} &{\scriptsize 14~}{16 }{\scriptsize 18} &{\scriptsize 12~}{15 }{\scriptsize 17} &{\scriptsize 13~}{16 }{\scriptsize 18} \tabularnewline
+CDR-SB&1911&{\scriptsize 0.0~}{0.0 }{\scriptsize 0.0} &{\scriptsize 1.0~}{1.5 }{\scriptsize 2.5} &{\scriptsize 3.5~}{4.5 }{\scriptsize 6.0} &{\scriptsize 0.0~}{1.5 }{\scriptsize 3.0} \tabularnewline
+ADAS~13&1895&{\scriptsize  5.67~}{ 8.67 }{\scriptsize 12.33} &{\scriptsize 14.67~}{19.33 }{\scriptsize 24.33} &{\scriptsize 24.67~}{30.00 }{\scriptsize 35.33} &{\scriptsize 10.67~}{18.00 }{\scriptsize 25.33} \tabularnewline
+MMSE&1917&{\scriptsize 29~}{29 }{\scriptsize 30} &{\scriptsize 25~}{27 }{\scriptsize 29} &{\scriptsize 20~}{23 }{\scriptsize 25} &{\scriptsize 25~}{27 }{\scriptsize 29} \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\noindent {\scriptsize $a$\ }{$b$\ }{\scriptsize $c$\ } represent the lower quartile $a$, the median $b$, and the upper quartile $c$\ for continuous variables.\\$N$\ is the number of non--missing values.\\Numbers after percents are frequencies.\end{table*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Prepares the data in a the form needed for plotting.
-<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-all_data  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data.csv.gz'))
-all_data_mean  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data_mean.csv.gz'))
-@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}
-<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-prep', fig.width=7, fig.height=7>>=
-stderr <- function(x) sqrt(var(x)/length(x))
-ggplot(subset(all_data_mean, (templates.mb*atlases) %% 2 == 1),
-       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
-  facet_grid(reg_method~method.mb) + 
-  stat_summary(fun.y=mean,geom='line',
-               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
-  stat_summary(fun.y=mean,
-               fun.ymin=function (x) mean(x) - stderr(x),
-               fun.ymax=function (x) mean(x) + stderr(x), 
-               geom='errorbar',
-               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
-  #stat_smooth(method='lm', formula=y~log(x)) + 
-  scale_colour_hue(name="Number of Atlases") +
-  xlab( "Number of Templates" ) + 
-  ylab( "Mean DSC" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text = element_text(size = 8))
-@
-  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with 
-  SNT \gs segmentations for 69 ADNI1 subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method. Points above zero
-  indicate an \mb parameter settings yielding a higher mean DSC score than
-  multi-atlas segmentation using the same number of atlases.
-  \label{fig:ADNI1-xval-k-mean}}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/fig_ADNI1-xval-k-mean}
+
+}
+
+\end{knitrout}
+
+  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with manual
+  segmentations (SNT) for 69 ADNI1 subjects vs. atlas and template library size,
+  registration algorithm, and label fusion method. Points above zero indicate an
+  \mb parameter settings yielding a higher mean DSC score than multi-atlas
+  segmentation using the same number of atlases. \label{fig:ADNI1-xval-k-mean}}
 \end{figure}
@@
 \begin{figure}
-<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-prep'>>=
-ggplot(subset(all_data_mean, reg_method=="ANTS" & 
-                method.mb=="Majority Vote" & (templates.mb*atlases) %% 2 == 1), 
-       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
-  #facet_grid(reg_method~method.y) + 
-  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
-               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
-  stat_summary(fun.y=mean, geom="point", 
-               aes(y=k_diff,colour=as.factor(atlases))) + 
-  stat_summary(fun.y=mean, geom="line", 
-               aes(y=k_diff,colour=as.factor(atlases))) + 
-  #stat_smooth(method=lm, formula=y~log(x)) + 
-  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
-  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
-  scale_colour_hue(name="Number of Atlases") +
-  scale_alpha_continuous(guide = "none") + 
-  scale_linetype_discrete(guide = "none") + 
-  xlab( "Number of Templates" ) + 
-  ylab( "Difference in mean DSC" ) +
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/fig_ADN11-xval-k-diff}
+
+}
+
+\end{knitrout}
+
   \caption{Difference in mean Dice's Similarity Coefficient of \mb and
-  multi-atlas segmentations with \gs (SNT) segmentations for a range of atlas
+  multi-atlas segmentations with manual (SNT) segmentations for a range of atlas
   and template library sizes. \label{fig:ADNI1-xval-k-diff}}
@@
 \begin{figure}
-<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ants = subset(all_data, 
-                     reg_method=="ANTS" & 
-                     method.mb=="Majority Vote" &
-                    templates.mb %in% seq(1,20,2))
-
-ants.stats = ddply(ants, c("subject", "label", "atlases","templates.mb"), 
-                   function (df) {
-                     data.frame(
-                       MA.sd  =  sd(df$k.ma),
-                       MB.sd  =  sd(df$k.mb),
-                       MA.var = var(df$k.ma),
-                       MB.var = var(df$k.mb)
-                       )
-                   })
-
-stats.tests = ddply(ants.stats, c("atlases","templates.mb"), function (df) {
-  t.var = t.test(df$MA.var, df$MB.var)
-  t.sd  = t.test(df$MA.sd , df$MB.sd)
-  diff_mean_var = mean(df$MA.var - df$MB.var)
-  diff_mean_sd  = mean(df$MA.sd -  df$MB.sd)
-
-  # test statistic is positive if MA mean is larger than MB
-  # i.e. that MA has a greater mean of variances/SDs
-  # i.e. that MB has a smaller mean, and so is doing "better"
-  data.frame(
-    stat      = c("Variance", "Standard Deviation"),
-    statistic = c(t.var$statistic, t.sd$statistic),
-    p.value   = c(t.var$p.value, t.sd$p.value),
-    direction = c(diff_mean_var, diff_mean_sd)
-  )
-})
-
-ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
-  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
-  #facet_grid( . ~ stat) + 
-  geom_line(size=1) +
-  geom_point(size=3) + 
-  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
-  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
-  scale_x_continuous(breaks=seq(1,20,by=2)) + 
-  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
-  xlab( "Number of Templates" ) + 
-  ylab( "Difference of variance during validation (p-value)") +
-  scale_colour_hue(name="Atlases") +
-  scale_alpha_continuous(guide='none') + 
-  scale_linetype_discrete(guide='none') + 
-  scale_size_continuous(guide='none') + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/ADN1-Xval-variability}
+
+}
+
+\end{knitrout}
+
   \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-This experiment explores \mb segmentation of hippocampal subfields. To achieve
-this, a leave-one-out validation was conducted in which lower-resolution images
-($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
-remaining Winterburn atlases.
-
-In general, across hippocampal subregions the percent error in volume of \mb
-segmentations compares favourably to the error resulting from image resampling
-(Figure \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
+This experiment explores \mb segmentation of hippocampal subfields. To
+achieve this, a leave-one-out validation is conducted in which lower-resolution
+images ($0.9mm^3$ voxels) of each Winterburn atlas subject are segmented using 
+the remaining Winterburn atlases.
+
+In general, across hippocampal subregions the percent error in volume of
+\mb segmentations compares
+favourably to the error resulting from image resampling (Figure
+\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
 subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
@@
 \begin{figure}[h!]
-<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
-data = subset(read.csv('data/WAval.csv'))
-casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
-casted = rename(casted, c("variable"="region"))
-
-error = ddply(casted, c("file", "region"), function(df) {
-  with(df, 
-  data.frame(
-    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
-
-levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
-                            "Subiculum"="X2","Subiculum"="X102", 
-                            "CA4/DG"="X4","CA4/DG"="X104", 
-                            "CA2/CA3"="X5","CA2/CA3"="X105", 
-                            "SR/SL/SM"="X6","SR/SL/SM"="X106")
-melted = melt(error,id.vars=c("file", "region"))
-melted = rename(melted, c("variable"="measure", "value"="percenterr"))
-levels(melted$measure) <- list("Subsampled"="resampled",
-                               "MAGeT + Subsampled" = "mb_0.9mm",
-                               "MAGeT + BRAVO"="mb_bravo")
-
-ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
-  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
-  geom_boxplot() +
-  labs(y=expression(paste("Percent error in volume relative to 0.3", mm^3, " Winterburn Atlases")), x="Region") +
-  scale_colour_discrete("") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/WAval-vol-boxplot}
+
+}
+
+\end{knitrout}
+
   \caption{{\bf Percent error in segmentation volume by hippocampus subregion.} 
@@
 21 subject images selected at random.  Expert manual whole hippocampal
-segmentations are used as a gold standard.
+segmentations are used as gold standards.
 
 \begin{figure}
-<<FEP-volumes,cache=FALSE,dependson="setup">>=
-SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
-mean_sz_vols = ddply(SZ_volumes, c('Subject', 'Method'), function (df) {
-  data.frame(Volume = mean(df$Volume))})
-
-casted = dcast(mean_sz_vols, Subject ~ Method, value.var = "Volume")
-lm = lm(MAGeT ~ Manual, casted)
-ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
-  geom_smooth(method="lm", formula=y~x) + 
-  geom_point() + 
-  geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
-  xlab(expression("Mean Manual hippocampus volume " (mm^3))) + 
-  ylab(expression("Mean MAGeT hippocampus volume "  (mm^3))) + 
-  scale_size_continuous(guide="none") 
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/FEP-volumes}
+
+}
+
+\end{knitrout}
+
   \caption{Subject mean hippocampal volumes by \mb vs. manually segmented
@@
 
-<<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
-  fep.cor.test = cor.test(casted$Manual, casted$MAGeT)
-@
 \mb produces hippocampus segmentation volumes that are highly correlated with 
-manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
-$t = \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001,
-fep.cor.test$p.value)}$; Figure \ref{fig:SZ-volumes}).
+manual segmentation volumes (Pearson $r = 0.877$, 
+$t = 16.244$, $p < 0.001$; Figure \ref{fig:SZ-volumes}).
 
@@
 subject images (7 from each disease class) and used majority vote label fusion.
-The entire \adnidataset dataset was segmented by \mb, and the resulting volumes
-compared with those obtained by expert semi-automated segmentation (SNT), and
-three other automated segmentation techniques: MAPER, FreeSurfer, and FSL.
-Table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total
-count of segmentations available, including a count of those which have failed
-a quality control inspection.  A total of \Sexpr{length(means.complete$RID)}
-images are included in the following analysis, having met quality criteria and
-having segmentations from every method.
-
-We found a close relationship in total bilateral hippocampal volume between all
-methods and SNT \gs volumes (Figure \ref{fig:ADNI-volumes-plot}).  Volumes are
-correlated with Pearson $r > 0.78$ for all methods across disease categories.
-Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}), \mb is
-consistently well correlated to manual volumes (Pearson $r > 0.85$), but appears
-to slightly over-estimate the volume of the AD hippocampus.
-
-To investigate the level of agreement with SNT \gs hippocampal
+The entire \adnidataset dataset was segmented by \mb and we now compare the
+resulting volumes with those obtained by manual segmentation (SNT), and other
+automated segmentation techniques (MAPER, FreeSurfer, and FSL).  Table
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
+segmentations available, including a count of those which have failed a quality
+control inspection. Quality control for FreeSurfer segmentations was provided along
+with the segmentations from ADNI \cite{Hartig2010}. One of the authors (MP) 
+performed visual quality inspection for MAGeT and FSL segmentations using similar 
+quality control guidelines (if either hippocampus was under or over segmented
+by greater than 10mm in three or more slices then the segmentation did not pass).
+
+Only those images which had quality segmentations from each method are 
+included in the following analysis (a total of 361
+images).
+
+We find a close relationship in total bilateral hippocampal volume between
+all methods and manually segmented volumes (Figure \ref{fig:ADNI-volumes-plot}). 
+Volumes are correlated with Pearson $r > 0.78$ for all methods across disease
+categories.  Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}),
+\mb is consistently well correlated to manual volumes (Pearson $r > 0.85$), but
+appears to slightly over-estimate the volume of the AD hippocampus.
+
+To investigate the level of agreement with manually segmented hippocampal
 volumes, we constructed Bland-Altman plots for each method (Figure
@@
 Figure \ref{fig:ADNI-subfields} shows a qualitative comparison of \mb and 
-SNT \gs hippocampal segmentations for 10 randomly selected subjects in 
+manual (SNT) hippocampal segmentations for 10 randomly selected subjects in 
 each disease category, and illustrates some of the common errors found during 
@@
 hippocampal sulcus and, although not anatomically incorrect, \mb under-estimates 
-the hippocampal body in comparison to the SNT \gs segmentation.
+the hippocampal body in comparison to the manual (SNT) segmentation.
 
@@
 %                        
-<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
-package_totals$Total <- NULL
-latex(package_totals, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
-    caption="\\textbf{Number of segmented images and quality control failures of 
-             \\adnidataset dataset by method.}", 
-    caption.loc = c("top"), rowname=NULL, 
-    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
-@
+% latex.default(package_totals, file = "", size = "scriptsize",      landscape = FALSE, ctable = TRUE, caption = "\\textbf{Number of segmented images and quality control failures of \n             \\adnidataset dataset by method.}",      caption.loc = c("top"), rowname = NULL, label = "tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
+%
+{\scriptsize\ctable[caption={\textbf{Number of segmented images and quality control failures of 
+             \adnidataset dataset by method.}},label=tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals,pos=!tbp,]{llrlrr}{}{\FL
+\multicolumn{1}{c}{X}&\multicolumn{1}{c}{SNT}&\multicolumn{1}{c}{MAGeT}&\multicolumn{1}{c}{MAPER}&\multicolumn{1}{c}{FSL}&\multicolumn{1}{c}{FS}\ML
+Images&1445&$1909$&636&$1876$&$1530$\NN
+Failures&--&$  34$&--&$  27$&$ 304$\LL
+}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[h]
-<<ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
-melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
-              variable.name="Method", value.name="Volume")
-correlations = ddply(melted, c("DX","Method"), function (df) {
-  data.frame(
-    pearson = cor(df$SNT, df$Volume)
-  )
-})
-m2 = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT","SNT"), variable.name="Method")
-levels(m2$Method) = c("FS"    = "FreeSurfer", 
-                      "FSL"   = "FSL", 
-                      "MAPER" = "MAPER", 
-                      "MAGeT" = "MAGeT", 
-                      "SNT"   = "SNT \\gs")
-m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
-
-qplot(DX,value,data=m2,
-      colour=Method,geom="boxplot") + 
-      #geom_text(aes(y=-Inf,x=c(.7,.85,1,1.15, 1.7,1.85,2,2.15, 2.7,2.85,3,3.15), vjust=-5, label=round(pearson,2)), 
-      #          colour = 'black', size=3, data=correlations) +
-      xlab("Diagnosis") + 
-      ylab(expression(paste("Hippocampal volume (", mm^3, ")"))) + 
-      theme(legend.direction = "horizontal", legend.position = "bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/ADNI-volumes-boxplot}
+
+}
+
+\end{knitrout}
+
   \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by FreeSurfer, FSL, MAPER, \mb, and SNT \gs vs. disease
+dataset by FreeSurfer, FSL, MAPER, \mb, and manual raters (SNT) vs. disease
 category.}
@@
 \begin{figure}[h]
-<<ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
-melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
-            variable.name = "Method", value.name = "volume")
-
-df = means.complete
-
-ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
-    geom_point(size=1) + 
-    geom_smooth(method="lm") + 
-    xlab(expression("SNT \\gs hippocampal volume " (mm^3))) + 
-    ylab(expression("Automated hippocampal volume " (mm^3))) +
-    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
-              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
-                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
-                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
-                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
-              hjust=1.1,vjust=1.5) +
-    theme(legend.direction = "horizontal", legend.position = "bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/ADNI-volumes-plot}
+
+}
+
+\end{knitrout}
+
   \caption{Subject mean hippocampal volume as measured in the \adnidataset
 dataset by each of the four automated methods investigated (FreeSurfer (FS),
-FSL, MAPER, \mb) vs. SNT \gs. Linear fit lines and pearson
+FSL, MAPER, \mb) vs. manual rating (SNT). Linear fit lines and pearson
 correlations are shown for each method.}
@@
 \begin{figure}
-<<ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
-names(means.complete)[8] = "FreeSurfer"
-melted=melt(means.complete, measure.vars=c("FreeSurfer", "FSL", "MAPER","MAGeT"), 
-            variable.name = "method", value.name = "volume")
-
-melted$diff = melted$SNT - melted$volume
-melted = subset(melted, (method == "FreeSurfer") | 
-                        (method == "FSL"  ) | # & diff > -5000 & diff < 900) | 
-                        (method == "MAPER") | # & diff > -500 & diff < 1000) | 
-                        (method == "MAGeT"))  # & diff > -1100))
-#melted = subset(melted, method == "MAGeT")
-melted$DX = factor(melted$DX, levels(melted$DX)[c(2,3,1)])
-melted$mean = ( melted$SNT + melted$volume ) / 2
-limits = ddply(melted, c("method"), function (df) { 
-  data.frame(
-    y = mean(df$diff) + c(-1.96,0,2) * sd(df$diff) 
-)})
-
-names(melted)[4] = "Diagnosis"
-ggplot(melted, aes(x= mean, y = diff, colour=Diagnosis)) + 
-  facet_wrap( ~ method, nrow=2, scales="fixed") + 
-  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=1) +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
-             data=limits, size=0.3, alpha=0.5) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits,
-            hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of SNT \\gs and automated hippocampal volume ' (mm^3))) + 
-  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
-  scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
-  scale_x_continuous(breaks=seq(0,10000, 1000)) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text.y = element_text(angle = 90, hjust=0.5), 
-        axis.text = element_text(size = 8))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/ADNI-Bland-Altman}
+
+}
+
+\end{knitrout}
+
   \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset dataset by SNT \gs segmentation and each of the
-four automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall
-mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
-dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
-Note, points below the mean difference indicate overestimation of the volume
-with respect to the \gs rating, and vice versa. }
+measured in the \adnidataset dataset by manual raters (SNT) and each of the four
+automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall mean
+difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
+horizontal lines. Linear fit lines are shown for each diagnosis group. Note,
+points below the mean difference indicate overestimation of the volume with
+respect to the manual rating, and vice versa. }
 
@@
   \end{centering}
-  \caption{SNT \\gs and MAGeT segmentations for 30 ADNI1 subjects (10
+  \caption{Manual and MAGeT segmentation results for 30 ADNI1 subjects (10
   subjects randomly selected from each disease category in the subject pool used
   in Experiment 1). Sagittal slices are shown for subject unlabelled T1-weighted
-  anatomical image, SNT \gs label (green), and \mb label (blue). Noted are
+  anatomical image, SNT manual label (green), and \mb label (blue). Noted are
   examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
-  {\em (b)} translated SNT \gs segmentation; 
+  {\em (b)} translated manual segmentation by SNT; 
   {\em (c)} under-estimation of hippocampal body and
@@
 %
-<<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
-# caption = 
-#   "\\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
-#    ADNI dataset.} For each method, the number of manually labelled atlases used for 
-#    training, the best Dice's overlap measure, the disease classes measured, and 
-#    the validation procedure are shown. LOOCV = Leave-one-out cross-validation. 
-#    Some studies of automated segmentation of ADNI images are excluded because they 
-#    do not not provide overlap measures for the hippocampus \\cite{Heckemann2011, Chupin2009}.
-#   "
-# 
-# tab = read.csv("data/ADNI-existing-work.csv")
-# tab$Notes <- NULL
-# 
-# adni_lit = subset(tab, Dataset=="ADNI")
-# adni_lit$Dataset <- NULL
-# 
-# 
-# latex(adni_lit, #file="", 
-#       size="scriptsize",
-#       caption=caption, 
-#       rowname=NULL,
-#       col.just=c("p{2in}","p{1in}","p{1in}","p{1in}","p{1in}"),
-#       title="")
-@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-<<mb-best-kappa, echo=F, cache=FALSE>>=
-ants.mean = subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &      templates.mb == 19 & atlases == 9)
-mb.best.kappa = mean(ants.mean$k.mb)
-@
-
-\begin{table}[!tbp]
+
+\begin{table*}[!tbp]
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy (overlap with SNT \gs labels)
-of the ADNI dataset.} For each method, the number of labelled atlases
-used for training, the best Dice's overlap measure, the disease classes
-measured, and the validation procedure are shown. Unless specified, validation
-datasets are composed equally of subjects diagnosed with Alzheimer's Disease
-(AD), Mild Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV =
-Leave-one-out cross-validation. 
+\caption{\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
+   ADNI dataset.} For each method, the number of manually labelled atlases used for 
+   training, the best Dice's overlap measure, the disease classes measured, and 
+   the validation procedure are shown. Unless specified, validation datasets are
+   composed equally of subjects diagnosed with Alzheimer's Disease (AD), Mild 
+   Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV = Leave-one-out 
+   cross-validation. 
    Some studies of automated segmentation of ADNI images are excluded because they 
@@
 \hline
-\mb                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
+\mb                     &9 & 0.838 & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
 LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
 available.  For example, other groups have used the STAPLE algorithm
-\citep{Warfield2004} (or variants of the STAPLE algorithm
-\citep{Robitaille2012}) which weights each segmentation based upon its
-estimated performance level with respect to the other available candidate
-segmentations.  Further, the sensitivity and specificity parameters could also
-be tuned to potentially improve segmentation accuracy and reliability.  It is
-likely that using more sophisticated voting methods would have a positive
-effect on the overall segmentation performance, as demonstrated by the STAPLE
-algorithm.
+\citep{Warfield2004} (or variants of the STAPLE algorithm \citep{Robitaille2012})
+which weights each segmentation based upon its estimated
+performance level with respect to the other available candidate segmentations.
+Further, the sensitivity and specificity parameters could also be tuned to
+potentially improve segmentation accuracy and reliability.  It is rather likely
+that using more sophisticated voting methods would have a positive effect on the
+overall segmentation performance.
 
 To this end, more work is required in order to determine the source of the
-slight decrease in segmentation performance when the number of templates are
-set to an even number.  Our initial concern was that this dip in performance
-was a by-product of the \mb algorithm itself.  However, this this pattern is
-also found in the results of the multi-atlas segmentations we used in our
-experiments.  We believe that our majority voting methodology is biased towards
-labels with the lowest numeric values when breaking ties (by way of the
-implementation of the \verb+mode+ function used to determine majority), thus
-causing the slight bias observed when using an even number of templates. This
-is another area where the voting scheme could be used to improve performance.
-However, it is worth noting that this limitation was previously identified by
+slight decrease in segmentation performance when the number of templates are set
+to an even number.  Our initial concern was that this dip in performance was a
+by-product of the \mb algorithm itself.  However, we determined that
+this pattern was also true in the analysis of multi-atlas segmentations we used
+in our experiments.  We believe that our majority voting methodology is biased
+towards labels with the lowest numeric values when breaking ties, thus causing
+the slight bias observed when using an even number of templates. This is another
+area where the voting scheme could be used to improve performance. However, it
+is worth noting that this limitation was previously identified by
 \citet{Heckemann2006a} and, subsequently, other groups have not even considered
@@
 multi-atlas/segmentation literature.  To the best of our knowledge, other groups
-using ADNI data for validation do at most a single round of
-leave-one-out-validation (Table \ref{tab:other-methods}), aside from
-\citep{Wang2011} who perform a similar validation as us). The thoroughness of
-our validation suggests that our results are reflective of a true average over
-the choice of parameter settings and are independent of atlas or template
-choice (provided the input atlases are properly segmented.
+using ADNI data for validation do at most a single round of leave-one-out-validation
+(Table \ref{tab:other-methods}), aside from \citep{Wang2011} who perform a 
+similar validation as us. The thoroughness of our validation suggests that
+our results are reflective of a true average over the choice of parameter
+settings and are independent of atlas or template choice (provided the input
+atlases are properly segmented).
 
@@
 probabilistic methodology using a cohort of 8 healthy controls and 15 epilepsy
-patients, and then use this method to segment an ADNI sample, with a
-hierarchical experimentation protocol that involved \todo{complete this
-sentence}.  These methods suggest that some post-processing of the final
-segmentations would improve accuracy of the segmentation. While that may be
-true, there is little consensus regarding how to achieve this.
+patients.  In subsequent methods, they retuned their methodology using the ADNI
+sample with a hierarchical experimentation protocol that involved.  These
+methods suggest that some post-processing of the final segmentations would
+improve accuracy of the segmentation.  While that may be true, there is little
+consensus regarding how to achieve this.
 
@@
 volumes we see that large hippocampi are over estimated while small hippocampi
-are under estimated. By comparison, \mb and MAPER are far more conservative in
-volume estimation, suggesting these methods may be better suited for estimating
-true-positives. However, in this analysis we have only compared methods by
-total hippocampal volume, and so more work is needed to understand the full
-extent to which these methods differ.
-
-Finally, we have also demonstrated that our algorithmic framework is
-appropriate for the segmentation of hippocampal subfields in standard
-T1-weighted data.  Subfield segmentation is a burgeoning topic in the
-literature although very few automated methods are available for the
-segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. However, the
-FreeSurfer implementation of subfield segmentation \todo{complete}. While
-recent work demonstrates that subfield segmentations can be used for
-classification of AD, MCI, and NC, there has been no explicit validation of the
-methodology based on accuracy or precision.  Although the initial hippocampal
-subfield segmentations from the Yushkevich group have been demonstrated to work
-on the ADNI population, there has also been no validation of their work.  In
-addition, their work requires some manual initialization to properly function.
-Our work demonstrates that we can reliably identify the CA1, subiculum, and CA4
-dentate with only modest amounts of error.  The fact that CA2/CA3 and molecular
-layers cannot be reasonably identified should not be surprising as these are
-extremely thin and spatially convoluted regions that originally required
-high-resolution MRI for identification.  It is likely that extents of these
-regions are well below the resolution offered by standard T1-weighted images
-and, in fact, many manual segmentation methodologies do not attempt to parse
-these regions either \citep{Wisse2012, Mueller2009}.
-
+are under estimated.  By comparison, our method and MAPER are far more
+conservative suggesting that these methods may be better suited for estimating
+true-positives.  While this comparison with these methods using only volume,
+more work is needed to better understand the differences between our method and
+other methods.
+
+Finally, we have also demonstrated that our algorithmic framework is appropriate
+for the segmentation of hippocampal subfields in standard T1-weighted data.
+This has started to become a burgeoning topic in the segmentation literature,
+although very few methods are available for the segmentation of 3T data
+\citep{Yushkevich2009,VanLeemput2009}. However, the FreeSurfer implementation of
+subfield segmentation. While recent work demonstrates that subfield
+segmentations can be used for classification of AD, MCI, and NC, there has been
+no explicit validation of the methodology based on accuracy or precision.
+Although the initial hippocampal subfield segmentations from the Yushkevich
+group have been demonstrated to work on the ADNI population, there has also been
+no validation of their work.  In addition, their work requires some manual
+initialization to properly function.  Our work demonstrates that we can reliably
+identify the CA1, subiculum, and CA4 dentate with only modest amounts of error.
+The fact that CA2/CA3 and molecular layers cannot be reasonably identified
+should not be surprising.  These are extremely thin and spatially convoluted
+regions that originally required high-resolution MRI for identification.  It is
+likely that they would be that extents of these regions are well below the
+resolution offered by standard T1-weighted images.  In fact many manual
+segmentation methodologies do not attempt to parse these regions either
+\citep{Wisse2012, Mueller2009}
+
+\subsection{Conclusion}
 In conclusion, we have presented a flexible multi-atlas framework.  It has
@@
 
-% \begin{verbatim}
-% Experiment 1: 
-% - address the absence of weighted-voting effects. 
-%   - hypothesis: the benefits of weighted voting only outweigh the resampling
-%     error effects when choosing from a large library.  In this experiment, the 
-%     template library (during weighted voting) consists of only 20 templates, so
-%     the most we could hope for is a "squashing" upwards of the curve towards 
-%     the 20-template limit (which is the same across all cases in this experiment).
-%   - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
-%     Perhaps in future experiments with larger template libraries, this could be
-%     explored further.
-% - Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
-%   - couldn't get STAPLE to work with our image formats
-%   - Idea: expect smoothness across range of templates (no even # dips, below)
-%   - perhaps more sophisticated fusion methods would boost results over majority 
-%     vote based techniques.
-% 
-% - why does MAGeT brain show a dip in average Kappa when using an even number
-%   templates. Does MA show this same pattern? (yes)
-%   - Hypothesis: our voting method is biased when breaking ties to choose the 
-%     label with the lowest numeric value.
-%   - If we compute the total number of labels fused (atlases * templates) then 
-%     both MB/MA perform worse when fusing an even number of labels. 
-% 
-% - we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
-%   using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
-%   this increase justify the extra effort involved in MB? 
-%   - decrease in variability
-%   - we are comparing "true" averages (see above)... does this make our criteria 
-%     for a worthy improvement less strict? 
-%     
-% 
-% - how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
-% - because of the extensive cross-validation in experiment 1, we are very likely
-%   showing results that approach the true average of MB and MA on that dataset
-%   (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
-%   each).
-% - even with this caveat, do we think we do well enough?  i.e. other than parameter 
-%   tuning, to what extent does this experiment tell us about how MB would do in 
-%   practice. 
-% 
-% - effects we are seeing is not only averaging effects (Chakravarty et al. 2012). 
-% 
-% - take aways: as in other studies (Aljabar, Heckemann) we find that performance
-%   scales with the number of inputs (approx. log(n)), and that a large enough
-%   template library can boost performance with a small number of atlases. The use 
-%   of templates /can/ have a negative impact on performance (Figure 2), that is 
-%   balanced by the improvement of growing the template library.  Likely the negative
-%   impact on peformance is as a result of resampling/mislabelling error. In other 
-%   words, tuning MAGeT brain involves balancing the tension between an improvement 
-%   in peformance due to increase neuroanatomical capture and decreased peformance
-%   due to resampling error. 
-% 
-% Experiment 2: 
-% - describe resampling error during downsampling (essentially partial volume effect 
-%   due to averaging/majority vote in nearest neighbour selection)
-% - Why would some structures have greater downsampling error than others? (i.e. what
-%   is it about a structure that would make it especially prone to resampling error)
-%   - since error is discretised to whole voxels, smaller regions will show larger
-%     percent error
-%   - average unilateral volume (approx; mm3): 
-%      region   volume   downsampling error
-%      CA1      800            2%
-%      CA4/DG   600          -30%
-%      SR/SL/SM 700           30%
-%      Sibculum 350            0%
-%      CA2/3    200           15%
-% 
-% - observation: when downsampling error is small (<10 percent), MB error is larger, 
-%   and vice versa when downsampling error is large.  Could this a case where the 
-%   inevitable MAGeT resampling error outweighs the small downsampling error?
-% - observation: MAGeT produces similar volumes for the downsampled and BRAVO images
-%   which demonstrates reliability. Additionally, MB's error is in the same *direction*
-%   as the downsampling error except for SR/SL/SM
-% - for MB to show less error than resampling means that voting across templates is 
-%   in aggregate performing better than local nearest neighbour fitting. presumably 
-%   we'd see the same improvement with basic multiatlas as well.
-% 
-% - take away: MB produces subregion volumes that are comparable or better than
-%   resampling error, except for the CA2/3 where error is near 25 percent (but 
-%   even then resampling error is ~15%)
-% 
-% Experiment 3: 
-% - what is an "acceptable" r^2 value? 
-% - take away: MB has proven to be robust with atlases derived from three different
-%   segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
-%   populations (older with AD progression; young and healthy; young and SZ). 
-%   
-% Experiment 4: 
-% - address the smaller difference in mean volume across disease classes. smaller 
-%   than every other method. 
-%   
-% - address the segmentation bias towards over-estimating smaller hippocampi, and 
-%   underestimating larger HC
-%   
-% - visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
-%   is lower than the other methods (but we may be biased. :-). 
-%   
-% - take away: MB, with the Winterburn atlases, produces HC volumes more inline with
-%   SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required. 
-% 
-% \end{verbatim}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{ADNI SNT Hippocampal Labels}
+\subsection{ADNI Manual Labels}
 % The following blurb is taken (except for the first sentence) verbatim 

### 2013-08-26 paper.Rnw
Jen's modifications

@@
 
-The hippocampus is a neuroanatomical structure situated in the medial temporal
-lobe of the brain, and has long been associated with learning and memory
-\citep{DenHeijer2012,Scoville2000}. In addition to its known functional roles,
-the hippocampus is of interest to neuroscientists because it is implicated in
-several forms of brain disfunction such as Alzheimer's disease
-\citep{Sabuncu2011} and schizophrenia \citep{Narr2004,Karnik-Henry2012}. In
-neuroimaging experiments, magnetic resonance images (MRI) are often used for the
-identification of the hippocampus. As such, accurate segmentation of the
-hippocampus and its subregions in MRI is a necessary first step to better
-understand the unique neuroanatomy of subjects. Typically, the gold standard for
-neuroanatomical segmentation is manual delineation by an expert human rater.
-However with the availability of increasingly large MRI datasets, the time and
-expertise required for manual segmentation becomes prohibitive
+The hippocampus is a brain structure situated in the medial temporal
+lobe, and has long been associated with learning and memory
+\citep{DenHeijer2012,Scoville2000}. The hippocampus is of interest to
+neuroscientists because it is implicated in many forms of brain disfunction,
+including Alzheimer's disease \citep{Sabuncu2011} and schizophrenia
+\citep{Narr2004,Karnik-Henry2012}. In neuroimaging studies, structural magnetic
+resonance images (MRI) are often used for the volumetric assessment of the
+hippocampus.  As such, accurate segmentation of the hippocampus and its
+subregions in MRI is a necessary first step to better understand the
+inter-individual variability of subject neuroanatomy. Typically, the gold
+standard for neuroanatomical segmentation is manual delineation by an expert
+human rater.  However, with the availability of increasingly large MRI datasets,
+the time and expertise required for manual segmentation becomes prohibitive
 \citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. Further, there is
-little agreement between researchers regarding how exactly the hippocampus
-should be identified in MRI images \citep{Geuze2004} and this has led to efforts
-to create an unified hippocampal segmentation protocol \citep{Jack2011}.
+significant variation between segmentation protocols regarding specific
+anatomical boundaries of the hippocampus \citep{Geuze2004} and this has led to
+efforts to create an unified hippocampal segmentation protocol \citep{Jack2011}.
 
 Automated segmentation techniques that are reliable, objective, and reproducible
-are a necessary alternative to manual segmentation. In the case of classical
+can be considered complementary to manual segmentation. In the case of classical
 model-based segmentation methods \citep{Haller1997,Csernansky1998}, an MRI atlas
@@
 
-To the best of our knowledge there are two other segmentation methods that
+To the best of our knowledge there are only two other segmentation methods that
 attempt to define the hippocampus and its subfields using standard T1-weighted
@@
 demonstrates that the applicability of hippocampal subfield segmentation using
-data from 10 subjects. In the second method, \citet{Yushkevich2009} acquired
-and labelled hippocampal subfields on high-resolution MRI data from post-mortem
+data from 10 subjects. In the second method, \citet{Yushkevich2009} hippocampal
+subfields were acquired and labelled on high-resolution MRI data from post-mortem
 medial temporal lobe samples. Using nonlinear registration guided by manually
-derived hippocampus masks and specific landmarks, they demonstrate accurate
-parcellation of hippocampal subfields in unlabelled MRI volumes.
+derived hippocampus masks and specific landmarks, the authors demonstrated
+accurate parcellation of hippocampal subfields in unlabelled MRI volumes.
 

### 2013-08-28 paper.Rnw
Jason's comments mostly responded to.

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 
-There are few methods that have attempted to perform multi-atlas segmentation
-with a limited number of input atlases. The LEAP algorithm is an elegant
-modification to the basic multi-atlas strategy \citep{Wolz2010} in which the
-atlas library is grown, beginning with a set of manually labelled atlases, and
-successively incorporating unlabelled target images after themselves being
-labelled using multi-atlas techniques. The sequence in which target images are
-labelled is chosen so that the similarity between the atlas images and the
-target images is minimised at each step, effectively allowing for deformations
-between very dissimilar images to be broken up into sequences of smaller
-deformations. Although \citet{Wolz2010} begin with an atlas library of 30 MR
-images, this method could theoretically work using a much smaller atlas library.
-In their validation, LEAP was used to segment the whole hippocampus in the
-ADNI-1 baseline dataset, achieving a mean Dice score of 0.85 with manual
-segmentations.
-
-To the best of our knowledge there are only two other segmentation methods that
-attempt to define the hippocampus and its subfields using standard T1-weighted
-data. The first is included with the FreeSurfer package \citep{VanLeemput2009}.
-This work is limited as it omits the tail of the hippocampus and the
-segmentation protocol has yet to be fully validated. Nonetheless, it
-demonstrates that the applicability of hippocampal subfield segmentation using
-data from 10 subjects. In the second method, \citet{Yushkevich2009} hippocampal
-subfields were acquired and labelled on high-resolution MRI data from post-mortem
-medial temporal lobe samples. Using nonlinear registration guided by manually
-derived hippocampus masks and specific landmarks, the authors demonstrated
-accurate parcellation of hippocampal subfields in unlabelled MRI volumes.
+Of special relevance to the present work is the LEAP algorithm \citep{Wolz2010}
+because of its focus on performing multi-atlas segmentation with a limited
+number of input atlases.  The LEAP algorithm is a clever modification to the
+basic multi-atlas strategy in which an atlas library is grown, beginning with a
+set of manually labelled atlases, by successively incorporating unlabelled
+target images once they themselves have been labelled using multi-atlas
+techniques. The sequence in which target images are labelled is chosen so that
+the similarity between the atlas images and the target images is minimised at
+each step, effectively allowing for deformations between very dissimilar images
+to be broken up into sequences of smaller deformations. Although
+\citet{Wolz2010} begin with an atlas library of 30 MR images, this method could
+theoretically work using a much smaller atlas library.  In their validation,
+LEAP was used to segment the whole hippocampus in the ADNI-1 baseline dataset,
+achieving a mean Dice score of 0.85 against \gs segmentations.
+
+To the best of our knowledge there are only two automated segmentation methods
+that attempt to define the hippocampus and its subfields using standard
+T1-weighted data. The first is included with the FreeSurfer package
+\citep{VanLeemput2009}.  This work is limited as it omits the tail of the
+hippocampus and the segmentation protocol has yet to be fully validated.
+Nonetheless, it demonstrates that the applicability of hippocampal subfield
+segmentation using data from 10 subjects. In the second method,
+\citet{Yushkevich2009} hippocampal subfields were acquired and labelled on
+high-resolution MRI data from post-mortem medial temporal lobe samples. Using
+nonlinear registration guided by manually derived hippocampus masks and specific
+landmarks, the authors demonstrated accurate parcellation of hippocampal
+subfields in unlabelled MRI volumes.
 
@@
 In this paper, we will use the term {\em label} to mean any segmentation (manual or 
-derived) of an MR image. {\em Label propagation}, or {\em labelling}, is the 
-process by which two images are registered and the resulting transformation 
-is applied to the labels from one image to bring them into alignment with the 
-other image. We will use the term {\em atlas} to mean a manually segmented image,
-and the term {\em template} to mean an automatically segmented image (i.e. 
-via label propagation). The terms {\em atlas library} and {\em template library} 
-describe any set of such images. Additionally, we will use the term {\em target} 
-to refer to an unlabelled image that is undergoing segmentation.
+derived) of an MR image. {\em Label propagation} is the process by which two
+images are registered and the resulting transformation is applied to the labels
+from one image to bring them into alignment with the other image. We will use
+the term {\em atlas} to mean a manually segmented image, and the term {\em
+template} to mean an automatically segmented image (i.e.  via label
+propagation). The terms {\em atlas library} and {\em template library} describe
+any set of such images. Additionally, we will use the term {\em target} to refer
+to an unlabelled image that is undergoing segmentation.
 
@@
 \citep{Heckemann2006, Heckemann2011}. This method is described in detail in
-other publications \citep{Collins2010,Heckemann2011,Aljabar2009}.
+other publications \citep{Collins2010,Heckemann2011,Aljabar2009}. When only a
+single atlas is used, basic multi-atlas segmentation degenerates into
+model-based segmentation: labels are propogated from the atlas to a target, and
+no label fusion is needed.
 
@@
 Label fusion is a term given to the process of combining the information from
-several candidate labellings for an intensity image into a single labelling.  In
+several candidate labels for an intensity image into a single labelling.  In
 this experiment we explore three fusion methods. 
@@
   output voxel is given the most frequent label at that voxel location amongst
-  all candidate labellings.
+  all candidate labels.
 
@@
 segmentations of five in-vivo $300\mu$ isotropic T1-weighted MR images. The
-segmentations include subfield segmentations for the cornus ammonis (CA) 1, CA4,
+segmentations include subfield segmentations for the cornu ammonis (CA) 1, CA4,
 dentate gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn
@@
 design in order to compare \mb segmentation accuracy with the accuracy of image
-resampling error. The five $300\mu$-isotropic voxel Winterburn atlas images and
+resampling error. The five $0.3mm$-isotropic voxel Winterburn atlas images and
 labels are used as the atlas library, and LOOCV is conducted using the five Winterburn
@@
 0.9mm-isotropic voxel resolution (referred to as the {\em Subsampled} dataset)
-as input subjects. Subsampling the subject images allows us to assess \mb in a
+as input subjects. Subsampling the target images allows us to assess \mb in a
 typical segmentation scenario (high-resolution atlases and lower-resolution
@@
 
-To validate that \mb works effectively in the context of other neuropsychiatric
+To validate that \mb works effectively in the context of other neurological
 disorders, in this experiment we use the Winterburn atlases to predict the
@@
 <<ADNI-volumes-prep,echo=F,cache=T>>=
-means.complete   <- read.csv(gzfile('data/cache/ADNI1:means.complete.csv.gz'))
-package_totals   <- read.csv(gzfile('data/cache/ADNI1:package_totals.csv.gz'))
+means.complete   <- read.csv('data/cache/ADNI1:qc.csv')
+package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
 @
@@
   indicate \mb parameter settings yielding a higher mean DSC score than
-  multi-atlas segmentation using the same number of atlases.
+  multi-atlas segmentation using the same number of atlases. Error bars
+  indictate standard error.
   \label{fig:ADNI1-xval-k-mean}}
@@
   multi-atlas segmentations with \gs (SNT) segmentations for a range of atlas
-  and template library sizes. \label{fig:ADNI1-xval-k-diff}}
+  and template library sizes. Error bars indicate standard error.
+  \label{fig:ADNI1-xval-k-diff}}
 \end{figure}
@@
 @
-  \caption{Subject mean hippocampal volumes by \mb vs. manually segmented
-   volumes from the First Episode Patients with Schizophrenia dataset. A linear fit
-   and standard error are shown.}
+  \caption{Mean hippocampus volume as measured by \mb  vs. manually segmented
+  volumes from the First Episode Patients with Schizophrenia dataset. A linear
+  fit line is shown, with shaded region showing standard error. }
   \label{fig:SZ-volumes}
@@
 
+levels(melted$Method) = c("FS"    = "FreeSurfer", 
+                          "FSL"   = "FSL", 
+                          "MAPER" = "MAPER", 
+                          "MAGeT" = "MAGeT")
+
 ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
@@
     geom_smooth(method="lm") + 
-    xlab(expression("SNT \\gs hippocampal volume " (mm^3))) + 
-    ylab(expression("Automated hippocampal volume " (mm^3))) +
+    xlab(expression("SNT \\gs mean hippocampus volume " (mm^3))) + 
+    ylab(expression("Automated mean hippocampal volume " (mm^3))) +
     annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,

### 2013-08-28 cache_analysis.R

@@ -75,6 +75,11 @@ write.csv(all_data_mean,'data/cache/ADNI-XVAL:all_data_mean.csv')
 # ADNI1-Complete
 # Prepares the data in a the form needed for plotting.
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+library(Hmisc)
+library(reshape2)
+library(ADNIMERGE)
+library(plyr)
+
 #bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
 mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
 maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
@@ -113,29 +118,37 @@ combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", al
 combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
 combined = merge(combined, qc          , by=c("Image.Data.ID", "RID"), all.x=TRUE)
 
-# Step 2: Of those, get counts for the numbers that fail QC
+# Step 2: Compute the mean hippocamus volume for each method
+attach(combined)
+means = data.frame(RID      = RID, 
+                   VISCODE  = VISCODE, 
+                   DX       = factor(DX.bl),
+                   MAGeT    = (MAGeT_L + MAGeT_R)/2, 
+                   MAPER    = (MAPER_L + MAPER_R)/2,
+                   SNT      = (SNT_L + SNT_R)/2, 
+                   FS       = (FS_L + FS_R/2),
+                   FSL      = (FSL_L + FSL_R)/2, 
+                   TEMPQC   = TEMPQC, 
+                   MAGeT_QC = MAGeT_QC, 
+                   FSL_QC   = FSL_QC)
+detach(combined)
+
+# Step 3: Filter images missing a measurement 
+complete = na.omit(means)
+
+# Step 4: Of those, get counts for the numbers that fail QC
+attach(complete)
 package_totals = data.frame(
-  Total     = c(length(subjects$VISCODE), 'N/A'),
-  SNT        = c(length(which(!is.na(combined$SNT_L)))    , "--"),
[…]

### 2013-08-31 paper.Rnw
Added Winterburn Atlas segmentation montage figure

@@
 \usepackage{authblk}             % Listing Author affiliations
+\usepackage{rotating}            % rotating figures
 \usepackage[hypcap]{caption}
@@
 
-\begin{figure}
+\begin{sidewaysfigure}
     \begin{centering}
-    \includegraphics[width=6in]{figure/subfield-montage.pdf}
+      \includegraphics[width=\linewidth]{figure/winterburn-atlas-montage/figure.pdf}
     \end{centering}
-  \caption{Sagittal slices from two subjects showing a comparison of the
-  original Winterburn atlas subfield segmentations (at 0.3mm-isotropic voxel
-  resolution), the subsampled Winterburn segmentations (at 0.9mm-isotropic
-  voxel resolution), and the \mb labels on the subsampled atlas image.}
+    \caption{Detailed subfield segmentation results of a single subject comparing the
+    original Winterburn atlas subfield segmentation (at 0.3mm-isotropic voxel
+    resolution), the Winterburn segmentation resampled (at 0.9mm-isotropic
+    voxel resolution), the \mb segmentation of the subsampled atlas image, and
+    the \mb segmentation of a separate T1 BRAVO acquisition of the subject.}
     \label{fig:subfield-montage}
-\end{figure}
-
-\todo[inline]{Figure \ref{fig:subfield-montage}. Show MAGeT segmentations on
-BRAVO images.  Also, show coronal/transverse slices and MNI coordinates}
+\end{sidewaysfigure}
 
@@
   \end{centering}
-  \caption{SNT \\gs and MAGeT segmentations for 30 ADNI1 subjects (10
+  \caption{SNT \gs and MAGeT segmentations for 30 ADNI1 subjects (10
   subjects randomly selected from each disease category in the subject pool used
@@
 \end{figure}
-
-\todo[inline]{Figure \ref{fig:ADNI-subfields}. Show coronal/transverse slices
-and MNI coordinates}
 
@@
 
-\todo[inline]{could we be seeing a ceiling effect... reaching the limit of what
-is achievable with a multi-atlas-based segmentation method}
-
 In comparison to other methodologies in the field \mb performs favourably. Table

### 2013-08-31 paper.Rnw
Montage labels now overlayed, not Plus'ed

@@
 \begin{figure}
-<<FEP-volumes,cache=FALSE,dependson="setup">>=
+<<FEP-volumes,cache=TRUE,dependson="setup">>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")

### 2013-09-01 paper.Rnw
Changes from Martin Lepage

@@
 % title and authors
-\title{Bootstrapping Multi-atlas Hippocampal Segmentation with \mb}
+\title{Bootstrapping Multi-atlas Hippocampal Subfield Segmentation by using
+Multiple Automatically Generated Templates}
 \author[1]{Jon Pipitone}
@@
 \author[1]{Julie Winterburn}
-\author[1,7]{Tristram A. Lett}
+\author[1,8]{Tristram A. Lett}
 \author[2,3]{Jason P. Lerch}
-\author[4]{Jens C. Pruessner}
+\author[4,5]{Jens C. Pruessner}
 \author[4]{Martin Lepage}
-\author[1,5,7]{Aristotle N. Voineskos}
-\author[1,5,6]{M. Mallar Chakravarty}
+\author[1,6,8]{Aristotle N. Voineskos}
+\author[1,6,7]{M. Mallar Chakravarty}
 \author[ ]{Alzheimer's Disease Neuroimaging Initiative}
@@
 Canada}
-\affil[4]{Brain Imaging Group, Douglas Hospital Research Centre, Verdun, QC, Canada}
-\affil[5]{Department of Psychiatry, University of Toronto, Toronto, ON, Canada}
-\affil[6]{Institute of Biomedical Engineering, University of Toronto, Toronto,
+\affil[4]{Douglas Mental Health University Institute, Verdun, QC, Canada}
+\affil[5]{Department of Psychiatry, McGill University, Montreal, QC, Canada}
+\affil[6]{Department of Psychiatry, University of Toronto, Toronto, ON, Canada}
+\affil[7]{Institute of Biomedical Engineering, University of Toronto, Toronto,
 ON, Canada}
-\affil[7]{Institute of Medical Science, University of Toronto, Toronto, ON, Canada}
+\affil[8]{Institute of Medical Science, University of Toronto, Toronto, ON, Canada}
 
@@
 intervention service at the Douglas Mental Health University Institute in
-Montreal, Canada. People aged 15 to 30 years from the local catchment area
+Montreal, Canada. People aged 14 to 35 years from the local catchment area
 suffering from either affective or non-affective psychosis who had not taken

### 2013-09-01 paper.Rnw
Added ADNI boilerplate

@@
 \author[1,6,7]{M. Mallar Chakravarty}
-\author[ ]{Alzheimer's Disease Neuroimaging Initiative}
+\author[ ]{and the Alzheimer's Disease Neuroimaging Initiative\footnote{Data
+used in preparation of this article were obtained from the Alzheimer's Disease
+Neuroimaging Initiative (ADNI) database (adni.loni.ucla.edu). As such, the
+investigators within the ADNI contributed to the design and implementation of
+ADNI and/or provided data but did not participate in analysis or writing of this
+report. A complete listing of ADNI investigators can be found at:
+http://adni.loni.ucla.edu/wp-content/uploads/how\_to\_apply/ADNI\_Acknowledgement\_List.pdf}}
 \affil[1]{Kimel Family Translational Imaging-Genetics Lab, Centre for Addiction and
@@
 subregions in MRI is a necessary first step to better understand the
-inter-individual variability of subject neuroanatomy. Typically, the gold
-standard for neuroanatomical segmentation is manual delineation by an expert
-human rater.  However, with the availability of increasingly large MRI datasets,
-the time and expertise required for manual segmentation becomes prohibitive
+inter-individual variability of subject neuroanatomy.
+
+Typically, the gold standard for neuroanatomical segmentation is manual
+delineation by an expert human rater.  However, with the availability of
+increasingly large MRI datasets, the time and expertise required for manual
+segmentation becomes prohibitive
 \citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. Further, there is
@@
 \citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited in
-accuracy due to inaccuracies in the estimated nonlinear transformation itself,
-partial volume effects in label resampling, and irreconcilable differences
-between the neuroanatomy represented within the atlas and target images.
+accuracy due to error in the estimated nonlinear transformation itself, partial
+volume effects in label resampling, and irreconcilable differences between the
+neuroanatomy represented within the atlas and target images.
 
@@
 
-More recently, many groups have been using multiple atlases to improve overall
+More recently, many groups have used multiple atlases to improve overall
 segmentation accuracy (i.e. multi-atlas segmentation) over model-based approaches
@@
 the hippocampus. Additionally, several methods have been explored for label
-fusion, including the STAPLE algorithm \citep{Warfield2004} that computes a
-probabilistic segmentation using an expectation maximization framework from an
-set of competing segmentations, or others where a subset of segmentations can be
-estimated using metrics such as the sum of squared differences in the regions of
-interest to be segmented \citep{Coupe2011}.
+fusion, including the STAPLE algorithm (Simultaneous Truth And Performance Level
+Estimation; \citet{Warfield2004}) that computes a probabilistic segmentation
+using an expectation-maximization framework from an set of competing
+segmentations, or others where a subset of segmentations can be estimated using
+metrics such as the sum of squared differences in the regions of interest to be
+segmented \citep{Coupe2011}.
 
@@
 to the hippocampus using more than a single input atlas, while simultaneously
-limiting the number of possible inputs used during segmentation, and
-demonstrating that accurate identification of the hippocampal subfields is
-indeed possible using this methodology.
-
-Of special relevance to the present work is the LEAP algorithm \citep{Wolz2010}
-because of its focus on performing multi-atlas segmentation with a limited
-number of input atlases.  The LEAP algorithm is a clever modification to the
-basic multi-atlas strategy in which an atlas library is grown, beginning with a
-set of manually labelled atlases, by successively incorporating unlabelled
-target images once they themselves have been labelled using multi-atlas
-techniques. The sequence in which target images are labelled is chosen so that
-the similarity between the atlas images and the target images is minimised at
-each step, effectively allowing for deformations between very dissimilar images
-to be broken up into sequences of smaller deformations. Although
-\citet{Wolz2010} begin with an atlas library of 30 MR images, this method could
-theoretically work using a much smaller atlas library.  In their validation,
-LEAP was used to segment the whole hippocampus in the ADNI-1 baseline dataset,
-achieving a mean Dice score of 0.85 against \gs segmentations.
+limiting the number of inputs used during segmentation, and demonstrating that
+accurate identification of the hippocampal subfields is indeed possible using
+this methodology.
+
+Of special relevance to the present work is the LEAP algorithm (Learning
+Embeddings for Atlas Propagation; \citet{Wolz2010}) because of its focus on
+performing multi-atlas segmentation with a limited number of input atlases.  The
+LEAP algorithm is a clever modification to the basic multi-atlas strategy in
+which an atlas library is grown, beginning with a set of manually labelled
+atlases, by successively incorporating unlabelled target images once they
+themselves have been labelled using multi-atlas techniques. The sequence in
+which target images are labelled is chosen so that the similarity between the
+atlas images and the target images is minimised at each step, effectively
+allowing for deformations between very dissimilar images to be broken up into
+sequences of smaller deformations. Although \citet{Wolz2010} begin with an atlas
+library of 30 MR images, this method could theoretically work using a much
+smaller atlas library.  In their validation, LEAP was used to segment the whole
+hippocampus in the ADNI-1 baseline dataset, achieving a mean Dice score of 0.85
+against \gs segmentations.
 
@@
 Third, the labels are combined into a single label by label fusion
-\citep{Heckemann2006, Heckemann2011}. This method is described in detail in
-other publications \citep{Collins2010,Heckemann2011,Aljabar2009}. When only a
-single atlas is used, basic multi-atlas segmentation degenerates into
-model-based segmentation: labels are propogated from the atlas to a target, and
-no label fusion is needed.
-
-\mb bootstraps the creation of a large template library given a limited input atlas 
-library, and then uses the template library in basic multi-atlas segmentation. 
-Images for the template library are selected from a set of input target images, 
-either arbitrarily or so as to reflect the neuroanatomy or demographics of the 
-target set as a whole (for instance, by sampling equally from cases and controls). 
-The template library images are automatically labelled by each of the atlases
-via label propagation. Basic multi-atlas segmentation is then conducted using
-the template library to segment the entire set of target images (including the
-targets whose images are used in the construction of the template library).
-Since each template library image has multiple labels (one from each atlas), the
-final number of labels to be fused for each target may be quite large (i.e. \#
-of atlas $\times$ \# of templates).
+\citep{Heckemann2006, Heckemann2011}. The basic multi-atlas segmentation method
+is described in detail in other publications
+\citep{Collins2010,Heckemann2011,Aljabar2009}. When only a single atlas is used,
+basic multi-atlas segmentation degenerates into model-based segmentation: labels
+are propogated from the atlas to a target, and no label fusion is needed.
+
+\mb ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
+T}emplates) bootstraps the creation of a large template library given a limited
+input atlas library, and then uses the template library in basic multi-atlas
+segmentation.  Images for the template library are selected from a set of input
+target images, either arbitrarily or so as to reflect the neuroanatomy or
+demographics of the target set as a whole (for instance, by sampling equally
+from cases and controls).  The template library images are automatically
+labelled by each of the atlases via label propagation.  Basic multi-atlas
+segmentation is then conducted using the template library to segment the entire
+set of target images (including the target images used in the construction of
+the template library).  Since each template library image has multiple labels
+(one from each atlas), the final number of labels to be fused for each target
+may be quite large (i.e. \# of atlas $\times$ \# of templates).
 
@@
 diseases: first episode schizophrenia (Experiment 3), and Alzheimer's disease
-(Experiment 4). Additionally, in experiment 4, we compare \mb segmentations with
+(Experiment 4). Additionally, in Experiment 4, we compare \mb segmentations with
 those of well-known automated and manual methods and assess segmentation bias.
@@
 \paragraph{\adnidataset dataset} 
+Data used in the preparation of this article were obtained from the Alzheimer’s
+Disease Neuroimaging Initiative (ADNI) database (adni.loni.ucla.edu). The ADNI
+was launched in 2003 by the National Institute on Aging (NIA), the National
+Institute of Biomedical Imaging and Bioengineering (NIBIB), the Food and Drug
+Administration (FDA), private pharmaceutical companies and non-profit
+organizations, as a $60 million, 5-year public- private partnership. The primary
+goal of ADNI has been to test whether serial magnetic resonance imaging (MRI),
+positron emission tomography (PET), other biological markers, and clinical and
+neuropsychological assessment can be combined to measure the progression of mild
+cognitive impairment (MCI) and early Alzheimer’s disease (AD).  Determination of
+sensitive and specific markers of very early AD progression is intended to aid
+researchers and clinicians to develop new treatments and monitor their
+effectiveness, as well as lessen the time and cost of clinical trials.
+
+The Principal Investigator of this initiative is Michael W. Weiner, MD, VA
+Medical Center and University of California – San Francisco. ADNI is the result
+of efforts of many co- investigators from a broad range of academic institutions
+and private corporations, and subjects have been recruited from over 50 sites
+across the U.S. and Canada. The initial goal of ADNI was to recruit 800 subjects
+but ADNI has been followed by ADNI-GO and ADNI-2.  To date these three protocols
+have recruited over 1500 adults, ages 55 to 90, to participate in the research,
+consisting of cognitively normal older individuals, people with early or late
+MCI, and people with early AD. The follow up duration of each group is specified
+in the protocols for ADNI-1, ADNI-2 and ADNI-GO. Subjects originally recruited
+for ADNI-1 and ADNI-GO had the option to be followed in ADNI-2. For up-to-date
+information, see www.adni-info.org.
+
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
@@
 \url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly preprocessed images
+\citep{Wyman2012}. This image collection contains uniformly pre-processed images
 which have been designated to be the "best" after quality control.  All images

### 2013-09-01 paper.Rnw
Some of Julie's edits

@@
 the use of multiple manually segmented atlases and probabilistic segmentation
-techniques, such as those found in the FreeSurfer package \citep{Fischl2002}.
-FreeSurfer uses a probabilistic atlas of anatomical and tissue classes along
+techniques, such as those found in the \freesurfer package \citep{Fischl2002}.
+\freesurfer uses a probabilistic atlas of anatomical and tissue classes along
 with spatial constraints for class labels encoded using a Markov random field
@@
 that attempt to define the hippocampus and its subfields using standard
-T1-weighted data. The first is included with the FreeSurfer package
+T1-weighted data. The first is included with the \freesurfer package
 \citep{VanLeemput2009}.  This work is limited as it omits the tail of the
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT \gs segmentations from the ADNI1 dataset. This form of
-cross-validation allows us to rigorously validate a large number of parameter
-settings of \mb (atlas and template library sizes, registration algorithm, and
-label fusion method) and select the best parameters to use in subsequent
-experiments.
+of images and SNT \gs segmentations from the ADNI1 dataset \citep{Hsu2002}. This
+form of cross-validation allows us to rigorously validate a large number of
+parameter settings of \mb (atlas and template library sizes, registration
+algorithm, and label fusion method) and select the best parameters to use in
+subsequent experiments.
 
@@
 Administration (FDA), private pharmaceutical companies and non-profit
-organizations, as a $60 million, 5-year public- private partnership. The primary
+organizations, as a \$60 million, 5-year public- private partnership. The primary
 goal of ADNI has been to test whether serial magnetic resonance imaging (MRI),
@@
 have recruited over 1500 adults, ages 55 to 90, to participate in the research,
-consisting of cognitively normal older individuals, people with early or late
-MCI, and people with early AD. The follow up duration of each group is specified
-in the protocols for ADNI-1, ADNI-2 and ADNI-GO. Subjects originally recruited
-for ADNI-1 and ADNI-GO had the option to be followed in ADNI-2. For up-to-date
-information, see www.adni-info.org.
+consisting of cognitively normal (CN) older individuals, people with early or
+late MCI, and people with early AD. The follow up duration of each group is
+specified in the protocols for ADNI-1, ADNI-2 and ADNI-GO. Subjects originally
+recruited for ADNI-1 and ADNI-GO had the option to be followed in ADNI-2. For
+up-to-date information, see www.adni-info.org.
 
@@
 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
-yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.
+yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.
 
@@
 segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and their \gs segmentations to be used as an
-atlas library, and a validation set consisting of the remaining images segmented
-by both \mb and multi-atlas. The computed segmentations were compared to the
-\gs segmentations (see below).
+training set consisting of images and \gs segmentations used as an atlas
+library, and a validation set consisting of the remaining images segmented by
+both \mb and multi-atlas. The computed segmentations were compared to the \gs
+segmentations (see below).
 
@@
 explored are: atlas library size (1-9), template library size (1-20),
-registration method (\ants or \animal), and label fusion method (majority vote,
+registration method (\ants or \animal, described below), and label fusion method (majority vote,
 cross-correlation weighted majority vote, and normalized mutual information
-weighted majority vote).  A total of $10 \times 69 \times 9 \times 20 \times 2
-\times 3 = \num{7452000}$ validation rounds were conducted, resulting in a total
-of $\num{1490400}$ segmentations analysed.
+weighted majority vote, described below).  A total of $10 \times 69 \times 9
+\times 20 \times 2 \times 3 = \num{7452000}$ validation rounds were conducted,
+resulting in a total of $\num{1490400}$ segmentations analysed.
 
@@
 estimates a 3D deformation field between two MR images. At first, large
-deformations are estimated using blurred version of the input data. These larger
+deformations are estimated using a blurred version of the input data. These larger
 deformations are then input to subsequent steps where the fit is refined by
@@
 Label fusion is a term given to the process of combining the information from
-several candidate labels for an intensity image into a single labelling.  In
-this experiment we explore three fusion methods. 
+several candidate labels for an image into a single labelling.  In this
+experiment we explore three fusion methods: 
 \begin{description}
@@
   from the EZMinc toolkit\footnote{\url{https://github.com/vfonov/EZminc}} is
-  used to calculate the normalised mutual information measure between to images.
+  used to calculate the normalised mutual information measure between two images.
 \end{description}
@@
 segmentations of five in-vivo $300\mu$ isotropic T1-weighted MR images. The
-segmentations include subfield segmentations for the cornu ammonis (CA) 1, CA4,
-dentate gyrus, subiculum, and CA 2 and 3 combined. Subjects in the Winterburn
-atlases range in age from 29-57 years (mean age of 37), and include two males
-and three females.
+segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
+and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR, strata
+lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
 
@@
 subject images are separately acquired T1 BRAVO images of four of the five
-Winterburn atlas subjects (referred to as the {\em BRAVO} dataset).  Figure
+Winterburn atlas subjects (referred to as the {\em T1 BRAVO} dataset).  Figure
 \ref{fig:LOOCV-schematic} illustrates the experimental set up of both parts of
@@
 To assess the \mb LOOCV segmentations we compute the relative percent error in 
-hippocampal volume with the full resolution Winterburn atlas 
+hippocampal volume with respect to the full resolution Winterburn atlas 
 segmentations. In addition, by computing the relative error in volume of the
 Winterburn atlas labels resampled (with nearest-neighbour interpolation) to
-0.9mm-isotropic voxels, we obtain a baseline error to assess against.
+0.9mm-isotropic voxels, we obtain a baseline error against which to assess \mb
+segmentations.
 
@@
 
-To validate that \mb works effectively in the context of other neurological
-disorders, in this experiment we use the Winterburn atlases to predict the
-hippocampal segmentations of patients with Schizophrenia. The resulting
-segmentations are assessed for quality by comparison with expert manual
-segmentations.
+To validate that the \mb works effectively in the context of other neurological
+disorders, in this experiment we use the Winterburn atlases to derive 
+whole hippocampal segmentations of a dataset of patients with Schizophrenia. The
+resulting segmentations are assessed for quality by comparison with expert
+manual segmentations.
 
@@
 
-Manual segmentation of each subject hippocampus is produced following a
+Manual segmentation of each subject whole hippocampus is produced following a
 validated segmentation protocol \citep{Pruessner2000}.
@@
 \mb is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2).  All images from the \FEPdataset dataset are segmented by \mb.  The 
-optimal size of template library, registration method, and label fusion method 
-found in Experiment 1 are used.
+(see Experiment 2) ignoring subfield delineations.  All images from the
+\FEPdataset dataset are segmented by \mb.  The optimal size of template library,
+registration method, and label fusion method found in Experiment 1 are used.
 
@@
 methods using different protocols. In place of an overlap metric, we can assess
-the degree of correlation in average hippocampal volume of the subjects produced
-by each method. Specifically, a Pearson correlation is used.
+the degree of correlation in average bilateral hippocampal volume of the
+subjects produced by each method. Specifically, a Pearson correlation is used.
 
@@
 
-To validate \mb segmentation quality with respect to other established
-automated methods, \mb was applied to large dataset from the ADNI project and
-the resulting segmentations were compared to those produced by FreeSurfer, FSL,
-MAPER and the SNT \gs segmentations.
+To validate \mb segmentation quality with respect to other established automated
+hippocampal segmentation methods, \mb was applied to large dataset from the ADNI
+project and the resulting segmentations were compared to those determined by
+\freesurfer, \fsl, \maper and the SNT \gs segmentation.
 
@@
 SNT and FreeSurfer hippocampal volumes for a subset of images were provided by
-ADNI, along with quality control data for each FreeSurfer segmentation
+ADNI, along with quality control data for each \freesurfer segmentation
 (guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
-the entire \adnidataset dataset are shown in Table \ref{tab:ADNI1-Dataset-Demographics}.
+the entire \adnidataset dataset are shown in Table
+\ref{tab:ADNI1-Dataset-Demographics}.
 
@@
 segmentations using similar quality control guidelines (if either hippocampus
-was under or over segmented by greater than 10mm in three or more slices then
+was under or over segmented by $10mm$ or greater in three or more slices then
 the segmentation did not pass).  Only images meeting the conditions of having
@@
 across subjects. We also computed the correlation in hippocampal volume between
-the existing, established automated segmentation methods: FSL, FreeSurfer, and
-MAPER. Additionally, we evaluate the volume-related fixed and proportional
-biases in all segmentation methods using Bland-Altman plots \citep{Bland1986}.
+existing, established automated segmentation methods -- \fsl, \freesurfer, and
+\maper, and SNT semi-automated segmentations. Additionally, we evaluate the
+volume-related fixed and proportional biases in all segmentation methods using
+Bland-Altman plots \citep{Bland1986}.
 
@@
 information weighted majority vote).  Hippocampal \mb-based segmentations using
-both \animal and \ants registration algorithm demonstrate good overlap with SNT
-\gs segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=3)}
+both \animal and \ants registration algorithms demonstrate good overlap with SNT
+\gs segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=2)}
 when using 9 atlases, 19 templates, ANTS registration, and majority vote label
@@
 @
-  \caption{Mean Dice's Similarity Coefficient of \mb segmentations with 
+  \caption{Mean Dice's Similarity Coefficient of \mb segmentations relative to
   SNT \gs segmentations for 69 ADNI1 subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method. Points above zero
-  indicate \mb parameter settings yielding a higher mean DSC score than
-  multi-atlas segmentation using the same number of atlases. Error bars
-  indictate standard error.
+  size, registration algorithm, and label fusion method. Error bars indictate
+  standard error.
   \label{fig:ADNI1-xval-k-mean}}
@@
 @
-  \caption{Difference in mean Dice's Similarity Coefficient of \mb and
-  multi-atlas segmentations with \gs (SNT) segmentations for a range of atlas
-  and template library sizes. Error bars indicate standard error.
+  \caption{Difference in mean Dice's Similarity Coefficient between \mb and
+  multi-atlas segmentations relative to \gs (SNT) segmentations vs. atlas and
+  template library size when using the \ants registration method, and
+  majority-vote label fusion. Error bars indicate standard error.
   \label{fig:ADNI1-xval-k-diff}}
@@
   accuracy.}
-  Variance of segmentation accuracy between \mb and multi-atlas segmentation 
-  is computed for each subject across all ten rounds of validation. Shown on the 
+  Variance of segmentation accuracy between \mb and multi-atlas segmentation is
+  computed for each subject across all ten rounds of validation. Shown on the
   y-axis (scaled logarithmically) is the  p-value resulting from a t-test
@@
   (atlas/template library size). Only points where \mb mean variability is lower
-  than multi-atlas are shown.}
+  than multi-atlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
   \label{ADNI1-xval-variability}
@@
 ($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
-remaining Winterburn atlases.
+remaining high-resolution Winterburn atlas subjects' images.
 
@@
 segmentations compares favourably to the error resulting from image resampling
-(Figure \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4, and Dentate
-subregions all show near or smaller percent errors. The Sibiculum and CA2/CA3
-subregions show distinctly larger error than is found through resampling.
+(Figure \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4/Dentate, and
+SR/SL/SM subregions all show near or smaller percent errors. The Subiculum and
+CA2/CA3 subregions show distinctly larger error than is found through
+resampling.
 
 Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb subfield
-segmentation.
+segmentations for a single subject.
 
@@
 levels(melted$measure) <- list("Subsampled"="resampled",
-                               "MAGeT + Subsampled" = "mb_0.9mm",
-                               "MAGeT + BRAVO"="mb_bravo")
+                               "MAGeT on Subsampled" = "mb_0.9mm",
+                               "MAGeT on T1 BRAVO"="mb_bravo")
 
@@
   {\bf Subsampled} are volumes of the manual segmentations of the Winterburn
-  atlases after resampling to $0.9mm^3$.  
-  {\bf MAGeT + Subsampled} volumes are \mb segmentations of the Winterburn atlas
+  atlases after resampling to $0.9mm^3$ voxels.  
+  {\bf MAGeT on Subsampled} volumes are \mb segmentations of the Winterburn atlas
   images after resampling to $0.9mm^3$ voxels.
-  {\bf MAGeT + BRAVO} volumes are \mb segmentations of T1 BRAVO images
+  {\bf MAGeT on T1 BRAVO} volumes are \mb segmentations of T1 BRAVO images
   ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
@@
   geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
-  xlab(expression("Mean Manual hippocampus volume " (mm^3))) + 
-  ylab(expression("Mean MAGeT hippocampus volume "  (mm^3))) + 
+  xlab(expression("Mean bilateral manual hippocampus volume " (mm^3))) + 
+  ylab(expression("Mean bilateral MAGeT hippocampus volume "  (mm^3))) + 
   scale_size_continuous(guide="none") 
 @
-  \caption{Mean hippocampus volume as measured by \mb  vs. manually segmented
-  volumes from the First Episode Patients with Schizophrenia dataset. A linear
-  fit line is shown, with shaded region showing standard error. }
+  \caption{Mean bilateral hippocampus volume as measured by \mb  vs. manually
+  segmented volumes from the First Episode Patients with Schizophrenia dataset.
+  A linear fit line is shown, with shaded region showing standard error. }
   \label{fig:SZ-volumes}
@@
 @
-\mb produces hippocampus segmentation volumes that are highly correlated with 
+\mb produces hippocampal volumes that are highly correlated with 
 manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
@@
 experiment \mb was configured with a template library of 21 randomly chosen
-subject images (7 from each disease class) and used majority vote label fusion.
+subject images (7 from each disease class) and majority vote label fusion.
 The entire \adnidataset dataset was segmented by \mb, and the resulting volumes
@@
 dataset by each of the four automated methods investigated (FreeSurfer (FS),
-FSL, MAPER, \mb) vs. SNT \gs. Linear fit lines and pearson
+FSL, MAPER, \mb) vs. SNT \gs. Linear fit lines and Pearson
 correlations are shown for each method.}
@@
   \end{centering}
-  \caption{SNT \gs and MAGeT segmentations for 30 ADNI1 subjects (10
+  \caption{SNT \gs and \mb segmentations for 30 ADNI1 subjects (10
   subjects randomly selected from each disease category in the subject pool used
-  in Experiment 1). Sagittal slices are shown for subject unlabelled T1-weighted
-  anatomical image, SNT \gs label (green), and \mb label (blue). Noted are
-  examples of common segmentation idiosyncrasies: 
+  in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
+  anatomical image. SNT \gs labels appear in green, and \mb labels appear in
+  blue. Noted are examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
@@
    Some studies of automated segmentation of ADNI images are excluded because they 
-   do not not provide overlap measures for the hippocampus \cite{Heckemann2011,Chupin2009}.
+   do not not provide overlap measures for the hippocampus \citep{Heckemann2011,Chupin2009}.
   \label{tab:other-methods}} 
@@
 \citep{Aljabar2009,Collins2010} has found that cross-correlation and normalized
-mutual information based weighted label fusion improves segmentation accuracy
+mutual information-based weighted label fusion improves segmentation accuracy
 over simple majority vote label fusion. Selectively filtering out atlases with
@@
 
-Although, the goal of this manuscript was to not exhaustively test or validate
+Although, the goal of this manuscript was not to exhaustively test or validate
 multiple different voting strategies in the context of our segmentation
@@
 estimated performance level with respect to the other available candidate
-segmentations.  Further, the sensitivity and specificity parameters could also
-be tuned to potentially improve segmentation accuracy and reliability.  It is
-likely that using more sophisticated voting methods would have a positive
-effect on the overall segmentation performance, as demonstrated by the STAPLE
+segmentations.  Further, the sensitivity and specificity parameters can also be
+tuned to potentially improve segmentation accuracy and reliability.  It is
+likely that using more sophisticated voting methods would have a positive effect
+on the overall segmentation performance, as demonstrated by the STAPLE
 algorithm.

### 2013-09-02 paper.Rnw
The rest of Julie's edits

@@
 %  - check tense: we are is have has 
-%  - update montages with TAL coordinates
 %  - add MAGeT brain diagram, and remove pseudocode
@@
 
-Figure \ref{alg:MAGeT} describes the \mb algorithm in pseudocode. Source code 
+Figure \ref{fig:MAGeT} illustrates the \mb algorithm graphically. Source code
 for \mb can be found at \url{http://github.com/pipitone/MAGeTbrain}.
 
-\begin{algorithm}
-  \scriptsize
-  \caption{Pseudocode for the \mb algorithm}
-  \label{alg:MAGeT}
-  \begin{algorithmic}
-    \Function{BasicMultiAtlasSegmentation}{Templates, Subjects}
-      \ForAll{$target$}
-        \ForAll{$template$}
-          \State propagate all labels for template to target space
-          \State store target labels
-        \EndFor
-        \State fuse target labels
-      \EndFor
-    \EndFunction
-    \\
-    \Function {MAGeTBrain}{Subjects, Atlases, n}
-      \For{$i = 1 \to n$}
-        \State choose a target to be used as a template
-        \State propagate labels from each atlas to template space
-        \State store the template with all of its labels
-      \EndFor
-      \State MultiAtlas(Templates, Subjects)
-    \EndFunction
-  \end{algorithmic}
-\end{algorithm}
+\begin{figure}
+  \caption{The \mb algorithm \label{fig:MAGeT}}
+\end{figure}
 
@@
 
-Figure \ref{fig:ADNI-subfields} shows a qualitative comparison of \mb and 
+Figure \ref{fig:ADNI-segmentations} shows a qualitative comparison of \mb and 
 SNT \gs hippocampal segmentations for 10 randomly selected subjects in 
@@
   {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb.}
-  \label{fig:ADNI-subfields}
+  \label{fig:ADNI-segmentations}
 \end{figure}
@@
 \mb framework -- a methodology that requires very few input atlases in
-order to provide accurate and reliable segmentations.  We have demonstrated that
-our methodology robustly provides accurate and consistent segmentations in
-populations with different ageing and neuropsychiatric characteristics.
-Further, we have demonstrated that algorithmic performance is not dependent on a
-single definition of the hippocampus but is effective with differing hippocampal
-definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}.  Finally, we also demonstrate
-that \mb provides accurate automatic identification of the hippocampal
-subfields despite contrast and resolution limitations in standard T1-weighted
-image volumes.
+order to provide accurate and reliable segmentations.  Experiment 1 establishes
+\mb as effective, and as comparable to multi-atlas segmentation. This experiment
+also rigorously characterizes the behaviour of \mb whole hippocampal
+segmentation under various parameter settings allowing us to choose an optimal
+setting for subsequent experiments. Experiment 2 demonstrates the accuracy of
+\mb hippocampal subfield identification despite contrast and resolution
+limitations in standard T1-weighted image volumes.  Experiments 2 and 3 validate
+\mb whole hippocampal segmentation accuracy and consistency on populations with
+different ageing and neuropsychiatric characteristics.  Furthermore, taken
+together, these experiments demonstrate that algorithmic performance is not
+dependent on a single definition of the hippocampus but is effective with
+differing hippocampal definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}.
 
@@
 leave-one-out-validation (Table \ref{tab:other-methods}), aside from
-\citep{Wang2011} who perform a similar validation as us). The thoroughness of
-our validation suggests that our results are reflective of a true average over
-the choice of parameter settings and are independent of atlas or template
-choice (provided the input atlases are properly segmented.
+\citep{Wang2011} who perform a similarily rigorous validation as us). The
+thoroughness of our validation suggests that our results are reflective of a
+true average over the choice of parameter settings and are independent of atlas
+or template choice (provided the input atlases are consistently segmented).
+
+On that note, it is worth observing that we found the SNT segmentations to be
+poor in general. As previously noted in Figure \ref{fig:ADNI-segmentations}, an
+expert manual rater (JW) identified many inconsistencies: occurances of over-
+and under-estimation, as well as misalignments of the entire segmentation
+volume.  Therefore, the use of the SNT segmentations as gold standards in
+Experiments 1 and 4 presents the possibility that the DSC scores and
+correlations we obtained are lower than what would be found with a more accurate
+and consistent gold standard.
 
@@
 In comparison to other methodologies in the field \mb performs favourably. Table
-\ref{tab:other-methods} surveys some of the most recent reported kappa values on
-the ADNI dataset using the same SNT labels for populating their template library
-and as gold standards for evaluation.  While it is difficult to compare
-segmentation results across studies, gold-standards, evaluation metrics, and
-algorithms it is worth noting that the methods summarized in Table
-\ref{tab:other-methods} require more atlases (between 16-55) than our \mb
-implementation with the Winterburn atlases \citep{Winterburn2013}.
+\ref{tab:other-methods} surveys some of the most recent reported DSC values
+reported on ADNI dataset, using SNT segmentations for the atlas library and as
+gold standards for evaluation.  While it is difficult to compare segmentation
+results across studies, gold standards, evaluation metrics, and algorithms it is
+worth noting that the methods summarized require more atlases (between 16-55)
+than our \mb implementation with the Winterburn atlases \citep{Winterburn2013}.
 
@@
 heuristically at the time of segmentation (this work also reports some of the
-highest kappas for the segmentation of ADNI data).  While this method is
+highest DSC scores for the segmentation of ADNI data). While this method is
 effective for the ADNI dataset (which is partially homogenized with respect to
@@
 then use this method to segment an ADNI sample, with a hierarchical
-experimentation protocol that involved \todo{complete this sentence}.  These
-methods suggest that some post-processing of the final segmentations would
-improve accuracy of the segmentation. While that may be true, there is little
-consensus regarding how to achieve this.
+experimentation protocol. These methods suggest that some post-processing of the
+final segmentations would improve accuracy of the segmentation. While that may
+be true, there is little consensus regarding how to achieve this.
 
@@
 
-In comparison to the methods that we compared and were available through the
-ADNI database (\freesurfer, \maper) and \fsl-FIRST, we find extremely variable
+Among the automated segmentation methods we compared in this paper
+(\freesurfer, \maper, \fsl-FIRST), we find extremely variable
 performance of all methods.  With the exception of \fsl all methods correlate
@@
 in comparison to the other methods.  Further, when estimating bias of these
-methods relative to SNT hippocampal volumes we see that large hippocampi are
+methods relative to SNT \gs hippocampal volumes we see that large hippocampi are
 over estimated while small hippocampi are under estimated. By comparison, \mb
 and MAPER are far more conservative in volume estimation, suggesting these
-methods may be better suited for estimating true-positives. However, in this
-analysis we have only compared methods by total hippocampal volume, and so more
-work is needed to understand the full extent to which these methods differ.
+methods may be better suited for estimating true-positives, especially in
+neurodegenerative disease subjects featuring smaller overall hippocampi.
+However, in this analysis we have only compared methods by total hippocampal
+volume, and so more work is needed to understand the full extent to which these
+methods differ.
 
@@
 recent work demonstrates that subfield segmentations can be used for
-classification of AD, MCI, and NC, there has been no explicit validation of the
+classification of AD, MCI, and CN, there has been no explicit validation of the
 methodology based on accuracy or precision.  Although the initial hippocampal
@@
 addition, their work requires some manual initialization to properly function.
-Our work demonstrates that we can reliably identify the CA1, subiculum, and CA4
-dentate with only modest amounts of error.  The fact that CA2/CA3 and molecular
-layers cannot be reasonably identified should not be surprising as these are
-extremely thin and spatially convoluted regions that originally required
-high-resolution MRI for identification.  It is likely that extents of these
-regions are well below the resolution offered by standard T1-weighted images
-and, in fact, many manual segmentation methodologies do not attempt to parse
-these regions either \citep{Wisse2012, Mueller2009}.
+Our work demonstrates that we can reliably identify the CA1, subiculum, and
+CA4/dentate with only modest amounts of error.  The fact that CA2/CA3 and
+molecular layers cannot be reasonably identified should not be surprising as
+these are extremely thin and spatially convoluted regions that originally
+required high-resolution MRI for identification.  It is likely that the extents of
+these regions are well below the resolution offered by standard T1-weighted
+images and, in fact, many manual segmentation methodologies do not attempt to
+parse these regions either \citep{Wisse2012, Mueller2009}.
 

### 2013-09-02 paper.Rnw
Added MAGeT schematic, and fixed a table

@@
 \begin{abstract}
-%Neuroimaging research often relies on automated anatomical segmentations of MR
-%images of the brain. Current multi-atlas based approaches provide accurate
-%segmentations of brain images by propagating manually derived segmentations of
-%specific neuroanatomical structures to unlabelled data. These approaches often
-%rely on a large number of such manually segmented atlases that take significant
-%time and expertise to produce. We present an algorithm for the automatic
-%segmentation of the hippocampus that minimizes the number of atlases needed
-%while still achieving similar accuracy to multi-atlas approaches.
+Neuroimaging research often relies on automated anatomical segmentations of MR
+images of the brain. Current multi-atlas based approaches provide accurate
+segmentations of brain images by propagating manually derived segmentations of
+specific neuroanatomical structures to unlabelled data. These approaches often
+rely on a large number of such manually segmented atlases that take significant
+time and expertise to produce. We present an algorithm for the automatic
+segmentation of the hippocampus that minimizes the number of atlases needed
+while still achieving similar accuracy to multi-atlas approaches.
 \end{abstract}
@@
 \begin{figure}
-  \caption{The \mb algorithm \label{fig:MAGeT}}
+  \begin{centering}
+    \includegraphics[width=\textwidth]{figure/MA-MAGeTBrain-Schematic}
+  \end{centering}
+  \caption{A schematic illustration of basic multi-atlas segmentation and \mb
+  segmentation \label{fig:MAGeT}}
 \end{figure}
@@
 master = read.csv('data/DT_study.csv')
-tab <- summary(~Age + Sex + Education + Handedness, 
-               data = master, method="reverse", na.rm = TRUE, test=FALSE)
+tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
+    method="reverse", na.rm = TRUE, test=FALSE)
 caption = 

### 2013-09-02 paper.Rnw
Abstract complete, todo's removed, and abbrv commands applied

@@
 %  - check tense: we are is have has 
-%  - add MAGeT brain diagram, and remove pseudocode
 % Sweave/Knitr init
@@
 \begin{abstract}
-Neuroimaging research often relies on automated anatomical segmentations of MR
-images of the brain. Current multi-atlas based approaches provide accurate
-segmentations of brain images by propagating manually derived segmentations of
-specific neuroanatomical structures to unlabelled data. These approaches often
-rely on a large number of such manually segmented atlases that take significant
-time and expertise to produce. We present an algorithm for the automatic
-segmentation of the hippocampus that minimizes the number of atlases needed
-while still achieving similar accuracy to multi-atlas approaches.
+
+\textbf{Introduction:} Neuroimaging research often relies on automated anatomical
+segmentations of MR images of the brain. Current multi-atlas based approaches
+provide accurate segmentations of brain images by propagating manually derived
+segmentations of specific neuroanatomical structures to unlabelled data. These
+approaches often rely on a large number of such manually segmented atlases (e.g.
+30-80) that take significant time and expertise to produce. We present an
+algorithm, \mb, for the automatic segmentation of the hippocampus that minimizes
+the number of atlases needed while still achieving similar accuracy to
+multi-atlas approaches.\\
+\textbf{Method:} \mb works by propagating atlas segmentations to a template library,
+formed from a subset of target images, via transformations estimated by non-linear
+image registration. The resultant segmentations are then propagated to each
+target image and fused using a label fusion method. We conduct a 10-fold Monte
+Carlo cross-validation of whole hippocampal segmentation on a pool of 69 ADNI
+subjects over a range of parameter settings, and a leave-one-out
+cross-validation (LOOCV) of hippocampal subfield segmentation using five
+high-resolution atlases. Two final experiments assess \mb when applied to
+Schizophrenia and Alzheimer's disease populations, and \mb segmentations are
+compared with existing automated methods (\fsl, \freesurfer, \maper) and biases are
+explored.\\
+\textbf{Results:} Using nine atlases and 19 template images, \mb achieves a
+mean Dice's Similarity Coefficient (DSC) of 0.84 over 10-folds of Monte Carlo
+cross-validation on 69 subjects relative to semi-automated \gs segmentations,
+and shows significantly lower variability in DSC than multi-atlas segmentation.
+In a LOOCV, \mb reproduces the volumes of the CA1, CA4/DG, and SR/SL/SM
+hippocampal subregions with a percent error in volume that is at or lower than
+that produced by image resampling. \mb produces hippocampal volumes in a
+Schizophrenia patient population that are highly correlated with expert manual
+segmentation volumes (Pearson $r = 0.877, t = 16.244, p < 0.001$). Compared to
+\fsl and \freesurfer, \mb shows much smaller fixed volume bias (within $250
+mm^3$ on average) to \gs segmentations, as well as a conservative, rather than
+exaggerated, proportional volume bias.\\
+\textbf{Conclusion:} We demonstrate that \mb produces accurate hippocampal
+segmentations using only 5 atlases over different hippocampal definitions,
+disease populations, and acquisition types, as well as showing that accurate
+identification of the hippocampal subfields is possible.
 \end{abstract}
@@
 settings offering the best performance for use in subsequent experiments.
-Experiment 2 tests hippocampal subfield segmentation quality. The last two
-experiments assess the validity of the \mb algorithm when applied to different
-diseases: first episode schizophrenia (Experiment 3), and Alzheimer's disease
-(Experiment 4). Additionally, in Experiment 4, we compare \mb segmentations with
-those of well-known automated and manual methods and assess segmentation bias.
+Experiment 2 assesses hippocampal subfield segmentation quality in a
+leave-one-out cross-validation design. The last two experiments assess the
+validity of the \mb algorithm when applied to different diseases: first episode
+schizophrenia (Experiment 3), and Alzheimer's disease (Experiment 4).
+Additionally, in Experiment 4, we compare \mb segmentations with those of
+well-known automated methods and assess segmentation bias.
 
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT \gs segmentations from the ADNI1 dataset \citep{Hsu2002}. This
+of images and SNT \gs segmentations from the ADNI dataset \citep{Hsu2002}. This
 form of cross-validation allows us to rigorously validate a large number of
@@
 segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
-and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR, strata
+and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
 lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
@@
 The {\em \adnidataset} standardized dataset contains $1919$ images in total.
-SNT and FreeSurfer hippocampal volumes for a subset of images were provided by
+SNT and \freesurfer hippocampal volumes for a subset of images were provided by
 ADNI, along with quality control data for each \freesurfer segmentation
@@
 registration method, and label fusion method were determined by the optimal
-settings found in Experiment 1.  The FSL FIRST segmentation method was used via
+settings found in Experiment 1.  The \freesurfer segmentation method was used via
 the \verb+run_first_all+ script according to the FIRST user guide
@@
 
-One author (MP) performed visual quality inspection for MAGeT and FSL
+One author (MP) performed visual quality inspection for \mb and \freesurfer 
 segmentations using similar quality control guidelines (if either hippocampus
@@
 the segmentation did not pass).  Only images meeting the conditions of having
-segmentations from all methods (SNT, MAPER, FreeSurfer, FSL, and \mb) and also
+segmentations from all methods (SNT, \maper, \freesurfer, \fsl, and \mb) and also
 passing quality control inspection were included in the analysis.
@@
 In general, across hippocampal subregions the percent error in volume of \mb
-segmentations compares favourably to the error resulting from image resampling
-(Figure \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4/Dentate, and
-SR/SL/SM subregions all show near or smaller percent errors. The Subiculum and
-CA2/CA3 subregions show distinctly larger error than is found through
-resampling.
+segmentations (relative to the full-resolution Winterburn atlas segmentation)
+compares favourably to the error resulting from image resampling (Figure
+\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4/DG, and SR/SL/SM
+subregions all show a percent error in volume that is at or lower than
+resampling error. The Subiculum and CA2/CA3 subregions show distinctly larger
+percent error in volume than is found through resampling.
 
@@
 
-Based on the results from the ADNI1 Cross-Validation experiment, in this
-experiment \mb was configured with a template library of 21 randomly chosen
-subject images (7 from each disease class) and majority vote label fusion.
-The entire \adnidataset dataset was segmented by \mb, and the resulting volumes
-compared with those obtained by expert semi-automated segmentation (SNT), and
-three other automated segmentation techniques: MAPER, FreeSurfer, and FSL.
-Table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total
-count of segmentations available, including a count of those which have failed
-a quality control inspection.  A total of \Sexpr{length(means.complete$RID)}
-images are included in the following analysis, having met quality criteria and
-having segmentations from every method.
+Based on the results from Experiment 1, in this experiment \mb was configured
+with a template library of 21 randomly chosen subject images (7 from each
+disease class) and majority vote label fusion.  The entire \adnidataset dataset
+was segmented by \mb, and the resulting volumes compared with those obtained by
+expert semi-automated segmentation (SNT), and three other automated segmentation
+techniques: \maper, \freesurfer, and \fsl. Table
+\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
+segmentations available, including a count of those which have failed a quality
+control inspection.  A total of \Sexpr{length(means.complete$RID)} images are
+included in the following analysis, having met quality criteria and having
+segmentations from every method.
 
@@
 correlation can be driven by a large range in true values, for instance).  All
-methods show an obvious proportional bias: FreeSurfer and FSL markedly
+methods show an obvious proportional bias: \freesurfer and \fsl markedly
 under-estimate smaller hippocampi and over-estimate large hippocampi, whereas
-MAPER and \mb show a reverse, conservative bias (Figure
+\maper and \mb show a reverse, conservative bias (Figure
 \ref{fig:ADNI-Bland-Altman}).  Additionally, all methods show a fixed volume
-bias, with FreeSurfer and FSL most dramatically over-estimating hippocampal
-volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and MAPER and
+bias, with \freesurfer and \fsl most dramatically over-estimating hippocampal
+volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and \maper and
 \mb within $250 mm^3$ on average.
@@
   \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by FreeSurfer, FSL, MAPER, \mb, and SNT \gs vs. disease
+dataset by \freesurfer, \fsl, \maper, \mb, and SNT \gs vs. disease
 category.}
@@
   \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by each of the four automated methods investigated (FreeSurfer (FS),
-FSL, MAPER, \mb) vs. SNT \gs. Linear fit lines and Pearson
-correlations are shown for each method.}
+dataset by each of the four automated methods investigated (\freesurfer (FS), \fsl,
+\maper, \mb) vs. SNT \gs. Linear fit lines and Pearson correlations are shown
+for each method.}
   \label{fig:ADNI-volumes-plot}
@@
 measured in the \adnidataset dataset by SNT \gs segmentation and each of the
-four automated methods investigated (FreeSurfer, FSL, MAPER, \mb). The overall
+four automated methods investigated (\freesurfer, \fsl, \maper, \mb). The overall
 mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
@@
 order to provide accurate and reliable segmentations.  Experiment 1 establishes
-\mb as effective, and as comparable to multi-atlas segmentation. This experiment
+\mb as effective and as comparable to multi-atlas segmentation. This experiment
 also rigorously characterizes the behaviour of \mb whole hippocampal
-segmentation under various parameter settings allowing us to choose an optimal
+segmentation under various parameter settings, allowing us to choose an optimal
 setting for subsequent experiments. Experiment 2 demonstrates the accuracy of
@@
 leave-one-out-validation (Table \ref{tab:other-methods}), aside from
-\citep{Wang2011} who perform a similarily rigorous validation as us). The
+\citep{Wang2011} who perform a similarily rigorous validation as us. The
 thoroughness of our validation suggests that our results are reflective of a
@@
 other work \citep{Heckemann2006}.
-\todo[inline]{time complexity seems suspicious}
 
@@
 over estimated while small hippocampi are under estimated. By comparison, \mb
-and MAPER are far more conservative in volume estimation, suggesting these
+and \maper are far more conservative in volume estimation, suggesting these
 methods may be better suited for estimating true-positives, especially in
@@
 literature although very few automated methods are available for the
-segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. However, the
-FreeSurfer implementation of subfield segmentation \todo{complete}. While
+segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. While
 recent work demonstrates that subfield segmentations can be used for

### 2013-09-03 paper.Rnw

@@
 % TODO: 
-%  - check tense: we are is have has 
+%  - check tense: we are is have has  use/used find/found
 % Sweave/Knitr init
@@
 \author[1,6,7]{M. Mallar Chakravarty}
-\author[ ]{and the Alzheimer's Disease Neuroimaging Initiative\footnote{Data
+\author[ ]{the Alzheimer's Disease Neuroimaging Initiative\footnote{Data
 used in preparation of this article were obtained from the Alzheimer's Disease
@@
 \citep{DenHeijer2012,Scoville2000}. The hippocampus is of interest to
-neuroscientists because it is implicated in many forms of brain disfunction,
+neuroscientists because it is implicated in many forms of brain dysfunction,
 including Alzheimer's disease \citep{Sabuncu2011} and schizophrenia
@@
 
-In this paper we undertake a thorough validation of the \mb algorithm for
-segmentation of the hippocampus. First, we address the feasibility and accuracy
-of whole hippocampus segmentation with a limited number of input atlases by
-performing a multi-fold experiment using a subset of the Alzheimer's Disease
-Neuroimaging Initiative (ADNI) 1 dataset. Based on the parameters we find in our
-experiment, we then validate our algorithm using all of the data available in
-the ADNI Complete 1Yr sample and compare our segmentations to the other
-segmentations that are available through the ADNI informatics portal. To ensure
-that we have not over-fit our parameters to the aging or neurodegenerative
-brain, we also apply our segmentations to a dataset of individuals suffering
-from first episode psychosis. Finally, we perform a leave-one-out validation
-experiment to determine if the subfields can be accurately identified using our
-multi-atlas framework.
+In this paper we undertook a thorough validation of the \mb algorithm for
+segmentation of the hippocampus. First, we addressed the feasibility and
+accuracy of whole hippocampus segmentation with a limited number of input
+atlases by performing a multi-fold experiment using a subset of the Alzheimer's
+Disease Neuroimaging Initiative (ADNI) dataset over a range of parameters.  We
+then performed a leave-one-out validation to determine if hippocampal subfields
+can be accurately identified using our multi-atlas framework.  To ensure that we
+have not over-fit our parameters to the aging or neurodegenerative brain, we
+also applied our segmentations to a dataset of individuals suffering from first
+episode psychosis.  Finally, validated our algorithm using all of the data
+available in the \adnidataset sample and compare our segmentations to the other
+segmentations available through the ADNI informatics portal.
 
@@
 \subsection{The \mb Algorithm}
-In this paper, we will use the term {\em label} to mean any segmentation (manual or 
+In this paper, we use the term {\em label} to mean any segmentation (manual or
 derived) of an MR image. {\em Label propagation} is the process by which two
 images are registered and the resulting transformation is applied to the labels
-from one image to bring them into alignment with the other image. We will use
-the term {\em atlas} to mean a manually segmented image, and the term {\em
-template} to mean an automatically segmented image (i.e.  via label
-propagation). The terms {\em atlas library} and {\em template library} describe
-any set of such images. Additionally, we will use the term {\em target} to refer
-to an unlabelled image that is undergoing segmentation.
-
-The simplest form of multi-atlas segmentation, (which we will call {\em basic 
-multi-atlas segmentation}), involves three steps. First, each labelled 
+from one image to bring them into alignment with the other image. We use the
+term {\em atlas} to mean a manually segmented image, and the term {\em template}
+to mean an automatically segmented image (i.e.  via label propagation). The
+terms {\em atlas library} and {\em template library} describe any set of such
+images. Additionally, we use the term {\em target} to refer to an unlabelled
+image that is undergoing segmentation.
+
+The simplest form of multi-atlas segmentation, which we call {\em basic 
+multi-atlas segmentation}, involves three steps. First, each labelled 
 input image (i.e. atlas or template) is registered to an unlabelled target image.
@@
 basic multi-atlas segmentation degenerates into model-based segmentation: labels
-are propogated from the atlas to a target, and no label fusion is needed.
+are propagated from the atlas to a target, and no label fusion is needed.
 
@@
 schizophrenia (Experiment 3), and Alzheimer's disease (Experiment 4).
-Additionally, in Experiment 4, we compare \mb segmentations with those of
-well-known automated methods and assess segmentation bias.
+Additionally, in Experiment 4, we compared \mb segmentations with those of
+well-known automated methods and assessed segmentation bias.
 
@@
 Materials for detailed discussion of the segmentation process used). Throughout
-this paper, we will use these segmentations as the \gs for comparison.
+this paper, we use these segmentations as the \gs for comparison.
 
@@
   \caption{Mean Dice's Similarity Coefficient of \mb segmentations relative to
-  SNT \gs segmentations for 69 ADNI1 subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method. Error bars indictate
+  SNT \gs segmentations for 69 ADNI subjects vs. atlas and template library
+  size, registration algorithm, and label fusion method. Error bars indicate
   standard error.
@@
 each disease category, and illustrates some of the common errors found during 
-visual inspection. Mostly frequently, we find \mb improperly includes the vestigial
+visual inspection. Mostly frequently, we found \mb improperly includes the vestigial
 hippocampal sulcus and, although not anatomically incorrect, \mb under-estimates 
@@
   \end{centering}
-  \caption{SNT \gs and \mb segmentations for 30 ADNI1 subjects (10
+  \caption{SNT \gs and \mb segmentations for 30 ADNI subjects (10
   subjects randomly selected from each disease category in the subject pool used
@@
 \citet{Heckemann2006a} and, subsequently, other groups have not even considered
-the potential pitfalls of an even number of candidate labels (eg.
+the potential pitfalls of an even number of candidate labels (e.g.
 \citet{Leung2010}).
@@
 leave-one-out-validation (Table \ref{tab:other-methods}), aside from
-\citep{Wang2011} who perform a similarily rigorous validation as us. The
+\citep{Wang2011} who perform a similarly rigorous validation as us. The
 thoroughness of our validation suggests that our results are reflective of a
@@
 poor in general. As previously noted in Figure \ref{fig:ADNI-segmentations}, an
-expert manual rater (JW) identified many inconsistencies: occurances of over-
+expert manual rater (JW) identified many inconsistencies: occurrences of over-
 and under-estimation, as well as misalignments of the entire segmentation
@@
 
-\section*{R Session Information}
-<<results = "asis">>=
-toLatex(sessionInfo(), locale=FALSE)
-@
+% \section*{R Session Information}
+% <<results = "asis">>=
+% toLatex(sessionInfo(), locale=FALSE)
+% @
 \end{document}

### 2013-09-03 paper.Rnw
Fixes for Mallar's recent edits

@@
 \renewcommand\Affilfont{\itshape\small}  %authblk
+\linespread{1.3} % one and a half line spacing
 
 % shortcuts 
-\newcommand{\mb}{MAGeT-Brain }
-\newcommand{\ants}{ANTS }
-\newcommand{\animal}{ANIMAL }
-\newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T }
-\newcommand{\FEPdataset}{SZ-FEP }
-\newcommand{\gs}{Gold Standard }
-\newcommand{\fsl}{FSL }
-\newcommand{\freesurfer}{FreeSurfer }
-\newcommand{\maper}{MAPER }
+\newcommand{\mb}{MAGeT-Brain}
+\newcommand{\ants}{ANTS}
+\newcommand{\animal}{ANIMAL}
+\newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T}
+\newcommand{\FEPdataset}{SZ-FEP}
+\newcommand{\gs}{Gold Standard}
+\newcommand{\fsl}{FSL}
+\newcommand{\freesurfer}{FreeSurfer}
+\newcommand{\maper}{MAPER}
 
 % title and authors
-\title{Bootstrapping Multi-atlas Hippocampal Subfield Segmentation by using
-Multiple Automatically Generated Templates}
+%\title{Bootstrapping Multi-atlas Hippocampal Subfield Segmentation by using
+%Multiple Automatically Generated Templates}
+\title{Bootstrapping Multi-atlas Segmentation Using Multiple Automatically
+Generated Templates for the Segmentation of the Whole Hippocampus and
+Subfields}
 \author[1]{Jon Pipitone}
@@
 \author[1]{Julie Winterburn}
-\author[1,8]{Tristram A. Lett}
+\author[1,9]{Tristram A. Lett}
 \author[2,3]{Jason P. Lerch}
@@
 \author[4]{Martin Lepage}
-\author[1,6,8]{Aristotle N. Voineskos}
-\author[1,6,7]{M. Mallar Chakravarty}
+\author[1,6,9]{Aristotle N. Voineskos}
+\author[1,6,8]{M. Mallar Chakravarty}
 \author[ ]{the Alzheimer's Disease Neuroimaging Initiative\footnote{Data
@@
 Mental Health, Toronto, ON, Canada}
-\affil[2]{Neurosciences and Mental Health, Hospital for Sick Children, Toronto,
-ON,Canada}
+\affil[2]{Neurosciences and Mental Health Laboratory, Hospital for Sick
+Children, Toronto, ON, Canada}
 \affil[3]{Department of Medical Biophysics, University of Toronto, Toronto, ON,
@@
 \affil[6]{Department of Psychiatry, University of Toronto, Toronto, ON, Canada}
-\affil[7]{Institute of Biomedical Engineering, University of Toronto, Toronto,
-ON, Canada}
-\affil[8]{Institute of Medical Science, University of Toronto, Toronto, ON, Canada}
+\affil[7]{Institute of Biomaterials and Biomedical Engineering, University of
+Toronto, Toronto, ON, Canada}
+\affil[8]{Rotman Research Institute, Baycrest, Toronto, ON, Canada}
+\affil[9]{Institute of Medical Science, University of Toronto, Toronto, ON, Canada}
 
@@
 
-\textbf{Introduction:} Neuroimaging research often relies on automated anatomical
-segmentations of MR images of the brain. Current multi-atlas based approaches
-provide accurate segmentations of brain images by propagating manually derived
-segmentations of specific neuroanatomical structures to unlabelled data. These
-approaches often rely on a large number of such manually segmented atlases (e.g.
-30-80) that take significant time and expertise to produce. We present an
-algorithm, \mb, for the automatic segmentation of the hippocampus that minimizes
+\textbf{Introduction:} 
+Advances in image segmentation of magnetic resonance images (MRI) have
+demonstrated that multi-atlas approaches improve segmentation accuracy and
+precision over regular atlas-based approaches.  These approaches often rely on
+a large number of such manually segmented atlases (e.g.  30-80) that take
+significant time and expertise to produce. We present an algorithm, \mb (\mb
+({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
+T}emplates), for the automatic segmentation of the hippocampus that minimizes
 the number of atlases needed while still achieving similar accuracy to
-multi-atlas approaches.\\
+multi-atlas approaches. Thus, our method acts as an accurate multi-atlas
+approach when using special, hard-to-define atlases that are laborious to
+construct. \\
 \textbf{Method:} \mb works by propagating atlas segmentations to a template library,
@@
 high-resolution atlases. Two final experiments assess \mb when applied to
-Schizophrenia and Alzheimer's disease populations, and \mb segmentations are
-compared with existing automated methods (\fsl, \freesurfer, \maper) and biases are
-explored.\\
+first episode psychosis and Alzheimer's disease populations, and \mb
+segmentations are compared with existing automated methods (\fsl, \freesurfer,
+\maper) and biases are explored.\\
 \textbf{Results:} Using nine atlases and 19 template images, \mb achieves a
@@
 and shows significantly lower variability in DSC than multi-atlas segmentation.
-In a LOOCV, \mb reproduces the volumes of the CA1, CA4/DG, and SR/SL/SM
-hippocampal subregions with a percent error in volume that is at or lower than
-that produced by image resampling. \mb produces hippocampal volumes in a
-Schizophrenia patient population that are highly correlated with expert manual
-segmentation volumes (Pearson $r = 0.877, t = 16.244, p < 0.001$). Compared to
-\fsl and \freesurfer, \mb shows much smaller fixed volume bias (within $250
-mm^3$ on average) to \gs segmentations, as well as a conservative, rather than
-exaggerated, proportional volume bias.\\
-\textbf{Conclusion:} We demonstrate that \mb produces accurate hippocampal
-segmentations using only 5 atlases over different hippocampal definitions,
-disease populations, and acquisition types, as well as showing that accurate
-identification of the hippocampal subfields is possible.
+In a LOOCV, \mb reproduces the volumes of the cornu ammonis (CA) 1; CA4/dentate
+gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
+moleculare (SM) hippocampal subregions with a percent error in volume that is
+at or lower than that produced by image resampling. \mb produces hippocampal
+volumes in a first episode psychosis patient population that are highly
+correlated with expert manual segmentation volumes (Pearson $r = 0.877, t =
+16.244, p < 0.001$). Compared to \fsl and \freesurfer, \mb shows much smaller
+fixed volume bias (within $250 mm^3$ on average) to \gs segmentations available
+from ADNI, as well as a conservative, rather than exaggerated, proportional
+volume bias.\\ \textbf{Conclusion:} We demonstrate that \mb produces accurate
+hippocampal segmentations using only 5 atlases over different hippocampal
+definitions, disease populations, and acquisition types, as well as showing
+that accurate identification of the hippocampal subfields is possible.
 \end{abstract}
@@
 \citep{DenHeijer2012,Scoville2000}. The hippocampus is of interest to
-neuroscientists because it is implicated in many forms of brain dysfunction,
-including Alzheimer's disease \citep{Sabuncu2011} and schizophrenia
-\citep{Narr2004,Karnik-Henry2012}. In neuroimaging studies, structural magnetic
-resonance images (MRI) are often used for the volumetric assessment of the
-hippocampus.  As such, accurate segmentation of the hippocampus and its
-subregions in MRI is a necessary first step to better understand the
-inter-individual variability of subject neuroanatomy.
-
-Typically, the gold standard for neuroanatomical segmentation is manual
-delineation by an expert human rater.  However, with the availability of
-increasingly large MRI datasets, the time and expertise required for manual
-segmentation becomes prohibitive
-\citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. Further, there is
-significant variation between segmentation protocols regarding specific
-anatomical boundaries of the hippocampus \citep{Geuze2004} and this has led to
-efforts to create an unified hippocampal segmentation protocol \citep{Jack2011}.
+clinical neuroscientists because it is implicated in many forms of brain
+dysfunction, including Alzheimer's disease \citep{Sabuncu2011} and
+schizophrenia \citep{Narr2004,Karnik-Henry2012}. In neuroimaging studies,
+structural magnetic resonance images (MRI) are often used for the volumetric
+assessment of the hippocampus.  As such, accurate segmentation of the
+hippocampus and its subregions in MRI is a necessary first step to better
+understand the inter-individual variability of subject neuroanatomy.
+
+The gold standard for neuroanatomical segmentation is manual delineation by an
+expert human rater.  However, with the availability of increasingly large MRI
+datasets, the time and expertise required for manual segmentation becomes
+prohibitive \citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. This
+effort is complicated by the fact that there is significant variation between
+segmentation protocols with respect to specific anatomical boundaries of the
+hippocampus \citep{Geuze2004} and this has led to efforts to create an unified
+hippocampal segmentation protocol \citep{Jack2011}.  In addition, there is
+controversy over the appropriate manual segmentation protocol to use in a
+particular imaging study \citep{Nestor2012}.  Thus, a segmentation algorithm
+that can easily adapt to different manual segmentation definitions would be of
+significant benefit to the neuroimaging community.
 
@@
 
-To the best of our knowledge there are only two automated segmentation methods
-that attempt to define the hippocampus and its subfields using standard
-T1-weighted data. The first is included with the \freesurfer package
+Also of interest to this manuscript are methods that attempt to define
+hippocampal subfields using standard T1-weighted data.  To the best of our
+knowledge, there are only two automated segmentation algorithms that attempt
+this problem. The first is included with the \freesurfer package
 \citep{VanLeemput2009}. This work is limited as it omits the tail of the
@@
 high-resolution MRI data from post-mortem medial temporal lobe samples. Using
-nonlinear registration guided by manually derived hippocampus masks and specific
-landmarks, the authors demonstrated accurate parcellation of hippocampal
-subfields in unlabelled MRI volumes.
-
-In this paper we undertook a thorough validation of the \mb algorithm for
-segmentation of the hippocampus. First, we addressed the feasibility and
+nonlinear registration guided by manually derived hippocampus masks and
+specific landmarks, the authors demonstrated accurate parcellation of
+hippocampal subfields in unlabelled MRI volumes.
+
+In this paper we have describe a thorough validation of the \mb algorithm for
+segmentation of the hippocampus and its subfields. First, we addressed the feasibility and
 accuracy of whole hippocampus segmentation with a limited number of input
-atlases by performing a multi-fold experiment using a subset of the Alzheimer's
-Disease Neuroimaging Initiative (ADNI) dataset over a range of parameters.  We
-then performed a leave-one-out validation to determine if hippocampal subfields
-can be accurately identified using our multi-atlas framework.  To ensure that we
-have not over-fit our parameters to the aging or neurodegenerative brain, we
-also applied our segmentations to a dataset of individuals suffering from first
-episode psychosis.  Finally, validated our algorithm using all of the data
-available in the \adnidataset sample and compare our segmentations to the other
+atlases \citep{MallarChakravarty2012} by performing a multi-fold experiment
+using a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
+dataset over a range of parameters.  We then performed a leave-one-out
+validation to determine if hippocampal subfields can be accurately identified
+using our multi-atlas framework.  To ensure that we have not over-fit our
+parameters to the aging or neurodegenerative brain, we also applied our
+segmentations to a dataset of individuals suffering from first episode
+psychosis.  Finally, validated our algorithm using all of the data available in
+the \adnidataset sample and compare our segmentations to the other
 segmentations available through the ADNI informatics portal.
@@
 described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
-2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240 x
+2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
@@
 atlases, we also obtained separate 3T T1 BRAVO acquisitions (0.9mm-isotropic
-voxels) of four of the five Winterburn atlas subjects.
+voxels) of four of the five Winterburn atlas subjects.  Images were acquired on
+a 3T GE Discovery MR 750 system (General Electric, Milwaukee, WI) using an
+8-channel head coil with the enhanced fast gradient recalled echo 3-dimensional
+acquisition protocol, EFGRE-BRAVO, with the following parameters: $TE/TR/TI =
+3.0ms/6.7ms/650ms$, flip angle $=8^\circ$, $FOV=15.8cm$, slice
+thickness$=0.9mm$, 176 in-plane steps for an approximate isotropic resolution
+of 0.9mm dimension voxels.
 
@@
 typical segmentation scenario (high-resolution atlases and lower-resolution
-subjects).  The template library is composed of all five subject images, plus an
-additional set of 3T T1 images (0.9mm-isotropic voxels) of healthy subjects
-acquired separately (Table \ref{tab:WAval-healthy-demographics}). The optimal
-size of template library, registration method, and label fusion method found in
-Experiment 1 are used.  As with traditional LOOCV, each subject image is
-segmented by \mb with that subject's image excluded from the atlas library.  In
-part 2b of this experiment, we carry out a similar validation in which the
-subject images are separately acquired T1 BRAVO images of four of the five
-Winterburn atlas subjects (referred to as the {\em T1 BRAVO} dataset).  Figure
-\ref{fig:LOOCV-schematic} illustrates the experimental set up of both parts of
-this experiment.
+subjects).  The template library is composed of all five subject images, plus
+an additional set of 16 of healthy subjects 3T T1 images acquired separately
+(Table \ref{tab:WAval-healthy-demographics}).  Images were acquired on a 3T GE
+Discovery MR 750 system (General Electric, Milwaukee, WI) using an 8-channel
+head coil with the enhanced fast gradient recalled echo 3-dimensional
+acquisition protocol, EFGRE-BRAVO, with the following parameters: $TE/TR/TI =
+3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution
+of 0.9mm dimension voxels.
+
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used.  As with traditional LOOCV, each subject
+image is segmented by \mb with that subject's image excluded from the atlas
+library.  In part 2b of this experiment, we carry out a similar validation in
+which the subject images are separately acquired T1 BRAVO images of four of the
+five Winterburn atlas subjects (referred to as the {\em T1 BRAVO} dataset).
+Figure \ref{fig:LOOCV-schematic} illustrates the experimental set up of both
+parts of this experiment.
 
@@
 caption = 
-  "\\textbf{Subfield validation healthy control demographics.}
-  CN     - Cognitively Normal. 
-  LMCI   - Late-onset Mild Cognitive Impairment. 
-  AD     - Alzheimer's Disease.  
-  Hisp   - Hispanic. 
-  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
-  ADAS   - Alzheimer's Disease Assessment Scale.  
-  MMSE   - Mini-Mental State Examination."
+  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
+control subject sample.} Education is shown in years."
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
@@
 with sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
-angle=$30{\circ}$, 180 1mm contiguous sagittal slices). The rectangular
+angle=$30^{\circ}$, 180 1mm contiguous sagittal slices). The rectangular
 field-of-view for the images was 256mm (SI)$\times$204mm (AP).  Subject
@@
 methods using different protocols. In place of an overlap metric, we can assess
-the degree of correlation in average bilateral hippocampal volume of the
-subjects produced by each method. Specifically, a Pearson correlation is used.
+the degree of (Pearson) correlation in average bilateral hippocampal volume of the
+subjects produced by each method.
 
@@
 (Figure \ref{ADNI1-xval-variability}).  The size of template library 
-necessary to reach a significant ($p < 0.05$) decrease in variance and standard 
-deviation grows with the size of atlas library used.  A template library of 19
-images is sufficient to show significant decrease in variance and standard 
-deviation for 3-7 atlases.
+necessary to reach a decrease ($p < 0.05$) in variance and standard deviation
+grows with the size of atlas library used.  A template library of 19 images is
+sufficient to show significant decrease in variance and standard deviation for
+3-7 atlases.
 
@@
     \end{centering}
-    \caption{Detailed subfield segmentation results of a single subject comparing the
-    original Winterburn atlas subfield segmentation (at 0.3mm-isotropic voxel
-    resolution), the Winterburn segmentation resampled (at 0.9mm-isotropic
-    voxel resolution), the \mb segmentation of the subsampled atlas image, and
-    the \mb segmentation of a separate T1 BRAVO acquisition of the subject.}
-    \label{fig:subfield-montage}
+    \caption{Detailed subfield segmentation results for a single subject. In
+    the upper left corner is the original high-resolution Winterburn atlas manual
+    subfield segmentation; in the upper right corner is the Winterburn atlas
+    segmentation resampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
+    corner is the \mb segmentation of the resampled Winterburn atlas images; in the
+    lower right corner is the \mb segmentation of a separately acquired T1 BRAVO
+    image of the same subject. In each segmentation, slices from the left
+    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
+    slices from inferior to superior; the second row shows sagittal slices from
+    lateral to medial; the third row shows coronal slices from anterior to
+    posterior.  \label{fig:subfield-montage}}
 \end{sidewaysfigure}
@@
 manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
-$t = \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001,
-fep.cor.test$p.value)}$; Figure \ref{fig:SZ-volumes}).
+$t = \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$; 
+Figure \ref{fig:SZ-volumes}).
 
@@
 
-On that note, it is worth observing that we found the SNT segmentations to be
-poor in general. As previously noted in Figure \ref{fig:ADNI-segmentations}, an
-expert manual rater (JW) identified many inconsistencies: occurrences of over-
-and under-estimation, as well as misalignments of the entire segmentation
-volume.  Therefore, the use of the SNT segmentations as gold standards in
+On that note, one author (JW), an expert manual rater, identified regular
+inconsistencies in the SNT segmentations: occurrences of over- and
+under-estimation, as well as misalignments of the entire segmentation volume
+(Figure \ref{fig:ADNI-segmentations}). Although the SNT segmentations are used
+as gold standards for validation in many other studies (Table
+\ref{tab:other-methods}) providing a benchmark for comparison, their use in
 Experiments 1 and 4 presents the possibility that the DSC scores and
-correlations we obtained are lower than what would be found with a more accurate
-and consistent gold standard.
+correlations we obtained are lower than what would be found with a more
+accurate and consistent gold standard.
 
@@
 literature although very few automated methods are available for the
-segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. While
-recent work demonstrates that subfield segmentations can be used for
-classification of AD, MCI, and CN, there has been no explicit validation of the
-methodology based on accuracy or precision.  Although the initial hippocampal
-subfield segmentations from the Yushkevich group have been demonstrated to work
-on the ADNI population, there has also been no validation of their work. In
-addition, their work requires some manual initialization to properly function.
-Our work demonstrates that we can reliably identify the CA1, subiculum, and
-CA4/dentate with only modest amounts of error.  The fact that CA2/CA3 and
-molecular layers cannot be reasonably identified should not be surprising as
-these are extremely thin and spatially convoluted regions that originally
-required high-resolution MRI for identification.  It is likely that the extents of
-these regions are well below the resolution offered by standard T1-weighted
-images and, in fact, many manual segmentation methodologies do not attempt to
-parse these regions either \citep{Wisse2012, Mueller2009}.
+segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. While recent
+work demonstrates that subfield segmentations can be used for classification of
+AD, MCI, and CN, there has been no explicit validation of the methodology based
+on accuracy or precision.  Although the initial hippocampal subfield
+segmentations from the Yushkevich group have been demonstrated to work on the
+ADNI population, there has also been no validation of their work. In addition,
+their work requires some manual initialization to properly function.  Our work
+demonstrates that we can reliably identify the CA1, subiculum, and CA4/dentate
+with only modest amounts of error.  The fact that CA2/CA3 and molecular layers
+cannot be reasonably identified should not be surprising as these are extremely
+thin and spatially convoluted regions that originally required high-resolution
+MRI for identification.  It is likely that the extents of these regions are
+well below the resolution offered by standard T1-weighted images and, in fact,
+many manual segmentation methodologies do not attempt to parse these regions
+either \citep{Wisse2012, Mueller2009}.
 

### 2013-09-03 paper.Rnw
Fixed up uses of newcommands

@@
 a large number of such manually segmented atlases (e.g.  30-80) that take
-significant time and expertise to produce. We present an algorithm, \mb (\mb
+significant time and expertise to produce. We present an algorithm, \mb{} (
 ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
@@
 construct. \\
-\textbf{Method:} \mb works by propagating atlas segmentations to a template library,
+\textbf{Method:} \mb{} works by propagating atlas segmentations to a template library,
 formed from a subset of target images, via transformations estimated by non-linear
@@
 cross-validation (LOOCV) of hippocampal subfield segmentation using five
-high-resolution atlases. Two final experiments assess \mb when applied to
-first episode psychosis and Alzheimer's disease populations, and \mb
-segmentations are compared with existing automated methods (\fsl, \freesurfer,
-\maper) and biases are explored.\\
-\textbf{Results:} Using nine atlases and 19 template images, \mb achieves a
+high-resolution atlases. Two final experiments assess \mb{} when applied to
+first episode psychosis and Alzheimer's disease populations, and \mb{}
+segmentations are compared with existing automated methods (\fsl{}, \freesurfer{},
+\maper{}) and biases are explored.\\
+\textbf{Results:} Using nine atlases and 19 template images, \mb{} achieves a
 mean Dice's Similarity Coefficient (DSC) of 0.84 over 10-folds of Monte Carlo
-cross-validation on 69 subjects relative to semi-automated \gs segmentations,
+cross-validation on 69 subjects relative to semi-automated \gs{} segmentations,
 and shows significantly lower variability in DSC than multi-atlas segmentation.
-In a LOOCV, \mb reproduces the volumes of the cornu ammonis (CA) 1; CA4/dentate
+In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1; CA4/dentate
 gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
 moleculare (SM) hippocampal subregions with a percent error in volume that is
-at or lower than that produced by image resampling. \mb produces hippocampal
+at or lower than that produced by image resampling. \mb{} produces hippocampal
 volumes in a first episode psychosis patient population that are highly
 correlated with expert manual segmentation volumes (Pearson $r = 0.877, t =
-16.244, p < 0.001$). Compared to \fsl and \freesurfer, \mb shows much smaller
-fixed volume bias (within $250 mm^3$ on average) to \gs segmentations available
+16.244, p < 0.001$). Compared to \fsl{} and \freesurfer{}, \mb{} shows much smaller
+fixed volume bias (within $250 mm^3$ on average) to \gs{} segmentations available
 from ADNI, as well as a conservative, rather than exaggerated, proportional
-volume bias.\\ \textbf{Conclusion:} We demonstrate that \mb produces accurate
+volume bias.\\ \textbf{Conclusion:} We demonstrate that \mb{} produces accurate
 hippocampal segmentations using only 5 atlases over different hippocampal
@@
 the use of multiple manually segmented atlases and probabilistic segmentation
-techniques, such as those found in the \freesurfer package \citep{Fischl2002}.
-\freesurfer uses a probabilistic atlas of anatomical and tissue classes along
+techniques, such as those found in the \freesurfer{} package \citep{Fischl2002}.
+\freesurfer{} uses a probabilistic atlas of anatomical and tissue classes along
 with spatial constraints for class labels encoded using a Markov random field
@@
 hippocampus in the ADNI-1 baseline dataset, achieving a mean Dice score of 0.85
-against \gs segmentations.
+against \gs{} segmentations.
 
@@
 knowledge, there are only two automated segmentation algorithms that attempt
-this problem. The first is included with the \freesurfer package
+this problem. The first is included with the \freesurfer{} package
 \citep{VanLeemput2009}. This work is limited as it omits the tail of the
@@
 
-In this paper we have describe a thorough validation of the \mb algorithm for
+In this paper we have describe a thorough validation of the \mb{} algorithm for
 segmentation of the hippocampus and its subfields. First, we addressed the feasibility and
@@
 psychosis.  Finally, validated our algorithm using all of the data available in
-the \adnidataset sample and compare our segmentations to the other
+the \adnidataset{} sample and compare our segmentations to the other
 segmentations available through the ADNI informatics portal.
@@
 
-\subsection{The \mb Algorithm}
+\subsection{The \mb{} Algorithm}
 In this paper, we use the term {\em label} to mean any segmentation (manual or
@@
 
-\mb ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
+\mb{} ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
 T}emplates) bootstraps the creation of a large template library given a limited
@@
 
-Figure \ref{fig:MAGeT} illustrates the \mb algorithm graphically. Source code
-for \mb can be found at \url{http://github.com/pipitone/MAGeTbrain}.
+Figure \ref{fig:MAGeT} illustrates the \mb{} algorithm graphically. Source code
+for \mb{} can be found at \url{http://github.com/pipitone/MAGeTbrain}.
 
@@
   \end{centering}
-  \caption{A schematic illustration of basic multi-atlas segmentation and \mb
+  \caption{A schematic illustration of basic multi-atlas segmentation and \mb{}
   segmentation \label{fig:MAGeT}}
@@
 The following section describes for experiments conducted to assess the
-segmentation quality of the \mb algorithm. The first two experiments assess the
-validity of \mb using a cross-validation design. Experiment 1 investigates the
-accuracy of \mb whole hippocampus segmentation over a wide range of parameter
+segmentation quality of the \mb{} algorithm. The first two experiments assess the
+validity of \mb{} using a cross-validation design. Experiment 1 investigates the
+accuracy of \mb{} whole hippocampus segmentation over a wide range of parameter
 settings. The results of this experiment enable us to choose the parameter
@@
 leave-one-out cross-validation design. The last two experiments assess the
-validity of the \mb algorithm when applied to different diseases: first episode
+validity of the \mb{} algorithm when applied to different diseases: first episode
 schizophrenia (Experiment 3), and Alzheimer's disease (Experiment 4).
-Additionally, in Experiment 4, we compared \mb segmentations with those of
+Additionally, in Experiment 4, we compared \mb{} segmentations with those of
 well-known automated methods and assessed segmentation bias.
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT \gs segmentations from the ADNI dataset \citep{Hsu2002}. This
+of images and SNT \gs{} segmentations from the ADNI dataset \citep{Hsu2002}. This
 form of cross-validation allows us to rigorously validate a large number of
-parameter settings of \mb (atlas and template library sizes, registration
+parameter settings of \mb{} (atlas and template library sizes, registration
 algorithm, and label fusion method) and select the best parameters to use in
@@
 
-\paragraph{\adnidataset dataset} 
+\paragraph{\adnidataset{} dataset} 
 Data used in the preparation of this article were obtained from the Alzheimer’s
@@
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
-\adnidataset} standardized dataset. 23 subjects were chosen from each disease
+\adnidataset{}} standardized dataset. 23 subjects were chosen from each disease
 category: cognitively normal (CN), mild cognitive impairment (MCI) and
@@
 Materials for detailed discussion of the segmentation process used). Throughout
-this paper, we use these segmentations as the \gs for comparison.
+this paper, we use these segmentations as the \gs{} for comparison.
 
@@
 authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset used was the "\adnidataset" standardized dataset
+August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
 available from ADNI \footnote{
@@
 
-In this experiment, our dataset consists of 69 1.5T images and corresponding \gs
+In this experiment, our dataset consists of 69 1.5T images and corresponding \gs{}
 segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and \gs segmentations used as an atlas
+training set consisting of images and \gs{} segmentations used as an atlas
 library, and a validation set consisting of the remaining images segmented by
-both \mb and multi-atlas. The computed segmentations were compared to the \gs
+both \mb{} and multi-atlas. The computed segmentations were compared to the \gs{}
 segmentations (see below).
@@
 explored are: atlas library size (1-9), template library size (1-20),
-registration method (\ants or \animal, described below), and label fusion method (majority vote,
+registration method (\ants{} or \animal{}, described below), and label fusion method (majority vote,
 cross-correlation weighted majority vote, and normalized mutual information
@@
 \subparagraph{Automatic Normalization and Image Matching and Anatomical
-Labeling (\animal)}
-
-The \animal algorithm carries out image registration in two phases. In the
+Labeling (\animal{})}
+
+The \animal{} algorithm carries out image registration in two phases. In the
 first, a 12-parameter linear transformation (3 translations, rotations, scales,
@@
 brain \citep{Collins}. In the second phase, nonlinear registration is completed
-using the \animal algorithm \citep{Collins1995}: an iterative procedure that
+using the \animal{} algorithm \citep{Collins1995}: an iterative procedure that
 estimates a 3D deformation field between two MR images. At first, large
@@
 \scriptsize
-\caption{\textbf{\animal registration parameters.} \label{tab:ANIMAL-params}}
+\caption{\textbf{\animal{} registration parameters.} \label{tab:ANIMAL-params}}
 \begin{center}
@@
 
-\subparagraph{Automatic Normalization Tools (\ants)}
-
-\ants is a diffeomorphic registration algorithm which provides great flexibility
+\subparagraph{Automatic Normalization Tools (\ants{})}
+
+\ants{} is a diffeomorphic registration algorithm which provides great flexibility
 over the choice of transformation model, objective function, and the consistency
@@
 the objective function are regularized with a Gaussian kernel at each level of
-the hierarchy. The \ants algorithm is freely available
-\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the \ants
+the hierarchy. The \ants{} algorithm is freely available
+\url{http://www.picsl.upenn.edu/ANTS/}. We used an implementation of the \ants{}
 algorithm compatible with the MINC data format, mincANTS
@@
 
-We used the following command line when running \ants:
+We used the following command line when running \ants{}:
 {\small
@@
 These settings were adapted from the "reasonable starting point" given in the
-\ants manual 
+\ants{} manual 
 \footnote{\url{https://sourceforge.net/projects/advants/files/Documentation/}}.
@@
 
-  The {\tt xcorr\_vol} utility from the \animal toolkit is used to calculate the
+  The {\tt xcorr\_vol} utility from the \animal{} toolkit is used to calculate the
   cross-correlation similarity measure.
@@
 
-The automatically produces labels are compared to the \gs SNT labels provided by
+The automatically produces labels are compared to the \gs{} SNT labels provided by
 ADNI. The segmentation accuracy reported for a parameter setting is the average
@@
 
-In order to investigate the performance of \mb in a real world setting in which 
+In order to investigate the performance of \mb{} in a real world setting in which 
 only one size of atlas and template libraries are used, we explore the
@@
 rounds per subject. The distribution of these statistics across all subjects is
-then compared between \mb and multi-atlas using a Student's t-test. A
+then compared between \mb{} and multi-atlas using a Student's t-test. A
 significant difference between distributions is taken to show either a larger or
@@
 \subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
-In this experiment, the accuracy of the \mb algorithm on hippocampal subregion
+In this experiment, the accuracy of the \mb{} algorithm on hippocampal subregion
 segmentation is tested using a leave-one-out cross-validation (LOOCV) design on
-two image sets. The optimal parameter settings for \mb found in Experiment 1 are
+two image sets. The optimal parameter settings for \mb{} found in Experiment 1 are
 used in this experiment.
@@
 In part 2a of this experiment, we have modified the LOOCV experimental
-design in order to compare \mb segmentation accuracy with the accuracy of image
+design in order to compare \mb{} segmentation accuracy with the accuracy of image
 resampling error. The five $0.3mm$-isotropic voxel Winterburn atlas images and
@@
 0.9mm-isotropic voxel resolution (referred to as the {\em Subsampled} dataset)
-as input subjects. Subsampling the target images allows us to assess \mb in a
+as input subjects. Subsampling the target images allows us to assess \mb{} in a
 typical segmentation scenario (high-resolution atlases and lower-resolution
@@
 method found in Experiment 1 are used.  As with traditional LOOCV, each subject
-image is segmented by \mb with that subject's image excluded from the atlas
+image is segmented by \mb{} with that subject's image excluded from the atlas
 library.  In part 2b of this experiment, we carry out a similar validation in
@@
 \paragraph{Evaluation method}  
-To assess the \mb LOOCV segmentations we compute the relative percent error in 
+To assess the \mb{} LOOCV segmentations we compute the relative percent error in 
 hippocampal volume with respect to the full resolution Winterburn atlas 
@@
 Winterburn atlas labels resampled (with nearest-neighbour interpolation) to
-0.9mm-isotropic voxels, we obtain a baseline error against which to assess \mb
+0.9mm-isotropic voxels, we obtain a baseline error against which to assess \mb{}
 segmentations.
@@
 %%%%%                                                                     %%%%%
-%%%%%         Experiment 3: Application of \mb to the segmentation        %%%%%
+%%%%%         Experiment 3: Application of \mb{} to the segmentation        %%%%%
 %%%%%                       of schizophrenia patients                     %%%%%
@@
 
-To validate that the \mb works effectively in the context of other neurological
+To validate that the \mb{} works effectively in the context of other neurological
 disorders, in this experiment we use the Winterburn atlases to derive 
@@
 
-\paragraph{\FEPdataset dataset}
+\paragraph{\FEPdataset{} dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 \paragraph{Experiment details} 
-\mb is configured with an atlas library composed of the Winterburn T1 atlases
+\mb{} is configured with an atlas library composed of the Winterburn T1 atlases
 (see Experiment 2) ignoring subfield delineations.  All images from the
-\FEPdataset dataset are segmented by \mb.  The optimal size of template library,
+\FEPdataset{} dataset are segmented by \mb{}.  The optimal size of template library,
 registration method, and label fusion method found in Experiment 1 are used.
@@
 %%%%%                                                                     %%%%%
-%%%%%         EXPERIMENT 4: Application of \mb to the segmentation        %%%%%
+%%%%%         EXPERIMENT 4: Application of \mb{} to the segmentation        %%%%%
 %%%%%                       of ADNI subjects                              %%%%%
@@
 
-To validate \mb segmentation quality with respect to other established automated
-hippocampal segmentation methods, \mb was applied to large dataset from the ADNI
+To validate \mb{} segmentation quality with respect to other established automated
+hippocampal segmentation methods, \mb{} was applied to large dataset from the ADNI
 project and the resulting segmentations were compared to those determined by
-\freesurfer, \fsl, \maper and the SNT \gs segmentation.
+\freesurfer{}, \fsl{}, \maper{} and the SNT \gs{} segmentation.
 
 \paragraph{ADNI1 dataset revisited} 
-The {\em \adnidataset} standardized dataset contains $1919$ images in total.
-SNT and \freesurfer hippocampal volumes for a subset of images were provided by
-ADNI, along with quality control data for each \freesurfer segmentation
+The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
+SNT and \freesurfer{} hippocampal volumes for a subset of images were provided by
+ADNI, along with quality control data for each \freesurfer{} segmentation
 (guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
-the entire \adnidataset dataset are shown in Table
+the entire \adnidataset{} dataset are shown in Table
 \ref{tab:ADNI1-Dataset-Demographics}.
@@
 \paragraph{Experiment details} 
-\mb was configured with an atlas library composed of the five Winterburn T1
+\mb{} was configured with an atlas library composed of the five Winterburn T1
 atlases (described in Experiment 2), and size of template library,
 registration method, and label fusion method were determined by the optimal
-settings found in Experiment 1.  The \freesurfer segmentation method was used via
+settings found in Experiment 1.  The \freesurfer{} segmentation method was used via
 the \verb+run_first_all+ script according to the FIRST user guide
 \footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
-in the \adnidataset dataset were segmented by both methods.
-
-One author (MP) performed visual quality inspection for \mb and \freesurfer 
+in the \adnidataset{} dataset were segmented by both methods.
+
+One author (MP) performed visual quality inspection for \mb{} and \freesurfer{} 
 segmentations using similar quality control guidelines (if either hippocampus
@@
 the segmentation did not pass).  Only images meeting the conditions of having
-segmentations from all methods (SNT, \maper, \freesurfer, \fsl, and \mb) and also
+segmentations from all methods (SNT, \maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also
 passing quality control inspection were included in the analysis.
@@
 \paragraph{Evaluation method}
-As in Experiment 3, the SNT \gs and Winterburn hippocampal segmentation
+As in Experiment 3, the SNT \gs{} and Winterburn hippocampal segmentation
 protocols differ in the neuroanatomical features delineated, and so we 
-assessed \mb by the degree of (Pearson) correlation of average hippocampal volume
+assessed \mb{} by the degree of (Pearson) correlation of average hippocampal volume
 across subjects. We also computed the correlation in hippocampal volume between
-existing, established automated segmentation methods -- \fsl, \freesurfer, and
-\maper, and SNT semi-automated segmentations. Additionally, we evaluate the
+existing, established automated segmentation methods -- \fsl{}, \freesurfer{}, and
+\maper{}, and SNT semi-automated segmentations. Additionally, we evaluate the
 volume-related fixed and proportional biases in all segmentation methods using
@@
 
-In this experiment we conducted 10 rounds of \mb and multi-atlas segmentation of
+In this experiment we conducted 10 rounds of \mb{} and multi-atlas segmentation of
 of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
-images respectively, two registration algorithms (\ants or \animal), and three
+images respectively, two registration algorithms (\ants{} or \animal{}), and three
 label fusion techniques (unweighted, cross-correlation, and normalised mutual
-information weighted majority vote).  Hippocampal \mb-based segmentations using
-both \animal and \ants registration algorithms demonstrate good overlap with SNT
-\gs segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=2)}
+information weighted majority vote).  Hippocampal \mb{}-based segmentations using
+both \animal{} and \ants{} registration algorithms demonstrate good overlap with SNT
+\gs{} segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=2)}
 when using 9 atlases, 19 templates, ANTS registration, and majority vote label
-fusion); Figure \ref{fig:ADNI1-xval-k-mean}).  Qualitatively, both \animal and
-\ants-based segmentations demonstrate trend overlap accuracy that increases with
+fusion); Figure \ref{fig:ADNI1-xval-k-mean}).  Qualitatively, both \animal{} and
+\ants{}-based segmentations demonstrate trend overlap accuracy that increases with
 the size of atlas library and template library. Improvement in accuracy plateaus
@@
 
-The use of \mb with \ants registration shows a pronounced increase
-in segmentation accuracy over \mb with \animal registration, across all
+The use of \mb{} with \ants{} registration shows a pronounced increase
+in segmentation accuracy over \mb{} with \animal{} registration, across all
 other variable settings we tested. Additionally, by itself, using a weighted
@@
 the findings of \citet{Aljabar2009} in basic multi-atlas segmentation. Given
-these findings, in the remainder of this section only results using the \ants
+these findings, in the remainder of this section only results using the \ants{}
 registration algorithm and majority vote fusion will be shown.
 
-With an increasing number of templates, \mb shows improvement over
+With an increasing number of templates, \mb{} shows improvement over
 multi-atlas-based segmentation in overlap accuracy when using the same number of
 atlases and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The two methods
-converge in accuracy when using seven atlases.  Peak improvement in \mb accuracy
+converge in accuracy when using seven atlases.  Peak improvement in \mb{} accuracy
 (~0.02 DSC) is found when one atlas is used with a template library of 19
@@
 In addition to an improvement in accuracy over multi-atlas-based segmentation,
-\mb also shows a decrease in the variability of segmentation accuracy 
+\mb{} also shows a decrease in the variability of segmentation accuracy 
 (Figure \ref{ADNI1-xval-variability}).  The size of template library 
@@
 @
-  \caption{Mean Dice's Similarity Coefficient of \mb segmentations relative to
-  SNT \gs segmentations for 69 ADNI subjects vs. atlas and template library
+  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
+  SNT \gs{} segmentations for 69 ADNI subjects vs. atlas and template library
   size, registration algorithm, and label fusion method. Error bars indicate
@@
 @
-  \caption{Difference in mean Dice's Similarity Coefficient between \mb and
-  multi-atlas segmentations relative to \gs (SNT) segmentations vs. atlas and
-  template library size when using the \ants registration method, and
+  \caption{Difference in mean Dice's Similarity Coefficient between \mb{} and
+  multi-atlas segmentations relative to \gs{} (SNT) segmentations vs. atlas and
+  template library size when using the \ants{} registration method, and
   majority-vote label fusion. Error bars indicate standard error.
@@
 @
-  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
+  \caption{{\bf Difference in variability of \mb{} vs. multi-atlas segmentation 
   accuracy.}
-  Variance of segmentation accuracy between \mb and multi-atlas segmentation is
+  Variance of segmentation accuracy between \mb{} and multi-atlas segmentation is
   computed for each subject across all ten rounds of validation. Shown on the
@@
   comparing the distribution of variances at each parameter setting
-  (atlas/template library size). Only points where \mb mean variability is lower
+  (atlas/template library size). Only points where \mb{} mean variability is lower
   than multi-atlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-This experiment explores \mb segmentation of hippocampal subfields. To achieve
+This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
 this, a leave-one-out validation was conducted in which lower-resolution images
@@
 
-In general, across hippocampal subregions the percent error in volume of \mb
+In general, across hippocampal subregions the percent error in volume of \mb{}
 segmentations (relative to the full-resolution Winterburn atlas segmentation)
@@
 
-Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb subfield
+Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb{} subfield
 segmentations for a single subject.
@@
   atlases after resampling to $0.9mm^3$ voxels.  
-  {\bf MAGeT on Subsampled} volumes are \mb segmentations of the Winterburn atlas
+  {\bf MAGeT on Subsampled} volumes are \mb{} segmentations of the Winterburn atlas
   images after resampling to $0.9mm^3$ voxels.
-  {\bf MAGeT on T1 BRAVO} volumes are \mb segmentations of T1 BRAVO images
+  {\bf MAGeT on T1 BRAVO} volumes are \mb{} segmentations of T1 BRAVO images
   ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
@@
     segmentation resampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
-    corner is the \mb segmentation of the resampled Winterburn atlas images; in the
-    lower right corner is the \mb segmentation of a separately acquired T1 BRAVO
+    corner is the \mb{} segmentation of the resampled Winterburn atlas images; in the
+    lower right corner is the \mb{} segmentation of a separately acquired T1 BRAVO
     image of the same subject. In each segmentation, slices from the left
@@
 %%%%%                                                                     %%%%%
-%%%%%    RESULTS: Experiment 3, Application of \mb to the segmentation    %%%%%
+%%%%%    RESULTS: Experiment 3, Application of \mb{} to the segmentation    %%%%%
 %%%%%                       of schizophrenia patients                     %%%%%
@@
 \subsection{Experiment 3 Results: Application to the segmentation of first episode
-\\ \mbox{schizophrenia patients}}
-
-In this experiment \mb is applied to a dataset of images of first episode
+\\ \mb{}ox{schizophrenia patients}}
+
+In this experiment \mb{} is applied to a dataset of images of first episode
 schizophrenia patients, using the Winterburn atlases and a template library of
@@
 \begin{figure}
-<<FEP-volumes,cache=TRUE,dependson="setup">>=
+<<FEP-volumes,cache=FALSE,dependson="setup",width=5,height=5>>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
@@
 @
-  \caption{Mean bilateral hippocampus volume as measured by \mb  vs. manually
+  \caption{Mean bilateral hippocampus volume as measured by \mb{}  vs. manually
   segmented volumes from the First Episode Patients with Schizophrenia dataset.
-  A linear fit line is shown, with shaded region showing standard error. }
-  \label{fig:SZ-volumes}
+  A linear fit line is shown, with shaded region showing standard error. 
+  \label{fig:SZ-volumes}}
 \end{figure}
@@
 @
-\mb produces hippocampal volumes that are highly correlated with 
+\mb{} produces hippocampal volumes that are highly correlated with 
 manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
@@
 
-Based on the results from Experiment 1, in this experiment \mb was configured
+Based on the results from Experiment 1, in this experiment \mb{} was configured
 with a template library of 21 randomly chosen subject images (7 from each
-disease class) and majority vote label fusion.  The entire \adnidataset dataset
-was segmented by \mb, and the resulting volumes compared with those obtained by
+disease class) and majority vote label fusion.  The entire \adnidataset{} dataset
+was segmented by \mb{}, and the resulting volumes compared with those obtained by
 expert semi-automated segmentation (SNT), and three other automated segmentation
-techniques: \maper, \freesurfer, and \fsl. Table
+techniques: \maper{}, \freesurfer{}, and \fsl{}. Table
 \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
@@
 We found a close relationship in total bilateral hippocampal volume between all
-methods and SNT \gs volumes (Figure \ref{fig:ADNI-volumes-plot}).  Volumes are
+methods and SNT \gs{} volumes (Figure \ref{fig:ADNI-volumes-plot}).  Volumes are
 correlated with Pearson $r > 0.78$ for all methods across disease categories.
-Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}), \mb is
+Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}), \mb{} is
 consistently well correlated to manual volumes (Pearson $r > 0.85$), but appears
@@
 
-To investigate the level of agreement with SNT \gs hippocampal
+To investigate the level of agreement with SNT \gs{} hippocampal
 volumes, we constructed Bland-Altman plots for each method (Figure
@@
 correlation can be driven by a large range in true values, for instance).  All
-methods show an obvious proportional bias: \freesurfer and \fsl markedly
+methods show an obvious proportional bias: \freesurfer{} and \fsl{} markedly
 under-estimate smaller hippocampi and over-estimate large hippocampi, whereas
-\maper and \mb show a reverse, conservative bias (Figure
+\maper{} and \mb{} show a reverse, conservative bias (Figure
 \ref{fig:ADNI-Bland-Altman}).  Additionally, all methods show a fixed volume
-bias, with \freesurfer and \fsl most dramatically over-estimating hippocampal
-volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and \maper and
-\mb within $250 mm^3$ on average.
-
-Figure \ref{fig:ADNI-segmentations} shows a qualitative comparison of \mb and 
-SNT \gs hippocampal segmentations for 10 randomly selected subjects in 
+bias, with \freesurfer{} and \fsl{} most dramatically over-estimating hippocampal
+volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and \maper{} and
+\mb{} within $250 mm^3$ on average.
+
+Figure \ref{fig:ADNI-segmentations} shows a qualitative comparison of \mb{} and 
+SNT \gs{} hippocampal segmentations for 10 randomly selected subjects in 
 each disease category, and illustrates some of the common errors found during 
-visual inspection. Mostly frequently, we found \mb improperly includes the vestigial
-hippocampal sulcus and, although not anatomically incorrect, \mb under-estimates 
-the hippocampal body in comparison to the SNT \gs segmentation.
+visual inspection. Mostly frequently, we found \mb{} improperly includes the vestigial
+hippocampal sulcus and, although not anatomically incorrect, \mb{} under-estimates 
+the hippocampal body in comparison to the SNT \gs{} segmentation.
 
@@
     caption="\\textbf{Number of segmented images and quality control failures of 
-             \\adnidataset dataset by method.}", 
+             \\adnidataset{} dataset by method.}", 
     caption.loc = c("top"), rowname=NULL, 
@@
                       "MAGeT" = "MAGeT", 
-                      "SNT"   = "SNT \\gs")
+                      "SNT"   = "SNT \\gs{}")
 m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
@@
 @
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by \freesurfer, \fsl, \maper, \mb, and SNT \gs vs. disease
+  \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
+dataset by \freesurfer{}, \fsl{}, \maper{}, \mb{}, and SNT \gs{} vs. disease
 category.}
@@
     geom_smooth(method="lm") + 
-    xlab(expression("SNT \\gs mean hippocampus volume " (mm^3))) + 
+    xlab(expression("SNT \\gs{} mean hippocampus volume " (mm^3))) + 
     ylab(expression("Automated mean hippocampal volume " (mm^3))) +
@@
 @
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset
-dataset by each of the four automated methods investigated (\freesurfer (FS), \fsl,
-\maper, \mb) vs. SNT \gs. Linear fit lines and Pearson correlations are shown
+  \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
+dataset by each of the four automated methods investigated (\freesurfer{} (FS), \fsl{},
+\maper{}, \mb{}) vs. SNT \gs{}. Linear fit lines and Pearson correlations are shown
 for each method.}
@@
             hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of SNT \\gs and automated hippocampal volume ' (mm^3))) + 
+  xlab(expression('Mean of SNT \\gs{} and automated hippocampal volume ' (mm^3))) + 
   ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
@@
   \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset dataset by SNT \gs segmentation and each of the
-four automated methods investigated (\freesurfer, \fsl, \maper, \mb). The overall
+measured in the \adnidataset{} dataset by SNT \gs{} segmentation and each of the
+four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
 mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
@@
 Note, points below the mean difference indicate overestimation of the volume
-with respect to the \gs rating, and vice versa. }
+with respect to the \gs{} rating, and vice versa. }
 
@@
   \end{centering}
-  \caption{SNT \gs and \mb segmentations for 30 ADNI subjects (10
+  \caption{SNT \gs{} and \mb{} segmentations for 30 ADNI subjects (10
   subjects randomly selected from each disease category in the subject pool used
   in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
-  anatomical image. SNT \gs labels appear in green, and \mb labels appear in
+  anatomical image. SNT \gs{} labels appear in green, and \mb{} labels appear in
   blue. Noted are examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
-  {\em (b)} translated SNT \gs segmentation; 
+  {\em (b)} translated SNT \gs{} segmentation; 
   {\em (c)} under-estimation of hippocampal body and
-  {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb.}
+  {\em (d)} improper inclusion of the vestigial hippocampal sulcus by \mb{}.}
   \label{fig:ADNI-segmentations}
@@
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy (overlap with SNT \gs labels)
+\caption{\textbf{Automated segmentation accuracy (overlap with SNT \gs{} labels)
 of the ADNI dataset.} For each method, the number of labelled atlases
@@
 \hline
-\mb                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
+\mb{}                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
 LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
@@
 In this manuscript we have presented the implementation and validation of the
-\mb framework -- a methodology that requires very few input atlases in
+\mb{} framework -- a methodology that requires very few input atlases in
 order to provide accurate and reliable segmentations.  Experiment 1 establishes
-\mb as effective and as comparable to multi-atlas segmentation. This experiment
-also rigorously characterizes the behaviour of \mb whole hippocampal
+\mb{} as effective and as comparable to multi-atlas segmentation. This experiment
+also rigorously characterizes the behaviour of \mb{} whole hippocampal
 segmentation under various parameter settings, allowing us to choose an optimal
 setting for subsequent experiments. Experiment 2 demonstrates the accuracy of
-\mb hippocampal subfield identification despite contrast and resolution
+\mb{} hippocampal subfield identification despite contrast and resolution
 limitations in standard T1-weighted image volumes.  Experiments 2 and 3 validate
-\mb whole hippocampal segmentation accuracy and consistency on populations with
+\mb{} whole hippocampal segmentation accuracy and consistency on populations with
 different ageing and neuropsychiatric characteristics.  Furthermore, taken
@@
 Throughout the cross validation in Experiment 1 (10-fold Monte Carlo cross
-validation in the \adnidataset dataset subsample) we find that two parameter
+validation in the \adnidataset{} dataset subsample) we find that two parameter
 choices improve segmentation accuracy: increasing the number of atlases, and the
@@
 neighbour image resampling, and neuroanatomical mismatch between atlases and
-subjects. That \mb does not see the same boost in performance from weighted
+subjects. That \mb{} does not see the same boost in performance from weighted
 voting may suggest that the neuroanatomical variability of a template library
 constructed from study subjects more closely matches any particular subject and
-thereby leaving less error to filter. From our previous work on the \mb
+thereby leaving less error to filter. From our previous work on the \mb{}
 algorithm we have shown that the reduction in error is not simply a smoothing or
@@
 even number.  Our initial concern was that this dip in performance was a
-by-product of the \mb algorithm itself.  However, this this pattern is also
+by-product of the \mb{} algorithm itself.  However, this this pattern is also
 found in the results of the multi-atlas segmentations we used in our
@@
 
-Another concern is the moderate-to-small improvement observed in \mb in
+Another concern is the moderate-to-small improvement observed in \mb{} in
 comparison to multi-atlas segmentation when using the same number of atlases.
-The actual benefit in using \mb is consistency of the labelling regardless of
+The actual benefit in using \mb{} is consistency of the labelling regardless of
 atlas or template choice. This is an important consideration that few have
@@
 
-In comparison to other methodologies in the field \mb performs favourably. Table
+In comparison to other methodologies in the field \mb{} performs favourably. Table
 \ref{tab:other-methods} surveys some of the most recent reported DSC values
@@
 worth noting that the methods summarized require more atlases (between 16-55)
-than our \mb implementation with the Winterburn atlases \citep{Winterburn2013}.
+than our \mb{} implementation with the Winterburn atlases \citep{Winterburn2013}.
 
@@
 Among the automated segmentation methods we compared in this paper
-(\freesurfer, \maper, \fsl-FIRST), we find extremely variable
-performance of all methods.  With the exception of \fsl all methods correlate
-well with the SNT volumes provided in the ADNI database.  However, \freesurfer
-and \fsl provide radically different definitions of the size of the hippocampus
+(\freesurfer{}, \maper{}, \fsl{}-FIRST), we find extremely variable
+performance of all methods.  With the exception of \fsl{} all methods correlate
+well with the SNT volumes provided in the ADNI database.  However, \freesurfer{}
+and \fsl{} provide radically different definitions of the size of the hippocampus
 in comparison to the other methods.  Further, when estimating bias of these
-methods relative to SNT \gs hippocampal volumes we see that large hippocampi are
-over estimated while small hippocampi are under estimated. By comparison, \mb
-and \maper are far more conservative in volume estimation, suggesting these
+methods relative to SNT \gs{} hippocampal volumes we see that large hippocampi are
+over estimated while small hippocampi are under estimated. By comparison, \mb{}
+and \maper{} are far more conservative in volume estimation, suggesting these
 methods may be better suited for estimating true-positives, especially in

### 2013-09-03 letters/neuroimage.tex
Crap draft of the neuroimage letter

\documentclass{letter}
\usepackage{hyperref}
\signature{Jon Pipitone \\
Kimel Family Translational Imaging-Genetics Laboratory \\
Research Imaging Centre \\
Centre for Addiction and Mental Health \\
Toronto, Ontario}
\address{}
\begin{document}

\begin{letter}{Editorial Board, Neuroimage}
\opening{Dear Sir or Madam:}

Please find enclosed our manuscript entitled, {\em Bootstrapping Multi-atlas
Segmentation Using Multiple Automatically Generated Templates for the
Segmentation of the Whole Hippocampus and Subfields} by Jon Pipitone and
colleagues.

In this manuscript we describe a novel automated MRI hippocampal segmentation
algorithm optimised to perform well with only a small number of manually
segmented training images. We believe this is an important contribution since
the expertise and effort needed to perform manual segmentations can be
prohibitive for many clinicians and researchers, and yet when one-of-a-kind or
modifications to existing segmentations need to be made, existing automated
methods require 30-80 training inputs.

Our automated segmentation algorithm, mischievously named MAGeT-Brain, takes
advantage of the neuroanatomical variability that exists in a population that
is being studied in order to automatically identify the hippocampus and its
subfields. Our results demonstrate consistent accuracy in the identification of
the whole hippocampus when compared to "gold-standard" manual segmentations.
Finally, we have made the our algorithm available publically online for use by
other groups, and are pursuing the the contribution of our segmentations to the
Alzheimer's Disease Neuroimaging Initiative image database.

We believe that the techniques we have developed and our findings are a
significant contribution to the neuroimaging community, specifically for those
researchers interested in large scale studies of the hippocampus in the context
of normal brain function and different forms of brain dysfunction.

We hope you find the enclosed manuscript meets the high standards of
NeuroImage.

\closing{Sincerely,}

%\ps{P.S. You can find the full text of GFDL license at}
\encl{Manuscript}

\end{letter}
\end{document}

### 2013-09-04 letters/neuroimage.tex
Finished letter, minor fixes to the paper

@@
-\documentclass{letter}
+\documentclass[12pt]{letter}
 \usepackage{hyperref}
+
+% page counting, header/footer
+\usepackage{fancyhdr}
+\usepackage{lastpage}
+\usepackage{graphicx}
+
+\pagestyle{fancy}
+\fancyhead[RO,RE]{\includegraphics[width=3cm]{letters/camh_logo.png}}
+%\renewcommand{\headheight}{24pt}
+%\renewcommand{\footrulewidth}{0.4pt}
+
 \signature{Jon Pipitone \\
@@
 \opening{Dear Sir or Madam:}
+\thispagestyle{fancy}
 
@@
 algorithm optimised to perform well with only a small number of manually
-segmented training images. We believe this is an important contribution since
-the expertise and effort needed to perform manual segmentations can be
-prohibitive for many clinicians and researchers, and yet when one-of-a-kind or
-modifications to existing segmentations need to be made, existing automated
-methods require 30-80 training inputs.
+segmented training images. We believe this is an important contribution because
+the expertise and effort needed to perform manual segmentation can be
+prohibitive for many clinicians and researchers, and yet existing automated
+methods generally require between 30 to 80 training segmentations. This
+situation makes it infeasible to use these methods for segmentations based on
+histological-based digital segmentations (because of their rarity),
+high-resolution digital atlases (because of the time needed to segment these
+images), or when exploring the effect of variations on a segmentation protocol.
 
-Our automated segmentation algorithm, mischievously named MAGeT-Brain, takes
-advantage of the neuroanatomical variability that exists in a population that
-is being studied in order to automatically identify the hippocampus and its
-subfields. Our results demonstrate consistent accuracy in the identification of
-the whole hippocampus when compared to "gold-standard" manual segmentations.
-Finally, we have made the our algorithm available publically online for use by
-other groups, and are pursuing the the contribution of our segmentations to the
-Alzheimer's Disease Neuroimaging Initiative image database.
+It is for these reasons that we developed the automated segmentation algorithm,
+mischievously named MAGeT-Brain, which takes advantage of the neuroanatomical
+variability that exists in the target population being studied to bootstrap a
+large template library from a small set of manually segmented images. In this
+manuscript we rigorously validate this approach on multiple disease
+populations, and compare our segmentations with existing popular methods (e.g.
+FSL and FreeSurfer).  Our results demonstrate consistent accuracy in the
+identification of the whole hippocampus when compared to "gold-standard" manual
+segmentations.  Finally, we have made our algorithm available publically online
+for use by other groups, and are pursuing the contribution of our segmentations
+to the Alzheimer's Disease Neuroimaging Initiative image database.
 
-We believe that the techniques we have developed and our findings are a
+We believe that the technique we have developed and our findings are a
 significant contribution to the neuroimaging community, specifically for those
@@
 
-%\ps{P.S. You can find the full text of GFDL license at}
 \encl{Manuscript}

### 2013-09-04 paper.Rnw
Finished letter, minor fixes to the paper

@@
 \author[1]{Jon Pipitone}
-\author[1]{Matt T. Park}
+\author[1]{Min Tae Park}
 \author[1]{Julie Winterburn}
@@
 \author[2,3]{Jason P. Lerch}
-\author[4,5]{Jens C. Pruessner}
-\author[4]{Martin Lepage}
+\author[4]{Jens C. Pruessner}
+\author[4,5]{Martin Lepage}
 \author[1,6,9]{Aristotle N. Voineskos}
-\author[1,6,8]{M. Mallar Chakravarty}
+\author[1,6,7,8]{M. Mallar Chakravarty}
 \author[ ]{the Alzheimer's Disease Neuroimaging Initiative\footnote{Data
@@
 The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
-SNT and \freesurfer{} hippocampal volumes for a subset of images were provided by
-ADNI, along with quality control data for each \freesurfer{} segmentation
-(guidelines described in \citep{Hartig2010}).  Clinical and demographic data for
-the entire \adnidataset{} dataset are shown in Table
+SNT, \maper, \freesurfer{} hippocampal volumes for a subset of images were
+provided by ADNI, along with quality control data for each \freesurfer{}
+segmentation (guidelines described in \citep{Hartig2010}).  Clinical and
+demographic data for the entire \adnidataset{} dataset are shown in Table
 \ref{tab:ADNI1-Dataset-Demographics}.
@@
 registration method, and label fusion method were determined by the optimal
-settings found in Experiment 1.  The \freesurfer{} segmentation method was used via
+settings found in Experiment 1.  The \fsl{} segmentation method was used via
 the \verb+run_first_all+ script according to the FIRST user guide
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 3 Results: Application to the segmentation of first episode
-\\ \mb{}ox{schizophrenia patients}}
+\subsection{Experiment 3 Results: Application to the segmentation of first
+episode schizophrenia patients}
 
@@
 
+\section{Acknowledgements}
+We wish acknowledge support from the CAMH Foundation, thanks to Michael and
+Sonja Koerner, the Kimel Family, and the Paul E. Garfinkel New Investigator
+Catalyst Award. MMC is funded by the W. Garfield Weston Foundation and ANV is
+funded by the Canadian Institutes of Health Research, Ontario Mental Health
+Foundation, NARSAD, and the National Institute of Mental Health (R01MH099167).
+
+Computations were performed on the gpc supercomputer at the SciNet HPC
+Consortium. SciNet is funded by: the Canada Foundation for Innovation under the
+auspices of Compute Canada; the Government of Ontario; Ontario Research Fund -
+Research Excellence; and the University of Toronto
+
+ADNI Acknowledgements: Data collection and sharing for this project was funded
+by the Alzheimer's Disease Neuroimaging Initiative (ADNI) (National Institutes
+of Health Grant U01 AG024904). ADNI is funded by the National Institute on
+Aging, the National Institute of Biomedical Imaging and Bioengineering, and
+through generous contributions from the following: Abbott; Alzheimer's
+Association; Alzheimer's Drug Discovery Foundation; Amorfix Life Sciences Ltd.;
+AstraZeneca; Bayer HealthCare; BioClinica, Inc.; Biogen Idec Inc.;
+Bristol-Myers Squibb Company; Eisai Inc.; Elan Pharmaceuticals Inc.; Eli Lilly
+and Company; F. Hoffmann-La Roche Ltd and its affiliated company Genentech,
+Inc.; GE Healthcare; Innogenetics, N.V.; IXICO Ltd.; Janssen Alzheimer
+Immunotherapy Research  Development, LLC.; Johnson \& Johnson Pharmaceutical
+Research  Development LLC.; Medpace, Inc.; Merck \& Co., Inc.; Meso Scale
+Diagnostics, LLC.; Novartis Pharmaceuticals Corporation; Pfizer Inc.; Servier;
+Synarc Inc.; and Takeda Pharmaceutical Company. The Canadian Institutes of
+Health Research is providing funds to support ADNI clinical sites in Canada.
+Private sector contributions are facilitated by the Foundation for the National
+Institutes of Health (www.fnih.org). The grantee organization is the Northern
+California Institute for Research and Education, and the study is Rev March 26,
+2012 coordinated by the Alzheimer's disease Cooperative Study at the University
+of California, San Diego. ADNI data are disseminated by the Laboratory for
+NeuroImaging at the University of California, Los Angeles. This research was
+also supported by NIH grants P30 AG010129 and K01 AG030514.
+
 \section{Supplementary Materials}

### 2013-12-05 paper.Rnw
Acknowledgements to F-delic, and neuromimage submission cool stuff
Added a header for Draft mode
First pass at adding JENS-XVAL into the mix

@@
 \setlength{\marginparwidth}{1.2in}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhead[RO,RE]{draft - f9776d7 - September, 2013}
 
@@
 \author[1]{Jon Pipitone}
-\author[1]{Min Tae Park}
+\author[1]{Matt Min Tae Park}
 \author[1]{Julie Winterburn}
@@
 <<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
+jens_RIDs = read.csv('data/jens_RID.csv')
+
 adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
@@
                  1217,29,56,1188,1293,149,382,431,1202)
+
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+               jens_RIDs$RID, method = "reverse", test=FALSE, overall=TRUE)
 caption = 
@@
 <<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-all_data  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data.csv.gz'))
-all_data_mean  <- read.csv(gzfile('data/cache/ADNI-XVAL:all_data_mean.csv.gz'))
-
-mb.best.kappa = mean(subset(all_data_mean, reg_method   == "ANTS" & 
+jens_xval_data  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
+jens_xval_mean  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
+
+mb.best.kappa = mean(subset(jens_xval_mean, reg_method   == "ANTS" & 
                                   method.mb    == "Majority Vote" &
@@
 stderr <- function(x) sqrt(var(x)/length(x))
-ggplot(subset(all_data_mean, (templates.mb*atlases) %% 2 == 1),
+ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1),
        aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
@@
   #stat_smooth(method='lm', formula=y~log(x)) + 
+  scale_y_continuous(breaks=seq(0,1,0.05)) + 
   scale_colour_hue(name="Number of Atlases") +
@@
   ylab( "Mean DSC" ) + 
-  opts(legend.direction = "horizontal", legend.position = "bottom", 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
         axis.text = element_text(size = 8))
@@
 <<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-prep'>>=
-ggplot(subset(all_data_mean, reg_method=="ANTS" & 
+ggplot(subset(jens_xval_mean, reg_method=="ANTS" & 
                 method.mb=="Majority Vote" & (templates.mb*atlases) %% 2 == 1), 
@@
   ylab( "Difference in mean DSC" ) +
-  opts(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
 <<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-ants = subset(all_data, 
+ants = subset(jens_xval_data, 
                      reg_method=="ANTS" & 
@@
 auspices of Compute Canada; the Government of Ontario; Ontario Research Fund -
-Research Excellence; and the University of Toronto
+Research Excellence; and the University of Toronto.
+
+In addition, computations were performed on the CAMH Specialized Computing 
+Cluster. The SCC is funded by: The Canada Foundation for Innovation, Research 
+Hospital Fund.
 
@@
 
+We would also like to thank G. Clinton, E. Hazel, and B. Worrell for inspiring
+this work.
+
 \section{Supplementary Materials}
@@
 
+\subsection{Experiment 1a: Whole Hippocampus Cross-Validation with SNT 
+\gs Segmentations }
+
+Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
+of images and SNT \gs{} segmentations from the ADNI dataset \citep{Hsu2002}. This
+form of cross-validation allows us to rigorously validate a large number of
+parameter settings of \mb{} (atlas and template library sizes, registration
+algorithm, and label fusion method) and select the best parameters to use in
+subsequent experiments.
+
+\subsubsection{Methods}
+\paragraph{\adnidataset{} dataset} 
+69 1.5T images were arbitrarily selected from the baseline scans in the {\em
+\adnidataset{}} standardized dataset. 23 subjects were chosen from each disease
+category: cognitively normal (CN), mild cognitive impairment (MCI) and
+Alzheimer's disease (AD). Demographics for this subset are shown in Table
+\ref{tab:ADNI1-xvalidation-demographics}.  Semi-automated segmentations of the
+left and right whole hippocampi are made available with a subset of ADNI images
+\citep{Hsu2002}. These labels have been generated using the SNT tool from
+Medtronic Surgical Navigation Technologies, Louisville, CO (see Supplementary
+Materials for detailed discussion of the segmentation process used). Throughout
+this paper, we use these segmentations as the \gs{} for comparison.
+
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
+available from ADNI \footnote{
+\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly pre-processed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
+2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
+240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% ADNI1 X-Validation Dataset Demographics Table
+%
+<<ADNI-SNT-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
+adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
+                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
+                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
+                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
+                 1217,29,56,1188,1293,149,382,431,1202)
+
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
+               MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
+               adni_69_RIDs, method = "reverse", test=FALSE, overall=TRUE)
+caption = 
+  "\\textbf{ADNI1 SNT cross-validation subset demographics.} 
+  CN     - Cognitively Normal. 
+  LMCI   - Late-onset Mild Cognitive Impairment. 
+  AD     - Alzheimer's Disease.  
+  Hisp   - Hispanic. 
+  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
+  ADAS   - Alzheimer's Disease Assessment Scale.  
+  MMSE   - Mini-Mental State Examination."
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      label="tab:ADNI1-SNT-xvalidation-demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Experiment details}
+
+Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
+cross-validation, consists of repeated rounds of validation conducted on a fixed
+dataset \citep{Shao1993}. In each round, the dataset is randomly partitioned into
+a training set and a validation set. The method to be validated is then given
+the training data, and its output is compared with the validation set.
+
+In this experiment, our dataset consists of 69 1.5T images and corresponding \gs{}
+segmentations. In each validation round, the dataset is partitioned into a
+training set consisting of images and \gs{} segmentations used as an atlas
+library, and a validation set consisting of the remaining images segmented by
+both \mb{} and multi-atlas. The computed segmentations were compared to the \gs{}
+segmentations (see below).
+
+A total of ten validation rounds were performed on each subject in the dataset,
+over each combination of parameter settings. The parameter settings
+explored are: atlas library size (1-9), template library size (1-20),
+registration method (\ants{} or \animal{}, described below), and label fusion method (majority vote,
+cross-correlation weighted majority vote, and normalized mutual information
+weighted majority vote, described below).  A total of $10 \times 69 \times 9
+\times 20 \times 2 \times 3 = \num{7452000}$ validation rounds were conducted,
+resulting in a total of $\num{1490400}$ segmentations analysed.
+
+Before registration, all images underwent preprocessing with the N3 algorithm
+\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we
+compared two non-linear image registration methods, three label fusion methods.
+
+The automatically produces labels are compared to the \gs{} SNT labels provided by
+ADNI. The segmentation accuracy reported for a parameter setting is the average
+over ten validation rounds.
+
+% ######################
+\subsubsection{Results}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: prep
+% Prepares the data in a the form needed for plotting.
+<<ADNI1-SNT-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
+snt_xval_data  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
+snt_xval_mean  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data_mean.csv.gz'))
+
+snt_xval_mb.best = mean(subset(snt_xval_data, reg_method   == "ANTS" & 
+                                  method.mb    == "Majority Vote" &
+                                  templates.mb == 19 & 
+                                  atlases      == 9)$k.mb)
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this experiment we conducted 10 rounds of \mb{} and multi-atlas segmentation of
+of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
+images respectively, two registration algorithms (\ants{} or \animal{}), and three
+label fusion techniques (unweighted, cross-correlation, and normalised mutual
+information weighted majority vote).  Hippocampal \mb{}-based segmentations using
+both \animal{} and \ants{} registration algorithms demonstrate good overlap with SNT
+\gs{} segmentations (maximum mean DSC of \Sexpr{round(snt_xval_mb.best, digits=2)}
+when using 9 atlases, 19 templates, ANTS registration, and majority vote label
+fusion); Figure \ref{fig:ADNI1-xval-k-mean}).  Qualitatively, both \animal{} and
+\ants{}-based segmentations demonstrate trend overlap accuracy that increases with
+the size of atlas library and template library. Improvement in accuracy plateaus
+with template libraries larger than ten images.
+
+The use of \mb{} with \ants{} registration shows a pronounced increase
+in segmentation accuracy over \mb{} with \animal{} registration, across all
+other variable settings we tested. Additionally, by itself, using a weighted
+voting strategy did not significantly improve segmentation accuracy, contrary to
+the findings of \citet{Aljabar2009} in basic multi-atlas segmentation. Given
+these findings, in the remainder of this section only results using the \ants{}
+registration algorithm and majority vote fusion will be shown.
+
+With an increasing number of templates, \mb{} shows improvement over
+multi-atlas-based segmentation in overlap accuracy when using the same number of
+atlases and voting method (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). The two methods
+converge in accuracy when using seven atlases.  Peak improvement in \mb{} accuracy
+(~0.02 DSC) is found when one atlas is used with a template library of 19
+images.
+
+In addition to an improvement in accuracy over multi-atlas-based segmentation,
+\mb{} also shows a decrease in the variability of segmentation accuracy 
+(Figure \ref{ADNI1-SNT-xval-variability}).  The size of template library 
+necessary to reach a decrease ($p < 0.05$) in variance and standard deviation
+grows with the size of atlas library used.  A template library of 19 images is
+sufficient to show significant decrease in variance and standard deviation for
+3-7 atlases.
+
+We have omitted results obtained when using an even number of atlases or
+templates since with this configuration we found significantly decreased
+performance. We believe this is as a result of an inherent bias in the majority
+vote fusion method used (see Discussion).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
+\begin{figure}
+<<fig:ADNI1-SNT-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-prep', fig.width=5, fig.height=5>>=
+stderr <- function(x) sqrt(var(x)/length(x))
+ggplot(subset(snt_xval_mean, (templates.mb*atlases) %% 2 == 1),
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
+  facet_grid(reg_method~method.mb) + 
+  stat_summary(fun.y=mean,geom='line',
+               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+  stat_summary(fun.y=mean,
+               fun.ymin=function (x) mean(x) - stderr(x),
+               fun.ymax=function (x) mean(x) + stderr(x), 
+               geom='errorbar',
+               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
+  #stat_smooth(method='lm', formula=y~log(x)) + 
+  scale_y_continuous(breaks=seq(0,1,0.05)) + 
+  scale_colour_hue(name="Number of Atlases") +
+  xlab( "Number of Templates" ) + 
+  ylab( "Mean DSC" ) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text = element_text(size = 8))
+@
+  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
+  SNT \gs{} segmentations for 69 ADNI subjects vs. atlas and template library
+  size, registration algorithm, and label fusion method. Error bars indicate
+  standard error.
+  \label{fig:ADNI1-SNT-xval-k-mean}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: "increase" in mean kappa over multi-atlas
+\begin{figure}
+<<fig:ADN11-SNT-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-prep', fig.width=5, fig.height=5>>=
+ggplot(subset(snt_xval_mean, reg_method=="ANTS" & 
+                method.mb=="Majority Vote" & (templates.mb*atlases) %% 2 == 1), 
+       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
+  #facet_grid(reg_method~method.y) + 
+  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
+               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
+  stat_summary(fun.y=mean, geom="point", 
+               aes(y=k_diff,colour=as.factor(atlases))) + 
+  stat_summary(fun.y=mean, geom="line", 
+               aes(y=k_diff,colour=as.factor(atlases))) + 
+  #stat_smooth(method=lm, formula=y~log(x)) + 
+  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
+  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
+  scale_colour_hue(name="Number of Atlases") +
+  scale_alpha_continuous(guide = "none") + 
+  scale_linetype_discrete(guide = "none") + 
+  xlab( "Number of Templates" ) + 
+  ylab( "Difference in mean DSC" ) +
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{Difference in mean Dice's Similarity Coefficient between \mb{} and
+  multi-atlas segmentations relative to \gs{} (SNT) segmentations vs. atlas and
+  template library size when using the \ants{} registration method, and
+  majority-vote label fusion. Error bars indicate standard error.
+  \label{fig:ADNI1-SNT-xval-k-diff}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-SNT-XVAL: show variability of MAGeT over parameters
+\begin{figure}
+<<ADN1-SNT-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=5, fig.width=5>>=
+ants = subset(snt_xval_data, 
+                     reg_method=="ANTS" & 
+                     method.mb=="Majority Vote" &
+                    templates.mb %in% seq(1,20,2))
+
+ants.stats = ddply(ants, c("subject", "label", "atlases","templates.mb"), 
+                   function (df) {
+                     data.frame(
+                       MA.sd  =  sd(df$k.ma),
+                       MB.sd  =  sd(df$k.mb),
+                       MA.var = var(df$k.ma),
+                       MB.var = var(df$k.mb)
+                       )
+                   })
+
+stats.tests = ddply(ants.stats, c("atlases","templates.mb"), function (df) {
+  t.var = t.test(df$MA.var, df$MB.var)
+  t.sd  = t.test(df$MA.sd , df$MB.sd)
+  diff_mean_var = mean(df$MA.var - df$MB.var)
+  diff_mean_sd  = mean(df$MA.sd -  df$MB.sd)
+
+  # test statistic is positive if MA mean is larger than MB
+  # i.e. that MA has a greater mean of variances/SDs
+  # i.e. that MB has a smaller mean, and so is doing "better"
+  data.frame(
+    stat      = c("Variance", "Standard Deviation"),
+    statistic = c(t.var$statistic, t.sd$statistic),
+    p.value   = c(t.var$p.value, t.sd$p.value),
+    direction = c(diff_mean_var, diff_mean_sd)
+  )
+})
+
+ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
+  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
+  #facet_grid( . ~ stat) + 
+  geom_line(size=1) +
+  geom_point(size=3) + 
+  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
+  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
+  scale_x_continuous(breaks=seq(1,20,by=2)) + 
+  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
+  xlab( "Number of Templates" ) + 
+  ylab( "Difference of variance during validation (p-value)") +
+  scale_colour_hue(name="Atlases") +
+  scale_alpha_continuous(guide='none') + 
+  scale_linetype_discrete(guide='none') + 
+  scale_size_continuous(guide='none') + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Difference in variability of \mb{} vs. multi-atlas segmentation 
+  accuracy using SNT \gs.}
+  Variance of segmentation accuracy between \mb{} and multi-atlas segmentation is
+  computed for each subject across all ten rounds of validation. Shown on the
+  y-axis (scaled logarithmically) is the  p-value resulting from a t-test
+  comparing the distribution of variances at each parameter setting
+  (atlas/template library size). Only points where \mb{} mean variability is lower
+  than multi-atlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
+  \label{ADNI1-SNT-xval-variability}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{abbrvnat}

### 2013-12-05 cache_analysis.R

@@ -7,69 +7,81 @@ library(plyr)
 # ADNI1-XVAL: prep
 # Prepares the data in a the form needed for plotting.
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-# all data from the validation trials
-a_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
-t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz'))
-diagnoses    = read.csv('data/a2a_diagnoses.csv')
+prep <- function(ants_data_file, tracc_data_file) {
+  # all data from the validation trials
+  a_data       = read.csv(gzfile(ants_data_file))
+  t_data       = read.csv(gzfile(tracc_data_file))
+  diagnoses    = read.csv('data/a2a_diagnoses.csv')
+  
+  # simplify the data
+  t_data       = subset(t_data, atlases %% 2 == 1)
+  a_data       = subset(a_data, atlases %% 2 == 1)
+  t_data$se    = NULL
+  t_data$sn    = NULL
+  t_data$j     = NULL
+  a_data$se    = NULL
+  a_data$sn    = NULL
+  a_data$j     = NULL       
+  
+  # add registration method column 
+  t_data       = cbind(t_data, reg_method = "ANIMAL")
+  a_data       = cbind(a_data, reg_method = "ANTS")
+  a_data$volume=0
+  
+  # separate MA majvote from MB so we can compute differences
+  # we only compare multiatlas using majority vote
+  t_ma = subset(t_data, approach %in% c("ma", "multiatlas") & method=="majvote")
+  a_ma = subset(a_data, approach %in% c("ma", "multiatlas") & method=="majvote")
+  t_mb = subset(t_data, approach=="mb")
+  a_mb = subset(a_data, approach=="mb")
+  t_ma$approach = "ma"
+  a_ma$approach = "ma"
+  ma = rbind(t_ma,a_ma)
+  mb = rbind(t_mb,a_mb)
+  
+  # set up a few equivalences for multi-atlas, to make nomenclature simpler
+  ma$templates =  ma$atlases  # because no template library
+  mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n 
+  
+  # aggregate over batches
+  ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases +
[…]

### 2013-12-11 paper.Rnw
Updated results using Pruessner segmentations, partial re-write

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: prep
+% Prepares the data in a the form needed for plotting.
+<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
+jens_atlases = read.csv('data/jens_atlases.csv')
+jens_xval_data  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
+jens_xval_mean  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
+jens_xval_mean$reg_method = factor(jens_xval_mean$reg_method, c("ANTS", "ANIMAL"))
+jens_xval_mean$method.mb = factor(jens_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
+mb.best.kappa = mean(subset(jens_xval_mean, reg_method   == "ANTS" & 
+                                  method.mb    == "Majority Vote" &
+                                  templates.mb == 19 & 
+                                  atlases      == 9)$k.mb)
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-COMPLETE
+% Prepares the data in the form needed for plotting.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+<<ADNI-volumes-prep,echo=F,cache=T>>=
+means.complete   <- read.csv('data/cache/ADNI1:qc.csv')
+package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
+@
 
@@
 <<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
-jens_RIDs = read.csv('data/jens_RID.csv')
-
 adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
@@
                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-               jens_RIDs$RID, method = "reverse", test=FALSE, overall=TRUE)
+               jens_atlasess$RID, method = "reverse", test=FALSE, overall=TRUE)
 caption = 
@@
 Bland-Altman plots \citep{Bland1986}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-COMPLETE
-% Prepares the data in the form needed for plotting.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<ADNI-volumes-prep,echo=F,cache=T>>=
-means.complete   <- read.csv('data/cache/ADNI1:qc.csv')
-package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
-@
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: prep
-% Prepares the data in a the form needed for plotting.
-<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
-jens_xval_data  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
-jens_xval_mean  <- read.csv(gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
-
-mb.best.kappa = mean(subset(jens_xval_mean, reg_method   == "ANTS" & 
-                                  method.mb    == "Majority Vote" &
-                                  templates.mb == 19 & 
-                                  atlases      == 9)$k.mb)
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 In this experiment we conducted 10 rounds of \mb{} and multi-atlas segmentation of
@@
 
-% \begin{figure}
-% <<ADN1-library-composition, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
-% # These composition datasheets describe the subjects and subject DX in the 
-% # template and atlas libraries for each voting run
-% atlas_comp = read.csv('data/a2a_ants/atlas_library_composition.csv')
-% templ_comp = read.csv('data/a2a_ants/template_library_composition.csv')
-% a2a.best = subset(all_data, atlases == 9 &
-%                     templates.mb == 19 &
-%                     reg_method=='ANTS' &
-%                     method.mb =='Majority Vote')
-% 
-% cumulative_dist = 
-%   ggplot(a2a.best, aes(x=k.mb)) + stat_ecdf() + scale_y_continuous(breaks=seq(0,1,0.1))
-% 
-% a2a.best = subset(a2a.best, k.mb > 0.75) # drop the guys that suck
-% 
-% a2a.best = merge(a2a.best, atlas_comp, by=c("timestamp","atlases"))
-% a2a.best = merge(a2a.best, templ_comp, by.x=c("timestamp","templates.mb"), by.y=c("timestamp","templates"))
-% a2a.best$templates_AD_percent = (a2a.best$templates_AD / 19) * 100
-% a2a.best$templates_MCI_percent = (a2a.best$templates_MCI / 19) * 100
-% a2a.best.melted = melt(a2a.best, measure.vars=c("templates_AD_percent", "templates_MCI_percent"))
-% 
-% k_by_composition = 
-%   ggplot(a2a.best.melted, aes(x=value, y=k.mb, colour=DX)) + facet_grid(variable ~ DX) + 
-%   geom_boxplot(aes(group=as.factor(value), colour=DX)) +  geom_smooth(method="lm") 
-% 
-% snt.unilateral.vols = read.csv('data/a2a_snt_volumes.csv')
-% a2a.best = merge(a2a.best, snt.unilateral.vols)
-% 
-% # bland-altman plots
-% bland_mb_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.mb, DX) + 
-%   geom_smooth(method="lm") + 
-%   scale_y_continuous(breaks=seq(-500, 500, 50)))
-% 
-% bland_ma_overall = with(a2a.best, bland_altman_plot(volume.snt, volume.ma, DX) + 
-%   geom_smooth(method="lm") + 
-%   scale_y_continuous(breaks=seq(-500, 500, 50)))
-% 
-% # just the trials with a high percentage of MCI in the template library
-% a2a.best.highMCI = subset(a2a.best, templates_MCI_percent > 50)
-% 
-% bland_mb_highMCI = with(a2a.best.highMCI, bland_altman_plot(volume.snt, volume.ma, DX) + 
-%   geom_smooth(method="lm") + 
-%   scale_y_continuous(breaks=seq(-500, 500, 50)))
-% 
-%   
-% @
-% \end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 stderr <- function(x) sqrt(var(x)/length(x))
-ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1),
-       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
-  facet_grid(reg_method~method.mb) + 
+ggplot(subset(jens_xval_mean, (templates.mb*atlases) %% 2 == 1 & templates.mb < 21),
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
+  facet_grid(.~method.mb) + 
   stat_summary(fun.y=mean,geom='line',
-               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
+  stat_summary(fun.y=mean,geom='point',
+               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases)))) + 
   stat_summary(fun.y=mean,
@@
                geom='errorbar',
-               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
+               aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases))), linetype='solid', width=0.5) + 
   #stat_smooth(method='lm', formula=y~log(x)) + 
-  scale_y_continuous(breaks=seq(0,1,0.05)) + 
+  scale_y_continuous(breaks=seq(0,1,0.01)) + 
   scale_colour_hue(name="Number of Atlases") +
+  scale_linetype(name="Registration Method") +
   xlab( "Number of Templates" ) + 
@@
   \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
-  SNT \gs{} segmentations for 69 ADNI subjects vs. atlas and template library
+  Pruessner \gs{} segmentations for 69 ADNI subjects vs. atlas and template library
   size, registration algorithm, and label fusion method. Error bars indicate
@@
   scale_x_continuous(breaks=seq(1,20,by=2)) + 
-  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
+  scale_y_log10(limits=c(0.0002,1), breaks=c(0.01,0.05,0.1,1.0)) + 
   xlab( "Number of Templates" ) + 
@@
   than multi-atlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
-  \label{ADNI1-xval-variability}
+  \label{fig:ADNI1-xval-variability}
 \end{figure}

### 2013-12-11 cache_analysis.R

@@ -26,7 +26,6 @@ prep <- function(ants_data_file, tracc_data_file) {
   # add registration method column 
   t_data       = cbind(t_data, reg_method = "ANIMAL")
   a_data       = cbind(a_data, reg_method = "ANTS")
-  a_data$volume=0
   
   # separate MA majvote from MB so we can compute differences
   # we only compare multiatlas using majority vote
@@ -72,16 +71,14 @@ prep <- function(ants_data_file, tracc_data_file) {
 }
 
 # Jens label XVAL
-d = prep('data/a2a_ants/results_2013_11_27_jens.csv.gz', 'data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz')
+d = prep('data/a2a_ants/results-adni-xval-jens-ants-2013-12-10.csv.gz', 'data/a2a_tracc/results-adni-xval-jens-tracc-2013-12-09.csv.gz')
 write.csv(d$raw ,gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
 write.csv(d$mean,gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
 
 # SNT label XVAL
-d = prep('data/a2a_ants/results_2013_01_03.csv.gz', 'data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz')
+d = prep('data/a2a_ants/results-adni-xval-snt-ants-2013-01-03.csv.gz', 'data/a2a_tracc/results-adni-xval-tracc-2013-04-04-xfmjoin.csv.gz')
 write.csv(d$raw ,gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
 write.csv(d$mean,gzfile('data/cache/ADNI-SNT-XVAL:all_data_mean.csv.gz'))
-
-quit()
 
 #%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 # ADNI1-Complete
@@ -163,4 +160,5 @@ detach(complete)
 qc = subset(complete, TEMPQC != "Fail" & MAGeT_QC == 1 & FSL_QC == 1 )
 
 write.csv(qc            ,  'data/cache/ADNI1:qc.csv')
+write.csv(means         ,  'data/cache/ADNI1:all_volumes.csv')
 write.csv(package_totals,  'data/cache/ADNI1:package_totals.csv')

### 2013-12-12 paper.Rnw
Updated Experiment 2 methods to address reviews

@@
 gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
-moleculare (SM) hippocampal subregions with a percent error in volume that is
+moleculare (SM) hippocampal subfields with a percent error in volume that is
 at or lower than that produced by image resampling. \mb{} produces hippocampal
@@
 assessment of the hippocampus.  As such, accurate segmentation of the
-hippocampus and its subregions in MRI is a necessary first step to better
+hippocampus and its subfields in MRI is a necessary first step to better
 understand the inter-individual variability of subject neuroanatomy.
@@
 \paragraph{\adnidataset{} dataset} 
-Data used in the preparation of this article were obtained from the Alzheimer’s
+Data used in the preparation of this article were obtained from the Alzheimer's
 Disease Neuroimaging Initiative (ADNI) database (adni.loni.ucla.edu). The ADNI
@@
 Administration (FDA), private pharmaceutical companies and non-profit
-organizations, as a \$60 million, 5-year public- private partnership. The primary
+organizations, as a \$60 million, 5-year public-private partnership. The primary
 goal of ADNI has been to test whether serial magnetic resonance imaging (MRI),
@@
 neuropsychological assessment can be combined to measure the progression of mild
-cognitive impairment (MCI) and early Alzheimer’s disease (AD).  Determination of
+cognitive impairment (MCI) and early Alzheimer's disease (AD).  Determination of
 sensitive and specific markers of very early AD progression is intended to aid
@@
 The Principal Investigator of this initiative is Michael W. Weiner, MD, VA
-Medical Center and University of California – San Francisco. ADNI is the result
+Medical Center and University of California San Francisco. ADNI is the result
 of efforts of many co- investigators from a broad range of academic institutions
@@
 \subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
-In this experiment, the accuracy of the \mb{} algorithm on hippocampal subregion
-segmentation is tested using a leave-one-out cross-validation (LOOCV) design on
-two image sets. The optimal parameter settings for \mb{} found in Experiment 1 are
-used in this experiment.
-
-\paragraph{Winterburn Atlases dataset} 
+The previous experiment assesses \mb{} peformance on whole hippocampus
+segmentation. In this experiment, we assess \mb{} hippocampal subfield
+segmentation of standard 3T T1-weighted images using a leave-one-out
+cross-validation (LOOCV) design.
+
+\paragraph{Winterburn Atlases} 
 The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-segmentations of five in-vivo $300\mu$ isotropic T1-weighted MR images. The
+segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
 segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
@@
 
-In addition to the high-resolution scans distributed as part of the Winterburn
-atlases, we also obtained separate 3T T1 BRAVO acquisitions (0.9mm-isotropic
-voxels) of four of the five Winterburn atlas subjects.  Images were acquired on
-a 3T GE Discovery MR 750 system (General Electric, Milwaukee, WI) using an
-8-channel head coil with the enhanced fast gradient recalled echo 3-dimensional
+\paragraph{Experiment details} 
+Leave-one-out cross-validation (LOOCV) is a validation approach in which an
+algorithm is given all but one item in a dataset as training data (in our case,
+atlas images and labels) and then is applied to the left-out item. This is done,
+in turn, for each item in the dataset and the output across all items is
+evaluated.
+
+In this experiment, the high-resolution $0.3mm^3$ voxel Winterburn atlases are
+used as the \mb{} atlas library, but in each round of LOOCV a $0.9mm^3$ voxel
+target image corresponding to the left-out atlas is used so as to evaluate \mb{}
+on standard 3T T1-weighted resolution images.
+
+Specifically, two sets of $0.9mm^3$ voxel target images are used in this
+experiment: separately acquired T1 acquisitions of four of the five Winterburn
+atlas subjects (referred to as the {\em BRAVO} set because of the scan
+parameters used, described below), and computationally subsampled versions of
+the Winterburn atlas images (the {\em Subsampled} set).
+
+The 3T T1 BRAVO acquisitions of four of the five Winterburn atlas subjects were
+separately obtained within a short time of the original atlas image
+acquisitions.  Images were acquired on a 3T GE Discovery MR 750 system (General
+Electric, Milwaukee, WI) using an 8-channel head coil with the enhanced fast
+gradient recalled echo 3-dimensional acquisition protocol, EFGRE-BRAVO, with the
+following parameters: $TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle $=8^\circ$,
+$FOV=15.8cm$, slice thickness$=0.9mm$, 176 in-plane steps for an approximate
+isotropic resolution of 0.9mm dimension voxels.
+
+Image subsampling of the Winterburn atlases was performed using trilinear
+interpolation techniques.
+
+The template library is composed of all $0.9mm^3$ voxel target images
+(either {\em BRAVO} or {\em Subsampled}), plus an additional set of 16 of
+healthy subjects 3T T1 images acquired separately (Table
+\ref{tab:WAval-healthy-demographics}). These images were acquired on a 3T GE
+Discovery MR 750 system (General Electric, Milwaukee, WI) using an 8-channel
+head coil with the enhanced fast gradient recalled echo 3-dimensional
 acquisition protocol, EFGRE-BRAVO, with the following parameters: $TE/TR/TI =
-3.0ms/6.7ms/650ms$, flip angle $=8^\circ$, $FOV=15.8cm$, slice
-thickness$=0.9mm$, 176 in-plane steps for an approximate isotropic resolution
-of 0.9mm dimension voxels.
-
-\paragraph{Experiment details} 
-Leave-one-out cross-validation (LOOCV) is an approach in which the method to be
-validated is given all but one item in a dataset as training data (in our case,
-atlas library images and labels), and the output accuracy is measured on the
-left-out item. This is done, in turn, for each item in the dataset.
+3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
+0.9mm dimension voxels.
+
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used. Figure \ref{fig:LOOCV-schematic}
+illustrates schematically the experimental set up.
 
@@
 \end{figure}
-
-In part 2a of this experiment, we have modified the LOOCV experimental
-design in order to compare \mb{} segmentation accuracy with the accuracy of image
-resampling error. The five $0.3mm$-isotropic voxel Winterburn atlas images and
-labels are used as the atlas library, and LOOCV is conducted using the five Winterburn
-atlas images subsampled (using trilinear interpolation) to
-0.9mm-isotropic voxel resolution (referred to as the {\em Subsampled} dataset)
-as input subjects. Subsampling the target images allows us to assess \mb{} in a
-typical segmentation scenario (high-resolution atlases and lower-resolution
-subjects).  The template library is composed of all five subject images, plus
-an additional set of 16 of healthy subjects 3T T1 images acquired separately
-(Table \ref{tab:WAval-healthy-demographics}).  Images were acquired on a 3T GE
-Discovery MR 750 system (General Electric, Milwaukee, WI) using an 8-channel
-head coil with the enhanced fast gradient recalled echo 3-dimensional
-acquisition protocol, EFGRE-BRAVO, with the following parameters: $TE/TR/TI =
-3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
-thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution
-of 0.9mm dimension voxels.
-
-The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used.  As with traditional LOOCV, each subject
-image is segmented by \mb{} with that subject's image excluded from the atlas
-library.  In part 2b of this experiment, we carry out a similar validation in
-which the subject images are separately acquired T1 BRAVO images of four of the
-five Winterburn atlas subjects (referred to as the {\em T1 BRAVO} dataset).
-Figure \ref{fig:LOOCV-schematic} illustrates the experimental set up of both
-parts of this experiment.
 
@@
 \paragraph{Evaluation method}  
-To assess the \mb{} LOOCV segmentations we compute the relative percent error in 
-hippocampal volume with respect to the full resolution Winterburn atlas 
-segmentations. In addition, by computing the relative error in volume of the
-Winterburn atlas labels resampled (with nearest-neighbour interpolation) to
-0.9mm-isotropic voxels, we obtain a baseline error against which to assess \mb{}
-segmentations.
+In this experiment we must evaluate \mb{} on a task which is ill-defined since
+there are no segmentation protocols of the hippocampal subfields for T1 images
+at $0.9mm^3$ voxels (in fact, whether subfields can even be differentiated at
+this resolution is debated \citep{who?}). Therefore, we evaluate the {\it
+precision} with \mb{} produces hippocampal subfields that correspond to the
+segmentation protocol used by the given atlas library images.
+
+Specifically, we compute the relative percent error in volume of the \mb{}-produced
+hippocampal subfields with respect to the full-resolution Winterburn atlas
+segmentations. In addtion, by directly resampling the Winterburn atlas
+segmentations to $0.9mm^3$ voxel images (using nearest-neighbour interpolation)
+we can obtain a baseline relative percent error to compare \mb{} against.
 
@@
 
-In general, across hippocampal subregions the percent error in volume of \mb{}
+In general, across hippocampal subfields the percent error in volume of \mb{}
 segmentations (relative to the full-resolution Winterburn atlas segmentation)
@@
 \ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4/DG, and SR/SL/SM
-subregions all show a percent error in volume that is at or lower than
-resampling error. The Subiculum and CA2/CA3 subregions show distinctly larger
+subfields all show a percent error in volume that is at or lower than
+resampling error. The Subiculum and CA2/CA3 subfields show distinctly larger
 percent error in volume than is found through resampling.
@@
 % 
-% - take away: MB produces subregion volumes that are comparable or better than
+% - take away: MB produces subfields volumes that are comparable or better than
 %   resampling error, except for the CA2/3 where error is near 25 percent (but 

### 2013-12-13 cache_analysis.R
moved a2a_ results into main data folder

@@ -71,12 +71,12 @@ prep <- function(ants_data_file, tracc_data_file) {
 }
 
 # Jens label XVAL
-d = prep('data/a2a_ants/results-adni-xval-jens-ants-2013-12-10.csv.gz', 'data/a2a_tracc/results-adni-xval-jens-tracc-2013-12-09.csv.gz')
+d = prep('data/results-adni-xval-jens-ants-2013-12-10.csv.gz', 'data/results-adni-xval-jens-tracc-2013-12-09.csv.gz')
 write.csv(d$raw ,gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
 write.csv(d$mean,gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
 
 # SNT label XVAL
-d = prep('data/a2a_ants/results-adni-xval-snt-ants-2013-01-03.csv.gz', 'data/a2a_tracc/results-adni-xval-tracc-2013-04-04-xfmjoin.csv.gz')
+d = prep('data/results-adni-xval-snt-ants-2013-01-03.csv.gz', 'data/results-adni-xval-tracc-2013-04-04-xfmjoin.csv.gz')
 write.csv(d$raw ,gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
 write.csv(d$mean,gzfile('data/cache/ADNI-SNT-XVAL:all_data_mean.csv.gz'))
 

### 2013-12-17 paper.Rnw
cleaned up use of SNT and Gold Standard

@@
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %  - check tense: we are is have has  use/used find/found
-% Sweave/Knitr init
+%  - remove 'accuracy' unless specifically talking relative to ground truth (and
+%    even then, qualify)
+%  - overlap/similarity --> agreement? choose one. 
+%  - check that chart labels are sensible 
+%  - in particular, check that Experiment 5 match Experiment 1
 <<setup, include=FALSE, cache=FALSE>>=
@@
 % Prepares the data in a the form needed for plotting.
-<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
+<<ADNI1-xval-load,include=FALSE,cache=TRUE>>=
 jens_atlases = read.csv('data/jens_atlases.csv')
@@
 jens_xval_mean$method.mb = factor(jens_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
-mb.best.kappa = mean(subset(jens_xval_mean, reg_method   == "ANTS" & 
+mb.best.kappa = mean(subset(jens_xval_mean, 
+                                  reg_method   == "ANTS" & 
                                   method.mb    == "Majority Vote" &
@@
 \newcommand{\mb}{MAGeT-Brain}
+\newcommand{\multiatlas}{multi-atlas}
 \newcommand{\ants}{ANTS}
@@
 \newcommand{\FEPdataset}{SZ-FEP}
-\newcommand{\gs}{Gold Standard}
 \newcommand{\fsl}{FSL}
@@
 \newcommand{\maper}{MAPER}
+\newcommand{\gs}{Gold Standard}
+\newcommand{\silver}{silver standard}
+\newcommand{\prsnr}{Pruessner}
+\newcommand{\pseg}{\prsnr-protocol manual segmentation}
+\newcommand{\psegs}{\prsnr-protocol manual segmentations}
+\newcommand{\snt}{SNT}
 
@@
 \author[1]{Jon Pipitone}
-\author[1]{Matt Min Tae Park}
+\author[1]{Min Tae M. Park}
 \author[1]{Julie Winterburn}
@@
 Advances in image segmentation of magnetic resonance images (MRI) have
-demonstrated that multi-atlas approaches improve segmentation accuracy and
+demonstrated that \multiatlas approaches improve segmentation accuracy and
 precision over regular atlas-based approaches.  These approaches often rely on
@@
 the number of atlases needed while still achieving similar accuracy to
-multi-atlas approaches. Thus, our method acts as an accurate multi-atlas
+\multiatlas approaches. Thus, our method acts as an accurate \multiatlas
 approach when using special, hard-to-define atlases that are laborious to
@@
 target image and fused using a label fusion method. We conduct a 10-fold Monte
-Carlo cross-validation of whole hippocampal segmentation on a pool of 69 ADNI
+Carlo cross-validation of whole hippocampal segmentation on a pool of 60 ADNI
 subjects over a range of parameter settings, and a leave-one-out
@@
 \textbf{Results:} Using nine atlases and 19 template images, \mb{} achieves a
-mean Dice's Similarity Coefficient (DSC) of 0.84 over 10-folds of Monte Carlo
-cross-validation on 69 subjects relative to semi-automated \gs{} segmentations,
-and shows significantly lower variability in DSC than multi-atlas segmentation.
+mean Dice's Similarity Coefficient (DSC) of 0.84 over 10-folds of Monte Carlo % TODO: update
+cross-validation on 60 subjects relative to manual segmentations,
+and shows significantly lower variability in DSC than \multiatlas segmentation.
 In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1; CA4/dentate
@@
 16.244, p < 0.001$). Compared to \fsl{} and \freesurfer{}, \mb{} shows much smaller
-fixed volume bias (within $250 mm^3$ on average) to \gs{} segmentations available
-from ADNI, as well as a conservative, rather than exaggerated, proportional
-volume bias.\\ \textbf{Conclusion:} We demonstrate that \mb{} produces accurate
-hippocampal segmentations using only 5 atlases over different hippocampal
-definitions, disease populations, and acquisition types, as well as showing
-that accurate identification of the hippocampal subfields is possible.
+fixed volume bias (within $250 mm^3$ on average) to semi-automated (\snt)
+segmentations available from ADNI, as well as a conservative, rather than
+exaggerated, proportional volume bias.\\ \textbf{Conclusion:} We demonstrate
+that \mb{} produces accurate hippocampal segmentations using only 5 atlases over
+different hippocampal definitions, disease populations, and acquisition types,
+as well as showing that accurate identification of the hippocampal subfields is
+possible.
 \end{abstract}
@@
 More recently, many groups have used multiple atlases to improve overall
-segmentation accuracy (i.e. multi-atlas segmentation) over model-based approaches
+segmentation accuracy (i.e. \multiatlas segmentation) over model-based approaches
 \citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
@@
 derived from serial histological data is possible \citep{MallarChakravarty2012}.
-The novelty of this manuscript is the extension of our multi-atlas methodology
+The novelty of this manuscript is the extension of our \multiatlas methodology
 to the hippocampus using more than a single input atlas, while simultaneously
@@
 Embeddings for Atlas Propagation; \citet{Wolz2010}) because of its focus on
-performing multi-atlas segmentation with a limited number of input atlases.  The
-LEAP algorithm is a clever modification to the basic multi-atlas strategy in
+performing \multiatlas segmentation with a limited number of input atlases.  The
+LEAP algorithm is a clever modification to the basic \multiatlas strategy in
 which an atlas library is grown, beginning with a set of manually labelled
 atlases, by successively incorporating unlabelled target images once they
-themselves have been labelled using multi-atlas techniques. The sequence in
+themselves have been labelled using \multiatlas techniques. The sequence in
 which target images are labelled is chosen so that the similarity between the
@@
 hippocampus in the ADNI-1 baseline dataset, achieving a mean Dice score of 0.85
-against \gs{} segmentations.
+against \gs{} segmentations. %TODO: which segmentations were used? SNT?
 
@@
 validation to determine if hippocampal subfields can be accurately identified
-using our multi-atlas framework.  To ensure that we have not over-fit our
+using our \multiatlas framework.  To ensure that we have not over-fit our
 parameters to the aging or neurodegenerative brain, we also applied our
@@
 
-The simplest form of multi-atlas segmentation, which we call {\em basic 
-multi-atlas segmentation}, involves three steps. First, each labelled 
+The simplest form of \multiatlas segmentation, which we call {\em basic 
+\multiatlas segmentation}, involves three steps. First, each labelled 
 input image (i.e. atlas or template) is registered to an unlabelled target image.
@@
 Third, the labels are combined into a single label by label fusion
-\citep{Heckemann2006, Heckemann2011}. The basic multi-atlas segmentation method
+\citep{Heckemann2006, Heckemann2011}. The basic \multiatlas segmentation method
 is described in detail in other publications
 \citep{Collins2010,Heckemann2011,Aljabar2009}. When only a single atlas is used,
-basic multi-atlas segmentation degenerates into model-based segmentation: labels
+basic \multiatlas segmentation degenerates into model-based segmentation: labels
 are propagated from the atlas to a target, and no label fusion is needed.
@@
 T}emplates) bootstraps the creation of a large template library given a limited
-input atlas library, and then uses the template library in basic multi-atlas
+input atlas library, and then uses the template library in basic \multiatlas
 segmentation.  Images for the template library are selected from a set of input
@@
 from cases and controls).  The template library images are automatically
-labelled by each of the atlases via label propagation.  Basic multi-atlas
+labelled by each of the atlases via label propagation.  Basic \multiatlas
 segmentation is then conducted using the template library to segment the entire
@@
   \end{centering}
-  \caption{A schematic illustration of basic multi-atlas segmentation and \mb{}
+  \caption{A schematic illustration of basic \multiatlas segmentation and \mb{}
   segmentation \label{fig:MAGeT}}
@@
 segmentation quality of the \mb{} algorithm. The first two experiments assess the
-validity of \mb{} using a cross-validation design. Experiment 1 investigates the
+validity of \mb{} using cross-validation designs. Experiment 1 investigates the
 accuracy of \mb{} whole hippocampus segmentation over a wide range of parameter
@@
 schizophrenia (Experiment 3), and Alzheimer's disease (Experiment 4).
-Additionally, in Experiment 4, we compared \mb{} segmentations with those of
+Additionally, in Experiment 4, we compare \mb{} segmentations with those of
 well-known automated methods and assessed segmentation bias.
@@
 
-Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT \gs{} segmentations from the ADNI dataset \citep{Hsu2002}. This
-form of cross-validation allows us to rigorously validate a large number of
+Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using images
+from the ADNI dataset \citep{Hsu2002}, and manual whole hippocampus
+segmentations following the protocol laid out in \citep{Pruessner2000}.
+
+This form of cross-validation allows us to rigorously validate a large number of
 parameter settings of \mb{} (atlas and template library sizes, registration
@@
 
-69 1.5T images were arbitrarily selected from the baseline scans in the {\em
-\adnidataset{}} standardized dataset. 23 subjects were chosen from each disease
+60 1.5T images were arbitrarily selected from the baseline scans in the {\em
+\adnidataset{}} standardized dataset. 20 subjects were chosen from each disease
 category: cognitively normal (CN), mild cognitive impairment (MCI) and
 Alzheimer's disease (AD). Demographics for this subset are shown in Table
-\ref{tab:ADNI1-xvalidation-demographics}.  Semi-automated segmentations of the
-left and right whole hippocampi are made available with a subset of ADNI images
-\citep{Hsu2002}. These labels have been generated using the SNT tool from
-Medtronic Surgical Navigation Technologies, Louisville, CO (see Supplementary
-Materials for detailed discussion of the segmentation process used). Throughout
-this paper, we use these segmentations as the \gs{} for comparison.
+\ref{tab:ADNI1-xvalidation-demographics}. Fully manual segmentations of the
+left and right whole hippocampi in these images were provided by one author
+(JCP) according to the segmentation protocol specified in \citet{Pruessner2000}.
 
@@
 %
-<<ADNI-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
-adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
-                 546,553,572,602,610,814,1341,675,681,731,1130,41,68,70,101,
-                 249,293,316,414,698,1206,1222,1339,751,1030,3,5,10,16,22,53,
-                 168,183,241,1205,328,1095,991,213,159,337,343,642,753,1109,
-                 1217,29,56,1188,1293,149,382,431,1202)
-
+<<ADNI-xval-demographics, echo=F, dependson='ADNI1-xval-load', results="asis",cache=FALSE>>=
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
                MMSE, data = adnimerge, subset = VISCODE == 'bl' & RID %in%
-               jens_atlasess$RID, method = "reverse", test=FALSE, overall=TRUE)
+               jens_atlases$RID, method = "reverse", test=FALSE, overall=TRUE)
 caption = 
@@
 
-In this experiment, our dataset consists of 69 1.5T images and corresponding \gs{}
-segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and \gs{} segmentations used as an atlas
-library, and a validation set consisting of the remaining images segmented by
-both \mb{} and multi-atlas. The computed segmentations were compared to the \gs{}
-segmentations (see below).
+In this experiment, our dataset consists of 60 1.5T images and corresponding
+\psegs{}. In each validation round, the dataset is partitioned into
+a training set consisting of images and manual segmentations used as an atlas
+library, and a validation set consisting of the remaining images to be segmented
+by both \mb{} and \multiatlas{}. The computed segmentations are compared to the
+manual segmentations (see Evaluation below).
 
 A total of ten validation rounds were performed on each subject in the dataset,
-over each combination of parameter settings. The parameter settings
-explored are: atlas library size (1-9), template library size (1-20),
-registration method (\ants{} or \animal{}, described below), and label fusion method (majority vote,
+over each combination of parameter settings. The parameter settings explored
+are: atlas library size (1-9), template library size (1-20), registration method
+(\ants{} or \animal{}, described below), and label fusion method (majority vote,
 cross-correlation weighted majority vote, and normalized mutual information
-weighted majority vote, described below).  A total of $10 \times 69 \times 9
-\times 20 \times 2 \times 3 = \num{7452000}$ validation rounds were conducted,
-resulting in a total of $\num{1490400}$ segmentations analysed.
+weighted majority vote, described below).  In each validation round, both a
+\mb{} and \multiatlas segmentation is produced.  A total of $10 \times 60 \times
+9 \times 20 \times 2 \times 3 = \Sexpr{10*60*9*20*2*3}$ validation rounds
+were conducted and resulting segmentations analysed.
 
@@
 
-The automatically produces labels are compared to the \gs{} SNT labels provided by
-ADNI. The segmentation accuracy reported for a parameter setting is the average
-over ten validation rounds.
-
-In order to investigate the performance of \mb{} in a real world setting in which 
-only one size of atlas and template libraries are used, we explore the
-variability in label agreement at fixed parameter settings under varying  
-atlas and template library compositions. This is achieved by first computing the
-standard deviation and variance of DSC scores in each block of ten validation
+The labels produced by \mb{} and \multiatlas segmentation are compared to the
+manual labels using the Dice similarity coefficient, and the recorded value for
+each subject at each parameter setting explored in this experiment is the
+average over ten validation rounds.
+
+Additionally, the sensitivity of \mb and \multiatlas to atlas and template
+library composition is evaluated by comparing the variability in Dice scores
+over all validation rounds at fixed parameter settings.  This is achieved by
+first computing the variance of DSC scores in each block of ten validation
 rounds per subject. The distribution of these statistics across all subjects is
-then compared between \mb{} and multi-atlas using a Student's t-test. A
+then compared between \mb{} and \multiatlas using a Student's t-test. A
 significant difference between distributions is taken to show either a larger or
@@
 
-To validate \mb{} segmentation quality with respect to other established automated
-hippocampal segmentation methods, \mb{} was applied to large dataset from the ADNI
-project and the resulting segmentations were compared to those determined by
-\freesurfer{}, \fsl{}, \maper{} and the SNT \gs{} segmentation.
+To validate \mb{} segmentation quality with respect to other established
+automated hippocampal segmentation methods, \mb{} was applied to large dataset
+from the ADNI project and the resulting segmentations were compared to those
+produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated
+hippocampal segmentations (SNT) provided by ADNI.
 
@@
 \ref{tab:ADNI1-Dataset-Demographics}.
+
+Semi-automated segmentations of the left and right whole hippocampi are made
+available with a subset of ADNI images \citep{Hsu2002}. These labels have been
+generated using the SNT tool from Medtronic Surgical Navigation Technologies,
+Louisville, CO (see Supplementary Materials for detailed discussion of the
+segmentation process used).
 
@@
 \paragraph{Evaluation method}
-As in Experiment 3, the SNT \gs{} and Winterburn hippocampal segmentation
-protocols differ in the neuroanatomical features delineated, and so we 
-assessed \mb{} by the degree of (Pearson) correlation of average hippocampal volume
-across subjects. We also computed the correlation in hippocampal volume between
-existing, established automated segmentation methods -- \fsl{}, \freesurfer{}, and
-\maper{}, and SNT semi-automated segmentations. Additionally, we evaluate the
-volume-related fixed and proportional biases in all segmentation methods using
-Bland-Altman plots \citep{Bland1986}.
+As in Experiment 3, the SNT and Winterburn hippocampal segmentation protocols
+differ in the neuroanatomical features delineated, and so we assessed \mb{} by
+the degree of (Pearson) correlation of average hippocampal volume across
+subjects. We also computed the correlation in hippocampal volume between
+existing, established automated segmentation methods -- \fsl{}, \freesurfer{},
+and \maper{}, and SNT semi-automated segmentations. Additionally, we evaluate
+the volume-related fixed and proportional biases in all segmentation methods
+using Bland-Altman plots \citep{Bland1986}.
 
@@
 
-In this experiment we conducted 10 rounds of \mb{} and multi-atlas segmentation of
-of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
+In this experiment we conduct 10 rounds of \mb{} and \multiatlas segmentation of
+of 60 subjects using atlas and template library sizes varying from 1-9 and 1-20
 images respectively, two registration algorithms (\ants{} or \animal{}), and three
@@
 information weighted majority vote).  Hippocampal \mb{}-based segmentations using
-both \animal{} and \ants{} registration algorithms demonstrate good overlap with SNT
-\gs{} segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=2)}
+both \animal{} and \ants{} registration algorithms demonstrate good overlap with
+manual segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=2)}
 when using 9 atlases, 19 templates, ANTS registration, and majority vote label
@@
 voting strategy did not significantly improve segmentation accuracy, contrary to
-the findings of \citet{Aljabar2009} in basic multi-atlas segmentation. Given
+the findings of \citet{Aljabar2009} in basic \multiatlas segmentation. Given
 these findings, in the remainder of this section only results using the \ants{}
@@
 With an increasing number of templates, \mb{} shows improvement over
-multi-atlas-based segmentation in overlap accuracy when using the same number of
+\multiatlas-based segmentation in overlap accuracy when using the same number of
 atlases and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The two methods
@@
 
-In addition to an improvement in accuracy over multi-atlas-based segmentation,
+In addition to an improvement in accuracy over \multiatlas-based segmentation,
 \mb{} also shows a decrease in the variability of segmentation accuracy 
@@
 \begin{figure}
-<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-prep', fig.width=7, fig.height=7>>=
+<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
 stderr <- function(x) sqrt(var(x)/length(x))
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Mean DSC" ) + 
+  ylab( "Mean similarity (DSC)" ) + 
   theme(legend.direction = "horizontal", legend.position = "bottom", 
@@
   \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
-  Pruessner \gs{} segmentations for 69 ADNI subjects vs. atlas and template library
+  \psegs{} for 60 ADNI subjects vs. atlas and template library
   size, registration algorithm, and label fusion method. Error bars indicate
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: "increase" in mean kappa over multi-atlas
+% ADNI1-XVAL: "increase" in mean kappa over \multiatlas
 \begin{figure}
-<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-prep'>>=
+<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-load'>>=
 ggplot(subset(jens_xval_mean, reg_method=="ANTS" & 
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Difference in mean DSC" ) +
+  ylab( "Increase in mean similarity (DSC)" ) +
   theme(legend.direction = "horizontal", legend.position = "bottom")
 @
-  \caption{Difference in mean Dice's Similarity Coefficient between \mb{} and
-  multi-atlas segmentations relative to \gs{} (SNT) segmentations vs. atlas and
-  template library size when using the \ants{} registration method, and
-  majority-vote label fusion. Error bars indicate standard error.
+  \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
+  \multiatlas{} segmentations vs. atlas and template library size when using the
+  \ants{} registration method, and majority-vote label fusion.
+  Segmentation similarity is computed against \psegs{}.  
+  Error bars indicate standard error.
   \label{fig:ADNI1-xval-k-diff}}
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Difference of variance during validation (p-value)") +
+  ylab( "Difference of variance during validation (log(p))") +
   scale_colour_hue(name="Atlases") +
@@
 @
-  \caption{{\bf Difference in variability of \mb{} vs. multi-atlas segmentation 
-  accuracy.}
-  Variance of segmentation accuracy between \mb{} and multi-atlas segmentation is
+  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas segmentation 
+  accuracy with respect to manual segmentation.}
+  Variance of segmentation accuracy between \mb{} and \multiatlas segmentation is
   computed for each subject across all ten rounds of validation. Shown on the
@@
   (atlas/template library size). Only points where \mb{} mean variability is lower
-  than multi-atlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
+  than \multiatlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
   \label{fig:ADNI1-xval-variability}
@@
   scale_colour_discrete("") + 
-  opts(legend.direction = "horizontal", legend.position = "bottom")
+  theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
 \begin{figure}
-<<FEP-volumes,cache=FALSE,dependson="setup",width=5,height=5>>=
+<<FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
@@
 We found a close relationship in total bilateral hippocampal volume between all
-methods and SNT \gs{} volumes (Figure \ref{fig:ADNI-volumes-plot}).  Volumes are
-correlated with Pearson $r > 0.78$ for all methods across disease categories.
-Within disease categories (Figure \ref{fig:ADNI-volumes-boxplot}), \mb{} is
-consistently well correlated to manual volumes (Pearson $r > 0.85$), but appears
-to slightly over-estimate the volume of the AD hippocampus.
-
-To investigate the level of agreement with SNT \gs{} hippocampal
-volumes, we constructed Bland-Altman plots for each method (Figure
-\ref{fig:ADNI-Bland-Altman}).  As \citet{Bland1986} noted, high correlation
-amongst measures of the same quantity does not necessarily imply agreement (as
-correlation can be driven by a large range in true values, for instance).  All
-methods show an obvious proportional bias: \freesurfer{} and \fsl{} markedly
-under-estimate smaller hippocampi and over-estimate large hippocampi, whereas
-\maper{} and \mb{} show a reverse, conservative bias (Figure
-\ref{fig:ADNI-Bland-Altman}).  Additionally, all methods show a fixed volume
-bias, with \freesurfer{} and \fsl{} most dramatically over-estimating hippocampal
-volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and \maper{} and
-\mb{} within $250 mm^3$ on average.
+methods and the SNT semi-automated label volumes (Figure
+\ref{fig:ADNI-volumes-plot}).  Volumes are correlated with Pearson $r > 0.78$
+for all methods across disease categories.  Within disease categories (Figure
+\ref{fig:ADNI-volumes-boxplot}), \mb{} is consistently well correlated to manual
+volumes (Pearson $r > 0.85$), but appears to slightly over-estimate the volume
+of the AD hippocampus.
+
+Bland-Altman plots illustrate the level of agreement of each method with \snt
+segmentation hippocampal volumes (Figure \ref{fig:ADNI-Bland-Altman}).  As
+\citet{Bland1986} noted, high correlation amongst measures of the same quantity
+does not necessarily imply agreement (as correlation can be driven by a large
+range in true values, for instance).  All methods show an obvious proportional
+bias: \freesurfer{} and \fsl{} markedly under-estimate smaller hippocampi and
+over-estimate large hippocampi, whereas \maper{} and \mb{} show a reverse,
+conservative bias (Figure \ref{fig:ADNI-Bland-Altman}).  Additionally, all
+methods show a fixed volume bias, with \freesurfer{} and \fsl{} most
+dramatically over-estimating hippocampal volume by $2600 mm^3$ and $2800 mm^3$
+on average, respectively, and \maper{} and \mb{} within $250 mm^3$ on average.
 
 Figure \ref{fig:ADNI-segmentations} shows a qualitative comparison of \mb{} and
-SNT \gs{} hippocampal segmentations for 10 randomly selected subjects in 
-each disease category, and illustrates some of the common errors found during 
-visual inspection. Mostly frequently, we found \mb{} improperly includes the vestigial
-hippocampal sulcus and, although not anatomically incorrect, \mb{} under-estimates 
-the hippocampal body in comparison to the SNT \gs{} segmentation.
+SNT hippocampal segmentations for 10 randomly selected subjects in each disease
+category, and illustrates some of the common errors found during visual
+inspection. Mostly frequently, we found \mb{} improperly includes the vestigial
+hippocampal sulcus and, although not anatomically incorrect, \mb{}
+under-estimates the hippocampal body in comparison to the SNT segmentation.
 
@@
                       "MAGeT" = "MAGeT", 
-                      "SNT"   = "SNT \\gs{}")
+                      "SNT"   = "SNT")
 m2$DX = factor(m2$DX, levels(m2$DX)[c(2,3,1)])
@@
       ylab(expression(paste("Hippocampal volume (", mm^3, ")"))) + 
-      opts(legend.direction = "horizontal", legend.position = "bottom")
+      theme(legend.direction = "horizontal", legend.position = "bottom")
 @
   \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
-dataset by \freesurfer{}, \fsl{}, \maper{}, \mb{}, and SNT \gs{} vs. disease
+dataset by \freesurfer{}, \fsl{}, \maper{}, \mb{}, and SNT vs. disease
 category.}
@@
     geom_smooth(method="lm") + 
-    xlab(expression("SNT \\gs{} mean hippocampus volume " (mm^3))) + 
+    xlab(expression("SNT mean hippocampus volume " (mm^3))) + 
     ylab(expression("Automated mean hippocampal volume " (mm^3))) +
@@
               hjust=1.1,vjust=1.5) +
-    opts(legend.direction = "horizontal", legend.position = "bottom")
+    theme(legend.direction = "horizontal", legend.position = "bottom")
 @
@@
 dataset by each of the four automated methods investigated (\freesurfer{} (FS), \fsl{},
-\maper{}, \mb{}) vs. SNT \gs{}. Linear fit lines and Pearson correlations are shown
+\maper{}, \mb{}) vs. SNT. Linear fit lines and Pearson correlations are shown
 for each method.}
@@
             hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of SNT \\gs{} and automated hippocampal volume ' (mm^3))) + 
+  xlab(expression('Mean of SNT and automated hippocampal volume ' (mm^3))) + 
   ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
@@
   \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset{} dataset by SNT \gs{} segmentation and each of the
+measured in the \adnidataset{} dataset by SNT segmentation and each of the
 four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
@@
 Note, points below the mean difference indicate overestimation of the volume
-with respect to the \gs{} rating, and vice versa. }
+with respect to the SNT volume, and vice versa. }
 
@@
   \end{centering}
-  \caption{SNT \gs{} and \mb{} segmentations for 30 ADNI subjects (10
+  \caption{SNT and \mb{} segmentations for 30 ADNI subjects (10
   subjects randomly selected from each disease category in the subject pool used
   in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
-  anatomical image. SNT \gs{} labels appear in green, and \mb{} labels appear in
+  anatomical image. SNT labels appear in green, and \mb{} labels appear in
   blue. Noted are examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
-  {\em (b)} translated SNT \gs{} segmentation; 
+  {\em (b)} translated segmentation (seen in SNT segmentations only); 
   {\em (c)} under-estimation of hippocampal body and
@@
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy (overlap with SNT \gs{} labels)
+\caption{\textbf{Automated segmentation accuracy (overlap with SNT labels)
 of the ADNI dataset.} For each method, the number of labelled atlases
@@
 \hline
-\mb{}                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline
+\mb{}                     &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10
+rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
 LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
@@
 order to provide accurate and reliable segmentations.  Experiment 1 establishes
-\mb{} as effective and as comparable to multi-atlas segmentation. This experiment
+\mb{} as effective and as comparable to \multiatlas segmentation. This experiment
 also rigorously characterizes the behaviour of \mb{} whole hippocampal
@@
 
-Previous work on multi-atlas segmentation methods
+Previous work on \multiatlas segmentation methods
 \citep{Aljabar2009,Collins2010} has found that cross-correlation and normalized
@@
 by-product of the \mb{} algorithm itself.  However, this this pattern is also
-found in the results of the multi-atlas segmentations we used in our
+found in the results of the \multiatlas segmentations we used in our
 experiments.  We believe that our majority voting methodology is biased towards
@@
 Another concern is the moderate-to-small improvement observed in \mb{} in
-comparison to multi-atlas segmentation when using the same number of atlases.
+comparison to \multiatlas segmentation when using the same number of atlases.
 The actual benefit in using \mb{} is consistency of the labelling regardless of
@@
 that we present in Experiment 1 is amongst one of the most stringent performed
-in the multi-atlas/segmentation literature. To the best of our knowledge, other
+in the \multiatlas/segmentation literature. To the best of our knowledge, other
 groups using ADNI data for validation do at most a single round of
@@
 in comparison to the other methods.  Further, when estimating bias of these
-methods relative to SNT \gs{} hippocampal volumes we see that large hippocampi are
+methods relative to SNT hippocampal volumes we see that large hippocampi are
 over estimated while small hippocampi are under estimated. By comparison, \mb{}
@@
 methods differ.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Tabulation of subfield segmentation protocols
+%
+%
+\begin{table}[!tbp]
+\scriptsize
+\caption{\textbf{Summary of labelled subfields of the Hippocampus from recent
+MRI segmentation protocols.} \label{tab:subfield-segmentation-protocols}} 
+\begin{center}
+\begin{tabular}{ l l }
+\hline\hline
+Protocol & Labelled Subfields  \tabularnewline
+\hline
+\citet{Winterburn2013} &  CA1,  CA2/CA3,  CA4/dentate gyrus,  strata
+                          radiatum/lacunosum/moleculare, subiculum \tabularnewline
+\citet{Wisse2012}      &  CA1,  CA2,  CA3,  CA4/DG, subiculum,  entorhinal
+                          cortex \tabularnewline
+\citet{VanLeemput2009} &  CA1,  CA2/CA3,  CA4/DG, presubiculum, subiculum,
+                          hippocampal fissure,  fimbria,  hippocampal tail,
+                          inferior lateral ventricle, choroid plexus \tabularnewline
+\citet{Yushkevich2009} &  CA1, CA2/CA3, dentate gyrus (hilus), dentate gyrus
+                          (stratum moleculare), strata
+                          radiatum/lacunosom/moleculare/vestigial hippocampal
+                          sulcus \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-In conclusion, we have presented a flexible multi-atlas framework that has
+In conclusion, we have presented a flexible \multiatlas framework that has
 considerable advantages over other methods since only a small number of atlases
@@
 identification of the hippocampal subfields is possible.
-
-% \begin{verbatim}
-% Experiment 1: 
-% - address the absence of weighted-voting effects. 
-%   - hypothesis: the benefits of weighted voting only outweigh the resampling
-%     error effects when choosing from a large library.  In this experiment, the 
-%     template library (during weighted voting) consists of only 20 templates, so
-%     the most we could hope for is a "squashing" upwards of the curve towards 
-%     the 20-template limit (which is the same across all cases in this experiment).
-%   - we do see a slight "squashing" effect, with XCORR more strongly than NMI. 
-%     Perhaps in future experiments with larger template libraries, this could be
-%     explored further.
-% - Why don't we use STAPLE? What would we expect from STAPLE or other fusion methods? 
-%   - couldn't get STAPLE to work with our image formats
-%   - Idea: expect smoothness across range of templates (no even # dips, below)
-%   - perhaps more sophisticated fusion methods would boost results over majority 
-%     vote based techniques.
-% 
-% - why does MAGeT brain show a dip in average Kappa when using an even number
-%   templates. Does MA show this same pattern? (yes)
-%   - Hypothesis: our voting method is biased when breaking ties to choose the 
-%     label with the lowest numeric value.
-%   - If we compute the total number of labels fused (atlases * templates) then 
-%     both MB/MA perform worse when fusing an even number of labels. 
-% 
-% - we only show a +0.02 increase in mean Kappa over multi-atlas, and this is when
-%   using 1 atlas (which we know from Figure 2, is when we perform worst). Why does
-%   this increase justify the extra effort involved in MB? 
-%   - decrease in variability
-%   - we are comparing "true" averages (see above)... does this make our criteria 
-%     for a worthy improvement less strict? 
-%     
-% 
-% - how does MAGeT stack up, Kappa-wise, to other methods, in an absolute sense.
-% - because of the extensive cross-validation in experiment 1, we are very likely
-%   showing results that approach the true average of MB and MA on that dataset
-%   (i.e. our mid 0.8 range result is a mean across 69 subjects and 10 repetitions
-%   each).
-% - even with this caveat, do we think we do well enough?  i.e. other than parameter 
-%   tuning, to what extent does this experiment tell us about how MB would do in 
-%   practice. 
-% 
-% - effects we are seeing is not only averaging effects (Chakravarty et al. 2012). 
-% 
-% - take aways: as in other studies (Aljabar, Heckemann) we find that performance
-%   scales with the number of inputs (approx. log(n)), and that a large enough
-%   template library can boost performance with a small number of atlases. The use 
-%   of templates /can/ have a negative impact on performance (Figure 2), that is 
-%   balanced by the improvement of growing the template library.  Likely the negative
-%   impact on peformance is as a result of resampling/mislabelling error. In other 
-%   words, tuning MAGeT brain involves balancing the tension between an improvement 
-%   in peformance due to increase neuroanatomical capture and decreased peformance
-%   due to resampling error. 
-% 
-% Experiment 2: 
-% - describe resampling error during downsampling (essentially partial volume effect 
-%   due to averaging/majority vote in nearest neighbour selection)
-% - Why would some structures have greater downsampling error than others? (i.e. what
-%   is it about a structure that would make it especially prone to resampling error)
-%   - since error is discretised to whole voxels, smaller regions will show larger
-%     percent error
-%   - average unilateral volume (approx; mm3): 
-%      region   volume   downsampling error
-%      CA1      800            2%
-%      CA4/DG   600          -30%
-%      SR/SL/SM 700           30%
-%      Sibculum 350            0%
-%      CA2/3    200           15%
-% 
-% - observation: when downsampling error is small (<10 percent), MB error is larger, 
-%   and vice versa when downsampling error is large.  Could this a case where the 
-%   inevitable MAGeT resampling error outweighs the small downsampling error?
-% - observation: MAGeT produces similar volumes for the downsampled and BRAVO images
-%   which demonstrates reliability. Additionally, MB's error is in the same *direction*
-%   as the downsampling error except for SR/SL/SM
-% - for MB to show less error than resampling means that voting across templates is 
-%   in aggregate performing better than local nearest neighbour fitting. presumably 
-%   we'd see the same improvement with basic multiatlas as well.
-% 
-% - take away: MB produces subfields volumes that are comparable or better than
-%   resampling error, except for the CA2/3 where error is near 25 percent (but 
-%   even then resampling error is ~15%)
-% 
-% Experiment 3: 
-% - what is an "acceptable" r^2 value? 
-% - take away: MB has proven to be robust with atlases derived from three different
-%   segmentation protocols (SNT, Winterburn, and now Pruessner) on three different
-%   populations (older with AD progression; young and healthy; young and SZ). 
-%   
-% Experiment 4: 
-% - address the smaller difference in mean volume across disease classes. smaller 
-%   than every other method. 
-%   
-% - address the segmentation bias towards over-estimating smaller hippocampi, and 
-%   underestimating larger HC
-%   
-% - visual inspection/QC reveals MB segmentations are satisfactory (failure rate 
-%   is lower than the other methods (but we may be biased. :-). 
-%   
-% - take away: MB, with the Winterburn atlases, produces HC volumes more inline with
-%   SNT than FSL/FS.  Comparable to MAPER but with far fewer atlases required. 
-% 
-% \end{verbatim}
 
@@
 
-\subsection{Experiment 1a: Whole Hippocampus Cross-Validation with SNT 
-\gs Segmentations }
+\subsection{Experiment 5: Whole Hippocampus Segmentation Cross-Validation with SNT 
+Segmentations}
 
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT \gs{} segmentations from the ADNI dataset \citep{Hsu2002}. This
-form of cross-validation allows us to rigorously validate a large number of
-parameter settings of \mb{} (atlas and template library sizes, registration
-algorithm, and label fusion method) and select the best parameters to use in
-subsequent experiments.
+of images and SNT semi-automated segmentations from the ADNI dataset
+\citep{Hsu2002}. This form of cross-validation allows us to rigorously validate
+a large number of parameter settings of \mb{} (atlas and template library sizes,
+registration algorithm, and label fusion method) and select the best parameters
+to use in subsequent experiments.
 
@@
 Alzheimer's disease (AD). Demographics for this subset are shown in Table
-\ref{tab:ADNI1-xvalidation-demographics}.  Semi-automated segmentations of the
-left and right whole hippocampi are made available with a subset of ADNI images
-\citep{Hsu2002}. These labels have been generated using the SNT tool from
-Medtronic Surgical Navigation Technologies, Louisville, CO (see Supplementary
-Materials for detailed discussion of the segmentation process used). Throughout
-this paper, we use these segmentations as the \gs{} for comparison.
+\ref{tab:ADNI1-xvalidation-demographics}.
 
@@
 
-In this experiment, our dataset consists of 69 1.5T images and corresponding \gs{}
+In this experiment, our dataset consists of 69 1.5T images and corresponding SNT
 segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and \gs{} segmentations used as an atlas
+training set consisting of images and SNT segmentations used as an atlas
 library, and a validation set consisting of the remaining images segmented by
-both \mb{} and multi-atlas. The computed segmentations were compared to the \gs{}
-segmentations (see below).
+both \mb{} and \multiatlas{}. The computed segmentations were compared to the
+SNT segmentations (see below).
 
@@
 
-The automatically produces labels are compared to the \gs{} SNT labels provided by
+The automatically produces labels are compared to the SNT labels provided by
 ADNI. The segmentation accuracy reported for a parameter setting is the average
@@
 
-snt_xval_mb.best = mean(subset(snt_xval_data, reg_method   == "ANTS" & 
+snt_xval_mb.best = mean(subset(snt_xval_data, 
+                                  reg_method   == "ANTS" & 
                                   method.mb    == "Majority Vote" &
@@
 
-In this experiment we conducted 10 rounds of \mb{} and multi-atlas segmentation of
+In this experiment we conduct 10 rounds of \mb{} and \multiatlas segmentation of
 of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
@@
 information weighted majority vote).  Hippocampal \mb{}-based segmentations using
-both \animal{} and \ants{} registration algorithms demonstrate good overlap with SNT
-\gs{} segmentations (maximum mean DSC of \Sexpr{round(snt_xval_mb.best, digits=2)}
+both \animal{} and \ants{} registration algorithms demonstrate good agreement with SNT
+segmentations (maximum mean DSC of \Sexpr{round(snt_xval_mb.best, digits=2)}
 when using 9 atlases, 19 templates, ANTS registration, and majority vote label
@@
 With an increasing number of templates, \mb{} shows improvement over
-multi-atlas-based segmentation in overlap accuracy when using the same number of
+\multiatlas-based segmentation in overlap accuracy when using the same number of
 atlases and voting method (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). The two methods
@@
 \begin{figure}
-<<fig:ADNI1-SNT-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-prep', fig.width=5, fig.height=5>>=
+<<fig:ADNI1-SNT-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=5, fig.height=5>>=
 stderr <- function(x) sqrt(var(x)/length(x))
@@
   \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
-  SNT \gs{} segmentations for 69 ADNI subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method. Error bars indicate
-  standard error.
+  SNT segmentations for 69 ADNI subjects vs. atlas and template library
+  size, registration algorithm, and label fusion method. 
+  Error bars indicate standard error.
   \label{fig:ADNI1-SNT-xval-k-mean}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: "increase" in mean kappa over multi-atlas
+% ADNI1-XVAL: "increase" in mean kappa over \multiatlas
 \begin{figure}
-<<fig:ADN11-SNT-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-prep', fig.width=5, fig.height=5>>=
+<<fig:ADN11-SNT-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-load', fig.width=5, fig.height=5>>=
 ggplot(subset(snt_xval_mean, reg_method=="ANTS" & 
@@
   xlab( "Number of Templates" ) + 
-  ylab( "Difference in mean DSC" ) +
+  ylab( "Increase in mean similarity (DSC)" ) +
   theme(legend.direction = "horizontal", legend.position = "bottom")
 @
-  \caption{Difference in mean Dice's Similarity Coefficient between \mb{} and
-  multi-atlas segmentations relative to \gs{} (SNT) segmentations vs. atlas and
-  template library size when using the \ants{} registration method, and
-  majority-vote label fusion. Error bars indicate standard error.
+  \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
+  \multiatlas segmentations vs. atlas and template library size when using the
+  \ants{} registration method, and majority-vote label fusion. 
+  Segmentation similarity is computed against SNT segmentations.  
+  Error bars indicate standard error.
   \label{fig:ADNI1-SNT-xval-k-diff}}
@@
 @
-  \caption{{\bf Difference in variability of \mb{} vs. multi-atlas segmentation 
-  accuracy using SNT \gs.}
-  Variance of segmentation accuracy between \mb{} and multi-atlas segmentation is
+  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas segmentation 
+  with respect to SNT segmentations.}
+  Variance of segmentation accuracy between \mb{} and \multiatlas segmentation is
   computed for each subject across all ten rounds of validation. Shown on the
@@
   (atlas/template library size). Only points where \mb{} mean variability is lower
-  than multi-atlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
+  than \multiatlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
   \label{ADNI1-SNT-xval-variability}

### 2013-12-17 cache_analysis.R
Switch to using SCC derived xval-jens-ants dataset

@@ -71,7 +71,7 @@ prep <- function(ants_data_file, tracc_data_file) {
 }
 
 # Jens label XVAL
-d = prep('data/results-adni-xval-jens-ants-2013-12-10.csv.gz', 'data/results-adni-xval-jens-tracc-2013-12-09.csv.gz')
+d = prep('data/results-adni-xval-jens-ants-2013-12-13.csv.gz', 'data/results-adni-xval-jens-tracc-2013-12-09.csv.gz')
 write.csv(d$raw ,gzfile('data/cache/ADNI-JENS-XVAL:all_data.csv.gz'))
 write.csv(d$mean,gzfile('data/cache/ADNI-JENS-XVAL:all_data_mean.csv.gz'))
 

### 2013-12-17 paper.Rnw
Fixed variability plots

@@
 %  - in particular, check that Experiment 5 match Experiment 1
+%  - fix figure 2 schematic language for the experiments (not 2a 2b). 
+% 
 <<setup, include=FALSE, cache=FALSE>>=
@@
 
-lm_eqn = function(m) {
-  l <- list(a = format(coef(m)[1], digits = 2),
-      b = format(abs(coef(m)[2]), digits = 2),
-      r2 = format(summary(m)$r.squared, digits = 3));
-  if (coef(m)[2] >= 0)  {
-    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
-  } else {
-    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
-  }
-
-  as.character(as.expression(eq))
-}
 @
@@
 \pagestyle{fancy}
-\fancyhead[RO,RE]{draft - f9776d7 - September, 2013}
+\fancyhead[RO,RE]{draft - December, 2013}
 
@@
 \begin{figure}
-<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
+<<ADN1-Xval-variability,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=5,fig.height=5>>=
 ants = subset(jens_xval_data, 
@@
   #facet_grid( . ~ stat) + 
-  geom_line(size=1) +
-  geom_point(size=3) + 
+  geom_line(size=0.5) +
+  geom_point(size=2) + 
   geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
@@
   scale_x_continuous(breaks=seq(1,20,by=2)) + 
-  scale_y_log10(limits=c(0.0002,1), breaks=c(0.01,0.05,0.1,1.0)) + 
   xlab( "Number of Templates" ) + 
-  ylab( "Difference of variance during validation (log(p))") +
+  ylab( "Variability (p)") +
   scale_colour_hue(name="Atlases") +
@@
   \caption{{\bf Difference in variability of \mb{} vs. \multiatlas segmentation
-  accuracy with respect to manual segmentation.}
-  Variance of segmentation accuracy between \mb{} and \multiatlas segmentation is
-  computed for each subject across all ten rounds of validation. Shown on the
-  y-axis (scaled logarithmically) is the  p-value resulting from a t-test
-  comparing the distribution of variances at each parameter setting
-  (atlas/template library size). Only points where \mb{} mean variability is lower
-  than \multiatlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
+  with respect to manual segmentation.}
+  Variance of segmentation accuracy between \mb{} and \multiatlas segmentation
+  is computed for each subject across all ten rounds of validation. 
+  Shown on the y-axis is the  p-value resulting from a t-test comparing the
+  distribution of variances at each parameter setting (atlas/template library
+  size). 
+  Only points where \mb{} mean variability is lower than \multiatlas are shown.
+  Dashed lines indicate a p-value of 0.05 and 0.01.}
   \label{fig:ADNI1-xval-variability}
@@
 <<FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
+lm_eqn = function(m) {
+  l <- list(a = format(coef(m)[1], digits = 2),
+      b = format(abs(coef(m)[2]), digits = 2),
+      r2 = format(summary(m)$r.squared, digits = 3));
+  if (coef(m)[2] >= 0)  {
+    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
+  } else {
+    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
+  }
+
+  as.character(as.expression(eq))
+}
+
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
@@
 \begin{center}
-\begin{tabular}{ l l }
+\begin{tabular}{ l p{4in} }
 \hline\hline
@@
   scale_x_continuous(breaks=seq(1,20,by=2)) + 
-  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
+  #  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
   xlab( "Number of Templates" ) + 
-  ylab( "Difference of variance during validation (p-value)") +
+  ylab( "Variability (p)") +
   scale_colour_hue(name="Atlases") +
@@
   computed for each subject across all ten rounds of validation. Shown on the
-  y-axis (scaled logarithmically) is the  p-value resulting from a t-test
-  comparing the distribution of variances at each parameter setting
-  (atlas/template library size). Only points where \mb{} mean variability is lower
-  than \multiatlas are shown. Dashed lines indicate a p-value of 0.05 and 0.01.}
+  y-axis is the p-value from a t-test comparing the distribution of variances at
+  each parameter setting (atlas/template library size). 
+  Only points where \mb{} mean variability is lower than \multiatlas are shown.
+  Dashed lines indicate a p-value of 0.05 and 0.01.}
   \label{ADNI1-SNT-xval-variability}

### 2013-12-20 paper.Rnw
Added subregion volume p-values

@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
-this, a leave-one-out validation was conducted in which lower-resolution images
-($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
-remaining high-resolution Winterburn atlas subjects' images.
-
-In general, across hippocampal subfields the percent error in volume of \mb{}
-segmentations (relative to the full-resolution Winterburn atlas segmentation)
-compares favourably to the error resulting from image resampling (Figure
-\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4/DG, and SR/SL/SM
-subfields all show a percent error in volume that is at or lower than
-resampling error. The Subiculum and CA2/CA3 subfields show distinctly larger
-percent error in volume than is found through resampling.
-
-Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb{} subfield
-segmentations for a single subject.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Winterburn-XVAL: compare resampling
-\begin{figure}[h!]
-<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
+<<WAval-prep, cache=T>>=
 data = subset(read.csv('data/WAval.csv'))
@@
 melted = rename(melted, c("variable"="measure", "value"="percenterr"))
-levels(melted$measure) <- list("Subsampled"="resampled",
-                               "MAGeT on Subsampled" = "mb_0.9mm",
-                               "MAGeT on T1 BRAVO"="mb_bravo")
-
-ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
+levels(melted$measure) <- list("Subsampled atlas labels"="resampled",
+                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
+                               "MAGeT on T1 BRAVO images"="mb_bravo")
+
+error.t.tests = ddply(error, c("region"), function (df) { 
+  data.frame( 
+    bravo = round(t.test(df$resampled, df$mb_bravo, paired=T)$p.value,3), 
+    atlas = round(t.test(df$resampled, df$mb_0.9mm, paired=T)$p.value,3)
+  )}) 
+@
+This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
+this, a leave-one-out validation was conducted in which lower-resolution images
+($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
+remaining high-resolution Winterburn atlas subjects' images.
+
+In general, across hippocampal subfields the percent error in volume of \mb{}
+segmentations (relative to the full-resolution Winterburn atlas segmentation)
+compares favourably to the error resulting from image resampling (Figure
+\ref{fig:WAval-vol-boxplot}). In particular, there is no significant difference 
+between percent error of \mb{} volumes compared to image resampling error for the 
+CA1, Subiculum, and CA2/CA3 subregions. The \mb{} volumes of CA4/DG and SR/SL/SM 
+subregions show both significantly different (lower) percent error (p < 0.001).
+
+%the CA1, CA4/DG, and SR/SL/SM
+%subfields all show a percent error in volume that is at or lower than
+%resampling error. The Subiculum and CA2/CA3 subfields show distinctly larger
+%percent error in volume than is found through resampling.
+
+Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb{} subfield
+segmentations for a single subject.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Winterburn-XVAL: compare resampling
+\begin{figure}[h!]
+<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
+ggplot(melted, aes(y=percenterr,x=measure,colour=measure)) +
+  facet_grid(.~region) + 
   geom_hline(aes(yintercept=0), colour="grey", size=1) + 
   geom_boxplot() + 
-  labs(y="Percent error in volume", x="Subfield") +
+  geom_point(group=1, position="dodge") + 
+  labs(y="Unilateral subregion volume percent error", x="Subfield") +
+  theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + 
   scale_colour_discrete("") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
-  \caption{{\bf Average percent error in hemispheric subfield volume.} 
+  theme(legend.direction = "vertical", legend.position = "bottom")
+@
+  \caption{{\bf Percent error in unilateral subfield volume.} 
   Percent error is measured against the volumes of the unmodified Winterburn
   atlas subfield segmentations.
-  {\bf Subsampled} are volumes of the manual segmentations of the Winterburn
+  {\bf Subsampled atlas labels} volumes of the manual segmentations of the Winterburn
   atlases after resampling to $0.9mm^3$ voxels.  
-  {\bf MAGeT on Subsampled} volumes are \mb{} segmentations of the Winterburn atlas
+  {\bf MAGeT on subsampled atlas images} volumes are \mb{} segmentations of the Winterburn atlas
   images after resampling to $0.9mm^3$ voxels.
@@
 \end{sidewaysfigure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1 1yr Complete Dataset Package Totals & Failures
+%                        
+<<ADNI-xvals-pvalue, echo=F, results="asis",cache=FALSE>>=
+latex(error.t.tests, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
+    caption="\\textbf{t-test p-values whem comparing to resampled percent error.}", 
+    caption.loc = c("top"), rowname=NULL, 
+    label="tab:resampled-percent-error")
+@
 

### 2013-12-21 paper.Rnw
Adding basic WA-volume plot

@@
 theme_set(theme_bw(base_size = 12))
+lm_eqn = function(m) {
+  l <- list(a = format(coef(m)[1], digits = 2),
+      b = format(abs(coef(m)[2]), digits = 2),
+      r2 = format(summary(m)$r.squared, digits = 3));
+  if (coef(m)[2] >= 0)  {
+    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
+  } else {
+    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
+  }
+
+  as.character(as.expression(eq))
+}
 
@@
 \mb{} and \multiatlas segmentation is produced.  A total of $10 \times 60 \times
-9 \times 20 \times 2 \times 3 = \Sexpr{10*60*9*20*2*3}$ validation rounds
+9 \times 20 \times 2 \times 3 =$ \Sexpr{10*60*9*20*2*3} validation rounds
 were conducted and resulting segmentations analysed.
@@
 <<WAval-prep, cache=T>>=
-data = subset(read.csv('data/WAval.csv'))
-casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
-casted = rename(casted, c("variable"="region"))
-
-error = ddply(casted, c("file", "region"), function(df) {
-  with(df, 
-  data.frame(
-    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
-
-levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
+wa.xval.data = subset(read.csv('data/WAval.csv'))
+wa.xval.casted = dcast(melt(wa.xval.data, id.vars=c("file","version")), file + variable ~ version)
+wa.xval.casted = rename(wa.xval.casted, c("variable"="region"))
+levels(wa.xval.casted$region) <-list("CA1"="X1", "CA1"="X101", 
                             "Subiculum"="X2","Subiculum"="X102", 
@@
                             "SR/SL/SM"="X6","SR/SL/SM"="X106")
-melted = melt(error,id.vars=c("file", "region"))
-melted = rename(melted, c("variable"="measure", "value"="percenterr"))
-levels(melted$measure) <- list("Subsampled atlas labels"="resampled",
-                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
-                               "MAGeT on T1 BRAVO images"="mb_bravo")
-
-error.t.tests = ddply(error, c("region"), function (df) { 
-  data.frame( 
-    bravo = round(t.test(df$resampled, df$mb_bravo, paired=T)$p.value,3), 
-    atlas = round(t.test(df$resampled, df$mb_0.9mm, paired=T)$p.value,3)
-  )}) 
 @
@@
 <<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
-ggplot(melted, aes(y=percenterr,x=measure,colour=measure)) +
+wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
+  with(df, 
+  data.frame(
+    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
+    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
+    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
+
+wa.xval.error.melted = melt(wa.xval.error,id.vars=c("file", "region"))
+wa.xval.error.melted = rename(wa.xval.error.melted, c("variable"="measure", "value"="percenterr"))
+levels(wa.xval.error.melted$measure) <- list("Subsampled atlas labels"="resampled",
+                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
+                               "MAGeT on T1 BRAVO images"="mb_bravo")
+
+ggplot(wa.xval.error.melted, aes(y=percenterr,x=measure,colour=measure)) +
   facet_grid(.~region) + 
@@
   \label{fig:WAval-vol-boxplot}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Winterburn-XVAL: volume plots
+\begin{figure}[h!]
+<<WAval-vol-plots,cache=T,fig.width=6,fig.height=6>>=
+wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
+wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
+levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
+                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
+                               "MAGeT on T1 BRAVO images"="mb_bravo")
+
+wa.xval.vols.lm = ddply(wa.xval.volumes, c("region", "measure"), function(df) {
+   model = lm(volume ~ gold_0.3mm, df)
+   pearson = cor(df$volume, df$gold_0.3mm, use='pairwise.complete.obs')
+   data.frame(
+     eqn  = lm_eqn(model), 
+     corr = pearson)})
+
+ggplot(wa.xval.volumes, aes(x=gold_0.3mm, y=volume, colour=measure)) + 
+  facet_wrap(~region, ncol=3, scales='free') + 
+  geom_point() + 
+  geom_smooth(method='lm')  + 
+  coord_equal() + 
+  ylab(expression("Automated Unilateral Volume" (mm^3))) +
+  xlab(expression("Manual Unilateral Volume" (mm^3))) +
+  geom_text(aes(x=Inf-200, y=Inf-200, label=eqn), parse=TRUE, data=wa.xval.vols.lm, position="dodge") +
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+
+@
+  \label{fig:WAval-volumes}
 \end{figure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1 1yr Complete Dataset Package Totals & Failures
-%                        
-<<ADNI-xvals-pvalue, echo=F, results="asis",cache=FALSE>>=
-latex(error.t.tests, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
-    caption="\\textbf{t-test p-values whem comparing to resampled percent error.}", 
-    caption.loc = c("top"), rowname=NULL, 
-    label="tab:resampled-percent-error")
-@
-
 \FloatBarrier
@@
 <<FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
-lm_eqn = function(m) {
-  l <- list(a = format(coef(m)[1], digits = 2),
-      b = format(abs(coef(m)[2]), digits = 2),
-      r2 = format(summary(m)$r.squared, digits = 3));
-  if (coef(m)[2] >= 0)  {
-    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
-  } else {
-    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
-  }
-
-  as.character(as.expression(eq))
-}
-
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")

### 2013-12-21 paper.Rnw
WAval plot

@@
 \usepackage[hypcap]{caption}
+\usepackage{lineno}
 
@@
 \maketitle
+\linenumbers
 
@@
 In this paper we have describe a thorough validation of the \mb{} algorithm for
-segmentation of the hippocampus and its subfields. First, we addressed the feasibility and
+segmentation of the hippocampus and its subfields. First, we address the feasibility and
 accuracy of whole hippocampus segmentation with a limited number of input
@@
 using a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
-dataset over a range of parameters.  We then performed a leave-one-out
-validation to determine if hippocampal subfields can be accurately identified
-using our \multiatlas framework.  To ensure that we have not over-fit our
-parameters to the aging or neurodegenerative brain, we also applied our
-segmentations to a dataset of individuals suffering from first episode
-psychosis.  Finally, validated our algorithm using all of the data available in
-the \adnidataset{} sample and compare our segmentations to the other
-segmentations available through the ADNI informatics portal.
+dataset over a range of parameters. We then perform a leave-one-out
+validation to determine if hippocampal subfields can be accurately identified,
+using the best parameters discovered in the first experiment.  To ensure that we
+have not overfit our parameters to the aging or neurodegenerative brain, we also
+apply \mb{} to a dataset of individuals suffering from first episode psychosis.
+Finally, we validat our algorithm using all of the data available in the
+\adnidataset{} sample and compare our segmentations to other popular
+segmentation algorithms.
 
@@
 \begin{figure}[h!]
-<<WAval-vol-plots,cache=T,fig.width=6,fig.height=6>>=
+<<WAval-vol-plots,cache=T,fig.width=7,fig.height=6>>=
 wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
@@
      corr = pearson)})
+wa.xval.vols.lm$yoffset = c( 1225 - 0:2 * 40,
+                             465  - 0:2 * 15, 
+                             750  - 0:2 * 30,
+                             325  - 0:2 * 18, 
+                             1100 - 0:2 * 40 )
+
+wa.xval.vols.lm$xoffset = c( rep(700, 3),
+                             rep(250, 3),
+                             rep(490, 3),
+                             rep(140, 3), 
+                             rep(570, 3) )
 
@@
   geom_point() + 
+  coord_equal() + 
   geom_smooth(method='lm')  + 
-  coord_equal() + 
+  xlab(expression("Manual Unilateral Volume" (mm^3))) +
   ylab(expression("Automated Unilateral Volume" (mm^3))) +
-  xlab(expression("Manual Unilateral Volume" (mm^3))) +
-  geom_text(aes(x=Inf-200, y=Inf-200, label=eqn), parse=TRUE, data=wa.xval.vols.lm, position="dodge") +
+  geom_text(aes(x=xoffset, y=yoffset, label=eqn), parse=TRUE,
+    data=wa.xval.vols.lm, size=2,hjust=0) +
   theme(legend.direction = "horizontal", legend.position = "bottom")
@@
        aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
-  facet_grid(reg_method~method.mb) + 
+  facet_grid(.~method.mb) + 
   stat_summary(fun.y=mean,geom='line',

### 2013-12-21 paper.Rnw
Integrated Mallar's comments up to methods

@@
 Advances in image segmentation of magnetic resonance images (MRI) have
-demonstrated that \multiatlas approaches improve segmentation accuracy and
+demonstrated that \multiatlas{} approaches improve segmentation accuracy and
 precision over regular atlas-based approaches.  These approaches often rely on
 a large number of such manually segmented atlases (e.g.  30-80) that take
-significant time and expertise to produce. We present an algorithm, \mb{} (
+significant time and expertise to produce. We present an algorithm, \mb{} 
 ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
@@
 the number of atlases needed while still achieving similar accuracy to
-\multiatlas approaches. Thus, our method acts as an accurate \multiatlas
+\multiatlas{} approaches. Thus, our method acts as an accurate \multiatlas{}
 approach when using special, hard-to-define atlases that are laborious to
 construct. \\
-\textbf{Method:} \mb{} works by propagating atlas segmentations to a template library,
-formed from a subset of target images, via transformations estimated by non-linear
-image registration. The resultant segmentations are then propagated to each
-target image and fused using a label fusion method. We conduct a 10-fold Monte
-Carlo cross-validation of whole hippocampal segmentation on a pool of 60 ADNI
-subjects over a range of parameter settings, and a leave-one-out
-cross-validation (LOOCV) of hippocampal subfield segmentation using five
-high-resolution atlases. Two final experiments assess \mb{} when applied to
-first episode psychosis and Alzheimer's disease populations, and \mb{}
-segmentations are compared with existing automated methods (\fsl{}, \freesurfer{},
-\maper{}) and biases are explored.\\
+\textbf{Method:} \mb{} works by propagating atlas segmentations to a template
+library, formed from a subset of target images, via transformations estimated by
+non-linear image registration. The resultant segmentations are then propagated
+to each target image and fused using a label fusion method.  To explore the
+effect of atlases and templates library size, registation and label
+fusion method, we conduct a 10-fold Monte Carlo cross-validation of \mb{} of whole
+hippocampal segmentation on 60 ADNI subjects, manually segmented using the
+\prsnr{}-protocol, over a range of parameter settings. Using the best settings
+found, we then conduct a leave-one-out cross-validation (LOOCV) of hippocampal
+subfield segmentation using five high-resolution manually segmented atlases.
+Two final experiments assess \mb{} when applied to first episode psychosis and
+Alzheimer's disease populations, and \mb{} segmentations are compared with
+existing automated methods (\fsl{}, \freesurfer{}, \maper{}) and biases are
+explored.\\
 \textbf{Results:} Using nine atlases and 19 template images, \mb{} achieves a
-mean Dice's Similarity Coefficient (DSC) of 0.84 over 10-folds of Monte Carlo % TODO: update
-cross-validation on 60 subjects relative to manual segmentations,
-and shows significantly lower variability in DSC than \multiatlas segmentation.
-In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1; CA4/dentate
+mean Dice's Similarity Coefficient (DSC) of 0.88 (to \psegs{}) over 10-folds of
+Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
+significantly lower variability in DSC than \multiatlas{} segmentation.
+In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1;
+CA4/dentate % TODO: update
 gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
@@
 16.244, p < 0.001$). Compared to \fsl{} and \freesurfer{}, \mb{} shows much smaller
-fixed volume bias (within $250 mm^3$ on average) to semi-automated (\snt)
+fixed volume bias (within $250 mm^3$ on average) to semi-automated (\snt{})
 segmentations available from ADNI, as well as a conservative, rather than
@@
 hippocampus \citep{Geuze2004} and this has led to efforts to create an unified
-hippocampal segmentation protocol \citep{Jack2011}.  In addition, there is
-controversy over the appropriate manual segmentation protocol to use in a
-particular imaging study \citep{Nestor2012}.  Thus, a segmentation algorithm
-that can easily adapt to different manual segmentation definitions would be of
-significant benefit to the neuroimaging community.
+hippocampal segmentation protocol \citep{Jack2011,Boccardi2013,Boccardi2013a}.
+In addition, there is controversy over the appropriate manual segmentation
+protocol to use in a particular imaging study \citep{Nestor2012}. Thus, a
+segmentation algorithm that can easily adapt to different manual segmentation
+definitions would be of significant benefit to the neuroimaging community.
 
@@
 transformation is applied to the manual labels (i.e. {\em label propagation}) to
-move them into the target image space. While this methodology has been used
+warp them into the target image space. While this methodology has been used
 successfully in several contexts
@@
 More recently, many groups have used multiple atlases to improve overall
-segmentation accuracy (i.e. \multiatlas segmentation) over model-based approaches
+segmentation accuracy (i.e. \multiatlas{} segmentation) over model-based approaches
 \citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
@@
 segmentation of the striatum, pallidum, and thalamus using a single atlas
-derived from serial histological data is possible \citep{MallarChakravarty2012}.
-The novelty of this manuscript is the extension of our \multiatlas methodology
+derived from serial histological data is possible \citep{Chakravarty2013}.
+The novelty of this manuscript is the extension of our \multiatlas{} methodology
 to the hippocampus using more than a single input atlas, while simultaneously
@@
 
-Of special relevance to the present work is the LEAP algorithm (Learning
+Of particular relevance to the present work is the LEAP algorithm (Learning
 Embeddings for Atlas Propagation; \citet{Wolz2010}) because of its focus on
-performing \multiatlas segmentation with a limited number of input atlases.  The
-LEAP algorithm is a clever modification to the basic \multiatlas strategy in
+performing \multiatlas{} segmentation with a limited number of input atlases.  The
+LEAP algorithm is a clever modification to the basic \multiatlas{} strategy in
 which an atlas library is grown, beginning with a set of manually labelled
 atlases, by successively incorporating unlabelled target images once they
-themselves have been labelled using \multiatlas techniques. The sequence in
+themselves have been labelled using \multiatlas{} techniques. The sequence in
 which target images are labelled is chosen so that the similarity between the
@@
 segmentation using data from 10 subjects. In the second method,
-\citet{Yushkevich2009} hippocampal subfields were acquired and labelled on
-high-resolution MRI data from post-mortem medial temporal lobe samples. Using
-nonlinear registration guided by manually derived hippocampus masks and
-specific landmarks, the authors demonstrated accurate parcellation of
-hippocampal subfields in unlabelled MRI volumes.
+\citet{Yushkevich2009} hippocampal subfields were labelled on highly-anistropic
+($0.380mm$ in-plane resolution, $0.8mm$ coronal slice thickness) T2-weighted MR
+images acquired from post-mortem medial temporal lobe samples. Using nonlinear
+registration guided by manually derived hippocampus masks and specific
+landmarks, the authors demonstrate accurate parcellation of hippocampal
+subfields in unlabelled MRI volumes.
 
 In this paper we have describe a thorough validation of the \mb{} algorithm for
-segmentation of the hippocampus and its subfields. First, we address the feasibility and
-accuracy of whole hippocampus segmentation with a limited number of input
-atlases \citep{MallarChakravarty2012} by performing a multi-fold experiment
-using a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
-dataset over a range of parameters. We then perform a leave-one-out
-validation to determine if hippocampal subfields can be accurately identified,
-using the best parameters discovered in the first experiment.  To ensure that we
-have not overfit our parameters to the aging or neurodegenerative brain, we also
-apply \mb{} to a dataset of individuals suffering from first episode psychosis.
-Finally, we validat our algorithm using all of the data available in the
-\adnidataset{} sample and compare our segmentations to other popular
-segmentation algorithms.
+segmentation of the hippocampus and its subfields. First, we address the
+feasibility of whole hippocampus segmentation with a limited number of input
+atlases \citep{Chakravarty2013} by performing a multi-fold validation experiment using a
+subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI) dataset
+manually segmented using the \prsnr{}-protocol, over a
+range of atlas and template library sizes, registration and label fusion
+methods. We then perform a leave-one-out validation to determine if
+hippocampal subfields can be accurately identified, using the best parameters
+discovered in the first experiment.  To ensure that we have not overfit our
+parameters to the aging or neurodegenerative brain, we also apply \mb{} to a
+dataset of individuals suffering from first episode psychosis.  Finally, we
+validate our algorithm using all of the data available in the \adnidataset{}
+sample and compare our segmentations to other popular segmentation algorithms.
 
@@
 
-The simplest form of \multiatlas segmentation, which we call {\em basic 
-\multiatlas segmentation}, involves three steps. First, each labelled 
+The simplest form of \multiatlas{} segmentation, which we call {\em basic 
+\multiatlas{} segmentation}, involves three steps. First, each labelled 
 input image (i.e. atlas or template) is registered to an unlabelled target image.
@@
 Third, the labels are combined into a single label by label fusion
-\citep{Heckemann2006, Heckemann2011}. The basic \multiatlas segmentation method
+\citep{Heckemann2006, Heckemann2011}. The basic \multiatlas{} segmentation method
 is described in detail in other publications
 \citep{Collins2010,Heckemann2011,Aljabar2009}. When only a single atlas is used,
-basic \multiatlas segmentation degenerates into model-based segmentation: labels
+basic \multiatlas{} segmentation degenerates into model-based segmentation: labels
 are propagated from the atlas to a target, and no label fusion is needed.
@@
   \end{centering}
-  \caption{A schematic illustration of basic \multiatlas segmentation and \mb{}
+  \caption{A schematic illustration of basic \multiatlas{} segmentation and \mb{}
   segmentation \label{fig:MAGeT}}
@@
 weighted majority vote, described below).  In each validation round, both a
-\mb{} and \multiatlas segmentation is produced.  A total of $10 \times 60 \times
+\mb{} and \multiatlas{} segmentation is produced.  A total of $10 \times 60 \times
 9 \times 20 \times 2 \times 3 =$ \Sexpr{10*60*9*20*2*3} validation rounds
@@
   atlas labels after linear registration to the template, dilated by three voxels
-  \citep{MallarChakravarty2012}.  The number of top ranked template library image labels
+  \citep{Chakravarty2013}.  The number of top ranked template library image labels
   is a configurable parameter and displayed as the size of the template library
@@
 
-The labels produced by \mb{} and \multiatlas segmentation are compared to the
+The labels produced by \mb{} and \multiatlas{} segmentation are compared to the
 manual labels using the Dice similarity coefficient, and the recorded value for
@@
 
-Additionally, the sensitivity of \mb and \multiatlas to atlas and template
+Additionally, the sensitivity of \mb{} and \multiatlas{} to atlas and template
 library composition is evaluated by comparing the variability in Dice scores
@@
 rounds per subject. The distribution of these statistics across all subjects is
-then compared between \mb{} and \multiatlas using a Student's t-test. A
+then compared between \mb{} and \multiatlas{} using a Student's t-test. A
 significant difference between distributions is taken to show either a larger or
@@
 produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated
-hippocampal segmentations (SNT) provided by ADNI.
+hippocampal segmentations (\snt{}) provided by ADNI.
 
@@
 The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
-SNT, \maper, \freesurfer{} hippocampal volumes for a subset of images were
+\snt{}, \maper, \freesurfer{} hippocampal volumes for a subset of images were
 provided by ADNI, along with quality control data for each \freesurfer{}
@@
 available with a subset of ADNI images \citep{Hsu2002}. These labels have been
-generated using the SNT tool from Medtronic Surgical Navigation Technologies,
+generated using the \snt{} tool from Medtronic Surgical Navigation Technologies,
 Louisville, CO (see Supplementary Materials for detailed discussion of the
@@
 the segmentation did not pass).  Only images meeting the conditions of having
-segmentations from all methods (SNT, \maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also
+segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also
 passing quality control inspection were included in the analysis.
@@
 \paragraph{Evaluation method}
-As in Experiment 3, the SNT and Winterburn hippocampal segmentation protocols
+As in Experiment 3, the \snt{} and Winterburn hippocampal segmentation protocols
 differ in the neuroanatomical features delineated, and so we assessed \mb{} by
@@
 existing, established automated segmentation methods -- \fsl{}, \freesurfer{},
-and \maper{}, and SNT semi-automated segmentations. Additionally, we evaluate
+and \maper{}, and \snt{} semi-automated segmentations. Additionally, we evaluate
 the volume-related fixed and proportional biases in all segmentation methods
@@
 
-In this experiment we conduct 10 rounds of \mb{} and \multiatlas segmentation of
+In this experiment we conduct 10 rounds of \mb{} and \multiatlas{} segmentation of
 of 60 subjects using atlas and template library sizes varying from 1-9 and 1-20
@@
 voting strategy did not significantly improve segmentation accuracy, contrary to
-the findings of \citet{Aljabar2009} in basic \multiatlas segmentation. Given
+the findings of \citet{Aljabar2009} in basic \multiatlas{} segmentation. Given
 these findings, in the remainder of this section only results using the \ants{}
@@
 With an increasing number of templates, \mb{} shows improvement over
-\multiatlas-based segmentation in overlap accuracy when using the same number of
+\multiatlas{}-based segmentation in overlap accuracy when using the same number of
 atlases and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The two methods
@@
 
-In addition to an improvement in accuracy over \multiatlas-based segmentation,
+In addition to an improvement in accuracy over \multiatlas{}-based segmentation,
 \mb{} also shows a decrease in the variability of segmentation accuracy 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: "increase" in mean kappa over \multiatlas
+% ADNI1-XVAL: "increase" in mean kappa over \multiatlas{}
 \begin{figure}
@@
 @
-  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas segmentation
+  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation
   with respect to manual segmentation.}
-  Variance of segmentation accuracy between \mb{} and \multiatlas segmentation
+  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
   is computed for each subject across all ten rounds of validation. 
@@
   size). 
-  Only points where \mb{} mean variability is lower than \multiatlas are shown.
+  Only points where \mb{} mean variability is lower than \multiatlas{} are shown.
   Dashed lines indicate a p-value of 0.05 and 0.01.}
@@
 was segmented by \mb{}, and the resulting volumes compared with those obtained by
-expert semi-automated segmentation (SNT), and three other automated segmentation
+expert semi-automated segmentation (\snt{}), and three other automated segmentation
 techniques: \maper{}, \freesurfer{}, and \fsl{}. Table
@@
 We found a close relationship in total bilateral hippocampal volume between all
-methods and the SNT semi-automated label volumes (Figure
+methods and the \snt{} semi-automated label volumes (Figure
 \ref{fig:ADNI-volumes-plot}).  Volumes are correlated with Pearson $r > 0.78$
@@
 
-Bland-Altman plots illustrate the level of agreement of each method with \snt
+Bland-Altman plots illustrate the level of agreement of each method with \snt{}
 segmentation hippocampal volumes (Figure \ref{fig:ADNI-Bland-Altman}).  As
@@
 Figure \ref{fig:ADNI-segmentations} shows a qualitative comparison of \mb{} and
-SNT hippocampal segmentations for 10 randomly selected subjects in each disease
+\snt{} hippocampal segmentations for 10 randomly selected subjects in each disease
 category, and illustrates some of the common errors found during visual
@@
 hippocampal sulcus and, although not anatomically incorrect, \mb{}
-under-estimates the hippocampal body in comparison to the SNT segmentation.
+under-estimates the hippocampal body in comparison to the \snt{} segmentation.
 
@@
   \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
-dataset by \freesurfer{}, \fsl{}, \maper{}, \mb{}, and SNT vs. disease
+dataset by \freesurfer{}, \fsl{}, \maper{}, \mb{}, and \snt{} vs. disease
 category.}
@@
     geom_smooth(method="lm") + 
-    xlab(expression("SNT mean hippocampus volume " (mm^3))) + 
+    xlab(expression("\\snt mean hippocampus volume " (mm^3))) + 
     ylab(expression("Automated mean hippocampal volume " (mm^3))) +
@@
 dataset by each of the four automated methods investigated (\freesurfer{} (FS), \fsl{},
-\maper{}, \mb{}) vs. SNT. Linear fit lines and Pearson correlations are shown
+\maper{}, \mb{}) vs. \snt{}. Linear fit lines and Pearson correlations are shown
 for each method.}
@@
   \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset{} dataset by SNT segmentation and each of the
+measured in the \adnidataset{} dataset by \snt{} segmentation and each of the
 four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
@@
 Note, points below the mean difference indicate overestimation of the volume
-with respect to the SNT volume, and vice versa. }
+with respect to the \snt{} volume, and vice versa. }
 
@@
   \end{centering}
-  \caption{SNT and \mb{} segmentations for 30 ADNI subjects (10
+  \caption{\snt{} and \mb{} segmentations for 30 ADNI subjects (10
   subjects randomly selected from each disease category in the subject pool used
   in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
-  anatomical image. SNT labels appear in green, and \mb{} labels appear in
+  anatomical image. \snt{} labels appear in green, and \mb{} labels appear in
   blue. Noted are examples of common segmentation idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
-  {\em (b)} translated segmentation (seen in SNT segmentations only); 
+  {\em (b)} translated segmentation (seen in \snt{} segmentations only); 
   {\em (c)} under-estimation of hippocampal body and
@@
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy (overlap with SNT labels)
+\caption{\textbf{Automated segmentation accuracy (overlap with \snt{} labels)
 of the ADNI dataset.} For each method, the number of labelled atlases
@@
 order to provide accurate and reliable segmentations.  Experiment 1 establishes
-\mb{} as effective and as comparable to \multiatlas segmentation. This experiment
+\mb{} as effective and as comparable to \multiatlas{} segmentation. This experiment
 also rigorously characterizes the behaviour of \mb{} whole hippocampal
@@
 
-Previous work on \multiatlas segmentation methods
+Previous work on \multiatlas{} segmentation methods
 \citep{Aljabar2009,Collins2010} has found that cross-correlation and normalized
@@
 algorithm we have shown that the reduction in error is not simply a smoothing or
-averaging effect \citep{MallarChakravarty2012}.
+averaging effect \citep{Chakravarty2013}.
 
@@
 by-product of the \mb{} algorithm itself.  However, this this pattern is also
-found in the results of the \multiatlas segmentations we used in our
+found in the results of the \multiatlas{} segmentations we used in our
 experiments.  We believe that our majority voting methodology is biased towards
@@
 Another concern is the moderate-to-small improvement observed in \mb{} in
-comparison to \multiatlas segmentation when using the same number of atlases.
+comparison to \multiatlas{} segmentation when using the same number of atlases.
 The actual benefit in using \mb{} is consistency of the labelling regardless of
@@
 On that note, one author (JW), an expert manual rater, identified regular
-inconsistencies in the SNT segmentations: occurrences of over- and
+inconsistencies in the \snt{} segmentations: occurrences of over- and
 under-estimation, as well as misalignments of the entire segmentation volume
-(Figure \ref{fig:ADNI-segmentations}). Although the SNT segmentations are used
+(Figure \ref{fig:ADNI-segmentations}). Although the \snt{} segmentations are used
 as gold standards for validation in many other studies (Table
@@
 \ref{tab:other-methods} surveys some of the most recent reported DSC values
-reported on ADNI dataset, using SNT segmentations for the atlas library and as
+reported on ADNI dataset, using \snt{} segmentations for the atlas library and as
 gold standards for evaluation.  While it is difficult to compare segmentation
@@
 performance of all methods.  With the exception of \fsl{} all methods correlate
-well with the SNT volumes provided in the ADNI database.  However, \freesurfer{}
+well with the \snt{} volumes provided in the ADNI database.  However, \freesurfer{}
 and \fsl{} provide radically different definitions of the size of the hippocampus
 in comparison to the other methods.  Further, when estimating bias of these
-methods relative to SNT hippocampal volumes we see that large hippocampi are
+methods relative to \snt{} hippocampal volumes we see that large hippocampi are
 over estimated while small hippocampi are under estimated. By comparison, \mb{}
@@
 
-In conclusion, we have presented a flexible \multiatlas framework that has
+In conclusion, we have presented a flexible \multiatlas{} framework that has
 considerable advantages over other methods since only a small number of atlases
@@
 
-\subsection{ADNI SNT Hippocampal Labels}
+\subsection{ADNI \snt{} Hippocampal Labels}
 % The following blurb is taken (except for the first sentence) verbatim 
@@
 
-\subsection{Experiment 5: Whole Hippocampus Segmentation Cross-Validation with SNT 
+\subsection{Experiment 5: Whole Hippocampus Segmentation Cross-Validation with \snt{} 
 Segmentations}
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and SNT semi-automated segmentations from the ADNI dataset
+of images and \snt{} semi-automated segmentations from the ADNI dataset
 \citep{Hsu2002}. This form of cross-validation allows us to rigorously validate
@@
 caption = 
-  "\\textbf{ADNI1 SNT cross-validation subset demographics.} 
+  "\\textbf{ADNI1 \\snt{} cross-validation subset demographics.} 
   CN     - Cognitively Normal. 
@@
 
-In this experiment, our dataset consists of 69 1.5T images and corresponding SNT
+In this experiment, our dataset consists of 69 1.5T images and corresponding \snt{} 
 segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and SNT segmentations used as an atlas
+training set consisting of images and \snt{}  segmentations used as an atlas
 library, and a validation set consisting of the remaining images segmented by
 both \mb{} and \multiatlas{}. The computed segmentations were compared to the
-SNT segmentations (see below).
+\snt{} segmentations (see below).
 
@@
 
-The automatically produces labels are compared to the SNT labels provided by
+The automatically produces labels are compared to the \snt{} labels provided by
 ADNI. The segmentation accuracy reported for a parameter setting is the average
@@
 
-In this experiment we conduct 10 rounds of \mb{} and \multiatlas segmentation of
+In this experiment we conduct 10 rounds of \mb{} and \multiatlas{} segmentation of
 of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
@@
 information weighted majority vote).  Hippocampal \mb{}-based segmentations using
-both \animal{} and \ants{} registration algorithms demonstrate good agreement with SNT
+both \animal{} and \ants{} registration algorithms demonstrate good agreement with \snt{}
 segmentations (maximum mean DSC of \Sexpr{round(snt_xval_mb.best, digits=2)}
@@
 With an increasing number of templates, \mb{} shows improvement over
-\multiatlas-based segmentation in overlap accuracy when using the same number of
+\multiatlas{}-based segmentation in overlap accuracy when using the same number of
 atlases and voting method (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). The two methods
 converge in accuracy when using seven atlases.  Peak improvement in \mb{} accuracy
-(~0.02 DSC) is found when one atlas is used with a template library of 19
-images.
-
-In addition to an improvement in accuracy over \multiatlas-based segmentation,
+($+0.02$ DSC) is found when one atlas is used with a template
+library of 19 images.
+
+In addition to an improvement in accuracy over \multiatlas{}-based segmentation,
 \mb{} also shows a decrease in the variability of segmentation accuracy 
@@
   \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
-  SNT segmentations for 69 ADNI subjects vs. atlas and template library
+  \snt{} segmentations for 69 ADNI subjects vs. atlas and template library
   size, registration algorithm, and label fusion method. 
@@
   \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
-  \multiatlas segmentations vs. atlas and template library size when using the
+  \multiatlas{} segmentations vs. atlas and template library size when using the
   \ants{} registration method, and majority-vote label fusion. 
-  Segmentation similarity is computed against SNT segmentations.  
+  Segmentation similarity is computed against \snt{} segmentations.  
   Error bars indicate standard error.
@@
 @
-  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas segmentation 
-  with respect to SNT segmentations.}
-  Variance of segmentation accuracy between \mb{} and \multiatlas segmentation is
+  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation 
+  with respect to \snt{} segmentations.}
+  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation is
   computed for each subject across all ten rounds of validation. Shown on the
@@
   each parameter setting (atlas/template library size). 
-  Only points where \mb{} mean variability is lower than \multiatlas are shown.
+  Only points where \mb{} mean variability is lower than \multiatlas{} are shown.
   Dashed lines indicate a p-value of 0.05 and 0.01.}

### 2013-12-22 paper.Rnw
Figure comments, WA validation schematic update, various fixes

@@
 \newcommand{\FEPdataset}{SZ-FEP}
-\newcommand{\fsl}{FSL}
+\newcommand{\fsl}{FSL FIRST}
 \newcommand{\freesurfer}{FreeSurfer}
@@
 explored.\\
-\textbf{Results:} Using nine atlases and 19 template images, \mb{} achieves a
-mean Dice's Similarity Coefficient (DSC) of 0.88 (to \psegs{}) over 10-folds of
-Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
+\textbf{Results:} Using 9 atlases and 19 template images, \mb{} achieves a
+mean Dice's Similarity Coefficient (DSC) of \Sexpr{round(mb.best.kappa, digits=3)} 
+(to \psegs{}) over 10-folds of Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
 significantly lower variability in DSC than \multiatlas{} segmentation.
@@
   \end{centering}
-  \caption{A schematic illustration of basic \multiatlas{} segmentation and \mb{}
-  segmentation \label{fig:MAGeT}}
+  \caption{A schematic illustration of basic \multiatlas{} segmentation and
+\mb{} segmentation. In multi-atlas segmentation, manual labels from atlas images
+are warped (propagated) into subject space by applying the transformations
+estimated from non-linear image registration. The resulting candidate labels
+from all atlas images are then fused to create a final segmentation. In \mb{}
+segmentation, a template library is created by sampling (either randomly or
+representatively) from the subject images. Atlas labels are propagated to all
+template images and then to each subject image (including those used in the template
+library). The candidate the labels for a subject are then fused into a final
+segmentation.  \label{fig:MAGeT}}
 \end{figure}
@@
 where $A$ and $B$ are the regions being compared, and the cardinality is the
-volume measured in voxels.
-
-The labels produced by \mb{} and \multiatlas{} segmentation are compared to the
-manual labels using the Dice similarity coefficient, and the recorded value for
-each subject at each parameter setting explored in this experiment is the
-average over ten validation rounds.
+volume measured in voxels. The labels produced by \mb{} and \multiatlas{}
+segmentation are compared to the manual labels using the Dice similarity
+coefficient, and the recorded value for each subject at each parameter setting
+explored in this experiment is the average over ten validation rounds.
 
@@
 
-Specifically, two sets of $0.9mm^3$ voxel target images are used in this
-experiment: separately acquired T1 acquisitions of four of the five Winterburn
-atlas subjects (referred to as the {\em BRAVO} set because of the scan
-parameters used, described below), and computationally subsampled versions of
-the Winterburn atlas images (the {\em Subsampled} set).
+Specifically, this experiments has two parts, each using different sets of
+$0.9mm^3$ voxel target images: trilinear subsampled versions of the Winterburn
+atlas images (Experiment 2a), and separately acquired T1 BRAVO acquisitions of
+four of the five Winterburn atlas subjects (Experiment 2b).  Image subsampling
+of the Winterburn atlas images was performed using trilinear subsampling
+techniques.
 
@@
 
-Image subsampling of the Winterburn atlases was performed using trilinear
-interpolation techniques.
-
-The template library is composed of all $0.9mm^3$ voxel target images
-(either {\em BRAVO} or {\em Subsampled}), plus an additional set of 16 of
+The template library is composed of all $0.9mm^3$ voxel target images (either
+the {\em BRAVO} or {\em subsampled} sets), plus an additional set of 16 of
 healthy subjects 3T T1 images acquired separately (Table
@@
   \begin{centering}
+    % TODO: fix schematic
     \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
@@
   "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
-control subject sample.} Education is shown in years."
+control subject sample used in the template library (excluding the Winterburn
+atlas subjects).} Education is shown in years."
 latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
@@
 \paragraph{Evaluation method}  
-In this experiment we must evaluate \mb{} on a task which is ill-defined since
-there are no segmentation protocols of the hippocampal subfields for T1 images
-at $0.9mm^3$ voxels (in fact, whether subfields can even be differentiated at
-this resolution is debated \citep{who?}). Therefore, we evaluate the {\it
-precision} with \mb{} produces hippocampal subfields that correspond to the
-segmentation protocol used by the given atlas library images.
-
-Specifically, we compute the relative percent error in volume of the \mb{}-produced
-hippocampal subfields with respect to the full-resolution Winterburn atlas
-segmentations. In addtion, by directly resampling the Winterburn atlas
-segmentations to $0.9mm^3$ voxel images (using nearest-neighbour interpolation)
-we can obtain a baseline relative percent error to compare \mb{} against.
+Evaluating the accuracy of hippocampal subfields for T1 images at
+$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
+protocols for segmentation at this resolution to use as a reference. Instead, we
+can evaluate the {\it precision} with which \mb{} produces hippocampal subfields
+that correspond to the segmentation protocol used by the given atlas library
+images.  Specifically, we compute the relative percent error in volume of the
+\mb{}-produced hippocampal subfields with respect to the full-resolution
+Winterburn atlas segmentations. In addtion, by directly resampling the
+Winterburn atlas segmentations to $0.9mm^3$ voxels (using standard
+nearest-neighbour image resampling techniques) we obtain a scaled down version
+of the labels preserving the original segmentation protocol within the limits of
+error from rounding and interpolation. The relative percent error in volume of
+these labels provides a baseline to compare \mb{} segmentations against.
 
@@
 
-One author (MP) performed visual quality inspection for \mb{} and \freesurfer{} 
+One author (MP) performed visual quality inspection for \mb{} and \fsl{} 
 segmentations using similar quality control guidelines (if either hippocampus
@@
 \mb{} also shows a decrease in the variability of segmentation accuracy 
-(Figure \ref{ADNI1-xval-variability}).  The size of template library 
+(Figure \ref{fig:ADNI1-xval-variability}).  The size of template library 
 necessary to reach a decrease ($p < 0.05$) in variance and standard deviation
@@
 sufficient to show significant decrease in variance and standard deviation for
-3-7 atlases.
+1-9 atlases.
 
@@
     subfield segmentation; in the upper right corner is the Winterburn atlas
-    segmentation resampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
-    corner is the \mb{} segmentation of the resampled Winterburn atlas images; in the
+    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
+    corner is the \mb{} segmentation of the subsampled Winterburn atlas images; in the
     lower right corner is the \mb{} segmentation of a separately acquired T1 BRAVO
@@
 \hline
+LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
+ACM (AdaBoost-based)    &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
 \mb{}                   &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10
 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
-LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
-ACM (AdaBoost-based)    &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
 Patch-based label fusion&16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
@@
 % Prepares the data in a the form needed for plotting.
-<<ADNI1-SNT-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
+<<ADNI1-SNT-xval-prep,include=FALSE,cache=TRUE>>=
 snt_xval_data  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
 snt_xval_mean  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data_mean.csv.gz'))
+snt_xval_mean$reg_method = factor(snt_xval_mean$reg_method, c("ANTS", "ANIMAL"))
+snt_xval_mean$method.mb = factor(snt_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
 

### 2013-12-22 paper.Rnw
Significant changes to wording in response to reviewer and Mallar comments

@@
 jens_xval_mean$method.mb = factor(jens_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
-mb.best.kappa = mean(subset(jens_xval_mean, 
+mb.kappa.best = mean(subset(jens_xval_mean, 
                                  reg_method   == "ANTS" & 
@@
                                  atlases      == 9)$k.mb)
+mb.diff.best = mean(subset(jens_xval_mean, 
+                                 reg_method   == "ANTS" & 
+                                 method.mb    == "Majority Vote" &
+                                 templates.mb == 19 & 
+                                 atlases      == 1)$k_diff)
 @
@@
 \textbf{Results:} Using 9 atlases and 19 template images, \mb{} achieves a
-mean Dice's Similarity Coefficient (DSC) of \Sexpr{round(mb.best.kappa, digits=3)} 
+mean Dice's Similarity Coefficient (DSC) of \Sexpr{round(mb.kappa.best, digits=3)} 
 (to \psegs{}) over 10-folds of Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
@@
 produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated
-hippocampal segmentations (\snt{}) provided by ADNI.
-
-\paragraph{ADNI1 dataset revisited} 
+whole hippocampal segmentations (\snt{}) provided by ADNI.
+
+\paragraph{\adnidataset dataset} 
 The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
@@
 provided by ADNI, along with quality control data for each \freesurfer{}
-segmentation (guidelines described in \citep{Hartig2010}).  Clinical and
-demographic data for the entire \adnidataset{} dataset are shown in Table
+segmentation (guidelines described in \citep{Hartig2010}).
+
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
+available from ADNI \footnote{
+\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly pre-processed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
+2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
+240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.  Clinical
+and demographic data for the entire \adnidataset{} dataset are shown in Table
 \ref{tab:ADNI1-Dataset-Demographics}.
@@
 Louisville, CO (see Supplementary Materials for detailed discussion of the
-segmentation process used).
+segmentation process).
 
@@
 label fusion techniques (unweighted, cross-correlation, and normalised mutual
-information weighted majority vote).  Hippocampal \mb{}-based segmentations using
-both \animal{} and \ants{} registration algorithms demonstrate good overlap with
-manual segmentations (maximum mean DSC of \Sexpr{round(mb.best.kappa, digits=2)}
-when using 9 atlases, 19 templates, ANTS registration, and majority vote label
-fusion); Figure \ref{fig:ADNI1-xval-k-mean}).  Qualitatively, both \animal{} and
-\ants{}-based segmentations demonstrate trend overlap accuracy that increases with
-the size of atlas library and template library. Improvement in accuracy plateaus
-with template libraries larger than ten images.
-
-The use of \mb{} with \ants{} registration shows a pronounced increase
-in segmentation accuracy over \mb{} with \animal{} registration, across all
-other variable settings we tested. Additionally, by itself, using a weighted
+information weighted majority vote).  Computed segmentations are compared to
+\psegs using Dice's Similarity Coefficient (DSC).
+
+We find that for \mb{} segmentations, similarity score increases
+as atlas and template library size is increased, although with diminishing returns and
+an eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}).  A maximum
+similarity score of \Sexpr{round(mb.kappa.best, digits=3)} is found when using 9
+atlases, 19 templates, ANTS registration, and majority vote label fusion).  The
+\ants{} registration method consistently out-performs \animal{} registration
+over all variable settings we tested. Additionally, by itself, using a weighted
 voting strategy did not significantly improve segmentation accuracy, contrary to
-the findings of \citet{Aljabar2009} in basic \multiatlas{} segmentation. Given
-these findings, in the remainder of this section only results using the \ants{}
-registration algorithm and majority vote fusion will be shown.
-
-With an increasing number of templates, \mb{} shows improvement over
-\multiatlas{}-based segmentation in overlap accuracy when using the same number of
-atlases and voting method (Figure \ref{fig:ADNI1-xval-k-diff}). The two methods
-converge in accuracy when using seven atlases.  Peak improvement in \mb{} accuracy
-(~0.02 DSC) is found when one atlas is used with a template library of 19
-images.
-
-In addition to an improvement in accuracy over \multiatlas{}-based segmentation,
-\mb{} also shows a decrease in the variability of segmentation accuracy 
-(Figure \ref{fig:ADNI1-xval-variability}).  The size of template library 
-necessary to reach a decrease ($p < 0.05$) in variance and standard deviation
-grows with the size of atlas library used.  A template library of 19 images is
-sufficient to show significant decrease in variance and standard deviation for
-1-9 atlases.
+the findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation.
+Given these findings, in the remainder of this section only results using the
+\ants{} registration algorithm and majority vote fusion will be shown.
+
+With at least five templates, \mb{} shows a mean improvement in similarity score
+over \multiatlas{} segmentation when using the same size of atlas library and
+majority vote label fusion (Figure \ref{fig:ADNI1-xval-k-diff}). The magnitude
+of improvement increases with the size of template library, and shows
+diminishing returns with larger atlas libraries. Peak improvement
+(+\Sexpr{round(snt.mb.diff.best, digits=3)} DSC) is found with a single atlas
+and template library of 19 images.
+
+In addition to a mean increase in similarity score over \multiatlas{}-based
+segmentation, \mb{} also shows more consistency in similarity scores across all
+subjects and validation folds (Figure \ref{ADNI1-xval-variability}). A
+template library of at least 13 images is sufficient to show significant ($p <
+0.05$) decrease in variance for all sizes of atlas library tested (1-9 images).
 
@@
 ACM (AdaBoost-based)    &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
-\mb{}                   &9 & \Sexpr{round(mb.best.kappa, digits=3)} & & 10
+\mb{}                   &9 & \Sexpr{round(mb.kappa.best, digits=3)} & & 10
 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
@@
 
-\subsection{ADNI \snt{} Hippocampal Labels}
+\subsection{\snt{} Hippocampal Labels}
 % The following blurb is taken (except for the first sentence) verbatim 
@@
 
-Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
-of images and \snt{} semi-automated segmentations from the ADNI dataset
-\citep{Hsu2002}. This form of cross-validation allows us to rigorously validate
-a large number of parameter settings of \mb{} (atlas and template library sizes,
-registration algorithm, and label fusion method) and select the best parameters
-to use in subsequent experiments.
+This experiment is a replication of Experiment 1 using a pool of 69 images and
+\snt{} semi-automated segmentations from the ADNI dataset \citep{Hsu2002}. See
+Experiment 1 for full details on the ADNI dataset, and validation process.
 
 \subsubsection{Methods}
-\paragraph{\adnidataset{} dataset} 
+\paragraph{Dataset} 
 69 1.5T images were arbitrarily selected from the baseline scans in the {\em
@@
 Alzheimer's disease (AD). Demographics for this subset are shown in Table
-\ref{tab:ADNI1-xvalidation-demographics}.
-
-Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
-available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly pre-processed images
-which have been designated to be the "best" after quality control.  All images
-were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
-Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
-2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
-240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
-yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.
+\ref{tab:ADNI1-xvalidation-demographics}.  Each image has a corresponding
+semi-automated segmentation of the left and right whole hippocampus made
+available with ADNI images (\snt; see Supplementary Materials).
 
@@
 \paragraph{Experiment details} 
-
-Monte Carlo Cross-Validation (MCCV), also known as repeated random sub-sampling
-cross-validation, consists of repeated rounds of validation conducted on a fixed
-dataset \citep{Shao1993}. In each round, the dataset is randomly partitioned into
-a training set and a validation set. The method to be validated is then given
-the training data, and its output is compared with the validation set.
-
-In this experiment, our dataset consists of 69 1.5T images and corresponding \snt{} 
-segmentations. In each validation round, the dataset is partitioned into a
-training set consisting of images and \snt{}  segmentations used as an atlas
-library, and a validation set consisting of the remaining images segmented by
-both \mb{} and \multiatlas{}. The computed segmentations were compared to the
-\snt{} segmentations (see below).
-
 A total of ten validation rounds were performed on each subject in the dataset,
-over each combination of parameter settings. The parameter settings
-explored are: atlas library size (1-9), template library size (1-20),
-registration method (\ants{} or \animal{}, described below), and label fusion method (majority vote,
-cross-correlation weighted majority vote, and normalized mutual information
-weighted majority vote, described below).  A total of $10 \times 69 \times 9
-\times 20 \times 2 \times 3 = \num{7452000}$ validation rounds were conducted,
-resulting in a total of $\num{1490400}$ segmentations analysed.
-
-Before registration, all images underwent preprocessing with the N3 algorithm
-\citep{Sled1998} to minimize intensity nonuniformity. In this experiment we
-compared two non-linear image registration methods, three label fusion methods.
-
-The automatically produces labels are compared to the \snt{} labels provided by
-ADNI. The segmentation accuracy reported for a parameter setting is the average
-over ten validation rounds.
+for each combination of parameter settings: atlas library size (1-9), template
+library size (1-20), registration method (\ants{} or \animal{}), and label
+fusion method (majority vote, cross-correlation weighted majority vote, and
+normalized mutual information weighted majority vote).  A total of $10 \times 69
+\times 9 \times 20 \times 2 \times 3 =$ \Sexpr{10*69*9*20*2*3} validation rounds
+are conducted.  The computed segmentations for a subject are compared to the
+\snt{} labels provided by ADNI using Dice's Similarity Coefficient and the score
+is averaged over the validation rounds.
 
@@
 
-snt_xval_mb.best = mean(subset(snt_xval_data, 
+snt.mb.kappa.best = mean(subset(snt_xval_mean, 
                                  reg_method   == "ANTS" & 
@@
                                  atlases      == 9)$k.mb)
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this experiment we conduct 10 rounds of \mb{} and \multiatlas{} segmentation of
-of 69 subjects using atlas and template library sizes varying from 1-9 and 1-20
-images respectively, two registration algorithms (\ants{} or \animal{}), and three
-label fusion techniques (unweighted, cross-correlation, and normalised mutual
-information weighted majority vote).  Hippocampal \mb{}-based segmentations using
+snt.mb.diff.best = mean(subset(snt_xval_mean, 
+                                 reg_method   == "ANTS" & 
+                                 method.mb    == "Majority Vote" &
+                                 templates.mb == 19 & 
+                                 atlases      == 1)$k_diff)
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\mb{}-based segmentations using
 both \animal{} and \ants{} registration algorithms demonstrate good agreement with \snt{}
-segmentations (maximum mean DSC of \Sexpr{round(snt_xval_mb.best, digits=2)}
+segmentations (maximum mean DSC of \Sexpr{round(snt.mb.kappa.best, digits=2)}
 when using 9 atlases, 19 templates, ANTS registration, and majority vote label
@@
 
-With an increasing number of templates, \mb{} shows improvement over
-\multiatlas{}-based segmentation in overlap accuracy when using the same number of
-atlases and voting method (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). The two methods
-converge in accuracy when using seven atlases.  Peak improvement in \mb{} accuracy
-($+0.02$ DSC) is found when one atlas is used with a template
-library of 19 images.
-
-In addition to an improvement in accuracy over \multiatlas{}-based segmentation,
-\mb{} also shows a decrease in the variability of segmentation accuracy 
-(Figure \ref{ADNI1-SNT-xval-variability}).  The size of template library 
-necessary to reach a decrease ($p < 0.05$) in variance and standard deviation
-grows with the size of atlas library used.  A template library of 19 images is
-sufficient to show significant decrease in variance and standard deviation for
-3-7 atlases.
-
-We have omitted results obtained when using an even number of atlases or
-templates since with this configuration we found significantly decreased
-performance. We believe this is as a result of an inherent bias in the majority
-vote fusion method used (see Discussion).
+With at least five templates, \mb{} shows a mean improvement in similarity score
+over \multiatlas{}-based segmentation for the same size of atlas library with
+majority vote label fusion (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). The
+magnitude of improvement increases with the size of template library, and shows
+diminishing returns with larger atlas libraries. Peak improvement
+(+\Sexpr{round(snt.mb.diff.best, digits=3)} DSC) is found with a single atlas
+and template library of 19 images.
+
+In addition to a mean increase in similarity score over \multiatlas{}-based
+segmentation, \mb{} also shows more consistency in similarity scores across all
+subjects and validation folds (Figure \ref{ADNI1-SNT-xval-variability}). A
+template library of at least 13 images is sufficient to show significant ($p <
+0.05$) decrease in variance for all sizes of atlas library tested (1-9 images).
 

### 2013-12-22 paper.Rnw
Experiment 5 write up overhauled. WAval bland-altman plot

@@
       b = format(abs(coef(m)[2]), digits = 2),
-      r2 = format(summary(m)$r.squared, digits = 3));
+      pearson = format(sqrt(summary(m)$r.squared), digits = 3));
   if (coef(m)[2] >= 0)  {
-    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
+    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)~"="~pearson,l)
   } else {
-    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
+    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)~"="~pearson,l)    
   }
@@
 <<WAval-prep, cache=T>>=
-wa.xval.data = subset(read.csv('data/WAval.csv'))
-wa.xval.casted = dcast(melt(wa.xval.data, id.vars=c("file","version")), file + variable ~ version)
+wa.xval.vols = subset(read.csv('data/WAval.csv'))
+wa.xval.kappa = subset(read.csv('data/WAval-kappa.csv'))
+wa.xval.casted = dcast(melt(wa.xval.vols, id.vars=c("file","version")), file + variable ~ version)
 wa.xval.casted = rename(wa.xval.casted, c("variable"="region"))
@@
     data=wa.xval.vols.lm, size=2,hjust=0) +
+  scale_color_discrete(name="") + 
   theme(legend.direction = "horizontal", legend.position = "bottom")
@@
   \label{fig:WAval-volumes}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Winterburn-XVAL: bland-altman 
+\begin{figure}[h!]
+<<WAval-bland,cache=T,fig.width=7,fig.height=6>>=
+wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
+wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
+levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
+                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
+                               "MAGeT on T1 BRAVO images"="mb_bravo")
+
+wa.xval.volumes$diff = wa.xval.volumes$gold_0.3mm - wa.xval.volumes$volume 
+wa.xval.volumes$mean = (wa.xval.volumes$gold_0.3mm + wa.xval.volumes$volume)/2
+wa.xval.bland.limits = ddply(wa.xval.volumes,c("region"),function(df) {
+  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
+})
+
+ggplot(wa.xval.volumes, aes(x=mean, y=diff, colour=measure)) + 
+  facet_wrap(~region, ncol=3, scales='free') + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=1) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=wa.xval.bland.limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
+             data = wa.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of manual and automated HC volume ' (mm^3))) + 
+  ylab(expression('Difference of manaul and automated HC volume ' (mm^3))) + 
+  scale_color_discrete(name="") + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+
+@
+  \label{fig:WAval-bland}
 \end{figure}
@@
 
-\mb{}-based segmentations using
-both \animal{} and \ants{} registration algorithms demonstrate good agreement with \snt{}
-segmentations (maximum mean DSC of \Sexpr{round(snt.mb.kappa.best, digits=2)}
-when using 9 atlases, 19 templates, ANTS registration, and majority vote label
-fusion); Figure \ref{fig:ADNI1-xval-k-mean}).  Qualitatively, both \animal{} and
-\ants{}-based segmentations demonstrate trend overlap accuracy that increases with
-the size of atlas library and template library. Improvement in accuracy plateaus
-with template libraries larger than ten images.
-
-The use of \mb{} with \ants{} registration shows a pronounced increase
-in segmentation accuracy over \mb{} with \animal{} registration, across all
-other variable settings we tested. Additionally, by itself, using a weighted
-voting strategy did not significantly improve segmentation accuracy, contrary to
-the findings of \citet{Aljabar2009} in basic \multiatlas{} segmentation. Given
-these findings, in the remainder of this section only results using the \ants{}
-registration algorithm and majority vote fusion will be shown.
-
-With at least five templates, \mb{} shows a mean improvement in similarity score
-over \multiatlas{}-based segmentation for the same size of atlas library with
-majority vote label fusion (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). The
-magnitude of improvement increases with the size of template library, and shows
-diminishing returns with larger atlas libraries. Peak improvement
-(+\Sexpr{round(snt.mb.diff.best, digits=3)} DSC) is found with a single atlas
-and template library of 19 images.
-
-In addition to a mean increase in similarity score over \multiatlas{}-based
-segmentation, \mb{} also shows more consistency in similarity scores across all
-subjects and validation folds (Figure \ref{ADNI1-SNT-xval-variability}). A
-template library of at least 13 images is sufficient to show significant ($p <
-0.05$) decrease in variance for all sizes of atlas library tested (1-9 images).
+As when comparing against manual labels in Experiment 1, we find similar
+behaviour when comparing \mb{} segmentations to \snt{} labels: similarity scores
+increase with increasing numbers of atlases and templates, with diminishing
+increases in improvement trending towards a plateau (Figure
+\ref{fig:ADNI1-xval-k-mean}).  As in Experiment 1, using \ants{} registration
+leads to significantly increased similarity scores, and there is no significant
+difference in scores from any of the label fusion methods.  Mean DSC score peaks
+at \Sexpr{round(snt.mb.kappa.best, digits=3)} when using 9 atlases, 19
+templates, ANTS registration, and majority vote label fusion).  Compared to
+\multiatlas segmentations, we find \mb{} segmentations show increasing
+improvement with larger atlas and template libraries when using more than 9
+templates and 5 or fewer atlases (Figure \ref{fig:ADNI1-SNT-xval-k-diff}). Peak
+improvement (+\Sexpr{round(snt.mb.diff.best, digits=3)} DSC) is found with a
+single atlas and template library of 19 images.  In addition to a mean increase
+in similarity score over \multiatlas{}-based segmentation, \mb{} also shows more
+consistency in similarity scores across all subjects and validation folds
+(Figure \ref{ADNI1-SNT-xval-variability}) with a large enough template library.
 
@@
 ggplot(subset(snt_xval_mean, (templates.mb*atlases) %% 2 == 1),
-       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + #as.factor((atlases*templates.mb) %% 2 ))) + 
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
   facet_grid(.~method.mb) + 
   stat_summary(fun.y=mean,geom='line',
-               aes(y=k.mb, weight=1, group=as.factor(atlases))) + 
+               aes(y=k.mb, weight=1, group=interaction(reg_method,atlases))) + 
   stat_summary(fun.y=mean,
@@
                geom='errorbar',
-               aes(y=k.mb, weight=1, group=as.factor(atlases)), width=0.5) + 
+               aes(y=k.mb, weight=1, group=interaction(reg_method,atlases)), width=0.5) + 
   #stat_smooth(method='lm', formula=y~log(x)) + 

### 2013-12-23 paper.Rnw
Discussion complete

@@
 \newcommand{\maper}{MAPER}
-\newcommand{\gs}{Gold Standard}
-\newcommand{\silver}{silver standard}
 \newcommand{\prsnr}{Pruessner}
@@
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using images
-from the ADNI dataset \citep{Hsu2002}, and manual whole hippocampus
-segmentations following the protocol laid out in \citep{Pruessner2000}.
-
-This form of cross-validation allows us to rigorously validate a large number of
+from the ADNI dataset \citep{Jack2008}, and manual whole hippocampus
+segmentations following the protocol laid out in \citep{Pruessner2000}.  This
+form of cross-validation allows us to rigorously validate a large number of
 parameter settings of \mb{} (atlas and template library sizes, registration
@@
 subsequent experiments.
+
+In the Supplementary Materials we have replicated this experiment using the \snt
+semi-automated labels included with the ADNI dataset.
 
@@
 information weighted majority vote).  Computed segmentations are compared to
-\psegs using Dice's Similarity Coefficient (DSC).
+\psegs using Dice's Similarity Coefficient (DSC).  In the Supplementary
+Materials we have replicated this experiment using the \snt semi-automated
+labels included with the ADNI dataset.
 
@@
                                "MAGeT on T1 BRAVO images"="mb_bravo")
-
 wa.xval.vols.lm = ddply(wa.xval.volumes, c("region", "measure"), function(df) {
@@
                              1100 - 0:2 * 40 )
-
 wa.xval.vols.lm$xoffset = c( rep(700, 3),
@@
                              rep(570, 3) )
-
 ggplot(wa.xval.volumes, aes(x=gold_0.3mm, y=volume, colour=measure)) + 
@@
 In this manuscript we have presented the implementation and validation of the
-\mb{} framework -- a methodology that requires very few input atlases in
-order to provide accurate and reliable segmentations.  Experiment 1 establishes
-\mb{} as effective and as comparable to \multiatlas{} segmentation. This experiment
-also rigorously characterizes the behaviour of \mb{} whole hippocampal
-segmentation under various parameter settings, allowing us to choose an optimal
-setting for subsequent experiments. Experiment 2 demonstrates the accuracy of
-\mb{} hippocampal subfield identification despite contrast and resolution
-limitations in standard T1-weighted image volumes.  Experiments 2 and 3 validate
-\mb{} whole hippocampal segmentation accuracy and consistency on populations with
-different ageing and neuropsychiatric characteristics.  Furthermore, taken
-together, these experiments demonstrate that algorithmic performance is not
-dependent on a single definition of the hippocampus but is effective with
-differing hippocampal definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}.
-
-Throughout the cross validation in Experiment 1 (10-fold Monte Carlo cross
-validation in the \adnidataset{} dataset subsample) we find that two parameter
-choices improve segmentation accuracy: increasing the number of atlases, and the
-number of templates. However, after setting the parameters to 5 atlases and 15
-templates there are diminishing returns with respect to this improvement.
-
-Previous work on \multiatlas{} segmentation methods
+\mb{} framework -- a methodology that requires very few input atlases in order
+to provide accurate and reliable segmentations.  Experiment 1 compares \mb{} to
+basic-\multiatlas{} segmentation by characterising the behaviour of each
+approach with varying parameter settings, and allows us to choose an optimal
+setting for subsequent experiments. Experiment 2 demonstrates the reliability of
+\mb{} in producing subfield segmentations which match the segmentation protocol
+of the input atlases despite contrast and resolution limitations in standard
+T1-weighted image volumes. Experiments 2 and 3 validate whole hippocampal
+segmentation precision and consistency on populations with different ageing and
+neuropsychiatric characteristics. All of these experiments together demonstrate
+that \mb{} algorithmic performance is not dependent on a single definition of
+the hippocampus but is effective with differing hippocampal definitions
+\citep{Winterburn2013,Pruessner2000,Hsu2002}.
+
+The core claim the \mb{} method is based on, that we can meaningfully bootstrap
+a template library from a small set of labelled atlas images, is validated in
+the cross validation conducted in Experiment 1.  We find that both increasing
+the number of atlases and the number of templates used improves \mb{}
+segmentation over and above basic-\multiatlas segmentations using the same
+number of atlas images. That is, by taking the extra step of generating a
+template library from a small atlas set \mb{} is able to improve the final
+segmentation similarity. The magnitude of this improvement is greatest with a
+small number of atlases, but even with larger atlas libraries we have found that
+generating a template library reduces the variability in segmentations (i.e.
+\mb{} more consistently produces high quality segmentations than does
+basic-\multiatlas).
+
+Interestingly, previous work on \multiatlas{} segmentation methods
 \citep{Aljabar2009,Collins2010} has found that cross-correlation and normalized
 mutual information-based weighted label fusion improves segmentation accuracy
-over simple majority vote label fusion. Selectively filtering out atlases with
-lower image similarity is believed to reduce sources of error from estimating
-deformations via nonlinear registration, partial volume effects from nearest
-neighbour image resampling, and neuroanatomical mismatch between atlases and
-subjects. That \mb{} does not see the same boost in performance from weighted
-voting may suggest that the neuroanatomical variability of a template library
-constructed from study subjects more closely matches any particular subject and
-thereby leaving less error to filter. From our previous work on the \mb{}
-algorithm we have shown that the reduction in error is not simply a smoothing or
-averaging effect \citep{Chakravarty2013}.
+over simple majority vote label fusion, and yet we did not see a significant
+indication of this effect in the \mb{} segmentations. Selectively filtering out
+atlases with lower image similarity is believed to reduce sources of error from
+estimating deformations via nonlinear registration, partial volume effects from
+nearest neighbour image resampling, and neuroanatomical mismatch between atlases
+and subjects. That \mb{} does not see the same boost in performance from
+weighted voting may suggest that the neuroanatomical variability of a template
+library constructed from study subjects more closely matches any particular
+subject and thereby leaving less error to filter. From our previous work on the
+\mb{} algorithm we have shown that the reduction in error is not simply a
+smoothing or averaging effect \citep{Chakravarty2013}.
 
@@
 \citep{Warfield2004} (or variants of the STAPLE algorithm
-\citep{Robitaille2012}) which weights each segmentation based upon its
-estimated performance level with respect to the other available candidate
-segmentations.  Further, the sensitivity and specificity parameters can also be
-tuned to potentially improve segmentation accuracy and reliability.  It is
-likely that using more sophisticated voting methods would have a positive effect
-on the overall segmentation performance, as demonstrated by the STAPLE
-algorithm.
-
-To this end, more work is required to determine the source of the slight
-decrease in segmentation performance when the number of templates are set to an
-even number.  Our initial concern was that this dip in performance was a
-by-product of the \mb{} algorithm itself.  However, this this pattern is also
-found in the results of the \multiatlas{} segmentations we used in our
-experiments.  We believe that our majority voting methodology is biased towards
-labels with the lowest numeric values when breaking ties (by way of the
-implementation of the \verb+mode+ function used to determine majority), thus
-causing the slight bias observed when using an even number of templates. This is
-another area where the voting scheme could be used to improve performance.
-However, it is worth noting that this limitation was previously identified by
-\citet{Heckemann2006a} and, subsequently, other groups have not even considered
-the potential pitfalls of an even number of candidate labels (e.g.
-\citet{Leung2010}).
-
-Another concern is the moderate-to-small improvement observed in \mb{} in
-comparison to \multiatlas{} segmentation when using the same number of atlases.
-The actual benefit in using \mb{} is consistency of the labelling regardless of
-atlas or template choice. This is an important consideration that few have
-touched on previously. The 10-fold Monte Carlo leave-one-out cross-validation
-that we present in Experiment 1 is amongst one of the most stringent performed
-in the \multiatlas/segmentation literature. To the best of our knowledge, other
-groups using ADNI data for validation do at most a single round of
-leave-one-out-validation (Table \ref{tab:other-methods}), aside from
-\citep{Wang2011} who perform a similarly rigorous validation as us. The
-thoroughness of our validation suggests that our results are reflective of a
+\citep{Robitaille2012}) which weights each segmentation based upon its estimated
+performance level with respect to the other available candidate segmentations.
+Further, the sensitivity and specificity parameters can also be tuned to
+potentially improve segmentation accuracy and reliability.  It is likely that
+using more sophisticated voting methods would have a positive effect on the
+overall segmentation performance, as demonstrated by the STAPLE algorithm.
+
+More work is required to determine the source of the slight decrease in
+segmentation performance when the number of templates are set to an even number.
+Our initial concern was that this dip in performance was a by-product of the
+\mb{} algorithm itself.  However, this this pattern is also found in the results
+of the \multiatlas{} segmentations we used in our experiments.  We believe that
+our majority voting methodology is biased towards labels with the lowest numeric
+values when breaking ties (by way of the implementation of the \verb+mode+
+function used to determine majority), thus causing the slight bias observed when
+using an even number of templates. This is another area where the voting scheme
+could be used to improve performance.  However, it is worth noting that this
+limitation was previously identified by \citet{Heckemann2006a} and,
+subsequently, other groups have not even considered the potential pitfalls of an
+even number of candidate labels (e.g.  \citet{Leung2010}).
+
+Another concern is the moderate improvement observed in \mb{} in comparison to
+\multiatlas{} segmentation when using the same number of atlases.  The actual
+benefit in using \mb{} is consistency of the labelling regardless of atlas or
+template choice, as mentioned above. This is an important consideration that few
+have touched on previously. The 10-fold Monte Carlo cross-validation that we
+present in Experiment 1 is amongst one of the most stringent performed in the
+\multiatlas/segmentation literature. To the best of our knowledge, with the
+exception of \citep{Wang2011}, other groups using ADNI data for validation do at
+most a single round of leave-one-out-validation (Table \ref{tab:other-methods}).
+The thoroughness of our validation suggests that our results are reflective of a
 true average over the choice of parameter settings and are independent of atlas
@@
 under-estimation, as well as misalignments of the entire segmentation volume
-(Figure \ref{fig:ADNI-segmentations}). Although the \snt{} segmentations are used
-as gold standards for validation in many other studies (Table
-\ref{tab:other-methods}) providing a benchmark for comparison, their use in
-Experiments 1 and 4 presents the possibility that the DSC scores and
-correlations we obtained are lower than what would be found with a more
-accurate and consistent gold standard.
-
-% Validation of ADNI in other works
-% Chupin 2009: LOOCV, 1-fold validation
-% Heckemann 2011: QC of segmentations, comparison across field strengths, 
-%                 agreement between two different atlas sets
-% Coupe 2011e: LOOCV
-% Lotjonen 2010: LOOCV, 1-fold validation
-% Leung 2010: 1-fold, QC, classification
-% Wolz 2010: 1-fold over different leap settings, classification
-% Collins 2010: 1-fold validation, QC
-% Morra 2008: LOOCV with 21 atlases, 2 manual raters
-% Wang 2011: 10 rounds of MCCV with 20 atlases and 20 subjects, chosen from 
-%            a pool of 139 images
-
-In comparison to other methodologies in the field \mb{} performs favourably. Table
-\ref{tab:other-methods} surveys some of the most recent reported DSC values
-reported on ADNI dataset, using \snt{} segmentations for the atlas library and as
-gold standards for evaluation.  While it is difficult to compare segmentation
-results across studies, gold standards, evaluation metrics, and algorithms it is
-worth noting that the methods summarized require more atlases (between 16-55)
-than our \mb{} implementation with the Winterburn atlases \citep{Winterburn2013}.
+(Figure \ref{fig:ADNI-segmentations}). Although the \snt{} segmentations are
+used as benchmarks for validation in many other studies (Table
+\ref{tab:other-methods}), these segmentation inconsistencies present the
+possibility that a more accurate and consistent benchmark segmentation protocol
+ought to be used in order to truly understand the results of such validations.
+Indeed, our replication of the 10-fold cross-validation using \snt{}
+segmentations (Experiment 5) shows noticeably poorer mean similarity scores for
+both \mb{} and \multiatlas.
+
+Thus, in comparison to other methodologies in the field \mb{} performs
+favourably. Table \ref{tab:other-methods} surveys some of the most recent
+reported DSC values reported on ADNI dataset, using \snt{} segmentations for the
+atlas library and as gold standards for evaluation.  While it is difficult to
+compare segmentation results across studies, gold standards, evaluation metrics,
+and algorithms it is worth noting that the methods summarized require more
+atlases (between 16-55) than our \mb{} implementation with the Winterburn
+atlases \citep{Winterburn2013}.
 
@@
 Among the automated segmentation methods we compared in this paper
-(\freesurfer{}, \maper{}, \fsl{}-FIRST), we find extremely variable
+(\freesurfer{}, \maper{}, \fsl{}), we find extremely variable
 performance of all methods.  With the exception of \fsl{} all methods correlate
@@
 
-Finally, we have also demonstrated that our algorithmic framework is
+Finally, we have provide evidence that using the Winterburn high-resolution
+hippocampal subfield atlases \citep{Winterburn2013} our algorithmic framework is
 appropriate for the segmentation of hippocampal subfields in standard
-T1-weighted data.  Subfield segmentation is a burgeoning topic in the
-literature although very few automated methods are available for the
-segmentation of 3T data \citep{Yushkevich2009,VanLeemput2009}. While recent
-work demonstrates that subfield segmentations can be used for classification of
-AD, MCI, and CN, there has been no explicit validation of the methodology based
-on accuracy or precision.  Although the initial hippocampal subfield
-segmentations from the Yushkevich group have been demonstrated to work on the
-ADNI population, there has also been no validation of their work. In addition,
-their work requires some manual initialization to properly function.  Our work
-demonstrates that we can reliably identify the CA1, subiculum, and CA4/dentate
-with only modest amounts of error.  The fact that CA2/CA3 and molecular layers
-cannot be reasonably identified should not be surprising as these are extremely
+T1-weighted data. Subfield segmentation is a burgeoning topic in the literature
+although very few automated methods are available for the segmentation of 3T
+data \citep{Yushkevich2009,VanLeemput2009}. While recent work demonstrates that
+subfield segmentations can be used for classification of AD, MCI, and CN, there
+has been no explicit validation of the segmentations being produced other than
+indirectly via classification accuracy (for instance, the semi-automated
+subfield segmentations produced in \citet{Yushkevich2009}).  Our work
+demonstrates that we can reliably produce CA1, subiculum, and CA4/dentate
+subfield segmentations that correspond to those in the input atlas with only 
+only modest amounts of error in volume.  The fact that CA2/CA3 and molecular layers
+cannot be reasonably reproduced should not be surprising as these are extremely
 thin and spatially convoluted regions that originally required high-resolution
-MRI for identification.  It is likely that the extents of these regions are
-well below the resolution offered by standard T1-weighted images and, in fact,
-many manual segmentation methodologies do not attempt to parse these regions
-either \citep{Wisse2012, Mueller2009}.
-
-In conclusion, we have presented a flexible \multiatlas{} framework that has
+MRI for identification and so it is likely that the extents of these regions are
+well below the resolution and contrast offered by standard T1-weighted images.
+In addition, some manual subfield segmentation methodologies do not attempt to parse
+these regions even on high-resolution scans (\citep{Wisse2012, Mueller2009}; see
+also Table \label{tab:subfield-segmentation-protocols}). Note, it is because of
+both these reasons (the inability to discern subregions in standard T1-weighted
+images, and the definitional uncertainty in subfield segmentation at the present
+moment) that we opted to speak of the precision or reliability of \mb{} subfield
+segmentation rather than speak of accuracy.
+
+In conclusion, we have presented a flexible \multiatlas{}-based framework that has
 considerable advantages over other methods since only a small number of atlases

### 2013-12-23 paper.Rnw
Fixed up all plots

@@
 hippocampus in the ADNI-1 baseline dataset, achieving a mean Dice score of 0.85
-against \gs{} segmentations. %TODO: which segmentations were used? SNT?
+against semi-automated segmentations.
 
@@
 diminishing returns with larger atlas libraries. Peak improvement
-(+\Sexpr{round(snt.mb.diff.best, digits=3)} DSC) is found with a single atlas
+(+\Sexpr{round(mb.diff.best, digits=3)} DSC) is found with a single atlas
 and template library of 19 images.
@@
 
-%the CA1, CA4/DG, and SR/SL/SM
-%subfields all show a percent error in volume that is at or lower than
-%resampling error. The Subiculum and CA2/CA3 subfields show distinctly larger
-%percent error in volume than is found through resampling.
-
 Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb{} subfield
@@
 \begin{figure}[h!]
-<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
+<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=
 wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
@@
 <<WAval-bland,cache=T,fig.width=7,fig.height=6>>=
-wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
+wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm', 'gold_0.9mm'))
 wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
@@
 
-wa.xval.volumes$diff = wa.xval.volumes$gold_0.3mm - wa.xval.volumes$volume 
-wa.xval.volumes$mean = (wa.xval.volumes$gold_0.3mm + wa.xval.volumes$volume)/2
+wa.xval.volumes$diff = wa.xval.volumes$gold_0.9mm - wa.xval.volumes$volume 
+wa.xval.volumes$mean = (wa.xval.volumes$gold_0.9mm + wa.xval.volumes$volume)/2
 wa.xval.bland.limits = ddply(wa.xval.volumes,c("region"),function(df) {
@@
 
+adni.volumes.lm = ddply(melted,c("Method"), function (df) {
+  data.frame(eqn=lm_eqn(lm(volume ~ SNT, df)))})
+adni.volumes.lm$y = 6450-0:3*300
+
 ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
@@
     geom_smooth(method="lm") + 
-    xlab(expression("\\snt mean hippocampus volume " (mm^3))) + 
+    xlab(expression("SNT mean hippocampus volume " (mm^3))) + 
     ylab(expression("Automated mean hippocampal volume " (mm^3))) +
-    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
-              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
-                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
-                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
-                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
-              hjust=1.1,vjust=1.5) +
+    geom_text(aes(x=850,y=y,label=eqn),parse=T,data=adni.volumes.lm,
+      size=2.5,hjust=0) +
     theme(legend.direction = "horizontal", legend.position = "bottom")
+
+#    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
+#              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
+#                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
+#                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
+#                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
+#              hjust=1.1,vjust=1.5) +
 @
@@
 \hline
-LEAP                    &30&0.848                 &\cite{Wolz2010}  &Segmentation of 60 subjects\tabularnewline
+LEAP                    &30&0.848                 &\cite{Wolz2010}
+&Segmentation of 182 subjects\tabularnewline
 ACM (AdaBoost-based)    &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline

### 2013-12-23 paper.Rnw
subfield segmentation plots

@@
 %
-<<ADNI-xval-demographics, echo=F, dependson='ADNI1-xval-load', results="asis",cache=FALSE>>=
+<<tab:ADNI-xval-demographics, echo=F, dependson='ADNI1-xval-load', results="asis",cache=FALSE>>=
 tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
@@
 %                                     
-<<WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
+<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
 master = read.csv('data/DT_study.csv')
@@
 %
-<<FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
+<<tab:FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
 caption = 
@@
 %                                                                             
-<<ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
+<<tab:ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
 # RID < 2000 ensures only ADNI1 patients
@@
 \begin{figure}
-<<ADN1-Xval-variability,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=5,fig.height=5>>=
+<<fig:ADN1-Xval-variability,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=5,fig.height=5>>=
 ants = subset(jens_xval_data, 
@@
 <<WAval-prep, cache=T>>=
-wa.xval.vols = subset(read.csv('data/WAval.csv'))
-wa.xval.kappa = subset(read.csv('data/WAval-kappa.csv'))
+wa.xval.vols = na.omit(read.csv('data/WAval.csv'))
 wa.xval.casted = dcast(melt(wa.xval.vols, id.vars=c("file","version")), file + variable ~ version)
@@
                             "SR/SL/SM"="X6","SR/SL/SM"="X106")
-@
-This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
-this, a leave-one-out validation was conducted in which lower-resolution images
-($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
-remaining high-resolution Winterburn atlas subjects' images.
-
-In general, across hippocampal subfields the percent error in volume of \mb{}
-segmentations (relative to the full-resolution Winterburn atlas segmentation)
-compares favourably to the error resulting from image resampling (Figure
-\ref{fig:WAval-vol-boxplot}). In particular, there is no significant difference 
-between percent error of \mb{} volumes compared to image resampling error for the 
-CA1, Subiculum, and CA2/CA3 subregions. The \mb{} volumes of CA4/DG and SR/SL/SM 
-subregions show both significantly different (lower) percent error (p < 0.001).
-
-Figure \ref{fig:subfield-montage} shows a qualitative comparison of \mb{} subfield
-segmentations for a single subject.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Winterburn-XVAL: compare resampling
-\begin{figure}[h!]
-<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=
+
 wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
@@
 
+wa.s = ddply(wa.xval.error, c('region'), function(df) {
+  with(df,data.frame(
+    m_res = mean(resampled,na.rm=T), sd_res = sd(resampled,na.rm=T),
+    m_bravo = mean(mb_bravo,na.rm=T), sd_bravo = sd(mb_bravo,na.rm=T),
+    m_mb_res = mean(mb_0.9mm,na.rm=T), sd_mb_res = sd(mb_0.9mm,na.rm=T)))})
+@
+
+In this experiment the \mb{} segmentation of hippocampal subfields is assessed
+through a leave-one-out validation in which resampled images ($0.9mm^3$
+voxels) of each Winterburn atlas subject are segmented using the remaining
+high-resolution Winterburn atlas subjects' images.
+
+Figure \ref{fig:WAval-vol-boxplot}) shows the mean relative percent error in volume, 
+with respect to the full-resolution Winterburn atlas, of each hippocampal subfield
+of the resampled Winterburn labels, and of the \mb{} segmentations of the resampled
+Winterburn atlas images and T1 BRAVO Winterburn subject acquisitions.  A
+paired-samples t-test was conducted to compare the distribution of relative
+percent error in the \mb{} segmentation volumes and the resampled segmentation
+volumes. There was no significant difference found for the CA1, Subiculum, and
+CA2/CA3 subregions. The CA4/DG subregion shows significantly lower percent error 
+  ($M_{res}=$\Sexpr{wa.s$m_res[wa.s$region=='CA4/DG']},
+  $SD_{res}=$\Sexpr{wa.s$sd_res[wa.s$region=='CA4/DG']};
+  $M_{mb-bravo}=$\Sexpr{wa.s$m_bravo[wa.s$region=='CA4/DG']},    
+  $SD_{mb-bravo}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='CA4/DG']};    
+  $M_{mb-res}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='CA4/DG']},    
+  $SD_{mb-res}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='CA4/DG']};    
+  $p < 0.001$), and the SR/SL/SM shows significantly different percent error in the opposite
+direction 
+  ($M_{res}=$\Sexpr{wa.s$m_res[wa.s$region=='SR/SL/SM']},
+  $SD_{res}=$\Sexpr{wa.s$sd_res[wa.s$region=='SR/SL/SM']};
+  $M_{mb-bravo}=$\Sexpr{wa.s$m_bravo[wa.s$region=='SR/SL/SM']},    
+  $SD_{mb-bravo}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='SR/SL/SM']};    
+  $M_{mb-res}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='SR/SL/SM']},    
+  $SD_{mb-res}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='SR/SL/SM']};    
+  $p < 0.001$). Bland-Altman plots comparing \mb{} to subsampled segmentation
+volume reveals a proportional bias towards underestimating volumes in all
+subregions, and a noticeable fixed bias in the CA4/DG and SR/SL/SM subregions.
+
+Figure \ref{fig:subfield-montage} shows slices subfield segmentations for a
+single subject for qualitative inspection.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Winterburn-XVAL: compare resampling
+\begin{figure}[h!]
+<<fig:WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=
+
 wa.xval.error.melted = melt(wa.xval.error,id.vars=c("file", "region"))
@@
   geom_point(group=1, position="dodge") + 
-  labs(y="Unilateral subregion volume percent error", x="Subfield") +
+  labs(y="Percent error in volume", x="Subfield") +
   theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + 
@@
 @
-  \caption{{\bf Percent error in unilateral subfield volume.} 
+  \caption{{\bf Percent error in computed unilateral hippocampal subfield
+segmentation volume relative to manual segmentation volume.} 
   Percent error is measured against the volumes of the unmodified Winterburn
@@
   ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
-  subjects.} 
-  \label{fig:WAval-vol-boxplot}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Winterburn-XVAL: volume plots
-\begin{figure}[h!]
-<<WAval-vol-plots,cache=T,fig.width=7,fig.height=6>>=
-wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm'))
-wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
-levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
-                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
-                               "MAGeT on T1 BRAVO images"="mb_bravo")
-wa.xval.vols.lm = ddply(wa.xval.volumes, c("region", "measure"), function(df) {
-   model = lm(volume ~ gold_0.3mm, df)
-   pearson = cor(df$volume, df$gold_0.3mm, use='pairwise.complete.obs')
-   data.frame(
-     eqn  = lm_eqn(model), 
-     corr = pearson)})
-wa.xval.vols.lm$yoffset = c( 1225 - 0:2 * 40,
-                             465  - 0:2 * 15, 
-                             750  - 0:2 * 30,
-                             325  - 0:2 * 18, 
-                             1100 - 0:2 * 40 )
-wa.xval.vols.lm$xoffset = c( rep(700, 3),
-                             rep(250, 3),
-                             rep(490, 3),
-                             rep(140, 3), 
-                             rep(570, 3) )
-ggplot(wa.xval.volumes, aes(x=gold_0.3mm, y=volume, colour=measure)) + 
-  facet_wrap(~region, ncol=3, scales='free') + 
-  geom_point() + 
-  coord_equal() + 
-  geom_smooth(method='lm')  + 
-  xlab(expression("Manual Unilateral Volume" (mm^3))) +
-  ylab(expression("Automated Unilateral Volume" (mm^3))) +
-  geom_text(aes(x=xoffset, y=yoffset, label=eqn), parse=TRUE,
-    data=wa.xval.vols.lm, size=2,hjust=0) +
-  scale_color_discrete(name="") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-
-@
-  \label{fig:WAval-volumes}
+  subjects.\label{fig:WAval-vol-boxplot}}
 \end{figure}
@@
 \begin{figure}[h!]
-<<WAval-bland,cache=T,fig.width=7,fig.height=6>>=
+<<fig:WAval-bland,cache=T,fig.width=7,fig.height=6>>=
 wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm', 'gold_0.9mm'))
@@
              data = wa.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of manual and automated HC volume ' (mm^3))) + 
-  ylab(expression('Difference of manaul and automated HC volume ' (mm^3))) + 
+  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
+  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
   scale_color_discrete(name="") + 
@@
 @
-  \label{fig:WAval-bland}
+  \caption{{\bf Bland-Altman plots comparing unilateral \mb{} subfield
+segmentation volume with resampled subfield segmentation volume.} The overall
+mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
+dashed horizontal lines. Linear fit lines are shown for each target image type.
+Note, points below the mean difference indicate overestimation of the volume
+with respect to the resampled volume, and vice versa. 
+  \label{fig:WAval-bland}}
 \end{figure}
@@
 \begin{figure}
-<<FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
+<<fig:FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
 SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
@@
 @
-\mb{} produces hippocampal volumes that are highly correlated with 
-manual segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, 
-$t = \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$; 
+\mb{} produces hippocampal volumes that are highly correlated with manual
+segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, $t =
+\Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$;
 Figure \ref{fig:SZ-volumes}).
@@
 %                        
-<<ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
+<<tab:ADNI-seg-package-totals, echo=F, results="asis",cache=FALSE>>=
 package_totals$Total <- NULL
@@
 \begin{figure}[h]
-<<ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
+<<fig:ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
 melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
@@
 \begin{figure}[h]
-<<ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
 melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
@@
 \begin{figure}
-<<ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
+<<fig:ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
 names(means.complete)[8] = "FreeSurfer"
@@
 %
-<<ADNI-SNT-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
+<<tab:ADNI-SNT-Xval-demographics, echo=F, results="asis",cache=FALSE>>=
 adni_69_RIDs = c(295,413,619,685,729,782,938,954,1018,1155,907,981,222,324,448,
@@
 \begin{figure}
-<<ADN1-SNT-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=5, fig.width=5>>=
+<<fig:ADN1-SNT-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=5, fig.width=5>>=
 ants = subset(snt_xval_data, 

### 2013-12-24 paper.Rnw
FEP with FS/FIRST

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
+\usepackage{algorithm, color, algpseudocode, amsmath, lrl, geometry, ctable}
 \usepackage{titlesec, siunitx, graphicx}
@@
 \newcommand{\adnidataset}{ADNI1:Complete 1Yr 1.5T}
-\newcommand{\FEPdataset}{SZ-FEP}
 \newcommand{\fsl}{FSL FIRST}
@@
 library, formed from a subset of target images, via transformations estimated by
-non-linear image registration. The resultant segmentations are then propagated
+nonlinear image registration. The resultant segmentations are then propagated
 to each target image and fused using a label fusion method.  To explore the
@@
 Also of interest to this manuscript are methods that attempt to define
-hippocampal subfields using standard T1-weighted data.  To the best of our
+hippocampal subfields using standard T1- or T2-weighted data.  To the best of our
 knowledge, there are only two automated segmentation algorithms that attempt
@@
 segmentation using data from 10 subjects. In the second method,
-\citet{Yushkevich2009} hippocampal subfields were labelled on highly-anistropic
-($0.380mm$ in-plane resolution, $0.8mm$ coronal slice thickness) T2-weighted MR
-images acquired from post-mortem medial temporal lobe samples. Using nonlinear
-registration guided by manually derived hippocampus masks and specific
-landmarks, the authors demonstrate accurate parcellation of hippocampal
-subfields in unlabelled MRI volumes.
+\citet{Yushkevich2009} hippocampal subfields were manually labelled on
+high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
+resolution voxels) T2-weighted MR images acquired from five post-mortem medial
+temporal lobe samples. Using nonlinear registration guided by shape-based models
+of the subfield segmentations, and manually derived hippocampus masks of the
+target images, the authors demonstrate accurate parcellation of hippocampal
+subfields in clinical 3T T1-weighted MRI volumes. This method is not fully
+automatic as it requires whole hippocampal parcellation of the target images.
+%TODO: ref other Yush paper
 
 In this paper we have describe a thorough validation of the \mb{} algorithm for
-segmentation of the hippocampus and its subfields. First, we address the
-feasibility of whole hippocampus segmentation with a limited number of input
-atlases \citep{Chakravarty2013} by performing a multi-fold validation experiment using a
-subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI) dataset
-manually segmented using the \prsnr{}-protocol, over a
-range of atlas and template library sizes, registration and label fusion
-methods. We then perform a leave-one-out validation to determine if
-hippocampal subfields can be accurately identified, using the best parameters
-discovered in the first experiment.  To ensure that we have not overfit our
-parameters to the aging or neurodegenerative brain, we also apply \mb{} to a
-dataset of individuals suffering from first episode psychosis.  Finally, we
-validate our algorithm using all of the data available in the \adnidataset{}
-sample and compare our segmentations to other popular segmentation algorithms.
+the automatic segmentation of the hippocampus and its subfields. First, we
+address the feasibility of whole hippocampus segmentation with a limited number
+of input atlases \citep{Chakravarty2013} by performing a multi-fold validation
+experiment using a subset of the Alzheimer's Disease Neuroimaging Initiative
+(ADNI) dataset manually segmented using the \prsnr{}-protocol, over a range of
+atlas and template library sizes, registration and label fusion methods. We then
+perform a leave-one-out validation to determine if hippocampal subfields can be
+accurately identified, using the best parameters discovered in the first
+experiment.  To ensure that we have not overfit our parameters to the aging or
+neurodegenerative brain, we also apply \mb{} to a dataset of individuals
+suffering from first episode psychosis.  Finally, we validate our algorithm
+using all of the data available in the \adnidataset{} sample and compare our
+segmentations to other popular segmentation algorithms.
 
@@
 from cases and controls).  The template library images are automatically
-labelled by each of the atlases via label propagation.  Basic \multiatlas
-segmentation is then conducted using the template library to segment the entire
-set of target images (including the target images used in the construction of
-the template library).  Since each template library image has multiple labels
-(one from each atlas), the final number of labels to be fused for each target
-may be quite large (i.e. \# of atlas $\times$ \# of templates).
+labelled by each of the atlases via label propagation. Effectively, basic
+\multiatlas{} segmentation is then conducted using the template library to
+segment the entire set of target images (including the target images used in the
+construction of the template library). Since each template library image has
+multiple labels (one from each atlas), the final number of labels to be fused
+for each target may be quite large (i.e. \# of atlas $\times$ \# of templates).
 
@@
 are warped (propagated) into subject space by applying the transformations
-estimated from non-linear image registration. The resulting candidate labels
+estimated from nonlinear image registration. The resulting candidate labels
 from all atlas images are then fused to create a final segmentation. In \mb{}
@@
 
-The following section describes for experiments conducted to assess the
+The following section describes experiments conducted to assess the
 segmentation quality of the \mb{} algorithm. The first two experiments assess the
@@
 
-In the Supplementary Materials we have replicated this experiment using the \snt
-semi-automated labels included with the ADNI dataset.
+In the Supplementary Materials we have replicated this experiment using the
+\snt{} semi-automated labels included with the ADNI dataset.
 
@@
 Data used in the preparation of this article were obtained from the Alzheimer's
-Disease Neuroimaging Initiative (ADNI) database (adni.loni.ucla.edu). The ADNI
+Disease Neuroimaging Initiative (ADNI) database (adni.loni.usc.edu). The ADNI
 was launched in 2003 by the National Institute on Aging (NIA), the National
@@
 Medical Center and University of California San Francisco. ADNI is the result
-of efforts of many co- investigators from a broad range of academic institutions
+of efforts of many co-investigators from a broad range of academic institutions
 and private corporations, and subjects have been recruited from over 50 sites
@@
 available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\url{http://adni.loni.usc.edu/methods/mri-analysis/adni-standardized-data/}}
 \citep{Wyman2012}. This image collection contains uniformly pre-processed images
@@
 \citep{Sled1998} to minimize intensity nonuniformity. In this experiment we
-compared two non-linear image registration methods:
+compared two nonlinear image registration methods:
 
@@
   \begin{centering}
-    % TODO: fix schematic
     \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
@@
 disorders, in this experiment we use the Winterburn atlases to derive 
-whole hippocampal segmentations of a dataset of patients with Schizophrenia. The
+whole hippocampal segmentations of a dataset of patients with schizophrenia. The
 resulting segmentations are assessed for quality by comparison with expert
@@
 
-\paragraph{\FEPdataset{} dataset}
+\paragraph{First Episode Schizophrenia Dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 (see Experiment 2) ignoring subfield delineations.  All images from the
-\FEPdataset{} dataset are segmented by \mb{}.  The optimal size of template library,
-registration method, and label fusion method found in Experiment 1 are used.
+first episode dataset are segmented by \mb{}.  The optimal size of template
+library, registration method, and label fusion method found in Experiment 1 are
+used.
 
@@
 segmentation, \mb{} also shows more consistency in similarity scores across all
-subjects and validation folds (Figure \ref{ADNI1-xval-variability}). A
+subjects and validation folds (Figure \ref{fig:ADNI1-xval-variability}). A
 template library of at least 13 images is sufficient to show significant ($p <
@@
 
+We find similar behaviour with respect to optimal parameter settings and
+increased consistency of \mb{} segmentations in the replication of this
+experiment (Experiment 5, Supplementary Materials) where atlases use a different
+hippocampal definition is used. This strongly suggests that these results are
+independant of the segmentation protocol used and are, instead, features of the
+\mb{} algorithm.
+
 We have omitted results obtained when using an even number of atlases or
-templates since with this configuration we found significantly decreased
-performance. We believe this is as a result of an inherent bias in the majority
+templates since with these configurations we found significantly decreased
+performance. We believe this is results from an inherent bias in the majority
 vote fusion method used (see Discussion).
@@
 
+fmt_tstat = function(a, b) {
+  test = t.test(a,b, paired=T)
+  p = sprintf("%0.3f", max(0.001,round(test$p.value,3)))
+  t = sprintf("%0.3f", round(test$statistic,3))
+  paste(t," (",p,")")
+}
+
+wa.p = ddply(wa.xval.casted, c("region"), function (df) {
+  with(df, {
+       data.frame(
+         res = fmt_tstat(gold_0.3mm, gold_0.9mm),
+         bra = fmt_tstat(gold_0.3mm, mb_bravo),
+         mbr = fmt_tstat(gold_0.3mm, mb_0.9mm)
+       )})
+  })
+
+wa.p = rename(wa.p, c(
+  'region'='Subregion','res'='Resampled',
+  'bra'='MAGeT on BRAVO','mbr'='MAGeT on Resampled'))
+
 wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
@@
 
-Figure \ref{fig:WAval-vol-boxplot}) shows the mean relative percent error in volume, 
+Figure \ref{fig:WAval-vol-boxplot} shows the mean relative percent error in volume, 
 with respect to the full-resolution Winterburn atlas, of each hippocampal subfield
@@
 CA2/CA3 subregions. The CA4/DG subregion shows significantly lower percent error 
-  ($M_{res}=$\Sexpr{wa.s$m_res[wa.s$region=='CA4/DG']},
-  $SD_{res}=$\Sexpr{wa.s$sd_res[wa.s$region=='CA4/DG']};
-  $M_{mb-bravo}=$\Sexpr{wa.s$m_bravo[wa.s$region=='CA4/DG']},    
-  $SD_{mb-bravo}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='CA4/DG']};    
-  $M_{mb-res}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='CA4/DG']},    
-  $SD_{mb-res}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='CA4/DG']};    
+  ($M_{resampled}=$\Sexpr{wa.s$m_res[wa.s$region=='CA4/DG']},
+  $SD_{resampled}=$\Sexpr{wa.s$sd_res[wa.s$region=='CA4/DG']};
+  $M_{MAGeT(BRAVO)}=$\Sexpr{wa.s$m_bravo[wa.s$region=='CA4/DG']},    
+  $SD_{MAGeT(BRAVO)}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='CA4/DG']};    
+  $M_{MAGeT(resampled)}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='CA4/DG']},    
+  $SD_{MAGeT(resampled)}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='CA4/DG']};    
   $p < 0.001$), and the SR/SL/SM shows significantly different percent error in the opposite
 direction 
-  ($M_{res}=$\Sexpr{wa.s$m_res[wa.s$region=='SR/SL/SM']},
-  $SD_{res}=$\Sexpr{wa.s$sd_res[wa.s$region=='SR/SL/SM']};
-  $M_{mb-bravo}=$\Sexpr{wa.s$m_bravo[wa.s$region=='SR/SL/SM']},    
-  $SD_{mb-bravo}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='SR/SL/SM']};    
-  $M_{mb-res}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='SR/SL/SM']},    
-  $SD_{mb-res}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='SR/SL/SM']};    
+  ($M_{resampled}=$\Sexpr{wa.s$m_res[wa.s$region=='SR/SL/SM']},
+  $SD_{resampled}=$\Sexpr{wa.s$sd_res[wa.s$region=='SR/SL/SM']};
+  $M_{MAGeT(BRAVO)}=$\Sexpr{wa.s$m_bravo[wa.s$region=='SR/SL/SM']},    
+  $SD_{MAGeT(BRAVO)}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='SR/SL/SM']};    
+  $M_{MAGeT(resampled)}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='SR/SL/SM']},    
+  $SD_{MAGeT(resampled)}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='SR/SL/SM']};    
   $p < 0.001$). Bland-Altman plots comparing \mb{} to subsampled segmentation
 volume reveals a proportional bias towards underestimating volumes in all
-subregions, and a noticeable fixed bias in the CA4/DG and SR/SL/SM subregions.
+subregions, and prominant fixed biases in the volumes of the CA4/DG and SR/SL/SM subregions 
+congruent with significant differences seen in percent error in volume
+associated with subsampling alone.
 
@@
 single subject for qualitative inspection.  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% WAxval - t.test comparing Manual volumes and Computed volumes of subregions 
+%          across folds
+<<tab:WAxval-t-test-manual-computed, echo=F, results="asis",cache=FALSE>>=
+latex(wa.p, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
+    caption="\\textbf{Paired-sample t-tests comparing manually segmented
+             Winterburn atlas subregion volumes and computed volumes (reported
+             as $t (p)$). }", 
+    col.just=c("l","r","r","r"),
+    caption.loc = c("bottom"), rowname=NULL, 
+    label="tab:WAxval-t-test-manual-computed")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% % Winterburn-XVAL: boxplot showing raw volumes of all methods and all subjects
+%% \begin{figure}[h!]
+%% <<fig:WAval-volumes,cache=T,fig.width=6,fig.height=5>>=
+%% m=melt(wa.xval.casted, id.vars=c('region','file'))
+%% levels(m$variable) <- list("Original atlas images"="gold_0.3mm",
+%%                            "Subsampled atlas images"="gold_0.9mm",
+%%                            "MAGeT on subsampled atlas images" = "mb_0.9mm",
+%%                            "MAGeT on T1 BRAVO images"="mb_bravo")
+%% m = rename(m, c(
+%%   'region'='Subregion','value'='Volume',
+%%   'variable'='Method'))
+%% 
+%% ggplot(m,aes(y=Volume,x=Subregion,colour=Method))  + 
+%%   geom_boxplot() + 
+%%   geom_point(position=position_dodge(width=0.75)) +
+%%   ylab(expression('Volume ' (mm^3))) + 
+%%   theme(legend.direction = "horizontal", legend.position = "bottom")
+%% @
+%%   \caption{{\bf Manual and computed unilateral hippocampal subfield
+%% volumes.}\label{fig:WAval-volumes}}
+%% \end{figure}
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 <<fig:WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=
-
 wa.xval.error.melted = melt(wa.xval.error,id.vars=c("file", "region"))
@@
 <<fig:FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
-SZ_volumes=read.csv("data/SZ_volumes_by_method.csv", sep="\t")
-mean_sz_vols = ddply(SZ_volumes, c('Subject', 'Method'), function (df) {
-  data.frame(Volume = mean(df$Volume))})
-
-casted = dcast(mean_sz_vols, Subject ~ Method, value.var = "Volume")
-lm = lm(MAGeT ~ Manual, casted)
-ggplot(data=casted, aes(x = Manual, y = MAGeT)) + 
+fep.raw.volumes=read.csv('data/SZ_vol_by_method.csv')
+fep.raw.volumes=subset(fep.raw.volumes, SID <= 81)
+fep.raw.volumes=melt(fep.raw.volumes, id.vars=c('SID','method')) 
+fep.volumes = dcast(fep.raw.volumes, SID~method, value.var="value", fun.aggregate=mean)
+fep.volumes = melt(fep.volumes, id.vars=c('SID', 'manual'), variable.name='method', value.name='volume')
+
+fep.lm = ddply(fep.volumes,c("method"), function (df) {
+  data.frame(eqn=lm_eqn(lm(manual ~ volume, df)))})
+fep.lm$y = 6800-0:3*150
+
+ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=method)) + 
   geom_smooth(method="lm", formula=y~x) + 
   geom_point() + 
-  geom_text(aes(x=3500, y=5000, label=lm_eqn(lm),hjust=0,size=1), parse=TRUE, data=data.frame()) +
+  geom_text(aes(x=3350,y=y,label=eqn),parse=T,data=fep.lm,
+      size=3,hjust=0) +
   xlab(expression("Mean bilateral manual hippocampus volume " (mm^3))) + 
-  ylab(expression("Mean bilateral MAGeT hippocampus volume "  (mm^3))) + 
+  ylab(expression("Mean bilateral computed hippocampus volume "  (mm^3))) + 
   scale_size_continuous(guide="none") 
@@
 
+\begin{figure}
+<<fig:FEP-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
+fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
+fep.volumes$mean = (fep.volumes$manual + fep.volumes$volume)/2
+limits = ddply(fep.volumes, c("method"), function (df) { 
+  data.frame(
+    y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T) 
+)})
+
+ggplot(fep.volumes, aes(x= mean, y = diff)) + 
+  facet_wrap( ~ method, nrow=2, scales="fixed") + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=1) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", data = limits,
+            hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of SNT and automated hippocampal volume ' (mm^3))) + 
+  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
+  scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
+  scale_x_continuous(breaks=seq(0,10000, 1000)) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+@
+  \caption{Bland-Altman plots comparing subject mean hippocampal volume as
+measured in the FEP dataset by \snt{} segmentation and each of the
+four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
+mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
+dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
+Note, points below the mean difference indicate overestimation of the volume
+with respect to the \snt{} volume, and vice versa. }
+  \label{fig:FEP-Bland-Altman}
+\end{figure}
+
 <<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
-  fep.cor.test = cor.test(casted$Manual, casted$MAGeT)
+  fep.cor.test = cor.test(fep.volumes$manual,fep.volumes$MAGeT_Winterburn)
 @
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[h]
+<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
+            variable.name = "Method", value.name = "volume")
+
+df = means.complete
+
+levels(melted$Method) = c("FS"    = "FreeSurfer", 
+                          "FSL"   = "FSL", 
+                          "MAPER" = "MAPER", 
+                          "MAGeT" = "MAGeT")
+
+adni.volumes.lm = ddply(melted,c("Method"), function (df) {
+  data.frame(eqn=lm_eqn(lm(volume ~ SNT, df)))})
+adni.volumes.lm$y = 6450-0:3*300
+
+ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
+    geom_point(size=1) + 
+    geom_smooth(method="lm") + 
+    xlab(expression("SNT mean hippocampus volume " (mm^3))) + 
+    ylab(expression("Automated mean hippocampal volume " (mm^3))) +
+    geom_text(aes(x=850,y=y,label=eqn),parse=T,data=adni.volumes.lm,
+      size=2.5,hjust=0) +
+    theme(legend.direction = "horizontal", legend.position = "bottom")
+
+#    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
+#              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
+#                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
+#                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
+#                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
+#              hjust=1.1,vjust=1.5) +
+@
+  \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
+dataset by each of the four automated methods investigated (\freesurfer{} (FS), \fsl{},
+\maper{}, \mb{}) vs. \snt{}. Linear fit lines and Pearson correlations are shown
+for each method.}
+  \label{fig:ADNI-volumes-plot}
+\end{figure}
 
@@
   \label{fig:ADNI-volumes-boxplot}
-\end{figure}
-
-\begin{figure}[h]
-<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
-melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
-            variable.name = "Method", value.name = "volume")
-
-df = means.complete
-
-levels(melted$Method) = c("FS"    = "FreeSurfer", 
-                          "FSL"   = "FSL", 
-                          "MAPER" = "MAPER", 
-                          "MAGeT" = "MAGeT")
-
-adni.volumes.lm = ddply(melted,c("Method"), function (df) {
-  data.frame(eqn=lm_eqn(lm(volume ~ SNT, df)))})
-adni.volumes.lm$y = 6450-0:3*300
-
-ggplot(data=melted, aes(x=SNT, y=volume, colour=Method)) + 
-    geom_point(size=1) + 
-    geom_smooth(method="lm") + 
-    xlab(expression("SNT mean hippocampus volume " (mm^3))) + 
-    ylab(expression("Automated mean hippocampal volume " (mm^3))) +
-    geom_text(aes(x=850,y=y,label=eqn),parse=T,data=adni.volumes.lm,
-      size=2.5,hjust=0) +
-    theme(legend.direction = "horizontal", legend.position = "bottom")
-
-#    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
-#              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
-#                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
-#                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
-#                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
-#              hjust=1.1,vjust=1.5) +
-@
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
-dataset by each of the four automated methods investigated (\freesurfer{} (FS), \fsl{},
-\maper{}, \mb{}) vs. \snt{}. Linear fit lines and Pearson correlations are shown
-for each method.}
-  \label{fig:ADNI-volumes-plot}
 \end{figure}
@@
 \hline
-LEAP                    &30&0.848                 &\cite{Wolz2010}
-&Segmentation of 182 subjects\tabularnewline
+\mb{}                        &9 & \Sexpr{round(mb.kappa.best, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
+LEAP                         &30&0.848                 &\cite{Wolz2010}  &Segmentation of 182 subjects\tabularnewline
 ACM (AdaBoost-based)         &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
-\mb{}                   &9 & \Sexpr{round(mb.kappa.best, digits=3)} & & 10
-rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
-Patch-based label fusion&16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
+Patch-based label fusion     &16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
 Multi-atlas                  &30&0.885                 &\cite{Lotjonen2010}&Segmentation of 60 subjects\tabularnewline
@@
 a template library from a small set of labelled atlas images, is validated in
-the cross validation conducted in Experiment 1.  We find that both increasing
-the number of atlases and the number of templates used improves \mb{}
-segmentation over and above basic-\multiatlas segmentations using the same
-number of atlas images. That is, by taking the extra step of generating a
-template library from a small atlas set \mb{} is able to improve the final
-segmentation similarity. The magnitude of this improvement is greatest with a
-small number of atlases, but even with larger atlas libraries we have found that
-generating a template library reduces the variability in segmentations (i.e.
-\mb{} more consistently produces high quality segmentations than does
-basic-\multiatlas).
+the cross validation conducted in Experiment 1 (and the replication in
+Experiment 5, Supplementary Materials). We find that both increasing the number
+of atlases and the number of templates used improves \mb{} segmentation over and
+above basic-\multiatlas segmentations using the same number of atlas images.
+That is, by taking the extra step of generating a template library from a small
+atlas set \mb{} is able to improve the final segmentation similarity. The
+magnitude of this improvement is greatest with a small number of atlases, but
+even with larger atlas libraries we have found that generating a template
+library reduces the variability in segmentation precision (i.e.  \mb{} more
+consistently produces high quality segmentations than does basic-\multiatlas{}
+segmentation). These effects do not appear dependant on the hippocampal
+segmentation protocol used.
 
@@
 
-On that note, one author (JW), an expert manual rater, identified regular
-inconsistencies in the \snt{} segmentations: occurrences of over- and
-under-estimation, as well as misalignments of the entire segmentation volume
-(Figure \ref{fig:ADNI-segmentations}). Although the \snt{} segmentations are
-used as benchmarks for validation in many other studies (Table
+On that note, one author (JW), an expert manual rater (citep{Winterburn2013),
+identified regular inconsistencies in the \snt{} segmentations: occurrences of
+over- and under-estimation, as well as misalignments of the entire segmentation
+volume (Figure \ref{fig:ADNI-segmentations}). Although the \snt{} segmentations
+are used as benchmarks for validation in many other studies (Table
 \ref{tab:other-methods}), these segmentation inconsistencies present the
@@
 Indeed, our replication of the 10-fold cross-validation using \snt{}
-segmentations (Experiment 5) shows noticeably poorer mean similarity scores for
-both \mb{} and \multiatlas.
+segmentations (Experiment 5, Supplementary Materials) shows noticeably poorer
+mean similarity scores for both \mb{} and \multiatlas.
 
@@
                           radiatum/lacunosum/moleculare, subiculum \tabularnewline
-\citet{Wisse2012}      &  CA1,  CA2,  CA3,  CA4/DG, subiculum,  entorhinal
-                          cortex \tabularnewline
-\citet{VanLeemput2009} &  CA1,  CA2/CA3,  CA4/DG, presubiculum, subiculum,
+\citet{Wisse2012}      &  CA1,  CA2,  CA3,  CA4/dentate gyrus, subiculum,
+                          entorhinal cortex \tabularnewline
+\citet{VanLeemput2009} &  CA1,  CA2/CA3,  CA4/dentate gyrus, presubiculum, subiculum,
                           hippocampal fissure,  fimbria,  hippocampal tail,
@@
                           sulcus \tabularnewline
+\citet{Mueller2007}    &  CA1, CA2, CA3/CA4 \& dentate gyrus, Sibiculum,
+                          entorhinal cortex \tabularnewline
 \hline
@@
 subfield segmentations produced in \citet{Yushkevich2009}).  Our work
-demonstrates that we can reliably produce CA1, subiculum, and CA4/dentate
-subfield segmentations that correspond to those in the input atlas with only 
-only modest amounts of error in volume.  The fact that CA2/CA3 and molecular layers
-cannot be reasonably reproduced should not be surprising as these are extremely
-thin and spatially convoluted regions that originally required high-resolution
-MRI for identification and so it is likely that the extents of these regions are
-well below the resolution and contrast offered by standard T1-weighted images.
-In addition, some manual subfield segmentation methodologies do not attempt to parse
-these regions even on high-resolution scans (\citep{Wisse2012, Mueller2009}; see
-also Table \label{tab:subfield-segmentation-protocols}). Note, it is because of
-both these reasons (the inability to discern subregions in standard T1-weighted
-images, and the definitional uncertainty in subfield segmentation at the present
-moment) that we opted to speak of the precision or reliability of \mb{} subfield
-segmentation rather than speak of accuracy.
+demonstrates that we can reliably reproduce segmentations, corresponding in volumes 
+to the input atlases used, with only modest error for the CA1, subiculum, and 
+CA4/dentate subregions.
+
+The fact that CA2/CA3 and molecular layers cannot be reasonably reproduced
+should not be surprising as these are extremely thin and spatially convoluted
+regions that originally required high-resolution MRI for identification and so
+it is likely that the extents of these regions are well below the resolution and
+contrast offered by standard T1-weighted images. In addition, some manual
+subfield segmentation methodologies do not attempt to parse these regions even
+on high-resolution scans. See Table \ref{ab:subfield-segmentation-protocols} for
+a comparison of manual subfield segmentation methodologies. Further complicating
+this, is that that different research groups have different operational
+definitions for these different subfields. The disagreement in the community has
+led to an international working group devoted to normalizing the ontology and
+segmentation rules for the hippocampal subfields
+(\url{http://www.hippocampalsubfields.com/}). The definitional and operational
+disagreements make evaluating the accuracy of automated methodologies with
+respect to a "ground truth" (e.g. using overlap similarity metrics, such as
+Dice's Kappa) not particularly meaningful until these disagreements have been
+resolved.  Thus, because of this, and the inability to discern subregions in
+standard T1-weighted images, we opted to speak of the precision or reliability
+of \mb{} subfield segmentation rather than speak of accuracy.
 

### 2013-12-28 paper.Rnw
FEP Bland-altman and volume comparison charts

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{algorithm, color, algpseudocode, amsmath, lrl, geometry, ctable}
+\usepackage{algorithm, color, algpseudocode, url, amsmath, geometry, ctable}
 \usepackage{titlesec, siunitx, graphicx}
@@
       size=3,hjust=0) +
-  xlab(expression("Mean bilateral manual hippocampus volume " (mm^3))) + 
-  ylab(expression("Mean bilateral computed hippocampus volume "  (mm^3))) + 
-  scale_size_continuous(guide="none") 
-@
-  \caption{Mean bilateral hippocampus volume as measured by \mb{}  vs. manually
-  segmented volumes from the First Episode Patients with Schizophrenia dataset.
-  A linear fit line is shown, with shaded region showing standard error. 
+  xlab(expression("Mean manual hippocampus volume " (mm^3))) + 
+  ylab(expression("Mean computed hippocampus volume "  (mm^3))) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{Mean bilateral hippocampus volume as measured by \mb{}, \fsl{}, and
+\freesurfer{} vs. manually segmented volumes from the First Episode Patients with
+Schizophrenia dataset. \mb{} is run two ways: with an atlas library consistening
+of the five Winterburn atlases, or with five manually segmented FEP subjects as
+atlases (not included as subjects). Linear fit lines are shown, with shaded
+region showing standard error. 
   \label{fig:SZ-volumes}}
@@
 \begin{figure}
-<<fig:FEP-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
+<<fig:FEP-Bland-Altman, cache=TRUE, dependson='fig:FEP-volumes', fig.width=7, fig.height=7>>=
 fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
@@
             hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of SNT and automated hippocampal volume ' (mm^3))) + 
-  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
+  xlab(expression('Mean manual and computed volume ' (mm^3))) + 
+  ylab(expression('Difference in manual and computed volume ' (mm^3))) + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
@@
   \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the FEP dataset by \snt{} segmentation and each of the
-four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
+measured in the FEP dataset by manual segmentation and each of the
+four automated methods investigated (\freesurfer{}, \fsl{}, \mb{} with
+Winterburn atlases, \mb{} with FEP manual atlases). The overall
 mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
@@
 Note, points below the mean difference indicate overestimation of the volume
-with respect to the \snt{} volume, and vice versa. }
+with respect to the manual volume, and vice versa. }
   \label{fig:FEP-Bland-Altman}
@@
 
-<<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
-  fep.cor.test = cor.test(fep.volumes$manual,fep.volumes$MAGeT_Winterburn)
-@
-\mb{} produces hippocampal volumes that are highly correlated with manual
-segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, $t =
-\Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$;
-Figure \ref{fig:SZ-volumes}).
+% <<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
+%   fep.cor.test = cor.test(fep.volumes$manual,fep.volumes$MAGeT_Winterburn)
+% @
+% \mb{} produces hippocampal volumes that are highly correlated with manual
+% segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, $t =
+% \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$;
+% Figure \ref{fig:SZ-volumes}). 
+%
 

### 2013-12-29 paper.Rnw
Added latest FEP xval results

@@
 jens_xval_mean$method.mb = factor(jens_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
-mb.kappa.best = mean(subset(jens_xval_mean, 
+mb.kappa.best = round(mean(subset(jens_xval_mean, 
                                  reg_method   == "ANTS" & 
@@
                                  templates.mb == 19 & 
-                                 atlases      == 9)$k.mb)
+                                 atlases      == 9)$k.mb), digits=3)
 mb.diff.best = mean(subset(jens_xval_mean, 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: prep
+% Prepares the data in a the form needed for plotting.
+<<ADNI1-SNT-xval-prep,include=FALSE,cache=TRUE>>=
+snt_xval_data  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
+snt_xval_mean  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data_mean.csv.gz'))
+snt_xval_mean$reg_method = factor(snt_xval_mean$reg_method, c("ANTS", "ANIMAL"))
+snt_xval_mean$method.mb = factor(snt_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
+
+snt.mb.kappa.best =round(mean(subset(snt_xval_mean, 
+                                 reg_method   == "ANTS" & 
+                                 method.mb    == "Majority Vote" &
+                                 templates.mb == 19 & 
+                                 atlases      == 9)$k.mb), digits=3)
+snt.mb.diff.best = mean(subset(snt_xval_mean, 
+                                 reg_method   == "ANTS" & 
+                                 method.mb    == "Majority Vote" &
+                                 templates.mb == 19 & 
+                                 atlases      == 1)$k_diff)
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \documentclass{article}
 \usepackage[margin=1in]{geometry}
-\usepackage{algorithm, color, algpseudocode, url, amsmath, geometry, ctable}
+\usepackage{color, url, amsmath, geometry, ctable}
 \usepackage{titlesec, siunitx, graphicx}
@@
 \usepackage[hypcap]{caption}
-\usepackage{lineno}
-
-%draft mode
-\usepackage[colorinlistoftodos, textwidth=4cm, shadow]{todonotes}
-\usepackage[displaymath, tightpage]{preview}
+\usepackage{lineno}              % line numbers
+\usepackage{multirow,rotating}   % vertical text in tables
+\usepackage[displaymath, tightpage]{preview} %draft mode
 \setlength{\marginparwidth}{1.2in}
@@
 \textbf{Results:} Using 9 atlases and 19 template images, \mb{} achieves a
-mean Dice's Similarity Coefficient (DSC) of \Sexpr{round(mb.kappa.best, digits=3)} 
+mean Dice's Similarity Coefficient (DSC) of \Sexpr{mb.kappa.best} 
 (to \psegs{}) over 10-folds of Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
@@
 an eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}).  A maximum
-similarity score of \Sexpr{round(mb.kappa.best, digits=3)} is found when using 9
+similarity score of \Sexpr{mb.kappa.best} is found when using 9
 atlases, 19 templates, ANTS registration, and majority vote label fusion).  The
@@
   \psegs{} for 60 ADNI subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method. Error bars indicate
-  standard error.
+  size, registration algorithm, and label fusion method over 10 folds of Monte Carlo       
+  cross validation. 
+  Error bars indicate standard error.
   \label{fig:ADNI1-xval-k-mean}}
@@
   \multiatlas{} segmentations vs. atlas and template library size when using the
-  \ants{} registration method, and majority-vote label fusion.
+  \ants{} registration method, and majority-vote label fusion over 10 folds of Monte 
+  Carlo cross validation. . 
   Segmentation similarity is computed against \psegs{}.  
@@
   Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
-  is computed for each subject across all ten rounds of validation. 
+  is computed for each subject across over 10 folds of Monte Carlo       
+  cross validation. . 
   Shown on the y-axis is the  p-value resulting from a t-test comparing the
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
+\begin{figure}
+<<fig:FEP-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
+fep.xval <- subset(read.csv('data/results-fep-xval-ants-2013-12-28-020512.csv'),
+                   k > 0.5)
+fep.xval$atlases <- as.factor(fep.xval$atlases)
+fep.xval.mean <- ddply(fep.xval, .(subject,atlases,templates,label), function (df) {
+  data.frame(k=mean(df$k),volume=mean(df$volume))
+})
+stderr <- function(x) sqrt(var(x)/length(x))
+ggplot(fep.xval.mean, aes(x=templates,y=k, colour=atlases)) + 
+  stat_summary(fun.y=mean,geom='line', aes(y=k,weight=1,group=atlases)) + 
+  stat_summary(fun.y=mean,geom='point',aes(y=k,weight=1,group=atlases)) + 
+  stat_summary(fun.y=mean,
+               fun.ymin=function (x) mean(x) - stderr(x),
+               fun.ymax=function (x) mean(x) + stderr(x), 
+               geom='errorbar',aes(y=k, weight=1, group=atlases), 
+               linetype='solid', width=0.5) + 
+  #stat_smooth(method='lm', formula=y~log(x)) + 
+  scale_y_continuous(breaks=seq(0,1,0.01)) + 
+  scale_colour_hue(name="Number of Atlases") +
+  scale_linetype(name="Registration Method") +
+  xlab( "Number of Templates" ) + 
+  ylab( "Mean similarity (DSC)" ) + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text = element_text(size = 8))
+@
+  \caption{Mean Dice's Similarity Coefficient of \mb{} unilateral hippocampal 
+  segmentations relative to
+  \psegs{} for 81 FEP subjects vs. atlas and template library
+  size using \ants registration algorithm and majority vote label fusion method 
+  over 3 folds of Monte Carlo cross validation. 
+  Error bars indicate standard error.
+  \label{fig:FEP-xval-k-mean}}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}
@@
 \freesurfer{} vs. manually segmented volumes from the First Episode Patients with
-Schizophrenia dataset. \mb{} is run two ways: with an atlas library consistening
+Schizophrenia dataset. \mb{} is run two ways: with an atlas library consisting
 of the five Winterburn atlases, or with five manually segmented FEP subjects as
@@
 %
-\begin{table}[!tbp]
+\begin{sidewaystable}[!tbp]
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy (overlap with \snt{} labels)
-of the ADNI dataset.} For each method, the number of labelled atlases
-used for training, the best Dice's overlap measure, the disease classes
-measured, and the validation procedure are shown. Unless specified, validation
-datasets are composed equally of subjects diagnosed with Alzheimer's Disease
-(AD), Mild Cognitive Impairment (MCI), and Cognitively Normal (CN). LOOCV =
-Leave-one-out cross-validation. 
-   Some studies of automated segmentation of ADNI images are excluded because they 
-   do not not provide overlap measures for the hippocampus \citep{Heckemann2011,Chupin2009}.
+\caption{\textbf{Automated segmentation accuracy of the Hippocampus} 
+This table shows the best published Dice's overlap measure between automated and
+ground truth segmentations of various methods.  Unless otherwise specified,
+validation datasets are composed equally of cases and control subjects, and use
+manual segmenation labels as ground truth in computing DSC scores. AD =
+Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
+(CN); LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
+cross-validation; SNT = Surgical Medtronic Navigation Technologies
+semi-automated labels.  Some studies of automated segmentation of ADNI images
+are excluded because they do not not provide overlap measures for the
+hippocampus \citep{Heckemann2011,Chupin2009}.
   \label{tab:other-methods}} 
 \begin{center}
-\begin{tabular}{ l  c  p{1in}  l  p{2in}  }
+\begin{tabular}{ p{1.2in}  c  p{1.1in} l  p{3.2in} p{1in} }
 \hline\hline
-Method & Atlases & DSC & Reference & Validation \tabularnewline
+Method & Atlases & DSC \newline {\tiny mean (AD; MCI; CN)} & Reference & Validation & Dataset (Truth)\tabularnewline
 \hline
-\mb{}                        &9 & \Sexpr{round(mb.kappa.best, digits=3)} & & 10 rounds of Monte Carlo CV on a pool of 69 subjects \tabularnewline % TODO: update
-LEAP                         &30&0.848                 &\cite{Wolz2010}  &Segmentation of 182 subjects\tabularnewline
-ACM (AdaBoost-based)         &21&0.862                 &\cite{Morra2008} &LOOCV on atlases\tabularnewline
-Patch-based label fusion     &16&0.883 (CN) \newline 0.838 (AD)&\cite{Coupe2011e}&LOOCV on atlases\tabularnewline
-Multi-atlas                  &30&0.885                 &\cite{Lotjonen2010}&Segmentation of 60 subjects\tabularnewline
-Multi-atlas + weighted fusion&20&0.898 (CN) \newline 0.798 (left HC, MCI) &\cite{Wang2011}  &10 rounds of Monte Carlo CV on 20 subjects, pool of 139 (CN/MCI)\tabularnewline
-Multi-atlas (MAPS)           &55&0.890                   &\cite{Leung2010}&Segmentation of 30 subjects (10 AD, MCI, and CN)\tabularnewline
+%%  
+\mb{}                     &9 & \Sexpr{snt.mb.kappa.best} & \cite{Pipitone2014}
+& 10-fold MCCV on 69 subjects 
+& ADNI (SNT) 
+\tabularnewline 
+%%  
+Patch-based label fusion  &16&0.861 {\tiny (0.838; ---; 0.883)} &\cite{Coupe2011e}
+& LOOCV 
+& ADNI (SNT) 
+\tabularnewline
+%%  
+Multi-atlas               &20&0.848 {\tiny (---; 0.798, 0.898)} &\cite{Wang2011}  
+& 10-fold MCCV on 20 of 139 subjects
+& ADNI (SNT) 
+\tabularnewline
+%%  
+ACM (AdaBoost-based)      &21&0.862                           &\cite{Morra2008} 
+& LOOCV                                          
+& ADNI (SNT)                                           
+\tabularnewline                                  
+%%                                               
+LEAP                      &30&0.848                           &\cite{Wolz2010}  
+& Segmentation of 182 subjects                   
+& ADNI (SNT)                                           
+\tabularnewline                                  
+%%                                               
+Multi-atlas               &30&0.885                           &\cite{Lotjonen2010}
+& Segmentation of 60 subjects                    
+& ADNI (SNT)                                           
+\tabularnewline                                  
+%%                                               
+Multi-atlas (MAPS)        &55&0.890                           &\cite{Leung2010}
+& Segmentation of 30 subjects (10 AD, MCI, and CN)
+& ADNI (SNT) 
+\tabularnewline
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hline
+%& & {\tiny mean (right; left)} & & & \tabularnewline
+%\hline
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%
+\mb{}                      &9 & \Sexpr{mb.kappa.best}       &\cite{Pipitone2014} 
+& 10-fold MCCV on 60 subjects 
+& ADNI (\prsnr)
+\tabularnewline 
+%%
+Neural nets                &10&0.740   &\cite{Powell2008} % {\tiny (0.72, 0.75)} 
+& Segmentation of 5 subjects 
+& controls %Other
+\tabularnewline
+%%
+Probablistic atlas         &11&0.852   &\cite{VanderLijn2008}
+& 11 atlases used in 100 rounds of LOOCV on 20 elderly subjects
+& elderly controls %Other
+\tabularnewline
+%%
+Probabilistic Atlas        &16&0.860   &\cite{Chupin2009}
+& LOOCV 
+& AD subjects %Other
+\tabularnewline
+%%
+Anatomically-guided EM     &17&0.812   &\cite{Pohl2007} %{\tiny (0.815, 0.808)} 
+& LOOCV on 17 controls, segmentation of 33 mixed subjects
+& mixed diagnosis %Other
+\tabularnewline
+%%
+Multi-atlas                &30&0.820   &\cite{Heckemann2006} %{\tiny (0.83, 0.81)}    
+& LOOCV
+& controls %Other
+\tabularnewline
+%%
+Multi-atlas                &30&0.880   &\cite{Gousias2008} %{\tiny (0.88,---)}    
+& 30 adult atlas used, segmentation of 33 2yr old subjects 
+& 2yr old controls  %Other
+\tabularnewline
+%%
+Multi-atlas                &80&0.890   &\cite{Collins2010}
+& LOOCV                 
+& controls %Other       
+\tabularnewline
+%%                      
+Multi-atlas                &55&0.860   &\cite{Barnes2008}%{\tiny (---,0.86)}    
+& LOOCV 
+& controls and AD %Other
+\tabularnewline
+%%
+Multi-atlas               &275&0.835   &\cite{Aljabar2009}% {\tiny (0.832, 0.837)} 
+& LOOCV 
+& controls %Other
+\tabularnewline
 \hline
@@
 \end{center}
-\end{table}
+\end{sidewaystable}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: prep
-% Prepares the data in a the form needed for plotting.
-<<ADNI1-SNT-xval-prep,include=FALSE,cache=TRUE>>=
-snt_xval_data  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
-snt_xval_mean  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data_mean.csv.gz'))
-snt_xval_mean$reg_method = factor(snt_xval_mean$reg_method, c("ANTS", "ANIMAL"))
-snt_xval_mean$method.mb = factor(snt_xval_mean$method.mb, c("Majority Vote", "Cross-correlation Vote", "NMI Vote"))
-
-snt.mb.kappa.best = mean(subset(snt_xval_mean, 
-                                 reg_method   == "ANTS" & 
-                                 method.mb    == "Majority Vote" &
-                                 templates.mb == 19 & 
-                                 atlases      == 9)$k.mb)
-snt.mb.diff.best = mean(subset(snt_xval_mean, 
-                                 reg_method   == "ANTS" & 
-                                 method.mb    == "Majority Vote" &
-                                 templates.mb == 19 & 
-                                 atlases      == 1)$k_diff)
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 As when comparing against manual labels in Experiment 1, we find similar
@@
 difference in scores from any of the label fusion methods.  Mean DSC score peaks
-at \Sexpr{round(snt.mb.kappa.best, digits=3)} when using 9 atlases, 19
+at \Sexpr{snt.mb.kappa.best} when using 9 atlases, 19
 templates, ANTS registration, and majority vote label fusion).  Compared to

### 2013-12-29 paper.Rnw
Added latest FEP xval results

@@
 <<fig:FEP-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
-fep.xval <- subset(read.csv('data/results-fep-xval-ants-2013-12-28-020512.csv'),
-                   k > 0.5)
+fep.xval <-subset(read.csv('data/results-fep-xval-ants-2013-12-28-223825.csv'), k > 0.5)
 fep.xval$atlases <- as.factor(fep.xval$atlases)

### 2013-12-30 paper.Rnw
Updated FEP figures; updated references

@@
 \usepackage[hypcap]{caption}
+\usepackage{subcaption}
+\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}
 \usepackage{lineno}              % line numbers
@@
 \begin{figure}
-<<fig:ADNI1-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
+<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load'>>=
 stderr <- function(x) sqrt(var(x)/length(x))
@@
 \begin{figure}
-<<fig:ADN11-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-load'>>=
+<<fig:ADN11-xval-k-diff,dependson='ADNI1-xval-load'>>=
 ggplot(subset(jens_xval_mean, reg_method=="ANTS" & 
@@
 \begin{figure}
-<<fig:ADN1-Xval-variability,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=5,fig.height=5>>=
+<<fig:ADN1-Xval-variability,dependson='ADNI1-Xval-prep'>>=
 ants = subset(jens_xval_data, 
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-<<WAval-prep, cache=T>>=
+<<WAval-prep>>=
 wa.xval.vols = na.omit(read.csv('data/WAval.csv'))
@@
 
-%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% % Winterburn-XVAL: boxplot showing raw volumes of all methods and all subjects
-%% \begin{figure}[h!]
-%% <<fig:WAval-volumes,cache=T,fig.width=6,fig.height=5>>=
-%% m=melt(wa.xval.casted, id.vars=c('region','file'))
-%% levels(m$variable) <- list("Original atlas images"="gold_0.3mm",
-%%                            "Subsampled atlas images"="gold_0.9mm",
-%%                            "MAGeT on subsampled atlas images" = "mb_0.9mm",
-%%                            "MAGeT on T1 BRAVO images"="mb_bravo")
-%% m = rename(m, c(
-%%   'region'='Subregion','value'='Volume',
-%%   'variable'='Method'))
-%% 
-%% ggplot(m,aes(y=Volume,x=Subregion,colour=Method))  + 
-%%   geom_boxplot() + 
-%%   geom_point(position=position_dodge(width=0.75)) +
-%%   ylab(expression('Volume ' (mm^3))) + 
-%%   theme(legend.direction = "horizontal", legend.position = "bottom")
-%% @
-%%   \caption{{\bf Manual and computed unilateral hippocampal subfield
-%% volumes.}\label{fig:WAval-volumes}}
-%% \end{figure}
-%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[h!]
-<<fig:WAval-vol-boxplot,cache=T,fig.width=6,fig.height=5>>=
+<<fig:WAval-vol-boxplot>>=
 wa.xval.error.melted = melt(wa.xval.error,id.vars=c("file", "region"))
@@
 \begin{figure}[h!]
-<<fig:WAval-bland,cache=T,fig.width=7,fig.height=6>>=
+<<fig:WAval-bland>>=
 wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm', 'gold_0.9mm'))
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
-\begin{figure}
-<<fig:FEP-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=7, fig.height=7>>=
-fep.xval <-subset(read.csv('data/results-fep-xval-ants-2013-12-28-223825.csv'), k > 0.5)
+% FEP-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
+<<fig:FEP-xval,fig.show='hide'>>=
+  fep.xval <-read.csv('data/results-fep-xval-ants-2013-12-29-191914.csv')
   fep.xval$atlases <- as.factor(fep.xval$atlases)
@@
     ylab( "Mean similarity (DSC)" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text = element_text(size = 8))
-@
-  \caption{Mean Dice's Similarity Coefficient of \mb{} unilateral hippocampal 
-  segmentations relative to
-  \psegs{} for 81 FEP subjects vs. atlas and template library
-  size using \ants registration algorithm and majority vote label fusion method 
-  over 3 folds of Monte Carlo cross validation. 
-  Error bars indicate standard error.
-  \label{fig:FEP-xval-k-mean}}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}
-<<fig:FEP-volumes,cache=FALSE,dependson="setup",fig.width=5,fig.height=5>>=
+    theme(legend.direction = "horizontal", legend.position = "bottom") 
+@
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FEP-volumes
+<<fig:FEP-vols,fig.show='hide'>>=
   fep.raw.volumes=read.csv('data/SZ_vol_by_method.csv')
   fep.raw.volumes=subset(fep.raw.volumes, SID <= 81)
-fep.raw.volumes=melt(fep.raw.volumes, id.vars=c('SID','method')) 
-fep.volumes = dcast(fep.raw.volumes, SID~method, value.var="value", fun.aggregate=mean)
-fep.volumes = melt(fep.volumes, id.vars=c('SID', 'manual'), variable.name='method', value.name='volume')
-
-fep.lm = ddply(fep.volumes,c("method"), function (df) {
+  fep.raw.volumes=melt(fep.raw.volumes, id.vars=c('SID','Method')) 
+  fep.volumes = dcast(fep.raw.volumes, SID~Method, value.var="value", fun.aggregate=mean)
+  fep.volumes = melt(fep.volumes, id.vars=c('SID', 'manual'), variable.name='Method', value.name='volume')
+
+  fep.lm = ddply(fep.volumes,c("Method"), function (df) {
     data.frame(eqn=lm_eqn(lm(manual ~ volume, df)))})
-fep.lm$y = 6800-0:3*150
-
-ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=method)) + 
+  fep.lm$y = 6900-0:3*150
+
+  ggplot(data=fep.volumes, aes(x=manual, y=volume, colour=Method)) + 
     geom_smooth(method="lm", formula=y~x) + 
@@
 @
-  \caption{Mean bilateral hippocampus volume as measured by \mb{}, \fsl{}, and
-\freesurfer{} vs. manually segmented volumes from the First Episode Patients with
-Schizophrenia dataset. \mb{} is run two ways: with an atlas library consisting
-of the five Winterburn atlases, or with five manually segmented FEP subjects as
-atlases (not included as subjects). Linear fit lines are shown, with shaded
-region showing standard error. 
-  \label{fig:SZ-volumes}}
-\end{figure}
-
-\begin{figure}
-<<fig:FEP-Bland-Altman, cache=TRUE, dependson='fig:FEP-volumes', fig.width=7, fig.height=7>>=
+
+<<fig:FEP-Bland-Altman,dependson='fig:FEP-vols',fig.width=7,fig.height=3,fig.show='hide'>>=
 fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
 fep.volumes$mean = (fep.volumes$manual + fep.volumes$volume)/2
-limits = ddply(fep.volumes, c("method"), function (df) { 
+limits = ddply(fep.volumes, c("Method"), function (df) { 
   data.frame(
@@
 ggplot(fep.volumes, aes(x= mean, y = diff)) + 
-  facet_wrap( ~ method, nrow=2, scales="fixed") + 
+  facet_wrap( ~ Method, nrow=1, scales="fixed") + 
   geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
@@
   xlab(expression('Mean manual and computed volume ' (mm^3))) + 
-  ylab(expression('Difference in manual and computed volume ' (mm^3))) + 
+  ylab(expression('manual - computed volume ' (mm^3))) + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
@@
 @
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.48\textwidth} 
+    \includegraphics[width=\textwidth]{figure/fig:FEP-xval}
+    \caption{Mean DSC overlap score vs. atlas and template library size in
+    a 5-fold Monte Carlo cross-validation (MCCV) of \mb. Error bars
+    indicate standard error. }
+    \label{fig:FEP-xval}
+  \end{subfigure}%
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:FEP-vols} 
+    \caption{Computed vs. manually segmented mean hippocampus volume. Linear fit
+      lines are shown, with the shaded region showing standard error. }
+    \label{fig:FEP-vols}
+  \end{subfigure}
+  \begin{subfigure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:FEP-Bland-Altman}
     \caption{Bland-Altman plots comparing subject mean hippocampal volume as
@@
     \label{fig:FEP-Bland-Altman}
+  \end{subfigure}
+  \caption{\textbf{First Episode Patient dataset automated segmentation.}
+Manual segmentation of all 81 subjects is completed using the \prsnr-protocol.
+\mb{} is configured using \ants{} registration algorithm and majority vote label
+fusion. In Figure \ref{fig:FEP-vols} \mb{} is run both with an atlas library
+consisting of the five Winterburn atlases (MAGeT-Winterburn), or with an atlas
+library of five manually segmented FEP subjects as atlases (MAGeT-FEP).}
 \end{figure}

### 2014-01-01 paper.Rnw
Added Yushkevich paper reviews

@@
 %  - fix figure 2 schematic language for the experiments (not 2a 2b). 
-% 
+%  - ADNI-1 or ADNI1, standardize
+
 <<setup, include=FALSE, cache=FALSE>>=
@@
 
-Also of interest to this manuscript are methods that attempt to define
-hippocampal subfields using standard T1- or T2-weighted data.  To the best of our
-knowledge, there are only two automated segmentation algorithms that attempt
-this problem. The first is included with the \freesurfer{} package
-\citep{VanLeemput2009}. This work is limited as it omits the tail of the
-hippocampus and the segmentation protocol has yet to be fully validated.
-Nonetheless, it demonstrates that the applicability of hippocampal subfield
-segmentation using data from 10 subjects. In the second method,
-\citet{Yushkevich2009} hippocampal subfields were manually labelled on
-high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
-resolution voxels) T2-weighted MR images acquired from five post-mortem medial
-temporal lobe samples. Using nonlinear registration guided by shape-based models
-of the subfield segmentations, and manually derived hippocampus masks of the
-target images, the authors demonstrate accurate parcellation of hippocampal
-subfields in clinical 3T T1-weighted MRI volumes. This method is not fully
-automatic as it requires whole hippocampal parcellation of the target images.
-%TODO: ref other Yush paper
-
-In this paper we have describe a thorough validation of the \mb{} algorithm for
-the automatic segmentation of the hippocampus and its subfields. First, we
+Also of interest to this manuscript are the methods that attempt to define
+hippocampal subfields using standard T1- or T2-weighted data, of which there are
+few.  \citet{VanLeemput2009} demonstrate that the applicability of hippocampal
+subfield segmentation in T1-weighted images by using Markov random field priors
+learned from 10 manual segmentations.  This work, available as part of the
+\freesurfer{} package, is limited as the segmentation omits the tail of the
+hippocampus and the protocol has yet to be fully validated.
+\citet{Yushkevich2009} manually segment hippocampal subfields on high-resolution
+(either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$ resolution
+voxels) T2-weighted MR images acquired from five post-mortem medial temporal
+lobe samples. Then, using nonlinear registration guided by shape-based models of
+the subfield segmentations, and manually derived hippocampus masks of the target
+images, the authors demonstrate accurate parcellation of hippocampal subfields
+in clinical 3T T1-weighted MRI volumes.  Using \multiatlas with bias correction
+techniques, \citet{Yushkevich2010} demonstrate a semi-automated method of
+subfield segmentation on in vivo focal T2-weighted MR acquisitions of the
+temporal lobe. Manual input is only needed to mark mark divisions between the
+head, body and tail of the hippocampus on target images.
+
+In this paper we describe a thorough validation of the \mb{} algorithm for the
+fully automatic segmentation of the hippocampus and its subfields. First, we
 address the feasibility of whole hippocampus segmentation with a limited number
@@
                             "SR/SL/SM"="X6","SR/SL/SM"="X106")
+
+d = ddply(na.omit(wa.xval.casted), c('region'), function (df) {
+  icc = ICC(df[4:5])
+  data.frame(
+    icc  = icc$results$ICC[3],
+    f    = icc$results$F[3],
+    lower= icc$results$"lower bound"[3],
+    upper= icc$results$"upper bound"[3],
+  )
+})
 

### 2014-01-02 paper.Rnw
Adding updated WA xval experiment

@@
     m_mb_res = mean(mb_0.9mm,na.rm=T), sd_mb_res = sd(mb_0.9mm,na.rm=T)))})
-@
+
+wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
+wa.n.vol = read.csv('data/winterburn.xval.volumes.csv')
+winterburn = read.csv('data/winterburn.volumes.csv')
+
+wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
+winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
+winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
+wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
+#wa.n.vols.mean = ddply(wa.n.all.vols,c('subject','variable'),function (df) {
+#  data.frame(
+#    mb=mean(df$value),
+#    X.3=mean(df$'0.3'),
+#    X.9=mean(df$'0.9'))})
+
+levels(wa.n.all.vols$variable) <-list(
+  "CA1"="X1", "CA1"="X101", "Subiculum"="X2","Subiculum"="X102", 
+  "CA4/DG"="X4","CA4/DG"="X104", "CA2/CA3"="X5","CA2/CA3"="X105", 
+  "SR/SL/SM"="X6","SR/SL/SM"="X106")
+
+wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
+wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2
+
+wa.n.xval.bland.limits = ddply(wa.n.all.vols,c("variable"),function(df) {
+  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
+})
+@
+
+\begin{figure}[h!]
+<<fig:WAxval-kappa>>=
+wa.n.sim$label = as.factor(wa.n.sim$label)
+levels(wa.n.sim$label) <-list(
+  "CA1"=1, "CA1"=101, "Subiculum"=2,"Subiculum"=102, 
+  "CA4/DG"=4,"CA4/DG"=104, "CA2/CA3"=5,"CA2/CA3"=105, 
+  "SR/SL/SM"=6,"SR/SL/SM"=106)
+
+# kappa vs. region
+ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
+  geom_boxplot() + 
+  xlab("Subregion") + 
+  ylab("Dice's Similarity Coefficient (DSC)")
+@
+  \caption{Similarity of \mb{} segmentation of subregions and resampled Winterburn atlas labels at $0.9mm^3$ voxel resolution over all folds of a leave-one-out cross validation.}
+\end{figure}
+
+\begin{figure}[h!]
+<<fig:WAxval-Bland-Altman>>=
+ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
+  facet_wrap(~variable, ncol=3, scales='free') + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=1) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
+             data = wa.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
+  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
+  scale_color_discrete(name="") + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+@
+  \caption{Bland-Altman plots across all folds of Winterburn atlas cross validation.}
+\end{figure}
 

### 2014-01-03 paper.Rnw
Nice updated winterburn xval plots and montage

--- 2014-01-01 paper.Rnw
@@
 \begin{figure}
-  \begin{centering}
+  \centering
     \includegraphics[width=\textwidth]{figure/MA-MAGeTBrain-Schematic}
-  \end{centering}
   \caption{A schematic illustration of basic \multiatlas{} segmentation and
@@
 \begin{figure}
-  \begin{centering}
+  \centering
   \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
-  \end{centering}
   \caption{A schematic illustration of the experimental set up of the
@@
 
-d = ddply(na.omit(wa.xval.casted), c('region'), function (df) {
-  icc = ICC(df[4:5])
-  data.frame(
-    icc  = icc$results$ICC[3],
-    f    = icc$results$F[3],
-    lower= icc$results$"lower bound"[3],
-    upper= icc$results$"upper bound"[3],
-  )
-})
+#d = ddply(na.omit(wa.xval.casted), c('region'), function (df) {
+#  icc = ICC(df[4:5])
+#  data.frame(
+#    icc  = icc$results$ICC[3],
+#    f    = icc$results$F[3],
+#    lower= icc$results$"lower bound"[3],
+#    upper= icc$results$"upper bound"[3],
+#  )
+#})
 
@@
 @
+
+\begin{figure}[h!]
+<<fig:WAxval-kappa>>=
+wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
+wa.n.sim$label = as.factor(wa.n.sim$label)
+levels(wa.n.sim$label) <-list(
+  "CA1"=1, "CA1"=101, "Subiculum"=2,"Subiculum"=102, 
+  "CA4/DG"=4,"CA4/DG"=104, "CA2/CA3"=5,"CA2/CA3"=105, 
+  "SR/SL/SM"=6,"SR/SL/SM"=106)
+
+# kappa vs. region
+ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
+  geom_boxplot(fill='lightblue') + 
+  xlab("Subregion") + 
+  ylab("Dice's Similarity Coefficient (DSC)")
+@
+  \caption{Similarity of \mb{} segmentation of subregions and resampled
+  Winterburn atlas labels at $0.9mm^3$ voxel resolution over all folds of a
+  leave-one-out cross validation.}
+\end{figure}
+
+\begin{figure}[h!]
+<<fig:WAxval-Bland-Altman>>=
+wa.n.vol = read.csv('data/winterburn.xval.volumes.csv')
+winterburn = read.csv('data/winterburn.volumes.csv')
+
+wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
+winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
+winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
+wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
+#wa.n.vols.mean = ddply(wa.n.all.vols,c('subject','variable'),function (df) {
+#  data.frame(
+#    mb=mean(df$value),
+#    X.3=mean(df$'0.3'),
+#    X.9=mean(df$'0.9'))})
+
+levels(wa.n.all.vols$variable) <-list(
+  "CA1"="X1", "CA1"="X101", "Subiculum"="X2","Subiculum"="X102", 
+  "CA4/DG"="X4","CA4/DG"="X104", "CA2/CA3"="X5","CA2/CA3"="X105", 
+  "SR/SL/SM"="X6","SR/SL/SM"="X106")
+
+wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
+wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2
+
+wa.n.xval.bland.limits = ddply(wa.n.all.vols,c("variable"),function(df) {
+  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
+})
+ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
+  facet_wrap(~variable, ncol=3, scales='free_x') + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=1) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
+             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
+  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
+  scale_color_discrete(name="") + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+@
+  \caption{Bland-Altman plots across all folds of Winterburn atlas cross validation.}
+\end{figure}
 
@@
 
-\begin{sidewaysfigure}
-    \begin{centering}
-      \includegraphics[width=\linewidth]{figure/winterburn-atlas-montage/figure.pdf}
-    \end{centering}
-    \caption{Detailed subfield segmentation results for a single subject. In
+\begin{figure}
+    \centering
+    \includegraphics[width=5in]{figure/winterburn-atlas-montage/figure.pdf}
+    \caption{\textbf{Detailed subfield segmentation results for a single subject.} In
     the upper left corner is the original high-resolution Winterburn atlas manual
@@
     segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
-    corner is the \mb{} segmentation of the subsampled Winterburn atlas images; in the
-    lower right corner is the \mb{} segmentation of a separately acquired T1 BRAVO
-    image of the same subject. In each segmentation, slices from the left
+    corner is the \mb{} segmentation of the subsampled Winterburn atlas image
+    from a single fold of the cross-validation. In each segmentation, slices from the left
     hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
@@
     posterior. \label{fig:subfield-montage}}
-\end{sidewaysfigure}
+\end{figure}
 
@@
 \begin{figure}
-  \begin{centering}
+  \centering
   \includegraphics[width=6in]{figure/ADNI1_SNT_MB_montage/montage.pdf}
-  \end{centering}
   \caption{\snt{} and \mb{} segmentations for 30 ADNI subjects (10

### 2014-01-03 paper.Rnw
Updated text to match WA xval figures, and reworked discussion

@@
 estimates a 3D deformation field between two MR images. At first, large
-deformations are estimated using a blurred version of the input data. These larger
-deformations are then input to subsequent steps where the fit is refined by
-estimating smaller deformations on data blurred with a Gaussian kernel with a
+deformations are estimated using a blurred version of the input data. These
+larger deformations are then input to subsequent steps where the fit is refined
+by estimating smaller deformations on data blurred with a Gaussian kernel with a
 smaller full width at half maximum (FWHM). The final transformation is a set of
@@
 The previous experiment assesses \mb{} peformance on whole hippocampus
-segmentation. In this experiment, we assess \mb{} hippocampal subfield
-segmentation of standard 3T T1-weighted images using a leave-one-out
-cross-validation (LOOCV) design.
+segmentation. In this experiment, we aim to evaluate \mb{} hippocampal subfield
+segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
+use a modified leave-one-out cross-validation (LOOCV) design.
 
@@
 algorithm is given all but one item in a dataset as training data (in our case,
-atlas images and labels) and then is applied to the left-out item. This is done,
-in turn, for each item in the dataset and the output across all items is
-evaluated.
-
-In this experiment, the high-resolution $0.3mm^3$ voxel Winterburn atlases are
-used as the \mb{} atlas library, but in each round of LOOCV a $0.9mm^3$ voxel
-target image corresponding to the left-out atlas is used so as to evaluate \mb{}
-on standard 3T T1-weighted resolution images.
-
-Specifically, this experiments has two parts, each using different sets of
-$0.9mm^3$ voxel target images: trilinear subsampled versions of the Winterburn
-atlas images (Experiment 2a), and separately acquired T1 BRAVO acquisitions of
-four of the five Winterburn atlas subjects (Experiment 2b).  Image subsampling
-of the Winterburn atlas images was performed using trilinear subsampling
-techniques.
-
-The 3T T1 BRAVO acquisitions of four of the five Winterburn atlas subjects were
-separately obtained within a short time of the original atlas image
-acquisitions.  Images were acquired on a 3T GE Discovery MR 750 system (General
-Electric, Milwaukee, WI) using an 8-channel head coil with the enhanced fast
-gradient recalled echo 3-dimensional acquisition protocol, EFGRE-BRAVO, with the
-following parameters: $TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle $=8^\circ$,
-$FOV=15.8cm$, slice thickness$=0.9mm$, 176 in-plane steps for an approximate
-isotropic resolution of 0.9mm dimension voxels.
-
-The template library is composed of all $0.9mm^3$ voxel target images (either
-the {\em BRAVO} or {\em subsampled} sets), plus an additional set of 16 of
-healthy subjects 3T T1 images acquired separately (Table
-\ref{tab:WAval-healthy-demographics}). These images were acquired on a 3T GE
-Discovery MR 750 system (General Electric, Milwaukee, WI) using an 8-channel
-head coil with the enhanced fast gradient recalled echo 3-dimensional
-acquisition protocol, EFGRE-BRAVO, with the following parameters: $TE/TR/TI =
-3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+atlas images and labels) and then the algorithm is applied to the left-out item.
+This is done, in turn, for each item in the dataset and the output across all
+items is evaluated together.
+
+In this experiment, the Winterburn atlases are resampled to $0.9mm^3$-isotropic
+voxel images to simulate standard 3T T1-weighted resolution images. Image
+subsampling is performed using trilinear subsampling techniques.
+
+In each round of LOOCV, one atlas is selected and segmented with \mb{} once
+using each possible triplet of atlas images from the remaining four unselected
+atlases. We chose to use an atlas library with an odd number of images so as to
+ensure an unbiased label fusion when using majority voting (see Discussion).
+
+The template library used has a total of 19 images composed of all $0.9mm^3$
+voxel target images plus an additional set of 14 of healthy subject 3T T1 images
+acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
+were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
+WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
+3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
+$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
 thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
@@
 The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used. Figure \ref{fig:LOOCV-schematic}
-illustrates schematically the experimental set up.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figure/WinterburnSubfieldLOOCVSchematic}
-  \caption{A schematic illustration of the experimental set up of the
-  leave-one-out cross-validation (LOOCV) subfield experiment.}
-  \label{fig:LOOCV-schematic}
-\end{figure}
+method found in Experiment 1 are used.
 
@@
 \subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
-<<WAval-prep>>=
-wa.xval.vols = na.omit(read.csv('data/WAval.csv'))
-wa.xval.casted = dcast(melt(wa.xval.vols, id.vars=c("file","version")), file + variable ~ version)
-wa.xval.casted = rename(wa.xval.casted, c("variable"="region"))
-levels(wa.xval.casted$region) <-list("CA1"="X1", "CA1"="X101", 
-                            "Subiculum"="X2","Subiculum"="X102", 
-                            "CA4/DG"="X4","CA4/DG"="X104", 
-                            "CA2/CA3"="X5","CA2/CA3"="X105", 
-                            "SR/SL/SM"="X6","SR/SL/SM"="X106")
-
-#d = ddply(na.omit(wa.xval.casted), c('region'), function (df) {
-#  icc = ICC(df[4:5])
-#  data.frame(
-#    icc  = icc$results$ICC[3],
-#    f    = icc$results$F[3],
-#    lower= icc$results$"lower bound"[3],
-#    upper= icc$results$"upper bound"[3],
-#  )
-#})
-
-fmt_tstat = function(a, b) {
-  test = t.test(a,b, paired=T)
-  p = sprintf("%0.3f", max(0.001,round(test$p.value,3)))
-  t = sprintf("%0.3f", round(test$statistic,3))
-  paste(t," (",p,")")
-}
-
-wa.p = ddply(wa.xval.casted, c("region"), function (df) {
-  with(df, {
-       data.frame(
-         res = fmt_tstat(gold_0.3mm, gold_0.9mm),
-         bra = fmt_tstat(gold_0.3mm, mb_bravo),
-         mbr = fmt_tstat(gold_0.3mm, mb_0.9mm)
-       )})
-  })
-
-wa.p = rename(wa.p, c(
-  'region'='Subregion','res'='Resampled',
-  'bra'='MAGeT on BRAVO','mbr'='MAGeT on Resampled'))
-
-wa.xval.error = ddply(wa.xval.casted, c("file", "region"), function(df) {
-  with(df, 
-  data.frame(
-    resampled = (gold_0.9mm - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_bravo  = (mb_bravo   - gold_0.3mm) / gold_0.3mm * 100, 
-    mb_0.9mm  = (mb_0.9mm   - gold_0.3mm) / gold_0.3mm * 100))})
-
-wa.s = ddply(wa.xval.error, c('region'), function(df) {
-  with(df,data.frame(
-    m_res = mean(resampled,na.rm=T), sd_res = sd(resampled,na.rm=T),
-    m_bravo = mean(mb_bravo,na.rm=T), sd_bravo = sd(mb_bravo,na.rm=T),
-    m_mb_res = mean(mb_0.9mm,na.rm=T), sd_mb_res = sd(mb_0.9mm,na.rm=T)))})
-@
-
-\begin{figure}[h!]
+
+\begin{figure}
 <<fig:WAxval-kappa>>=
@@
   ylab("Dice's Similarity Coefficient (DSC)")
-@
-  \caption{Similarity of \mb{} segmentation of subregions and resampled
-  Winterburn atlas labels at $0.9mm^3$ voxel resolution over all folds of a
-  leave-one-out cross validation.}
+
+CA1   = subset(wa.n.sim,label=='CA1')$k
+CA23  = subset(wa.n.sim,label=='CA2/CA3')$k
+CA4DG = subset(wa.n.sim,label=='CA4/DG')$k
+Sub   = subset(wa.n.sim,label=='Subiculum')$k
+SRLM  = subset(wa.n.sim,label=='SR/SL/SM')$k
+@
+  \caption{Similarity of \mb{} segmentation of subregions and the resampled
+  Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
+  leave-one-out cross validation. Overlap score for each hemisphere is measured
+  separately. \label{fig:WAxval-kappa}}
 \end{figure}
 
-\begin{figure}[h!]
+In this experiment the \mb{} segmentation of hippocampal subfields is assessed
+through a leave-one-out validation in which resampled images ($0.9mm^3$
+voxels) of each Winterburn atlas subject are segmented using the remaining
+high-resolution Winterburn atlas subjects' images as atlases.
+
+Figure \ref{fig:WAval-kappa} shows the overlap similarity scores between the
+\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
+subregion across all subjects and folds of the validation. Mean (standard
+deviation) DSC scores of the subregions are: 
+  \Sexpr{mean(CA1)}   ($\pm$\Sexpr{sd(CA1)}) for the CA1 region,  
+  \Sexpr{mean(CA23)}  ($\pm$\Sexpr{sd(CA23)}) for the CA2/CA3 regions,  
+  \Sexpr{mean(CA4DG)} ($\pm$\Sexpr{sd(CA4DG)}) for the CA4/DG regions,  
+  \Sexpr{mean(Sub)}   ($\pm$\Sexpr{sd(Sub)}) for the Subiculum region, and
+  \Sexpr{mean(SRLM)}  ($\pm$\Sexpr{sd(SRLM)}) for the moleculare regions.
+Bland-Altman plots comparing \mb{} to resampled subregion volumes reveal a
+proportional bias towards underestimating volumes in all subregions (Figure
+\ref{fig:WAxval-Bland-Altman}). Prominent fixed biases in the volumes of the
+CA4/DG and SR/SL/SM subregions are congruent with the lower similarity scores
+seen in these regions.
+
+\begin{figure}
 <<fig:WAxval-Bland-Altman>>=
@@
 wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
-#wa.n.vols.mean = ddply(wa.n.all.vols,c('subject','variable'),function (df) {
-#  data.frame(
-#    mb=mean(df$value),
-#    X.3=mean(df$'0.3'),
-#    X.9=mean(df$'0.9'))})
 
@@
 @
-  \caption{Bland-Altman plots across all folds of Winterburn atlas cross validation.}
+  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
+across all folds of Winterburn atlas cross validation.} The overall mean
+difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
+horizontal lines. Linear fit lines are shown. Note, points below the mean
+difference indication overestimation of the volume with respect to the resampled
+volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
 \end{figure}
-
-In this experiment the \mb{} segmentation of hippocampal subfields is assessed
-through a leave-one-out validation in which resampled images ($0.9mm^3$
-voxels) of each Winterburn atlas subject are segmented using the remaining
-high-resolution Winterburn atlas subjects' images.
-
-Figure \ref{fig:WAval-vol-boxplot} shows the mean relative percent error in volume, 
-with respect to the full-resolution Winterburn atlas, of each hippocampal subfield
-of the resampled Winterburn labels, and of the \mb{} segmentations of the resampled
-Winterburn atlas images and T1 BRAVO Winterburn subject acquisitions.  A
-paired-samples t-test was conducted to compare the distribution of relative
-percent error in the \mb{} segmentation volumes and the resampled segmentation
-volumes. There was no significant difference found for the CA1, Subiculum, and
-CA2/CA3 subregions. The CA4/DG subregion shows significantly lower percent error 
-  ($M_{resampled}=$\Sexpr{wa.s$m_res[wa.s$region=='CA4/DG']},
-  $SD_{resampled}=$\Sexpr{wa.s$sd_res[wa.s$region=='CA4/DG']};
-  $M_{MAGeT(BRAVO)}=$\Sexpr{wa.s$m_bravo[wa.s$region=='CA4/DG']},    
-  $SD_{MAGeT(BRAVO)}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='CA4/DG']};    
-  $M_{MAGeT(resampled)}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='CA4/DG']},    
-  $SD_{MAGeT(resampled)}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='CA4/DG']};    
-  $p < 0.001$), and the SR/SL/SM shows significantly different percent error in the opposite
-direction 
-  ($M_{resampled}=$\Sexpr{wa.s$m_res[wa.s$region=='SR/SL/SM']},
-  $SD_{resampled}=$\Sexpr{wa.s$sd_res[wa.s$region=='SR/SL/SM']};
-  $M_{MAGeT(BRAVO)}=$\Sexpr{wa.s$m_bravo[wa.s$region=='SR/SL/SM']},    
-  $SD_{MAGeT(BRAVO)}=$\Sexpr{wa.s$sd_bravo[wa.s$region=='SR/SL/SM']};    
-  $M_{MAGeT(resampled)}=$\Sexpr{wa.s$m_mb_res[wa.s$region=='SR/SL/SM']},    
-  $SD_{MAGeT(resampled)}=$\Sexpr{wa.s$sd_mb_res[wa.s$region=='SR/SL/SM']};    
-  $p < 0.001$). Bland-Altman plots comparing \mb{} to subsampled segmentation
-volume reveals a proportional bias towards underestimating volumes in all
-subregions, and prominant fixed biases in the volumes of the CA4/DG and SR/SL/SM subregions 
-congruent with significant differences seen in percent error in volume
-associated with subsampling alone.
 
@@
 single subject for qualitative inspection.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% WAxval - t.test comparing Manual volumes and Computed volumes of subregions 
-%          across folds
-<<tab:WAxval-t-test-manual-computed, echo=F, results="asis",cache=FALSE>>=
-latex(wa.p, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
-    caption="\\textbf{Paired-sample t-tests comparing manually segmented
-             Winterburn atlas subregion volumes and computed volumes (reported
-             as $t (p)$). }", 
-    col.just=c("l","r","r","r"),
-    caption.loc = c("bottom"), rowname=NULL, 
-    label="tab:WAxval-t-test-manual-computed")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Winterburn-XVAL: compare resampling
-\begin{figure}[h!]
-<<fig:WAval-vol-boxplot>>=
-wa.xval.error.melted = melt(wa.xval.error,id.vars=c("file", "region"))
-wa.xval.error.melted = rename(wa.xval.error.melted, c("variable"="measure", "value"="percenterr"))
-levels(wa.xval.error.melted$measure) <- list("Subsampled atlas labels"="resampled",
-                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
-                               "MAGeT on T1 BRAVO images"="mb_bravo")
-
-ggplot(wa.xval.error.melted, aes(y=percenterr,x=measure,colour=measure)) +
-  facet_grid(.~region) + 
-  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
-  geom_boxplot() + 
-  geom_point(group=1, position="dodge") + 
-  labs(y="Percent error in volume", x="Subfield") +
-  theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + 
-  scale_colour_discrete("") + 
-  theme(legend.direction = "vertical", legend.position = "bottom")
-@
-  \caption{{\bf Percent error in computed unilateral hippocampal subfield
-segmentation volume relative to manual segmentation volume.} 
-  Percent error is measured against the volumes of the unmodified Winterburn
-  atlas subfield segmentations.
-  {\bf Subsampled atlas labels} volumes of the manual segmentations of the Winterburn
-  atlases after resampling to $0.9mm^3$ voxels.  
-  {\bf MAGeT on subsampled atlas images} volumes are \mb{} segmentations of the Winterburn atlas
-  images after resampling to $0.9mm^3$ voxels.
-  {\bf MAGeT on T1 BRAVO} volumes are \mb{} segmentations of T1 BRAVO images
-  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
-  subjects.\label{fig:WAval-vol-boxplot}}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Winterburn-XVAL: bland-altman 
-\begin{figure}[h!]
-<<fig:WAval-bland>>=
-wa.xval.volumes = melt(wa.xval.casted, id.vars=c('file', 'region', 'gold_0.3mm', 'gold_0.9mm'))
-wa.xval.volumes = rename(wa.xval.volumes, c("variable"="measure", "value"="volume"))
-levels(wa.xval.volumes$measure) <- list("Subsampled atlas labels"="gold_0.9mm",
-                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
-                               "MAGeT on T1 BRAVO images"="mb_bravo")
-
-wa.xval.volumes$diff = wa.xval.volumes$gold_0.9mm - wa.xval.volumes$volume 
-wa.xval.volumes$mean = (wa.xval.volumes$gold_0.9mm + wa.xval.volumes$volume)/2
-wa.xval.bland.limits = ddply(wa.xval.volumes,c("region"),function(df) {
-  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
-})
-
-ggplot(wa.xval.volumes, aes(x=mean, y=diff, colour=measure)) + 
-  facet_wrap(~region, ncol=3, scales='free') + 
-  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=1) +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
-             data=wa.xval.bland.limits, size=0.3, alpha=0.5) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
-             data = wa.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
-  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
-  scale_color_discrete(name="") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text.y = element_text(angle = 90, hjust=0.5), 
-        axis.text = element_text(size = 8))
-
-@
-  \caption{{\bf Bland-Altman plots comparing unilateral \mb{} subfield
-segmentation volume with resampled subfield segmentation volume.} The overall
-mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
-dashed horizontal lines. Linear fit lines are shown for each target image type.
-Note, points below the mean difference indicate overestimation of the volume
-with respect to the resampled volume, and vice versa. 
-  \label{fig:WAval-bland}}
-\end{figure}
 
@@
 \section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Tabulation of segmentation accuracies of existing methods (ADNI)
-%
-<<ADNI1-literature, echo=F, results="asis",cache=FALSE>>=
-# caption = 
-#   "\\textbf{Automated segmentation accuracy (overlap with manual labels) of the 
-#    ADNI dataset.} For each method, the number of manually labelled atlases used for 
-#    training, the best Dice's overlap measure, the disease classes measured, and 
-#    the validation procedure are shown. LOOCV = Leave-one-out cross-validation. 
-#    Some studies of automated segmentation of ADNI images are excluded because they 
-#    do not not provide overlap measures for the hippocampus \\cite{Heckemann2011, Chupin2009}.
-#   "
-# 
-# tab = read.csv("data/ADNI-existing-work.csv")
-# tab$Notes <- NULL
-# 
-# adni_lit = subset(tab, Dataset=="ADNI")
-# adni_lit$Dataset <- NULL
-# 
-# 
-# latex(adni_lit, #file="", 
-#       size="scriptsize",
-#       caption=caption, 
-#       rowname=NULL,
-#       col.just=c("p{2in}","p{1in}","p{1in}","p{1in}","p{1in}"),
-#       title="")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The core claim the \mb{} method is based on, that we can meaningfully bootstrap
-a template library from a small set of labelled atlas images, is validated in
+The core claim the \mb{} method is based on -- that we can meaningfully bootstrap
+a template library from a small set of labelled atlas images -- is validated in
 the cross validation conducted in Experiment 1 (and the replication in
@@
 
-Finally, we have provide evidence that using the Winterburn high-resolution
+Finally, we have provided evidence that using the Winterburn high-resolution
 hippocampal subfield atlases \citep{Winterburn2013} our algorithmic framework is
@@
 subfield segmentations produced in \citet{Yushkevich2009}).  Our work
-demonstrates that we can reliably reproduce segmentations, corresponding in volumes 
-to the input atlases used, with only modest error for the CA1, subiculum, and 
-CA4/dentate subregions.
-
-The fact that CA2/CA3 and molecular layers cannot be reasonably reproduced
-should not be surprising as these are extremely thin and spatially convoluted
-regions that originally required high-resolution MRI for identification and so
-it is likely that the extents of these regions are well below the resolution and
-contrast offered by standard T1-weighted images. In addition, some manual
-subfield segmentation methodologies do not attempt to parse these regions even
-on high-resolution scans. See Table \ref{ab:subfield-segmentation-protocols} for
-a comparison of manual subfield segmentation methodologies. Further complicating
-this, is that that different research groups have different operational
-definitions for these different subfields. The disagreement in the community has
-led to an international working group devoted to normalizing the ontology and
-segmentation rules for the hippocampal subfields
+demonstrates that we can reliably reproduce segmentations for the CA1,
+subiculum, and CA4/dentate subregions.  That the CA2/CA3 and molecular
+layers are less well reproduced should not be surprising as these are extremely
+thin and spatially convoluted regions that originally required high-resolution
+MRI for identification and so it is likely that the extents of these regions are
+well below the resolution and contrast offered by standard T1-weighted images.
+
+This points to a larger issue of how to truly validate subfield segmentations,
+both in high resolution images and in standard T1-weighted images.  There are
+several manual subfield segmentation methodologies, and they do not agree on
+which regions can be differentiated, even on high-resolution scans. See Table
+\ref{ab:subfield-segmentation-protocols} for a comparison of manual subfield
+segmentation methodologies. A further complication is that different
+researchers have differing operational definitions for the subfields and how
+they ought to be parcellated. The disagreement in the community has led to an
+international working group devoted to normalizing the ontology and segmentation
+rules for the hippocampal subfields
 (\url{http://www.hippocampalsubfields.com/}). The definitional and operational
@@
 Dice's Kappa) not particularly meaningful until these disagreements have been
-resolved.  Thus, because of this, and the inability to discern subregions in
-standard T1-weighted images, we opted to speak of the precision or reliability
-of \mb{} subfield segmentation rather than speak of accuracy.
+resolved.  Thus, because of these disagreements and the difficulting discerning
+subregions in standard T1-weighted images, we opted to speak of the precision or
+reliability of \mb{} subfield segmentation rather than speak of accuracy.
 
@@
 identification of the hippocampal subfields is possible. 
+% TODO: say that we have shown the viability of template generation, and as such
+% it may also serve as a useful addition to existing segmentation techniques
 

### 2014-01-07 paper.Rnw
Adding WA xval perturb tests

@@
 lobe, and has long been associated with learning and memory
-\citep{DenHeijer2012,Scoville2000}. The hippocampus is of interest to
-clinical neuroscientists because it is implicated in many forms of brain
-dysfunction, including Alzheimer's disease \citep{Sabuncu2011} and
+\citep{DenHeijer2012,Jeneson2012,Wixted2011,Scoville2000}. The hippocampus is of
+interest to clinical neuroscientists because it is implicated in many forms of
+brain dysfunction, including Alzheimer's disease \citep{Sabuncu2011} and
 schizophrenia \citep{Narr2004,Karnik-Henry2012}. In neuroimaging studies,
@@
 to the hippocampus using more than a single input atlas, while simultaneously
-limiting the number of inputs used during segmentation, and demonstrating that
-accurate identification of the hippocampal subfields is indeed possible using
-this methodology.
+limiting the number of inputs used during segmentation, and so demonstrating that
+reliable application of the atlases' subfield segmentation to an unlabelled
+target image is indeed possible using this methodology.
 
@@
 few.  \citet{VanLeemput2009} demonstrate that the applicability of hippocampal
-subfield segmentation in T1-weighted images by using Markov random field priors
-learned from 10 manual segmentations.  This work, available as part of the
-\freesurfer{} package, is limited as the segmentation omits the tail of the
-hippocampus and the protocol has yet to be fully validated.
-\citet{Yushkevich2009} manually segment hippocampal subfields on high-resolution
-(either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$ resolution
-voxels) T2-weighted MR images acquired from five post-mortem medial temporal
-lobe samples. Then, using nonlinear registration guided by shape-based models of
-the subfield segmentations, and manually derived hippocampus masks of the target
-images, the authors demonstrate accurate parcellation of hippocampal subfields
-in clinical 3T T1-weighted MRI volumes.  Using \multiatlas with bias correction
-techniques, \citet{Yushkevich2010} demonstrate a semi-automated method of
-subfield segmentation on in vivo focal T2-weighted MR acquisitions of the
-temporal lobe. Manual input is only needed to mark mark divisions between the
-head, body and tail of the hippocampus on target images.
+subfield segmentation in T1-weighted images by Bayesian techniques using Markov
+random field shape priors learned from 10 manual segmentations.  This work,
+available as part of the \freesurfer{} package, is limited as the segmentation
+omits the tail of the hippocampus and the protocol has yet to be fully
+validated.  \citet{Yushkevich2009} manually segment hippocampal subfields on
+high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
+resolution voxels) T2-weighted MR images acquired from five post-mortem medial
+temporal lobe samples. Then, using nonlinear registration guided by shape-based
+models of the subfield segmentations, and manually derived hippocampus masks of
+the target images, the authors demonstrate accurate parcellation of hippocampal
+subfields in clinical 3T T1-weighted MRI volumes.  Using \multiatlas with bias
+correction techniques, \citet{Yushkevich2010} demonstrate a semi-automated
+method of subfield segmentation on in vivo focal T2-weighted MR acquisitions of
+the temporal lobe. Manual input is only needed to mark mark divisions between
+the head, body and tail of the hippocampus on target images.
 
@@
 The previous experiment assesses \mb{} peformance on whole hippocampus
-segmentation. In this experiment, we aim to evaluate \mb{} hippocampal subfield
+segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
 segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
@@
 
-In this experiment, the Winterburn atlases are resampled to $0.9mm^3$-isotropic
+In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
 voxel images to simulate standard 3T T1-weighted resolution images. Image
-subsampling is performed using trilinear subsampling techniques.
-
-In each round of LOOCV, one atlas is selected and segmented with \mb{} once
-using each possible triplet of atlas images from the remaining four unselected
-atlases. We chose to use an atlas library with an odd number of images so as to
-ensure an unbiased label fusion when using majority voting (see Discussion).
+subsampling is performed using trilinear subsampling techniques.  In each round
+of LOOCV, a single atlas is selected and segmented with \mb{} once using each
+possible triplet of atlas images from the remaining four unselected atlases.
+Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
+image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
+with an odd number of images so as to ensure an unbiased label fusion when using
+majority voting (see Discussion).
 
@@
 levels(wa.n.sim$label) <-list(
-  "CA1"=1, "CA1"=101, "Subiculum"=2,"Subiculum"=102, 
-  "CA4/DG"=4,"CA4/DG"=104, "CA2/CA3"=5,"CA2/CA3"=105, 
+  "CA1"=1, "CA1"=101, 
+  "CA2/CA3"=5,"CA2/CA3"=105, 
+  "CA4/DG"=4,"CA4/DG"=104, 
+  "Subiculum"=2,"Subiculum"=102, 
   "SR/SL/SM"=6,"SR/SL/SM"=106)
@@
 \mb{} segmentations and the resampled Winterburn atlases for each hippocampal
-subregion across all subjects and folds of the validation. Mean (standard
-deviation) DSC scores of the subregions are: 
-  \Sexpr{mean(CA1)}   ($\pm$\Sexpr{sd(CA1)}) for the CA1 region,  
-  \Sexpr{mean(CA23)}  ($\pm$\Sexpr{sd(CA23)}) for the CA2/CA3 regions,  
-  \Sexpr{mean(CA4DG)} ($\pm$\Sexpr{sd(CA4DG)}) for the CA4/DG regions,  
-  \Sexpr{mean(Sub)}   ($\pm$\Sexpr{sd(Sub)}) for the Subiculum region, and
-  \Sexpr{mean(SRLM)}  ($\pm$\Sexpr{sd(SRLM)}) for the moleculare regions.
+subregion across all subjects and folds of the validation. Mean and standard
+deviation DSC scores of the subregions are: 
+  \Sexpr{mean(CA1)}  $\pm$\Sexpr{sd(CA1)} for the CA1 region,  
+  \Sexpr{mean(CA23)} $\pm$\Sexpr{sd(CA23)} for the CA2/CA3 regions,  
+  \Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)} for the CA4/DG regions,  
+  \Sexpr{mean(Sub)}  $\pm$\Sexpr{sd(Sub)} for the Subiculum region, and
+  \Sexpr{mean(SRLM)} $\pm$\Sexpr{sd(SRLM)} for the moleculare regions.
+
 Bland-Altman plots comparing \mb{} to resampled subregion volumes reveal a
@@
 seen in these regions.
+
+<<tab:WAxval-perturb-kappa, echo=F,results="asis">>=
+wa.perturb = read.csv('data/winterburn.perturb.kappa.csv')
+wa.perturb = wa.perturb[c("subject","label","operation","k")]
+wa.perturb$label = as.factor(wa.perturb$label)
+levels(wa.perturb$label) <-list(
+  "CA1"="1", "CA1"="101",  "CA2/CA3"="5","CA2/CA3"="105",
+  "CA4/DG"="4","CA4/DG"="104",  "Subiculum"="2","Subiculum"="102", 
+  "SR/SL/SM"="6", "SR/SL/SM"="106")
+mean_range = function (v) {
+  sprintf("$%0.2f \\pm %0.2f$", mean(v), sd(v))
+}
+wa.perturb.table = dcast(wa.perturb, label ~ operation, fun.aggregate=mean_range)
+wa.sim.sum = dcast(wa.n.sim, label ~ ., value.var = 'k', fun.aggregate=mean_range)
+names(wa.sim.sum)[2] = 'MAGeT'
+wa.perturb.table = merge(wa.perturb.table, wa.sim.sum)
+
+wa.perturb.table=wa.perturb.table[c('label','MAGeT', 'translate','contract','expand')]
+names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")
+
+caption="Overlap similarity results for the each of the subfields of the hippocampus. 
+Simulated overlap similarity results are also given for resampled labels that were 
+translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
+by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
+Values are given as mean Dice's Kappa coefficient \\pm stddev."
+latex(wa.perturb.table, #file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      label="tab:WAxval-perturb-kappa", rowname=NULL)
+@
 
@@
 %%  
-\mb{}                     &9 & \Sexpr{snt.mb.kappa.best} & \cite{Pipitone2014}
+\mb{}                     &9 & \Sexpr{snt.mb.kappa.best} &
 & 10-fold MCCV on 69 subjects 
@@
 %%
-\mb{}                      &9 & \Sexpr{mb.kappa.best}       &\cite{Pipitone2014} 
+\mb{}                      &9 & \Sexpr{mb.kappa.best}       &
 & 10-fold MCCV on 60 subjects 
@@
 although very few automated methods are available for the segmentation of 3T
-data \citep{Yushkevich2009,VanLeemput2009}. While recent work demonstrates that
-subfield segmentations can be used for classification of AD, MCI, and CN, there
-has been no explicit validation of the segmentations being produced other than
-indirectly via classification accuracy (for instance, the semi-automated
-subfield segmentations produced in \citet{Yushkevich2009}).  Our work
-demonstrates that we can reliably reproduce segmentations for the CA1,
-subiculum, and CA4/dentate subregions.  That the CA2/CA3 and molecular
-layers are less well reproduced should not be surprising as these are extremely
-thin and spatially convoluted regions that originally required high-resolution
-MRI for identification and so it is likely that the extents of these regions are
-well below the resolution and contrast offered by standard T1-weighted images.
+data \citep{Yushkevich2009,Yushkevich2010,VanLeemput2009}.  Table
+\ref{tab:subfield-segmentation-kappa} compares segmentation accuracy from some
+of these methods and \mb{}. The overlap DSC scores for \mb{} subregions are
+notably lower but a direct comparison of overlap values must be done cautiously.
+In the present work, our overlap scores are computed on $0.9mm$-isotropic voxel
+resolution images, whereas \citet{Yushkevich2010} uses focal
+$0.4\times0.5\times2.0mm$ voxel resolution images, and \citet{VanLeemput2009}
+use supersampled $0.380mm$-isotropic voxel resolution images. The larger voxel
+images we use necessarily entail a greater change in DSC for each incorrectly
+labelled voxel.  In addition, our automated segmentations are compared to manual
+segmentations resampled from $0.3mm$-isotropic voxel labels; the resampling
+process inevitably introduces noise which may lower overlap scores.  Lastly, as
+our method is aimed specifically at situations when manually produced atlases
+are scarce, in our cross validation we are forced to use three rather than all
+five of the Winterburn atlases which, based on our findings with whole
+hippocampal segmentation, would lead to improved overlap similarity.
+Nevertheless, the advantage of evaluating \mb{} on standard 3T T1-weighted
+resolution MR images with a publically available atlas library is that our
+results reflect typical usage scenarios of researchers and clinicians.
+
+Our overlap DSC values demonstrates that we can reliably reproduce segmentations
+for the CA1, subiculum, and CA4/dentate subregions.  That the CA2/CA3 and
+molecular layers are less well reproduced (DSC < 0.5) should not be surprising
+as these are extremely thin and spatially convoluted regions that originally
+required high-resolution MRI for identification and so it is likely that the
+extents of these regions are well below the resolution and contrast offered by
+standard T1-weighted images.
+
+\begin{table}[!tbp]
+\scriptsize
+\caption{\textbf{A comparison of subfield segmentation overlap similarity with
+manual raters.} \label{tab:subfield-segmentation-kappa}} 
+\begin{center}
+\begin{tabular}{ l c c c }
+\hline\hline
+Subfield             & \mb{}               & \citet{VanLeemput2009} & \citet{Yushkevich2010}  \tabularnewline
+\hline                                   
+CA1                  &\Sexpr{mean(CA1)}    & 0.62 & 0.875                                     \tabularnewline
+CA2/3                &\Sexpr{mean(CA23)}   & 0.74 & $CA2=0.538, CA3=0.618$                    \tabularnewline
+CA4/DG               &\Sexpr{mean(CA4DG)}  & 0.68 & $DG=0.873$                                \tabularnewline
+presubiculum         & ---                 & 0.68 & ---                                       \tabularnewline
+subiculum            &\Sexpr{mean(Sub)}    & 0.74 & 0.770                                     \tabularnewline
+hippocampal fissure  & ---                 & 0.53 & ---                                       \tabularnewline
+SR/SL/SM             &\Sexpr{mean(SRLM)}   & ---  & ---                                       \tabularnewline
+fimbria              & ---                 & 0.51 & ---                                       \tabularnewline
+head                 & ---                 & ---  & 0.902                                     \tabularnewline
+tail                 & ---                 & ---  & 0.863                                     \tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2014-01-10 paper.Rnw
Fixes to the WA xval presentation

@@
 %  - ADNI-1 or ADNI1, standardize
+%  - Don't use "Experiment X", instead use some short form description, and then
+%  reference (Experiment X).
 
@@
 protocols for segmentation at this resolution to use as a reference. Instead, we
-can evaluate the {\it precision} with which \mb{} produces hippocampal subfields
-that correspond to the segmentation protocol used by the given atlas library
-images.  Specifically, we compute the relative percent error in volume of the
-\mb{}-produced hippocampal subfields with respect to the full-resolution
-Winterburn atlas segmentations. In addtion, by directly resampling the
-Winterburn atlas segmentations to $0.9mm^3$ voxels (using standard
-nearest-neighbour image resampling techniques) we obtain a scaled down version
-of the labels preserving the original segmentation protocol within the limits of
-error from rounding and interpolation. The relative percent error in volume of
-these labels provides a baseline to compare \mb{} segmentations against.
+can evaluate the reliability, or {\it precision}, with which \mb{} produces
+hippocampal subfields that correspond to the segmentation protocol used by the
+given atlas library images. By directly resampling the Winterburn atlas
+segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
+resampling techniques) we obtain a scaled down version of the labels which
+preserve the original segmentation protocol within the limits of error from
+rounding and interpolation. Therefore, using the resampled Winterburn
+segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
+of \mb{} segmentations using using DSC overlap scores and evaluate consistency
+across the range of hippocampal sizes using Bland-Altman plots of subfield
+volumes.
 
@@
 In this experiment the \mb{} segmentation of hippocampal subfields is assessed
-through a leave-one-out validation in which resampled images ($0.9mm^3$
-voxels) of each Winterburn atlas subject are segmented using the remaining
-high-resolution Winterburn atlas subjects' images as atlases.
+through a leave-one-out validation in which resampled images ($0.9mm^3$ voxels)
+of each Winterburn atlas subject are segmented using the remaining
+high-resolution Winterburn atlases.
 
@@
 subregion across all subjects and folds of the validation. Mean and standard
-deviation DSC scores of the subregions are: 
-  \Sexpr{mean(CA1)}  $\pm$\Sexpr{sd(CA1)} for the CA1 region,  
-  \Sexpr{mean(CA23)} $\pm$\Sexpr{sd(CA23)} for the CA2/CA3 regions,  
-  \Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)} for the CA4/DG regions,  
-  \Sexpr{mean(Sub)}  $\pm$\Sexpr{sd(Sub)} for the Subiculum region, and
-  \Sexpr{mean(SRLM)} $\pm$\Sexpr{sd(SRLM)} for the moleculare regions.
-
-Bland-Altman plots comparing \mb{} to resampled subregion volumes reveal a
-proportional bias towards underestimating volumes in all subregions (Figure
+deviation DSC scores of the subregions are shown in Table
+\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
+segmentations when perturbed slightly and compared to the originals.  We find
+that the CA4/DG subregion shows the highest mean DSC score of
+\Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
+subregions having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
+\Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively.  Both the CA4/DG and
+molecular regions score below 0.5.  These scores may seem low but not when taken
+in context and compared to existing (semi-)automated methods (see Discussion).
+
+\begin{figure}
+<<fig:WAxval-Bland-Altman>>=
+wa.n.vol = read.csv('data/winterburn.xval.volumes.csv')
+winterburn = read.csv('data/winterburn.volumes.csv')
+subfield_labels <-list(
+  "CA1"="X1", "CA1"="X101", 
+  "CA2/CA3"="X5","CA2/CA3"="X105", 
+  "CA4/DG"="X4","CA4/DG"="X104",
+  "Subiculum"="X2","Subiculum"="X102",
+  "SR/SL/SM"="X6","SR/SL/SM"="X106")
+hemispheres <- list(
+  "Right"="X1","Left"="X101", 
+  "Right"="X5","Left"="X105", 
+  "Right"="X4","Left"="X104",
+  "Right"="X2","Left"="X102",
+  "Right"="X6","Left"="X106")
+
+wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
+winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
+winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
+wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
+
+levels(wa.n.all.vols$subject) <- list(
+  "1"="anusha", 
+  "2"="jon",
+  "3"="mallar",
+  "4"="sofia",
+  "5"="nancy")
+
+# mean per subfield/hemisphere across all 4 validation folds
+wa.n.mean.vols = ddply(wa.n.all.vols, c("variable", "subject", "res"), function(df) {
+  data.frame(mb = mean(df$value),
+             high  = mean(df$'0.3'),
+             low   = mean(df$'0.9'))
+})
+wa.n.mean.vols$diff = wa.n.mean.vols$low - wa.n.mean.vols$mb
+wa.n.mean.vols$mean = (wa.n.mean.vols$low + wa.n.mean.vols$mb)/2
+
+wa.n.mean.vols$hemisphere <- wa.n.mean.vols$variable
+levels(wa.n.all.vols$variable)   <- subfield_labels
+levels(wa.n.mean.vols$variable)  <- subfield_labels
+levels(wa.n.all.vols$hemisphere) <- hemispheres
+
+wa.n.xval.bland.limits = ddply(wa.n.mean.vols,c("variable"),function(df) {
+  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
+})
+
+wa.n.xval.bland.cor = ddply(wa.n.mean.vols,c("variable"),function(df) {
+  cor.t = cor.test(df$low,df$mb)
+  data.frame(
+    pearson = sprintf("r=%0.2f, t=%0.2f, p < %0.2f", 
+                cor.t$estimate, cor.t$statistic, cor.t$p.value)
+  )
+})
+
+ggplot(wa.n.mean.vols, aes(x=mean, y=diff)) + 
+  facet_wrap(~variable, ncol=3, scales='free_x') + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=2, aes(colour=hemisphere, shape=subject)) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
+             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
+  geom_text(aes(y = Inf, x=-Inf, label=pearson), colour="black", 
+             data = wa.n.xval.bland.cor, hjust=-0.1,vjust=2.5, size=2.5) + 
+  xlab(expression('Mean of resampled and MAGeT-Brain subregion volume ' (mm^3))) + 
+  ylab(expression('Difference of manual and MAGeT-Brain subregion volume ' (mm^3))) + 
+  #scale_x_continuous(limits=c(0,1000)) + 
+  #scale_color_discrete(name="") + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+@
+  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
+across all folds of Winterburn atlas cross validation.} The overall mean
+difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
+horizontal lines. Linear fit lines are shown. Note, points below the mean
+difference indication overestimation of the volume with respect to the resampled
+volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
+\end{figure}
+
+We find that in each subfield, \mb{} shows a proportional bias towards
+underestimating volumes in all subregions (Figure
 \ref{fig:WAxval-Bland-Altman}).  Prominent fixed biases in the volumes of the
@@
 by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
-Values are given as mean Dice's Kappa coefficient \\pm stddev."
-latex(wa.perturb.table, #file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+Values are given as mean Dice's Kappa coefficient $\\pm$ stddev."
+latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
       caption=caption, caption.loc = c("top"),
+      col.just=rep('c',5),  
       label="tab:WAxval-perturb-kappa", rowname=NULL)
 @
-
-\begin{figure}
-<<fig:WAxval-Bland-Altman>>=
-wa.n.vol = read.csv('data/winterburn.xval.volumes.csv')
-winterburn = read.csv('data/winterburn.volumes.csv')
-
-wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
-winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
-winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
-wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
-
-levels(wa.n.all.vols$variable) <-list(
-  "CA1"="X1", "CA1"="X101", "Subiculum"="X2","Subiculum"="X102", 
-  "CA4/DG"="X4","CA4/DG"="X104", "CA2/CA3"="X5","CA2/CA3"="X105", 
-  "SR/SL/SM"="X6","SR/SL/SM"="X106")
-
-wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
-wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2
-
-wa.n.xval.bland.limits = ddply(wa.n.all.vols,c("variable"),function(df) {
-  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
-})
-ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
-  facet_wrap(~variable, ncol=3, scales='free_x') + 
-  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=1) +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
-             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
-             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of resampled and \\mb subregion volume ' (mm^3))) + 
-  ylab(expression('Difference between manual and \\mb subregion volume ' (mm^3))) + 
-  scale_color_discrete(name="") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text.y = element_text(angle = 90, hjust=0.5), 
-        axis.text = element_text(size = 8))
-@
-  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
-across all folds of Winterburn atlas cross validation.} The overall mean
-difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
-horizontal lines. Linear fit lines are shown. Note, points below the mean
-difference indication overestimation of the volume with respect to the resampled
-volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
-\end{figure}
 
@@
 results reflect typical usage scenarios of researchers and clinicians. 
+%TODO: mention that DSC by it's nature penalizes structures with high surface
+%area-to-vol ratios, so necessarily penalise all subfield errors.
+%TODO: discuss perturbed segmentation kappa values in relation to MAGeT error
 
@@
 
-In conclusion, we have presented a flexible \multiatlas{}-based framework that has
-considerable advantages over other methods since only a small number of atlases
-is required to initialize the algorithm. We demonstrate that our method works
-robustly over hippocampal definitions, different disease populations, and
-different acquisition types. Finally, we also demonstrate that accurate
-identification of the hippocampal subfields is possible. 
-% TODO: say that we have shown the viability of template generation, and as such
-% it may also serve as a useful addition to existing segmentation techniques
+In conclusion, we have demonstrated the viability of leveraging a small
+number of input atlases to bootstrap a large template library and thereby
+improve segmentation reliability when using \multiatlas{} methods.  We
+demonstrated that this method works robustly over hippocampal definitions,
+different disease populations, and different acquisition types. Finally, we also
+demonstrate that accurate identification of the hippocampal subfields in
+standard 3T T1-weighted images is possible.
 
@@
 })
-
-ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
-  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
-  #facet_grid( . ~ stat) + 
-  geom_line(size=0.5) +
-  geom_point(size=2) + 
-  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
-  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
-  scale_x_continuous(breaks=seq(1,20,by=2)) + 
-  #  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
-  xlab( "Number of Templates" ) + 
-  ylab( "Variability (p)") +
-  scale_colour_hue(name="Atlases") +
-  scale_alpha_continuous(guide='none') + 
-  scale_linetype_discrete(guide='none') + 
-  scale_size_continuous(guide='none') + 
-  theme(legend.direction = "horizontal", legend.position = "bottom")
-@
-  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation 
-  with respect to \snt{} segmentations.}
-  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation is
-  computed for each subject across all ten rounds of validation. Shown on the
   y-axis is the p-value from a t-test comparing the distribution of variances at

### 2014-01-11 paper.Rnw
reorganised

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                               METHODS
+%%%%%                                                                      %%%%%
+%%%%%                     MAGeT Brain Algorithm                            %%%%%
+%%%%%                                                                      %%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-
-%%%%%                     MAGeT Brain Algorithm                            %%%%%
-
-\subsection{The \mb{} Algorithm}
+\section{The \mb{} Algorithm}
 In this paper, we use the term {\em label} to mean any segmentation (manual or
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiments}
+\section{Experiments}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 1: Whole Hippocampus Cross-Validation}
+\subsection{Experiment 1: Whole Hippocampus Cross-Validation}
 
@@
 
+\subsubsection{Materials and Methods}
 \paragraph{\adnidataset{} dataset} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%        Experiment 2: Hippocampal Subfield Cross-Validation          %%%%%
-%%%%%                                                                     %%%%%
+%%%%%     RESULTS: Experiment 1: Whole Hippocampus Cross-Validation       %%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 2: Hippocampal Subfield Cross-Validation}
-The previous experiment assesses \mb{} peformance on whole hippocampus
-segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
-segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
-use a modified leave-one-out cross-validation (LOOCV) design.
-
-\paragraph{Winterburn Atlases} 
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
-segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
-and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
-lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
-range in age from 29-57 years (mean age of 37), and include two males and three
-females.
-
-\paragraph{Experiment details} 
-Leave-one-out cross-validation (LOOCV) is a validation approach in which an
-algorithm is given all but one item in a dataset as training data (in our case,
-atlas images and labels) and then the algorithm is applied to the left-out item.
-This is done, in turn, for each item in the dataset and the output across all
-items is evaluated together.
-
-In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
-voxel images to simulate standard 3T T1-weighted resolution images. Image
-subsampling is performed using trilinear subsampling techniques.  In each round
-of LOOCV, a single atlas is selected and segmented with \mb{} once using each
-possible triplet of atlas images from the remaining four unselected atlases.
-Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
-image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
-with an odd number of images so as to ensure an unbiased label fusion when using
-majority voting (see Discussion).
-
-The template library used has a total of 19 images composed of all $0.9mm^3$
-voxel target images plus an additional set of 14 of healthy subject 3T T1 images
-acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
-were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
-WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
-3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
-$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
-thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
-0.9mm dimension voxels.
-
-The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Subfield Validation Template library demographics (healthy controls)
-%                                     
-<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
-master = read.csv('data/DT_study.csv')
-tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
-    method="reverse", na.rm = TRUE, test=FALSE)
-caption = 
-  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
-control subject sample used in the template library (excluding the Winterburn
-atlas subjects).} Education is shown in years."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      title="tab:WAval-healthy-demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{Evaluation method}  
-Evaluating the accuracy of hippocampal subfields for T1 images at
-$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
-protocols for segmentation at this resolution to use as a reference. Instead, we
-can evaluate the reliability, or {\it precision}, with which \mb{} produces
-hippocampal subfields that correspond to the segmentation protocol used by the
-given atlas library images. By directly resampling the Winterburn atlas
-segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
-resampling techniques) we obtain a scaled down version of the labels which
-preserve the original segmentation protocol within the limits of error from
-rounding and interpolation. Therefore, using the resampled Winterburn
-segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
-of \mb{} segmentations using using DSC overlap scores and evaluate consistency
-across the range of hippocampal sizes using Bland-Altman plots of subfield
-volumes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%         Experiment 3: Application of \mb{} to the segmentation      %%%%%
-%%%%%                       of schizophrenia patients                     %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 3: Application to the segmentation of first episode
-schizophrenia patients}
-
-To validate that the \mb{} works effectively in the context of other neurological
-disorders, in this experiment we use the Winterburn atlases to derive 
-whole hippocampal segmentations of a dataset of patients with schizophrenia. The
-resulting segmentations are assessed for quality by comparison with expert
-manual segmentations.
-
-\paragraph{First Episode Schizophrenia Dataset}
-All patients were recruited and treated through the Prevention and Early
-Intervention Program for Psychoses (PEPP-Montreal), a specialized early
-intervention service at the Douglas Mental Health University Institute in
-Montreal, Canada. People aged 14 to 35 years from the local catchment area
-suffering from either affective or non-affective psychosis who had not taken
-antipsychotic medication for more than one month with an IQ above 70 were
-consecutively admitted as either in- or out-patients. Of those treated at PEPP,
-only patients aged 18 to 30 years with no previous history of neurological
-disease or head trauma causing loss of consciousness were eligible for the
-neuroimaging study; only those suffering from schizophrenia spectrum disorders
-were considered for this analysis.  For complete program details see
-\citet{Malla2003}.
-
-Scanning of 81 subjects was carried out at the Montreal Neurological Institute
-on a 1.5-T Siemens whole body MRI system.  Structural T1 volumes were acquired
-for each participant using a three-dimensional (3D) gradient echo pulse sequence
-with sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
-angle=$30^{\circ}$, 180 1mm contiguous sagittal slices). The rectangular
-field-of-view for the images was 256mm (SI)$\times$204mm (AP).  Subject
-demographics are shown in Table \ref{tab:SZFEP-Demographics}.
-
-Manual segmentation of each subject whole hippocampus is produced following a
-validated segmentation protocol \citep{Pruessner2000}.
-
-\paragraph{Experiment details} 
-\mb{} is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2) ignoring subfield delineations.  All images from the
-first episode dataset are segmented by \mb{}.  The optimal size of template
-library, registration method, and label fusion method found in Experiment 1 are
-used.
-
-\paragraph{Evaluation method}
-The Pruessner and Winterburn hippocampal segmentation protocols differ slightly in
-the neuroanatomical features that are delineated \citep{Winterburn2013}. This
-difference poses a problem for evaluation by measuring overlap.  That is, since
-different protocols will necessarily produce segmentations that do not perfectly
-overlap, the degree of overlap cannot be solely used to compare segmentation
-methods using different protocols. In place of an overlap metric, we can assess
-the degree of (Pearson) correlation in average bilateral hippocampal volume of the
-subjects produced by each method.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% First Episode SZ Demographics Table
-%
-<<tab:FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
-caption = 
-  "\\textbf{Schizophrenia First Episode Patient Demographics.}
-  ambi - ambidextrous. 
-  SES  - Socioeconomic Status score. 
-  FSIQ - Full Scale IQ.
-"
-fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
-fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
-tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
-               data=fep_sz_demographics, method = "reverse", test = FALSE)
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      title="tab:SZFEP-Demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%         EXPERIMENT 4: Application of \mb{} to the segmentation      %%%%%
-%%%%%                       of ADNI subjects                              %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Experiment 4: Application to the segmentation of Alzheimer's
-disease patients}
-
-To validate \mb{} segmentation quality with respect to other established
-automated hippocampal segmentation methods, \mb{} was applied to large dataset
-from the ADNI project and the resulting segmentations were compared to those
-produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated
-whole hippocampal segmentations (\snt{}) provided by ADNI.
-
-\paragraph{\adnidataset dataset} 
-The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
-\snt{}, \maper, \freesurfer{} hippocampal volumes for a subset of images were
-provided by ADNI, along with quality control data for each \freesurfer{}
-segmentation (guidelines described in \citep{Hartig2010}).
-
-Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
-available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly pre-processed images
-which have been designated to be the "best" after quality control.  All images
-were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
-Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
-2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
-240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
-yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.  Clinical
-and demographic data for the entire \adnidataset{} dataset are shown in Table
-\ref{tab:ADNI1-Dataset-Demographics}.
-
-Semi-automated segmentations of the left and right whole hippocampi are made
-available with a subset of ADNI images \citep{Hsu2002}. These labels have been
-generated using the \snt{} tool from Medtronic Surgical Navigation Technologies,
-Louisville, CO (see Supplementary Materials for detailed discussion of the
-segmentation process).
-
-\paragraph{Experiment details} 
-\mb{} was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 2), and size of template library,
-registration method, and label fusion method were determined by the optimal
-settings found in Experiment 1.  The \fsl{} segmentation method was used via
-the \verb+run_first_all+ script according to the FIRST user guide
-\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
-in the \adnidataset{} dataset were segmented by both methods.
-
-One author (MP) performed visual quality inspection for \mb{} and \fsl{} 
-segmentations using similar quality control guidelines (if either hippocampus
-was under or over segmented by $10mm$ or greater in three or more slices then
-the segmentation did not pass).  Only images meeting the conditions of having
-segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also
-passing quality control inspection were included in the analysis.
-
-\paragraph{Evaluation method}
-As in Experiment 3, the \snt{} and Winterburn hippocampal segmentation protocols
-differ in the neuroanatomical features delineated, and so we assessed \mb{} by
-the degree of (Pearson) correlation of average hippocampal volume across
-subjects. We also computed the correlation in hippocampal volume between
-existing, established automated segmentation methods -- \fsl{}, \freesurfer{},
-and \maper{}, and \snt{} semi-automated segmentations. Additionally, we evaluate
-the volume-related fixed and proportional biases in all segmentation methods
-using Bland-Altman plots \citep{Bland1986}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                                                             
-% ADNI1 Complete 1Yr Dataset Demographics Table 
-%                                                                             
-<<tab:ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
-# RID < 2000 ensures only ADNI1 patients
-yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
-yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
-yr1 = subset(yr1, !grepl("Scaled_2", Description))
-yr1 = yr1[,c("RID","VISCODE")]
-yr1 = 
-adnimerge.yr1 = merge(adnimerge, yr1) 
-tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
-               MMSE, data = adnimerge.yr1, 
-               #subset = RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
-               method = "reverse", test=FALSE, overall=TRUE)
-caption = 
-  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} 
-  CN     - Cognitively Normal. 
-  LMCI   - Late-onset Mild Cognitive Impairment. 
-  AD     - Alzheimer's Disease.  
-  Hisp   - Hispanic. 
-  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
-  ADAS   - Alzheimer's Disease Assessment Scale.  
-  MMSE   - Mini-Mental State Examination."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-    caption=caption, caption.loc = c("top"),
-    title="tab:ADNI1-Dataset-Demographics",
-    label="tab:ADNI1-Dataset-Demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                               RESULTS                                        %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%     RESULTS: Experiment 1: Whole Hippocampus Cross-Validation       %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 1 Results: Whole Hippocampus Cross-Validation}
+\subsubsection{Results}
 
@@
   is computed for each subject across over 10 folds of Monte Carlo       
-  cross validation. . 
+  cross validation.  
   Shown on the y-axis is the  p-value resulting from a t-test comparing the
@@
 %%%%%                                                                     %%%%%
-%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
+%%%%%        Experiment 2: Hippocampal Subfield Cross-Validation          %%%%%
 %%%%%                                                                     %%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
+\subsection{Experiment 2: Hippocampal Subfield Cross-Validation}
+
+The previous experiment assesses \mb{} peformance on whole hippocampus
+segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
+segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
+use a modified leave-one-out cross-validation (LOOCV) design.
+
+\subsubsection{Materials and Methods}
+\paragraph{Winterburn Atlases} 
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
+segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
+and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
+lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
+
+\paragraph{Experiment details} 
+Leave-one-out cross-validation (LOOCV) is a validation approach in which an
+algorithm is given all but one item in a dataset as training data (in our case,
+atlas images and labels) and then the algorithm is applied to the left-out item.
+This is done, in turn, for each item in the dataset and the output across all
+items is evaluated together.
+
+In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
+voxel images to simulate standard 3T T1-weighted resolution images. Image
+subsampling is performed using trilinear subsampling techniques.  In each round
+of LOOCV, a single atlas is selected and segmented with \mb{} once using each
+possible triplet of atlas images from the remaining four unselected atlases.
+Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
+image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
+with an odd number of images so as to ensure an unbiased label fusion when using
+majority voting (see Discussion).
+
+The template library used has a total of 19 images composed of all $0.9mm^3$
+voxel target images plus an additional set of 14 of healthy subject 3T T1 images
+acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
+were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
+WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
+3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
+$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
+0.9mm dimension voxels.
+
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Subfield Validation Template library demographics (healthy controls)
+%                                     
+<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
+master = read.csv('data/DT_study.csv')
+tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
+    method="reverse", na.rm = TRUE, test=FALSE)
+caption = 
+  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
+control subject sample used in the template library (excluding the Winterburn
+atlas subjects).} Education is shown in years."
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      title="tab:WAval-healthy-demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Evaluation method}  
+Evaluating the accuracy of hippocampal subfields for T1 images at
+$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
+protocols for segmentation at this resolution to use as a reference. Instead, we
+can evaluate the reliability, or {\it precision}, with which \mb{} produces
+hippocampal subfields that correspond to the segmentation protocol used by the
+given atlas library images. By directly resampling the Winterburn atlas
+segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
+resampling techniques) we obtain a scaled down version of the labels which
+preserve the original segmentation protocol within the limits of error from
+rounding and interpolation. Therefore, using the resampled Winterburn
+segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
+of \mb{} segmentations using using DSC overlap scores and evaluate consistency
+across the range of hippocampal sizes using Bland-Altman plots of subfield
+volumes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Results}
 
@@
 %%%%%                                                                     %%%%%
-%%%%%    RESULTS: Experiment 3, Application of \mb{} to the segmentation  %%%%%
+%%%%%         Experiment 3: Application of \mb{} to the segmentation      %%%%%
 %%%%%                       of schizophrenia patients                     %%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 3 Results: Application to the segmentation of first
-episode schizophrenia patients}
+\subsection{Experiment 3: Application to the segmentation of first episode
+schizophrenia patients}
+
+To validate that the \mb{} works effectively in the context of other neurological
+disorders, in this experiment we use the Winterburn atlases to derive 
+whole hippocampal segmentations of a dataset of patients with schizophrenia. The
+resulting segmentations are assessed for quality by comparison with expert
+manual segmentations.
+
+\subsubsection{Materials and Methods}
+\paragraph{First Episode Schizophrenia Dataset}
+All patients were recruited and treated through the Prevention and Early
+Intervention Program for Psychoses (PEPP-Montreal), a specialized early
+intervention service at the Douglas Mental Health University Institute in
+Montreal, Canada. People aged 14 to 35 years from the local catchment area
+suffering from either affective or non-affective psychosis who had not taken
+antipsychotic medication for more than one month with an IQ above 70 were
+consecutively admitted as either in- or out-patients. Of those treated at PEPP,
+only patients aged 18 to 30 years with no previous history of neurological
+disease or head trauma causing loss of consciousness were eligible for the
+neuroimaging study; only those suffering from schizophrenia spectrum disorders
+were considered for this analysis.  For complete program details see
+\citet{Malla2003}.
+
+Scanning of 81 subjects was carried out at the Montreal Neurological Institute
+on a 1.5-T Siemens whole body MRI system.  Structural T1 volumes were acquired
+for each participant using a three-dimensional (3D) gradient echo pulse sequence
+with sagittal volume excitation (repetition time=22ms, echo time=9.2ms, flip
+angle=$30^{\circ}$, 180 1mm contiguous sagittal slices). The rectangular
+field-of-view for the images was 256mm (SI)$\times$204mm (AP).  Subject
+demographics are shown in Table \ref{tab:SZFEP-Demographics}.
+
+Manual segmentation of each subject whole hippocampus is produced following a
+validated segmentation protocol \citep{Pruessner2000}.
+
+\paragraph{Experiment details} 
+\mb{} is configured with an atlas library composed of the Winterburn T1 atlases
+(see Experiment 2) ignoring subfield delineations.  All images from the
+first episode dataset are segmented by \mb{}.  The optimal size of template
+library, registration method, and label fusion method found in Experiment 1 are
+used.
+
+\paragraph{Evaluation method}
+The Pruessner and Winterburn hippocampal segmentation protocols differ slightly in
+the neuroanatomical features that are delineated \citep{Winterburn2013}. This
+difference poses a problem for evaluation by measuring overlap.  That is, since
+different protocols will necessarily produce segmentations that do not perfectly
+overlap, the degree of overlap cannot be solely used to compare segmentation
+methods using different protocols. In place of an overlap metric, we can assess
+the degree of (Pearson) correlation in average bilateral hippocampal volume of the
+subjects produced by each method.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% First Episode SZ Demographics Table
+%
+<<tab:FEP-SZ-demographics, echo=F, results="asis",cache=FALSE>>=
+caption = 
+  "\\textbf{Schizophrenia First Episode Patient Demographics.}
+  ambi - ambidextrous. 
+  SES  - Socioeconomic Status score. 
+  FSIQ - Full Scale IQ.
+"
+fep_sz_demographics = read.csv("data/SZ_demographics.csv", na.strings=c("","."))
+fep_sz_demographics = subset(fep_sz_demographics, Anon <= 81)
+tab <- summary(DX ~ Age + Gender + Handedness + Education + SES + FSIQ,
+               data=fep_sz_demographics, method = "reverse", test = FALSE)
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      title="tab:SZFEP-Demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\FloatBarrier
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%    RESULTS: Experiment 3, Application of \mb{} to the segmentation  %%%%%
+%%%%%                       of schizophrenia patients                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Results}
 
@@
 %%%%%                                                                     %%%%%
-%%%%%                           RESULTS                                   %%%%%
-%%%%%                                                                     %%%%%
 %%%%%         EXPERIMENT 4: Application of \mb{} to the segmentation      %%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 4 Results: Application to the segmentation of Alzheimer's
+\subsection{Experiment 4: Application to the segmentation of Alzheimer's
 disease patients}
+
+To validate \mb{} segmentation quality with respect to other established
+automated hippocampal segmentation methods, \mb{} was applied to large dataset
+from the ADNI project and the resulting segmentations were compared to those
+produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated
+whole hippocampal segmentations (\snt{}) provided by ADNI.
+
+\subsubsection{Materials and Methods}
+\paragraph{\adnidataset dataset} 
+The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
+\snt{}, \maper, \freesurfer{} hippocampal volumes for a subset of images were
+provided by ADNI, along with quality control data for each \freesurfer{}
+segmentation (guidelines described in \citep{Hartig2010}).
+
+Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
+authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
+August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
+available from ADNI \footnote{
+\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
+\citep{Wyman2012}. This image collection contains uniformly pre-processed images
+which have been designated to be the "best" after quality control.  All images
+were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
+Systems or Siemens Medical Solutions) at multiple sites using the protocol
+described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR =
+2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
+240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
+yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.  Clinical
+and demographic data for the entire \adnidataset{} dataset are shown in Table
+\ref{tab:ADNI1-Dataset-Demographics}.
+
+Semi-automated segmentations of the left and right whole hippocampi are made
+available with a subset of ADNI images \citep{Hsu2002}. These labels have been
+generated using the \snt{} tool from Medtronic Surgical Navigation Technologies,
+Louisville, CO (see Supplementary Materials for detailed discussion of the
+segmentation process).
+
+\paragraph{Experiment details} 
+\mb{} was configured with an atlas library composed of the five Winterburn T1
+atlases (described in Experiment 2), and size of template library,
+registration method, and label fusion method were determined by the optimal
+settings found in Experiment 1.  The \fsl{} segmentation method was used via
+the \verb+run_first_all+ script according to the FIRST user guide
+\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
+in the \adnidataset{} dataset were segmented by both methods.
+
+One author (MP) performed visual quality inspection for \mb{} and \fsl{} 
+segmentations using similar quality control guidelines (if either hippocampus
+was under or over segmented by $10mm$ or greater in three or more slices then
+the segmentation did not pass).  Only images meeting the conditions of having
+segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also
+passing quality control inspection were included in the analysis.
+
+\paragraph{Evaluation method}
+As in Experiment 3, the \snt{} and Winterburn hippocampal segmentation protocols
+differ in the neuroanatomical features delineated, and so we assessed \mb{} by
+the degree of (Pearson) correlation of average hippocampal volume across
+subjects. We also computed the correlation in hippocampal volume between
+existing, established automated segmentation methods -- \fsl{}, \freesurfer{},
+and \maper{}, and \snt{} semi-automated segmentations. Additionally, we evaluate
+the volume-related fixed and proportional biases in all segmentation methods
+using Bland-Altman plots \citep{Bland1986}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                             
+% ADNI1 Complete 1Yr Dataset Demographics Table 
+%                                                                             
+<<tab:ADNI1-Dataset-Demographics, echo=F, results="asis",cache=FALSE>>=
+# RID < 2000 ensures only ADNI1 patients
+yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
+yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
+yr1 = subset(yr1, !grepl("Scaled_2", Description))
+yr1 = yr1[,c("RID","VISCODE")]
+yr1 = 
+adnimerge.yr1 = merge(adnimerge, yr1) 
+tab <- summary(DX.bl ~ AGE + PTGENDER + PTEDUCAT + CDRSB + ADAS13 +
+               MMSE, data = adnimerge.yr1, 
+               #subset = RID %in% yr1$RID & VISCODE %in% c('bl','m06','m12'),   
+               method = "reverse", test=FALSE, overall=TRUE)
+caption = 
+  "\\textbf{ADNI1 1.5T Complete 1Yr dataset demographics.} 
+  CN     - Cognitively Normal. 
+  LMCI   - Late-onset Mild Cognitive Impairment. 
+  AD     - Alzheimer's Disease.  
+  Hisp   - Hispanic. 
+  CDR-SB - Clinical Dementia Rating-Sum of Boxes.  
+  ADAS   - Alzheimer's Disease Assessment Scale.  
+  MMSE   - Mini-Mental State Examination."
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+    caption=caption, caption.loc = c("top"),
+    title="tab:ADNI1-Dataset-Demographics",
+    label="tab:ADNI1-Dataset-Demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\FloatBarrier
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                           RESULTS                                   %%%%%
+%%%%%         EXPERIMENT 4: Application of \mb{} to the segmentation      %%%%%
+%%%%%                       of ADNI subjects                              %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Results}
 
@@
 
+\FloatBarrier
 \subsection{Experiment 5: Whole Hippocampus Segmentation Cross-Validation with \snt{} 
@@
 })
-  y-axis is the p-value from a t-test comparing the distribution of variances at
-  each parameter setting (atlas/template library size). 
-  Only points where \mb{} mean variability is lower than \multiatlas{} are shown.
-  Dashed lines indicate a p-value of 0.05 and 0.01.}
+
+ggplot(subset(stats.tests, direction > 0 & stat == "Variance"), 
+  aes(x=templates.mb, y=p.value, colour=as.factor(atlases))) + 
+  #facet_grid( . ~ stat) + 
+  geom_line(size=0.5) +
+  geom_point(size=2) + 
+  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
+  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
+  scale_x_continuous(breaks=seq(1,20,by=2)) + 
+  xlab( "Number of Templates" ) + 
+  ylab( "Variability (p)") +
+  scale_colour_hue(name="Atlases") +
+  scale_alpha_continuous(guide='none') + 
+  scale_linetype_discrete(guide='none') + 
+  scale_size_continuous(guide='none') + 
+  theme(legend.direction = "horizontal", legend.position = "bottom")
+@
+  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation
+  with respect to \snt segmentation.}
+  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
+  is computed for each subject across over 10 folds of Monte Carlo       
+  cross validation.  
+  Shown on the y-axis is the p-value from a t-test comparing the distribution of
+  variances at each parameter setting (atlas/template library size).  Only points
+  where \mb{} mean variability is lower than \multiatlas{} are shown.  Dashed
+  lines indicate a p-value of 0.05 and 0.01.}
   \label{ADNI1-SNT-xval-variability}

### 2014-01-13 paper.Rnw
XVal updated

@@
 %  reference (Experiment X). 
+%  - subfield or subregion -- choose one
 
@@
 
-The gold standard for neuroanatomical segmentation is manual delineation by an
+THe gold standard for neuroanatomical segmentation is manual delineation by an
 expert human rater.  However, with the availability of increasingly large MRI
@@
 \section{Experiments}
+\label{sec:experiments}
 
 The following section describes experiments conducted to assess the
-segmentation quality of the \mb{} algorithm. The first two experiments assess the
-validity of \mb{} using cross-validation designs. Experiment 1 investigates the
-accuracy of \mb{} whole hippocampus segmentation over a wide range of parameter
-settings. The results of this experiment enable us to choose the parameter
-settings offering the best performance for use in subsequent experiments.
-Experiment 2 assesses hippocampal subfield segmentation quality in a
-leave-one-out cross-validation design. The last two experiments assess the
-validity of the \mb{} algorithm when applied to different diseases: first episode
-schizophrenia (Experiment 3), and Alzheimer's disease (Experiment 4).
-Additionally, in Experiment 4, we compare \mb{} segmentations with those of
-well-known automated methods and assessed segmentation bias.
+segmentation quality of the \mb{} algorithm:  
+\begin{itemize}
+\item Experiment 1 investigates the accuracy of \mb{} whole hippocampus
+segmentation of aging and Alzhiemer's diseased subjects over a wide range of
+parameter settings using a Monte Carlo cross-validation design. The results of
+this experiment enable us to choose the parameter settings offering the best
+performance for use in subsequent experiments.  
+\item Experiment 2 consists of a similar cross-validation design to verify these
+parameter settings on brain images of young, first episode Schizophrenia
+patients. 
+\item Experiment 3 assesses hippocampal subfield segmentation quality in a
+leave-one-out cross-validation design. 
+\item Experiment 4 compares \mb{} segmentations with those of well-known
+automated methods on the entire \adnidataset image dataset consisting of 
+\Sexpr{length(means.complete$RID)} brain images of subjects diagnosed as
+cognitively normal, having mild cognitive impairment, or Alzheimer's disease.
+\end{itemize}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 1: Whole Hippocampus Cross-Validation}
-
+\subsection{Experiment 1: Whole Hippocampus Segmentation --- Cross-Validation on 
+Alzheimer's Diseased Subjects}
+\label{exp:adni-xval}
+
+% TODO: properly introduce
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using images
@@
 
-In the Supplementary Materials we have replicated this experiment using the
+% TODO: fix wording, relate to pruessner labels
+Note, in the Supplementary Materials we have replicated this experiment using the
 \snt{} semi-automated labels included with the ADNI dataset.
@@
 \FloatBarrier
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%                                                                     %%%%%
-%%%%%        Experiment 2: Hippocampal Subfield Cross-Validation          %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 2: Hippocampal Subfield Cross-Validation}
-
-The previous experiment assesses \mb{} peformance on whole hippocampus
-segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
-segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
-use a modified leave-one-out cross-validation (LOOCV) design.
-
-\subsubsection{Materials and Methods}
-\paragraph{Winterburn Atlases} 
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
-segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
-and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
-lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
-range in age from 29-57 years (mean age of 37), and include two males and three
-females.
-
-\paragraph{Experiment details} 
-Leave-one-out cross-validation (LOOCV) is a validation approach in which an
-algorithm is given all but one item in a dataset as training data (in our case,
-atlas images and labels) and then the algorithm is applied to the left-out item.
-This is done, in turn, for each item in the dataset and the output across all
-items is evaluated together.
-
-In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
-voxel images to simulate standard 3T T1-weighted resolution images. Image
-subsampling is performed using trilinear subsampling techniques.  In each round
-of LOOCV, a single atlas is selected and segmented with \mb{} once using each
-possible triplet of atlas images from the remaining four unselected atlases.
-Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
-image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
-with an odd number of images so as to ensure an unbiased label fusion when using
-majority voting (see Discussion).
-
-The template library used has a total of 19 images composed of all $0.9mm^3$
-voxel target images plus an additional set of 14 of healthy subject 3T T1 images
-acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
-were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
-WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
-3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
-$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
-thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
-0.9mm dimension voxels.
-
-The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Subfield Validation Template library demographics (healthy controls)
-%                                     
-<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
-master = read.csv('data/DT_study.csv')
-tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
-    method="reverse", na.rm = TRUE, test=FALSE)
-caption = 
-  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
-control subject sample used in the template library (excluding the Winterburn
-atlas subjects).} Education is shown in years."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      title="tab:WAval-healthy-demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{Evaluation method}  
-Evaluating the accuracy of hippocampal subfields for T1 images at
-$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
-protocols for segmentation at this resolution to use as a reference. Instead, we
-can evaluate the reliability, or {\it precision}, with which \mb{} produces
-hippocampal subfields that correspond to the segmentation protocol used by the
-given atlas library images. By directly resampling the Winterburn atlas
-segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
-resampling techniques) we obtain a scaled down version of the labels which
-preserve the original segmentation protocol within the limits of error from
-rounding and interpolation. Therefore, using the resampled Winterburn
-segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
-of \mb{} segmentations using using DSC overlap scores and evaluate consistency
-across the range of hippocampal sizes using Bland-Altman plots of subfield
-volumes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results}
-
-\begin{figure}
-<<fig:WAxval-kappa>>=
-wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
-wa.n.sim$label = as.factor(wa.n.sim$label)
-levels(wa.n.sim$label) <-list(
-  "CA1"=1, "CA1"=101, 
-  "CA2/CA3"=5,"CA2/CA3"=105, 
-  "CA4/DG"=4,"CA4/DG"=104, 
-  "Subiculum"=2,"Subiculum"=102, 
-  "SR/SL/SM"=6,"SR/SL/SM"=106)
-
-# kappa vs. region
-ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
-  geom_boxplot(fill='lightblue') + 
-  xlab("Subregion") + 
-  ylab("Dice's Similarity Coefficient (DSC)")
-
-CA1   = subset(wa.n.sim,label=='CA1')$k
-CA23  = subset(wa.n.sim,label=='CA2/CA3')$k
-CA4DG = subset(wa.n.sim,label=='CA4/DG')$k
-Sub   = subset(wa.n.sim,label=='Subiculum')$k
-SRLM  = subset(wa.n.sim,label=='SR/SL/SM')$k
-@
-  \caption{Similarity of \mb{} segmentation of subregions and the resampled
-  Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
-  leave-one-out cross validation. Overlap score for each hemisphere is measured
-  separately. \label{fig:WAxval-kappa}}
-\end{figure}
-
-In this experiment the \mb{} segmentation of hippocampal subfields is assessed
-through a leave-one-out validation in which resampled images ($0.9mm^3$ voxels)
-of each Winterburn atlas subject are segmented using the remaining
-high-resolution Winterburn atlases.
-
-Figure \ref{fig:WAval-kappa} shows the overlap similarity scores between the
-\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
-subregion across all subjects and folds of the validation. Mean and standard
-deviation DSC scores of the subregions are shown in Table
-\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
-segmentations when perturbed slightly and compared to the originals.  We find
-that the CA4/DG subregion shows the highest mean DSC score of
-\Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
-subregions having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
-\Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively.  Both the CA4/DG and
-molecular regions score below 0.5.  These scores may seem low but not when taken
-in context and compared to existing (semi-)automated methods (see Discussion).
-
-\begin{figure}
-<<fig:WAxval-Bland-Altman>>=
-wa.n.vol = read.csv('data/winterburn.xval.volumes.csv')
-winterburn = read.csv('data/winterburn.volumes.csv')
-subfield_labels <-list(
-  "CA1"="X1", "CA1"="X101", 
-  "CA2/CA3"="X5","CA2/CA3"="X105", 
-  "CA4/DG"="X4","CA4/DG"="X104",
-  "Subiculum"="X2","Subiculum"="X102",
-  "SR/SL/SM"="X6","SR/SL/SM"="X106")
-hemispheres <- list(
-  "Right"="X1","Left"="X101", 
-  "Right"="X5","Left"="X105", 
-  "Right"="X4","Left"="X104",
-  "Right"="X2","Left"="X102",
-  "Right"="X6","Left"="X106")
-
-wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
-winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
-winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
-wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
-
-levels(wa.n.all.vols$subject) <- list(
-  "1"="anusha", 
-  "2"="jon",
-  "3"="mallar",
-  "4"="sofia",
-  "5"="nancy")
-
-# mean per subfield/hemisphere across all 4 validation folds
-wa.n.mean.vols = ddply(wa.n.all.vols, c("variable", "subject", "res"), function(df) {
-  data.frame(mb = mean(df$value),
-             high  = mean(df$'0.3'),
-             low   = mean(df$'0.9'))
-})
-wa.n.mean.vols$diff = wa.n.mean.vols$low - wa.n.mean.vols$mb
-wa.n.mean.vols$mean = (wa.n.mean.vols$low + wa.n.mean.vols$mb)/2
-
-wa.n.mean.vols$hemisphere <- wa.n.mean.vols$variable
-levels(wa.n.all.vols$variable)   <- subfield_labels
-levels(wa.n.mean.vols$variable)  <- subfield_labels
-levels(wa.n.all.vols$hemisphere) <- hemispheres
-
-wa.n.xval.bland.limits = ddply(wa.n.mean.vols,c("variable"),function(df) {
-  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
-})
-
-wa.n.xval.bland.cor = ddply(wa.n.mean.vols,c("variable"),function(df) {
-  cor.t = cor.test(df$low,df$mb)
-  data.frame(
-    pearson = sprintf("r=%0.2f, t=%0.2f, p < %0.2f", 
-                cor.t$estimate, cor.t$statistic, cor.t$p.value)
-  )
-})
-
-ggplot(wa.n.mean.vols, aes(x=mean, y=diff)) + 
-  facet_wrap(~variable, ncol=3, scales='free_x') + 
-  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=2, aes(colour=hemisphere, shape=subject)) +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
-             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
-             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
-  geom_text(aes(y = Inf, x=-Inf, label=pearson), colour="black", 
-             data = wa.n.xval.bland.cor, hjust=-0.1,vjust=2.5, size=2.5) + 
-  xlab(expression('Mean of resampled and MAGeT-Brain subregion volume ' (mm^3))) + 
-  ylab(expression('Difference of manual and MAGeT-Brain subregion volume ' (mm^3))) + 
-  #scale_x_continuous(limits=c(0,1000)) + 
-  #scale_color_discrete(name="") + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text.y = element_text(angle = 90, hjust=0.5), 
-        axis.text = element_text(size = 8))
-@
-  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
-across all folds of Winterburn atlas cross validation.} The overall mean
-difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
-horizontal lines. Linear fit lines are shown. Note, points below the mean
-difference indication overestimation of the volume with respect to the resampled
-volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
-\end{figure}
-
-We find that in each subfield, \mb{} shows a proportional bias towards
-underestimating volumes in all subregions (Figure
-\ref{fig:WAxval-Bland-Altman}).  Prominent fixed biases in the volumes of the
-CA4/DG and SR/SL/SM subregions are congruent with the lower similarity scores
-seen in these regions.
-
-<<tab:WAxval-perturb-kappa, echo=F,results="asis">>=
-wa.perturb = read.csv('data/winterburn.perturb.kappa.csv')
-wa.perturb = wa.perturb[c("subject","label","operation","k")]
-wa.perturb$label = as.factor(wa.perturb$label)
-levels(wa.perturb$label) <-list(
-  "CA1"="1", "CA1"="101",  "CA2/CA3"="5","CA2/CA3"="105",
-  "CA4/DG"="4","CA4/DG"="104",  "Subiculum"="2","Subiculum"="102", 
-  "SR/SL/SM"="6", "SR/SL/SM"="106")
-mean_range = function (v) {
-  sprintf("$%0.2f \\pm %0.2f$", mean(v), sd(v))
-}
-wa.perturb.table = dcast(wa.perturb, label ~ operation, fun.aggregate=mean_range)
-wa.sim.sum = dcast(wa.n.sim, label ~ ., value.var = 'k', fun.aggregate=mean_range)
-names(wa.sim.sum)[2] = 'MAGeT'
-wa.perturb.table = merge(wa.perturb.table, wa.sim.sum)
-
-wa.perturb.table=wa.perturb.table[c('label','MAGeT', 'translate','contract','expand')]
-names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")
-
-caption="Overlap similarity results for the each of the subfields of the hippocampus. 
-Simulated overlap similarity results are also given for resampled labels that were 
-translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
-by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
-Values are given as mean Dice's Kappa coefficient $\\pm$ stddev."
-latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
-      caption=caption, caption.loc = c("top"),
-      col.just=rep('c',5),  
-      label="tab:WAxval-perturb-kappa", rowname=NULL)
-@
-
-Figure \ref{fig:subfield-montage} shows slices subfield segmentations for a
-single subject for qualitative inspection.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=5in]{figure/winterburn-atlas-montage/figure.pdf}
-    \caption{\textbf{Detailed subfield segmentation results for a single subject.} In
-    the upper left corner is the original high-resolution Winterburn atlas manual
-    subfield segmentation; in the upper right corner is the Winterburn atlas
-    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
-    corner is the \mb{} segmentation of the subsampled Winterburn atlas image
-    from a single fold of the cross-validation. In each segmentation, slices from the left
-    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
-    slices from inferior to superior; the second row shows sagittal slices from
-    lateral to medial; the third row shows coronal slices from anterior to
-    posterior. \label{fig:subfield-montage}}
-\end{figure}
-
-\FloatBarrier
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%         Experiment 3: Application of \mb{} to the segmentation      %%%%%
+%%%%%         Experiment 2: Application of \mb{} to the segmentation      %%%%%
 %%%%%                       of schizophrenia patients                     %%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 3: Application to the segmentation of first episode
-schizophrenia patients}
-
+\subsection{Experiment 2: Whole Hippocampus Segmentation --- Cross-validation
+on First Episode Schizophrenia Subjects}
+\label{exp:fep-xval}
+%TODO: update
 To validate that the \mb{} works effectively in the context of other neurological
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%    RESULTS: Experiment 3, Application of \mb{} to the segmentation  %%%%%
+%%%%%    RESULTS: Experiment 2, Application of \mb{} to the segmentation  %%%%%
 %%%%%                       of schizophrenia patients                     %%%%%
@@
 \FloatBarrier
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%        Experiment 3: Hippocampal Subfield Cross-Validation          %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment 3: Hippocampal Subfield Segmentation --- Cross-Validation
+on the Winterburn Atlases}
+\label{exp:winterburn-xval}
+The previous experiment assesses \mb{} peformance on whole hippocampus
+segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
+segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
+use a modified leave-one-out cross-validation (LOOCV) design.
+
+\subsubsection{Materials and Methods}
+\paragraph{Winterburn Atlases} 
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
+segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
+and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
+lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
+
+\paragraph{Experiment details} 
+Leave-one-out cross-validation (LOOCV) is a validation approach in which an
+algorithm is given all but one item in a dataset as training data (in our case,
+atlas images and labels) and then the algorithm is applied to the left-out item.
+This is done, in turn, for each item in the dataset and the output across all
+items is evaluated together.
+
+In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
+voxel images to simulate standard 3T T1-weighted resolution images. Image
+subsampling is performed using trilinear subsampling techniques.  In each round
+of LOOCV, a single atlas is selected and segmented with \mb{} once using each
+possible triplet of atlas images from the remaining four unselected atlases.
+Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
+image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
+with an odd number of images so as to ensure an unbiased label fusion when using
+majority voting (see Discussion).
+
+The template library used has a total of 19 images composed of all $0.9mm^3$
+voxel target images plus an additional set of 14 of healthy subject 3T T1 images
+acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
+were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
+WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
+3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
+$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
+0.9mm dimension voxels.
+
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Subfield Validation Template library demographics (healthy controls)
+%                                     
+<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
+master = read.csv('data/DT_study.csv')
+tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
+    method="reverse", na.rm = TRUE, test=FALSE)
+caption = 
+  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
+control subject sample used in the template library (excluding the Winterburn
+atlas subjects).} Education is shown in years."
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      title="tab:WAval-healthy-demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Evaluation method}  
+Evaluating the accuracy of hippocampal subfields for T1 images at
+$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
+protocols for segmentation at this resolution to use as a reference. Instead, we
+can evaluate the reliability, or {\it precision}, with which \mb{} produces
+hippocampal subfields that correspond to the segmentation protocol used by the
+given atlas library images. By directly resampling the Winterburn atlas
+segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
+resampling techniques) we obtain a scaled down version of the labels which
+preserve the original segmentation protocol within the limits of error from
+rounding and interpolation. Therefore, using the resampled Winterburn
+segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
+of \mb{} segmentations using using DSC overlap scores and evaluate consistency
+across the range of hippocampal sizes using Bland-Altman plots of subfield
+volumes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%    RESULTS: Experiment 3: Hippocampal Subfield Cross-Validation     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Results}
+
+\begin{figure}
+<<fig:WAxval-kappa>>=
+wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
+wa.n.sim.whole = read.csv('data/winterburn.xval.similarity.whole.csv')
+wa.n.sim.whole$label = as.factor(wa.n.sim.whole$label)
+levels(wa.n.sim.whole$label) <- list(
+  "Whole"=1,
+  "Whole"=2)
+wa.n.sim.whole$res='0.9'
+wa.n.sim.whole=wa.n.sim.whole[c('atlases','subject','res','label','k','se','sn','j')]
+
+wa.n.sim$label = as.factor(wa.n.sim$label)
+levels(wa.n.sim$label) <-list(
+  "CA1"=1, "CA1"=101, 
+  "CA2/CA3"=5,"CA2/CA3"=105, 
+  "CA4/DG"=4,"CA4/DG"=104, 
+  "Subiculum"=2,"Subiculum"=102, 
+  "SR/SL/SM"=6,"SR/SL/SM"=106)
+
+wa.n.sim = rbind(wa.n.sim, wa.n.sim.whole)
+
+# kappa vs. region
+ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
+  geom_boxplot(fill='lightblue') + 
+  scale_y_continuous(breaks=seq(0,1,0.1)) +
+  xlab("Subregion") + 
+  ylab("Dice's Similarity Coefficient (DSC)")
+
+CA1   = subset(wa.n.sim,label=='CA1')$k
+CA23  = subset(wa.n.sim,label=='CA2/CA3')$k
+CA4DG = subset(wa.n.sim,label=='CA4/DG')$k
+Sub   = subset(wa.n.sim,label=='Subiculum')$k
+SRLM  = subset(wa.n.sim,label=='SR/SL/SM')$k
+Whole = subset(wa.n.sim,label=='Whole')$k
+@
+  \caption{Similarity of \mb{} segmentation of subregions and the resampled
+  Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
+  leave-one-out cross validation. Overlap score for each hemisphere is measured
+  separately. \label{fig:WAxval-kappa}}
+\end{figure}
+
+In this experiment the \mb{} segmentation of hippocampal subfields is assessed
+through a leave-one-out validation in which resampled images ($0.9mm^3$ voxels)
+of each Winterburn atlas subject are segmented using the remaining
+high-resolution Winterburn atlases.
+
+Figure \ref{fig:WAxval-kappa} shows the overlap similarity scores between the
+\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
+subregion across all subjects and folds of the validation. Mean and standard
+deviation DSC scores of the subregions are shown in Table
+\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
+segmentations when perturbed slightly and compared to the originals.  We find
+that the CA4/DG subregion shows the highest mean DSC score of
+\Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
+subregions having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
+\Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively.  Both the CA4/DG and
+molecular regions score below 0.5.  These scores may seem low but not when taken
+in context and compared to existing (semi-)automated methods (see Discussion). The whole hippocampus is segmented with a mean DSC score of\Sexpr{mean(Whole)} $\pm$\Sexpr{sd(Whole)}.
+
+\begin{figure}
+<<fig:WAxval-Bland-Altman,fig.height=5,fig.width=7>>=
+wa.n.vol = read.csv('data/winterburn.xval.volumes.csv') # xval volumes
+winterburn = read.csv('data/winterburn.volumes.csv')    # resampled volumes
+subfield_labels <-list(
+  "CA1"="X1", "CA1"="X101", 
+  "CA2/CA3"="X5","CA2/CA3"="X105", 
+  "CA4/DG"="X4","CA4/DG"="X104",
+  "Subiculum"="X2","Subiculum"="X102",
+  "SR/SL/SM"="X6","SR/SL/SM"="X106", 
+  "Whole" = "left", "Whole" = "right")
+hemispheres <- list(
+  "Right"="X1","Left"="X101", 
+  "Right"="X5","Left"="X105", 
+  "Right"="X4","Left"="X104",
+  "Right"="X2","Left"="X102",
+  "Right"="X6","Left"="X106",
+  "Right"="right","Left"="left")
+
+# compute whole hippocampal volumes (we do this the stupid way)
+wa.n.vol = within(wa.n.vol, {
+  right = X1+X2+X4+X5+X6
+  left  = X101+X102+X104+X105+X106
+  })
+winterburn = within(winterburn, {
+  right = X1+X2+X4+X5+X6
+  left  = X101+X102+X104+X105+X106
+  })
+
+wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
+winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
+winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
+wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
+
+levels(wa.n.all.vols$subject) <- list(
+  "1"="anusha", 
+  "2"="jon",
+  "3"="mallar",
+  "4"="sofia",
+  "5"="nancy")
+
+# mean per subfield/hemisphere across all 4 validation folds
+wa.n.mean.vols = ddply(wa.n.all.vols, c("variable", "subject", "res"), function(df) {
+  data.frame(mb    = mean(df$value),
+             high  = mean(df$'0.3'),
+             low   = mean(df$'0.9'))
+})
+wa.n.mean.vols$diff = wa.n.mean.vols$low - wa.n.mean.vols$mb
+wa.n.mean.vols$mean = (wa.n.mean.vols$low + wa.n.mean.vols$mb)/2
+wa.n.mean.vols$hemisphere        <- wa.n.mean.vols$variable
+levels(wa.n.mean.vols$variable)  <- subfield_labels
+
+wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
+wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2
+wa.n.all.vols$Hemisphere        <- wa.n.all.vols$variable
+levels(wa.n.all.vols$variable)   <- subfield_labels
+levels(wa.n.all.vols$Hemisphere) <- hemispheres
+
+wa.n.xval.bland.limits = ddply(wa.n.mean.vols,c("variable"),function(df) {
+  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
+})
+
+ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
+  facet_wrap(~variable, ncol=3, scales='free_x') + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=2) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
+             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of resampled and MAGeT-Brain subregion volume ' (mm^3))) + 
+  ylab(expression('Difference of manual and MAGeT-Brain subregion volume ' (mm^3))) + 
+  #scale_x_continuous(limits=c(0,1000)) + 
+  #scale_color_discrete(name="") + 
+  scale_shape_discrete(name='Subject') + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+@
+  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
+across all folds of Winterburn atlas cross validation.} The overall mean
+difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
+horizontal lines. Linear fit lines are shown. Note, points below the mean
+difference indication overestimation of the volume with respect to the resampled
+volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
+\end{figure}
+
+Figure \ref{fig:WAxval-Bland-Altman} contains Bland-Altman plots comparing \mb{}
+volumes with manual volumes across all validation folds.  \mb{} displays a
+conservative proportional bias --- small hippocampi are overestimated in volume,
+and larger hippocampi are underestimated 
+(a mean maximum difference of approximately 200mm^3 across all subfields). 
+\mb{} display a slight conservative fixed bias, tending to underestimate all  
+subfields except CA4/DG (mean underestimation: $CA1=76mm^3$, $CA2/3=56mm^3$,
+$CA4/DG=-16mm^3$, $Subiculum=48mm^3$, $SR/SL/SM=96mm^3$).
+
+<<tab:WAxval-perturb-kappa, echo=F,results="asis">>=
+wa.perturb = read.csv('data/winterburn.perturb.kappa.csv')
+wa.perturb = wa.perturb[c("subject","label","operation","k")]
+wa.perturb$label = as.factor(wa.perturb$label)
+levels(wa.perturb$label) <-list(
+  "CA1"="1", "CA1"="101",  "CA2/CA3"="5","CA2/CA3"="105",
+  "CA4/DG"="4","CA4/DG"="104",  "Subiculum"="2","Subiculum"="102", 
+  "SR/SL/SM"="6", "SR/SL/SM"="106")
+mean_range = function (v) {
+  sprintf("$%0.2f \\pm %0.2f$", mean(v), sd(v))
+}
+wa.perturb.table = dcast(wa.perturb, label ~ operation, fun.aggregate=mean_range)
+wa.sim.sum = dcast(wa.n.sim, label ~ ., value.var = 'k', fun.aggregate=mean_range)
+names(wa.sim.sum)[2] = 'MAGeT'
+wa.perturb.table = merge(wa.perturb.table, wa.sim.sum)
+
+wa.perturb.table=wa.perturb.table[c('label','MAGeT', 'translate','contract','expand')]
+names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")
+
+caption="Overlap similarity results for the each of the subfields of the hippocampus. 
+Simulated overlap similarity results are also given for resampled labels that were 
+translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
+by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
+Values are given as mean Dice's Kappa coefficient $\\pm$ stddev."
+latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
+      caption=caption, caption.loc = c("top"),
+      col.just=rep('c',5),  
+      label="tab:WAxval-perturb-kappa", rowname=NULL)
+@
+
+Figure \ref{fig:subfield-montage} shows slices subfield segmentations for a
+single subject for qualitative inspection.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=5in]{figure/winterburn-atlas-montage/figure.pdf}
+    \caption{\textbf{Detailed subfield segmentation results for a single subject.} In
+    the upper left corner is the original high-resolution Winterburn atlas manual
+    subfield segmentation; in the upper right corner is the Winterburn atlas
+    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
+    corner is the \mb{} segmentation of the subsampled Winterburn atlas image
+    from a single fold of the cross-validation. In each segmentation, slices from the left
+    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
+    slices from inferior to superior; the second row shows sagittal slices from
+    lateral to medial; the third row shows coronal slices from anterior to
+    posterior. \label{fig:subfield-montage}}
+\end{figure}
+
+\FloatBarrier
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 disease patients}
-
+\label{exp:adni}
 To validate \mb{} segmentation quality with respect to other established
@@
 \mb{} was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 2), and size of template library,
+atlases (described in Experiment 3, section \ref{exp:winterburn-xval}), and size of template library,
 registration method, and label fusion method were determined by the optimal
-settings found in Experiment 1.  The \fsl{} segmentation method was used via
+settings found in Experiment 1 (Section \ref{exp:adni-xval}).  The \fsl{} segmentation method was used via
 the \verb+run_first_all+ script according to the FIRST user guide
@@
 \paragraph{Evaluation method}
-As in Experiment 3, the \snt{} and Winterburn hippocampal segmentation protocols
-differ in the neuroanatomical features delineated, and so we assessed \mb{} by
+As in previous experiments, the \snt{} and Winterburn hippocampal segmentation protocols
+differ in the neuroanatomical features delineated, and so we assess \mb{} by
 the degree of (Pearson) correlation of average hippocampal volume across
@@
 
-Based on the results from Experiment 1, in this experiment \mb{} was configured
+Based on the results from Experiment 1 (Section \ref{exp:adni-xval}), in this experiment \mb{} was configured
 with a template library of 21 randomly chosen subject images (7 from each
@@
   \includegraphics[width=6in]{figure/ADNI1_SNT_MB_montage/montage.pdf}
-  \caption{\snt{} and \mb{} segmentations for 30 ADNI subjects (10
+  \caption{\snt{} and \mb{} segmentations for 30 ADNI subjects --- 10
   subjects randomly selected from each disease category in the subject pool used
-  in Experiment 1). Sagittal slices are shown for each unlabelled T1-weighted
+  in Experiment 1 (Section \ref{exp:adni-xval}. Sagittal slices are shown for each unlabelled T1-weighted
   anatomical image. \snt{} labels appear in green, and \mb{} labels appear in
@@
 This table shows the best published Dice's overlap measure between automated and
-ground truth segmentations of various methods.  Unless otherwise specified,
+ground truth segmentations. Unless otherwise specified,
 validation datasets are composed equally of cases and control subjects, and use
@@
 
+%TODO: 
 %%
@@
 \mb{} framework -- a methodology that requires very few input atlases in order
-to provide accurate and reliable segmentations.  Experiment 1 compares \mb{} to
+to provide accurate and reliable segmentations.  Experiment 1 (Section
+\ref{exp:adni-xval}) compares \mb{} to
 basic-\multiatlas{} segmentation by characterising the behaviour of each
-approach with varying parameter settings, and allows us to choose an optimal
-setting for subsequent experiments. Experiment 2 demonstrates the reliability of
+approach with varying parameter settings, and allows us to choose optimal
+settings for subsequent experiments. 
+%TODO: fix refs 
+Experiment 2 (Section \ref{exp:winterburn-xval}) emonstrates the reliability of
 \mb{} in producing subfield segmentations which match the segmentation protocol
 of the input atlases despite contrast and resolution limitations in standard
-T1-weighted image volumes. Experiments 2 and 3 validate whole hippocampal
+T1-weighted image volumes.
+
+Experiments 2 and 3 validate whole hippocampal
 segmentation precision and consistency on populations with different ageing and
@@
 which regions can be differentiated, even on high-resolution scans. See Table
-\ref{ab:subfield-segmentation-protocols} for a comparison of manual subfield
+\ref{tab:subfield-segmentation-protocols} for a comparison of manual subfield
 segmentation methodologies. A further complication is that different
@@
 disagreements make evaluating the accuracy of automated methodologies with
-respect to a "ground truth" (e.g. using overlap similarity metrics, such as
+respect to a ``ground truth" (e.g. using overlap similarity metrics, such as
 Dice's Kappa) not particularly meaningful until these disagreements have been

### 2014-01-14 paper.Rnw
Merged figures, exposed TODO items

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% TODO: 
-%  - check tense: we are is have has  use/used find/found
-%  - remove 'accuracy' unless specifically talking relative to ground truth (and
-%    even then, qualify)
-%  - overlap/similarity --> agreement? choose one. 
-%  - check that chart labels are sensible 
-%  - in particular, check that Experiment 5 match Experiment 1
-%  - fix figure 2 schematic language for the experiments (not 2a 2b). 
-%  - ADNI-1 or ADNI1, standardize
-%  - Don't use "Experiment X", instead use some short form description, and then
-%  reference (Experiment X). 
-%  - subfield or subregion -- choose one
-
 <<setup, include=FALSE, cache=FALSE>>=
@@
 explored.\\
-\textbf{Results:} Using 9 atlases and 19 template images, \mb{} achieves a
+\textbf{Results:} \textbf{[TODO: update with FEP-XVAL and subfield XVAL results.]} 
+Using 9 atlases and 19 template images, \mb{} achieves a
 mean Dice's Similarity Coefficient (DSC) of \Sexpr{mb.kappa.best} 
-(to \psegs{}) over 10-folds of Monte Carlo % TODO: update cross-validation on 60 subjects, and shows
+(to \psegs{}) over 10-folds of Monte Carlo 
 significantly lower variability in DSC than \multiatlas{} segmentation.
-In a LOOCV, \mb{} reproduces the volumes of the cornu ammonis (CA) 1;
-CA4/dentate % TODO: update
-gyrus (DG); and strata radiatum (SR), strata lacunosum (SL), and strata
-moleculare (SM) hippocampal subfields with a percent error in volume that is
-at or lower than that produced by image resampling. \mb{} produces hippocampal
+In a LOOCV, \mb{} accurately reproduces the subfields... .
+
+\mb{} produces hippocampal
 volumes in a first episode psychosis patient population that are highly
@@
 jon.pipitone@camh.ca; mallar.chakravarty@camh.ca}
+
+\parbox{7in}{
+\subsection{TODO:}
+\begin{enumerate}
+\itemsep-0.5em
+\item check tense: we are is have has use/used find/found
+\item remove 'accuracy' unless specifically talking relative to ground truth (and
+    even then, qualify)
+\item change overlap/similarity --> agreement? choose one. 
+\item check that chart labels are sensible 
+\item check that Experiment 5 matches Experiment 1
+\item fix figure 2 schematic language for the experiments (not 2a 2b). 
+\item ADNI-1 or ADNI1: standardize
+\item experiment references formatted as Experiment 2 (Section 3.2)
+\item subfield or subregion -- choose one
+\end{enumerate}}
 
@@
 
-% TODO: properly introduce
+\textbf{[TODO: cut, re-introduce as above]}
 Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using images
@@
 
-% TODO: fix wording, relate to pruessner labels
+\textbf{[TODO: fix wording to relate to pruessner labels]}
 Note, in the Supplementary Materials we have replicated this experiment using the
@@
 
-In this experiment we conduct 10 rounds of \mb{} and \multiatlas{} segmentation of
-of 60 subjects using atlas and template library sizes varying from 1-9 and 1-20
-images respectively, two registration algorithms (\ants{} or \animal{}), and three
-label fusion techniques (unweighted, cross-correlation, and normalised mutual
-information weighted majority vote).  Computed segmentations are compared to
-\psegs using Dice's Similarity Coefficient (DSC).  In the Supplementary
-Materials we have replicated this experiment using the \snt semi-automated
-labels included with the ADNI dataset.
-
-We find that for \mb{} segmentations, similarity score increases
-as atlas and template library size is increased, although with diminishing returns and
-an eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}).  A maximum
-similarity score of \Sexpr{mb.kappa.best} is found when using 9
-atlases, 19 templates, ANTS registration, and majority vote label fusion).  The
-\ants{} registration method consistently out-performs \animal{} registration
-over all variable settings we tested. Additionally, by itself, using a weighted
-voting strategy did not significantly improve segmentation accuracy, contrary to
-the findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation.
-Given these findings, in the remainder of this section only results using the
-\ants{} registration algorithm and majority vote fusion will be shown.
+% In this experiment we conduct 10 rounds of \mb{} and \multiatlas{} segmentation of
+% of 60 subjects using atlas and template library sizes varying from 1-9 and 1-20
+% images respectively, two registration algorithms (\ants{} or \animal{}), and three
+% label fusion techniques (unweighted, cross-correlation, and normalised mutual
+% information weighted majority vote).  Computed segmentations are compared to
+% \psegs using Dice's Similarity Coefficient (DSC).  In the Supplementary
+% Materials we have replicated this experiment using the \snt semi-automated
+% labels included with the ADNI dataset.
+
+We find that for \mb{} segmentations, similarity score increases as atlas and
+template library size is increased, although with diminishing returns and an
+eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}).  A
+maximum similarity score of \Sexpr{mb.kappa.best} is found when using 9 atlases,
+19 templates, ANTS registration, and majority vote label fusion).  The \ants{}
+registration method consistently out-performs \animal{} registration over all
+variable settings we tested. Additionally, by itself, using a weighted voting
+strategy did not significantly improve segmentation accuracy, contrary to the
+findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation.  Given
+these findings, in the remainder of this section only results using the \ants{}
+registration algorithm and majority vote fusion will be shown.
 
@@
 % ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
-\begin{figure}
-<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load'>>=
+<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load',fig.show='hide'>>=
 stderr <- function(x) sqrt(var(x)/length(x))
@@
 @
-  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
-  \psegs{} for 60 ADNI subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method over 10 folds of Monte Carlo       
-  cross validation. 
-  Error bars indicate standard error.
-  \label{fig:ADNI1-xval-k-mean}}
-\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ADNI1-XVAL: "increase" in mean kappa over \multiatlas{}
-\begin{figure}
-<<fig:ADN11-xval-k-diff,dependson='ADNI1-xval-load'>>=
+<<fig:ADNI1-xval-k-diff,dependson='ADNI1-xval-load',fig.show='hide'>>=
 ggplot(subset(jens_xval_mean, reg_method=="ANTS" & 
@@
 @
-  \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
-  \multiatlas{} segmentations vs. atlas and template library size when using the
-  \ants{} registration method, and majority-vote label fusion over 10 folds of Monte 
-  Carlo cross validation. . 
-  Segmentation similarity is computed against \psegs{}.  
-  Error bars indicate standard error.
-  \label{fig:ADNI1-xval-k-diff}}
-\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % ADNI1-XVAL: show variability of MAGeT over parameters
-\begin{figure}
-<<fig:ADN1-Xval-variability,dependson='ADNI1-Xval-prep'>>=
+<<fig:ADNI1-xval-variability,dependson='ADNI1-Xval-prep',fig.show='hide'>>=
 ants = subset(jens_xval_data, 
@@
 @
-  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation
-  with respect to manual segmentation.}
-  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
-  is computed for each subject across over 10 folds of Monte Carlo       
-  cross validation.  
-  Shown on the y-axis is the  p-value resulting from a t-test comparing the
-  distribution of variances at each parameter setting (atlas/template library
-  size). 
-  Only points where \mb{} mean variability is lower than \multiatlas{} are shown.
-  Dashed lines indicate a p-value of 0.05 and 0.01.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-mean}
+  \caption{Dice's similarity score vs. atlas and template library size.  Dice's
+  similarity score of \mb{} segmentations and \psegs{} for 60 ADNI subjects over a
+  range of atlas and template library size, registration algorithm, and label
+  fusion method on 10 folds of Monte Carlo cross validation.  Error bars indicate
+  standard error.} 
+  \label{fig:ADNI1-xval-k-mean}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-diff}
+    \caption{Increase in similarity score over \multiatlas} 
+    \label{fig:ADNI1-xval-k-diff}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-variability}
+    \caption{Difference in variability} 
     \label{fig:ADNI1-xval-variability}
+  \end{subfigure}
+  \caption{\textbf{Whole hippocampus segmentation cross-validation on
+  Alzheimer's diseased subjects.}
+(\ref{fig:ADNI1-xval-k-diff}) shows increase in mean Dice's Similarity Coefficient of \mb{} over
+  \multiatlas{} segmentations vs. atlas and template library size when using the
+  \ants{} registration method, and majority-vote label fusion over 10 folds of
+  Monte Carlo cross validation.  Segmentation similarity is computed against
+  \psegs{}.  Error bars indicate standard error.
+(\ref{fig:ADNI1-xval-variability}) compares the variability in Dice scores of \mb{} vs.
+  \multiatlas{} segmentations in the validation.  Variance of segmentation
+  accuracy between \mb{} and \multiatlas{} segmentation is computed for each
+  subject across over 10 folds of Monte Carlo cross validation.  Shown on
+  the y-axis is the p-value resulting from a t-test comparing the distribution of
+  variances at each parameter setting (atlas/template library size).  Only points
+  where \mb{} mean variability is lower than \multiatlas{} are shown.  Dashed
+  lines indicate a p-value of 0.05 and 0.01.}
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \label{exp:fep-xval}
-%TODO: update
+\textbf{[TODO: tie with exp 1]}
 To validate that the \mb{} works effectively in the context of other neurological
@@
 \subsubsection{Results}
-
-In this experiment \mb{} is applied to a dataset of images of first episode
-schizophrenia patients, using the Winterburn atlases and a template library of
-21 subject images selected at random.  Expert manual whole hippocampal
-segmentations are used as a gold standard.
+\textbf{[TODO: move up results to here.]}
+
+% In this experiment \mb{} is applied to a dataset of images of first episode
+% schizophrenia patients, using the Winterburn atlases and a template library of
+% 21 subject images selected at random.  Expert manual whole hippocampal
+% segmentations are used as a gold standard.
 
@@
   })
+
+  fep.mb.kappa.best = round(mean(subset(fep.xval.mean, 
+                                 templates == 19 & 
+                                 atlases      == 9)$k), digits=3)
   stderr <- function(x) sqrt(var(x)/length(x))
@@
     \includegraphics[width=\textwidth]{figure/fig:FEP-xval}
-    \caption{Mean DSC overlap score vs. atlas and template library size in
-    a 5-fold Monte Carlo cross-validation (MCCV) of \mb. Error bars
-    indicate standard error. }
+    \caption{Dice's similarity score vs. atlas and template library size} 
     \label{fig:FEP-xval}
@@
     \includegraphics[width=\textwidth]{figure/fig:FEP-vols} 
-    \caption{Computed vs. manually segmented mean hippocampus volume. Linear fit
-      lines are shown, with the shaded region showing standard error. }
+    \caption{Computed vs. manual hippocampus volume} 
     \label{fig:FEP-vols}
@@
     \includegraphics[width=\textwidth]{figure/fig:FEP-Bland-Altman}
-    \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-      measured in the FEP dataset by manual segmentation and each of the
-      four automated methods investigated (\freesurfer{}, \fsl{}, \mb{} with
-      Winterburn atlases, \mb{} with FEP manual atlases). The overall
-      mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
-      dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
-      Note, points below the mean difference indicate overestimation of the volume
-      with respect to the manual volume, and vice versa. }
+    \caption{Bland-Altman plots of computed vs. manual hippocampus volume} 
     \label{fig:FEP-Bland-Altman}
   \end{subfigure}
-  \caption{\textbf{First Episode Patient dataset automated segmentation.}
+  \caption{\textbf{First Episode Patient dataset validation.}
 Manual segmentation of all 81 subjects is completed using the \prsnr-protocol.
 \mb{} is configured using \ants{} registration algorithm and majority vote label
-fusion. In Figure \ref{fig:FEP-vols} \mb{} is run both with an atlas library
-consisting of the five Winterburn atlases (MAGeT-Winterburn), or with an atlas
-library of five manually segmented FEP subjects as atlases (MAGeT-FEP).}
+fusion. 
+(\ref{fig:FEP-xval}) shows mean DSC overlap similarity score of \mb{}
+  segmentations as they vary with atlas and template library size over 5-fold
+  validation. Error bars indicate standard error. 
+(\ref{fig:FEP-vols}) shows the volumes of the 81 subjects as determined by \fsl,
+  \freesurfer, and \mb{} configured in two ways: with an atlas library
+  of the five Winterburn atlases (MAGeT-Winterburn), or with an atlas
+  library of five manually segmented FEP subjects as atlases (MAGeT-FEP). Linear
+  fit lines are shown, with the shaded region showing standard error.
+(\ref{fig:FEP-Bland-Altman}) shows the agreement between computed volumes and
+  manually segmented volume. The overall mean difference in volume, and limits of
+  agreement ($\pm 1.96SD$) are shown by dashed horizontal lines. Linear fit lines
+  are shown for each diagnosis group.  Note, points below the mean difference
+  indicate overestimation of the volume with respect to the manual volume, and
+  vice versa.}
 \end{figure}
-
-% <<FEP-t-test,cache=TRUE,dependson="FEP-volumes">>=
-%   fep.cor.test = cor.test(fep.volumes$manual,fep.volumes$MAGeT_Winterburn)
-% @
-% \mb{} produces hippocampal volumes that are highly correlated with manual
-% segmentation volumes (Pearson $r = \Sexpr{fep.cor.test$estimate}$, $t =
-% \Sexpr{fep.cor.test$statistic}$, $p < \Sexpr{max(0.001, fep.cor.test$p.value)}$;
-% Figure \ref{fig:SZ-volumes}). 
-%
 
@@
 0.9mm dimension voxels.
-
+\textbf{[TODO: "What is the full voxel resolution of the EFGRE-BRAVO
+sequence?"]}
 The optimal size of template library, registration method, and label fusion
@@
 
-In this experiment the \mb{} segmentation of hippocampal subfields is assessed
-through a leave-one-out validation in which resampled images ($0.9mm^3$ voxels)
-of each Winterburn atlas subject are segmented using the remaining
-high-resolution Winterburn atlases.
-
 Figure \ref{fig:WAxval-kappa} shows the overlap similarity scores between the
@@
 molecular regions score below 0.5.  These scores may seem low but not when taken
-in context and compared to existing (semi-)automated methods (see Discussion). The whole hippocampus is segmented with a mean DSC score of\Sexpr{mean(Whole)} $\pm$\Sexpr{sd(Whole)}.
+in context and compared to existing (semi-)automated methods (see Discussion).
+The whole hippocampus is segmented with a mean DSC score of\Sexpr{mean(Whole)}
+$\pm$\Sexpr{sd(Whole)}.
 
@@
 and larger hippocampi are underestimated 
-(a mean maximum difference of approximately 200mm^3 across all subfields). 
+(a mean maximum difference of approximately $200mm^3$ across all subfields).
 \mb{} display a slight conservative fixed bias, tending to underestimate all
@@
 
-Based on the results from Experiment 1 (Section \ref{exp:adni-xval}), in this experiment \mb{} was configured
-with a template library of 21 randomly chosen subject images (7 from each
-disease class) and majority vote label fusion.  The entire \adnidataset{} dataset
-was segmented by \mb{}, and the resulting volumes compared with those obtained by
-expert semi-automated segmentation (\snt{}), and three other automated segmentation
-techniques: \maper{}, \freesurfer{}, and \fsl{}. Table
-\ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the total count of
-segmentations available, including a count of those which have failed a quality
-control inspection.  A total of \Sexpr{length(means.complete$RID)} images are
-included in the following analysis, having met quality criteria and having
-segmentations from every method.
+Based on the results from Experiment 1 (Section \ref{exp:adni-xval}), in this
+experiment \mb{} was configured with a template library of 21 randomly chosen
+subject images (7 from each disease class) and majority vote label fusion.  The
+entire \adnidataset{} dataset was segmented by \mb{}, and the resulting volumes
+compared with those obtained by expert semi-automated segmentation (\snt{}), and
+three other automated segmentation techniques: \maper{}, \freesurfer{}, and
+\fsl{}. Table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the
+total count of segmentations available, including a count of those which have
+failed a quality control inspection.  A total of
+\Sexpr{length(means.complete$RID)} images are included in the following
+analysis, having met quality criteria and having segmentations from every
+method.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\textbf{[TODO: combine volume plot/box-plot into a single figure]}
 
@@
 Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
-(CN); LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
+(CN); SZ = Schizophrenia; LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
 cross-validation; SNT = Surgical Medtronic Navigation Technologies
@@
 
-%TODO: 
 %%
@@
 & ADNI (\prsnr)
+\tabularnewline 
+\mb{}                      &9 & \Sexpr{fep.mb.kappa.best}   &
+& 5-fold MCCV on 81 subjects 
+& first episode SZ
 \tabularnewline 
@@
 
+\textbf{[TODO: fix experiment references]} 
 In this manuscript we have presented the implementation and validation of the
@@
 to provide accurate and reliable segmentations.  Experiment 1 (Section
-\ref{exp:adni-xval}) compares \mb{} to
-basic-\multiatlas{} segmentation by characterising the behaviour of each
-approach with varying parameter settings, and allows us to choose optimal
-settings for subsequent experiments. 
-%TODO: fix refs 
-Experiment 2 (Section \ref{exp:winterburn-xval}) emonstrates the reliability of
-\mb{} in producing subfield segmentations which match the segmentation protocol
-of the input atlases despite contrast and resolution limitations in standard
-T1-weighted image volumes.
-
-Experiments 2 and 3 validate whole hippocampal
+\ref{exp:adni-xval}) compares \mb{} to basic-\multiatlas{} segmentation by
+characterising the behaviour of each approach with varying parameter settings,
+and allows us to choose optimal settings for subsequent experiments.  Experiment
+2 (Section \ref{exp:winterburn-xval}) emonstrates the reliability of \mb{} in
+producing subfield segmentations which match the segmentation protocol of the
+input atlases despite contrast and resolution limitations in standard
+T1-weighted image volumes.  Experiments 2 and 3 validate whole hippocampal
 segmentation precision and consistency on populations with different ageing and
@@
 results reflect typical usage scenarios of researchers and clinicians. 
-%TODO: mention that DSC by it's nature penalizes structures with high surface
-%area-to-vol ratios, so necessarily penalise all subfield errors.
-%TODO: discuss perturbed segmentation kappa values in relation to MAGeT error
+\textbf{[TODO: mention that DSC by it's nature penalizes structures with high surface
+area-to-vol ratios, so necessarily penalise all subfield errors.]}
+\textbf{[TODO: discuss perturbed segmentation kappa values in relation to MAGeT
+error]}
 
@@
 Research Excellence; and the University of Toronto.
+\textbf{[TODO: cite scinet paper]}
 
@@
 
+\textbf{[TODO: Merge figures]}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-01-27 paper.Rnw
Moved subfield exp to last, added experiment headings on subsubsections, FEP = psychosis fix.

@@
 \pagestyle{fancy}
-\fancyhead[RO,RE]{draft - December, 2013}
+\fancyhead[RO,RE]{draft - January, 2014}
 
@@
 \newcommand{\snt}{SNT}
+\newcommand{\expadnixval}{Whole Hippocampus Segmentation Cross-Validation --- Alzheimer's Disease} 
+\newcommand{\expfepxval}{Whole Hippocampus Segmentation Cross-Validation --- First Episode of Psychosis} 
+\newcommand{\expadni}{Whole Hippocampus Segmentation Comparision --- ADNI1 Complete 1Yr} 
+\newcommand{\expsubfield}{Hippocampal Subfield Segmentation Cross-Validation} 
+\newcommand{\expadnixvalsnt}{Whole Hippocampus Segmentation Cross-Validation --- Alzheimer's Disease, \snt{} Segmentations}
 
@@
 \item change overlap/similarity --> agreement? choose one. 
-\item check that chart labels are sensible 
+\item check that all chart labels are sensible 
 \item check that Experiment 5 matches Experiment 1
+\item check that all Experiment references resolve correctly
 \item fix figure 2 schematic language for the experiments (not 2a 2b). 
 \item ADNI-1 or ADNI1: standardize
-\item experiment references formatted as Experiment 2 (Section 3.2)
+\item check that experiment references are formatted as Experiment 2 (Section 3.2)
 \item subfield or subregion -- choose one
@@
 (ADNI) dataset manually segmented using the \prsnr{}-protocol, over a range of
-atlas and template library sizes, registration and label fusion methods. We then
-perform a leave-one-out validation to determine if hippocampal subfields can be
-accurately identified, using the best parameters discovered in the first
-experiment.  To ensure that we have not overfit our parameters to the aging or
-neurodegenerative brain, we also apply \mb{} to a dataset of individuals
-suffering from first episode psychosis.  Finally, we validate our algorithm
-using all of the data available in the \adnidataset{} sample and compare our
+atlas and template library sizes, registration and label fusion methods. This
+experiment is then replicated with a first episode psychosis patient dataset
+to determine how the behaviour of \mb{} changes when segmenting younger and
+differently diseased subjects, and to ensure that we have not overfit our
+parameters to the aging or neurodegenerative brain. A leave-one-out validation
+of subfield segmentation is performed on the \citet{Winterburn2013} manually
+segmented high-resolution MR atlases, using the optimal \mb{} parameters found
+from the previous experiments. Finally, we validate our algorithm using
+all of the data available in the \adnidataset{} sample and compare \mb{}
 segmentations to other popular segmentation algorithms.
@@
 \item Experiment 2 consists of a similar cross-validation design to verify these
-parameter settings on brain images of young, first episode Schizophrenia
+parameter settings on brain images of young, first episode psychosis
 patients. 
-\item Experiment 3 assesses hippocampal subfield segmentation quality in a
+\item Experiment 3 compares \mb{} whole hippocampus segmentations with those of
+well-known automated methods on the entire \adnidataset image dataset consisting
+of 
+\item Experiment 4 assesses hippocampal subfield segmentation quality in a
 leave-one-out cross-validation design. 
-\item Experiment 4 compares \mb{} segmentations with those of well-known
-automated methods on the entire \adnidataset image dataset consisting of 
 \Sexpr{length(means.complete$RID)} brain images of subjects diagnosed as
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 1: Whole Hippocampus Segmentation --- Cross-Validation on 
-Alzheimer's Diseased Subjects}
+\subsection{Experiment 1: \expadnixval}
 \label{exp:adni-xval}
@@
 
-\subsubsection{Materials and Methods}
+\subsubsection{Materials and Methods: \expadnixval}
 \paragraph{\adnidataset{} dataset} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results}
+\subsubsection{Results: \expadnixval}
 
@@
 %%%%%         Experiment 2: Application of \mb{} to the segmentation      %%%%%
-%%%%%                       of schizophrenia patients                     %%%%%
+%%%%%                       of Psychosis patients                         %%%%%
 %%%%%                                                                     %%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 2: Whole Hippocampus Segmentation --- Cross-validation
-on First Episode Schizophrenia Subjects}
+\subsection{Experiment 2: \expfepxval}
 \label{exp:fep-xval}
@@
 disorders, in this experiment we use the Winterburn atlases to derive 
-whole hippocampal segmentations of a dataset of patients with schizophrenia. The
-resulting segmentations are assessed for quality by comparison with expert
-manual segmentations.
-
-\subsubsection{Materials and Methods}
-\paragraph{First Episode Schizophrenia Dataset}
+whole hippocampal segmentations of a dataset of patients having had a single
+episode of psychosis. The resulting segmentations are assessed for quality by
+comparison with expert manual segmentations.
+
+\subsubsection{Materials and Methods: \expfepxval}
+\paragraph{First Episode Psychosis Dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 %
-% First Episode SZ Demographics Table
+% First Episode Psychosis Demographics Table
 %
@@
 caption = 
-  "\\textbf{Schizophrenia First Episode Patient Demographics.}
+  "\\textbf{First Episode Psychosis Subject Demographics.}
   ambi - ambidextrous. 
@@
 %%%%%    RESULTS: Experiment 2, Application of \mb{} to the segmentation  %%%%%
-%%%%%                       of schizophrenia patients                     %%%%%
+%%%%%                       of psychosis patients                         %%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results}
+\subsubsection{Results: \expfepxval}
 \textbf{[TODO: move up results to here.]}
@@
 % In this experiment \mb{} is applied to a dataset of images of first episode
-% schizophrenia patients, using the Winterburn atlases and a template library of
+% psychosis patients, using the Winterburn atlases and a template library of
 % 21 subject images selected at random.  Expert manual whole hippocampal
@@
 \FloatBarrier
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%                                                                     %%%%%
-%%%%%        Experiment 3: Hippocampal Subfield Cross-Validation          %%%%%
-%%%%%                                                                     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 3: Hippocampal Subfield Segmentation --- Cross-Validation
-on the Winterburn Atlases}
-\label{exp:winterburn-xval}
-The previous experiment assesses \mb{} peformance on whole hippocampus
-segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
-segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
-use a modified leave-one-out cross-validation (LOOCV) design.
-
-\subsubsection{Materials and Methods}
-\paragraph{Winterburn Atlases} 
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
-segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
-and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
-lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
-range in age from 29-57 years (mean age of 37), and include two males and three
-females.
-
-\paragraph{Experiment details} 
-Leave-one-out cross-validation (LOOCV) is a validation approach in which an
-algorithm is given all but one item in a dataset as training data (in our case,
-atlas images and labels) and then the algorithm is applied to the left-out item.
-This is done, in turn, for each item in the dataset and the output across all
-items is evaluated together.
-
-In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
-voxel images to simulate standard 3T T1-weighted resolution images. Image
-subsampling is performed using trilinear subsampling techniques.  In each round
-of LOOCV, a single atlas is selected and segmented with \mb{} once using each
-possible triplet of atlas images from the remaining four unselected atlases.
-Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
-image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
-with an odd number of images so as to ensure an unbiased label fusion when using
-majority voting (see Discussion).
-
-The template library used has a total of 19 images composed of all $0.9mm^3$
-voxel target images plus an additional set of 14 of healthy subject 3T T1 images
-acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
-were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
-WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
-3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
-$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
-thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
-0.9mm dimension voxels.
-\textbf{[TODO: "What is the full voxel resolution of the EFGRE-BRAVO
-sequence?"]}
-The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Subfield Validation Template library demographics (healthy controls)
-%                                     
-<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
-master = read.csv('data/DT_study.csv')
-tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
-    method="reverse", na.rm = TRUE, test=FALSE)
-caption = 
-  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
-control subject sample used in the template library (excluding the Winterburn
-atlas subjects).} Education is shown in years."
-latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
-      caption=caption, caption.loc = c("top"),
-      title="tab:WAval-healthy-demographics")
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{Evaluation method}  
-Evaluating the accuracy of hippocampal subfields for T1 images at
-$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
-protocols for segmentation at this resolution to use as a reference. Instead, we
-can evaluate the reliability, or {\it precision}, with which \mb{} produces
-hippocampal subfields that correspond to the segmentation protocol used by the
-given atlas library images. By directly resampling the Winterburn atlas
-segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
-resampling techniques) we obtain a scaled down version of the labels which
-preserve the original segmentation protocol within the limits of error from
-rounding and interpolation. Therefore, using the resampled Winterburn
-segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
-of \mb{} segmentations using using DSC overlap scores and evaluate consistency
-across the range of hippocampal sizes using Bland-Altman plots of subfield
-volumes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%    RESULTS: Experiment 3: Hippocampal Subfield Cross-Validation     %%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results}
-
-\begin{figure}
-<<fig:WAxval-kappa>>=
-wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
-wa.n.sim.whole = read.csv('data/winterburn.xval.similarity.whole.csv')
-wa.n.sim.whole$label = as.factor(wa.n.sim.whole$label)
-levels(wa.n.sim.whole$label) <- list(
-  "Whole"=1,
-  "Whole"=2)
-wa.n.sim.whole$res='0.9'
-wa.n.sim.whole=wa.n.sim.whole[c('atlases','subject','res','label','k','se','sn','j')]
-
-wa.n.sim$label = as.factor(wa.n.sim$label)
-levels(wa.n.sim$label) <-list(
-  "CA1"=1, "CA1"=101, 
-  "CA2/CA3"=5,"CA2/CA3"=105, 
-  "CA4/DG"=4,"CA4/DG"=104, 
-  "Subiculum"=2,"Subiculum"=102, 
-  "SR/SL/SM"=6,"SR/SL/SM"=106)
-
-wa.n.sim = rbind(wa.n.sim, wa.n.sim.whole)
-
-# kappa vs. region
-ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
-  geom_boxplot(fill='lightblue') + 
-  scale_y_continuous(breaks=seq(0,1,0.1)) +
-  xlab("Subregion") + 
-  ylab("Dice's Similarity Coefficient (DSC)")
-
-CA1   = subset(wa.n.sim,label=='CA1')$k
-CA23  = subset(wa.n.sim,label=='CA2/CA3')$k
-CA4DG = subset(wa.n.sim,label=='CA4/DG')$k
-Sub   = subset(wa.n.sim,label=='Subiculum')$k
-SRLM  = subset(wa.n.sim,label=='SR/SL/SM')$k
-Whole = subset(wa.n.sim,label=='Whole')$k
-@
-  \caption{Similarity of \mb{} segmentation of subregions and the resampled
-  Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
-  leave-one-out cross validation. Overlap score for each hemisphere is measured
-  separately. \label{fig:WAxval-kappa}}
-\end{figure}
-
-Figure \ref{fig:WAxval-kappa} shows the overlap similarity scores between the
-\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
-subregion across all subjects and folds of the validation. Mean and standard
-deviation DSC scores of the subregions are shown in Table
-\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
-segmentations when perturbed slightly and compared to the originals.  We find
-that the CA4/DG subregion shows the highest mean DSC score of
-\Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
-subregions having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
-\Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively.  Both the CA4/DG and
-molecular regions score below 0.5.  These scores may seem low but not when taken
-in context and compared to existing (semi-)automated methods (see Discussion).
-The whole hippocampus is segmented with a mean DSC score of\Sexpr{mean(Whole)}
-$\pm$\Sexpr{sd(Whole)}.
-
-\begin{figure}
-<<fig:WAxval-Bland-Altman,fig.height=5,fig.width=7>>=
-wa.n.vol = read.csv('data/winterburn.xval.volumes.csv') # xval volumes
-winterburn = read.csv('data/winterburn.volumes.csv')    # resampled volumes
-subfield_labels <-list(
-  "CA1"="X1", "CA1"="X101", 
-  "CA2/CA3"="X5","CA2/CA3"="X105", 
-  "CA4/DG"="X4","CA4/DG"="X104",
-  "Subiculum"="X2","Subiculum"="X102",
-  "SR/SL/SM"="X6","SR/SL/SM"="X106", 
-  "Whole" = "left", "Whole" = "right")
-hemispheres <- list(
-  "Right"="X1","Left"="X101", 
-  "Right"="X5","Left"="X105", 
-  "Right"="X4","Left"="X104",
-  "Right"="X2","Left"="X102",
-  "Right"="X6","Left"="X106",
-  "Right"="right","Left"="left")
-
-# compute whole hippocampal volumes (we do this the stupid way)
-wa.n.vol = within(wa.n.vol, {
-  right = X1+X2+X4+X5+X6
-  left  = X101+X102+X104+X105+X106
-  })
-winterburn = within(winterburn, {
-  right = X1+X2+X4+X5+X6
-  left  = X101+X102+X104+X105+X106
-  })
-
-wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
-winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
-winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
-wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
-
-levels(wa.n.all.vols$subject) <- list(
-  "1"="anusha", 
-  "2"="jon",
-  "3"="mallar",
-  "4"="sofia",
-  "5"="nancy")
-
-# mean per subfield/hemisphere across all 4 validation folds
-wa.n.mean.vols = ddply(wa.n.all.vols, c("variable", "subject", "res"), function(df) {
-  data.frame(mb    = mean(df$value),
-             high  = mean(df$'0.3'),
-             low   = mean(df$'0.9'))
-})
-wa.n.mean.vols$diff = wa.n.mean.vols$low - wa.n.mean.vols$mb
-wa.n.mean.vols$mean = (wa.n.mean.vols$low + wa.n.mean.vols$mb)/2
-wa.n.mean.vols$hemisphere        <- wa.n.mean.vols$variable
-levels(wa.n.mean.vols$variable)  <- subfield_labels
-
-wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
-wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2
-wa.n.all.vols$Hemisphere        <- wa.n.all.vols$variable
-levels(wa.n.all.vols$variable)   <- subfield_labels
-levels(wa.n.all.vols$Hemisphere) <- hemispheres
-
-wa.n.xval.bland.limits = ddply(wa.n.mean.vols,c("variable"),function(df) {
-  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
-})
-
-ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
-  facet_wrap(~variable, ncol=3, scales='free_x') + 
-  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=2) +  
-  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
-             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
-  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
-             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of resampled and MAGeT-Brain subregion volume ' (mm^3))) + 
-  ylab(expression('Difference of manual and MAGeT-Brain subregion volume ' (mm^3))) + 
-  #scale_x_continuous(limits=c(0,1000)) + 
-  #scale_color_discrete(name="") + 
-  scale_shape_discrete(name='Subject') + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
-        axis.text.y = element_text(angle = 90, hjust=0.5), 
-        axis.text = element_text(size = 8))
-@
-  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
-across all folds of Winterburn atlas cross validation.} The overall mean
-difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
-horizontal lines. Linear fit lines are shown. Note, points below the mean
-difference indication overestimation of the volume with respect to the resampled
-volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
-\end{figure}
-
-Figure \ref{fig:WAxval-Bland-Altman} contains Bland-Altman plots comparing \mb{}
-volumes with manual volumes across all validation folds.  \mb{} displays a
-conservative proportional bias --- small hippocampi are overestimated in volume,
-and larger hippocampi are underestimated 
-(a mean maximum difference of approximately $200mm^3$ across all subfields).
-\mb{} display a slight conservative fixed bias, tending to underestimate all
-subfields except CA4/DG (mean underestimation: $CA1=76mm^3$, $CA2/3=56mm^3$,
-$CA4/DG=-16mm^3$, $Subiculum=48mm^3$, $SR/SL/SM=96mm^3$).
-
-<<tab:WAxval-perturb-kappa, echo=F,results="asis">>=
-wa.perturb = read.csv('data/winterburn.perturb.kappa.csv')
-wa.perturb = wa.perturb[c("subject","label","operation","k")]
-wa.perturb$label = as.factor(wa.perturb$label)
-levels(wa.perturb$label) <-list(
-  "CA1"="1", "CA1"="101",  "CA2/CA3"="5","CA2/CA3"="105",
-  "CA4/DG"="4","CA4/DG"="104",  "Subiculum"="2","Subiculum"="102", 
-  "SR/SL/SM"="6", "SR/SL/SM"="106")
-mean_range = function (v) {
-  sprintf("$%0.2f \\pm %0.2f$", mean(v), sd(v))
-}
-wa.perturb.table = dcast(wa.perturb, label ~ operation, fun.aggregate=mean_range)
-wa.sim.sum = dcast(wa.n.sim, label ~ ., value.var = 'k', fun.aggregate=mean_range)
-names(wa.sim.sum)[2] = 'MAGeT'
-wa.perturb.table = merge(wa.perturb.table, wa.sim.sum)
-
-wa.perturb.table=wa.perturb.table[c('label','MAGeT', 'translate','contract','expand')]
-names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")
-
-caption="Overlap similarity results for the each of the subfields of the hippocampus. 
-Simulated overlap similarity results are also given for resampled labels that were 
-translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
-by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
-Values are given as mean Dice's Kappa coefficient $\\pm$ stddev."
-latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
-      caption=caption, caption.loc = c("top"),
-      col.just=rep('c',5),  
-      label="tab:WAxval-perturb-kappa", rowname=NULL)
-@
-
-Figure \ref{fig:subfield-montage} shows slices subfield segmentations for a
-single subject for qualitative inspection.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=5in]{figure/winterburn-atlas-montage/figure.pdf}
-    \caption{\textbf{Detailed subfield segmentation results for a single subject.} In
-    the upper left corner is the original high-resolution Winterburn atlas manual
-    subfield segmentation; in the upper right corner is the Winterburn atlas
-    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
-    corner is the \mb{} segmentation of the subsampled Winterburn atlas image
-    from a single fold of the cross-validation. In each segmentation, slices from the left
-    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
-    slices from inferior to superior; the second row shows sagittal slices from
-    lateral to medial; the third row shows coronal slices from anterior to
-    posterior. \label{fig:subfield-montage}}
-\end{figure}
-
-\FloatBarrier
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%                                                                     %%%%%
-%%%%%         EXPERIMENT 4: Application of \mb{} to the segmentation      %%%%%
+%%%%%         EXPERIMENT 3: Application of \mb{} to the segmentation      %%%%%
 %%%%%                       of ADNI subjects                              %%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 4: Application to the segmentation of Alzheimer's
-disease patients}
+\subsection{Experiment 3: \expadni}
 \label{exp:adni}
@@
 
-\subsubsection{Materials and Methods}
+\subsubsection{Materials and Methods: \expadni}
 \paragraph{\adnidataset dataset} 
@@
 \mb{} was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 3, section \ref{exp:winterburn-xval}), and size of template library,
+atlases (described in Experiment 4, section \ref{exp:winterburn-xval}), and size of template library,
 registration method, and label fusion method were determined by the optimal
@@
 %%%%%                           RESULTS                                   %%%%%
-%%%%%         EXPERIMENT 4: Application of \mb{} to the segmentation      %%%%%
+%%%%%         EXPERIMENT 3: Application of \mb{} to the segmentation      %%%%%
 %%%%%                       of ADNI subjects                              %%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results}
+\subsubsection{Results: \expadni}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%        Experiment 4: Hippocampal Subfield Cross-Validation          %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment 4: \expsubfield}
+\label{exp:winterburn-xval}
+The previous experiment assesses \mb{} peformance on whole hippocampus
+segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
+segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels.  We
+use a modified leave-one-out cross-validation (LOOCV) design.
+
+\subsubsection{Materials and Methods: \expsubfield}
+\paragraph{Winterburn Atlases} 
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
+segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
+and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
+lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
+
+\paragraph{Experiment details} 
+Leave-one-out cross-validation (LOOCV) is a validation approach in which an
+algorithm is given all but one item in a dataset as training data (in our case,
+atlas images and labels) and then the algorithm is applied to the left-out item.
+This is done, in turn, for each item in the dataset and the output across all
+items is evaluated together.
+
+In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
+voxel images to simulate standard 3T T1-weighted resolution images. Image
+subsampling is performed using trilinear subsampling techniques.  In each round
+of LOOCV, a single atlas is selected and segmented with \mb{} once using each
+possible triplet of atlas images from the remaining four unselected atlases.
+Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
+image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
+with an odd number of images so as to ensure an unbiased label fusion when using
+majority voting (see Discussion).
+
+The template library used has a total of 19 images composed of all $0.9mm^3$
+voxel target images plus an additional set of 14 of healthy subject 3T T1 images
+acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
+were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
+WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
+3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
+$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
+0.9mm dimension voxels.
+\textbf{[TODO: "What is the full voxel resolution of the EFGRE-BRAVO
+sequence?"]}
+The optimal size of template library, registration method, and label fusion
+method found in Experiment 1 are used.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Subfield Validation Template library demographics (healthy controls)
+%                                     
+<<tab:WAval-healthy-demographics, echo=F, results="asis",cache=FALSE>>=
+master = read.csv('data/DT_study.csv')
+tab <- summary( Diagnosis ~ Age + Sex + Education + Handedness, data = master,
+    method="reverse", na.rm = TRUE, test=FALSE)
+caption = 
+  "\\textbf{Demographics for the hippocampal subfield cross-validation healthy
+control subject sample used in the template library (excluding the Winterburn
+atlas subjects).} Education is shown in years."
+latex(tab, file="", size="scriptsize",landscape=FALSE, ctable=FALSE,
+      caption=caption, caption.loc = c("top"),
+      title="tab:WAval-healthy-demographics")
+@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Evaluation method}  
+Evaluating the accuracy of hippocampal subfields for T1 images at
+$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
+protocols for segmentation at this resolution to use as a reference. Instead, we
+can evaluate the reliability, or {\it precision}, with which \mb{} produces
+hippocampal subfields that correspond to the segmentation protocol used by the
+given atlas library images. By directly resampling the Winterburn atlas
+segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
+resampling techniques) we obtain a scaled down version of the labels which
+preserve the original segmentation protocol within the limits of error from
+rounding and interpolation. Therefore, using the resampled Winterburn
+segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
+of \mb{} segmentations using using DSC overlap scores and evaluate consistency
+across the range of hippocampal sizes using Bland-Altman plots of subfield
+volumes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%    RESULTS: Experiment 4: Hippocampal Subfield Cross-Validation     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Results: \expsubfield}
+
+\begin{figure}
+<<fig:WAxval-kappa>>=
+wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
+wa.n.sim.whole = read.csv('data/winterburn.xval.similarity.whole.csv')
+wa.n.sim.whole$label = as.factor(wa.n.sim.whole$label)
+levels(wa.n.sim.whole$label) <- list(
+  "Whole"=1,
+  "Whole"=2)
+wa.n.sim.whole$res='0.9'
+wa.n.sim.whole=wa.n.sim.whole[c('atlases','subject','res','label','k','se','sn','j')]
+
+wa.n.sim$label = as.factor(wa.n.sim$label)
+levels(wa.n.sim$label) <-list(
+  "CA1"=1, "CA1"=101, 
+  "CA2/CA3"=5,"CA2/CA3"=105, 
+  "CA4/DG"=4,"CA4/DG"=104, 
+  "Subiculum"=2,"Subiculum"=102, 
+  "SR/SL/SM"=6,"SR/SL/SM"=106)
+
+wa.n.sim = rbind(wa.n.sim, wa.n.sim.whole)
+
+# kappa vs. region
+ggplot(wa.n.sim, aes(x=as.factor(label),y=k)) + 
+  geom_boxplot(fill='lightblue') + 
+  scale_y_continuous(breaks=seq(0,1,0.1)) +
+  xlab("Subregion") + 
+  ylab("Dice's Similarity Coefficient (DSC)")
+
+CA1   = subset(wa.n.sim,label=='CA1')$k
+CA23  = subset(wa.n.sim,label=='CA2/CA3')$k
+CA4DG = subset(wa.n.sim,label=='CA4/DG')$k
+Sub   = subset(wa.n.sim,label=='Subiculum')$k
+SRLM  = subset(wa.n.sim,label=='SR/SL/SM')$k
+Whole = subset(wa.n.sim,label=='Whole')$k
+@
+  \caption{Similarity of \mb{} segmentation of subregions and the resampled
+  Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
+  leave-one-out cross validation. Overlap score for each hemisphere is measured
+  separately. \label{fig:WAxval-kappa}}
+\end{figure}
+
+Figure \ref{fig:WAxval-kappa} shows the overlap similarity scores between the
+\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
+subregion across all subjects and folds of the validation. Mean and standard
+deviation DSC scores of the subregions are shown in Table
+\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
+segmentations when perturbed slightly and compared to the originals.  We find
+that the CA4/DG subregion shows the highest mean DSC score of
+\Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
+subregions having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
+\Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively.  Both the CA4/DG and
+molecular regions score below 0.5.  These scores may seem low but not when taken
+in context and compared to existing (semi-)automated methods (see Discussion).
+The whole hippocampus is segmented with a mean DSC score of\Sexpr{mean(Whole)}
+$\pm$\Sexpr{sd(Whole)}.
+
+\begin{figure}
+<<fig:WAxval-Bland-Altman,fig.height=5,fig.width=7>>=
+wa.n.vol = read.csv('data/winterburn.xval.volumes.csv') # xval volumes
+winterburn = read.csv('data/winterburn.volumes.csv')    # resampled volumes
+subfield_labels <-list(
+  "CA1"="X1", "CA1"="X101", 
+  "CA2/CA3"="X5","CA2/CA3"="X105", 
+  "CA4/DG"="X4","CA4/DG"="X104",
+  "Subiculum"="X2","Subiculum"="X102",
+  "SR/SL/SM"="X6","SR/SL/SM"="X106", 
+  "Whole" = "left", "Whole" = "right")
+hemispheres <- list(
+  "Right"="X1","Left"="X101", 
+  "Right"="X5","Left"="X105", 
+  "Right"="X4","Left"="X104",
+  "Right"="X2","Left"="X102",
+  "Right"="X6","Left"="X106",
+  "Right"="right","Left"="left")
+
+# compute whole hippocampal volumes (we do this the stupid way)
+wa.n.vol = within(wa.n.vol, {
+  right = X1+X2+X4+X5+X6
+  left  = X101+X102+X104+X105+X106
+  })
+winterburn = within(winterburn, {
+  right = X1+X2+X4+X5+X6
+  left  = X101+X102+X104+X105+X106
+  })
+
+wa.n.vol.melt = melt(wa.n.vol, id.vars=c('atlases','subject','res'))
+winterburn.melt = melt(winterburn, id.vars=c('subject','res'))
+winterburn.cast = dcast(winterburn.melt, subject + variable ~ res)
+wa.n.all.vols = merge(wa.n.vol.melt, winterburn.cast, by=c('subject','variable'))
+
+levels(wa.n.all.vols$subject) <- list(
+  "1"="anusha", 
+  "2"="jon",
+  "3"="mallar",
+  "4"="sofia",
+  "5"="nancy")
+
+# mean per subfield/hemisphere across all 4 validation folds
+wa.n.mean.vols = ddply(wa.n.all.vols, c("variable", "subject", "res"), function(df) {
+  data.frame(mb    = mean(df$value),
+             high  = mean(df$'0.3'),
+             low   = mean(df$'0.9'))
+})
+wa.n.mean.vols$diff = wa.n.mean.vols$low - wa.n.mean.vols$mb
+wa.n.mean.vols$mean = (wa.n.mean.vols$low + wa.n.mean.vols$mb)/2
+wa.n.mean.vols$hemisphere        <- wa.n.mean.vols$variable
+levels(wa.n.mean.vols$variable)  <- subfield_labels
+
+wa.n.all.vols$diff = wa.n.all.vols$'0.9' - wa.n.all.vols$value
+wa.n.all.vols$mean = (wa.n.all.vols$'0.9' + wa.n.all.vols$value)/2
+wa.n.all.vols$Hemisphere        <- wa.n.all.vols$variable
+levels(wa.n.all.vols$variable)   <- subfield_labels
+levels(wa.n.all.vols$Hemisphere) <- hemispheres
+
+wa.n.xval.bland.limits = ddply(wa.n.mean.vols,c("variable"),function(df) {
+  data.frame(y = mean(df$diff,na.rm=T) + c(-1.96,0,2) * sd(df$diff,na.rm=T))
+})
+
+ggplot(wa.n.all.vols, aes(x=mean, y=diff)) + 
+  facet_wrap(~variable, ncol=3, scales='free_x') + 
+  geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
+  geom_point(size=2) +  
+  geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
+             data=wa.n.xval.bland.limits, size=0.3, alpha=0.5) + 
+  geom_text(aes(y = y, x=Inf, label=round(y,0)), colour="black", 
+             data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
+  xlab(expression('Mean of resampled and MAGeT-Brain subregion volume ' (mm^3))) + 
+  ylab(expression('Difference of manual and MAGeT-Brain subregion volume ' (mm^3))) + 
+  #scale_x_continuous(limits=c(0,1000)) + 
+  #scale_color_discrete(name="") + 
+  scale_shape_discrete(name='Subject') + 
+  theme(legend.direction = "horizontal", legend.position = "bottom", 
+        axis.text.y = element_text(angle = 90, hjust=0.5), 
+        axis.text = element_text(size = 8))
+@
+  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
+across all folds of Winterburn atlas cross validation.} The overall mean
+difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
+horizontal lines. Linear fit lines are shown. Note, points below the mean
+difference indication overestimation of the volume with respect to the resampled
+volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
+\end{figure}
+
+Figure \ref{fig:WAxval-Bland-Altman} contains Bland-Altman plots comparing \mb{}
+volumes with manual volumes across all validation folds.  \mb{} displays a
+conservative proportional bias --- small hippocampi are overestimated in volume,
+and larger hippocampi are underestimated 
+(a mean maximum difference of approximately $200mm^3$ across all subfields).
+\mb{} display a slight conservative fixed bias, tending to underestimate all
+subfields except CA4/DG (mean underestimation: $CA1=76mm^3$, $CA2/3=56mm^3$,
+$CA4/DG=-16mm^3$, $Subiculum=48mm^3$, $SR/SL/SM=96mm^3$).
+
+<<tab:WAxval-perturb-kappa, echo=F,results="asis">>=
+wa.perturb = read.csv('data/winterburn.perturb.kappa.csv')
+wa.perturb = wa.perturb[c("subject","label","operation","k")]
+wa.perturb$label = as.factor(wa.perturb$label)
+levels(wa.perturb$label) <-list(
+  "CA1"="1", "CA1"="101",  "CA2/CA3"="5","CA2/CA3"="105",
+  "CA4/DG"="4","CA4/DG"="104",  "Subiculum"="2","Subiculum"="102", 
+  "SR/SL/SM"="6", "SR/SL/SM"="106")
+mean_range = function (v) {
+  sprintf("$%0.2f \\pm %0.2f$", mean(v), sd(v))
+}
+wa.perturb.table = dcast(wa.perturb, label ~ operation, fun.aggregate=mean_range)
+wa.sim.sum = dcast(wa.n.sim, label ~ ., value.var = 'k', fun.aggregate=mean_range)
+names(wa.sim.sum)[2] = 'MAGeT'
+wa.perturb.table = merge(wa.perturb.table, wa.sim.sum)
+
+wa.perturb.table=wa.perturb.table[c('label','MAGeT', 'translate','contract','expand')]
+names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")
+
+caption="Overlap similarity results for the each of the subfields of the hippocampus. 
+Simulated overlap similarity results are also given for resampled labels that were 
+translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
+by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
+Values are given as mean Dice's Kappa coefficient $\\pm$ stddev."
+latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
+      caption=caption, caption.loc = c("top"),
+      col.just=rep('c',5),  
+      label="tab:WAxval-perturb-kappa", rowname=NULL)
+@
+
+Figure \ref{fig:subfield-montage} shows slices subfield segmentations for a
+single subject for qualitative inspection.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=5in]{figure/winterburn-atlas-montage/figure.pdf}
+    \caption{\textbf{Detailed subfield segmentation results for a single subject.} In
+    the upper left corner is the original high-resolution Winterburn atlas manual
+    subfield segmentation; in the upper right corner is the Winterburn atlas
+    segmentation subsampled from 0.3mm- to 0.9mm-isotropic voxels; in the lower left
+    corner is the \mb{} segmentation of the subsampled Winterburn atlas image
+    from a single fold of the cross-validation. In each segmentation, slices from the left
+    hemisphere are shown in Talairach-like ICBM152 space: the first row shows axial
+    slices from inferior to superior; the second row shows sagittal slices from
+    lateral to medial; the third row shows coronal slices from anterior to
+    posterior. \label{fig:subfield-montage}}
+\end{figure}
+
+\FloatBarrier
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
-(CN); SZ = Schizophrenia; LOOCV = Leave-one-out cross-validation; MCCV = Monte Carlo
-cross-validation; SNT = Surgical Medtronic Navigation Technologies
+(CN); FEP = First Episode of Psychosis; LOOCV = Leave-one-out cross-validation; MCCV
+= Monte Carlo cross-validation; SNT = Surgical Medtronic Navigation Technologies
 semi-automated labels.  Some studies of automated segmentation of ADNI images
@@
 & 5-fold MCCV on 81 subjects 
-& first episode SZ
+& FEP subjects 
 \tabularnewline 
@@
 \FloatBarrier
-\subsection{Experiment 5: Whole Hippocampus Segmentation Cross-Validation with \snt{} 
-Segmentations}
+\subsection{Experiment 5: \expadnixvalsnt}
 
@@
 
-\subsubsection{Methods}
+\subsubsection{Materials and Methods: \expadnixvalsnt}
 \paragraph{Dataset} 
@@
 % ######################
-\subsubsection{Results}
+\subsubsection{Results: \expadnixvalsnt}
 

### 2014-01-27 paper.Rnw
Fixed unclosed macros

@@
 
-THe gold standard for neuroanatomical segmentation is manual delineation by an
+The gold standard for neuroanatomical segmentation is manual delineation by an
 expert human rater.  However, with the availability of increasingly large MRI
-datasets, the time and expertise required for manual segmentation becomes
+datasets the time and expertise required for manual segmentation becomes
 prohibitive \citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. This
@@
 the target images, the authors demonstrate accurate parcellation of hippocampal
-subfields in clinical 3T T1-weighted MRI volumes.  Using \multiatlas with bias
+subfields in clinical 3T T1-weighted MRI volumes.  Using \multiatlas{} with bias
 correction techniques, \citet{Yushkevich2010} demonstrate a semi-automated
@@
 \item Experiment 3 compares \mb{} whole hippocampus segmentations with those of
-well-known automated methods on the entire \adnidataset image dataset consisting
+well-known automated methods on the entire \adnidataset{} image dataset consisting
 of 
@@
 \subsubsection{Results: \expadnixval}
-
-% In this experiment we conduct 10 rounds of \mb{} and \multiatlas{} segmentation of
-% of 60 subjects using atlas and template library sizes varying from 1-9 and 1-20
-% images respectively, two registration algorithms (\ants{} or \animal{}), and three
-% label fusion techniques (unweighted, cross-correlation, and normalised mutual
-% information weighted majority vote).  Computed segmentations are compared to
-% \psegs using Dice's Similarity Coefficient (DSC).  In the Supplementary
-% Materials we have replicated this experiment using the \snt semi-automated
-% labels included with the ADNI dataset.
 
@@
 \subsubsection{Materials and Methods: \expadni}
-\paragraph{\adnidataset dataset} 
+\paragraph{\adnidataset{} dataset} 
 The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
@@
 of atlases and the number of templates used improves \mb{} segmentation over and
-above basic-\multiatlas segmentations using the same number of atlas images.
+above basic-\multiatlas{} segmentations using the same number of atlas images.
 That is, by taking the extra step of generating a template library from a small
@@
 templates, ANTS registration, and majority vote label fusion).  Compared to
-\multiatlas segmentations, we find \mb{} segmentations show increasing
+\multiatlas{} segmentations, we find \mb{} segmentations show increasing
 improvement with larger atlas and template libraries when using more than 9
@@
 @
-  \caption{{\bf Difference in variability of \mb{} vs. \multiatlas{} segmentation
-  with respect to \snt segmentation.}
+  \caption{\textbf{Difference in variability of \mb{} vs. \multiatlas{} segmentation
+  with respect to \snt{} segmentation.}
   Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation

### 2014-01-27 paper.Rnw
Mallar's comments on Section 1 addressed

@@
 method of subfield segmentation on in vivo focal T2-weighted MR acquisitions of
-the temporal lobe. Manual input is only needed to mark mark divisions between
+the temporal lobe. Manual input is only needed to mark divisions between
 the head, body and tail of the hippocampus on target images.
@@
 experiment is then replicated with a first episode psychosis patient dataset
-to determine how the behaviour of \mb{} changes when segmenting younger and
-differently diseased subjects, and to ensure that we have not overfit our
-parameters to the aging or neurodegenerative brain. A leave-one-out validation
-of subfield segmentation is performed on the \citet{Winterburn2013} manually
-segmented high-resolution MR atlases, using the optimal \mb{} parameters found
-from the previous experiments. Finally, we validate our algorithm using
-all of the data available in the \adnidataset{} sample and compare \mb{}
-segmentations to other popular segmentation algorithms.
+to determine the behaviour of \mb{} when segmenting younger and differently
+diseased subjects, and to ensure that we have not also overfit our parameters to
+the aging or neurodegenerative brain. Next, we consider all of the images
+available in the \adnidataset{} sample, and compare \mb{} segmentations to other
+popular segmentation algorithms.  Finally, we conduct a leave-one-out validation
+of subfield segmentation using the \citet{Winterburn2013} manually segmented
+high-resolution MR atlases.
 

### 2014-01-29 paper.Rnw
Section 2 plus Experiment X references macro'd

@@
 \newcommand{\snt}{SNT}
-\newcommand{\expadnixval}{Whole Hippocampus Segmentation Cross-Validation --- Alzheimer's Disease} 
-\newcommand{\expfepxval}{Whole Hippocampus Segmentation Cross-Validation --- First Episode of Psychosis} 
-\newcommand{\expadni}{Whole Hippocampus Segmentation Comparision --- ADNI1 Complete 1Yr} 
-\newcommand{\expsubfield}{Hippocampal Subfield Segmentation Cross-Validation} 
-\newcommand{\expadnixvalsnt}{Whole Hippocampus Segmentation Cross-Validation --- Alzheimer's Disease, \snt{} Segmentations}
+% experiment #'s 
+\newcommand{\expadnixval}{Experiment 1} 
+\newcommand{\expfepxval}{Experiment 2} 
+\newcommand{\expadni}{Experiment 3}
+\newcommand{\expsubfield}{Experiment 4}
+\newcommand{\expadnixvalsnt}{Experiment 5}
+% experiment descriptions (suitable for titles)
+\newcommand{\expadnixvaldesc}{Whole Hippocampus Segmentation Cross-Validation --- Alzheimer's Disease} 
+\newcommand{\expfepxvaldesc}{Whole Hippocampus Segmentation Cross-Validation --- First Episode of Psychosis} 
+\newcommand{\expadnidesc}{Whole Hippocampus Segmentation Comparision --- ADNI1 Complete 1Yr} 
+\newcommand{\expsubfielddesc}{Hippocampal Subfield Segmentation Cross-Validation} 
+\newcommand{\expadnixvalsntdesc}{Whole Hippocampus Segmentation Cross-Validation --- Alzheimer's Disease, \snt{} Segmentations}
 
@@
 Alzheimer's disease populations, and \mb{} segmentations are compared with
-existing automated methods (\fsl{}, \freesurfer{}, \maper{}) and biases are
-explored.\\
+existing automated methods (\fsl{}, \freesurfer{}, \maper{}, and \snt) and
+biases are explored.\\
 \textbf{Results:} \textbf{[TODO: update with FEP-XVAL and subfield XVAL results.]} 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                 TODO 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \parbox{7in}{
@@
 fully automatic segmentation of the hippocampus and its subfields. First, we
-address the feasibility of whole hippocampus segmentation with a limited number
-of input atlases \citep{Chakravarty2013} by performing a multi-fold validation
-experiment using a subset of the Alzheimer's Disease Neuroimaging Initiative
-(ADNI) dataset manually segmented using the \prsnr{}-protocol, over a range of
-atlas and template library sizes, registration and label fusion methods. This
-experiment is then replicated with a first episode psychosis patient dataset
-to determine the behaviour of \mb{} when segmenting younger and differently
-diseased subjects, and to ensure that we have not also overfit our parameters to
-the aging or neurodegenerative brain. Next, we consider all of the images
-available in the \adnidataset{} sample, and compare \mb{} segmentations to other
-popular segmentation algorithms.  Finally, we conduct a leave-one-out validation
-of subfield segmentation using the \citet{Winterburn2013} manually segmented
-high-resolution MR atlases.
+address the very idea of boostrapping a template library from a limited number
+of input atlases \citep{Chakravarty2013} for whole hippocampus segmentation by
+conducting a multi-fold validation experiment over a range of atlas and template
+library sizes, registration and label fusion methods. This type of validation is
+done first on a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
+dataset with manual segmentations  \prsnr{}-protocol, and then replicated on a
+first episode psychosis patient dataset to determine the behaviour of \mb{} when
+segmenting younger and differently diseased subjects. Next, we compare \mb{}
+with other popular segmentation algorithms (\freesurfer{}, \fsl{}, \maper{}, and
+\snt{}) on all the images available in the \adnidataset{} sample.  Lastly, using
+the optimal parameter settings for \mb{} found from the previous experiments, we
+investigate hippocampal subfield segmentation by conducting a leave-one-out
+validation using the \citet{Winterburn2013} manually segmented high-resolution
+MR atlases.
 
@@
 segmentation quality of the \mb{} algorithm:
+
 \begin{itemize}
-\item Experiment 1 investigates the accuracy of \mb{} whole hippocampus
+  \item \expadnixval{} investigates the accuracy of \mb{} whole hippocampus
   segmentation of aging and Alzhiemer's diseased subjects over a wide range of
@@
   performance for use in subsequent experiments.
-\item Experiment 2 consists of a similar cross-validation design to verify these
-parameter settings on brain images of young, first episode psychosis
-patients. 
-\item Experiment 3 compares \mb{} whole hippocampus segmentations with those of
-well-known automated methods on the entire \adnidataset{} image dataset consisting
-of 
-\item Experiment 4 assesses hippocampal subfield segmentation quality in a
-leave-one-out cross-validation design. 
+
+  \item \expfepxval{} consists of a similar cross-validation to explore \mb{}
+  segmentations on the brain images of young, first episode psychosis patients,
+  and to establish parameter settings that do not overfit on a single dataset.
+
+  \item \expadni{} bridges \mb{} with the existing segmentation literature by
+  comparing \mb{} whole hippocampus segmentations with those of well-known
+  automated methods (\freesurfer, \fsl, \maper, \snt) on the entire
+  \adnidataset{} image dataset consistening of
   \Sexpr{length(means.complete$RID)} brain images of subjects diagnosed as
   cognitively normal, having mild cognitive impairment, or Alzheimer's disease.
+
+  \item \expsubfield{} assesses hippocampal subfield segmentation quality in a
+  leave-one-out cross-validation on the five high-resolution manually segmented
+  Winterburn MR atlases \citep(Winterburn2013).
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 1: \expadnixval}
+\subsection{\expadnixval{}: \expadnixvaldesc{}}
 \label{exp:adni-xval}
 
-\textbf{[TODO: cut, re-introduce as above]}
-Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using images
-from the ADNI dataset \citep{Jack2008}, and manual whole hippocampus
-segmentations following the protocol laid out in \citep{Pruessner2000}.  This
-form of cross-validation allows us to rigorously validate a large number of
-parameter settings of \mb{} (atlas and template library sizes, registration
-algorithm, and label fusion method) and select the best parameters to use in
-subsequent experiments.
-
-\textbf{[TODO: fix wording to relate to pruessner labels]}
-Note, in the Supplementary Materials we have replicated this experiment using the
-\snt{} semi-automated labels included with the ADNI dataset.
-
-\subsubsection{Materials and Methods: \expadnixval}
+In this experiment we explore the very idea of bootstrapping a template library
+for multi-atlas-based segmentation from a small number of input atlases. To do
+so, we conduct repeated cross-validations of \mb{} whilst varying the
+composition and sizes of the atlas and template libraries used, as well as
+varying the registration algorithm and label fusion method. The dataset used in
+this experiment is images from the ADNI dataset \citep{Jack2008} along with
+whole hippocampus labels manually segmented following the \citep{Pruessner2000}
+protocol.
+
+Note, in the Supplementary Materials we have replicated this experiment using
+the \snt{} semi-automated segmentations included as part of the the ADNI
+dataset.
+
+\subsubsection{Materials and Methods: \expadnixval{}}
+
 \paragraph{\adnidataset{} dataset} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expadnixval}
+\subsubsection{Results: \expadnixval{}}
 
@@
 increased consistency of \mb{} segmentations in the replication of this
-experiment (Experiment 5, Supplementary Materials) where atlases use a different
+experiment (\expadnixvalsnt{}, Supplementary Materials) where atlases use a different
 hippocampal definition is used. This strongly suggests that these results are
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 2: \expfepxval}
+\subsection{\expfepxval{}: \expfepxvaldesc{}}
 \label{exp:fep-xval}
@@
 
-\subsubsection{Materials and Methods: \expfepxval}
+\subsubsection{Materials and Methods: \expfepxval{}}
 \paragraph{First Episode Psychosis Dataset}
@@
 \mb{} is configured with an atlas library composed of the Winterburn T1 atlases
-(see Experiment 2) ignoring subfield delineations.  All images from the
+(see \expfepxval{}) ignoring subfield delineations.  All images from the
 first episode dataset are segmented by \mb{}.  The optimal size of template
-library, registration method, and label fusion method found in Experiment 1 are
+library, registration method, and label fusion method found in \expadnixval{} are
 used.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expfepxval}
+\subsubsection{Results: \expfepxval{}}
 \textbf{[TODO: move up results to here.]}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 3: \expadni}
+\subsection{\expadni{}: \expadnidesc{}}
 \label{exp:adni}
@@
 
-\subsubsection{Materials and Methods: \expadni}
+\subsubsection{Materials and Methods: \expadni{}}
 \paragraph{\adnidataset{} dataset} 
@@
 \mb{} was configured with an atlas library composed of the five Winterburn T1
-atlases (described in Experiment 4, section \ref{exp:winterburn-xval}), and size of template library,
+atlases (described in \expsubfield{}, section \ref{exp:winterburn-xval}), and size of template library,
 registration method, and label fusion method were determined by the optimal
-settings found in Experiment 1 (Section \ref{exp:adni-xval}).  The \fsl{} segmentation method was used via
+settings found in \expadnixval{} (Section \ref{exp:adni-xval}).  The \fsl{} segmentation method was used via
 the \verb+run_first_all+ script according to the FIRST user guide
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expadni}
-
-Based on the results from Experiment 1 (Section \ref{exp:adni-xval}), in this
+\subsubsection{Results: \expadni{}}
+
+Based on the results from \expadnixval{} (Section \ref{exp:adni-xval}), in this
 experiment \mb{} was configured with a template library of 21 randomly chosen
@@
   subjects randomly selected from each disease category in the subject pool used
-  in Experiment 1 (Section \ref{exp:adni-xval}. Sagittal slices are shown for each unlabelled T1-weighted
-  anatomical image. \snt{} labels appear in green, and \mb{} labels appear in
-  blue. Noted are examples of common segmentation idiosyncrasies: 
+  in \expadnixval{} (Section \ref{exp:adni-xval}. Sagittal slices are shown for
+  each unlabelled T1-weighted anatomical image. \snt{} labels appear in green,
+  and \mb{} labels appear in blue. Noted are examples of common segmentation
+  idiosyncrasies: 
   {\em (a)} over-estimation of hippocampal head and  
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment 4: \expsubfield}
+\subsection{\expsubfield{}: \expsubfielddesc{}}
 \label{exp:winterburn-xval}
@@
 
-\subsubsection{Materials and Methods: \expsubfield}
+\subsubsection{Materials and Methods: \expsubfield{}}
 \paragraph{Winterburn Atlases} 
@@
 The optimal size of template library, registration method, and label fusion
-method found in Experiment 1 are used.
+method found in \expadnixval{} are used.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expsubfield}
+\subsubsection{Results: \expsubfield{}}
 
@@
 \mb{} framework -- a methodology that requires very few input atlases in order
-to provide accurate and reliable segmentations.  Experiment 1 (Section
+to provide accurate and reliable segmentations.  \expadnixval{} (Section
 \ref{exp:adni-xval}) compares \mb{} to basic-\multiatlas{} segmentation by
 characterising the behaviour of each approach with varying parameter settings,
-and allows us to choose optimal settings for subsequent experiments.  Experiment
-2 (Section \ref{exp:winterburn-xval}) emonstrates the reliability of \mb{} in
-producing subfield segmentations which match the segmentation protocol of the
-input atlases despite contrast and resolution limitations in standard
-T1-weighted image volumes.  Experiments 2 and 3 validate whole hippocampal
+and allows us to choose optimal settings for subsequent experiments.
+\expfepxval{} (Section \ref{exp:winterburn-xval}) demonstrates the reliability of
+\mb{} in producing subfield segmentations which match the segmentation protocol
+of the input atlases despite contrast and resolution limitations in standard
+T1-weighted image volumes.  \expadnixval{} and \expfepxval{} validate whole hippocampal
 segmentation precision and consistency on populations with different ageing and
@@
 \citep{Winterburn2013,Pruessner2000,Hsu2002}.  
+\textbf{[TOOD: are all experiments referred to here?]}
 
@@
 a template library from a small set of labelled atlas images -- is validated in
-the cross validation conducted in Experiment 1 (and the replication in
-Experiment 5, Supplementary Materials). We find that both increasing the number
-of atlases and the number of templates used improves \mb{} segmentation over and
-above basic-\multiatlas{} segmentations using the same number of atlas images.
-That is, by taking the extra step of generating a template library from a small
-atlas set \mb{} is able to improve the final segmentation similarity. The
-magnitude of this improvement is greatest with a small number of atlases, but
-even with larger atlas libraries we have found that generating a template
-library reduces the variability in segmentation precision (i.e.  \mb{} more
-consistently produces high quality segmentations than does basic-\multiatlas{}
-segmentation). These effects do not appear dependant on the hippocampal
-segmentation protocol used.
+the cross validation conducted in \expadnixval{} (and the replication in in
+\expfepxval{} and \expadnixvalsnt{}, Supplementary Materials). We find that both
+increasing the number of atlases and the number of templates used improves \mb{}
+segmentation over and above basic-\multiatlas{} segmentations using the same
+number of atlas images.  That is, by taking the extra step of generating a
+template library from a small atlas set \mb{} is able to improve the final
+segmentation similarity. The magnitude of this improvement is greatest with a
+small number of atlases, but even with larger atlas libraries we have found that
+generating a template library reduces the variability in segmentation precision
+(i.e.  \mb{} more consistently produces high quality segmentations than does
+basic-\multiatlas{} segmentation). These effects do not appear dependant on the
+hippocampal segmentation protocol used.
 
@@
 have touched on previously. The 10-fold Monte Carlo cross-validation that we
-present in Experiment 1 is amongst one of the most stringent performed in the
-\multiatlas/segmentation literature. To the best of our knowledge, with the
-exception of \citep{Wang2011}, other groups using ADNI data for validation do at
-most a single round of leave-one-out-validation (Table \ref{tab:other-methods}).
-The thoroughness of our validation suggests that our results are reflective of a
-true average over the choice of parameter settings and are independent of atlas
-or template choice (provided the input atlases are consistently segmented).
+present in \expadnixval{} and \expfepxval{} is amongst one of the most stringent
+performed in the \multiatlas{} segmentation literature. To the best of our
+knowledge, with the exception of \citep{Wang2011}, other groups using ADNI data
+for validation do at most a single round of leave-one-out-validation (Table
+\ref{tab:other-methods}).  The thoroughness of our validation suggests that our
+results are reflective of a true average over the choice of parameter settings
+and are independent of atlas or template choice (provided the input atlases are
+consistently segmented).
 
@@
 Indeed, our replication of the 10-fold cross-validation using \snt{}
-segmentations (Experiment 5, Supplementary Materials) shows noticeably poorer
+segmentations (\expadnixvalsnt{}, Supplementary Materials) shows noticeably poorer
 mean similarity scores for both \mb{} and \multiatlas.
@@
 \FloatBarrier
-\subsection{Experiment 5: \expadnixvalsnt}
-
-This experiment is a replication of Experiment 1 using a pool of 69 images and
+\subsection{\expadnixvalsnt{}: \expadnixvalsntdesc{}}
+
+This experiment is a replication of \expadnixval{} using a pool of 69 images and
 \snt{} semi-automated segmentations from the ADNI dataset \citep{Hsu2002}. See
-Experiment 1 for full details on the ADNI dataset, and validation process.
-
-\subsubsection{Materials and Methods: \expadnixvalsnt}
+\expadnixval{} for full details on the ADNI dataset, and validation process. 
+\textbf{[TODO: use reference here]}
+
+\subsubsection{Materials and Methods: \expadnixvalsnt{}}
 \paragraph{Dataset} 
@@
 caption = 
-  "\\textbf{ADNI1 \\snt{} cross-validation subset demographics.} 
+  "\\textbf{\\adnidataset{} \\snt{} cross-validation subset demographics.} 
   CN     - Cognitively Normal. 
@@
 % ######################
-\subsubsection{Results: \expadnixvalsnt}
-
-As when comparing against manual labels in Experiment 1, we find similar
+\subsubsection{Results: \expadnixvalsnt{}}
+
+As when comparing against manual labels in \expadnixval{}, we find similar
 behaviour when comparing \mb{} segmentations to \snt{} labels: similarity scores
@@
 increases in improvement trending towards a plateau (Figure
-\ref{fig:ADNI1-xval-k-mean}).  As in Experiment 1, using \ants{} registration
+\ref{fig:ADNI1-xval-k-mean}).  As in \expadnixval{}, using \ants{} registration
 leads to significantly increased similarity scores, and there is no significant

### 2014-01-30 paper.Rnw
Fixed FEP experiment evaluation and methods description

@@
 
-\subsubsection{Materials and Methods: \expadnixval{}}
+\subsubsection{\expadnixval{}: Materials and Methods}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expadnixval{}}
+\subsubsection{\expadnixval{}: Results}
 
@@
 findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation.  Given
-these findings, in the remainder of this section only results using the \ants{}
+these findings, in the remainder of our experiments only results using the \ants{}
 registration algorithm and majority vote fusion will be shown.
@@
 increased consistency of \mb{} segmentations in the replication of this
-experiment (\expadnixvalsnt{}, Supplementary Materials) where atlases use a different
-hippocampal definition is used. This strongly suggests that these results are
-independant of the segmentation protocol used and are, instead, features of the
-\mb{} algorithm.
+experiment (\expadnixvalsnt{}, Supplementary Materials) where a different
+hippocampal definition is used (\snt labels available with the ADNI datasets).
+This strongly suggests that these results are independant of the segmentation
+protocol used and are, instead, features of the \mb{} algorithm.
 
@@
 To validate that the \mb{} works effectively in the context of other neurological
-disorders, in this experiment we use the Winterburn atlases to derive 
-whole hippocampal segmentations of a dataset of patients having had a single
-episode of psychosis. The resulting segmentations are assessed for quality by
-comparison with expert manual segmentations.
-
-\subsubsection{Materials and Methods: \expfepxval{}}
+disorders, in this experiment we use the Winterburn atlases
+\citep{Winterburn2013} to derive whole hippocampal segmentations of a dataset of
+patients having had a single episode of psychosis. The resulting segmentations
+are assessed for quality by comparison with expert manual segmentations.
+
+\subsubsection{\expfepxval{}: Materials and Methods}
 \paragraph{First Episode Psychosis Dataset}
@@
 
+\paragraph{Winterburn Atlases} 
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
+segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
+and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
+lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
+range in age from 29-57 years (mean age of 37), and include two males and three
+females.
+
 \paragraph{Experiment details} 
-\mb{} is configured with an atlas library composed of the Winterburn T1 atlases
-(see \expfepxval{}) ignoring subfield delineations.  All images from the
-first episode dataset are segmented by \mb{}.  The optimal size of template
-library, registration method, and label fusion method found in \expadnixval{} are
-used.
+The same overall design as \expadnixval{} is followed in this experiment: a
+Monte Carlo cross-validation (MCCV) is conducted using the pool of 81 first
+episode psychosis subject brain images and corresponding \prsnr-protocol manual
+segmentations. Five rounds of validation are conducted for each subject, and
+each atlas and template library size combination (1-9 atlases, 1-19 templates).
+In each round, images and their manual labels are randomly selected from the
+pool, and the remaining images are segmented using \mb{} with a random subset of
+the unlabelled images also serve as template images. Majority vote fusion, and
+the \ants{} registration algorithm are used, as these have shown to behave
+favourably in previous experiments.
+
+In addition to the MCCV, we segment the entire first episode psychosis dataset
+using \mb{} and two other popular segmentation packages, \fsl{} and
+\freesurfer{}. Firstly, \mb{} is used with the five Winterburn atlas images and labels as atlases, and a
+randomly selected subset of 19 psychosis dataset images as templates. Secondly,
+\mb{} is run using the same template images, but instead as atlases we use five
+additional five first episode psychosis subjects and corresponding manual
+segmentations (not included above). Finally, \fsl{} and \freesurfer{} are run
+with their default settings to segment the entire dataset. The \fsl
+\verb+run_first_all+ script was used according to the FIRST user guide
+\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}, and
+\freesurfer was run using \verb+recon-all+. \textbf{
 
 \paragraph{Evaluation method}
-The Pruessner and Winterburn hippocampal segmentation protocols differ slightly in
-the neuroanatomical features that are delineated \citep{Winterburn2013}. This
-difference poses a problem for evaluation by measuring overlap.  That is, since
-different protocols will necessarily produce segmentations that do not perfectly
-overlap, the degree of overlap cannot be solely used to compare segmentation
-methods using different protocols. In place of an overlap metric, we can assess
-the degree of (Pearson) correlation in average bilateral hippocampal volume of the
-subjects produced by each method.
+Manual and automated segmentations are directly compared using Dice's similarity
+coefficent (DSC). In the MCCV, the per-subject DSC value is computed as the
+average value over the five rounds of validation for a given atlas and template
+library size. The reported average DSC value per given atlas and template
+library size is the average DSC value over all subjects segmented.
+
+The Pruessner segmentation protocol differs slightly from the Winterburn
+protocol those used by \freesurfer and \fsl in the neuroanatomical features that
+are delineated (see \citet{Winterburn2013}, and Table
+\label{tab:subfield-segmentation-protocols} in the Discussion below).  This
+variation in protocol poses a problem if an overlap measure is used for
+evaluation: since different protocols will necessarily produce segmentations
+that do not perfectly overlap, the degree of overlap cannot be solely used to
+compare segmentation methods using different protocols. In place of an overlap
+metric, we assess the degree of (Pearson) correlation in average bilateral
+hippocampal volume produced by each method. Additionally, we evaluate the
+volume-related fixed and proportional biases in all segmentation methods using
+Bland-Altman plots \citep{Bland1986}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expfepxval{}}
+\subsubsection{\expfepxval{}: Results}
 \textbf{[TODO: move up results to here.]}
@@
 
-\subsubsection{Materials and Methods: \expadni{}}
+\subsubsection{\expadni{}: Materials and Methods}
 \paragraph{\adnidataset{} dataset} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expadni{}}
+\subsubsection{\expadni{}: Results}
 
@@
 
-\subsubsection{Materials and Methods: \expsubfield{}}
-\paragraph{Winterburn Atlases} 
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images. The
-segmentations include subfield segmentations for the cornu ammonis (CA) 1; CA2
-and CA3; CA4 and dentate gyrus; subiculum; and strata radiatum (SR), strata
-lacunosum (SL), and strata moleculare (SM). Subjects in the Winterburn atlases
-range in age from 29-57 years (mean age of 37), and include two males and three
-females.
-
+\subsubsection{\expsubfield{}: Materials and Methods}
+%^ TODO: winterburn atlases described in expfepxval
 \paragraph{Experiment details} 
@@
 segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
-of \mb{} segmentations using using DSC overlap scores and evaluate consistency
+of \mb{} segmentations using DSC overlap scores and evaluate consistency
 across the range of hippocampal sizes using Bland-Altman plots of subfield
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Results: \expsubfield{}}
+\subsubsection{\expsubfield{}: Results}
 
@@
 
-\subsubsection{Materials and Methods: \expadnixvalsnt{}}
+\subsubsection{\expadnixvalsnt{}: Materials and Methods}
 \paragraph{Dataset} 
@@
 % ######################
-\subsubsection{Results: \expadnixvalsnt{}}
+\subsubsection{\expadnixvalsnt{}: Results}
 

### 2014-02-01 paper.Rnw
Responded to all of mallar's comments

@@
 T}emplates), for the automatic segmentation of the hippocampus that minimizes
-the number of atlases needed while still achieving similar accuracy to
+the number of atlases needed while still achieving similar aggreement to
 \multiatlas{} approaches. Thus, our method acts as an accurate \multiatlas{}
@@
 \item change overlap/similarity --> agreement? choose one. 
-\item check that all chart labels are sensible 
-\item check that Experiment 5 matches Experiment 1
-\item check that all Experiment references resolve correctly
 \item fix figure 2 schematic language for the experiments (not 2a 2b). 
-\item ADNI-1 or ADNI1: standardize
 \item check that experiment references are formatted as Experiment 2 (Section 3.2)
-\item subfield or subregion -- choose one
 \end{enumerate}}
@@
 smaller atlas library. In their validation, LEAP was used to segment the whole
-hippocampus in the ADNI-1 baseline dataset, achieving a mean Dice score of 0.85
+hippocampus in the ADNI1 baseline dataset, achieving a mean Dice score of 0.85
 against semi-automated segmentations.
@@
 \begin{itemize}
-  \item \expadnixval{} investigates the accuracy of \mb{} whole hippocampus
+  \item \expadnixval{} investigates \mb{} whole hippocampus
   segmentation of aging and Alzhiemer's diseased subjects over a wide range of
@@
 caption = 
-  "\\textbf{ADNI-1 cross-validation subset demographics.} 
+  "\\textbf{ADNI1 cross-validation subset demographics.} 
   CN     - Cognitively Normal. 
@@
 agreement between two segmentations. It is one of the most widely used measures
-of segmentation accuracy, and we use it as the basis of comparison in this
+of segmentation agreement, and we use it as the basis of comparison in this
 experiment.
@@
 variable settings we tested. Additionally, by itself, using a weighted voting
-strategy did not significantly improve segmentation accuracy, contrary to the
+strategy did not significantly improve segmentation agreement, contrary to the
 findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation. Given
@@
 experiment (\expadnixvalsnt{}, Supplementary Materials) where a different
-hippocampal definition is used (\snt labels available with the ADNI datasets).
+hippocampal definition is used (\snt{} labels available with the ADNI datasets).
 This strongly suggests that these results are independant of the segmentation
@@
   \multiatlas{} segmentations in the validation. Variance of segmentation
-  accuracy between \mb{} and \multiatlas{} segmentation is computed for each
+  agreement between \mb{} and \multiatlas{} segmentation is computed for each
   subject across over 10 folds of Monte Carlo cross validation. Shown on
@@
 \end{figure}
+
+%
+%-------------------------------------------------------------------------------
+%
 
@@
 \label{exp:fep-xval}
-\textbf{[TODO: tie with exp 1]}
 To validate that the \mb{} works effectively in the context of other neurological
-disorders, in this experiment we use the Winterburn atlases
-\citep{Winterburn2013} to derive whole hippocampal segmentations of a dataset of
-patients having had a single episode of psychosis. The resulting segmentations
-are assessed for quality by comparison with expert manual segmentations.
+disorders, in this experiment we replicate the cross-validation done in
+\expadnixval{} a dataset of patients having had a single episode of psychosis. We
+also compare \mb{} segmentations with those of two well-known automated
+segmentation methods, \fsl{} and \freesurfer{}.
 
 \subsubsection{\expfepxval{}: Materials and Methods}
-\paragraph{First Episode Psychosis Dataset}
+\paragraph{First Episode Psychosis (FEP) Dataset}
 All patients were recruited and treated through the Prevention and Early
@@
 
-Manual segmentation of each subject whole hippocampus is produced following a
-validated segmentation protocol \citep{Pruessner2000}.
+Export whole hippocampal manual segmentation of each subject is produced
+following a validated segmentation protocol \citep{Pruessner2000}.
 
@@
 using \mb{} and two other popular segmentation packages, \fsl{} and
-\freesurfer{}. Firstly, \mb{} is used with the five Winterburn atlas images and labels as atlases, and a
-randomly selected subset of 19 psychosis dataset images as templates. Secondly,
-\mb{} is run using the same template images, but instead as atlases we use five
-additional five first episode psychosis subjects and corresponding manual
-segmentations (not included above). Finally, \fsl{} and \freesurfer{} are run
-with their default settings to segment the entire dataset. The \fsl
-\verb+run_first_all+ script was used according to the FIRST user guide
+\freesurfer{}. Firstly, \mb{} is used with the five Winterburn atlas images and
+labels as atlases, and a randomly selected subset of 19 psychosis dataset images
+as templates. Secondly, \mb{} is run using the same template images, but instead
+as atlases we use five additional five first episode psychosis subjects and
+corresponding manual segmentations (not included above). Finally, \fsl{} and
+\freesurfer{} are run with their default settings to segment the entire dataset.
+The \fsl{} \verb+run_first_all+ script was used according to the FIRST user guide
 \footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}, and
-\freesurfer was run using \verb+recon-all+. \textbf{
+\freesurfer{} was run using \verb+recon-all -all+.
 
@@
 The Pruessner segmentation protocol differs slightly from the Winterburn
-protocol those used by \freesurfer and \fsl in the neuroanatomical features that
-are delineated (see \citet{Winterburn2013}, and Table
-\label{tab:subfield-segmentation-protocols} in the Discussion below).  This
+protocol those used by \freesurfer{} and \fsl{} in the neuroanatomical features
+that are delineated (see \citet{Winterburn2013}, and Table
+\ref{tab:subfield-segmentation-protocols} in the Discussion below). This
 variation in protocol poses a problem if an overlap measure is used for
@@
 
+%
+%-------------------------------------------------------------------------------
+%
+
 \FloatBarrier
@@
 \subsubsection{\expfepxval{}: Results}
-\textbf{[TODO: move up results to here.]}
-
-% In this experiment \mb{} is applied to a dataset of images of first episode
-% psychosis patients, using the Winterburn atlases and a template library of
-% 21 subject images selected at random.  Expert manual whole hippocampal
-% segmentations are used as a gold standard.
+
+As in \expadnixval, we find that similarity score increases with a greater
+number of atlases or templates but quickly plateaus (Figure
+\label{fig:FEP-xval}). A maximum similarity score of \Sexpr{snt.mb.kappa.best}
+is found when using 9 atlases, 19 templates, ANTS registration, and majority
+vote label fusion.
+
+We found a close relationship in average hippocampal volume between the manual
+label volumes and \mb{} when using the Winterburn atlases, or manually segmented
+FEP subjects as atlases (Figure \label{fig:FEP-vols}). Both sets of volumes are
+correlated with Pearson $r > 0.88$. \freesurfer{} and \fsl{} volumes are both
+correlated with manual volumes at Pearson $r > 0.7$.
+
+As \citet{Bland1986} noted, high correlation amongst measures of the same
+quantity does not necessarily imply agreement (as correlation can be driven by a
+large range in true values, for instance). Figure \label{fig:FEP-Bland-Altman}
+shows Bland-Altman plots illustrating the level of agreement of each method with
+manual volumes. All methods show an obvious proportional bias: \freesurfer{}
+and \fsl{} markedly underestimate smaller hippocampi and over-estimate large
+hippocampi (the limits of agreement are between $-2482^3mm$ and $-784^3mm$, and
+between $-1653^3mm$ and $79^3mm$, respectively), whereas both \mb{} methods show
+a much less exaggerated, but conservative bias (limits of agreement between
+$-67^3mm$ and $766^3mm$ when using FEP atlases, and between $-333^3mm$ and
+$504^3mm$ when using Winterburn atlases). On average, \freesurfer{} and \fsl{}
+overestimate hippocampal volume by about $1600 mm^3$ and $800 mm^3$,
+respectively. In contrast, on average \mb{} underestimates volumes by about $300
+mm^3$ when using FEP atlases, and by about $80mm^3$ when using Winterburn
+atlases.
 
@@
 
-<<fig:FEP-Bland-Altman,dependson='fig:FEP-vols',fig.width=7,fig.height=3,fig.show='hide'>>=
+<<fig:FEP-Bland-Altman,dependson='fig:FEP-vols',fig.width=10,fig.height=3,fig.show='hide'>>=
 fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
@@
 To validate \mb{} segmentation quality with respect to other established
-automated hippocampal segmentation methods, \mb{} was applied to large dataset
+automated hippocampal segmentation methods, \mb{} was applied to a large dataset
 from the ADNI project and the resulting segmentations were compared to those
-produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated
-whole hippocampal segmentations (\snt{}) provided by ADNI.
-
+produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated whole
+hippocampal segmentations (\snt{}) provided by ADNI.
+
+%%%%%%%%%%
 \subsubsection{\expadni{}: Materials and Methods}
-\paragraph{\adnidataset{} dataset} 
+
+\paragraph{\adnidataset{} dataset}  %%%%%%%%%%%
 The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
-\snt{}, \maper, \freesurfer{} hippocampal volumes for a subset of images were
+\snt{}, \maper{}, \freesurfer{} hippocampal volumes for a subset of images were
 provided by ADNI, along with quality control data for each \freesurfer{}
@@
 
-\paragraph{Experiment details} 
-\mb{} was configured with an atlas library composed of the five Winterburn T1
-atlases (described in \expsubfield{}, section \ref{exp:winterburn-xval}), and size of template library,
-registration method, and label fusion method were determined by the optimal
-settings found in \expadnixval{} (Section \ref{exp:adni-xval}).  The \fsl{} segmentation method was used via
-the \verb+run_first_all+ script according to the FIRST user guide
-\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}.  All images
-in the \adnidataset{} dataset were segmented by both methods.
+\paragraph{Winterburn Atlases}  %%%%%%%%%%
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images, described
+more fully in \expfepxval, section \ref{exp:fep-xval}.
+
+\paragraph{Experiment details} %%%%%%%%%%
+\mb{} was configured with an atlas library composed of the five Winterburn and
+segmentations. A template library of 19 images were randomly selected from the target
+dataset of ADNI subjects, and \ants{} registration and majority vote label fusion
+were used as these were found to perform favourably in earlier experiments.
+
+\fsl{} segmentation was performed using the \verb+run_first_all+ script
+according to the FIRST user guide
+\footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}. All images in
+the \adnidataset{} dataset were segmented by both methods.
 
@@
 the segmentation did not pass). Only images meeting the conditions of having
-segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also
-passing quality control inspection were included in the analysis.
-
-\paragraph{Evaluation method}
-As in previous experiments, the \snt{} and Winterburn hippocampal segmentation protocols
-differ in the neuroanatomical features delineated, and so we assess \mb{} by
-the degree of (Pearson) correlation of average hippocampal volume across
-subjects. We also computed the correlation in hippocampal volume between
+segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and
+\mb{}) and also passing similary quality control inspection were included in the
+analysis.
+
+\paragraph{Evaluation method} %%%%%%%%%%
+As in previous experiments, the Winterburn hippocampal segmentation protocol
+differs in the delineated neuroanatomical features (\citet{Winterburn2013}, and
+Table \ref{tab:subfield-segmentation-protocols}, Discussion) and so we assess
+\mb{} by the degree of (Pearson) correlation of average hippocampal volume
+across subjects. We also computed the correlation in hippocampal volume between
 existing, established automated segmentation methods -- \fsl{}, \freesurfer{},
@@
 
-Based on the results from \expadnixval{} (Section \ref{exp:adni-xval}), in this
-experiment \mb{} was configured with a template library of 21 randomly chosen
-subject images (7 from each disease class) and majority vote label fusion.  The
-entire \adnidataset{} dataset was segmented by \mb{}, and the resulting volumes
-compared with those obtained by expert semi-automated segmentation (\snt{}), and
-three other automated segmentation techniques: \maper{}, \freesurfer{}, and
-\fsl{}. Table \ref{tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals} shows the
-total count of segmentations available, including a count of those which have
-failed a quality control inspection.  A total of
-\Sexpr{length(means.complete$RID)} images are included in the following
-analysis, having met quality criteria and having segmentations from every
-method.
-
 We found a close relationship in total bilateral hippocampal volume between all
@@
 Bland-Altman plots illustrate the level of agreement of each method with \snt{}
-segmentation hippocampal volumes (Figure \ref{fig:ADNI-Bland-Altman}).  As
-\citet{Bland1986} noted, high correlation amongst measures of the same quantity
-does not necessarily imply agreement (as correlation can be driven by a large
-range in true values, for instance).  All methods show an obvious proportional
-bias: \freesurfer{} and \fsl{} markedly under-estimate smaller hippocampi and
-over-estimate large hippocampi, whereas \maper{} and \mb{} show a reverse,
-conservative bias (Figure \ref{fig:ADNI-Bland-Altman}).  Additionally, all
-methods show a fixed volume bias, with \freesurfer{} and \fsl{} most
-dramatically over-estimating hippocampal volume by $2600 mm^3$ and $2800 mm^3$
-on average, respectively, and \maper{} and \mb{} within $250 mm^3$ on average.
+segmentation hippocampal volumes (Figure \ref{fig:ADNI-Bland-Altman}). All
+methods show an obvious proportional bias: \freesurfer{} and \fsl{} markedly
+under-estimate smaller hippocampi and over-estimate large hippocampi, whereas
+\maper{} and \mb{} show a reverse, conservative bias (Figure
+\ref{fig:ADNI-Bland-Altman}). Additionally, all methods show a fixed volume
+bias, with \freesurfer{} and \fsl{} most dramatically over-estimating
+hippocampal volume by $2600 mm^3$ and $2800 mm^3$ on average, respectively, and
+\maper{} and \mb{} within $250 mm^3$ on average.
 
@@
 @
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\textbf{[TODO: combine volume plot/box-plot into a single figure]}
-
-\begin{figure}[h]
-<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',cache=T,fig.width=6,fig.height=4>>=
+
+%%%%%%%%%
+% ADNI volumes by automated method
+%%%%%%%%%
+<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',fig.show='hide'>>=
 melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
@@
     theme(legend.direction = "horizontal", legend.position = "bottom")
-
-#    annotate("text", y = c(2000,1500,1000,500), x=Inf-100, size=2.5,
-#              label = c(paste("FS r =", round(cor(df$SNT, df$FS),2)),
-#                        paste("FSL r =", round(cor(df$SNT, df$FSL),2)),
-#                        paste("MAPER r =", round(cor(df$SNT, df$MAPER),2)),
-#                        paste("MAGeT r =", round(cor(df$SNT, df$MAGeT),2))),
-#              hjust=1.1,vjust=1.5) +
 @
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
-dataset by each of the four automated methods investigated (\freesurfer{} (FS), \fsl{},
-\maper{}, \mb{}) vs. \snt{}. Linear fit lines and Pearson correlations are shown
-for each method.}
-  \label{fig:ADNI-volumes-plot}
-\end{figure}
-
-\begin{figure}[h]
-<<fig:ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',cache=T,fig.width=7,fig.height=4>>=
+
+%%%%%%%%%
+% ADNI volumes by disease group and method
+%%%%%%%%%
+<<fig:ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',fig.show='hide'>>=
 melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
@@
 @
-  \caption{Subject mean hippocampal volume as measured in the \adnidataset{}
-dataset by \freesurfer{}, \fsl{}, \maper{}, \mb{}, and \snt{} vs. disease
-category.}
-  \label{fig:ADNI-volumes-boxplot}
-\end{figure}
-
-\begin{figure}
-<<fig:ADNI-Bland-Altman, cache=TRUE, dependson='ADNI-volumes-prep', fig.width=7, fig.height=7>>=
+
+%%%%%%%%%
+% ADNI Bland-Altman plots
+%%%%%%%%%
+<<fig:ADNI-Bland-Altman,dependson='ADNI-volumes-prep',fig.show='hide',fig.height=4,fig.width=10>>=
 names(means.complete)[8] = "FreeSurfer"
@@
 ggplot(melted, aes(x= mean, y = diff, colour=Diagnosis)) + 
-  facet_wrap( ~ method, nrow=2, scales="fixed") + 
+  facet_wrap( ~ method, nrow=1, scales="fixed") + 
   geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
@@
             hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of SNT and automated hippocampal volume ' (mm^3))) + 
-  ylab(expression('Difference in manual and automated hippocampal volume ' (mm^3))) + 
+  xlab(expression('Mean of SNT and automated volume ' (mm^3))) + 
+  ylab(expression('SNT - automated volume ' (mm^3))) + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
@@
 @
-  \caption{Bland-Altman plots comparing subject mean hippocampal volume as
-measured in the \adnidataset{} dataset by \snt{} segmentation and each of the
-four automated methods investigated (\freesurfer{}, \fsl{}, \maper{}, \mb{}). The overall
-mean difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by
-dashed horizontal lines. Linear fit lines are shown for each diagnosis group.
-Note, points below the mean difference indicate overestimation of the volume
-with respect to the \snt{} volume, and vice versa. }
-
+
+%%%%%%%%%
+% Big figure combining the above three ADNI plots
+%%%%%%%%%
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.48\textwidth} 
+    \includegraphics[width=\textwidth]{figure/fig:ADNI-volumes-plot}
+    \caption{Computed vs. semi-automated (\snt) segmentation volume} 
+    \label{fig:ADNI-volumes-plot}
+  \end{subfigure}%
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI-volumes-boxplot} 
+    \caption{Hippocampal volume by diagnosis group and segmentation method} 
+    \label{fig:ADNI-volumes-boxplot}
+  \end{subfigure}
+  \begin{subfigure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI-Bland-Altman}
+    \caption{Bland-Altman plots of computed vs. \snt{} hippocampus volume} 
     \label{fig:ADNI-Bland-Altman}
+  \end{subfigure}
+  \caption{\textbf{\adnidataset{} dataset segmentation.}
+(\ref{fig:ADNI-volumes-plot}) Subject mean hippocampal volume as measured by
+  each of the four automated methods (\freesurfer{} (FS), \fsl{},
+  \maper{}, \mb{}) versus the semi-automated \snt{} segmentation volumes. Linear
+  fit lines and Pearson correlations with \snt{} labels are shown for each method.
+(\ref{fig:ADNI-volumes-boxplot}) Mean hippocampal volume by method and disease
+  category. AD = Alzheimer's disease, LMCI = late-onset mild cognitive
+  impairment, and CN = cognitively normal.
+(\ref{fig:ADNI-Bland-Altman}) Bland-Altman plots comparing subject mean
+  hippocampal volume as measured in the \adnidataset{} dataset by \snt{}
+  segmentation and each automated methods. The overall mean difference in volume,
+  and limits of agreement ($\pm 1.96SD$) are shown by dashed horizontal lines.
+  Linear fit lines are shown for each diagnosis group. Note, points below the
+  mean difference indicate overestimation of the volume with respect to the \snt{}
+  volume, and vice versa.}
 \end{figure}
@@
 
+%
+%-------------------------------------------------------------------------------
+%
+
 \FloatBarrier
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection{\expsubfield{}: Materials and Methods}
-%^ TODO: winterburn atlases described in expfepxval
+
+\paragraph{Winterburn Atlases}  %%%%%%%%%%
+The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
+subfield segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images,
+described more fully in \expfepxval, section \ref{exp:fep-xval}.
+
+\paragraph{Healthy Control Dataset}  %%%%%%%%%%
+T1 MR images of 14 subjects were acquired as a part of an ongoing study
+at the Centre for Addiction and Mental Health (Table
+\ref{tab:WAval-healthy-demographics}). Subjects were known to be free of
+neuropsychiatric disorders and gave informed consent. These images were
+acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee, WI)
+using an 8-channel head coil with the enhanced fast gradient recalled echo
+3-dimensional acquisition protocol, FGRE-BRAVO, with the following parameters:
+$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
+thickness$=0.9mm$, 170 in-plane steps for an approximate 0.9mm isotropic voxel
+resolution.
+
 \paragraph{Experiment details} 
@@
 In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
-voxel images to simulate standard 3T T1-weighted resolution images. Image
+voxel resolution to simulate standard 3T T1-weighted resolution images. Image
 subsampling is performed using trilinear subsampling techniques. In each round
-of LOOCV, a single atlas is selected and segmented with \mb{} once using each
-possible triplet of atlas images from the remaining four unselected atlases.
-Thus, a total of ${3 \choose 4}=4$ segmentations are evaluated for each atlas
-image, for $5\times4=20$ segmentations evaluated in total.  We chose an atlas library
-with an odd number of images so as to ensure an unbiased label fusion when using
-majority voting (see Discussion).
-
-The template library used has a total of 19 images composed of all $0.9mm^3$
-voxel target images plus an additional set of 14 of healthy subject 3T T1 images
-acquired separately (Table \ref{tab:WAval-healthy-demographics}). These images
-were acquired on a 3T GE Discovery MR 750 system (General Electric, Milwaukee,
-WI) using an 8-channel head coil with the enhanced fast gradient recalled echo
-3-dimensional acquisition protocol, EFGRE-BRAVO, with the following parameters:
-$TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
-thickness$=0.9mm$, 170 in-plane steps for an approximate isotropic resolution of
-0.9mm dimension voxels.
-\textbf{[TODO: "What is the full voxel resolution of the EFGRE-BRAVO
-sequence?"]}
-The optimal size of template library, registration method, and label fusion
-method found in \expadnixval{} are used.
+of LOOCV, a single atlas image is selected and treated as a target image to be
+segmented by \mb{}. So as to have an odd-sized atlas library, atlas image is
+segmented once using each possible triple of atlas images, and corresponding
+manual segmentations, from the remaining four unselected atlases. Thus, for
+each of the five atlases, a total of ${3 \choose 4}=4$ segmentations are
+evaluated, resulting in a combined total of $5\times4=20$ segmentations
+evaluated overall. We chose an atlas library with an odd number of images so as
+to ensure unbiased label fusion when using majority voting (see Discussion).
+
+The template library used has a total of 19 images composed of all five
+resampled atlas images plus the additional 14 images from the healthy control
+dataset. The \ants{} registration algorithm was used for image registration, and
+majority voting was used for label fusion, as these methods proved most
+favourable in the previous whole hippocampal validation experiments.
 
@@
 \paragraph{Evaluation method}  
-Evaluating the accuracy of hippocampal subfields for T1 images at
-$0.9mm^3$-isotropic voxels is inherently ill-defined since there are no manual
-protocols for segmentation at this resolution to use as a reference. Instead, we
-can evaluate the reliability, or {\it precision}, with which \mb{} produces
-hippocampal subfields that correspond to the segmentation protocol used by the
-given atlas library images. By directly resampling the Winterburn atlas
-segmentations to $0.9mm^3$ voxels (using standard nearest-neighbour image
-resampling techniques) we obtain a scaled down version of the labels which
-preserve the original segmentation protocol within the limits of error from
-rounding and interpolation. Therefore, using the resampled Winterburn
-segmentations as definitive for the $0.9mm^3$ resolution we evaluate reliability
-of \mb{} segmentations using DSC overlap scores and evaluate consistency
-across the range of hippocampal sizes using Bland-Altman plots of subfield
-volumes.
+Evaluating the agreement of automated hippocampal subfield segmentations with
+manual segmetnations for T1 images at $0.9mm^3$-isotropic voxels is inherently
+ill-defined since there are no manual protocols for segmentation at this
+resolution. Instead, we must evaluate the reliability, or {\it precision}, with
+which \mb{} produces hippocampal subfields segmentations at this resolution that
+correspond in form to the segmentation protocol used by the given
+high-resolution atlas library images.
+
+By directly resampling the Winterburn atlas segmentations to $0.9mm^3$ voxels
+(using standard nearest-neighbour image resampling techniques) we obtain a
+subsampled version of the labels which preserve the original segmentation
+protocol within the limits of error from rounding and interpolation. Therefore,
+using the resampled Winterburn segmentations as definitive for the $0.9mm^3$
+resolution we evaluate reliability of \mb{} segmentations using DSC overlap
+scores and evaluate consistency across the range of hippocampal sizes using
+Bland-Altman plots of subfield volumes.
 
@@
 @
-  \caption{Similarity of \mb{} segmentation of subregions and the resampled
+  \caption{Similarity of \mb{} segmentation of subfields and the resampled
   Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
@@
 \mb{} segmentations and the resampled Winterburn atlases for each hippocampal
-subregion across all subjects and folds of the validation. Mean and standard
-deviation DSC scores of the subregions are shown in Table
+subfield across all subjects and folds of the validation. Mean and standard
+deviation DSC scores of the subfields are shown in Table
 \ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
 segmentations when perturbed slightly and compared to the originals. We find
-that the CA4/DG subregion shows the highest mean DSC score of
+that the CA4/DG subfield shows the highest mean DSC score of
 \Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
-subregions having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
+subfields having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
 \Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively. Both the CA4/DG and
@@
              data = wa.n.xval.bland.limits, hjust=1.1,vjust=-0.8, size=2.5) + 
-  xlab(expression('Mean of resampled and MAGeT-Brain subregion volume ' (mm^3))) + 
-  ylab(expression('Difference of manual and MAGeT-Brain subregion volume ' (mm^3))) + 
+  xlab(expression('Mean of resampled and MAGeT-Brain subfield volume ' (mm^3))) + 
+  ylab(expression('Difference of manual and MAGeT-Brain subfield volume ' (mm^3))) + 
   #scale_x_continuous(limits=c(0,1000)) + 
@@
 @
-  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subregion volume
+  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subfield volume
 across all folds of Winterburn atlas cross validation.} The overall mean
@@
 names(wa.perturb.table) = c("Subfield", "MAGeT", "0.9mm translation", "0.99 scaling", "1.01 scaling")
+wa.perturb.table=wa.perturb.table[c(1,2,3)] # remove scaling results
 
@@
 Simulated overlap similarity results are also given for resampled labels that were 
-translated by one voxel (i.e.: $0.9mm$ in all directions), scaled by 0.99, and scaled
-by 1.01 (representing a 1\\% shrinkage or scale in all three cardinal directions).
-Values are given as mean Dice's Kappa coefficient $\\pm$ stddev."
+translated by one voxel (i.e.: $0.9mm$) in all directions. Values are given as
+mean Dice's Similarity Coefficient (DSC) $\\pm$ standard deviation"
 latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
@@
 
+%
+%-------------------------------------------------------------------------------
+%
+
 \FloatBarrier
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\textbf{[TODO: fix experiment references]} 
 In this manuscript we have presented the implementation and validation of the
@@
 number of atlas images. That is, by taking the extra step of generating a
-template library from a small atlas set \mb{} is able to improve the final
-segmentation similarity. The magnitude of this improvement is greatest with a
+template library using target images, \mb{} is able to improve the overlap
+between the automatically generated segmentations and manually generated ``gold
+standard'' segmentations. The magnitude of this improvement is greatest with a
 small number of atlases, but even with larger atlas libraries we have found that
@@
 overall segmentation performance, as demonstrated by the STAPLE algorithm.
+However, it is also important to note that even in the absence of a more
+sophisticated label fusion algorithm, MAGeT Brain performs reasonably well in
+comparison to other groups that have tested new segmentation algorithm with
+Alzhiemer disease, mild cognitive impairment, and cognitively normal data from
+the ADNI database (Table \ref{tab:other-methods}. In addition, our validation
+in \expfepxval{} (with the first episode psychosis subjects) yields DSC's that
+are amongst the highest reported. Thus, more work is required to determine the
+extent to which label-fusion will improve the accuracy of our algorithm.
 
@@
 Our initial concern was that this dip in performance was a by-product of the
-\mb{} algorithm itself.  However, this this pattern is also found in the results
+\mb{} algorithm itself. However, this pattern is also found in the results
 of the \multiatlas{} segmentations we used in our experiments. We believe that
@@
 
-On that note, one author (JW), an expert manual rater (citep{Winterburn2013),
+On that note, one author (JW), an expert manual rater \citep{Winterburn2013},
 identified regular inconsistencies in the \snt{} segmentations: occurrences of
@@
 \ref{tab:subfield-segmentation-kappa} compares segmentation accuracy from some
-of these methods and \mb{}. The overlap DSC scores for \mb{} subregions are
+of these methods and \mb{}. The overlap DSC scores for \mb{} subfields are
 notably lower but a direct comparison of overlap values must be done cautiously.
@@
 are scarce, in our cross validation we are forced to use three rather than all
-five of the Winterburn atlases which, based on our findings with whole
-hippocampal segmentation, would lead to improved overlap similarity.
-Nevertheless, the advantage of evaluating \mb{} on standard 3T T1-weighted
-resolution MR images with a publically available atlas library is that our
-results reflect typical usage scenarios of researchers and clinicians. 
-\textbf{[TODO: mention that DSC by it's nature penalizes structures with high surface
-area-to-vol ratios, so necessarily penalise all subfield errors.]}
-\textbf{[TODO: discuss perturbed segmentation kappa values in relation to MAGeT
-error]}
+five of the Winterburn atlases (which, based on our findings with whole
+hippocampal segmentation, would lead to improved overlap similarity).
+Although having more atlases would be ideal in this context, these atlases are
+incredibly time consuming to generate \citep{Winterburn2013}. Nevertheless, the
+advantage of evaluating \mb{} on standard 3T T1-weighted resolution MR images
+with a publically available atlas library is that our results reflect typical
+usage scenarios of researchers and clinicians. \textbf{[TODO: mention that DSC
+by it's nature penalizes structures with high surface area-to-vol ratios, so
+necessarily penalise all subfield errors.]} \textbf{[TODO: discuss perturbed
+segmentation kappa values in relation to MAGeT error]}
 
 Our overlap DSC values demonstrates that we can reliably reproduce segmentations
-for the CA1, subiculum, and CA4/dentate subregions.  That the CA2/CA3 and
-molecular layers are less well reproduced (DSC < 0.5) should not be surprising
-as these are extremely thin and spatially convoluted regions that originally
-required high-resolution MRI for identification and so it is likely that the
-extents of these regions are well below the resolution and contrast offered by
+for the CA1, subiculum, and CA4/dentate subfields. %TODO: fix?
+That the CA2/CA3 and
+molecular layers are less well reproduced (DSC < 0.5) %TODO: fix!
+should not be surprising as these are extremely thin and spatially convoluted
+regions that originally required high-resolution MRI for identification and so
+it is likely that the extents of these regions are well below the resolution and
+contrast offered by
 standard T1-weighted images.
@@
 \ref{tab:subfield-segmentation-protocols} for a comparison of manual subfield
-segmentation methodologies. A further complication is that different
-researchers have differing operational definitions for the subfields and how
-they ought to be parcellated. The disagreement in the community has led to an
-international working group devoted to normalizing the ontology and segmentation
-rules for the hippocampal subfields
-(\url{http://www.hippocampalsubfields.com/}). The definitional and operational
-disagreements make evaluating the accuracy of automated methodologies with
-respect to a ``ground truth" (e.g. using overlap similarity metrics, such as
-Dice's Kappa) not particularly meaningful until these disagreements have been
-resolved.  Thus, because of these disagreements and the difficulting discerning
-subregions in standard T1-weighted images, we opted to speak of the precision or
-reliability of \mb{} subfield segmentation rather than speak of accuracy.
+segmentation methodologies. A further complication is that different researchers
+have differing operational definitions for the subfields and how they ought to
+be parcellated. The disagreement in the community has led to an international
+working group devoted to normalizing the ontology and segmentation rules for the
+hippocampal subfields (\url{http://www.hippocampalsubfields.com/}). Both of
+these disagreements suggest that direct comparison of automated methods using
+``ground truth''-based overlap similarity metrics, such as Dice's Similarity
+Coefficient, are not possible without carefully taking into account the
+differences in underlying segmentation protocols and image characteristics.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                               Conclusion 
+%                              Acknowledgements 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%
+%-------------------------------------------------------------------------------
+%
+
 \FloatBarrier
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%                                                                     %%%%%
+%%%%%       Experiment 5: SNT Hippocampal Subfield Cross-Validation       %%%%%
+%%%%%                                                                     %%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{\expadnixvalsnt{}: \expadnixvalsntdesc{}}
@@
 at \Sexpr{snt.mb.kappa.best} when using 9 atlases, 19
-templates, ANTS registration, and majority vote label fusion).  Compared to
+templates, ANTS registration, and majority vote label fusion. Compared to
 \multiatlas{} segmentations, we find \mb{} segmentations show increasing
@@
   with respect to \snt{} segmentation.}
-  Variance of segmentation accuracy between \mb{} and \multiatlas{} segmentation
+  Variance of segmentation agreement between \mb{} and \multiatlas{} segmentation
   is computed for each subject across over 10 folds of Monte Carlo       

### 2014-02-01 paper.Rnw
Condensed a few figures, and other small but significant changes

@@
 theme_set(theme_bw(base_size = 12))
+theme_update(
+  legend.key = element_rect(fill='white',colour='white'))
 lm_eqn = function(m) {
@@
 % ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
-<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load',fig.show='hide'>>=
+<<fig:ADNI1-xval-k-mean,dependson='ADNI1-xval-load',fig.show='hide',fig.width=10>>=
 stderr <- function(x) sqrt(var(x)/length(x))
@@
   ylab( "Mean similarity (DSC)" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
+  theme(legend.key.size = unit(0.7, 'lines'),
+        legend.position = "bottom", 
+        legend.box = "horizontal", 
+        legend.text = element_text(size=rel(0.7)),
         axis.text = element_text(size = 8))
@@
 \begin{figure}
+  \centering
+  \begin{subfigure}[t]{\textwidth}
     \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-mean}
-  \caption{Dice's similarity score vs. atlas and template library size. Dice's
-  similarity score of \mb{} segmentations and \psegs{} for 60 ADNI subjects over a
-  range of atlas and template library size, registration algorithm, and label
-  fusion method on 10 folds of Monte Carlo cross validation. Error bars indicate
-  standard error.} 
+    \caption{DSC vs. atlas and template library size} 
     \label{fig:ADNI1-xval-k-mean}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{0.48\textwidth}
+  \end{subfigure}%
+  \hfill
+  \begin{subfigure}[b]{0.48\textwidth}
     \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-k-diff}
@@
     \label{fig:ADNI1-xval-k-diff}
-  \end{subfigure}
+  \end{subfigure}%
   \hfill
-  \begin{subfigure}[t]{0.48\textwidth}
+  \begin{subfigure}[b]{0.48\textwidth}
     \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-variability}
@@
   \caption{\textbf{Whole hippocampus segmentation cross-validation on
-  Alzheimer's diseased subjects.}
-(\ref{fig:ADNI1-xval-k-diff}) shows increase in mean Dice's Similarity Coefficient of \mb{} over
-  \multiatlas{} segmentations vs. atlas and template library size when using the
-  \ants{} registration method, and majority-vote label fusion over 10 folds of
-  Monte Carlo cross validation. Segmentation similarity is computed against
-  \psegs{}. Error bars indicate standard error.
-(\ref{fig:ADNI1-xval-variability}) compares the variability in Dice scores of \mb{} vs.
-  \multiatlas{} segmentations in the validation. Variance of segmentation
-  agreement between \mb{} and \multiatlas{} segmentation is computed for each
-  subject across over 10 folds of Monte Carlo cross validation. Shown on
-  the y-axis is the p-value resulting from a t-test comparing the distribution of
-  variances at each parameter setting (atlas/template library size). Only points
-  where \mb{} mean variability is lower than \multiatlas{} are shown. Dashed
-  lines indicate a p-value of 0.05 and 0.01.}
+  Alzheimer's diseased subjects with \psegs{}.}
+(\ref{fig:ADNI1-xval-k-mean}) Average DSC score of \mb{} with manual
+  segmentations for 60 ADNI subjects taken over 10 folds of
+  cross-validation at each parameter setting. Error bars indicate standard
+  error. 
+(\ref{fig:ADNI1-xval-k-diff}) Increase in DSC of \mb{} over \multiatlas{}
+  segmentations. 
+(\ref{fig:ADNI1-xval-variability}) shows the significance of t-tests comparing
+  the variability in DSC scores of \mb{} and \multiatlas{} across validation
+  folds. Only points where \mb{} mean variability is lower than \multiatlas{} are
+  shown. Dashed lines indicate p-values of 0.05 and 0.01.}
 \end{figure}

### 2014-02-01 paper.Rnw
More figure fixes, and the last of the TODOs

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                 TODO 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\parbox{7in}{
-\subsection{TODO:}
-\begin{enumerate}
-\itemsep-0.5em
-\item check tense: we are is have has use/used find/found
-\item remove 'accuracy' unless specifically talking relative to ground truth (and
-    even then, qualify)
-\item change overlap/similarity --> agreement? choose one. 
-\item fix figure 2 schematic language for the experiments (not 2a 2b). 
-\item check that experiment references are formatted as Experiment 2 (Section 3.2)
-\end{enumerate}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                               INTRODUCTION
@@
                aes(y=k.mb, weight=1, group=interaction(reg_method, as.factor(atlases))), linetype='solid', width=0.5) + 
-  #stat_smooth(method='lm', formula=y~log(x)) + 
   scale_y_continuous(breaks=seq(0,1,0.01)) + 
@@
     \includegraphics[width=\textwidth]{figure/fig:ADNI1-xval-variability}
-    \caption{Difference in variability} 
+    \caption{Difference in variability with \multiatlas} 
     \label{fig:ADNI1-xval-variability}
@@
   \caption{\textbf{Whole hippocampus segmentation cross-validation on
-  Alzheimer's diseased subjects with \psegs{}.}
+  ADNI subjects with \psegs{}.}
 (\ref{fig:ADNI1-xval-k-mean}) Average DSC score of \mb{} with manual
@@
 
-\begin{figure}
-<<fig:WAxval-kappa>>=
+%%%%%%%%%%
+% Winterburn xValidation Kappas
+%%%%%%%%%%
+<<fig:WAxval-kappa,fig.show='hide'>>=
 wa.n.sim = read.csv('data/winterburn.xval.similarity.csv')
@@
 @
-  \caption{Similarity of \mb{} segmentation of subfields and the resampled
-  Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all folds of a
-  leave-one-out cross validation. Overlap score for each hemisphere is measured
-  separately. \label{fig:WAxval-kappa}}
-\end{figure}
 
@@
 that the CA4/DG subfield shows the highest mean DSC score of
-\Sexpr{mean(CA4DG)} $\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
-subfields having scores of \Sexpr{mean(CA1)} $\pm$\Sexpr{sd(CA1)} and
-\Sexpr{mean(Sub)} $\pm$\Sexpr{sd(Sub)}, respectively. Both the CA4/DG and
+\Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
+subfields having scores of \Sexpr{mean(CA1)}$\pm$\Sexpr{sd(CA1)} and
+\Sexpr{mean(Sub)}$\pm$\Sexpr{sd(Sub)}, respectively. Both the CA4/DG and
 molecular regions score below 0.5. These scores may seem low but not when taken
 in context and compared to existing (semi-)automated methods (see Discussion).
-The whole hippocampus is segmented with a mean DSC score of\Sexpr{mean(Whole)}
-$\pm$\Sexpr{sd(Whole)}.
-
-\begin{figure}
-<<fig:WAxval-Bland-Altman,fig.height=5,fig.width=7>>=
+The whole hippocampus is segmented with a mean DSC score
+of\Sexpr{mean(Whole)}$\pm$\Sexpr{sd(Whole)}.
+
+%%%%%%%%%%
+% Winterburn Bland-Altman
+%%%%%%%%%%
+<<fig:WAxval-Bland-Altman,fig.width=10,fig.show='hide'>>=
 wa.n.vol = read.csv('data/winterburn.xval.volumes.csv') # xval volumes
@@
 @
-  \caption{\textbf{Bland-Altman plots comparing unilateral hippocampal subfield volume
-across all folds of Winterburn atlas cross validation.} The overall mean
-difference in volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed
-horizontal lines. Linear fit lines are shown. Note, points below the mean
-difference indication overestimation of the volume with respect to the resampled
-volume, and vice versa. \label{fig:WAxval-Bland-Altman}}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{\textwidth}
+    \centering
+    \includegraphics[height=3in]{figure/fig:WAxval-kappa}
+    \caption{DSC score by subfield}
+    \label{fig:WAxval-kappa}
+  \end{subfigure}\hfill
+  \begin{subfigure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:WAxval-Bland-Altman}
+    \caption{Bland-Altman plots of computed vs. manual subfield volumes}
+    \label{fig:WAxval-Bland-Altman}
+  \end{subfigure}
+  \caption{\textbf{Hippocampal subfield cross-validation.}
+(\ref{fig:WAxval-kappa}) Similarity of \mb{} segmentation of subfields and the
+  resampled Winterburn atlas segmentations at $0.9mm^3$ voxel resolution, over all
+  validaiton folds. Overlap score for each hemisphere is measured separately.
+(\ref{fig:WAxval-Bland-Altman}) shows the agreement, by subfield, of computed
+  and manual volumes across all validation folds. The overall mean difference in
+  volume, and limits of agreement ($\pm 1.96SD$) are shown by dashed horizontal
+  lines. Linear fit lines are shown. Note, points below the mean difference
+  indication overestimation of the volume with respect to the resampled volume,
+  and vice versa.}
 \end{figure}
@@
 with a publically available atlas library is that our results reflect typical
-usage scenarios of researchers and clinicians. \textbf{[TODO: mention that DSC
-by it's nature penalizes structures with high surface area-to-vol ratios, so
-necessarily penalise all subfield errors.]} \textbf{[TODO: discuss perturbed
-segmentation kappa values in relation to MAGeT error]}
+usage scenarios of researchers and clinicians.
+
+One further complication common to all subfield segmentation evaluation is that,
+by its nature, the Dice's Similarity Coefficient score penalizes structures with
+high surface area to volume ratios. Therefore subfield DSC scores will generally
+be lower than whole hippocampal segmentations. We attempted to put this
+effect into perspective by comparing \mb{} subfield segmentation agreement with
+the agreement of voxel-shifted manual segmentations (Table
+\ref{tab:WAxval-perturb-kappa}). The results of this exercise show conclusively,
+despite the very limited number of atlases we had to work with, that \mb{}
+subfield segmentations are well within the bounds of error of a $0.3mm$ shift.
 
 Our overlap DSC values demonstrates that we can reliably reproduce segmentations
-for the CA1, subiculum, and CA4/dentate subfields. %TODO: fix?
-That the CA2/CA3 and
-molecular layers are less well reproduced (DSC < 0.5) %TODO: fix!
-should not be surprising as these are extremely thin and spatially convoluted
-regions that originally required high-resolution MRI for identification and so
-it is likely that the extents of these regions are well below the resolution and
-contrast offered by
-standard T1-weighted images.
+for the CA1, subiculum, and CA4/dentate subfields (DSC > 0.5). That the CA2/CA3
+and molecular layers are less well reproduced (DSC < 0.5) should not be
+surprising as these are extremely thin and spatially convoluted regions that
+originally required high-resolution MRI for identification and so it is likely
+that the extents of these regions are well below the resolution and contrast
+offered by standard T1-weighted images.
 
@@
 Computations were performed on the gpc supercomputer at the SciNet HPC
-Consortium. SciNet is funded by: the Canada Foundation for Innovation under the
-auspices of Compute Canada; the Government of Ontario; Ontario Research Fund -
-Research Excellence; and the University of Toronto.
-\textbf{[TODO: cite scinet paper]}
+Consortium \cite{Loken2010}. SciNet is funded by: the Canada Foundation for
+Innovation under the auspices of Compute Canada; the Government of Ontario;
+Ontario Research Fund - Research Excellence; and the University of Toronto.
 
@@
 \expadnixval{} for full details on the ADNI dataset, and validation process.
-\textbf{[TODO: use reference here]}
 
@@
 
-\textbf{[TODO: Merge figures]}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
-\begin{figure}
-<<fig:ADNI1-SNT-xval-k-mean, cache=TRUE, dependson='ADNI1-xval-load', fig.width=5, fig.height=5>>=
+%%%%%%%%%%%%%
+% ADNI1-SNT-XVAL: mean Kappa plot by settings 
+%%%%%%%%%%%%%
+<<fig:ADNI1-SNT-xval-k-mean,dependson='ADNI1-xval-load',fig.width=10,fig.show='hide'>>=
 stderr <- function(x) sqrt(var(x)/length(x))
 ggplot(subset(snt_xval_mean, (templates.mb*atlases) %% 2 == 1),
-       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + #as.factor((atlases*templates.mb) %% 2 ))) + 
+       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases), linetype=reg_method)) + 
+      #as.factor((atlases*templates.mb) %% 2 ))) + 
   facet_grid(.~method.mb) + 
@@
                geom='errorbar',
-               aes(y=k.mb, weight=1, group=interaction(reg_method,atlases)), width=0.5) + 
-  #stat_smooth(method='lm', formula=y~log(x)) + 
-  scale_y_continuous(breaks=seq(0,1,0.05)) + 
+               aes(y=k.mb, weight=1, 
+                 group=interaction(reg_method,atlases)), width=0.5) + 
+  scale_y_continuous(breaks=seq(0,1,0.01)) + 
   scale_colour_hue(name="Number of Atlases") +
+  scale_linetype(name="Registration Method") +
   xlab( "Number of Templates" ) + 
   ylab( "Mean similarity (DSC)" ) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
+  theme(legend.key.size = unit(0.7, 'lines'),
+        legend.position = "bottom", 
+        legend.box = "horizontal", 
+        legend.text = element_text(size=rel(0.7)),
         axis.text = element_text(size = 8))
 @
-  \caption{Mean Dice's Similarity Coefficient of \mb{} segmentations relative to
-  \snt{} segmentations for 69 ADNI subjects vs. atlas and template library
-  size, registration algorithm, and label fusion method. 
-  Error bars indicate standard error.
-  \label{fig:ADNI1-SNT-xval-k-mean}}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: "increase" in mean kappa over \multiatlas
-\begin{figure}
-<<fig:ADN11-SNT-xval-k-diff,cache=TRUE,dependson='ADNI1-xval-load', fig.width=5, fig.height=5>>=
+
+%%%%%%%%%%%%%
+% ADNI1-SNT-XVAL: "increase" in mean kappa over \multiatla
+s%%%%%%%%%%%%%
+<<fig:ADNI1-SNT-xval-k-diff,dependson='ADNI1-xval-load',fig.show='hide'>>=
 ggplot(subset(snt_xval_mean, reg_method=="ANTS" & 
@@
 @
-  \caption{Increase in mean Dice's Similarity Coefficient of \mb{} over
-  \multiatlas{} segmentations vs. atlas and template library size when using the
-  \ants{} registration method, and majority-vote label fusion. 
-  Segmentation similarity is computed against \snt{} segmentations. 
-  Error bars indicate standard error.
-  \label{fig:ADNI1-SNT-xval-k-diff}}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-SNT-XVAL: show variability of MAGeT over parameters
-\begin{figure}
-<<fig:ADN1-SNT-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=5, fig.width=5>>=
+
+%%%%%%%%%%%%%
+% ADNI1-SNT-XVAL: variability
+%%%%%%%%%%%%%
+<<fig:ADNI1-SNT-xval-variability,dependson='ADNI1-Xval-prep',fig.show='hide'>>=
 ants = subset(snt_xval_data, 
@@
 @
-  \caption{\textbf{Difference in variability of \mb{} vs. \multiatlas{} segmentation
-  with respect to \snt{} segmentation.}
-  Variance of segmentation agreement between \mb{} and \multiatlas{} segmentation
-  is computed for each subject across over 10 folds of Monte Carlo       
-  cross validation. 
-  Shown on the y-axis is the p-value from a t-test comparing the distribution of
-  variances at each parameter setting (atlas/template library size). Only points
-  where \mb{} mean variability is lower than \multiatlas{} are shown. Dashed
-  lines indicate a p-value of 0.05 and 0.01.}
-  \label{ADNI1-SNT-xval-variability}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI1-SNT-xval-k-mean}
+    \caption{DSC vs. atlas and template library size} 
+    \label{fig:ADNI1-SNT-xval-k-mean}
+  \end{subfigure}%
+  \hfill
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI1-SNT-xval-k-diff}
+    \caption{Increase in similarity score over \multiatlas} 
+    \label{fig:ADNI1-SNT-xval-k-diff}
+  \end{subfigure}%
+  \hfill
+  \begin{subfigure}[b]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{figure/fig:ADNI1-SNT-xval-variability}
+    \caption{Difference in variability with \multiatlas} 
+    \label{fig:ADNI1-SNT-xval-variability}
+  \end{subfigure}
+  \caption{\textbf{Whole hippocampus segmentation cross-validation on
+  ADNI subjects with \snt{} segmentations.}
+(\ref{fig:ADNI1-SNT-xval-k-mean}) Average DSC score of \mb{} with \snt{} 
+  segmentations for 69 ADNI subjects taken over 10 folds of cross-validation at
+  each parameter setting. Error bars indicate standard error. 
+(\ref{fig:ADNI1-SNT-xval-k-diff}) Increase in DSC of \mb{} over \multiatlas{}
+  segmentations. 
+(\ref{fig:ADNI1-SNT-xval-variability}) shows the significance of t-tests comparing
+  the variability in DSC scores of \mb{} and \multiatlas{} across validation
+  folds. Only points where \mb{} mean variability is lower than \multiatlas{} are
+  shown. Dashed lines indicate p-values of 0.05 and 0.01.}
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2014-02-01 paper.Rnw
Typo fixes from Ainsley

@@
 \newcommand{\expfepxvaldesc}{Whole Hippocampus Segmentation Cross-Validation --- First Episode of Psychosis} 
-\newcommand{\expadnidesc}{Whole Hippocampus Segmentation Comparision --- ADNI1 Complete 1Yr} 
+\newcommand{\expadnidesc}{Whole Hippocampus Segmentation Comparison --- ADNI1 Complete 1Yr} 
 \newcommand{\expsubfielddesc}{Hippocampal Subfield Segmentation Cross-Validation} 
@@
 
-One methodology that can be used to mitigate these sources of errors involves
+One methodology that can be used to mitigate these sources of error involves
 the use of multiple manually segmented atlases and probabilistic segmentation
@@
 from a library of 80 atlases, were required to achieve accurate segmentations of
-the hippocampus. Additionally, several methods have been explored for label
-fusion, including the STAPLE algorithm (Simultaneous Truth And Performance Level
-Estimation; \citet{Warfield2004}) that computes a probabilistic segmentation
-using an expectation-maximization framework from an set of competing
-segmentations, or others where a subset of segmentations can be estimated using
-metrics such as the sum of squared differences in the regions of interest to be
-segmented \citep{Coupe2011}.
+the hippocampus. Also, several methods have been explored for label
+fusion. For example, the STAPLE algorithm (Simultaneous Truth And Performance
+Level Estimation; \citet{Warfield2004}) uses an expectation-maximization
+framework to compute a probabilistic segmentation from a set of competing
+regmentations, or the work of \citet{Coupe2011} who show that a subset of
+segmentations can be estimated using metrics, such as the sum of squared
+differences in the regions of interest to be segmented.
 
@@
 The novelty of this manuscript is the extension of our \multiatlas{} methodology
-to the hippocampus using more than a single input atlas, while simultaneously
-limiting the number of inputs used during segmentation, and so demonstrating that
-reliable application of the atlases' subfield segmentation to an unlabelled
-target image is indeed possible using this methodology.
+to the segmentation of hippocampus. Additionally, in this paper we rigorously explore the
+effects of using multiple input atlases, of varying the size of the template
+library constructed, and registration and label fusion methods. As a result, we
+aim to demonstrate that it is indeed possible to reliably apply the segmentation
+represented in a very small set of segmented input atlases to an unlabelled
+target image set.
 
@@
 T}emplates) bootstraps the creation of a large template library given a limited
-input atlas library, and then uses the template library in basic \multiatlas
+input atlas library, and then uses the template library in basic \multiatlas{}
 segmentation. Images for the template library are selected from a set of input
@@
 template images and then to each subject image (including those used in the template
-library). The candidate the labels for a subject are then fused into a final
+library). The candidate labels for a subject are then fused into a final
 segmentation. \label{fig:MAGeT}}
@@
   \item \expadnixval{} investigates \mb{} whole hippocampus
-  segmentation of aging and Alzhiemer's diseased subjects over a wide range of
+  segmentation of aging and Alzheimer's diseased subjects over a wide range of
   parameter settings using a Monte Carlo cross-validation design. The results of
@@
   automated methods (\freesurfer, \fsl, \maper, \snt) on the entire
-  \adnidataset{} image dataset consistening of
-  \Sexpr{length(means.complete$RID)} brain images of subjects diagnosed as
-  cognitively normal, having mild cognitive impairment, or Alzheimer's disease.
+  \adnidataset{} image dataset consisting of \Sexpr{length(means.complete$RID)}
+  brain images of subjects diagnosed as cognitively normal, having mild cognitive
+  impairment, or Alzheimer's disease.
 
@@
   leave-one-out cross-validation on the five high-resolution manually segmented
-  Winterburn MR atlases \citep(Winterburn2013).
+  Winterburn MR atlases \citep{Winterburn2013}.
 \end{itemize}
@@
 this experiment is images from the ADNI dataset \citep{Jack2008} along with
-whole hippocampus labels manually segmented following the \citep{Pruessner2000}
-protocol.
+whole hippocampus labels manually segmented following the \prsnr-protocol
+\citep{Pruessner2000}.
 
 Note, in the Supplementary Materials we have replicated this experiment using
-the \snt{} semi-automated segmentations included as part of the the ADNI
+the \snt{} semi-automated segmentations included as part of the ADNI
 dataset.
@@
 \paragraph{\adnidataset{} dataset} 
+\label{adnidataset}
 Data used in the preparation of this article were obtained from the Alzheimer's
@@
 
-60 1.5T images were arbitrarily selected from the baseline scans in the {\em
-\adnidataset{}} standardized dataset. 20 subjects were chosen from each disease
+Sixty 1.5T images were arbitrarily selected from the baseline scans in the {\em
+\adnidataset{}} standardized dataset. Twenty subjects were chosen from each disease
 category: cognitively normal (CN), mild cognitive impairment (MCI) and
@@
 authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
+August 2012. The image dataset used was the \adnidataset{} standardized dataset
 available from ADNI \footnote{
@@
 \citep{Wyman2012}. This image collection contains uniformly pre-processed images
-which have been designated to be the "best" after quality control. All images
+which have been designated to be the ``best'' after quality control. All images
 were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
 Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008}. Representative 1.5T imaging parameters were TR =
+described in \citet{Jack2008}. Representative 1.5T imaging parameters were TR =
 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
-yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.
+yielding voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$.
 
@@
   AD     - Alzheimer's Disease. 
-  Hisp   - Hispanic. 
   CDR-SB - Clinical Dementia Rating-Sum of Boxes. 
@@
   similarity is calculated by the normalised mutual information score over the
-  region of interest\citep{Studholme2001}. The {\tt itk\_similarity} utility
+  region of interest \citep{Studholme2001}. The {\tt itk\_similarity} utility
   from the EZMinc toolkit\footnote{\url{https://github.com/vfonov/EZminc}} is
@@
 maximum similarity score of \Sexpr{mb.kappa.best} is found when using 9 atlases,
-19 templates, ANTS registration, and majority vote label fusion). The \ants{}
+19 templates, ANTS registration, and majority vote label fusion. The \ants{}
 registration method consistently out-performs \animal{} registration over all
@@
 hippocampal definition is used (\snt{} labels available with the ADNI datasets).
-This strongly suggests that these results are independant of the segmentation
+This strongly suggests that these results are independent of the segmentation
 protocol used and are, instead, features of the \mb{} algorithm.
@@
 templates since with these configurations we found significantly decreased
-performance. We believe this is results from an inherent bias in the majority
-vote fusion method used (see Discussion).
+performance. We believe this results from an inherent bias in the majority vote
+fusion method used (see Discussion).
 
@@
 \label{exp:fep-xval}
-To validate that the \mb{} works effectively in the context of other neurological
-disorders, in this experiment we replicate the cross-validation done in
-\expadnixval{} a dataset of patients having had a single episode of psychosis. We
-also compare \mb{} segmentations with those of two well-known automated
-segmentation methods, \fsl{} and \freesurfer{}.
+To validate that the \mb{} works effectively in the context of other
+neurological disorders, in this experiment we replicate the cross-validation
+done in \expadnixval{} with a dataset of patients having had a single episode of
+psychosis. We also compare \mb{} segmentations with those of two well-known
+automated segmentation methods, \fsl{} and \freesurfer{}.
 
@@
 
-Export whole hippocampal manual segmentation of each subject is produced
+Expert whole hippocampal manual segmentation of each subject is produced
 following a validated segmentation protocol \citep{Pruessner2000}.
@@
 pool, and the remaining images are segmented using \mb{} with a random subset of
-the unlabelled images also serve as template images. Majority vote fusion, and
+the unlabelled images also serving as template images. Majority vote fusion, and
 the \ants{} registration algorithm are used, as these have shown to behave
@@
 as templates. Secondly, \mb{} is run using the same template images, but instead
-as atlases we use five additional five first episode psychosis subjects and
+as atlases we use five additional first episode psychosis subjects and
 corresponding manual segmentations (not included above). Finally, \fsl{} and
@@
 
-The Pruessner segmentation protocol differs slightly from the Winterburn
-protocol those used by \freesurfer{} and \fsl{} in the neuroanatomical features
-that are delineated (see \citet{Winterburn2013}, and Table
-\ref{tab:subfield-segmentation-protocols} in the Discussion below). This
-variation in protocol poses a problem if an overlap measure is used for
-evaluation: since different protocols will necessarily produce segmentations
-that do not perfectly overlap, the degree of overlap cannot be solely used to
-compare segmentation methods using different protocols. In place of an overlap
-metric, we assess the degree of (Pearson) correlation in average bilateral
-hippocampal volume produced by each method. Additionally, we evaluate the
-volume-related fixed and proportional biases in all segmentation methods using
-Bland-Altman plots \citep{Bland1986}.
+The \prsnr{} segmentation protocol differs slightly from the Winterburn
+protocol, and those used by \freesurfer{} and \fsl{}, in the inclusion of
+neuroanatomical features and the manner they are delineated (see
+\citet{Winterburn2013}, and Table \ref{tab:subfield-segmentation-protocols} in
+the Discussion below). This variation in protocol poses a problem if an overlap
+measure is used for evaluation: since different protocols will necessarily
+produce segmentations that do not perfectly overlap, the degree of overlap
+cannot be solely used to compare segmentation methods using different protocols.
+In place of an overlap metric, we assess the degree of (Pearson) correlation in
+average bilateral hippocampal volume produced by each method. Additionally, we
+evaluate the volume-related fixed and proportional biases in all segmentation
+methods using Bland-Altman plots \citep{Bland1986}.
 
@@
 As in \expadnixval, we find that similarity score increases with a greater
-number of atlases or templates but quickly plateaus (Figure
-\label{fig:FEP-xval}). A maximum similarity score of \Sexpr{snt.mb.kappa.best}
-is found when using 9 atlases, 19 templates, ANTS registration, and majority
-vote label fusion.
+number of atlases or templates but quickly plateaus (Figure \ref{fig:FEP-xval}).
+A maximum similarity score of \Sexpr{snt.mb.kappa.best} is found when using 9
+atlases, 19 templates, ANTS registration, and majority vote label fusion.
 
@@
 label volumes and \mb{} when using the Winterburn atlases, or manually segmented
-FEP subjects as atlases (Figure \label{fig:FEP-vols}). Both sets of volumes are
+FEP subjects as atlases (Figure \ref{fig:FEP-vols}). Both sets of volumes are
 correlated with Pearson $r > 0.88$. \freesurfer{} and \fsl{} volumes are both
@@
 quantity does not necessarily imply agreement (as correlation can be driven by a
-large range in true values, for instance). Figure \label{fig:FEP-Bland-Altman}
+large range in true values, for instance). Figure \ref{fig:FEP-Bland-Altman}
 shows Bland-Altman plots illustrating the level of agreement of each method with
@@
 The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
-\snt{}, \maper{}, \freesurfer{} hippocampal volumes for a subset of images were
+\snt{}, \maper{}, and \freesurfer{} hippocampal volumes for a subset of images were
 provided by ADNI, along with quality control data for each \freesurfer{}
@@
 
-Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
-August 2012. The image dataset used was the "\adnidataset{}" standardized dataset
-available from ADNI \footnote{
-\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
-\citep{Wyman2012}. This image collection contains uniformly pre-processed images
-which have been designated to be the "best" after quality control. All images
-were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
-Systems or Siemens Medical Solutions) at multiple sites using the protocol
-described in \citep{Jack2008}. Representative 1.5T imaging parameters were TR =
-2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8^{\circ}$, field of view = 240 x
-240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
-yielding a voxel dimensions of $1.25mm \times 1.25mm \times 1.2mm$. Clinical
-and demographic data for the entire \adnidataset{} dataset are shown in Table
-\ref{tab:ADNI1-Dataset-Demographics}.
+See Section \ref{adnidataset} for study details, inclusion criteria and imaging
+characteristics.
 
@@
 \paragraph{Experiment details} %%%%%%%%%%
-\mb{} was configured with an atlas library composed of the five Winterburn and
-segmentations. A template library of 19 images were randomly selected from the target
-dataset of ADNI subjects, and \ants{} registration and majority vote label fusion
-were used as these were found to perform favourably in earlier experiments.
+\mb{} was configured with an atlas library composed of the five Winterburn atlas
+images and segmentations. A template library of 19 images were randomly selected
+from the target dataset of ADNI subjects, and \ants{} registration and majority
+vote label fusion were used as these were found to perform favourably in earlier
+experiments.
 
@@
 segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and
-\mb{}) and also passing similary quality control inspection were included in the
+\mb{}) and also passing similar quality control inspection were included in the
 analysis.
@@
 \label{exp:winterburn-xval}
-The previous experiment assesses \mb{} peformance on whole hippocampus
+The previous experiment assesses \mb{} performance on whole hippocampus
 segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
@@
 Evaluating the agreement of automated hippocampal subfield segmentations with
-manual segmetnations for T1 images at $0.9mm^3$-isotropic voxels is inherently
+manual segmentations for T1 images at $0.9mm^3$-isotropic voxels is inherently
 ill-defined since there are no manual protocols for segmentation at this
@@
 validation datasets are composed equally of cases and control subjects, and use
-manual segmenation labels as ground truth in computing DSC scores. AD =
+manual segmentation labels as ground truth in computing DSC scores. AD =
 Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
@@
 \mb{} framework -- a methodology that requires very few input atlases in order
-to provide accurate and reliable segmentations. \expadnixval{} (Section
-\ref{exp:adni-xval}) compares \mb{} to basic-\multiatlas{} segmentation by
-characterising the behaviour of each approach with varying parameter settings,
-and allows us to choose optimal settings for subsequent experiments.
-\expfepxval{} (Section \ref{exp:winterburn-xval}) demonstrates the reliability of
-\mb{} in producing subfield segmentations which match the segmentation protocol
-of the input atlases despite contrast and resolution limitations in standard
-T1-weighted image volumes. \expadnixval{} and \expfepxval{} validate whole hippocampal
-segmentation precision and consistency on populations with different ageing and
-neuropsychiatric characteristics. All of these experiments together demonstrate
-that \mb{} algorithmic performance is not dependent on a single definition of
-the hippocampus but is effective with differing hippocampal definitions
-\citep{Winterburn2013,Pruessner2000,Hsu2002}. 
-\textbf{[TOOD: are all experiments referred to here?]}
+to provide accurate and reliable segmentations. Both \expadnixval{} (Section
+\ref{exp:adni-xval}) and \expfepxval (Section \ref{exp:fepi-xval}) compare \mb{}
+to basic-\multiatlas{} segmentation by characterising the change in segmentation
+quality with varying parameter settings (atlas and template library sizes,
+registration method, and label fusion method) and differing age and
+neuropsychiatric populations. Together, these experiments allow us to choose
+optimal \mb{} parameter settings for use in subsequent experiments. \expadni{}
+(Section \ref{exp:adni}) demonstrates that across
+\Sexpr{length(means.complete$RID)} images from the \adnidataset{} dataset, \mb{}
+performs as well as, or better, than other established and popular methods.
+Finally, \expsubfield{} (Section \ref{exp:winterburn-xval}) demonstrates the
+reliability of \mb{} in producing subfield segmentations which match the
+segmentation protocol of the input atlases despite contrast and resolution
+limitations in standard T1-weighted image volumes.  All of these experiments
+together demonstrate that \mb's algorithmic performance is not dependent on a
+single definition of the hippocampus but is effective with differing hippocampal
+definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}, across image types,
+and subject populations.
 

### 2014-02-04 paper.Rnw
Abstract mods

@@
 library, formed from a subset of target images, via transformations estimated by
-nonlinear image registration. The resultant segmentations are then propagated
-to each target image and fused using a label fusion method. To explore the
-effect of atlases and templates library size, registation and label
-fusion method, we conduct a 10-fold Monte Carlo cross-validation of \mb{} of whole
-hippocampal segmentation on 60 ADNI subjects, manually segmented using the
-\prsnr{}-protocol, over a range of parameter settings. Using the best settings
-found, we then conduct a leave-one-out cross-validation (LOOCV) of hippocampal
-subfield segmentation using five high-resolution manually segmented atlases.
-Two final experiments assess \mb{} when applied to first episode psychosis and
-Alzheimer's disease populations, and \mb{} segmentations are compared with
-existing automated methods (\fsl{}, \freesurfer{}, \maper{}, and \snt) and
-biases are explored.\\
+nonlinear image registration. The resulting segmentations are then propagated
+to each target image and fused using a label fusion method.
+
+We conduct two separate Monte Carlo cross-validation experiments comparing \mb{}
+and \multiatlas{} whole hippocampal segmentation using differing atlas and
+template library sizes, and registation and label fusion methods. The first
+experiment is a 10-fold validation (per parameter setting) over 60 subjects
+taken from the ADNI Alzheimer's disease imaging database, and the second is a
+5-fold validation over 81 subjects having had a first episode of psychosis. In
+both cases, automated segmentations are compared with manual segmentions
+following the \prsnr{}-protocol. Using the best settings found from these
+experiments, we segment \Sexpr{length(means.complete$RID)} images of the
+\adnidataset{} dataset and compare these with segmentations from existing
+automated methods: \fsl{}, \freesurfer{}, \maper{}, and \snt.  Finally, we
+conduct a leave-one-out cross-validation (LOOCV) of hippocampal subfield
+segmentation using five high-resolution manually segmented atlases
+\citep{Winterburn2013}.
+
 \textbf{Results:} \textbf{[TODO: update with FEP-XVAL and subfield XVAL results.]} 

### 2014-02-08 paper.Rnw
safety checkin'

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: prep
-% Prepares the data in a the form needed for plotting.
-<<ADNI1-xval-load,include=FALSE,cache=TRUE>>=
+<<load,include=FALSE,cache=TRUE>>=
+
+########################################
+# ADNI1-XVAL: prep
+# Prepares the data in a the form needed for plotting.
 jens_atlases = read.csv('data/jens_atlases.csv')
@@
                                  atlases      == 1)$k_diff)
-@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-COMPLETE
-% Prepares the data in the form needed for plotting.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-<<ADNI-volumes-prep,echo=F,cache=T>>=
+
+########################################
+# ADNI1-COMPLETE
+# Prepares the data in the form needed for plotting.
+########################################
 means.complete   <- read.csv('data/cache/ADNI1:qc.csv')
 package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
-@
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ADNI1-XVAL: prep
-% Prepares the data in a the form needed for plotting.
-<<ADNI1-SNT-xval-prep,include=FALSE,cache=TRUE>>=
+
+########################################
+# ADNI1-SNT-XVAL: prep
+# Prepares the data in a the form needed for plotting.
 snt_xval_data  <- read.csv(gzfile('data/cache/ADNI-SNT-XVAL:all_data.csv.gz'))
@@
                                  atlases      == 1)$k_diff)
+
+########################################
+# ADNI1-FEP-XVAL: prep
+# Prepares the data in a the form needed for plotting.
+fep.xval <-read.csv('data/results-fep-xval-ants-2013-12-29-191914.csv')
+fep.xval$atlases <- as.factor(fep.xval$atlases)
+fep.xval.mean <- ddply(fep.xval, .(subject,atlases,templates,label), function (df) {
+  data.frame(k=mean(df$k),volume=mean(df$volume))
+})
+
+fep.mb.kappa.best = round(mean(subset(fep.xval.mean, 
+                               templates == 19 & 
+                               atlases   == 9)$k), digits=3)
 @
@@
 \usepackage[hypcap]{caption}
-\usepackage{subcaption}
-\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}
+\usepackage{setspace}            % line spacing control
+\usepackage{subcaption}          % subfigures
+%\newcommand{\subfloat}[2][need a sub-caption]{\subcaptionbox{#1}{#2}}
 \usepackage{lineno}              % line numbers
 \usepackage{multirow,rotating}   % vertical text in tables
-\usepackage[displaymath, tightpage]{preview} %draft mode
-\setlength{\marginparwidth}{1.2in}
+%\usepackage[displaymath, tightpage]{preview} %draft mode
 \usepackage{fancyhdr}
 \pagestyle{fancy}
-\fancyhead[RO,RE]{draft - January, 2014}
+\fancyhead[RO,RE]{}
 
@@
 \renewcommand\Affilfont{\itshape\small}  %authblk
-\linespread{1.3} % one and a half line spacing
+\onehalfspacing
 
@@
 used in preparation of this article were obtained from the Alzheimer's Disease
-Neuroimaging Initiative (ADNI) database (adni.loni.ucla.edu). As such, the
+Neuroimaging Initiative (ADNI) database (adni.loni.usc.edu). As such, the
 investigators within the ADNI contributed to the design and implementation of
@@
 report. A complete listing of ADNI investigators can be found at:
-http://adni.loni.ucla.edu/wp-content/uploads/how\_to\_apply/ADNI\_Acknowledgement\_List.pdf}}
+http://adni.loni.usc.edu/wp-content/uploads/how\_to\_apply/ADNI\_Acknowledgement\_List.pdf}}
 \affil[1]{Kimel Family Translational Imaging-Genetics Lab, Centre for Addiction and
@@
 experiment is a 10-fold validation (per parameter setting) over 60 subjects
-taken from the ADNI Alzheimer's disease imaging database, and the second is a
-5-fold validation over 81 subjects having had a first episode of psychosis. In
-both cases, automated segmentations are compared with manual segmentions
+taken from the Alzheimer's Disease Neuroimaging Database (ADNI), and the second
+is a five-fold validation over 81 subjects having had a first episode of psychosis.
+In both cases, automated segmentations are compared with manual segmentions
 following the \prsnr{}-protocol. Using the best settings found from these
@@
 
-  \item \expfepxval{} consists of a similar cross-validation to explore \mb{}
-  segmentations on the brain images of young, first episode psychosis patients,
-  and to establish parameter settings that do not overfit on a single dataset.
+  \item \expfepxval{} is a similar cross-validation to explore \mb{}
+  segmentations on the brain images of young, first episode psychosis patients. 
+  In addition, \mb{} segmentations with two different atlas segmentation
+  protocols are compared to automated segmentations by the \fsl and
+  \freesurfer algorithms.  The results of this experiment combined with the
+  previous experiment establishes parameter settings that do not overfit to the
+  neuroanatomical features of a specific patient cohort.
 
   \item \expadni{} bridges \mb{} with the existing segmentation literature by
-  comparing \mb{} whole hippocampus segmentations with those of well-known
+  comparing \mb{} whole hippocampus segmentations with those of several well-known
   automated methods (\freesurfer, \fsl, \maper, \snt) on the entire
@@
 Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
-authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
+authors from the ADNI database (adni.loni.usc.edu) between March 2012 and
 August 2012. The image dataset used was the \adnidataset{} standardized dataset
@@
 % ADNI1-XVAL: show variability of MAGeT over parameters
-<<fig:ADNI1-xval-variability,dependson='ADNI1-Xval-prep',fig.show='hide'>>=
+<<fig:ADNI1-xval-variability,dependson='load',fig.show='hide'>>=
 ants = subset(jens_xval_data, 
@@
 In addition to the MCCV, we segment the entire first episode psychosis dataset
-using \mb{} and two other popular segmentation packages, \fsl{} and
-\freesurfer{}. Firstly, \mb{} is used with the five Winterburn atlas images and
-labels as atlases, and a randomly selected subset of 19 psychosis dataset images
-as templates. Secondly, \mb{} is run using the same template images, but instead
-as atlases we use five additional first episode psychosis subjects and
-corresponding manual segmentations (not included above). Finally, \fsl{} and
-\freesurfer{} are run with their default settings to segment the entire dataset.
-The \fsl{} \verb+run_first_all+ script was used according to the FIRST user guide
+using \mb{} using two different atlases, as well as two popular automated
+segmentation packages, \fsl{} and \freesurfer{}. Specifically, \mb{} is run once
+with the five Winterburn atlas images and labels as atlases and a randomly selected
+subset of 19 target images as templates. \mb{} is run a second time using
+the same template images, but we using five additional first
+episode psychosis subjects and corresponding manual segmentations (not included
+above) as atlases. \fsl{} and \freesurfer{} are run with the default settings:
+\fsl{} \verb+run_first_all+ script was used according to the FIRST user guide
 \footnote{http://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FIRST/UserGuide}, and
-\freesurfer{} was run using \verb+recon-all -all+.
+\freesurfer{} was run with the command \verb+recon-all -all+.
 
@@
 number of atlases or templates but quickly plateaus (Figure \ref{fig:FEP-xval}).
-A maximum similarity score of \Sexpr{snt.mb.kappa.best} is found when using 9
+A maximum similarity score of \Sexpr{fep.mb.kappa.best} is found when using 9
 atlases, 19 templates, ANTS registration, and majority vote label fusion.
@@
 and \fsl{} markedly underestimate smaller hippocampi and over-estimate large
-hippocampi (the limits of agreement are between $-2482^3mm$ and $-784^3mm$, and
-between $-1653^3mm$ and $79^3mm$, respectively), whereas both \mb{} methods show
+hippocampi (the limits of agreement are between $-2482mm^3$ and $-784mm^3$, and
+between $-1653mm^3$ and $79mm^3$, respectively), whereas both \mb{} methods show
 a much less exaggerated, but conservative bias (limits of agreement between
-$-67^3mm$ and $766^3mm$ when using FEP atlases, and between $-333^3mm$ and
-$504^3mm$ when using Winterburn atlases). On average, \freesurfer{} and \fsl{}
+$-67mm^3$ and $766mm^3$ when using FEP atlases, and between $-333mm^3$ and
+$504mm^3$ when using Winterburn atlases). On average, \freesurfer{} and \fsl{}
 overestimate hippocampal volume by about $1600 mm^3$ and $800 mm^3$,
-respectively. In contrast, on average \mb{} underestimates volumes by about $300
-mm^3$ when using FEP atlases, and by about $80mm^3$ when using Winterburn
-atlases.
+respectively. In contrast, on average \mb{} underestimates volumes by about
+$300mm^3$ when using FEP atlases and by about $80mm^3$ when using Winterburn
+atlases (compared to the \psegs).
 
@@
 % FEP-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
-<<fig:FEP-xval,fig.show='hide'>>=
-  fep.xval <-read.csv('data/results-fep-xval-ants-2013-12-29-191914.csv')
-  fep.xval$atlases <- as.factor(fep.xval$atlases)
-  fep.xval.mean <- ddply(fep.xval, .(subject,atlases,templates,label), function (df) {
-    data.frame(k=mean(df$k),volume=mean(df$volume))
-  })
-
-  fep.mb.kappa.best = round(mean(subset(fep.xval.mean, 
-                                 templates == 19 & 
-                                 atlases      == 9)$k), digits=3)
+<<fig:FEP-xval,fig.show='hide',dependson='load'>>=
   stderr <- function(x) sqrt(var(x)/length(x))
@@
 
-<<fig:FEP-Bland-Altman,dependson='fig:FEP-vols',fig.width=10,fig.height=3,fig.show='hide'>>=
+<<fig:FEP-Bland-Altman,dependson='fig:FEP-vols',fig.width=10,fig.height=6,fig.show='hide'>>=
 fep.volumes$diff = fep.volumes$manual - fep.volumes$volume
@@
 ggplot(fep.volumes, aes(x= mean, y = diff)) + 
-  facet_wrap( ~ Method, nrow=1, scales="fixed") + 
+  facet_wrap( ~ Method, nrow=2, scales="fixed") + 
   geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  geom_point(size=1) +  
+  scale_shape_discrete(solid=F,legend=F) + 
+  geom_point(size=2,mapping=aes(shape=1)) +  
   geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
@@
   \caption{\textbf{First Episode Patient dataset validation.}
-Manual segmentation of all 81 subjects is completed using the \prsnr-protocol.
-\mb{} is configured using \ants{} registration algorithm and majority vote label
-fusion. 
-(\ref{fig:FEP-xval}) shows mean DSC overlap similarity score of \mb{}
-  segmentations as they vary with atlas and template library size over 5-fold
-  validation. Error bars indicate standard error. 
-(\ref{fig:FEP-vols}) shows the volumes of the 81 subjects as determined by \fsl,
-  \freesurfer, and \mb{} configured in two ways: with an atlas library
-  of the five Winterburn atlases (MAGeT-Winterburn), or with an atlas
-  library of five manually segmented FEP subjects as atlases (MAGeT-FEP). Linear
+  All manual segmentation of the 81 subjects is done with the \prsnr-protocol.
+  \mb{} uses \ants{} registration and majority vote label fusion. 
+(\ref{fig:FEP-xval}) shows mean DSC score of \mb{} segmentations, as atlas and
+  template library size is varied over a 5-fold validation. Error bars indicate
+  standard error. 
+(\ref{fig:FEP-vols}) shows segmentation volumes from \fsl,
+  \freesurfer, \mb{} using the five Winterburn atlases (MAGeT-Winterburn), and
+  \mb{} using five manually segmented FEP subjects as atlases (MAGeT-FEP). Linear
   fit lines are shown, with the shaded region showing standard error.
-(\ref{fig:FEP-Bland-Altman}) shows the agreement between computed volumes and
-  manually segmented volume. The overall mean difference in volume, and limits of
+(\ref{fig:FEP-Bland-Altman}) shows the agreement between computed and
+  manually volumes. The overall mean difference in volume, and limits of
   agreement ($\pm 1.96SD$) are shown by dashed horizontal lines. Linear fit lines
@@
 \paragraph{\adnidataset{} dataset}  %%%%%%%%%%%
-The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
-\snt{}, \maper{}, and \freesurfer{} hippocampal volumes for a subset of images were
-provided by ADNI, along with quality control data for each \freesurfer{}
-segmentation (guidelines described in \citep{Hartig2010}).
-
-See Section \ref{adnidataset} for study details, inclusion criteria and imaging
+The {\em \adnidataset{}} standardized dataset contains $1919$ images of subjects
+diagnosed with Alzheimer's disease (AD), late-onset mild cognitive impairment
+(LMCI), or who are identified as cognitively normal (CN). See Section
+\ref{adnidataset} for study details, inclusion criteria and imaging
 characteristics.
@@
 
-\paragraph{Winterburn Atlases}  %%%%%%%%%%
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images, described
-more fully in \expfepxval, section \ref{exp:fep-xval}.
+\maper{} -- a \multiatlas segmentation tool , and \freesurfer{} hippocampal volumes for a subset of images were also
+provided by ADNI, along with quality control data for each \freesurfer{}
+segmentation (guidelines described in \citep{Hartig2010}).
 
@@
 \mb{} was configured with an atlas library composed of the five Winterburn atlas
-images and segmentations. A template library of 19 images were randomly selected
-from the target dataset of ADNI subjects, and \ants{} registration and majority
-vote label fusion were used as these were found to perform favourably in earlier
-experiments.
+images (\expfepxval, section \ref{exp:fep-xval}) and segmentations. A template
+library of 19 images were randomly selected from the target dataset of ADNI
+subjects, and \ants{} registration and majority vote label fusion were used as
+these were found to perform favourably in earlier experiments.
 
@@
 %%%%%%%%%
-<<fig:ADNI-volumes-plot,dependson='ADNI-volumes-prep',fig.show='hide'>>=
+<<fig:ADNI-volumes-plot,dependson='load',fig.show='hide'>>=
 melted=melt(means.complete, measure.vars=c("FS", "FSL", "MAPER","MAGeT"), 
@@
 %%%%%%%%%
-<<fig:ADNI-volumes-boxplot,dependson='ADNI-volumes-prep',fig.show='hide'>>=
+<<fig:ADNI-volumes-boxplot,dependson='load',fig.show='hide'>>=
 melted = melt(means.complete,measure.vars=c("FS", "FSL", "MAPER","MAGeT"),
@@
 %%%%%%%%%
-<<fig:ADNI-Bland-Altman,dependson='ADNI-volumes-prep',fig.show='hide',fig.height=4,fig.width=10>>=
+<<fig:ADNI-Bland-Altman,dependson='load',fig.show='hide',fig.height=4,fig.width=10>>=
 names(means.complete)[8] = "FreeSurfer"
@@
 are amongst the highest reported. Thus, more work is required to determine the
-extent to which label-fusion will improve the accuracy of our algorithm.
+extent to which label fusion will improve the accuracy of our algorithm.
 
@@
 despite the very limited number of atlases we had to work with, that \mb{}
-subfield segmentations are well within the bounds of error of a $0.3mm$ shift.
+subfield segmentations are well within the bounds of error of a $0.3mm$ voxel shift.
 
@@
 %%%%%%%%%%%%%
-<<fig:ADNI1-SNT-xval-variability,dependson='ADNI1-Xval-prep',fig.show='hide'>>=
+<<fig:ADNI1-SNT-xval-variability,dependson='load',fig.show='hide'>>=
 ants = subset(snt_xval_data, 

### 2014-02-10 paper.Rnw
MMC changes

@@
   In addition, \mb{} segmentations with two different atlas segmentation
-  protocols are compared to automated segmentations by the \fsl and
-  \freesurfer algorithms.  The results of this experiment combined with the
+  protocols are compared to automated segmentations by the \fsl{} and
+  \freesurfer{} algorithms.  The results of this experiment combined with the
   previous experiment establishes parameter settings that do not overfit to the
@@
   comparing \mb{} whole hippocampus segmentations with those of several well-known
-  automated methods (\freesurfer, \fsl, \maper, \snt) on the entire
+  automated methods (\freesurfer{}, \fsl{}, \maper{}, \snt{}) on the entire
   \adnidataset{} image dataset consisting of \Sexpr{length(means.complete$RID)}
@@
   geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
-  scale_shape_discrete(solid=F,legend=F) + 
-  geom_point(size=2,mapping=aes(shape=1)) +  
+  geom_point(size=2) +  
   geom_hline(aes(yintercept = y), linetype="dashed", colour="brown",
@@
   ylab(expression('manual - computed volume ' (mm^3))) + 
+  scale_shape_discrete(solid=F,guide='none') + 
   scale_y_continuous(breaks=seq(-10000,4000, 1000)) + 
   scale_x_continuous(breaks=seq(0,10000, 1000)) + 
-  theme(legend.direction = "horizontal", legend.position = "bottom", 
+  theme(legend.position = "none", 
         axis.text.y = element_text(angle = 90, hjust=0.5), 
@@
 
-Semi-automated segmentations of the left and right whole hippocampi are made
-available with a subset of ADNI images \citep{Hsu2002}. These labels have been
-generated using the \snt{} tool from Medtronic Surgical Navigation Technologies,
-Louisville, CO (see Supplementary Materials for detailed discussion of the
-segmentation process).
-
-\maper{} -- a \multiatlas segmentation tool , and \freesurfer{} hippocampal volumes for a subset of images were also
-provided by ADNI, along with quality control data for each \freesurfer{}
-segmentation (guidelines described in \citep{Hartig2010}).
+For a subset of the ADNI images, semi-automated segmentations of the left and
+right whole hippocampi generated using the \snt{} tool from Medtronic Surgical
+Navigation Technologies, Louisville, CO (see Supplementary Materials for
+detailed discussion of the segmentation process) are made available
+\citep{Hsu2002}. These labels are used as the reference labels in several other
+studies of (semi-)automated segmentation methods (see Discussion).  In addition,
+ADNI also distributes hippocampal segmentations and volumes determined using
+\maper{} \citep{Heckemann2011}, a \multiatlas{} segmentation tool, and the
+\freesurfer{} tool (including quality control data, with guidelines described in
+\citet{Hartig2010}).
 
@@
 One author (MP) performed visual quality inspection for \mb{} and \fsl{}
-segmentations using similar quality control guidelines (if either hippocampus
-was under or over segmented by $10mm$ or greater in three or more slices then
-the segmentation did not pass). Only images meeting the conditions of having
-segmentations from all methods (\snt{}, \maper{}, \freesurfer{}, \fsl{}, and
-\mb{}) and also passing similar quality control inspection were included in the
-analysis.
+segmentations using similar quality control guidelines to those used by
+\freesurfer. If either hippocampus was under or over segmented by $10mm$ or
+greater in three or more slices then the segmentation did not pass. Only images
+meeting the conditions of having segmentations from all methods (\snt{},
+\maper{}, \freesurfer{}, \fsl{}, and \mb{}) and also passing quality control
+inspection were included in the analysis.
 
@@
 volumes (Pearson $r > 0.85$), but appears to slightly over-estimate the volume
-of the AD hippocampus.
+of the AD hippocampus compared to the \snt{} segmentations.
 
@@
 %%%%%%%%%
-<<fig:ADNI-Bland-Altman,dependson='load',fig.show='hide',fig.height=4,fig.width=10>>=
+<<fig:ADNI-Bland-Altman,dependson='load',fig.show='hide',fig.height=6,fig.width=10>>=
 names(means.complete)[8] = "FreeSurfer"
@@
 ggplot(melted, aes(x= mean, y = diff, colour=Diagnosis)) + 
-  facet_wrap( ~ method, nrow=1, scales="fixed") + 
+  facet_wrap( ~ method, nrow=2, scales="fixed") + 
   geom_smooth(method="lm", formula=y~x, fullrange=T, se=F) + 
@@
   impairment, and CN = cognitively normal.
-(\ref{fig:ADNI-Bland-Altman}) Bland-Altman plots comparing subject mean
-  hippocampal volume as measured in the \adnidataset{} dataset by \snt{}
-  segmentation and each automated methods. The overall mean difference in volume,
+(\ref{fig:ADNI-Bland-Altman}) Bland-Altman plots shows the agreement between
+  computed and \snt{} hippocampus volume. The overall mean difference in volume,
   and limits of agreement ($\pm 1.96SD$) are shown by dashed horizontal lines.
@@
 
-\paragraph{Winterburn Atlases}  %%%%%%%%%%
-The Winterburn atlases \citep{Winterburn2013} are digital hippocampal
-subfield segmentations of five in-vivo $0.3mm$-isotropic T1-weighted MR images,
-described more fully in \expfepxval, section \ref{exp:fep-xval}.
-
 \paragraph{Healthy Control Dataset}  %%%%%%%%%%
@@
 
-In this experiment, the Winterburn atlases are resampled to $0.9mm$-isotropic
+In this experiment, the Winterburn atlases (\expfepxval, section
+\ref{exp:fep-xval}) are resampled to $0.9mm$-isotropic
 voxel resolution to simulate standard 3T T1-weighted resolution images. Image
@@
 
+Additionally, by shifting the original manual $0.3mm^3$-isotropic voxel
+segmentations by one voxel in the x, y, and z direction and then resampling it
+to $0.9mm^3$-isotropic voxels we obtain a simulated manual segmentation having a
+small amount of error. We can compare the DSC overlap score of the shifted
+labels (relative to the directly resampled labels) with the DSC score of the
+\mb{} generated labels in order to evaluate their relevance.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 deviation DSC scores of the subfields are shown in Table
-\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
-segmentations when perturbed slightly and compared to the originals. We find
-that the CA4/DG subfield shows the highest mean DSC score of
+\ref{tab:WAxval-perturb-kappa}. We find that the CA4/DG subfield shows the highest mean DSC score of
 \Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
@@
 The whole hippocampus is segmented with a mean DSC score
-of\Sexpr{mean(Whole)}$\pm$\Sexpr{sd(Whole)}.
+of\Sexpr{mean(Whole)}$\pm$\Sexpr{sd(Whole)}. Table
+\ref{tab:WAxval-perturb-kappa} also shows mean DSC overlap scores for simulated
+manual segmentations with error introduced by a one-voxel translation in all
+directions. For all subfields, \mb{} DSC scores are well within the error bounds
+set by these segmentations.
 
@@
   ylab(expression('Difference of manual and MAGeT-Brain subfield volume ' (mm^3))) + 
-  #scale_x_continuous(limits=c(0,1000)) + 
-  #scale_color_discrete(name="") + 
-  scale_shape_discrete(name='Subject') + 
   theme(legend.direction = "horizontal", legend.position = "bottom", 
@@
 caption="Overlap similarity results for the each of the subfields of the hippocampus. 
-Simulated overlap similarity results are also given for resampled labels that were 
-translated by one voxel (i.e.: $0.9mm$) in all directions. Values are given as
-mean Dice's Similarity Coefficient (DSC) $\\pm$ standard deviation"
+Simulated overlap similarity results are also given for manual labels that were 
+translated by one voxel (i.e.: $0.3mm$) in all directions and then resampled.
+Values are given as mean Dice's Similarity Coefficient (DSC) $\\pm$ standard
+deviation."
 latex(wa.perturb.table, file="", size="scriptsize",landscape=FALSE, ctable=F,
@@
 to provide accurate and reliable segmentations. Both \expadnixval{} (Section
-\ref{exp:adni-xval}) and \expfepxval (Section \ref{exp:fepi-xval}) compare \mb{}
+\ref{exp:adni-xval}) and \expfepxval{} (Section \ref{exp:fep-xval}) compare \mb{}
 to basic-\multiatlas{} segmentation by characterising the change in segmentation
@@
 \Sexpr{length(means.complete$RID)} images from the \adnidataset{} dataset, \mb{}
-performs as well as, or better, than other established and popular methods.
-Finally, \expsubfield{} (Section \ref{exp:winterburn-xval}) demonstrates the
-reliability of \mb{} in producing subfield segmentations which match the
-segmentation protocol of the input atlases despite contrast and resolution
-limitations in standard T1-weighted image volumes.  All of these experiments
-together demonstrate that \mb's algorithmic performance is not dependent on a
-single definition of the hippocampus but is effective with differing hippocampal
+performs as well as, or better, than other established and popular methods, and
+has a much more conservative proportional bias in segmentation volume.  Finally,
+\expsubfield{} (Section \ref{exp:winterburn-xval}) demonstrates the reliability
+of \mb{} in producing subfield segmentations which match the segmentation
+protocol of the input atlases despite contrast and resolution limitations in
+standard T1-weighted image volumes.  All of these experiments together
+demonstrate that \mb's algorithmic performance is not dependent on a single
+definition of the hippocampus but is effective with differing hippocampal
 definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}, across image types,
@@
 a template library from a small set of labelled atlas images -- is validated in
-the cross validation conducted in \expadnixval{} (and the replication in in
+the cross validation conducted in \expadnixval{} (and the replication in
 \expfepxval{} and \expadnixvalsnt{}, Supplementary Materials). We find that both
@@
 
-Another concern is the moderate improvement observed in \mb{} in comparison to
-\multiatlas{} segmentation when using the same number of atlases. The actual
-benefit in using \mb{} is consistency of the labelling regardless of atlas or
-template choice, as mentioned above. This is an important consideration that few
-have touched on previously. The 10-fold Monte Carlo cross-validation that we
-present in \expadnixval{} and \expfepxval{} is amongst one of the most stringent
-performed in the \multiatlas{} segmentation literature. To the best of our
-knowledge, with the exception of \citep{Wang2011}, other groups using ADNI data
-for validation do at most a single round of leave-one-out-validation (Table
-\ref{tab:other-methods}). The thoroughness of our validation suggests that our
-results are reflective of a true average over the choice of parameter settings
-and are independent of atlas or template choice (provided the input atlases are
-consistently segmented).
+Despite \mb{} achieving segmentation results which are competitive with the rest
+of the field (Table \ref{tab:other-methods}), a concern may be raised over the
+modest improvement in segmentation agreement observed using \mb{} over
+\multiatlas{}, with the same number of atlases (\expadnixval{}). As we have
+shown in that same experiment, the benefit in using \mb{} is both an increase in
+the overlap agreement and also in the improved consistency of the labelling
+regardless of atlas or template choice. Reducing the variability in segmentation
+accuracy is an important consideration that few have touched on previously. In
+addition, the Monte Carlo cross-validations that we present in \expadnixval{}
+and \expfepxval{} are amongst the most stringent performed in the \multiatlas{}
+segmentation literature. To the best of our knowledge, with the exception of
+\citep{Wang2011}, other groups do at most a single round of
+leave-one-out-validation (Table \ref{tab:other-methods}).  Thus, the
+thoroughness of our validation suggests that our results are reflective of a
+true average over the choice of parameter settings and are independent of atlas
+or template choice.
 
@@
 Indeed, our replication of the 10-fold cross-validation using \snt{}
-segmentations (\expadnixvalsnt{}, Supplementary Materials) shows noticeably poorer
-mean similarity scores for both \mb{} and \multiatlas.
+segmentations (\expadnixvalsnt{}, Supplementary Materials) shows noticeably
+poorer mean similarity scores for both \mb{} and \multiatlas.
 
@@
 by its nature, the Dice's Similarity Coefficient score penalizes structures with
-high surface area to volume ratios. Therefore subfield DSC scores will generally
+high surface area-to-volume ratios. Therefore subfield DSC scores will generally
 be lower than whole hippocampal segmentations. We attempted to put this
@@
 despite the very limited number of atlases we had to work with, that \mb{}
-subfield segmentations are well within the bounds of error of a $0.3mm$ voxel shift.
+subfield segmentations are well within the bounds of error of a $0.3mm^3$ voxel
+shift.
 
 Our overlap DSC values demonstrates that we can reliably reproduce segmentations
-for the CA1, subiculum, and CA4/dentate subfields (DSC > 0.5). That the CA2/CA3
-and molecular layers are less well reproduced (DSC < 0.5) should not be
+for the CA1, subiculum, and CA4/dentate subfields ($DSC > 0.5$). That the CA2/CA3
+and molecular layers are less well reproduced ($DSC < 0.5$) should not be
 surprising as these are extremely thin and spatially convoluted regions that
@@
 consistency in similarity scores across all subjects and validation folds
-(Figure \ref{ADNI1-SNT-xval-variability}) with a large enough template library.
+(Figure \ref{fig:ADNI1-SNT-xval-variability}) with a large enough template library.
 

### 2014-02-12 paper.Rnw
Several fixes

@@
 
+# http://druedin.com/2013/02/16/p-level-in-sweave/
+plevel <- function (x, strict=FALSE) {
+  # levels of p-values, for Sweave
+  # strict cuts at 0.05, otherwise cuts at 0.1
+  if (x>0.1 & strict==FALSE) p <- "p>0.1"  # not significant
+  if (x>0.1 & strict==TRUE)  p <- "p>0.05" # not significant
+  if (x<=0.1 & strict==FALSE) p <- "p<0.1"  # significant
+  if (x<=0.1 & strict==TRUE)  p <- "p>0.05" # not significant
+  if (x<=0.01)  p <- "p<0.01"  # significant
+  if (x<=0.05)  p <- "p<0.05"  # significant
+  if (x<=0.001) p <- "p<0.001" # significant
+  return(p)
+  }
 @
@@
 
-\textbf{Results:} \textbf{[TODO: update with FEP-XVAL and subfield XVAL results.]} 
-Using 9 atlases and 19 template images, \mb{} achieves a
-mean Dice's Similarity Coefficient (DSC) of \Sexpr{mb.kappa.best} 
-(to \psegs{}) over 10-folds of Monte Carlo 
-significantly lower variability in DSC than \multiatlas{} segmentation.
+\textbf{Results:} 
+ADNI-XVAL
+In the ADNI cross-validation, using 9 atlases \mb{} achieves a mean Dice's
+Similarity Coefficient (DSC) score of \Sexpr{mb.kappa.best} with respect to
+manual whole hippocampus segmentations, and also exhibits significantly lower
+variability in DSC scores than \multiatlas{} segmentation.  In the younger,
+psychosis dataset, \mb{} achieves a mean DSC score of \Sexpr{fep.mb.kappa.best}
+and produces volumes which agree with manual segmentation volumes better than
+those produced by the \freesurfer and \fsl methods (mean difference in volume:
+$80mm^3$, $1600mm^3$, and $800mm^3$, respectively). Across the \adnidataset
+dataset,
+
+ADNI-1YR
+Subfields
+
 In a LOOCV, \mb{} accurately reproduces the subfields... .
@@
 segmentations available from ADNI, as well as a conservative, rather than
-exaggerated, proportional volume bias.\\ \textbf{Conclusion:} We demonstrate
-that \mb{} produces accurate hippocampal segmentations using only 5 atlases over
-different hippocampal definitions, disease populations, and acquisition types,
-as well as showing that accurate identification of the hippocampal subfields is
-possible.
+exaggerated, proportional volume bias.
+
+\textbf{Conclusion:} We demonstrate that \mb{} produces accurate hippocampal
+segmentations using only 5 atlases over different hippocampal definitions,
+disease populations, and acquisition types, as well as showing that accurate
+identification of the hippocampal subfields is possible.
 \end{abstract}
@@
 
+<<ADNI1-xval-scores,include=FALSE>>=
+k.1t = round(mean(subset(jens_xval_mean, 
+                           reg_method   == "ANTS" & 
+                           method.mb    == "Majority Vote" &
+                           templates.mb == 1 & 
+                           atlases      == 9)$k.mb), digits=3)
+k.5t = round(mean(subset(jens_xval_mean, 
+                           reg_method   == "ANTS" & 
+                           method.mb    == "Majority Vote" &
+                           templates.mb == 5 & 
+                           atlases      == 9)$k.mb), digits=3)
+k.9t = round(mean(subset(jens_xval_mean, 
+                           reg_method   == "ANTS" & 
+                           method.mb    == "Majority Vote" &
+                           templates.mb == 9 & 
+                           atlases      == 9)$k.mb), digits=3)
+k.17t = round(mean(subset(jens_xval_mean, 
+                           reg_method   == "ANTS" & 
+                           method.mb    == "Majority Vote" &
+                           templates.mb == 17 & 
+                           atlases      == 9)$k.mb), digits=3)
+k.9t.animal = round(mean(subset(jens_xval_mean, 
+                                 reg_method   == "ANIMAL" & 
+                                 method.mb    == "Majority Vote" &
+                                 templates.mb == 9 & 
+                                 atlases      == 9)$k.mb), digits=3)
+
+# Does weighted voting significantly improve scores? 
+# Let's correlate DSC scores at (a, t) = (1, 5) and (1, 15), (9, 5) and (9, 15) 
+ants   = subset(jens_xval_mean, reg_method=='ANTS')
+animal = subset(jens_xval_mean, reg_method=='ANIMAL')
+mavote = subset(ants, method.mb == 'Majority Vote')
+mivote = subset(ants, method.mb == 'NMI Vote')
+ants.stats = ddply(ants, c("atlases", "templates.mb"), 
+                   function (df) {
+                     test = cor.test(subset(df, method.mb == 'Majority Vote')$k.mb, 
+                                     subset(df, method.mb == 'NMI Vote')$k.mb)
+                     data.frame(
+                       estimate = test$estimate, 
+                       p.value  = test$p.value
+                       )
+                   })
+@
+
 We find that for \mb{} segmentations, similarity score increases as atlas and
 template library size is increased, although with diminishing returns and an
-eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}). A
-maximum similarity score of \Sexpr{mb.kappa.best} is found when using 9 atlases,
-19 templates, ANTS registration, and majority vote label fusion. The \ants{}
-registration method consistently out-performs \animal{} registration over all
-variable settings we tested. Additionally, by itself, using a weighted voting
-strategy did not significantly improve segmentation agreement, contrary to the
-findings of \citet{Aljabar2009} using basic \multiatlas{} segmentation. Given
-these findings, in the remainder of our experiments only results using the \ants{}
-registration algorithm and majority vote fusion will be shown.
-
-With at least five templates, \mb{} shows a mean improvement in similarity score
-over \multiatlas{} segmentation when using the same size of atlas library and
-majority vote label fusion (Figure \ref{fig:ADNI1-xval-k-diff}). The magnitude
-of improvement increases with the size of template library, and shows
-diminishing returns with larger atlas libraries. Peak improvement
-(+\Sexpr{round(mb.diff.best, digits=3)} DSC) is found with a single atlas
-and template library of 19 images.
+eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}). For
+instance, with 9 atlases and using \ants for registration and majority vote
+fusion, the mean DSC scores for 1, 5, 9 and 17 templates are \Sexpr{k.1t},
+\Sexpr{k.5t}, \Sexpr{k.9t}, \Sexpr{k.17t}, respectively.  A maximum
+similarity score of \Sexpr{mb.kappa.best} is found when using 9 atlases, 19
+templates, ANTS registration, and majority vote label fusion.
+
+The \ants{} registration method consistently outperforms \animal{} registration
+over all variable settings we tested (mean increase in DSC is
+\Sexpr{mean(ants$k.mb-animal$k.mb)}).  Pearson correlations of \mb{} DSC scores
+when using weighted voting and when using non-weighted majority vote label
+fusion (with \ants{} registration) for all combinations of atlases and templates
+are $r > \Sexpr{min(ants.stats$estimate)}$,
+\Sexpr{plevel(max(ants.stats$p.value))}, with a mean difference in DSC score of
+\Sexpr{mean(mivote$k.mb - mavote$k.mb)}.  This result suggests that using a
+weighted voting strategy does not significantly improve \mb{} segmentation
+agreement, contrary to the findings of \citet{Aljabar2009} for basic
+\multiatlas{} segmentation. Thus, in the remainder of our experiments only
+results using the \ants{} registration algorithm and majority vote fusion will
+be shown.
+
+With at least five templates, \mb{} consistently shows a higher DSC score than
+\multiatlas{} segmentation wth the same number of atlases: 
+  $r = \Sexpr{round(cor.test(ants$k.mb, ants$k.ma)$estimate, 2)}$, 
+       $\Sexpr{plevel(cor.test(ants$k.mb, ants$k.ma)$p.value)}$, 
+mean DSC increase = \Sexpr{mean(ants$k.mb - ants$k.ma)} (Figure
+\ref{fig:ADNI1-xval-k-diff}). The magnitude of DSC increase grows with template
+library size but shows diminishing returns with larger atlas libraries.  Peak
+increase (+\Sexpr{mb.diff.best} DSC) is found with a single atlas and template
+library of 19 images.
 
@@
 To validate \mb{} segmentation quality with respect to other established
-automated hippocampal segmentation methods, \mb{} was applied to a large dataset
-from the ADNI project and the resulting segmentations were compared to those
+automated hippocampal segmentation methods, we apply \mb{} to a large dataset
+from the ADNI project. The resulting segmentations are compared to those
 produced by \freesurfer{}, \fsl{}, \maper{}, as well as semi-automated whole
@@
 \paragraph{\adnidataset{} dataset}  %%%%%%%%%%%
-The {\em \adnidataset{}} standardized dataset contains $1919$ images of subjects
-diagnosed with Alzheimer's disease (AD), late-onset mild cognitive impairment
-(LMCI), or who are identified as cognitively normal (CN). See Section
+The {\em \adnidataset{}} standardized dataset contains $1919$ images in total.
+\snt{}, \maper{}, and \freesurfer{} hippocampal volumes for a subset of images were
+provided by ADNI, along with quality control data for each \freesurfer{}
+segmentation (guidelines described in \citep{Hartig2010}).  See Section
 \ref{adnidataset} for study details, inclusion criteria and imaging
@@
 In this experiment, the Winterburn atlases (\expfepxval, section
-\ref{exp:fep-xval}) are resampled to $0.9mm$-isotropic
-voxel resolution to simulate standard 3T T1-weighted resolution images. Image
-subsampling is performed using trilinear subsampling techniques. In each round
-of LOOCV, a single atlas image is selected and treated as a target image to be
-segmented by \mb{}. So as to have an odd-sized atlas library, atlas image is
-segmented once using each possible triple of atlas images, and corresponding
-manual segmentations, from the remaining four unselected atlases. Thus, for
-each of the five atlases, a total of ${3 \choose 4}=4$ segmentations are
-evaluated, resulting in a combined total of $5\times4=20$ segmentations
-evaluated overall. We chose an atlas library with an odd number of images so as
-to ensure unbiased label fusion when using majority voting (see Discussion).
+\ref{exp:fep-xval}) are resampled to $0.9mm$-isotropic voxel resolution to
+simulate standard 3T T1-weighted resolution images. Image subsampling is
+performed using trilinear subsampling techniques. In each round of LOOCV, a
+single atlas image is selected and treated as a target image to be segmented by
+\mb{}. So as to have an odd-sized atlas library, atlas image is segmented once
+using each possible triple of atlas images, and corresponding manual
+segmentations, from the remaining four unselected atlases. Thus, for each of the
+five atlases, a total of ${3 \choose 4}=4$ segmentations are evaluated,
+resulting in a combined total of $5\times4=20$ segmentations evaluated overall.
+We chose an atlas library with an odd number of images so as to ensure unbiased
+label fusion when using majority voting (see Discussion).
 
@@
 @
-
-Figure \ref{fig:WAxval-kappa} shows the overlap similarity scores between the
-\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
-subfield across all subjects and folds of the validation. Mean and standard
-deviation DSC scores of the subfields are shown in Table
-\ref{tab:WAxval-perturb-kappa}. We find that the CA4/DG subfield shows the highest mean DSC score of
-\Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
-subfields having scores of \Sexpr{mean(CA1)}$\pm$\Sexpr{sd(CA1)} and
-\Sexpr{mean(Sub)}$\pm$\Sexpr{sd(Sub)}, respectively. Both the CA4/DG and
-molecular regions score below 0.5. These scores may seem low but not when taken
-in context and compared to existing (semi-)automated methods (see Discussion).
-The whole hippocampus is segmented with a mean DSC score
-of\Sexpr{mean(Whole)}$\pm$\Sexpr{sd(Whole)}. Table
-\ref{tab:WAxval-perturb-kappa} also shows mean DSC overlap scores for simulated
-manual segmentations with error introduced by a one-voxel translation in all
-directions. For all subfields, \mb{} DSC scores are well within the error bounds
-set by these segmentations.
 
@@
 @
+
+Figure \ref{fig:WAxval-kappa} shows the overlap similarity scores between the
+\mb{} segmentations and the resampled Winterburn atlases for each hippocampal
+subfield across all subjects and folds of the validation. Mean and standard
+deviation DSC scores of the subfields are shown in Table
+\ref{tab:WAxval-perturb-kappa}, along with DSC scores for the resampled atlas
+segmentations when perturbed slightly and compared to the originals. We find
+that the CA4/DG subfield shows the highest mean DSC score of
+\Sexpr{mean(CA4DG)}$\pm$\Sexpr{sd(CA4DG)}, followed by the Subiculum and CA1
+subfields having scores of \Sexpr{mean(CA1)}$\pm$\Sexpr{sd(CA1)} and
+\Sexpr{mean(Sub)}$\pm$\Sexpr{sd(Sub)}, respectively. Both the CA4/DG and
+molecular regions score below 0.5. These scores may seem low but not when taken
+in context and compared to existing (semi-)automated methods (see Discussion).
+The whole hippocampus is segmented with a mean DSC score
+of\Sexpr{mean(Whole)}$\pm$\Sexpr{sd(Whole)}.
 

### 2014-02-12 paper.Rnw
Draft complete, sending to Mallar for final comments.

@@
 construct. \\
+
 \textbf{Method:} \mb{} works by propagating atlas segmentations to a template
@@
 conduct a leave-one-out cross-validation (LOOCV) of hippocampal subfield
-segmentation using five high-resolution manually segmented atlases
-\citep{Winterburn2013}.
-
-\textbf{Results:} 
-ADNI-XVAL
-In the ADNI cross-validation, using 9 atlases \mb{} achieves a mean Dice's
-Similarity Coefficient (DSC) score of \Sexpr{mb.kappa.best} with respect to
-manual whole hippocampus segmentations, and also exhibits significantly lower
-variability in DSC scores than \multiatlas{} segmentation.  In the younger,
-psychosis dataset, \mb{} achieves a mean DSC score of \Sexpr{fep.mb.kappa.best}
-and produces volumes which agree with manual segmentation volumes better than
-those produced by the \freesurfer and \fsl methods (mean difference in volume:
-$80mm^3$, $1600mm^3$, and $800mm^3$, respectively). Across the \adnidataset
-dataset,
-
-ADNI-1YR
-Subfields
-
-In a LOOCV, \mb{} accurately reproduces the subfields... .
-
-\mb{} produces hippocampal
-volumes in a first episode psychosis patient population that are highly
-correlated with expert manual segmentation volumes (Pearson $r = 0.877, t =
-16.244, p < 0.001$). Compared to \fsl{} and \freesurfer{}, \mb{} shows much smaller
-fixed volume bias (within $250 mm^3$ on average) to semi-automated (\snt{})
-segmentations available from ADNI, as well as a conservative, rather than
-exaggerated, proportional volume bias.
-
-\textbf{Conclusion:} We demonstrate that \mb{} produces accurate hippocampal
-segmentations using only 5 atlases over different hippocampal definitions,
-disease populations, and acquisition types, as well as showing that accurate
-identification of the hippocampal subfields is possible.
+segmentation in standard 3T T1 images, using five high-resolution manually
+segmented atlases \citep{Winterburn2013}.
+
+\textbf{Results:} In the ADNI cross-validation, using 9 atlases \mb{} achieves a
+mean Dice's Similarity Coefficient (DSC) score of \Sexpr{mb.kappa.best} with
+respect to manual whole hippocampus segmentations, and also exhibits
+significantly lower variability in DSC scores than \multiatlas{} segmentation.
+In the younger, psychosis dataset, \mb{} achieves a mean DSC score of
+\Sexpr{fep.mb.kappa.best} and produces volumes which agree with manual
+segmentation volumes better than those produced by the \freesurfer{} and \fsl{}
+methods (mean difference in volume: $80mm^3$, $1600mm^3$, and $800mm^3$,
+respectively). Similarily, in the \adnidataset{} dataset, \mb{} produces
+hippocampal segmentations well correlated ($r > 0.85$) with \snt{} semi-automated
+reference volumes within disease categories, and shows a conservative bias and a
+mean difference in volume of $250 mm^3$ across the entire dataset, compared with
+\freesurfer{} and \fsl{} which both overesimate volume differences by $2600
+mm^3$ and $2800 mm^3$ on average, respectively. Finally, \mb{} segments the CA1,
+CA4/DG and subiculum subfields with an DSC overlap scores of 0.56, 0.65, 0.58,
+respectively.
+
+\textbf{Conclusion:} We demonstrate that \mb{} produces accurate whole
+hippocampal segmentations using at most only 9 atlases with various hippocampal
+definitions, disease populations, and image acquisition types, as well we show
+that accurate identification of the hippocampal subfields is possible.
 \end{abstract}
@@
 eventual trend towards a plateau (Figure \ref{fig:ADNI1-xval-k-mean}). For
-instance, with 9 atlases and using \ants for registration and majority vote
+instance, with 9 atlases and using \ants{} for registration and majority vote
 fusion, the mean DSC scores for 1, 5, 9 and 17 templates are \Sexpr{k.1t},
@@
 methods and the \snt{} semi-automated label volumes (Figure
-\ref{fig:ADNI-volumes-plot}). Volumes are correlated with Pearson $r > 0.78$
-for all methods across disease categories. Within disease categories (Figure
-\ref{fig:ADNI-volumes-boxplot}), \mb{} is consistently well correlated to manual
-volumes (Pearson $r > 0.85$), but appears to slightly over-estimate the volume
-of the AD hippocampus compared to the \snt{} segmentations.
+\ref{fig:ADNI-volumes-plot}). Volumes are well correlated ($r > 0.78$)c for all
+methods, and across disease categories. Within disease categories (Figure
+\ref{fig:ADNI-volumes-boxplot}), \mb{} is consistently well correlated to \snt{}
+volumes ($r > 0.85$), but appears to slightly over-estimate the volume of the AD
+hippocampus compared to the \snt{} segmentations.
 
@@
 Computations were performed on the gpc supercomputer at the SciNet HPC
-Consortium \cite{Loken2010}. SciNet is funded by: the Canada Foundation for
+Consortium \citep{Loken2010}. SciNet is funded by: the Canada Foundation for
 Innovation under the auspices of Compute Canada; the Government of Ontario;

### 2014-02-13 paper.Rnw
Mallar's comments

@@
 conduct a leave-one-out cross-validation (LOOCV) of hippocampal subfield
-segmentation in standard 3T T1 images, using five high-resolution manually
-segmented atlases \citep{Winterburn2013}.
+segmentation in standard 3T T1-weighted images, using five high-resolution
+manually segmented atlases \citep{Winterburn2013}.
 
@@
 mm^3$ and $2800 mm^3$ on average, respectively. Finally, \mb{} segments the CA1,
-CA4/DG and subiculum subfields with an DSC overlap scores of 0.56, 0.65, 0.58,
-respectively.
+CA4/DG and subiculum subfields on standard 3T T1-weighted resolution images with
+DSC overlap scores of 0.56, 0.65, and 0.58, respectively, relative to manual
+segmentations.  %TODO: use Sexpr
 
 \textbf{Conclusion:} We demonstrate that \mb{} produces accurate whole
-hippocampal segmentations using at most only 9 atlases with various hippocampal
-definitions, disease populations, and image acquisition types, as well we show
-that accurate identification of the hippocampal subfields is possible.
+hippocampal segmentations using only 9 atlases, or fewer, with various
+hippocampal definitions, disease populations, and image acquisition types.
+Additionally, we show that \mb{} identifies hippocampal subfields in standard 3T
+T1-weighted images with overlap scores comparable to competing methods.
 \end{abstract}
@@
 five of the Winterburn atlases (which, based on our findings with whole
-hippocampal segmentation, would lead to improved overlap similarity).
+hippocampal segmentation, would have resulted in improved overlap similarity).
 Although having more atlases would be ideal in this context, these atlases are
-incredibly time consuming to generate \citep{Winterburn2013}. Nevertheless, the
+very time consuming to generate \citep{Winterburn2013}. Nevertheless, the
 advantage of evaluating \mb{} on standard 3T T1-weighted resolution MR images
@@
 which regions can be differentiated, even on high-resolution scans. See Table
-\ref{tab:subfield-segmentation-protocols} for a comparison of manual subfield
-segmentation methodologies. A further complication is that different researchers
-have differing operational definitions for the subfields and how they ought to
-be parcellated. The disagreement in the community has led to an international
-working group devoted to normalizing the ontology and segmentation rules for the
-hippocampal subfields (\url{http://www.hippocampalsubfields.com/}). Both of
-these disagreements suggest that direct comparison of automated methods using
-``ground truth''-based overlap similarity metrics, such as Dice's Similarity
-Coefficient, are not possible without carefully taking into account the
-differences in underlying segmentation protocols and image characteristics.
+\ref{tab:subfield-segmentation-protocols} for a comparison of MRI-based manual
+subfield segmentation methodologies. A further complication is that different
+researchers have differing operational definitions for the subfields and how
+they ought to be parcellated. The disagreement in the community has led to an
+international working group devoted to normalizing the ontology and segmentation
+rules for the hippocampal subfields
+(\url{http://www.hippocampalsubfields.com/}).  In addition, there have been
+recent advances from the Yushkevich group to revise their MRI subfield
+segmentation protocol based on anatomy discerned from serial histological
+acquisitions \citep{Adler2014}.  The definitional and operational disagreements
+suggest that direct comparison across automated methods using ``ground
+truth''-based overlap similarity metrics, such as Dice's Similarity Coefficient,
+are not possible without carefully taking into account the differences in
+underlying segmentation protocols and image characteristics.
 

### 2014-02-18 paper.Rnw
Mallar's comments and fixes based on the reviews

@@
 validated. \citet{Yushkevich2009} manually segment hippocampal subfields on
-high-resolution (either $0.2mm^3$ isotropic or $0.2mm \times 0.3mm \times 0.2mm$
+high-resolution (either $0.2mm$-isotropic or $0.2mm \times 0.3mm \times 0.2mm$
 resolution voxels) T2-weighted MR images acquired from five post-mortem medial
@@
 segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
-segmentation of standard 3T T1-weighted images at $0.9m^3$-isotropic voxels. We
+segmentation of standard 3T T1-weighted images at $0.9mm$-isotropic voxels. We
 use a modified leave-one-out cross-validation (LOOCV) design.
@@
 $TE/TR/TI = 3.0ms/6.7ms/650ms$, flip angle=$8^\circ$ , $FOV = 15.3cm$, slice
-thickness$=0.9mm$, 170 in-plane steps for an approximate 0.9mm isotropic voxel
+thickness$=0.9mm$, 170 in-plane steps for an approximate $0.9mm$-isotropic voxel
 resolution.
@@
 Evaluating the agreement of automated hippocampal subfield segmentations with
-manual segmentations for T1 images at $0.9mm^3$-isotropic voxels is inherently
+manual segmentations for T1 images at $0.9mm$-isotropic voxels is inherently
 ill-defined since there are no manual protocols for segmentation at this
@@
 
-Additionally, by shifting the original manual $0.3mm^3$-isotropic voxel
+Additionally, by shifting the original manual $0.3mm$-isotropic voxel
 segmentations by one voxel in the x, y, and z direction and then resampling it
-to $0.9mm^3$-isotropic voxels we obtain a simulated manual segmentation having a
+to $0.9mm$-isotropic voxels we obtain a simulated manual segmentation having a
 small amount of error. We can compare the DSC overlap score of the shifted
@@
 regardless of atlas or template choice. Reducing the variability in segmentation
-accuracy is an important consideration that few have touched on previously. In
+agreement is an important consideration that few have touched on previously. In
 addition, the Monte Carlo cross-validations that we present in \expadnixval{}
@@
 experimentation protocol. These methods suggest that some post-processing of the
-final segmentations would improve accuracy of the segmentation. While that may
+final segmentations would improve agreement of the segmentation. While that may
 be true, there is little consensus regarding how to achieve this.
@@
 through the template library. However it should be noted that the increased
-accuracy that follows increasing the number of atlases and templates comes at an
+agreement that follows increasing the number of atlases and templates comes at an
 increased computational cost ($O(log(n))$), as previously mentioned in
@@
 data \citep{Yushkevich2009,Yushkevich2010,VanLeemput2009}. Table
-\ref{tab:subfield-segmentation-kappa} compares segmentation accuracy from some
+\ref{tab:subfield-segmentation-kappa} compares segmentation agreement from some
 of these methods and \mb{}. The overlap DSC scores for \mb{} subfields are
@@
 usage scenarios of researchers and clinicians.
+
+Experiments 1, 2, and 5 have demonstrated that our algorithm flexibly
+accommodates different whole hippocampus manual segmentation methodologies. We
+have not explicitly evaluated a subfield definition other than the Winterburn
+protocol, and therefore it is possible that using an alternate subfield
+definition could improve the reliability of our automated subfield definitions.
+For example, established definitions such as those from \citet{Mueller2007}
+could be a prime candidate for further exploration. In addition, the
+conservative nature of the Mueller definition (labelling of the 5 slices in the
+hippocampus body only) would likely further aid in reliability measurement.
+However, there are two main logistical problems that we would have to overcome
+prior to implementation. The first is that these definitions were developed for
+data that is highly anisotropic ($0.4mm \times 0.5mm \times 2mm$), and it is
+unclear how our algorithms would deal with such atlases used as input. The
+second is that, since these atlases are not publicly available, we would be have
+to re-implement the protocol using our atlases. At the present time it is
+unclear how we would adapt these protocol to data that we used, where subfield
+segmentations are defined on $0.3mm^3$ voxels. However, the impact of subfield
+definitions in the context of our work is an important one and should be
+considered in subsequent studies.
 
@@
 different disease populations, and different acquisition types. Finally, we also
-demonstrate that accurate identification of the hippocampal subfields in
+demonstrate that reliable reproduction of hippocampal subfield segmentations in
 standard 3T T1-weighted images is possible.

### 2014-04-09 paper.Rnw
Addressed reviewer comment about the use of 'accuracy'

@@
 %Multiple Automatically Generated Templates}
-\title{Bootstrapping Multi-atlas Segmentation Using Multiple Automatically
-Generated Templates for the Segmentation of the Whole Hippocampus and
-Subfields}
+\title{Multi-atlas Segmentation of the Whole Hippocampus and Subfields
+Using Multiple Automatically Generated Templates}
 \author[1]{Jon Pipitone}
@@
 Advances in image segmentation of magnetic resonance images (MRI) have
-demonstrated that \multiatlas{} approaches improve segmentation accuracy and
-precision over regular atlas-based approaches. These approaches often rely on
-a large number of such manually segmented atlases (e.g. 30-80) that take
-significant time and expertise to produce. We present an algorithm, \mb{} 
-({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
-T}emplates), for the automatic segmentation of the hippocampus that minimizes
-the number of atlases needed while still achieving similar aggreement to
-\multiatlas{} approaches. Thus, our method acts as an accurate \multiatlas{}
-approach when using special, hard-to-define atlases that are laborious to
-construct. \\
+demonstrated that \multiatlas{} approaches improve segmentation over regular
+atlas-based approaches. These approaches often rely on a large number of such
+manually segmented atlases (e.g. 30-80) that take significant time and
+expertise to produce. We present an algorithm, \mb{} ({\textbf M}ultiple
+{\textbf A}utomatically {\textbf Ge}nerated {\textbf T}emplates), for the
+automatic segmentation of the hippocampus that minimizes the number of atlases
+needed while still achieving similar aggreement to \multiatlas{} approaches.
+Thus, our method acts as an reliable \multiatlas{} approach when using special
+or hard-to-define atlases that are laborious to construct. \\
 
@@
 
-\textbf{Conclusion:} We demonstrate that \mb{} produces accurate whole
+\textbf{Conclusion:} We demonstrate that \mb{} produces consistent whole
 hippocampal segmentations using only 9 atlases, or fewer, with various
@@
 structural magnetic resonance images (MRI) are often used for the volumetric
-assessment of the hippocampus. As such, accurate segmentation of the
-hippocampus and its subfields in MRI is a necessary first step to better
+assessment of the hippocampus. As such, reliable and faithful segmentation of
+the hippocampus and its subfields in MRI is a necessary first step to better
 understand the inter-individual variability of subject neuroanatomy.
 
-The gold standard for neuroanatomical segmentation is manual delineation by an
-expert human rater. However, with the availability of increasingly large MRI
-datasets the time and expertise required for manual segmentation becomes
+The gold standard for neuroanatomical image segmentation is manual delineation
+by an expert human rater. However, with the availability of increasingly large
+MRI datasets the time and expertise required for manual segmentation becomes
 prohibitive \citep{Mazziotta1995,Mazziotta2001,Mazziotta,Pausova2007}. This
@@
 successfully in several contexts
-\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited in
-accuracy due to error in the estimated nonlinear transformation itself, partial
-volume effects in label resampling, and irreconcilable differences between the
+\citep{Chakravarty2008,Chakravarty2009,Collins1995,Haller1997}, it is limited
+by the error in the estimated nonlinear transformation itself, partial volume
+effects in label resampling, and irreconcilable differences between the
 neuroanatomy represented within the atlas and target images.
@@
 More recently, many groups have used multiple atlases to improve overall
-segmentation accuracy (i.e. \multiatlas{} segmentation) over model-based approaches
+segmentation reliability (i.e. \multiatlas{} segmentation) over model-based approaches
 \citep{Heckemann2006,Heckemann2011,Collins2010,Lotjonen2010,Aljabar2009,Leung2010,Wolz2010}.
@@
 neuroanatomy as evaluated by normalized mutual information \citep{Studholme1999}
-from a library of 80 atlases, were required to achieve accurate segmentations of
+from a library of 80 atlases, were required to achieve favourable segmentations of
 the hippocampus. Also, several methods have been explored for label
@@
 subfields on standard T1-weighted images. Our group recently demonstrated that
-through use of an intermediary automated segmentation stage, robust and accurate
+through use of an intermediary automated segmentation stage, robust and reliable
 segmentation of the striatum, pallidum, and thalamus using a single atlas
@@
 temporal lobe samples. Then, using nonlinear registration guided by shape-based
-models of the subfield segmentations, and manually derived hippocampus masks of
+models of the subfield segmentations and manually derived hippocampus masks of
 the target images, the authors demonstrate accurate parcellation of hippocampal
-subfields in clinical 3T T1-weighted MRI volumes. Using \multiatlas{} with bias
-correction techniques, \citet{Yushkevich2010} demonstrate a semi-automated
-method of subfield segmentation on in vivo focal T2-weighted MR acquisitions of
-the temporal lobe. Manual input is only needed to mark divisions between
-the head, body and tail of the hippocampus on target images.
+subfields, with respect to manual segmentations, in clinical 3T T1-weighted MRI
+volumes. Using \multiatlas{} with bias correction techniques,
+\citet{Yushkevich2010} demonstrate a semi-automated method of subfield
+segmentation on in vivo focal T2-weighted MR acquisitions of the temporal lobe.
+Manual input is only needed to mark divisions between the head, body and tail
+of the hippocampus on target images.
 
@@
   \item[Cross-correlation Weighted Majority Vote]
-  An optimal combination of targets from the template library has previously been
-  shown to improve segmentation accuracy \citep{Aljabar2009,Collins2010}. In this
-  method, each template library image is ranked in similarity to each unlabelled 
-  image by the normalized cross-correlation (CC) of image intensities after linear
+  An optimal combination of targets from the template library has previously
+  been shown to improve segmentation accuracy with respect to manual
+  segmentations \citep{Aljabar2009,Collins2010}. In this method, each template
+  library image is ranked in similarity to each unlabelled image by the
+  normalized cross-correlation (CC) of image intensities after linear
   registration, over a region of interest (ROI) generously encompassing the
   hippocampus. Only the top ranked template library image labels are used in a
-  voxel-wise majority vote. The ROI is heuristically defined as the extent of all
-  atlas labels after linear registration to the template, dilated by three voxels
-  \citep{Chakravarty2013}. The number of top ranked template library image labels
-  is a configurable parameter and displayed as the size of the template library
-  in the rest of the paper.
+  voxel-wise majority vote. The ROI is heuristically defined as the extent of
+  all atlas labels after linear registration to the template, dilated by three
+  voxels \citep{Chakravarty2013}. The number of top ranked template library
+  image labels is a configurable parameter and displayed as the size of the
+  template library in the rest of the paper.
 
@@
 \scriptsize
-\caption{\textbf{Automated segmentation accuracy of the Hippocampus} 
-This table shows the best published Dice's overlap measure between automated and
-ground truth segmentations. Unless otherwise specified,
+\caption{\textbf{Summary of automated segmentation methods of the Hippocampus.} 
+This table summarizes published Dice's overlap measure between automated
+and manual segmentations of the hippocampus. Unless otherwise specified,
 validation datasets are composed equally of cases and control subjects, and use
@@
 Alzheimer's Disease; MCI = Mild Cognitive Impairment; CN = Cognitively Normal
-(CN); FEP = First Episode of Psychosis; LOOCV = Leave-one-out cross-validation; MCCV
-= Monte Carlo cross-validation; SNT = Surgical Medtronic Navigation Technologies
-semi-automated labels. Some studies of automated segmentation of ADNI images
-are excluded because they do not not provide overlap measures for the
-hippocampus \citep{Heckemann2011,Chupin2009}.
+(CN); FEP = First Episode of Psychosis; LOOCV = Leave-one-out cross-validation;
+MCCV = Monte Carlo cross-validation; SNT = Surgical Medtronic Navigation
+Technologies semi-automated labels. Some studies of automated segmentation of
+ADNI images are excluded because they do not not provide overlap measures for
+the hippocampus \citep{Heckemann2011,Chupin2009}.
   \label{tab:other-methods}} 
@@
 \mb{} framework -- a methodology that requires very few input atlases in order
-to provide accurate and reliable segmentations. Both \expadnixval{} (Section
-\ref{exp:adni-xval}) and \expfepxval{} (Section \ref{exp:fep-xval}) compare \mb{}
-to basic-\multiatlas{} segmentation by characterising the change in segmentation
-quality with varying parameter settings (atlas and template library sizes,
-registration method, and label fusion method) and differing age and
-neuropsychiatric populations. Together, these experiments allow us to choose
-optimal \mb{} parameter settings for use in subsequent experiments. \expadni{}
-(Section \ref{exp:adni}) demonstrates that across
-\Sexpr{length(means.complete$RID)} images from the \adnidataset{} dataset, \mb{}
-performs as well as, or better, than other established and popular methods, and
-has a much more conservative proportional bias in segmentation volume.  Finally,
-\expsubfield{} (Section \ref{exp:winterburn-xval}) demonstrates the reliability
-of \mb{} in producing subfield segmentations which match the segmentation
-protocol of the input atlases despite contrast and resolution limitations in
-standard T1-weighted image volumes.  All of these experiments together
-demonstrate that \mb's algorithmic performance is not dependent on a single
-definition of the hippocampus but is effective with differing hippocampal
-definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}, across image types,
-and subject populations.
+to provide accurate and reliable segmentations with respect to manual
+segmentations. Both \expadnixval{} (Section \ref{exp:adni-xval}) and
+\expfepxval{} (Section \ref{exp:fep-xval}) compare \mb{} to basic-\multiatlas{}
+segmentation by characterising the change in segmentation quality with varying
+parameter settings (atlas and template library sizes, registration method, and
+label fusion method) and differing age and neuropsychiatric populations.
+Together, these experiments allow us to choose optimal \mb{} parameter settings
+for use in subsequent experiments. \expadni{} (Section \ref{exp:adni})
+demonstrates that across \Sexpr{length(means.complete$RID)} images from the
+\adnidataset{} dataset, \mb{} performs as well as, or better, than other
+established and popular methods, and has a much more conservative proportional
+bias in segmentation volume.  Finally, \expsubfield{} (Section
+\ref{exp:winterburn-xval}) demonstrates the reliability of \mb{} in producing
+subfield segmentations which match the segmentation protocol of the input
+atlases despite contrast and resolution limitations in standard T1-weighted
+image volumes.  All of these experiments together demonstrate that \mb's
+algorithmic performance is not dependent on a single definition of the
+hippocampus but is effective with differing hippocampal definitions
+\citep{Winterburn2013,Pruessner2000,Hsu2002}, across image types, and subject
+populations.
 
@@
 \citep{Aljabar2009,Collins2010} has found that cross-correlation and normalized
-mutual information-based weighted label fusion improves segmentation accuracy
+mutual information-based weighted label fusion improves segmentation reliability
 over simple majority vote label fusion, and yet we did not see a significant
@@
 Further, the sensitivity and specificity parameters can also be tuned to
-potentially improve segmentation accuracy and reliability. It is likely that
+potentially improve segmentation reliability. It is likely that
 using more sophisticated voting methods would have a positive effect on the
@@
 are amongst the highest reported. Thus, more work is required to determine the
-extent to which label fusion will improve the accuracy of our algorithm.
+extent to which label fusion will improve the reliability of our algorithm.
 
@@
 atlases than our implementation with the Winterburn atlases.
-\citet{Lotjonen2010} achieve highly accurate segmentation but correct their
-segmentations using classifications derived using an expectation maximization
-framework. In their initial work, \citet{Chupin2009} develop their probabilistic
-methodology using a cohort of 8 healthy controls and 15 epilepsy patients, and
-then use this method to segment an ADNI sample, with a hierarchical
-experimentation protocol. These methods suggest that some post-processing of the
-final segmentations would improve agreement of the segmentation. While that may
-be true, there is little consensus regarding how to achieve this.
+\citet{Lotjonen2010} produced segmentations which strongly agree with manual
+segmentations by way of post-processing corrections using classifications
+derived using an expectation maximization framework. In their initial work,
+\citet{Chupin2009} develop their probabilistic methodology using a cohort of 8
+healthy controls and 15 epilepsy patients, and then use this method to segment
+an ADNI sample, with a hierarchical experimentation protocol. These methods
+suggest that some post-processing of the final segmentations would improve
+agreement of the segmentation. While that may be true, there is little
+consensus regarding how to achieve this.
 

### 2014-04-09 paper.Rnw
Other reviewer comments from R2

@@
 In this paper we describe a thorough validation of the \mb{} algorithm for the
-fully automatic segmentation of the hippocampus and its subfields. First, we
-address the very idea of boostrapping a template library from a limited number
-of input atlases \citep{Chakravarty2013} for whole hippocampus segmentation by
-conducting a multi-fold validation experiment over a range of atlas and template
-library sizes, registration and label fusion methods. This type of validation is
-done first on a subset of the Alzheimer's Disease Neuroimaging Initiative (ADNI)
-dataset with manual segmentations  \prsnr{}-protocol, and then replicated on a
-first episode psychosis patient dataset to determine the behaviour of \mb{} when
-segmenting younger and differently diseased subjects. Next, we compare \mb{}
-with other popular segmentation algorithms (\freesurfer{}, \fsl{}, \maper{}, and
-\snt{}) on all the images available in the \adnidataset{} sample. Lastly, using
-the optimal parameter settings for \mb{} found from the previous experiments, we
+fully automatic segmentation of the hippocampus and a proof-of-concept
+validation of its application to the segmentation of hippocampal subfields in
+standard T1-weighted images. First, we address the very idea of generating a
+template library from a limited number of input atlases \citep{Chakravarty2013}
+for whole hippocampus segmentation by conducting a multi-fold validation
+experiment over a range of atlas and template library sizes, registration and
+label fusion methods. This type of validation is done first on a subset of the
+Alzheimer's Disease Neuroimaging Initiative (ADNI) dataset with manual
+segmentations  \prsnr{}-protocol, and then replicated on a first episode
+psychosis patient dataset to determine the behaviour of \mb{} when segmenting
+younger and differently diseased subjects. Next, we compare \mb{} with other
+popular segmentation algorithms (\freesurfer{}, \fsl{}, \maper{}, and \snt{})
+on all the images available in the \adnidataset{} sample. Lastly, using the
+optimal parameter settings for \mb{} found from the previous experiments, we
 investigate hippocampal subfield segmentation by conducting a leave-one-out
@@
 
-\mb{} ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
-T}emplates) bootstraps the creation of a large template library given a limited
-input atlas library, and then uses the template library in basic \multiatlas{}
-segmentation. Images for the template library are selected from a set of input
-target images, either arbitrarily or so as to reflect the neuroanatomy or
-demographics of the target set as a whole (for instance, by sampling equally
-from cases and controls). The template library images are automatically
-labelled by each of the atlases via label propagation. Effectively, basic
-\multiatlas{} segmentation is then conducted using the template library to
-segment the entire set of target images (including the target images used in the
-construction of the template library). Since each template library image has
-multiple labels (one from each atlas), the final number of labels to be fused
-for each target may be quite large (i.e. \# of atlas $\times$ \# of templates).
+The \mb{} ({\textbf M}ultiple {\textbf A}utomatically {\textbf Ge}nerated {\textbf
+T}emplates) algorithm creates a large template library given a much smaller
+sized input atlas library and then uses this template library in basic
+\multiatlas{} segmentation to segment a set of input target images. The images
+used in the template library are selected from the input images, either
+arbitrarily or so as to reflect the neuroanatomy or demographics of the target
+set as a whole (for instance, by sampling equally from cases and controls). The
+template library images are automatically labelled by each of the atlases via
+label propagation.  Effectively, basic \multiatlas{} segmentation is then
+conducted using the template library to segment the entire set of target images
+(including the target images used in the construction of the template library).
+Since each template library image has multiple labels (one from each atlas),
+the final number of labels to be fused for each target may be quite large (i.e.
+\# of atlas $\times$ \# of templates).
 
@@
 
-In this experiment we explore the very idea of bootstrapping a template library
+In this experiment we explore the very idea of generating a template library
 for multi-atlas-based segmentation from a small number of input atlases. To do
@@
 The previous experiment assesses \mb{} performance on whole hippocampus
-segmentation. In this experiment, we evaluate \mb{} hippocampal subfield
-segmentation of standard 3T T1-weighted images at $0.9mm$-isotropic voxels. We
-use a modified leave-one-out cross-validation (LOOCV) design.
+segmentation. In this experiment, we conduct a proof-of-concept evaluatation of
+\mb{} hippocampal subfield segmentation of standard 3T T1-weighted images at
+$0.9mm$-isotropic voxels. We use a modified leave-one-out cross-validation
+(LOOCV) design.
 
@@
 ill-defined since there are no manual protocols for segmentation at this
-resolution. Instead, we must evaluate the reliability, or {\it precision}, with
-which \mb{} produces hippocampal subfields segmentations at this resolution that
-correspond in form to the segmentation protocol used by the given
-high-resolution atlas library images.
-
-By directly resampling the Winterburn atlas segmentations to $0.9mm^3$ voxels
-(using standard nearest-neighbour image resampling techniques) we obtain a
-subsampled version of the labels which preserve the original segmentation
-protocol within the limits of error from rounding and interpolation. Therefore,
-using the resampled Winterburn segmentations as definitive for the $0.9mm^3$
-resolution we evaluate reliability of \mb{} segmentations using DSC overlap
-scores and evaluate consistency across the range of hippocampal sizes using
-Bland-Altman plots of subfield volumes.
+resolution. Instead, we must evaluate how well the lower-resolution \mb{}
+hippocampal subfield segmentations correspond in form to the segmentation
+protocol used in the high-resolution images. By directly resampling the
+Winterburn atlas segmentations to $0.9mm^3$ voxels (using standard
+nearest-neighbour image resampling techniques) we obtain a subsampled version
+of the labels which preserve the original segmentation protocol within the
+limits of error from rounding and interpolation. Therefore, using the resampled
+Winterburn segmentations as definitive for the $0.9mm^3$ resolution we evaluate
+agreement of \mb{} segmentations using DSC overlap scores and evaluate
+consistency across the range of hippocampal sizes using Bland-Altman plots of
+subfield volumes.
 
@@
 bias in segmentation volume.  Finally, \expsubfield{} (Section
-\ref{exp:winterburn-xval}) demonstrates the reliability of \mb{} in producing
-subfield segmentations which match the segmentation protocol of the input
-atlases despite contrast and resolution limitations in standard T1-weighted
-image volumes.  All of these experiments together demonstrate that \mb's
-algorithmic performance is not dependent on a single definition of the
-hippocampus but is effective with differing hippocampal definitions
-\citep{Winterburn2013,Pruessner2000,Hsu2002}, across image types, and subject
-populations.
-
-The core claim the \mb{} method is based on -- that we can meaningfully bootstrap
-a template library from a small set of labelled atlas images -- is validated in
-the cross validation conducted in \expadnixval{} (and the replication in
-\expfepxval{} and \expadnixvalsnt{}, Supplementary Materials). We find that both
-increasing the number of atlases and the number of templates used improves \mb{}
-segmentation over and above basic-\multiatlas{} segmentations using the same
-number of atlas images. That is, by taking the extra step of generating a
+\ref{exp:winterburn-xval}) is a proof-of-concept validation demonstrating the
+reliability of \mb{} in producing subfield segmentations which match the
+segmentation protocol of the input atlases despite contrast and resolution
+limitations in standard T1-weighted image volumes.  All of these experiments
+together demonstrate that \mb's algorithmic performance is not dependent on a
+single definition of the hippocampus but is effective with differing
+hippocampal definitions \citep{Winterburn2013,Pruessner2000,Hsu2002}, across
+image types, and subject populations.
+
+The core claim the \mb{} method is based on -- that a useful template library
+can be generated from a small set of labelled atlas images -- is validated in
+the cross-validation conducted in \expadnixval{} (and the replication in
+\expfepxval{} and \expadnixvalsnt{}, Supplementary Materials).  We find that
+both increasing the number of atlases and the number of templates used improves
+\mb{} segmentation over and above basic-\multiatlas{} segmentations using the
+same number of atlas images. That is, by taking the extra step of generating a
 template library using target images, \mb{} is able to improve the overlap
@@
 standard'' segmentations. The magnitude of this improvement is greatest with a
-small number of atlases, but even with larger atlas libraries we have found that
-generating a template library reduces the variability in segmentation precision
-(i.e. \mb{} more consistently produces high quality segmentations than does
-basic-\multiatlas{} segmentation). These effects do not appear dependant on the
+small number of atlases, but even with larger atlas libraries we have found
+that generating a template library reduces the variability in segmentation
+agreement (i.e. \mb{} more consistently produces segmentations in greater
+agreement with manual segmentations than does basic-\multiatlas{} method, over
+repeated randomized trials). These effects do not appear dependant on the
 hippocampal segmentation protocol used.
@@
 performance of all methods. With the exception of \fsl{} all methods correlate
-well with the \snt{} volumes provided in the ADNI database. However, \freesurfer{}
-and \fsl{} provide radically different definitions of the size of the hippocampus
-in comparison to the other methods. Further, when estimating bias of these
-methods relative to \snt{} hippocampal volumes we see that large hippocampi are
-over estimated while small hippocampi are under estimated. By comparison, \mb{}
-and \maper{} are far more conservative in volume estimation, suggesting these
-methods may be better suited for estimating true-positives, especially in
-neurodegenerative disease subjects featuring smaller overall hippocampi.
-However, in this analysis we have only compared methods by total hippocampal
-volume, and so more work is needed to understand the full extent to which these
-methods differ.
+well with the semi-automated \snt{} volumes provided in the ADNI database.
+However, the \freesurfer{} and \fsl{} hippocampal segmentations are on average
+about twice the volume of those from all other methods. Furthermore, when
+estimating the bias of \freesurfer{} and \fsl{} relative to the \snt{}
+hippocampal volumes we see that large hippocampi are over estimated while small
+hippocampi are under estimated. By comparison, \mb{} and \maper{} are far more
+conservative in volume estimation, suggesting these methods may be better
+suited for estimating true-positives, especially in neurodegenerative disease
+subjects featuring smaller overall hippocampi.  However, in this analysis we
+have only compared methods by total hippocampal volume, and so more work is
+needed to understand the full extent to which these methods differ.
 
@@
 In conclusion, we have demonstrated the viability of leveraging a small
-number of input atlases to bootstrap a large template library and thereby
+number of input atlases to generate a large template library and thereby
 improve segmentation reliability when using \multiatlas{} methods. We
